### 2020-10-15 note/SIR_notes.Rnw
Begin

\documentclass{article}
\usepackage{natbib}
\usepackage{hyperref}
\bibliographystyle{apalike}

\title{SIR Model with Testing and Isolation Mechanisms}

\begin{document}
\maketitle

\section{Method}

\citep{arino2020simple} developed a SLIAR compartmental model to study the spread of an epidemic, specifically COVID-19, in a population. The model incorporates an Erlang distribution of times of sojourn inincubating, symptomatically and asymptomatically infectious compartments. Basic reproduction number is derived. Also, sensitivity analysis with respect to the underlying parameters for the following two outputs was carried out; (i) the number of observable cases during the course of the epidemic and at the peak, and (ii) the timing of the peak of the outbreak. Sensitivity analysis is performed using the R package multisensi.

\section{Literature Review}
\cite{bergstrom2020frequency}
(1) Model, assumptions: They developed a function, namely expected expoisour $E(C,\tau)$, to approximate trade-offs between the frequency of testing, n, the sensitivity of testing, q, and the delay between
testing and results, d. This function is explicitly derived and was connected the effective reproduction number $R=R_0 S$, where $S$ is the proportion of population susciptable.
assumption that transmission rates are a step function: individuals who
have COVID go from non-infectious to fully infectious instantaneously,
and remain fully infectious until they are no longer able to transmit disease. Test sensitivity takes the same form over the course of infection.
More sophisticated models could allow varying infectiousness and varying
sensitivity over time, as in 
\citep{larremore2020test}.

\bibliography{../SIRlibrary}
\end{document}

### 2020-10-15 codes/ali_test_sir.Rmd
initial code

---
title: "SIR Model With Testing Flows and Isolation"
author: "Ali"
date: "15/08/2020"
output: html_document
# output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Notes on other references: (1) [testing_flow.md](testing_flow.md)(I am using the notation in this doc)

# On Math and Weighting

## Model and Definitions

- Assume there are 3 groups of individuals: $S$, $I$ and $R$.
- Each group splits into 4 subgroups to incorporate the testing mechanism, so overall the model framework has 12 compartments as follows. $S_u,$ $S_n$, $S_p$, $S_t$, $I_u$, $I_n$, $I_p$, $I_t$, $R_u$, $R_n$, $R_p$, $R_t$.
- Further, assume that (1) a perfectly specific test, i.e. no false positives, then $S_p$ and $S_t$ are always zero (2) there are two 'integrator' or 'accumulator' compartments, $N$ and $P$, to collect cumulative reported negative or positive tests.

## Per Capita Rate of Flow from Untested to Positive or Negative Compartments

- $t$: testing intensity with dimension of $1/time$.
- $\omega$: the rate of *onward flow* from the awaiting positive compartment, $p$, to reported/tested compartment, $t$, or from awaiting negative compartment, $n$, back to $u$.  It has units of $1/time$.

- Note: The complicated part is the flow rate from $Z_u$ to $Z_n$, $Z_p$.  The (total, not per capita) flow from $Z_u$ to $Z_n$ is $t (1-P_Z) F(W_Z) Z_u$, from $Z_u$ to $Z_p$ is $t P_Z F(W_Z) Z_u$. Where, $P_Z$ is the probability of positive test.

- At the simplest $F(W_Z)=W_Z$.  The flows from $Z_n$ to $Z_u$ and $Z_p$ to $Z_t$ are $\omega Z_n$ and $\omega Z_p$ respectively. We are using a dimensionless scaling where $F(W_Z) = (W_Z \cdot N_0)/\sum_{Z'} \left( W_{Z'} {Z'}_{u} \right)$, i.e. the denominator is the sum of the product of the weights and the current occupancy of the untested compartments.

## Model

![](../inst/pix/sir_test1.jpg)

Let the force of infection $\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_t I_t)}{N_0}$, where $\eta$ is the isolation parameter with the following assumptions:

(1) All susceptible individuals are exposed to all infectious individuals regardless of the group factor. 
(2) $\eta_t<\eta_w$, i.e., the awaiting individuals for test results have higher transmission probability than the reported individuals. 
(3) The scaling parameter is defined as $sc = t N_0/(W_s S_u + W_i I_u + W_r R_u)$ and $F_s = sc \times W_s$, $F_i=sc \times W_i$ and $F_r = sc\times W_r$.

The model is
\begin{align}
\label{mod:1}
1)\ d S_u/dt &= -\Lambda S_u - F_s S_u + \omega S_n \\
2)\ d S_n/dt &= -\Lambda S_n + F_s S_u - \omega S_n \\
3)\ d I_u/dt &= \Lambda S_u - F_i I_u + \omega I_n  - \gamma I_u  \\
4)\ d I_n/dt &= \Lambda S_n + (1-P_i) F_i I_u - \omega I_n -\gamma I_n \\
5)\ d I_p/dt &= P_i F_i I_u - \omega I_p -\gamma I_p \\
6)\ d I_t/dt &= \omega I_p - \gamma I_t  \\
7)\ d R_u/dt &= \gamma I_u - F_r R_u + \omega R_n \\
8)\ d R_n/dt &= \gamma I_n + (1-P_r) F_r R_u - \omega R_n  \\
9)\ d R_p/dt &= \gamma I_p + P_r F_r R_u  - \omega R_p  \\
10)\ d R_t/dt&= \gamma I_t + \omega R_p  \\
11)\ dN/dt &= \omega (S_n + I_n + R_n)   \\
12)\ dP/dt &= \omega(I_p + R_p).
\end{align}

```{r packages, message=FALSE}
library(deSolve)
library(ggplot2)
library(tidyr)
library(McMasterPandemic)
library(directlabels)
unpack <- McMasterPandemic::unpack

```

```{r, model formulation}
sir.model <- function(time,state,params){
    unpack(as.list(c(state,params)))
    ## Force of Infection  
    Lambda <- beta * (I_u + eta_w*(I_n+I_p) + eta_t*I_t)/N0
    ## scaling the weights
    sc <- t*N0/(W_s*S_u+W_i*I_u+W_r*R_u)
    # sc <- t*N0
    F_s <- sc*W_s
    F_i <- sc*W_i
    F_r <- sc*W_r

    dS_u.dt <- -Lambda*S_u - (1-P_s) * F_s * S_u + omega * S_n
    dS_n.dt <- -Lambda*S_n + (1-P_s) * F_s * S_u - omega * S_n 
    dI_u.dt <-  Lambda*S_u + omega * I_n - F_i * I_u - gamma * I_u
    dI_n.dt <-  Lambda*S_n + (1-P_i) * F_i * I_u  - omega * I_n - gamma * I_n
    dI_p.dt <-  P_i * F_i * I_u - omega * I_p - gamma * I_p 
    dI_t.dt <-  omega * I_p - gamma * I_t 
    dR_u.dt <- gamma * I_u + omega * R_n - F_r * R_u  
    dR_n.dt <- gamma * I_n + (1-P_r) * F_r * R_u - omega * R_n
    dR_p.dt <- gamma * I_p + P_r * F_r * R_u - omega * R_p
    dR_t.dt <- gamma * I_t + omega * R_p
    dN.dt <- omega * (S_n + I_n + R_n)
    dP.dt <- omega *(I_p + R_p)

    # return the rate of change
  dxdt <- c(dS_u.dt,dS_n.dt,dI_u.dt,dI_n.dt,dI_p.dt,dI_t.dt,dR_u.dt,dR_n.dt,dR_p.dt,dR_t.dt,dN.dt,dP.dt)
    ## }
    return(list(dxdt))
}

```

```{r model inputs, message=FALSE}
params <- c(N0=1000000, beta=1,gamma=1/3, omega=0.25, t=0.01, 
             W_s=0.01, W_i=1, W_r=0.01, #weights
             P_s=0, P_i=0.5, P_r=0.5, #prob of waiting for being tested positive
            eta_w=0.02, eta_t=0.01 #isolation parameter   
            )
class(params) <- "params_pansim"
state_init <- c(S_u=params[["N0"]], S_n=0,
                I_u=1,I_n=0,I_p=0,I_t=0,
                R_u=0,R_n=0,R_p=0,R_t=0,
                N=0,P=0)
# Disease free equilibrium
Sn_dfe <-function(params){
  unpack(as.list(params))
  return((t*(1-P_s)*N0)/omega)
}
Su_dfe<- params[["N0"]]-Sn_dfe(params)

state_dfe <- c(S_u=Su_dfe, S_n=Sn_dfe(params),
                I_u=0,I_n=0,I_p=0,I_t=0,
                R_u=0,R_n=0,R_p=0,R_t=0,
                N=0,P=0)

d <- seq(0,200,by=0.1)
# Conditional stop for desolver
rootfun <- function (time, state,params) { 
  unpack(as.list(c(state,params)))
  return(I_u - 1) }
```

Test gradient
```{r}
g1 <- sir.model(time=0,state=state_init, params=params)
all.equal(sum(g1[[1]]),0)
## debug(sir.model)
## g1 <- sir.model(time=0,state=state_init, params=params)
```

```{r model_run, message=FALSE}
out <- as.data.frame(
  ode(
    func=sir.model,
    y=state_init,
    times= d, 
    parms=update(params,beta=2),
    # atol = 1e+1, rtol = 1e+1
    # rootfun = rootfun,
    # method="lsodar"
  )
)

out2 <- out %>%
  pivot_longer(c(S_u,S_n,I_u,I_n,I_p,I_t, R_u,R_n,R_p,R_t), names_to = "compartment", values_to = "value")
```

```{r plot_run1, warning=FALSE}
gg1 <- (ggplot(data=out2, aes(x=time,y=value,col=compartment))
    + geom_line()
    + scale_y_log10(limits=c(0.1,params[["N0"]]))
##    + scale_x_continuous(limits=c(0,240))
)
direct.label(gg1,"last.bumpup")
```

```{r run2, message=FALSE}
d_long <- seq(0,300,by=0.1)
## capturing badly behaved integrator warning messages
cc <- capture.output(
    out_long <- as.data.frame(
        ode(
    func=sir.model,
    y=state_init,
    times= d_long, 
    parms=update(params,beta=2)
    # rootfun = rootfun,
    # method="lsodar"
  )
  ))
out_long2 <- out_long %>%
    pivot_longer(c(S_u,S_n,I_u,I_n,I_p,I_t, R_u,R_n,R_p,R_t),
                 names_to = "compartment", values_to = "value")

```

```{r plot_long, warning=FALSE}
direct.label(gg1 %+% out_long2,"last.bumpup")
```

### Steady States

The Disease-Free Equilibrium, namely $E_1$, for the above SIR model is
\label{exp:dfe} $S_n^*= \frac{t}{\omega} N_0, S_u^*= N_0-S_n^*$, and $I_j=R_j=0 \ \forall j$.

### Basic Reproduction Number

The next generation matrix, $G = F V^{-1}$, is in the lower triangular form with $F$ and $G$ as follows.

$$R_0= (A \times S_u^* + B \times S_n^*) \times C,$$ where
$A=\gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_t P_i) F_i$,
$B=\big(\omega+(F_i+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_t\omega) \omega P_i F_i }{\omega+\gamma}$ and 
$C=\frac{\beta/\gamma}{N_0 (\gamma(\omega+\gamma)+F_i(\gamma+\omega P_i))}$.

\begin{equation}\label{FV}
F= \beta/N_0 \left[ \begin {array}{cccc} 
S_u&\eta_w\,S_u&\eta_w\,S_u&\eta_t\,S_u\\
S_n&\eta_w\,S_n&\eta_w\,S_n&\eta_t\,S_n\\ 
0&0&0&0\\
0&0&0&0
 \end {array} \right],
 V=
 \left[ \begin {array}{cccc}  
F_i+\gamma&-\omega&0&0\\
-(1-P_i)F_i&\omega+\gamma&0&0\\
-P_i F_i&0&\omega+\gamma&0\\
0&0&-\omega&\gamma
\end {array} \right],
\end{equation}

\begin{equation}
G = \left[ \begin {array}{cc}
G_{11}&G_{12}\\
0&0
\end {array} \right], \text{ where } \\
G_{11} =C
\left[\begin {array}{cc}
A\,S_u & B\,S_u\\
A\,S_n & B\,S_n
\end {array}\right],
\end{equation}

The Eigenvalues of $G_{11}$, are 0 and $R_0$.
```{r R0}
F_i <- function(state,params){
   unpack(as.list(c(state,params)))
   return(t*N0*W_i/(W_s*S_u+W_i*I_u+W_r*R_u))
 }
 # just a test
  F_i(state=state_dfe,params=update(params,t=0))

R0<-function(params){
    unpack(as.list(params))
    Sn_dfe<- Sn_dfe(params)
    Su_dfe<- N0-Sn_dfe
    F_i <-t*N0*W_i/(W_s*Su_dfe) #at DFE
    # F_i <-t*N0*W_i
    A <-gamma*(omega+gamma+eta_w*F_i) + omega*eta_t*P_i*F_i
    B <-(omega+eta_w*(F_i+gamma))*gamma+(eta_w*gamma+eta_t*omega)*(omega*P_i*F_i)/(omega+gamma)

    C <- beta/(N0*gamma*(gamma*(omega+gamma)+F_i*(gamma+omega*P_i)))
    return((A*Su_dfe+B*Sn_dfe)*C)
}

# R0<1:
R0(params=params)
# R0>1:
R0(params=update(params,beta=2))
```

Can we recover $R_0=\beta/\gamma$ of the simple SIR? -yes
```{r checkings}
# checking if simple R_0 for SIR can be recovered?
# Case #1: t=0
all.equal(R0(params=update(params,t=0)), params[["beta"]]/params[["gamma"]])
# Case #2: 
all.equal(R0(params=update(params,eta_t=1,eta_w=1)), params[["beta"]]/params[["gamma"]])
```

### On Behaviour of Testing at DFE Boundry
(1) Note that once the only untested people are susceptibles, the FoI will be $\Lambda=0$, testing rate $F_s=t N_0/S_u$. Thus, eq(1) of the model will be $d S_u/dt = - t N_0 + \omega S_n$ which is a linear rate of leaving the $S_u$ compartment. That means that the testing rate will be no longer dependent on $S_u$.

### Contour plot of $R_0=1$ and interpretations 
```{r contour_R0, message=FALSE,warning=FALSE }
 # specify the ranges
 beta  <- 0.5  ## set beta high enough to allow R0>1 in worst case
 eta_w <- seq(0,1, length.out=5) 
 eta_t <- seq(0,1, length.out=5) 
 t     <- seq(0,0.01, length.out=10)
 omega <- seq(0.1,1 , length.out=10) #note omega must be > t

 df1 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,t=t,
                    W_s=params[["W_s"]],W_i=params[["W_i"]],W_r=params[["W_r"]],
                    P_s=params[["P_s"]],P_i=params[["P_i"]],P_r=params[["P_r"]],
                    eta_w=eta_w,eta_t=eta_t)

df2<- data.frame(df1,R0=apply(df1,1,function(params_in)R0(params=params_in)),
                      Fi= apply(df1,1,function(params_in)F_i(state=state_dfe,params=params_in)))

which(df2[]<0)
```

```{r}
## hacked (not very robustly) to chain label_both() and label_parsed() ...
label_special <- function (labels, multi_line=FALSE, sep = "== ") {
    value <- ggplot2:::label_value(labels, multi_line = multi_line)
    variable <- ggplot2:::label_variable(labels, multi_line = multi_line)
    variable[[1]] <- gsub("_(.)$","[\\1]",variable[[1]])
    ## not using multiple faceting variables on each margin, so forget paste
    out <- Map(paste, variable, value, sep = sep)
    out[[1]] <- lapply(out[[1]], function(expr) c(parse(text=expr)))
    out
}
```

```{r show_contours, width=24,height=24,message=FALSE,warning=FALSE}
p1 <- ggplot(df2,aes(x=omega,y=t,z=R0))+ theme_bw() +
    theme(panel.spacing=grid::unit(0,"lines"))
(p1
    + geom_contour_filled(breaks=pretty(unique(df2$R0),5),
                          alpha=0.5) ## breaks=0.2)
    + geom_contour(breaks=1,alpha=0.5,colour="black")
    + facet_grid(eta_w~eta_t, labeller=label_special)
    + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_fill_viridis_d(name=parse(text="R[0]"))
)
```

Let call the $F_i$ the critical testing rate in the above plot.

**Propositions**

1. $\partial{R_0}/\partial{\eta_t} \geq 0$ and $\partial{R_0}/\partial{\eta_w} \geq 0$.
2. $\partial{R_0}/\partial{t} \leq 0$; this is concluded from the Taylor expansion of $R_0$ around $t=0$ which is
$R_0 \approx \beta/\gamma + \frac{\beta t}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta_w-1)(\gamma W_s+\omega W_i) + (\eta_t -1)P_iW_i \omega^2 \Big) + \mathcal{O}(t^2)$.
3. Based on the above Taylor expansion around $t=0$,
$$\partial{R_0}/\partial{\omega}=  \frac{-\beta t}{\gamma W_s\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c), $$
where $a=(\eta_w-1)W_i-(\eta_t-1)P_iW_i$, $b=2(\eta_w-1)\gamma W_s$ and $c=(\eta_w-1)\gamma^2 W_s$.
Given that $\eta_t\leq \eta_w$, it is inferred that $a\geq 0$, $b\leq 0$ and $c \leq 0$.
Let $\omega_1<0$ and $\omega_2>0$ be the roots of the quadratic expression in $\partial{R_0}/\partial{\omega}$. Thus, $\partial{R_0}/\partial{\omega}>0$ for $0<\omega<\omega_2$ and  $\partial{R_0}/\partial{\omega}<0$ for $\omega>\omega_2$.

**Interpretation of the R0 contour plot**

Note that $\eta_t\leq \eta_w$, i.e., reported people have a lower transmission probability than awaiting individuals. The perfect isolation/quarantine occurs when $\eta_w=\eta_t=0$ (see the upper left corner panel). Thus, we only consider the lower-triangle panels in the plot [not sure how to extract them in ggplot?].
(1) Looking at the first column of panels, there seems to be a compensatory relation between $t$ and $\omega$. i.e., higher $t$ requires smaller $\omega$ to keep R0 lower than 1. Also, for a fixed $\omega$, increasing $\eta_w$ requires higher $t$ to achive the critical R0.

### 2020-10-19 codes/ali_test_sir.Rmd
pix was added
adding the parametrization of s

@@
 # On Math and Weighting
+
+## Objectives
+
+1- What combination of testing intensity, $t$, and isolation, $\eta_w$ and $\eta_t$, will result in $R_0 <1$?
 
@@
 
-![](../inst/pix/sir_test1.jpg)
+![](../pix/sir_test1.jpg)
 
@@
              P_s=0, P_i=0.5, P_r=0.5, #prob of waiting for being tested positive
-            eta_w=0.02, eta_t=0.01 #isolation parameter   
+            eta_w=0.02, eta_t=0.01, #isolation parameter 
+            s=2 #eta_w/eta_t
             )
@@
 
+# R0 expression with eta_w/eta_t factored
+R02<-function(params){
+    unpack(as.list(params))
+    Sn_dfe<- Sn_dfe(params)
+    Su_dfe<- N0-Sn_dfe
+    F_i <-t*N0*W_i/(W_s*Su_dfe) #at DFE
+
+    A2 <- gamma*(P_i*F_i*omega*Sn_dfe+(omega+gamma)*(gamma*Sn_dfe+F_i*N0))
+    B2 <- P_i*F_i*omega*(gamma*Su_dfe+omega*N0)
+    D <- gamma*(omega+gamma)*(gamma*Su_dfe+omega*N0)
+    C2 <- beta/((N0*gamma*(gamma*(omega+gamma)+F_i*(gamma+omega*P_i)))*(omega+gamma))
+    # s <- eta_w/eta_t
+
+    # return((eta_w*A2+eta_t*B2+D)*C2)
+    return((eta_t*(s*A2+B2)+D)*C2)
+}
+
 # R0<1:
@@
 R0(params=update(params,beta=2))
+
+# check
+all.equal(R0(params=params),R02(params=params))
 ```
@@
 
-### Contour plot of $R_0=1$ and interpretations 
-```{r contour_R0, message=FALSE,warning=FALSE }
+### Contour plot of $R_0=1$ and interpretations approache 1
+```{r contour_R0_1, message=FALSE,warning=FALSE }
  # specify the ranges
@@
  eta_w <- seq(0,1, length.out=5)
- eta_t <- seq(0,1, length.out=5) 
+ eta_t <- seq(0.001,0.5, length.out=5) 
  t     <- seq(0,0.01, length.out=10)
  omega <- seq(0.1,1 , length.out=10) #note omega must be > t
+# s <- seq(1,2.5,length.out=10)
 
@@
                     P_s=params[["P_s"]],P_i=params[["P_i"]],P_r=params[["P_r"]],
-                    eta_w=eta_w,eta_t=eta_t)
+                    eta_w=eta_w,eta_t=eta_t,s=params[["s"]])
 
@@
 
-which(df2[]<0)
+# which(df2[]<0)
 ```
@@
 
-```{r show_contours, width=24,height=24,message=FALSE,warning=FALSE}
+```{r show_contours_1, width=24,height=24,message=FALSE,warning=FALSE}
 p1 <- ggplot(df2,aes(x=omega,y=t,z=R0))+ theme_bw() +
@@
     + facet_grid(eta_w~eta_t, labeller=label_special)
+    + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
+    + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
+    + scale_fill_viridis_d(name=parse(text="R[0]"))
+)
+```
+
+### Contour plot of $R_0=1$ and interpretations approache 2, parametrization $s=\eta_w/\eta_t$
+
+```{r contour_R0_2, message=FALSE,warning=FALSE }
+ # specify the ranges
+ beta  <- 0.5  ## set beta high enough to allow R0>1 in worst case
+ # eta_w <- seq(0,1, length.out=5) 
+ eta_t <- seq(0.001,0.5, length.out=5) 
+ t     <- seq(0,0.01, length.out=10)
+ omega <- seq(0.1,1 , length.out=10) #note omega must be > t
+s <- seq(1,2.5,length.out=10)
+
+ df11 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,t=t,
+                    W_s=params[["W_s"]],W_i=params[["W_i"]],W_r=params[["W_r"]],
+                    P_s=params[["P_s"]],P_i=params[["P_i"]],P_r=params[["P_r"]],
+                    eta_w=params[["eta_w"]],eta_t=eta_t,s=s)
+
+df22<- data.frame(df11,R0=apply(df11,1,function(params_in)R02(params=params_in)),
+                      Fi= apply(df1,1,function(params_in)F_i(state=state_dfe,params=params_in)))
+
+```
+
+```{r show_contours_2, width=24,height=24,message=FALSE,warning=FALSE}
+p11 <- ggplot(df22,aes(x=omega,y=t,z=R0))+ theme_bw() +
+    theme(panel.spacing=grid::unit(0,"lines"))
+(p11
+    + geom_contour_filled(breaks=pretty(unique(df22$R0),5),
+                          alpha=0.5) ## breaks=0.2)
+    + geom_contour(breaks=1,alpha=0.5,colour="black")
+    + facet_grid(s~eta_t, labeller=label_special)
     + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
@@
 i.e., higher $t$ requires smaller $\omega$ to keep R0 lower than 1.
-Also, for a fixed $\omega$, increasing $\eta_w$ requires higher $t$ to achive the critical R0.
+Also, for a fixed $\omega$, increasing $\eta_w$ requires higher $t$ to achieve the critical R0.
+(2) A sensible result that we are expecting is $\partial{R_0}/\partial{\omega}<0$.

### 2020-10-21 codes/ali_test_sir.Rmd
P_i changed to 1

@@
              W_s=0.01, W_i=1, W_r=0.01, #weights
-             P_s=0, P_i=0.5, P_r=0.5, #prob of waiting for being tested positive
+             P_s=0, P_i=1, P_r=0.5, #prob of waiting for being tested positive
             eta_w=0.02, eta_t=0.01, #isolation parameter 

### 2020-10-21 note/R0_wrt_omega.tex
Add files via upload

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{xcolor}

\newcommand{\fady}[1]{\textcolor{cyan}{$\langle${\slshape{\bfseries Fady:} #1 }$\rangle$}}
\newcommand{\Rnum}{\mathcal{R}_0}
\setlength{\parindent}{0pt} 
\usepackage{amsmath}
\usepackage{xspace}
\usepackage[nameinlink,capitalize]{cleveref}
\crefname{appendix}{Appendix}{Appendices}
\Crefname{appendix}{Appendix}{Appendices}
\title{SIRtesting}
\setlength{\parskip}{1em}
\begin{document}

\maketitle

\section{Intrdoctuion}

Much of this document concerns the effect of $\omega$ (the rate at which test results are returned) on $\Rnum$. As shown by Ali, when the overall testing rate $t$ is small (i.e., $t \approx 0$), the sign of $\partial{\Rnum}/\partial{\omega}$ is determined by a quadratic.

For a relatively clean description of these roots, see starting from \cref{eq:defQ}. We show that there is always two real roots; one negative and one positive. For $w \in (0, \infty)$, there is a global maximum at the second root. For a necessary and sufficient condition (inequality) for having $\partial{\Rnum}/\partial{\omega} < 0$, see \cref{eq:necsuf}. Finally, we give a brief discussion on finding a local minimum of $\Rnum$ with respect to $w$ for $w \in (0,\infty)$. However, our analysis of this is complicated by the fact that we are analyzing the linearization of $\Rnum$ near $t = 0$, rather than the exact form of $\Rnum$. \fady{In section 3, I expand a bit on this and basically ask if it would be a good idea for me to try working with the exact form of $\Rnum$ to see if we can generalize some of our analysis for $t$ not close to 0.}

\section{Behaviour of $\Rnum$ for $t \approx 0$}

From Ali's analysis we know:

the Taylor expansion of $\Rnum$ around $t=0$ is
\begin{equation}
\Rnum \approx \beta/\gamma + \frac{\beta t}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta_w-1)(\gamma W_s+\omega W_i) + (\eta_t -1)P_iW_i \omega^2 \Big) + \mathcal{O}(t^2). 
\end{equation}

Based on the above Taylor expansion around $t=0$, $$\partial{\Rnum}/\partial{\omega}=  \frac{-\beta t}{\gamma W_s\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c), $$

where $a=(\eta_w-1)W_i-(\eta_t-1)P_iW_i$, $b=2(\eta_w-1)\gamma W_s$ and $c=(\eta_w-1)\gamma^2 W_s$. Given that $\eta_t\leq \eta_w$, it is inferred that $a\geq 0$, $b\leq 0$ and $c \leq 0$. Let $\omega_1<0$ and $\omega_2>0$ be the roots of the quadratic expression in $\partial{\Rnum}/\partial{\omega}$. Thus, $\partial{\Rnum}/\partial{\omega}>0$ for $0<\omega<\omega_2$ and $\partial{\Rnum}/\partial{\omega}<0$ for $\omega>\omega_2$.

\hrulefill

Defining the following quantity, $Q$, will help us write the roots of $\partial{\Rnum}/\partial{\omega}$ neatly. 
\begin{align}\label{eq:defQ}
    Q =& \frac{W_i}{W_s}\left(1-\frac{n_{t}-1}{n_{w}-1}P_{i}\right) \\
\end{align}

Since $n_w > n_t$ and $P_i \approx 1$, we know that $Q < 0$.

With that in mind, we can write the roots of $\partial{\Rnum}/\partial{\omega}$ as

\begin{align}
    \omega_1 =& \frac{\gamma}{-\sqrt{1-Q}-1} \\
    \omega_2 =& \frac{\gamma}{\sqrt{1-Q}-1}
\end{align}

It is easy to confirm that $\omega_1 < 0$ by looking at the denominator. To see that $\omega_2 > 0$, recall that $Q < 0$, so $\sqrt{1-Q} > 1$ and so $\sqrt{1-Q} -1 > 0$. Knowing that $\omega_1 < 0$, our only root of interest (i.e., biologically relevant) is $\omega_2$.

With some analysis, we can prove that $\partial{\Rnum}/\partial{\omega} > 0$ when $\omega \in (0,\omega_2)$ and $\partial{\Rnum}/\partial{\omega} < 0$ when $\omega \in (\omega_2,\infty)$. So we know that $\Rnum$ has a global maximum with respect to $\omega$ at $\omega = \omega_2$.

Before we look for strategies to minimize $\Rnum$ wrt to $\omega$, let us first discuss the behaviour of $\omega_2$ first. This may give us biological insight into when it is beneficial (i.e., lowers $\Rnum$) to return test results faster, and when it is not. \textbf{It is extremely important to note that all of this analysis is done with the linearization of $\Rnum$ around $t = 0$ as shown above, thus, this analysis holds only for small $t$.}

Suppose that $\omega > \omega_2$. By the previous analysis, this is equivalent to saying that increases in $\omega$ cause decreases in $\Rnum$. For what parameter regions is this true? To understand this, we examine the inequality

\begin{align}\label{eq:necsuf}
    &\omega > \omega_2 \nonumber \\
    &\omega > \frac{\gamma}{\sqrt{1-Q}-1} \nonumber \\
    &\vdots \nonumber \\
    &\frac{1-n_{t}}{1-n_{w}}P_{i}>\frac{W_{s}}{W_{i}}\left(\frac{\gamma}{\omega}+1\right)^{2}
\end{align}

The previous inequality is a necessary and sufficient condition for $\Rnum$ to be \emph{decreasing} with respect to $\omega$. In other words, it tells us exactly under what circumstances tests should be returned more rapidly and when they should not be. We can give a biological interpretation as follows:

The left hand side is a measure of the relative lowering in contact for the groups whose tests results are confirmed to be positive, compared to those whose test results are pending. As the difference between these two becomes extreme (i.e., as $\eta_t \ll \eta_{w}$), the LHS becomes larger, which means that returning test results faster may become beneficial. The whole left term is scaled by $P_i$, which is the sensitivity of the test. All of this makes sense with our biological intuition; it is useful to return test results quickly when the test is sensitive, and the benefit to positive individuals finding out they are positive outweighs the risk of negative individuals increasing their contact rate upon finding out they do not have the virus.

\fady{The RHS doesn't have as clean an interpretation, I think. But I'll add in more stuff here later.}

Now we focus our attention on minimizing $\Rnum$ wrt to $\omega$. Since $\Rnum(\omega)$ has a global (for $\omega \in [0,\infty)$) maximum at $\omega_2$, we know that $\Rnum$ attains its smallest values either near $\omega = 0$ or as $\omega$ approaches $+\infty$. We can study the behaviour of $\Rnum(\omega)$ on $(0,\infty)$, but in reality, there must be a bounded range of realistic values of $\omega$ that we should optimize on. However, optimizing on $(0,\infty)$ and this bounded range should be equally easy, given that $\Rnum(\omega)$ has only one critical point; so we know the minimum must occur at the boundaries of the domain.

If $\Rnum$ attains its smallest values near $\omega = 0$, then testing results should never be returned (probably not a practical implementation, but it does tell us something interesting). On the other hand, if $\Rnum$ attains its lowest values as $\omega$ tends to infinity, then test results should be returned as quickly as possible. Of course, in reality, if test results were never returned, people may not continue to comply with being careful while ``awaiting" their test result; since it never comes.

Before doing any analysis, notice that ``never returning results" may be \fady{I'm not 100 \% sure} dynamically equivalent to having an ``SIR with lockdown model", except the lockdown period is permanent. This is because some people are lowering their contact rate while ``waiting" for their results (which will never come). In this case, the lowering of contact is given by the parameter $\eta_s$.

Looking at the limit as $\omega \rightarrow \infty$ we have
\begin{align}
&\lim_{w \rightarrow \infty} \beta/\gamma + \frac{\beta t}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta_w-1)(\gamma W_s+\omega W_i) + (\eta_t -1)P_iW_i \omega^2 \Big) \nonumber \\
&= \frac{\beta}{\gamma} - \frac{\beta t\left(1-t_n\right)P_{i}W_{i}}{\gamma^{2}W_{s}}
\end{align}

Naturally, we would now want to evaluate the limit as $\omega \rightarrow 0$. However, if we were to naively compute this limit assuming $t$ is constant, we would get $- \infty$. I suspect \fady {I'm not sure about this} the reason for this is because we're working with a linearization where $t \approx 0$, thus, we can only evaluate the limit as another variable (e.g., $\omega$) goes to 0 only if we insist that $t$ also goes to 0. But then, the value of the limit depends on the relative decay of $t$ and $\omega$ towards 0. \fady{I will not write down analysis here until I've talked to someone about this.}

\section{Rough thoughts: what about when $t$ is not close to 0?}

Initially, I thought the exact formula for $\Rnum$ was linear with respect to $t$, which would have allowed us to generalize the results in the previous section. However, I remembered that during the meeting, someone (I think it was Jonathan) was asking about a quadratic term. I figured that someone had noticed that if $\Rnum$ was linear we would've been able to generalize the analysis above, so looking harder, I noticed that indeed $\Rnum$ is not linear wrt to $t$ because $F_i$ contain $t$'s that then get multiplied with the one explicit $t$ in the expression. So indeed it's not linear.

I could try spending some time \emph{trying} to work on the exact $\Rnum$ function to see if I can get anywhere with it, if that's of interest.

\end{document}

### 2020-10-21 smoothing.tex
smoothing notes source

We don't want to test people from the \_u compartments if they're not there.

There are two natural ways to do this. One is with a discontinuity: we define 
the weighted number of people available for tests: $W = W_s S_u + W_i I_u + W_r R_u$ and then define the scaling parameter (can we not call it ``sc'', which looks like a product?) as $sc = \textrm{H}(W) tN_0/W$, where H is the Heaviside function. This makes the system well-defined, but is still ugly in a few ways.

The second natural way is to define a time scale $\tau$ for the maximum rate of testing the whole untested population. If this is fast (e.g., 1d), it should have minimal effects on the system when we're in the usual domain. A simple way to do this is by using $sc = \frac{\tau tN_0}{\tau W + t N_0}$. Since $\tau \gg t$, this generally collapses to the original form. When $W$ is super-small, however, it collapses instead to $\tau$.

My suggested notation for this equation is $$\sigma = \frac{\tau \rho N_0}{\tau W + \rho N_0}$$

### 2020-10-22 note/SIR_notes.Rnw
P_i changed to 1
updated literature review

@@
 
+% %%%%%%%
 \section{Method}
 
-\citep{arino2020simple} developed a SLIAR compartmental model to study the spread of an epidemic, specifically COVID-19, in a population.
-The model incorporates an Erlang distribution of times of sojourn inincubating, symptomatically and asymptomatically infectious compartments.
-Basic reproduction number is derived.
-Also, sensitivity analysis with respect to the underlying parameters for the following two outputs was carried out; (i) the number of observable cases during the course of the epidemic and at the peak, and (ii) the timing of the peak of the outbreak.
-Sensitivity analysis is performed using the R package multisensi.
+% %%%%%%%
+\section{Analysis}
 
 \section{Literature Review}
+
+% %%%%%%%
+\subsection{Explicit models of TTI (trace/test/isolate) based on network or agent-based models}
+
+\citep{endo2020implication} [Ali: It seems to me that this is just a statistical model to estimate the parent-offspring of an infected index, not sure if it fits into agent-based group!] Used simulation on a branching process model to assess the forward and backward contact tracing efficiency.
+Assuming a negative-binomial branching process with a mean R, reproduction number, and overdispersion parameter k, the mean total number of generation G3 and averted G3 are estimated.
+The effectiveness of TTI is defined as the ratio of averted to the mean.
+
+\citep{jenness2020modeling} developed a network-based transmission model for SARS-CoV-2 on the Diamond Princess outbreak to characterize transmission dynamics and to estimate the epidemiological impact of outbreak control and prevention measures.
+
+\citep{elbanna2020entry} [seems similar to MacPan model!]
+% %%%%%%%
+\subsection{Models of repeated random testing of isolated populations}
 \cite{bergstrom2020frequency}
@@
 
+\citep{lopman2020model} Used a Deterministic SEIR model, incorporated TTI, applicable to a university setting. They assumed a fairly high reproductive number that is not reduced through social
+distancing measures. They found that community-introduction of SARS-CoV-2 infection onto campus can be
+relatively controlled with effective testing, isolation, contract tracing and quarantine.
+
+% %%%%%%%
+\subsection{Other maybe-related works}
+\citep{arino2020simple} developed a SLIAR compartmental model to study the spread of an epidemic, specifically COVID-19, in a population.
+The model incorporates an Erlang distribution of times of sojourn inincubating, symptomatically and asymptomatically infectious compartments.
+Basic reproduction number is derived.
+Also, sensitivity analysis with respect to the underlying parameters for the following two outputs was carried out; (i) the number of observable cases during the course of the epidemic and at the peak, and (ii) the timing of the peak of the outbreak.
+Sensitivity analysis is performed using the R package multisensi.
+% %%%%%%%
 \bibliography{../SIRlibrary}

### 2020-10-22 codes/ali_test_sir.Rmd
testing function is updated, code is updated not the text

@@
     ## scaling the weights
-    sc <- t*N0/(W_s*S_u+W_i*I_u+W_r*R_u)
-    # sc <- t*N0
+    sc <- tau*t*N0/(tau*(W_s*S_u+W_i*I_u+W_r*R_u)+t*N0)
     F_s <- sc*W_s
@@
 ```{r model inputs, message=FALSE}
-params <- c(N0=1000000, beta=1,gamma=1/3, omega=0.25, t=0.01, 
+params <- c(N0=1000000, beta=0.8,gamma=1/3, omega=0.25, t=0.01, 
              W_s=0.01, W_i=1, W_r=0.01, #weights
@@
             eta_w=0.02, eta_t=0.01, #isolation parameter 
-            s=2 #eta_w/eta_t
+            s=2, #eta_w/eta_t
+            tau=1
             )
@@
                 N=0,P=0)
-# Disease free equilibrium
-Sn_dfe <-function(params){
-  unpack(as.list(params))
-  return((t*(1-P_s)*N0)/omega)
-}
-Su_dfe<- params[["N0"]]-Sn_dfe(params)
-
-state_dfe <- c(S_u=Su_dfe, S_n=Sn_dfe(params),
-                I_u=0,I_n=0,I_p=0,I_t=0,
-                R_u=0,R_n=0,R_p=0,R_t=0,
-                N=0,P=0)
-
 d <- seq(0,200,by=0.1)
@@
   return(I_u - 1) }
+```
+
+```{r DFE, message=FALSE}
+# DFE; numerical approach
+library(rootSolve)
+DFE <- function(S,params){
+  S_u <- S[1]
+  S_n <- S[2]
+  I_u <-0
+  R_u<-0
+  unpack(as.list(params))
+  sc <- tau*t*N0/(tau*(W_s*S_u+W_i*I_u+W_r*R_u)+t*N0)
+  F_s <- sc*W_s
+
+ return(c(eq1=S_u+S_n-N0,
+   eq2=-F_s*S_u+omega*S_n))
+}
+
+roots_dfe <- multiroot(DFE, c(1, 1), parms = params, positive = T)
+roots_dfe$root
+
+all.equal(sum(roots_dfe$root),params[["N0"]])
+
+# DFE; analytic approach
+Su_dfe <-function(params){
+  unpack(as.list(params))
+  a <- t
+  b <- -t+(1-t/omega)*tau*W_s
+  c <- -tau*W_s
+  d <- b^2-4*a*c
+
+  return(N0*2*a/(-b+sqrt(d)))
+}
+
+Sn_dfe<- params[["N0"]]-Su_dfe(params)
+
+state_dfe <- c(S_u=Su_dfe(params), S_n=Sn_dfe,
+                I_u=0,I_n=0,I_p=0,I_t=0,
+                R_u=0,R_n=0,R_p=0,R_t=0,
+                N=0,P=0)
+
+all.equal(c(state_dfe[["S_u"]],state_dfe[["S_n"]]),roots_dfe$root)
 ```
@@
    unpack(as.list(c(state,params)))
-   return(t*N0*W_i/(W_s*S_u+W_i*I_u+W_r*R_u))
+  sc <- tau*t*N0/(tau*(W_s*S_u+W_i*I_u+W_r*R_u)+t*N0)
+   return(sc*W_i)
  }
@@
     unpack(as.list(params))
-    Sn_dfe<- Sn_dfe(params)
-    Su_dfe<- N0-Sn_dfe
-    F_i <-t*N0*W_i/(W_s*Su_dfe) #at DFE
-    # F_i <-t*N0*W_i
+    Su_dfe<- Su_dfe(params)
+    Sn_dfe<- N0-Su_dfe
+    F_i <- tau*t*N0*W_i/(tau*(W_s*Su_dfe)+t*N0) #at DFE
+
     A <-gamma*(omega+gamma+eta_w*F_i) + omega*eta_t*P_i*F_i
@@
     unpack(as.list(params))
-    Sn_dfe<- Sn_dfe(params)
-    Su_dfe<- N0-Sn_dfe
-    F_i <-t*N0*W_i/(W_s*Su_dfe) #at DFE
+    Su_dfe<- Su_dfe(params)
+    Sn_dfe<- N0-Su_dfe
+    F_i <-tau*t*N0*W_i/(tau*(W_s*Su_dfe)+t*N0) #at DFE
 
@@
                     P_s=params[["P_s"]],P_i=params[["P_i"]],P_r=params[["P_r"]],
-                    eta_w=eta_w,eta_t=eta_t,s=params[["s"]])
+                    eta_w=eta_w,eta_t=eta_t,s=params[["s"]],tau=params[["tau"]])
 
@@
                     P_s=params[["P_s"]],P_i=params[["P_i"]],P_r=params[["P_r"]],
-                    eta_w=params[["eta_w"]],eta_t=eta_t,s=s)
+                    eta_w=params[["eta_w"]],eta_t=eta_t,s=s,tau=params[["tau"]])
 

### 2020-10-27 codes/ali_test_sir.Rmd
updated analysis and sim

@@
 
-1- What combination of testing intensity, $t$, and isolation, $\eta_w$ and $\eta_t$, will result in $R_0 <1$?
+1- What combination of testing intensity, $\rho$, and isolation, $\eta_w$ and $\eta_t$, will result in $R_0 <1$?
 
@@
 
+### Assumptions
 - Assume there are 3 groups of individuals: $S$, $I$ and $R$.
@@
 no false positives, then $S_p$ and $S_t$ are always zero (2) there are two 'integrator' or 'accumulator' compartments, $N$ and $P$, to collect cumulative reported negative or positive tests.
+- There is a single FoI, $\Lambda$, all groups being exposed to all others. That is, where a susceptible individual move to does not depend on who infected it.
 
-## Per Capita Rate of Flow from Untested to Positive or Negative Compartments
+### Parameters
 
-- $t$: testing intensity with dimension of $1/time$.
-- $\omega$: the rate of *onward flow* from the awaiting positive compartment, $p$, to reported/tested compartment, $t$, or from awaiting negative compartment, $n$, back to $u$.
+```{r, out.width = "400px"}
+knitr::include_graphics("../pix/sir_comp2.png")
+```
+
+- $\Lambda$: force of infection defined as $$\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_t I_t)}{N_0},$$ where $\beta$ is transmission rate, $\eta_w$ and $\eta_t$ are the isolation parameters for awaiting and reported individuals, respectively.
+We assume that $\eta_t<\eta_w$, i.e., the awaiting individuals for test results have a higher transmission probability than the reported individuals.
+- $\omega$: the rate of *onward flow* from the awaiting positive compartment, $p$, to reported/tested compartment, $\rho$, or from awaiting negative compartment, $n$, back to $u$.
 It has units of $1/time$.
+- $\gamma$: recovery rate ($1/time$).
+- $\rho$: per capita testing intensity ($1/time$).
+- $\tau$: the maximum rate of testing the whole untested population ($1/time$).
+- $W$: weighted number of people available for tests, defined as $W = W_s S_u + W_i I_u + W_r R_u$.
+- $\sigma$: scaling parameter for testing defined as $\sigma = \frac{\tau \rho N_0}{\tau W + \rho N_0}$.
+- $F_z$: Weighted testing rate defined as $F_z=\sigma W_z$. That is, $F_s = \sigma W_s$, $F_i=\sigma W_i$ and $F_r = \sigma W_r$.
 
-- Note: The complicated part is the flow rate from $Z_u$ to $Z_n$, $Z_p$.
-The (total, not per capita) flow from $Z_u$ to $Z_n$ is $t (1-P_Z) F(W_Z) Z_u$, from $Z_u$ to $Z_p$ is $t P_Z F(W_Z) Z_u$.
-Where, $P_Z$ is the probability of positive test.
+**Notes:** 
+- The complicated part is the flow rate from $Z_u$ to $Z_n$, $Z_p$.
+The (total, not per capita) flow from $Z_u$ to $Z_n$ is $\rho (1-P_z) F(W_z) Z_u$, from $Z_u$ to $Z_p$ is $\rho P_Z F(W_Z) Z_u$.
+Where, $P_z$ is the probability of positive test.
+Thus, $P_i$ represents the test accuracy.
+We assume $P_s=0$.
+- On testing and scaling; at the simplest $F_z=W_z$.  
+- JD: If this $\tau$ is fast (e.g., 1d), it should have minimal effects on the system when we're in the usual domain.
+Since $\tau \gg \rho$, this generally collapses to the original form.
+When $W$ is super-small, however, it collapses instead to $\tau$.
 
-- At the simplest $F(W_Z)=W_Z$.
-The flows from $Z_n$ to $Z_u$ and $Z_p$ to $Z_t$ are $\omega Z_n$ and $\omega Z_p$ respectively.
-We are using a dimensionless scaling where $F(W_Z) = (W_Z \cdot N_0)/\sum_{Z'} \left( W_{Z'} {Z'}_{u} \right)$, i.e.
-the denominator is the sum of the product of the weights and the current occupancy of the untested compartments.
-
-## Model
-
-![](../pix/sir_test1.jpg)
-
-Let the force of infection $\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_t I_t)}{N_0}$, where $\eta$ is the isolation parameter with the following assumptions:
-
-(1) All susceptible individuals are exposed to all infectious individuals regardless of the group factor. 
-(2) $\eta_t<\eta_w$, i.e., the awaiting individuals for test results have higher transmission probability than the reported individuals. 
-(3) The scaling parameter is defined as $sc = t N_0/(W_s S_u + W_i I_u + W_r R_u)$ and $F_s = sc \times W_s$, $F_i=sc \times W_i$ and $F_r = sc\times W_r$.
-
+### Model
 The model is
@@
     ## scaling the weights
-    sc <- tau*t*N0/(tau*(W_s*S_u+W_i*I_u+W_r*R_u)+t*N0)
-    F_s <- sc*W_s
-    F_i <- sc*W_i
-    F_r <- sc*W_r
+    sigma <- tau*rho*N0/(tau*(W_s*S_u+W_i*I_u+W_r*R_u)+rho*N0)
+    F_s <- sigma*W_s
+    F_i <- sigma*W_i
+    F_r <- sigma*W_r
 
@@
 ```{r model inputs, message=FALSE}
-params <- c(N0=1000000, beta=0.8,gamma=1/3, omega=0.25, t=0.01, 
+params <- c(N0=1000000, beta=0.8,gamma=1/3, omega=0.25, rho=0.01, 
              W_s=0.01, W_i=1, W_r=0.01, #weights
@@
   unpack(as.list(params))
-  sc <- tau*t*N0/(tau*(W_s*S_u+W_i*I_u+W_r*R_u)+t*N0)
-  F_s <- sc*W_s
+  sigma <- tau*rho*N0/(tau*(W_s*S_u+W_i*I_u+W_r*R_u)+rho*N0)
+  F_s <- sigma*W_s
 
@@
 
-all.equal(sum(roots_dfe$root),params[["N0"]])
+# DFE; analytic approach
 
-# DFE; analytic approach
 Su_dfe <-function(params){
   unpack(as.list(params))
-  a <- t
-  b <- -t+(1-t/omega)*tau*W_s
+  a <- rho
+  b <- -rho+(1-rho/omega)*tau*W_s
   c <- -tau*W_s
@@
 
-  return(N0*2*a/(-b+sqrt(d)))
+  if(rho){return(N0*2*a/(-b+sqrt(d)))}
+    else {return(N0)}
 }
@@
     times= d, 
-    parms=update(params,beta=2),
+    parms=update(params,beta=2)
     # atol = 1e+1, rtol = 1e+1
@@
 
-The Disease-Free Equilibrium, namely $E_1$, for the above SIR model is
-\label{exp:dfe} $S_n^*= \frac{t}{\omega} N_0, S_u^*= N_0-S_n^*$, and $I_j=R_j=0 \ \forall j$.
+The Disease-Free Equilibrium, namely $E_1$, for the above SIR model is given by the solution of the following system:
+$S_u+S_n=N0$ and $F_s S_u-\omega S_n=0$.
+Let $\tilde{S_u}=N_0/S_u$, the positive solution of the quadratic $$a \tilde{S_u}+b \tilde{S_u}^2+c=0,$$
+where $a=\rho$, $b= -\rho+(1-\rho/\omega) \tau W_s$ and $c=-\tau W_s$. Let
+  $d = b^2-4ac$, the DFE of the model is
+
+$$S_u^* =N_0 2 a/(-b+\sqrt d), \\ S_n^*=N_0-S_u^* \\ I_j=R_j=0 \ \forall j$$
 
@@
 ```{r R0}
-F_i <- function(state,params){
-   unpack(as.list(c(state,params)))
-  sc <- tau*t*N0/(tau*(W_s*S_u+W_i*I_u+W_r*R_u)+t*N0)
-   return(sc*W_i)
+
+F_i <- function(params){
+   # Returns F_i at DFE 
+  unpack(as.list(params))
+  Su_dfe<- Su_dfe(params)
+  Sn_dfe<- N0-Su_dfe
+  sigma <- tau*rho*N0/(tau*(W_s*Su_dfe)+rho*N0) #at DFE 
+  return(sigma*W_i)
 }
+
  # just a test
-  F_i(state=state_dfe,params=update(params,t=0))
+  F_i(params=update(params,rho=0))
 
@@
     Sn_dfe<- N0-Su_dfe
-    F_i <- tau*t*N0*W_i/(tau*(W_s*Su_dfe)+t*N0) #at DFE
+    # print(c(Su_dfe,Sn_dfe,Su_dfe+Sn_dfe))
+    F_i <- F_i(params) #at DFE
 
@@
     Sn_dfe<- N0-Su_dfe
-    F_i <-tau*t*N0*W_i/(tau*(W_s*Su_dfe)+t*N0) #at DFE
+    F_i <-F_i(params)  #at DFE
 
@@
 # checking if simple R_0 for SIR can be recovered?
-# Case #1: t=0
-all.equal(R0(params=update(params,t=0)), params[["beta"]]/params[["gamma"]])
+# Case #1: rho=0
+all.equal(R0(params=update(params,rho=0)), params[["beta"]]/params[["gamma"]])
 # Case #2: 
@@
 
-### On Behaviour of Testing at DFE Boundry
-(1) Note that once the only untested people are susceptibles, the FoI will be $\Lambda=0$, testing rate $F_s=t N_0/S_u$.
-Thus, eq(1) of the model will be $d S_u/dt = - t N_0 + \omega S_n$ which is a linear rate of leaving the $S_u$ compartment.
-That means that the testing rate will be no longer dependent on $S_u$.
-
 ### Contour plot of $R_0=1$ and interpretations approache 1
@@
  # specify the ranges
- beta  <- 0.5  ## set beta high enough to allow R0>1 in worst case
+ beta  <- 0.4  ## set beta high enough to allow R0>1 in worst case
  eta_w <- seq(0,1, length.out=5)
- eta_t <- seq(0.001,0.5, length.out=5) 
- t     <- seq(0,0.01, length.out=10)
- omega <- seq(0.1,1 , length.out=10) #note omega must be > t
-# s <- seq(1,2.5,length.out=10)
+ eta_t <- seq(0.001,0.5, length.out=5) #0.001 to 0.5 
+ rho     <- seq(0,0.01, length.out=10)
+ omega <- seq(0.001,1 , length.out=10) #note omega must be > rho, was 0.1
 
- df1 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,t=t,
+ df1 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
                     W_s=params[["W_s"]],W_i=params[["W_i"]],W_r=params[["W_r"]],
@@
 
+# Need to be corrected: in F_i, state_dfe should be changed according to the df1!
 df2<- data.frame(df1,R0=apply(df1,1,function(params_in)R0(params=params_in)),
-                      Fi= apply(df1,1,function(params_in)F_i(state=state_dfe,params=params_in)))
+                      Fi= apply(df1,1,function(params_in)F_i(params=params_in)))
 
@@
 ```{r show_contours_1, width=24,height=24,message=FALSE,warning=FALSE}
-p1 <- ggplot(df2,aes(x=omega,y=t,z=R0))+ theme_bw() +
+p1 <- ggplot(df2,aes(x=omega,y=rho,z=R0))+ theme_bw() +
     theme(panel.spacing=grid::unit(0,"lines"))
@@
  # specify the ranges
- beta  <- 0.5  ## set beta high enough to allow R0>1 in worst case
+ beta  <- 0.4  
  # eta_w <- seq(0,1, length.out=5) 
- eta_t <- seq(0.001,0.5, length.out=5) 
- t     <- seq(0,0.01, length.out=10)
- omega <- seq(0.1,1 , length.out=10) #note omega must be > t
-s <- seq(1,2.5,length.out=10)
+ eta_t <- seq(0.001,0.5, length.out=5) #0.001 
+ rho     <- seq(0,0.01, length.out=5)
+ omega <- seq(0.001,1 , length.out=10) #note omega must be > 0, was 0.1
+s <- seq(1,2,length.out=5)
 
- df11 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,t=t,
+ df11 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
                     W_s=params[["W_s"]],W_i=params[["W_i"]],W_r=params[["W_r"]],
@@
 df22<- data.frame(df11,R0=apply(df11,1,function(params_in)R02(params=params_in)),
-                      Fi= apply(df1,1,function(params_in)F_i(state=state_dfe,params=params_in)))
-
+                      Fi= apply(df11,1,function(params_in)F_i(params=params_in)),
+                       Su=apply(df11,1,function(params_in)Su_dfe(params = params_in)))
+min(df22$Su)
+which(df22[]<0)          
 ```
@@
 ```{r show_contours_2, width=24,height=24,message=FALSE,warning=FALSE}
-p11 <- ggplot(df22,aes(x=omega,y=t,z=R0))+ theme_bw() +
+p11 <- ggplot(df22,aes(x=omega,y=rho,z=R0))+ theme_bw() +
     theme(panel.spacing=grid::unit(0,"lines"))
@@
 1. $\partial{R_0}/\partial{\eta_t} \geq 0$ and $\partial{R_0}/\partial{\eta_w} \geq 0$.
-2. $\partial{R_0}/\partial{t} \leq 0$; this is concluded from the Taylor expansion of $R_0$ around $t=0$ which is
-$R_0 \approx \beta/\gamma + \frac{\beta t}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta_w-1)(\gamma W_s+\omega W_i) + (\eta_t -1)P_iW_i \omega^2 \Big) + \mathcal{O}(t^2)$.
-3. Based on the above Taylor expansion around $t=0$,
-$$\partial{R_0}/\partial{\omega}=  \frac{-\beta t}{\gamma W_s\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c), $$
+2. $\partial{R_0}/\partial{\rho} \leq 0$; this is concluded from the Taylor expansion of $R_0$ around $\rho=0$ which is
+$R_0 \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta_w-1)(\gamma W_s+\omega W_i) + (\eta_t -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2)$.
+3. Based on the above Taylor expansion around $\rho=0$,
+$$\partial{R_0}/\partial{\omega}=  \frac{-\beta \rho}{\gamma W_s\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c), $$
 where $a=(\eta_w-1)W_i-(\eta_t-1)P_iW_i$, $b=2(\eta_w-1)\gamma W_s$ and $c=(\eta_w-1)\gamma^2 W_s$.
@@
 Thus, we only consider the lower-triangle panels in the plot [not sure how to extract them in ggplot?].
-(1) Looking at the first column of panels, there seems to be a compensatory relation between $t$ and $\omega$.
-i.e., higher $t$ requires smaller $\omega$ to keep R0 lower than 1.
-Also, for a fixed $\omega$, increasing $\eta_w$ requires higher $t$ to achieve the critical R0.
+(1) Looking at the first column of panels, there seems to be a compensatory relation between $\rho$ and $\omega$.
+i.e., higher $\rho$ requires smaller $\omega$ to keep R0 lower than 1.
+Also, for a fixed $\omega$, increasing $\eta_w$ requires higher $\rho$ to achieve the critical R0.
 (2) A sensible result that we are expecting is $\partial{R_0}/\partial{\omega}<0$.

### 2020-10-27 codes/ali_test_sir.Rmd
update dimensions for contour plots

@@
 ```{r contour_R0_1, message=FALSE,warning=FALSE }
+n_out <- 5 ## facets (rows/cols)
+n_in <- 41 ## grid N within facets
  # specify the ranges
  beta  <- 0.4  ## set beta high enough to allow R0>1 in worst case
- eta_w <- seq(0,1, length.out=5)
- eta_t <- seq(0.001,0.5, length.out=5) #0.001 to 0.5 
- rho     <- seq(0,0.01, length.out=10)
- omega <- seq(0.001,1 , length.out=10) #note omega must be > rho, was 0.1
+ eta_w <- seq(0,1, length.out=n_out)
+ eta_t <- seq(0.001,0.5, length.out=n_out) #0.001 to 0.5 
+ rho     <- seq(0,0.01, length.out=n_in)
+ omega <- seq(0.001,1 , length.out=n_in) #note omega must be > rho, was 0.1
 
@@
  # eta_w <- seq(0,1, length.out=5) 
- eta_t <- seq(0.001,0.5, length.out=5) #0.001 
- rho     <- seq(0,0.01, length.out=5)
- omega <- seq(0.001,1 , length.out=10) #note omega must be > 0, was 0.1
-s <- seq(1,2,length.out=5)
+ eta_t <- seq(0.001,0.5, length.out=n_out) #0.001 
+ rho     <- seq(0,0.01, length.out=n_in)
+ omega <- seq(0.001,1 , length.out=n_in) #note omega must be > 0, was 0.1
+ s <- seq(1,2,length.out=n_out)
 

### 2020-10-28 smoothing.tex
some question on suggested sigma

@@
 My suggested notation for this equation is $$\sigma = \frac{\tau \rho N_0}{\tau W + \rho N_0}$$
+
+Questions/Clarifications:
+(1) We have two testing rates; (i) $\tau$ and (ii) $\rho$.
+$\tau$ is the testing rate when $W \approx 0$ so no one is left in the untested compartments.
+(Here I am not sure why "a time scale $\tau$ for the maximum rate of testing the whole untested population"?).
+On the other hand $\rho$ is the testing rate at the begining of the pandemic when $W \approx N_0$. The latter requires $\tau \gg \rho$. Am I thinking right here?  
+(2) $\tau$ has the same dimension as $\rho$ which is testing rate, thus 1/time per capita?
+(3) In the simmulation, I picked $\tau=1$. This made sense to me if the answeres to all above questions are "yes". Not sure what to pick if I was not thinking right in the above question.

### 2020-10-28 codes/ali_test_sir.Rmd
comments and responses

@@
 ### Assumptions
+JD: Why are we assuming structurally that the test is perfectly specific? I guess it's better to put all of the boxes in and just sometimes assume that that $P_S=0$ 
 - Assume there are 3 groups of individuals: $S$, $I$ and $R$.
@@
 - There is a single FoI, $\Lambda$, all groups being exposed to all others. That is, where a susceptible individual move to does not depend on who infected it.
+
+JD: Can we do a better job with capitalization?
+When $I$ in a subscript refers to the infected class, for example, it should be capitalized.
+Also, I guess it would be good to always use $P$ for positive (thus $N$ for negative), and then use $p$ for the probabilities.
+I guess the concern is that $p_S$ looks awkward?
+But I think clarity is more important.
 

### 2020-10-28 smoothing.tex
comments and responses

@@
 On the other hand $\rho$ is the testing rate at the begining of the pandemic when $W \approx N_0$. The latter requires $\tau \gg \rho$. Am I thinking right here?
+
+JD: Yes, you are right that $\tau \gg \rho$.
+But it's not exactly about the beginning of the pandemic.
+In general, we want to test at a rate of $\rho$ across the whole population.
+This won't always be possible.
+So we impose a maximum rate of $\tau$ per testable person.
+It is consistent to think of both $\tau$ and $\rho$ as pure rates, but it might be clearer to think of $\rho$ as tests per capita per unit time, and $\tau$ as tests per testable person per unit time.
+It's not that we're switching through time, it's that we're imposing both of these as limitations.
+At the beginning, we expect the answer to be close to $\rho N_0$ since $\tau$ should be very fast.
+Once $W$ becomes small, the limitation imposed by $\tau W$ will become important.
+
 (2) $\tau$ has the same dimension as $\rho$ which is testing rate, thus 1/time per capita?
+
+JD: See above.
+Note that if you are thinking about ``people'' in the denominator as above, you need to remember there are also people in the numerator (you can choose to be specific, or to let them cancel).
+But it's a bit off to have people only in the denominator.
+
 (3) In the simmulation, I picked $\tau=1$. This made sense to me if the answeres to all above questions are "yes". Not sure what to pick if I was not thinking right in the above question.
+
+JD: Yes, but you need units. $\tau = 1 \mathrm{day}$ is exactly what I meant by a time scale of 1 day.

### 2020-10-28 smoothing.tex
more questions

@@
 JD: Yes, but you need units. $\tau = 1 \mathrm{day}$ is exactly what I meant by a time scale of 1 day. 
+AG: Sorry, I am not sure here about the units.
+
+(4) $\sigma$ is the testing rate with unit of 1/time. $\sigma$ has two limits
+(i) When $W \approx N0$;
+If we assume that $\tau \gg \rho$, $\sigma=\rho \frac{\tau N0}{(\tau+\rho)N_0} \approx \rho \frac{\tau N0}{(\tau)N_0} = \rho$.
+
+(ii) When $W \approx 0$;
+$\sigma \approx \tau$.
+
+(5) Right now there are no restictions on the weights, $W_S, W_I$ and $W_R$.
+These are representing different testing strategies.
+For example if $W_S=W_I=W_R$, it represents the random testing from the whole population with no tracing.
+(i) Does it make sense $W_S+W_I+W_R=1$? thus if we increase the W_I, the other two weights will decrease.
+(ii) I am thinking of somehow connect $W_*$ and testing rate of the whole population $\rho$.
+With the new definition of the $\sigma$ that JD suggested above, this is already incorporated in the rate of testing but not in the testing strategies.

### 2020-10-30 smoothing.tex
meeting with Ali
clean up

@@
 
-There are two natural ways to do this. One is with a discontinuity: we define 
-the weighted number of people available for tests: $W = W_s S_u + W_i I_u + W_r R_u$ and then define the scaling parameter (can we not call it ``sc'', which looks like a product?) as $sc = \textrm{H}(W) tN_0/W$, where H is the Heaviside function.
-This makes the system well-defined, but is still ugly in a few ways.
+- The initial approache and the singularity issue;
+Initial approach and singularity issue (not sure if I call it structural or numerical singularity?); 
+$\sigma = \frac{\rho N_0}{W}$ was used initially.
+The issue was that the population in $S$ compartments appeared to blow up when the DFE is achieved ( in the simulation when time, $t>200$).
+This is  once the only untested people are susceptibles, the FoI will become $\Lambda=0$, testing rate $F_s=t N_0/S_u$.
+Thus, eq(1) of the model will be $d S_u/dt = - t N_0 + \omega S_n$ which is no longer dependent on $S_u$ and a linear rate of leaving the $S_u$ compartment.
 
-The second natural way is to define a time scale $\tau$ for the maximum rate of testing the whole untested population.
-If this is fast (e.g., 1d), it should have minimal effects on the system when we're in the usual domain.
-A simple way to do this is by using $sc = \frac{\tau tN_0}{\tau W + t N_0}$.
-Since $\tau \gg t$, this generally collapses to the original form.
+- Solution;
+It was suggested that $$\sigma = \frac{\tau \rho N_0}{\tau W + \rho N_0}.$$
+Note that $F_Z=\sigma * W_Z$.
+One natural way is to define a rate $\tau$ ($\mathrm{day}$) for the maximum rate of testing the whole untested population.
+In general, since $\tau \gg t$, this generally collapses to the original form.
 When $W$ is super-small, however, it collapses instead to $\tau$.
-
-My suggested notation for this equation is $$\sigma = \frac{\tau \rho N_0}{\tau W + \rho N_0}$$
-
-Questions/Clarifications:
-(1) We have two testing rates; (i) $\tau$ and (ii) $\rho$.
-$\tau$ is the testing rate when $W \approx 0$ so no one is left in the untested compartments.
-(Here I am not sure why "a time scale $\tau$ for the maximum rate of testing the whole untested population"?).
-On the other hand $\rho$ is the testing rate at the begining of the pandemic when $W \approx N_0$. The latter requires $\tau \gg \rho$. Am I thinking right here?
-
-JD: Yes, you are right that $\tau \gg \rho$.
-But it's not exactly about the beginning of the pandemic.
 In general, we want to test at a rate of $\rho$ across the whole population.
@@
 
-(2) $\tau$ has the same dimension as $\rho$ which is testing rate, thus 1/time per capita?
-
-JD: See above.
-Note that if you are thinking about ``people'' in the denominator as above, you need to remember there are also people in the numerator (you can choose to be specific, or to let them cancel).
-But it's a bit off to have people only in the denominator.
-
-(3) In the simmulation, I picked $\tau=1$. This made sense to me if the answeres to all above questions are "yes". Not sure what to pick if I was not thinking right in the above question.
-
-JD: Yes, but you need units. $\tau = 1 \mathrm{day}$ is exactly what I meant by a time scale of 1 day. 
-AG: Sorry, I am not sure here about the units.
-
-(4) $\sigma$ is the testing rate with unit of 1/time. $\sigma$ has two limits
-(i) When $W \approx N0$;
-If we assume that $\tau \gg \rho$, $\sigma=\rho \frac{\tau N0}{(\tau+\rho)N_0} \approx \rho \frac{\tau N0}{(\tau)N_0} = \rho$.
-
-(ii) When $W \approx 0$;
-$\sigma \approx \tau$.
+To summerize, in general $\tau \gg \rho$, thus $\sigma=\rho \frac{\tau N0}{(\tau+\rho)N_0} \approx \rho \frac{\tau N0}{(\tau)N_0} = \rho$. When $W \approx 0$, $\sigma \approx \tau$.
 
@@
 For example if $W_S=W_I=W_R$, it represents the random testing from the whole population with no tracing.
-(i) Does it make sense $W_S+W_I+W_R=1$? thus if we increase the W_I, the other two weights will decrease.
+(i) Does it make sense $W_S+W_I+W_R=1$? thus if we increase the $W_I$, the other two weights will decrease.
 (ii) I am thinking of somehow connect $W_*$ and testing rate of the whole population $\rho$.

### 2020-10-30 smoothing.tex
clean up and adding testing strategies

@@
-We don't want to test people from the \_u compartments if they're not there.
+Note that the equation for the untested susciptable compartment is $d S_u/dt = -\Lambda S_u - F_S S_u + \omega S_n$, where $\Lambda$ is FoI, $F_Z$ in general form is the weighted testing rate defined as $F_Z=\sigma W_Z$, and $\sigma$ is the testing rate.
+Let $W = W_S S_u + W_I I_u + W_R R_u$ be the weighted number of people available for tests.
+
+\section{On testing rate}
+
+The testing rate, $\sigma$, should be formulated such that people from the \_u compartments will not be tested if they're not there.
 
 - The initial approache and the singularity issue;
-Initial approach and singularity issue (not sure if I call it structural or numerical singularity?); 
 $\sigma = \frac{\rho N_0}{W}$ was used initially.
-The issue was that the population in $S$ compartments appeared to blow up when the DFE is achieved ( in the simulation when time, $t>200$).
-This is  once the only untested people are susceptibles, the FoI will become $\Lambda=0$, testing rate $F_s=t N_0/S_u$.
-Thus, eq(1) of the model will be $d S_u/dt = - t N_0 + \omega S_n$ which is no longer dependent on $S_u$ and a linear rate of leaving the $S_u$ compartment.
+The issue was that the population in $S$ compartments appeared to blow up when the DFE is achieved.
+This is  once the only untested people are susceptibles, the FoI will become $\Lambda=0$, testing rate $F_s=\rho N_0/S_u$.
+Thus, eq(1) of the model will be $d S_u/dt = - \rho N_0 + \omega S_n$ which is no longer dependent on $S_u$ and a linear rate of leaving the $S_u$ compartment.
 
-- Solution;
-It was suggested that $$\sigma = \frac{\tau \rho N_0}{\tau W + \rho N_0}.$$
-Note that $F_Z=\sigma * W_Z$.
-One natural way is to define a rate $\tau$ ($\mathrm{day}$) for the maximum rate of testing the whole untested population.
-In general, since $\tau \gg t$, this generally collapses to the original form.
-When $W$ is super-small, however, it collapses instead to $\tau$.
+- The secondary approache;
+It was suggested that
+
+One natural way is to define $$\sigma = \frac{\tau \rho N_0}{\tau W + \rho N_0}.$$
+Where  $\tau$ ($1/ \mathrm{day}$) is a rate for the maximum rate of testing the whole untested population.
 In general, we want to test at a rate of $\rho$ across the whole population.
@@
 It's not that we're switching through time, it's that we're imposing both of these as limitations.
-At the beginning, we expect the answer to be close to $\rho N_0$ since $\tau$ should be very fast.
-Once $W$ becomes small, the limitation imposed by $\tau W$ will become important.
+Note that, in general, since $\tau \gg \rho$, thus $\sigma=\rho \frac{\tau N0}{(\tau+\rho)N_0} \approx \rho \frac{\tau N0}{(\tau)N_0} = \rho$.
+This generally collapses to the original form.
+On the other hand, when $W$ is super-small, however, it collapses instead to $\tau$.
+  In other words, at the beginning, we expect the answer to be close to $\rho N_0$ since $\tau$ should be very fast. Once $W$ becomes small, the limitation imposed by $\tau W$ will become important.
 
-To summerize, in general $\tau \gg \rho$, thus $\sigma=\rho \frac{\tau N0}{(\tau+\rho)N_0} \approx \rho \frac{\tau N0}{(\tau)N_0} = \rho$. When $W \approx 0$, $\sigma \approx \tau$.
+\section{ On the testing strategies}
 
-(5) Right now there are no restictions on the weights, $W_S, W_I$ and $W_R$.
-These are representing different testing strategies.
-For example if $W_S=W_I=W_R$, it represents the random testing from the whole population with no tracing.
-(i) Does it make sense $W_S+W_I+W_R=1$? thus if we increase the $W_I$, the other two weights will decrease.
-(ii) I am thinking of somehow connect $W_*$ and testing rate of the whole population $\rho$.
-With the new definition of the $\sigma$ that JD suggested above, this is already incorporated in the rate of testing but not in the testing strategies.
+- The testing weights, $W_S, W_I$ and $W_R$, are representing different testing strategies.
+- The only thing that matters is the relative proportions of W's.
+- Right now there are no restictions on the weights. If we impose $\sum_Z W_Z=1$, we are just reducing a degree of freedom. i.e., if we increase the $W_I$, the other two weights will decrease.
+
+- Two things to explore are: 
+(i) Random testing; this scenariocan be explored when $W_S=W_I=W_R$.
+
+(ii) Testing and tracing; this scenariocan be explored when $W_S \ll W_I$ and for example $W_I=W_R=1$.

### 2020-11-02 codes/ali_test_sir.Rmd
updated testing strategies analysis

@@
 
-1- What combination of testing intensity, $\rho$, and isolation, $\eta_w$ and $\eta_t$, will result in $R_0 <1$?
+- An exploration of simple mathematical/computational models of COVID-19 to gain insight into the effects of different levels of testing and isolation on $R_t$/growth rates. The goal is insight.
 
@@
 **Notes:**
+
 - The complicated part is the flow rate from $Z_u$ to $Z_n$, $Z_p$.
@@
 We assume $P_s=0$.
+
 - On testing and scaling; at the simplest $F_z=W_z$.
+
 - JD: If this $\tau$ is fast (e.g., 1d), it should have minimal effects on the system when we're in the usual domain.
@@
 When $W$ is super-small, however, it collapses instead to $\tau$.
+
+### On Testing
+
+- Initial approach and singularity issue (not sure if I call it structural or numerical singularity?); 
+$\sigma = \frac{\rho N_0}{W}$ was used initially.
+The issue was that the population in $S$ compartments appeared to blow up when the DFE is achieved ( in the simulation when time, $t>200$).
+This is  once the only untested people are susceptibles, the FoI will become $\Lambda=0$, testing rate $F_s=t N_0/S_u$.
+Thus, eq(1) of the model will be $d S_u/dt = - t N_0 + \omega S_n$ which is no longer dependent on $S_u$ and a linear rate of leaving the $S_u$ compartment.
 
@@
 params <- c(N0=1000000, beta=0.8,gamma=1/3, omega=0.25, rho=0.01, 
-             W_s=0.01, W_i=1, W_r=0.01, #weights
+             W_s=1, W_i=1, W_r=1, #weights
              P_s=0, P_i=1, P_r=0.5, #prob of waiting for being tested positive
@@
 ```{r run2, message=FALSE}
-d_long <- seq(0,300,by=0.1)
+d_long <- seq(0,400,by=0.1)
 ## capturing badly behaved integrator warning messages
@@
     # print(c(Su_dfe,Sn_dfe,Su_dfe+Sn_dfe))
-    F_i <- F_i(params) #at DFE
+    Fi <- F_i(params) #at DFE
 
-    A <-gamma*(omega+gamma+eta_w*F_i) + omega*eta_t*P_i*F_i
-    B <-(omega+eta_w*(F_i+gamma))*gamma+(eta_w*gamma+eta_t*omega)*(omega*P_i*F_i)/(omega+gamma)
+    A <-gamma*(omega+gamma+eta_w*Fi) + omega*eta_t*P_i*Fi
+    B <-(omega+eta_w*(Fi+gamma))*gamma+(eta_w*gamma+eta_t*omega)*(omega*P_i*Fi)/(omega+gamma)
 
-    C <- beta/(N0*gamma*(gamma*(omega+gamma)+F_i*(gamma+omega*P_i)))
+    C <- beta/(N0*gamma*(gamma*(omega+gamma)+Fi*(gamma+omega*P_i)))
     return((A*Su_dfe+B*Sn_dfe)*C)
@@
  # specify the ranges
- beta  <- 0.4  ## set beta high enough to allow R0>1 in worst case
+ beta  <- 0.5  ## set beta high enough to allow R0>1 in worst case
  eta_w <- seq(0,1, length.out=n_out)
@@
  # specify the ranges
- beta  <- 0.4  
+ beta  <- 0.5
  # eta_w <- seq(0,1, length.out=5) 
- eta_t <- seq(0.001,0.5, length.out=n_out) #0.001 
+ eta_t <- seq(0.0001,0.5, length.out=n_out) #0.001 
  rho     <- seq(0,0.01, length.out=n_in)
@@
                        Su=apply(df11,1,function(params_in)Su_dfe(params = params_in)))
-min(df22$Su)
-which(df22[]<0)          
+# min(df22$Su)
+# which(df22[]<0)          
 ```
@@
 
+4. Taylor expansion gives $R_0|_{(\rho=0,W_s=1)} = \beta/\gamma$.
+
+5.
+Taylor expansion gives $R_0|_{(\rho=0,W_s=0)} = \beta/\gamma \frac{\gamma \tau(\omega+\gamma)W_s+\tau\rho(\gamma\eta_w+\omega\eta_tP_i)W_i+ \gamma \rho(\gamma+\omega)}{\gamma \tau(\omega+\gamma)W_s+\tau\rho(\gamma+\omega P_i)W_i+ \gamma \rho(\gamma+\omega)}$.
+
 **Interpretation of the R0 contour plot**

### 2020-11-02 note/SIR_notes.Rnw
testing function is updated, code is updated not the text
updated testing strategies analysis

@@
 
+\citep{tuite2020mathematical} used an age-structured compartmental model of COVID-19 transmission in the population of Ontario, Canada.
+We compared a base case with limited testing, isolation and quarantine to different scenarios.
+
 % %%%%%%%
@@
 Sensitivity analysis is performed using the R package multisensi.
+
+\citep{ruszkiewicz2020diagnosis} novel with-in-a-minute breath testing with 80\% accuracy. 
 % %%%%%%%

### 2020-11-03 codes/ali_test_sir.Rmd
some questions in smoothing

@@
 params <- c(N0=1000000, beta=0.8,gamma=1/3, omega=0.25, rho=0.01, 
-             W_s=1, W_i=1, W_r=1, #weights
+             W_s=.01, W_i=1, W_r=1, #weights
              P_s=0, P_i=1, P_r=0.5, #prob of waiting for being tested positive
@@
 ```{r run2, message=FALSE}
-d_long <- seq(0,400,by=0.1)
+d_long <- seq(0,300,by=0.1)
 ## capturing badly behaved integrator warning messages
@@
 ```{r R0}
-
 F_i <- function(params){
@@
  # specify the ranges
- beta  <- 0.5
+ # beta  <- 0.5
  # eta_w <- seq(0,1, length.out=5) 
- eta_t <- seq(0.0001,0.5, length.out=n_out) #0.001 
- rho     <- seq(0,0.01, length.out=n_in)
- omega <- seq(0.001,1 , length.out=n_in) #note omega must be > 0, was 0.1
+ # eta_t <- seq(0.001,0.5, length.out=n_out) #0.001 
+ # rho     <- seq(0,0.01, length.out=n_in)
+ # omega <- seq(0.001,1 , length.out=n_in) #note omega must be > 0, was 0.1
  s <- seq(1,2,length.out=n_out)
@@
                        Su=apply(df11,1,function(params_in)Su_dfe(params = params_in)))
-# min(df22$Su)
-# which(df22[]<0)          
 ```

### 2020-11-03 smoothing.tex
some questions in smoothing

@@
 - The initial approache and the singularity issue;
+
 $\sigma = \frac{\rho N_0}{W}$ was used initially.
@@
 - The secondary approache;
-It was suggested that
 
@@
 (ii) Testing and tracing; this scenariocan be explored when $W_S \ll W_I$ and for example $W_I=W_R=1$.
+
+\section{Questions};
+
+1- When $\sigma=\rho$?  
+  $\frac{\tau N0}{\tau W+ \rho N_0} = 1$. Mathematically, this can be achieved when $W \approx N_0$ and   $\tau \gg \rho$.
+
+2- What is the most interesting thing in the testing strategies?
+We intuitively know testing/tracing/isolation (TTI) strategy is more effective than random testing in the sense that it lowers the $R_0$. Obvious things to ask are
+
+(i)  How much TTI is lowering $R_0$? 
+(ii) How fast TTI is lowering $R_0$?

### 2020-11-05 codes/ali_test_sir.Rmd
back to initial sigma

@@
 ### Assumptions
-JD: Why are we assuming structurally that the test is perfectly specific? I guess it's better to put all of the boxes in and just sometimes assume that that $P_S=0$ 
+JD: Why are we assuming structurally that the test is perfectly specific? I guess it's better to put all of the boxes in and just sometimes assume that that $p_S=0$ 
 - Assume there are 3 groups of individuals: $S$, $I$ and $R$.
@@
 - $\gamma$: recovery rate ($1/time$).
-- $\rho$: per capita testing intensity ($1/time$).
-- $\tau$: the maximum rate of testing the whole untested population ($1/time$).
-- $W$: weighted number of people available for tests, defined as $W = W_s S_u + W_i I_u + W_r R_u$.
-- $\sigma$: scaling parameter for testing defined as $\sigma = \frac{\tau \rho N_0}{\tau W + \rho N_0}$.
-- $F_z$: Weighted testing rate defined as $F_z=\sigma W_z$. That is, $F_s = \sigma W_s$, $F_i=\sigma W_i$ and $F_r = \sigma W_r$.
+- $\rho$: per capita testing intensity across the whole population ($1/time$).
+- $W$: weighted number of people available for tests, defined as $W = W_S S_u + W_I I_u + W_R R_u$.
+- $\sigma$: scaling parameter for testing defined as $\sigma = \frac{\rho N_0}{W}$.
+- $F_Z$: Weighted testing rate defined as $F_Z=\sigma W_Z$. That is, $F_S = \sigma W_S$, $F_I=\sigma W_I$ and $F_R = \sigma W_R$.
 
@@
 - The complicated part is the flow rate from $Z_u$ to $Z_n$, $Z_p$.
-The (total, not per capita) flow from $Z_u$ to $Z_n$ is $\rho (1-P_z) F(W_z) Z_u$, from $Z_u$ to $Z_p$ is $\rho P_Z F(W_Z) Z_u$.
-Where, $P_z$ is the probability of positive test.
-Thus, $P_i$ represents the test accuracy.
-We assume $P_s=0$.
+The (total, not per capita) flow from $Z_u$ to $Z_n$ is $\rho (1-P_Z) F(W_Z) Z_u$, from $Z_u$ to $Z_p$ is $\rho P_Z F(W_Z) Z_u$.
+Where, $P_Z$ is the probability of positive test.
+Thus, $p_I$ represents the test accuracy.
+We assume $p_S=0$.
 
-- On testing and scaling; at the simplest $F_z=W_z$.
-
-- JD: If this $\tau$ is fast (e.g., 1d), it should have minimal effects on the system when we're in the usual domain.
-Since $\tau \gg \rho$, this generally collapses to the original form.
-When $W$ is super-small, however, it collapses instead to $\tau$.
+- On testing and scaling; at the simplest $F_Z=W_Z$.
 
@@
 The issue was that the population in $S$ compartments appeared to blow up when the DFE is achieved ( in the simulation when time, $t>200$).
-This is  once the only untested people are susceptibles, the FoI will become $\Lambda=0$, testing rate $F_s=t N_0/S_u$.
-Thus, eq(1) of the model will be $d S_u/dt = - t N_0 + \omega S_n$ which is no longer dependent on $S_u$ and a linear rate of leaving the $S_u$ compartment.
+This is  once the only untested people are susceptibles, the FoI will become $\Lambda=0$, testing rate $F_S=\rho N_0/S_u$.
+Thus, eq(1) of the model will be $d S_u/dt = - \rho N_0 + \omega S_n$ which is no longer dependent on $S_u$ and a linear rate of leaving the $S_u$ compartment.
 
@@
 \label{mod:1}
-1)\ d S_u/dt &= -\Lambda S_u - F_s S_u + \omega S_n \\
-2)\ d S_n/dt &= -\Lambda S_n + F_s S_u - \omega S_n \\
-3)\ d I_u/dt &= \Lambda S_u - F_i I_u + \omega I_n  - \gamma I_u  \\
-4)\ d I_n/dt &= \Lambda S_n + (1-P_i) F_i I_u - \omega I_n -\gamma I_n \\
-5)\ d I_p/dt &= P_i F_i I_u - \omega I_p -\gamma I_p \\
+1)\ d S_u/dt &= -\Lambda S_u - F_S S_u + \omega S_n \\
+2)\ d S_n/dt &= -\Lambda S_n + F_S S_u - \omega S_n \\
+3)\ d I_u/dt &= \Lambda S_u - F_I I_u + \omega I_n  - \gamma I_u  \\
+4)\ d I_n/dt &= \Lambda S_n + (1-p_I) F_I I_u - \omega I_n -\gamma I_n \\
+5)\ d I_p/dt &= p_I F_I I_u - \omega I_p -\gamma I_p \\
 6)\ d I_t/dt &= \omega I_p - \gamma I_t  \\
-7)\ d R_u/dt &= \gamma I_u - F_r R_u + \omega R_n \\
-8)\ d R_n/dt &= \gamma I_n + (1-P_r) F_r R_u - \omega R_n  \\
-9)\ d R_p/dt &= \gamma I_p + P_r F_r R_u  - \omega R_p  \\
+7)\ d R_u/dt &= \gamma I_u - F_R R_u + \omega R_n \\
+8)\ d R_n/dt &= \gamma I_n + (1-p_R) F_R R_u - \omega R_n  \\
+9)\ d R_p/dt &= \gamma I_p + p_R F_R R_u  - \omega R_p  \\
 10)\ d R_t/dt&= \gamma I_t + \omega R_p  \\
@@
     ## scaling the weights
-    sigma <- tau*rho*N0/(tau*(W_s*S_u+W_i*I_u+W_r*R_u)+rho*N0)
-    F_s <- sigma*W_s
-    F_i <- sigma*W_i
-    F_r <- sigma*W_r
+    W <- W_S*S_u+W_I*I_u+W_R*R_u
+    sigma <- rho*N0/W
+    F_S <- sigma*W_S
+    F_I <- sigma*W_I
+    F_R <- sigma*W_R
 
-    dS_u.dt <- -Lambda*S_u - (1-P_s) * F_s * S_u + omega * S_n
-    dS_n.dt <- -Lambda*S_n + (1-P_s) * F_s * S_u - omega * S_n 
-    dI_u.dt <-  Lambda*S_u + omega * I_n - F_i * I_u - gamma * I_u
-    dI_n.dt <-  Lambda*S_n + (1-P_i) * F_i * I_u  - omega * I_n - gamma * I_n
-    dI_p.dt <-  P_i * F_i * I_u - omega * I_p - gamma * I_p 
+    dS_u.dt <- -Lambda*S_u - (1-p_S) * F_S * S_u + omega * S_n
+    dS_n.dt <- -Lambda*S_n + (1-p_S) * F_S * S_u - omega * S_n 
+    dI_u.dt <-  Lambda*S_u + omega * I_n - F_I * I_u - gamma * I_u
+    dI_n.dt <-  Lambda*S_n + (1-p_I) * F_I * I_u  - omega * I_n - gamma * I_n
+    dI_p.dt <-  p_I * F_I * I_u - omega * I_p - gamma * I_p 
     dI_t.dt <-  omega * I_p - gamma * I_t 
-    dR_u.dt <- gamma * I_u + omega * R_n - F_r * R_u  
-    dR_n.dt <- gamma * I_n + (1-P_r) * F_r * R_u - omega * R_n
-    dR_p.dt <- gamma * I_p + P_r * F_r * R_u - omega * R_p
+    dR_u.dt <- gamma * I_u + omega * R_n - F_R * R_u  
+    dR_n.dt <- gamma * I_n + (1-p_R) * F_R * R_u - omega * R_n
+    dR_p.dt <- gamma * I_p + p_R * F_R * R_u - omega * R_p
     dR_t.dt <- gamma * I_t + omega * R_p
@@
 ```{r model inputs, message=FALSE}
-params <- c(N0=1000000, beta=0.8,gamma=1/3, omega=0.25, rho=0.01, 
-             W_s=.01, W_i=1, W_r=1, #weights
-             P_s=0, P_i=1, P_r=0.5, #prob of waiting for being tested positive
+params <- c(N0=1000000, beta=0.5,gamma=1/3, omega=0.25, rho=0.01, 
+             W_S=1/3, W_I=1/3, W_R=1/3, #weights
+             p_S=0, p_I=1, p_R=0.5, #prob of waiting for being tested positive
             eta_w=0.02, eta_t=0.01, #isolation parameter 
-            s=2, #eta_w/eta_t
-            tau=1
+            s=2 #eta_w/eta_t
             )
@@
 # Conditional stop for desolver
-rootfun <- function (time, state,params) { 
-  unpack(as.list(c(state,params)))
-  return(I_u - 1) }
+# rootfun <- function (time, state,params) { 
+#   unpack(as.list(c(state,params)))
+#   return(I_u - 1) }
 ```
@@
   I_u <-0
-  R_u<-0
+  R_u <-0
   unpack(as.list(params))
-  sigma <- tau*rho*N0/(tau*(W_s*S_u+W_i*I_u+W_r*R_u)+rho*N0)
-  F_s <- sigma*W_s
+  W <- W_S*S_u+W_I*I_u+W_R*R_u
+  sigma <- rho*N0/W
+  F_S <- sigma*W_S
 
  return(c(eq1=S_u+S_n-N0,
-   eq2=-F_s*S_u+omega*S_n))
+          eq2=-F_S*S_u+omega*S_n))
 }
@@
 # DFE; analytic approach
-
-Su_dfe <-function(params){
+Sn_dfe <-function(params){
   unpack(as.list(params))
-  a <- rho
-  b <- -rho+(1-rho/omega)*tau*W_s
-  c <- -tau*W_s
-  d <- b^2-4*a*c
-
-  if(rho){return(N0*2*a/(-b+sqrt(d)))}
-    else {return(N0)}
+  return((rho*(1-p_S)*N0)/omega)
   }
 
-Sn_dfe<- params[["N0"]]-Su_dfe(params)
+Su_dfe<- params[["N0"]]-Sn_dfe(params)
 
-state_dfe <- c(S_u=Su_dfe(params), S_n=Sn_dfe,
+state_dfe <- c(S_u=Su_dfe, S_n=Sn_dfe(params),
                 I_u=0,I_n=0,I_p=0,I_t=0,
@@
                 N=0,P=0)
-
+# check:
 all.equal(c(state_dfe[["S_u"]],state_dfe[["S_n"]]),roots_dfe$root)
@@
 all.equal(sum(g1[[1]]),0)
-## debug(sir.model)
-## g1 <- sir.model(time=0,state=state_init, params=params)
 ```
@@
 The Disease-Free Equilibrium, namely $E_1$, for the above SIR model is given by the solution of the following system:
-$S_u+S_n=N0$ and $F_s S_u-\omega S_n=0$.
-Let $\tilde{S_u}=N_0/S_u$, the positive solution of the quadratic $$a \tilde{S_u}+b \tilde{S_u}^2+c=0,$$
-where $a=\rho$, $b= -\rho+(1-\rho/\omega) \tau W_s$ and $c=-\tau W_s$. Let
-  $d = b^2-4ac$, the DFE of the model is
+$S_u+S_n=N0$ and $F_S S_u-\omega S_n=0$. The DFE is
 
-$$S_u^* =N_0 2 a/(-b+\sqrt d), \\ S_n^*=N_0-S_u^* \\ I_j=R_j=0 \ \forall j$$
+$S_n^*= \frac{\rho}{\omega} N_0, S_u^*= N_0-S_n^*$, and $I_j=R_j=0 \ \forall j$.
 
@@
 $$R_0= (A \times S_u^* + B \times S_n^*) \times C,$$ where
-$A=\gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_t P_i) F_i$,
-$B=\big(\omega+(F_i+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_t\omega) \omega P_i F_i }{\omega+\gamma}$ and 
-$C=\frac{\beta/\gamma}{N_0 (\gamma(\omega+\gamma)+F_i(\gamma+\omega P_i))}$.
+$A=\gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_t p_I) F_I$,
+$B=\big(\omega+(F_I+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_t\omega) \omega p_I F_I }{\omega+\gamma}$ and 
+$C=\frac{\beta/\gamma}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}$.
 
@@
  \left[ \begin {array}{cccc}  
-F_i+\gamma&-\omega&0&0\\
--(1-P_i)F_i&\omega+\gamma&0&0\\
--P_i F_i&0&\omega+\gamma&0\\
+F_I+\gamma&-\omega&0&0\\
+-(1-p_I)F_I&\omega+\gamma&0&0\\
+-p_I F_I&0&\omega+\gamma&0\\
 0&0&-\omega&\gamma
@@
 ```{r R0}
-F_i <- function(params){
-   # Returns F_i at DFE 
+F_I <- function(params){
+   # Returns F_I at DFE 
   unpack(as.list(params))
-  Su_dfe<- Su_dfe(params)
-  Sn_dfe<- N0-Su_dfe
-  sigma <- tau*rho*N0/(tau*(W_s*Su_dfe)+rho*N0) #at DFE 
-  return(sigma*W_i)
+  Sn<- Sn_dfe(params)
+  Su<- N0-Sn
+  sigma <- rho*N0/(W_S*Su) #at DFE 
+  return(sigma*W_I)
 }
@@
  # just a test
-  F_i(params=update(params,rho=0))
+  F_I(params=update(params,rho=0))
 
@@
     unpack(as.list(params))
-    Su_dfe<- Su_dfe(params)
-    Sn_dfe<- N0-Su_dfe
+    Sn <- Sn_dfe(params)
+    Su <- N0-Sn
     # print(c(Su_dfe,Sn_dfe,Su_dfe+Sn_dfe))
-    Fi <- F_i(params) #at DFE
-
-    A <-gamma*(omega+gamma+eta_w*Fi) + omega*eta_t*P_i*Fi
-    B <-(omega+eta_w*(Fi+gamma))*gamma+(eta_w*gamma+eta_t*omega)*(omega*P_i*Fi)/(omega+gamma)
-
-    C <- beta/(N0*gamma*(gamma*(omega+gamma)+Fi*(gamma+omega*P_i)))
-    return((A*Su_dfe+B*Sn_dfe)*C)
+    Fi <- F_I(params) #at DFE
+    A <-gamma*(omega+gamma+eta_w*Fi) + omega*eta_t*p_I*Fi
+    B <-(omega+eta_w*(Fi+gamma))*gamma+(eta_w*gamma+eta_t*omega)*(omega*p_I*Fi)/(omega+gamma)
+    C <- beta/(N0*gamma*(gamma*(omega+gamma)+Fi*(gamma+omega*p_I)))
+    return((A*Su+B*Sn)*C)
 }
@@
     unpack(as.list(params))
-    Su_dfe<- Su_dfe(params)
-    Sn_dfe<- N0-Su_dfe
-    F_i <-F_i(params)  #at DFE
+    Sn<- Sn_dfe(params)
+    Su<- N0-Sn
+    F_I <-F_I(params)  #at DFE
 
-    A2 <- gamma*(P_i*F_i*omega*Sn_dfe+(omega+gamma)*(gamma*Sn_dfe+F_i*N0))
-    B2 <- P_i*F_i*omega*(gamma*Su_dfe+omega*N0)
-    D <- gamma*(omega+gamma)*(gamma*Su_dfe+omega*N0)
-    C2 <- beta/((N0*gamma*(gamma*(omega+gamma)+F_i*(gamma+omega*P_i)))*(omega+gamma))
+    A2 <- gamma*(p_I*F_I*omega*Sn+(omega+gamma)*(gamma*Sn+F_I*N0))
+    B2 <- p_I*F_I*omega*(gamma*Su+omega*N0)
+    D <- gamma*(omega+gamma)*(gamma*Su+omega*N0)
+    C2 <- beta/((N0*gamma*(gamma*(omega+gamma)+F_I*(gamma+omega*p_I)))*(omega+gamma))
     # s <- eta_w/eta_t
-
-    # return((eta_w*A2+eta_t*B2+D)*C2)
     return((eta_t*(s*A2+B2)+D)*C2)
@@
  eta_w <- seq(0,1, length.out=n_out)
- eta_t <- seq(0.001,0.5, length.out=n_out) #0.001 to 0.5 
+ eta_t <- seq(0,1, length.out=n_out) #0.001 to 0.5 
  rho     <- seq(0,0.01, length.out=n_in)
- omega <- seq(0.001,1 , length.out=n_in) #note omega must be > rho, was 0.1
+ omega <- seq(0.011,1 , length.out=n_in) #note omega must be > rho, was 0.1
 
  df1 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
-                    W_s=params[["W_s"]],W_i=params[["W_i"]],W_r=params[["W_r"]],
-                    P_s=params[["P_s"]],P_i=params[["P_i"]],P_r=params[["P_r"]],
-                    eta_w=eta_w,eta_t=eta_t,s=params[["s"]],tau=params[["tau"]])
+                    W_S=params[["W_S"]],W_I=params[["W_I"]],W_R=params[["W_R"]],
+                    p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
+                    eta_w=eta_w,eta_t=eta_t,s=params[["s"]])
 
-# Need to be corrected: in F_i, state_dfe should be changed according to the df1!
+# Need to be corrected: in F_I, state_dfe should be changed according to the df1!
 df2<- data.frame(df1,R0=apply(df1,1,function(params_in)R0(params=params_in)),
-                      Fi= apply(df1,1,function(params_in)F_i(params=params_in)))
+                      Fi= apply(df1,1,function(params_in)F_I(params=params_in)))
 
-# which(df2[]<0)
+which(df2[]<0)
 ```
@@
  # specify the ranges
- # beta  <- 0.5
- # eta_w <- seq(0,1, length.out=5) 
- # eta_t <- seq(0.001,0.5, length.out=n_out) #0.001 
- # rho     <- seq(0,0.01, length.out=n_in)
- # omega <- seq(0.001,1 , length.out=n_in) #note omega must be > 0, was 0.1
+eta_t <- seq(0.001,0.5, length.out=n_out)  
 s <- seq(1,2,length.out=n_out)
@@
  df11 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
-                    W_s=params[["W_s"]],W_i=params[["W_i"]],W_r=params[["W_r"]],
-                    P_s=params[["P_s"]],P_i=params[["P_i"]],P_r=params[["P_r"]],
-                    eta_w=params[["eta_w"]],eta_t=eta_t,s=s,tau=params[["tau"]])
+                    W_S=params[["W_S"]],W_I=params[["W_I"]],W_R=params[["W_R"]],
+                    p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
+                    eta_w=params[["eta_w"]],eta_t=eta_t,s=s)
 
 df22<- data.frame(df11,R0=apply(df11,1,function(params_in)R02(params=params_in)),
-                      Fi= apply(df11,1,function(params_in)F_i(params=params_in)),
-                       Su=apply(df11,1,function(params_in)Su_dfe(params = params_in)))
+                      Fi= apply(df11,1,function(params_in)F_I(params=params_in)),
+                       Sn=apply(df11,1,function(params_in)Sn_dfe(params = params_in)))
 ```
@@
 
-Let call the $F_i$ the critical testing rate in the above plot.
+Let call the $F_I$ the critical testing rate in the above plot.
 
@@
 2. $\partial{R_0}/\partial{\rho} \leq 0$; this is concluded from the Taylor expansion of $R_0$ around $\rho=0$ which is
-$R_0 \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta_w-1)(\gamma W_s+\omega W_i) + (\eta_t -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2)$.
+$R_0 \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_S} \Big(\gamma(\eta_w-1)(\gamma W_S+\omega W_I) + (\eta_t -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2)$.
 3. Based on the above Taylor expansion around $\rho=0$,
-$$\partial{R_0}/\partial{\omega}=  \frac{-\beta \rho}{\gamma W_s\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c), $$
-where $a=(\eta_w-1)W_i-(\eta_t-1)P_iW_i$, $b=2(\eta_w-1)\gamma W_s$ and $c=(\eta_w-1)\gamma^2 W_s$.
+$$\partial{R_0}/\partial{\omega}=  \frac{-\beta \rho}{\gamma W_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c), $$
+where $a=(\eta_w-1)W_I-(\eta_t-1)P_iW_i$, $b=2(\eta_w-1)\gamma W_S$ and $c=(\eta_w-1)\gamma^2 W_S$.
 Given that $\eta_t\leq \eta_w$, it is inferred that $a\geq 0$, $b\leq 0$ and $c \leq 0$.
@@
 
-4. Taylor expansion gives $R_0|_{(\rho=0,W_s=1)} = \beta/\gamma$.
+4. Taylor expansion gives $R_0|_{(\rho=0,W_S=1)} = \beta/\gamma$.
 
 5.
-Taylor expansion gives $R_0|_{(\rho=0,W_s=0)} = \beta/\gamma \frac{\gamma \tau(\omega+\gamma)W_s+\tau\rho(\gamma\eta_w+\omega\eta_tP_i)W_i+ \gamma \rho(\gamma+\omega)}{\gamma \tau(\omega+\gamma)W_s+\tau\rho(\gamma+\omega P_i)W_i+ \gamma \rho(\gamma+\omega)}$.
+Taylor expansion gives $R_0|_{(\rho=0,W_S=0)} = \beta/\gamma \frac{\gamma \tau(\omega+\gamma)W_S+\tau\rho(\gamma\eta_w+\omega\eta_tP_i)W_I+ \gamma \rho(\gamma+\omega)}{\gamma \tau(\omega+\gamma)W_S+\tau\rho(\gamma+\omega p_I)W_I+ \gamma \rho(\gamma+\omega)}$.
 

### 2020-11-05 codes/ali_test_sir.Rmd
fix parameter range W_s \neq W_I

@@
 ```{r model inputs, message=FALSE}
-params <- c(N0=1000000, beta=0.5,gamma=1/3, omega=0.25, rho=0.01, 
-             W_S=1/3, W_I=1/3, W_R=1/3, #weights
+params <- c(N0=1000000, beta=.8,gamma=1/3, omega=0.25, rho=0.01, 
+             W_S=0.5, W_I=1, W_R=1, #weights
              p_S=0, p_I=1, p_R=0.5, #prob of waiting for being tested positive

### 2020-11-05 codes/ali_test_sir.Rmd
sim when W_s=W_I

@@
 params <- c(N0=1000000, beta=.8,gamma=1/3, omega=0.25, rho=0.01, 
-             W_S=0.5, W_I=1, W_R=1, #weights
+             W_S=1, W_I=1, W_R=1, #weights
              p_S=0, p_I=1, p_R=0.5, #prob of waiting for being tested positive
@@
  # specify the ranges
- beta  <- 0.5  ## set beta high enough to allow R0>1 in worst case
+ beta  <- 0.4  ## set beta high enough to allow R0>1 in worst case
  eta_w <- seq(0,1, length.out=n_out)
@@
  rho     <- seq(0,0.01, length.out=n_in)
- omega <- seq(0.011,1 , length.out=n_in) #note omega must be > rho, was 0.1
+ omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, was 0.011
 
@@
  # specify the ranges
-eta_t <- seq(0.001,0.5, length.out=n_out)  
+beta  <- 0.35  ## set beta high enough to allow R0>1 in worst case
+ eta_w <- seq(0,1, length.out=n_out)
+ rho     <- seq(0,0.01, length.out=n_in)
+ omega <- seq(0.1,1.5 , length.out=n_in) #note omega must be > rho, was 0.011
+
+eta_t <- seq(0.001,0.01, length.out=n_out)  
 s <- seq(1,2,length.out=n_out)

### 2020-11-06 codes/ali_test_sir.Rmd
update analysis of R0 wrt omega

@@
 ```{r model inputs, message=FALSE}
-params <- c(N0=1000000, beta=.8,gamma=1/3, omega=0.25, rho=0.01, 
+params <- c(N0=1000000, beta=.8,gamma=1/3, omega=0.25, rho=0.01, #0.8, 1/3
              W_S=1, W_I=1, W_R=1, #weights
@@
 $$\partial{R_0}/\partial{\omega}=  \frac{-\beta \rho}{\gamma W_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c), $$
-where $a=(\eta_w-1)W_I-(\eta_t-1)P_iW_i$, $b=2(\eta_w-1)\gamma W_S$ and $c=(\eta_w-1)\gamma^2 W_S$.
+where $a=(\eta_w-1)W_I-(\eta_t-1)P_I W_I = ((s-p_I)\eta_t + (p_I-1)) W_I$, $b=2(\eta_w-1)\gamma W_S$ and $c=(\eta_w-1)\gamma^2 W_S$.
 Given that $\eta_t\leq \eta_w$, it is inferred that $a\geq 0$, $b\leq 0$ and $c \leq 0$.
@@
 
-4. Taylor expansion gives $R_0|_{(\rho=0,W_S=1)} = \beta/\gamma$.
+In general, $a \geq 0$ is not true.
+$$a \geq 0 \iff (s-p_I) \eta_t \geq (1-p_I),$$ where $s=\frac{\eta_w}{\eta_t} \geq 1$.
+As an example, in the case of a very accurate testing regime,  i.e., $P_I=1$, $a \geq 0$ is achieved.
 
-5.
-Taylor expansion gives $R_0|_{(\rho=0,W_S=0)} = \beta/\gamma \frac{\gamma \tau(\omega+\gamma)W_S+\tau\rho(\gamma\eta_w+\omega\eta_tP_i)W_I+ \gamma \rho(\gamma+\omega)}{\gamma \tau(\omega+\gamma)W_S+\tau\rho(\gamma+\omega p_I)W_I+ \gamma \rho(\gamma+\omega)}$.
+**Perfect isolation**; When $\eta_w=0$ and consequently $\eta_t=0$.
+This leads to $a \leq 0$, $b<0$, $c<0$, thus $\partial{R_0}/\partial{\omega} \geq 0$.
+Intuitively, returning the test results back to the awaiting individuals while they are in a perfect isolation results in an increase of $R_0$.
 
@@
 (2) A sensible result that we are expecting is $\partial{R_0}/\partial{\omega}<0$.
+
+```{r  chnages of R0 wrt omega,message=FALSE,warning=FALSE }
+R0_omega <- function(params){
+  unpack(as.list(params))
+         # a <- (eta_w-1)*W_I-(eta_t-1)*p_I*W_I
+         a <- ((s-p_I)*eta_t + (p_I-1))*W_I  #involving s
+         b <- 2*(eta_w-1)*gamma*W_S
+         c <- b*gamma/2
+
+         d <- (-beta*rho)/(gamma*W_S*omega^2*(gamma+omega)^2)*(a*omega^2 + b*omega + c)
+         return(d)
+}
+
+R0_omega(params)
+
+```
+
+```{r }
+ # specify the ranges
+if(F){
+beta  <- 0.5  ## set beta 
+gamma <- seq(0.1,0.5, length.out=n_out)
+p_I <- 0.9
+
+# eta_w <- seq(0,1, length.out=n_out)
+eta_t <- seq(0.001,1, length.out=n_out)
+s <- seq(1,2,length.out=n_out)
+
+rho <- seq(0,0.01, length.out=n_in)
+omega <- seq(0.011,2 , length.out=n_in) #note omega must be > rho, was 0.011
+
+W_S <- 1/100
+W_I <- 1/3
+W_R <- 1/3
+
+ df3_1 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=gamma,omega=omega,rho=rho,
+                    W_S=W_S,W_I=W_I,W_R=W_R,
+                    p_S=params[["p_S"]],p_I=p_I,p_R=params[["p_R"]],
+                    eta_w=params[["eta_w"]],eta_t=eta_t,s=s)
+
+df3_2 <- data.frame(df3_1,R0=apply(df3_1,1,function(params_in)R02(params=params_in)),
+                      R0_omega= apply(df3_1,1,function(params_in)R0_omega(params=params_in))
+                       )
+}
+```
+
+```{r show_contours_3, width=24,height=24,message=FALSE,warning=FALSE}
+if(F){
+p3 <- ggplot(df3_2,aes(x=omega,y=gamma,z=R0_omega))+ theme_bw() +
+    theme(panel.spacing=grid::unit(0,"lines"))
+(p3
+    + geom_contour_filled(breaks=pretty(unique(df3_2$R0_omega),5),
+                          alpha=0.5) ## breaks=0.2)
+    + geom_contour(breaks=0,alpha=0.5,colour="black")
+    + facet_grid(s~eta_t, labeller=label_special)
+    + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
+    + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
+    + scale_fill_viridis_d(name=parse(text="dR[0]"))
+)
+}
+```

### 2020-11-08 note/R0_wrt_omega.tex
Add files via upload

@@
 
-\section{Intrdoctuion}
+\section{Introduction}
 
@@
 For a necessary and sufficient condition (inequality) for having $\partial{\Rnum}/\partial{\omega} < 0$, see \cref{eq:necsuf}.
-Finally, we give a brief discussion on finding a local minimum of $\Rnum$ with respect to $w$ for $w \in (0,\infty)$.
-However, our analysis of this is complicated by the fact that we are analyzing the linearization of $\Rnum$ near $t = 0$, rather than the exact form of $\Rnum$.
-\fady{In section 3, I expand a bit on this and basically ask if it would be a good idea for me to try working with the exact form of $\Rnum$ to see if we can generalize some of our analysis for $t$ not close to 0.}
 
-\section{Behaviour of $\Rnum$ for $t \approx 0$}
+\section{Effect of speed of returning test results, $\omega$}
 
@@
 
-Now we focus our attention on minimizing $\Rnum$ wrt to $\omega$.
-Since $\Rnum(\omega)$ has a global (for $\omega \in [0,\infty)$) maximum at $\omega_2$, we know that $\Rnum$ attains its smallest values either near $\omega = 0$ or as $\omega$ approaches $+\infty$.
-We can study the behaviour of $\Rnum(\omega)$ on $(0,\infty)$, but in reality, there must be a bounded range of realistic values of $\omega$ that we should optimize on.
-However, optimizing on $(0,\infty)$ and this bounded range should be equally easy, given that $\Rnum(\omega)$ has only one critical point; so we know the minimum must occur at the boundaries of the domain.
+\section{Few high sensitivity tests vs. many low sensitivity tests}
 
-If $\Rnum$ attains its smallest values near $\omega = 0$, then testing results should never be returned (probably not a practical implementation, but it does tell us something interesting).
-On the other hand, if $\Rnum$ attains its lowest values as $\omega$ tends to infinity, then test results should be returned as quickly as possible.
-Of course, in reality, if test results were never returned, people may not continue to comply with being careful while ``awaiting" their test result; since it never comes.
+Let us consider the linearization of $\Rnum$ around $\Rnum \approx 0$.
+\fady{As Ali pointed out to me, working with the linearization is okay since $\rho$ is the \text{daily per capita testing rate}, so it is expected to be fairly small.
+I was somehow thinking values of $\rho$ weren't gonna be small, but now I understand.}
 
-Before doing any analysis, notice that ``never returning results" may be \fady{I'm not 100 \% sure} dynamically equivalent to having an ``SIR with lockdown model", except the lockdown period is permanent.
-This is because some people are lowering their contact rate while ``waiting" for their results (which will never come).
-In this case, the lowering of contact is given by the parameter $\eta_s$.
+\begin{equation}
+\Rnum \approx \beta/\gamma + \frac{\beta t}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta_w-1)(\gamma W_s+\omega W_i) + (\eta_t -1)P_iW_i \omega^2 \Big). 
+\end{equation}
 
-Looking at the limit as $\omega \rightarrow \infty$ we have
-\begin{align}
-&\lim_{w \rightarrow \infty} \beta/\gamma + \frac{\beta t}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta_w-1)(\gamma W_s+\omega W_i) + (\eta_t -1)P_iW_i \omega^2 \Big) \nonumber \\
-&= \frac{\beta}{\gamma} - \frac{\beta t\left(1-t_n\right)P_{i}W_{i}}{\gamma^{2}W_{s}}
+One can conceptualize two extremes of testing strategies: use only expensive high-sensitivity tests, but have a low testing rate as a result vs.
+use cheap low-sensitivity tests (thus, risking many false negatives), but have a high testing rate as a result.
+
+As a first step, consider a test that allows us to test at rate $\rho_1$ and has sensitivity $P_{i,1}$, and another test that allows us to test at $\rho_2$ and has sensitivity $P_{i,2}$.
+Under the assumption that $\rho_1 > \rho_2$, \cref{eq:rho1vsrho2} tells us exactly when the second test causes a lower $\Rnum$ compared to the first.
+\begin{align}\label{eq:rho1vsrho2}
+    &\rho_1\left(\gamma(\eta_w-1)(\gamma W_s + \omega W_i) + (\eta_t-1)P_{i, 1}W_i\omega^2\right) - \rho_2\left(\gamma(\eta_w-1)(\gamma W_s + \omega W_i) + (\eta_t-1)P_{i, 2}W_i\omega^2\right) > 0 \nonumber \\
+    &\vdots \nonumber \\
+    &\frac{\rho_2P_{i, 2}-\rho_1P_{i, 1} }{\rho_1-\rho_2} > \frac{1-\eta_w}{1-\eta_t}\cdot \frac{\gamma(\gamma W_s + \omega W_i)}{\omega^2 W_i}
 \end{align}
 
-Naturally, we would now want to evaluate the limit as $\omega \rightarrow 0$.
-However, if we were to naively compute this limit assuming $t$ is constant, we would get $- \infty$.
-I suspect \fady {I'm not sure about this} the reason for this is because we're working with a linearization where $t \approx 0$, thus, we can only evaluate the limit as another variable (e.g., $\omega$) goes to 0 only if we insist that $t$ also goes to 0.
-But then, the value of the limit depends on the relative decay of $t$ and $\omega$ towards 0.
-\fady{I will not write down analysis here until I've talked to someone about this.}
+The previous inequality tells us precisely when a test corresponding to $\rho_2, P_{i,2}$ will yield a lower $\Rnum$ than a test corresponding to $\rho_1, P_{i,1}$, where $\rho_1 > \rho_2$.
+It is then natural to ask, what hypothetical test corresponding to $\rho^*, P_{i,*}$ will \textit{minimize} $\Rnum$?
+In order to examine this question, we can consider $P_i$ (which is the sensitivity of the test) as a function of the testing rate, $\rho$.
+So let $P_i(\rho)$.
+We expect that $P_i(\rho)$ is decreasing.
 
-\section{Rough thoughts: what about when $t$ is not close to 0?}
+We can then compute $\frac{\partial \Rnum}{\partial \rho}$ using the chain rule. Doing so, we obtain:
+\begin{equation}
+\frac{\partial \Rnum}{\partial \rho} = \dfrac{\beta\left(\left(\eta_\text{t}-1\right)\omega^2W_iP_i't+\left(\eta_t-1\right)\omega^2W_iP_i+\left(\eta_w-1\right)W_s\gamma^2+\left(\eta_w-1\right)\omega W_i\gamma\right)}{\omega W_s\gamma^2\left(\gamma+\omega\right)}\,,
+\end{equation}
 
-Initially, I thought the exact formula for $\Rnum$ was linear with respect to $t$, which would have allowed us to generalize the results in the previous section.
-However, I remembered that during the meeting, someone (I think it was Jonathan) was asking about a quadratic term.
-I figured that someone had noticed that if $\Rnum$ was linear we would've been able to generalize the analysis above, so looking harder, I noticed that indeed $\Rnum$ is not linear wrt to $t$ because $F_i$ contain $t$'s that then get multiplied with the one explicit $t$ in the expression.
-So indeed it's not linear.
+Setting $\frac{\partial \Rnum}{\partial \rho} = 0$:
 
-I could try spending some time \emph{trying} to work on the exact $\Rnum$ function to see if I can get anywhere with it, if that's of interest.
+\begin{align}\label{eq:r0rho=0}
+    &\frac{\partial \Rnum(\rho^*)}{\partial \rho} = 0 \nonumber \\
+    &\Longleftrightarrow P_i'(\rho^*) = \frac{1}{\rho^*} \left( \frac{1-\eta_w}{1-\eta_t}\cdot\frac{\gamma}{\omega}\cdot\big(\frac{\gamma}{\omega}+\frac{W_i}{W_s}\big) - P_i(\rho^*)\right)
+\end{align}
+
+In order to check if this critical point is a minimum, we can check the sign of the second derivative.
+For convenience, let $Q = \frac{1-\eta_w}{1-\eta_t}\cdot\frac{\gamma}{\omega}\cdot\big(\frac{\gamma}{\omega}+\frac{W_i}{W_s}\big)$
+
+\begin{align}
+    P_i''(\rho^*) &= \frac{-1}{{\rho^*}^2}(Q-P_i(\rho^*)) -\frac{1}{{\rho^*}^2}P_i'(\rho^*) \\
+    &= -\frac{1}{{\rho^*}}P_i'(\rho^*) -\frac{1}{{\rho^*}^2}P_i'(\rho^*) > 0
+\end{align}
+Where the final inequality follows since $\rho^* > 0$ and $P_i'(\rho^*) < 0$. So indeed, the critical point we found is a minimum.
+
+To gain some intuition, let us assume that $P_i(\rho)$ is of the form $1-k\rho$ for some $k > 0$. \fady{to be continued...}
 

### 2020-11-09 note/R0_wrt_omega.tex
updated notes

@@
 
-Let us consider the linearization of $\Rnum$ around $\Rnum \approx 0$.
+Let us consider the linearization of $\Rnum$ around $\rho \approx 0$.
 \fady{As Ali pointed out to me, working with the linearization is okay since $\rho$ is the \text{daily per capita testing rate}, so it is expected to be fairly small.
@@
 For convenience, let $Q = \frac{1-\eta_w}{1-\eta_t}\cdot\frac{\gamma}{\omega}\cdot\big(\frac{\gamma}{\omega}+\frac{W_i}{W_s}\big)$
+AGh: Call $Q$ something else to avoid confussion with the Q in eq 2.
 

### 2020-11-09 codes/ali_test_sir.Rmd
some insights updated

@@
  # specify the ranges
- beta  <- 0.4  ## set beta high enough to allow R0>1 in worst case
+ beta  <- 0.339  ## set beta high enough to allow R0>1 in worst case
  eta_w <- seq(0,1, length.out=n_out)
@@
  omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, was 0.011
+ W_S <- .3 #choose 0.3 for TTI testing plan and 1 for the random testing.
 
  df1 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
-                    W_S=params[["W_S"]],W_I=params[["W_I"]],W_R=params[["W_R"]],
+                    W_S=W_S,W_I=params[["W_I"]],W_R=params[["W_R"]],
                     p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
@@
  # specify the ranges
-beta  <- 0.35  ## set beta high enough to allow R0>1 in worst case
- eta_w <- seq(0,1, length.out=n_out)
+beta  <- 0.339  ## set beta high enough to allow R0>1 in worst case 
 rho     <- seq(0,0.01, length.out=n_in)
- omega <- seq(0.1,1.5 , length.out=n_in) #note omega must be > rho, was 0.011
-
-eta_t <- seq(0.001,0.01, length.out=n_out)  
-s <- seq(1,2,length.out=n_out)
+omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, 
+eta_t <- seq(0.001,0.1, length.out=n_out)
+s <- seq(1,10,length.out=n_out)
 
  df11 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
-                    W_S=params[["W_S"]],W_I=params[["W_I"]],W_R=params[["W_R"]],
+                    W_S=W_S,W_I=params[["W_I"]],W_R=params[["W_R"]],
                     p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
@@
 1. $\partial{R_0}/\partial{\eta_t} \geq 0$ and $\partial{R_0}/\partial{\eta_w} \geq 0$.
+That is, the lower isolation the higher $R_0$. Note that perfect isolation occurs at $\eta_w = 0$. 
 2. $\partial{R_0}/\partial{\rho} \leq 0$; this is concluded from the Taylor expansion of $R_0$ around $\rho=0$ which is
@@
 
-**Interpretation of the R0 contour plot**
+**Insights from the R0-contour plots**
 
-Note that $\eta_t\leq \eta_w$, i.e., reported people have a lower transmission probability than awaiting individuals.
-The perfect isolation/quarantine occurs when $\eta_w=\eta_t=0$ (see the upper left corner panel).
-Thus, we only consider the lower-triangle panels in the plot [not sure how to extract them in ggplot?].
-(1) Looking at the first column of panels, there seems to be a compensatory relation between $\rho$ and $\omega$.
-i.e., higher $\rho$ requires smaller $\omega$ to keep R0 lower than 1.
-Also, for a fixed $\omega$, increasing $\eta_w$ requires higher $\rho$ to achieve the critical R0.
-(2) A sensible result that we are expecting is $\partial{R_0}/\partial{\omega}<0$.
+1. TTI requires lower testing rate than random testing to achieve $R_0=1$.
+
+2. When random testing regime is applied, $W_S=W_I=W_R$,
+
+(i) in the case of perfect isolation of the reported individuals, $\eta_t=0$, the followings are inferred (see the first few columns of the panel plot); 
+ - variations of the $\eta_w$ has a negligible effect on the critical testing.
+ - for low $\omega$, $R_0$ increases as $\omega$ increases but saturates.
+This means that when the reporting process is very very slow, increasing the speed of the reporting causes higher $R_0$.
+Maybe because more people will move to $I_u$ [?, not very sure here.]
+ (ii) in case of the non-perfect isolation (see the last column), when $\eta_w$ increases (more awaiting people follow the isolation rules) faster reporting reduces $R_0$.  
+Note that it is sensible to expect that $\partial{R_0}/\partial{\omega}<0$. Here we showed that this can be less true.
+
+3. When TTI regime is applied,
 

### 2020-11-10 codes/ali_test_sir.Rmd
update plots and prep notes

@@
  omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, was 0.011
- W_S <- .3 #choose 0.3 for TTI testing plan and 1 for the random testing.
+ W_S <- 0.3 #choose 0.3 for TTI testing plan and 1 for the random testing.
 
@@
     theme(panel.spacing=grid::unit(0,"lines"))
-(p1
+p1_temp <- (p1
     + geom_contour_filled(breaks=pretty(unique(df2$R0),5),
@@
     + scale_fill_viridis_d(name=parse(text="R[0]"))
+    + ggtitle("W[S]=0.3, W[I]=W[R]=1")
 )
+p1_temp
+# filename = "R0contour_TTI.png"
+# filename = "R0contour_random.png"
+ggsave(p1_temp,filename = "R0contour_TTI.png" , path="../pix/")
 ```
@@
     + scale_fill_viridis_d(name=parse(text="R[0]"))
+    + ggtitle("W[S]=0.3, W[I]=W[R]=1")
 )

### 2020-11-10 note/SIR_notes.Rnw
update analysis of R0 wrt omega
update plots and prep notes

@@
 \usepackage{hyperref}
+\usepackage{graphicx}
+\usepackage{subcaption}
 \bibliographystyle{apalike}
@@
 \section{Method}
+- $\Lambda$: force of infection defined as $$\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_t I_t)}{N_0},$$ where $\beta$ is transmission rate, $\eta_w$ and $\eta_t$ are the isolation parameters for awaiting and reported individuals, respectively.
+We assume that $\eta_t<\eta_w$, i.e., the awaiting individuals for test results have a higher transmission probability than the reported individuals.
+- $\omega$: the rate of *onward flow* from the awaiting positive compartment, $p$, to reported/tested compartment, $\rho$, or from awaiting negative compartment, $n$, back to $u$.
+It has units of $1/time$.
+- $\gamma$: recovery rate ($1/time$).
+- $\rho$: per capita testing intensity across the whole population ($1/time$).
+- $W$: weighted number of people available for tests, defined as $W = W_S S_u + W_I I_u + W_R R_u$.
+- $\sigma$: scaling parameter for testing defined as $\sigma = \frac{\rho N_0}{W}$.
+- $F_Z$: Weighted testing rate defined as $F_Z=\sigma W_Z$. That is, $F_S = \sigma W_S$, $F_I=\sigma W_I$ and $F_R = \sigma W_R$.
+
+\includegraphics[width=\linewidth]{../pix/sir_comp2.png}
 
 % %%%%%%%
-\section{Analysis}
+\section{Results}
 
+\newpage
+\begin{figure}
+\centering
+\begin{subfigure}[t]{.45\textwidth}
+\centering
+\includegraphics[width=\linewidth]{../pix/R0contour_random.png}
+        \caption{}\label{fig:fig_a}
+\end{subfigure}
+%
+\begin{subfigure}[t]{.45\textwidth}
+\centering
+\includegraphics[width=\linewidth]{../pix/R0contour_TTI.png}
+\caption{}\label{fig:fig_b}
+\end{subfigure}
+\end{figure}
+
+% %%%%%%%
+\section{Thoughts/Objectives}
+
+\begin{enumerate}
+\item aa
+\end{enumerate}
+
+% %%%%%%%
 \section{Literature Review}
 
-% %%%%%%%
 \subsection{Explicit models of TTI (trace/test/isolate) based on network or agent-based models}
@@
 \citep{elbanna2020entry} [seems similar to MacPan model!]
+
+\citep{de2020influenza} Was discussed in the Math 747 
+SEIR Asymptomatic and symptomatic $I_1, I_2$. Used linear chain trick 
+Stringency index as a control force lowering $\beta$.
+
+\citep{rice2020effect} Effect of school closures on mortality. Reproduce Report 9 results by spatial agent based CovidSim. 
 % %%%%%%%

### 2020-11-10 codes/ali_test_sir.Rmd
updated notes

@@
  omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, was 0.011
- W_S <- 0.3 #choose 0.3 for TTI testing plan and 1 for the random testing.
+ W_S <- 1 #choose 0.3 for TTI testing plan and 1 for the random testing.
 
@@
     theme(panel.spacing=grid::unit(0,"lines"))
-(p11
+p2_temp <- (p11
     + geom_contour_filled(breaks=pretty(unique(df22$R0),5),
@@
     + scale_fill_viridis_d(name=parse(text="R[0]"))
-    + ggtitle("W[S]=0.3, W[I]=W[R]=1")
+    + ggtitle("W[S]=1, W[I]=W[R]=1")
 )
+p2_temp
+ggsave(p2_temp,filename = "R0contour_random_s.png" , path="../pix/")
 ```

### 2020-11-10 note/SIR_notes.Rnw
updated notes

@@
 \usepackage{subcaption}
+\usepackage{amsmath}
 \bibliographystyle{apalike}
@@
 \section{Method}
-- $\Lambda$: force of infection defined as $$\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_t I_t)}{N_0},$$ where $\beta$ is transmission rate, $\eta_w$ and $\eta_t$ are the isolation parameters for awaiting and reported individuals, respectively.
+
+\subsection{model and parameters}
+\begin{itemize}
+\item $\Lambda$: force of infection defined as $$\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_t I_t)}{N_0},$$ where $\beta$ is transmission rate, $\eta_w$ and $\eta_t$ are the isolation parameters for awaiting and reported individuals, respectively.
 We assume that $\eta_t<\eta_w$, i.e., the awaiting individuals for test results have a higher transmission probability than the reported individuals.
-- $\omega$: the rate of *onward flow* from the awaiting positive compartment, $p$, to reported/tested compartment, $\rho$, or from awaiting negative compartment, $n$, back to $u$.
+\item $\omega$: the rate of *onward flow* from the awaiting positive compartment, $p$, to reported/tested compartment, $\rho$, or from awaiting negative compartment, $n$, back to $u$.
 It has units of $1/time$.
-- $\gamma$: recovery rate ($1/time$).
-- $\rho$: per capita testing intensity across the whole population ($1/time$).
-- $W$: weighted number of people available for tests, defined as $W = W_S S_u + W_I I_u + W_R R_u$.
-- $\sigma$: scaling parameter for testing defined as $\sigma = \frac{\rho N_0}{W}$.
-- $F_Z$: Weighted testing rate defined as $F_Z=\sigma W_Z$. That is, $F_S = \sigma W_S$, $F_I=\sigma W_I$ and $F_R = \sigma W_R$.
+\item $\gamma$: recovery rate ($1/time$).
+\item $\rho$: per capita testing intensity across the whole population ($1/time$).
+\item $W$: weighted number of people available for tests, defined as $W = W_S S_u + W_I I_u + W_R R_u$.
+\item $\sigma$: scaling parameter for testing defined as $\sigma = \frac{\rho N_0}{W}$.
+\item $F_Z$: Weighted testing rate defined as $F_Z=\sigma W_Z$. That is, $F_S = \sigma W_S$, $F_I=\sigma W_I$ and $F_R = \sigma W_R$.
+\end{itemize}
 
 \includegraphics[width=\linewidth]{../pix/sir_comp2.png}
+
+The model is
+
+\begin{align}
+\label{mod:1}
+ d S_u/dt &= -\Lambda S_u - F_S S_u + \omega S_n \\
+ d S_n/dt &= -\Lambda S_n + F_S S_u - \omega S_n \\
+ d I_u/dt &= \Lambda S_u - F_I I_u + \omega I_n  - \gamma I_u  \\
+ d I_n/dt &= \Lambda S_n + (1-p_I) F_I I_u - \omega I_n -\gamma I_n \\
+ d I_p/dt &= p_I F_I I_u - \omega I_p -\gamma I_p \\
+ d I_t/dt &= \omega I_p - \gamma I_t  \\
+ d R_u/dt &= \gamma I_u - F_R R_u + \omega R_n \\
+ d R_n/dt &= \gamma I_n + (1-p_R) F_R R_u - \omega R_n  \\
+ d R_p/dt &= \gamma I_p + p_R F_R R_u  - \omega R_p  \\
+ d R_t/dt&= \gamma I_t + \omega R_p  \\
+ dN/dt &= \omega (S_n + I_n + R_n)   \\
+ dP/dt &= \omega(I_p + R_p).
+\end{align}
 
@@
 
+\begin{figure}
+\centering
+\begin{subfigure}[t]{.45\textwidth}
+\centering
+\includegraphics[width=\linewidth]{../pix/R0contour_random_s.png}
+        \caption{}\label{fig:fig_a}
+\end{subfigure}
+%
+\begin{subfigure}[t]{.45\textwidth}
+\centering
+\includegraphics[width=\linewidth]{../pix/R0contour_TTI_s.png}
+\caption{}\label{fig:fig_b}
+\end{subfigure}
+\end{figure}
+
 % %%%%%%%
@@
 \begin{enumerate}
-\item aa
+\item 
+\begin{itemize}
+\item analysis of the basic SIR model with testing: analytical results as far as possible, supplemented by numerical results as necessary/for illustration.
+Heuristic explanations/insights of what we learn from this.
+\item Maybe??
+order-of-magnitude/back-of-the-envelope calculations of the magnitudes of testing and isolation necessary.
+(However, the model may be too simplified for even back-of-the-envelope calculations to be meaningful.)
+\item Maybe?? comparison with strength-and-speed framework (since T/T/I is essentially a speed-based intervention)
+\end{itemize}
+
+\item
+\begin{itemize}
+\item extension to a SEPAIR model (i.e. including latent/exposed, presymptomatic, asymptomatic)
+  (i) analytical results will probably be out of reach
+	(ii) but it may be possible to extend some of the insights gained in part 1
+	(iii) in any case, some numerical results (the space of relevant/interesting parameters will be larger: figure out a sensible way to show results from this parameter space.
+Are there useful dimensionless parameters that could help frame this?)
+\item more detailed quantitative exploration: in particular, *either* Latin hypercube *or* (nearly the same!) a sample over estimated ('prior') distributions of the parameters.
+Goal: answer the general question "how likely is TTI to be able to control COVID?"
+\end{itemize}
 \end{enumerate}
+
+% \item extension to a SEPAIR model (i.e. including latent/exposed, presymptomatic, asymptomatic)
+%     * analytical results will probably be out of reach
+% 	* but it _may_ be possible to extend some of the insights gained in part 1
+% 	* in any case, some numerical results (the space of relevant/interesting parameters will be larger: figure out a sensible way to show results from this parameter space.
+Are there useful dimensionless parameters that could help frame this?)
+% \item more detailed quantitative exploration: in particular, *either* Latin hypercube *or* (nearly the same!) a sample over estimated ('prior') distributions of the parameters.
+Goal: answer the general question "how likely is TTI to be able to control COVID?"
 

### 2020-11-10 note/SIR_notes.Rnw

@@
 \end{figure}
+
+**Insights from the R0-contour plots**
+
+1. TTI requires lower testing rate than random testing to achieve $R_0=1$.
+
+2. When random testing regime is applied, $W_S=W_I=W_R$,
+
+(i) in the case of perfect isolation of the reported individuals, $\eta_t=0$, the followings are inferred (see the first few columns of the panel plot); 
+ - variations of the $\eta_w$ has a negligible effect on the critical testing.
+ - for low $\omega$, $R_0$ increases as $\omega$ increases but saturates.
+This means that when the reporting process is very very slow, increasing the speed of the reporting causes higher $R_0$.
+Maybe because more people will move to $I_u$ [?, not very sure here.]
+
+ (ii) in case of the non-perfect isolation (see the last column), when $\eta_w$ increases (more awaiting people follow the isolation rules) faster reporting reduces $R_0$.  
+Note that it is sensible to expect that $\partial{R_0}/\partial{\omega}<0$. Here we showed that this can be less true.
+
+3. When TTI regime is applied,
 

### 2020-11-10 note/SIR_notes.Rnw
bib cleanup, spelling, etc

@@
 \item $\omega$: the rate of *onward flow* from the awaiting positive compartment, $p$, to reported/tested compartment, $\rho$, or from awaiting negative compartment, $n$, back to $u$.
-It has units of $1/time$.
-\item $\gamma$: recovery rate ($1/time$).
-\item $\rho$: per capita testing intensity across the whole population ($1/time$).
+It has units of 1/time.
+\item $\gamma$: recovery rate (1/time).
+\item $\rho$: per capita testing intensity across the whole population (1/time).
 \item $W$: weighted number of people available for tests, defined as $W = W_S S_u + W_I I_u + W_R R_u$.
@@
 \cite{bergstrom2020frequency}
-(1) Model, assumptions: They developed a function, namely expected expoisour $E(C,\tau)$, to approximate trade-offs between the frequency of testing, n, the sensitivity of testing, q, and the delay between
-testing and results, d. This function is explicitly derived and was connected the effective reproduction number $R=R_0 S$, where $S$ is the proportion of population susciptable.
+(1) Model, assumptions: They developed a function, namely expected exposure $E(C,\tau)$, to approximate trade-offs between the frequency of testing, n, the sensitivity of testing, q, and the delay between
+testing and results, d. This function is explicitly derived and was connected the effective reproduction number $R=R_0 S$, where $S$ is the proportion of population susceptible.
 assumption that transmission rates are a step function: individuals who
@@
 \citep{arino2020simple} developed a SLIAR compartmental model to study the spread of an epidemic, specifically COVID-19, in a population.
-The model incorporates an Erlang distribution of times of sojourn inincubating, symptomatically and asymptomatically infectious compartments.
+The model incorporates an Erlang distribution of times of sojourn in incubating, symptomatically and asymptomatically infectious compartments.
 Basic reproduction number is derived.

### 2020-11-12 SIR_manuscript.tex
attemp to use Make!

--- 2020-11-10 note/SIR_notes.Rnw
@@
 \maketitle
-
-% %%%%%%%
-\section{Method}
-
-\subsection{model and parameters}
-\begin{itemize}
-\item $\Lambda$: force of infection defined as $$\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_t I_t)}{N_0},$$ where $\beta$ is transmission rate, $\eta_w$ and $\eta_t$ are the isolation parameters for awaiting and reported individuals, respectively.
-We assume that $\eta_t<\eta_w$, i.e., the awaiting individuals for test results have a higher transmission probability than the reported individuals.
-\item $\omega$: the rate of *onward flow* from the awaiting positive compartment, $p$, to reported/tested compartment, $\rho$, or from awaiting negative compartment, $n$, back to $u$.
-It has units of $1/time$.
-\item $\gamma$: recovery rate ($1/time$).
-\item $\rho$: per capita testing intensity across the whole population ($1/time$).
-\item $W$: weighted number of people available for tests, defined as $W = W_S S_u + W_I I_u + W_R R_u$.
-\item $\sigma$: scaling parameter for testing defined as $\sigma = \frac{\rho N_0}{W}$.
-\item $F_Z$: Weighted testing rate defined as $F_Z=\sigma W_Z$. That is, $F_S = \sigma W_S$, $F_I=\sigma W_I$ and $F_R = \sigma W_R$.
-\end{itemize}
-
-\includegraphics[width=\linewidth]{../pix/sir_comp2.png}
-
-The model is
-
-\begin{align}
-\label{mod:1}
- d S_u/dt &= -\Lambda S_u - F_S S_u + \omega S_n \\
- d S_n/dt &= -\Lambda S_n + F_S S_u - \omega S_n \\
- d I_u/dt &= \Lambda S_u - F_I I_u + \omega I_n  - \gamma I_u  \\
- d I_n/dt &= \Lambda S_n + (1-p_I) F_I I_u - \omega I_n -\gamma I_n \\
- d I_p/dt &= p_I F_I I_u - \omega I_p -\gamma I_p \\
- d I_t/dt &= \omega I_p - \gamma I_t  \\
- d R_u/dt &= \gamma I_u - F_R R_u + \omega R_n \\
- d R_n/dt &= \gamma I_n + (1-p_R) F_R R_u - \omega R_n  \\
- d R_p/dt &= \gamma I_p + p_R F_R R_u  - \omega R_p  \\
- d R_t/dt&= \gamma I_t + \omega R_p  \\
- dN/dt &= \omega (S_n + I_n + R_n)   \\
- dP/dt &= \omega(I_p + R_p).
-\end{align}
-
-% %%%%%%%
-\section{Results}
-
-\newpage
-\begin{figure}
-\centering
-\begin{subfigure}[t]{.45\textwidth}
-\centering
-\includegraphics[width=\linewidth]{../pix/R0contour_random.png}
-        \caption{}\label{fig:fig_a}
-\end{subfigure}
-%
-\begin{subfigure}[t]{.45\textwidth}
-\centering
-\includegraphics[width=\linewidth]{../pix/R0contour_TTI.png}
-\caption{}\label{fig:fig_b}
-\end{subfigure}
-\end{figure}
-
-\begin{figure}
-\centering
-\begin{subfigure}[t]{.45\textwidth}
-\centering
-\includegraphics[width=\linewidth]{../pix/R0contour_random_s.png}
-        \caption{}\label{fig:fig_a}
-\end{subfigure}
-%
-\begin{subfigure}[t]{.45\textwidth}
-\centering
-\includegraphics[width=\linewidth]{../pix/R0contour_TTI_s.png}
-\caption{}\label{fig:fig_b}
-\end{subfigure}
-\end{figure}
-
-% %%%%%%%
-\section{Thoughts/Objectives}
-
-\begin{enumerate}
-\item 
-\begin{itemize}
-\item analysis of the basic SIR model with testing: analytical results as far as possible, supplemented by numerical results as necessary/for illustration.
-Heuristic explanations/insights of what we learn from this.
-\item Maybe??
-order-of-magnitude/back-of-the-envelope calculations of the magnitudes of testing and isolation necessary.
-(However, the model may be too simplified for even back-of-the-envelope calculations to be meaningful.)
-\item Maybe?? comparison with strength-and-speed framework (since T/T/I is essentially a speed-based intervention)
-\end{itemize}
-
-\item
-\begin{itemize}
-\item extension to a SEPAIR model (i.e. including latent/exposed, presymptomatic, asymptomatic)
-  (i) analytical results will probably be out of reach
-	(ii) but it may be possible to extend some of the insights gained in part 1
-	(iii) in any case, some numerical results (the space of relevant/interesting parameters will be larger: figure out a sensible way to show results from this parameter space.
-Are there useful dimensionless parameters that could help frame this?)
-\item more detailed quantitative exploration: in particular, *either* Latin hypercube *or* (nearly the same!) a sample over estimated ('prior') distributions of the parameters.
-Goal: answer the general question "how likely is TTI to be able to control COVID?"
-\end{itemize}
-\end{enumerate}
-
-% \item extension to a SEPAIR model (i.e. including latent/exposed, presymptomatic, asymptomatic)
-%     * analytical results will probably be out of reach
-% 	* but it _may_ be possible to extend some of the insights gained in part 1
-% 	* in any case, some numerical results (the space of relevant/interesting parameters will be larger: figure out a sensible way to show results from this parameter space.
-Are there useful dimensionless parameters that could help frame this?)
-% \item more detailed quantitative exploration: in particular, *either* Latin hypercube *or* (nearly the same!) a sample over estimated ('prior') distributions of the parameters.
-Goal: answer the general question "how likely is TTI to be able to control COVID?"
+test $x^2$
 
@@
 
-\citep{endo2020implication} [Ali: It seems to me that this is just a statistical model to estimate the parent-offspring of an infected index, not sure if it fits into agent-based group!] Used simulation on a branching process model to assess the forward and backward contact tracing efficiency.
+% \citep{endo2020implication} [Ali: It seems to me that this is just a statistical model to estimate the parent-offspring of an infected index, not sure if it fits into agent-based group!] Used simulation on a branching process model to assess the forward and backward contact tracing efficiency.
 Assuming a negative-binomial branching process with a mean R, reproduction number, and overdispersion parameter k, the mean total number of generation G3 and averted G3 are estimated.
@@
 
-\citep{jenness2020modeling} developed a network-based transmission model for SARS-CoV-2 on the Diamond Princess outbreak to characterize transmission dynamics and to estimate the epidemiological impact of outbreak control and prevention measures.
-
-\citep{elbanna2020entry} [seems similar to MacPan model!]
-
-\citep{de2020influenza} Was discussed in the Math 747 
-SEIR Asymptomatic and symptomatic $I_1, I_2$. Used linear chain trick 
-Stringency index as a control force lowering $\beta$.
-
-\citep{rice2020effect} Effect of school closures on mortality. Reproduce Report 9 results by spatial agent based CovidSim. 
-% %%%%%%%
-\subsection{Models of repeated random testing of isolated populations}
-\cite{bergstrom2020frequency}
-(1) Model, assumptions: They developed a function, namely expected expoisour $E(C,\tau)$, to approximate trade-offs between the frequency of testing, n, the sensitivity of testing, q, and the delay between
-testing and results, d. This function is explicitly derived and was connected the effective reproduction number $R=R_0 S$, where $S$ is the proportion of population susciptable.
-assumption that transmission rates are a step function: individuals who
-have COVID go from non-infectious to fully infectious instantaneously,
-and remain fully infectious until they are no longer able to transmit disease. Test sensitivity takes the same form over the course of infection.
-More sophisticated models could allow varying infectiousness and varying
-sensitivity over time, as in 
-\citep{larremore2020test}.
-
-\citep{lopman2020model} Used a Deterministic SEIR model, incorporated TTI, applicable to a university setting. They assumed a fairly high reproductive number that is not reduced through social
-distancing measures. They found that community-introduction of SARS-CoV-2 infection onto campus can be
-relatively controlled with effective testing, isolation, contract tracing and quarantine.
-
-\citep{tuite2020mathematical} used an age-structured compartmental model of COVID-19 transmission in the population of Ontario, Canada.
-We compared a base case with limited testing, isolation and quarantine to different scenarios.
-
-% %%%%%%%
-\subsection{Other maybe-related works}
-\citep{arino2020simple} developed a SLIAR compartmental model to study the spread of an epidemic, specifically COVID-19, in a population.
-The model incorporates an Erlang distribution of times of sojourn inincubating, symptomatically and asymptomatically infectious compartments.
-Basic reproduction number is derived.
-Also, sensitivity analysis with respect to the underlying parameters for the following two outputs was carried out; (i) the number of observable cases during the course of the epidemic and at the peak, and (ii) the timing of the peak of the outbreak.
-Sensitivity analysis is performed using the R package multisensi.
-
-\citep{ruszkiewicz2020diagnosis} novel with-in-a-minute breath testing with 80\% accuracy. 
 % %%%%%%%
 \bibliography{../SIRlibrary}
+
 \end{document}

### 2020-11-12 SIR_manuscript.tex
trying to get Make work

@@
 
-% \citep{endo2020implication} [Ali: It seems to me that this is just a statistical model to estimate the parent-offspring of an infected index, not sure if it fits into agent-based group!] Used simulation on a branching process model to assess the forward and backward contact tracing efficiency.
+\citep{endo2020implication} [Ali: It seems to me that this is just a statistical model to estimate the parent-offspring of an infected index, not sure if it fits into agent-based group!] Used simulation on a branching process model to assess the forward and backward contact tracing efficiency.
 Assuming a negative-binomial branching process with a mean R, reproduction number, and overdispersion parameter k, the mean total number of generation G3 and averted G3 are estimated.

### 2020-11-13 SIR_manuscript.tex
texdeps rules and debugging
updated manuscript

@@
 
-\title{SIR Model with Testing and Isolation Mechanisms}
+\title{Testing Efficacy; Insights from a Simple Epidemic Model }
 
@@
 \maketitle
-test $x^2$
+
+% %%%%%%%
+\section{Introduction}
+
+- Why non-random testing/testing-tracing-isolation (TTI) is more efficent in controling an epidemic event?
+- Literature reviwe, evidences of efficacy of TTI
+- Simple ODE models provide useful insights (some classic examples, Kermack-Makendric,...)
+Historically, Ronald Ross (I think) developed a series of mathematical models to study the spread of malaria and implementation of control. His work contributed to the modern theory of the control of
+infectious disease (ref: Ross R. The Prevention of Malaria, 2nd edn. London: Murray, 1911).
+
+- Covid-19 and chalanges, questions to be addressed with a simple model.
+This study provides a framework/platform to study a pandamic in a population with a determinstic system of differential equations with testing mechinisms and isolation.
+
+- Our work; we developed a deterministic compartmental model which incorporates testing strategies and isolation mechanism. 
+- The basic reproduction number was derived analytically as a threshold of the dynamics. 
+- Our simulation shows ...
 
@@
 \subsection{Explicit models of TTI (trace/test/isolate) based on network or agent-based models}
-
 \citep{endo2020implication} [Ali: It seems to me that this is just a statistical model to estimate the parent-offspring of an infected index, not sure if it fits into agent-based group!] Used simulation on a branching process model to assess the forward and backward contact tracing efficiency.
@@
 
+\citep{jenness2020modeling} developed a network-based transmission model for SARS-CoV-2 on the Diamond Princess outbreak to characterize transmission dynamics and to estimate the epidemiological impact of outbreak control and prevention measures.
+
+\citep{elbanna2020entry} [seems similar to MacPan model!]
+
+\citep{de2020influenza} Was discussed in the Math 747 
+SEIR Asymptomatic and symptomatic $I_1, I_2$. Used linear chain trick 
+Stringency index as a control force lowering $\beta$.
+
+\citep{rice2020effect} Effect of school closures on mortality. Reproduce Report 9 results by spatial agent based CovidSim. 
 % %%%%%%%
-\bibliography{../SIRlibrary}
+\subsection{Models of repeated random testing of isolated populations}
+\cite{bergstrom2020frequency}
+(1) Model, assumptions: They developed a function, namely expected exposure $E(C,\tau)$, to approximate trade-offs between the frequency of testing, n, the sensitivity of testing, q, and the delay between
+testing and results, d. This function is explicitly derived and was connected the effective reproduction number $R=R_0 S$, where $S$ is the proportion of population susceptible.
+assumption that transmission rates are a step function: individuals who
+have COVID go from non-infectious to fully infectious instantaneously,
+and remain fully infectious until they are no longer able to transmit disease. Test sensitivity takes the same form over the course of infection.
+More sophisticated models could allow varying infectiousness and varying
+sensitivity over time, as in 
+\citep{larremore2020test}.
+
+\citep{lopman2020model} Used a Deterministic SEIR model, incorporated TTI, applicable to a university setting. They assumed a fairly high reproductive number that is not reduced through social
+distancing measures. They found that community-introduction of SARS-CoV-2 infection onto campus can be
+relatively controlled with effective testing, isolation, contract tracing and quarantine.
+
+\citep{tuite2020mathematical} used an age-structured compartmental model of COVID-19 transmission in the population of Ontario, Canada.
+We compared a base case with limited testing, isolation and quarantine to different scenarios.
+% %%%%%%%
+\subsection{Other maybe-related works}
+\citep{arino2020simple} developed a SLIAR compartmental model to study the spread of an epidemic, specifically COVID-19, in a population.
+The model incorporates an Erlang distribution of times of sojourn in incubating, symptomatically and asymptomatically infectious compartments.
+Basic reproduction number is derived.
+Also, sensitivity analysis with respect to the underlying parameters for the following two outputs was carried out; (i) the number of observable cases during the course of the epidemic and at the peak, and (ii) the timing of the peak of the outbreak.
+Sensitivity analysis is performed using the R package multisensi.
+
+\citep{ruszkiewicz2020diagnosis} novel with-in-a-minute breath testing with 80\% accuracy.
+
+% %%%%%%%
+\section{Method}
+\begin{itemize}
+\item $\Lambda$: force of infection defined as $$\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_t I_t)}{N_0},$$ where $\beta$ is transmission rate, $\eta_w$ and $\eta_t$ are the isolation parameters for awaiting and reported individuals, respectively.
+We assume that $\eta_t \leq \eta_w$, i.e., the awaiting individuals for test results have a higher transmission probability than the reported individuals.
+\item $\omega$: the rate of onward flow from the awaiting positive compartment, $p$, to reported/tested compartment, $\rho$, or from awaiting negative compartment, $n$, back to $u$.
+It has units of 1/time.
+\item $\gamma$: recovery rate (1/time).
+\item $\rho$: per capita testing intensity across the whole population (1/time).
+\item $W$: weighted number of people available for tests, defined as $W = W_S S_u + W_I I_u + W_R R_u$.
+\item $\sigma$: scaling parameter for testing defined as $\sigma = \frac{\rho N_0}{W}$.
+\item $F_Z$: Weighted testing rate defined as $F_Z=\sigma W_Z$. That is, $F_S = \sigma W_S$, $F_I=\sigma W_I$ and $F_R = \sigma W_R$.
+\end{itemize}
+
+\includegraphics[width=\linewidth]{./pix/sir_comp.png}
+
+The model is
+
+\begin{align}
+\label{model}
+ d S_u/dt &= -\Lambda S_u - F_S S_u + \omega S_n \\
+ d S_n/dt &= -\Lambda S_n + (1-p_S) F_S S_u - \omega S_n \\
+ d S_p/dt &= -\Lambda S_p + p_S F_S S_u - \omega S_p \\
+ d S_t/dt &= -\Lambda S_t + \omega S_p \\
+ d I_u/dt &= \Lambda S_u - F_I I_u + \omega I_n  - \gamma I_u  \\
+ d I_n/dt &= \Lambda S_n + (1-p_I) F_I I_u - \omega I_n -\gamma I_n \\
+ d I_p/dt &= p_I F_I I_u - \omega I_p -\gamma I_p \\
+ d I_t/dt &= \omega I_p - \gamma I_t  \\
+ d R_u/dt &= \gamma I_u - F_R R_u + \omega R_n \\
+ d R_n/dt &= \gamma I_n + (1-p_R) F_R R_u - \omega R_n  \\
+ d R_p/dt &= \gamma I_p + p_R F_R R_u  - \omega R_p  \\
+ d R_t/dt&= \gamma I_t + \omega R_p  \\
+ dN/dt &= \omega (S_n + I_n + R_n)   \\
+ dP/dt &= \omega(I_p + R_p).
+\end{align}
+
+% %%%%%%%
+\section{Results}
+
+% %%%%%%%
+\section{Discussion}
+
+- May be something about the testing rate/intensity, now the per capita testing intensity is very low ($\rho \approx 0$).
+In near future new test kits may be widely accessible, our model provide insights in this case (what are those?)
+
+- "delay negative results" result
+
+% %%%%%%%
+\bibliography{SIRlibrary.bib}
 

### 2020-11-13 SIR_manuscript.tex
model was updated

@@
  d I_n/dt &= \Lambda S_n + (1-p_I) F_I I_u - \omega I_n -\gamma I_n \\
- d I_p/dt &= p_I F_I I_u - \omega I_p -\gamma I_p \\
- d I_t/dt &= \omega I_p - \gamma I_t  \\
+ d I_p/dt &= \Lambda S_p + p_I F_I I_u - \omega I_p -\gamma I_p \\
+ d I_t/dt &= \Lambda S_t + \omega I_p - \gamma I_t  \\
  d R_u/dt &= \gamma I_u - F_R R_u + \omega R_n \\

### 2020-11-15 SIR_manuscript.tex
Add files via upload

@@
 \bibliographystyle{apalike}
+
+%%Packages and commands added by Fady
+\usepackage{xcolor}
+\newcommand{\fady}[1]{\textcolor{cyan}{$\langle${\slshape{\bfseries Fady:} #1 }$\rangle$}}
+\newcommand{\Rnum}{\mathcal{R}_0}
+\usepackage{xspace}
+\usepackage[nameinlink,capitalize]{cleveref}
 
@@
 
+\subsection{rate of returning tests}
+The linearization of $\Rnum$ around $\rho=0$ is
+\begin{equation}\label{linearization}
+\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta_w-1)(\gamma W_s+\omega W_i) + (\eta_t -1)P_iW_i \omega^2 \Big). 
+\end{equation}
+
+So when $\rho \approx 0$ we have $$\partial{\Rnum}/\partial{\omega} \approx  \frac{-\beta \rho}{\gamma W_s\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c).$$
+
+where $a=(\eta_w-1)W_i-(\eta_t-1)P_iW_i$, $b=2(\eta_w-1)\gamma W_s$ and $c=(\eta_w-1)\gamma^2 W_s$.
+
+Perhaps counter-intuitively, the equation above does not predict that $\Rnum$ is monotone decreasing with respect to $\omega$.
+In other words; our model does not predict that returning test results more rapidly \textit{always} lower $\Rnum$.
+In order to gain insight into this intriguing behaviour, we examine the zeroes of $\frac{\partial{\Rnum}}{\partial{\omega}}(\omega)$.
+
+Defining the following quantity, $Q$, will help us write the roots of $\partial{\Rnum}/\partial{\omega}$ neatly. 
+\begin{align}\label{eq:defQ}
+    Q =& \frac{W_i}{W_s}\left(1-\frac{n_{t}-1}{n_{w}-1}P_{i}\right) \\
+\end{align}
+
+With that in mind, we can write the roots of $\partial{\Rnum}/\partial{\omega}$ as
+
+\begin{align}
+    \omega_1 =& \frac{\gamma}{-\sqrt{1-Q}-1} \\
+    \omega_2 =& \frac{\gamma}{\sqrt{1-Q}-1}
+\end{align}
+
+Note that the zeroes are real if and only if $Q < 1$.
+Note that have $\eta_t < \eta_w$, so if $P_i \approx 1$, we will have $Q < 0 < 1$.
+Thus, if we assume near-perfect test sensitivity, $\omega_1$ and $\omega_2$ will be real.
+
+Assuming $\omega_1, \omega_2$ are real, it is easy to confirm that $\omega_1 < 0$ by looking at the denominator.
+To see that $\omega_2 > 0$, recall that $Q < 0$, so $\sqrt{1-Q} > 1$ and so $\sqrt{1-Q} -1 > 0$.
+Knowing that $\omega_1 < 0$, the only root of interest (i.e., biologically relevant quantity) is $\omega_2$.
+
+We can prove that $\partial{\Rnum}/\partial{\omega} > 0$ when $\omega \in (0,\omega_2)$ and $\partial{\Rnum}/\partial{\omega} < 0$ when $\omega \in (\omega_2,\infty)$ by computing the limits of $\partial{\Rnum}/\partial{\omega}$ at $0$ and $\infty$ respectively.
+So it follows that $\Rnum$ has a global maximum with respect to $\omega$ at $\omega = \omega_2$.
+
+Now we want to characterize the parameter regions on which $\partial{\Rnum}/\partial{\omega} < 0$ (i.e., the conditions under which returning test results more rapidly is favourable).
+By the previous analysis, this is equivalent to solving for $\omega > \omega_2$.
+So
+
+\begin{align}\label{eq:necsuf}
+    &\omega > \omega_2 \nonumber \\
+    &\omega > \frac{\gamma}{\sqrt{1-Q}-1} \nonumber \\
+    &\vdots \nonumber \\
+    &\frac{1-n_{t}}{1-n_{w}}P_{i}>\frac{W_{s}}{W_{i}}\left(\frac{\gamma}{\omega}+1\right)^{2}
+\end{align}
+
+This inequality quantifies the exact relationships between model parameters that result in returning tests more rapidly being favourable.
+To give a biological interpretation, we describe the qualitative \textit{trends} predicted by the inequality.
+Returning test results more rapidly \textit{tends} to be more favourable when
+
+\begin{itemize}
+    \item individuals who have tested positive lower their contact much more than individuals who are waiting for their test results (i.e., $\eta_t \ll \eta_w$).
+    \item the test is sensitive.
+    \item testing of infected individuals is more intense than testing of susceptible individuals.
+\end{itemize}
+An in-depth discussion of these results is presented in the discussion section.
+
+\subsection{Expensive vs. cheap tests}
+
+The use of tests cheaper than RT-PCR has been proposed as a potential strategy for containing the COVID-19 pandemic.
+While cheaper tests may be less sensitive and reliable than RT-PCR, they allow for broader and more intense testing.
+In the analysis below, we compare the $\Rnum$ predicted by our model depending on the testing strategy.
+
+Consider a test that allows us to test at rate $\rho_1$ and has sensitivity $P_{i,1}$, and another test that allows us to test at $\rho_2$ and has sensitivity $P_{i,2}$.
+Suppose that $\rho_1 > \rho_2$.
+Recall that the linearization of $\Rnum$ around $\rho \approx 0$ is given by $$\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta_w-1)(\gamma W_s+\omega W_i) + (\eta_t -1)P_iW_i \omega^2 \Big).$$
+
+Treating $\Rnum$ as a function of $\rho$ and $P_i$,we can reduce the inequality $$\Rnum(\rho_2, P_{i,2}) < \Rnum(\rho_1, P_{i,1})$$ into
+
+\begin{align}\label{eq:rho1vsrho2}
+    &\rho_1\left(\gamma(\eta_w-1)(\gamma W_s + \omega W_i) + (\eta_t-1)P_{i, 1}W_i\omega^2\right) - \rho_2\left(\gamma(\eta_w-1)(\gamma W_s + \omega W_i) + (\eta_t-1)P_{i, 2}W_i\omega^2\right) > 0 \nonumber \\
+    &\vdots \nonumber \\
+    &\frac{\rho_2P_{i, 2}-\rho_1P_{i, 1} }{\rho_1-\rho_2} > \frac{1-\eta_w}{1-\eta_t}\cdot \frac{\gamma(\gamma W_s + \omega W_i)}{\omega^2 W_i}
+\end{align}
+
+Note that the RHS is positive, thus a necessary condition for the inequality above to hold is that $\rho_2P_{i,2} > \rho_1P_{i,1}$, equivalently
+
+\begin{equation}
+\frac{P_{i,2}}{P_{i,1}} > \frac{\rho_1}{\rho_2}.
+\end{equation}
+
+To state an example of this, if test $A$ is three times as expensive as test $B$ (and hence one can test three times as many people with test $B$), using test $A$ rather than $B$ will be favourable only if test $A$ is at least 3 times more sensitive than test $B$.
+Note that this is a necessary but not sufficient condition, so even if test $A$ is three times more sensitive, it is still possible for test $B$ to be more effective.
+
+\cref{eq:rho1vsrho2} tells us precisely when a test corresponding to $\rho_2, P_{i,2}$ will yield a lower $\Rnum$ than a test corresponding to $\rho_1, P_{i,1}$, where $\rho_1 > \rho_2$.
+Some of the qualitative \textit{trends} that favour test 2 (the higher-sensitivity test) include
+
+\begin{itemize}
+    \item individuals who test positive self-isolate much more than individuals who are waiting for their test result.
+    \item the time it takes to return tests is much shorter than the mean infectious period.
+    \item the testing intensity is much greater for infected individuals than susceptible individuals.
+\end{itemize}
+
 % %%%%%%%
@@
 
-- "delay negative results" result
+\subsection{Rate of returning tests}
+\fady{needs to be expanded}
+When the difference in self-isolation between individuals waiting for their tests and individuals who are confirmed positive is small, returning tests rapidly is not beneficial because the benefit in doing so (i.e., the extra self-isolating when an individual tests positive) is minimal and outweighed by the fact that those who test negative may increase their contacts.
+Similarly, if the test is not sensitive, many infected individuals will test negative, and thus may increase their contacts under the false belief that they are not infectious, hence returning test results more quickly in this case may not be beneficial.
+
+When testing of infected individuals is more intense than susceptible individuals, returning test results rapidly is beneficial because those infected individuals will be informed and self-isolate.
+Conversely, if testing of susceptible individuals is much greater, very few tests be positive, and the benefit of informing those few positive cases will be outweighed by the potential increase in contacts of the individuals testing negative.
+
+\subsection{Expensive vs. cheap tests}
+\fady{needs to be written \dots}
 

### 2020-11-15 note/SIR_notes.Rnw
minor updates

@@
 
-\includegraphics[width=\linewidth]{../pix/sir_comp2.png}
+\includegraphics[width=\linewidth]{../pix/sir_comp.png}
 
@@
 \begin{align}
-\label{mod:1}
+\label{model}
  d S_u/dt &= -\Lambda S_u - F_S S_u + \omega S_n \\

### 2020-11-16 SIR_manuscript.tex
Add files via upload
minor updates
clean up and organized
updated panel plot and manuscript

@@
 \usepackage{amsmath}
-\bibliographystyle{apalike}
-
-%%Packages and commands added by Fady
 \usepackage{xcolor}
@@
 \usepackage[nameinlink,capitalize]{cleveref}
+
+\bibliographystyle{apalike}
 
@@
 
-- Our work; we developed a deterministic compartmental model which incorporates testing strategies and isolation mechanism. 
+- What did we do? Why novel? 
+we developed a deterministic compartmental model which incorporates testing strategies and isolation mechanism. 
 - The basic reproduction number was derived analytically as a threshold of the dynamics. 
@@
 \section{Method}
-\begin{itemize}
-\item $\Lambda$: force of infection defined as $$\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_t I_t)}{N_0},$$ where $\beta$ is transmission rate, $\eta_w$ and $\eta_t$ are the isolation parameters for awaiting and reported individuals, respectively.
-We assume that $\eta_t \leq \eta_w$, i.e., the awaiting individuals for test results have a higher transmission probability than the reported individuals.
-\item $\omega$: the rate of onward flow from the awaiting positive compartment, $p$, to reported/tested compartment, $\rho$, or from awaiting negative compartment, $n$, back to $u$.
-It has units of 1/time.
-\item $\gamma$: recovery rate (1/time).
-\item $\rho$: per capita testing intensity across the whole population (1/time).
-\item $W$: weighted number of people available for tests, defined as $W = W_S S_u + W_I I_u + W_R R_u$.
-\item $\sigma$: scaling parameter for testing defined as $\sigma = \frac{\rho N_0}{W}$.
-\item $F_Z$: Weighted testing rate defined as $F_Z=\sigma W_Z$. That is, $F_S = \sigma W_S$, $F_I=\sigma W_I$ and $F_R = \sigma W_R$.
-\end{itemize}
 
-\includegraphics[width=\linewidth]{./pix/sir_comp.png}
+We used a SIR (Susceptible-Infectious-Recovered) deterministic model.
+In order to collect cumulative reported negative or positive tests, two 'integrator' or 'accumulator' compartments, $N$ and $P$, were incorporated.
+The flowchart is
+
+\begin{figure}[!h] 
+\begin{center} 
+\includegraphics[scale=0.3]{./pix/sir_comp.png}
+\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model. Say more here to make it stand on its own.
+\label{fig:flowchart}} 
+\end{center} 
+\end{figure}
 
 The model is
-
 \begin{align}
@@
  dN/dt &= \omega (S_n + I_n + R_n)   \\
- dP/dt &= \omega(I_p + R_p).
+ dP/dt &= \omega(I_p + R_p),
 \end{align}
+
+where $\beta$, (1/day), is the disease transmission rate, $\eta_w$ and $\eta_t$ are the isolation parameters for awaiting and reported individuals, respectively, $\Lambda$ is the force of infection and defined as
+\begin{equation}
+\label{Lambda}
+\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_t I_t)}{N_0},
+\end{equation}
+where $N_0$ is the total population size.
+Further, $\omega$, (1/day), is the rate of onward flow from the awaiting positive or negative compartments to reported or untested compartments, respectively, $\gamma$, (1/day), is the recovery rate, $\rho$, (1/day), is the per capita testing intensity across the whole population.
+Further, various testing strategies are incorporated through the compartment-specific relative testing weights $W_S$, $W_I$ and $W_R$.
+The weighted number of people available for tests is defined as $W = W_S S_u + W_I I_u + W_R R_u$, and the scaling parameter for testing defined as \begin{equation}
+\label{sigma}
+\sigma = \frac{\rho N_0}{W}.
+\end{equation}
+Thus, the weighted testing rate defined as $F_Z=\sigma W_Z$, for $Z \in \{S,I,R\}$.
+
+- In this model we assumed: (i) all groups being exposed to all others, (ii) there is a single force of infection, $\Lambda$, that is, where a susceptible individual move to does not depend on who infected it, (iii) for simplicity of the analysis, a perfectly specific test, $p_S=0$, was assumed, (iv)  $\eta_t \leq \eta_w$, i.e., the awaiting individuals for test results have a higher transmission probability than the reported individuals.
+-why not bith and death are included?
 
@@
 \section{Results}
+
+The Disease-Free Equilibrium, DFE hearafter, for the above SIR model is given by the solution of the following system $S_u+S_n=N0$ and $F_S S_u-\omega S_n=0$. The DFE is
+
+\begin{equation}
+\label{dfe}
+S_n^*= \frac{\rho}{\omega} N_0, S_u^*= N_0-S_n^*, \text{and} I_j=R_j=0 \ \forall j.
+\end{equation}
+
+The Basic reproduction number, $\Rnum$, was calculated by using the method of next generation matrix \cite{van2002reproduction}. See the Appendix for details.
+
+\begin{equation}
+\label{R0}
+\Rnum= (A \times S_u^* + B \times S_n^*) \times C, 
+\end{equation}
+where
+$A=\gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_t p_I) F_I$,
+$B=\big(\omega+(F_I+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_t\omega) \omega p_I F_I }{\omega+\gamma}$ and 
+$C=\frac{\beta/\gamma}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}$.
+
+\begin{figure}[h!]
+\centering
+\begin{subfigure}[t]{.45\textwidth}
+\centering
+\includegraphics[width=\linewidth]{./pix/R0contour_random.png}
+        \caption{}\label{fig:res1_a}
+\end{subfigure}
+%
+\begin{subfigure}[t]{.45\textwidth}
+\centering
+\includegraphics[width=\linewidth]{./pix/R0contour_TTI.png}
+\caption{}\label{fig:res1_b}
+\end{subfigure}
+\end{figure}
 
@@
 
+- "delay negative results" result
+
 \subsection{Rate of returning tests}
@@
 \subsection{Expensive vs. cheap tests}
-\fady{needs to be written \dots}
+\fady{needs to be written \dots must discuss the fact that our analysis holds only when $\rho$ is relatively small, so if there is a test so cheap that $\rho$ does not stay small, our results do not apply.}
 
@@
 
+% %%%%%%%
+\section{supplementary material}
+
+The next generation matrix, $G = F V^{-1}$, is in the lower triangular form with $F$ and $G$ as follows.
+
+\begin{equation}\label{FV}
+F= \beta/N_0 \left[ \begin {array}{cccc} 
+S_u&\eta_w\,S_u&\eta_w\,S_u&\eta_t\,S_u\\
+S_n&\eta_w\,S_n&\eta_w\,S_n&\eta_t\,S_n\\ 
+0&0&0&0\\
+0&0&0&0
+ \end {array} \right],
+ V=
+ \left[ \begin {array}{cccc}  
+F_I+\gamma&-\omega&0&0\\
+-(1-p_I)F_I&\omega+\gamma&0&0\\
+-p_I F_I&0&\omega+\gamma&0\\
+0&0&-\omega&\gamma
+\end {array} \right],
+\end{equation}
+
+\begin{equation}
+G = \left[ \begin {array}{cc}
+G_{11}&G_{12}\\
+0&0
+\end {array} \right], \text{ where } \\
+G_{11} =C
+\left[\begin {array}{cc}
+A\,S_u & B\,S_u\\
+A\,S_n & B\,S_n
+\end {array}\right],
+\end{equation}
+
+The Eigenvalues of $G_{11}$, are 0 and $\Rnum$.
+
 \end{document}

### 2020-11-16 codes/ali_test_sir.Rmd
updated panel plot and manuscript

--- 2020-11-10 codes/ali_test_sir.Rmd
@@
 ```{r, out.width = "400px"}
-knitr::include_graphics("../pix/sir_comp2.png")
+knitr::include_graphics("../pix/sir_comp.png")
 ```
@@
 library(directlabels)
+library(latex2exp)
 unpack <- McMasterPandemic::unpack
@@
 
+df2 <- df2 %>% dplyr::mutate(R0_sub = ifelse(eta_w<eta_t, NA, R0))
+
 which(df2[]<0)
@@
 ```{r show_contours_1, width=24,height=24,message=FALSE,warning=FALSE}
-p1 <- ggplot(df2,aes(x=omega,y=rho,z=R0))+ theme_bw() +
+p1 <- ggplot(df2,aes(x=omega,y=rho,z=R0_sub))+ theme_bw() +
+  xlab(TeX('$\\omega$, Returning test result rate (1/day)')) +
+  ylab(TeX('$\\rho$, testing intensity (1/day per capita)')) +
     theme(panel.spacing=grid::unit(0,"lines"))
 p1_temp <- (p1
-    + geom_contour_filled(breaks=pretty(unique(df2$R0),5),
+    + geom_contour_filled(breaks=pretty(unique(df2$R0_sub),5),
                           alpha=0.5) ## breaks=0.2)
@@
     + scale_fill_viridis_d(name=parse(text="R[0]"))
-    + ggtitle("W[S]=0.3, W[I]=W[R]=1")
+    + ggtitle(TeX(sprintf("W_S=%.1f, W_I=W_R=1",W_S)))
 )
 p1_temp
-# filename = "R0contour_TTI.png"
-# filename = "R0contour_random.png"
+# ggsave(p1_temp,filename = "R0contour_random.png" , path="../pix/")
 ggsave(p1_temp,filename = "R0contour_TTI.png" , path="../pix/")
+
 ```
@@
     + scale_fill_viridis_d(name=parse(text="R[0]"))
-    + ggtitle("W[S]=0.3, W[I]=W[R]=1")
+    + ggtitle("W[S]=1, W[I]=W[R]=1")
 )
+p2_temp
+# ggsave(p2_temp,filename = "R0contour_random_s.png" , path="../pix/")
 ```

### 2020-11-17 SIR_manuscript.tex
more updates

@@
 \usepackage{subcaption}
-\usepackage{amsmath}
+\usepackage{amssymb,amsmath,amsthm}
 \usepackage{xcolor}
-\newcommand{\fady}[1]{\textcolor{cyan}{$\langle${\slshape{\bfseries Fady:} #1 }$\rangle$}}
-\newcommand{\Rnum}{\mathcal{R}_0}
 \usepackage{xspace}
@@
 
+\newcommand{\fady}[1]{\textcolor{cyan}{$\langle${\slshape{\bfseries Fady:} #1 }$\rangle$}}
+\newcommand{\Rnum}{\mathcal{R}_0}
+\theoremstyle{definition} % amsthm only
+\newtheorem{proposition}{Proposition}
+\newtheorem{theorem}{Theorem}
+
 \bibliographystyle{apalike}
 
-\title{Testing Efficacy; Insights from a Simple Epidemic Model }
+\title{Testing and Isolation Efficacy; Insights from a Simple Epidemic Model }
 
@@
 
-- Why non-random testing/testing-tracing-isolation (TTI) is more efficent in controling an epidemic event?
+- Non-random testing/testing-tracing-isolation (TTI) is more efficent in controling an epidemic event.
+- Potential advantage of slow reporting of test result. 
 - Literature reviwe, evidences of efficacy of TTI
@@
 - Covid-19 and chalanges, questions to be addressed with a simple model.
-This study provides a framework/platform to study a pandamic in a population with a determinstic system of differential equations with testing mechinisms and isolation.
+This study provides a simple deterministic model/tool in which testing mechinisms and isolation have been modeled to study a pandamic in a population.
 
 - What did we do? Why novel? 
-we developed a deterministic compartmental model which incorporates testing strategies and isolation mechanism. 
+(i) We developed a deterministic compartmental model which incorporates testing strategies and isolation mechanism to get insights.
+(ii) the potential advantage of slow reporting of test result on the spread of the infection was addressed.
 - The basic reproduction number was derived analytically as a threshold of the dynamics. 
-- Our simulation shows ...
+- Our results include...
 
 % %%%%%%%
-\section{Literature Review}
+\subsection{Literature Review}
 
-\subsection{Explicit models of TTI (trace/test/isolate) based on network or agent-based models}
+\subsubsection{Explicit models of TTI (trace/test/isolate) based on network or agent-based models}
 \citep{endo2020implication} [Ali: It seems to me that this is just a statistical model to estimate the parent-offspring of an infected index, not sure if it fits into agent-based group!] Used simulation on a branching process model to assess the forward and backward contact tracing efficiency.
@@
 % %%%%%%%
-\subsection{Models of repeated random testing of isolated populations}
+\subsubsection{Models of repeated random testing of isolated populations}
 \cite{bergstrom2020frequency}
@@
 % %%%%%%%
-\subsection{Other maybe-related works}
+\subsubsection{Other maybe-related works}
 \citep{arino2020simple} developed a SLIAR compartmental model to study the spread of an epidemic, specifically COVID-19, in a population.
@@
 \end{equation}
-Thus, the weighted testing rate defined as $F_Z=\sigma W_Z$, for $Z \in \{S,I,R\}$.
+Thus, the weighted testing rate is defined as $F_Z=\sigma W_Z$, for $Z \in \{S,I,R\}$.
 
-- In this model we assumed: (i) all groups being exposed to all others, (ii) there is a single force of infection, $\Lambda$, that is, where a susceptible individual move to does not depend on who infected it, (iii) for simplicity of the analysis, a perfectly specific test, $p_S=0$, was assumed, (iv)  $\eta_t \leq \eta_w$, i.e., the awaiting individuals for test results have a higher transmission probability than the reported individuals.
+- In this model we assumed: (i) all groups being exposed to all others, (ii) there is a single force of infection, $\Lambda$, that is, where a susceptible individual move to does not depend on who infected it, (iii) a perfectly specific test, $p_S=0$, for simplicity in analysis, (iv)  $\eta_t \leq \eta_w$, i.e., the awaiting individuals for test results have a higher transmission probability than the reported individuals.
 -why not bith and death are included?
 
-% %%%%%%%
-\section{Results}
-
-The Disease-Free Equilibrium, DFE hearafter, for the above SIR model is given by the solution of the following system $S_u+S_n=N0$ and $F_S S_u-\omega S_n=0$. The DFE is
+The Disease-Free Equilibrium, DFE hearafter, for the SIR model \ref{model} is given by the solution of the following system including $S_u+S_n=N_0$ and $F_S S_u-\omega S_n=0$. The DFE is
 
@@
 \label{dfe}
-S_n^*= \frac{\rho}{\omega} N_0, S_u^*= N_0-S_n^*, \text{and} I_j=R_j=0 \ \forall j.
+S_n^*= \frac{\rho}{\omega} N_0, \ S_u^*= N_0-S_n^*, \text{and} I_j=R_j=0 \ \text{for all j}.
 \end{equation}
 
-The Basic reproduction number, $\Rnum$, was calculated by using the method of next generation matrix \cite{van2002reproduction}. See the Appendix for details.
-
+The Basic reproduction number, $\Rnum$, is 
 \begin{equation}
@@
 
+$\Rnum$ was calculated by using the generation matrix metod developed by \cite{van2002reproduction}. Further details are provided in the Appendix.
+
+% %%%%%%%
+\section{Results}
+
+The following propositions are restlted from the expression of $\Rnum$ and its partial derivatives with respect to the underlying parameters.
+
+\begin{proposition}
+\label{prop1}
+Using the expression of $\Rnum$ \eqref{R0},
+\begin{enumerate}
+\item \label{p1:eta}
+$\partial{R_0}/\partial{\eta_t} \geq 0$ and $\partial{R_0}/\partial{\eta_w} \geq 0$. 
+\item \label{p1:rho}
+$\partial{\Rnum}/\partial{\rho} \leq 0$ for $\rho \approx 0$.
+\item \label{p1:omega}
+$\partial{R_0}/\partial{\omega}$ can be positive or negative when $\rho \approx 0$.
+\end{enumerate}
+\end{proposition}
+
+[Thinking of putting most of these stuff in the Appendix] Given that the perfect isolation occurs when $\eta_t=\eta_w = 0$, \ref{p1:eta} means that lifting isolation results in greater $\Rnum$ and consequently greater number of infected individuals.
+\ref{p1:rho} and \ref{p1:omega} are concluded from the following Taylor approximation of $\Rnum$ at $\rho=0$ given by 
+\begin{equation}
+\label{eq:R0appr}
+R_0 \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_S} \Big(\gamma(\eta_w-1)(\gamma W_S+\omega W_I) + (\eta_t -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2).
+\end{equation}
+
+It is straight forward to have
+\begin{equation}
+\label{eq:R0wrtOm}
+\partial{R_0}/\partial{\omega}=  \frac{-\beta \rho}{\gamma W_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c),
+\end{equation}
+
+where $a=(\eta_w-1)W_I-(\eta_t-1)P_I W_I = ((s-p_I)\eta_t + (p_I-1)) W_I$, $b=2(\eta_w-1)\gamma W_S$ and $c=(\eta_w-1)\gamma^2 W_S$.
+Given that $0 \leq \eta_t\leq \eta_w \leq 1 $, one can easily derive $b\leq 0$ and $c \leq 0$.
+Note that in general, the necessary and sufficient condition for $a \geq 0$ is $(s-p_I) \eta_t \geq (1-p_I)$, where $s=\frac{\eta_w}{\eta_t} \geq 1$.
+As an example, in the case of a very accurate testing regime,  i.e., $P_I=1$, $a \geq 0$ is achieved.
+If $a\geq 0$, the quadratic expression in \eqref{eq:R0wrtOm}, has Real roots.
+Assuming that $\omega_1<0$ and $\omega_2>0$ be the roots of the quadratic expression in $\partial{R_0}/\partial{\omega}$.
+Thus, $\partial{R_0}/\partial{\omega}>0$ for $0<\omega<\omega_2$ and  $\partial{R_0}/\partial{\omega}<0$ for $\omega>\omega_2$.
+
+Note that in case of ``Perfect isolation``, i.e., when $\eta_w=0$ and consequently $\eta_t=0$, it is straight forward to see that $a \leq 0$, $b<0$ and $c<0$.
+Thus, $\partial{\Rnum}/\partial{\omega} \geq 0$.
+Intuitively, this means that returning the test results back to the awaiting individuals while they were following a perfect isolation, results in an increase of $\Rnum$ and consequently an increase in the number of infected individuals (see Figure \ref{fig:panels} at lower $\omega$).
+
 \begin{figure}[h!]
@@
 \includegraphics[width=\linewidth]{./pix/R0contour_random.png}
-        \caption{}\label{fig:res1_a}
 \end{subfigure}
@@
 \includegraphics[width=\linewidth]{./pix/R0contour_TTI.png}
-\caption{}\label{fig:res1_b}
 \end{subfigure}
+\caption{Behaviour of the basic reproduction number, $\Rnum$, with respect to changes in the underlying parameters; remind the reader about the parameters and their values in this simulations.}
+\label{fig:panels}
 \end{figure}
@@
 
+- JD's email:
+We're basically looking at the potential advantage of slow tests:
+people waiting for test results may be more careful than those who
+have received negatives. These advantages are real, and neglected. We
+also compare to an individual-level advantage of fast tests: people
+who test positive may be even more careful. But we are missing out on
+community-level advantages of fast testing: better assessment of the
+situation, identification of hot spots, contact-tracing, etc.
+
 - May be something about the testing rate/intensity, now the per capita testing intensity is very low ($\rho \approx 0$).

### 2020-11-18 SIR_manuscript.tex
intro updated

@@
 
-- Non-random testing/testing-tracing-isolation (TTI) is more efficent in controling an epidemic event.
-- Potential advantage of slow reporting of test result. 
-- Literature reviwe, evidences of efficacy of TTI
-- Simple ODE models provide useful insights (some classic examples, Kermack-Makendric,...)
-Historically, Ronald Ross (I think) developed a series of mathematical models to study the spread of malaria and implementation of control. His work contributed to the modern theory of the control of
-infectious disease (ref: Ross R. The Prevention of Malaria, 2nd edn. London: Murray, 1911).
+One of the real chalanges interpreting COVID-19 dynamics is to unravel/investigate information gained from data.
+One of the strongest singnals we get to see is the daily case reports which are driven by a combination of two processes: (i) epidemic processes and (ii) testing processes.
+The real chalange here is to seperate these two processes.
+Simple mathematical models have a long history in providing useful insights facing with real epidemiological chalanges \citep{ross1911prevention} [maybe add influenza work, ideas?].
+We developed a model to incorporate mechanistic processes of epidemic processes and testing.
+We used the model to study the potential effect of testing strategies and slow-fast reporting testing result.
 
-- Covid-19 and chalanges, questions to be addressed with a simple model.
-This study provides a simple deterministic model/tool in which testing mechinisms and isolation have been modeled to study a pandamic in a population.
+There are different testing strategies with test-specific efficacy, the time required for the results to be reported and the cost of the test.
+Testing strategies can be grouped into two broad cathegories including (i) random testing across the whole population and (ii) Non-random/bios testing.
+Note that contact tracing is an example of non-random testing strategies.
+Each of these strategies has been used as a tool to study COVID-19; see (refs) for group (i) and (refs) for group (ii).
+The advantage of non-random testing comparing to random testing has been investigated specifically during COVID-19 pandemic (refs).
+Further, lots of efforts have been invested in developing widely accessible and fast tests in order to metigate the spread of the pandemic (refs).
+As a comparison of slow versus fast test, the viral PCR test, which is the most accessible test and takes about four days, is very slow relative to the potential tests such as breath-based test \citep{ruszkiewicz2020diagnosis}, which is still in trial stage and takes a few seconds.
+The point to be made is that there can be a potential advantage of slow test.
 
-- What did we do? Why novel? 
-(i) We developed a deterministic compartmental model which incorporates testing strategies and isolation mechanism to get insights.
-(ii) the potential advantage of slow reporting of test result on the spread of the infection was addressed.
-- The basic reproduction number was derived analytically as a threshold of the dynamics. 
-- Our results include...
-
-% %%%%%%%
-\subsection{Literature Review}
-
-\subsubsection{Explicit models of TTI (trace/test/isolate) based on network or agent-based models}
-\citep{endo2020implication} [Ali: It seems to me that this is just a statistical model to estimate the parent-offspring of an infected index, not sure if it fits into agent-based group!] Used simulation on a branching process model to assess the forward and backward contact tracing efficiency.
-Assuming a negative-binomial branching process with a mean R, reproduction number, and overdispersion parameter k, the mean total number of generation G3 and averted G3 are estimated.
-The effectiveness of TTI is defined as the ratio of averted to the mean.
-
-\citep{jenness2020modeling} developed a network-based transmission model for SARS-CoV-2 on the Diamond Princess outbreak to characterize transmission dynamics and to estimate the epidemiological impact of outbreak control and prevention measures.
-
-\citep{elbanna2020entry} [seems similar to MacPan model!]
-
-\citep{de2020influenza} Was discussed in the Math 747 
-SEIR Asymptomatic and symptomatic $I_1, I_2$. Used linear chain trick 
-Stringency index as a control force lowering $\beta$.
-
-\citep{rice2020effect} Effect of school closures on mortality. Reproduce Report 9 results by spatial agent based CovidSim. 
-% %%%%%%%
-\subsubsection{Models of repeated random testing of isolated populations}
-\cite{bergstrom2020frequency}
-(1) Model, assumptions: They developed a function, namely expected exposure $E(C,\tau)$, to approximate trade-offs between the frequency of testing, n, the sensitivity of testing, q, and the delay between
-testing and results, d. This function is explicitly derived and was connected the effective reproduction number $R=R_0 S$, where $S$ is the proportion of population susceptible.
-assumption that transmission rates are a step function: individuals who
-have COVID go from non-infectious to fully infectious instantaneously,
-and remain fully infectious until they are no longer able to transmit disease. Test sensitivity takes the same form over the course of infection.
-More sophisticated models could allow varying infectiousness and varying
-sensitivity over time, as in 
-\citep{larremore2020test}.
-
-\citep{lopman2020model} Used a Deterministic SEIR model, incorporated TTI, applicable to a university setting. They assumed a fairly high reproductive number that is not reduced through social
-distancing measures. They found that community-introduction of SARS-CoV-2 infection onto campus can be
-relatively controlled with effective testing, isolation, contract tracing and quarantine.
-
-\citep{tuite2020mathematical} used an age-structured compartmental model of COVID-19 transmission in the population of Ontario, Canada.
-We compared a base case with limited testing, isolation and quarantine to different scenarios.
-% %%%%%%%
-\subsubsection{Other maybe-related works}
-\citep{arino2020simple} developed a SLIAR compartmental model to study the spread of an epidemic, specifically COVID-19, in a population.
-The model incorporates an Erlang distribution of times of sojourn in incubating, symptomatically and asymptomatically infectious compartments.
-Basic reproduction number is derived.
-Also, sensitivity analysis with respect to the underlying parameters for the following two outputs was carried out; (i) the number of observable cases during the course of the epidemic and at the peak, and (ii) the timing of the peak of the outbreak.
-Sensitivity analysis is performed using the R package multisensi.
-
-\citep{ruszkiewicz2020diagnosis} novel with-in-a-minute breath testing with 80\% accuracy.
+In this work we developed a model to incorporate epidemic processes and testing processes mechanisticly to study the potential effect of testing strategies and slow-fast testing.
+The analysis of the deterministic model included deriving the basic reproduction number, $\Rnum$, as the index of threshold dynamics and the effect of underlying parameters on this key index.
+We found that the non-random testing strategy is more effective than the random testing.
+Specifically, the required critical testing intensity to keep $\Rnum$ lower than the threshold value of one, is much lower in the non-random testing than the random one.
+Further, we show that there is a potential advantage of slow reporting of test result on the spread of the infection.
 
@@
 
+[comments about the figure:]
+1- use a single legend, change the range manually.
 \begin{figure}[h!]
@@
 
+% %%%%%%%
+\section{Literature Review}
+
+\subsection{Explicit models of TTI (trace/test/isolate) based on network or agent-based models}
+\citep{endo2020implication} [Ali: It seems to me that this is just a statistical model to estimate the parent-offspring of an infected index, not sure if it fits into agent-based group!] Used simulation on a branching process model to assess the forward and backward contact tracing efficiency.
+Assuming a negative-binomial branching process with a mean R, reproduction number, and overdispersion parameter k, the mean total number of generation G3 and averted G3 are estimated.
+The effectiveness of TTI is defined as the ratio of averted to the mean.
+
+\citep{jenness2020modeling} developed a network-based transmission model for SARS-CoV-2 on the Diamond Princess outbreak to characterize transmission dynamics and to estimate the epidemiological impact of outbreak control and prevention measures.
+
+\citep{elbanna2020entry} [seems similar to MacPan model!]
+
+\citep{de2020influenza} Was discussed in the Math 747 
+SEIR Asymptomatic and symptomatic $I_1, I_2$. Used linear chain trick 
+Stringency index as a control force lowering $\beta$.
+
+\citep{rice2020effect} Effect of school closures on mortality. Reproduce Report 9 results by spatial agent based CovidSim. 
+% %%%%%%%
+\subsection{Models of repeated random testing of isolated populations}
+\cite{bergstrom2020frequency}
+(1) Model, assumptions: They developed a function, namely expected exposure $E(C,\tau)$, to approximate trade-offs between the frequency of testing, n, the sensitivity of testing, q, and the delay between
+testing and results, d. This function is explicitly derived and was connected the effective reproduction number $R=R_0 S$, where $S$ is the proportion of population susceptible.
+assumption that transmission rates are a step function: individuals who
+have COVID go from non-infectious to fully infectious instantaneously,
+and remain fully infectious until they are no longer able to transmit disease. Test sensitivity takes the same form over the course of infection.
+More sophisticated models could allow varying infectiousness and varying
+sensitivity over time, as in 
+\citep{larremore2020test}.
+
+\citep{lopman2020model} Used a Deterministic SEIR model, incorporated TTI, applicable to a university setting. They assumed a fairly high reproductive number that is not reduced through social
+distancing measures. They found that community-introduction of SARS-CoV-2 infection onto campus can be
+relatively controlled with effective testing, isolation, contract tracing and quarantine.
+
+\citep{tuite2020mathematical} used an age-structured compartmental model of COVID-19 transmission in the population of Ontario, Canada.
+We compared a base case with limited testing, isolation and quarantine to different scenarios.
+% %%%%%%%
+\subsection{Other maybe-related works}
+\citep{arino2020simple} developed a SLIAR compartmental model to study the spread of an epidemic, specifically COVID-19, in a population.
+The model incorporates an Erlang distribution of times of sojourn in incubating, symptomatically and asymptomatically infectious compartments.
+Basic reproduction number is derived.
+Also, sensitivity analysis with respect to the underlying parameters for the following two outputs was carried out; (i) the number of observable cases during the course of the epidemic and at the peak, and (ii) the timing of the peak of the outbreak.
+Sensitivity analysis is performed using the R package multisensi.
+
+\citep{ruszkiewicz2020diagnosis} novel with-in-a-minute breath testing with 80\% accuracy.
+
 \end{document}

### 2020-11-18 SIR_manuscript.tex
update tex

@@
 % %%%%%%%
+\section{Abstract}
+
+% %%%%%%%
 \section{Introduction}
 
-One of the real chalanges interpreting COVID-19 dynamics is to unravel/investigate information gained from data.
-One of the strongest singnals we get to see is the daily case reports which are driven by a combination of two processes: (i) epidemic processes and (ii) testing processes.
-The real chalange here is to seperate these two processes.
-Simple mathematical models have a long history in providing useful insights facing with real epidemiological chalanges \citep{ross1911prevention} [maybe add influenza work, ideas?].
+One of the real challenges interpreting COVID-19 dynamics is to unravel/investigate information gained from data.
+One of the strongest signals we get to see is the daily case reports which are driven by a combination of two processes: (i) epidemic processes and (ii) testing processes.
+The real challenge here is to separate these two processes.
+Simple mathematical models have a long history in providing useful insights facing with real epidemiological challenges \citep{ross1911prevention} [maybe add influenza work, ideas?].
 We developed a model to incorporate mechanistic processes of epidemic processes and testing.
@@
 There are different testing strategies with test-specific efficacy, the time required for the results to be reported and the cost of the test.
-Testing strategies can be grouped into two broad cathegories including (i) random testing across the whole population and (ii) Non-random/bios testing.
+Testing strategies can be grouped into two broad categories including (i) random testing across the whole population and (ii) Non-random testing.
 Note that contact tracing is an example of non-random testing strategies.
@@
 The advantage of non-random testing comparing to random testing has been investigated specifically during COVID-19 pandemic (refs).
-Further, lots of efforts have been invested in developing widely accessible and fast tests in order to metigate the spread of the pandemic (refs).
-As a comparison of slow versus fast test, the viral PCR test, which is the most accessible test and takes about four days, is very slow relative to the potential tests such as breath-based test \citep{ruszkiewicz2020diagnosis}, which is still in trial stage and takes a few seconds.
+
+Further, lots of efforts have been invested in developing widely accessible and fast tests in order to mitigate the spread of the pandemic (refs).
+In order to have a sense and as a comparison of slow versus fast test, we provide an example.
+The viral PCR test, which is the most accessible test and takes about four days, is very slow relative to the potential tests such as breath-based test \citep{ruszkiewicz2020diagnosis}, which is still in trial stage and takes a few seconds.
 The point to be made is that there can be a potential advantage of slow test.
 
-In this work we developed a model to incorporate epidemic processes and testing processes mechanisticly to study the potential effect of testing strategies and slow-fast testing.
+In this work we developed a model to incorporate epidemic processes and testing processes mechanistically to study the potential effect of testing strategies and slow-fast testing.
 The analysis of the deterministic model included deriving the basic reproduction number, $\Rnum$, as the index of threshold dynamics and the effect of underlying parameters on this key index.
@@
 
-We used a SIR (Susceptible-Infectious-Recovered) deterministic model.
-In order to collect cumulative reported negative or positive tests, two 'integrator' or 'accumulator' compartments, $N$ and $P$, were incorporated.
-The flowchart is
+We developed a deterministic model as a platform of our study.
+A generic individual is identified in a model compartment based on the disease-specific and test-specific criteria.
+The former criterion includes Susceptible, Infectious and Recovered, thus SIR, and the latter criterion group people in either untested, awaiting for the test result with expectation of positive or negative outcome, or reported positive, thus $u$, $p$, $n$ and $t$, respectively, reflected as subscript.
+Also, two 'integrator' or 'accumulator' compartments, $N$ and $P$, were incorporated in the model in order to collect cumulative reported negative or positive tests.
+The flowchart model is as follows
 
@@
 \includegraphics[scale=0.3]{./pix/sir_comp.png}
-\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model. Say more here to make it stand on its own.
+\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model. [Fady] Say more about the compartments and parameters here to make it stand on its own.
 \label{fig:flowchart}} 
@@
 \label{model}
- d S_u/dt &= -\Lambda S_u - F_S S_u + \omega S_n \\
- d S_n/dt &= -\Lambda S_n + (1-p_S) F_S S_u - \omega S_n \\
- d S_p/dt &= -\Lambda S_p + p_S F_S S_u - \omega S_p \\
- d S_t/dt &= -\Lambda S_t + \omega S_p \\
- d I_u/dt &= \Lambda S_u - F_I I_u + \omega I_n  - \gamma I_u  \\
- d I_n/dt &= \Lambda S_n + (1-p_I) F_I I_u - \omega I_n -\gamma I_n \\
- d I_p/dt &= \Lambda S_p + p_I F_I I_u - \omega I_p -\gamma I_p \\
- d I_t/dt &= \Lambda S_t + \omega I_p - \gamma I_t  \\
- d R_u/dt &= \gamma I_u - F_R R_u + \omega R_n \\
- d R_n/dt &= \gamma I_n + (1-p_R) F_R R_u - \omega R_n  \\
- d R_p/dt &= \gamma I_p + p_R F_R R_u  - \omega R_p  \\
- d R_t/dt&= \gamma I_t + \omega R_p  \\
- dN/dt &= \omega (S_n + I_n + R_n)   \\
+ d S_u/dt &= -\Lambda S_u - F_S S_u + \omega S_n, \\
+ d S_n/dt &= -\Lambda S_n + (1-p_S) F_S S_u - \omega S_n, \\
+ d S_p/dt &= -\Lambda S_p + p_S F_S S_u - \omega S_p, \\
+ d S_t/dt &= -\Lambda S_t + \omega S_p, \\
+ d I_u/dt &= \Lambda S_u - F_I I_u + \omega I_n  - \gamma I_u,  \\
+ d I_n/dt &= \Lambda S_n + (1-p_I) F_I I_u - \omega I_n -\gamma I_n, \\
+ d I_p/dt &= \Lambda S_p + p_I F_I I_u - \omega I_p -\gamma I_p, \\
+ d I_t/dt &= \Lambda S_t + \omega I_p - \gamma I_t,  \\
+ d R_u/dt &= \gamma I_u - F_R R_u + \omega R_n, \\
+ d R_n/dt &= \gamma I_n + (1-p_R) F_R R_u - \omega R_n,  \\
+ d R_p/dt &= \gamma I_p + p_R F_R R_u  - \omega R_p,  \\
+ d R_t/dt&= \gamma I_t + \omega R_p,  \\
+ dN/dt &= \omega (S_n + I_n + R_n),   \\
  dP/dt &= \omega(I_p + R_p),
@@
 where $\beta$, (1/day), is the disease transmission rate, $\eta_w$ and $\eta_t$ are the isolation parameters for awaiting and reported individuals, respectively, $\Lambda$ is the force of infection and defined as
+
 \begin{equation}
@@
 \end{equation}
+
 where $N_0$ is the total population size.
@@
 Further, various testing strategies are incorporated through the compartment-specific relative testing weights $W_S$, $W_I$ and $W_R$.
-The weighted number of people available for tests is defined as $W = W_S S_u + W_I I_u + W_R R_u$, and the scaling parameter for testing defined as \begin{equation}
+What matters in these weight is their reltive proportions.
+For example, $W_I/W_S=2$ means that, the probability that infected individuals get tested is as twice as that of susiptibles.
+The weighted number of people available for tests is defined as $W = W_S S_u + W_I I_u + W_R R_u$, and the scaling parameter for testing defined as
+
+\begin{equation}
 \label{sigma}
@@
 \end{equation}
-Thus, the weighted testing rate is defined as $F_Z=\sigma W_Z$, for $Z \in \{S,I,R\}$.
+Thus, the weighted testing rate is defined as $F_Z=\sigma W_Z$, where $Z \in \{S,I,R\}$.
+Note that the particular form of $\sigma$ \eqref{sigma} in the case when no one is left untested, $W \approx 0$, leads to a singularity issue in the simulation.
+We refer the reader to the Appendix for further details and the solution.
 
-- In this model we assumed: (i) all groups being exposed to all others, (ii) there is a single force of infection, $\Lambda$, that is, where a susceptible individual move to does not depend on who infected it, (iii) a perfectly specific test, $p_S=0$, for simplicity in analysis, (iv)  $\eta_t \leq \eta_w$, i.e., the awaiting individuals for test results have a higher transmission probability than the reported individuals.
--why not bith and death are included?
+Some of the underlying assumptions of the model \eqref{model} are as follows: (i) all groups being exposed to all others, (ii) there is a single force of infection, $\Lambda$, that is, where a susceptible individual move to does not depend on who infected it, (iii) a perfectly specific test, $p_S=0$, for simplicity in analysis, (iv)  $\eta_t \leq \eta_w$, i.e., the awaiting individuals for test results have a higher transmission probability than the reported individuals.
+- [Q] why not birth and death are included?
 
-The Disease-Free Equilibrium, DFE hearafter, for the SIR model \ref{model} is given by the solution of the following system including $S_u+S_n=N_0$ and $F_S S_u-\omega S_n=0$. The DFE is
+The Disease-Free Equilibrium, DFE hereafter, for the SIR model \ref{model} is given by solving the coupled system including $S_u+S_n=N_0$ and $F_S S_u-\omega S_n=0$. The DFE is
 
@@
 
-The Basic reproduction number, $\Rnum$, is 
+The Basic reproduction number, $\Rnum$, was calculated by using the next generation matrix method developed by \cite{van2002reproduction}. $\Rnum$ is
+
 \begin{equation}
@@
 where
-$A=\gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_t p_I) F_I$,
-$B=\big(\omega+(F_I+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_t\omega) \omega p_I F_I }{\omega+\gamma}$ and 
-$C=\frac{\beta/\gamma}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}$.
+\begin{align*}
+A=& \gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_t p_I) F_I, \\
+B=& \big(\omega+(F_I+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_t\omega) \omega p_I F_I }{\omega+\gamma}, \\ 
+C=& \frac{\beta/\gamma}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}.
+\end{align*}
+Further details are provided in the Appendix.
 
-$\Rnum$ was calculated by using the generation matrix metod developed by \cite{van2002reproduction}. Further details are provided in the Appendix.
+ [Fady?] 
+ Something needs to be said here about simulation part which was done by using R, using the expression on $\Rnum$ and not the approximation, which give the panel plot in Figure \ref{pan}.
+The analytical calculation of the next generation matrix, $G$, and the derivation of $\Rnum$ was carried out in Maple (ref) by using simple linear Algebra package.
+The numerical simulation of $\Rnum$ in the specified parameter space and plotting the contours of $\Rnum$ was carried out in R by using ggplot (ref?).
+It is notable that in the simulation part, the explicit expression of $\Rnum$ was used and not the approximation \eqref{eq:R0appr}.
 
@@
 
-The following propositions are restlted from the expression of $\Rnum$ and its partial derivatives with respect to the underlying parameters.
+The explicit expression of the basic reproduction number, $\Rnum$ \eqref{R0}, provids an oppotunity to study the influence of changes in the underlying parameters on the critical index of epidemic dynamics.
+We are interested in understanding the effect of parameters that can be realistically controled by changing in isolation, testing intensity and test resulting, i.e., $\eta_t$ and $\eta_w$, $\rho$ and $\omega$, respectively.
+The following Proposition is the direct result of taking partial drivative of $\Rnum$ \eqref{R0} with respect to selected parameters.
 
@@
 \item \label{p1:rho}
-$\partial{\Rnum}/\partial{\rho} \leq 0$ for $\rho \approx 0$.
+$\partial{\Rnum}/\partial{\rho} \leq 0$ when $\rho \approx 0$.
 \item \label{p1:omega}
@@
 
-[Thinking of putting most of these stuff in the Appendix] Given that the perfect isolation occurs when $\eta_t=\eta_w = 0$, \ref{p1:eta} means that lifting isolation results in greater $\Rnum$ and consequently greater number of infected individuals.
-\ref{p1:rho} and \ref{p1:omega} are concluded from the following Taylor approximation of $\Rnum$ at $\rho=0$ given by 
-\begin{equation}
-\label{eq:R0appr}
-R_0 \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_S} \Big(\gamma(\eta_w-1)(\gamma W_S+\omega W_I) + (\eta_t -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2).
-\end{equation}
+ Given that the perfect isolation occurs at $\eta_t=\eta_w = 0$, Prop.
+\ref{p1:eta} means that lifting the isolation from awaiting group results in an increased $\Rnum$ and consequently a greater number of infected individuals.
+We used Taylor approximation of $\Rnum$ at $\rho=0$ due to the complexity of the expression and thus inconclusive.
+See the Appendix for details.
+It is straight forward to derive \ref{p1:rho} and \ref{p1:omega} from the linear approximation of $\Rnum$ at $\rho=0$.
 
-It is straight forward to have
-\begin{equation}
-\label{eq:R0wrtOm}
-\partial{R_0}/\partial{\omega}=  \frac{-\beta \rho}{\gamma W_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c),
-\end{equation}
+ Prop.
+\ref{p1:rho} indicates that increasing testing intensity reduces the $\Rnum$.
+This is sensible since as people are moved to test compartments, $I_p$ or $I_n$, the higher probability of being subject to isolation and the lower the transmission \eqref{Lambda} will become.
 
-where $a=(\eta_w-1)W_I-(\eta_t-1)P_I W_I = ((s-p_I)\eta_t + (p_I-1)) W_I$, $b=2(\eta_w-1)\gamma W_S$ and $c=(\eta_w-1)\gamma^2 W_S$.
-Given that $0 \leq \eta_t\leq \eta_w \leq 1 $, one can easily derive $b\leq 0$ and $c \leq 0$.
-Note that in general, the necessary and sufficient condition for $a \geq 0$ is $(s-p_I) \eta_t \geq (1-p_I)$, where $s=\frac{\eta_w}{\eta_t} \geq 1$.
-As an example, in the case of a very accurate testing regime,  i.e., $P_I=1$, $a \geq 0$ is achieved.
-If $a\geq 0$, the quadratic expression in \eqref{eq:R0wrtOm}, has Real roots.
-Assuming that $\omega_1<0$ and $\omega_2>0$ be the roots of the quadratic expression in $\partial{R_0}/\partial{\omega}$.
-Thus, $\partial{R_0}/\partial{\omega}>0$ for $0<\omega<\omega_2$ and  $\partial{R_0}/\partial{\omega}<0$ for $\omega>\omega_2$.
+ Prop.
+\ref{p1:omega} is the most intresting and somehow the unexpected one.
+It shows that the influence of the test reporting rate, $\omega$, on the epidemic dynamics is dependant on a combination of test reporting rate itself, testing processes/strategies ($W$'s), testing efficacy ($P_I$) and isolation ($\eta$'s).
+Specifically, the $\partial{R_0}/\partial{\omega}>0$ suggests that reproduction number may increase as the test reporting proccess become faster.
+This can be easily noticed from expression \eqref{Rom} in the case of ``perfect isolation``, when $\eta_w=0$ and consequently $\eta_t=0$.
+This provides insight that returning the test results back to the awaiting individuals while they were following a perfect isolation, results in an increase of $\Rnum$ and consequently an increase in the number of infected individuals (see Figure \ref{pan} at lower $\omega$).
+This property also provides a tool to quantify the amount of delay required in the test reporting process as a strategy to reduce $\Rnum$ and consequently control the epidemy.
 
-Note that in case of ``Perfect isolation``, i.e., when $\eta_w=0$ and consequently $\eta_t=0$, it is straight forward to see that $a \leq 0$, $b<0$ and $c<0$.
-Thus, $\partial{\Rnum}/\partial{\omega} \geq 0$.
-Intuitively, this means that returning the test results back to the awaiting individuals while they were following a perfect isolation, results in an increase of $\Rnum$ and consequently an increase in the number of infected individuals (see Figure \ref{fig:panels} at lower $\omega$).
+We also carried out simulation to support the analytical results in Proposition \ref{prop1}.
 
@@
 \caption{Behaviour of the basic reproduction number, $\Rnum$, with respect to changes in the underlying parameters; remind the reader about the parameters and their values in this simulations.}
-\label{fig:panels}
+\label{pan}
 \end{figure}
 
+When random testing regime is applied, $W_S=W_I=W_R$,
+
+(i) in the case of perfect isolation of the reported individuals, $\eta_t=0$, the followings are inferred (see the first few columns of the panel plot); 
+ - variations of the $\eta_w$ has a negligible effect on the critical testing.
+ - for low $\omega$, $R_0$ increases as $\omega$ increases but saturates.
+This means that when the reporting process is very very slow, increasing the speed of the reporting causes higher $R_0$.
+Maybe because more people will move to $I_u$ [?, not very sure here.]
+ (ii) in case of the non-perfect isolation (see the last column), when $\eta_w$ increases (more awaiting people follow the isolation rules) faster reporting reduces $R_0$.  
+Note that it is sensible to expect that $\partial{R_0}/\partial{\omega}<0$. Here we showed that this can be less true. 
+%%%%%%%%%%%%
+[Fady: this subsection and the next belong to the Appendix since they are very detailed. Would you 
+(i) clean up the writing, there are some repeated expressions that can be referenced, (ii) write a paragraph for each to reflect what the message/application is?]
 \subsection{rate of returning tests}
@@
 In other words; our model does not predict that returning test results more rapidly \textit{always} lower $\Rnum$.
-In order to gain insight into this intriguing behaviour, we examine the zeroes of $\frac{\partial{\Rnum}}{\partial{\omega}}(\omega)$.
+In order to gain insight into this intriguing behavior, we examine the zeroes of $\frac{\partial{\Rnum}}{\partial{\omega}}(\omega)$.
 
@@
 
-Now we want to characterize the parameter regions on which $\partial{\Rnum}/\partial{\omega} < 0$ (i.e., the conditions under which returning test results more rapidly is favourable).
+Now we want to characterize the parameter regions on which $\partial{\Rnum}/\partial{\omega} < 0$ (i.e., the conditions under which returning test results more rapidly is favorable).
 By the previous analysis, this is equivalent to solving for $\omega > \omega_2$.
@@
 
-This inequality quantifies the exact relationships between model parameters that result in returning tests more rapidly being favourable.
+This inequality quantifies the exact relationships between model parameters that result in returning tests more rapidly being favorable.
 To give a biological interpretation, we describe the qualitative \textit{trends} predicted by the inequality.
-Returning test results more rapidly \textit{tends} to be more favourable when
+Returning test results more rapidly \textit{tends} to be more favorable when
 
@@
 
-To state an example of this, if test $A$ is three times as expensive as test $B$ (and hence one can test three times as many people with test $B$), using test $A$ rather than $B$ will be favourable only if test $A$ is at least 3 times more sensitive than test $B$.
+To state an example of this, if test $A$ is three times as expensive as test $B$ (and hence one can test three times as many people with test $B$), using test $A$ rather than $B$ will be favorable only if test $A$ is at least 3 times more sensitive than test $B$.
 Note that this is a necessary but not sufficient condition, so even if test $A$ is three times more sensitive, it is still possible for test $B$ to be more effective.
@@
 \cref{eq:rho1vsrho2} tells us precisely when a test corresponding to $\rho_2, P_{i,2}$ will yield a lower $\Rnum$ than a test corresponding to $\rho_1, P_{i,1}$, where $\rho_1 > \rho_2$.
-Some of the qualitative \textit{trends} that favour test 2 (the higher-sensitivity test) include
+Some of the qualitative \textit{trends} that favor test 2 (the higher-sensitivity test) include
 
@@
 
-The next generation matrix, $G = F V^{-1}$, is in the lower triangular form with $F$ and $G$ as follows.
+\subsection{Calculation of $\Rnum$}
 
-\begin{equation}\label{FV}
-F= \beta/N_0 \left[ \begin {array}{cccc} 
+The next generation matrix is $G = F V^{-1}$, where matrix $F$ represents the inflow of new infection to the infected compartments and matrix $V$ represents the flow in the infected compartments when the population is totally susiptable.
+For model \eqref{model}, $F$ and $V$ are
+
+\begin{align}
+\label{FV}
+F =& \beta/N_0 \left[ \begin {array}{cccc} 
 S_u&\eta_w\,S_u&\eta_w\,S_u&\eta_t\,S_u\\
@@
 0&0&0&0
- \end {array} \right],
- V=
+ \end {array} \right], \\
+  V =&
  \left[ \begin {array}{cccc}  
@@
 0&0&-\omega&\gamma
-\end {array} \right],
-\end{equation}
+\end {array} \right].
+\end{align}
 
+The particular form of $F$ with two rows of zeros at the bottom, simplifies $G$ as 
 \begin{equation}
@@
 A\,S_n & B\,S_n
-\end {array}\right],
+\end {array}\right].
 \end{equation}
 
-The Eigenvalues of $G_{11}$, are 0 and $\Rnum$.
+Note that the block matrix $G_{12}$ does not influence $\Rnum$ defined as the spectral radius of $G$. All matters here are the eigenvalues of $G_{11}$, which are 0 and $\Rnum$ \eqref{R0}.
 
 % %%%%%%%
-\section{Literature Review}
+\subsection{On Testing Rate and Numerical Singularity}
+The testing rate, $\sigma$, should be formulated such that people from the untested compartments will not be tested if they are not there.
+The numerical singularity issue with the chosen $\sigma$ \eqref{sigma} is that the population in $S$ compartments appeared to blow up when the DFE is achieved.
+This is once the only untested people are susceptibles, the FOI will become $\Lambda=0$, testing rate $F_s=\rho N_0/S_u$.
+Thus, eq(1) of the model will be $d S_u/dt = - \rho N_0 + \omega S_n$ which is no longer dependent on $S_u$ and a linear rate of leaving the $S_u$ compartment.
 
-\subsection{Explicit models of TTI (trace/test/isolate) based on network or agent-based models}
+One way to fix this issue, is to consider a maximum testing rate, $\tau$ (1/day).
+In general, we want to test at a rate of $\rho$ across the whole population.
+This won't always be possible, so we impose a maximum rate of $\tau$ per testable person and redefine $\sigma = \frac{\tau \rho N_0}{\tau W + \rho N_0}$, with the assumption that $\tau \gg \rho$.
+This alteration in $\sigma$, does not change any results related to $\Rnum$, thus we only impose it in the simulation of the epidemic dynamic.
+
+% %%%%%%%
+\subsection{Taylor Approximation of $\Rnum$ at $\rho=0$ }
+
+The basic reproduction number, $\Rnum$, close to $\rho=0$ can be approximated linearly in $\rho$ by using Taylor approximation. it follows
+
+\begin{equation}
+\label{eq:R0appr}
+R_0 \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_S} \Big(\gamma(\eta_w-1)(\gamma W_S+\omega W_I) + (\eta_t -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2).
+\end{equation}
+
+Also, to analyze the influence of $\omega$ on $\Rnum$, the approximation \eqref{eq:R0appr} was used. Then it is straight forward to have
+
+\begin{equation}
+\label{Rom}
+\partial{R_0}/\partial{\omega}=  \frac{-\beta \rho}{\gamma W_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c),
+\end{equation}
+
+where $a=(\eta_w-1)W_I-(\eta_t-1)P_I W_I = ((s-p_I)\eta_t + (p_I-1)) W_I$, $b=2(\eta_w-1)\gamma W_S$ and $c=(\eta_w-1)\gamma^2 W_S$.
+Given that $0 \leq \eta_t\leq \eta_w \leq 1 $, one can easily derive $b\leq 0$ and $c \leq 0$.
+
+Note that in general, the necessary and sufficient condition for $a \geq 0$ is $(s-p_I) \eta_t \geq (1-p_I)$, where $s=\frac{\eta_w}{\eta_t} \geq 1$.
+
+Note that in case of ``perfect isolation``, i.e., when $\eta_w=0$ and consequently $\eta_t=0$, it is straight forward to see that $a \leq 0$, $b<0$ and $c<0$.
+Thus, $\partial{\Rnum}/\partial{\omega} \geq 0$.
+As an example, in the case of a very accurate testing regime,  i.e., $P_I=1$, $a \geq 0$ is achieved.
+If $a\geq 0$, the quadratic expression in \eqref{Rom}, has Real roots.
+Assuming that $\omega_1<0$ and $\omega_2>0$ be the roots of the quadratic expression in $\partial{R_0}/\partial{\omega}$.
+Thus, $\partial{R_0}/\partial{\omega}>0$ for $0<\omega<\omega_2$ and  $\partial{R_0}/\partial{\omega}<0$ for $\omega>\omega_2$.
+
+% %%%%%%%
+\subsection{Literature Review}
+
+\subsubsection{Explicit models of TTI (trace/test/isolate) based on network or agent-based models}
 \citep{endo2020implication} [Ali: It seems to me that this is just a statistical model to estimate the parent-offspring of an infected index, not sure if it fits into agent-based group!] Used simulation on a branching process model to assess the forward and backward contact tracing efficiency.

### 2020-11-19 SIR_manuscript.tex
version 0

@@
 It is notable that in the simulation part, the explicit expression of $\Rnum$ was used and not the approximation \eqref{eq:R0appr}.
+This simulation reflects the behavior of $\Rnum$ with respect to the selected parameters for two different testing strategies: (i) random testing, represented by all testing weights to be the same, $W_S=W_I=W_R$, and (ii) non-random testing, when testing weight are not equal.
+For simmulation purposes we chose $W_S=W_I=W_R=1$ for random testing, and $W_S=0.3$ and $W_I=W_R=1$ for non-random testing strategy.
+Note that the critical contour of $\Rnum=1$ is plotted in solid line in Figure \ref{pan}.
 
@@
 
-We also carried out simulation to support the analytical results in Proposition \ref{prop1}.
-
 [comments about the figure:]
@@
 \includegraphics[width=\linewidth]{./pix/R0contour_random.png}
+\caption{}\label{p.a}
 \end{subfigure}
@@
 \includegraphics[width=\linewidth]{./pix/R0contour_TTI.png}
+\caption{}\label{p.b}
 \end{subfigure}
-\caption{Behaviour of the basic reproduction number, $\Rnum$, with respect to changes in the underlying parameters; remind the reader about the parameters and their values in this simulations.}
+\caption{Behaviour of the basic reproduction number, $\Rnum$, with respect to changes in the underlying parameters; remind the reader about the parameters and their values in this simulations.
+the solid line in each panel represents the critical contour of $\Rnum=1$.}
 \label{pan}
@@
 
-When random testing regime is applied, $W_S=W_I=W_R$,
+Our numerical simulation in support of the analytical results in Proposition \ref{prop1} are presented in Figure \ref{pan}.
+It can be infered that when random testing strategy is applied, Figure \ref{p.a}, comparing to the non-random testing strategy, Figure \ref{p.a}, the critical testing intensity of the coresponding panels are lower in non-random testing.
+Also, it appears that speeding the test reporting, i.e., increasing $\omega$, does not significantly lower $\Rnum$ when awaiting people follow the isolation perfectly, i.e., $\eta_w$ is closer to 0.
+However, speeding the test reporting reduces the epidemy more when the awaiting people less follow the isolation, i.e., $\eta_w$ is closer to 1.
 
-(i) in the case of perfect isolation of the reported individuals, $\eta_t=0$, the followings are inferred (see the first few columns of the panel plot); 
- - variations of the $\eta_w$ has a negligible effect on the critical testing.
- - for low $\omega$, $R_0$ increases as $\omega$ increases but saturates.
-This means that when the reporting process is very very slow, increasing the speed of the reporting causes higher $R_0$.
-Maybe because more people will move to $I_u$ [?, not very sure here.]
- (ii) in case of the non-perfect isolation (see the last column), when $\eta_w$ increases (more awaiting people follow the isolation rules) faster reporting reduces $R_0$.  
-Note that it is sensible to expect that $\partial{R_0}/\partial{\omega}<0$. Here we showed that this can be less true. 
-%%%%%%%%%%%%
-[Fady: this subsection and the next belong to the Appendix since they are very detailed. Would you 
-(i) clean up the writing, there are some repeated expressions that can be referenced, (ii) write a paragraph for each to reflect what the message/application is?]
-\subsection{rate of returning tests}
-The linearization of $\Rnum$ around $\rho=0$ is
-\begin{equation}\label{linearization}
-\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta_w-1)(\gamma W_s+\omega W_i) + (\eta_t -1)P_iW_i \omega^2 \Big). 
-\end{equation}
-
-So when $\rho \approx 0$ we have $$\partial{\Rnum}/\partial{\omega} \approx  \frac{-\beta \rho}{\gamma W_s\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c).$$
-
-where $a=(\eta_w-1)W_i-(\eta_t-1)P_iW_i$, $b=2(\eta_w-1)\gamma W_s$ and $c=(\eta_w-1)\gamma^2 W_s$.
-
-Perhaps counter-intuitively, the equation above does not predict that $\Rnum$ is monotone decreasing with respect to $\omega$.
-In other words; our model does not predict that returning test results more rapidly \textit{always} lower $\Rnum$.
-In order to gain insight into this intriguing behavior, we examine the zeroes of $\frac{\partial{\Rnum}}{\partial{\omega}}(\omega)$.
-
-Defining the following quantity, $Q$, will help us write the roots of $\partial{\Rnum}/\partial{\omega}$ neatly. 
-\begin{align}\label{eq:defQ}
-    Q =& \frac{W_i}{W_s}\left(1-\frac{n_{t}-1}{n_{w}-1}P_{i}\right) \\
-\end{align}
-
-With that in mind, we can write the roots of $\partial{\Rnum}/\partial{\omega}$ as
-
-\begin{align}
-    \omega_1 =& \frac{\gamma}{-\sqrt{1-Q}-1} \\
-    \omega_2 =& \frac{\gamma}{\sqrt{1-Q}-1}
-\end{align}
-
-Note that the zeroes are real if and only if $Q < 1$.
-Note that have $\eta_t < \eta_w$, so if $P_i \approx 1$, we will have $Q < 0 < 1$.
-Thus, if we assume near-perfect test sensitivity, $\omega_1$ and $\omega_2$ will be real.
-
-Assuming $\omega_1, \omega_2$ are real, it is easy to confirm that $\omega_1 < 0$ by looking at the denominator.
-To see that $\omega_2 > 0$, recall that $Q < 0$, so $\sqrt{1-Q} > 1$ and so $\sqrt{1-Q} -1 > 0$.
-Knowing that $\omega_1 < 0$, the only root of interest (i.e., biologically relevant quantity) is $\omega_2$.
-
-We can prove that $\partial{\Rnum}/\partial{\omega} > 0$ when $\omega \in (0,\omega_2)$ and $\partial{\Rnum}/\partial{\omega} < 0$ when $\omega \in (\omega_2,\infty)$ by computing the limits of $\partial{\Rnum}/\partial{\omega}$ at $0$ and $\infty$ respectively.
-So it follows that $\Rnum$ has a global maximum with respect to $\omega$ at $\omega = \omega_2$.
-
-Now we want to characterize the parameter regions on which $\partial{\Rnum}/\partial{\omega} < 0$ (i.e., the conditions under which returning test results more rapidly is favorable).
-By the previous analysis, this is equivalent to solving for $\omega > \omega_2$.
-So
-
-\begin{align}\label{eq:necsuf}
-    &\omega > \omega_2 \nonumber \\
-    &\omega > \frac{\gamma}{\sqrt{1-Q}-1} \nonumber \\
-    &\vdots \nonumber \\
-    &\frac{1-n_{t}}{1-n_{w}}P_{i}>\frac{W_{s}}{W_{i}}\left(\frac{\gamma}{\omega}+1\right)^{2}
-\end{align}
-
-This inequality quantifies the exact relationships between model parameters that result in returning tests more rapidly being favorable.
+Further, we derived an expression that quantifies the exact relationships between model parameters that result in returning tests more rapidly being favorable \eqref{eq:necsuf}.
 To give a biological interpretation, we describe the qualitative \textit{trends} predicted by the inequality.
@@
 
-\subsection{Expensive vs. cheap tests}
-
-The use of tests cheaper than RT-PCR has been proposed as a potential strategy for containing the COVID-19 pandemic.
-While cheaper tests may be less sensitive and reliable than RT-PCR, they allow for broader and more intense testing.
-In the analysis below, we compare the $\Rnum$ predicted by our model depending on the testing strategy.
-
-Consider a test that allows us to test at rate $\rho_1$ and has sensitivity $P_{i,1}$, and another test that allows us to test at $\rho_2$ and has sensitivity $P_{i,2}$.
-Suppose that $\rho_1 > \rho_2$.
-Recall that the linearization of $\Rnum$ around $\rho \approx 0$ is given by $$\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta_w-1)(\gamma W_s+\omega W_i) + (\eta_t -1)P_iW_i \omega^2 \Big).$$
-
-Treating $\Rnum$ as a function of $\rho$ and $P_i$,we can reduce the inequality $$\Rnum(\rho_2, P_{i,2}) < \Rnum(\rho_1, P_{i,1})$$ into
-
-\begin{align}\label{eq:rho1vsrho2}
-    &\rho_1\left(\gamma(\eta_w-1)(\gamma W_s + \omega W_i) + (\eta_t-1)P_{i, 1}W_i\omega^2\right) - \rho_2\left(\gamma(\eta_w-1)(\gamma W_s + \omega W_i) + (\eta_t-1)P_{i, 2}W_i\omega^2\right) > 0 \nonumber \\
-    &\vdots \nonumber \\
-    &\frac{\rho_2P_{i, 2}-\rho_1P_{i, 1} }{\rho_1-\rho_2} > \frac{1-\eta_w}{1-\eta_t}\cdot \frac{\gamma(\gamma W_s + \omega W_i)}{\omega^2 W_i}
-\end{align}
-
-Note that the RHS is positive, thus a necessary condition for the inequality above to hold is that $\rho_2P_{i,2} > \rho_1P_{i,1}$, equivalently
-
-\begin{equation}
-\frac{P_{i,2}}{P_{i,1}} > \frac{\rho_1}{\rho_2}.
-\end{equation}
-
-To state an example of this, if test $A$ is three times as expensive as test $B$ (and hence one can test three times as many people with test $B$), using test $A$ rather than $B$ will be favorable only if test $A$ is at least 3 times more sensitive than test $B$.
-Note that this is a necessary but not sufficient condition, so even if test $A$ is three times more sensitive, it is still possible for test $B$ to be more effective.
-
-\cref{eq:rho1vsrho2} tells us precisely when a test corresponding to $\rho_2, P_{i,2}$ will yield a lower $\Rnum$ than a test corresponding to $\rho_1, P_{i,1}$, where $\rho_1 > \rho_2$.
-Some of the qualitative \textit{trends} that favor test 2 (the higher-sensitivity test) include
-
-\begin{itemize}
-    \item individuals who test positive self-isolate much more than individuals who are waiting for their test result.
-    \item the time it takes to return tests is much shorter than the mean infectious period.
-    \item the testing intensity is much greater for infected individuals than susceptible individuals.
-\end{itemize}
+[Fady, would you sumerize the results of "Expensive vs. cheap tests" into a paragraph and leave the Mathy stuff in the Appendix.]
+%%%%%%%%%%%%
 
@@
 % %%%%%%%
+[Fady: this subsection and most of the next belong to the Appendix since they are very detailed. Would you clean up the writing, there are some repeated expressions that can be referenced?]
+
+\subsection{rate of returning tests}
+The linearization of $\Rnum$ around $\rho=0$ is
+\begin{equation}\label{linearization}
+\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta_w-1)(\gamma W_s+\omega W_i) + (\eta_t -1)P_iW_i \omega^2 \Big). 
+\end{equation}
+
+So when $\rho \approx 0$ we have $$\partial{\Rnum}/\partial{\omega} \approx  \frac{-\beta \rho}{\gamma W_s\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c).$$
+
+where $a=(\eta_w-1)W_i-(\eta_t-1)P_iW_i$, $b=2(\eta_w-1)\gamma W_s$ and $c=(\eta_w-1)\gamma^2 W_s$.
+
+Perhaps counter-intuitively, the equation above does not predict that $\Rnum$ is monotone decreasing with respect to $\omega$.
+In other words; our model does not predict that returning test results more rapidly \textit{always} lower $\Rnum$.
+In order to gain insight into this intriguing behavior, we examine the zeroes of $\frac{\partial{\Rnum}}{\partial{\omega}}(\omega)$.
+
+Defining the following quantity, $Q$, will help us write the roots of $\partial{\Rnum}/\partial{\omega}$ neatly. 
+\begin{align}\label{eq:defQ}
+    Q =& \frac{W_i}{W_s}\left(1-\frac{n_{t}-1}{n_{w}-1}P_{i}\right) \\
+\end{align}
+
+With that in mind, we can write the roots of $\partial{\Rnum}/\partial{\omega}$ as
+
+\begin{align}
+    \omega_1 =& \frac{\gamma}{-\sqrt{1-Q}-1} \\
+    \omega_2 =& \frac{\gamma}{\sqrt{1-Q}-1}
+\end{align}
+
+Note that the zeroes are real if and only if $Q < 1$.
+Note that have $\eta_t < \eta_w$, so if $P_i \approx 1$, we will have $Q < 0 < 1$.
+Thus, if we assume near-perfect test sensitivity, $\omega_1$ and $\omega_2$ will be real.
+
+Assuming $\omega_1, \omega_2$ are real, it is easy to confirm that $\omega_1 < 0$ by looking at the denominator.
+To see that $\omega_2 > 0$, recall that $Q < 0$, so $\sqrt{1-Q} > 1$ and so $\sqrt{1-Q} -1 > 0$.
+Knowing that $\omega_1 < 0$, the only root of interest (i.e., biologically relevant quantity) is $\omega_2$.
+
+We can prove that $\partial{\Rnum}/\partial{\omega} > 0$ when $\omega \in (0,\omega_2)$ and $\partial{\Rnum}/\partial{\omega} < 0$ when $\omega \in (\omega_2,\infty)$ by computing the limits of $\partial{\Rnum}/\partial{\omega}$ at $0$ and $\infty$ respectively.
+So it follows that $\Rnum$ has a global maximum with respect to $\omega$ at $\omega = \omega_2$.
+
+Now we want to characterize the parameter regions on which $\partial{\Rnum}/\partial{\omega} < 0$ (i.e., the conditions under which returning test results more rapidly is favorable).
+By the previous analysis, this is equivalent to solving for $\omega > \omega_2$.
+So
+
+\begin{align}\label{eq:necsuf}
+    &\omega > \omega_2 \nonumber \\
+    &\omega > \frac{\gamma}{\sqrt{1-Q}-1} \nonumber \\
+    &\vdots \nonumber \\
+    &\frac{1-n_{t}}{1-n_{w}}P_{i}>\frac{W_{s}}{W_{i}}\left(\frac{\gamma}{\omega}+1\right)^{2}
+\end{align}
+
+\subsection{Expensive vs. cheap tests}
+
+The use of tests cheaper than RT-PCR has been proposed as a potential strategy for containing the COVID-19 pandemic.
+While cheaper tests may be less sensitive and reliable than RT-PCR, they allow for broader and more intense testing.
+In the analysis below, we compare the $\Rnum$ predicted by our model depending on the testing strategy.
+
+Consider a test that allows us to test at rate $\rho_1$ and has sensitivity $P_{i,1}$, and another test that allows us to test at $\rho_2$ and has sensitivity $P_{i,2}$.
+Suppose that $\rho_1 > \rho_2$.
+Recall that the linearization of $\Rnum$ around $\rho \approx 0$ is given by $$\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta_w-1)(\gamma W_s+\omega W_i) + (\eta_t -1)P_iW_i \omega^2 \Big).$$
+
+Treating $\Rnum$ as a function of $\rho$ and $P_i$,we can reduce the inequality $$\Rnum(\rho_2, P_{i,2}) < \Rnum(\rho_1, P_{i,1})$$ into
+
+\begin{align}\label{eq:rho1vsrho2}
+    &\rho_1\left(\gamma(\eta_w-1)(\gamma W_s + \omega W_i) + (\eta_t-1)P_{i, 1}W_i\omega^2\right) - \rho_2\left(\gamma(\eta_w-1)(\gamma W_s + \omega W_i) + (\eta_t-1)P_{i, 2}W_i\omega^2\right) > 0 \nonumber \\
+    &\vdots \nonumber \\
+    &\frac{\rho_2P_{i, 2}-\rho_1P_{i, 1} }{\rho_1-\rho_2} > \frac{1-\eta_w}{1-\eta_t}\cdot \frac{\gamma(\gamma W_s + \omega W_i)}{\omega^2 W_i}
+\end{align}
+
+Note that the RHS is positive, thus a necessary condition for the inequality above to hold is that $\rho_2P_{i,2} > \rho_1P_{i,1}$, equivalently
+
+\begin{equation}
+\frac{P_{i,2}}{P_{i,1}} > \frac{\rho_1}{\rho_2}.
+\end{equation}
+
+To state an example of this, if test $A$ is three times as expensive as test $B$ (and hence one can test three times as many people with test $B$), using test $A$ rather than $B$ will be favorable only if test $A$ is at least 3 times more sensitive than test $B$.
+Note that this is a necessary but not sufficient condition, so even if test $A$ is three times more sensitive, it is still possible for test $B$ to be more effective.
+
+\cref{eq:rho1vsrho2} tells us precisely when a test corresponding to $\rho_2, P_{i,2}$ will yield a lower $\Rnum$ than a test corresponding to $\rho_1, P_{i,1}$, where $\rho_1 > \rho_2$.
+Some of the qualitative \textit{trends} that favor test 2 (the higher-sensitivity test) include
+
+\begin{itemize}
+    \item individuals who test positive self-isolate much more than individuals who are waiting for their test result.
+    \item the time it takes to return tests is much shorter than the mean infectious period.
+    \item the testing intensity is much greater for infected individuals than susceptible individuals.
+\end{itemize}
+
+% %%%%%%%
 \subsection{Literature Review}

### 2020-11-23 SIR_manuscript.tex
Add files via upload

@@
 
-One of the real challenges interpreting COVID-19 dynamics is to unravel/investigate information gained from data.
-One of the strongest signals we get to see is the daily case reports which are driven by a combination of two processes: (i) epidemic processes and (ii) testing processes.
-The real challenge here is to separate these two processes.
+One of the main challenges in interpreting COVID-19 dynamics is to unravel/investigate information gained from data.
+The daily case count is a highly accessible form of data where observations are driven by two clusters of processes: (i) epidemic processes and (ii) testing processes.
+The challenge is then to separate these processes and understand how they interact.
 Simple mathematical models have a long history in providing useful insights facing with real epidemiological challenges \citep{ross1911prevention} [maybe add influenza work, ideas?].
-We developed a model to incorporate mechanistic processes of epidemic processes and testing.
+We developed a model that incorporates mechanistic processes of epidemic processes and testing.
 We used the model to study the potential effect of testing strategies and slow-fast reporting testing result.
 
-There are different testing strategies with test-specific efficacy, the time required for the results to be reported and the cost of the test.
+There are different testing strategies with different test-specific efficacy, the time required for the results to be reported and the cost of the test.
 Testing strategies can be grouped into two broad categories including (i) random testing across the whole population and (ii) Non-random testing.
@@
 Each of these strategies has been used as a tool to study COVID-19; see (refs) for group (i) and (refs) for group (ii).
-The advantage of non-random testing comparing to random testing has been investigated specifically during COVID-19 pandemic (refs).
+The advantage of non-random testing comparing to random testing has been investigated specifically during the COVID-19 pandemic (refs).
 
-Further, lots of efforts have been invested in developing widely accessible and fast tests in order to mitigate the spread of the pandemic (refs).
+Much effort has been invested in developing widely accessible and fast tests in order to mitigate the spread of the pandemic (refs).
 In order to have a sense and as a comparison of slow versus fast test, we provide an example.
-The viral PCR test, which is the most accessible test and takes about four days, is very slow relative to the potential tests such as breath-based test \citep{ruszkiewicz2020diagnosis}, which is still in trial stage and takes a few seconds.
+The viral PCR test, which is the most commonly used test and takes about four days, is very slow relative to the potential tests such as breath-based test \citep{ruszkiewicz2020diagnosis}, which is still in trial stage and takes a few seconds.
 The point to be made is that there can be a potential advantage of slow test.
@@
 We found that the non-random testing strategy is more effective than the random testing.
-Specifically, the required critical testing intensity to keep $\Rnum$ lower than the threshold value of one, is much lower in the non-random testing than the random one.
-Further, we show that there is a potential advantage of slow reporting of test result on the spread of the infection.
+Specifically, the required critical testing intensity to keep $\Rnum$ lower than the threshold value of one is much lower in the non-random testing than the random one.
+Further, we show that for certain parameter combinations, there may be a potential advantage to slow reporting of test results on the spread of the infection.
 
@@
 A generic individual is identified in a model compartment based on the disease-specific and test-specific criteria.
-The former criterion includes Susceptible, Infectious and Recovered, thus SIR, and the latter criterion group people in either untested, awaiting for the test result with expectation of positive or negative outcome, or reported positive, thus $u$, $p$, $n$ and $t$, respectively, reflected as subscript.
+The former criteria include Susceptible, Infectious and Recovered, thus SIR, and the latter criteria group people in either untested, awaiting for the test result with expectation of positive or negative outcome, or reported positive, thus $u$, $p$, $n$ and $t$, respectively, reflected as subscript.
 Also, two 'integrator' or 'accumulator' compartments, $N$ and $P$, were incorporated in the model in order to collect cumulative reported negative or positive tests.
@@
 
-where $\beta$, (1/day), is the disease transmission rate, $\eta_w$ and $\eta_t$ are the isolation parameters for awaiting and reported individuals, respectively, $\Lambda$ is the force of infection and defined as
+where $\beta$, (in units of 1/day), is the disease transmission rate, $\eta_w$ and $\eta_t$ are the isolation parameters for awaiting and reported individuals, respectively.
+$\Lambda$ is the force of infection and is defined as
 
@@
 Further, various testing strategies are incorporated through the compartment-specific relative testing weights $W_S$, $W_I$ and $W_R$.
-What matters in these weight is their reltive proportions.
-For example, $W_I/W_S=2$ means that, the probability that infected individuals get tested is as twice as that of susiptibles.
+What matters in these weight is their relative proportions.
+For example, $W_I/W_S=2$ means that, the probability that infected individuals get tested is as twice as that of susceptible individuals.
 The weighted number of people available for tests is defined as $W = W_S S_u + W_I I_u + W_R R_u$, and the scaling parameter for testing defined as
@@
 
-Some of the underlying assumptions of the model \eqref{model} are as follows: (i) all groups being exposed to all others, (ii) there is a single force of infection, $\Lambda$, that is, where a susceptible individual move to does not depend on who infected it, (iii) a perfectly specific test, $p_S=0$, for simplicity in analysis, (iv)  $\eta_t \leq \eta_w$, i.e., the awaiting individuals for test results have a higher transmission probability than the reported individuals.
+Some of the underlying assumptions of the model \eqref{model} are as follows: (i) all groups are being exposed to all others, (ii) there is a single force of infection, $\Lambda$, that is, where a susceptible individual moves to does not depend on who infected it, (iii) a perfectly specific test, $p_S=0$, for simplicity in analysis, (iv)  $\eta_t \leq \eta_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
 - [Q] why not birth and death are included?
@@
 
- [Fady?] 
+ [Fady? \fady{I didn't do any simulations though}] 
  Something needs to be said here about simulation part which was done by using R, using the expression on $\Rnum$ and not the approximation, which give the panel plot in Figure \ref{pan}.
@@
 
-The explicit expression of the basic reproduction number, $\Rnum$ \eqref{R0}, provids an oppotunity to study the influence of changes in the underlying parameters on the critical index of epidemic dynamics.
-We are interested in understanding the effect of parameters that can be realistically controled by changing in isolation, testing intensity and test resulting, i.e., $\eta_t$ and $\eta_w$, $\rho$ and $\omega$, respectively.
-The following Proposition is the direct result of taking partial drivative of $\Rnum$ \eqref{R0} with respect to selected parameters.
+The explicit formula for the basic reproduction number, $\Rnum$ \eqref{R0}, provides an opportunity to study the influence of changes in the underlying parameters on the critical index of epidemic dynamics.
+We are interested in understanding the effect of parameters that can be realistically controlled by changing in isolation, testing intensity and test resulting, i.e., $\eta_t$ and $\eta_w$, $\rho$ and $\omega$, respectively.
+The following Proposition is the direct result of taking partial derivative of $\Rnum$ \eqref{R0} with respect to selected parameters.
 
@@
 See the Appendix for details.
-It is straight forward to derive \ref{p1:rho} and \ref{p1:omega} from the linear approximation of $\Rnum$ at $\rho=0$.
+It is straightforward to derive \ref{p1:rho} and \ref{p1:omega} from the linear approximation of $\Rnum$ at $\rho=0$.
 
@@
 \ref{p1:rho} indicates that increasing testing intensity reduces the $\Rnum$.
-This is sensible since as people are moved to test compartments, $I_p$ or $I_n$, the higher probability of being subject to isolation and the lower the transmission \eqref{Lambda} will become.
+This is sensible since as people are moved to test compartments ($I_p$, $I_n$ \fady{the susceptibles also reduce contact and transmission}) the higher probability of being subject to isolation and the lower the transmission \eqref{Lambda} will become.
 
- Prop.
-\ref{p1:omega} is the most intresting and somehow the unexpected one.
-It shows that the influence of the test reporting rate, $\omega$, on the epidemic dynamics is dependant on a combination of test reporting rate itself, testing processes/strategies ($W$'s), testing efficacy ($P_I$) and isolation ($\eta$'s).
+ Of these results, prop.
+\ref{p1:omega} is the most surprising.
+It states that returning test results more rapidly (i.e., increasing $\omega$) does not necessarily lower $\Rnum$, rather, whether increasing $\omega$ lowers $\Rnum$ is dependant on the precise combination of model parameters (test reporting rate itself, testing processes/strategies ($W$'s), testing efficacy ($P_I$) and isolation ($\eta$'s).
 Specifically, the $\partial{R_0}/\partial{\omega}>0$ suggests that reproduction number may increase as the test reporting proccess become faster.
-This can be easily noticed from expression \eqref{Rom} in the case of ``perfect isolation``, when $\eta_w=0$ and consequently $\eta_t=0$.
-This provides insight that returning the test results back to the awaiting individuals while they were following a perfect isolation, results in an increase of $\Rnum$ and consequently an increase in the number of infected individuals (see Figure \ref{pan} at lower $\omega$).
-This property also provides a tool to quantify the amount of delay required in the test reporting process as a strategy to reduce $\Rnum$ and consequently control the epidemy.
+This can be seen from expression \eqref{Rom} in the case of ``perfect isolation``, when $\eta_w=0$ and $\eta_t=0$.
+The underlying mechanism for this result is that individuals may take more precautionary behaviours (e.g., physical distancing) when they are awaiting test results compared to when they are untested or their test has returned negative.
+An example of a situation that would favour slow return of tests is if the test being employed produces many false negatives, because many infected individuals will believe they are negative and thus may unknowingly spread the virus to many others.
+For an example of the behaviour of $\Rnum$ see Figure \ref{pan} at varying values of $\omega$.
+This property also provides a tool to quantify the amount of delay required in the test reporting process as a strategy to reduce $\Rnum$ and consequently control the epidemic.
 
@@
 
-Further, we derived an expression that quantifies the exact relationships between model parameters that result in returning tests more rapidly being favorable \eqref{eq:necsuf}.
+Furthermore, we derived an inequality that quantifies the exact relationships between model parameters that result in returning tests more rapidly being favorable \eqref{eq:necsuf}.
 To give a biological interpretation, we describe the qualitative \textit{trends} predicted by the inequality.
@@
 
-[Fady, would you sumerize the results of "Expensive vs. cheap tests" into a paragraph and leave the Mathy stuff in the Appendix.]
+The use of tests cheaper than RT-PCR has been proposed as a potential strategy for containing the COVID-19 pandemic.
+While cheaper tests may be less sensitive and reliable than RT-PCR, they allow for broader and more intense testing.
+Using our Taylor approximation of $\Rnum$ near $\rho = 0$, we examined what circumstances (i.e., model parameters) make the use of one test more favourable than another, and give a complete description through inequality \ref{eq:rho1vsrho2}.
+In general, we found that the expensive test \textit{tends} to more effectively lower $\Rnum$ when
+
+\begin{itemize}
+    \item individuals who test positive self-isolate much more than individuals who are waiting for their test result.
+    \item the time it takes to return tests is much shorter than the mean infectious period.
+    \item the testing intensity is much greater for infected individuals than susceptible individuals.
+\end{itemize}
+
+Our model also predicts an asymmetry between cheap and expensive tests: if a cheap test can identify on average more infected individuals as an expensive test, then our model predicts that the cheap test will lower $\Rnum$ more.
+In contrast, if an expensive test can identify on average more infected individuals, it will not necessarily lower $\Rnum$ more than the cheap test.
+\fady{we can state this as a formal proposition in equations rather than in words, if that would be better.}
+
 %%%%%%%%%%%%
@@
 \subsection{Rate of returning tests}
-\fady{needs to be expanded}
-When the difference in self-isolation between individuals waiting for their tests and individuals who are confirmed positive is small, returning tests rapidly is not beneficial because the benefit in doing so (i.e., the extra self-isolating when an individual tests positive) is minimal and outweighed by the fact that those who test negative may increase their contacts.
-Similarly, if the test is not sensitive, many infected individuals will test negative, and thus may increase their contacts under the false belief that they are not infectious, hence returning test results more quickly in this case may not be beneficial.
+
+Our model predicts that when the difference in self-isolation between individuals waiting for their tests and individuals who are confirmed positive is small, returning tests rapidly is not beneficial.
+The underlying mechanism for this prediction is that the benefit in returning tests rapidly so (i.e., the extra self-isolation an individual does when they test positive) is minimal and outweighed by the fact that those who test negative may increase their contacts.
+Thus, in such a case, returning tests more rapidly is expected to increase $\Rnum$.
+Similarly, if the test is not sensitive, many infected individuals will test negative, and thus may increase their contacts under the false belief that they are not infectious, so returning tests rapidly in this case may also contribute to increased transmission.
 
 When testing of infected individuals is more intense than susceptible individuals, returning test results rapidly is beneficial because those infected individuals will be informed and self-isolate.
-Conversely, if testing of susceptible individuals is much greater, very few tests be positive, and the benefit of informing those few positive cases will be outweighed by the potential increase in contacts of the individuals testing negative.
+Conversely, if testing of susceptible individuals is much greater, very few tests will be positive, and the benefit of informing those few positive cases will be outweighed by the potential increase in contacts of the individuals testing negative.
 
 \subsection{Expensive vs. cheap tests}
-\fady{needs to be written \dots must discuss the fact that our analysis holds only when $\rho$ is relatively small, so if there is a test so cheap that $\rho$ does not stay small, our results do not apply.}
+\fady{I think we should discuss this before I write it. I'm a little hesitant to really expand on these results because of the points JD brought up over email.}
 

### 2020-11-23 SIR_manuscript.tex
some updates

@@
 \usepackage[nameinlink,capitalize]{cleveref}
+
+\usepackage{xspace}
+\newcommand{\Rlogo}{\protect\includegraphics[height=2ex,keepaspectratio]
+{pix/Rlogo.pdf}\xspace}
 
@@
 
-Some of the underlying assumptions of the model \eqref{model} are as follows: (i) all groups are being exposed to all others, (ii) there is a single force of infection, $\Lambda$, that is, where a susceptible individual moves to does not depend on who infected it, (iii) a perfectly specific test, $p_S=0$, for simplicity in analysis, (iv)  $\eta_t \leq \eta_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
+The classical SIR model is based on the following implicite assumptions (see a ref for details); well-mixed population, homogeneity of the population, exponentially distributed duration of infection and large population size (may be refer the reader to \cite{keeling2011modeling}).
+In addition to these standard assumptions, our model \eqref{model} assumes: (i) there is a single force of infection, $\Lambda$, that is, where a susceptible individual moves to does not depend on who infected it, (ii) a perfectly specific test, $p_S=0$, for simplicity in analysis, (iii)  $\eta_t \leq \eta_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
 - [Q] why not birth and death are included?
+say it or not?
 
-The Disease-Free Equilibrium, DFE hereafter, for the SIR model \ref{model} is given by solving the coupled system including $S_u+S_n=N_0$ and $F_S S_u-\omega S_n=0$. The DFE is
+The Disease-Free Equilibrium (Ali, not sure if this is the right term?), DFE hereafter, for the SIR model \ref{model} is given by solving the coupled system including $S_u+S_n=N_0$ and $F_S S_u-\omega S_n=0$.
+The DFE is
 
@@
 
- [Fady? \fady{I didn't do any simulations though}] 
- Something needs to be said here about simulation part which was done by using R, using the expression on $\Rnum$ and not the approximation, which give the panel plot in Figure \ref{pan}.
+  [to be completed!
+Something needs to be said here about simulation part].
 The analytical calculation of the next generation matrix, $G$, and the derivation of $\Rnum$ was carried out in Maple (ref) by using simple linear Algebra package.
-The numerical simulation of $\Rnum$ in the specified parameter space and plotting the contours of $\Rnum$ was carried out in R by using ggplot (ref?).
+  We used \Rlogo for the simulation part which included using the explicite expression of $\Rnum$ \ref{R0} as a function of the underlying parameters, specifying the parameters a realistic range, calculating the corresponding $\Rnum$ and plotting the contours of $\Rnum$ by using ggplot package in \Rlogo.
+The related result is presented in Figure \ref{pan}, where  panel \eqref{p.a} represents the random testing, and panel \eqref{p.b} is representing non-random testing.
+The parameter values or ranges are as follows (I like using a table here, with the param, their units and values).
 It is notable that in the simulation part, the explicit expression of $\Rnum$ was used and not the approximation \eqref{eq:R0appr}.

### 2020-11-26 SIR_manuscript.tex
Some edits, thoughts and comments. Much more to do.

@@
-\documentclass{article}
+\documentclass[12pt]{article}
 \usepackage{natbib}
@@
 
-One of the main challenges in interpreting COVID-19 dynamics is to unravel/investigate information gained from data.
+One of the main challenges in understanding COVID-19 dynamics is to unravel information available from data.
 The daily case count is a highly accessible form of data where observations are driven by two clusters of processes: (i) epidemic processes and (ii) testing processes.
@@
 We developed a model that incorporates mechanistic processes of epidemic processes and testing.
-We used the model to study the potential effect of testing strategies and slow-fast reporting testing result.
+We used the model to study the potential effect of testing strategies on disease dynamics.
 
+[[JD: There is a gap here between unravelling data and what we actually do.]]
+
+[[JD: This \P\ is confusing. The topic sentence seems to be about kinds of tests, and then there is a jump to sampling strategies.]]
 There are different testing strategies with different test-specific efficacy, the time required for the results to be reported and the cost of the test.
 Testing strategies can be grouped into two broad categories including (i) random testing across the whole population and (ii) Non-random testing.
-Note that contact tracing is an example of non-random testing strategies.
+[[JD: Random testing is not a broad category, and non-random testing is very vague.
+What are we trying to convey with this \P?]] Note that contact tracing is an example of non-random testing strategies.
 Each of these strategies has been used as a tool to study COVID-19; see (refs) for group (i) and (refs) for group (ii).
@@
 Much effort has been invested in developing widely accessible and fast tests in order to mitigate the spread of the pandemic (refs).
-In order to have a sense and as a comparison of slow versus fast test, we provide an example.
-The viral PCR test, which is the most commonly used test and takes about four days, is very slow relative to the potential tests such as breath-based test \citep{ruszkiewicz2020diagnosis}, which is still in trial stage and takes a few seconds.
-The point to be made is that there can be a potential advantage of slow test.
+For example, PCR tests are commonly used, and very sensitive, but take longer compared to more recent tests, such as a breath-based test currently in trial stage, which can provide a result in seconds.
 
-In this work we developed a model to incorporate epidemic processes and testing processes mechanistically to study the potential effect of testing strategies and slow-fast testing.
-The analysis of the deterministic model included deriving the basic reproduction number, $\Rnum$, as the index of threshold dynamics and the effect of underlying parameters on this key index.
-We found that the non-random testing strategy is more effective than the random testing.
-Specifically, the required critical testing intensity to keep $\Rnum$ lower than the threshold value of one is much lower in the non-random testing than the random one.
-Further, we show that for certain parameter combinations, there may be a potential advantage to slow reporting of test results on the spread of the infection.
+Here, we develop a model to incorporate epidemic processes and testing processes mechanistically to study the potential effect of testing strategies and test speed on epidemic control.
+We investigate the effect of underlying parameters on the basic reproduction number, $\Rnum$.
+While focused testing strategies are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is viral spread is expected to be stronger for a slow test than for a fast test.
+This counter-intuitive effect can occur when people are cautious when awaiting a test result, and may not be robust to second-order effects of fast testing (such as better contact tracing).
 
 % %%%%%%%
-\section{Method}
+\section{Methods}
 
-We developed a deterministic model as a platform of our study.
-A generic individual is identified in a model compartment based on the disease-specific and test-specific criteria.
-The former criteria include Susceptible, Infectious and Recovered, thus SIR, and the latter criteria group people in either untested, awaiting for the test result with expectation of positive or negative outcome, or reported positive, thus $u$, $p$, $n$ and $t$, respectively, reflected as subscript.
-Also, two 'integrator' or 'accumulator' compartments, $N$ and $P$, were incorporated in the model in order to collect cumulative reported negative or positive tests.
+We developed a deterministic model which groups individuals based on disease status and testing status.
+The former includes Susceptible, Infectious and Recovered, thus SIR, and the latter categorizes people as \emph untested, waiting-for-\emph positive, waiting-for-\emph negative, or reported posi{}\emph tive [[JD: would it be better to use $c$ confirmed?]].
+Also, two 'accumulator' compartments, $N$ and $P$, were incorporated in the model in order to collect cumulative reported negative or positive tests.
 The flowchart model is as follows
@@
 
+[[JD: It would be nice to modularize the model somehow, but probably not worth anyone's time.
+It would also be nice to put the model into Supp.
+Do we need these equations in the main text for any reason?]]
 The model is
@@
 Thus, the weighted testing rate is defined as $F_Z=\sigma W_Z$, where $Z \in \{S,I,R\}$.
-Note that the particular form of $\sigma$ \eqref{sigma} in the case when no one is left untested, $W \approx 0$, leads to a singularity issue in the simulation.
-We refer the reader to the Appendix for further details and the solution.
+Note that requiring a fixed testing rate as in \eqref{sigma} can lead to a singularity when no one is left untested.
+We discuss this issue, and a practical solution, in the Appendix.
 
-The classical SIR model is based on the following implicite assumptions (see a ref for details); well-mixed population, homogeneity of the population, exponentially distributed duration of infection and large population size (may be refer the reader to \cite{keeling2011modeling}).
+The classical SIR model is based on the following implicit assumptions (see a ref for details); well-mixed population, homogeneity of the population, exponentially distributed duration of infection and large population size (may be refer the reader to \cite{keeling2011modeling}).
 In addition to these standard assumptions, our model \eqref{model} assumes: (i) there is a single force of infection, $\Lambda$, that is, where a susceptible individual moves to does not depend on who infected it, (ii) a perfectly specific test, $p_S=0$, for simplicity in analysis, (iii)  $\eta_t \leq \eta_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
@@
 say it or not?
+[[JD: The way perfect specificity is presented is confusing here.
+We need to distinguish between the model we first develop (12 boxes) and the model we analyze (10 boxes)]]
 
-The Disease-Free Equilibrium (Ali, not sure if this is the right term?), DFE hereafter, for the SIR model \ref{model} is given by solving the coupled system including $S_u+S_n=N_0$ and $F_S S_u-\omega S_n=0$.
-The DFE is
+The Disease-Free Equilibrium (DFE) for the SIR model \ref{model} is given by solving the coupled system including $S_u+S_n=N_0$ and $F_S S_u-\omega S_n=0$. The DFE is
 
@@
 The analytical calculation of the next generation matrix, $G$, and the derivation of $\Rnum$ was carried out in Maple (ref) by using simple linear Algebra package.
-  We used \Rlogo for the simulation part which included using the explicite expression of $\Rnum$ \ref{R0} as a function of the underlying parameters, specifying the parameters a realistic range, calculating the corresponding $\Rnum$ and plotting the contours of $\Rnum$ by using ggplot package in \Rlogo.
+  We used \Rlogo for the simulation part which included using the explicit expression of $\Rnum$ \ref{R0} as a function of the underlying parameters, specifying the parameters a realistic range, calculating the corresponding $\Rnum$ and plotting the contours of $\Rnum$ by using ggplot package in \Rlogo.
 The related result is presented in Figure \ref{pan}, where  panel \eqref{p.a} represents the random testing, and panel \eqref{p.b} is representing non-random testing.

### 2020-11-26 SIR_manuscript.tex
minor fixes

@@
 
-We developed a deterministic model which groups individuals based on disease status and testing status.
-The former includes Susceptible, Infectious and Recovered, thus SIR, and the latter categorizes people as \emph untested, waiting-for-\emph positive, waiting-for-\emph negative, or reported posi{}\emph tive [[JD: would it be better to use $c$ confirmed?]].
+We developed a deterministic model \eqref{model} which groups individuals based on disease status and testing status.
+The former includes Susceptible, Infectious and Recovered, thus SIR, and the latter categorizes people as \emph untested, waiting-for-\emph positive, waiting-for-\emph negative, or reported posi{}\emph tive [[JD: would it be better to use $c$ confirmed?
+Ali: yes]].
 Also, two 'accumulator' compartments, $N$ and $P$, were incorporated in the model in order to collect cumulative reported negative or positive tests.
+The model \ref{model} and details of calculation of $\Rnum$ are presented in the Supplimentary Materials.
 The flowchart model is as follows
@@
 \includegraphics[scale=0.3]{./pix/sir_comp.png}
-\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model. [Fady] Say more about the compartments and parameters here to make it stand on its own.
+\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model \eqref{model}.
+[Fady] Say more about the compartments and parameters here to make it stand on its own.
+The equations of the model are in Supplimentary Materials.
 \label{fig:flowchart}}
@@
 
-[[JD: It would be nice to modularize the model somehow, but probably not worth anyone's time.
-It would also be nice to put the model into Supp.
-Do we need these equations in the main text for any reason?]]
-The model is
-\begin{align}
-\label{model}
- d S_u/dt &= -\Lambda S_u - F_S S_u + \omega S_n, \\
- d S_n/dt &= -\Lambda S_n + (1-p_S) F_S S_u - \omega S_n, \\
- d S_p/dt &= -\Lambda S_p + p_S F_S S_u - \omega S_p, \\
- d S_t/dt &= -\Lambda S_t + \omega S_p, \\
- d I_u/dt &= \Lambda S_u - F_I I_u + \omega I_n  - \gamma I_u,  \\
- d I_n/dt &= \Lambda S_n + (1-p_I) F_I I_u - \omega I_n -\gamma I_n, \\
- d I_p/dt &= \Lambda S_p + p_I F_I I_u - \omega I_p -\gamma I_p, \\
- d I_t/dt &= \Lambda S_t + \omega I_p - \gamma I_t,  \\
- d R_u/dt &= \gamma I_u - F_R R_u + \omega R_n, \\
- d R_n/dt &= \gamma I_n + (1-p_R) F_R R_u - \omega R_n,  \\
- d R_p/dt &= \gamma I_p + p_R F_R R_u  - \omega R_p,  \\
- d R_t/dt&= \gamma I_t + \omega R_p,  \\
- dN/dt &= \omega (S_n + I_n + R_n),   \\
- dP/dt &= \omega(I_p + R_p),
-\end{align}
+[[JD: It would be nice to modularize the model somehow, but probably not worth anyone's time.]]
 
-where $\beta$, (in units of 1/day), is the disease transmission rate, $\eta_w$ and $\eta_t$ are the isolation parameters for awaiting and reported individuals, respectively.
-$\Lambda$ is the force of infection and is defined as
-
-\begin{equation}
-\label{Lambda}
-\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_t I_t)}{N_0},
-\end{equation}
-
-where $N_0$ is the total population size.
-Further, $\omega$, (1/day), is the rate of onward flow from the awaiting positive or negative compartments to reported or untested compartments, respectively, $\gamma$, (1/day), is the recovery rate, $\rho$, (1/day), is the per capita testing intensity across the whole population.
+[Ali: not sure if these should go to the Supp material?] In our model the following parameters are used.
+$N_0$ is the total population size, $\omega$, (1/day), is the rate of onward flow from the awaiting positive or negative compartments to reported or untested compartments, respectively, $\gamma$, (1/day), is the recovery rate, $\rho$, (1/day), is the per capita testing intensity across the whole population.
 Further, various testing strategies are incorporated through the compartment-specific relative testing weights $W_S$, $W_I$ and $W_R$.
@@
 \end{equation}
-where
-\begin{align*}
-A=& \gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_t p_I) F_I, \\
-B=& \big(\omega+(F_I+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_t\omega) \omega p_I F_I }{\omega+\gamma}, \\ 
-C=& \frac{\beta/\gamma}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}.
-\end{align*}
 Further details are provided in the Appendix.
@@
 \ref{p1:omega} is the most surprising.
-It states that returning test results more rapidly (i.e., increasing $\omega$) does not necessarily lower $\Rnum$, rather, whether increasing $\omega$ lowers $\Rnum$ is dependant on the precise combination of model parameters (test reporting rate itself, testing processes/strategies ($W$'s), testing efficacy ($P_I$) and isolation ($\eta$'s).
-Specifically, the $\partial{R_0}/\partial{\omega}>0$ suggests that reproduction number may increase as the test reporting proccess become faster.
+It states that returning test results more rapidly (i.e., increasing $\omega$) does not necessarily lower $\Rnum$, rather, whether increasing $\omega$ lowers $\Rnum$ is defendant on the precise combination of model parameters (test reporting rate itself, testing processes/strategies ($W$'s), testing efficacy ($P_I$) and isolation ($\eta$'s).
+Specifically, the $\partial{R_0}/\partial{\omega}>0$ suggests that reproduction number may increase as the test reporting process become faster.
 This can be seen from expression \eqref{Rom} in the case of ``perfect isolation``, when $\eta_w=0$ and $\eta_t=0$.
-The underlying mechanism for this result is that individuals may take more precautionary behaviours (e.g., physical distancing) when they are awaiting test results compared to when they are untested or their test has returned negative.
+The underlying mechanism for this result is that individuals may take more precautionary behaviors (e.g., physical distancing) when they are awaiting test results compared to when they are untested or their test has returned negative.
 An example of a situation that would favour slow return of tests is if the test being employed produces many false negatives, because many infected individuals will believe they are negative and thus may unknowingly spread the virus to many others.
@@
 % %%%%%%%
-\section{supplementary material}
+\section{Supplementary Material}
 
-\subsection{Calculation of $\Rnum$}
+\subsection{Model and Calculation of $\Rnum$}
+
+The model is 
+\begin{align}
+\label{model}
+ d S_u/dt &= -\Lambda S_u - F_S S_u + \omega S_n, \\
+ d S_n/dt &= -\Lambda S_n + (1-p_S) F_S S_u - \omega S_n, \\
+ d S_p/dt &= -\Lambda S_p + p_S F_S S_u - \omega S_p, \\
+ d S_t/dt &= -\Lambda S_t + \omega S_p, \\
+ d I_u/dt &= \Lambda S_u - F_I I_u + \omega I_n  - \gamma I_u,  \\
+ d I_n/dt &= \Lambda S_n + (1-p_I) F_I I_u - \omega I_n -\gamma I_n, \\
+ d I_p/dt &= \Lambda S_p + p_I F_I I_u - \omega I_p -\gamma I_p, \\
+ d I_t/dt &= \Lambda S_t + \omega I_p - \gamma I_t,  \\
+ d R_u/dt &= \gamma I_u - F_R R_u + \omega R_n, \\
+ d R_n/dt &= \gamma I_n + (1-p_R) F_R R_u - \omega R_n,  \\
+ d R_p/dt &= \gamma I_p + p_R F_R R_u  - \omega R_p,  \\
+ d R_t/dt&= \gamma I_t + \omega R_p,  \\
+ dN/dt &= \omega (S_n + I_n + R_n),   \\
+ dP/dt &= \omega(I_p + R_p),
+\end{align}
+
+where $\beta$, (in units of 1/day), is the disease transmission rate, $\eta_w$ and $\eta_t$ are the isolation parameters for awaiting and reported individuals, respectively.
+$\Lambda$ is the force of infection and is defined as
+
+\begin{equation}
+\label{Lambda}
+\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_t I_t)}{N_0},
+\end{equation}
+
+% %
 
@@
 0&0&-\omega&\gamma
+\end {array} \right], \text{thus}\\
+V^{-1} =&
+\left[ \begin {array}{cccc}
+\frac {\omega+\gamma}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&\frac {\omega}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&0&0\\
+\noalign{\medskip}
+\frac{(1-p_I) F_I}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&
+\frac{F_I+\gamma}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&0&0\\
+\noalign{\medskip}
+\frac{p_I F_I}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&
+\frac{\omega p_I F_I}{(\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I)(\omega+\gamma)}& \frac{1}{\omega+\gamma}&0 \\
+\noalign{\medskip}
+\frac{\omega\,F_I\,p_I}{(\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I) \gamma}& 
+\frac{\omega^2\,F_I\,p_I}{(\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I)(\omega+\gamma) \gamma}&
+\frac{\omega}{(\omega+\gamma) \gamma}&
+\frac{1}{\gamma}
 \end {array} \right].
@@
 
-Note that the block matrix $G_{12}$ does not influence $\Rnum$ defined as the spectral radius of $G$. All matters here are the eigenvalues of $G_{11}$, which are 0 and $\Rnum$ \eqref{R0}.
+Note that the block matrix $G_{12}$ does not influence $\Rnum$ defined as the spectral radius of $G$. All matters here are the eigenvalues of $G_{11}$, which are 0 and $\Rnum$ \eqref{R0}, 
+where
+\begin{align*}
+A=& \gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_t p_I) F_I, \\
+B=& \big(\omega+(F_I+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_t\omega) \omega p_I F_I }{\omega+\gamma}, \\ 
+C=& \frac{\beta/\gamma}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}.
+\end{align*}
 

### 2020-11-30 SIR_manuscript.tex
updated the tex, code, pix

@@
 \usepackage[nameinlink,capitalize]{cleveref}
+\usepackage{cleveref}
+\usepackage{geometry}
+\usepackage{pdflscape}
 
@@
 There are different testing strategies with different test-specific efficacy, the time required for the results to be reported and the cost of the test.
-Testing strategies can be grouped into two broad categories including (i) random testing across the whole population and (ii) Non-random testing.
+Testing strategies can be grouped into two categories including (i) random testing across the whole population and (ii) Non-random testing.
 [[JD: Random testing is not a broad category, and non-random testing is very vague.
-What are we trying to convey with this \P?]] Note that contact tracing is an example of non-random testing strategies.
+What are we trying to convey with this \P?
+Ali: I am thinking of the weighted testing parameters W's in our model, that reflects testing strategies.(not sure the best replacement for "non-random" testing?)]] Note that contact tracing is an example of non-random testing strategies.
 Each of these strategies has been used as a tool to study COVID-19; see (refs) for group (i) and (refs) for group (ii).
@@
 
-We developed a deterministic model \eqref{model} which groups individuals based on disease status and testing status.
-The former includes Susceptible, Infectious and Recovered, thus SIR, and the latter categorizes people as \emph untested, waiting-for-\emph positive, waiting-for-\emph negative, or reported posi{}\emph tive [[JD: would it be better to use $c$ confirmed?
-Ali: yes]].
+We developed a deterministic model, \crefrange{eq1}{eq12}, which groups individuals based on disease status and testing status.
+The former includes Susceptible, Infectious and Recovered, thus SIR, and the latter categorizes people as \emph untested, waiting-for-\emph positive, waiting-for-\emph negative, or \emph confirmed positive.
 Also, two 'accumulator' compartments, $N$ and $P$, were incorporated in the model in order to collect cumulative reported negative or positive tests.
-The model \ref{model} and details of calculation of $\Rnum$ are presented in the Supplimentary Materials.
+The model and details of calculation of $\Rnum$ are presented in the Supplimentary Materials.
 The flowchart model is as follows
@@
 \includegraphics[scale=0.3]{./pix/sir_comp.png}
-\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model \eqref{model}.
-[Fady] Say more about the compartments and parameters here to make it stand on its own.
-The equations of the model are in Supplimentary Materials.
+\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model, \crefrange{eq1}{eq12}. For the description of the parameters see \cref{tab:params}.
 \label{fig:flowchart}}
@@
 
-[Ali: not sure if these should go to the Supp material?] In our model the following parameters are used.
+[Ali: not sure if these should go to the Supp material?] In our model the following parameters are used (for a summery of parameters' description, see Table \ref{tab:params}).
 $N_0$ is the total population size, $\omega$, (1/day), is the rate of onward flow from the awaiting positive or negative compartments to reported or untested compartments, respectively, $\gamma$, (1/day), is the recovery rate, $\rho$, (1/day), is the per capita testing intensity across the whole population.
@@
 The classical SIR model is based on the following implicit assumptions (see a ref for details); well-mixed population, homogeneity of the population, exponentially distributed duration of infection and large population size (may be refer the reader to \cite{keeling2011modeling}).
-In addition to these standard assumptions, our model \eqref{model} assumes: (i) there is a single force of infection, $\Lambda$, that is, where a susceptible individual moves to does not depend on who infected it, (ii) a perfectly specific test, $p_S=0$, for simplicity in analysis, (iii)  $\eta_t \leq \eta_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
+In addition to these standard assumptions, our model, \crefrange{eq1}{eq12}, assumes: (i) there is a single force of infection, $\Lambda$, that is, where a susceptible individual moves to does not depend on who infected it, (ii) a perfectly specific test, $p_S=0$, for simplicity in analysis, (iii)  $\eta_c \leq \eta_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
 - [Q] why not birth and death are included?
@@
 We need to distinguish between the model we first develop (12 boxes) and the model we analyze (10 boxes)]]
+[[Ali: If we use the full model, with 12 compartments including $S_p$ and $S_c$, the DFE will be; $S_n=S_p=0$, $S_u+S_c=N_0$ (it can be assumed that $S_u=N_0$ and $S_c=0$) and all $I$ and $R$ compartments to be zero.
+This simplifies the expression of $\Rnum$ since $F_I=\rho$.]]
 
-The Disease-Free Equilibrium (DFE) for the SIR model \ref{model} is given by solving the coupled system including $S_u+S_n=N_0$ and $F_S S_u-\omega S_n=0$. The DFE is
+\newgeometry{margin=1cm} % modify this if you need even more space
+\begin{landscape}
+\begin{table}[htp]
+\centering
+{\tiny %
+\begin{tabular}{|c|ccc|} \hline
+  Symbol & Description & Unit & Value \\ \hline
+  $N_0$     & Total population size & people & $10^6$ \\ \hline
+  $\omega$  & Rate of onward flow from the awaiting to reported or untested compartments  & 1/day & - \\ \hline
+  $\gamma$ & Recovery rate & 1/day & 1/3 \\ \hline 
+  $\rho$   & Per capita testing intensity & 1/day & 0.01 \\ \hline 
+  $\eta_w$  & Relative probability of transmission for isolated awaiting individuals & - & - \\ \hline
+  $\eta_c$  & Relative probability of transmission for isolated confirmed individuals & - & -  \\ \hline
+  $\Lambda$ & Force of infection & 1/time & - \\ \hline
+  $p_S$ & Probability of false positive for susciptibles& - & 0 \\ \hline
+  $p_I$ & Probability of being infected and tested positive & - & 1 \\ \hline
+  $p_R$ & Probability of being recovered and tested positive & - & 0.5 \\ \hline
+  $W_S, W_I, W_R$ & Relative testing weight & - & 
+  \begin{minipage}[t]{0.21\columnwidth}%
+ Random testing: $W_S=W_I=W_R=1$, \\Non-random testing: $W_S=0.3, W_I=W_R=1$
+\end{minipage} \\ \hline
+
+  \end{tabular}
+  }%
+\caption{\label{tab:params} The underlying parameters of model, \crefrange{eq1}{eq12}.}
+\end{table}
+\end{landscape}
+\restoregeometry
+
+The Disease-Free Equilibrium (DFE) for the SIR model, \crefrange{eq1}{eq12}, is given by solving the coupled system including $S_u+S_n=N_0$ and $F_S S_u-\omega S_n=0$. The DFE is
 
@@
 The explicit formula for the basic reproduction number, $\Rnum$ \eqref{R0}, provides an opportunity to study the influence of changes in the underlying parameters on the critical index of epidemic dynamics.
-We are interested in understanding the effect of parameters that can be realistically controlled by changing in isolation, testing intensity and test resulting, i.e., $\eta_t$ and $\eta_w$, $\rho$ and $\omega$, respectively.
+We are interested in understanding the effect of parameters that can be realistically controlled by changing in isolation, testing intensity and test resulting, i.e., $\eta_c$ and $\eta_w$, $\rho$ and $\omega$, respectively.
 The following Proposition is the direct result of taking partial derivative of $\Rnum$ \eqref{R0} with respect to selected parameters.
@@
 \item \label{p1:eta}
-$\partial{R_0}/\partial{\eta_t} \geq 0$ and $\partial{R_0}/\partial{\eta_w} \geq 0$. 
+$\partial{R_0}/\partial{\eta_c} \geq 0$ and $\partial{R_0}/\partial{\eta_w} \geq 0$. 
 \item \label{p1:rho}
@@
 
- Given that the perfect isolation occurs at $\eta_t=\eta_w = 0$, Prop.
+ Given that the perfect isolation occurs at $\eta_c=\eta_w = 0$, Prop.
 \ref{p1:eta} means that lifting the isolation from awaiting group results in an increased $\Rnum$ and consequently a greater number of infected individuals.
@@
 Specifically, the $\partial{R_0}/\partial{\omega}>0$ suggests that reproduction number may increase as the test reporting process become faster.
-This can be seen from expression \eqref{Rom} in the case of ``perfect isolation``, when $\eta_w=0$ and $\eta_t=0$.
+This can be seen from expression \eqref{Rom} in the case of ``perfect isolation``, when $\eta_w=0$ and $\eta_c=0$.
 The underlying mechanism for this result is that individuals may take more precautionary behaviors (e.g., physical distancing) when they are awaiting test results compared to when they are untested or their test has returned negative.
@@
 \begin{itemize}
-    \item individuals who have tested positive lower their contact much more than individuals who are waiting for their test results (i.e., $\eta_t \ll \eta_w$).
+    \item individuals who have tested positive lower their contact much more than individuals who are waiting for their test results (i.e., $\eta_c \ll \eta_w$).
     \item the test is sensitive.
@@
 \begin{align}
-\label{model}
- d S_u/dt &= -\Lambda S_u - F_S S_u + \omega S_n, \\
- d S_n/dt &= -\Lambda S_n + (1-p_S) F_S S_u - \omega S_n, \\
- d S_p/dt &= -\Lambda S_p + p_S F_S S_u - \omega S_p, \\
- d S_t/dt &= -\Lambda S_t + \omega S_p, \\
- d I_u/dt &= \Lambda S_u - F_I I_u + \omega I_n  - \gamma I_u,  \\
- d I_n/dt &= \Lambda S_n + (1-p_I) F_I I_u - \omega I_n -\gamma I_n, \\
- d I_p/dt &= \Lambda S_p + p_I F_I I_u - \omega I_p -\gamma I_p, \\
- d I_t/dt &= \Lambda S_t + \omega I_p - \gamma I_t,  \\
- d R_u/dt &= \gamma I_u - F_R R_u + \omega R_n, \\
- d R_n/dt &= \gamma I_n + (1-p_R) F_R R_u - \omega R_n,  \\
- d R_p/dt &= \gamma I_p + p_R F_R R_u  - \omega R_p,  \\
- d R_t/dt&= \gamma I_t + \omega R_p,  \\
- dN/dt &= \omega (S_n + I_n + R_n),   \\
- dP/dt &= \omega(I_p + R_p),
+% \label{model}
+ d S_u/dt &= -\Lambda S_u - F_S S_u + \omega S_n, \label{eq1}\\
+ d S_n/dt &= -\Lambda S_n + (1-p_S) F_S S_u - \omega S_n, \label{eq2}\\
+ d S_p/dt &= -\Lambda S_p + p_S F_S S_u - \omega S_p, \label{eq3}\\
+ d S_c/dt &= -\Lambda S_c + \omega S_p, \label{eq4}\\
+ d I_u/dt &= \Lambda S_u - F_I I_u + \omega I_n  - \gamma I_u,  \label{eq5}\\
+ d I_n/dt &= \Lambda S_n + (1-p_I) F_I I_u - \omega I_n -\gamma I_n, \label{eq6}\\
+ d I_p/dt &= \Lambda S_p + p_I F_I I_u - \omega I_p -\gamma I_p, \label{eq7}\\
+ d I_c/dt &= \Lambda S_c + \omega I_p - \gamma I_c,  \label{eq8}\\
+ d R_u/dt &= \gamma I_u - F_R R_u + \omega R_n, \label{eq9}\\
+ d R_n/dt &= \gamma I_n + (1-p_R) F_R R_u - \omega R_n,  \label{eq10}\\
+ d R_p/dt &= \gamma I_p + p_R F_R R_u  - \omega R_p,  \label{eq11}\\
+ d R_c/dt&= \gamma I_c + \omega R_p,  \label{eq12}\\
+ dN/dt &= \omega (S_n + I_n + R_n),   \label{eq13}\\
+ dP/dt &= \omega(I_p + R_p) \label{eq14},
 \end{align}
 
-where $\beta$, (in units of 1/day), is the disease transmission rate, $\eta_w$ and $\eta_t$ are the isolation parameters for awaiting and reported individuals, respectively.
+where $\beta$, (in units of 1/day), is the disease transmission rate, $\eta_w$ and $\eta_c$ are the isolation parameters for awaiting and reported individuals, respectively.
 $\Lambda$ is the force of infection and is defined as
@@
 \label{Lambda}
-\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_t I_t)}{N_0},
+\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_c I_c)}{N_0},
 \end{equation}
@@
 The next generation matrix is $G = F V^{-1}$, where matrix $F$ represents the inflow of new infection to the infected compartments and matrix $V$ represents the flow in the infected compartments when the population is totally susiptable.
-For model \eqref{model}, $F$ and $V$ are
+For model, \crefrange{eq1}{eq12}, $F$ and $V$ are
 
@@
 F =& \beta/N_0 \left[ \begin {array}{cccc} 
-S_u&\eta_w\,S_u&\eta_w\,S_u&\eta_t\,S_u\\
-S_n&\eta_w\,S_n&\eta_w\,S_n&\eta_t\,S_n\\ 
+S_u&\eta_w\,S_u&\eta_w\,S_u&\eta_c\,S_u\\
+S_n&\eta_w\,S_n&\eta_w\,S_n&\eta_c\,S_n\\ 
 0&0&0&0\\
@@
 \begin{align*}
-A=& \gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_t p_I) F_I, \\
-B=& \big(\omega+(F_I+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_t\omega) \omega p_I F_I }{\omega+\gamma}, \\ 
+A=& \gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_c p_I) F_I, \\
+B=& \big(\omega+(F_I+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_c\omega) \omega p_I F_I }{\omega+\gamma}, \\ 
 C=& \frac{\beta/\gamma}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}.
@@
 \label{eq:R0appr}
-R_0 \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_S} \Big(\gamma(\eta_w-1)(\gamma W_S+\omega W_I) + (\eta_t -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2).
+R_0 \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_S} \Big(\gamma(\eta_w-1)(\gamma W_S+\omega W_I) + (\eta_c -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2).
 \end{equation}
@@
 
-where $a=(\eta_w-1)W_I-(\eta_t-1)P_I W_I = ((s-p_I)\eta_t + (p_I-1)) W_I$, $b=2(\eta_w-1)\gamma W_S$ and $c=(\eta_w-1)\gamma^2 W_S$.
-Given that $0 \leq \eta_t\leq \eta_w \leq 1 $, one can easily derive $b\leq 0$ and $c \leq 0$.
+where $a=(\eta_w-1)W_I-(\eta_c-1)P_I W_I = ((s-p_I)\eta_c + (p_I-1)) W_I$, $b=2(\eta_w-1)\gamma W_S$ and $c=(\eta_w-1)\gamma^2 W_S$.
+Given that $0 \leq \eta_c\leq \eta_w \leq 1 $, one can easily derive $b\leq 0$ and $c \leq 0$.
 
-Note that in general, the necessary and sufficient condition for $a \geq 0$ is $(s-p_I) \eta_t \geq (1-p_I)$, where $s=\frac{\eta_w}{\eta_t} \geq 1$.
+Note that in general, the necessary and sufficient condition for $a \geq 0$ is $(s-p_I) \eta_c \geq (1-p_I)$, where $s=\frac{\eta_w}{\eta_c} \geq 1$.
 
-Note that in case of ``perfect isolation``, i.e., when $\eta_w=0$ and consequently $\eta_t=0$, it is straight forward to see that $a \leq 0$, $b<0$ and $c<0$.
+Note that in case of ``perfect isolation``, i.e., when $\eta_w=0$ and consequently $\eta_c=0$, it is straight forward to see that $a \leq 0$, $b<0$ and $c<0$.
 Thus, $\partial{\Rnum}/\partial{\omega} \geq 0$.
@@
 \begin{equation}\label{linearization}
-\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta_w-1)(\gamma W_s+\omega W_i) + (\eta_t -1)P_iW_i \omega^2 \Big). 
+\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta_w-1)(\gamma W_s+\omega W_i) + (\eta_c -1)P_iW_i \omega^2 \Big). 
 \end{equation}
@@
 
-where $a=(\eta_w-1)W_i-(\eta_t-1)P_iW_i$, $b=2(\eta_w-1)\gamma W_s$ and $c=(\eta_w-1)\gamma^2 W_s$.
+where $a=(\eta_w-1)W_i-(\eta_c-1)P_iW_i$, $b=2(\eta_w-1)\gamma W_s$ and $c=(\eta_w-1)\gamma^2 W_s$.
 
@@
 Note that the zeroes are real if and only if $Q < 1$.
-Note that have $\eta_t < \eta_w$, so if $P_i \approx 1$, we will have $Q < 0 < 1$.
+Note that have $\eta_c < \eta_w$, so if $P_i \approx 1$, we will have $Q < 0 < 1$.
 Thus, if we assume near-perfect test sensitivity, $\omega_1$ and $\omega_2$ will be real.
@@
 Suppose that $\rho_1 > \rho_2$.
-Recall that the linearization of $\Rnum$ around $\rho \approx 0$ is given by $$\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta_w-1)(\gamma W_s+\omega W_i) + (\eta_t -1)P_iW_i \omega^2 \Big).$$
+Recall that the linearization of $\Rnum$ around $\rho \approx 0$ is given by $$\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta_w-1)(\gamma W_s+\omega W_i) + (\eta_c -1)P_iW_i \omega^2 \Big).$$
 
@@
 \begin{align}\label{eq:rho1vsrho2}
-    &\rho_1\left(\gamma(\eta_w-1)(\gamma W_s + \omega W_i) + (\eta_t-1)P_{i, 1}W_i\omega^2\right) - \rho_2\left(\gamma(\eta_w-1)(\gamma W_s + \omega W_i) + (\eta_t-1)P_{i, 2}W_i\omega^2\right) > 0 \nonumber \\
+    &\rho_1\left(\gamma(\eta_w-1)(\gamma W_s + \omega W_i) + (\eta_c-1)P_{i, 1}W_i\omega^2\right) - \rho_2\left(\gamma(\eta_w-1)(\gamma W_s + \omega W_i) + (\eta_c-1)P_{i, 2}W_i\omega^2\right) > 0 \nonumber \\
     &\vdots \nonumber \\
-    &\frac{\rho_2P_{i, 2}-\rho_1P_{i, 1} }{\rho_1-\rho_2} > \frac{1-\eta_w}{1-\eta_t}\cdot \frac{\gamma(\gamma W_s + \omega W_i)}{\omega^2 W_i}
+    &\frac{\rho_2P_{i, 2}-\rho_1P_{i, 1} }{\rho_1-\rho_2} > \frac{1-\eta_w}{1-\eta_c}\cdot \frac{\gamma(\gamma W_s + \omega W_i)}{\omega^2 W_i}
 \end{align}

### 2020-11-30 codes/ali_test_sir.Rmd
updated the tex, code, pix

@@
 
-which(df2[]<0)
+df2_fake <- df2 %>% dplyr::select(eta_t,eta_w)
+df2_fake <- unique(df2_fake) %>% dplyr::mutate(flag=ifelse(eta_t<=eta_w,1,0))
 ```
@@
     + ggtitle(TeX(sprintf("W_S=%.1f, W_I=W_R=1",W_S)))
+    + geom_rect(data=df2,
+                          fill=ifelse(df2$eta_t>df2$eta_w,"white","NA" ),
+                          color= NA,
+                          ymin=-1,
+                          ymax=10,
+                          xmin=-1,
+                          xmax=2)
     )
 p1_temp
+
+# ## hack (By Dr. Ben Bolker) white rectangle to cover grid
+#   ggplot() + 
+#   scale_x_continuous(name="x") + 
+#   scale_y_continuous(name="y") +
+#   facet_grid(eta_w~eta_t, labeller=label_special) +
+#     geom_rect(data=df2_fake,
+#                           fill=ifelse(df2_fake$eta_t>df2_fake$eta_w,"white","black"),
+#                           color= NA,
+#                           ymin=-1,
+#                           ymax=10,
+#                           xmin=-1,
+#                           xmax=2)
+
 # ggsave(p1_temp,filename = "R0contour_random.png" , path="../pix/")
-ggsave(p1_temp,filename = "R0contour_TTI.png" , path="../pix/")
+# ggsave(p1_temp,filename = "R0contour_TTI.png" , path="../pix/")
 

### 2020-12-01 codes/ali_test_sir.Rmd
panel plot updated

--- 2020-11-16 codes/ali_test_sir.Rmd
@@
 
-- $\Lambda$: force of infection defined as $$\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_t I_t)}{N_0},$$ where $\beta$ is transmission rate, $\eta_w$ and $\eta_t$ are the isolation parameters for awaiting and reported individuals, respectively.
-We assume that $\eta_t<\eta_w$, i.e., the awaiting individuals for test results have a higher transmission probability than the reported individuals.
+- $\Lambda$: force of infection defined as $$\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_c I_t)}{N_0},$$ where $\beta$ is transmission rate, $\eta_w$ and $\eta_c$ are the isolation parameters for awaiting and reported individuals, respectively.
+We assume that $\eta_c<\eta_w$, i.e., the awaiting individuals for test results have a higher transmission probability than the reported individuals.
 - $\omega$: the rate of *onward flow* from the awaiting positive compartment, $p$, to reported/tested compartment, $\rho$, or from awaiting negative compartment, $n$, back to $u$.
@@
     ## Force of Infection  
-    Lambda <- beta * (I_u + eta_w*(I_n+I_p) + eta_t*I_t)/N0
+    Lambda <- beta * (I_u + eta_w*(I_n+I_p) + eta_c*I_t)/N0
     ## scaling the weights
@@
              p_S=0, p_I=1, p_R=0.5, #prob of waiting for being tested positive
-            eta_w=0.02, eta_t=0.01, #isolation parameter 
-            s=2 #eta_w/eta_t
+            eta_w=0.02, eta_c=0.01, #isolation parameter 
+            s=2 #eta_w/eta_c
             )
@@
 $$R_0= (A \times S_u^* + B \times S_n^*) \times C,$$ where
-$A=\gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_t p_I) F_I$,
-$B=\big(\omega+(F_I+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_t\omega) \omega p_I F_I }{\omega+\gamma}$ and 
+$A=\gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_c p_I) F_I$,
+$B=\big(\omega+(F_I+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_c\omega) \omega p_I F_I }{\omega+\gamma}$ and 
 $C=\frac{\beta/\gamma}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}$.
@@
 F= \beta/N_0 \left[ \begin {array}{cccc} 
-S_u&\eta_w\,S_u&\eta_w\,S_u&\eta_t\,S_u\\
-S_n&\eta_w\,S_n&\eta_w\,S_n&\eta_t\,S_n\\ 
+S_u&\eta_w\,S_u&\eta_w\,S_u&\eta_c\,S_u\\
+S_n&\eta_w\,S_n&\eta_w\,S_n&\eta_c\,S_n\\ 
 0&0&0&0\\
@@
     Fi <- F_I(params) #at DFE
-    A <-gamma*(omega+gamma+eta_w*Fi) + omega*eta_t*p_I*Fi
-    B <-(omega+eta_w*(Fi+gamma))*gamma+(eta_w*gamma+eta_t*omega)*(omega*p_I*Fi)/(omega+gamma)
+    A <-gamma*(omega+gamma+eta_w*Fi) + omega*eta_c*p_I*Fi
+    B <-(omega+eta_w*(Fi+gamma))*gamma+(eta_w*gamma+eta_c*omega)*(omega*p_I*Fi)/(omega+gamma)
     C <- beta/(N0*gamma*(gamma*(omega+gamma)+Fi*(gamma+omega*p_I)))
@@
 
-# R0 expression with eta_w/eta_t factored
+# R0 expression with eta_w/eta_c factored
 R02<-function(params){
@@
     C2 <- beta/((N0*gamma*(gamma*(omega+gamma)+F_I*(gamma+omega*p_I)))*(omega+gamma))
-    # s <- eta_w/eta_t
-    return((eta_t*(s*A2+B2)+D)*C2)
+    # s <- eta_w/eta_c
+    return((eta_c*(s*A2+B2)+D)*C2)
 }
@@
 # Case #2: 
-all.equal(R0(params=update(params,eta_t=1,eta_w=1)), params[["beta"]]/params[["gamma"]])
+all.equal(R0(params=update(params,eta_c=1,eta_w=1)), params[["beta"]]/params[["gamma"]])
 ```
@@
  eta_w <- seq(0,1, length.out=n_out)
- eta_t <- seq(0,1, length.out=n_out) #0.001 to 0.5 
+ eta_c <- seq(0,1, length.out=n_out) #0.001 to 0.5 
  rho     <- seq(0,0.01, length.out=n_in)
  omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, was 0.011
- W_S <- 0.3 #choose 0.3 for TTI testing plan and 1 for the random testing.
+ W_S <- 1 #choose 0.3 for TTI testing plan and 1 for the random testing.
 
@@
                     p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
-                    eta_w=eta_w,eta_t=eta_t,s=params[["s"]])
+                    eta_w=eta_w,eta_c=eta_c,s=params[["s"]])
 
@@
 
-df2 <- df2 %>% dplyr::mutate(R0_sub = ifelse(eta_w<eta_t, NA, R0))
+df2 <- df2 %>% dplyr::mutate(R0_sub = ifelse(eta_w<eta_c, NA, R0))
 
-which(df2[]<0)
+# df2_fake <- df2 %>% dplyr::select(eta_c,eta_w)
+# df2_fake <- unique(df2_fake) %>% dplyr::mutate(flag=ifelse(eta_c<=eta_w,1,0))
+
 ```
@@
   ylab(TeX('$\\rho$, testing intensity (1/day per capita)')) +
-    theme(panel.spacing=grid::unit(0,"lines"))
+    # theme(panel.spacing=grid::unit(0,"lines")) # initial theme, comment it for TTI testing plot
+    theme(panel.spacing=grid::unit(0,"lines"),legend.position = "none") #AG, uncomment for Random Case
+# p1_temp <- (p1
+#     + geom_contour_filled(breaks=pretty(unique(df2$R0_sub),5),
+#                           alpha=0.5) ## breaks=0.2)
+#     + geom_contour(breaks=1,alpha=0.5,colour="black")
+#     + facet_grid(eta_w~eta_c, labeller=label_special)
+#     + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
+#     + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
+#     + scale_fill_viridis_d(name=parse(text="R[0]"))
+#     + ggtitle(TeX(sprintf("W_S=%.1f, W_I=W_R=1",W_S)))
+#     + geom_rect(data=df2,
+#                           fill=ifelse(df2$eta_c>df2$eta_w,"white","NA" ),
+#                           color= NA,
+#                           ymin=-1,
+#                           ymax=10,
+#                           xmin=-1,
+#                           xmax=2)
+#     )
+
 p1_temp <- (p1
-    + geom_contour_filled(breaks=pretty(unique(df2$R0_sub),5),
-                          alpha=0.5) ## breaks=0.2)
+    + geom_contour_filled(breaks=seq(0.80,1.05,by=0.05)
+                          )
     + geom_contour(breaks=1,alpha=0.5,colour="black")
-    + facet_grid(eta_w~eta_t, labeller=label_special)
+    + facet_grid(eta_w~eta_c, labeller=label_special)
     + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
@@
     + ggtitle(TeX(sprintf("W_S=%.1f, W_I=W_R=1",W_S)))
+    + geom_rect(data=df2,
+                          fill=ifelse(df2$eta_c>df2$eta_w,"white","NA" ),
+                          color= NA,
+                          ymin=-1,
+                          ymax=10,
+                          xmin=-1,
+                          xmax=2)
     )
 p1_temp
-# ggsave(p1_temp,filename = "R0contour_random.png" , path="../pix/")
-ggsave(p1_temp,filename = "R0contour_TTI.png" , path="../pix/")
+# ## hack (By Dr. Ben Bolker) white rectangle to cover grid
+#   ggplot() + 
+#   scale_x_continuous(name="x") + 
+#   scale_y_continuous(name="y") +
+#   facet_grid(eta_w~eta_c, labeller=label_special) +
+#     geom_rect(data=df2_fake,
+#                           fill=ifelse(df2_fake$eta_c>df2_fake$eta_w,"white","black"),
+#                           color= NA,
+#                           ymin=-1,
+#                           ymax=10,
+#                           xmin=-1,
+#                           xmax=2)
+
+ggsave(p1_temp,filename = "R0contour_random.png" , path="../pix/")
+# ggsave(p1_temp,filename = "R0contour_TTI.png" , path="../pix/")
 
@@
 
-### Contour plot of $R_0=1$ and interpretations approache 2, parametrization $s=\eta_w/\eta_t$
+### Contour plot of $R_0=1$ and interpretations approache 2, parametrization $s=\eta_w/\eta_c$
 
@@
 omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, 
-eta_t <- seq(0.001,0.1, length.out=n_out)
+eta_c <- seq(0.001,0.1, length.out=n_out)
 s <- seq(1,10,length.out=n_out)
@@
                     p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
-                    eta_w=params[["eta_w"]],eta_t=eta_t,s=s)
+                    eta_w=params[["eta_w"]],eta_c=eta_c,s=s)
 
@@
     + geom_contour(breaks=1,alpha=0.5,colour="black")
-    + facet_grid(s~eta_t, labeller=label_special)
+    + facet_grid(s~eta_c, labeller=label_special)
     + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
@@
 
-1. $\partial{R_0}/\partial{\eta_t} \geq 0$ and $\partial{R_0}/\partial{\eta_w} \geq 0$.
+1. $\partial{R_0}/\partial{\eta_c} \geq 0$ and $\partial{R_0}/\partial{\eta_w} \geq 0$.
 That is, the lower isolation the higher $R_0$. Note that perfect isolation occurs at $\eta_w = 0$. 
 2. $\partial{R_0}/\partial{\rho} \leq 0$; this is concluded from the Taylor expansion of $R_0$ around $\rho=0$ which is
-$R_0 \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_S} \Big(\gamma(\eta_w-1)(\gamma W_S+\omega W_I) + (\eta_t -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2)$.
+$R_0 \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_S} \Big(\gamma(\eta_w-1)(\gamma W_S+\omega W_I) + (\eta_c -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2)$.
 3. Based on the above Taylor expansion around $\rho=0$,
 $$\partial{R_0}/\partial{\omega}=  \frac{-\beta \rho}{\gamma W_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c), $$
-where $a=(\eta_w-1)W_I-(\eta_t-1)P_I W_I = ((s-p_I)\eta_t + (p_I-1)) W_I$, $b=2(\eta_w-1)\gamma W_S$ and $c=(\eta_w-1)\gamma^2 W_S$.
-Given that $\eta_t\leq \eta_w$, it is inferred that $a\geq 0$, $b\leq 0$ and $c \leq 0$.
+where $a=(\eta_w-1)W_I-(\eta_c-1)P_I W_I = ((s-p_I)\eta_c + (p_I-1)) W_I$, $b=2(\eta_w-1)\gamma W_S$ and $c=(\eta_w-1)\gamma^2 W_S$.
+Given that $\eta_c\leq \eta_w$, it is inferred that $a\geq 0$, $b\leq 0$ and $c \leq 0$.
 Let $\omega_1<0$ and $\omega_2>0$ be the roots of the quadratic expression in $\partial{R_0}/\partial{\omega}$.
@@
 In general, $a \geq 0$ is not true.
-$$a \geq 0 \iff (s-p_I) \eta_t \geq (1-p_I),$$ where $s=\frac{\eta_w}{\eta_t} \geq 1$.
+$$a \geq 0 \iff (s-p_I) \eta_c \geq (1-p_I),$$ where $s=\frac{\eta_w}{\eta_c} \geq 1$.
 As an example, in the case of a very accurate testing regime,  i.e., $P_I=1$, $a \geq 0$ is achieved.
 
-**Perfect isolation**; When $\eta_w=0$ and consequently $\eta_t=0$.
+**Perfect isolation**; When $\eta_w=0$ and consequently $\eta_c=0$.
 This leads to $a \leq 0$, $b<0$, $c<0$, thus $\partial{R_0}/\partial{\omega} \geq 0$.
@@
 
-(i) in the case of perfect isolation of the reported individuals, $\eta_t=0$, the followings are inferred (see the first few columns of the panel plot); 
+(i) in the case of perfect isolation of the reported individuals, $\eta_c=0$, the followings are inferred (see the first few columns of the panel plot); 
  - variations of the $\eta_w$ has a negligible effect on the critical testing.
@@
   unpack(as.list(params))
-         # a <- (eta_w-1)*W_I-(eta_t-1)*p_I*W_I
-         a <- ((s-p_I)*eta_t + (p_I-1))*W_I  #involving s
+         # a <- (eta_w-1)*W_I-(eta_c-1)*p_I*W_I
+         a <- ((s-p_I)*eta_c + (p_I-1))*W_I  #involving s
          b <- 2*(eta_w-1)*gamma*W_S
@@
 # eta_w <- seq(0,1, length.out=n_out)
-eta_t <- seq(0.001,1, length.out=n_out)
+eta_c <- seq(0.001,1, length.out=n_out)
 s <- seq(1,2,length.out=n_out)
@@
                     p_S=params[["p_S"]],p_I=p_I,p_R=params[["p_R"]],
-                    eta_w=params[["eta_w"]],eta_t=eta_t,s=s)
+                    eta_w=params[["eta_w"]],eta_c=eta_c,s=s)
 
@@
     + geom_contour(breaks=0,alpha=0.5,colour="black")
-    + facet_grid(s~eta_t, labeller=label_special)
+    + facet_grid(s~eta_c, labeller=label_special)
     + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)

### 2020-12-01 codes/ali_test_sir.Rmd
bmb notes

@@
 ---
+
+**TO DO** (for contour plots):
+
+- save as PDF
+- separate code for these plots into a separate .R file that can be run with a single batch command
+- consider using `tikzDevice`
 
@@
 ```{r show_contours_1, width=24,height=24,message=FALSE,warning=FALSE}
+brks_vec <- seq(0.8,1.05,by=0.05)
 p1 <- ggplot(df2,aes(x=omega,y=rho,z=R0_sub))+ theme_bw() +
@@
 p1_temp <- (p1
-    + geom_contour_filled(breaks=seq(0.80,1.05,by=0.05)
-                          )
+    + geom_contour_filled(breaks=brks_vec)
     + geom_contour(breaks=1,alpha=0.5,colour="black")
@@
     + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
-    + scale_fill_viridis_d(name=parse(text="R[0]"))
+    + scale_fill_viridis_d(name=parse(text="R[0]"), breaks=brks_vec)
     + ggtitle(TeX(sprintf("W_S=%.1f, W_I=W_R=1",W_S)))
@@
     + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
-    + scale_fill_viridis_d(name=parse(text="R[0]"))
+    + scale_fill_viridis_d(name=parse(text="R[0]"),breaks=brks_vec)
     + ggtitle("W[S]=1, W[I]=W[R]=1")
@@
 p2_temp
-# ggsave(p2_temp,filename = "R0contour_random_s.png" , path="../pix/")
+## ggsave(p2_temp,filename = "R0contour_random_s.png" , path="../pix/")
+ggsave(p2_temp,filename = "R0contour_random_s.pdf" , path="../pix/")
 ```

### 2020-12-02 codes/ali_test_sir.Rmd
panel plot update

@@
 library(latex2exp)
+library(rootSolve)
 unpack <- McMasterPandemic::unpack
@@
 # DFE; numerical approach
-library(rootSolve)
 DFE <- function(S,params){
@@
  omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, was 0.011
- W_S <- 1 #choose 0.3 for TTI testing plan and 1 for the random testing.
+ W_S <- 0.3 #choose 0.3 for TTI testing plan and 1 for the random testing.
 
@@
 df2 <- df2 %>% dplyr::mutate(R0_sub = ifelse(eta_w<eta_c, NA, R0))  
-
-# df2_fake <- df2 %>% dplyr::select(eta_c,eta_w)
-# df2_fake <- unique(df2_fake) %>% dplyr::mutate(flag=ifelse(eta_c<=eta_w,1,0))
-
 ```
@@
 ```{r show_contours_1, width=24,height=24,message=FALSE,warning=FALSE}
-brks_vec <- seq(0.8,1.05,by=0.05)
+brks_vec <- seq(0.8,1.05,by=0.05) # Break vector for unifying the legends in Random and TTI testing cases.
+
 p1 <- ggplot(df2,aes(x=omega,y=rho,z=R0_sub))+ theme_bw() +
@@
   ylab(TeX('$\\rho$, testing intensity (1/day per capita)')) +
-    # theme(panel.spacing=grid::unit(0,"lines")) # initial theme, comment it for TTI testing plot
-    theme(panel.spacing=grid::unit(0,"lines"),legend.position = "none") #AG, uncomment for Random Case
-# p1_temp <- (p1
-#     + geom_contour_filled(breaks=pretty(unique(df2$R0_sub),5),
-#                           alpha=0.5) ## breaks=0.2)
-#     + geom_contour(breaks=1,alpha=0.5,colour="black")
-#     + facet_grid(eta_w~eta_c, labeller=label_special)
-#     + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
-#     + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
-#     + scale_fill_viridis_d(name=parse(text="R[0]"))
-#     + ggtitle(TeX(sprintf("W_S=%.1f, W_I=W_R=1",W_S)))
-#     + geom_rect(data=df2,
-#                           fill=ifelse(df2$eta_c>df2$eta_w,"white","NA" ),
-#                           color= NA,
-#                           ymin=-1,
-#                           ymax=10,
-#                           xmin=-1,
-#                           xmax=2)
-#     )
-
+  theme(panel.spacing=grid::unit(0,"lines"),legend.position = c(0.89, 0.73)) # initial theme, comment it for TTI testing plot
+  # theme(panel.spacing=grid::unit(0,"lines"),legend.position = "none") #AG, uncomment for Random Case
 p1_temp <- (p1
@@
     + geom_contour(breaks=1,alpha=0.5,colour="black")
+    # + scale_fill_manual(palette=mycolors(brks_vec),drop=FALSE)
     + facet_grid(eta_w~eta_c, labeller=label_special)
@@
     + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
-    + scale_fill_viridis_d(name=parse(text="R[0]"), breaks=brks_vec)
+    + scale_fill_viridis_d(name=parse(text="R[0]"),drop=FALSE)
     + ggtitle(TeX(sprintf("W_S=%.1f, W_I=W_R=1",W_S)))
-    + geom_rect(data=df2,
-                          fill=ifelse(df2$eta_c>df2$eta_w,"white","NA" ),
+    + geom_rect(data=df2, fill=ifelse(df2$eta_c>df2$eta_w,"grey90","NA" ),
                           color= NA,
@@
 p1_temp
-# ## hack (By Dr. Ben Bolker) white rectangle to cover grid
-#   ggplot() + 
-#   scale_x_continuous(name="x") + 
-#   scale_y_continuous(name="y") +
-#   facet_grid(eta_w~eta_c, labeller=label_special) +
-#     geom_rect(data=df2_fake,
-#                           fill=ifelse(df2_fake$eta_c>df2_fake$eta_w,"white","black"),
-#                           color= NA,
-#                           ymin=-1,
-#                           ymax=10,
-#                           xmin=-1,
-#                           xmax=2)
 
-ggsave(p1_temp,filename = "R0contour_random.png" , path="../pix/")
-# ggsave(p1_temp,filename = "R0contour_TTI.png" , path="../pix/")
+# ggsave(p1_temp,filename = "R0contour_random.pdf" , path="../pix/")
+ggsave(p1_temp,filename = "R0contour_TTI.pdf" , path="../pix/")
 
@@
 ```{r contour_R0_2, message=FALSE,warning=FALSE }
- # specify the ranges
-beta  <- 0.339  ## set beta high enough to allow R0>1 in worst case 
-rho     <- seq(0,0.01, length.out=n_in)
-omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, 
-eta_c <- seq(0.001,0.1, length.out=n_out)
-s <- seq(1,10,length.out=n_out)
+# This part is done to carry our simulation when s=eta_c/eta_w is involved. Also to fill the upper triangle panels in the panel plot. 
+#  # specify the ranges
+# beta  <- 0.339  ## set beta high enough to allow R0>1 in worst case 
+# rho     <- seq(0,0.01, length.out=n_in)
+# omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, 
+# eta_c <- seq(0.001,0.1, length.out=n_out)
+# s <- seq(1,10,length.out=n_out) 
+# 
+#  df11 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
+#                     W_S=W_S,W_I=params[["W_I"]],W_R=params[["W_R"]],
+#                     p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
+#                     eta_w=params[["eta_w"]],eta_c=eta_c,s=s)
+# 
+# df22<- data.frame(df11,R0=apply(df11,1,function(params_in)R02(params=params_in)),
+#                       Fi= apply(df11,1,function(params_in)F_I(params=params_in)),
+#                        Sn=apply(df11,1,function(params_in)Sn_dfe(params = params_in)))
 
- df11 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
-                    W_S=W_S,W_I=params[["W_I"]],W_R=params[["W_R"]],
-                    p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
-                    eta_w=params[["eta_w"]],eta_c=eta_c,s=s)
-
-df22<- data.frame(df11,R0=apply(df11,1,function(params_in)R02(params=params_in)),
-                      Fi= apply(df11,1,function(params_in)F_I(params=params_in)),
-                       Sn=apply(df11,1,function(params_in)Sn_dfe(params = params_in)))
 ```
@@
 ```{r show_contours_2, width=24,height=24,message=FALSE,warning=FALSE}
-p11 <- ggplot(df22,aes(x=omega,y=rho,z=R0))+ theme_bw() +
-    theme(panel.spacing=grid::unit(0,"lines"))
-p2_temp <- (p11
-    + geom_contour_filled(breaks=pretty(unique(df22$R0),5),
-                          alpha=0.5) ## breaks=0.2)
-    + geom_contour(breaks=1,alpha=0.5,colour="black")
-    + facet_grid(s~eta_c, labeller=label_special)
-    + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
-    + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
-    + scale_fill_viridis_d(name=parse(text="R[0]"),breaks=brks_vec)
-    + ggtitle("W[S]=1, W[I]=W[R]=1")
-)
-p2_temp
-## ggsave(p2_temp,filename = "R0contour_random_s.png" , path="../pix/")
-ggsave(p2_temp,filename = "R0contour_random_s.pdf" , path="../pix/")
+# p11 <- ggplot(df22,aes(x=omega,y=rho,z=R0))+ theme_bw() +
+#     theme(panel.spacing=grid::unit(0,"lines"))
+# p2_temp <- (p11
+#     + geom_contour_filled(breaks=pretty(unique(df22$R0),5),
+#                           alpha=0.5) ## breaks=0.2)
+#     + geom_contour(breaks=1,alpha=0.5,colour="black")
+#     + facet_grid(s~eta_c, labeller=label_special)
+#     + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
+#     + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
+#     + scale_fill_viridis_d(name=parse(text="R[0]"),breaks=brks_vec)
+#     + ggtitle("W[S]=1, W[I]=W[R]=1")
+# )
+# p2_temp
+# ## ggsave(p2_temp,filename = "R0contour_random_s.png" , path="../pix/")
+# # ggsave(p2_temp,filename = "R0contour_random_s.pdf" , path="../pix/")
 ```
@@
 ```{r  chnages of R0 wrt omega,message=FALSE,warning=FALSE }
-R0_omega <- function(params){
-  unpack(as.list(params))
-         # a <- (eta_w-1)*W_I-(eta_c-1)*p_I*W_I
-         a <- ((s-p_I)*eta_c + (p_I-1))*W_I  #involving s
-         b <- 2*(eta_w-1)*gamma*W_S
-         c <- b*gamma/2
-
-         d <- (-beta*rho)/(gamma*W_S*omega^2*(gamma+omega)^2)*(a*omega^2 + b*omega + c)
-         return(d)
-}
-
-R0_omega(params)
+# R0_omega <- function(params){
+#   unpack(as.list(params))
+#          # a <- (eta_w-1)*W_I-(eta_c-1)*p_I*W_I
+#          a <- ((s-p_I)*eta_c + (p_I-1))*W_I  #involving s
+#          b <- 2*(eta_w-1)*gamma*W_S
+#          c <- b*gamma/2
+#          
+#          d <- (-beta*rho)/(gamma*W_S*omega^2*(gamma+omega)^2)*(a*omega^2 + b*omega + c)
+#          return(d)
+# }
+# 
+# R0_omega(params)
 
@@
  # specify the ranges
-if(F){
-beta  <- 0.5  ## set beta 
-gamma <- seq(0.1,0.5, length.out=n_out)
-p_I <- 0.9
-
-# eta_w <- seq(0,1, length.out=n_out)
-eta_c <- seq(0.001,1, length.out=n_out)
-s <- seq(1,2,length.out=n_out)
-
-rho <- seq(0,0.01, length.out=n_in)
-omega <- seq(0.011,2 , length.out=n_in) #note omega must be > rho, was 0.011
-
-W_S <- 1/100
-W_I <- 1/3
-W_R <- 1/3
-
- df3_1 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=gamma,omega=omega,rho=rho,
-                    W_S=W_S,W_I=W_I,W_R=W_R,
-                    p_S=params[["p_S"]],p_I=p_I,p_R=params[["p_R"]],
-                    eta_w=params[["eta_w"]],eta_c=eta_c,s=s)
-
-df3_2 <- data.frame(df3_1,R0=apply(df3_1,1,function(params_in)R02(params=params_in)),
-                      R0_omega= apply(df3_1,1,function(params_in)R0_omega(params=params_in))
-                       )
-}
+# if(F){
+# beta  <- 0.5  ## set beta 
+# gamma <- seq(0.1,0.5, length.out=n_out)
+# p_I <- 0.9
+# 
+# # eta_w <- seq(0,1, length.out=n_out)
+# eta_c <- seq(0.001,1, length.out=n_out)
+# s <- seq(1,2,length.out=n_out)
+# 
+# rho <- seq(0,0.01, length.out=n_in)
+# omega <- seq(0.011,2 , length.out=n_in) #note omega must be > rho, was 0.011
+# 
+# W_S <- 1/100
+# W_I <- 1/3
+# W_R <- 1/3
+# 
+#  df3_1 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=gamma,omega=omega,rho=rho,
+#                     W_S=W_S,W_I=W_I,W_R=W_R,
+#                     p_S=params[["p_S"]],p_I=p_I,p_R=params[["p_R"]],
+#                     eta_w=params[["eta_w"]],eta_c=eta_c,s=s)
+# 
+# df3_2 <- data.frame(df3_1,R0=apply(df3_1,1,function(params_in)R02(params=params_in)),
+#                       R0_omega= apply(df3_1,1,function(params_in)R0_omega(params=params_in))
+#                        )
+# }
 ```
@@
 ```{r show_contours_3, width=24,height=24,message=FALSE,warning=FALSE}
-if(F){
-p3 <- ggplot(df3_2,aes(x=omega,y=gamma,z=R0_omega))+ theme_bw() +
-    theme(panel.spacing=grid::unit(0,"lines"))
-(p3
-    + geom_contour_filled(breaks=pretty(unique(df3_2$R0_omega),5),
-                          alpha=0.5) ## breaks=0.2)
-    + geom_contour(breaks=0,alpha=0.5,colour="black")
-    + facet_grid(s~eta_c, labeller=label_special)
-    + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
-    + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
-    + scale_fill_viridis_d(name=parse(text="dR[0]"))
-)
-}
+# if(F){
+# p3 <- ggplot(df3_2,aes(x=omega,y=gamma,z=R0_omega))+ theme_bw() +
+#     theme(panel.spacing=grid::unit(0,"lines"))
+# (p3
+#     + geom_contour_filled(breaks=pretty(unique(df3_2$R0_omega),5),
+#                           alpha=0.5) ## breaks=0.2)
+#     + geom_contour(breaks=0,alpha=0.5,colour="black")
+#     + facet_grid(s~eta_c, labeller=label_special)
+#     + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
+#     + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
+#     + scale_fill_viridis_d(name=parse(text="dR[0]"))
+# )
+# }
 ```

### 2020-12-03 SIR_manuscript.tex
panel plot update
Intro updated

@@
 The advantage of non-random testing comparing to random testing has been investigated specifically during the COVID-19 pandemic (refs).
+
+There is useful comment from Troy Gayle [?, Ben?] on the purposes of testing:"The importance of the relative roles of at least four different, and sometimes conflicting, purposes of testing: (i) diagnosis and treatment, (ii) screening people for access to things like flights, surgery, LTCs etc, (iii) surveillance, and (iv) mitigation and control of disease spread." In this context, the random testing, the simplest form of testing strategy, is typically carried out for the purpose of surveillance.
+It is notable that, by random testing we mainly refer to PCR test given that there is almost negligable virological testing for survilance purposes.
+Furthermore, all of the accessible testing data is being used for one of the other three purposes rather than surveillance.
+These are non-random testing strategies with a different bios.
+Specifically, when the purpose is "(i) diagnosis and treatment" or "(iv) mitigation and control of disease spread", the bios is towards infectious individuals more strongly for (i) than (iv).
+When people are tested for "(ii) screening", the bios is towards the group-specific charactristics [I am making this up for now!].
+For example, people who are getting on flights may/will be more mobile than people awaiting for a surgery and are probably going to stay in long-term care homes and consequently more isolated to begin with.
 
@@
 \centering
-\includegraphics[width=\linewidth]{./pix/R0contour_random.png}
+\includegraphics[width=\linewidth]{./pix/R0contour_random.pdf}
 \caption{}\label{p.a}
@@
 \centering
-\includegraphics[width=\linewidth]{./pix/R0contour_TTI.png}
+\includegraphics[width=\linewidth]{./pix/R0contour_TTI.pdf}
 \caption{}\label{p.b}

### 2020-12-04 SIR_manuscript.tex
a few more comments…

@@
 
-\newcommand{\fady}[1]{\textcolor{cyan}{$\langle${\slshape{\bfseries Fady:} #1 }$\rangle$}}
+\newcommand{\comment}{\showcomment}
+%% \newcommand{\comment}{\nocomment}
+
+\newcommand{\showcomment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}
+\newcommand{\nocomment}[3]{}
+
+\newcommand{\fady}[1]{\comment{cyan}{Fady}{#1}}
+\newcommand{\ali}[1]{\comment{magenta}{Ali}{#1}}
+\newcommand{\jd}[1]{\comment{blue}{JD}{#1}}
+\newcommand{\djde}[1]{\comment{red}{DJDE}{#1}}
+\newcommand{\bmb}[1]{\comment{red}{BMB}{#1}}
+\newcommand{\todo}[1]{\comment{red}{TODO}{#1}}
+
 \newcommand{\Rnum}{\mathcal{R}_0}
@@
 
-[[JD: There is a gap here between unravelling data and what we actually do.]]
+\jd{There is a gap here between unravelling data and what we actually do.}
 
-[[JD: This \P\ is confusing. The topic sentence seems to be about kinds of tests, and then there is a jump to sampling strategies.]]
+\jd{This \P\ is confusing. The topic sentence seems to be about kinds of tests, and then there is a jump to sampling strategies.}
 There are different testing strategies with different test-specific efficacy, the time required for the results to be reported and the cost of the test.
 Testing strategies can be grouped into two categories including (i) random testing across the whole population and (ii) Non-random testing.
-[[JD: Random testing is not a broad category, and non-random testing is very vague.
-What are we trying to convey with this \P?
-Ali: I am thinking of the weighted testing parameters W's in our model, that reflects testing strategies.(not sure the best replacement for "non-random" testing?)]] Note that contact tracing is an example of non-random testing strategies.
+\jd{Random testing is not a broad category, and non-random testing is very vague.
+What are we trying to convey with this \P?} \ali{I am thinking of the weighted testing parameters W's in our model, that reflects testing strategies.(not sure the best replacement for "non-random" testing?)} Note that contact tracing is an example of non-random testing strategies.
 Each of these strategies has been used as a tool to study COVID-19; see (refs) for group (i) and (refs) for group (ii).
@@
 
-[[JD: It would be nice to modularize the model somehow, but probably not worth anyone's time.]]
+\jd{It would be nice to modularize the model somehow, but probably not worth anyone's time.}
 
-[Ali: not sure if these should go to the Supp material?] In our model the following parameters are used (for a summery of parameters' description, see Table \ref{tab:params}).
+\ali{not sure if these should go to the Supp material?} In our model the following parameters are used (for a summery of parameters' description, see Table \ref{tab:params}).
 $N_0$ is the total population size, $\omega$, (1/day), is the rate of onward flow from the awaiting positive or negative compartments to reported or untested compartments, respectively, $\gamma$, (1/day), is the recovery rate, $\rho$, (1/day), is the per capita testing intensity across the whole population.
@@
 say it or not?
-[[JD: The way perfect specificity is presented is confusing here.
-We need to distinguish between the model we first develop (12 boxes) and the model we analyze (10 boxes)]]
-[[Ali: If we use the full model, with 12 compartments including $S_p$ and $S_c$, the DFE will be; $S_n=S_p=0$, $S_u+S_c=N_0$ (it can be assumed that $S_u=N_0$ and $S_c=0$) and all $I$ and $R$ compartments to be zero.
-This simplifies the expression of $\Rnum$ since $F_I=\rho$.]]
+\jd{The way perfect specificity is presented is confusing here.
+We need to distinguish between the model we first develop (12 boxes) and the model we analyze (10 boxes)}
+\ali{If we use the full model, with 12 compartments including $S_p$ and $S_c$, the DFE will be; $S_n=S_p=0$, $S_u+S_c=N_0$ (it can be assumed that $S_u=N_0$ and $S_c=0$) and all $I$ and $R$ compartments to be zero.
+This simplifies the expression of $\Rnum$ since $F_I=\rho$.}
+\jd{I was asking you to clarify the MS -- you never explain when you switch to assuming that the false-positive rate is zero -- not explain to me in notes.
+This explanation here looks wrong, though (compare with \eqref{dfe}).}
 
@@
 
-The Basic reproduction number, $\Rnum$, was calculated by using the next generation matrix method developed by \cite{van2002reproduction}. $\Rnum$ is
+The basic reproduction number, $\Rnum$, was calculated by using the next generation matrix method developed by \cite{van2002reproduction}. $\Rnum$ is
 
@@
 \end{equation}
+\jd{It's hard to any point to presenting this equation in the main text if A, B C are not also in the main text.}
 Further details are provided in the Appendix.
 
-  [to be completed!
-Something needs to be said here about simulation part].
-The analytical calculation of the next generation matrix, $G$, and the derivation of $\Rnum$ was carried out in Maple (ref) by using simple linear Algebra package.
+\todo{Something needs to be said here about simulation part.} The analytical calculation of the next generation matrix, $G$, and the derivation of $\Rnum$ was carried out in Maple (ref) by using simple linear Algebra package.
   We used \Rlogo for the simulation part which included using the explicit expression of $\Rnum$ \ref{R0} as a function of the underlying parameters, specifying the parameters a realistic range, calculating the corresponding $\Rnum$ and plotting the contours of $\Rnum$ by using ggplot package in \Rlogo.
@@
 \centering
-\includegraphics[width=\linewidth]{./pix/R0contour_random.pdf}
+\includegraphics[width=\linewidth]{./pix/R0contour_random.png}
 \caption{}\label{p.a}
@@
 \centering
-\includegraphics[width=\linewidth]{./pix/R0contour_TTI.pdf}
+\includegraphics[width=\linewidth]{./pix/R0contour_TTI.png}
 \caption{}\label{p.b}
@@
 \subsubsection{Explicit models of TTI (trace/test/isolate) based on network or agent-based models}
-\citep{endo2020implication} [Ali: It seems to me that this is just a statistical model to estimate the parent-offspring of an infected index, not sure if it fits into agent-based group!] Used simulation on a branching process model to assess the forward and backward contact tracing efficiency.
+\citep{endo2020implication} \ali{It seems to me that this is just a statistical model to estimate the parent-offspring of an infected index, not sure if it fits into agent-based group!} Used simulation on a branching process model to assess the forward and backward contact tracing efficiency.
 Assuming a negative-binomial branching process with a mean R, reproduction number, and overdispersion parameter k, the mean total number of generation G3 and averted G3 are estimated.

### 2020-12-04 SIR_manuscript.tex
method updated

@@
 The classical SIR model is based on the following implicit assumptions (see a ref for details); well-mixed population, homogeneity of the population, exponentially distributed duration of infection and large population size (may be refer the reader to \cite{keeling2011modeling}).
-In addition to these standard assumptions, our model, \crefrange{eq1}{eq12}, assumes: (i) there is a single force of infection, $\Lambda$, that is, where a susceptible individual moves to does not depend on who infected it, (ii) a perfectly specific test, $p_S=0$, for simplicity in analysis, (iii)  $\eta_c \leq \eta_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
-- [Q] why not birth and death are included?
-say it or not?
-\jd{The way perfect specificity is presented is confusing here.
-We need to distinguish between the model we first develop (12 boxes) and the model we analyze (10 boxes)}
-\ali{If we use the full model, with 12 compartments including $S_p$ and $S_c$, the DFE will be; $S_n=S_p=0$, $S_u+S_c=N_0$ (it can be assumed that $S_u=N_0$ and $S_c=0$) and all $I$ and $R$ compartments to be zero.
-This simplifies the expression of $\Rnum$ since $F_I=\rho$.}
-\jd{I was asking you to clarify the MS -- you never explain when you switch to assuming that the false-positive rate is zero -- not explain to me in notes.
-This explanation here looks wrong, though (compare with \eqref{dfe}).}
+In addition to these standard assumptions, our model, \crefrange{eq1}{eq12}, assumes: (i) there is a single force of infection, $\Lambda$, that is, where a susceptible individual moves to does not depend on who infected it, (ii) $\eta_c \leq \eta_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals, (iii) a perfectly specific test, $p_S=0$, for simplicity in analysis.
+Note that the latter assumption combined with the assumption that the total population was initially distributed between untested and awaiting for negative test, that is $S_u+S_n=0$, reduce the model to 10 equations with equations \ref{eq3} and \ref{eq4} are eliminated.
+Specifically, when the population is at steady states and the number of infected people is negligable, the assumption of perfect specificity results in $S_p=0$ in \cref{eq3} and $d S_c/dt=0$ from \cref{eq4}.
+Also, since we assumed $S_u+S_n=0$, initially $S_c=0$ and stays in this state.
 
@@
 \end{equation}
-\jd{It's hard to any point to presenting this equation in the main text if A, B C are not also in the main text.}
+where
+\begin{align*}
+A=& \gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_c p_I) F_I, \\
+B=& \big(\omega+(F_I+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_c\omega) \omega p_I F_I }{\omega+\gamma}, \\ 
+C=& \frac{\beta/\gamma}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}.
+\end{align*}
+
 Further details are provided in the Appendix.
@@
 
-Note that the block matrix $G_{12}$ does not influence $\Rnum$ defined as the spectral radius of $G$. All matters here are the eigenvalues of $G_{11}$, which are 0 and $\Rnum$ \eqref{R0}, 
-where
-\begin{align*}
-A=& \gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_c p_I) F_I, \\
-B=& \big(\omega+(F_I+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_c\omega) \omega p_I F_I }{\omega+\gamma}, \\ 
-C=& \frac{\beta/\gamma}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}.
-\end{align*}
+Note that the block matrix $G_{12}$ does not influence $\Rnum$ defined as the spectral radius of $G$. All matters here are the eigenvalues of $G_{11}$, which are 0 and $\Rnum$ \eqref{R0}.
 

### 2020-12-04 SIR_manuscript.tex
more edits on Intro section

@@
 \jd{This \P\ is confusing. The topic sentence seems to be about kinds of tests, and then there is a jump to sampling strategies.}
-There are different testing strategies with different test-specific efficacy, the time required for the results to be reported and the cost of the test.
-Testing strategies can be grouped into two categories including (i) random testing across the whole population and (ii) Non-random testing.
-\jd{Random testing is not a broad category, and non-random testing is very vague.
-What are we trying to convey with this \P?} \ali{I am thinking of the weighted testing parameters W's in our model, that reflects testing strategies.(not sure the best replacement for "non-random" testing?)} Note that contact tracing is an example of non-random testing strategies.
-Each of these strategies has been used as a tool to study COVID-19; see (refs) for group (i) and (refs) for group (ii).
-The advantage of non-random testing comparing to random testing has been investigated specifically during the COVID-19 pandemic (refs).
 
-There is useful comment from Troy Gayle [?, Ben?] on the purposes of testing:"The importance of the relative roles of at least four different, and sometimes conflicting, purposes of testing: (i) diagnosis and treatment, (ii) screening people for access to things like flights, surgery, LTCs etc, (iii) surveillance, and (iv) mitigation and control of disease spread." In this context, the random testing, the simplest form of testing strategy, is typically carried out for the purpose of surveillance.
+It is often the purpose of testing, apart from the testing infrastructure, that dictates/defines \ali{[better word?]} the testing strategy.
+There are at least four different, and sometimes conflicting, purposes of testing: (i) diagnosis and treatment, (ii) screening people for access to things like flights, surgery, LTCs etc, (iii) surveillance, and (iv) mitigation and control of disease spread (ref to Troy Gayle?).
+Further, the testing strategies can be grouped into two categories including random testing across the whole population and Non-random/biosed testing.
+Random testing is the simplest form of testing strategy and is typically carried out for the purpose of surveillance.
 It is notable that, by random testing we mainly refer to PCR test given that there is almost negligable virological testing for survilance purposes.
@@
 
+-Examples of random versus biosed testing strategies;
+Each of these strategies has been used as a tool to study COVID-19; see (refs) for group (i) and (refs) for group (ii).
+The advantage of non-random testing comparing to random testing has been investigated specifically during the COVID-19 pandemic (refs).
+
+- Speed of test; 
 Much effort has been invested in developing widely accessible and fast tests in order to mitigate the spread of the pandemic (refs).

### 2020-12-08 codes/sir_macpan.Rnw
an excercise with MacPan

\documentclass{article}

\begin{document}
By Ali Gharouni,
<<date,echo=F>>=
Sys.time()
@
This is my a test for MacPan package.

\section{Install and Play}

<<install the package, echo=F  >>=
# install remotes package if necessary:
if (!require(remotes)) {
install.packages("remotes")
}
## install development version of bbmle:
if (!require("bbmle") || packageVersion("bbmle") < "1.0.23.5") {
remotes::install_github("bbolker/bbmle")
}
## install the target package and all its dependencies:
remotes::install_github("bbolker/McMasterPandemic",
dependencies = TRUE,
build_vignettes = TRUE
)
library(McMasterPandemic)
library(ggplot2); theme_set(theme_bw())
library(cowplot)

library(directlabels)

# params1 <- read_params("ICU1.csv")
# folder <- system.file("params", package="McMasterPandemic")
# list.files(folder)
# knitr::kable(describe_params(params1))

# knitr::kable(round(t(summary(params1)),2))
# knitr::kable(round(t(get_R0(params1, components=TRUE)),2))
@

<<6: MacPan Params , echo=F>>=
library(zoo)
library(tidyverse)

params <- read_params("ICU1.csv")

knitr::kable(round(t(summary(params)),2))
knitr::kable(round(t(get_R0(params, components=TRUE)),2)) 
@

<<Ali:SIR , echo=F>>=
# A simple SIR model
library(deSolve)
unpack <- McMasterPandemic::unpack

sir.mod <- function(time,state,params){
    unpack(as.list(c(state,params)))
    dS.dt <- -beta*S*I/N0
    dI.dt <-  beta*S*I/N0-gamma*I
    dR.dt <-  gamma*I
    # return the rate of change
    dxdt <- c(dS.dt,dI.dt,dR.dt)
    return(list(dxdt))
}
sir.params <- c(N0=10^6, beta=0.5,gamma=1/7)
class(sir.params) <- "params_pansim"
# Set the initial state
sir.state_init <- c(S=sir.params[["N0"]], 
                I=1,
                R=0)
#set the number of days for SIR simulation as simulation time of MacPan 
d <- seq(0,as.numeric(enddate-startdate),by=1)

sir.R0 <- sir.params[["beta"]]/sir.params[["gamma"]]

sir.out <- as.data.frame(
  ode(
    func=sir.mod,
    y=sir.state_init,
    times= d, 
    parms=update(sir.params)
    ))

sir.out2 <- sir.out %>%
  pivot_longer(c(S,I,R), names_to = "compartment", values_to = "value")

gg1 <- (ggplot(data=sir.out2, aes(x=time,y=value,col=compartment))
    + geom_line()
    # + scale_y_log10(limits=c(0.1,sir.params[["N0"]]))
)
direct.label(gg1,"last.bumpup")

@

<<Try MacPan to recover the SIR result, echo=FALSE>>=
# update MacPan's params according to the SIR params
params <- update(params, c(beta0=sir.params[["beta"]],
                           gamma_a=sir.params[["gamma"]],
                           gamma_m=sir.params[["gamma"]],
                           gamma_s=sir.params[["gamma"]],
                           gamma_p=sir.params[["gamma"]]))

pp <- fix_pars(params, target = c(R0 = sir.R0)) #not sure if I need to update Gbar=6?
# Q? should I update beta0 again? 
state <- make_state(params=pp)
# state <- update(state,c(S=sir.params[["N0"]],E=0,Ip=0,Im=0))

sim0 <- run_sim(pp,state,start_date=startdate,end_date=enddate)

gg0 <- (ggplot(sim0,aes(x=date))
+ geom_point(aes(y=incidence))
)
print(gg0)

# combining the SIR and MacPan results also calculating incidences fr the SIR model
sim_combined <- sim0 %>% mutate(sir.S=sir.out[["S"]],sir.I=sir.out[["I"]],sir.S=sir.out[["R"]],
                                sir.incidence=sir.params[["beta"]]*sir.out[["S"]]*sir.out[["I"]])

gg01 <- (ggplot(sim_combined,aes(x=date))
+ geom_point(aes(y=I))
+ geom_point(aes(y=sir.I), color='red')
)
print(gg01)

@

\end{document}

### 2020-12-08 SIR_manuscript.tex
working on ms

@@
 \jd{There is a gap here between unravelling data and what we actually do.}
+\bmb{As JD suggests, the focus of the paper is not on unraveling data (although that is something we \emph{could} do, and are trying to do, with this modeling framework as well).
+I would say that the ``bullet points'' to start the paper are (1) the processes driving the observed dynamics of the COVID-19 epidemic include both epidemiological processes (infection, recovery, etc.) and testing processes (``processes'' is repeated here, could rephrase); (2) in addition to determining what is actually observed (case reports), testing processes can also feed back to affect the epidemiological dynamics through isolation and quarantine; (3) your last sentence in the \P\ above.}
 
@@
 
+\bmb{It's fine for this \P\ to be about testing, but the topic sentence still doesn't say what the paragraph is trying to do.
+You should probably give the general idea of how isolation and quarantine affect epi dynamics (i.e.
+individuals who in the process of testing may partially or fully self-isolate, and individuals who have tested positive are highly likely to self-isolate), and say that your goal is to figure out the effects of isolation and testing: when can isolation and testing control an epidemic?
+You might even want to do this in the \emph{previous} paragraph.
+Then you can say that since isolation applies to people who have been tested, that the epi effects will depend on \emph{who gets tested}, i.e.
+how tests are distributed across compartments, or for a given testing intensity how many people in each epidemiological compartment get tested.
+Then you can go through some examples, i.e.
+start by saying that in the case where tests are randomly assigned within the population (as might happen, for example, when disease surveillance is the purpose of testing), the tests will be weighted equally across all compartments \ldots}
+
 It is often the purpose of testing, apart from the testing infrastructure, that dictates/defines \ali{[better word?]} the testing strategy.
-There are at least four different, and sometimes conflicting, purposes of testing: (i) diagnosis and treatment, (ii) screening people for access to things like flights, surgery, LTCs etc, (iii) surveillance, and (iv) mitigation and control of disease spread (ref to Troy Gayle?).
-Further, the testing strategies can be grouped into two categories including random testing across the whole population and Non-random/biosed testing.
+There are at least four different, and sometimes conflicting, purposes of testing: (i) diagnosis and treatment, (ii) screening people for access to things like flights, surgery, LTCs etc, (iii) surveillance, and (iv) mitigation and control of disease spread.
+
+Further, the testing strategies can be grouped into two categories including random testing across the whole population and Non-random/biased testing.
 Random testing is the simplest form of testing strategy and is typically carried out for the purpose of surveillance.
-It is notable that, by random testing we mainly refer to PCR test given that there is almost negligable virological testing for survilance purposes.
+It is notable that, by random testing we mainly refer to PCR test given that there is almost negligible virological testing for surveillance purposes.
 Furthermore, all of the accessible testing data is being used for one of the other three purposes rather than surveillance.
-These are non-random testing strategies with a different bios.
-Specifically, when the purpose is "(i) diagnosis and treatment" or "(iv) mitigation and control of disease spread", the bios is towards infectious individuals more strongly for (i) than (iv).
-When people are tested for "(ii) screening", the bios is towards the group-specific charactristics [I am making this up for now!].
+These are non-random testing strategies with a different bias.
+Specifically, when the purpose is "(i) diagnosis and treatment" or "(iv) mitigation and control of disease spread", the bias is towards infectious individuals more strongly for (i) than (iv).
+When people are tested for "(ii) screening", the bias is towards the group-specific characteristics [I am making this up for now!].
 For example, people who are getting on flights may/will be more mobile than people awaiting for a surgery and are probably going to stay in long-term care homes and consequently more isolated to begin with.
 
--Examples of random versus biosed testing strategies;
+-Examples of random versus biased testing strategies;
 Each of these strategies has been used as a tool to study COVID-19; see (refs) for group (i) and (refs) for group (ii).
@@
 Also, two 'accumulator' compartments, $N$ and $P$, were incorporated in the model in order to collect cumulative reported negative or positive tests.
-The model and details of calculation of $\Rnum$ are presented in the Supplimentary Materials.
+The model and details of calculation of $\Rnum$ are presented in the Supplementary Materials.
 The flowchart model is as follows
@@
 Note that the latter assumption combined with the assumption that the total population was initially distributed between untested and awaiting for negative test, that is $S_u+S_n=0$, reduce the model to 10 equations with equations \ref{eq3} and \ref{eq4} are eliminated.
-Specifically, when the population is at steady states and the number of infected people is negligable, the assumption of perfect specificity results in $S_p=0$ in \cref{eq3} and $d S_c/dt=0$ from \cref{eq4}.
+Specifically, when the population is at steady states and the number of infected people is negligible, the assumption of perfect specificity results in $S_p=0$ in \cref{eq3} and $d S_c/dt=0$ from \cref{eq4}.
 Also, since we assumed $S_u+S_n=0$, initially $S_c=0$ and stays in this state.
@@
   $\Lambda$ & Force of infection & 1/time & - \\ \hline
-  $p_S$ & Probability of false positive for susciptibles& - & 0 \\ \hline
+  $p_S$ & Probability of false positive for susceptible& - & 0 \\ \hline
   $p_I$ & Probability of being infected and tested positive & - & 1 \\ \hline
@@
 This simulation reflects the behavior of $\Rnum$ with respect to the selected parameters for two different testing strategies: (i) random testing, represented by all testing weights to be the same, $W_S=W_I=W_R$, and (ii) non-random testing, when testing weight are not equal.
-For simmulation purposes we chose $W_S=W_I=W_R=1$ for random testing, and $W_S=0.3$ and $W_I=W_R=1$ for non-random testing strategy.
+For simulation purposes we chose $W_S=W_I=W_R=1$ for random testing, and $W_S=0.3$ and $W_I=W_R=1$ for non-random testing strategy.
 Note that the critical contour of $\Rnum=1$ is plotted in solid line in Figure \ref{pan}.
@@
 Our numerical simulation in support of the analytical results in Proposition \ref{prop1} are presented in Figure \ref{pan}.
-It can be infered that when random testing strategy is applied, Figure \ref{p.a}, comparing to the non-random testing strategy, Figure \ref{p.a}, the critical testing intensity of the coresponding panels are lower in non-random testing.
+It can be inferred that when random testing strategy is applied, Figure \ref{p.a}, comparing to the non-random testing strategy, Figure \ref{p.a}, the critical testing intensity of the corresponding panels are lower in non-random testing.
 Also, it appears that speeding the test reporting, i.e., increasing $\omega$, does not significantly lower $\Rnum$ when awaiting people follow the isolation perfectly, i.e., $\eta_w$ is closer to 0.
-However, speeding the test reporting reduces the epidemy more when the awaiting people less follow the isolation, i.e., $\eta_w$ is closer to 1.
+However, speeding the test reporting reduces the epidemic more when the awaiting people less follow the isolation, i.e., $\eta_w$ is closer to 1.
 
@@
 
-The next generation matrix is $G = F V^{-1}$, where matrix $F$ represents the inflow of new infection to the infected compartments and matrix $V$ represents the flow in the infected compartments when the population is totally susiptable.
+The next generation matrix is $G = F V^{-1}$, where matrix $F$ represents the inflow of new infection to the infected compartments and matrix $V$ represents the flow in the infected compartments when the population is totally susceptible.
 For model, \crefrange{eq1}{eq12}, $F$ and $V$ are

### 2020-12-08 codes/sir_macpan.Rnw
macpan example

@@
 # install remotes package if necessary:
-if (!require(remotes)) {
+while (!require(remotes)) {
     install.packages("remotes")
@@
                 R=0)
-#set the number of days for SIR simulation as simulation time of MacPan 
+##set the number of days for SIR simulation as simulation time of MacPan
+startdate <- as.Date("01-03-2020")
+enddate <- as.Date("01-09-2020")
 d <- seq(0,as.numeric(enddate-startdate),by=1)
@@
 params <- update(params, c(beta0=sir.params[["beta"]],
+                           alpha=0, ## no asymptomatic
+                           sigma=3, ## very short latent period
                            gamma_a=sir.params[["gamma"]],
@@
 
-pp <- fix_pars(params, target = c(R0 = sir.R0)) #not sure if I need to update Gbar=6?
+pp <- params
+## pp <- fix_pars(params, target = c(R0 = sir.R0)) #not sure if I need to update Gbar=6?
 # Q? should I update beta0 again? 
 state <- make_state(params=pp)
+state[] <- 0
+state[["S"]] <- 1e6
+state[["Im"]] <- 1
 # state <- update(state,c(S=sir.params[["N0"]],E=0,Ip=0,Im=0))
 
-sim0 <- run_sim(pp,state,start_date=startdate,end_date=enddate)
+summary(pp)
+sim0 <- run_sim(pp,state,
+                start_date=startdate,
+                end_date=enddate,
+                use_ode=TRUE)
 
@@
 # combining the SIR and MacPan results also calculating incidences fr the SIR model
-sim_combined <- sim0 %>% mutate(sir.S=sir.out[["S"]],sir.I=sir.out[["I"]],sir.S=sir.out[["R"]],
+sim_combined <- sim0 %>% mutate(sir.S=sir.out[["S"]],
+                                sir.I=sir.out[["I"]],
+                                sir.S=sir.out[["R"]],
                                 sir.incidence=sir.params[["beta"]]*sir.out[["S"]]*sir.out[["I"]])
@@
 print(gg01)
-
+print(gg01 + scale_y_log10())
 @

### 2020-12-08 codes/sir_macpan.Rnw
cleaned up the code

@@
 library(directlabels)
-
-# params1 <- read_params("ICU1.csv")
-# folder <- system.file("params", package="McMasterPandemic")
-# list.files(folder)
-# knitr::kable(describe_params(params1))
-
-# knitr::kable(round(t(summary(params1)),2))
-# knitr::kable(round(t(get_R0(params1, components=TRUE)),2))
+library(zoo)
+library(tidyverse)
 @
@@
 <<6: MacPan Params , echo=F>>=
-library(zoo)
-library(tidyverse)
-
 params <- read_params("ICU1.csv")
@@
 ## pp <- fix_pars(params, target = c(R0 = sir.R0)) #not sure if I need to update Gbar=6?
-# Q? should I update beta0 again? 
 state <- make_state(params=pp)
@@
 state[["Im"]] <- 1
-# state <- update(state,c(S=sir.params[["N0"]],E=0,Ip=0,Im=0))
 

### 2020-12-08 codes/sir_macpan.Rnw
macpan/SIR equivalence

--- 2020-12-08 codes/sir_macpan.Rnw
@@
 <<Try MacPan to recover the SIR result, echo=FALSE>>=
-# update MacPan's params according to the SIR params
+## update MacPan's params according to the SIR params
+##
+big_rate <- 20
 params <- update(params, c(beta0=sir.params[["beta"]],
                            alpha=0, ## no asymptomatic
-                           sigma=3, ## very short latent period
-                           gamma_a=sir.params[["gamma"]],
+                           sigma=big_rate, ## very short latent period
+                           ## would like this to -> Inf but can't
+                           gamma_a=0,
+                           mu=1,  ## all cases mild
                            gamma_m=sir.params[["gamma"]],
-                           gamma_s=sir.params[["gamma"]],
-                           gamma_p=sir.params[["gamma"]]))
+                           gamma_s=0,
+                           ## try to skip presymptomatic period
+                           gamma_p=big_rate
+                           ))
 
@@
 + geom_point(aes(y=I))
-+ geom_point(aes(y=sir.I), color='red')
++ geom_lines(aes(y=sir.I), color='red')
 )

### 2020-12-09 codes/ali_test_sir.Rmd
working on ms
seperating functions

@@
 ### Model
+
 The model is
@@
 
-```
-
-```{r, model formulation}
-sir.model <- function(time,state,params){
-    unpack(as.list(c(state,params)))
-    ## Force of Infection  
-    Lambda <- beta * (I_u + eta_w*(I_n+I_p) + eta_c*I_t)/N0
-    ## scaling the weights
-    W <- W_S*S_u+W_I*I_u+W_R*R_u
-    sigma <- rho*N0/W
-    F_S <- sigma*W_S
-    F_I <- sigma*W_I
-    F_R <- sigma*W_R
-
-    dS_u.dt <- -Lambda*S_u - (1-p_S) * F_S * S_u + omega * S_n
-    dS_n.dt <- -Lambda*S_n + (1-p_S) * F_S * S_u - omega * S_n 
-    dI_u.dt <-  Lambda*S_u + omega * I_n - F_I * I_u - gamma * I_u
-    dI_n.dt <-  Lambda*S_n + (1-p_I) * F_I * I_u  - omega * I_n - gamma * I_n
-    dI_p.dt <-  p_I * F_I * I_u - omega * I_p - gamma * I_p 
-    dI_t.dt <-  omega * I_p - gamma * I_t 
-    dR_u.dt <- gamma * I_u + omega * R_n - F_R * R_u  
-    dR_n.dt <- gamma * I_n + (1-p_R) * F_R * R_u - omega * R_n
-    dR_p.dt <- gamma * I_p + p_R * F_R * R_u - omega * R_p
-    dR_t.dt <- gamma * I_t + omega * R_p
-    dN.dt <- omega * (S_n + I_n + R_n)
-    dP.dt <- omega *(I_p + R_p)
-
-    # return the rate of change
-  dxdt <- c(dS_u.dt,dS_n.dt,dI_u.dt,dI_n.dt,dI_p.dt,dI_t.dt,dR_u.dt,dR_n.dt,dR_p.dt,dR_t.dt,dN.dt,dP.dt)
-    ## }
-    return(list(dxdt))
-}
-
+source("SIRfunctions.R")
 ```
@@
 d <- seq(0,200,by=0.1)
-# Conditional stop for desolver
-# rootfun <- function (time, state,params) { 
-#   unpack(as.list(c(state,params)))
-#   return(I_u - 1) }
 ```
@@
 # DFE; numerical approach
-DFE <- function(S,params){
-  S_u <- S[1]
-  S_n <- S[2]
-  I_u <-0
-  R_u <-0
-  unpack(as.list(params))
-  W <- W_S*S_u+W_I*I_u+W_R*R_u
-  sigma <- rho*N0/W
-  F_S <- sigma*W_S
-
- return(c(eq1=S_u+S_n-N0,
-          eq2=-F_S*S_u+omega*S_n))
-}
-
 roots_dfe <- multiroot(DFE, c(1, 1), parms = params, positive = T)
@@
 # DFE; analytic approach
-Sn_dfe <-function(params){
-  unpack(as.list(params))
-  return((rho*(1-p_S)*N0)/omega)
-  }
+Su <- Su_dfe(params)
+Sn <- Sn_dfe(params)
 
-Su_dfe<- params[["N0"]]-Sn_dfe(params)
-
-state_dfe <- c(S_u=Su_dfe, S_n=Sn_dfe(params),
+state_dfe <- c(S_u=Su, S_n=Sn,
                 I_u=0,I_n=0,I_p=0,I_t=0,
@@
 
-# ggsave(p1_temp,filename = "R0contour_random.pdf" , path="../pix/")
+## BMB: FIXME: probably wrong but ... ???
+## ggsave(p1_temp,filename = "R0contour_random.pdf" , path="../pix/")
 ggsave(p1_temp,filename = "R0contour_TTI.pdf" , path="../pix/")
@@
 ```{r  chnages of R0 wrt omega,message=FALSE,warning=FALSE }
-# R0_omega <- function(params){
-#   unpack(as.list(params))
-#          # a <- (eta_w-1)*W_I-(eta_c-1)*p_I*W_I
-#          a <- ((s-p_I)*eta_c + (p_I-1))*W_I  #involving s
-#          b <- 2*(eta_w-1)*gamma*W_S
-#          c <- b*gamma/2
-#          
-#          d <- (-beta*rho)/(gamma*W_S*omega^2*(gamma+omega)^2)*(a*omega^2 + b*omega + c)
-#          return(d)
-# }
-# 
-# R0_omega(params)
 
 ```
-
-```{r }
- # specify the ranges
-# if(F){
-# beta  <- 0.5  ## set beta 
-# gamma <- seq(0.1,0.5, length.out=n_out)
-# p_I <- 0.9
-# 
-# # eta_w <- seq(0,1, length.out=n_out)
-# eta_c <- seq(0.001,1, length.out=n_out)
-# s <- seq(1,2,length.out=n_out)
-# 
-# rho <- seq(0,0.01, length.out=n_in)
-# omega <- seq(0.011,2 , length.out=n_in) #note omega must be > rho, was 0.011
-# 
-# W_S <- 1/100
-# W_I <- 1/3
-# W_R <- 1/3
-# 
-#  df3_1 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=gamma,omega=omega,rho=rho,
-#                     W_S=W_S,W_I=W_I,W_R=W_R,
-#                     p_S=params[["p_S"]],p_I=p_I,p_R=params[["p_R"]],
-#                     eta_w=params[["eta_w"]],eta_c=eta_c,s=s)
-# 
-# df3_2 <- data.frame(df3_1,R0=apply(df3_1,1,function(params_in)R02(params=params_in)),
-#                       R0_omega= apply(df3_1,1,function(params_in)R0_omega(params=params_in))
-#                        )
-# }
-```
-
-```{r show_contours_3, width=24,height=24,message=FALSE,warning=FALSE}
-# if(F){
-# p3 <- ggplot(df3_2,aes(x=omega,y=gamma,z=R0_omega))+ theme_bw() +
-#     theme(panel.spacing=grid::unit(0,"lines"))
-# (p3
-#     + geom_contour_filled(breaks=pretty(unique(df3_2$R0_omega),5),
-#                           alpha=0.5) ## breaks=0.2)
-#     + geom_contour(breaks=0,alpha=0.5,colour="black")
-#     + facet_grid(s~eta_c, labeller=label_special)
-#     + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
-#     + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
-#     + scale_fill_viridis_d(name=parse(text="dR[0]"))
-# )
-# }
-```

### 2020-12-09 codes/SIRfunctions.R

@@ -1,0 +1,90 @@
+# Functions for the SIR model
+# By: A Gharouni
+
+# install remotes package if necessary:
+while (!require(remotes)) {
+  install.packages("remotes")
+}
+## install development version of bbmle:
+if (!require("bbmle") || packageVersion("bbmle") < "1.0.23.5") {
+  remotes::install_github("bbolker/bbmle")
+}
+## install the target package and all its dependencies:
+remotes::install_github("bbolker/McMasterPandemic",
+                        dependencies = TRUE,
+                        build_vignettes = TRUE
+)
+
+
+# SIR model
+sir.model <- function(time,state,params){
+  unpack(as.list(c(state,params)))
+  ## Force of Infection  
+  Lambda <- beta * (I_u + eta_w*(I_n+I_p) + eta_c*I_t)/N0
+  ## scaling the weights
+  W <- W_S*S_u+W_I*I_u+W_R*R_u
+  sigma <- rho*N0/W
+  # testing intensity
+  F_S <- sigma*W_S
+  F_I <- sigma*W_I
+  F_R <- sigma*W_R
+  # Equations
+  dS_u.dt <- -Lambda*S_u - (1-p_S) * F_S * S_u + omega * S_n
+  dS_n.dt <- -Lambda*S_n + (1-p_S) * F_S * S_u - omega * S_n 
+  dI_u.dt <-  Lambda*S_u + omega * I_n - F_I * I_u - gamma * I_u
+  dI_n.dt <-  Lambda*S_n + (1-p_I) * F_I * I_u  - omega * I_n - gamma * I_n
+  dI_p.dt <-  p_I * F_I * I_u - omega * I_p - gamma * I_p 
+  dI_t.dt <-  omega * I_p - gamma * I_t 
+  dR_u.dt <- gamma * I_u + omega * R_n - F_R * R_u  
+  dR_n.dt <- gamma * I_n + (1-p_R) * F_R * R_u - omega * R_n
+  dR_p.dt <- gamma * I_p + p_R * F_R * R_u - omega * R_p
+  dR_t.dt <- gamma * I_t + omega * R_p
+  dN.dt <- omega * (S_n + I_n + R_n)
+  dP.dt <- omega *(I_p + R_p)
+  
+  # return the rate of change
[…]

### 2020-12-09 codes/ali_test_sir.Rmd
update

@@
 ```{r packages, message=FALSE}
-library(deSolve)
 library(ggplot2)
 library(tidyr)
-library(McMasterPandemic)
 library(directlabels)
@@
 library(rootSolve)
-unpack <- McMasterPandemic::unpack
 
@@
                 N=0,P=0)
-d <- seq(0,200,by=0.1)
+sim_time <- seq(0,200,by=0.1)
 ```
@@
 Sn <- Sn_dfe(params)
-
 state_dfe <- c(S_u=Su, S_n=Sn,
@@
 ```{r model_run, message=FALSE}
-out <- as.data.frame(
-  ode(
-    func=sir.model,
-    y=state_init,
-    times= d, 
-    parms=update(params,beta=2)
-    # atol = 1e+1, rtol = 1e+1
-    # rootfun = rootfun,
-    # method="lsodar"
-  )
-)
+out <- run.sir(model=sir.model,
+               params=update(params,beta=2),
+               state=state_init,
+               sim_time=sim_time)
 
@@
 ```{r run2, message=FALSE}
-d_long <- seq(0,300,by=0.1)
+sim_time_long <- seq(0,300,by=0.1)
 ## capturing badly behaved integrator warning messages
 cc <- capture.output(
-    out_long <- as.data.frame(
-        ode(
-    func=sir.model,
-    y=state_init,
-    times= d_long, 
-    parms=update(params,beta=2)
-    # rootfun = rootfun,
-    # method="lsodar"
+   out_long <- run.sir(model=sir.model,
+               params=update(params,beta=2),
+               state=state_init,
+               sim_time=sim_time_long)
 )
-  ))
+
 out_long2 <- out_long %>%
@@
 The Eigenvalues of $G_{11}$, are 0 and $R_0$.
-```{r R0}
-F_I <- function(params){
-   # Returns F_I at DFE 
-  unpack(as.list(params))
-  Sn<- Sn_dfe(params)
-  Su<- N0-Sn
-  sigma <- rho*N0/(W_S*Su) #at DFE 
-  return(sigma*W_I)
-}
 
- # just a test
-  F_I(params=update(params,rho=0))
-
-R0<-function(params){
-    unpack(as.list(params))
-    Sn <- Sn_dfe(params)
-    Su <- N0-Sn
-    # print(c(Su_dfe,Sn_dfe,Su_dfe+Sn_dfe))
-    Fi <- F_I(params) #at DFE
-    A <-gamma*(omega+gamma+eta_w*Fi) + omega*eta_c*p_I*Fi
-    B <-(omega+eta_w*(Fi+gamma))*gamma+(eta_w*gamma+eta_c*omega)*(omega*p_I*Fi)/(omega+gamma)
-    C <- beta/(N0*gamma*(gamma*(omega+gamma)+Fi*(gamma+omega*p_I)))
-    return((A*Su+B*Sn)*C)
-}
-
-# R0 expression with eta_w/eta_c factored
-R02<-function(params){
-    unpack(as.list(params))
-    Sn<- Sn_dfe(params)
-    Su<- N0-Sn
-    F_I <-F_I(params)  #at DFE
-
-    A2 <- gamma*(p_I*F_I*omega*Sn+(omega+gamma)*(gamma*Sn+F_I*N0))
-    B2 <- p_I*F_I*omega*(gamma*Su+omega*N0)
-    D <- gamma*(omega+gamma)*(gamma*Su+omega*N0)
-    C2 <- beta/((N0*gamma*(gamma*(omega+gamma)+F_I*(gamma+omega*p_I)))*(omega+gamma))
-    # s <- eta_w/eta_c
-    return((eta_c*(s*A2+B2)+D)*C2)
-}
+```{r R0, echo=FALSE}
+F_I(state=state_dfe,params=params)
 

### 2020-12-09 codes/SIRfunctions.R

@@ -15,6 +15,16 @@ remotes::install_github("bbolker/McMasterPandemic",
                         build_vignettes = TRUE
 )
 
+library(McMasterPandemic)
+unpack <- McMasterPandemic::unpack
+
+# Testing functions
+sigma <- function(state,params){
+  unpack(as.list(c(state,params)))
+  W <- W_S*S_u+W_I*I_u+W_R*R_u
+  testing_rate <- rho*N0/W
+  return(testing_rate)
+}
 
 # SIR model
 sir.model <- function(time,state,params){
@@ -22,8 +32,10 @@ sir.model <- function(time,state,params){
   ## Force of Infection  
   Lambda <- beta * (I_u + eta_w*(I_n+I_p) + eta_c*I_t)/N0
   ## scaling the weights
-  W <- W_S*S_u+W_I*I_u+W_R*R_u
-  sigma <- rho*N0/W
+  # W <- W_S*S_u+W_I*I_u+W_R*R_u 
+  # sigma <- rho*N0/W
+  sigma <- sigma(state,params)
+  
   # testing intensity
   F_S <- sigma*W_S
   F_I <- sigma*W_I
@@ -48,15 +60,16 @@ sir.model <- function(time,state,params){
   return(list(dxdt))
 }
 
-
+# TODO: fix the sigma the issue is when DFE(state,params), the multiroot() gives error.
 DFE <- function(S,params){
+  unpack(as.list(params))
   # S: Susceptibles
   S_u <- S[1]
   S_n <- S[2]
   I_u <-0
   R_u <-0
-  unpack(as.list(params))
-  W <- W_S*S_u+W_I*I_u+W_R*R_u #Weighted untested people
+    #Weighted untested people
+  W <- W_S*S_u+W_I*I_u+W_R*R_u
   sigma <- rho*N0/W
   F_S <- sigma*W_S
   
@@ -76,6 +89,60 @@ Su_dfe <-function(params){
   return(N0-Sn_dfe(params))
 }
 
+# Model simulation function
+run.sir <- function(model, params,state,sim_time){
+  # use update(params,beta=2)
+  library(deSolve)
+  unpack(as.list(c(state,params)))
+  out <- as.data.frame(
+    ode(
+    func=model,
+    y=state,
+    times= sim_time, 
[…]

### 2020-12-10 codes/ali_test_sir.Rmd
seperating functions

@@
 ```{r packages, message=FALSE}
-library(ggplot2)
-library(tidyr)
-library(directlabels)
-library(latex2exp)
-library(rootSolve)
-
 source("SIRfunctions.R")
@@
 R0(params=update(params,beta=2))
-
 # check
@@
 # Case #1: rho=0
-all.equal(R0(params=update(params,rho=0)), params[["beta"]]/params[["gamma"]])
+R0_base <-params[["beta"]]/params[["gamma"]]
+all.equal(R0(params=update(params,rho=0)), R0_base)
 # Case #2: 
-all.equal(R0(params=update(params,eta_c=1,eta_w=1)), params[["beta"]]/params[["gamma"]])
+all.equal(R0(params=update(params,eta_c=1,eta_w=1)), R0_base)
 ```

### 2020-12-10 codes/SIRfunctions.R

@@ -15,6 +15,11 @@ remotes::install_github("bbolker/McMasterPandemic",
                         build_vignettes = TRUE
 )
 
+library(ggplot2)
+library(tidyr)
+library(directlabels)
+library(latex2exp)
+library(rootSolve)
 library(McMasterPandemic)
 unpack <- McMasterPandemic::unpack
 
@@ -115,29 +120,29 @@ F_I <- function(state,params){
 }
 
 # Basic Reproduction Number
-R0<-function(params){
+R0<-function(state=state_dfe,params){
   unpack(as.list(c(state,params)))
   Sn <- Sn_dfe(params)
-  Su <- N0-Sn
-  # print(c(Su_dfe,Sn_dfe,Su_dfe+Sn_dfe))
-  Fi <- F_I(params) #at DFE
-  A <-gamma*(omega+gamma+eta_w*Fi) + omega*eta_c*p_I*Fi
-  B <-(omega+eta_w*(Fi+gamma))*gamma+(eta_w*gamma+eta_c*omega)*(omega*p_I*Fi)/(omega+gamma)
+  Su <- Su_dfe(params)
+  Fi <- F_I(state,params) #at DFE
+  
+  A <- gamma*(omega+gamma+eta_w*Fi) + omega*eta_c*p_I*Fi
+  B <- (omega+eta_w*(Fi+gamma))*gamma+(eta_w*gamma+eta_c*omega)*(omega*p_I*Fi)/(omega+gamma)
   C <- beta/(N0*gamma*(gamma*(omega+gamma)+Fi*(gamma+omega*p_I)))
   return((A*Su+B*Sn)*C)
 }
 
 # R0 expression with eta_w/eta_c factored
-R02<-function(params){
-  unpack(as.list(params))
-  Sn<- Sn_dfe(params)
-  Su<- N0-Sn
-  F_I <-F_I(params)  #at DFE
+R02 <- function(state=state_dfe,params){
+  unpack(as.list(c(state,params)))
+  Sn <- Sn_dfe(params)
+  Su <- Su_dfe(params)
+  Fi <- F_I(state,params) #at DFE
   
-  A2 <- gamma*(p_I*F_I*omega*Sn+(omega+gamma)*(gamma*Sn+F_I*N0))
-  B2 <- p_I*F_I*omega*(gamma*Su+omega*N0)
[…]

### 2020-12-10 codes/ali_test_sir.Rmd
modulate the code, functions, plotting

@@
 ```{r model inputs, message=FALSE}
-params <- c(N0=1000000, beta=.8,gamma=1/3, omega=0.25, rho=0.01, #0.8, 1/3
-             W_S=1, W_I=1, W_R=1, #weights
-             p_S=0, p_I=1, p_R=0.5, #prob of waiting for being tested positive
-            eta_w=0.02, eta_c=0.01, #isolation parameter 
-            s=2 #eta_w/eta_c
-            )
-class(params) <- "params_pansim"
-state_init <- c(S_u=params[["N0"]], S_n=0,
-                I_u=1,I_n=0,I_p=0,I_t=0,
-                R_u=0,R_n=0,R_p=0,R_t=0,
-                N=0,P=0)
-sim_time <- seq(0,200,by=0.1)
+source("params")
 ```
@@
 # DFE; analytic approach
-Su <- Su_dfe(params)
-Sn <- Sn_dfe(params)
-state_dfe <- c(S_u=Su, S_n=Sn,
-                I_u=0,I_n=0,I_p=0,I_t=0,
-                R_u=0,R_n=0,R_p=0,R_t=0,
-                N=0,P=0)
+# Su <- Su_dfe(params)
+# Sn <- Sn_dfe(params)
+# state_dfe <- c(S_u=Su, S_n=Sn,
+#                 I_u=0,I_n=0,I_p=0,I_t=0,
+#                 R_u=0,R_n=0,R_p=0,R_t=0,
+#                 N=0,P=0)
 # check:
@@
 ```{r model_run, message=FALSE}
+sim_time <- seq(0,200,by=0.1)
+
 out <- run.sir(model=sir.model,
@@
 ## ggsave(p1_temp,filename = "R0contour_random.pdf" , path="../pix/")
-ggsave(p1_temp,filename = "R0contour_TTI.pdf" , path="../pix/")
 

### 2020-12-10 codes/SIRfunctions.R

@@ -10,10 +10,13 @@ if (!require("bbmle") || packageVersion("bbmle") < "1.0.23.5") {
   remotes::install_github("bbolker/bbmle")
 }
 ## install the target package and all its dependencies:
-remotes::install_github("bbolker/McMasterPandemic",
-                        dependencies = TRUE,
-                        build_vignettes = TRUE
-)
+while (!require(McMasterPandemic)) {
+  remotes::install_github("bbolker/McMasterPandemic",
+                          dependencies = TRUE,
+                          build_vignettes = TRUE
+  )
+}
+
 
 library(ggplot2)
 library(tidyr)
@@ -147,7 +150,29 @@ R02 <- function(state=state_dfe,params){
   return((eta_c*(s*A2+B2)+D)*C2)
 }
 
-
+# make grid and calc R0 for plotting
+eval_R0 <- function(state=state_dfe,params,filename){
+  # input the params and their range, this function makes a grid dataframe, calls R0 and outputs the csv file
+  unpack(as.list(c(state,params)))
+  # specify the ranges, FIXME, not to be hard coded!
+  beta  <- 0.339  ## set beta high enough to allow R0>1 in worst case
+  eta_w <- seq(0,1, length.out=n_out)
+  eta_c <- seq(0,1, length.out=n_out) #0.001 to 0.5 
+  rho <- seq(0,0.01, length.out=n_in)
+  omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, was 0.011
+  
+  df1 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
+                     W_S=W_S,W_I=params[["W_I"]],W_R=params[["W_R"]],
+                     p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
+                     eta_w=eta_w,eta_c=eta_c)
+  df2<- data.frame(df1,
+                   R0=apply(df1,1,function(params_in)R0(params=params_in)),
+                   Fi= apply(df1,1,function(params_in)F_I(state=state, params=params_in)))
+  # This is for plotting purposes:
+  df2 <- df2 %>% dplyr::mutate(R0_sub = ifelse(eta_w<eta_c, NA, R0)) 
+  # Output
+  write.csv(df2,file=filename, row.names = FALSE)
+}
 
 
 

### 2020-12-10 codes/sir_plot.R

@@ -1,0 +1,61 @@
+
+source("SIRfunctions.R")
+source("params")
+
+# R0 contour plots
+n_out <- 5 ## facets (rows/cols)
+n_in <- 41 ## grid N within facets
+
+#choose weight W_S 0.3 for TTI testing plan and 1 for the random testing.
+W_S_random <- 1
+W_S_targeted <- 0.3 
+
+eval_R0(params = update(params,W_S=W_S_random),filename="random_test_df.csv")
+eval_R0(params = update(params,W_S=W_S_targeted),filename="targeted_test_df.csv")
+
+
+
+## hacked (not very robustly) to chain label_both() and label_parsed() ...
+label_special <- function (labels, multi_line=FALSE, sep = "== ") {
+  value <- ggplot2:::label_value(labels, multi_line = multi_line)
+  variable <- ggplot2:::label_variable(labels, multi_line = multi_line)
+  variable[[1]] <- gsub("_(.)$","[\\1]",variable[[1]])
+  ## not using multiple faceting variables on each margin, so forget paste
+  out <- Map(paste, variable, value, sep = sep)
+  out[[1]] <- lapply(out[[1]], function(expr) c(parse(text=expr)))
+  out
+}
+
+
+# show_contours_1, width=24,height=24,message=FALSE,warning=FALSE}
+brks_vec <- seq(0.8,1.05,by=0.05) # Break vector for unifying the legends in Random and TTI testing cases. 
+
+p1 <- ggplot(df2,aes(x=omega,y=rho,z=R0_sub))+ theme_bw() +
+  xlab(TeX('$\\omega$, Returning test result rate (1/day)')) +
+  ylab(TeX('$\\rho$, testing intensity (1/day per capita)')) +
+  theme(panel.spacing=grid::unit(0,"lines"),legend.position = c(0.89, 0.73)) # initial theme, comment it for TTI testing plot
+# theme(panel.spacing=grid::unit(0,"lines"),legend.position = "none") #AG, uncomment for Random Case
+p1_temp <- (p1
+            + geom_contour_filled(breaks=brks_vec)
+            + geom_contour(breaks=1,alpha=0.5,colour="black")
+            # + scale_fill_manual(palette=mycolors(brks_vec),drop=FALSE)
+            + facet_grid(eta_w~eta_c, labeller=label_special)
+            + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
[…]

### 2020-12-11 codes/sir_plot.R
plot single panle

@@ -2,6 +2,9 @@
 source("SIRfunctions.R")
 source("params")
 
+# #################################
+# Making Dataframe part
+# #################################
 # R0 contour plots
 n_out <- 5 ## facets (rows/cols)
 n_in <- 41 ## grid N within facets
@@ -10,11 +13,20 @@ n_in <- 41 ## grid N within facets
 W_S_random <- 1
 W_S_targeted <- 0.3 
 
+# make dataframe
 eval_R0(params = update(params,W_S=W_S_random),filename="random_test_df.csv")
 eval_R0(params = update(params,W_S=W_S_targeted),filename="targeted_test_df.csv")
 
+# #################################
+# Load the data
+# #################################
+df_random <- read.csv(file = 'random_test_df.csv')
+df_targeted <- read.csv(file = 'targeted_test_df.csv')
 
-
+# #################################
+# Plotting Part
+# #################################
+# Setting the plots
 ## hacked (not very robustly) to chain label_both() and label_parsed() ...
 label_special <- function (labels, multi_line=FALSE, sep = "== ") {
   value <- ggplot2:::label_value(labels, multi_line = multi_line)
@@ -26,25 +38,31 @@ label_special <- function (labels, multi_line=FALSE, sep = "== ") {
   out
 }
 
+# verify the break range in brks_vec (FIXME)
+mn <- min(df_random$R0_sub,df_targeted$R0_sub, na.rm = T)
+mx <- max(df_random$R0_sub,df_targeted$R0_sub, na.rm = T)
 
 # show_contours_1, width=24,height=24,message=FALSE,warning=FALSE}
 brks_vec <- seq(0.8,1.05,by=0.05) # Break vector for unifying the legends in Random and TTI testing cases. 
 
-p1 <- ggplot(df2,aes(x=omega,y=rho,z=R0_sub))+ theme_bw() +
+# #################################
+# 1. Plot the Random Testing Scenario:
+# #################################
+df_temp <- df_random
+p1 <- ggplot(df_temp,aes(x=omega,y=rho,z=R0_sub))+ theme_bw() +
   xlab(TeX('$\\omega$, Returning test result rate (1/day)')) +
   ylab(TeX('$\\rho$, testing intensity (1/day per capita)')) +
[…]

### 2020-12-12 codes/plots4talk.R
plotting stuff

@@ -1,0 +1,103 @@
+# Making plots for presentations/slides etc in the "path"
+path <- "~/projects/aliworkstation/references/presentations/epimodeling/pix/"
+# R code is : ~/projects/SIRmodel/codes/
+# setwd(path)
+
+source("SIRfunctions.R")
+source("params")
+
+# #################################
+# Load the data
+# #################################
+df_random <- read.csv(file = 'random_test_df.csv') #random testing scenario
+df_targeted <- read.csv(file = 'targeted_test_df.csv') #targeted testing scenario
+#choose weight W_S 0.3 for TTI testing plan and 1 for the random testing.
+W_S_random <- unique(df_random$W_S)
+W_S_targeted <- unique(df_targeted$W_S)
+
+# #################################
+# Plotting Part
+# #################################
+# Setting the plots
+## hacked (not very robustly) to chain label_both() and label_parsed() ...
+label_special <- function (labels, multi_line=FALSE, sep = "== ") {
+  value <- ggplot2:::label_value(labels, multi_line = multi_line)
+  variable <- ggplot2:::label_variable(labels, multi_line = multi_line)
+  variable[[1]] <- gsub("_(.)$","[\\1]",variable[[1]])
+  ## not using multiple faceting variables on each margin, so forget paste
+  out <- Map(paste, variable, value, sep = sep)
+  out[[1]] <- lapply(out[[1]], function(expr) c(parse(text=expr)))
+  out
+}
+
+# verify the break range in brks_vec (FIXME)
+mn <- min(df_random$R0_sub,df_targeted$R0_sub, na.rm = T)
+mx <- max(df_random$R0_sub,df_targeted$R0_sub, na.rm = T)
+
+# show_contours_1, width=24,height=24,message=FALSE,warning=FALSE}
+brks_vec <- seq(0.8,1.05,by=0.05) # Break vector for unifying the legends in Random and TTI testing cases. 
+
+# #################################
+#  SubPlots the Random Testing Scenario:
+# #################################
+# Random Testing, Perfect isolation
+
+# FIXME: df <- paste("df",scenario,sep="_")
+# scenario <-"random"
+# df <- df_random
+scenario <-"targeted"
+df <- df_targeted
+
+# isolation for awaiting people
+eta_w_val <-0
[…]

### 2020-12-12 codes/sir_plot.R

@@ -107,29 +107,4 @@ ggsave(p2_temp,filename = "R0contour_TTI.pdf" ,
        width = 14, height = 14, units = "cm",
        path="../pix/")
 
-# #################################
-# 3. SubPlots the Random Testing Scenario:
-# #################################
 
-df_temp <- df_random %>% dplyr::filter(eta_w==0 & eta_c==0) 
-
-p1 <- ggplot(df_temp,aes(x=omega,y=rho,z=R0_sub))+ theme_bw() +
-  xlab(TeX('$\\omega$, Returning test result rate (1/day)')) +
-  ylab(TeX('$\\rho$, testing intensity (1/day per capita)')) +
-  theme(panel.spacing=grid::unit(0,"lines"),legend.position = "right") 
-p1_temp <- (p1
-            + geom_contour_filled(breaks=brks_vec)
-            + geom_contour(breaks=1,alpha=0.5,colour="black")
-            + facet_grid(eta_w~eta_c, labeller=label_special)
-            + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
-            + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
-            + scale_fill_viridis_d(name=parse(text="R[0]"),drop=FALSE)
-            + ggtitle(TeX(sprintf("W_S=W_I=W_R=1")))
-            + geom_rect(data=df_temp, fill=ifelse(df_temp$eta_c > df_temp$eta_w,"grey90","NA" ),
-                        color= NA,
-                        ymin=-1,
-                        ymax=10,
-                        xmin=-1,
-                        xmax=2)
-)
-p1_temp

### 2020-12-15 codes/ali_test_sir.Rmd
seperate code and plottings

@@
 
-# DFE; analytic approach
-# Su <- Su_dfe(params)
-# Sn <- Sn_dfe(params)
-# state_dfe <- c(S_u=Su, S_n=Sn,
-#                 I_u=0,I_n=0,I_p=0,I_t=0,
-#                 R_u=0,R_n=0,R_p=0,R_t=0,
-#                 N=0,P=0)
-# check:
+# check if the DFE is the same in numerical and analytic:
 all.equal(c(state_dfe[["S_u"]],state_dfe[["S_n"]]),roots_dfe$root)
@@
 gg1 <- (ggplot(data=out2, aes(x=time,y=value,col=compartment))
+            + theme_bw()
             + geom_line()
+            + ylab(label="Prevalence")
+            + xlab(label="Time (days)")
             + scale_y_log10(limits=c(0.1,params[["N0"]]))
-##    + scale_x_continuous(limits=c(0,240))
+            + xlim(0,310)
 )
@@
 ```{r plot_long, warning=FALSE}
-direct.label(gg1 %+% out_long2,"last.bumpup")
+p_temp <- direct.label(gg1 %+% out_long2,"last.bumpup")
+p_temp
+ggsave(p_temp,filename = "sir_sing.pdf" ,
+       width = 14, height = 14, units = "cm",
+       path="../pix/")
 ```
@@
 ### Contour plot of $R_0=1$ and interpretations approache 1
-```{r contour_R0_1, message=FALSE,warning=FALSE }
-n_out <- 5 ## facets (rows/cols)
-n_in <- 41 ## grid N within facets
- # specify the ranges
- beta  <- 0.339  ## set beta high enough to allow R0>1 in worst case
- eta_w <- seq(0,1, length.out=n_out)
- eta_c <- seq(0,1, length.out=n_out) #0.001 to 0.5 
- rho     <- seq(0,0.01, length.out=n_in)
- omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, was 0.011
- W_S <- 0.3 #choose 0.3 for TTI testing plan and 1 for the random testing.
-
- df1 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
-                    W_S=W_S,W_I=params[["W_I"]],W_R=params[["W_R"]],
-                    p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
-                    eta_w=eta_w,eta_c=eta_c,s=params[["s"]])
-
-# Need to be corrected: in F_I, state_dfe should be changed according to the df1!
-df2<- data.frame(df1,R0=apply(df1,1,function(params_in)R0(params=params_in)),
-                      Fi= apply(df1,1,function(params_in)F_I(params=params_in)))
-
-df2 <- df2 %>% dplyr::mutate(R0_sub = ifelse(eta_w<eta_c, NA, R0))  
-```
-
-```{r}
-## hacked (not very robustly) to chain label_both() and label_parsed() ...
-label_special <- function (labels, multi_line=FALSE, sep = "== ") {
-    value <- ggplot2:::label_value(labels, multi_line = multi_line)
-    variable <- ggplot2:::label_variable(labels, multi_line = multi_line)
-    variable[[1]] <- gsub("_(.)$","[\\1]",variable[[1]])
-    ## not using multiple faceting variables on each margin, so forget paste
-    out <- Map(paste, variable, value, sep = sep)
-    out[[1]] <- lapply(out[[1]], function(expr) c(parse(text=expr)))
-    out
-}
-```
-
-```{r show_contours_1, width=24,height=24,message=FALSE,warning=FALSE}
-brks_vec <- seq(0.8,1.05,by=0.05) # Break vector for unifying the legends in Random and TTI testing cases.
-
-p1 <- ggplot(df2,aes(x=omega,y=rho,z=R0_sub))+ theme_bw() +
-  xlab(TeX('$\\omega$, Returning test result rate (1/day)')) +
-  ylab(TeX('$\\rho$, testing intensity (1/day per capita)')) +
-  theme(panel.spacing=grid::unit(0,"lines"),legend.position = c(0.89, 0.73)) # initial theme, comment it for TTI testing plot
-  # theme(panel.spacing=grid::unit(0,"lines"),legend.position = "none") #AG, uncomment for Random Case
-p1_temp <- (p1
-    + geom_contour_filled(breaks=brks_vec)
-    + geom_contour(breaks=1,alpha=0.5,colour="black")
-    # + scale_fill_manual(palette=mycolors(brks_vec),drop=FALSE)
-    + facet_grid(eta_w~eta_c, labeller=label_special)
-    + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
-    + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
-    + scale_fill_viridis_d(name=parse(text="R[0]"),drop=FALSE)
-    + ggtitle(TeX(sprintf("W_S=%.1f, W_I=W_R=1",W_S)))
-    + geom_rect(data=df2, fill=ifelse(df2$eta_c>df2$eta_w,"grey90","NA" ),
-                          color= NA,
-                          ymin=-1,
-                          ymax=10,
-                          xmin=-1,
-                          xmax=2)
-    )
-p1_temp
-
-## BMB: FIXME: probably wrong but ... ???
-## ggsave(p1_temp,filename = "R0contour_random.pdf" , path="../pix/")
-
-```
-
-### Contour plot of $R_0=1$ and interpretations approache 2, parametrization $s=\eta_w/\eta_c$
-
-```{r contour_R0_2, message=FALSE,warning=FALSE }
-# This part is done to carry our simulation when s=eta_c/eta_w is involved. Also to fill the upper triangle panels in the panel plot. 
-#  # specify the ranges
-# beta  <- 0.339  ## set beta high enough to allow R0>1 in worst case 
-# rho     <- seq(0,0.01, length.out=n_in)
-# omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, 
-# eta_c <- seq(0.001,0.1, length.out=n_out)
-# s <- seq(1,10,length.out=n_out) 
-# 
-#  df11 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
-#                     W_S=W_S,W_I=params[["W_I"]],W_R=params[["W_R"]],
-#                     p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
-#                     eta_w=params[["eta_w"]],eta_c=eta_c,s=s)
-# 
-# df22<- data.frame(df11,R0=apply(df11,1,function(params_in)R02(params=params_in)),
-#                       Fi= apply(df11,1,function(params_in)F_I(params=params_in)),
-#                        Sn=apply(df11,1,function(params_in)Sn_dfe(params = params_in)))
-
-```
-
-```{r show_contours_2, width=24,height=24,message=FALSE,warning=FALSE}
-# p11 <- ggplot(df22,aes(x=omega,y=rho,z=R0))+ theme_bw() +
-#     theme(panel.spacing=grid::unit(0,"lines"))
-# p2_temp <- (p11
-#     + geom_contour_filled(breaks=pretty(unique(df22$R0),5),
-#                           alpha=0.5) ## breaks=0.2)
-#     + geom_contour(breaks=1,alpha=0.5,colour="black")
-#     + facet_grid(s~eta_c, labeller=label_special)
-#     + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
-#     + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
-#     + scale_fill_viridis_d(name=parse(text="R[0]"),breaks=brks_vec)
-#     + ggtitle("W[S]=1, W[I]=W[R]=1")
-# )
-# p2_temp
-# ## ggsave(p2_temp,filename = "R0contour_random_s.png" , path="../pix/")
-# # ggsave(p2_temp,filename = "R0contour_random_s.pdf" , path="../pix/")
-```
+ <!-- Run sir.plot.R -->
 
@@
 3. When TTI regime is applied,
-
-```{r  chnages of R0 wrt omega,message=FALSE,warning=FALSE }
-
-```

### 2020-12-15 codes/sir_plot.R

@@ -13,9 +13,10 @@ n_in <- 41 ## grid N within facets
 W_S_random <- 1
 W_S_targeted <- 0.3 
 
-# make dataframe
-eval_R0(params = update(params,W_S=W_S_random),filename="random_test_df.csv")
-eval_R0(params = update(params,W_S=W_S_targeted),filename="targeted_test_df.csv")
+# make dataframe and save it. 
+# FIXME: I feel that this part needs to be separated from here?
+# eval_R0(params = update(params,W_S=W_S_random),filename="random_test_df.csv")
+# eval_R0(params = update(params,W_S=W_S_targeted),filename="targeted_test_df.csv")
 
 # #################################
 # Load the data

### 2020-12-15 codes/sir_plot.R
tweak plotting code

@@ -1,6 +1,6 @@
 
 source("SIRfunctions.R")
-source("params")
+source("params.R")
 
 # #################################
 # Making Dataframe part
@@ -46,9 +46,6 @@ mx <- max(df_random$R0_sub,df_targeted$R0_sub, na.rm = T)
 # show_contours_1, width=24,height=24,message=FALSE,warning=FALSE}
 brks_vec <- seq(0.8,1.05,by=0.05) # Break vector for unifying the legends in Random and TTI testing cases. 
 
-# #################################
-# 1. Plot the Random Testing Scenario:
-# #################################
 df_temp <- df_random
 p1 <- ggplot(df_temp,aes(x=omega,y=rho,z=R0_sub))+ theme_bw() +
   xlab(TeX('$\\omega$, Returning test result rate (1/day)')) +
@@ -62,7 +59,6 @@ p1_temp <- (p1
             + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
             + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
             + scale_fill_viridis_d(name=parse(text="R[0]"),drop=FALSE)
-            + ggtitle(TeX(sprintf("W_S=W_I=W_R=1")))
             + geom_rect(data=df_temp, fill=ifelse(df_temp$eta_c > df_temp$eta_w,"grey90","NA" ),
                         color= NA,
                         ymin=-1,
@@ -70,42 +66,27 @@ p1_temp <- (p1
                         xmin=-1,
                         xmax=2)
 )
-p1_temp
 
-ggsave(p1_temp,filename = "R0contour_random.pdf" ,
+# #################################
+# 1. Plot the Random Testing Scenario:
+# #################################
+
+ggsave(p1_temp + ggtitle(TeX("W_S=W_I=W_R=1")),
+       filename = "R0contour_random.pdf" ,
        width = 14, height = 14, units = "cm",
        path="../pix/")
 
-
 # #################################
-# 2. Plot the Random Testing Scenario:
+# 2. Plot the Targeted Testing Scenario:
 # #################################
-df_temp <- df_targeted
 
-p2 <- ggplot(df_temp,aes(x=omega,y=rho,z=R0_sub))+ theme_bw() +
-  xlab(TeX('$\\omega$, Returning test result rate (1/day)')) +
-  ylab(TeX('$\\rho$, testing intensity (1/day per capita)')) +
[…]

### 2020-12-16 codes/sir_macpan.tex
update the plots for AARMS slides

\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
% maxwidth is the original width if it is less than linewidth
% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
By Ali Gharouni,
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{verbatim}
## [1] "2020-12-13 08:16:24 EST"
\end{verbatim}
\end{kframe}
\end{knitrout}
This is my a test for MacPan package.

\section{Install and Play}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# install remotes package if necessary:}
\hlkwa{while} \hlstd{(}\hlopt{!}\hlkwd{require}\hlstd{(remotes)) \{}
    \hlkwd{install.packages}\hlstd{(}\hlstr{"remotes"}\hlstd{)}
\hlstd{\}}
\end{alltt}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Loading required package: remotes}}\begin{alltt}
\hlcom{## install development version of bbmle:}
\hlkwa{if} \hlstd{(}\hlopt{!}\hlkwd{require}\hlstd{(}\hlstr{"bbmle"}\hlstd{)} \hlopt{||} \hlkwd{packageVersion}\hlstd{(}\hlstr{"bbmle"}\hlstd{)} \hlopt{<} \hlstr{"1.0.23.5"}\hlstd{) \{}
    \hlstd{remotes}\hlopt{::}\hlkwd{install_github}\hlstd{(}\hlstr{"bbolker/bbmle"}\hlstd{)}
\hlstd{\}}
\end{alltt}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Loading required package: bbmle}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Loading required package: stats4}}\begin{alltt}
\hlcom{## install the target package and all its dependencies:}
\hlstd{remotes}\hlopt{::}\hlkwd{install_github}\hlstd{(}\hlstr{"bbolker/McMasterPandemic"}\hlstd{,}
                        \hlkwc{dependencies} \hlstd{=} \hlnum{TRUE}\hlstd{,}
                        \hlkwc{build_vignettes} \hlstd{=} \hlnum{TRUE}
\hlstd{)}
\end{alltt}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Skipping install of 'McMasterPandemic' from a github remote, the SHA1 (54095dc4) has not changed since last install.\\\#\#\ \  Use `force = TRUE` to force installation}}\begin{alltt}
\hlkwd{library}\hlstd{(McMasterPandemic)}
\hlkwd{library}\hlstd{(ggplot2);} \hlkwd{theme_set}\hlstd{(}\hlkwd{theme_bw}\hlstd{())}
\hlkwd{library}\hlstd{(cowplot)}

\hlkwd{library}\hlstd{(directlabels)}
\hlkwd{library}\hlstd{(zoo)}
\end{alltt}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# \\\#\# Attaching package: 'zoo'}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# The following objects are masked from 'package:base':\\\#\# \\\#\#\ \ \ \  as.Date, as.Date.numeric}}\begin{alltt}
\hlkwd{library}\hlstd{(tidyverse)}
\end{alltt}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Attaching packages --------------------------------------- tidyverse 1.3.0 --}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# v tibble\ \ 3.0.4\ \ \ \  v dplyr\ \  1.0.2\\\#\# v tidyr\ \  1.1.2\ \ \ \  v stringr 1.4.0\\\#\# v readr\ \  1.3.1\ \ \ \  v forcats 0.5.0\\\#\# v purrr\ \  0.3.4}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# -- Conflicts ------------------------------------------ tidyverse\_conflicts() --\\\#\# x dplyr::filter() masks stats::filter()\\\#\# x dplyr::lag()\ \ \ \ masks stats::lag()\\\#\# x dplyr::slice()\ \ masks bbmle::slice()\\\#\# x tidyr::unpack() masks McMasterPandemic::unpack()}}\end{kframe}
\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{params} \hlkwb{<-} \hlkwd{read_params}\hlstd{(}\hlstr{"ICU1.csv"}\hlstd{)}

\hlstd{knitr}\hlopt{::}\hlkwd{kable}\hlstd{(}\hlkwd{round}\hlstd{(}\hlkwd{t}\hlstd{(}\hlkwd{summary}\hlstd{(params)),}\hlnum{2}\hlstd{))}
\end{alltt}
\end{kframe}
\begin{tabular}{r|r|r|r|r}
\hline
r0 & R0 & Gbar & CFR\_gen & dbl\_time\\
\hline
0.23 & 6.52 & 12.19 & 0.04 & 3.04\\
\hline
\end{tabular}

\begin{kframe}\begin{alltt}
\hlstd{knitr}\hlopt{::}\hlkwd{kable}\hlstd{(}\hlkwd{round}\hlstd{(}\hlkwd{t}\hlstd{(}\hlkwd{get_R0}\hlstd{(params,} \hlkwc{components}\hlstd{=}\hlnum{TRUE}\hlstd{)),}\hlnum{2}\hlstd{))}
\end{alltt}
\end{kframe}
\begin{tabular}{r|r|r|r}
\hline
asymptomatic & pre-symptomatic & mild & severe\\
\hline
1.56 & 0.33 & 4.46 & 0.17\\
\hline
\end{tabular}

\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# A simple SIR model}
\hlkwd{library}\hlstd{(deSolve)}
\hlstd{unpack} \hlkwb{<-} \hlstd{McMasterPandemic}\hlopt{::}\hlstd{unpack}

\hlstd{sir.mod} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{time}\hlstd{,}\hlkwc{state}\hlstd{,}\hlkwc{params}\hlstd{)\{}
    \hlkwd{unpack}\hlstd{(}\hlkwd{as.list}\hlstd{(}\hlkwd{c}\hlstd{(state,params)))}
    \hlstd{dS.dt} \hlkwb{<-} \hlopt{-}\hlstd{beta}\hlopt{*}\hlstd{S}\hlopt{*}\hlstd{I}\hlopt{/}\hlstd{N0}
    \hlstd{dI.dt} \hlkwb{<-}  \hlstd{beta}\hlopt{*}\hlstd{S}\hlopt{*}\hlstd{I}\hlopt{/}\hlstd{N0}\hlopt{-}\hlstd{gamma}\hlopt{*}\hlstd{I}
    \hlstd{dR.dt} \hlkwb{<-}  \hlstd{gamma}\hlopt{*}\hlstd{I}
    \hlcom{# return the rate of change}
    \hlstd{dxdt} \hlkwb{<-} \hlkwd{c}\hlstd{(dS.dt,dI.dt,dR.dt)}
    \hlkwd{return}\hlstd{(}\hlkwd{list}\hlstd{(dxdt))}
\hlstd{\}}
\hlstd{sir.params} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlkwc{N0}\hlstd{=}\hlnum{10}\hlopt{^}\hlnum{6}\hlstd{,} \hlkwc{beta}\hlstd{=}\hlnum{0.5}\hlstd{,}\hlkwc{gamma}\hlstd{=}\hlnum{1}\hlopt{/}\hlnum{7}\hlstd{)}
\hlkwd{class}\hlstd{(sir.params)} \hlkwb{<-} \hlstr{"params_pansim"}
\hlcom{# Set the initial state}
\hlstd{sir.state_init} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlkwc{S}\hlstd{=sir.params[[}\hlstr{"N0"}\hlstd{]],}
                \hlkwc{I}\hlstd{=}\hlnum{1}\hlstd{,}
                \hlkwc{R}\hlstd{=}\hlnum{0}\hlstd{)}
\hlcom{##set the number of days for SIR simulation as simulation time of MacPan}
\hlstd{startdate} \hlkwb{<-} \hlkwd{as.Date}\hlstd{(}\hlstr{"01-03-2020"}\hlstd{)}
\hlstd{enddate} \hlkwb{<-} \hlkwd{as.Date}\hlstd{(}\hlstr{"01-09-2020"}\hlstd{)}
\hlstd{d} \hlkwb{<-} \hlkwd{seq}\hlstd{(}\hlnum{0}\hlstd{,}\hlkwd{as.numeric}\hlstd{(enddate}\hlopt{-}\hlstd{startdate),}\hlkwc{by}\hlstd{=}\hlnum{1}\hlstd{)}

\hlstd{sir.R0} \hlkwb{<-} \hlstd{sir.params[[}\hlstr{"beta"}\hlstd{]]}\hlopt{/}\hlstd{sir.params[[}\hlstr{"gamma"}\hlstd{]]}

\hlstd{sir.out} \hlkwb{<-} \hlkwd{as.data.frame}\hlstd{(}
  \hlkwd{ode}\hlstd{(}
    \hlkwc{func}\hlstd{=sir.mod,}
    \hlkwc{y}\hlstd{=sir.state_init,}
    \hlkwc{times}\hlstd{= d,}
    \hlkwc{parms}\hlstd{=}\hlkwd{update}\hlstd{(sir.params)}
    \hlstd{))}

\hlstd{sir.out2} \hlkwb{<-} \hlstd{sir.out} \hlopt{%>%}
  \hlkwd{pivot_longer}\hlstd{(}\hlkwd{c}\hlstd{(S,I,R),} \hlkwc{names_to} \hlstd{=} \hlstr{"compartment"}\hlstd{,} \hlkwc{values_to} \hlstd{=} \hlstr{"value"}\hlstd{)}

\hlstd{gg1} \hlkwb{<-} \hlstd{(}\hlkwd{ggplot}\hlstd{(}\hlkwc{data}\hlstd{=sir.out2,} \hlkwd{aes}\hlstd{(}\hlkwc{x}\hlstd{=time,}\hlkwc{y}\hlstd{=value,}\hlkwc{col}\hlstd{=compartment))}
    \hlopt{+} \hlkwd{geom_line}\hlstd{()}
    \hlcom{# + scale_y_log10(limits=c(0.1,sir.params[["N0"]]))}
\hlstd{)}
\hlkwd{direct.label}\hlstd{(gg1,}\hlstr{"last.bumpup"}\hlstd{)}
\end{alltt}
\end{kframe}
\includegraphics[width=\maxwidth]{figure/Ali:SIR_-1}

\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{## update MacPan's params according to the SIR params}
\hlcom{##}
\hlstd{big_rate} \hlkwb{<-} \hlnum{20}
\hlstd{params} \hlkwb{<-} \hlkwd{update}\hlstd{(params,} \hlkwd{c}\hlstd{(}\hlkwc{beta0}\hlstd{=sir.params[[}\hlstr{"beta"}\hlstd{]],}
                           \hlkwc{alpha}\hlstd{=}\hlnum{0}\hlstd{,} \hlcom{## no asymptomatic}
                           \hlkwc{sigma}\hlstd{=big_rate,} \hlcom{## very short latent period}
                           \hlcom{## would like this to -> Inf but can't}
                           \hlkwc{gamma_a}\hlstd{=}\hlnum{0}\hlstd{,}
                           \hlkwc{mu}\hlstd{=}\hlnum{1}\hlstd{,}  \hlcom{## all cases mild}
                           \hlkwc{gamma_m}\hlstd{=sir.params[[}\hlstr{"gamma"}\hlstd{]],}
                           \hlkwc{gamma_s}\hlstd{=}\hlnum{0}\hlstd{,}
                           \hlcom{## try to skip presymptomatic period}
                           \hlkwc{gamma_p}\hlstd{=big_rate}
                           \hlstd{))}

\hlstd{pp} \hlkwb{<-} \hlstd{params}
\hlcom{## pp <- fix_pars(params, target = c(R0 = sir.R0)) #not sure if I need to update Gbar=6?}
\hlstd{state} \hlkwb{<-} \hlkwd{make_state}\hlstd{(}\hlkwc{params}\hlstd{=pp)}
\hlstd{state[]} \hlkwb{<-} \hlnum{0}
\hlstd{state[[}\hlstr{"S"}\hlstd{]]} \hlkwb{<-} \hlnum{1e6}
\hlstd{state[[}\hlstr{"Im"}\hlstd{]]} \hlkwb{<-} \hlnum{1}

\hlkwd{summary}\hlstd{(pp)}
\end{alltt}

{\ttfamily\noindent\color{warningcolor}{\#\# Warning in log(r\_last/r\_nextlast): NaNs produced}}\begin{verbatim}
##       r0       R0     Gbar  CFR_gen dbl_time 
##      NaN      NaN      NaN        0      NaN
\end{verbatim}
\begin{alltt}
\hlstd{sim0} \hlkwb{<-} \hlkwd{run_sim}\hlstd{(pp,state,}
                \hlkwc{start_date}\hlstd{=startdate,}
                \hlkwc{end_date}\hlstd{=enddate,}
                \hlkwc{use_ode}\hlstd{=}\hlnum{TRUE}\hlstd{)}

\hlstd{gg0} \hlkwb{<-} \hlstd{(}\hlkwd{ggplot}\hlstd{(sim0,}\hlkwd{aes}\hlstd{(}\hlkwc{x}\hlstd{=date))}
\hlopt{+} \hlkwd{geom_point}\hlstd{(}\hlkwd{aes}\hlstd{(}\hlkwc{y}\hlstd{=incidence))}
\hlstd{)}
\hlkwd{print}\hlstd{(gg0)}
\end{alltt}
\end{kframe}
\includegraphics[width=\maxwidth]{figure/Try_MacPan_to_recover_the_SIR_result-1} 
\begin{kframe}\begin{alltt}
\hlcom{# combining the SIR and MacPan results also calculating incidences fr the SIR model}
\hlstd{sim_combined} \hlkwb{<-} \hlstd{sim0} \hlopt{%>%} \hlkwd{mutate}\hlstd{(}\hlkwc{sir.S}\hlstd{=sir.out[[}\hlstr{"S"}\hlstd{]],}
                                \hlkwc{sir.I}\hlstd{=sir.out[[}\hlstr{"I"}\hlstd{]],}
                                \hlkwc{sir.S}\hlstd{=sir.out[[}\hlstr{"R"}\hlstd{]],}
                                \hlkwc{sir.incidence}\hlstd{=sir.params[[}\hlstr{"beta"}\hlstd{]]}\hlopt{*}\hlstd{sir.out[[}\hlstr{"S"}\hlstd{]]}\hlopt{*}\hlstd{sir.out[[}\hlstr{"I"}\hlstd{]])}

\hlstd{gg01} \hlkwb{<-} \hlstd{(}\hlkwd{ggplot}\hlstd{(sim_combined,}\hlkwd{aes}\hlstd{(}\hlkwc{x}\hlstd{=date))}
\hlopt{+} \hlkwd{geom_point}\hlstd{(}\hlkwd{aes}\hlstd{(}\hlkwc{y}\hlstd{=I))}
\hlopt{+} \hlkwd{geom_line}\hlstd{(}\hlkwd{aes}\hlstd{(}\hlkwc{y}\hlstd{=sir.I),} \hlkwc{color}\hlstd{=}\hlstr{'red'}\hlstd{)}
\hlstd{)}
\hlkwd{print}\hlstd{(gg01)}
\end{alltt}
\end{kframe}
\includegraphics[width=\maxwidth]{figure/Try_MacPan_to_recover_the_SIR_result-2} 
\begin{kframe}\begin{alltt}
\hlkwd{print}\hlstd{(gg01} \hlopt{+} \hlkwd{scale_y_log10}\hlstd{())}
\end{alltt}
\end{kframe}
\includegraphics[width=\maxwidth]{figure/Try_MacPan_to_recover_the_SIR_result-3}

\end{knitrout}

\end{document}

### 2020-12-16 codes/plots4talk.R

@@ -1,10 +1,10 @@
 # Making plots for presentations/slides etc in the "path"
 path <- "~/projects/aliworkstation/references/presentations/epimodeling/pix/"
-# R code is : ~/projects/SIRmodel/codes/
-# setwd(path)
+# R code is at: ~/projects/SIRmodel/codes/
+# setwd("~/projects/SIRmodel/codes/")
 
 source("SIRfunctions.R")
-source("params")
+source("params.R")
 
 # #################################
 # Load the data
@@ -63,8 +63,8 @@ title_set <- ifelse(scenario=="targeted",
                     TeX(sprintf("Targeted Testing; W_S=%.1f, W_I=W_R=1",W_S_targeted )),
                     TeX(sprintf("Random Testing; W_S=W_I=W_R=1")))
   
-p1 <- ggplot(df_temp,aes(x=omega,y=rho,z=R0_sub))+ theme_bw(base_size = 14) +
-  xlab(TeX('$\\omega$, Returning test result rate (1/day)')) +
+p1 <- ggplot(df_temp,aes(x=omega,y=rho,z=R0_sub))+ theme_bw(base_size = 12) +
+  xlab(TeX('$\\omega$, rate of test return  (1/day)')) +
   ylab(TeX('$\\rho$, testing intensity (1/day per capita)')) +
   theme(panel.spacing=grid::unit(0,"lines"),legend.position = leg_pos)
          
@@ -90,13 +90,14 @@ p1_temp
 
 ggsave(p1_temp,
        filename = paste("R0_",scenario,"_","w",eta_w_val,"c",eta_c_val,".pdf",sep = "") ,
-       width = 14, height = 14, units = "cm",
+       width = 10, height = 10, units = "cm",
        path=path)
 
-# ggsave(p1_temp,
-#        filename = paste("R0_",scenario,"_","w","all","c",eta_c_val,".pdf",sep = "") ,
-#        width = 14, height = 14, units = "cm",
-#        path=path)
+# Saving the column panels
+ggsave(p1_temp,
+       filename = paste("R0_",scenario,"_","w","all","c",eta_c_val,".pdf",sep = "") ,
+       width = 12, height = 12, units = "cm",
+       path=path)
 
 
 

### 2020-12-16 codes/sir_plot.R

@@ -1,7 +1,8 @@
 
 source("SIRfunctions.R")
 source("params.R")
-
+# where to save the output:
+# path="../pix/"
 # #################################
 # Making Dataframe part
 # #################################
@@ -47,10 +48,11 @@ mx <- max(df_random$R0_sub,df_targeted$R0_sub, na.rm = T)
 brks_vec <- seq(0.8,1.05,by=0.05) # Break vector for unifying the legends in Random and TTI testing cases. 
 
 df_temp <- df_random
+# df_temp <- df_targeted
 p1 <- ggplot(df_temp,aes(x=omega,y=rho,z=R0_sub))+ theme_bw() +
-  xlab(TeX('$\\omega$, Returning test result rate (1/day)')) +
+  xlab(TeX('$\\omega$, rate of test return (1/day)')) +
   ylab(TeX('$\\rho$, testing intensity (1/day per capita)')) +
-  # theme(panel.spacing=grid::unit(0,"lines"),legend.position = c(0.89, 0.73)) # initial theme, comment it for TTI testing plot
+  # theme(panel.spacing=grid::unit(0,"lines"),legend.position = c(0.8, 0.75)) # initial theme, comment it for TTI testing plot
   theme(panel.spacing=grid::unit(0,"lines"),legend.position = "none") #AG, uncomment for Random Case
 p1_temp <- (p1
             + geom_contour_filled(breaks=brks_vec)
@@ -73,8 +75,8 @@ p1_temp <- (p1
 
 ggsave(p1_temp + ggtitle(TeX("W_S=W_I=W_R=1")),
        filename = "R0contour_random.pdf" ,
-       width = 14, height = 14, units = "cm",
-       path="../pix/")
+       width = 12, height = 12, units = "cm",
+       path=path)
 
 # #################################
 # 2. Plot the Targeted Testing Scenario:
@@ -83,8 +85,8 @@ ggsave(p1_temp + ggtitle(TeX("W_S=W_I=W_R=1")),
 ggsave((p1_temp %+% df_targeted) +
         ggtitle(TeX(sprintf("W_S=%.1f, W_I=W_R=1",W_S_targeted ))),
     filename = "R0contour_TTI.pdf" ,
-    width = 14, height = 14, units = "cm",
-    path="../pix/")
+    width = 12, height = 12, units = "cm",
+    path=path)
 
 
 

### 2020-12-17 note/ProofOfRho.tex
initial proof of diff R0 wrt rho

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{xcolor}
\newtheorem{theorem}[]{Theorem}
\newcommand{\Rnum}{\mathcal{R}_0}
\title{ProofOfRho}
\date{December 2020}

\newcommand{\comment}{\showcomment}
\newcommand{\showcomment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}
\newcommand{\nocomment}[3]{}

\newcommand{\fady}[1]{\comment{cyan}{Fady}{#1}}
\newcommand{\ali}[1]{\comment{magenta}{Ali}{#1}}

\begin{document}

\maketitle
[Fady: for the proof below, it's probably good to include the matrix $V^{-1}$ in the proof. You have it on slide 39/49 from your talk today.]
\begin{theorem}
$\Rnum$ is a decreasing function of $\rho$
\begin{proof}
\ali{Test}We know that $FV^{-1}$ is given by

\begin{equation}
FV^{-1} = \left[ \begin {array}{cc}
G_{11}&G_{12}\\ 
0&0
\end {array} \right], \text{ where } \\
G_{11} =C
\left[\begin {array}{cc}
A\,S_u & B\,S_u\\
A\,S_n & B\,S_n
\end {array}\right]\nonumber \, ,
\end{equation}

The spectral radius of $FV^{-1}$ (i.e., $\Rnum$) is the nonzero eigenvalue of $G_{11}$. Equivalently, the spectral radius of $FV^{-1}$ is the trace of $G_{11}$, hence, it is sufficient to prove that all entries of $G_{11}$ decrease whenever $\rho$ increases.

Recall that

[Fady: we should insert the matrix $V^{-1}$ here. I found it on slide 39 in your talk]

notice that the columns of $V^{-1}$ always sum to $\gamma$. Furthermore, notice that the nonzero entries of the first row of $V^{-1}$ are strictly decreasing with respect to $\rho$.  Hence, the effect of increasing $\rho$ is to ''shift weight" from the entries in the first row to entries in lower rows. Now recall that

\begin{equation}
F = \beta/N_0 \left[ \begin {array}{cccc} 
S_u&\eta_w\,S_u&\eta_w\,S_u&\eta_t\,S_u\\
S_n&\eta_w\,S_n&\eta_w\,S_n&\eta_t\,S_n\\ 
0&0&0&0\\
0&0&0&0
\end {array} \right]\,,
\end{equation}

Since $\eta_w < 1$, $\eta_t < 1$, we know that $S_u$ is the largest entry in the first row of $F$, and $S_n$ is the largest entry in the second row of $F$. Thus, whenever $\rho$ increases, the dot product between row vectors of $F$ and column vectors of $V^{-1}$ decreases and hence entries of $FV^{-1}$ decrease. So in particular, the trace of $G_{11}$ (and hence $\Rnum$) decreases.

\end{proof}
\end{theorem}

\end{document}

### 2020-12-20 note/SIR_notes.Rnw
panel plot updated
more on understanding R0

@@
 \usepackage{subcaption}
-\usepackage{amsmath}
-\bibliographystyle{apalike}
+\usepackage{amssymb,amsmath,amsthm}
+\usepackage{xcolor}
+\usepackage{xspace}
+\usepackage[nameinlink,capitalize]{cleveref}
+\usepackage{cleveref}
+\usepackage{geometry}
+\usepackage{pdflscape}
+
+\usepackage{xspace}
+\newcommand{\Rlogo}{\protect\includegraphics[height=2ex,keepaspectratio]
+{pix/Rlogo.pdf}\xspace}
+
+\newcommand{\comment}{\showcomment}
+\newcommand{\showcomment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}
+\newcommand{\nocomment}[3]{}
+
+\newcommand{\fady}[1]{\comment{cyan}{Fady}{#1}}
+\newcommand{\ali}[1]{\comment{magenta}{Ali}{#1}}
+\newcommand{\jd}[1]{\comment{blue}{JD}{#1}}
+\newcommand{\djde}[1]{\comment{red}{DJDE}{#1}}
+\newcommand{\bmb}[1]{\comment{red}{BMB}{#1}}
+\newcommand{\todo}[1]{\comment{red}{TODO}{#1}}
+
+\newcommand{\Rnum}{\mathcal{R}_0}
+\theoremstyle{definition} % amsthm only
+\newtheorem{proposition}{Proposition}
+\newtheorem{note}{Note}
+\newtheorem{theorem}{Theorem}
 
@@
 
-\includegraphics[width=\linewidth]{../pix/sir_comp.png}
+\includegraphics[width=0.5\linewidth]{../pix/sir_comp.png}
 
 The model is 
+\begin{align}
+% \label{model}
+ d S_u/dt &= -\Lambda S_u - F_S S_u + \omega S_n, \label{eq1}\\
+ d S_n/dt &= -\Lambda S_n + (1-p_S) F_S S_u - \omega S_n, \label{eq2}\\
+ d S_p/dt &= -\Lambda S_p + p_S F_S S_u - \omega S_p, \label{eq3}\\
+ d S_c/dt &= -\Lambda S_c + \omega S_p, \label{eq4}\\
+ d I_u/dt &= \Lambda S_u - F_I I_u + \omega I_n  - \gamma I_u,  \label{eq5}\\
+ d I_n/dt &= \Lambda S_n + (1-p_I) F_I I_u - \omega I_n -\gamma I_n, \label{eq6}\\
+ d I_p/dt &= \Lambda S_p + p_I F_I I_u - \omega I_p -\gamma I_p, \label{eq7}\\
+ d I_c/dt &= \Lambda S_c + \omega I_p - \gamma I_c,  \label{eq8}\\
+ d R_u/dt &= \gamma I_u - F_R R_u + \omega R_n, \label{eq9}\\
+ d R_n/dt &= \gamma I_n + (1-p_R) F_R R_u - \omega R_n,  \label{eq10}\\
+ d R_p/dt &= \gamma I_p + p_R F_R R_u  - \omega R_p,  \label{eq11}\\
+ d R_c/dt&= \gamma I_c + \omega R_p,  \label{eq12}\\
+ dN/dt &= \omega (S_n + I_n + R_n),   \label{eq13}\\
+ dP/dt &= \omega(I_p + R_p) \label{eq14},
+\end{align}
+
+\newgeometry{margin=1cm} % modify this if you need even more space
+\begin{landscape}
+\begin{table}[htp]
+\centering
+{\tiny %
+\begin{tabular}{|c|ccc|} \hline
+  Symbol & Description & Unit & Value \\ \hline
+  $N_0$     & Total population size & people & $10^6$ \\ \hline
+  $\omega$  & Rate of onward flow from the awaiting to reported or untested compartments  & 1/day & - \\ \hline
+  $\gamma$ & Recovery rate & 1/day & 1/3 \\ \hline 
+  $\rho$   & Per capita testing intensity & 1/day & 0.01 \\ \hline 
+  $\eta_w$  & Relative probability of transmission for isolated awaiting individuals & - & - \\ \hline
+  $\eta_c$  & Relative probability of transmission for isolated confirmed individuals & - & -  \\ \hline
+  $\Lambda$ & Force of infection & 1/day & - \\ \hline
+  $p_S$ & Probability of false positive for susceptible& - & 0 \\ \hline
+  $p_I$ & Probability of being infected and tested positive & - & 1 \\ \hline
+  $p_R$ & Probability of being recovered and tested positive & - & 0.5 \\ \hline
+  $W_S, W_I, W_R$ & Relative testing weight & - & 
+  \begin{minipage}[t]{0.21\columnwidth}%
+ Random testing: $W_S=W_I=W_R=1$, \\Non-random testing: $W_S=0.3, W_I=W_R=1$
+\end{minipage} \\ \hline
+
+  \end{tabular}
+  }%
+\caption{\label{tab:params} The underlying parameters of model, \crefrange{eq1}{eq12}.}
+\end{table}
+\end{landscape}
+\restoregeometry
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Results}
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{On Calculation of $\Rnum$}
 
 \begin{align}
-\label{model}
- d S_u/dt &= -\Lambda S_u - F_S S_u + \omega S_n \\
- d S_n/dt &= -\Lambda S_n + F_S S_u - \omega S_n \\
- d I_u/dt &= \Lambda S_u - F_I I_u + \omega I_n  - \gamma I_u  \\
- d I_n/dt &= \Lambda S_n + (1-p_I) F_I I_u - \omega I_n -\gamma I_n \\
- d I_p/dt &= p_I F_I I_u - \omega I_p -\gamma I_p \\
- d I_t/dt &= \omega I_p - \gamma I_t  \\
- d R_u/dt &= \gamma I_u - F_R R_u + \omega R_n \\
- d R_n/dt &= \gamma I_n + (1-p_R) F_R R_u - \omega R_n  \\
- d R_p/dt &= \gamma I_p + p_R F_R R_u  - \omega R_p  \\
- d R_t/dt&= \gamma I_t + \omega R_p  \\
- dN/dt &= \omega (S_n + I_n + R_n)   \\
- dP/dt &= \omega(I_p + R_p).
+\label{FV}
+F =& \beta/N_0 \left[ \begin {array}{cccc} 
+S_u&\eta_w\,S_u&\eta_w\,S_u&\eta_c\,S_u\\
+S_n&\eta_w\,S_n&\eta_w\,S_n&\eta_c\,S_n\\ 
+0&0&0&0\\
+0&0&0&0
+ \end {array} \right], \\
+  V =&
+ \left[ \begin {array}{cccc}  
+F_I+\gamma&-\omega&0&0\\
+-(1-p_I)F_I&\omega+\gamma&0&0\\
+-p_I F_I&0&\omega+\gamma&0\\
+0&0&-\omega&\gamma
+\end {array} \right], \text{thus}\\
+V^{-1} =&
+\left[ \begin {array}{cccc}
+\frac {\omega+\gamma}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&\frac {\omega}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&0&0\\
+\noalign{\medskip}
+\frac{(1-p_I) F_I}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&
+\frac{F_I+\gamma}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&0&0\\
+\noalign{\medskip}
+\frac{p_I F_I}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&
+\frac{\omega p_I F_I}{(\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I)(\omega+\gamma)}& \frac{1}{\omega+\gamma}&0 \\
+\noalign{\medskip}
+\frac{\omega\,F_I\,p_I}{(\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I) \gamma}& 
+\frac{\omega^2\,F_I\,p_I}{(\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I)(\omega+\gamma) \gamma}&
+\frac{\omega}{(\omega+\gamma) \gamma}&
+\frac{1}{\gamma}
+\end {array} \right].
 \end{align}
 
-% %%%%%%%
-\section{Results}
+The particular form of $F$ with two rows of zeros at the bottom, simplifies $G$ as 
+\begin{equation}
+G = \left[ \begin {array}{cc}
+G_{11}&G_{12}\\
+0&0
+\end {array} \right], \text{ where } \\
+G_{11} =C
+\left[\begin {array}{cc}
+A\,S_u & B\,S_u\\
+A\,S_n & B\,S_n
+\end {array}\right].
+\end{equation}
 
-\newpage
-\begin{figure}
-\centering
-\begin{subfigure}[t]{.45\textwidth}
-\centering
-\includegraphics[width=\linewidth]{../pix/R0contour_random.png}
-        \caption{}\label{fig:fig_a}
-\end{subfigure}
-%
-\begin{subfigure}[t]{.45\textwidth}
-\centering
-\includegraphics[width=\linewidth]{../pix/R0contour_TTI.png}
-\caption{}\label{fig:fig_b}
-\end{subfigure}
-\end{figure}
+The Disease-Free Equilibrium (DFE) for the SIR model, \crefrange{eq1}{eq12}, is given by solving the coupled system including $S_u+S_n=N_0$ and $F_S S_u-\omega S_n=0$. The DFE is
 
-\begin{figure}
-\centering
-\begin{subfigure}[t]{.45\textwidth}
-\centering
-\includegraphics[width=\linewidth]{../pix/R0contour_random_s.png}
-        \caption{}\label{fig:fig_a}
-\end{subfigure}
-%
-\begin{subfigure}[t]{.45\textwidth}
-\centering
-\includegraphics[width=\linewidth]{../pix/R0contour_TTI_s.png}
-\caption{}\label{fig:fig_b}
-\end{subfigure}
-\end{figure}
+\begin{equation}
+\label{dfe}
+S_n^*= \frac{\rho}{\omega} N_0, \ S_u^*= N_0-S_n^*, \text{and} I_j=R_j=0 \ \text{for all j}.
+\end{equation}
 
-**Insights from the R0-contour plots**
+The basic reproduction number, $\Rnum$, was calculated by using the next generation matrix method developed by \cite{van2002reproduction}. $\Rnum$ is
 
-1. TTI requires lower testing rate than random testing to achieve $R_0=1$.
+\begin{equation}
+\label{R0}
+\Rnum= (A \times S_u^* + B \times S_n^*) \times C, 
+\end{equation}
+where
+\begin{align*}
+A=& \gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_c p_I) F_I, \\
+B=& \big(\omega+(F_I+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_c\omega) \omega p_I F_I }{\omega+\gamma}, \\ 
+C=& \frac{\beta/\gamma}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}.
+\end{align*}
 
-2. When random testing regime is applied, $W_S=W_I=W_R$,
+Note that the block matrix $G_{12}$ does not influence $\Rnum$ defined as the spectral radius of $G$. All matters here are the eigenvalues of $G_{11}$, which are 0 and $\Rnum$ \eqref{R0}.
 
-(i) in the case of perfect isolation of the reported individuals, $\eta_t=0$, the followings are inferred (see the first few columns of the panel plot); 
- - variations of the $\eta_w$ has a negligible effect on the critical testing.
- - for low $\omega$, $R_0$ increases as $\omega$ increases but saturates.
-This means that when the reporting process is very very slow, increasing the speed of the reporting causes higher $R_0$.
-Maybe because more people will move to $I_u$ [?, not very sure here.]
+% %%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Sensitivity of $\Rnum$ with respect to the underlying parameters}
 
- (ii) in case of the non-perfect isolation (see the last column), when $\eta_w$ increases (more awaiting people follow the isolation rules) faster reporting reduces $R_0$.  
-Note that it is sensible to expect that $\partial{R_0}/\partial{\omega}<0$. Here we showed that this can be less true.
+\begin{note}
+Notice that matrix $F$ is a rank one matrix, thus $FV^{-1}$ is of a rank one.
+This is why $FV^{-1}$ has only one non-zero eigenvalue.
+Specifically, $F$ is a rank one means that it can be written as matrix product of a column vector and a row vector as follows
+$$
+F = \beta/N_0 \left[\begin{array}{c} S_u\\  S_n\\ 0\\ 0 \end {array} \right]
+\left[\begin{array}{cccc} 1 \eta_w \eta_w \eta_c \end {array} \right].
+$$
 
-3. When TTI regime is applied,
+\end{note}
 
-% %%%%%%%
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Thoughts/Objectives}
@@
 \end{enumerate}
-
-% \item extension to a SEPAIR model (i.e. including latent/exposed, presymptomatic, asymptomatic)
-%     * analytical results will probably be out of reach
-% 	* but it _may_ be possible to extend some of the insights gained in part 1
-% 	* in any case, some numerical results (the space of relevant/interesting parameters will be larger: figure out a sensible way to show results from this parameter space.
-Are there useful dimensionless parameters that could help frame this?)
-% \item more detailed quantitative exploration: in particular, *either* Latin hypercube *or* (nearly the same!) a sample over estimated ('prior') distributions of the parameters.
-Goal: answer the general question "how likely is TTI to be able to control COVID?"
 

### 2020-12-30 SIR_manuscript.tex

@@
   $\eta_c$  & Relative probability of transmission for isolated confirmed individuals & - & -  \\ \hline
-  $\Lambda$ & Force of infection & 1/time & - \\ \hline
+  $\Lambda$ & Force of infection & 1/day & - \\ \hline
   $p_S$ & Probability of false positive for susceptible& - & 0 \\ \hline

### 2020-12-30 codes/sir_macpan.Rnw
seperating functions

--- 2020-12-08 codes/sir_macpan.Rnw
@@
 
-<<install the package, echo=F  >>=
+<<install the package, echo=T  >>=
 # install remotes package if necessary:
@@
 
-<<6: MacPan Params , echo=F>>=
+<<6: MacPan Params , echo=T>>=
 params <- read_params("ICU1.csv")
@@
 
-<<Ali:SIR , echo=F>>=
+<<Ali:SIR , echo=T>>=
 # A simple SIR model
@@
 
-<<Try MacPan to recover the SIR result, echo=FALSE>>=
-# update MacPan's params according to the SIR params
+<<Try MacPan to recover the SIR result, echo=T>>=
+## update MacPan's params according to the SIR params
+##
+big_rate <- 20
 params <- update(params, c(beta0=sir.params[["beta"]],
                            alpha=0, ## no asymptomatic
-                           sigma=3, ## very short latent period
-                           gamma_a=sir.params[["gamma"]],
+                           sigma=big_rate, ## very short latent period
+                           ## would like this to -> Inf but can't
+                           gamma_a=0,
+                           mu=1,  ## all cases mild
                            gamma_m=sir.params[["gamma"]],
-                           gamma_s=sir.params[["gamma"]],
-                           gamma_p=sir.params[["gamma"]]))
+                           gamma_s=0,
+                           ## try to skip presymptomatic period
+                           gamma_p=big_rate
+                           ))
 
@@
 + geom_point(aes(y=I))
-+ geom_point(aes(y=sir.I), color='red')
++ geom_line(aes(y=sir.I), color='red')
 )
@@
 print(gg01 + scale_y_log10())
+
 @

### 2020-12-30 note/ProofOfRho.tex

@@
 \begin{proof}
-\ali{Test}We know that $FV^{-1}$ is given by
+We know that $FV^{-1}$ is given by
 
@@
 
-notice that the columns of $V^{-1}$ always sum to $\gamma$.
+notice that the columns of $V^{-1}$ always sum to $\gamma$ \ali{$1/\gamma$}.
 Furthermore, notice that the nonzero entries of the first row of $V^{-1}$ are strictly decreasing with respect to $\rho$.
@@
 Since $\eta_w < 1$, $\eta_t < 1$, we know that $S_u$ is the largest entry in the first row of $F$, and $S_n$ is the largest entry in the second row of $F$.
-Thus, whenever $\rho$ increases, the dot product between row vectors of $F$ and column vectors of $V^{-1}$ decreases and hence entries of $FV^{-1}$ decrease.
+Thus, whenever $\rho$ increases, \ali{the next sentence sounds odd} the dot product between row vectors of $F$ and column vectors of $V^{-1}$ decreases and hence entries of $FV^{-1}$ decrease.
 So in particular, the trace of $G_{11}$ (and hence $\Rnum$) decreases.

### 2020-12-30 SIR_manuscript.tex
working on intro

@@
 
-One of the main challenges in understanding COVID-19 dynamics is to unravel information available from data.
+The observed dynamics of the COVID-19 epidemic is dirven by the integrated processes including both epidemiological processes (infection, recovery, etc.) and testing processes.
+In addition to determining what is actually observed (case reports), testing processes can also feed back to affect the epidemiological dynamics through isolation and quarantine.
+Specifically, individuals who are in the process of testing may partially or fully self-isolate, and individuals who have tested positive are highly likely to self-isolate.
+We developed a model that incorporates mechanistic epidemic processes and testing to explore the effects of testing and isolation on the control of the epidemic.
+
+Spereating the epidemiological processes from testing processes through a mathematical modeling excercise provides a tool to investigate the effect of the testing on the epidemic dynamics.
+Although, this effect is intuitively through isolation and quarantine practices, the sensitivity of the epidemic dynamics to these practices needs to be studied further.
+During COVID-19 epidemic, the isolation has applied to people who have been tested.
+Thus, the epidemic dynamics will depend on testing strategies which determines how tests are distributed across epidemiological compartments, or for a given testing intensity how many people in each compartment get tested.
+Specifically, in the case where tests are randomly assigned within the population (as might happen, for example, when disease surveillance is the purpose of testing), the tests will be distributed equally across all compartments.
+
+that the epi effects will depend on \emph{who gets tested}, i.e.
+how tests are distributed across compartments, or for a given testing intensity how many people in each epidemiological compartment get tested.
+Then you can go through some examples, i.e.
+start by saying that in the case where tests are randomly assigned within the population (as might happen, for example, when disease surveillance is the purpose of testing), the tests will be weighted equally across all compartments \ldots}
+
+% One of the main challenges in understanding COVID-19 dynamics is to unravel information available from data.
 The daily case count is a highly accessible form of data where observations are driven by two clusters of processes: (i) epidemic processes and (ii) testing processes.
@@
 
-\jd{There is a gap here between unravelling data and what we actually do.}
-\bmb{As JD suggests, the focus of the paper is not on unraveling data (although that is something we \emph{could} do, and are trying to do, with this modeling framework as well).
+% \jd{There is a gap here between unravelling data and what we actually do.}
+% \bmb{As JD suggests, the focus of the paper is not on unraveling data (although that is something we \emph{could} do, and are trying to do, with this modeling framework as well).
 I would say that the ``bullet points'' to start the paper are (1) the processes driving the observed dynamics of the COVID-19 epidemic include both epidemiological processes (infection, recovery, etc.) and testing processes (``processes'' is repeated here, could rephrase); (2) in addition to determining what is actually observed (case reports), testing processes can also feed back to affect the epidemiological dynamics through isolation and quarantine; (3) your last sentence in the \P\ above.}

### 2020-12-31 SIR_manuscript.tex
working on intro

@@
 
-The observed dynamics of the COVID-19 epidemic is dirven by the integrated processes including both epidemiological processes (infection, recovery, etc.) and testing processes.
+The observed dynamics of the COVID-19 epidemic is dirven by the underlying processes including both epidemiological processes (infection, recovery, etc.) and testing processes.
 In addition to determining what is actually observed (case reports), testing processes can also feed back to affect the epidemiological dynamics through isolation and quarantine.
 Specifically, individuals who are in the process of testing may partially or fully self-isolate, and individuals who have tested positive are highly likely to self-isolate.
-We developed a model that incorporates mechanistic epidemic processes and testing to explore the effects of testing and isolation on the control of the epidemic.
+We developed a model that incorporates mechanistic epidemic processes and testing to explore the effects of testing and isolation on the epidemic dynamics, and providing insights for the epidemic management/control.
 
-Spereating the epidemiological processes from testing processes through a mathematical modeling excercise provides a tool to investigate the effect of the testing on the epidemic dynamics.
-Although, this effect is intuitively through isolation and quarantine practices, the sensitivity of the epidemic dynamics to these practices needs to be studied further.
-During COVID-19 epidemic, the isolation has applied to people who have been tested.
-Thus, the epidemic dynamics will depend on testing strategies which determines how tests are distributed across epidemiological compartments, or for a given testing intensity how many people in each compartment get tested.
-Specifically, in the case where tests are randomly assigned within the population (as might happen, for example, when disease surveillance is the purpose of testing), the tests will be distributed equally across all compartments.
+Spereating the epidemiological processes from testing processes in a mechanistic modeling framework provides a tool to investigate the effect of the testing on the epidemic dynamics.
+Intutively, this effect is through isolation and quarantine practices.
+The informative/useful part is the sensitivity of the epidemic dynamics to testing, isolation and quarantine.
+Specifically, during COVID-19 epidemic, the isolation has applied to people who have been tested.
+Thus, the epidemic dynamics will depend on testing processes.
+The latter determines how tests are distributed across epidemiological compartments, or for a given testing intensity how many people in each compartment get tested.
+Specifically, in the case where tests are randomly assigned within the population (as might happen, for example, when disease surveillance is the purpose of testing), the tests will be weighted equally across all compartments.
+However, in the case where tests are targeted certain clusters of the population (as might happen, for example, for diagnosis and treatment, screening people for access to things like flights, surgery, long-term-care facilities, etc., and for mitigation and control of the disease spread) the tests will be weighted differently, where the greater the testing weight for a compartment, the more likely people in that compartment get tested.
+\todo{edit the last sentence, it is very long!}
 
-that the epi effects will depend on \emph{who gets tested}, i.e.
-how tests are distributed across compartments, or for a given testing intensity how many people in each epidemiological compartment get tested.
-Then you can go through some examples, i.e.
-start by saying that in the case where tests are randomly assigned within the population (as might happen, for example, when disease surveillance is the purpose of testing), the tests will be weighted equally across all compartments \ldots}
-
-% One of the main challenges in understanding COVID-19 dynamics is to unravel information available from data.
-The daily case count is a highly accessible form of data where observations are driven by two clusters of processes: (i) epidemic processes and (ii) testing processes.
-The challenge is then to separate these processes and understand how they interact.
-Simple mathematical models have a long history in providing useful insights facing with real epidemiological challenges \citep{ross1911prevention} [maybe add influenza work, ideas?].
-We developed a model that incorporates mechanistic processes of epidemic processes and testing.
-We used the model to study the potential effect of testing strategies on disease dynamics.
-
-% \jd{There is a gap here between unravelling data and what we actually do.}
 % \bmb{As JD suggests, the focus of the paper is not on unraveling data (although that is something we \emph{could} do, and are trying to do, with this modeling framework as well).
 I would say that the ``bullet points'' to start the paper are (1) the processes driving the observed dynamics of the COVID-19 epidemic include both epidemiological processes (infection, recovery, etc.) and testing processes (``processes'' is repeated here, could rephrase); (2) in addition to determining what is actually observed (case reports), testing processes can also feed back to affect the epidemiological dynamics through isolation and quarantine; (3) your last sentence in the \P\ above.}
-
-\jd{This \P\ is confusing. The topic sentence seems to be about kinds of tests, and then there is a jump to sampling strategies.}
-
-\bmb{It's fine for this \P\ to be about testing, but the topic sentence still doesn't say what the paragraph is trying to do.
-You should probably give the general idea of how isolation and quarantine affect epi dynamics (i.e.
-individuals who in the process of testing may partially or fully self-isolate, and individuals who have tested positive are highly likely to self-isolate), and say that your goal is to figure out the effects of isolation and testing: when can isolation and testing control an epidemic?
-You might even want to do this in the \emph{previous} paragraph.
-Then you can say that since isolation applies to people who have been tested, that the epi effects will depend on \emph{who gets tested}, i.e.
-how tests are distributed across compartments, or for a given testing intensity how many people in each epidemiological compartment get tested.
-Then you can go through some examples, i.e.
-start by saying that in the case where tests are randomly assigned within the population (as might happen, for example, when disease surveillance is the purpose of testing), the tests will be weighted equally across all compartments \ldots}
 
@@
 
-Further, the testing strategies can be grouped into two categories including random testing across the whole population and Non-random/biased testing.
-Random testing is the simplest form of testing strategy and is typically carried out for the purpose of surveillance.
-It is notable that, by random testing we mainly refer to PCR test given that there is almost negligible virological testing for surveillance purposes.
+It is notable that, by random testing we mainly refer to the PCR test given that there is almost negligible virological testing for surveillance purposes.
 Furthermore, all of the accessible testing data is being used for one of the other three purposes rather than surveillance.
-These are non-random testing strategies with a different bias.
+These are targeted testing strategies with a different bias.
 Specifically, when the purpose is "(i) diagnosis and treatment" or "(iv) mitigation and control of disease spread", the bias is towards infectious individuals more strongly for (i) than (iv).
@@
 For example, people who are getting on flights may/will be more mobile than people awaiting for a surgery and are probably going to stay in long-term care homes and consequently more isolated to begin with.
-
--Examples of random versus biased testing strategies;
-Each of these strategies has been used as a tool to study COVID-19; see (refs) for group (i) and (refs) for group (ii).
-The advantage of non-random testing comparing to random testing has been investigated specifically during the COVID-19 pandemic (refs).
 

### 2021-01-05 SIR_manuscript.tex
working on intro
updated intro

@@
 
-The observed dynamics of the COVID-19 epidemic is dirven by the underlying processes including both epidemiological processes (infection, recovery, etc.) and testing processes.
+The observed dynamics of the COVID-19 epidemic is dirven by the underlying processes including both epidemiological processes (infection, recovery, etc.) and testing processes (who gets tested and intensity).
 In addition to determining what is actually observed (case reports), testing processes can also feed back to affect the epidemiological dynamics through isolation and quarantine.
 Specifically, individuals who are in the process of testing may partially or fully self-isolate, and individuals who have tested positive are highly likely to self-isolate.
-We developed a model that incorporates mechanistic epidemic processes and testing to explore the effects of testing and isolation on the epidemic dynamics, and providing insights for the epidemic management/control.
+We developed a model that incorporates mechanistic epidemic processes and testing to explore the effects of testing and isolation on the epidemic dynamics, and providing insights for the population-level epidemic control.
 
-Spereating the epidemiological processes from testing processes in a mechanistic modeling framework provides a tool to investigate the effect of the testing on the epidemic dynamics.
-Intutively, this effect is through isolation and quarantine practices.
-The informative/useful part is the sensitivity of the epidemic dynamics to testing, isolation and quarantine.
-Specifically, during COVID-19 epidemic, the isolation has applied to people who have been tested.
-Thus, the epidemic dynamics will depend on testing processes.
-The latter determines how tests are distributed across epidemiological compartments, or for a given testing intensity how many people in each compartment get tested.
-Specifically, in the case where tests are randomly assigned within the population (as might happen, for example, when disease surveillance is the purpose of testing), the tests will be weighted equally across all compartments.
-However, in the case where tests are targeted certain clusters of the population (as might happen, for example, for diagnosis and treatment, screening people for access to things like flights, surgery, long-term-care facilities, etc., and for mitigation and control of the disease spread) the tests will be weighted differently, where the greater the testing weight for a compartment, the more likely people in that compartment get tested.
-\todo{edit the last sentence, it is very long!}
+During COVID-19 epidemic, since isolation has applied to people including, but not limited to, those who have been tested, the epidemic dynamics will depend on who gets tested.
+That is, how tests are distributed across epidemiological compartments for a given testing intensity.
+Specifically, in the case where tests are randomly assigned within the population, the tests will be weighted equally across all compartments.
+Random testing is the simplest form of testing regime and might happen, for example, when disease surveillance is the purpose of testing.
+However, in the case where targeted testing is administered within the population, the testing will be more weighted towards the compartments with people who have higher probability of getting tested.
+Targeted testing might happen, for example, for different purposes including diagnosis and treatment, screening people for access to things like flights, surgery, long-term-care facilities, and for mitigation and control of the disease spread.
+Specifically, when the purpose is "diagnosis and treatment" or "mitigation and control of disease spread", the bias is skewed towards infectious individuals with even more weights for the diagnosis and treatment than the mitigation and control.
+Also, when people are tested for "screening", the bias is towards the group-specific characteristics \ali{better wording needed}.
+For example, people who are getting on flights may or will be more mobile than people awaiting for a surgery and are probably going to stay in a long-term care facility and consequently more isolated to begin with.
+
+Furthermore, test-specific sensitivities, intensity and sample-to-answer reporting time are some components of the testing processes that influence the epidemic dynamic.
+We model these components and explore the trade-offs between them in achieving the epidemic control objective.
+
+R0
+
+Frequent proactive testing of asymptomatic individuals offers an additional way to reduce transmission, and is among the most powerful and least disruptive of non-pharmaceutical interventions such as face masks, social distancing.
 
@@
 
-It is often the purpose of testing, apart from the testing infrastructure, that dictates/defines \ali{[better word?]} the testing strategy.
-There are at least four different, and sometimes conflicting, purposes of testing: (i) diagnosis and treatment, (ii) screening people for access to things like flights, surgery, LTCs etc, (iii) surveillance, and (iv) mitigation and control of disease spread.
-
-It is notable that, by random testing we mainly refer to the PCR test given that there is almost negligible virological testing for surveillance purposes.
-Furthermore, all of the accessible testing data is being used for one of the other three purposes rather than surveillance.
-These are targeted testing strategies with a different bias.
-Specifically, when the purpose is "(i) diagnosis and treatment" or "(iv) mitigation and control of disease spread", the bias is towards infectious individuals more strongly for (i) than (iv).
-When people are tested for "(ii) screening", the bias is towards the group-specific characteristics [I am making this up for now!].
-For example, people who are getting on flights may/will be more mobile than people awaiting for a surgery and are probably going to stay in long-term care homes and consequently more isolated to begin with.
+\ali{discussion maybe: all of the accessible testing data is being used for one of the other three purposes rather than surveillance. These are targeted testing strategies with a different bias.}
 

### 2021-01-05 SIR_manuscript.tex
intro updated

@@
 
-The observed dynamics of the COVID-19 epidemic is dirven by the underlying processes including both epidemiological processes (infection, recovery, etc.) and testing processes (who gets tested and intensity).
+The observed dynamics of the COVID-19 epidemic is dirven by the underlying processes including both epidemiological processes (infection, recovery, etc.) and testing processes (including who gets tested for a given testing intensity, test sensitivity, etc.
+\ali{right?}).
 In addition to determining what is actually observed (case reports), testing processes can also feed back to affect the epidemiological dynamics through isolation and quarantine.
@@
 During COVID-19 epidemic, since isolation has applied to people including, but not limited to, those who have been tested, the epidemic dynamics will depend on who gets tested.
-That is, how tests are distributed across epidemiological compartments for a given testing intensity.
+That is, how tests are distributed across epidemiological compartments (susceptible-infectious-recovered) for a given testing intensity.
 Specifically, in the case where tests are randomly assigned within the population, the tests will be weighted equally across all compartments.
-Random testing is the simplest form of testing regime and might happen, for example, when disease surveillance is the purpose of testing.
+Random testing strategy is the simplest form of testing and might happen, for example, when disease surveillance is the purpose of testing \citep{foddai2020base}.
 However, in the case where targeted testing is administered within the population, the testing will be more weighted towards the compartments with people who have higher probability of getting tested.
-Targeted testing might happen, for example, for different purposes including diagnosis and treatment, screening people for access to things like flights, surgery, long-term-care facilities, and for mitigation and control of the disease spread.
-Specifically, when the purpose is "diagnosis and treatment" or "mitigation and control of disease spread", the bias is skewed towards infectious individuals with even more weights for the diagnosis and treatment than the mitigation and control.
-Also, when people are tested for "screening", the bias is towards the group-specific characteristics \ali{better wording needed}.
-For example, people who are getting on flights may or will be more mobile than people awaiting for a surgery and are probably going to stay in a long-term care facility and consequently more isolated to begin with.
+Targeted testing might happen, for example, for different purposes including diagnosis and treatment, screening people for access to things like flights, surgery, long-term-care facilities, and for the purpose of mitigation and control of the disease spread.
+Specifically, when the purpose is "diagnosis and treatment" or "mitigation and control of disease spread", the weights of testing is skewed towards infectious compartment rather than susceptible and recovered compartments.
+Further, heavier testing weight may be put on the infected compartment in the case of diagnosis and treatment than the case of mitigation and control \ali{why?
+refs?}.
+Also, when people are tested for "screening", the individuals with potential higher mobility, eg.
+people who are getting on flights, get more tested and thus the coresponding heavier testing weight is assigned than people awaiting for a surgery and are probably going to stay in a long-term care facility and consequently less mobile and more isolated to begin with.
 
-Furthermore, test-specific sensitivities, intensity and sample-to-answer reporting time are some components of the testing processes that influence the epidemic dynamic.
-We model these components and explore the trade-offs between them in achieving the epidemic control objective.
+In addition to the efect of compartment-specific testing weights on the isolation, and thus epidemic dynamics, other influential components of testing processes on isolation include test-specific sensitivities, intensity and sample-to-answer reporting time.
+They determine how accurately and how fast people are put in the isolation.
+We model these components and explore their effects on the epidemic dynamics.
+Specifically, we study the trade-offs between the testing strategies (random and targeted) and the discussed testing components on a single epidemiological-dynamics index, the ``basic reproduction number'' $\Rnum$.
 
-R0
-
-Frequent proactive testing of asymptomatic individuals offers an additional way to reduce transmission, and is among the most powerful and least disruptive of non-pharmaceutical interventions such as face masks, social distancing.
-
-% \bmb{As JD suggests, the focus of the paper is not on unraveling data (although that is something we \emph{could} do, and are trying to do, with this modeling framework as well).
-I would say that the ``bullet points'' to start the paper are (1) the processes driving the observed dynamics of the COVID-19 epidemic include both epidemiological processes (infection, recovery, etc.) and testing processes (``processes'' is repeated here, could rephrase); (2) in addition to determining what is actually observed (case reports), testing processes can also feed back to affect the epidemiological dynamics through isolation and quarantine; (3) your last sentence in the \P\ above.}
-
-\ali{discussion maybe: all of the accessible testing data is being used for one of the other three purposes rather than surveillance. These are targeted testing strategies with a different bias.}
-
-- Speed of test; 
-Much effort has been invested in developing widely accessible and fast tests in order to mitigate the spread of the pandemic (refs).
-For example, PCR tests are commonly used, and very sensitive, but take longer compared to more recent tests, such as a breath-based test currently in trial stage, which can provide a result in seconds.
+The early dynamics of an epidemic are determined by $\Rnum$.
+This index is defined as the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}.
+In the early stages of an epidemic the number of infected individuals is expected grow exponentially over time when $\Rnum>1$.
+To control the spread of disease in the population, $\Rnum$ is required to be below unity.
 
@@
 This counter-intuitive effect can occur when people are cautious when awaiting a test result, and may not be robust to second-order effects of fast testing (such as better contact tracing).
+
+% Frequent proactive testing of asymptomatic individuals offers an additional way to reduce transmission, and is among the most powerful and least disruptive of non-pharmaceutical interventions such as face masks, social distancing.
+% \bmb{As JD suggests, the focus of the paper is not on unraveling data (although that is something we \emph{could} do, and are trying to do, with this modeling framework as well).
+I would say that the ``bullet points'' to start the paper are (1) the processes driving the observed dynamics of the COVID-19 epidemic include both epidemiological processes (infection, recovery, etc.) and testing processes (``processes'' is repeated here, could rephrase); (2) in addition to determining what is actually observed (case reports), testing processes can also feed back to affect the epidemiological dynamics through isolation and quarantine; (3) your last sentence in the \P\ above.}
+
+% - Speed of test; 
+% Much effort has been invested in developing widely accessible and fast tests in order to mitigate the spread of the pandemic (refs).
+For example, PCR tests are commonly used, and very sensitive, but take longer compared to more recent tests, such as a breath-based test currently in trial stage, which can provide a result in seconds.
+\ali{discussion maybe: all of the accessible testing data is being used for one of the other three purposes rather than surveillance. These are targeted testing strategies with a different bias.}
 

### 2021-01-05 codes/sir_plot.R
tweak path, makefile

@@ -2,7 +2,7 @@
 source("SIRfunctions.R")
 source("params.R")
 # where to save the output:
-# path="../pix/"
+path <- "../pix/"
 # #################################
 # Making Dataframe part
 # #################################

### 2021-01-05 SIR_manuscript.tex
updated intro

@@
 To control the spread of disease in the population, $\Rnum$ is required to be below unity.
+\ali{now may be say what we expect from testing and isolation to reduce R0 and the suprise that the reporting time increase R0.
+}
 
-Here, we develop a model to incorporate epidemic processes and testing processes mechanistically to study the potential effect of testing strategies and test speed on epidemic control.
-We investigate the effect of underlying parameters on the basic reproduction number, $\Rnum$.
-While focused testing strategies are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is viral spread is expected to be stronger for a slow test than for a fast test.
+Here, we develop a compartmental model to incorporate epidemic processes and testing processes mechanistically to study the potential effect of testing strategies, testing intensity, test sensitivity and specificity, reporting time and isolation on epidemic dynamic.
+We investigate the effect of underlying parameters of the model on the basic reproduction number, $\Rnum$.
+While targeted testing strategies are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is viral spread is expected to be stronger for a slow test than for a fast test.
 This counter-intuitive effect can occur when people are cautious when awaiting a test result, and may not be robust to second-order effects of fast testing (such as better contact tracing).

### 2021-01-12 codes/sir_macpan.Rnw
minutes added

@@
 @
-This is my a test for MacPan package.
+This is my a test for MacPan package in 2 steps:
+- step 1: simulate a simple SIR model and recover the results by specifying the parameters of MacPan.
+- step 2: Read COVID19-Canada data from Mike's repo, estimate the transmision rate beta(t) and do projection.
 
@@
 @
+
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Step 1, SIR and MacPan}
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Step 2, Data and MacPan}
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+<<Read Data from Mike's repo, echo=T>>=
+library(readr)
+
+covid_url <- "https://raw.githubusercontent.com/wzmli/COVID19-Canada/master/git_push/clean.Rout.csv"
+df_covid <- read_csv(covid_url)
+
+plot_pos_prov <- ggplot(df_covid,aes(x=Date,y=confirmed_positive,col=Province))+
+                  geom_point()
+
+direct.label(plot_pos_prov,"last.bumpup")
+
+@
+
 \end{document}

### 2021-01-12 note/minutes.Rmd
minutes added

---
title: "minutes of meetings"
author: "Ali"
date: "06/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dr. Ben Bolker

### date: Jan 5, 2021
- targeted testing, 
  - no complete reference including all we have said in the SIR paper
  - A way to think about testing weights is a "table of conditional probabilities", i.e.,  
  -  Example: lets say (i) we only test symptomatic people, and we know the fraction of I/N, and prob of one has symptoms if you have covid and the prob of symptoms if one has no covid, then we should know what the test positivity will be. May be not!?
  - More on this:
    - suppose we know the conditional prob(tested|asymptomatic)  (ie not infected) = 0
    - the prob(tested|symptomatic) is independent of whether you are infected or not. 
    - example suppose we have a population of people, unknown number, with symptoms and we test 20% of all those with symp, then test pos is the prob(infected|symptomatic). Also
    - If we randomly test people with symptoms, then W_Zs are proportional to conditional 7jdsm probabilities of being in those compartments given that you have symptoms.
  - Lets just think of only 4 compartments, S and I with and without symptoms. let's 3% of pop is I, 60% of I's show symp, 1.8% I and Symp and 1.2% I and Asymp. The other 97% of pop are in S, 2% S and symp and 95% is S ans Asyp. Thus, 3.8% of the whole pop is symptomatic and of that 1.8% were infected results in 
  test positivity is 9/19 ~ 50%. We are assuming also that the test is perfect sensitive and specific. that is, if we only testing symp people, the intensity doesn't matter, we are pulling them from 3.8% of the pop. This says, with symptom fraction in S and I and I/S, this tells us what the weights are, how many I's we sample to test for every S to sample to test. That is not quite right, that is only when we rescale the weights. In this example these are the ratios at which we test not the per capitas. Some of the effects of pop size will cancel out, I think the frac of pop of S and I will cancel out, and what we get as W_Z is the rel rates of showing symptoms of S and I. In our example, the testing weights before scaling will be 2/62 and 60/62.      
  -  More thinking: We know that not all symptomatic people are the same. people know sth about their exposure. Let's say we don't know about symptoms. We know that there is a bias towards testing infected people, due to having some knowledge of risk. It is harder to quantify that than quantifying symptom/no symptom but we can say that would shift if we have a policy of screening people of high risk areas will push things towards infected people. Also testing people of known exposures will increase W_I, this is exactly testing and tracing. It is much harder to quantify that. In principle if we knew enough of the case report, hospitalizations, test positivity, death, we could make an estimate of W_z's. W's don't stay the same because it is all bound up with capacity because we tend to bias towards testing people who are more likely be infected and the more testing capacity we have the more we are willing to test more people less likely to be infected. Thus there is a relationship between testing capacity and the W's because our testing policy changes as the testing capacity changes. 
    - Except for the symptomatic example, it is very hard to quantify the effect of a particular testing policy on the W's. We can certainly make qualitative statements about when testing is biased towards people who are likely to be more infected then W_I will go up.

- Next project may be "Test positivity tells a lot about test Weights"; It is hard to know exactly how much information there is and how much we can estimate that. 
  - If we have a perfect test, then we know that the rescaled weights (not the percapita) are equal to positivity which is equal to per capita weights and scale them by the fraction of infected i.e., ~ W_S *S/N. But we have a single number and 2 unknowns (a weight and a prevalence). Example let say we have 10% test positivity, that is consistent with random sampling and a 10% prevelence. It also consistent with perfectly bias sampling when we know how to only test infected people and exactly 10% of the population is infected [?].
  N=1000, we do 100 tests and 10 of our samples are infected. If we sample it random, we think that 100 people in the population are infected.  If W_I is perfect, then there might be no more than 10 infected people in the entire population, that would mean that the prevalence is 1%. We also know how manu tests we did, but doesn't resolve it. One solution is bounding it; that we know that the minimum prevalence is equal to observed positive tests divided by the population size and the max prevalence is N - (the # of neg tests we did)/N =1-min prevalence. Because if we got 10 pos samples in example above,  there might be I=10 or I= 910.

<!-- %%%%%%%%%%%%%%%%%%%%%%%%%%%%% -->
## Dr. Jonathan Dushoff

### Date: 28 Dec, 2020
- We agreed that the previous proof of R0 decreasing with testing intensity rho was not correct.
- It is actually hard to prove that R0 is decreasing wrt testing intensity, it is a biological problem not mathematical. Specifically, if one goes in the awaiting compartment, it is harder to get a confirmed test because we are assuming that the test may come back negative.
- next step is to show/see if there is a place where increasing rho can increase R0? Is it at the beginning or later on at the pandemic. prove it!

### 2021-01-12 codes/sir_plot.R

@@ -88,7 +88,29 @@ ggsave((p1_temp %+% df_targeted) +
     width = 12, height = 12, units = "cm",
     path=path)
 
+# #################################
+# 3. Plot R0 as a function of rho (testing intensity):
+# #################################
+eta_w <- 0.5
+eta_c <- eta_w
+gamma <- 1/4
+W_S <- 1
+omega <- 10^6
+rho <- seq(0.5,1, length.out=100)
+f <- function(params){
+  unpack(as.list(params))
+  return(gamma*(W_S/W_I)*1/(sqrt(eta_w/gamma)-1))
+}
+f(params = update(params,c(eta_w=eta_w,eta_c=eta_c,
+                           gamma=gamma)))
 
-
-
-
+df1 <- expand.grid(N0=params[["N0"]],beta=params[["beta"]],gamma=gamma,
+                   omega=omega,rho=rho,
+                   W_S=W_S,W_I=params[["W_I"]],W_R=params[["W_R"]],
+                   p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
+                   eta_w=eta_w,eta_c=eta_c)
+df2<- data.frame(df1, R0=apply(df1,1,function(params_in)R0(params=params_in)),
+                      f=apply(df1,1,function(params_in)f(params = params_in)),
+                      S_u=apply(df1,1,function(params_in)Su_dfe(params = params_in))
+                )
+ggplot(df2,aes(x=rho,y=R0))+geom_point()

### 2021-01-15 codes/sir_macpan.Rnw
minor updates

@@
 <<Read Data from Mike's repo, echo=T>>=
+library(McMasterPandemic)
 library(readr)
+library(dplyr)
+library(ggplot2); theme_set(theme_bw())
+library(directlabels)
 
@@
 
-plot_pos_prov <- ggplot(df_covid,aes(x=Date,y=confirmed_positive,col=Province))+
+## BMB: the fit was using *cumulative* cases before, 
+## Ali I used confirmed_positive before
+plot_pos_prov <- ggplot(df_covid,aes(x=Date,y=newConfirmations,col=Province))+
   geom_point()
@@
 
+# Assume we are just looking at Ontario's epidemic case:
+
+startdate <- as.Date("2020-12-11")
+enddate <- as.Date("2021-01-12")
+## BMB: if we use the original, short time series
+##   Error in stats::filter(i, kern, sides = 1) : 
+##     'filter' is longer than time series
+
+df_covid_on <- (df_covid 
+                %>% filter(Province =="ON", Date >= startdate) # Ontario
+                %>% mutate(value= round(newConfirmations), var= "report", date=Date)
+                %>% select(date, value, var)
+                %>% na.omit()
+)
+## BMB: check length, beginning
+nrow(df_covid_on)
+head(df_covid_on,1)
+
+## beta0 (and E0) the only parameter we're going to optimize:
+params <- read_params("ICU1.csv")
+## adjust to ON pop size, approx starting value:
+## about 1800 cases reported
+## adjust ascertainment rate (fraction of cases reported) to 1/3
+## (was 0.1, which was reasonable at beginning of epidemic)
+## (will just rescale results)
+## so 1800 cases -> prevalence of approx 3*1800
+## initial conditions automatically spread E0 over the eigenvector
+params <- update(params, N=15e6,E0=3*1800, c_prop=1/3)
+
+## set up optimization parameters
+opt_pars <- get_opt_pars(params, vars="report")
+
+params1obs <- update(params, obs_disp=20)
+
+## document: bounds for DE optimization???
+
+plot(run_sim(
+    params = params1obs,
+    start_date=startdate,
+    end_date=enddate),
+    keep_states="report", log=TRUE)
+
+## fit beta0 based on the report data:
+fitted.mod <- calibrate(
+  data = df_covid_on
+  , start_date = startdate
+  ## skip breaks that are present by default:
+, time_args = list(break_dates = NULL)
+  , base_params = params1obs
+, opt_pars = opt_pars
+ , debug_plot = TRUE # instructive plotting during optimization
+)
+
+## plot the resulting fit
+plot(fitted.mod,data=df_covid_on)
+
+## FIXME in macpan: check for correct structure in opt_pars
+## check length of time series?
+## FIXME: shouldn't need to specify time_args ...
+
 @

### 2021-01-17 SIR_manuscript.tex
Faddy updated

@@
 This counter-intuitive effect can occur when people are cautious when awaiting a test result, and may not be robust to second-order effects of fast testing (such as better contact tracing).
+\ali{discussion maybe: all of the accessible testing data is being used for one of the other three purposes rather than surveillance. These are targeted testing strategies with a different bias.}
 
+\fady{Not sure if interesting: \\
+one point we can make when discussing the relative weight of testing in different compartments is to discuss pre-testing screening tools, such as surveys or questionnaires.
+If we have a quantitative description of how the testing intensities affect the dynamics, we can make statements like ``our results suggest that employing a pre-testing screening tool can help target infected individuals more effectively.
+In particular, doubling the sensitivity of the pre-screening tool would *do something* to $\Rnum$.}
 % Frequent proactive testing of asymptomatic individuals offers an additional way to reduce transmission, and is among the most powerful and least disruptive of non-pharmaceutical interventions such as face masks, social distancing.

### 2021-01-17 codes/sir_macpan.Rnw
Faddy updated

@@
 startdate <- as.Date("2020-12-11")
-enddate <- as.Date("2021-01-12")
+enddate <- as.Date("2021-01-14")
 ## BMB: if we use the original, short time series
@@
 ## plot the resulting fit
-plot(fitted.mod,data=df_covid_on)
+plot(fitted.mod,data=df_covid_on) # all
+plot(fitted.mod,data=df_covid_on, # just report data
+     predict_args=list(keep_vars=c("report")))
+
+coef(fitted.mod) ## output all fitted params
+coef(fitted.mod,"fitted") ##output just beta0
+## Note that that the value of \code{beta0} used for the simulation was 1
 
@@
 
+# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+## Forecasting
+# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+## Here, an example involving changing the transmission rate ($\beta$) in the future due to lockdown ...
+
+## Goal: to forcast after the date of lockdown
+lockdown <- as.Date("2021-01-14")
+enddate_f <- as.Date("2021-01-31") ## for forecasting
+
 @

### 2021-01-18 codes/sir_macpan.Rnw
forcast updated

@@
 # %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-## Here, an example involving changing the transmission rate ($\beta$) in the future due to lockdown ...
+## Here, an example involving changing the transmission rate ($\beta$) in the future due to lockdown
+
+pp <- coef(fitted.mod, "all") # all the fitted params in the form of ICU1
+# pp <- fix_pars(params, target = c(R0 = 1.3, Gbar=6)) # or you could specify your R0 and Gbar
+state <- make_state(params=pp)
+
+startdate <- as.Date("2020-12-11")
+enddate <- as.Date("2021-01-30")
+
+sim0 <- run_sim(pp,state,start_date=startdate,end_date=enddate)
+gg0 <- (ggplot(sim0,aes(x=date))
+	+ geom_point(aes(y=report))
+)
+print(gg0)
+
+## Question: the runsim gives reported cases starrting from 2465 cases and grows, it doesn't match the reported data! I think I am missing some scaling here?
 
 ## Goal: to forcast after the date of lockdown
+## We want a dataframe that includes the time varying relative $\beta$ at relative $\beta$ to drop by a factor of 2 (linearly) between 14 Jan 2021 on at each saved time point.
 lockdown <- as.Date("2021-01-14")
-enddate_f <- as.Date("2021-01-31") ## for forecasting
+
+time_pars <- data.frame(Date=as.Date(startdate:enddate)
+	, Symbol="beta0"
+	, Relative_value= pp[["beta0"]] ## Here I am using the caliberated Beta0
+)
+	# , stringsAsFactors=FALSE)
+
+time_pars2 <- 
+    data.frame(Date=as.Date(startdate:enddate)
+             , Symbol="beta0"
+             , Relative_value = 
+                   c(rep(pp[["beta0"]], length(startdate:lockdown)-1)
+                   , seq(pp[["beta0"]],pp[["beta0"]]-0.1,length.out = length(lockdown:enddate))
+                     )
+               )
+print(time_pars2)
+
+sim0_t_reduce <- update(sim0, params_timevar=time_pars2)
+gg_rel_beta <- (ggplot(time_pars, aes(x=Date))
+	+ geom_point(aes(y=Relative_value))
+	+ geom_point(data=time_pars2, aes(x=Date, y=Relative_value), color="red")
+)
+print(gg_rel_beta)
+# We can now look at the relative value of $\beta$ in each scenario, and the corresponding forecasted epidemic curves.
+
+print(gg0
+		+ geom_point(data=sim0_t_reduce, aes(x=date,y=incidence), color="red")
+)
 

### 2021-01-18 codes/sir_macpan.Rnw
forcast updated
updated with BMB comments

@@
 % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Step 1, SIR and MacPan}
+\section{Step 1 SIR and MacPan}
 % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 library(directlabels)
+path <- "/home/ag/projects/SIRmodel/codes/"
+covid_url <- "https://raw.githubusercontent.com/wzmli/COVID19-Canada/master/git_push/clean.Rout.csv"
+covid_can <- read_csv(covid_url)
+fn_covid_can = paste0(path,"covid_can.csv")
+write_csv(covid_can,fn_covid_can)
 
-covid_url <- "https://raw.githubusercontent.com/wzmli/COVID19-Canada/master/git_push/clean.Rout.csv"
-df_covid <- read_csv(covid_url)
-
+## Ali I used confirmed_positive before
 ## BMB: the fit was using *cumulative* cases before, 
-## Ali I used confirmed_positive before
-plot_pos_prov <- ggplot(df_covid,aes(x=Date,y=newConfirmations,col=Province))+
+plot_pos_prov <- ggplot(covid_can,
+                        aes(x=Date,y=newConfirmations,col=Province))+
                   geom_point()
-
 direct.label(plot_pos_prov,"last.bumpup")
@@
 # Assume we are just looking at Ontario's epidemic case:
-
 startdate <- as.Date("2020-12-11")
@@
 
-df_covid_on <- (df_covid 
+df_covid_on <- (covid_can 
                 %>% filter(Province =="ON", Date >= startdate) # Ontario
@@
 head(df_covid_on,1)
+tail(df_covid_on)
 
@@
 ## initial conditions automatically spread E0 over the eigenvector
-params <- update(params, N=15e6,E0=3*1800, c_prop=1/3)
+params <- update(params, N=15e6,E0=exp(11.2), c_prop=1/3, beta=-1.5)
 
@@
 
-params1obs <- update(params, obs_disp=20)
+params1obs <- update(params, obs_disp=20) #set dispersion params to 20
 
@@
   , debug_plot = TRUE # instructive plotting during optimization
+  , use_DE= FALSE
+  , DE_lwr=c(6,-2,-2),
+  , DE_upr=c(12,2,2)
 )
@@
 print(gg0)
-
-## Question: the runsim gives reported cases starrting from 2465 cases and grows, it doesn't match the reported data! I think I am missing some scaling here?
+# Ben
+print(gg0 + geom_point(data=df_covid_on,aes(y=value),colour="red"
+                       ))
 
@@
 	, Symbol="beta0"
-	, Relative_value= pp[["beta0"]] ## Here I am using the caliberated Beta0
+	, Relative_value= 1 ## Here I am using the caliberated Beta0
 )
-	# , stringsAsFactors=FALSE)
 
 time_pars2 <- 
-    data.frame(Date=as.Date(startdate:enddate)
+    data.frame(Date=seq(lockdown,enddate,by=1)
              , Symbol="beta0"
              , Relative_value = 
-                   c(rep(pp[["beta0"]], length(startdate:lockdown)-1)
-                   , seq(pp[["beta0"]],pp[["beta0"]]-0.1,length.out = length(lockdown:enddate))
-                     )
-               )
+               seq(1,.5,length=enddate-lockdown+1) )
+
 print(time_pars2)
@@
 print(gg0
-		+ geom_point(data=sim0_t_reduce, aes(x=date,y=incidence), color="red")
+		+ geom_point(data=sim0_t_reduce, aes(x=date,y=report), color="blue")
+		+ geom_point(data=df_covid_on,aes(y=value),colour="red"  )
       )
+## the blue (black) curve is forecast with (without) lockdown impact
+## the red curve is actual data
+
+## Question: How to add confidence Int?
 

### 2021-01-22 codes/sir_macpan.Rnw
bib update

@@
   , debug_plot = TRUE # instructive plotting during optimization
-  , use_DE= FALSE
+  , use_DE= FALSE ## use differential evolution algorithm
   , DE_lwr=c(6,-2,-2),

### 2021-01-24 SIR_manuscript.tex
rewrite intro, notes elsewhere

@@
 
+% https://tex.stackexchange.com/questions/12703/how-to-create-fixed-width-table-columns-with-text-raggedright-centered-raggedlef
+\usepackage{array}
+\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+
 \usepackage{xspace}
-\newcommand{\Rlogo}{\protect\includegraphics[height=2ex,keepaspectratio]
-{pix/Rlogo.pdf}\xspace}
+%\newcommand{\Rlogo}{\protect\includegraphics[height=2ex,keepaspectratio]{pix/Rlogo.pdf}\xspace}
+% BMB: we don't need the weird graphical logo ...
+\newcommand{\Rlogo}{R\xspace}
 
@@
 
-The observed dynamics of the COVID-19 epidemic is dirven by the underlying processes including both epidemiological processes (infection, recovery, etc.) and testing processes (including who gets tested for a given testing intensity, test sensitivity, etc.
-\ali{right?}).
-In addition to determining what is actually observed (case reports), testing processes can also feed back to affect the epidemiological dynamics through isolation and quarantine.
-Specifically, individuals who are in the process of testing may partially or fully self-isolate, and individuals who have tested positive are highly likely to self-isolate.
-We developed a model that incorporates mechanistic epidemic processes and testing to explore the effects of testing and isolation on the epidemic dynamics, and providing insights for the population-level epidemic control.
+The observed dynamics of the COVID-19 epidemic are driven by both epidemiological processes (infection and recovery) and testing processes (testing and test reporting).
+In addition to determining what we actually observe about the epidemic (case reports), testing processes can also feed back to affect the epidemiological dynamics through isolation and quarantine.
+Specifically, individuals who are awaiting the results of a test may partially or fully self-isolate, and individuals who have tested positive are highly likely to self-isolate.
+We developed a model that incorporates mechanistic epidemic processes and testing in order to explore the effects of testing and isolation on epidemic dynamics.
 
-During COVID-19 epidemic, since isolation has applied to people including, but not limited to, those who have been tested, the epidemic dynamics will depend on who gets tested.
-That is, how tests are distributed across epidemiological compartments (susceptible-infectious-recovered) for a given testing intensity.
-Specifically, in the case where tests are randomly assigned within the population, the tests will be weighted equally across all compartments.
-Random testing strategy is the simplest form of testing and might happen, for example, when disease surveillance is the purpose of testing \citep{foddai2020base}.
-However, in the case where targeted testing is administered within the population, the testing will be more weighted towards the compartments with people who have higher probability of getting tested.
-Targeted testing might happen, for example, for different purposes including diagnosis and treatment, screening people for access to things like flights, surgery, long-term-care facilities, and for the purpose of mitigation and control of the disease spread.
-Specifically, when the purpose is "diagnosis and treatment" or "mitigation and control of disease spread", the weights of testing is skewed towards infectious compartment rather than susceptible and recovered compartments.
-Further, heavier testing weight may be put on the infected compartment in the case of diagnosis and treatment than the case of mitigation and control \ali{why?
-refs?}.
-Also, when people are tested for "screening", the individuals with potential higher mobility, eg.
-people who are getting on flights, get more tested and thus the coresponding heavier testing weight is assigned than people awaiting for a surgery and are probably going to stay in a long-term care facility and consequently less mobile and more isolated to begin with.
+Because individuals who have been tested or received a positive test are likely to self-isolate, thus reducing or eliminating  their potentially infectious contacts, the dynamics of an epidemic will depend on who gets tested and
+why. For a given testing intensity (tests performed per day), testing will have a stronger effect on infection rates
+the more strongly it is focused on people who are infectious for SARS-CoV-2, whether or not they are symptomatic.
+This level of focus depends in turn on the purpose and design of testing programs.
+(1) When testing is done for the purposes of disease surveillance \citep{foddai2020base}
+we might expect tests to be assigned randomly across the population (in an ideal world: some bias in selection for testing is probably inevitable), although a stratified sampling design would be preferred for statistical efficiency \citep{cleevely2020workable}.
+In the case where tests are randomly assigned within the population, the \emph{per capita} probability of being selected for testing will (by definition) be equal across all epidemiological compartments.
+Over the course of the COVID19 pandemic, however, the vast majority of testing has been done with other goals,
+primarily diagnosis (determining COVID19 infection status for clinical purposes) or control (determining COVID19 status
+in order to quickly isolate cases that have been found by contact tracing), which we characterize
+as \emph{targeted} testing strategies.
+In these cases, testing probabilities vary widely across epidemiological compartments; in our dynamical model,
+we will characterize these probabilities by assigning a \emph{per capita} testing weight to each compartment
+that determines the \emph{relative} probability that an individual in that compartment will be selected for
+testing (see Methods).
+When testing is used primarily for diagnosis it will focus on people with COVID19-like symptoms; thus the testing weights for people with vs.
+without SARS-CoV-2 infections will depend on the relative magnitudes of the probability of having COVID19-like symptoms when infected with SARS-CoV-2 (i.e.
+the relative proportions of presymptomatic or asymptomatic vs.
+symptomatic infections) and the probability of having COVID19-like symptoms when \emph{not} infected (e.g.
+due to other respiratory illnesses).
+Testing for epidemic control will focus on people who are known to have been in contact with known COVID19 cases; in this case the testing weights for infected vs.
+uninfected people will depend on the probability of infection given contact, as well as the thoroughness and effectiveness of the system for identifying suspicious contacts.
 
-In addition to the efect of compartment-specific testing weights on the isolation, and thus epidemic dynamics, other influential components of testing processes on isolation include test-specific sensitivities, intensity and sample-to-answer reporting time.
-They determine how accurately and how fast people are put in the isolation.
-We model these components and explore their effects on the epidemic dynamics.
-Specifically, we study the trade-offs between the testing strategies (random and targeted) and the discussed testing components on a single epidemiological-dynamics index, the ``basic reproduction number'' $\Rnum$.
+\bmb{save stuff about flights, surgery, LTC for Discussion?}
+
+In addition to the overall testing intensity, the testing weights for different compartments, and the degree of self-isolation of test-awaiting and positive-tested people,
+our models will also need to include test specificity and sensitivity, as well as the distribution of times
+from testing to reporting. For a parameter set that quantifies all of these phenomena,
+we can characterize the expected epidemiological dynamics.
+Specifically, we study the effect of different testing strategies (random vs.
+targeted), as well as testing intensity, degrees of self-isolation, and rate of test processing on a single index of epidemiological dynamics, the basic reproduction number $\Rnum$.
 
@@
 In the early stages of an epidemic the number of infected individuals is expected grow exponentially over time when $\Rnum>1$.
-To control the spread of disease in the population, $\Rnum$ is required to be below unity.
-\ali{now may be say what we expect from testing and isolation to reduce R0 and the suprise that the reporting time increase R0.
-}
+To control the spread of disease in the population, $\Rnum$ must be $<1$.
+Although the value of $\Rnum$ cannot completely characterize the dynamics of even the simplest epidemic model
+\citep{shaw2021what},
+it does give a simple and widely accepted index for the effectiveness of control, as well as some indication
+of the likely final size of an epidemic \citep{ma2006generality}. Thus, we analyze the model by deriving $\Rnum$ and describing its
+sensitivity to the parameters that describe the testing process.
 
-Here, we develop a compartmental model to incorporate epidemic processes and testing processes mechanistically to study the potential effect of testing strategies, testing intensity, test sensitivity and specificity, reporting time and isolation on epidemic dynamic.
+%% \ali{now may be say what we expect from testing and isolation to reduce R0 and the surprise that the reporting time increase R0.} \bmb{No, that belongs in the results.
+Don't give the results in the introduction!}
+
+%% Here, we develop a compartmental model to incorporate epidemic processes and testing processes mechanistically to study the potential effect of testing strategies, testing intensity, test sensitivity and specificity, reporting time and isolation on epidemic dynamic.
 We investigate the effect of underlying parameters of the model on the basic reproduction number, $\Rnum$.
@@
 This counter-intuitive effect can occur when people are cautious when awaiting a test result, and may not be robust to second-order effects of fast testing (such as better contact tracing).
-\ali{discussion maybe: all of the accessible testing data is being used for one of the other three purposes rather than surveillance. These are targeted testing strategies with a different bias.}
+%% \bmb{Most of this is either repetitive or belongs in the discussion.}
+%% \ali{discussion maybe: all of the accessible testing data is being used for one of the other three purposes rather than surveillance. These are targeted testing strategies with a different bias.}
 
@@
 In particular, doubling the sensitivity of the pre-screening tool would *do something* to $\Rnum$.}
+\bmb{possibly interesting, but \textbf{belongs in the Discussion} if anywhere.}
 % Frequent proactive testing of asymptomatic individuals offers an additional way to reduce transmission, and is among the most powerful and least disruptive of non-pharmaceutical interventions such as face masks, social distancing.
@@
 For example, PCR tests are commonly used, and very sensitive, but take longer compared to more recent tests, such as a breath-based test currently in trial stage, which can provide a result in seconds.
-\ali{discussion maybe: all of the accessible testing data is being used for one of the other three purposes rather than surveillance. These are targeted testing strategies with a different bias.}
 
@@
 We developed a deterministic model, \crefrange{eq1}{eq12}, which groups individuals based on disease status and testing status.
-The former includes Susceptible, Infectious and Recovered, thus SIR, and the latter categorizes people as \emph untested, waiting-for-\emph positive, waiting-for-\emph negative, or \emph confirmed positive.
-Also, two 'accumulator' compartments, $N$ and $P$, were incorporated in the model in order to collect cumulative reported negative or positive tests.
+Disease states include Susceptible, Infectious and Recovered (thus this is an SIR-type model), and testing status categorizes people as \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive} (Figure~\ref{fig:flowchart}).
+Two 'accumulator' compartments, $N$ and $P$, were also incorporated in the model in order to collect cumulative reported negative or positive tests.
 The model and details of calculation of $\Rnum$ are presented in the Supplementary Materials.
-The flowchart model is as follows
 
@@
 \jd{It would be nice to modularize the model somehow, but probably not worth anyone's time.}
+\bmb{Can you say what you mean by this? Just curious.}
 
-\ali{not sure if these should go to the Supp material?} In our model the following parameters are used (for a summery of parameters' description, see Table \ref{tab:params}).
-$N_0$ is the total population size, $\omega$, (1/day), is the rate of onward flow from the awaiting positive or negative compartments to reported or untested compartments, respectively, $\gamma$, (1/day), is the recovery rate, $\rho$, (1/day), is the per capita testing intensity across the whole population.
+\ali{not sure if these should go to the Supp material?}
+\bmb{I think Table 1 is sufficient, except for the details about the testing weights}
+Our model uses the following parameters (Table \ref{tab:params}): $N_0$ is the total population size, $\omega$, (1/day), is the rate of onward flow from the awaiting positive or negative compartments to reported or untested compartments, respectively, $\gamma$, (1/day), is the recovery rate, $\rho$, (1/day), is the per capita testing intensity across the whole population.
 Further, various testing strategies are incorporated through the compartment-specific relative testing weights $W_S$, $W_I$ and $W_R$.
@@
 For example, $W_I/W_S=2$ means that, the probability that infected individuals get tested is as twice as that of susceptible individuals.
-The weighted number of people available for tests is defined as $W = W_S S_u + W_I I_u + W_R R_u$, and the scaling parameter for testing defined as
+The weighted number of people available for tests is defined as $W = W_S S_u + W_I I_u + W_R R_u$, and the scaling parameter for testing is defined as
 
@@
 
-The classical SIR model is based on the following implicit assumptions (see a ref for details); well-mixed population, homogeneity of the population, exponentially distributed duration of infection and large population size (may be refer the reader to \cite{keeling2011modeling}).
+The classical SIR model is based on the following implicit assumptions; well-mixed population, homogeneity of the population, exponentially distributed duration of infection and large population size (may be refer the reader to \cite{keeling2011modeling}).
 In addition to these standard assumptions, our model, \crefrange{eq1}{eq12}, assumes: (i) there is a single force of infection, $\Lambda$, that is, where a susceptible individual moves to does not depend on who infected it, (ii) $\eta_c \leq \eta_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals, (iii) a perfectly specific test, $p_S=0$, for simplicity in analysis.
 Note that the latter assumption combined with the assumption that the total population was initially distributed between untested and awaiting for negative test, that is $S_u+S_n=0$, reduce the model to 10 equations with equations \ref{eq3} and \ref{eq4} are eliminated.
-Specifically, when the population is at steady states and the number of infected people is negligible, the assumption of perfect specificity results in $S_p=0$ in \cref{eq3} and $d S_c/dt=0$ from \cref{eq4}.
+Specifically, when the population is at steady state and the number of infected people is negligible, the assumption of perfect specificity results in $S_p=0$ in \cref{eq3} and $d S_c/dt=0$ from \cref{eq4}.
 Also, since we assumed $S_u+S_n=0$, initially $S_c=0$ and stays in this state.
 
-\newgeometry{margin=1cm} % modify this if you need even more space
-\begin{landscape}
+% \newgeometry{margin=1cm} % modify this if you need even more space
+% \begin{landscape}
 \begin{table}[htp]
 \centering
-{\tiny %
-\begin{tabular}{|c|ccc|} \hline
+{%\tiny %
+\begin{tabular}{|c|L{2in}cc|} \hline
   Symbol & Description & Unit & Value \\ \hline
   $N_0$     & Total population size & people & $10^6$ \\ \hline
-  $\omega$  & Rate of onward flow from the awaiting to reported or untested compartments  & 1/day & - \\ \hline
+  $\omega$  & Rate of onward flow from ``waiting'' to ``confirmed'' or ``untested'' compartments  & 1/day & - \\ \hline
   $\gamma$ & Recovery rate & 1/day & 1/3 \\ \hline 
   $\rho$   & Per capita testing intensity & 1/day & 0.01 \\ \hline 
-  $\eta_w$  & Relative probability of transmission for isolated awaiting individuals & - & - \\ \hline
-  $\eta_c$  & Relative probability of transmission for isolated confirmed individuals & - & -  \\ \hline
+  $\eta_w$  & Relative probability of transmission for ``waiting'' individuals & - & - \\ \hline
+  $\eta_c$  & Relative probability of transmission for ``confirmed positive'' individuals & - & -  \\ \hline
   $\Lambda$ & Force of infection & 1/day & - \\ \hline
-  $p_S$ & Probability of false positive for susceptible& - & 0 \\ \hline
-  $p_I$ & Probability of being infected and tested positive & - & 1 \\ \hline
-  $p_R$ & Probability of being recovered and tested positive & - & 0.5 \\ \hline
-  $W_S, W_I, W_R$ & Relative testing weight & - & 
+  $p_S$ & Probability of positive tests for $S$ ($= 1-\textrm{specificity}$) & - & 0 \\ \hline
+  $p_I$ & Probability of positive tests for $I$ ($= \textrm{sensitivity}$) & - & 1 \\ \hline
+  $p_R$ & Probability of positive tests for $R$ ($= 1-\textrm{specificity}$) & - & 0.5 \\ \hline
+  $W_S, W_I, W_R$ & Relative testing weights & - & 
   \begin{minipage}[t]{0.21\columnwidth}%
- Random testing: $W_S=W_I=W_R=1$, \\Non-random testing: $W_S=0.3, W_I=W_R=1$
+ Random: $\{1,1,1\}$ \\ Targeted: $\{0.3,1,1\}$
 \end{minipage} \\ \hline
@@
   }%
-\caption{\label{tab:params} The underlying parameters of model, \crefrange{eq1}{eq12}.}
+\caption{\label{tab:params} Model parameters for \crefrange{eq1}{eq12}.}
 \end{table}
-\end{landscape}
-\restoregeometry
+%\end{landscape}
+%\restoregeometry
 
@@
 
- Given that the perfect isolation occurs at $\eta_c=\eta_w = 0$, Prop.
-\ref{p1:eta} means that lifting the isolation from awaiting group results in an increased $\Rnum$ and consequently a greater number of infected individuals.
+ Given that the perfect isolation occurs at $\eta_c=\eta_w = 0$, Prop.~\ref{p1:eta} means that lifting the isolation from awaiting group results in an increased $\Rnum$ and consequently a greater number of infected individuals.
 We used Taylor approximation of $\Rnum$ at $\rho=0$ due to the complexity of the expression and thus inconclusive.
@@
  Prop.
-\ref{p1:rho} indicates that increasing testing intensity reduces the $\Rnum$.
+\ref{p1:rho} indicates that increasing testing intensity reduces $\Rnum$.
 This is sensible since as people are moved to test compartments ($I_p$, $I_n$ \fady{the susceptibles also reduce contact and transmission}) the higher probability of being subject to isolation and the lower the transmission \eqref{Lambda} will become.
 
- Of these results, prop.
-\ref{p1:omega} is the most surprising.
-It states that returning test results more rapidly (i.e., increasing $\omega$) does not necessarily lower $\Rnum$, rather, whether increasing $\omega$ lowers $\Rnum$ is defendant on the precise combination of model parameters (test reporting rate itself, testing processes/strategies ($W$'s), testing efficacy ($P_I$) and isolation ($\eta$'s).
+ \bmb{move to discussion?? this applies to this \P, also most of what's below here.
+   To a first approximation, any text that discusses the context and meaning of the results,
+   rather than the results themselves, can go in the discussion.}
+Of these results, prop.~\ref{p1:omega} is the most surprising.
+It states that returning test results more rapidly (i.e., increasing $\omega$) does not necessarily lower $\Rnum$; whether increasing $\omega$ lowers $\Rnum$ depends on the precise combination of model parameters (test reporting rate, testing processes/strategies ($W_*$), testing efficacy ($P_I$) and isolation ($\eta_*$).
 Specifically, the $\partial{R_0}/\partial{\omega}>0$ suggests that reproduction number may increase as the test reporting process become faster.
-This can be seen from expression \eqref{Rom} in the case of ``perfect isolation``, when $\eta_w=0$ and $\eta_c=0$.
+This can be seen from expression \eqref{Rom} in the case of perfect isolation, when $\eta_w=0$ and $\eta_c=0$.
 The underlying mechanism for this result is that individuals may take more precautionary behaviors (e.g., physical distancing) when they are awaiting test results compared to when they are untested or their test has returned negative.
@@
 \end{subfigure}
-\caption{Behaviour of the basic reproduction number, $\Rnum$, with respect to changes in the underlying parameters; remind the reader about the parameters and their values in this simulations.
-the solid line in each panel represents the critical contour of $\Rnum=1$.}
+\caption{Behaviour of the basic reproduction number, $\Rnum$, with respect to changes in the underlying parameters; \bmb{remind the reader about the parameters and their values}.
+The solid line in each panel represents the critical contour of $\Rnum=1$.
+\bmb{where did the legend go??
+also, we still have some collisions in the y-axis tick labels}}
 \label{pan}
@@
 
-Our numerical simulation in support of the analytical results in Proposition \ref{prop1} are presented in Figure \ref{pan}.
+Our numerical solutions in support of the analytical results in Proposition \ref{prop1} are presented in Figure \ref{pan}.
 It can be inferred that when random testing strategy is applied, Figure \ref{p.a}, comparing to the non-random testing strategy, Figure \ref{p.a}, the critical testing intensity of the corresponding panels are lower in non-random testing.
@@
 Furthermore, we derived an inequality that quantifies the exact relationships between model parameters that result in returning tests more rapidly being favorable \eqref{eq:necsuf}.
-To give a biological interpretation, we describe the qualitative \textit{trends} predicted by the inequality.
-Returning test results more rapidly \textit{tends} to be more favorable when
+To give a biological interpretation, we describe the qualitative trends predicted by the inequality.
+Returning test results more rapidly tends to be more favorable when
 
@@
 Using our Taylor approximation of $\Rnum$ near $\rho = 0$, we examined what circumstances (i.e., model parameters) make the use of one test more favourable than another, and give a complete description through inequality \ref{eq:rho1vsrho2}.
-In general, we found that the expensive test \textit{tends} to more effectively lower $\Rnum$ when
+In general, we found that the expensive test tends to more effectively lower $\Rnum$ when
 
@@
 
-- JD's email:
-We're basically looking at the potential advantage of slow tests:
+From JD's email:
+
+\begin{itemize}
+
+  \item We're basically looking at the potential advantage of slow tests:
 people waiting for test results may be more careful than those who
@@
 situation, identification of hot spots, contact-tracing, etc. 
+\item  Maybe something about the testing rate/intensity, now the per capita testing intensity is very low ($\rho \approx 0$).
+In near future new test kits may be widely accessible, our model provide insights in this case (what are those?)
+\item ``delay negative results" result
 
-- May be something about the testing rate/intensity, now the per capita testing intensity is very low ($\rho \approx 0$).
-In near future new test kits may be widely accessible, our model provide insights in this case (what are those?)
-
-- "delay negative results" result
+\end{itemize}
 
@@
 % %%%%%%%
-\bibliography{SIRlibrary.bib}
+\bibliography{SIRlibrary}
 

### 2021-01-24 SIR_manuscript.tex
a few more notes

@@
     Random: $\{1,1,1\}$ \\ Targeted: $\{0.3,1,1\}$
-\end{minipage} \\ \hline
+  \end{minipage} \\
+  \hline
 
@@
 
-Note that the block matrix $G_{12}$ does not influence $\Rnum$ defined as the spectral radius of $G$. All matters here are the eigenvalues of $G_{11}$, which are 0 and $\Rnum$ \eqref{R0}.
+The block matrix $G_{12}$ does not influence $\Rnum$ (defined as the spectral radius of $G$). All that matters here are the eigenvalues of $G_{11}$, which are 0 and $\Rnum$ \eqref{R0}.
 
@@
 \subsection{On Testing Rate and Numerical Singularity}
+
+\bmb{this should go in the main methods section, I think}
+
 The testing rate, $\sigma$, should be formulated such that people from the untested compartments will not be tested if they are not there.
@@
 
-Note that in case of ``perfect isolation``, i.e., when $\eta_w=0$ and consequently $\eta_c=0$, it is straight forward to see that $a \leq 0$, $b<0$ and $c<0$.
+In the case of ``perfect isolation'', i.e., when $\eta_w=0$ and consequently $\eta_c=0$, it is straightforward to see that $a \leq 0$, $b<0$ and $c<0$.
 Thus, $\partial{\Rnum}/\partial{\omega} \geq 0$.

### 2021-01-24 SIR_manuscript.tex
more miscellaneous ...

@@
 
-\ali{not sure if these should go to the Supp material?}
-\bmb{I think Table 1 is sufficient, except for the details about the testing weights}
-Our model uses the following parameters (Table \ref{tab:params}): $N_0$ is the total population size, $\omega$, (1/day), is the rate of onward flow from the awaiting positive or negative compartments to reported or untested compartments, respectively, $\gamma$, (1/day), is the recovery rate, $\rho$, (1/day), is the per capita testing intensity across the whole population.
-Further, various testing strategies are incorporated through the compartment-specific relative testing weights $W_S$, $W_I$ and $W_R$.
-What matters in these weight is their relative proportions.
-For example, $W_I/W_S=2$ means that, the probability that infected individuals get tested is as twice as that of susceptible individuals.
-The weighted number of people available for tests is defined as $W = W_S S_u + W_I I_u + W_R R_u$, and the scaling parameter for testing is defined as
-
+Table~\ref{tab:params} defines the model parameters, which are generally straightforward
+\emph{per capita} flows between compartments, or modifiers to these flow rates.
+The novel component of the model comes in through the compartment-specific relative testing weights $W_S$, $W_I$ and $W_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively.
+For example, $W_I/W_S=2$ means that infected individuals are tested at twice the \emph{per capita} rate of susceptible individuals.
+The weighted size of the testing pool is $W = W_S S_u + W_I I_u + W_R R_u$, and the scaling parameter for testing is
 \begin{equation}
@@
 \end{equation}
-Thus, the weighted testing rate is defined as $F_Z=\sigma W_Z$, where $Z \in \{S,I,R\}$.
-Note that requiring a fixed testing rate as in \eqref{sigma} can lead to a singularity when no one is left untested.
-We discuss this issue, and a practical solution, in the Appendix.
+Thus, \emph{per capita} testing rate for compartment $Z$ is $F_Z=\sigma W_Z$, where $Z \in \{S,I,R\}$.
+For a high-sensitivity test, infected people typically flow through to the ``confirmed positive'' ($I_c$/$R_c$) compartments and are thus unavailable for further testing.
+Over the course of the epidemic, a fixed testing rate as specified in \eqref{sigma} can
+(if large enough) exhaust the pool of people available for testing,
+leading to a singularity when no one is left untested.
+Although this phenomenon does not affect our analysis of $\Rnum$, it does sometimes occur in
+numerical solutions of the time dynamics: we discuss an adjustment to the model that solves
+this problem in the Appendix.
 
@@
 In addition to these standard assumptions, our model, \crefrange{eq1}{eq12}, assumes: (i) there is a single force of infection, $\Lambda$, that is, where a susceptible individual moves to does not depend on who infected it, (ii) $\eta_c \leq \eta_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals, (iii) a perfectly specific test, $p_S=0$, for simplicity in analysis.
-Note that the latter assumption combined with the assumption that the total population was initially distributed between untested and awaiting for negative test, that is $S_u+S_n=0$, reduce the model to 10 equations with equations \ref{eq3} and \ref{eq4} are eliminated.
-Specifically, when the population is at steady state and the number of infected people is negligible, the assumption of perfect specificity results in $S_p=0$ in \cref{eq3} and $d S_c/dt=0$ from \cref{eq4}.
-Also, since we assumed $S_u+S_n=0$, initially $S_c=0$ and stays in this state.
+The latter assumption, combined with the assumption that the total population was initially distributed between untested and awaiting for negative test, that is $S_u(0)+S_n(0)=N(0)$ \bmb{I don't think this is what we need: I think we need $S_p(0)=S_c(0)=0$?
+The previous assumption is awkward because we must have some infected individuals at time 0, and is more than we need (we could start with some $R$ without loss of generality)}, reduces the model to 10 equations (equations \ref{eq3} and \ref{eq4} are eliminated).
+Specifically, when the population is at steady state and the number of infected people is negligible \bmb{do we really need anything other than the previously stated conditions?}, the assumption of perfect specificity results in $S_p=0$ in \cref{eq3} and $d S_c/dt=0$ from \cref{eq4}.
+Also, since we assumed $S_u+S_n=0$, $S_c=0$ and stays in this state.
 

### 2021-01-25 SIR_manuscript.tex
Just a couple minor changes before an interruption

@@
 The observed dynamics of the COVID-19 epidemic are driven by both epidemiological processes (infection and recovery) and testing processes (testing and test reporting).
-In addition to determining what we actually observe about the epidemic (case reports), testing processes can also feed back to affect the epidemiological dynamics through isolation and quarantine.
-Specifically, individuals who are awaiting the results of a test may partially or fully self-isolate, and individuals who have tested positive are highly likely to self-isolate.
+In addition to shaping epidemic observations (via case reports), testing processes can also affect epidemiological dynamics.
+In particular, individuals who are awaiting the results of a test may partially or fully self-isolate, and individuals who have tested positive are highly likely to self-isolate.
 We developed a model that incorporates mechanistic epidemic processes and testing in order to explore the effects of testing and isolation on epidemic dynamics.

### 2021-01-25 SIR_manuscript.tex
Quickly through intro

@@
 why. For a given testing intensity (tests performed per day), testing will have a stronger effect on infection rates
-the more strongly it is focused on people who are infectious for SARS-CoV-2, whether or not they are symptomatic.
+the more strongly it is focused on people who are infectious for SARS-CoV-2.
 This level of focus depends in turn on the purpose and design of testing programs.
@@
 we might expect tests to be assigned randomly across the population (in an ideal world: some bias in selection for testing is probably inevitable), although a stratified sampling design would be preferred for statistical efficiency \citep{cleevely2020workable}.
-In the case where tests are randomly assigned within the population, the \emph{per capita} probability of being selected for testing will (by definition) be equal across all epidemiological compartments.
-Over the course of the COVID19 pandemic, however, the vast majority of testing has been done with other goals,
-primarily diagnosis (determining COVID19 infection status for clinical purposes) or control (determining COVID19 status
+\jd{This is a bit confusing; cleevely is specifically about control, and not just surveillance.}
+%% In the case where tests are randomly assigned within the population, the \emph{per capita} probability of being selected for testing will (by definition) be equal across all epidemiological compartments.
+\jd{This sentence seems unnecessary to me.}
+Over the course of the COVID19 pandemic, however, the vast majority of testing has been done with other goals --
+primarily diagnosis (determining COVID19 infection status for clinical purposes), or control (determining COVID19 status
 in order to quickly isolate cases that have been found by contact tracing), which we characterize
@@
 testing (see Methods).
+
+\jd{note that we have not yet introduced the term ``testing weights''.}
 When testing is used primarily for diagnosis it will focus on people with COVID19-like symptoms; thus the testing weights for people with vs.
@@
 
-In addition to the overall testing intensity, the testing weights for different compartments, and the degree of self-isolation of test-awaiting and positive-tested people,
-our models will also need to include test specificity and sensitivity, as well as the distribution of times
-from testing to reporting. For a parameter set that quantifies all of these phenomena,
-we can characterize the expected epidemiological dynamics.
-Specifically, we study the effect of different testing strategies (random vs.
-targeted), as well as testing intensity, degrees of self-isolation, and rate of test processing on a single index of epidemiological dynamics, the basic reproduction number $\Rnum$.
+In addition to the overall testing intensity, the testing weights for different compartments, and the degree of self-isolation of test-pending and test-confirmed people,
+our models will also need to include test specificity and sensitivity, as well as the distribution of times from testing to reporting. 
+\jd{This seems weak. Why are we studying \Rnum? What does it cover and what not?}
+Here, we focus on the the effect of testing intensity, different levels of testing ``focus'' (from random to highly targeted), and rate of test processing on a single index of epidemiological dynamics, the basic reproduction number $\Rnum$.
 
@@
 \citep{shaw2021what},
-it does give a simple and widely accepted index for the effectiveness of control, as well as some indication
-of the likely final size of an epidemic \citep{ma2006generality}. Thus, we analyze the model by deriving $\Rnum$ and describing its
-sensitivity to the parameters that describe the testing process.
+it does give a simple and widely accepted index for the difficulty of control, as well as some indication
+of the likely final size of an epidemic \citep{ma2006generality}.
 

### 2021-01-25 SIR_manuscript.tex
some progress with Methods

@@
 \newcommand{\Rlogo}{R\xspace}
+\newcommand{\percap}{\emph{per capita}\xspace}
 
@@
 \bmb{Can you say what you mean by this? Just curious.}
+\jd{I wrote this when the equations were here; I meant it might be good to use less equations, by writing summaries of the SIR process and the within-each-box testing process.}
 
@@
 For example, $W_I/W_S=2$ means that infected individuals are tested at twice the \emph{per capita} rate of susceptible individuals.
-The weighted size of the testing pool is $W = W_S S_u + W_I I_u + W_R R_u$, and the scaling parameter for testing is
+
+In order to link to more applied models, we constructed this model so we could specify  to be able to specify the total \percap testing rate.
+We do this by defining the weighted size of the testing pool $W = W_S S_u + W_I I_u + W_R R_u$, and calculating the scaling parameter for testing as:
 \begin{equation}
@@
 
-The classical SIR model is based on the following implicit assumptions; well-mixed population, homogeneity of the population, exponentially distributed duration of infection and large population size (may be refer the reader to \cite{keeling2011modeling}).
-In addition to these standard assumptions, our model, \crefrange{eq1}{eq12}, assumes: (i) there is a single force of infection, $\Lambda$, that is, where a susceptible individual moves to does not depend on who infected it, (ii) $\eta_c \leq \eta_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals, (iii) a perfectly specific test, $p_S=0$, for simplicity in analysis.
-The latter assumption, combined with the assumption that the total population was initially distributed between untested and awaiting for negative test, that is $S_u(0)+S_n(0)=N(0)$ \bmb{I don't think this is what we need: I think we need $S_p(0)=S_c(0)=0$?
-The previous assumption is awkward because we must have some infected individuals at time 0, and is more than we need (we could start with some $R$ without loss of generality)}, reduces the model to 10 equations (equations \ref{eq3} and \ref{eq4} are eliminated).
-Specifically, when the population is at steady state and the number of infected people is negligible \bmb{do we really need anything other than the previously stated conditions?}, the assumption of perfect specificity results in $S_p=0$ in \cref{eq3} and $d S_c/dt=0$ from \cref{eq4}.
+The classical SIR model is based on the following implicit assumptions; well-mixed population, homogeneity of the population, exponentially distributed duration of infection and large population size (see, e.g., \cite{keeling2011modeling}).
+In addition to these standard assumptions, our model, \crefrange{eq1}{eq12}, assumes: (i) there is a single force of infection, $\Lambda$, across all susceptible pools [\jd{This is de-clarifying here; let's drop it: that is, where a susceptible individual moves to does not depend on who infected it], (ii) $\eta_c \leq \eta_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
+For this analysis, we also assume a perfectly specific test,($p_S=0$).
+This last assumption
+%% The latter assumption, combined with the assumption that the total population was initially distributed between untested and awaiting for negative test, that is $S_u(0)+S_n(0)=N(0)$ \bmb{I don't think this is what we need: I think we need $S_p(0)=S_c(0)=0$?
+The previous assumption is awkward because we must have some infected individuals at time 0, and is more than we need (we could start with some $R$ without loss of generality)},
+reduces the model to 10 equations (equations \ref{eq3} and \ref{eq4} are eliminated).
+%% Specifically, when the population is at steady state and the number of infected people is negligible \bmb{do we really need anything other than the previously stated conditions?}, the assumption of perfect specificity results in $S_p=0$ in \cref{eq3} and $d S_c/dt=0$ from \cref{eq4}.
 Also, since we assumed $S_u+S_n=0$, $S_c=0$ and stays in this state.

### 2021-01-25 SIR_manuscript.tex
notes from BMB meeting (various small fixes)

--- 2021-01-25 SIR_manuscript.tex
@@
 The observed dynamics of the COVID-19 epidemic are driven by both epidemiological processes (infection and recovery) and testing processes (testing and test reporting).
-In addition to shaping epidemic observations (via case reports), testing processes can also affect epidemiological dynamics.
-In particular, individuals who are awaiting the results of a test may partially or fully self-isolate, and individuals who have tested positive are highly likely to self-isolate.
+In addition to determining what we actually observe about the epidemic (case reports), testing processes can also feed back to affect the epidemiological dynamics through isolation and quarantine.
+Specifically, individuals who are awaiting the results of a test may partially or fully self-isolate, and individuals who have tested positive are highly likely to self-isolate.
 We developed a model that incorporates mechanistic epidemic processes and testing in order to explore the effects of testing and isolation on epidemic dynamics.
@@
 We developed a deterministic model, \crefrange{eq1}{eq12}, which groups individuals based on disease status and testing status.
-Disease states include Susceptible, Infectious and Recovered (thus this is an SIR-type model), and testing status categorizes people as \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive} (Figure~\ref{fig:flowchart}).
-Two 'accumulator' compartments, $N$ and $P$, were also incorporated in the model in order to collect cumulative reported negative or positive tests.
+Disease states include Susceptible, Infectious and Recovered (thus this is an SIR-type model), and testing status categorizes people as \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive} (Figure~\ref{fig:flowchart}) \bmb{make sure we're using consistent terminology for the testing compartments}.
+Two `accumulator' compartments, $N$ and $P$, were also incorporated in the model in order to collect cumulative reported negative or positive tests.
 The model and details of calculation of $\Rnum$ are presented in the Supplementary Materials.
+
+\bmb{need a little bit more detail moved from the Supp.
+(1) If equations are in the supp we need to tag the equation numbers (e.g.
+eq S1-S14); (2) we might want the whole set of equations to have a single number (we could also sublabel S1(a), S1(b) etc., but as we never refer to the individual equations we probably don't need to bother); (3) at the very least, $\Lambda$ should be defined in the body of the text (we need it to understand the figure), and $\beta$ should be included in the parameter table}
 
@@
    rather than the results themselves, can go in the discussion.}
+
 Of these results, prop.~\ref{p1:omega} is the most surprising.
@@
 
-[comments about the figure:]
-1- use a single legend, change the range manually.
 \begin{figure}[h!]
@@
 The solid line in each panel represents the critical contour of $\Rnum=1$.
-\bmb{where did the legend go??
-also, we still have some collisions in the y-axis tick labels}}
+\bmb{we still have some collisions in the y-axis tick labels}}
 \label{pan}
@@
 
-\bmb{this should go in the main methods section, I think}
+\bmb{need a little bit of context here, saying that we didn't do any numerical solutions for the trajectories in our analysis,
+  but if you do try them you run into this problem ...}
 

### 2021-01-25 codes/sir_plot.R

@@ -49,11 +49,12 @@ brks_vec <- seq(0.8,1.05,by=0.05) # Break vector for unifying the legends in Ran
 
 df_temp <- df_random
 # df_temp <- df_targeted
-p1 <- ggplot(df_temp,aes(x=omega,y=rho,z=R0_sub))+ theme_bw() +
-  xlab(TeX('$\\omega$, rate of test return (1/day)')) +
-  ylab(TeX('$\\rho$, testing intensity (1/day per capita)')) +
-  # theme(panel.spacing=grid::unit(0,"lines"),legend.position = c(0.8, 0.75)) # initial theme, comment it for TTI testing plot
-  theme(panel.spacing=grid::unit(0,"lines"),legend.position = "none") #AG, uncomment for Random Case
+p1 <- (ggplot(df_temp,aes(x=omega,y=rho,z=R0_sub))
+    + theme_bw()
+    + xlab(TeX('$\\omega$, rate of test return (1/day)'))
+    + ylab(TeX('$\\rho$, testing intensity (1/day per capita)'))
+)
+
 p1_temp <- (p1
             + geom_contour_filled(breaks=brks_vec)
             + geom_contour(breaks=1,alpha=0.5,colour="black")
@@ -67,13 +68,15 @@ p1_temp <- (p1
                         ymax=10,
                         xmin=-1,
                         xmax=2)
+    + theme(panel.spacing=grid::unit(0,"lines"))
 )
 
 # #################################
 # 1. Plot the Random Testing Scenario:
 # #################################
 
-ggsave(p1_temp + ggtitle(TeX("W_S=W_I=W_R=1")),
+ggsave(p1_temp + ggtitle(TeX("W_S=W_I=W_R=1")) +
+       theme(legend.position = "none"),
        filename = "R0contour_random.pdf" ,
        width = 12, height = 12, units = "cm",
        path=path)
@@ -83,10 +86,11 @@ ggsave(p1_temp + ggtitle(TeX("W_S=W_I=W_R=1")),
 # #################################
 
 ggsave((p1_temp %+% df_targeted) +
-        ggtitle(TeX(sprintf("W_S=%.1f, W_I=W_R=1",W_S_targeted ))),
-    filename = "R0contour_TTI.pdf" ,
-    width = 12, height = 12, units = "cm",
-    path=path)
+       ggtitle(TeX(sprintf("W_S=%.1f, W_I=W_R=1",W_S_targeted))) +
+       theme(legend.position = c(0.8, 0.75)),
[…]

### 2021-01-25 SIR_manuscript.tex
fix tex breakage

@@
 our models will also need to include test specificity and sensitivity, as well as the distribution of times from testing to reporting. 
-\jd{This seems weak. Why are we studying \Rnum? What does it cover and what not?}
+\jd{This seems weak. Why are we studying $\Rnum$? What does it cover and what not?}
 Here, we focus on the the effect of testing intensity, different levels of testing ``focus'' (from random to highly targeted), and rate of test processing on a single index of epidemiological dynamics, the basic reproduction number $\Rnum$.
@@
 The classical SIR model is based on the following implicit assumptions; well-mixed population, homogeneity of the population, exponentially distributed duration of infection and large population size (see, e.g., \cite{keeling2011modeling}).
-In addition to these standard assumptions, our model, \crefrange{eq1}{eq12}, assumes: (i) there is a single force of infection, $\Lambda$, across all susceptible pools [\jd{This is de-clarifying here; let's drop it: that is, where a susceptible individual moves to does not depend on who infected it], (ii) $\eta_c \leq \eta_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
+In addition to these standard assumptions, our model, \crefrange{eq1}{eq12}, assumes: (i) there is a single force of infection, $\Lambda$, across all susceptible pools \jd{This is de-clarifying here; let's drop it: that is, where a susceptible individual moves to does not depend on who infected it}, (ii) $\eta_c \leq \eta_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
 For this analysis, we also assume a perfectly specific test,($p_S=0$).

### 2021-01-27 SIR_manuscript.tex
comments addressed in Method

@@
 \usepackage{xspace}
+
 %\newcommand{\Rlogo}{\protect\includegraphics[height=2ex,keepaspectratio]{pix/Rlogo.pdf}\xspace}
@@
 
-We developed a deterministic model, \crefrange{eq1}{eq12}, which groups individuals based on disease status and testing status.
-Disease states include Susceptible, Infectious and Recovered (thus this is an SIR-type model), and testing status categorizes people as \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive} (Figure~\ref{fig:flowchart}) \bmb{make sure we're using consistent terminology for the testing compartments}.
+We developed a deterministic model, Eqs.
+\eqref{model}, which groups individuals based on disease status and testing status.
+Disease states include Susceptible, Infectious and Recovered (thus this is an SIR-type model), and testing status categorizes people as \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive} (Figure~\ref{fig:flowchart}).
 Two `accumulator' compartments, $N$ and $P$, were also incorporated in the model in order to collect cumulative reported negative or positive tests.
 The model and details of calculation of $\Rnum$ are presented in the Supplementary Materials.
-
-\bmb{need a little bit more detail moved from the Supp.
-(1) If equations are in the supp we need to tag the equation numbers (e.g.
-eq S1-S14); (2) we might want the whole set of equations to have a single number (we could also sublabel S1(a), S1(b) etc., but as we never refer to the individual equations we probably don't need to bother); (3) at the very least, $\Lambda$ should be defined in the body of the text (we need it to understand the figure), and $\beta$ should be included in the parameter table}
 
@@
 \includegraphics[scale=0.3]{./pix/sir_comp.png}
-\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model, \crefrange{eq1}{eq12}. For the description of the parameters see \cref{tab:params}.
+\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model, \ref{model}. For the description of the parameters see \cref{tab:params}.
 \label{fig:flowchart}}
@@
 
-\jd{It would be nice to modularize the model somehow, but probably not worth anyone's time.}
-\bmb{Can you say what you mean by this? Just curious.}
-\jd{I wrote this when the equations were here; I meant it might be good to use less equations, by writing summaries of the SIR process and the within-each-box testing process.}
+Table~\ref{tab:params} defines the model parameters, which are generally straightforward
+\percap flows between compartments, or modifiers to these flow rates.
+The novel component of the model comes in through the compartment-specific relative testing weights $W_S$, $W_I$ and $W_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively.
+For example, $W_I/W_S=2$ means that infected individuals are tested at twice the \percap rate of susceptible individuals.
 
-Table~\ref{tab:params} defines the model parameters, which are generally straightforward
-\emph{per capita} flows between compartments, or modifiers to these flow rates.
-The novel component of the model comes in through the compartment-specific relative testing weights $W_S$, $W_I$ and $W_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively.
-For example, $W_I/W_S=2$ means that infected individuals are tested at twice the \emph{per capita} rate of susceptible individuals.
-
-In order to link to more applied models, we constructed this model so we could specify  to be able to specify the total \percap testing rate.
+In order to link to more applied models, we constructed this model so we could specify the total \percap testing rate.
 We do this by defining the weighted size of the testing pool $W = W_S S_u + W_I I_u + W_R R_u$, and calculating the scaling parameter for testing as:
@@
 \end{equation}
-Thus, \emph{per capita} testing rate for compartment $Z$ is $F_Z=\sigma W_Z$, where $Z \in \{S,I,R\}$.
+Thus, \percap testing rate for compartment $Z$ is $F_Z=\sigma W_Z$, where $Z \in \{S,I,R\}$.
 For a high-sensitivity test, infected people typically flow through to the ``confirmed positive'' ($I_c$/$R_c$) compartments and are thus unavailable for further testing.
@@
 numerical solutions of the time dynamics: we discuss an adjustment to the model that solves
-this problem in the Appendix.
+this problem in the Supplementary material.
 
 The classical SIR model is based on the following implicit assumptions; well-mixed population, homogeneity of the population, exponentially distributed duration of infection and large population size (see, e.g., \cite{keeling2011modeling}).
-In addition to these standard assumptions, our model, \crefrange{eq1}{eq12}, assumes: (i) there is a single force of infection, $\Lambda$, across all susceptible pools \jd{This is de-clarifying here; let's drop it: that is, where a susceptible individual moves to does not depend on who infected it}, (ii) $\eta_c \leq \eta_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
+In addition to these standard assumptions, our model, \ref{model}, assumes: (i) there is a single force of infection, $\Lambda$, defined as follows
+\begin{equation}
+\label{Lambda}
+\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_c I_c)}{N_0},
+\end{equation}
+across all susceptible pools (see Table~\ref{tab:params} for detail), (ii) $\eta_c \leq \eta_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
 For this analysis, we also assume a perfectly specific test,($p_S=0$).
-This last assumption
-%% The latter assumption, combined with the assumption that the total population was initially distributed between untested and awaiting for negative test, that is $S_u(0)+S_n(0)=N(0)$ \bmb{I don't think this is what we need: I think we need $S_p(0)=S_c(0)=0$?
-The previous assumption is awkward because we must have some infected individuals at time 0, and is more than we need (we could start with some $R$ without loss of generality)},
-reduces the model to 10 equations (equations \ref{eq3} and \ref{eq4} are eliminated).
-%% Specifically, when the population is at steady state and the number of infected people is negligible \bmb{do we really need anything other than the previously stated conditions?}, the assumption of perfect specificity results in $S_p=0$ in \cref{eq3} and $d S_c/dt=0$ from \cref{eq4}.
-Also, since we assumed $S_u+S_n=0$, $S_c=0$ and stays in this state.
+This last assumption combined with the assumption that no individual is in waiting-for-\emph{positive} or \emph{confirmed positive} compartments, i.e., $S_p(0)=S_c(0)=0$, reduces the model to 10 equations (equations c and d  in \eqref{model} are eliminated).
 
-% \newgeometry{margin=1cm} % modify this if you need even more space
-% \begin{landscape}
 \begin{table}[htp]
 \centering
-{%\tiny %
 \begin{tabular}{|c|L{2in}cc|} \hline
@@
   $\eta_c$  & Relative probability of transmission for ``confirmed positive'' individuals & - & -  \\ \hline
+  $\beta$ & Transmission rate & 1/day & 0.39 \\ \hline
   $\Lambda$ & Force of infection & 1/day & - \\ \hline
@@
   \hline
+  \end{tabular}
+\caption{\label{tab:params} Parameters of the model \eqref{model}.}
+\end{table}
 
-  \end{tabular}
-  }%
-\caption{\label{tab:params} Model parameters for \crefrange{eq1}{eq12}.}
-\end{table}
-%\end{landscape}
-%\restoregeometry
-
-The Disease-Free Equilibrium (DFE) for the SIR model, \crefrange{eq1}{eq12}, is given by solving the coupled system including $S_u+S_n=N_0$ and $F_S S_u-\omega S_n=0$. The DFE is
+The Disease-Free Equilibrium (DFE) for the SIR model, Eqs. \eqref{model}, is given by solving the coupled system including $S_u+S_n=N_0$ and $F_S S_u-\omega S_n=0$. The DFE is
 
@@
 
-Further details are provided in the Appendix.
+Further details are provided in the Supplementary material section.
 
-\todo{Something needs to be said here about simulation part.} The analytical calculation of the next generation matrix, $G$, and the derivation of $\Rnum$ was carried out in Maple (ref) by using simple linear Algebra package.
-  We used \Rlogo for the simulation part which included using the explicit expression of $\Rnum$ \ref{R0} as a function of the underlying parameters, specifying the parameters a realistic range, calculating the corresponding $\Rnum$ and plotting the contours of $\Rnum$ by using ggplot package in \Rlogo.
+The analytical calculation of the next generation matrix, $G$, and the derivation of $\Rnum$ was carried out in Maple (ref?) by using simple linear Algebra package. 
+  We used R for the simulation part which included using the explicit expression of $\Rnum$ \ref{R0} as a function of the underlying parameters, specifying the parameters a realistic range, calculating the corresponding $\Rnum$ and plotting the contours of $\Rnum$ by using ggplot package in R.
 The related result is presented in Figure \ref{pan}, where  panel \eqref{p.a} represents the random testing, and panel \eqref{p.b} is representing non-random testing.
@@
 
-% %%%%%%%
-\section{Supplementary Material}
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\clearpage
+% \widetext
+\begin{center}
+\textbf{\large Supplemental Materials: Title for main text}
+\end{center}
+%%%%%%%%%% Merge with supplemental materials %%%%%%%%%%
+%%%%%%%%%% Prefix a "S" to all equations, figures, tables and reset the counter %%%%%%%%%%
+\setcounter{equation}{0}
+\setcounter{figure}{0}
+\setcounter{table}{0}
+\setcounter{page}{1}
+\makeatletter
+\renewcommand{\theequation}{S\arabic{equation}}
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\bibnumfmt}[1]{[S#1]}
+\renewcommand{\citenumfont}[1]{S#1}
+%%%%%%%%%% Prefix a "S" to all equations, figures, tables and reset the counter %%%%%%%%%%
 
-\subsection{Model and Calculation of $\Rnum$}
+\subsection{Model and calculation of $\Rnum$}
 
 The model is 
+\begin{subequations}\label{model}
 \begin{align}
-% \label{model}
- d S_u/dt &= -\Lambda S_u - F_S S_u + \omega S_n, \label{eq1}\\
- d S_n/dt &= -\Lambda S_n + (1-p_S) F_S S_u - \omega S_n, \label{eq2}\\
- d S_p/dt &= -\Lambda S_p + p_S F_S S_u - \omega S_p, \label{eq3}\\
- d S_c/dt &= -\Lambda S_c + \omega S_p, \label{eq4}\\
- d I_u/dt &= \Lambda S_u - F_I I_u + \omega I_n  - \gamma I_u,  \label{eq5}\\
- d I_n/dt &= \Lambda S_n + (1-p_I) F_I I_u - \omega I_n -\gamma I_n, \label{eq6}\\
- d I_p/dt &= \Lambda S_p + p_I F_I I_u - \omega I_p -\gamma I_p, \label{eq7}\\
- d I_c/dt &= \Lambda S_c + \omega I_p - \gamma I_c,  \label{eq8}\\
- d R_u/dt &= \gamma I_u - F_R R_u + \omega R_n, \label{eq9}\\
- d R_n/dt &= \gamma I_n + (1-p_R) F_R R_u - \omega R_n,  \label{eq10}\\
- d R_p/dt &= \gamma I_p + p_R F_R R_u  - \omega R_p,  \label{eq11}\\
- d R_c/dt&= \gamma I_c + \omega R_p,  \label{eq12}\\
- dN/dt &= \omega (S_n + I_n + R_n),   \label{eq13}\\
- dP/dt &= \omega(I_p + R_p) \label{eq14},
+ d S_u/dt &= -\Lambda S_u - F_S S_u + \omega S_n, \\
+ d S_n/dt &= -\Lambda S_n + (1-p_S) F_S S_u - \omega S_n, \\
+ d S_p/dt &= -\Lambda S_p + p_S F_S S_u - \omega S_p, \\
+ d S_c/dt &= -\Lambda S_c + \omega S_p, \\
+ d I_u/dt &= \Lambda S_u - F_I I_u + \omega I_n  - \gamma I_u, \\
+ d I_n/dt &= \Lambda S_n + (1-p_I) F_I I_u - \omega I_n -\gamma I_n, \\
+ d I_p/dt &= \Lambda S_p + p_I F_I I_u - \omega I_p -\gamma I_p, \\
+ d I_c/dt &= \Lambda S_c + \omega I_p - \gamma I_c, \\
+ d R_u/dt &= \gamma I_u - F_R R_u + \omega R_n, \\
+ d R_n/dt &= \gamma I_n + (1-p_R) F_R R_u - \omega R_n, \\
+ d R_p/dt &= \gamma I_p + p_R F_R R_u  - \omega R_p, \\
+ d R_c/dt&= \gamma I_c + \omega R_p, \\
+ dN/dt &= \omega (S_n + I_n + R_n),  \\
+ dP/dt &= \omega(I_p + R_p) ,
 \end{align}
+\end{subequations}
 
 where $\beta$, (in units of 1/day), is the disease transmission rate, $\eta_w$ and $\eta_c$ are the isolation parameters for awaiting and reported individuals, respectively.
-$\Lambda$ is the force of infection and is defined as
-
-\begin{equation}
-\label{Lambda}
-\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_c I_c)}{N_0},
-\end{equation}
-
-% %
+$\Lambda$ is the force of infection defined in Eq.
+\eqref{Lambda}.
 

### 2021-02-01 SIR_manuscript.tex
working on results and discussion

@@
 
-%\newcommand{\Rlogo}{\protect\includegraphics[height=2ex,keepaspectratio]{pix/Rlogo.pdf}\xspace}
-% BMB: we don't need the weird graphical logo ...
 \newcommand{\Rlogo}{R\xspace}
 \newcommand{\percap}{\emph{per capita}\xspace}
-
+\newcommand{\Rnum}{\mathcal{R}_0}
 \newcommand{\comment}{\showcomment}
-%% \newcommand{\comment}{\nocomment}
-
 \newcommand{\showcomment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}
@@
 
-\newcommand{\Rnum}{\mathcal{R}_0}
 \theoremstyle{definition} % amsthm only
@@
 (1) When testing is done for the purposes of disease surveillance \citep{foddai2020base}
-we might expect tests to be assigned randomly across the population (in an ideal world: some bias in selection for testing is probably inevitable), although a stratified sampling design would be preferred for statistical efficiency \citep{cleevely2020workable}.
-\jd{This is a bit confusing; cleevely is specifically about control, and not just surveillance.}
-%% In the case where tests are randomly assigned within the population, the \emph{per capita} probability of being selected for testing will (by definition) be equal across all epidemiological compartments.
-\jd{This sentence seems unnecessary to me.}
-Over the course of the COVID19 pandemic, however, the vast majority of testing has been done with other goals --
-primarily diagnosis (determining COVID19 infection status for clinical purposes), or control (determining COVID19 status
-in order to quickly isolate cases that have been found by contact tracing), which we characterize
-as \emph{targeted} testing strategies.
-In these cases, testing probabilities vary widely across epidemiological compartments; in our dynamical model,
-we will characterize these probabilities by assigning a \emph{per capita} testing weight to each compartment
-that determines the \emph{relative} probability that an individual in that compartment will be selected for
-testing (see Methods).
+we might expect tests to be assigned randomly across the population (in an ideal world: some bias in selection for testing is probably inevitable), although a stratified sampling design would be preferred for statistical efficiency.
 
-\jd{note that we have not yet introduced the term ``testing weights''.}
+Over the course of the COVID19 \ali{consistency COVID-19 or COVID19?} pandemic, however, the vast majority of testing has been done with other goals --
+primarily diagnosis (determining COVID19 infection status for clinical purposes), or control (determining COVID19 status in order to quickly isolate cases that have been found by contact tracing), which we characterize as \emph{targeted} testing strategies.
+In these cases, testing probabilities vary widely across epidemiological compartments; in our dynamical model, we will characterize these probabilities by assigning a \percap testing weight to each compartment that determines the \emph{relative} probability that an individual in that compartment will be selected for testing (see Methods).
+
 When testing is used primarily for diagnosis it will focus on people with COVID19-like symptoms; thus the testing weights for people with vs.
@@
 
-\bmb{save stuff about flights, surgery, LTC for Discussion?}
-
-In addition to the overall testing intensity, the testing weights for different compartments, and the degree of self-isolation of test-pending and test-confirmed people,
-our models will also need to include test specificity and sensitivity, as well as the distribution of times from testing to reporting. 
-\jd{This seems weak. Why are we studying $\Rnum$? What does it cover and what not?}
+In addition to the overall testing intensity, the testing weights for different compartments, and the degree of self-isolation of test-pending and test-confirmed people, our models will also need to include test specificity and sensitivity, as well as the distribution of times from testing to reporting.
 Here, we focus on the the effect of testing intensity, different levels of testing ``focus'' (from random to highly targeted), and rate of test processing on a single index of epidemiological dynamics, the basic reproduction number $\Rnum$.
@@
 Although the value of $\Rnum$ cannot completely characterize the dynamics of even the simplest epidemic model
-\citep{shaw2021what},
-it does give a simple and widely accepted index for the difficulty of control, as well as some indication
-of the likely final size of an epidemic \citep{ma2006generality}.
-
-%% \ali{now may be say what we expect from testing and isolation to reduce R0 and the surprise that the reporting time increase R0.} \bmb{No, that belongs in the results.
-Don't give the results in the introduction!}
-
-%% Here, we develop a compartmental model to incorporate epidemic processes and testing processes mechanistically to study the potential effect of testing strategies, testing intensity, test sensitivity and specificity, reporting time and isolation on epidemic dynamic.
-We investigate the effect of underlying parameters of the model on the basic reproduction number, $\Rnum$.
-While targeted testing strategies are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is viral spread is expected to be stronger for a slow test than for a fast test.
-This counter-intuitive effect can occur when people are cautious when awaiting a test result, and may not be robust to second-order effects of fast testing (such as better contact tracing).
-%% \bmb{Most of this is either repetitive or belongs in the discussion.}
-%% \ali{discussion maybe: all of the accessible testing data is being used for one of the other three purposes rather than surveillance. These are targeted testing strategies with a different bias.}
-
-\fady{Not sure if interesting: \\
-  one point we can make when discussing the relative weight of testing in different compartments is to discuss pre-testing screening tools, such as surveys or questionnaires.
-If we have a quantitative description of how the testing intensities affect the dynamics, we can make statements like ``our results suggest that employing a pre-testing screening tool can help target infected individuals more effectively.
-In particular, doubling the sensitivity of the pre-screening tool would *do something* to $\Rnum$.}
-\bmb{possibly interesting, but \textbf{belongs in the Discussion} if anywhere.}
-% Frequent proactive testing of asymptomatic individuals offers an additional way to reduce transmission, and is among the most powerful and least disruptive of non-pharmaceutical interventions such as face masks, social distancing.
-% \bmb{As JD suggests, the focus of the paper is not on unraveling data (although that is something we \emph{could} do, and are trying to do, with this modeling framework as well).
-I would say that the ``bullet points'' to start the paper are (1) the processes driving the observed dynamics of the COVID-19 epidemic include both epidemiological processes (infection, recovery, etc.) and testing processes (``processes'' is repeated here, could rephrase); (2) in addition to determining what is actually observed (case reports), testing processes can also feed back to affect the epidemiological dynamics through isolation and quarantine; (3) your last sentence in the \P\ above.}
-
-% - Speed of test; 
-% Much effort has been invested in developing widely accessible and fast tests in order to mitigate the spread of the pandemic (refs).
-For example, PCR tests are commonly used, and very sensitive, but take longer compared to more recent tests, such as a breath-based test currently in trial stage, which can provide a result in seconds.
+\citep{shaw2021what}, it does give a simple and widely accepted index for the difficulty of control, as well as some indication of the likely final size of an epidemic \citep{ma2006generality}.
 
@@
 \label{dfe}
-S_n^*= \frac{\rho}{\omega} N_0, \ S_u^*= N_0-S_n^*, \text{and} I_j=R_j=0 \ \text{for all j}.
+S_n^*= \frac{\rho}{\omega} N_0, \ S_u^*= N_0-S_n^*, \text{~and~} I_j=R_j=0 \ \text{for all j}.
 \end{equation}
@@
 \end{align*}
-
-Further details are provided in the Supplementary material section.
+Further details of derivation of $\Rnum$ are provided in the Supplementary material section.
 
@@
 The explicit formula for the basic reproduction number, $\Rnum$ \eqref{R0}, provides an opportunity to study the influence of changes in the underlying parameters on the critical index of epidemic dynamics.
-We are interested in understanding the effect of parameters that can be realistically controlled by changing in isolation, testing intensity and test resulting, i.e., $\eta_c$ and $\eta_w$, $\rho$ and $\omega$, respectively.
+We are interested in understanding the effect of parameters that can be realistically controlled by changing in isolation, \percap testing intensity and test resulting, i.e., $\eta_c$ and $\eta_w$, $\rho$ and $\omega$, respectively.
 The following Proposition is the direct result of taking partial derivative of $\Rnum$ \eqref{R0} with respect to selected parameters.
+Note that we used Taylor approximation of $\Rnum$ at $\rho=0$ due to the complexity of the expression and thus inconclusive.
+See the Supplementary material for details.
 
@@
  Given that the perfect isolation occurs at $\eta_c=\eta_w = 0$, Prop.~\ref{p1:eta} means that lifting the isolation from awaiting group results in an increased $\Rnum$ and consequently a greater number of infected individuals.
-We used Taylor approximation of $\Rnum$ at $\rho=0$ due to the complexity of the expression and thus inconclusive.
-See the Appendix for details.
-It is straightforward to derive \ref{p1:rho} and \ref{p1:omega} from the linear approximation of $\Rnum$ at $\rho=0$.
-
- Prop.
-\ref{p1:rho} indicates that increasing testing intensity reduces $\Rnum$.
-This is sensible since as people are moved to test compartments ($I_p$, $I_n$ \fady{the susceptibles also reduce contact and transmission}) the higher probability of being subject to isolation and the lower the transmission \eqref{Lambda} will become.
-
- \bmb{move to discussion?? this applies to this \P, also most of what's below here.
-   To a first approximation, any text that discusses the context and meaning of the results,
-   rather than the results themselves, can go in the discussion.}
-
-Of these results, prop.~\ref{p1:omega} is the most surprising.
-It states that returning test results more rapidly (i.e., increasing $\omega$) does not necessarily lower $\Rnum$; whether increasing $\omega$ lowers $\Rnum$ depends on the precise combination of model parameters (test reporting rate, testing processes/strategies ($W_*$), testing efficacy ($P_I$) and isolation ($\eta_*$).
-Specifically, the $\partial{R_0}/\partial{\omega}>0$ suggests that reproduction number may increase as the test reporting process become faster.
-This can be seen from expression \eqref{Rom} in the case of perfect isolation, when $\eta_w=0$ and $\eta_c=0$.
-The underlying mechanism for this result is that individuals may take more precautionary behaviors (e.g., physical distancing) when they are awaiting test results compared to when they are untested or their test has returned negative.
-An example of a situation that would favour slow return of tests is if the test being employed produces many false negatives, because many infected individuals will believe they are negative and thus may unknowingly spread the virus to many others.
-For an example of the behaviour of $\Rnum$ see Figure \ref{pan} at varying values of $\omega$.
-This property also provides a tool to quantify the amount of delay required in the test reporting process as a strategy to reduce $\Rnum$ and consequently control the epidemic.
+Prop.~\ref{p1:rho} indicates that increasing the \percap testing intensity, $\rho$, reduces $\Rnum$.
+Lastly, Prop.~\ref{p1:omega} means that speeding up the test reporting may or may not reduce $\Rnum$, it will depend on other parameters.
+It is straightforward to derive \ref{p1:rho} and \ref{p1:omega} from the Taylor approximation of $\Rnum$ at $\rho=0$.
 
@@
 Our numerical solutions in support of the analytical results in Proposition \ref{prop1} are presented in Figure \ref{pan}.
-It can be inferred that when random testing strategy is applied, Figure \ref{p.a}, comparing to the non-random testing strategy, Figure \ref{p.a}, the critical testing intensity of the corresponding panels are lower in non-random testing.
-Also, it appears that speeding the test reporting, i.e., increasing $\omega$, does not significantly lower $\Rnum$ when awaiting people follow the isolation perfectly, i.e., $\eta_w$ is closer to 0.
+It can be inferred that when random testing strategy is applied, Figure \ref{p.a}, comparing to the targeted testing strategy, Figure \ref{p.b}, the critical \percap testing intensity of the corresponding panels are lower in non-random testing.
+Also, it appears that speeding the test reporting, i.e., increasing $\omega$, does not significantly lower $\Rnum$ when awaiting people (including people in $I_p$ and $I_n$ compartments) follow the isolation perfectly, i.e., $\eta_w$ is closer to 0.
 However, speeding the test reporting reduces the epidemic more when the awaiting people less follow the isolation, i.e., $\eta_w$ is closer to 1.
@@
 Furthermore, we derived an inequality that quantifies the exact relationships between model parameters that result in returning tests more rapidly being favorable \eqref{eq:necsuf}.
+
+%%%%%%%%%%%%
+\section{Discussion}
+
+Here, we develop a compartmental model to incorporate epidemic processes and testing processes mechanistically to study the potential effect of testing strategies, testing intensity, test sensitivity and specificity, reporting time and isolation on epidemic dynamic.
+We investigate the effect of underlying parameters of the model on the basic reproduction number, $\Rnum$.
+While targeted testing strategies are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is viral spread is expected to be stronger for a slow test than for a fast test.
+This counter-intuitive effect can occur when people are cautious when awaiting a test result, and may not be robust to second-order effects of fast testing (such as better contact tracing).
+
+{\bf \percap testing intensity, $\rho$};
+Prop.
+\ref{p1:rho} indicates that increasing testing intensity reduces $\Rnum$.
+This is sensible since as people are moved to test compartments, namely $I_p$ and $I_n$ in particular, may partially or fully self-isolate, and individuals who have tested positive, namely $I_c$ compartment, are highly likely to self-isolate.
+The higher probability of being subject to isolation, the lower the transmission \eqref{Lambda} will become and the lower $\Rnum$ will become.
+Our simulation supports this fact.
+However, it appeared to be hard to conclude this by directly differentiating the expression of $\Rnum$ due to its complexity.
+
+{\bf Delay in test reporting rate, $\omega$};
+Prop.~\ref{p1:omega} is the most surprising.
+It states that returning test results more rapidly (i.e., increasing $\omega$) does not necessarily lower $\Rnum$; whether increasing $\omega$ lowers $\Rnum$ depends on the precise combination of model parameters (test reporting rate, testing processes/strategies ($W_*$), testing efficacy ($P_I$) and isolation ($\eta_*$).
+Specifically, the $\partial{R_0}/\partial{\omega}>0$ suggests that reproduction number may increase as the test reporting process become faster.
+This can be seen from expression \eqref{Rom} in the case of perfect isolation, when $\eta_w=0$ and $\eta_c=0$.
+The underlying mechanism for this result is that individuals may take more precautionary behaviors (e.g., physical distancing) when they are awaiting test results compared to when they are untested or their test has returned negative.
+An example of a situation that would favour slow return of tests is if the test being employed produces many false negatives, because many infected individuals will believe they are negative and thus may unknowingly spread the virus to many others.
+For an example of the behaviour of $\Rnum$ see Figure \ref{pan} at varying values of $\omega$.
+This property also provides a tool to quantify the amount of delay required in the test reporting process as a strategy to reduce $\Rnum$ and consequently control the epidemic.
+
+{\bf The correspondence of testing strategies and \percap testing weights}; 
+-All of the accessible testing data is being used for one of the other three purposes rather than surveillance. These are targeted testing strategies with a different bias.
+- when people are tested for "screening", the individuals with potential higher mobility, eg.
+people who are getting on flights, get more tested and thus the coresponding heavier testing weight is assigned than people awaiting for a surgery and are probably going to stay in a long-term care facility and consequently less mobile and more isolated to begin with.
+
 To give a biological interpretation, we describe the qualitative trends predicted by the inequality.
-Returning test results more rapidly tends to be more favorable when
-
-\begin{itemize}
-\item individuals who have tested positive lower their contact much more than individuals who are waiting for their test results (i.e., $\eta_c \ll \eta_w$).
-\item the test is sensitive.
-\item testing of infected individuals is more intense than testing of susceptible individuals.
-\end{itemize}
+Returning test results more rapidly tends to be more favorable, i.e., reduces $\Rnum$, when (i) \emph{Confirmed-positive} individuals, lower their contact much more than individuals who are waiting for their test results (i.e., $\eta_c \ll \eta_w$), (ii) the test is highly sensitive, i.e., $p_I$ is close to 1, and (iii) the targeted testing strategy is practiced.
 An in-depth discussion of these results is presented in the discussion section.
 
-The use of tests cheaper than RT-PCR has been proposed as a potential strategy for containing the COVID-19 pandemic.
+- The use of tests cheaper than RT-PCR has been proposed as a potential strategy for containing the COVID-19 pandemic.
 While cheaper tests may be less sensitive and reliable than RT-PCR, they allow for broader and more intense testing.
 Using our Taylor approximation of $\Rnum$ near $\rho = 0$, we examined what circumstances (i.e., model parameters) make the use of one test more favourable than another, and give a complete description through inequality \ref{eq:rho1vsrho2}.
-In general, we found that the expensive test tends to more effectively lower $\Rnum$ when
+In general, we found that the expensive test tends to more effectively lower $\Rnum$ when (a) individuals who test positive self-isolate much more than individuals who are waiting for their test result, (b) the time it takes to return tests is much shorter than the mean infectious period, and (c) the testing intensity is much greater for infected individuals than susceptible individuals.
+- Our model also predicts an asymmetry between cheap and expensive tests: if a cheap test can identify on average more infected individuals as an expensive test, then our model predicts that the cheap test will lower $\Rnum$ more.
+In contrast, if an expensive test can identify on average more infected individuals, it will not necessarily lower $\Rnum$ more than the cheap test.
 
-\begin{itemize}
-    \item individuals who test positive self-isolate much more than individuals who are waiting for their test result.
-    \item the time it takes to return tests is much shorter than the mean infectious period.
-    \item the testing intensity is much greater for infected individuals than susceptible individuals.
-\end{itemize}
+{\bf Other points};
+one point we can make when discussing the relative weight of testing in different compartments is to discuss pre-testing screening tools, such as surveys or questionnaires.
+If we have a quantitative description of how the testing intensities affect the dynamics, we can make statements like ``our results suggest that employing a pre-testing screening tool can help target infected individuals more effectively.
+In particular, doubling the sensitivity of the pre-screening tool would *do something* to $\Rnum$.
 
-Our model also predicts an asymmetry between cheap and expensive tests: if a cheap test can identify on average more infected individuals as an expensive test, then our model predicts that the cheap test will lower $\Rnum$ more.
-In contrast, if an expensive test can identify on average more infected individuals, it will not necessarily lower $\Rnum$ more than the cheap test.
-\fady{we can state this as a formal proposition in equations rather than in words, if that would be better.}
-
-%%%%%%%%%%%%
-
-% %%%%%%%
-\section{Discussion}
-
-From JD's email:
-
-\begin{itemize}
-
-  \item We're basically looking at the potential advantage of slow tests:
-people waiting for test results may be more careful than those who
-have received negatives. These advantages are real, and neglected. We
-also compare to an individual-level advantage of fast tests: people
-who test positive may be even more careful. But we are missing out on
-community-level advantages of fast testing: better assessment of the
-situation, identification of hot spots, contact-tracing, etc. 
-\item  Maybe something about the testing rate/intensity, now the per capita testing intensity is very low ($\rho \approx 0$).
+% JD:
+% \begin{itemize}
+%  \item We're basically looking at the potential advantage of slow tests:
+% people waiting for test results may be more careful than those who
+% have received negatives. These advantages are real, and neglected. We
+% also compare to an individual-level advantage of fast tests: people
+% who test positive may be even more careful. But we are missing out on
+% community-level advantages of fast testing: better assessment of the
+% situation, identification of hot spots, contact-tracing, etc. 
+% \item  Maybe something about the testing rate/intensity, now the per capita testing intensity is very low ($\rho \approx 0$).
 In near future new test kits may be widely accessible, our model provide insights in this case (what are those?)
-\item ``delay negative results" result
-
-\end{itemize}
-
-\subsection{Rate of returning tests}
-
-Our model predicts that when the difference in self-isolation between individuals waiting for their tests and individuals who are confirmed positive is small, returning tests rapidly is not beneficial.
-The underlying mechanism for this prediction is that the benefit in returning tests rapidly so (i.e., the extra self-isolation an individual does when they test positive) is minimal and outweighed by the fact that those who test negative may increase their contacts.
-Thus, in such a case, returning tests more rapidly is expected to increase $\Rnum$.
-Similarly, if the test is not sensitive, many infected individuals will test negative, and thus may increase their contacts under the false belief that they are not infectious, so returning tests rapidly in this case may also contribute to increased transmission.
-
-When testing of infected individuals is more intense than susceptible individuals, returning test results rapidly is beneficial because those infected individuals will be informed and self-isolate.
-Conversely, if testing of susceptible individuals is much greater, very few tests will be positive, and the benefit of informing those few positive cases will be outweighed by the potential increase in contacts of the individuals testing negative.
-
-\subsection{Expensive vs. cheap tests}
-\fady{I think we should discuss this before I write it. I'm a little hesitant to really expand on these results because of the points JD brought up over email.}
+% \item ``delay negative results" result
+% \end{itemize}
 

### 2021-02-01 SIR_manuscript.tex
more on discussion and updated a ref in intro

@@
 (1) When testing is done for the purposes of disease surveillance \citep{foddai2020base}
-we might expect tests to be assigned randomly across the population (in an ideal world: some bias in selection for testing is probably inevitable), although a stratified sampling design would be preferred for statistical efficiency.
+we might expect tests to be assigned randomly across the population (in an ideal world: some bias in selection for testing is probably inevitable), although a stratified sampling design would be preferred for statistical efficiency \citep{graubard1996modelling} \ali{better ref?}.
 
@@
 
-Here, we develop a compartmental model to incorporate epidemic processes and testing processes mechanistically to study the potential effect of testing strategies, testing intensity, test sensitivity and specificity, reporting time and isolation on epidemic dynamic.
-We investigate the effect of underlying parameters of the model on the basic reproduction number, $\Rnum$.
+Mathematical modeling of infectious disease outbreaks provides insights on how testing processes influence the epidemiological processes through isolation. 
+Here, we develop a compartmental SIR-type model to study the potential effect of testing strategies, testing intensity, test sensitivity and specificity, test reporting time and isolation on epidemic dynamics.
 While targeted testing strategies are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is viral spread is expected to be stronger for a slow test than for a fast test.
@@
 
+We compared two testing strategies, random and \emph{tergeted}, by incorporating the compartment-specific relative testing weights $W_S$, $W_I$ and $W_R$.
+Although, in this study for the purpose of simplicity and illustraition, we assumed $W_I/W_S=3$ for the targeted strategy scenario, we have not specified a particular relative testing weights corresponding to a particular targeted testing scenario.
+Modeling different targeted testing strategies, thus test-specific testing weights, requires prior information of the conditional probabilities of getting tested for people in a given compartment.
+This part needs to be developed further in future work.
+This can be implied when we would like to quantify and compare the effect of different levels of test forcus for infectious people on the basic reproduction number $\Rnum$, and conclude about the disease spread management.
+For example, when people are tested for "screening", the individuals with potential higher mobility, eg.
+people who are getting on flights, get more tested and thus the coresponding heavier testing weight is assigned than people awaiting for a surgery and are probably going to stay in a long-term care facility and consequently less mobile and more isolated to begin with.
+With our model, we would be able to compare the sensitivity of the disease dynamics, through $\Rnum$, with respect to testing processes.
+
 {\bf \percap testing intensity, $\rho$};
-Prop.
-\ref{p1:rho} indicates that increasing testing intensity reduces $\Rnum$.
+Proposition \ref{prop1} part \ref{p1:rho} indicates that increasing testing intensity $\rho$ reduces $\Rnum$.
 This is sensible since as people are moved to test compartments, namely $I_p$ and $I_n$ in particular, may partially or fully self-isolate, and individuals who have tested positive, namely $I_c$ compartment, are highly likely to self-isolate.
@@
 {\bf Delay in test reporting rate, $\omega$};
-Prop.~\ref{p1:omega} is the most surprising.
-It states that returning test results more rapidly (i.e., increasing $\omega$) does not necessarily lower $\Rnum$; whether increasing $\omega$ lowers $\Rnum$ depends on the precise combination of model parameters (test reporting rate, testing processes/strategies ($W_*$), testing efficacy ($P_I$) and isolation ($\eta_*$).
+Proposition \ref{prop1} part \ref{p1:omega} is the most interesting result.
+It states that returning test results more rapidly, i.e., increasing $\omega$, does not necessarily lower $\Rnum$; whether increasing $\omega$ lowers $\Rnum$ depends on the precise combination of model parameters (test reporting rate, testing processes/strategies ($W_*$), testing efficacy ($P_I$) and isolation ($\eta_*$).
 Specifically, the $\partial{R_0}/\partial{\omega}>0$ suggests that reproduction number may increase as the test reporting process become faster.
@@
 This property also provides a tool to quantify the amount of delay required in the test reporting process as a strategy to reduce $\Rnum$ and consequently control the epidemic.
-
-{\bf The correspondence of testing strategies and \percap testing weights}; 
--All of the accessible testing data is being used for one of the other three purposes rather than surveillance. These are targeted testing strategies with a different bias.
-- when people are tested for "screening", the individuals with potential higher mobility, eg.
-people who are getting on flights, get more tested and thus the coresponding heavier testing weight is assigned than people awaiting for a surgery and are probably going to stay in a long-term care facility and consequently less mobile and more isolated to begin with.
 

### 2021-02-02 SIR_manuscript.tex
working on discussion

@@
 
-We compared two testing strategies, random and \emph{tergeted}, by incorporating the compartment-specific relative testing weights $W_S$, $W_I$ and $W_R$.
-Although, in this study for the purpose of simplicity and illustraition, we assumed $W_I/W_S=3$ for the targeted strategy scenario, we have not specified a particular relative testing weights corresponding to a particular targeted testing scenario.
+We incorporated the compartment-specific relative testing weights, $W_S$, $W_I$ and $W_R$, to model random testing and \emph{targeted} testing strategies.
+Here, in the case of targeted testing and for the simplicity and illustraition purposes, we assumed that infected and recovered individuals are tested at three times the \percap rate of susceptible individuals, i.e., $W_I/W_S=3$ and $W_R=W_I$.
+Note we have not specified a methodology to assigne particular relative testing weights corresponding to a particular targeted testing scenario.
+This part needs to be developed further in future work. 
 Modeling different targeted testing strategies, thus test-specific testing weights, requires prior information of the conditional probabilities of getting tested for people in a given compartment. 
-This part needs to be developed further in future work.
 This can be implied when we would like to quantify and compare the effect of different levels of test forcus for infectious people on the basic reproduction number $\Rnum$, and conclude about the disease spread management.
@@
 
-{\bf \percap testing intensity, $\rho$};
-Proposition \ref{prop1} part \ref{p1:rho} indicates that increasing testing intensity $\rho$ reduces $\Rnum$.
-This is sensible since as people are moved to test compartments, namely $I_p$ and $I_n$ in particular, may partially or fully self-isolate, and individuals who have tested positive, namely $I_c$ compartment, are highly likely to self-isolate.
-The higher probability of being subject to isolation, the lower the transmission \eqref{Lambda} will become and the lower $\Rnum$ will become.
-Our simulation supports this fact.
-However, it appeared to be hard to conclude this by directly differentiating the expression of $\Rnum$ due to its complexity.
+{\bf The \percap testing intensity, $\rho$;}
+Proposition \ref{prop1}, part \ref{p1:rho}, and also Figure \ref{pan} indicates that increasing testing intensity $\rho$ reduces $\Rnum$.
+This is sensible since as people are moved to test compartments, namely $Z_p$ and $Z_n$ for $Z \in \{S,I,R\}$, they may partially or fully self-isolate.
+Also individuals who have tested positive, namely $Z_c$ compartment for $Z \in \{S,I,R\}$, are highly likely to self-isolate.
+The higher probability of being subject to isolation, the lower the force of infection \eqref{Lambda} and the lower $\Rnum$ will become.
+While our simulation supported this result, analytically it appeared to be hard to conclude this due to the complexity of $\Rnum$ expression.
 

### 2021-02-02 SIR_manuscript.tex
discussion updated

@@
 We incorporated the compartment-specific relative testing weights, $W_S$, $W_I$ and $W_R$, to model random testing and \emph{targeted} testing strategies.
-Here, in the case of targeted testing and for the simplicity and illustraition purposes, we assumed that infected and recovered individuals are tested at three times the \percap rate of susceptible individuals, i.e., $W_I/W_S=3$ and $W_R=W_I$.
-Note we have not specified a methodology to assigne particular relative testing weights corresponding to a particular targeted testing scenario.
+Here, in the case of targeted testing and for the simplicity and illustraition purposes, we assumed that infected and recovered individuals are tested at three times the \percap rate of susceptible individuals, thus $W_I/W_S=3$ and $W_R=W_I$.
+Note that we have not specified a methodology to assigne particular relative testing weights corresponding to a particular targeted testing scenario.
 This part needs to be developed further in future work. 
-Modeling different targeted testing strategies, thus test-specific testing weights, requires prior information of the conditional probabilities of getting tested for people in a given compartment. 
+Modeling different targeted testing strategies, equivalently test-specific testing weights in our framework, requires prior information of the conditional probabilities of getting tested for people in a given compartment.
 This can be implied when we would like to quantify and compare the effect of different levels of test forcus for infectious people on the basic reproduction number $\Rnum$, and conclude about the disease spread management.
@@
 {\bf The \percap testing intensity, $\rho$;}
-Proposition \ref{prop1}, part \ref{p1:rho}, and also Figure \ref{pan} indicates that increasing testing intensity $\rho$ reduces $\Rnum$.
+Proposition \ref{prop1} part \ref{p1:rho} and also Figure \ref{pan} indicates that increasing testing intensity $\rho$ reduces $\Rnum$.
 This is sensible since as people are moved to test compartments, namely $Z_p$ and $Z_n$ for $Z \in \{S,I,R\}$, they may partially or fully self-isolate.
@@
 
-{\bf Delay in test reporting rate, $\omega$};
+{\bf Delay in test reporting rate, $\omega$;}
 Proposition \ref{prop1} part \ref{p1:omega} is the most interesting result.
-It states that returning test results more rapidly, i.e., increasing $\omega$, does not necessarily lower $\Rnum$; whether increasing $\omega$ lowers $\Rnum$ depends on the precise combination of model parameters (test reporting rate, testing processes/strategies ($W_*$), testing efficacy ($P_I$) and isolation ($\eta_*$).
-Specifically, the $\partial{R_0}/\partial{\omega}>0$ suggests that reproduction number may increase as the test reporting process become faster.
-This can be seen from expression \eqref{Rom} in the case of perfect isolation, when $\eta_w=0$ and $\eta_c=0$.
-The underlying mechanism for this result is that individuals may take more precautionary behaviors (e.g., physical distancing) when they are awaiting test results compared to when they are untested or their test has returned negative.
-An example of a situation that would favour slow return of tests is if the test being employed produces many false negatives, because many infected individuals will believe they are negative and thus may unknowingly spread the virus to many others.
-For an example of the behaviour of $\Rnum$ see Figure \ref{pan} at varying values of $\omega$.
-This property also provides a tool to quantify the amount of delay required in the test reporting process as a strategy to reduce $\Rnum$ and consequently control the epidemic.
+It states that returning test results more rapidly, i.e., increasing $\omega$, does not necessarily lower the reproduction number $\Rnum$; whether increasing $\omega$ lowers $\Rnum$ depends on the precise combination of model parameters (test reporting rate, testing strategies or test-specific testing weights, test sensitivity and specificity and isolation.
+Specifically, in the case of perfect isolation, i.e., when $\eta_w=0$ and $\eta_c=0$, $\Rnum$ may increase as the test reporting process become faster. This can be seen from expression \eqref{Rom}.
+The interpretation for this phenomenon is that individuals are highly likely to fully self-isolate when they are awaiting test results and when they are reported positive with a negligable influence on the force of infection, $\lambda$.
+Thus the faster the test reporting rate, the shorter they stay in these ``\emph {safe}'' awaiting compartments, and the more they get involved in the infection process.
+Another example of this situation, when the slow test reporting rate would be favorable is in the case when the test being employed produces many false negatives.
+Because many infected individuals will believe they are uninfectious and thus may unknowingly spread the virus to many others.
 
-To give a biological interpretation, we describe the qualitative trends predicted by the inequality.
-Returning test results more rapidly tends to be more favorable, i.e., reduces $\Rnum$, when (i) \emph{Confirmed-positive} individuals, lower their contact much more than individuals who are waiting for their test results (i.e., $\eta_c \ll \eta_w$), (ii) the test is highly sensitive, i.e., $p_I$ is close to 1, and (iii) the targeted testing strategy is practiced.
-An in-depth discussion of these results is presented in the discussion section.
+Our model enables us to quantify the amount of delay required in the test reporting process as a strategy to reduce $\Rnum$.
+To give a biological interpretation, we describe the qualitative trends predicted by the inequality \eqref{eq:necsuf} \todo{Fady please confirm}.
+To summarise, returning test results more rapidly tends to be more favorable, i.e., reduces $\Rnum$, when (i) \emph{Confirmed-positive} individuals, lower their contact much more than individuals who are waiting for their test results (i.e., $\eta_c \ll \eta_w$), (ii) the test is highly sensitive, i.e., $p_I$ is close to 1, and (iii) the targeted testing strategy is practiced.
 
-- The use of tests cheaper than RT-PCR has been proposed as a potential strategy for containing the COVID-19 pandemic.
+Our model also predicts an asymmetry between cheap and expensive tests: if a cheap test can identify on average more infected individuals as an expensive test, then our model predicts that the cheap test will lower $\Rnum$ more.
+In contrast, if an expensive test can identify on average more infected individuals, it will not necessarily lower $\Rnum$ more than the cheap test.
+The use of tests cheaper than RT-PCR has been proposed as a potential strategy for containing the COVID-19 pandemic.
 While cheaper tests may be less sensitive and reliable than RT-PCR, they allow for broader and more intense testing.
@@
 In general, we found that the expensive test tends to more effectively lower $\Rnum$ when (a) individuals who test positive self-isolate much more than individuals who are waiting for their test result, (b) the time it takes to return tests is much shorter than the mean infectious period, and (c) the testing intensity is much greater for infected individuals than susceptible individuals.
-- Our model also predicts an asymmetry between cheap and expensive tests: if a cheap test can identify on average more infected individuals as an expensive test, then our model predicts that the cheap test will lower $\Rnum$ more.
-In contrast, if an expensive test can identify on average more infected individuals, it will not necessarily lower $\Rnum$ more than the cheap test.
 
-{\bf Other points};
+{\bf Other points [?];}
 one point we can make when discussing the relative weight of testing in different compartments is to discuss pre-testing screening tools, such as surveys or questionnaires.
@@
 
+\todo{Fady, can you clean this up}
 \begin{align}\label{eq:necsuf}

### 2021-02-03 SIR_manuscript.tex
works on discussion and supp material

@@
 where
-\begin{align*}
+\begin{equation}
+\begin{aligned}
+\label{eq:abc}
 A=& \gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_c p_I) F_I, \\
@@
 C=& \frac{\beta/\gamma}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}.
-\end{align*}
+\end{aligned}
+\end{equation}
 Further details of derivation of $\Rnum$ are provided in the Supplementary material section.
 
-The analytical calculation of the next generation matrix, $G$, and the derivation of $\Rnum$ was carried out in Maple (ref?) by using simple linear Algebra package. 
-  We used R for the simulation part which included using the explicit expression of $\Rnum$ \ref{R0} as a function of the underlying parameters, specifying the parameters a realistic range, calculating the corresponding $\Rnum$ and plotting the contours of $\Rnum$ by using ggplot package in R.
+The analytical calculation of the next generation matrix, $G$, and the derivation of $\Rnum$ was carried out in Maple \citep{maple14} by using simple linear Algebra package. 
+  We used R \citep{r} for the simulation part which included using the explicit expression of $\Rnum$ \ref{R0} as a function of the underlying parameters, specifying the parameters a realistic range, calculating the corresponding $\Rnum$ and plotting the contours of $\Rnum$ by using ggplot package (ref?) in R.
 The related result is presented in Figure \ref{pan}, where  panel \eqref{p.a} represents the random testing, and panel \eqref{p.b} is representing non-random testing.
@@
 \end{subfigure}
-\caption{Behaviour of the basic reproduction number, $\Rnum$, with respect to changes in the underlying parameters; \bmb{remind the reader about the parameters and their values}.
-The solid line in each panel represents the critical contour of $\Rnum=1$.
+\caption{
+{\bf A comparison of the behaviour of the basic reproduction number, $\Rnum$, between random versus targeted testing strategies at differnt levels of testing and isolation.}
+We numerically evaluate $\Rnum$ \eqref{R0} using the following parameters (listed in Table \ref{tab:params}):
+$N_0=1 \times 10^6$, $\omega \in [0.1,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.01]$ 1/day, $\eta_w \in [0,1]$, $\eta_c \in [0,1]$, $\beta=0.39$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
+Contours of $\Rnum$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $W_S=W_I=W_R=1$ and (b) targeted testing where $W_S=0.3$ and $W_I=W_R=1$.
+The black solid line in each panel represents the critical contour of $\Rnum=1$.
 \bmb{we still have some collisions in the y-axis tick labels}}
@@
 
-The model is 
+The model in the form of a system of ordinary differential equations is 
 \begin{subequations}\label{model}
@@
 
-where $\beta$, (in units of 1/day), is the disease transmission rate, $\eta_w$ and $\eta_c$ are the isolation parameters for awaiting and reported individuals, respectively.
-$\Lambda$ is the force of infection defined in Eq.
-\eqref{Lambda}.
-
-The next generation matrix is $G = F V^{-1}$, where matrix $F$ represents the inflow of new infection to the infected compartments and matrix $V$ represents the flow in the infected compartments when the population is totally susceptible.
-For model, \crefrange{eq1}{eq12}, $F$ and $V$ are
+where parameters are specified in Table \ref{tab:params}.
+The next generation matrix for this model is $G = F V^{-1}$, where matrix $F$ represents the inflow of new infection to the infected compartments and matrix $V$ represents the flow in the infected compartments when the population is totally susceptible.
+Matrices $F$ and $V$ are
 
@@
 \begin{equation}
+\label{mat:G}
 G = \left[ \begin {array}{cc}
@@
 A\,S_n & B\,S_n
-\end {array}\right].
+\end {array}\right],
 \end{equation}
-
+where expressions $A$, $B$ and $C$ are specified in \eqref{eq:abc}.
 The block matrix $G_{12}$ does not influence $\Rnum$ (defined as the spectral radius of $G$). All that matters here are the eigenvalues of $G_{11}$, which are 0 and $\Rnum$ \eqref{R0}.
+Also note that, all $S_u$'s and $S_n$'s are evaluated at the DFE \eqref{dfe} in matrix $F$ \eqref{FV} and the block matrix $G_{11}$ \eqref{mat:G}.
 

### 2021-02-03 SIR_manuscript.tex
minor edits on supp material

@@
 
-\bmb{need a little bit of context here, saying that we didn't do any numerical solutions for the trajectories in our analysis,
-  but if you do try them you run into this problem ...}
-
-The testing rate, $\sigma$, should be formulated such that people from the untested compartments will not be tested if they are not there.
-The numerical singularity issue with the chosen $\sigma$ \eqref{sigma} is that the population in $S$ compartments appeared to blow up when the DFE is achieved.
+In this work, we didn't do any numerical solutions for the trajectories in our analysis. However, if one tries to do so there would be a singularity issue to deal with. 
+Specifically, the numerical singularity issue with the chosen $\sigma$ \eqref{sigma} is that the population in $S$ compartments appeared to blow up when the DFE is achieved.
 This is once the only untested people are susceptibles, the FOI will become $\Lambda=0$, testing rate $F_s=\rho N_0/S_u$.
-Thus, eq(1) of the model will be $d S_u/dt = - \rho N_0 + \omega S_n$ which is no longer dependent on $S_u$ and a linear rate of leaving the $S_u$ compartment.
-
+Thus, the first equation of the model \eqref{model} will become
+$d S_u/dt = - \rho N_0 + \omega S_n$. Thus changes in $S_u$ will be no longer dependent on $S_u$ with a linear rate of leaving the $S_u$ compartment.
+IN fact the testing rate, $\sigma$, should be formulated such that people from the untested compartments will not be tested if they are not there.
 One way to fix this issue, is to consider a maximum testing rate, $\tau$ (1/day).
@@
 \subsection{Taylor Approximation of $\Rnum$ at $\rho=0$ }
-
 The basic reproduction number, $\Rnum$, close to $\rho=0$ can be approximated linearly in $\rho$ by using Taylor approximation. it follows
-
 \begin{equation}
@@
 Also, to analyze the influence of $\omega$ on $\Rnum$, the approximation \eqref{eq:R0appr} was used. Then it is straight forward to have
-
 \begin{equation}
@@
 \end{equation}
-
 where $a=(\eta_w-1)W_I-(\eta_c-1)P_I W_I = ((s-p_I)\eta_c + (p_I-1)) W_I$, $b=2(\eta_w-1)\gamma W_S$ and $c=(\eta_w-1)\gamma^2 W_S$.
@@
 % %%%%%%%
-[Fady: this subsection and most of the next belong to the Appendix since they are very detailed. Would you clean up the writing, there are some repeated expressions that can be referenced?]
-
 \subsection{rate of returning tests}
@@
 With that in mind, we can write the roots of $\partial{\Rnum}/\partial{\omega}$ as
-
 \begin{align}
     \omega_1 =& \frac{\gamma}{-\sqrt{1-Q}-1} \\
-    \omega_2 =& \frac{\gamma}{\sqrt{1-Q}-1}
+    \omega_2 =& \frac{\gamma}{\sqrt{1-Q}-1}.
 \end{align}

### 2021-02-03 SIR_manuscript.tex
further updates on discussion

@@
 
-{\bf Delay in test reporting rate, $\omega$;}
-Proposition \ref{prop1} part \ref{p1:omega} is the most interesting result.
-It states that returning test results more rapidly, i.e., increasing $\omega$, does not necessarily lower the reproduction number $\Rnum$; whether increasing $\omega$ lowers $\Rnum$ depends on the precise combination of model parameters (test reporting rate, testing strategies or test-specific testing weights, test sensitivity and specificity and isolation.
-Specifically, in the case of perfect isolation, i.e., when $\eta_w=0$ and $\eta_c=0$, $\Rnum$ may increase as the test reporting process become faster. This can be seen from expression \eqref{Rom}.
-The interpretation for this phenomenon is that individuals are highly likely to fully self-isolate when they are awaiting test results and when they are reported positive with a negligable influence on the force of infection, $\lambda$.
-Thus the faster the test reporting rate, the shorter they stay in these ``\emph {safe}'' awaiting compartments, and the more they get involved in the infection process.
-Another example of this situation, when the slow test reporting rate would be favorable is in the case when the test being employed produces many false negatives.
-Because many infected individuals will believe they are uninfectious and thus may unknowingly spread the virus to many others.
+{\bf The potential advantage of slow test reporting, or favorable-delay-reporting;}
+Individuals are highly likely to fully self-isolate when they are either awaiting the test results or they are reported, thus reducing or eliminating their potentially infectious contacts.
+Thus, the faster the test reporting rate, $\omega$, the shorter these individuals stay in the ``\emph {safe}'' awaiting-confirmed compartments, namely $Z_n$, $Z_p$ and $Z_c$ for $Z \in \{S,I,R\}$, and the more they get involved in the infection process.
+This advantage of slow test reporting is real, and neglected. 
+We also compare to an individual-level advantage of fast tests: people who test positive may be even more careful.
+In our model analysis, Proposition \ref{prop1} part \ref{p1:omega} is describing this potential advantage. 
+It states that returning test results more rapidly, i.e., increasing the rate $\omega$, does not necessarily lower the reproduction number $\Rnum$; whether increasing $\omega$ lowers $\Rnum$ depends on the precise combination of model parameters  including test reporting rate, testing strategies represented by compartment-specific testing weights, test sensitivity and specificity, and the level of isolation.
+Specifically, in the case of perfect isolation, i.e., when $\eta_w=0$ and $\eta_c=0$, $\Rnum$ may increase as the test reporting process becomes faster. This can be seen from expression \eqref{Rom}.
+Another example of this favorable delay in $\omega$ could be when the test being employed produces many false negatives.
+Because many infected asymptomatic individuals will believe they are uninfected/uninfectious, thus may unknowingly spread the virus to many others.
+Agian the delay in the test reporting rate keeps these individuals in the ``\emph {safe}'' awaiting-confirmed compartments.
 
-Our model enables us to quantify the amount of delay required in the test reporting process as a strategy to reduce $\Rnum$.
-To give a biological interpretation, we describe the qualitative trends predicted by the inequality \eqref{eq:necsuf} \todo{Fady please confirm}.
-To summarise, returning test results more rapidly tends to be more favorable, i.e., reduces $\Rnum$, when (i) \emph{Confirmed-positive} individuals, lower their contact much more than individuals who are waiting for their test results (i.e., $\eta_c \ll \eta_w$), (ii) the test is highly sensitive, i.e., $p_I$ is close to 1, and (iii) the targeted testing strategy is practiced.
-
-Our model also predicts an asymmetry between cheap and expensive tests: if a cheap test can identify on average more infected individuals as an expensive test, then our model predicts that the cheap test will lower $\Rnum$ more.
+we are missing out on
+community-level advantages of fast testing: better assessment of the
+situation, identification of hot spots, contact-tracing, etc.
+The \percap testing intensity of CIVID-19 after about a year from the first case reported in December 2019, is still low (($\rho \approx 0$) in our model).
+In near future new test kits may be widely accessible, our model provide insights in this case.
+In particular, if a cheap test can identify on average more infected individuals as an expensive test, then our model predicts that the cheap test will lower $\Rnum$ more.
 In contrast, if an expensive test can identify on average more infected individuals, it will not necessarily lower $\Rnum$ more than the cheap test.
@@
 
-{\bf Other points [?];}
+\ali{ not sure if we want these 2 following parageraphs, or shrink it into afew short sentences?}
+In addition to the favorable-delay-reporting observation of our model \eqref{model}, the model enables us to quantify the amount of delay required in the test reporting process as a strategy to reduce $\Rnum$.
+To give a biological interpretation, we describe the qualitative trends predicted by the inequality \eqref{eq:necsuf} \todo{Fady please confirm}.
+To summarise, returning test results more rapidly tends to be more favorable, i.e., reduces $\Rnum$, when (i) \emph{Confirmed-positive} individuals, lower their contact much more than individuals who are waiting for their test results (i.e., $\eta_c \ll \eta_w$), (ii) the test is highly sensitive, i.e., $p_I$ is close to 1, and (iii) the targeted testing strategy is practiced.
+
 one point we can make when discussing the relative weight of testing in different compartments is to discuss pre-testing screening tools, such as surveys or questionnaires.
@@
 In particular, doubling the sensitivity of the pre-screening tool would *do something* to $\Rnum$.
-
-% JD:
-% \begin{itemize}
-%  \item We're basically looking at the potential advantage of slow tests:
-% people waiting for test results may be more careful than those who
-% have received negatives. These advantages are real, and neglected. We
-% also compare to an individual-level advantage of fast tests: people
-% who test positive may be even more careful. But we are missing out on
-% community-level advantages of fast testing: better assessment of the
-% situation, identification of hot spots, contact-tracing, etc. 
-% \item  Maybe something about the testing rate/intensity, now the per capita testing intensity is very low ($\rho \approx 0$).
-In near future new test kits may be widely accessible, our model provide insights in this case (what are those?)
-% \item ``delay negative results" result
-% \end{itemize}
 

### 2021-02-06 SIR_manuscript.tex
geometry and line numbers
comments and minor tweaks in introduction

@@
 \usepackage{cleveref}
-\usepackage{geometry}
+\usepackage[margin=1in]{geometry}
+\usepackage{lineno}\renewcommand\thelinenumber{\color{gray}\arabic{linenumber}}
 \usepackage{pdflscape}
@@
 \newcommand{\jd}[1]{\comment{blue}{JD}{#1}}
-\newcommand{\djde}[1]{\comment{red}{DJDE}{#1}}
+\newcommand{\david}[1]{\comment{red}{DJDE}{#1}}
 \newcommand{\bmb}[1]{\comment{red}{BMB}{#1}}
@@
 
-\title{Testing and Isolation Efficacy; Insights from a Simple Epidemic Model }
+\title{Testing and Isolation Efficacy: Insights from a Simple Epidemic Model}
 
@@
 \maketitle
+
+\linenumbers
 
@@
 In addition to shaping epidemic observations (via case reports), testing processes can also affect epidemiological dynamics.
-In particular, individuals who are awaiting the results of a test may partially or fully self-isolate, and individuals who have tested positive are highly likely to self-isolate.
+In particular, individuals who are awaiting the results of a test may partially or fully self-isolate, and individuals who have tested positive are more likely to self-isolate.
 We developed a model that incorporates mechanistic epidemic processes and testing in order to explore the effects of testing and isolation on epidemic dynamics.
 
-Because individuals who have been tested or received a positive test are likely to self-isolate, thus reducing or eliminating  their potentially infectious contacts, the dynamics of an epidemic will depend on who gets tested and
+Because individuals who have been tested or received a positive test are likely to self-isolate---reducing or eliminating  their potentially infectious contacts---the dynamics of an epidemic will depend on who gets tested and
 why. For a given testing intensity (tests performed per day), testing will have a stronger effect on infection rates
-the more strongly it is focused on people who are infectious for SARS-CoV-2.
+the more strongly it is focused on people who are infectious.\david{deleted `` for SARS-CoV-2''.
+Since this is an idealized model intended for insights, it is probably better to use more generic language most of the time and not say covid unless necessary in a specific context.}
 This level of focus depends in turn on the purpose and design of testing programs.
@@
 
-Over the course of the COVID19 \ali{consistency COVID-19 or COVID19?} pandemic, however, the vast majority of testing has been done with other goals --
+Over the course of the COVID19 \ali{consistency COVID-19 or COVID19?}\david{create a \texttt{\detokenize{\covid}} macro and use that.
+I suggest defining it as COVID-19 since that is more common.} pandemic, however, the vast majority of testing has been done with other goals --
 primarily diagnosis (determining COVID19 infection status for clinical purposes), or control (determining COVID19 status in order to quickly isolate cases that have been found by contact tracing), which we characterize as \emph{targeted} testing strategies.
@@
 
-When testing is used primarily for diagnosis it will focus on people with COVID19-like symptoms; thus the testing weights for people with vs.
-without SARS-CoV-2 infections will depend on the relative magnitudes of the probability of having COVID19-like symptoms when infected with SARS-CoV-2 (i.e.
-the relative proportions of presymptomatic or asymptomatic vs.
-symptomatic infections) and the probability of having COVID19-like symptoms when \emph{not} infected (e.g.
-due to other respiratory illnesses).
-Testing for epidemic control will focus on people who are known to have been in contact with known COVID19 cases; in this case the testing weights for infected vs.
-uninfected people will depend on the probability of infection given contact, as well as the thoroughness and effectiveness of the system for identifying suspicious contacts.
+When testing is used primarily for diagnosis it will focus on people with COVID19-like symptoms; thus the testing weights for people with \david{escape spaces after dots unless it is a full stop}vs.\ without SARS-CoV-2 infections will depend on the relative magnitudes of the probability of having COVID19-like symptoms when infected with SARS-CoV-2 (i.e., the relative proportions of presymptomatic or asymptomatic vs.\ symptomatic infections) and the probability of having COVID19-like symptoms when \emph{not} infected (e.g., due to other respiratory illnesses).
+Testing for epidemic control will focus on people who are known to have been in contact with known COVID19 cases; in this case the testing weights for infected vs.\ uninfected people will depend on the probability of infection given contact, as well as the thoroughness and effectiveness of the system for identifying suspicious contacts.
 
-In addition to the overall testing intensity, the testing weights for different compartments, and the degree of self-isolation of test-pending and test-confirmed people, our models will also need to include test specificity and sensitivity, as well as the distribution of times from testing to reporting.
+In addition to the overall testing intensity, the testing weights for different compartments, and the degree of self-isolation of test-pending and test-confirmed people, our models will also need to include test specificity and sensitivity\david{I presume you mean that models would ideally take account of these things, which should not be mentioned unless they are defined since the terms are often misunderstood}, as well as the distribution of times from testing to reporting.
 Here, we focus on the the effect of testing intensity, different levels of testing ``focus'' (from random to highly targeted), and rate of test processing on a single index of epidemiological dynamics, the basic reproduction number $\Rnum$.
@@
 In the early stages of an epidemic the number of infected individuals is expected grow exponentially over time when $\Rnum>1$. 
-To control the spread of disease in the population, $\Rnum$ must be $<1$.
+To control the spread of disease in the population, \david{the effective} $\Rnum$ must be $<1$.
 Although the value of $\Rnum$ cannot completely characterize the dynamics of even the simplest epidemic model
 \citep{shaw2021what}, it does give a simple and widely accepted index for the difficulty of control, as well as some indication of the likely final size of an epidemic \citep{ma2006generality}.
+\david{This paragraph does not flow from the rest of the intro, and is a weird way to end when the point of the paper is about testing.}
+
+\david{There is some back and forth between generic discussion and COVID-specific comments.  I think COVID should be used as motivation, but that the whole paper should be framed generically.}
 
@@
 
-\subsection{Expensive vs. cheap tests}
+\subsection{Expensive vs.\ cheap tests}
 

### 2021-02-07 SIR_manuscript.tex
more minor repairs and a bunch more comments

@@
 We developed a deterministic model, Eqs.
-\eqref{model}, which groups individuals based on disease status and testing status.
+\eqref{model}\david{relegating the equations to supp is not OK.
+You don't want to publish a paper in which the model is specified exactly only in online supplementary material.
+I don't think there is any need for supp material.
+You can have appendices, but I would keep everything as part of the paper.}, which groups individuals based on disease status and testing status.
 Disease states include Susceptible, Infectious and Recovered (thus this is an SIR-type model), and testing status categorizes people as \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive} (Figure~\ref{fig:flowchart}).
 Two `accumulator' compartments, $N$ and $P$, were also incorporated in the model in order to collect cumulative reported negative or positive tests.
-The model and details of calculation of $\Rnum$ are presented in the Supplementary Materials.
+The model and details of calculation of the basic reproduction number $\Rnum$ are presented in the Supplementary Materials.
 
@@
 \percap flows between compartments, or modifiers to these flow rates.
+\david{you need to define what you mean by ``testing intensity'' $\rho$}
 The novel component of the model comes in through the compartment-specific relative testing weights $W_S$, $W_I$ and $W_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively.
@@
 
-In order to link to more applied models, we constructed this model so we could specify the total \percap testing rate.
-We do this by defining the weighted size of the testing pool $W = W_S S_u + W_I I_u + W_R R_u$, and calculating the scaling parameter for testing as:
+In order to link to more applied models\david{what does that mean?
+what is an ``applied model''?}, we constructed this model so we could specify the total \percap testing rate.
+We do this by defining the weighted size of the testing pool $W = W_S S_u + W_I I_u + W_R R_u$, and calculating a scaling parameter for testing as:
 \begin{equation}
@@
 \end{equation}
-Thus, \percap testing rate for compartment $Z$ is $F_Z=\sigma W_Z$, where $Z \in \{S,I,R\}$.
-For a high-sensitivity test, infected people typically flow through to the ``confirmed positive'' ($I_c$/$R_c$) compartments and are thus unavailable for further testing.
-Over the course of the epidemic, a fixed testing rate as specified in \eqref{sigma} can
-(if large enough) exhaust the pool of people available for testing,
-leading to a singularity when no one is left untested.
-Although this phenomenon does not affect our analysis of $\Rnum$, it does sometimes occur in
-numerical solutions of the time dynamics: we discuss an adjustment to the model that solves
-this problem in the Supplementary material.
+Thus, the \percap testing rate for compartment $Z$ is $F_Z=\sigma W_Z$, where $Z \in \{S,I,R\}$.
+\david{For a generic compartment, I would suggest $X$ rather than $Z$, since $Z$ is often the final size.}  For a high-sensitivity test, infected people typically flow through to the ``confirmed positive'' ($I_c$, $R_c$) compartments and are thus unavailable for further testing.
+Over the course of the epidemic, a fixed testing rate as specified in \eqref{sigma} can (if large enough) exhaust the pool of people available for testing, leading to a singularity when no one is left untested.
+Although this phenomenon does not affect our analysis of $\Rnum$, it can affect the temporal dynamics (we discuss an adjustment to the model that solves this problem in the Supplementary material).
 
-The classical SIR model is based on the following implicit assumptions; well-mixed population, homogeneity of the population, exponentially distributed duration of infection and large population size (see, e.g., \cite{keeling2011modeling}).
+\david{A comment on notation:  You haven't actually defined $S_u$, $I_u$ etc.
+In any case, I urge you to use macros and use Roman font for subscripts that are textual as opposed to variables.
+So $S_{\rm u}$ and $I_{\rm p}$ etc.
+Then it is clear the subscript isn't a variable with a value.}
+
+The classical SIR model is based on the following implicit assumptions; well-mixed population, homogeneity of the population\david{What's the difference btw ``well-mixed'' and ``homoegeneous''?}, exponentially distributed duration of infection and large population size (see, e.g., \cite{keeling2011modeling}).
 In addition to these standard assumptions, our model, \ref{model}, assumes: (i) there is a single force of infection, $\Lambda$, defined as follows
@@
 \end{equation}
-across all susceptible pools (see Table~\ref{tab:params} for detail), (ii) $\eta_c \leq \eta_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
-For this analysis, we also assume a perfectly specific test,($p_S=0$).
+across all susceptible pools (see Table~\ref{tab:params} for detail\david{there are no details in the table; you need to do a better job of explaining what's going on}), (ii) $\eta_c \leq \eta_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
+For this analysis, we also assume a perfectly specific test ($p_S=0$).
 This last assumption combined with the assumption that no individual is in waiting-for-\emph{positive} or \emph{confirmed positive} compartments, i.e., $S_p(0)=S_c(0)=0$, reduces the model to 10 equations (equations c and d  in \eqref{model} are eliminated).
@@
 
-The Disease-Free Equilibrium (DFE) for the SIR model, Eqs. \eqref{model}, is given by solving the coupled system including $S_u+S_n=N_0$ and $F_S S_u-\omega S_n=0$. The DFE is
-
+The Disease-Free Equilibrium (DFE) for the SIR model, Eqs. \eqref{model}, is given by solving the coupled system including $S_u+S_n=N_0$ and $F_S S_u-\omega S_n=0$.
+\david{you mean by solving the algebraic equations obtain by setting the DEs to 0.  not clear why you bother to mention two of these algebraic eqns.}
+The DFE is
+\david{don't put blank lines around equations unless you are started a new paragraph.  comment out blank lines if want the .tex file to be more readable}
+%%
 \begin{equation}
 \label{dfe}
-S_n^*= \frac{\rho}{\omega} N_0, \ S_u^*= N_0-S_n^*, \text{~and~} I_j=R_j=0 \ \text{for all j}.
+S_n^*= \frac{\rho}{\omega} N_0, \ S_u^*= N_0-S_n^*, \text{~and~} I_j=R_j=0 \ \text{for all }j.
 \end{equation}
+\david{above you had j in Roman font even though it *is* a variable, so I changed to $j$}
 
@@
 \end{equation}
+\david{put $\beta/\gamma$ in front of the expression in eq \eqref{R0}, so it is clear that the rest is a modification of the the simplest expression.
+It would be worth spending some time trying to write the expression in a way that helps the reader understand where the different factors come from.}
 Further details of derivation of $\Rnum$ are provided in the Supplementary material section.
 
-The analytical calculation of the next generation matrix, $G$, and the derivation of $\Rnum$ was carried out in Maple \citep{maple14} by using simple linear Algebra package. 
+The analytical calculation of the next generation matrix, $G$\david{don't waste the symbol $G$ on a quantity we don't really need; $G$ is often used for generation interval}, and the derivation of $\Rnum$ was carried out in Maple \citep{maple14} by using simple linear Algebra package.
   We used R \citep{r} for the simulation part which included using the explicit expression of $\Rnum$ \ref{R0} as a function of the underlying parameters, specifying the parameters a realistic range, calculating the corresponding $\Rnum$ and plotting the contours of $\Rnum$ by using ggplot package (ref?) in R.
 The related result is presented in Figure \ref{pan}, where  panel \eqref{p.a} represents the random testing, and panel \eqref{p.b} is representing non-random testing.
-The parameter values or ranges are as follows (I like using a table here, with the param, their units and values).
-It is notable that in the simulation part, the explicit expression of $\Rnum$ was used and not the approximation \eqref{eq:R0appr}.
-This simulation reflects the behavior of $\Rnum$ with respect to the selected parameters for two different testing strategies: (i) random testing, represented by all testing weights to be the same, $W_S=W_I=W_R$, and (ii) non-random testing, when testing weight are not equal.
+The parameter values or ranges are as follows (I like using a table here, with the param, their units and values).\david{use comment macro for comments} It is notable that in the simulation part, the explicit expression of $\Rnum$ was used and not the approximation \eqref{eq:R0appr}.\david{that's the first mention of ``the approximation''} This simulation reflects the behavior of $\Rnum$ with respect to the selected parameters for two different testing strategies: (i) random testing, represented by all testing weights to be the same, $W_S=W_I=W_R$, and (ii) non-random testing, when testing weight are not equal.
 For simulation purposes we chose $W_S=W_I=W_R=1$ for random testing, and $W_S=0.3$ and $W_I=W_R=1$ for non-random testing strategy.
 Note that the critical contour of $\Rnum=1$ is plotted in solid line in Figure \ref{pan}.
+
+\david{Details such as the fact that you used ggplot are not important to state in the main text.
+On the other hand, there isn't a clear narrative here.
+The reader should feel that a story is being told.}
 
@@
 The following Proposition is the direct result of taking partial derivative of $\Rnum$ \eqref{R0} with respect to selected parameters.
-Note that we used Taylor approximation of $\Rnum$ at $\rho=0$ due to the complexity of the expression and thus inconclusive.
+Note that we used Taylor approximation of $\Rnum$ at $\rho=0$ due to the complexity of the expression and thus inconclusive.\david{in that case it would be good to establish numerically that the relationships given in the proposition are valid more generally, at least for reasonable parameter ranges}
 See the Supplementary material for details.
 
+\david{make sure to use $\Rnum$ not $R_0$ below (and often in the supp material section).
+More importantly, you really need to state in words what this proposition is saying.
+The reader has to check the param table and think it through themselves.}
 \begin{proposition}
@@
 
- Given that the perfect isolation occurs at $\eta_c=\eta_w = 0$, Prop.~\ref{p1:eta} means that lifting the isolation from awaiting group results in an increased $\Rnum$ and consequently a greater number of infected individuals.
+Given that the perfect isolation occurs at $\eta_c=\eta_w = 0$,
+\david{That's a perfect example of the kind of phrasing that really must be changed.
+You need to say something like ``occurs when no one transmits while waiting for test results (i.e., $\eta_c=\eta_w = 0$)''.
+It makes the text so much easier to process and doesn't require the reader to keep looking at the table to make sense of what you're saying.}
+Prop.~\ref{p1:eta} means that lifting the isolation from awaiting group results in an increased $\Rnum$ and consequently a greater number of infected individuals.
 Prop.~\ref{p1:rho} indicates that increasing the \percap testing intensity, $\rho$, reduces $\Rnum$.
@@
 It is straightforward to derive \ref{p1:rho} and \ref{p1:omega} from the Taylor approximation of $\Rnum$ at $\rho=0$.
+\david{can we gain some insight from the expressions for the partial derivatives in the proposition, as opposed to just their signs?}
 
@@
 Furthermore, we derived an inequality that quantifies the exact relationships between model parameters that result in returning tests more rapidly being favorable \eqref{eq:necsuf}.
+\david{What do we learn from this inequality?  Why is it relegated to an appendix?}
 
@@
 Here, we develop a compartmental SIR-type model to study the potential effect of testing strategies, testing intensity, test sensitivity and specificity, test reporting time and isolation on epidemic dynamics.
-While targeted testing strategies are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is viral spread is expected to be stronger for a slow test than for a fast test.
+While targeted testing strategies 
+\david{you need to be clearer: you mean targeting individuals who have contacted people who were infected.
+For example, here you could say something like ``While targeting the contacts of confirmed cases\dots''}
+are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is that viral spread
+is greater for a slow test than for a fast test.
 This counter-intuitive effect can occur when people are cautious when awaiting a test result, and may not be robust to second-order effects of fast testing (such as better contact tracing).
@@
 We incorporated the compartment-specific relative testing weights, $W_S$, $W_I$ and $W_R$, to model random testing and \emph{targeted} testing strategies.
-Here, in the case of targeted testing and for the simplicity and illustraition purposes, we assumed that infected and recovered individuals are tested at three times the \percap rate of susceptible individuals, thus $W_I/W_S=3$ and $W_R=W_I$.
-Note that we have not specified a methodology to assigne particular relative testing weights corresponding to a particular targeted testing scenario.
+Here, in the case of targeted testing and for the simplicity and illustration purposes, we assumed that infected and recovered individuals are tested at three times the \percap rate of susceptible individuals, thus $W_I/W_S=3$ and $W_R=W_I$.
+Note that we have not specified a methodology to assign particular relative testing weights corresponding to a particular targeted testing scenario.
 This part needs to be developed further in future work. 
 Modeling different targeted testing strategies, equivalently test-specific testing weights in our framework, requires prior information of the conditional probabilities of getting tested for people in a given compartment.
-This can be implied when we would like to quantify and compare the effect of different levels of test forcus for infectious people on the basic reproduction number $\Rnum$, and conclude about the disease spread management.
+This can be implied when we would like to quantify and compare the effect of different levels of test forcus \david{??} for infectious people on the basic reproduction number $\Rnum$, and conclude about the disease spread management.
 For example, when people are tested for "screening", the individuals with potential higher mobility, eg.
@@
 Because many infected asymptomatic individuals will believe they are uninfected/uninfectious, thus may unknowingly spread the virus to many others.
-Agian the delay in the test reporting rate keeps these individuals in the ``\emph {safe}'' awaiting-confirmed compartments.
+Again the delay in the test reporting rate keeps these individuals in the ``\emph {safe}'' awaiting-confirmed compartments.
 
@@
 In particular, doubling the sensitivity of the pre-screening tool would *do something* to $\Rnum$.
+
+\david{Some of this reads like notes for discussion rather than text for the paper, so I'm not trying to edit.}
 

### 2021-02-07 SIR_manuscript.tex
working on abstract, rm figures dir

@@
 \newcommand{\Rnum}{\mathcal{R}_0}
+\newcommand{\covid}{COVID-19\xspace}
 \newcommand{\comment}{\showcomment}
@@
 
+The potential advantage of slow test reporting on the epidemic dynamics, motivated by the \covid epidemic, is real and needs to be quantified.
+
+We investigated this advantage by developing a SIR-type model with testing and isolation components.
+The novel part of the model is compartment-specific relative testing weights which reflects the testing strategies.
+
+We also compare \covid to an individual-level advantage of fast tests: people who test positive
+may be even more careful. we are missing out on community-level advantages of fast testing: better assessment of the situation, identification of hot spots, contact-tracing, etc.
+
+The novel component of the model comes in through the compartment-specific relative testing weights $W_S$, $W_I$ and $W_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively.
+
 % %%%%%%%
@@
 
-The observed dynamics of the COVID-19 epidemic are driven by both epidemiological processes (infection and recovery) and testing processes (testing and test reporting).
+The observed dynamics of the \covid epidemic are driven by both epidemiological processes (infection and recovery) and testing processes (testing and test reporting).
 In addition to shaping epidemic observations (via case reports), testing processes can also affect epidemiological dynamics.
@@
 
-Over the course of the COVID19 \ali{consistency COVID-19 or COVID19?}\david{create a \texttt{\detokenize{\covid}} macro and use that.
-I suggest defining it as COVID-19 since that is more common.} pandemic, however, the vast majority of testing has been done with other goals --
+Over the course of the \covid pandemic, however, the vast majority of testing has been done with other goals --
 primarily diagnosis (determining COVID19 infection status for clinical purposes), or control (determining COVID19 status in order to quickly isolate cases that have been found by contact tracing), which we characterize as \emph{targeted} testing strategies.

### 2021-02-08 SIR_manuscript.tex
Resolved todo{Fady} comments

@@
 In addition to the favorable-delay-reporting observation of our model \eqref{model}, the model enables us to quantify the amount of delay required in the test reporting process as a strategy to reduce $\Rnum$.
-To give a biological interpretation, we describe the qualitative trends predicted by the inequality \eqref{eq:necsuf} \todo{Fady please confirm}.
-To summarise, returning test results more rapidly tends to be more favorable, i.e., reduces $\Rnum$, when (i) \emph{Confirmed-positive} individuals, lower their contact much more than individuals who are waiting for their test results (i.e., $\eta_c \ll \eta_w$), (ii) the test is highly sensitive, i.e., $p_I$ is close to 1, and (iii) the targeted testing strategy is practiced.
+To give a biological interpretation, we describe the qualitative trends predicted by the inequality \eqref{eq:necsuf} \todo{Fady please confirm}\fady{yes, looks good to me}.
+To summarise, returning test results more rapidly tends to be favourable (i.e., reduces $\Rnum$) when (i) \emph{Confirmed-positive} individuals, lower their contact much more than individuals who are waiting for their test results (i.e., $\eta_c \ll \eta_w$), (ii) the test is highly sensitive (i.e., $p_I$ is close to 1), and (iii) the targeted testing strategy is used (i.e., $W_i \gg W_s$).
 
@@
 
-\todo{Fady, can you clean this up}
-\begin{align}\label{eq:necsuf}
+\todo{Fady, can you clean this up}\fady{Done (I think)}
+\begin{align}
     &\omega > \omega_2 \nonumber \\
     &\omega > \frac{\gamma}{\sqrt{1-Q}-1} \nonumber \\
-    &\vdots \nonumber \\ 
-    &\frac{1-n_{t}}{1-n_{w}}P_{i}>\frac{W_{s}}{W_{i}}\left(\frac{\gamma}{\omega}+1\right)^{2}
+    &\sqrt{1-Q} > \frac{\gamma}{\omega}+1 \\
+    &1-Q > (\frac{\gamma}{\omega}+1)^2 \\
 \end{align}
+
+Substituting in $Q$ from \eqref{eq:defQ} we have
+
+\begin{align}
+    &1-\frac{W_i}{W_s}\left(1-\frac{n_{t}-1}{n_{\omega}-1}P_{i}\right)> \left(\frac{\gamma}{\omega}+1\right)^2 \\
+    &-\frac{W_i}{W_s}\left(1-\frac{n_{t}-1}{n_{\omega}-1}P_{i}\right)> \left(\frac{\gamma}{\omega}+1\right)^2+1 \\    
+    &\frac{W_i}{W_s}\left(\frac{1-n_{t}}{1-n_{\omega}}P_{i}-1\right)> \left(\frac{\gamma}{\omega}+1\right)^2+1\label{eq:necsuf}.
+\end{align}
+
+Since all steps in deriving \eqref{eq:necsuf} are reversible, \eqref{eq:necsuf} gives a necessary and sufficient condition for $\omega > \omega_2$, which characterizes when returning tests more rapidly would cause a decrease in $\Rnum$.
 

### 2021-02-08 SIR_manuscript.tex
abstract added

@@
 
-The potential advantage of slow test reporting on the epidemic dynamics, motivated by the \covid epidemic, is real and needs to be quantified.
-
-We investigated this advantage by developing a SIR-type model with testing and isolation components.
-The novel part of the model is compartment-specific relative testing weights which reflects the testing strategies.
-
-We also compare \covid to an individual-level advantage of fast tests: people who test positive
-may be even more careful. we are missing out on community-level advantages of fast testing: better assessment of the situation, identification of hot spots, contact-tracing, etc.
-
-The novel component of the model comes in through the compartment-specific relative testing weights $W_S$, $W_I$ and $W_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively.
+The potential advantage of slow rate of test return on the dynamics of an epidemic is real, often neglected and needs to be quantified.
+In the individual level, this advantage can be in the following sense: people who are awaiting the test results or have tested positive may partially or fully self-isolate, thus reducing or eliminating their potential infectious contacts.
+In the community-level (e.g., nursing homes, LTC facilities, etc.), the fast test reporting may be  more benefitial as it gives a fast assessment of the situation, hot spots and contact-tracing.  
+In this paper, we investigated this individual-level effect of testing processes on the epidemic dynamics by developing a SIR-type model.
+Although the model developement was motivated by the \covid epidemic, the model has generic  epidemiological and testing structures.
+The realistic components of the model include testing intensity, test sensitivity and specificity, rate of test return and isolation.
+The novel component is the compartment-specific relative testing weights which reflect the testing strategies determined by the purpose of testing-- surveillance, diagnosis or control.
+Here, we showed that the slow test reporting can be benefitial on the dynamics of an epidemic.
+In particular, the sensitivity of the basic reproduction number, $\Rnum$, with respect to the rate of test return can be positive.
+Further, we compare two testing strategies, random vs.
+targeted, in this framework.
+We conclude that the targeted testing strategy is more effective than the random one in the sense that a lower range of testing intensity is required to keep $\Rnum <1$.
+Also, targeted testing reduces the individual-level of advantage of slow rate of test.
 

### 2021-02-09 SIR_manuscript.tex
Some more edits

--- 2021-02-08 SIR_manuscript.tex
@@
 
-The potential advantage of slow test reporting on the epidemic dynamics, motivated by the \covid epidemic, is real and needs to be quantified.
+The potential advantage of a slow rate of test return on the dynamics of an epidemic is real, often neglected, and needs to be quantified.
+At the individual level, this advantage can manifest in the following sense: individuals awaiting their test results or who have tested positive may partially or fully self-isolate, thus reducing or eliminating their potential in the transmission process.
 
-We investigated this advantage by developing a SIR-type model with testing and isolation components.
-The novel part of the model is compartment-specific relative testing weights which reflects the testing strategies.
+At the community-level (e.g., nursing homes, LTC facilities, etc.), fast test reporting may be more beneficial.
+It gives a rapid assessment of the situation, identifies hot spots, and may enable rapid contact-tracing.
+In this paper, we investigated this individual-level effect of testing processes on the epidemic dynamics by developing a SIR-type model.
 
-We also compare \covid to an individual-level advantage of fast tests: people who test positive
-may be even more careful. we are missing out on community-level advantages of fast testing: better assessment of the situation, identification of hot spots, contact-tracing, etc.
+Although the model development was motivated by the \covid epidemic, the model has general epidemiological and testing structures.
+The realistic components of the model include testing intensity, test sensitivity and specificity, rate of test return, and isolation.
+The novel component is the compartment-specific relative testing weights, which reflect the testing strategies-- surveillance, diagnosis, or control.
+Here, we showed that the slow test reporting can be beneficial on the dynamics of an epidemic.
+In particular, it is possible for the basic reproduction number, $\Rnum$, to be increasing with respect to the rate of test return.
 
-The novel component of the model comes in through the compartment-specific relative testing weights $W_S$, $W_I$ and $W_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively.
+Further, we compare two testing strategies, random vs.
+targeted, in this framework.
+We conclude that the targeted testing strategy is more effective than the random one in the sense that a lower range of testing intensity is required to keep $\Rnum <1$.
+Also, targeted testing reduces the individual-level of advantage of a slow rate of test return.
+
+\fady{General thoughts on the abstract (take with a grain of salt): I think the abstract is eye-catching and interesting.
+I think leading in by highlighting potential advantages to a slow test return may be ``provocative" (which may not necessarily be bad, I'm not sure).
+I think the second paragraph in some sense undermines or clashes with the first paragraph, especially since we don't actually address these potential advantages in this model.
+Maybe one way to spin this around would be to lead in with a discussion of the advantages of a FAST test return (i.e., the second paragraph), and then say ``but situations where returning tests more rapidly is BAD have not been studied".
+I think that could have better harmony or coordination.}
 

### 2021-02-09 SIR_manuscript.tex
edited start of the discussion

@@
 
-Mathematical modeling of infectious disease outbreaks provides insights on how testing processes influence the epidemiological processes through isolation. 
-Here, we develop a compartmental SIR-type model to study the potential effect of testing strategies, testing intensity, test sensitivity and specificity, test reporting time and isolation on epidemic dynamics.
-While targeted testing strategies 
+Mathematical modelling of infectious disease outbreaks provides insights on how testing processes influence the epidemiological processes through isolation.
+Here, we develop a compartmental SIR-type model to study the potential effect of testing strategies, testing intensity, test sensitivity and specificity, test reporting time, and isolation on epidemic dynamics.
+
+In agreement with our intuition, targeted testing strategies (i.e., strategies that identify the most infectious cases) are always more effective than random testing at lowering $\Rnum$.
+This reinforces the benefit of contact tracing in slowing transmission since tracing contacts of positive cases increases the likelihood of identifying other positive cases.
+Another approach that may enhance targeting accuracy may be to use pre-testing selection criteria that must be met by individuals wishing to be tested (e.g., symptoms questionnaires).
+While such an approach may enhance targeting, it may carry significant downsides (e.g., lessening the surveillance breadth of testing), which must be judged against the benefits.
 \david{you need to be clearer: you mean targeting individuals who have contacted people who were infected.
-For example, here you could say something like ``While targeting the contacts of confirmed cases\dots''}
-are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is that viral spread
-is greater for a slow test than for a fast test.
-This counter-intuitive effect can occur when people are cautious when awaiting a test result, and may not be robust to second-order effects of fast testing (such as better contact tracing).
+For example, here you could say something like ``While targeting the contacts of confirmed cases\dots''}\fady{I moved this comment below the paragraph (so David is referring to the previous paragraph) and tried to resolve it.
+What do you think?}
+
+Less intuitively, our analysis suggests that there are cases where reporting test results more rapidly could increase, rather than decrease $\Rnum$.
+This can occur when individuals are much more cautious when awaiting a test result than when they believe they are not infected, and may not be robust to second-order effects of fast testing (such as better contact tracing).
 
 We incorporated the compartment-specific relative testing weights, $W_S$, $W_I$ and $W_R$, to model random testing and \emph{targeted} testing strategies.
-Here, in the case of targeted testing and for the simplicity and illustration purposes, we assumed that infected and recovered individuals are tested at three times the \percap rate of susceptible individuals, thus $W_I/W_S=3$ and $W_R=W_I$.
-Note that we have not specified a methodology to assign particular relative testing weights corresponding to a particular targeted testing scenario.
+In the case of targeted testing and for the simplicity and illustration purposes, we assumed that infected and recovered individuals are tested at three times the \percap rate of susceptible individuals (thus $W_I/W_S=3$ and $W_R=W_I$).
+Note that we have not specified a methodology to assign particular relative testing weights corresponding to a particular targeted testing scenario.\fady{I don't understand the previous sentence}
 This part needs to be developed further in future work. 
 Modeling different targeted testing strategies, equivalently test-specific testing weights in our framework, requires prior information of the conditional probabilities of getting tested for people in a given compartment.
-This can be implied when we would like to quantify and compare the effect of different levels of test forcus \david{??} for infectious people on the basic reproduction number $\Rnum$, and conclude about the disease spread management.
+This can be implied when we would like to quantify and compare the effect of different levels of test forcus \david{??}\fady{I don't get this part either.} for infectious people on the basic reproduction number $\Rnum$, and conclude about the disease spread management.
 For example, when people are tested for "screening", the individuals with potential higher mobility, eg.

### 2021-02-09 SIR_manuscript.tex
abstract revised, some David's comments were addressed

--- 2021-02-09 SIR_manuscript.tex
@@
 
-The potential advantage of a slow rate of test return on the dynamics of an epidemic is real, often neglected, and needs to be quantified.
+The effect of testing processes, including (testing and test reporting) on an epidemic dynamics (infection and recovery) can be studied at the individual level or the community level (e.g., nursing homes, LTC facilities, etc.).
+Gaining insights to determine the sensitivity of the epidemic dynamics with respect to the testing processes will depend on underlying factors including the level of focus (individual or community), assumptions (model), and the interplay between these factors.
+In particular, the fast test reporting may be beneficial at the community-level, supported by many studies, as it gives a rapid assessment of the situation, identifies hot spots, and may enable rapid contact-tracing.
+However, the potential advantage of a slow rate of test return on the dynamics of an epidemic is real, often neglected, and needs to be quantified.
 At the individual level, this advantage can manifest in the following sense: individuals awaiting their test results or who have tested positive may partially or fully self-isolate, thus reducing or eliminating their potential in the transmission process.
-
-At the community-level (e.g., nursing homes, LTC facilities, etc.), fast test reporting may be more beneficial.
-It gives a rapid assessment of the situation, identifies hot spots, and may enable rapid contact-tracing.
 In this paper, we investigated this individual-level effect of testing processes on the epidemic dynamics by developing a SIR-type model.
-
 Although the model development was motivated by the \covid epidemic, the model has general epidemiological and testing structures.
@@
 In particular, it is possible for the basic reproduction number, $\Rnum$, to be increasing with respect to the rate of test return.
-
 Further, we compare two testing strategies, random vs.
@@
 Also, targeted testing reduces the individual-level of advantage of a slow rate of test return.
-
-\fady{General thoughts on the abstract (take with a grain of salt): I think the abstract is eye-catching and interesting.
-I think leading in by highlighting potential advantages to a slow test return may be ``provocative" (which may not necessarily be bad, I'm not sure).
-I think the second paragraph in some sense undermines or clashes with the first paragraph, especially since we don't actually address these potential advantages in this model.
-Maybe one way to spin this around would be to lead in with a discussion of the advantages of a FAST test return (i.e., the second paragraph), and then say ``but situations where returning tests more rapidly is BAD have not been studied".
-I think that could have better harmony or coordination.}
 
@@
 why. For a given testing intensity (tests performed per day), testing will have a stronger effect on infection rates
-the more strongly it is focused on people who are infectious.\david{deleted `` for SARS-CoV-2''.
-Since this is an idealized model intended for insights, it is probably better to use more generic language most of the time and not say covid unless necessary in a specific context.}
+the more strongly it is focused on people who are infectious.
 This level of focus depends in turn on the purpose and design of testing programs.
 (1) When testing is done for the purposes of disease surveillance \citep{foddai2020base}
-we might expect tests to be assigned randomly across the population (in an ideal world: some bias in selection for testing is probably inevitable), although a stratified sampling design would be preferred for statistical efficiency \citep{graubard1996modelling} \ali{better ref?}.
+we might expect tests to be assigned randomly across the population (in an ideal world: some bias in selection for testing is probably inevitable), although a stratified sampling design would be preferred for statistical efficiency \citep{graubard1996modelling} \ali{a better ref?}.
 
 Over the course of the \covid pandemic, however, the vast majority of testing has been done with other goals --
-primarily diagnosis (determining COVID19 infection status for clinical purposes), or control (determining COVID19 status in order to quickly isolate cases that have been found by contact tracing), which we characterize as \emph{targeted} testing strategies.
+primarily diagnosis (determining the infection status for clinical purposes), or control (determining the infection status in order to quickly isolate cases that have been found by contact tracing), which we characterize as \emph{targeted} testing strategies.
 In these cases, testing probabilities vary widely across epidemiological compartments; in our dynamical model, we will characterize these probabilities by assigning a \percap testing weight to each compartment that determines the \emph{relative} probability that an individual in that compartment will be selected for testing (see Methods).
 
-When testing is used primarily for diagnosis it will focus on people with COVID19-like symptoms; thus the testing weights for people with \david{escape spaces after dots unless it is a full stop}vs.\ without SARS-CoV-2 infections will depend on the relative magnitudes of the probability of having COVID19-like symptoms when infected with SARS-CoV-2 (i.e., the relative proportions of presymptomatic or asymptomatic vs.\ symptomatic infections) and the probability of having COVID19-like symptoms when \emph{not} infected (e.g., due to other respiratory illnesses).
-Testing for epidemic control will focus on people who are known to have been in contact with known COVID19 cases; in this case the testing weights for infected vs.\ uninfected people will depend on the probability of infection given contact, as well as the thoroughness and effectiveness of the system for identifying suspicious contacts.
+When testing is used primarily for diagnosis it will focus on people with infection-like symptoms; thus the testing weights for people with vs.\ without the infections will depend on the relative magnitudes of the probability of having the infection-like symptoms when infected.
+Specifically, for \covid infection, the testing weights will depend on the relative proportions of presymptomatic or asymptomatic vs.\ symptomatic infections and the probability of having \covid-like symptoms when \emph{not} infected (e.g., due to other respiratory illnesses).
+Testing for epidemic control will focus on people who are known to have been in contact with known infected cases; in this case the testing weights for infected vs.\ uninfected people will depend on the probability of infection given contact, as well as the thoroughness and effectiveness of the system for identifying suspicious contacts.
 
-In addition to the overall testing intensity, the testing weights for different compartments, and the degree of self-isolation of test-pending and test-confirmed people, our models will also need to include test specificity and sensitivity\david{I presume you mean that models would ideally take account of these things, which should not be mentioned unless they are defined since the terms are often misunderstood}, as well as the distribution of times from testing to reporting.
-Here, we focus on the the effect of testing intensity, different levels of testing ``focus'' (from random to highly targeted), and rate of test processing on a single index of epidemiological dynamics, the basic reproduction number $\Rnum$.
-
-The early dynamics of an epidemic are determined by $\Rnum$.
+In order to understand the effect of testing processes on an epidemic dynamics, we deveopled a mechanistic model with epidemic and testing components.
+Here, we focus on the the effect of testing intensity, different levels of testing ``focus'' (from random to highly targeted), and rate of test return on a single index of epidemiological dynamics, the basic reproduction number $\Rnum$, which is useful to estimate the early dynamics of an epidemic.
 This index is defined as the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}.
@@
 \citep{shaw2021what}, it does give a simple and widely accepted index for the difficulty of control, as well as some indication of the likely final size of an epidemic \citep{ma2006generality}.
-\david{This paragraph does not flow from the rest of the intro, and is a weird way to end when the point of the paper is about testing.}
+Our model provides insights to the sensitivity of the epidemiological dynamics, through $\Rnum$, with respect to the undelying testing processes.
+\david{This paragraph does not flow from the rest of the intro, and is a weird way to end when the point of the paper is about testing.}\ali{edited, but still needs work}
 
-\david{There is some back and forth between generic discussion and COVID-specific comments.  I think COVID should be used as motivation, but that the whole paper should be framed generically.}
+\david{There is some back and forth between generic discussion and \covid-specific comments.  I think COVID should be used as motivation, but that the whole paper should be framed generically.}
 
@@
 We developed a deterministic model, Eqs.
-\eqref{model}\david{relegating the equations to supp is not OK.
-You don't want to publish a paper in which the model is specified exactly only in online supplementary material.
-I don't think there is any need for supp material.
-You can have appendices, but I would keep everything as part of the paper.}, which groups individuals based on disease status and testing status.
+\eqref{model}, which groups individuals based on disease status and testing status.
 Disease states include Susceptible, Infectious and Recovered (thus this is an SIR-type model), and testing status categorizes people as \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive} (Figure~\ref{fig:flowchart}).
 Two `accumulator' compartments, $N$ and $P$, were also incorporated in the model in order to collect cumulative reported negative or positive tests.
-The model and details of calculation of the basic reproduction number $\Rnum$ are presented in the Supplementary Materials.
+The model and details of calculation of the basic reproduction number $\Rnum$ are presented in the appendixs.
 
@@
 \end{equation}
-Thus, the \percap testing rate for compartment $Z$ is $F_Z=\sigma W_Z$, where $Z \in \{S,I,R\}$.
-\david{For a generic compartment, I would suggest $X$ rather than $Z$, since $Z$ is often the final size.}  For a high-sensitivity test, infected people typically flow through to the ``confirmed positive'' ($I_c$, $R_c$) compartments and are thus unavailable for further testing.
+Thus, the \percap testing rate for compartment $X$ is $F_X=\sigma W_X$, where $X \in \{S,I,R\}$. 
+For a high-sensitivity test, infected people typically flow through to the ``confirmed positive'' ($I_c$, $R_c$) compartments and are thus unavailable for further testing.
 Over the course of the epidemic, a fixed testing rate as specified in \eqref{sigma} can (if large enough) exhaust the pool of people available for testing, leading to a singularity when no one is left untested.
-Although this phenomenon does not affect our analysis of $\Rnum$, it can affect the temporal dynamics (we discuss an adjustment to the model that solves this problem in the Supplementary material).
+Although this phenomenon does not affect our analysis of $\Rnum$, it can affect the temporal dynamics (we discuss an adjustment to the model that solves this problem in the appendix).
 
@@
 So $S_{\rm u}$ and $I_{\rm p}$ etc.
-Then it is clear the subscript isn't a variable with a value.}
+Then it is clear the subscript isn't a variable with a value.}\ali{will do}
 
-The classical SIR model is based on the following implicit assumptions; well-mixed population, homogeneity of the population\david{What's the difference btw ``well-mixed'' and ``homoegeneous''?}, exponentially distributed duration of infection and large population size (see, e.g., \cite{keeling2011modeling}).
+The classical SIR model is based on the following implicit assumptions; well-mixed population, homogeneity of the population\david{What's the difference btw ``well-mixed'' and ``homoegeneous''?}\ali{ non structured eg age, ok?}, exponentially distributed duration of infection and large population size (see, e.g., \cite{keeling2011modeling}).
 In addition to these standard assumptions, our model, \ref{model}, assumes: (i) there is a single force of infection, $\Lambda$, defined as follows
@@
 \end{equation}
-across all susceptible pools (see Table~\ref{tab:params} for detail\david{there are no details in the table; you need to do a better job of explaining what's going on}), (ii) $\eta_c \leq \eta_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
+across all susceptible pools (see Table~\ref{tab:params} for details\david{there are no details in the table; you need to do a better job of explaining what's going on}), (ii) $\eta_c \leq \eta_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
 For this analysis, we also assume a perfectly specific test ($p_S=0$).
@@
 The DFE is
-\david{don't put blank lines around equations unless you are started a new paragraph.  comment out blank lines if want the .tex file to be more readable}
-%%
 \begin{equation}
@@
 The basic reproduction number, $\Rnum$, was calculated by using the next generation matrix method developed by \cite{van2002reproduction}. $\Rnum$ is
-
 \begin{equation}
 \label{R0}
-\Rnum= (A \times S_u^* + B \times S_n^*) \times C, 
+\Rnum= \frac{\beta}{\gamma} (A \times S_u^* + B \times S_n^*) \times C, 
 \end{equation}
@@
 B=& \big(\omega+(F_I+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_c\omega) \omega p_I F_I }{\omega+\gamma}, \\ 
-C=& \frac{\beta/\gamma}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}.
+C=& \frac{1}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}.
 \end{aligned}
@@
 It would be worth spending some time trying to write the expression in a way that helps the reader understand where the different factors come from.}
-Further details of derivation of $\Rnum$ are provided in the Supplementary material section.
+Further details of derivation of $\Rnum$ are provided in the appendix section.
 
@@
 Note that we used Taylor approximation of $\Rnum$ at $\rho=0$ due to the complexity of the expression and thus inconclusive.\david{in that case it would be good to establish numerically that the relationships given in the proposition are valid more generally, at least for reasonable parameter ranges}
-See the Supplementary material for details.
+See the appendix for details.
 
-\david{make sure to use $\Rnum$ not $R_0$ below (and often in the supp material section).
+\david{make sure to use $\Rnum$ not $\Rnum$ below (and often in the supp material section).
 More importantly, you really need to state in words what this proposition is saying.
@@
 \item \label{p1:eta}
-$\partial{R_0}/\partial{\eta_c} \geq 0$ and $\partial{R_0}/\partial{\eta_w} \geq 0$. 
+$\partial{\Rnum}/\partial{\eta_c} \geq 0$ and $\partial{\Rnum}/\partial{\eta_w} \geq 0$. 
 \item \label{p1:rho}
@@
 \item \label{p1:omega}
-$\partial{R_0}/\partial{\omega}$ can be positive or negative when $\rho \approx 0$.
+$\partial{\Rnum}/\partial{\omega}$ can be positive or negative when $\rho \approx 0$.
 \end{enumerate}
@@
 Modeling different targeted testing strategies, equivalently test-specific testing weights in our framework, requires prior information of the conditional probabilities of getting tested for people in a given compartment.
-This can be implied when we would like to quantify and compare the effect of different levels of test forcus \david{??} for infectious people on the basic reproduction number $\Rnum$, and conclude about the disease spread management.
+This can be implied when we would like to quantify and compare the effect of different levels of test focus for infectious people on the basic reproduction number $\Rnum$, and conclude about the disease spread management.
 For example, when people are tested for "screening", the individuals with potential higher mobility, eg.
@@
 Proposition \ref{prop1} part \ref{p1:rho} and also Figure \ref{pan} indicates that increasing testing intensity $\rho$ reduces $\Rnum$.
-This is sensible since as people are moved to test compartments, namely $Z_p$ and $Z_n$ for $Z \in \{S,I,R\}$, they may partially or fully self-isolate.
-Also individuals who have tested positive, namely $Z_c$ compartment for $Z \in \{S,I,R\}$, are highly likely to self-isolate.
+This is sensible since as people are moved to test compartments, namely $X_p$ and $X_n$ for $X \in \{S,I,R\}$, they may partially or fully self-isolate.
+Also individuals who have tested positive, namely $X_c$ compartment for $X \in \{S,I,R\}$, are highly likely to self-isolate.
 The higher probability of being subject to isolation, the lower the force of infection \eqref{Lambda} and the lower $\Rnum$ will become.
@@
 Individuals are highly likely to fully self-isolate when they are either awaiting the test results or they are reported, thus reducing or eliminating their potentially infectious contacts.
-Thus, the faster the test reporting rate, $\omega$, the shorter these individuals stay in the ``\emph {safe}'' awaiting-confirmed compartments, namely $Z_n$, $Z_p$ and $Z_c$ for $Z \in \{S,I,R\}$, and the more they get involved in the infection process.
+Thus, the faster the test reporting rate, $\omega$, the shorter these individuals stay in the ``\emph {safe}'' awaiting-confirmed compartments, namely $X_n$, $X_p$ and $X_c$ for $X \in \{S,I,R\}$, and the more they get involved in the infection process.
 This advantage of slow test reporting is real, and neglected. 
@@
 In contrast, if an expensive test can identify on average more infected individuals, it will not necessarily lower $\Rnum$ more than the cheap test.
-The use of tests cheaper than RT-PCR has been proposed as a potential strategy for containing the COVID-19 pandemic.
+The use of tests cheaper than RT-PCR has been proposed as a potential strategy for containing the COVID pandemic.
 While cheaper tests may be less sensitive and reliable than RT-PCR, they allow for broader and more intense testing.
@@
 In addition to the favorable-delay-reporting observation of our model \eqref{model}, the model enables us to quantify the amount of delay required in the test reporting process as a strategy to reduce $\Rnum$.
-To give a biological interpretation, we describe the qualitative trends predicted by the inequality \eqref{eq:necsuf} \todo{Fady please confirm}\fady{yes, looks good to me}.
+To give a biological interpretation, we describe the qualitative trends predicted by the inequality \eqref{eq:necsuf}.
 To summarise, returning test results more rapidly tends to be favourable (i.e., reduces $\Rnum$) when (i) \emph{Confirmed-positive} individuals, lower their contact much more than individuals who are waiting for their test results (i.e., $\eta_c \ll \eta_w$), (ii) the test is highly sensitive (i.e., $p_I$ is close to 1), and (iii) the targeted testing strategy is used (i.e., $W_i \gg W_s$).
@@
 \begin{center}
-\textbf{\large Supplemental Materials: Title for main text}
+\textbf{\large Appendix}
 \end{center}
@@
 \makeatletter
-\renewcommand{\theequation}{S\arabic{equation}}
-\renewcommand{\thefigure}{S\arabic{figure}}
-\renewcommand{\bibnumfmt}[1]{[S#1]}
-\renewcommand{\citenumfont}[1]{S#1}
+\renewcommand{\theequation}{A\arabic{equation}}
+\renewcommand{\thefigure}{A\arabic{figure}}
+\renewcommand{\bibnumfmt}[1]{[A#1]}
+\renewcommand{\citenumfont}[1]{A#1}
 %%%%%%%%%% Prefix a "S" to all equations, figures, tables and reset the counter %%%%%%%%%%
@@
 The next generation matrix for this model is $G = F V^{-1}$, where matrix $F$ represents the inflow of new infection to the infected compartments and matrix $V$ represents the flow in the infected compartments when the population is totally susceptible.
-Matrices $F$ and $V$ are
-
+Matrices $F$ and $V$ are \ali{all $S_u \& S_n$ should be $S_u^* \& S_n^*$}
 \begin{align}
@@
 \end {array} \right], \text{ where } \\
-G_{11} =C
+G_{11} = \frac{\beta}{\gamma}\ C
 \left[\begin {array}{cc}
@@
 \end{equation}
-where expressions $A$, $B$ and $C$ are specified in \eqref{eq:abc}.
+with expressions $A$, $B$ and $C$ are specified in \eqref{eq:abc}.
 The block matrix $G_{12}$ does not influence $\Rnum$ (defined as the spectral radius of $G$). All that matters here are the eigenvalues of $G_{11}$, which are 0 and $\Rnum$ \eqref{R0}.
@@
 \label{eq:R0appr}
-R_0 \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_S} \Big(\gamma(\eta_w-1)(\gamma W_S+\omega W_I) + (\eta_c -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2).
+\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_S} \Big(\gamma(\eta_w-1)(\gamma W_S+\omega W_I) + (\eta_c -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2).
 \end{equation}
@@
 \label{Rom}
-\partial{R_0}/\partial{\omega}=  \frac{-\beta \rho}{\gamma W_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c),
+\partial{\Rnum}/\partial{\omega}=  \frac{-\beta \rho}{\gamma W_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c),
 \end{equation}
@@
 If $a\geq 0$, the quadratic expression in \eqref{Rom}, has Real roots.
-Assuming that $\omega_1<0$ and $\omega_2>0$ be the roots of the quadratic expression in $\partial{R_0}/\partial{\omega}$.
-Thus, $\partial{R_0}/\partial{\omega}>0$ for $0<\omega<\omega_2$ and  $\partial{R_0}/\partial{\omega}<0$ for $\omega>\omega_2$.
+Assuming that $\omega_1<0$ and $\omega_2>0$ be the roots of the quadratic expression in $\partial{\Rnum}/\partial{\omega}$.
+Thus, $\partial{\Rnum}/\partial{\omega}>0$ for $0<\omega<\omega_2$ and  $\partial{\Rnum}/\partial{\omega}<0$ for $\omega>\omega_2$.
 
@@
 \subsection{rate of returning tests} 
+\ali{needs editig}
 The linearization of $\Rnum$ around $\rho=0$ is
@@
 \end{equation}
-
-So when $\rho \approx 0$ we have $$\partial{\Rnum}/\partial{\omega} \approx  \frac{-\beta \rho}{\gamma W_s\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c).$$
-
+So when $\rho \approx 0$ we have $$\partial{\Rnum}/\partial{\omega} \approx  \frac{-\beta \rho}{\gamma W_s\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c),$$
 where $a=(\eta_w-1)W_i-(\eta_c-1)P_iW_i$, $b=2(\eta_w-1)\gamma W_s$ and $c=(\eta_w-1)\gamma^2 W_s$.
@@
 In order to gain insight into this intriguing behavior, we examine the zeroes of $\frac{\partial{\Rnum}}{\partial{\omega}}(\omega)$.
-
-Defining the following quantity, $Q$, will help us write the roots of $\partial{\Rnum}/\partial{\omega}$ neatly. 
+Defining the following quantity, $Q$, will help us write the roots of $\partial{\Rnum}/\partial{\omega}$ neatly as follows.
 \begin{align}\label{eq:defQ}
-    Q =& \frac{W_i}{W_s}\left(1-\frac{n_{t}-1}{n_{w}-1}P_{i}\right) \\
+    Q =& \frac{W_i}{W_s}\left(1-\frac{n_{t}-1}{n_{w}-1}P_{i}\right).
 \end{align}
-
 With that in mind, we can write the roots of $\partial{\Rnum}/\partial{\omega}$ as
@@
 So
-
-\todo{Fady, can you clean this up}\fady{Done (I think)}
 \begin{align}
@@
     &\sqrt{1-Q} > \frac{\gamma}{\omega}+1 \\
-    &1-Q > (\frac{\gamma}{\omega}+1)^2 \\
+    &1-Q > (\frac{\gamma}{\omega}+1)^2.
 \end{align}
-
 Substituting in $Q$ from \eqref{eq:defQ} we have
-
 \begin{align}
@@
 \end{align}
-
 Since all steps in deriving \eqref{eq:necsuf} are reversible, \eqref{eq:necsuf} gives a necessary and sufficient condition for $\omega > \omega_2$, which characterizes when returning tests more rapidly would cause a decrease in $\Rnum$.
@@
 
-The use of tests cheaper than RT-PCR has been proposed as a potential strategy for containing the COVID-19 pandemic.
+The use of tests cheaper than RT-PCR has been proposed as a potential strategy for containing the \covid pandemic.
 While cheaper tests may be less sensitive and reliable than RT-PCR, they allow for broader and more intense testing.
@@
 (1) Model, assumptions: They developed a function, namely expected exposure $E(C,\tau)$, to approximate trade-offs between the frequency of testing, n, the sensitivity of testing, q, and the delay between
-testing and results, d. This function is explicitly derived and was connected the effective reproduction number $R=R_0 S$, where $S$ is the proportion of population susceptible.
+testing and results, d. This function is explicitly derived and was connected the effective reproduction number $R=\Rnum S$, where $S$ is the proportion of population susceptible.
 assumption that transmission rates are a step function: individuals who
-have COVID go from non-infectious to fully infectious instantaneously,
+have \covid go from non-infectious to fully infectious instantaneously,
 and remain fully infectious until they are no longer able to transmit disease. Test sensitivity takes the same form over the course of infection.
@@
 
-\citep{tuite2020mathematical} used an age-structured compartmental model of COVID-19 transmission in the population of Ontario, Canada.
+\citep{tuite2020mathematical} used an age-structured compartmental model of \covid transmission in the population of Ontario, Canada.
 We compared a base case with limited testing, isolation and quarantine to different scenarios.
@@
 \subsubsection{Other maybe-related works}
-\citep{arino2020simple} developed a SLIAR compartmental model to study the spread of an epidemic, specifically COVID-19, in a population.
+\citep{arino2020simple} developed a SLIAR compartmental model to study the spread of an epidemic, specifically COVID, in a population.
 The model incorporates an Erlang distribution of times of sojourn in incubating, symptomatically and asymptomatically infectious compartments.

### 2021-02-16 SIR_manuscript.tex
more David's comments were addresses, change of subscr to roman in \_

@@
 \newcommand{\covid}{COVID-19\xspace}
+
+\newcommand*\subtxt[1]{_{\textnormal{#1}}}
+\DeclareRobustCommand\_{\ifmmode\expandafter\subtxt\else\textunderscore\fi}
+
 \newcommand{\comment}{\showcomment}
@@
 Disease states include Susceptible, Infectious and Recovered (thus this is an SIR-type model), and testing status categorizes people as \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive} (Figure~\ref{fig:flowchart}).
-Two `accumulator' compartments, $N$ and $P$, were also incorporated in the model in order to collect cumulative reported negative or positive tests.
+Symbolically, the testing status of an individual in the disease compartment $X$, where $X \in \{S,I,R\}$ , is reflected in the subscript, namely $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
+Further, two `accumulator' compartments, $N$ and $P$, were also incorporated in the model in order to collect cumulative reported negative or positive tests.
 The model and details of calculation of the basic reproduction number $\Rnum$ are presented in the appendixs.
@@
 \includegraphics[scale=0.3]{./pix/sir_comp.png}
-\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model, \ref{model}. For the description of the parameters see \cref{tab:params}.
+\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model, \ref{model}. For the description of the parameters see \cref{tab:params}. \todo{change the subs to roman font}
 \label{fig:flowchart}}
@@
 Table~\ref{tab:params} defines the model parameters, which are generally straightforward
-\percap flows between compartments, or modifiers to these flow rates.
-\david{you need to define what you mean by ``testing intensity'' $\rho$}
+\percap flows between compartments, or modifiers to these flow rates. The \percap testing intensity is defined as the number of daily tests taken in a population of size $N_0$. 
 The novel component of the model comes in through the compartment-specific relative testing weights $W_S$, $W_I$ and $W_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively.
@@
 what is an ``applied model''?}, we constructed this model so we could specify the total \percap testing rate.
-We do this by defining the weighted size of the testing pool $W = W_S S_u + W_I I_u + W_R R_u$, and calculating a scaling parameter for testing as:
+We do this by defining the weighted size of the testing pool $W = W_S S\_u + W_I I\_u + W_R R\_u$, and calculating a scaling parameter for testing as:
 \begin{equation}
@@
 Thus, the \percap testing rate for compartment $X$ is $F_X=\sigma W_X$, where $X \in \{S,I,R\}$. 
-For a high-sensitivity test, infected people typically flow through to the ``confirmed positive'' ($I_c$, $R_c$) compartments and are thus unavailable for further testing.
+For a high-sensitivity test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus unavailable for further testing.
 Over the course of the epidemic, a fixed testing rate as specified in \eqref{sigma} can (if large enough) exhaust the pool of people available for testing, leading to a singularity when no one is left untested.
@@
 So $S_{\rm u}$ and $I_{\rm p}$ etc.
-Then it is clear the subscript isn't a variable with a value.}\ali{will do}
+Then it is clear the subscript isn't a variable with a value.}\ali{done, $p_S$ as is?}
 
@@
 \label{Lambda}
-\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_c I_c)}{N_0},
+\Lambda=\beta \frac{(I\_u+\eta\_w I\_n+\eta\_w I\_p+ \eta\_c I\_c)}{N_0},
 \end{equation}
-across all susceptible pools (see Table~\ref{tab:params} for details\david{there are no details in the table; you need to do a better job of explaining what's going on}), (ii) $\eta_c \leq \eta_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
+across all susceptible pools (see Table~\ref{tab:params} for details\david{there are no details in the table; you need to do a better job of explaining what's going on}\ali{help?}), (ii) $\eta\_c \leq \eta\_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
 For this analysis, we also assume a perfectly specific test ($p_S=0$).
-This last assumption combined with the assumption that no individual is in waiting-for-\emph{positive} or \emph{confirmed positive} compartments, i.e., $S_p(0)=S_c(0)=0$, reduces the model to 10 equations (equations c and d  in \eqref{model} are eliminated).
+This last assumption combined with the assumption that no individual is in waiting-for-\emph{positive} or \emph{confirmed positive} compartments, i.e., $S\_p(0)=S\_c(0)=0$, reduces the model to 10 equations (equations c and d  in \eqref{model} are eliminated).
 
@@
   $\rho$   & Per capita testing intensity & 1/day & 0.01 \\ \hline 
-  $\eta_w$  & Relative probability of transmission for ``waiting'' individuals & - & - \\ \hline
-  $\eta_c$  & Relative probability of transmission for ``confirmed positive'' individuals & - & -  \\ \hline
+  $\eta\_w$  & Relative probability of transmission for ``waiting'' individuals & - & - \\ \hline
+  $\eta\_c$  & Relative probability of transmission for ``confirmed positive'' individuals & - & -  \\ \hline
   $\beta$ & Transmission rate & 1/day & 0.39 \\ \hline
@@
 
-The Disease-Free Equilibrium (DFE) for the SIR model, Eqs. \eqref{model}, is given by solving the coupled system including $S_u+S_n=N_0$ and $F_S S_u-\omega S_n=0$.
-\david{you mean by solving the algebraic equations obtain by setting the DEs to 0.  not clear why you bother to mention two of these algebraic eqns.}
+The Disease-Free Equilibrium (DFE) for the SIR model, Eqs.
+\eqref{model}, is given by setting the differential equations to 0 and solving for the unknowns (\ali{is this clear or the "unkowns" needs to be defined?
+also * in sup for I and R?}.
 The DFE is
@@
 \label{dfe}
-S_n^*= \frac{\rho}{\omega} N_0, \ S_u^*= N_0-S_n^*, \text{~and~} I_j=R_j=0 \ \text{for all }j.
+S\_n^*= \frac{\rho}{\omega} N_0, \ S\_u^*= N_0-S\_n^*, \text{~and~} I_j=R_j=0 \ \text{for all }j.
 \end{equation}
-\david{above you had j in Roman font even though it *is* a variable, so I changed to $j$}
 
@@
 \label{R0}
-\Rnum= \frac{\beta}{\gamma} (A \times S_u^* + B \times S_n^*) \times C, 
+\Rnum= \frac{\beta}{\gamma} (A \times S\_u^* + B \times S\_n^*) \times C, 
 \end{equation}
@@
 \label{eq:abc}
-A=& \gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_c p_I) F_I, \\
-B=& \big(\omega+(F_I+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_c\omega) \omega p_I F_I }{\omega+\gamma}, \\ 
+A=& \gamma(\omega+\gamma) + (\gamma \eta\_w + \omega \eta\_c p_I) F_I, \\
+B=& \big(\omega+(F_I+\gamma)\eta\_w\big) \gamma+\frac{(\eta\_w \gamma+ \eta\_c\omega) \omega p_I F_I }{\omega+\gamma}, \\ 
 C=& \frac{1}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}.
@@
 \end{equation}
-\david{put $\beta/\gamma$ in front of the expression in eq \eqref{R0}, so it is clear that the rest is a modification of the the simplest expression.
-It would be worth spending some time trying to write the expression in a way that helps the reader understand where the different factors come from.}
 Further details of derivation of $\Rnum$ are provided in the appendix section.
 
-The analytical calculation of the next generation matrix, $G$\david{don't waste the symbol $G$ on a quantity we don't really need; $G$ is often used for generation interval}, and the derivation of $\Rnum$ was carried out in Maple \citep{maple14} by using simple linear Algebra package.
-  We used R \citep{r} for the simulation part which included using the explicit expression of $\Rnum$ \ref{R0} as a function of the underlying parameters, specifying the parameters a realistic range, calculating the corresponding $\Rnum$ and plotting the contours of $\Rnum$ by using ggplot package (ref?) in R.
+The analytical calculation of the next generation matrix and the derivation of $\Rnum$ was carried out in Maple \citep{maple14} by using simple linear Algebra package. 
+  We used R \citep{r} for the simulation part which included using the explicit expression of $\Rnum$ \ref{R0} as a function of the underlying parameters, specifying the parameters a realistic range, calculating the corresponding $\Rnum$ and plotting the contours of $\Rnum$ \eqref{R0}.
 The related result is presented in Figure \ref{pan}, where  panel \eqref{p.a} represents the random testing, and panel \eqref{p.b} is representing non-random testing.
-The parameter values or ranges are as follows (I like using a table here, with the param, their units and values).\david{use comment macro for comments} It is notable that in the simulation part, the explicit expression of $\Rnum$ was used and not the approximation \eqref{eq:R0appr}.\david{that's the first mention of ``the approximation''} This simulation reflects the behavior of $\Rnum$ with respect to the selected parameters for two different testing strategies: (i) random testing, represented by all testing weights to be the same, $W_S=W_I=W_R$, and (ii) non-random testing, when testing weight are not equal.
+The parameter values or ranges are presented in Table~\ref{tab:params}.
+This simulation reflects the behavior of $\Rnum$ with respect to the selected parameters for two different testing strategies: (i) random testing, represented by all testing weights to be the same, $W_S=W_I=W_R$, and (ii) non-random testing, when testing weight are not equal.
 For simulation purposes we chose $W_S=W_I=W_R=1$ for random testing, and $W_S=0.3$ and $W_I=W_R=1$ for non-random testing strategy.
@@
 On the other hand, there isn't a clear narrative here.
-The reader should feel that a story is being told.}
+The reader should feel that a story is being told.}\ali{partially addressed}
 
@@
 The explicit formula for the basic reproduction number, $\Rnum$ \eqref{R0}, provides an opportunity to study the influence of changes in the underlying parameters on the critical index of epidemic dynamics.
-We are interested in understanding the effect of parameters that can be realistically controlled by changing in isolation, \percap testing intensity and test resulting, i.e., $\eta_c$ and $\eta_w$, $\rho$ and $\omega$, respectively.
+We are interested in understanding the effect of parameters that can be realistically controlled by changing in isolation, \percap testing intensity and test resulting, i.e., $\eta\_c$ and $\eta\_w$, $\rho$ and $\omega$, respectively.
 The following Proposition is the direct result of taking partial derivative of $\Rnum$ \eqref{R0} with respect to selected parameters.
-Note that we used Taylor approximation of $\Rnum$ at $\rho=0$ due to the complexity of the expression and thus inconclusive.\david{in that case it would be good to establish numerically that the relationships given in the proposition are valid more generally, at least for reasonable parameter ranges}
+Note that we used Taylor approximation of $\Rnum$ at $\rho=0$ due to the complexity of the expression and thus inconclusive.\david{in that case it would be good to establish numerically that the relationships given in the proposition are valid more generally, at least for reasonable parameter ranges}\ali{help}
 See the appendix for details.
 
-\david{make sure to use $\Rnum$ not $\Rnum$ below (and often in the supp material section).
-More importantly, you really need to state in words what this proposition is saying.
-The reader has to check the param table and think it through themselves.}
+\david{More importantly, you really need to state in words what this proposition is saying.  The reader has to check the param table and think it through themselves.}
 \begin{proposition}
@@
 \item \label{p1:eta}
-$\partial{\Rnum}/\partial{\eta_c} \geq 0$ and $\partial{\Rnum}/\partial{\eta_w} \geq 0$. 
+$\partial{\Rnum}/\partial{\eta\_c} \geq 0$ and $\partial{\Rnum}/\partial{\eta\_w} \geq 0$. 
 \item \label{p1:rho}
@@
 
-Given that the perfect isolation occurs at $\eta_c=\eta_w = 0$,
-\david{That's a perfect example of the kind of phrasing that really must be changed.
-You need to say something like ``occurs when no one transmits while waiting for test results (i.e., $\eta_c=\eta_w = 0$)''.
-It makes the text so much easier to process and doesn't require the reader to keep looking at the table to make sense of what you're saying.}
-Prop.~\ref{p1:eta} means that lifting the isolation from awaiting group results in an increased $\Rnum$ and consequently a greater number of infected individuals.
+Given that the perfect isolation occurs when no one transmits while waiting for test results (i.e., $\eta\_c=\eta\_w = 0$), Prop.~\ref{p1:eta} means that lifting the isolation from awaiting group results in an increased $\Rnum$ and consequently a greater number of infected individuals.
 Prop.~\ref{p1:rho} indicates that increasing the \percap testing intensity, $\rho$, reduces $\Rnum$.
@@
 It is straightforward to derive \ref{p1:rho} and \ref{p1:omega} from the Taylor approximation of $\Rnum$ at $\rho=0$.
-\david{can we gain some insight from the expressions for the partial derivatives in the proposition, as opposed to just their signs?}
+\david{can we gain some insight from the expressions for the partial derivatives in the proposition, as opposed to just their signs?}\ali{help}
 
@@
 We numerically evaluate $\Rnum$ \eqref{R0} using the following parameters (listed in Table \ref{tab:params}):
-$N_0=1 \times 10^6$, $\omega \in [0.1,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.01]$ 1/day, $\eta_w \in [0,1]$, $\eta_c \in [0,1]$, $\beta=0.39$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
+$N_0=1 \times 10^6$, $\omega \in [0.1,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.01]$ 1/day, $\eta\_w \in [0,1]$, $\eta\_c \in [0,1]$, $\beta=0.39$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
 Contours of $\Rnum$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $W_S=W_I=W_R=1$ and (b) targeted testing where $W_S=0.3$ and $W_I=W_R=1$.
@@
 It can be inferred that when random testing strategy is applied, Figure \ref{p.a}, comparing to the targeted testing strategy, Figure \ref{p.b}, the critical \percap testing intensity of the corresponding panels are lower in non-random testing.
-Also, it appears that speeding the test reporting, i.e., increasing $\omega$, does not significantly lower $\Rnum$ when awaiting people (including people in $I_p$ and $I_n$ compartments) follow the isolation perfectly, i.e., $\eta_w$ is closer to 0.
-However, speeding the test reporting reduces the epidemic more when the awaiting people less follow the isolation, i.e., $\eta_w$ is closer to 1.
+Also, it appears that speeding the test reporting, i.e., increasing $\omega$, does not significantly lower $\Rnum$ when awaiting people (including people in $I\_p$ and $I\_n$ compartments) follow the isolation perfectly, i.e., $\eta\_w$ is closer to 0.
+However, speeding the test reporting reduces the epidemic more when the awaiting people less follow the isolation, i.e., $\eta\_w$ is closer to 1.
 
@@
 It states that returning test results more rapidly, i.e., increasing the rate $\omega$, does not necessarily lower the reproduction number $\Rnum$; whether increasing $\omega$ lowers $\Rnum$ depends on the precise combination of model parameters  including test reporting rate, testing strategies represented by compartment-specific testing weights, test sensitivity and specificity, and the level of isolation.
-Specifically, in the case of perfect isolation, i.e., when $\eta_w=0$ and $\eta_c=0$, $\Rnum$ may increase as the test reporting process becomes faster. This can be seen from expression \eqref{Rom}.
+Specifically, in the case of perfect isolation, i.e., when $\eta\_w=0$ and $\eta\_c=0$, $\Rnum$ may increase as the test reporting process becomes faster. This can be seen from expression \eqref{Rom}.
 Another example of this favorable delay in $\omega$ could be when the test being employed produces many false negatives.
@@
 To give a biological interpretation, we describe the qualitative trends predicted by the inequality \eqref{eq:necsuf}.
-To summarise, returning test results more rapidly tends to be favourable (i.e., reduces $\Rnum$) when (i) \emph{Confirmed-positive} individuals, lower their contact much more than individuals who are waiting for their test results (i.e., $\eta_c \ll \eta_w$), (ii) the test is highly sensitive (i.e., $p_I$ is close to 1), and (iii) the targeted testing strategy is used (i.e., $W_i \gg W_s$).
+To summarise, returning test results more rapidly tends to be favourable (i.e., reduces $\Rnum$) when (i) \emph{Confirmed-positive} individuals, lower their contact much more than individuals who are waiting for their test results (i.e., $\eta\_c \ll \eta\_w$), (ii) the test is highly sensitive (i.e., $p_I$ is close to 1), and (iii) the targeted testing strategy is used (i.e., $W_i \gg W_s$).
 
@@
 \begin{align}
- d S_u/dt &= -\Lambda S_u - F_S S_u + \omega S_n, \\
- d S_n/dt &= -\Lambda S_n + (1-p_S) F_S S_u - \omega S_n, \\
- d S_p/dt &= -\Lambda S_p + p_S F_S S_u - \omega S_p, \\
- d S_c/dt &= -\Lambda S_c + \omega S_p, \\
- d I_u/dt &= \Lambda S_u - F_I I_u + \omega I_n  - \gamma I_u, \\
- d I_n/dt &= \Lambda S_n + (1-p_I) F_I I_u - \omega I_n -\gamma I_n, \\
- d I_p/dt &= \Lambda S_p + p_I F_I I_u - \omega I_p -\gamma I_p, \\
- d I_c/dt &= \Lambda S_c + \omega I_p - \gamma I_c, \\
- d R_u/dt &= \gamma I_u - F_R R_u + \omega R_n, \\
- d R_n/dt &= \gamma I_n + (1-p_R) F_R R_u - \omega R_n, \\
- d R_p/dt &= \gamma I_p + p_R F_R R_u  - \omega R_p, \\
- d R_c/dt&= \gamma I_c + \omega R_p, \\
- dN/dt &= \omega (S_n + I_n + R_n),  \\
- dP/dt &= \omega(I_p + R_p) ,
+ d S\_u/dt &= -\Lambda S\_u - F_S S\_u + \omega S\_n, \\
+ d S\_n/dt &= -\Lambda S\_n + (1-p_S) F_S S\_u - \omega S\_n, \\
+ d S\_p/dt &= -\Lambda S\_p + p_S F_S S\_u - \omega S\_p, \\
+ d S\_c/dt &= -\Lambda S\_c + \omega S\_p, \\
+ d I\_u/dt &= \Lambda S\_u - F_I I\_u + \omega I\_n  - \gamma I\_u, \\
+ d I\_n/dt &= \Lambda S\_n + (1-p_I) F_I I\_u - \omega I\_n -\gamma I\_n, \\
+ d I\_p/dt &= \Lambda S\_p + p_I F_I I\_u - \omega I\_p -\gamma I\_p, \\
+ d I\_c/dt &= \Lambda S\_c + \omega I\_p - \gamma I\_c, \\
+ d R\_u/dt &= \gamma I\_u - F_R R\_u + \omega R\_n, \\
+ d R\_n/dt &= \gamma I\_n + (1-p_R) F_R R\_u - \omega R\_n, \\
+ d R\_p/dt &= \gamma I\_p + p_R F_R R\_u  - \omega R\_p, \\
+ d R\_c/dt&= \gamma I\_c + \omega R\_p, \\
+ dN/dt &= \omega (S\_n + I\_n + R\_n),  \\
+ dP/dt &= \omega(I\_p + R\_p) ,
 \end{align}
@@
 The next generation matrix for this model is $G = F V^{-1}$, where matrix $F$ represents the inflow of new infection to the infected compartments and matrix $V$ represents the flow in the infected compartments when the population is totally susceptible.
-Matrices $F$ and $V$ are \ali{all $S_u \& S_n$ should be $S_u^* \& S_n^*$}
+Matrices $F$ and $V$ are \ali{all $S\_u \& S\_n$ should be $S\_u^* \& S\_n^*$}
 \begin{align}
@@
 F =& \beta/N_0 \left[ \begin {array}{cccc} 
-S_u&\eta_w\,S_u&\eta_w\,S_u&\eta_c\,S_u\\
-S_n&\eta_w\,S_n&\eta_w\,S_n&\eta_c\,S_n\\ 
+S\_u&\eta\_w\,S\_u&\eta\_w\,S\_u&\eta\_c\,S\_u\\
+S\_n&\eta\_w\,S\_n&\eta\_w\,S\_n&\eta\_c\,S\_n\\ 
 0&0&0&0\\
@@
 \left[\begin {array}{cc}
-A\,S_u & B\,S_u\\
-A\,S_n & B\,S_n
+A\,S\_u & B\,S\_u\\
+A\,S\_n & B\,S\_n
 \end {array}\right],
@@
 The block matrix $G_{12}$ does not influence $\Rnum$ (defined as the spectral radius of $G$). All that matters here are the eigenvalues of $G_{11}$, which are 0 and $\Rnum$ \eqref{R0}.
-Also note that, all $S_u$'s and $S_n$'s are evaluated at the DFE \eqref{dfe} in matrix $F$ \eqref{FV} and the block matrix $G_{11}$ \eqref{mat:G}.
+Also note that, all $S\_u$'s and $S\_n$'s are evaluated at the DFE \eqref{dfe} in matrix $F$ \eqref{FV} and the block matrix $G_{11}$ \eqref{mat:G}.
 
@@
 Specifically, the numerical singularity issue with the chosen $\sigma$ \eqref{sigma} is that the population in $S$ compartments appeared to blow up when the DFE is achieved.
-This is once the only untested people are susceptibles, the FOI will become $\Lambda=0$, testing rate $F_s=\rho N_0/S_u$.
+This is once the only untested people are susceptibles, the FOI will become $\Lambda=0$, testing rate $F_s=\rho N_0/S\_u$.
 Thus, the first equation of the model \eqref{model} will become
-$d S_u/dt = - \rho N_0 + \omega S_n$. Thus changes in $S_u$ will be no longer dependent on $S_u$ with a linear rate of leaving the $S_u$ compartment.
+$d S\_u/dt = - \rho N_0 + \omega S\_n$. Thus changes in $S\_u$ will be no longer dependent on $S\_u$ with a linear rate of leaving the $S\_u$ compartment.
 IN fact the testing rate, $\sigma$, should be formulated such that people from the untested compartments will not be tested if they are not there.
@@
 \label{eq:R0appr}
-\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_S} \Big(\gamma(\eta_w-1)(\gamma W_S+\omega W_I) + (\eta_c -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2).
+\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_S} \Big(\gamma(\eta\_w-1)(\gamma W_S+\omega W_I) + (\eta\_c -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2).
 \end{equation}
@@
 \end{equation}
-where $a=(\eta_w-1)W_I-(\eta_c-1)P_I W_I = ((s-p_I)\eta_c + (p_I-1)) W_I$, $b=2(\eta_w-1)\gamma W_S$ and $c=(\eta_w-1)\gamma^2 W_S$.
-Given that $0 \leq \eta_c\leq \eta_w \leq 1 $, one can easily derive $b\leq 0$ and $c \leq 0$.
+where $a=(\eta\_w-1)W_I-(\eta\_c-1)P_I W_I = ((s-p_I)\eta\_c + (p_I-1)) W_I$, $b=2(\eta\_w-1)\gamma W_S$ and $c=(\eta\_w-1)\gamma^2 W_S$.
+Given that $0 \leq \eta\_c\leq \eta\_w \leq 1 $, one can easily derive $b\leq 0$ and $c \leq 0$.
 
-Note that in general, the necessary and sufficient condition for $a \geq 0$ is $(s-p_I) \eta_c \geq (1-p_I)$, where $s=\frac{\eta_w}{\eta_c} \geq 1$.
+Note that in general, the necessary and sufficient condition for $a \geq 0$ is $(s-p_I) \eta\_c \geq (1-p_I)$, where $s=\frac{\eta\_w}{\eta\_c} \geq 1$.
 
-In the case of ``perfect isolation'', i.e., when $\eta_w=0$ and consequently $\eta_c=0$, it is straightforward to see that $a \leq 0$, $b<0$ and $c<0$.
+In the case of ``perfect isolation'', i.e., when $\eta\_w=0$ and consequently $\eta\_c=0$, it is straightforward to see that $a \leq 0$, $b<0$ and $c<0$.
 Thus, $\partial{\Rnum}/\partial{\omega} \geq 0$.
@@
 \begin{equation}\label{linearization}
-\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta_w-1)(\gamma W_s+\omega W_i) + (\eta_c -1)P_iW_i \omega^2 \Big). 
+\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta\_w-1)(\gamma W_s+\omega W_i) + (\eta\_c -1)P_iW_i \omega^2 \Big). 
 \end{equation}
 So when $\rho \approx 0$ we have $$\partial{\Rnum}/\partial{\omega} \approx  \frac{-\beta \rho}{\gamma W_s\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c),$$
-where $a=(\eta_w-1)W_i-(\eta_c-1)P_iW_i$, $b=2(\eta_w-1)\gamma W_s$ and $c=(\eta_w-1)\gamma^2 W_s$.
+where $a=(\eta\_w-1)W_i-(\eta\_c-1)P_iW_i$, $b=2(\eta\_w-1)\gamma W_s$ and $c=(\eta\_w-1)\gamma^2 W_s$.
 
@@
 Note that the zeroes are real if and only if $Q < 1$.
-Note that have $\eta_c < \eta_w$, so if $P_i \approx 1$, we will have $Q < 0 < 1$.
+Note that have $\eta\_c < \eta\_w$, so if $P_i \approx 1$, we will have $Q < 0 < 1$.
 Thus, if we assume near-perfect test sensitivity, $\omega_1$ and $\omega_2$ will be real.
@@
 Suppose that $\rho_1 > \rho_2$.
-Recall that the linearization of $\Rnum$ around $\rho \approx 0$ is given by $$\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta_w-1)(\gamma W_s+\omega W_i) + (\eta_c -1)P_iW_i \omega^2 \Big).$$
+Recall that the linearization of $\Rnum$ around $\rho \approx 0$ is given by $$\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta\_w-1)(\gamma W_s+\omega W_i) + (\eta\_c -1)P_iW_i \omega^2 \Big).$$
 
@@
 \begin{align}\label{eq:rho1vsrho2}
-    &\rho_1\left(\gamma(\eta_w-1)(\gamma W_s + \omega W_i) + (\eta_c-1)P_{i, 1}W_i\omega^2\right) - \rho_2\left(\gamma(\eta_w-1)(\gamma W_s + \omega W_i) + (\eta_c-1)P_{i, 2}W_i\omega^2\right) > 0 \nonumber \\
+    &\rho_1\left(\gamma(\eta\_w-1)(\gamma W_s + \omega W_i) + (\eta\_c-1)P_{i, 1}W_i\omega^2\right) - \rho_2\left(\gamma(\eta\_w-1)(\gamma W_s + \omega W_i) + (\eta\_c-1)P_{i, 2}W_i\omega^2\right) > 0 \nonumber \\
     &\vdots \nonumber \\
-    &\frac{\rho_2P_{i, 2}-\rho_1P_{i, 1} }{\rho_1-\rho_2} > \frac{1-\eta_w}{1-\eta_c}\cdot \frac{\gamma(\gamma W_s + \omega W_i)}{\omega^2 W_i}
+    &\frac{\rho_2P_{i, 2}-\rho_1P_{i, 1} }{\rho_1-\rho_2} > \frac{1-\eta\_w}{1-\eta\_c}\cdot \frac{\gamma(\gamma W_s + \omega W_i)}{\omega^2 W_i}
 \end{align}
@@
 \citep{de2020influenza} Was discussed in the Math 747 
-SEIR Asymptomatic and symptomatic $I_1, I_2$. Used linear chain trick 
+SEIR Asymptomatic and symptomatic $I\_1, I\_2$. Used linear chain trick 
 Stringency index as a control force lowering $\beta$.

### 2021-02-22 codes/sir_macpan.Rnw
macpan minor update

@@
 )
+
+##this works remotes::install_github("bbolker/McMasterPandemic")
+
 library(McMasterPandemic)

### 2021-02-23 SIR_manuscript.tex
introduction was updated

@@
 This level of focus depends in turn on the purpose and design of testing programs. 
-(1) When testing is done for the purposes of disease surveillance \citep{foddai2020base}
-we might expect tests to be assigned randomly across the population (in an ideal world: some bias in selection for testing is probably inevitable), although a stratified sampling design would be preferred for statistical efficiency \citep{graubard1996modelling} \ali{a better ref?}.
+When testing is done for the purposes of disease surveillance \citep{foddai2020base}
+we might expect tests to be assigned randomly across the population, thus we characterize it as \emph{random} testing strategy.
+In an ideal world, some bias in selection for testing is probably inevitable, although a stratified sampling design would be preferred for statistical efficiency \citep{graubard1996modelling} \ali{a better ref everyone?}.
 
@@
 
-In order to understand the effect of testing processes on an epidemic dynamics, we deveopled a mechanistic model with epidemic and testing components.
-Here, we focus on the the effect of testing intensity, different levels of testing ``focus'' (from random to highly targeted), and rate of test return on a single index of epidemiological dynamics, the basic reproduction number $\Rnum$, which is useful to estimate the early dynamics of an epidemic.
-This index is defined as the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}.
-In the early stages of an epidemic the number of infected individuals is expected grow exponentially over time when $\Rnum>1$. 
-To control the spread of disease in the population, \david{the effective} $\Rnum$ must be $<1$.
+The main interest from the epidemiological point of view is to know whether the number of infected individuals goes through an exponential growth phase, following the introduction of an infection in a totally susciptable population, before the disease becomes extinct.
+This is determined by studying the basic reproduction number $\Rnum$.
+It is defined as the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}.
+In the early stages of an epidemic the number of infected individuals is expected to grow exponentially over time when $\Rnum>1$, and to decline over time when $\Rnum<1$. 
 Although the value of $\Rnum$ cannot completely characterize the dynamics of even the simplest epidemic model
 \citep{shaw2021what}, it does give a simple and widely accepted index for the difficulty of control, as well as some indication of the likely final size of an epidemic \citep{ma2006generality}.
-Our model provides insights to the sensitivity of the epidemiological dynamics, through $\Rnum$, with respect to the undelying testing processes.
-\david{This paragraph does not flow from the rest of the intro, and is a weird way to end when the point of the paper is about testing.}\ali{edited, but still needs work}
+
+In order to understand the effect of testing processes on an epidemic dynamics, we deveopled a mechanistic SIR-type model with epidemic and testing components.
+Here, we focus on the the effect of testing intensity, different levels of testing ``focus'' (from random to highly targeted), and rate of test return on the epidemic dynamics.
+Our model provides insights to the sensitivity of the epidemiological dynamics, through $\Rnum$, with respect to the undelying epidemic and testing parameters.
+\david{This paragraph does not flow from the rest of the intro, and is a weird way to end when the point of the paper is about testing.}\ali{major changes done}
 

### 2021-02-23 SIR_manuscript.tex
updated methods

@@
 Our model provides insights to the sensitivity of the epidemiological dynamics, through $\Rnum$, with respect to the undelying epidemic and testing parameters.
-\david{This paragraph does not flow from the rest of the intro, and is a weird way to end when the point of the paper is about testing.}\ali{major changes done}
+\david{This paragraph does not flow from the rest of the intro, and is a weird way to end when the point of the paper is about testing.}\ali{some edits/rewording are done, ok?}
 
@@
 Disease states include Susceptible, Infectious and Recovered (thus this is an SIR-type model), and testing status categorizes people as \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive} (Figure~\ref{fig:flowchart}).
-Symbolically, the testing status of an individual in the disease compartment $X$, where $X \in \{S,I,R\}$ , is reflected in the subscript, namely $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
+Symbolically, the testing status of an individual in the disease compartment $X$, where $X \in \{S,I,R\}$, is reflected in the subscript, namely $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
 Further, two `accumulator' compartments, $N$ and $P$, were also incorporated in the model in order to collect cumulative reported negative or positive tests.
@@
 \includegraphics[scale=0.3]{./pix/sir_comp.png}
-\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model, \ref{model}. For the description of the parameters see \cref{tab:params}. \todo{change the subs to roman font}
+\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model, \ref{model}.
+Here, the disease-based status of a compartment $X$, where $X \in \{S,I,R\}$, is combined with the testing-based status including $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
+Also,  $\Lambda$ is the force of infection with definition in Eq.~\eqref{Lambda}, $\gamma$ is the recovery rate, $\omega$ is the rate of test return, $F_X$ and $p_X$ represent the \percap testing rate and the probability of positive tests, respectively, for compartment $X$.
+For further description of the parameters see \cref{tab:params}.
+\ali{(i) will change the subs to roman font, (ii) not sure if ref color is helpful in u compartment?}
 \label{fig:flowchart}}
@@
 
-In order to link to more applied models\david{what does that mean?
-what is an ``applied model''?}, we constructed this model so we could specify the total \percap testing rate.
+In order to link to more applied models, i.e., including both epidemiological and testing processes, we constructed this model so we could specify the total \percap testing rate.
 We do this by defining the weighted size of the testing pool $W = W_S S\_u + W_I I\_u + W_R R\_u$, and calculating a scaling parameter for testing as:
@@
 
-\david{A comment on notation:  You haven't actually defined $S_u$, $I_u$ etc.
-In any case, I urge you to use macros and use Roman font for subscripts that are textual as opposed to variables.
-So $S_{\rm u}$ and $I_{\rm p}$ etc.
-Then it is clear the subscript isn't a variable with a value.}\ali{done, $p_S$ as is?}
-
-The classical SIR model is based on the following implicit assumptions; well-mixed population, homogeneity of the population\david{What's the difference btw ``well-mixed'' and ``homoegeneous''?}\ali{ non structured eg age, ok?}, exponentially distributed duration of infection and large population size (see, e.g., \cite{keeling2011modeling}).
-In addition to these standard assumptions, our model, \ref{model}, assumes: (i) there is a single force of infection, $\Lambda$, defined as follows
+The classical SIR model is based on the following implicit assumptions; well-mixed population, homogeneity of the population (i.e., all individuals are equaly susciptable and equaly infectious for the same length of time when infected), exponentially distributed duration of infection and large population size (see, e.g., \cite{keeling2011modeling}).
+In addition to these standard assumptions, our model, \ref{model}, assumes: (i) there is a single force of infection (new cases per unit time), $\Lambda$, defined as follows
 \begin{equation}
@@
 \end{equation}
-across all susceptible pools (see Table~\ref{tab:params} for details\david{there are no details in the table; you need to do a better job of explaining what's going on}\ali{help?}), (ii) $\eta\_c \leq \eta\_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
+across all susceptible pools with transmission rate $\beta$ and relative probability of transmission for ``waiting'' and \emph{confirmed positive} individuals, $\eta\_w$ and $\eta\_c$ respectively, (see Table~\ref{tab:params} for further details), (ii) $\eta\_c \leq \eta\_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
 For this analysis, we also assume a perfectly specific test ($p_S=0$).
@@
 On the other hand, there isn't a clear narrative here.
-The reader should feel that a story is being told.}\ali{partially addressed}
+The reader should feel that a story is being told.}\ali{partially addressed?, feedback is welcomed}
 

### 2021-02-23 SIR_manuscript.tex
a bit more on results

@@
 On the other hand, there isn't a clear narrative here.
-The reader should feel that a story is being told.}\ali{partially addressed?, feedback is welcomed}
+The reader should feel that a story is being told.}\ali{addressed?, feedback is welcomed}
 
@@
 The following Proposition is the direct result of taking partial derivative of $\Rnum$ \eqref{R0} with respect to selected parameters.
-Note that we used Taylor approximation of $\Rnum$ at $\rho=0$ due to the complexity of the expression and thus inconclusive.\david{in that case it would be good to establish numerically that the relationships given in the proposition are valid more generally, at least for reasonable parameter ranges}\ali{help}
+Note that we used Taylor approximation of $\Rnum$ at $\rho=0$ due to the complexity of the expression and thus inconclusive.\david{in that case it would be good to establish numerically that the relationships given in the proposition are valid more generally, at least for reasonable parameter ranges}\ali{will do the simulation for testing intensity $>>1$}
 See the appendix for details.
@@
 \label{prop1}
-Using the expression of $\Rnum$ \eqref{R0},
+For model \eqref{model}, and by using the expression of the basic reproduction number $\Rnum$ \eqref{R0} ,
 \begin{enumerate}
-\item \label{p1:eta}
-$\partial{\Rnum}/\partial{\eta\_c} \geq 0$ and $\partial{\Rnum}/\partial{\eta\_w} \geq 0$. 
-\item \label{p1:rho}
-$\partial{\Rnum}/\partial{\rho} \leq 0$ when $\rho \approx 0$.
-\item \label{p1:omega}
-$\partial{\Rnum}/\partial{\omega}$ can be positive or negative when $\rho \approx 0$.
+\item \label{p1:eta} increasing the relative probabilities of transmission for tested individuals, increases $\Rnum$;
+$$\partial{\Rnum}/\partial{\eta\_c} \geq 0 \text{~and~} \partial{\Rnum}/\partial{\eta\_w} \geq 0.$$ 
+\item \label{p1:rho} higher rate of testing reduces $\Rnum$;
+$$\partial{\Rnum}/\partial{\rho} \leq 0 \text{~when~} \rho \approx 0.$$
+\item \label{p1:omega} The rate of test return may result in the disease becomes extinct or not;
+$$\partial{\Rnum}/\partial{\omega} \text{~can be positive or negative when~} \rho \approx 0.$$
 \end{enumerate}

### 2021-02-28 SIR_manuscript.tex
all plots to pdf, flowchart modified

@@
 \begin{center} 
-\includegraphics[scale=0.3]{./pix/sir_comp.png}
+\includegraphics[scale=2]{./pix/sir_comp.pdf}
 \caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model, \ref{model}.
@@
 For further description of the parameters see \cref{tab:params}.
-\ali{(i) will change the subs to roman font, (ii) not sure if ref color is helpful in u compartment?}
 \label{fig:flowchart}}
@@
 \percap flows between compartments, or modifiers to these flow rates. The \percap testing intensity is defined as the number of daily tests taken in a population of size $N_0$. 
-The novel component of the model comes in through the compartment-specific relative testing weights $W_S$, $W_I$ and $W_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively.
-For example, $W_I/W_S=2$ means that infected individuals are tested at twice the \percap rate of susceptible individuals.
+The novel component of the model comes in through the compartment-specific relative testing weights $w_S$, $w_I$ and $w_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively.
+For example, $w_I/w_S=2$ means that infected individuals are tested at twice the \percap rate of susceptible individuals.
 
 In order to link to more applied models, i.e., including both epidemiological and testing processes, we constructed this model so we could specify the total \percap testing rate.
-We do this by defining the weighted size of the testing pool $W = W_S S\_u + W_I I\_u + W_R R\_u$, and calculating a scaling parameter for testing as:
+We do this by defining the weighted size of the testing pool $W = w_S S\_u + w_I I\_u + w_R R\_u$, and calculating a scaling parameter for testing as:
 \begin{equation}
@@
   $p_R$ & Probability of positive tests for $R$ ($= 1-\textrm{specificity}$) & - & 0.5 \\ \hline
-  $W_S, W_I, W_R$ & Relative testing weights & - &
+  $w_S, w_I, w_R$ & Relative testing weights & - &
   \begin{minipage}[t]{0.21\columnwidth}%
@@
 The parameter values or ranges are presented in Table~\ref{tab:params}.
-This simulation reflects the behavior of $\Rnum$ with respect to the selected parameters for two different testing strategies: (i) random testing, represented by all testing weights to be the same, $W_S=W_I=W_R$, and (ii) non-random testing, when testing weight are not equal.
-For simulation purposes we chose $W_S=W_I=W_R=1$ for random testing, and $W_S=0.3$ and $W_I=W_R=1$ for non-random testing strategy.
+This simulation reflects the behavior of $\Rnum$ with respect to the selected parameters for two different testing strategies: (i) random testing, represented by all testing weights to be the same, $w_S=w_I=w_R$, and (ii) non-random testing, when testing weight are not equal.
+For simulation purposes we chose $w_S=w_I=w_R=1$ for random testing, and $w_S=0.3$ and $w_I=w_R=1$ for non-random testing strategy.
 Note that the critical contour of $\Rnum=1$ is plotted in solid line in Figure \ref{pan}.
@@
 $N_0=1 \times 10^6$, $\omega \in [0.1,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.01]$ 1/day, $\eta\_w \in [0,1]$, $\eta\_c \in [0,1]$, $\beta=0.39$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
-Contours of $\Rnum$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $W_S=W_I=W_R=1$ and (b) targeted testing where $W_S=0.3$ and $W_I=W_R=1$.
+Contours of $\Rnum$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
 The black solid line in each panel represents the critical contour of $\Rnum=1$.
-\bmb{we still have some collisions in the y-axis tick labels}}
+\bmb{we still have some collisions in the y-axis tick labels}
+% \ali{plotr + expand_limits(y = 0), change 1/\omega on x-axis}
+}
 \label{pan}
@@
 
-We incorporated the compartment-specific relative testing weights, $W_S$, $W_I$ and $W_R$, to model random testing and \emph{targeted} testing strategies.
-Here, in the case of targeted testing and for the simplicity and illustration purposes, we assumed that infected and recovered individuals are tested at three times the \percap rate of susceptible individuals, thus $W_I/W_S=3$ and $W_R=W_I$.
+We incorporated the compartment-specific relative testing weights, $w_S$, $w_I$ and $w_R$, to model random testing and \emph{targeted} testing strategies.
+Here, in the case of targeted testing and for the simplicity and illustration purposes, we assumed that infected and recovered individuals are tested at three times the \percap rate of susceptible individuals, thus $w_I/w_S=3$ and $w_R=w_I$.
 Note that we have not specified a methodology to assign particular relative testing weights corresponding to a particular targeted testing scenario.
@@
 To give a biological interpretation, we describe the qualitative trends predicted by the inequality \eqref{eq:necsuf}.
-To summarise, returning test results more rapidly tends to be favourable (i.e., reduces $\Rnum$) when (i) \emph{Confirmed-positive} individuals, lower their contact much more than individuals who are waiting for their test results (i.e., $\eta\_c \ll \eta\_w$), (ii) the test is highly sensitive (i.e., $p_I$ is close to 1), and (iii) the targeted testing strategy is used (i.e., $W_i \gg W_s$).
+To summarise, returning test results more rapidly tends to be favourable (i.e., reduces $\Rnum$) when (i) \emph{Confirmed-positive} individuals, lower their contact much more than individuals who are waiting for their test results (i.e., $\eta\_c \ll \eta\_w$), (ii) the test is highly sensitive (i.e., $p_I$ is close to 1), and (iii) the targeted testing strategy is used (i.e., $w_I \gg w_S$).
 
@@
 \label{eq:R0appr}
-\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_S} \Big(\gamma(\eta\_w-1)(\gamma W_S+\omega W_I) + (\eta\_c -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2).
+\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(\eta\_w-1)(\gamma w_S+\omega w_I) + (\eta\_c -1)p_I w_I \omega^2 \Big) + \mathcal{O}(\rho^2).
 \end{equation}
@@
 \label{Rom}
-\partial{\Rnum}/\partial{\omega}=  \frac{-\beta \rho}{\gamma W_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c),
+\partial{\Rnum}/\partial{\omega}=  \frac{-\beta \rho}{\gamma w_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c),
 \end{equation}
-where $a=(\eta\_w-1)W_I-(\eta\_c-1)P_I W_I = ((s-p_I)\eta\_c + (p_I-1)) W_I$, $b=2(\eta\_w-1)\gamma W_S$ and $c=(\eta\_w-1)\gamma^2 W_S$.
+where $a=(\eta\_w-1)w_I-(\eta\_c-1)p_I w_I = ((s-p_I)\eta\_c + (p_I-1)) w_I$, $b=2(\eta\_w-1)\gamma w_S$ and $c=(\eta\_w-1)\gamma^2 w_S$.
 Given that $0 \leq \eta\_c\leq \eta\_w \leq 1 $, one can easily derive $b\leq 0$ and $c \leq 0$.
@@
 \begin{equation}\label{linearization}
-\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta\_w-1)(\gamma W_s+\omega W_i) + (\eta\_c -1)P_iW_i \omega^2 \Big). 
+\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(\eta\_w-1)(\gamma w_S+\omega w_I) + (\eta\_c -1)p_I w_I \omega^2 \Big). 
 \end{equation}
-So when $\rho \approx 0$ we have $$\partial{\Rnum}/\partial{\omega} \approx  \frac{-\beta \rho}{\gamma W_s\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c),$$
-where $a=(\eta\_w-1)W_i-(\eta\_c-1)P_iW_i$, $b=2(\eta\_w-1)\gamma W_s$ and $c=(\eta\_w-1)\gamma^2 W_s$.
+So when $\rho \approx 0$ we have $$\partial{\Rnum}/\partial{\omega} \approx  \frac{-\beta \rho}{\gamma w_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c),$$
+where $a=(\eta\_w-1)w_I-(\eta\_c-1)p_I w_I$, $b=2(\eta\_w-1)\gamma w_S$ and $c=(\eta\_w-1)\gamma^2 w_S$.
 
@@
 \begin{align}\label{eq:defQ}
-    Q =& \frac{W_i}{W_s}\left(1-\frac{n_{t}-1}{n_{w}-1}P_{i}\right).
+    Q =& \frac{w_I}{w_S}\left(1-\frac{n_t-1}{n_w-1}p_I \right).
 \end{align}
@@
 \begin{align}
-    &1-\frac{W_i}{W_s}\left(1-\frac{n_{t}-1}{n_{\omega}-1}P_{i}\right)> \left(\frac{\gamma}{\omega}+1\right)^2 \\
-    &-\frac{W_i}{W_s}\left(1-\frac{n_{t}-1}{n_{\omega}-1}P_{i}\right)> \left(\frac{\gamma}{\omega}+1\right)^2+1 \\    
-    &\frac{W_i}{W_s}\left(\frac{1-n_{t}}{1-n_{\omega}}P_{i}-1\right)> \left(\frac{\gamma}{\omega}+1\right)^2+1\label{eq:necsuf}.
+    &1-\frac{w_I}{w_S}\left(1-\frac{n_{t}-1}{n_{\omega}-1}P_{i}\right)> \left(\frac{\gamma}{\omega}+1\right)^2 \\
+    &-\frac{w_I}{w_S}\left(1-\frac{n_{t}-1}{n_{\omega}-1}P_{i}\right)> \left(\frac{\gamma}{\omega}+1\right)^2+1 \\    
+    &\frac{w_I}{w_S}\left(\frac{1-n_{t}}{1-n_{\omega}}P_{i}-1\right)> \left(\frac{\gamma}{\omega}+1\right)^2+1\label{eq:necsuf}.
 \end{align}
@@
 Suppose that $\rho_1 > \rho_2$.
-Recall that the linearization of $\Rnum$ around $\rho \approx 0$ is given by $$\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta\_w-1)(\gamma W_s+\omega W_i) + (\eta\_c -1)P_iW_i \omega^2 \Big).$$
+Recall that the linearization of $\Rnum$ around $\rho \approx 0$ is given by $$\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(\eta\_w-1)(\gamma w_S+\omega w_I) + (\eta\_c -1)p_I w_I \omega^2 \Big).$$
 
-Treating $\Rnum$ as a function of $\rho$ and $P_i$,we can reduce the inequality $$\Rnum(\rho_2, P_{i,2}) < \Rnum(\rho_1, P_{i,1})$$ into
+Treating $\Rnum$ as a function of $\rho$ and $P_i$,we can reduce the inequality $$\Rnum(\rho_2, p_{I,2}) < \Rnum(\rho_1, p_{I,1})$$ into
 
 \begin{align}\label{eq:rho1vsrho2}
-    &\rho_1\left(\gamma(\eta\_w-1)(\gamma W_s + \omega W_i) + (\eta\_c-1)P_{i, 1}W_i\omega^2\right) - \rho_2\left(\gamma(\eta\_w-1)(\gamma W_s + \omega W_i) + (\eta\_c-1)P_{i, 2}W_i\omega^2\right) > 0 \nonumber \\
+    &\rho_1\left(\gamma(\eta\_w-1)(\gamma w_S + \omega w_I) + (\eta\_c-1)p_{I,1}w_I\omega^2\right) - \rho_2\left(\gamma(\eta\_w-1)(\gamma w_S + \omega w_I) + (\eta\_c-1)p_{I, 2}w_I\omega^2\right) > 0 \nonumber \\
     &\vdots \nonumber \\
-    &\frac{\rho_2P_{i, 2}-\rho_1P_{i, 1} }{\rho_1-\rho_2} > \frac{1-\eta\_w}{1-\eta\_c}\cdot \frac{\gamma(\gamma W_s + \omega W_i)}{\omega^2 W_i}
+    &\frac{\rho_2P_{i, 2}-\rho_1P_{i, 1} }{\rho_1-\rho_2} > \frac{1-\eta\_w}{1-\eta\_c}\cdot \frac{\gamma(\gamma w_S + \omega w_I)}{\omega^2 w_I}
 \end{align}

### 2021-03-03 codes/sir_plot.R
panel plots changed to 1/omega, 1-eta, in the manuscript R0 is rewritten

@@ -1,4 +1,4 @@
-
+## setwd("/home/ag/projects/SIRmodel/codes/")
 source("SIRfunctions.R")
 source("params.R")
 # where to save the output:
@@ -14,17 +14,21 @@ n_in <- 41 ## grid N within facets
 W_S_random <- 1
 W_S_targeted <- 0.3 
 
-# make dataframe and save it. 
-# FIXME: I feel that this part needs to be separated from here?
-# eval_R0(params = update(params,W_S=W_S_random),filename="random_test_df.csv")
-# eval_R0(params = update(params,W_S=W_S_targeted),filename="targeted_test_df.csv")
+# make dataframe and save it, if it doesn't exists already.
+if (!file.exists("random_test_df.csv")) {
+  eval_R0(params = update(params,W_S=W_S_random),filename="random_test_df.csv")
+} else {
+  df_random <- read.csv(file = "random_test_df.csv")
+}
 
-# #################################
-# Load the data
-# #################################
-df_random <- read.csv(file = 'random_test_df.csv')
-df_targeted <- read.csv(file = 'targeted_test_df.csv')
-
+if (!file.exists("targeted_test_df.csv")) {
+  eval_R0(params = update(params,W_S=W_S_random),filename="targeted_test_df.csv")
+} else {
+  df_targeted <- read.csv(file = "targeted_test_df.csv")
+}
+## Add 2 columns
+df_random <- df_random %>% mutate(Eta_w=1-eta_w,Eta_c=1-eta_c )
+df_targeted <- df_targeted %>% mutate(Eta_w=1-eta_w,Eta_c=1-eta_c )
 # #################################
 # Plotting Part
 # #################################
@@ -49,33 +53,34 @@ brks_vec <- seq(0.8,1.05,by=0.05) # Break vector for unifying the legends in Ran
 
 df_temp <- df_random
 # df_temp <- df_targeted
-p1 <- (ggplot(df_temp,aes(x=omega,y=rho,z=R0_sub))
+p1 <- (ggplot(df_temp,aes(x=1/omega,y=rho,z=R0_sub))
     + theme_bw()
-    + xlab(TeX('$\\omega$, rate of test return (1/day)'))
+    + xlab(TeX('$\\1/omega$, mean test return time (day)'))
     + ylab(TeX('$\\rho$, testing intensity (1/day per capita)'))
 )
-
+## There might be a solution for the tick label collision with expand_limits(), but that's hard to do across facets
[…]

### 2021-03-04 note/targtedTesting.tex
starting review of  targeted testing modeling

--- 2020-11-15 note/SIR_notes.Rnw
@@
 \usepackage{subcaption}
-\usepackage{amsmath}
+\usepackage{amssymb,amsmath,amsthm}
+\usepackage{xcolor}
+\usepackage{xspace}
+\usepackage[nameinlink,capitalize]{cleveref}
+\usepackage{cleveref}
+\usepackage{geometry}
+\usepackage{pdflscape}
+
+\newcommand{\comment}{\showcomment}
+%% \newcommand{\comment}{\nocomment}
+
+\newcommand{\showcomment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}
+\newcommand{\nocomment}[3]{}
+
+\newcommand{\fady}[1]{\comment{cyan}{Fady}{#1}}
+\newcommand{\ali}[1]{\comment{magenta}{Ali}{#1}}
+\newcommand{\jd}[1]{\comment{blue}{JD}{#1}}
+\newcommand{\djde}[1]{\comment{red}{DJDE}{#1}}
+\newcommand{\bmb}[1]{\comment{red}{BMB}{#1}}
+\newcommand{\todo}[1]{\comment{red}{TODO}{#1}}
+
+\newcommand{\Rnum}{\mathcal{R}_0}
+\theoremstyle{definition} % amsthm only
+\newtheorem{prop}{Proposition}
+\newtheorem{note}{note}
+\newtheorem{theorem}{Theorem}
+
 \bibliographystyle{apalike}
 
-\title{SIR Model with Testing and Isolation Mechanisms}
+\title{Notes on modeling targeted testing }
 
@@
 % %%%%%%%
-\section{Method}
+\section{Review}
 
-\subsection{model and parameters}
-\begin{itemize}
-\item $\Lambda$: force of infection defined as $$\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_t I_t)}{N_0},$$ where $\beta$ is transmission rate, $\eta_w$ and $\eta_t$ are the isolation parameters for awaiting and reported individuals, respectively.
-We assume that $\eta_t<\eta_w$, i.e., the awaiting individuals for test results have a higher transmission probability than the reported individuals.
-\item $\omega$: the rate of *onward flow* from the awaiting positive compartment, $p$, to reported/tested compartment, $\rho$, or from awaiting negative compartment, $n$, back to $u$.
-It has units of 1/time.
-\item $\gamma$: recovery rate (1/time).
-\item $\rho$: per capita testing intensity across the whole population (1/time).
-\item $W$: weighted number of people available for tests, defined as $W = W_S S_u + W_I I_u + W_R R_u$.
-\item $\sigma$: scaling parameter for testing defined as $\sigma = \frac{\rho N_0}{W}$.
-\item $F_Z$: Weighted testing rate defined as $F_Z=\sigma W_Z$. That is, $F_S = \sigma W_S$, $F_I=\sigma W_I$ and $F_R = \sigma W_R$.
-\end{itemize}
+\cite{scarabel2021renewal} say: 
+Modelling contact tracing is challenging: a satisfactory mathematical description should account at least for concurrent screening/diagnosis programs that can initiate contact tracing, and the contacts between individuals that occurred prior to that moment.
+For these reasons,
+many mathematical models have been formulated as stochastic branching processes [6, 7, 8], or agent-based models [9, 10, 11, 12, 13], which allow to keep
+track of the epidemiological status of each individual in the population together
+with all their infectious contacts.
 
-\includegraphics[width=\linewidth]{../pix/sir_comp.png}
+The few deterministic models available in the literature [6, 14, 15, 16] are typically obtained as approximations of a corresponding stochastic model under simplifying
+assumptions.
 
-The model is
-
-\begin{align}
-\label{model}
- d S_u/dt &= -\Lambda S_u - F_S S_u + \omega S_n \\
- d S_n/dt &= -\Lambda S_n + F_S S_u - \omega S_n \\
- d I_u/dt &= \Lambda S_u - F_I I_u + \omega I_n  - \gamma I_u  \\
- d I_n/dt &= \Lambda S_n + (1-p_I) F_I I_u - \omega I_n -\gamma I_n \\
- d I_p/dt &= p_I F_I I_u - \omega I_p -\gamma I_p \\
- d I_t/dt &= \omega I_p - \gamma I_t  \\
- d R_u/dt &= \gamma I_u - F_R R_u + \omega R_n \\
- d R_n/dt &= \gamma I_n + (1-p_R) F_R R_u - \omega R_n  \\
- d R_p/dt &= \gamma I_p + p_R F_R R_u  - \omega R_p  \\
- d R_t/dt&= \gamma I_t + \omega R_p  \\
- dN/dt &= \omega (S_n + I_n + R_n)   \\
- dP/dt &= \omega(I_p + R_p).
-\end{align}
-
-% %%%%%%%
-\section{Results}
-
-\newpage
-\begin{figure}
-\centering
-\begin{subfigure}[t]{.45\textwidth}
-\centering
-\includegraphics[width=\linewidth]{../pix/R0contour_random.png}
-        \caption{}\label{fig:fig_a}
-\end{subfigure}
-%
-\begin{subfigure}[t]{.45\textwidth}
-\centering
-\includegraphics[width=\linewidth]{../pix/R0contour_TTI.png}
-\caption{}\label{fig:fig_b}
-\end{subfigure}
-\end{figure}
-
-\begin{figure}
-\centering
-\begin{subfigure}[t]{.45\textwidth}
-\centering
-\includegraphics[width=\linewidth]{../pix/R0contour_random_s.png}
-        \caption{}\label{fig:fig_a}
-\end{subfigure}
-%
-\begin{subfigure}[t]{.45\textwidth}
-\centering
-\includegraphics[width=\linewidth]{../pix/R0contour_TTI_s.png}
-\caption{}\label{fig:fig_b}
-\end{subfigure}
-\end{figure}
-
-**Insights from the R0-contour plots**
-
-1. TTI requires lower testing rate than random testing to achieve $R_0=1$.
-
-2. When random testing regime is applied, $W_S=W_I=W_R$,
-
-(i) in the case of perfect isolation of the reported individuals, $\eta_t=0$, the followings are inferred (see the first few columns of the panel plot); 
- - variations of the $\eta_w$ has a negligible effect on the critical testing.
- - for low $\omega$, $R_0$ increases as $\omega$ increases but saturates.
-This means that when the reporting process is very very slow, increasing the speed of the reporting causes higher $R_0$.
-Maybe because more people will move to $I_u$ [?, not very sure here.]
-
- (ii) in case of the non-perfect isolation (see the last column), when $\eta_w$ increases (more awaiting people follow the isolation rules) faster reporting reduces $R_0$.  
-Note that it is sensible to expect that $\partial{R_0}/\partial{\omega}<0$. Here we showed that this can be less true.
-
-3. When TTI regime is applied,
-
-% %%%%%%%
-\section{Thoughts/Objectives}
-
-\begin{enumerate}
-\item 
-\begin{itemize}
-\item analysis of the basic SIR model with testing: analytical results as far as possible, supplemented by numerical results as necessary/for illustration.
-Heuristic explanations/insights of what we learn from this.
-\item Maybe??
-order-of-magnitude/back-of-the-envelope calculations of the magnitudes of testing and isolation necessary.
-(However, the model may be too simplified for even back-of-the-envelope calculations to be meaningful.)
-\item Maybe?? comparison with strength-and-speed framework (since T/T/I is essentially a speed-based intervention)
-\end{itemize}
-
-\item
-\begin{itemize}
-\item extension to a SEPAIR model (i.e. including latent/exposed, presymptomatic, asymptomatic)
-  (i) analytical results will probably be out of reach
-	(ii) but it may be possible to extend some of the insights gained in part 1
-	(iii) in any case, some numerical results (the space of relevant/interesting parameters will be larger: figure out a sensible way to show results from this parameter space.
-Are there useful dimensionless parameters that could help frame this?)
-\item more detailed quantitative exploration: in particular, *either* Latin hypercube *or* (nearly the same!) a sample over estimated ('prior') distributions of the parameters.
-Goal: answer the general question "how likely is TTI to be able to control COVID?"
-\end{itemize}
-\end{enumerate}
-
-% \item extension to a SEPAIR model (i.e. including latent/exposed, presymptomatic, asymptomatic)
-%     * analytical results will probably be out of reach
-% 	* but it _may_ be possible to extend some of the insights gained in part 1
-% 	* in any case, some numerical results (the space of relevant/interesting parameters will be larger: figure out a sensible way to show results from this parameter space.
-Are there useful dimensionless parameters that could help frame this?)
-% \item more detailed quantitative exploration: in particular, *either* Latin hypercube *or* (nearly the same!) a sample over estimated ('prior') distributions of the parameters.
-Goal: answer the general question "how likely is TTI to be able to control COVID?"
-
-% %%%%%%%
-\section{Literature Review}
-
-\subsection{Explicit models of TTI (trace/test/isolate) based on network or agent-based models}
-
-\citep{endo2020implication} [Ali: It seems to me that this is just a statistical model to estimate the parent-offspring of an infected index, not sure if it fits into agent-based group!] Used simulation on a branching process model to assess the forward and backward contact tracing efficiency.
-Assuming a negative-binomial branching process with a mean R, reproduction number, and overdispersion parameter k, the mean total number of generation G3 and averted G3 are estimated.
-The effectiveness of TTI is defined as the ratio of averted to the mean.
-
-\citep{jenness2020modeling} developed a network-based transmission model for SARS-CoV-2 on the Diamond Princess outbreak to characterize transmission dynamics and to estimate the epidemiological impact of outbreak control and prevention measures.
-
-\citep{elbanna2020entry} [seems similar to MacPan model!]
-
-\citep{de2020influenza} Was discussed in the Math 747 
-SEIR Asymptomatic and symptomatic $I_1, I_2$. Used linear chain trick 
-Stringency index as a control force lowering $\beta$.
-
-\citep{rice2020effect} Effect of school closures on mortality. Reproduce Report 9 results by spatial agent based CovidSim. 
-% %%%%%%%
-\subsection{Models of repeated random testing of isolated populations}
-\cite{bergstrom2020frequency}
-(1) Model, assumptions: They developed a function, namely expected exposure $E(C,\tau)$, to approximate trade-offs between the frequency of testing, n, the sensitivity of testing, q, and the delay between
-testing and results, d. This function is explicitly derived and was connected the effective reproduction number $R=R_0 S$, where $S$ is the proportion of population susceptible.
-assumption that transmission rates are a step function: individuals who
-have COVID go from non-infectious to fully infectious instantaneously,
-and remain fully infectious until they are no longer able to transmit disease. Test sensitivity takes the same form over the course of infection.
-More sophisticated models could allow varying infectiousness and varying
-sensitivity over time, as in 
-\citep{larremore2020test}.
-
-\citep{lopman2020model} Used a Deterministic SEIR model, incorporated TTI, applicable to a university setting. They assumed a fairly high reproductive number that is not reduced through social
-distancing measures. They found that community-introduction of SARS-CoV-2 infection onto campus can be
-relatively controlled with effective testing, isolation, contract tracing and quarantine.
-
-\citep{tuite2020mathematical} used an age-structured compartmental model of COVID-19 transmission in the population of Ontario, Canada.
-We compared a base case with limited testing, isolation and quarantine to different scenarios.
-
-% %%%%%%%
-\subsection{Other maybe-related works}
-\citep{arino2020simple} developed a SLIAR compartmental model to study the spread of an epidemic, specifically COVID-19, in a population.
-The model incorporates an Erlang distribution of times of sojourn in incubating, symptomatically and asymptomatically infectious compartments.
-Basic reproduction number is derived.
-Also, sensitivity analysis with respect to the underlying parameters for the following two outputs was carried out; (i) the number of observable cases during the course of the epidemic and at the peak, and (ii) the timing of the peak of the outbreak.
-Sensitivity analysis is performed using the R package multisensi.
-
-\citep{ruszkiewicz2020diagnosis} novel with-in-a-minute breath testing with 80\% accuracy. 
 % %%%%%%%

### 2021-03-04 codes/SIRfunctions.R
pipe ing

@@ -1,3 +1,5 @@
+library(shellpipes)
+
 # Functions for the SIR model
 # By: A Gharouni
 
@@ -17,13 +19,6 @@ while (!require(McMasterPandemic)) {
   )
 }
 
-
-library(ggplot2)
-library(tidyr)
-library(directlabels)
-library(latex2exp)
-library(rootSolve)
-library(McMasterPandemic)
 unpack <- McMasterPandemic::unpack
 
 # Testing functions
@@ -195,5 +190,5 @@ eval_R0 <- function(state=state_dfe,params,filename){
 #   method="lsodar"
 # )
 
-
+saveEnvironment()
 

### 2021-03-04 codes/params.R

@@ -1,3 +1,8 @@
+library(shellpipes)
+rpcall("params.Rout params.R SIRfunctions.rda")
+
+loadEnvironments()
+
 # Parameters used in SIR model
 # Path: ~/projects/SIRmodel/codes
 
@@ -26,3 +31,5 @@ state_dfe <- c(S_u=Su_dfe(params), S_n=Sn_dfe(params),
                 I_u=0,I_n=0,I_p=0,I_t=0,
                 R_u=0,R_n=0,R_p=0,R_t=0,
                 N=0,P=0)
+
+saveEnvironment()

### 2021-03-04 codes/sir_plot.R

@@ -1,8 +1,22 @@
+library(ggplot2)
+library(dplyr)
+library(tidyr)
+library(directlabels)
+library(latex2exp)
+library(rootSolve)
+library(McMasterPandemic)
+
+library(shellpipes)
+
+rpcall("sir_plot.Rout sir_plot.R params.rda")
+
+loadEnvironments()
+
 ## setwd("/home/ag/projects/SIRmodel/codes/")
-source("SIRfunctions.R")
-source("params.R")
+
 # where to save the output:
-path <- "../pix/"
+path <- "./"
+
 # #################################
 # Making Dataframe part
 # #################################

### 2021-03-04 codes/plots4talk.R
Housekeeping for new, clean branch

@@ -1,104 +1,0 @@
-# Making plots for presentations/slides etc in the "path"
-path <- "~/projects/aliworkstation/references/presentations/epimodeling/pix/"
-# R code is at: ~/projects/SIRmodel/codes/
-# setwd("~/projects/SIRmodel/codes/")
-
-source("SIRfunctions.R")
-source("params.R")
-
-# #################################
-# Load the data
-# #################################
-df_random <- read.csv(file = 'random_test_df.csv') #random testing scenario
-df_targeted <- read.csv(file = 'targeted_test_df.csv') #targeted testing scenario
-#choose weight W_S 0.3 for TTI testing plan and 1 for the random testing.
-W_S_random <- unique(df_random$W_S)
-W_S_targeted <- unique(df_targeted$W_S)
-
-# #################################
-# Plotting Part
-# #################################
-# Setting the plots
-## hacked (not very robustly) to chain label_both() and label_parsed() ...
-label_special <- function (labels, multi_line=FALSE, sep = "== ") {
-  value <- ggplot2:::label_value(labels, multi_line = multi_line)
-  variable <- ggplot2:::label_variable(labels, multi_line = multi_line)
-  variable[[1]] <- gsub("_(.)$","[\\1]",variable[[1]])
-  ## not using multiple faceting variables on each margin, so forget paste
-  out <- Map(paste, variable, value, sep = sep)
-  out[[1]] <- lapply(out[[1]], function(expr) c(parse(text=expr)))
-  out
-}
-
-# verify the break range in brks_vec (FIXME)
-mn <- min(df_random$R0_sub,df_targeted$R0_sub, na.rm = T)
-mx <- max(df_random$R0_sub,df_targeted$R0_sub, na.rm = T)
-
-# show_contours_1, width=24,height=24,message=FALSE,warning=FALSE}
-brks_vec <- seq(0.8,1.05,by=0.05) # Break vector for unifying the legends in Random and TTI testing cases. 
-
-# #################################
-#  SubPlots the Random Testing Scenario:
-# #################################
-# Random Testing, Perfect isolation
-
-# FIXME: df <- paste("df",scenario,sep="_")
-# scenario <-"random"
-# df <- df_random
-scenario <-"targeted"
-df <- df_targeted
-
-# isolation for awaiting people
-eta_w_val <-0
[…]

### 2021-03-08 note/minutes.Rmd
more on the targeted testing, minutes are updated

@@
 ## Dr. Ben Bolker
+
+### Date: MArch 8
+ - There are papers by Friston 
+ Friston, Karl J., Thomas Parr, Peter Zeidman, Adeel Razi, Guillaume Flandin, Jean Daunizeau, Oliver J.
+Hulme, et al.
+“Dynamic Causal Modelling of COVID-19.” ArXiv:2004.04463 [q-Bio], April 9, 2020.
+http://arxiv.org/abs/2004.04463.
+
+ - in there model, they have a fig1 big model includes testing, don't know how they have testing based on symptoms (it seems), the thetas here represent just parameters , sigma is sigmoid function (we think), the 4 compartment ...
 

### 2021-03-10 SIR_manuscript.tex
panel plots changed to 1/omega, 1-eta, in the manuscript R0 is rewritten
New branch is working? for MS only, at least
amsync
Trying to understand ...

@@
 \usepackage{hyperref}
+\usepackage{grffile}
 \usepackage{graphicx}
@@
 \begin{center} 
-\includegraphics[scale=2]{./pix/sir_comp.pdf}
+\includegraphics[scale=2]{pix/sir_comp.pdf}
 \caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model, \ref{model}.
@@
 \label{R0}
-\Rnum= \frac{\beta}{\gamma} (A \times S\_u^* + B \times S\_n^*) \times C, 
+\Rnum= \frac{\beta}{\gamma} (1-\Delta), 
 \end{equation}
@@
 \label{eq:abc}
+1-\Delta=& (A~ S\_u^* + B~ S\_n^*)~ C\\
 A=& \gamma(\omega+\gamma) + (\gamma \eta\_w + \omega \eta\_c p_I) F_I, \\
@@
 \centering
-\includegraphics[width=\linewidth]{./pix/R0contour_random.pdf}
+\includegraphics[width=\linewidth]{codes/R0contour_random.pdf}
 \caption{}\label{p.a}
@@
 \centering
-\includegraphics[width=\linewidth]{./pix/R0contour_TTI.pdf}
+\includegraphics[width=\linewidth]{codes/R0contour_TTI.pdf}
 \caption{}\label{p.b}
@@
 \caption{
-{\bf A comparison of the behaviour of the basic reproduction number, $\Rnum$, between random versus targeted testing strategies at differnt levels of testing and isolation.}
+{\bf A comparison of the behaviour of the basic reproduction number, $\Rnum$, between random versus targeted testing strategies at different levels of testing and isolation.}
 We numerically evaluate $\Rnum$ \eqref{R0} using the following parameters (listed in Table \ref{tab:params}):

### 2021-03-10 codes/SIRfunctions.R
contours of Delta instead of R0 were added to the fig2

@@ -155,7 +155,7 @@ eval_R0 <- function(state=state_dfe,params,filename){
   # input the params and their range, this function makes a grid dataframe, calls R0 and outputs the csv file
   unpack(as.list(c(state,params)))
   # specify the ranges, FIXME, not to be hard coded!
-  beta  <- 0.339  ## set beta high enough to allow R0>1 in worst case
+  beta  <- params[["beta"]]  ## set beta high enough to allow R0>1 in worst case
   eta_w <- seq(0,1, length.out=n_out)
   eta_c <- seq(0,1, length.out=n_out) #0.001 to 0.5 
   rho <- seq(0,0.01, length.out=n_in)
@@ -169,7 +169,10 @@ eval_R0 <- function(state=state_dfe,params,filename){
                    R0=apply(df1,1,function(params_in)R0(params=params_in)),
                    Fi= apply(df1,1,function(params_in)F_I(state=state, params=params_in)))
   # This is for plotting purposes:
-  df2 <- df2 %>% dplyr::mutate(R0_sub = ifelse(eta_w<eta_c, NA, R0)) 
+  df2 <- (df2 %>% 
+            dplyr::mutate(R0_sub = ifelse(eta_w<eta_c, NA, R0), 
+                          Delta = ifelse(eta_w<eta_c, NA, 1-(R0*gamma/beta) )) 
+          )
   # Output
   write.csv(df2,file=filename, row.names = FALSE)
 }

### 2021-03-10 codes/params.R

@@ -3,7 +3,7 @@
 
 params <- c(
 N0=1000000, #total population size
-beta=.8, #transmission rate
+beta= 0.339, #transmission rate (was 0.8) 0.339
 gamma=1/3, #recovery rate
 omega=0.25, #test returning rate 
 rho=0.01, #testing rate percapita (also used 0.8, 1/3)

### 2021-03-10 codes/sir_plot.R

@@ -22,13 +22,20 @@ if (!file.exists("random_test_df.csv")) {
 }
 
 if (!file.exists("targeted_test_df.csv")) {
-  eval_R0(params = update(params,W_S=W_S_random),filename="targeted_test_df.csv")
+  eval_R0(params = update(params,W_S=W_S_targeted),filename="targeted_test_df.csv")
 } else {
   df_targeted <- read.csv(file = "targeted_test_df.csv")
 }
 ## Add 2 columns
-df_random <- df_random %>% mutate(Eta_w=1-eta_w,Eta_c=1-eta_c )
-df_targeted <- df_targeted %>% mutate(Eta_w=1-eta_w,Eta_c=1-eta_c )
+tol <- 1e-10
+df_random <- (df_random 
+             %>% dplyr::mutate(Eta_w=1-eta_w,Eta_c=1-eta_c,
+                              Delta= ifelse(abs(Delta)<tol,0,Delta) ))
+df_targeted <- (df_targeted 
+                %>% dplyr::mutate(Eta_w=1-eta_w,Eta_c=1-eta_c,
+                                  Delta= ifelse(abs(Delta)<tol,0,Delta) ))
+threshold <- 1-(params[["gamma"]]/params[["beta"]]) ## Or 0?  corresponding to R0=1 
+## note if we plot the contours of R0, the threshold =1
 # #################################
 # Plotting Part
 # #################################
@@ -45,15 +52,15 @@ label_special <- function (labels, multi_line=FALSE, sep = "== ") {
 }
 
 # verify the break range in brks_vec (FIXME)
-mn <- min(df_random$R0_sub,df_targeted$R0_sub, na.rm = T)
-mx <- max(df_random$R0_sub,df_targeted$R0_sub, na.rm = T)
+mn <- min(df_random$Delta,df_targeted$Delta, na.rm = T)
+mx <- max(df_random$Delta,df_targeted$Delta, na.rm = T)
 
 # show_contours_1, width=24,height=24,message=FALSE,warning=FALSE}
-brks_vec <- seq(0.8,1.05,by=0.05) # Break vector for unifying the legends in Random and TTI testing cases. 
+brks_vec <- seq(0,.165,by=0.02) # Break vector for unifying the legends in Random and TTI testing cases. 
 
 df_temp <- df_random
 # df_temp <- df_targeted
-p1 <- (ggplot(df_temp,aes(x=1/omega,y=rho,z=R0_sub))
[…]

### 2021-03-11 codes/SIRfunctions.R
Cleaning, fixing and merging with Ali

@@ -146,7 +146,7 @@ R02 <- function(state=state_dfe,params){
 }
 
 # make grid and calc R0 for plotting
-eval_R0 <- function(state=state_dfe,params,filename){
+eval_R0 <- function(state=state_dfe,params){
   # input the params and their range, this function makes a grid dataframe, calls R0 and outputs the csv file
   unpack(as.list(c(state,params)))
   # specify the ranges, FIXME, not to be hard coded!
@@ -169,7 +169,7 @@ eval_R0 <- function(state=state_dfe,params,filename){
                           Delta = ifelse(eta_w<eta_c, NA, 1-(R0*gamma/beta) )) 
           )
   # Output
-  write.csv(df2,file=filename, row.names = FALSE)
+  return(df2)
 }
 
 

### 2021-03-11 codes/params.R

@@ -32,4 +32,4 @@ state_dfe <- c(S_u=Su_dfe(params), S_n=Sn_dfe(params),
                 R_u=0,R_n=0,R_p=0,R_t=0,
                 N=0,P=0)
 
-saveEnvironment()
+saveVars(params, state_init, state_dfe)

### 2021-03-11 codes/sir_plot.R

@@ -14,9 +14,6 @@ loadEnvironments()
 
 ## setwd("/home/ag/projects/SIRmodel/codes/")
 
-# where to save the output:
-path <- "./"
-
 # #################################
 # Making Dataframe part
 # #################################
@@ -29,17 +26,9 @@ W_S_random <- 1
 W_S_targeted <- 0.3 
 
 # make dataframe and save it, if it doesn't exists already.
-if (!file.exists("random_test_df.csv")) {
-  eval_R0(params = update(params,W_S=W_S_random),filename="random_test_df.csv")
-} else {
-  df_random <- read.csv(file = "random_test_df.csv")
-}
+df_random <- eval_R0(params = update(params,W_S=W_S_random))
+df_targeted <- eval_R0(params = update(params,W_S=W_S_targeted))
 
-if (!file.exists("targeted_test_df.csv")) {
-  eval_R0(params = update(params,W_S=W_S_targeted),filename="targeted_test_df.csv")
-} else {
-  df_targeted <- read.csv(file = "targeted_test_df.csv")
-}
 ## Add 2 columns
 tol <- 1e-10
 df_random <- (df_random 
@@ -104,8 +93,7 @@ p1_temp
 ggsave(p1_temp + ggtitle(TeX("w_S=w_I=w_R=1")) +
        theme(legend.position = "none"),
        filename = "R0contour_random.pdf" ,
-       width = 12, height = 12, units = "cm",
-       path=path)
+       width = 12, height = 12, units = "cm")
 
 # #################################
 # 2. Plot the Targeted Testing Scenario:
@@ -115,8 +103,7 @@ ggsave((p1_temp %+% df_targeted) +
        ggtitle(TeX(sprintf("w_S=%.1f, w_I=w_R=1",W_S_targeted))) +
        theme(legend.position = c(0.2, 0.3)),
        filename = "R0contour_TTI.pdf" ,
-       width = 12, height = 12, units = "cm",
-       path=path)
+       width = 12, height = 12, units = "cm")
 
 # #################################
 # 3. Plot R0 as a function of rho (testing intensity):

### 2021-03-12 SIR_manuscript.tex
improved the eval_R0, more context added in the plot

@@
 {\bf A comparison of the behaviour of the basic reproduction number, $\Rnum$, between random versus targeted testing strategies at different levels of testing and isolation.}
-We numerically evaluate $\Rnum$ \eqref{R0} using the following parameters (listed in Table \ref{tab:params}):
+We numerically evaluate $\Delta$ \eqref{R0}, reflecting the reduction of $\Rnum$ with respect to testing and isolation. We use the following parameters (listed in Table \ref{tab:params}):
 $N_0=1 \times 10^6$, $\omega \in [0.1,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.01]$ 1/day, $\eta\_w \in [0,1]$, $\eta\_c \in [0,1]$, $\beta=0.39$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
-Contours of $\Rnum$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
-The black solid line in each panel represents the critical contour of $\Rnum=1$.
-\bmb{we still have some collisions in the y-axis tick labels}
-% \ali{plotr + expand_limits(y = 0), change 1/\omega on x-axis}
+Note we are plotting $\eta\_w'=1-\eta\_w$ and $\eta\_c'=1-\eta\_c$ in the facets to better reflect the effect of isolation (0 for no effect and 1 for full effect).
+Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
+The black solid line in each panel represents the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ corresponding to the threshold of $\Rnum=1$.
+\bmb{we still have some collisions in the y-axis tick labels} \ali{I am using $\eta'=1-\eta$, but in R, I used Eta which is shown as H, I don't know how to show the prime in the facets!
+Help!}
 }

### 2021-03-12 codes/SIRfunctions.R

@@ -149,6 +149,7 @@ R02 <- function(state=state_dfe,params){
 eval_R0 <- function(state=state_dfe,params){
   # input the params and their range, this function makes a grid dataframe, calls R0 and outputs the csv file
   unpack(as.list(c(state,params)))
+  tol <- 1e-10 ## to resolve the issue of very small numbers 
   # specify the ranges, FIXME, not to be hard coded!
   beta  <- params[["beta"]]  ## set beta high enough to allow R0>1 in worst case
   eta_w <- seq(0,1, length.out=n_out)
@@ -165,10 +166,13 @@ eval_R0 <- function(state=state_dfe,params){
                    Fi= apply(df1,1,function(params_in)F_I(state=state, params=params_in)))
   # This is for plotting purposes:
   df2 <- (df2 %>% 
-            dplyr::mutate(R0_sub = ifelse(eta_w<eta_c, NA, R0), 
-                          Delta = ifelse(eta_w<eta_c, NA, 1-(R0*gamma/beta) )) 
+            dplyr::mutate(R0_sub=ifelse(eta_w<eta_c, NA, R0), 
+                          Eta_w=1-eta_w,
+                          Eta_c=1-eta_c,
+                          Delta=ifelse(eta_w<eta_c, NA, 1-(R0*gamma/beta) ),
+                          Delta=ifelse(abs(Delta)<tol,0,Delta)
+                          ) 
           )
-  # Output
   return(df2)
 }
 

### 2021-03-12 codes/sir_plot.R

@@ -29,16 +29,8 @@ W_S_targeted <- 0.3
 df_random <- eval_R0(params = update(params,W_S=W_S_random))
 df_targeted <- eval_R0(params = update(params,W_S=W_S_targeted))
 
-## Add 2 columns
-tol <- 1e-10
-df_random <- (df_random 
-             %>% dplyr::mutate(Eta_w=1-eta_w,Eta_c=1-eta_c,
-                              Delta= ifelse(abs(Delta)<tol,0,Delta) ))
-df_targeted <- (df_targeted 
-                %>% dplyr::mutate(Eta_w=1-eta_w,Eta_c=1-eta_c,
-                                  Delta= ifelse(abs(Delta)<tol,0,Delta) ))
+##important contour, ie R0=1 thus threshold=1 when plotting R0 contours, or Delta(R0=1)
 threshold <- 1-(params[["gamma"]]/params[["beta"]]) ## Or 0?  corresponding to R0=1 
-## note if we plot the contours of R0, the threshold =1
 # #################################
 # Plotting Part
 # #################################
@@ -56,7 +48,7 @@ label_special <- function (labels, multi_line=FALSE, sep = "== ") {
 
 # verify the break range in brks_vec (FIXME)
 mn <- min(df_random$Delta,df_targeted$Delta, na.rm = T)
-mx <- max(df_random$Delta,df_targeted$Delta, na.rm = T)
+mx <- max(df_random$Delta,df_targeted$Delta, na.rm = T) ## take mn and mx and set the brks_vec, smarter way?
 
 # show_contours_1, width=24,height=24,message=FALSE,warning=FALSE}
 brks_vec <- seq(0,.165,by=0.02) # Break vector for unifying the legends in Random and TTI testing cases. 
@@ -73,10 +65,10 @@ p1 <- (ggplot(df_temp,aes(x=1/omega,y=rho,z=Delta))
 p1_temp <- (p1
             + geom_contour_filled(breaks=brks_vec)
             + geom_contour(breaks=threshold,alpha=0.5,colour="black")
-            + facet_grid(Eta_w~Eta_c, labeller=label_special)
+            + facet_grid(Eta_w~Eta_c, labeller=label_special) ## Eta looks like H in the text, better notation?
             + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
[…]

### 2021-03-18 SIR_manuscript.tex
Commenting, editing and thinking too much about algebra

@@
 In addition to shaping epidemic observations (via case reports), testing processes can also affect epidemiological dynamics.
-In particular, individuals who are awaiting the results of a test may partially or fully self-isolate, and individuals who have tested positive are more likely to self-isolate.
-We developed a model that incorporates mechanistic epidemic processes and testing in order to explore the effects of testing and isolation on epidemic dynamics.
+In particular, individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may do so also (possibly to a lesser extent).
+We developed a mechanistic model that incorporates epidemic processes and testing in order to explore the effects of testing and isolation on epidemic dynamics.
 
-Because individuals who have been tested or received a positive test are likely to self-isolate---reducing or eliminating  their potentially infectious contacts---the dynamics of an epidemic will depend on who gets tested and
-why. For a given testing intensity (tests performed per day), testing will have a stronger effect on infection rates
-the more strongly it is focused on people who are infectious.
+If testing influences behavior, then epidemic dynamics will depend on patterns of who gets tested.
+The impacts of testing will depend on intensity (tests performed per day), and on how strongly testing is focused on people who are infectious.
 This level of focus depends in turn on the purpose and design of testing programs. 
 When testing is done for the purposes of disease surveillance \citep{foddai2020base}
-we might expect tests to be assigned randomly across the population, thus we characterize it as \emph{random} testing strategy.
-In an ideal world, some bias in selection for testing is probably inevitable, although a stratified sampling design would be preferred for statistical efficiency \citep{graubard1996modelling} \ali{a better ref everyone?}.
+tests should be assigned randomly across the population, possibly with a stratified design for statistical efficiency \citep{graubard1996modelling} \ali{a better ref everyone?}.
 
@@
 
-When testing is used primarily for diagnosis it will focus on people with infection-like symptoms; thus the testing weights for people with vs.\ without the infections will depend on the relative magnitudes of the probability of having the infection-like symptoms when infected.
-Specifically, for \covid infection, the testing weights will depend on the relative proportions of presymptomatic or asymptomatic vs.\ symptomatic infections and the probability of having \covid-like symptoms when \emph{not} infected (e.g., due to other respiratory illnesses).
+When testing is used primarily for diagnosis it will focus on people with infection-like symptoms; thus the relative testing weights for infected people will depend on the relative probability of infected people having symptoms.
+For \covid infection, the testing weights will depend on the relative asymptomatic infections, time spent pre-sympomatic vs.\ symptomatic infections -- and also the incidence of \covid-like symptoms among people in the population \emph{not} infected with \covid.
 Testing for epidemic control will focus on people who are known to have been in contact with known infected cases; in this case the testing weights for infected vs.\ uninfected people will depend on the probability of infection given contact, as well as the thoroughness and effectiveness of the system for identifying suspicious contacts.
@@
 
-In order to link to more applied models, i.e., including both epidemiological and testing processes, we constructed this model so we could specify the total \percap testing rate.
+In order to link to more applied models, we constructed this model so we could specify the total \percap testing rate.
 We do this by defining the weighted size of the testing pool $W = w_S S\_u + w_I I\_u + w_R R\_u$, and calculating a scaling parameter for testing as:

### 2021-03-19 SIR_manuscript.tex
Autocommit

@@
 \end{equation}
-Thus, the \percap testing rate for compartment $X$ is $F_X=\sigma W_X$, where $X \in \{S,I,R\}$. 
+Thus, the \percap testing rate for compartment $X$ is $F_X=\sigma w_X$, where $X \in \{S,I,R\}$. 
 For a high-sensitivity test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus unavailable for further testing.
@@
 across all susceptible pools with transmission rate $\beta$ and relative probability of transmission for ``waiting'' and \emph{confirmed positive} individuals, $\eta\_w$ and $\eta\_c$ respectively, (see Table~\ref{tab:params} for further details), (ii) $\eta\_c \leq \eta\_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
+Note that for convential purposes in the interpretation of the results with respect to these isolation parameters, we define the transformed isolation parameters as $\theta\_w=1-\eta\_w$ and $\theta\_c=1-\eta\_c$.
+Thus, for instance when the awaiting people follow the isolation perfectly, $\theta\_w$ is closer to 1, while when they less follow the isolation, $\theta\_w$ is closer to 0.
 For this analysis, we also assume a perfectly specific test ($p_S=0$).
@@
   $\rho$   & Per capita testing intensity & 1/day & 0.01 \\ \hline 
-  $\eta\_w$  & Relative probability of transmission for ``waiting'' individuals & - & - \\ \hline
-  $\eta\_c$  & Relative probability of transmission for ``confirmed positive'' individuals & - & -  \\ \hline
+  $\eta\_w$ ($\theta\_w=1-\eta\_w$)  & Relative probability of transmission for ``waiting'' individuals & - & - \\ \hline
+  $\eta\_c$ ($\theta\_c=1-\eta\_c$) & Relative probability of transmission for ``confirmed positive'' individuals & - & -  \\ \hline
   $\beta$ & Transmission rate & 1/day & 0.39 \\ \hline
@@
   \end{tabular}
-\caption{\label{tab:params} Parameters of the model \eqref{model}.}
+\caption{\label{tab:params} Parameters of the model \eqref{model}.\todo{figure out the best way to put the new definition $\theta=1-\eta$} in the table.}
 \end{table}
 
-The Disease-Free Equilibrium (DFE) for the SIR model, Eqs.
+\todo{the algebra needs to be changed based on JD's comment} The Disease-Free Equilibrium (DFE) for the SIR model, Eqs.
 \eqref{model}, is given by setting the differential equations to 0 and solving for the unknowns (\ali{is this clear or the "unkowns" needs to be defined?
@@
 \label{eq:abc}
-1-\Delta=& (A~ S\_u^* + B~ S\_n^*)~ C\\
+\Delta=& 1-(A~ S\_u^* + B~ S\_n^*)~ C\\
 A=& \gamma(\omega+\gamma) + (\gamma \eta\_w + \omega \eta\_c p_I) F_I, \\
@@
 The black solid line in each panel represents the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ corresponding to the threshold of $\Rnum=1$.
-\bmb{we still have some collisions in the y-axis tick labels} \ali{I am using $\eta'=1-\eta$, but in R, I used Eta which is shown as H, I don't know how to show the prime in the facets!
-Help!}
+\bmb{we still have some collisions in the y-axis tick labels}
 }
@@
 
-% %%%%%%%
-\subsection{Literature Review}
-
-\subsubsection{Explicit models of TTI (trace/test/isolate) based on network or agent-based models}
-\citep{endo2020implication} \ali{It seems to me that this is just a statistical model to estimate the parent-offspring of an infected index, not sure if it fits into agent-based group!} Used simulation on a branching process model to assess the forward and backward contact tracing efficiency.
-Assuming a negative-binomial branching process with a mean R, reproduction number, and overdispersion parameter k, the mean total number of generation G3 and averted G3 are estimated.
-The effectiveness of TTI is defined as the ratio of averted to the mean.
-
-\citep{jenness2020modeling} developed a network-based transmission model for SARS-CoV-2 on the Diamond Princess outbreak to characterize transmission dynamics and to estimate the epidemiological impact of outbreak control and prevention measures.
-
-\citep{elbanna2020entry} [seems similar to MacPan model!]
-
-\citep{de2020influenza} Was discussed in the Math 747 
-SEIR Asymptomatic and symptomatic $I\_1, I\_2$. Used linear chain trick 
-Stringency index as a control force lowering $\beta$.
-
-\citep{rice2020effect} Effect of school closures on mortality. Reproduce Report 9 results by spatial agent based CovidSim. 
-% %%%%%%%
-\subsubsection{Models of repeated random testing of isolated populations}
-\cite{bergstrom2020frequency}
-(1) Model, assumptions: They developed a function, namely expected exposure $E(C,\tau)$, to approximate trade-offs between the frequency of testing, n, the sensitivity of testing, q, and the delay between
-testing and results, d. This function is explicitly derived and was connected the effective reproduction number $R=\Rnum S$, where $S$ is the proportion of population susceptible.
-assumption that transmission rates are a step function: individuals who
-have \covid go from non-infectious to fully infectious instantaneously,
-and remain fully infectious until they are no longer able to transmit disease. Test sensitivity takes the same form over the course of infection.
-More sophisticated models could allow varying infectiousness and varying
-sensitivity over time, as in 
-\citep{larremore2020test}.
-
-\citep{lopman2020model} Used a Deterministic SEIR model, incorporated TTI, applicable to a university setting. They assumed a fairly high reproductive number that is not reduced through social
-distancing measures. They found that community-introduction of SARS-CoV-2 infection onto campus can be
-relatively controlled with effective testing, isolation, contract tracing and quarantine.
-
-\citep{tuite2020mathematical} used an age-structured compartmental model of \covid transmission in the population of Ontario, Canada.
-We compared a base case with limited testing, isolation and quarantine to different scenarios.
-% %%%%%%%
-\subsubsection{Other maybe-related works}
-\citep{arino2020simple} developed a SLIAR compartmental model to study the spread of an epidemic, specifically COVID, in a population.
-The model incorporates an Erlang distribution of times of sojourn in incubating, symptomatically and asymptomatically infectious compartments.
-Basic reproduction number is derived.
-Also, sensitivity analysis with respect to the underlying parameters for the following two outputs was carried out; (i) the number of observable cases during the course of the epidemic and at the peak, and (ii) the timing of the peak of the outbreak.
-Sensitivity analysis is performed using the R package multisensi.
-
-\citep{ruszkiewicz2020diagnosis} novel with-in-a-minute breath testing with 80\% accuracy.
-
 \end{document}

### 2021-03-19 codes/SIRfunctions.R
Eta changed to theta in the plotting and dataframes

@@ -167,8 +167,8 @@ eval_R0 <- function(state=state_dfe,params){
   # This is for plotting purposes:
   df2 <- (df2 %>% 
             dplyr::mutate(R0_sub=ifelse(eta_w<eta_c, NA, R0), 
-                          Eta_w=1-eta_w,
-                          Eta_c=1-eta_c,
+                          theta_w=1-eta_w,
+                          theta_c=1-eta_c,
                           Delta=ifelse(eta_w<eta_c, NA, 1-(R0*gamma/beta) ),
                           Delta=ifelse(abs(Delta)<tol,0,Delta)
                           ) 

### 2021-03-19 codes/sir_plot.R

@@ -65,11 +65,11 @@ p1 <- (ggplot(df_temp,aes(x=1/omega,y=rho,z=Delta))
 p1_temp <- (p1
             + geom_contour_filled(breaks=brks_vec)
             + geom_contour(breaks=threshold,alpha=0.5,colour="black")
-            + facet_grid(Eta_w~Eta_c, labeller=label_special) ## Eta looks like H in the text, better notation?
+            + facet_grid(theta_w~theta_c, labeller=label_special) 
             + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
             + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
             + scale_fill_viridis_d(name=parse(text="Delta"),drop=FALSE)
-            + geom_rect(data=df_temp, fill=ifelse(df_temp$Eta_c < df_temp$Eta_w,"grey90","NA" ),
+            + geom_rect(data=df_temp, fill=ifelse(df_temp$theta_c < df_temp$theta_w,"grey90","NA" ),
                         color= NA,
                         ymin=-1,
                         ymax=10,

### 2021-03-23 SIR_manuscript.tex
fixed van den in bib

@@
 \end{equation}
+For simplification of the analysis, we adopt the following change $x\_u^* = S\_u^*/N_0$ and $x\_n^* = S\_n^*/N_0$.
 
@@
 \label{eq:abc}
-\Delta=& 1-(A~ S\_u^* + B~ S\_n^*)~ C\\
-A=& \gamma(\omega+\gamma) + (\gamma \eta\_w + \omega \eta\_c p_I) F_I, \\
-B=& \big(\omega+(F_I+\gamma)\eta\_w\big) \gamma+\frac{(\eta\_w \gamma+ \eta\_c\omega) \omega p_I F_I }{\omega+\gamma}, \\ 
-C=& \frac{1}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}.
+\Delta=& (A~ \theta\_w + B~ \theta\_c)~ C\\
+A=& \gamma \big((\omega+\gamma)+\omega p_I F_I \big) \big(x_n^*+\gamma(\omega+\gamma)F_I \big)\\
+B=& \omega p_I F_I (\gamma x_u^*+\omega) \\ 
+C=& \frac{1}{(\omega+\gamma) \big(\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I) \big)}.
 \end{aligned}
 \end{equation}
+% \begin{equation}
+% \label{R0}
+% \Rnum= \frac{\beta}{\gamma} (1-\Delta), 
+% \end{equation}
+% where
+% \begin{equation}
+% \begin{aligned}
+% \label{eq:abc}
+% \Delta=& 1-(A~ S\_u^* + B~ S\_n^*)~ C\\
+% A=& \gamma(\omega+\gamma) + (\gamma \eta\_w + \omega \eta\_c p_I) F_I, \\
+% B=& \big(\omega+(F_I+\gamma)\eta\_w\big) \gamma+\frac{(\eta\_w \gamma+ \eta\_c\omega) \omega p_I F_I }{\omega+\gamma}, \\ 
+% C=& \frac{1}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}.
+% \end{aligned}
+% \end{equation}
 Further details of derivation of $\Rnum$ are provided in the appendix section.

### 2021-03-24 codes/sir_plot.R
corrected A in Delta expression, white color threshold

@@ -51,7 +51,7 @@ mn <- min(df_random$Delta,df_targeted$Delta, na.rm = T)
 mx <- max(df_random$Delta,df_targeted$Delta, na.rm = T) ## take mn and mx and set the brks_vec, smarter way?
 
 # show_contours_1, width=24,height=24,message=FALSE,warning=FALSE}
-brks_vec <- seq(0,.165,by=0.02) # Break vector for unifying the legends in Random and TTI testing cases. 
+brks_vec <- seq(0,.165,by=0.03) # Break vector for unifying the legends in Random and TTI testing cases. 
 
 df_temp <- df_random
 # df_temp <- df_targeted
@@ -64,7 +64,7 @@ p1 <- (ggplot(df_temp,aes(x=1/omega,y=rho,z=Delta))
 ## see https://stackoverflow.com/questions/41575045/avoiding-axis-tick-label-collision-in-faceted-ggplots
 p1_temp <- (p1
             + geom_contour_filled(breaks=brks_vec)
-            + geom_contour(breaks=threshold,alpha=0.5,colour="black")
+            + geom_contour(breaks=threshold,alpha=0.5,colour="white")
             + facet_grid(theta_w~theta_c, labeller=label_special) 
             + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
             + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)

### 2021-03-25 SIR_manuscript.tex
corrected A in Delta expression, white color threshold
rewrite tex in theta, change the order of legend in plot

--- 2021-03-19 SIR_manuscript.tex
@@
 \label{Lambda}
-\Lambda=\beta \frac{(I\_u+\eta\_w I\_n+\eta\_w I\_p+ \eta\_c I\_c)}{N_0},
+\Lambda=\frac{\beta}{N_0} \big(I\_u + (1-\theta\_w)(I\_n+I\_p) + (1-\theta\_c)I\_c \big),
 \end{equation}
-across all susceptible pools with transmission rate $\beta$ and relative probability of transmission for ``waiting'' and \emph{confirmed positive} individuals, $\eta\_w$ and $\eta\_c$ respectively, (see Table~\ref{tab:params} for further details), (ii) $\eta\_c \leq \eta\_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
-Note that for convential purposes in the interpretation of the results with respect to these isolation parameters, we define the transformed isolation parameters as $\theta\_w=1-\eta\_w$ and $\theta\_c=1-\eta\_c$.
+across all susceptible pools with transmission rate $\beta$ and isolation efficacy in reduction of the probability of transmission for ``waiting'' and \emph{confirmed positive} individuals, $\theta\_w$ and $\theta\_c$ respectively, (see Table~\ref{tab:params} for further details), (ii) $\theta\_c \geq \theta\_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
 Thus, for instance when the awaiting people follow the isolation perfectly, $\theta\_w$ is closer to 1, while when they less follow the isolation, $\theta\_w$ is closer to 0.
@@
   $\rho$   & Per capita testing intensity & 1/day & 0.01 \\ \hline 
-  $\eta\_w$ ($\theta\_w=1-\eta\_w$)  & Relative probability of transmission for ``waiting'' individuals & - & - \\ \hline
-  $\eta\_c$ ($\theta\_c=1-\eta\_c$) & Relative probability of transmission for ``confirmed positive'' individuals & - & -  \\ \hline
+  $\theta\_w$ & Isolation efficacy in reduction of the probability of transmission for ``waiting'' individuals & - & - \\ \hline
+  $\theta\_c$ & Isolation efficacy in reduction of the probability of transmission for ``confirmed positive'' individuals & - & -  \\ \hline
   $\beta$ & Transmission rate & 1/day & 0.39 \\ \hline
@@
   \end{tabular}
-\caption{\label{tab:params} Parameters of the model \eqref{model}.\todo{figure out the best way to put the new definition $\theta=1-\eta$} in the table.}
+\caption{\label{tab:params} Parameters of the model \eqref{model}.}
 \end{table}
@@
 \label{eq:abc}
-\Delta=& 1-(A~ S\_u^* + B~ S\_n^*)~ C\\
-A=& \gamma(\omega+\gamma) + (\gamma \eta\_w + \omega \eta\_c p_I) F_I, \\
-B=& \big(\omega+(F_I+\gamma)\eta\_w\big) \gamma+\frac{(\eta\_w \gamma+ \eta\_c\omega) \omega p_I F_I }{\omega+\gamma}, \\ 
-C=& \frac{1}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}.
+\Delta=& (A~ \theta\_w + B~ \theta\_c)~ C\\
+A=& \gamma \big( (\gamma \omega+ \omega p_I F_I+\gamma^2) x_n^* + (\omega+\gamma) F_I \big)\\
+B=& \omega p_I F_I (\gamma x_u^*+\omega) \\ 
+C=& \frac{1}{(\omega+\gamma) \big(\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I) \big)}.
 \end{aligned}
 \end{equation}
+
 Further details of derivation of $\Rnum$ are provided in the appendix section.
@@
 The explicit formula for the basic reproduction number, $\Rnum$ \eqref{R0}, provides an opportunity to study the influence of changes in the underlying parameters on the critical index of epidemic dynamics.
-We are interested in understanding the effect of parameters that can be realistically controlled by changing in isolation, \percap testing intensity and test resulting, i.e., $\eta\_c$ and $\eta\_w$, $\rho$ and $\omega$, respectively.
+We are interested in understanding the effect of parameters that can be realistically controlled by changing in isolation, \percap testing intensity and test resulting, i.e., $\theta\_c$ and $\theta\_w$, $\rho$ and $\omega$, respectively.
 The following Proposition is the direct result of taking partial derivative of $\Rnum$ \eqref{R0} with respect to selected parameters.
@@
 \item \label{p1:eta} increasing the relative probabilities of transmission for tested individuals, increases $\Rnum$;
-$$\partial{\Rnum}/\partial{\eta\_c} \geq 0 \text{~and~} \partial{\Rnum}/\partial{\eta\_w} \geq 0.$$ 
+$$\partial{\Rnum}/\partial{\theta\_c} \leq 0 \text{~and~} \partial{\Rnum}/\partial{\theta\_w} \leq 0.$$ 
 \item \label{p1:rho} higher rate of testing reduces $\Rnum$;
@@
 
-Given that the perfect isolation occurs when no one transmits while waiting for test results (i.e., $\eta\_c=\eta\_w = 0$), Prop.~\ref{p1:eta} means that lifting the isolation from awaiting group results in an increased $\Rnum$ and consequently a greater number of infected individuals.
+Given that the perfect isolation occurs when no one transmits while waiting for test results (i.e., $\theta\_c=\theta\_w = 1$), Prop.~\ref{p1:eta} means that lifting the isolation from awaiting group results in an increased $\Rnum$ and consequently a greater number of infected individuals.
 Prop.~\ref{p1:rho} indicates that increasing the \percap testing intensity, $\rho$, reduces $\Rnum$.
@@
 We numerically evaluate $\Delta$ \eqref{R0}, reflecting the reduction of $\Rnum$ with respect to testing and isolation. We use the following parameters (listed in Table \ref{tab:params}):
-$N_0=1 \times 10^6$, $\omega \in [0.1,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.01]$ 1/day, $\eta\_w \in [0,1]$, $\eta\_c \in [0,1]$, $\beta=0.39$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
-Note we are plotting $\eta\_w'=1-\eta\_w$ and $\eta\_c'=1-\eta\_c$ in the facets to better reflect the effect of isolation (0 for no effect and 1 for full effect).
+$N_0=1 \times 10^6$, $\omega \in [0.1,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.01]$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.39$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
 Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
@@
 It can be inferred that when random testing strategy is applied, Figure \ref{p.a}, comparing to the targeted testing strategy, Figure \ref{p.b}, the critical \percap testing intensity of the corresponding panels are lower in non-random testing.
-Also, it appears that speeding the test reporting, i.e., increasing $\omega$, does not significantly lower $\Rnum$ when awaiting people (including people in $I\_p$ and $I\_n$ compartments) follow the isolation perfectly, i.e., $\eta\_w$ is closer to 0.
-However, speeding the test reporting reduces the epidemic more when the awaiting people less follow the isolation, i.e., $\eta\_w$ is closer to 1.
+Also, it appears that speeding the test reporting, i.e., increasing $\omega$, does not significantly lower $\Rnum$ when awaiting people (including people in $I\_p$ and $I\_n$ compartments) follow the isolation perfectly, i.e., $\theta\_w$ is closer to 1.
+However, speeding the test reporting reduces the epidemic more when the awaiting people less follow the isolation, i.e., $\theta\_w$ is closer to 0.
 

### 2021-03-25 codes/sir_plot.R

@@ -51,7 +51,7 @@ mn <- min(df_random$Delta,df_targeted$Delta, na.rm = T)
 mx <- max(df_random$Delta,df_targeted$Delta, na.rm = T) ## take mn and mx and set the brks_vec, smarter way?
 
 # show_contours_1, width=24,height=24,message=FALSE,warning=FALSE}
-brks_vec <- seq(0,.165,by=0.03) # Break vector for unifying the legends in Random and TTI testing cases. 
+brks_vec <- seq(.165,0,by=-0.03) # Break vector for unifying the legends in Random and TTI testing cases. 
 
 df_temp <- df_random
 # df_temp <- df_targeted
@@ -64,7 +64,7 @@ p1 <- (ggplot(df_temp,aes(x=1/omega,y=rho,z=Delta))
 ## see https://stackoverflow.com/questions/41575045/avoiding-axis-tick-label-collision-in-faceted-ggplots
 p1_temp <- (p1
             + geom_contour_filled(breaks=brks_vec)
-            + geom_contour(breaks=threshold,alpha=0.5,colour="white")
+            + geom_contour(breaks=threshold,alpha=0.5,colour="black")
             + facet_grid(theta_w~theta_c, labeller=label_special) 
             + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
             + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)

### 2021-03-25 SIR_manuscript.tex
replace eta by 1-theta in Appendix (still can be simplified more)

@@
 It states that returning test results more rapidly, i.e., increasing the rate $\omega$, does not necessarily lower the reproduction number $\Rnum$; whether increasing $\omega$ lowers $\Rnum$ depends on the precise combination of model parameters  including test reporting rate, testing strategies represented by compartment-specific testing weights, test sensitivity and specificity, and the level of isolation.
-Specifically, in the case of perfect isolation, i.e., when $\eta\_w=0$ and $\eta\_c=0$, $\Rnum$ may increase as the test reporting process becomes faster. This can be seen from expression \eqref{Rom}.
+Specifically, in the case of perfect isolation, i.e., when $\theta\_w=1$ and $\theta\_c=1$, $\Rnum$ may increase as the test reporting process becomes faster.
+This can be seen from expression \eqref{Rom}.
 Another example of this favorable delay in $\omega$ could be when the test being employed produces many false negatives.
@@
 To give a biological interpretation, we describe the qualitative trends predicted by the inequality \eqref{eq:necsuf}.
-To summarise, returning test results more rapidly tends to be favourable (i.e., reduces $\Rnum$) when (i) \emph{Confirmed-positive} individuals, lower their contact much more than individuals who are waiting for their test results (i.e., $\eta\_c \ll \eta\_w$), (ii) the test is highly sensitive (i.e., $p_I$ is close to 1), and (iii) the targeted testing strategy is used (i.e., $w_I \gg w_S$).
+To summarise, returning test results more rapidly tends to be favourable (i.e., reduces $\Rnum$) when (i) \emph{Confirmed-positive} individuals, lower their contact much more than individuals who are waiting for their test results (i.e., $\theta\_c \gg \theta\_w$), (ii) the test is highly sensitive (i.e., $p_I$ is close to 1), and (iii) the targeted testing strategy is used (i.e., $w_I \gg w_S$).
 
@@
 F =& \beta/N_0 \left[ \begin {array}{cccc} 
-S\_u&\eta\_w\,S\_u&\eta\_w\,S\_u&\eta\_c\,S\_u\\
-S\_n&\eta\_w\,S\_n&\eta\_w\,S\_n&\eta\_c\,S\_n\\ 
+S\_u&(1-\theta\_w)\,S\_u&(1-\theta\_w)\,S\_u&(1-\theta\_c)\,S\_u\\
+S\_n&(1-\theta\_w)\,S\_n&(1-\theta\_w)\,S\_n&(1-\theta\_c)\,S\_n\\ 
 0&0&0&0\\
@@
 \label{eq:R0appr}
-\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(\eta\_w-1)(\gamma w_S+\omega w_I) + (\eta\_c -1)p_I w_I \omega^2 \Big) + \mathcal{O}(\rho^2).
+\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(-\theta\_w)(\gamma w_S+\omega w_I) + (-\theta\_c)p_I w_I \omega^2 \Big) + \mathcal{O}(\rho^2).
 \end{equation}
@@
 \end{equation}
-where $a=(\eta\_w-1)w_I-(\eta\_c-1)p_I w_I = ((s-p_I)\eta\_c + (p_I-1)) w_I$, $b=2(\eta\_w-1)\gamma w_S$ and $c=(\eta\_w-1)\gamma^2 w_S$.
-Given that $0 \leq \eta\_c\leq \eta\_w \leq 1 $, one can easily derive $b\leq 0$ and $c \leq 0$.
+where $a=(-\theta\_w)w_I-(-\theta\_c)p_I w_I = ((s-p_I)(1-\theta\_c) + (p_I-1)) w_I$, $b=2(-\theta\_w)\gamma w_S$ and $c=(-\theta\_w)\gamma^2 w_S$.
+Given that $0 \leq \theta\_w \leq \theta\_c \leq 1 $, one can easily derive $b\leq 0$ and $c \leq 0$.
 
-Note that in general, the necessary and sufficient condition for $a \geq 0$ is $(s-p_I) \eta\_c \geq (1-p_I)$, where $s=\frac{\eta\_w}{\eta\_c} \geq 1$.
+Note that in general, the necessary and sufficient condition for $a \geq 0$ is $(s-p_I) (1-\theta\_c) \geq (1-p_I)$, where $s=\frac{(1-\theta\_w)}{(1-\theta\_c)} \geq 1$.
 
-In the case of ``perfect isolation'', i.e., when $\eta\_w=0$ and consequently $\eta\_c=0$, it is straightforward to see that $a \leq 0$, $b<0$ and $c<0$.
+In the case of ``perfect isolation'', i.e., when $\theta\_w=1$ and consequently $\theta\_c=1$, it is straightforward to see that $a \leq 0$, $b<0$ and $c<0$.
 Thus, $\partial{\Rnum}/\partial{\omega} \geq 0$.
@@
 \begin{equation}\label{linearization}
-\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(\eta\_w-1)(\gamma w_S+\omega w_I) + (\eta\_c -1)p_I w_I \omega^2 \Big). 
+\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(-\theta\_w)(\gamma w_S+\omega w_I) + (-\theta\_c)p_I w_I \omega^2 \Big). 
 \end{equation}
 So when $\rho \approx 0$ we have $$\partial{\Rnum}/\partial{\omega} \approx  \frac{-\beta \rho}{\gamma w_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c),$$
-where $a=(\eta\_w-1)w_I-(\eta\_c-1)p_I w_I$, $b=2(\eta\_w-1)\gamma w_S$ and $c=(\eta\_w-1)\gamma^2 w_S$.
+where $a=(-\theta\_w)w_I-(-\theta\_c)p_I w_I$, $b=2(-\theta\_w)\gamma w_S$ and $c=(-\theta\_w)\gamma^2 w_S$.
 
@@
 Note that the zeroes are real if and only if $Q < 1$.
-Note that have $\eta\_c < \eta\_w$, so if $P_i \approx 1$, we will have $Q < 0 < 1$.
+Note that have $\theta\_c > \theta\_w$, so if $p_I \approx 1$, we will have $Q < 0 < 1$.
 Thus, if we assume near-perfect test sensitivity, $\omega_1$ and $\omega_2$ will be real.
@@
 Suppose that $\rho_1 > \rho_2$.
-Recall that the linearization of $\Rnum$ around $\rho \approx 0$ is given by $$\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(\eta\_w-1)(\gamma w_S+\omega w_I) + (\eta\_c -1)p_I w_I \omega^2 \Big).$$
+Recall that the linearization of $\Rnum$ around $\rho \approx 0$ is given by $$\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(-\theta\_w)(\gamma w_S+\omega w_I) + (-\theta\_c)p_I w_I \omega^2 \Big).$$
 
@@
 \begin{align}\label{eq:rho1vsrho2}
-    &\rho_1\left(\gamma(\eta\_w-1)(\gamma w_S + \omega w_I) + (\eta\_c-1)p_{I,1}w_I\omega^2\right) - \rho_2\left(\gamma(\eta\_w-1)(\gamma w_S + \omega w_I) + (\eta\_c-1)p_{I, 2}w_I\omega^2\right) > 0 \nonumber \\
+    &\rho_1\left(\gamma(-\theta\_w)(\gamma w_S + \omega w_I) + (-\theta\_c)p_{I,1}w_I\omega^2\right) - \rho_2\left(\gamma(-\theta\_w)(\gamma w_S + \omega w_I) + (-\theta\_c)p_{I, 2}w_I\omega^2\right) > 0 \nonumber \\
     &\vdots \nonumber \\
-    &\frac{\rho_2P_{i, 2}-\rho_1P_{i, 1} }{\rho_1-\rho_2} > \frac{1-\eta\_w}{1-\eta\_c}\cdot \frac{\gamma(\gamma w_S + \omega w_I)}{\omega^2 w_I}
+    &\frac{\rho_2P_{i, 2}-\rho_1P_{i, 1} }{\rho_1-\rho_2} > \frac{\theta\_w}{\theta\_c}\cdot \frac{\gamma(\gamma w_S + \omega w_I)}{\omega^2 w_I}
 \end{align}

### 2021-03-25 SIR_manuscript.tex
Fiddling and thinking

--- 2021-03-25 SIR_manuscript.tex
@@
 
-The analytical calculation of the next generation matrix and the derivation of $\Rnum$ was carried out in Maple \citep{maple14} by using simple linear Algebra package. 
-  We used R \citep{r} for the simulation part which included using the explicit expression of $\Rnum$ \ref{R0} as a function of the underlying parameters, specifying the parameters a realistic range, calculating the corresponding $\Rnum$ and plotting the contours of $\Rnum$ \eqref{R0}.
-The related result is presented in Figure \ref{pan}, where  panel \eqref{p.a} represents the random testing, and panel \eqref{p.b} is representing non-random testing.
+The analytical calculation of the next-generation matrix and the derivation of $\Rnum$ was carried out in Maple \citep{maple14}. 
+We used R \citep{r} for simulations and for plotting $\Rnum$ as a function of the underlying parameters (using \ref{R0}).
+Results are shown in Figure \ref{pan}, where  panel \eqref{p.a} represents the random testing, and panel \eqref{p.b} is representing non-random testing.
 The parameter values or ranges are presented in Table~\ref{tab:params}.
@@
 
-\david{Details such as the fact that you used ggplot are not important to state in the main text.
-On the other hand, there isn't a clear narrative here.
-The reader should feel that a story is being told.}\ali{addressed?, feedback is welcomed}
-
 % %%%%%%%
@@
 The explicit formula for the basic reproduction number, $\Rnum$ \eqref{R0}, provides an opportunity to study the influence of changes in the underlying parameters on the critical index of epidemic dynamics.
-We are interested in understanding the effect of parameters that can be realistically controlled by changing in isolation, \percap testing intensity and test resulting, i.e., $\theta\_c$ and $\theta\_w$, $\rho$ and $\omega$, respectively.
-The following Proposition is the direct result of taking partial derivative of $\Rnum$ \eqref{R0} with respect to selected parameters.
-Note that we used Taylor approximation of $\Rnum$ at $\rho=0$ due to the complexity of the expression and thus inconclusive.\david{in that case it would be good to establish numerically that the relationships given in the proposition are valid more generally, at least for reasonable parameter ranges}\ali{will do the simulation for testing intensity $>>1$}
+We are interested in exploring the effect of parameters that could be manipulated by public-health policy, including the isolation parameters ($\theta\_c$ and $\theta\_w$) testing intensity ($\rho$) and test-return rate ($\omega$).
+
+Examining our formula for $\Rnum$ \eqref{R0} gives the following results. 
 See the appendix for details.
 
-\david{More importantly, you really need to state in words what this proposition is saying.  The reader has to check the param table and think it through themselves.}
-\begin{proposition}
-\label{prop1}
-For model \eqref{model}, and by using the expression of the basic reproduction number $\Rnum$ \eqref{R0} ,
 \begin{enumerate}
@@
 \end{enumerate}
-\end{proposition}
 

### 2021-03-31 SIR_manuscript.tex
change the brake vector to be automatic, plot panels with rho from 0 to 0.5

@@
 
+\subsection{Simulation with $\rho$ varies from 0 to 0.5 }
+
+\begin{figure}[h!]
+\centering
+\begin{subfigure}[t]{.45\textwidth}
+\centering
+\includegraphics[width=\linewidth]{codes/R0contour_random_rho.5.pdf}
+\caption{}\label{p.a}
+\end{subfigure}
+%
+\begin{subfigure}[t]{.45\textwidth}
+\centering
+\includegraphics[width=\linewidth]{codes/R0contour_TTI_rho.5.pdf}
+\caption{}\label{p.b}
+\end{subfigure}
+\caption{
+{\bf A comparison of the behaviour of the basic reproduction number, $\Rnum$, between random versus targeted testing strategies at different levels of testing and isolation.}
+We numerically evaluate $\Delta$ \eqref{R0}, reflecting the reduction of $\Rnum$ with respect to testing and isolation. We use the following parameters (listed in Table \ref{tab:params}):
+$N_0=1 \times 10^6$, $\omega \in [0.5,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.5]$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.39$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
+Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
+The black solid line in each panel represents the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ corresponding to the threshold of $\Rnum=1$.
+\bmb{we still have some collisions in the y-axis tick labels}
+}
+\label{pan}
+\end{figure}
+
 \end{document}

### 2021-03-31 codes/SIRfunctions.R

@@ -147,17 +147,17 @@ R02 <- function(state=state_dfe,params){
 
 # make grid and calc R0 for plotting
 eval_R0 <- function(state=state_dfe,params){
-  # input the params and their range, this function makes a grid dataframe, calls R0 and outputs the csv file
+  ## input the params and their range, this function makes a grid dataframe, calls R0 and outputs the csv file
+  ## the params are defaults, specify the range of favorite params here.
   unpack(as.list(c(state,params)))
   tol <- 1e-10 ## to resolve the issue of very small numbers 
   # specify the ranges, FIXME, not to be hard coded!
-  beta  <- params[["beta"]]  ## set beta high enough to allow R0>1 in worst case
   eta_w <- seq(0,1, length.out=n_out)
-  eta_c <- seq(0,1, length.out=n_out) #0.001 to 0.5 
-  rho <- seq(0,0.01, length.out=n_in)
-  omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, was 0.011
+  eta_c <- seq(0,1, length.out=n_out)  
+  rho <- seq(0,0.01, length.out=n_in) ## 0,0.01
+  omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, was 0.1 to 2
   
-  df1 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
+  df1 <- expand.grid(N0=params[["N0"]],beta=params[["beta"]],gamma=params[["gamma"]],omega=omega,rho=rho,
                      W_S=W_S,W_I=params[["W_I"]],W_R=params[["W_R"]],
                      p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
                      eta_w=eta_w,eta_c=eta_c)

### 2021-03-31 codes/params.R

@@ -14,7 +14,7 @@ omega=0.25, #test returning rate
 rho=0.01, #testing rate percapita (also used 0.8, 1/3)
 W_S=1, W_I=1, W_R=1, #Testing relative weights
 p_S=0, p_I=1, p_R=0.5, #test specificity, i.e., prob of waiting for being tested positive
-eta_w=0.02, eta_c=0.01, #isolation parameter (eta=0 is the perfect isolation) 
+eta_w=0.02, eta_c=0.01, #isolation parameter (eta=0 is the perfect isolation)
 s=2 # isolation ratio parameter, i.e., s=eta_w/eta_c
             )
             
@@ -33,3 +33,4 @@ state_dfe <- c(S_u=Su_dfe(params), S_n=Sn_dfe(params),
                 N=0,P=0)
 
 saveVars(params, state_init, state_dfe)
+

### 2021-03-31 codes/sir_plot.R

@@ -47,12 +47,18 @@ label_special <- function (labels, multi_line=FALSE, sep = "== ") {
 }
 
 # verify the break range in brks_vec (FIXME)
-mn <- min(df_random$Delta,df_targeted$Delta, na.rm = T)
-mx <- max(df_random$Delta,df_targeted$Delta, na.rm = T) ## take mn and mx and set the brks_vec, smarter way?
+ # mn <- min(df_random$Delta,df_targeted$Delta, na.rm = T)
+ # mx <- max(df_random$Delta,df_targeted$Delta, na.rm = T) ## take mn and mx and set the brks_vec, smarter way?
 
+n_contour <- 8 ## number of desired contours or bins
+bins <- cut(sort(unique(c(df_random$Delta,df_targeted$Delta))), 
+            n_contour)
+brks <- levels(bins)
+brks_vec <- unique(as.numeric(unlist(lapply(strsplit(brks, ","), function(x) gsub("\\(|]", "", x)))))
+brks_vec <- sort(ifelse(brks_vec[]< 0,0,brks_vec[]), decreasing = TRUE) ## replace neg with 0 and reorder
 # show_contours_1, width=24,height=24,message=FALSE,warning=FALSE}
-brks_vec <- seq(.165,0,by=-0.03) # Break vector for unifying the legends in Random and TTI testing cases. 
-
+# brks_vec <- seq(1,0,by=-0.2) # Break vector for unifying the legends in Random and TTI testing cases. 
+# seq(.165,0,by=-0.03)
 df_temp <- df_random
 # df_temp <- df_targeted
 p1 <- (ggplot(df_temp,aes(x=1/omega,y=rho,z=Delta))
@@ -63,6 +69,7 @@ p1 <- (ggplot(df_temp,aes(x=1/omega,y=rho,z=Delta))
 ## There might be a solution for the tick label collision with expand_limits(), but that's hard to do across facets
 ## see https://stackoverflow.com/questions/41575045/avoiding-axis-tick-label-collision-in-faceted-ggplots
 p1_temp <- (p1
+            # + geom_contour_filled(breaks=brks)
             + geom_contour_filled(breaks=brks_vec)
             + geom_contour(breaks=threshold,alpha=0.5,colour="black")
             + facet_grid(theta_w~theta_c, labeller=label_special) 

### 2021-04-01 SIR_manuscript.tex
proving that R0 decreases wrt rho without using Tylor appr

@@
 \label{eq:abc}
-\Delta=& (A~ \theta\_w + B~ \theta\_c)~ C\\
-A=& \gamma \big( (\gamma \omega+ \omega p_I F_I+\gamma^2) x_n^* + (\omega+\gamma) F_I \big)\\
+\Delta=& (A~ \theta\_w + B~ \theta\_c) / C\\
+A=& \gamma \Big( \big(\gamma(\omega+\gamma) + \omega p_I F_I \big) x_n^* + (\omega+\gamma) F_I \Big)\\
 B=& \omega p_I F_I (\gamma x_u^*+\omega) \\ 
-C=& \frac{1}{(\omega+\gamma) \big(\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I) \big)}.
+C=& (\omega+\gamma) \Big(\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I) \Big).
 \end{aligned}
@@
 \item \label{p1:rho} higher rate of testing reduces $\Rnum$;
-$$\partial{\Rnum}/\partial{\rho} \leq 0 \text{~when~} \rho \approx 0.$$
+$$\partial{\Rnum}/\partial{\rho} \leq 0.$$
 \item \label{p1:omega} The rate of test return may result in the disease becomes extinct or not;
-$$\partial{\Rnum}/\partial{\omega} \text{~can be positive or negative when~} \rho \approx 0.$$
+$$\partial{\Rnum}/\partial{\omega} \text{~can be positive or negative}.$$
 \end{enumerate}
@@
 % %%%%%%%
-\subsection{Taylor Approximation of $\Rnum$ at $\rho=0$ }
-The basic reproduction number, $\Rnum$, close to $\rho=0$ can be approximated linearly in $\rho$ by using Taylor approximation. it follows
-\begin{equation}
-\label{eq:R0appr}
-\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(-\theta\_w)(\gamma w_S+\omega w_I) + (-\theta\_c)p_I w_I \omega^2 \Big) + \mathcal{O}(\rho^2).
-\end{equation}
+\subsection{More testing rate reduces $\Rnum$}
 
-Also, to analyze the influence of $\omega$ on $\Rnum$, the approximation \eqref{eq:R0appr} was used. Then it is straight forward to have
-\begin{equation}
-\label{Rom}
-\partial{\Rnum}/\partial{\omega}=  \frac{-\beta \rho}{\gamma w_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c),
-\end{equation}
-where $a=(-\theta\_w)w_I-(-\theta\_c)p_I w_I = ((s-p_I)(1-\theta\_c) + (p_I-1)) w_I$, $b=2(-\theta\_w)\gamma w_S$ and $c=(-\theta\_w)\gamma^2 w_S$.
-Given that $0 \leq \theta\_w \leq \theta\_c \leq 1 $, one can easily derive $b\leq 0$ and $c \leq 0$.
-
-Note that in general, the necessary and sufficient condition for $a \geq 0$ is $(s-p_I) (1-\theta\_c) \geq (1-p_I)$, where $s=\frac{(1-\theta\_w)}{(1-\theta\_c)} \geq 1$.
-
-In the case of ``perfect isolation'', i.e., when $\theta\_w=1$ and consequently $\theta\_c=1$, it is straightforward to see that $a \leq 0$, $b<0$ and $c<0$.
-Thus, $\partial{\Rnum}/\partial{\omega} \geq 0$.
-As an example, in the case of a very accurate testing regime,  i.e., $P_I=1$, $a \geq 0$ is achieved.
-If $a\geq 0$, the quadratic expression in \eqref{Rom}, has Real roots.
-Assuming that $\omega_1<0$ and $\omega_2>0$ be the roots of the quadratic expression in $\partial{\Rnum}/\partial{\omega}$.
-Thus, $\partial{\Rnum}/\partial{\omega}>0$ for $0<\omega<\omega_2$ and  $\partial{\Rnum}/\partial{\omega}<0$ for $\omega>\omega_2$.
+ The expression of the basic reproduction number $\Rnum$, given in Eqs.
+\eqref{R0}, makes it straight  forward to conclude $\frac{\partial \Delta}{\partial \rho} > 0$ and thus $\frac{\partial \Rnum}{\partial \rho} < 0$.
+Given the fractional form of $\Delta$, it is easy to show that $\partial A/\partial \rho >0$, $\partial B/\partial \rho >0$ and $\partial C/\partial \rho <0$. Note that we use the fact that
+$\partial F_I/\partial \rho >0$, $\partial S\_n^*/\partial \rho >0$ and $\partial S\_u^*/\partial \rho <0$.
 

### 2021-04-01 SIR_manuscript.tex
Meeting with Ali

@@
 \newcommand{\percap}{\emph{per capita}\xspace}
-\newcommand{\Rnum}{\mathcal{R}_0}
+\newcommand{\Rnum}{\ensuremath{\mathcal{R}_0}}
 \newcommand{\covid}{COVID-19\xspace}
@@
 \end{equation}
-For simplification of the analysis, we adopt the following change $x\_u^* = S\_u^*/N_0$ and $x\_n^* = S\_n^*/N_0$.
+For brevity, we write $x\_u^* = S\_u^*/N_0$ and $x\_n^* = S\_n^*/N_0$.
+We can then write $\hat F_I = (\omega\rho/(\omega-\rho))w_I/w_S$ for the \percap testing rate of infectious individuals near the disease-free equilibrium.
 
@@
 \begin{enumerate}
-\item \label{p1:eta} increasing the relative probabilities of transmission for tested individuals, increases $\Rnum$;
+\item \label{p1:eta} increasing the isolation parameters for tested and confirmed individuals decreases \Rnum;
 $$\partial{\Rnum}/\partial{\theta\_c} \leq 0 \text{~and~} \partial{\Rnum}/\partial{\theta\_w} \leq 0.$$ 
-\item \label{p1:rho} higher rate of testing reduces $\Rnum$;
+\item \label{p1:rho} a higher rate of testing decreases $\Rnum$;
 $$\partial{\Rnum}/\partial{\rho} \leq 0.$$
-\item \label{p1:omega} The rate of test return may result in the disease becomes extinct or not;
+\item \label{p1:omega} A higher rate of of test return may increase or decrease \Rnum.
 $$\partial{\Rnum}/\partial{\omega} \text{~can be positive or negative}.$$
 \end{enumerate}
-
-Given that the perfect isolation occurs when no one transmits while waiting for test results (i.e., $\theta\_c=\theta\_w = 1$), Prop.~\ref{p1:eta} means that lifting the isolation from awaiting group results in an increased $\Rnum$ and consequently a greater number of infected individuals.
-Prop.~\ref{p1:rho} indicates that increasing the \percap testing intensity, $\rho$, reduces $\Rnum$.
-Lastly, Prop.~\ref{p1:omega} means that speeding up the test reporting may or may not reduce $\Rnum$, it will depend on other parameters.
-It is straightforward to derive \ref{p1:rho} and \ref{p1:omega} from the Taylor approximation of $\Rnum$ at $\rho=0$.
-\david{can we gain some insight from the expressions for the partial derivatives in the proposition, as opposed to just their signs?}\ali{help}
 
@@
 {\bf A comparison of the behaviour of the basic reproduction number, $\Rnum$, between random versus targeted testing strategies at different levels of testing and isolation.}
-We numerically evaluate $\Delta$ \eqref{R0}, reflecting the reduction of $\Rnum$ with respect to testing and isolation. We use the following parameters (listed in Table \ref{tab:params}):
+We numerically evaluate $\Delta$ \eqref{R0}, reflecting the reduction of $\Rnum$ with respect to testing and isolation. We use the following parameters (see Table \ref{tab:params}):
 $N_0=1 \times 10^6$, $\omega \in [0.1,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.01]$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.39$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
@@
 
-Our numerical solutions in support of the analytical results in Proposition \ref{prop1} are presented in Figure \ref{pan}.
+Our numerical results are shown in Figure \ref{pan}.
 It can be inferred that when random testing strategy is applied, Figure \ref{p.a}, comparing to the targeted testing strategy, Figure \ref{p.b}, the critical \percap testing intensity of the corresponding panels are lower in non-random testing.

### 2021-04-04 SIR_manuscript.tex
half way in dR0/d rho >0

@@
 \newcommand{\covid}{COVID-19\xspace}
+\newcommand{\pro}[1][]{\ensuremath{\frac{\partial #1}{\partial \rho}}}
 
@@
 
- The expression of the basic reproduction number $\Rnum$, given in Eqs.
-\eqref{R0}, makes it straight  forward to conclude $\frac{\partial \Delta}{\partial \rho} > 0$ and thus $\frac{\partial \Rnum}{\partial \rho} < 0$.
-Given the fractional form of $\Delta$, it is easy to show that $\partial A/\partial \rho >0$, $\partial B/\partial \rho >0$ and $\partial C/\partial \rho <0$. Note that we use the fact that
-$\partial F_I/\partial \rho >0$, $\partial S\_n^*/\partial \rho >0$ and $\partial S\_u^*/\partial \rho <0$.
+This is to prove that $\pro \Delta > 0$ with $\Delta$ defined in Eq. \eqref{eq:abc}, and thus $\pro \Rnum < 0$, where $\Rnum$ is the basic reproduction number, given in Eqs. \eqref{R0}.
+We will show that (a) $\pro (A/C)>0$ and (b) $\pro (B/C)>0$.
+Note that we will use the following information: 
+\begin{enumerate}
+\item
+$\pro \hat F_I >0$ for  all $\rho$ and specifically the feasible range, $\rho \in [0,\omega]$.
+Here $\hat F_I = (\omega\rho/(\omega-\rho))w_I/w_S$ for the \percap testing rate of infectious individuals near the disease-free equilibrium,
+\item $\pro \hat F_I x\_u^* >0$ since $\hat F_I x\_u^* = \rho w_I/w_S$,
+\item $\pro \hat F_I x\_n^* >0$ since $\hat F_I x\_n^* = \frac{\rho^2}{\omega-\rho} w_I/w_S$, 
+\item $\pro x\_n^* >0$ and $\pro x\_u^* <0$ given the DFE expression in \ref{dfe},
+\item
+\begin{align}
+\pro A &= 
+\gamma \Big(\gamma(\omega+\gamma)\pro x\_n^* + \omega p_I \pro(\hat F_I x\_n^*)+(\omega+\gamma) \pro(\hat F_I)
+\Big),\\
+\pro B &= \gamma \omega p_I \pro(\hat F_I x\_u^*)+\omega^2 p_I \pro(\hat F_I),\\
+\pro C &=(\omega+\gamma)(\omega p_I+\gamma) \pro(\hat F_I).
+\end{align}
+\end{enumerate}
 
+(a) $\pro (A/C)>0$;
+
+\begin{align}
+& \pro (A/C) = \pro{A} \dot C - \pro{C} \dot A \\
+&= \gamma (\omega+\gamma)^2(\omega p_I+\gamma) 
+\Big( 
+[\gamma \pro x\_n^* + \frac{\omega p_I}{(\omega+\gamma)} \pro(\hat F_I x\_n^*) + \pro\hat F_I]
+[\gamma\frac{(\omega+\gamma)}{(\omega p_I+\gamma)}+\hat F_I] - \\
+& \pro(\hat F_I)
+[\gamma x\_n^* + \frac{\omega p_I}{(\omega+\gamma)}(\hat F_I x\_n^*) + \hat F_I] 
+\Big) 
+\end{align}
+
+Since $\pro(x\_n^*) >0$ and $\pro (\hat F_I x\_n^*)>0$,  it is enough to show that the coefficient of 
+$\pro (\hat F_I)$ is summed up to a positive value.
+$\pro (\hat F_I)$ coefficient is as follows
+$$\big( \frac{\omega p_I}{(\omega+\gamma)} x\_n^* +1 \big) 
+\big( \gamma\frac{(\omega+\gamma)}{(\omega p_I+\gamma)} + \hat F_I \big) 
+-
+\big( \gamma x\_n^* + \frac{\omega p_I}{(\omega+\gamma)}(\hat F_I x\_n^*) + \hat F_I 
+\big).
+$$
+After simplification, we use the fact that the following expression is positive
+$$(\frac{(\omega+\gamma)}{(\omega p_I+\gamma)}) (1+ \frac{\omega p_I x\_n^*}{(\omega+\gamma)}) -1. $$
+This is because $(\frac{(\omega+\gamma)}{(\omega p_I+\gamma)})>1$ and  $(1+ \frac{\omega p_I x\_n^*}{(\omega+\gamma)}) >1$.
+
+(b) $\pro (B/C)>0$;
 % %%%%%%%

### 2021-04-08 RhoProof.tex

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newcommand{\Rnum}{\mathcal{R}_0}
\usepackage{xcolor}
\begin{document}

We want to prove that $\Rnum$ is decreasing wrt $\rho$. Recall that:

\begin{align}
    \Rnum = \frac{\beta}{\gamma}(1-\Delta)
\end{align}

where

\begin{align}
\Delta =& (A~ \theta_w + B~ \theta_c) C 
\\
A=& \gamma \Big( \big(\gamma(\omega+\gamma) + \omega p_I F_I \big) x_n^* + (\omega+\gamma) F_I \Big) \\
B=& \omega p_I F_I (\gamma x_u^*+\omega) \\ 
C=& (\omega+\gamma) \Big(\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I) \Big).
\end{align}

So, we want to prove that $\Delta$ is increasing wrt $\rho$. We will do so under model assumptions that $\rho < w$ and $\theta_w < \theta_c$. Consider the following lemmas.

\begin{lemma} we have:
\begin{itemize}
    \item $A\theta_w + B\theta_c = -U\rho^2 + V\rho$
    \item $C = \frac{1}{K\rho+W}$
\end{itemize} where:

\begin{align}
    U &= \Big((\theta_c-\theta_w)w\gamma p_IF_Ix^*_n\Big)\cdot \frac{1}{\rho^2} \geq 0\\
    V &= \Big(F_I\big(\gamma(w+\gamma)\theta_w+wp_I(w+\gamma)\theta_c\big) + x^*_n\gamma^2(w+\gamma)\theta_w\Big)\cdot \frac{1}{\rho} \geq 0 \\
    K &= F_I(w+\gamma)(\gamma+wp_I) \cdot \frac{1}{\rho} \geq 0 \\
    W &= \gamma(w+\gamma)^2 \geq 0
\end{align}
\begin{proof}
Note that $\rho$ appears only in $x^*_n$, $x^*_u$, and $F_I$, and is linear in all of them. The rest is brute force rewriting...
\end{proof}
\end{lemma}

\begin{lemma} $\dfrac{d}{d\rho} \, \dfrac{-U\rho^2+V\rho}{K\rho+W} > 0 \Longleftrightarrow V\rho > U\rho^2$
\begin{proof}
\begin{align}
    \dfrac{d}{d\rho} \, \dfrac{-U\rho^2+V\rho}{K\rho+W} &= (-U\rho+V)(K\rho+W) - K(U\rho^2+V\rho) \\
    &= (-U\rho+V)(K\rho+W) - K\rho(U\rho+V) \\
    &= (-U\rho+V)W
\end{align}

Since $W \geq 0$, we know that the final expression will be positive if and only if $V > U\rho$, which is equivalent to $V\rho > U\rho^2$.
\end{proof}
\end{lemma}

So, the lemma above tells us that we need to show that:
\begin{align}
    \Big({\color{red}F_I}\big(\gamma(w+\gamma)\theta_w+{\color{red}wp_I(w+\gamma)\theta_c}\big) + x^*_n\gamma^2(w+\gamma)\theta_w\Big) &> \Big((\theta_c-\theta_w)w\gamma p_IF_Ix^*_n\Big)
\end{align}

But since all terms on the LHS are nonnegative, it is sufficient to prove that

\begin{align}
    {\color{red}F_Iwp_I(w+\gamma)\theta_c} &> (\theta_c-\theta_w)w\gamma p_IF_Ix^*_n \\
    (w+\gamma)\theta_c &> \gamma(x^*_n)(\theta_c-\theta_w) \\
    \frac{(w+\gamma)}{\gamma} \cdot \frac{\theta_c}{\theta_c-\theta_w} &> \frac{\rho}{w} \\
\end{align}

but the final inequality is certainly true since the LHS is larger than 1, but the RHS is strictly less than 1 by the model assumption that $\rho < w$. Hence, $\Delta$ is increasing wrt to $\rho$, as required.
\end{document}

### 2021-04-08 SIR_manuscript.tex
simplify V^-1 to a user friendly one
more on d R0/d rho, remove old analysis
more Appendix

@@
 F =& \beta/N_0 \left[ \begin {array}{cccc} 
-S\_u&(1-\theta\_w)\,S\_u&(1-\theta\_w)\,S\_u&(1-\theta\_c)\,S\_u\\
-S\_n&(1-\theta\_w)\,S\_n&(1-\theta\_w)\,S\_n&(1-\theta\_c)\,S\_n\\ 
+S\_u^* & (1-\theta\_w)\,S\_u^* & (1-\theta\_w)\,S\_u^* & (1-\theta\_c)\,S\_u^*\\
+S\_n^* & (1-\theta\_w)\,S\_n^* & (1-\theta\_w)\,S\_n^* & (1-\theta\_c)\,S\_n^* \\ 
 0&0&0&0\\
 0&0&0&0
- \end {array} \right], \\
-  V =&
+ \end {array} \right] \\
+ =&\beta\left[ \begin {array}{c} x\_n^* \\ x\_u^* \\ 0 \\ 0 \end {array} \right]
+        \left[ \begin {array}{cccc} 1, \, 1-\theta\_w, \, 1-\theta\_w, \, 1-\theta\_c \end {array} \right], \text{and}\\  V=&
  \left[ \begin {array}{cccc}  
@@
 0&0&-\omega&\gamma
-\end {array} \right], \text{thus}\\
-V^{-1} =&
-\left[ \begin {array}{cccc}
-\frac {\omega+\gamma}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&\frac {\omega}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&0&0\\
-\noalign{\medskip}
-\frac{(1-p_I) F_I}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&
-\frac{F_I+\gamma}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&0&0\\
-\noalign{\medskip}
-\frac{p_I F_I}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&
-\frac{\omega p_I F_I}{(\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I)(\omega+\gamma)}& \frac{1}{\omega+\gamma}&0 \\
-\noalign{\medskip}
-\frac{\omega\,F_I\,p_I}{(\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I) \gamma}& 
-\frac{\omega^2\,F_I\,p_I}{(\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I)(\omega+\gamma) \gamma}&
-\frac{\omega}{(\omega+\gamma) \gamma}&
-\frac{1}{\gamma}
 \end {array} \right].
 \end{align}
+The matrix inverse of $V$ is 
+\begin{align}
+\label{vinv}
+V^{-1} =&
+\frac{1}{\gamma C}
+\left[ \begin {array}{cccc}
+\gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)&0&0\\ \noalign{\medskip}
+\gamma(\omega+\gamma)(1-p_I)F_I & \gamma(\omega+\gamma)(F_I+\gamma)&0&0 \\ \noalign{\medskip}
+\gamma(\omega+\gamma) p_I F_I & \gamma\omega p_I F_I & C \gamma/(\omega+\gamma) & 0 \\ \noalign{\medskip}
+\omega(\omega+\gamma) p_I F_I & \omega^2 p_I F_I & C \omega/(\omega+\gamma) & C
+\end {array} \right],
+\end{align}
+where $C= \big( \gamma(\omega+\gamma)+(\gamma+\omega p_I)F_I \big) (\omega+\gamma)$. Note that all the columns of matrix $V^{-1}$ summ up to $C$.
 
-The particular form of $F$ with two rows of zeros at the bottom, simplifies $G$ as 
+The particular form of $F$ with two rows of zeros at the bottom, simplifies $G$ as the following block matrix.
 \begin{equation}
@@
 0&0
-\end {array} \right], \text{ where } \\
-G_{11} = \frac{\beta}{\gamma}\ C
-\left[\begin {array}{cc}
-A\,S\_u & B\,S\_u\\
-A\,S\_n & B\,S\_n
-\end {array}\right],
+\end {array} \right],
 \end{equation}
-with expressions $A$, $B$ and $C$ are specified in \eqref{eq:abc}.
-The block matrix $G_{12}$ does not influence $\Rnum$ (defined as the spectral radius of $G$). All that matters here are the eigenvalues of $G_{11}$, which are 0 and $\Rnum$ \eqref{R0}.
-Also note that, all $S\_u$'s and $S\_n$'s are evaluated at the DFE \eqref{dfe} in matrix $F$ \eqref{FV} and the block matrix $G_{11}$ \eqref{mat:G}.
+where the 2 by 2 submatrix $G_{11}$ is 
+\begin{equation}
+\label{G11}
+G_{11} = \frac{\beta}{\gamma \, C} 
+\left[ \begin {array}{c}  x\_n^* \\ x\_u^* \\ 0 \\ 0  \end {array} \right]
+\left[ \begin {array}{cccc} 1, \, 1-\theta\_w, \, 1-\theta\_w, \, 1-\theta\_c \end {array} \right]
+\left[ \begin {array}{cc}
+\gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)\\ \noalign{\medskip}
+\gamma(\omega+\gamma)(1-p_I)F_I & \gamma(\omega+\gamma)(F_I+\gamma) \\ \noalign{\medskip}
+\gamma(\omega+\gamma) p_I F_I & \gamma\omega p_I F_I \\ \noalign{\medskip}
+\omega(\omega+\gamma) p_I F_I & \omega^2 p_I F_I
+\end {array} \right].
+\end{equation}
+
+It is notable that matrix $F$ \eqref{FV} and consequently $G_{11}$ have rank one.
+That is $G_{11}$ has only one non-zero eigenvalue which we refer to it the basic reproduction number $\Rnum$ (defined as the spectral radius of $G$).
+Also the block matrix $G_{12}$ does not influence $\Rnum$ and all $S\_u$'s and $S\_n$'s are evaluated at the DFE \eqref{dfe} in matrix $F$ \eqref{FV} and the block matrix $G_{11}$ \eqref{mat:G}.
+
+$\Rnum$ has a complicated expression and depend on the type of analysis, the expression can be simplifed and presented accordingly.
+For the purpose of simplicity we present $\Rnum$ in the manuscript in terms of expressions $A$, $B$ and $C$, specified in \eqref{eq:abc}.
+This is intended to show that $\Rnum$ decreases as the strength of isolation, shown by $\theta\_w$ and $\theta\_c$, increases.
+
+The hard thing to prove is to show the reduction of $\Rnum$ with respect to testing intensity, $\rho$, and the speed of the test return, $\omega$, for the feasible ranges of the parameters, i.e., $0<\omega$, $0 \leq \rho<\omega$, and $0 \leq \theta\_w \leq \theta\_c \leq 1$.
 

### 2021-04-08 SIR_manuscript.tex
old analysis of diff R0 wrt rho was wrong and ditched

@@
 
-The hard thing to prove is to show the reduction of $\Rnum$ with respect to testing intensity, $\rho$, and the speed of the test return, $\omega$, for the feasible ranges of the parameters, i.e., $0<\omega$, $0 \leq \rho<\omega$, and $0 \leq \theta\_w \leq \theta\_c \leq 1$.
+It remains hard to show the reduction of $\Rnum$ with respect to \percap testing intensity, $\rho$, and the speed of the test return, $\omega$, for the feasible ranges of the parameters, i.e., $0<\omega$, $0 \leq \rho<\omega$, and $0 \leq \theta\_w \leq \theta\_c \leq 1$.
+One easy way to do such an analysis is based on the fact that $\rho$ is \percap rate so for a large population size, $N_0 \gg 1$, $\rho$ is very small or close to 0, comparing to $N_0$.
+This provides a base to linearly approximate $\Rnum$ when $\rho$ is close to 0, and use this approximation to analyze the behaviour of $\Rnum$ with respect to $\omega$ (see section \ref{taylor}).
+In the next section we provide an equivalent representation of $\Rnum$ for our model to provide to follow proving that more testing intensity decreases $\Rnum$ for a general rane of parameters.
 
 % %%%%%%%
-\subsection{On Testing Rate and Numerical Singularity}
+\subsection{More testing intensity decreases $\Rnum$}
 
-In this work, we didn't do any numerical solutions for the trajectories in our analysis. However, if one tries to do so there would be a singularity issue to deal with. 
-Specifically, the numerical singularity issue with the chosen $\sigma$ \eqref{sigma} is that the population in $S$ compartments appeared to blow up when the DFE is achieved.
-This is once the only untested people are susceptibles, the FOI will become $\Lambda=0$, testing rate $F_s=\rho N_0/S\_u$.
-Thus, the first equation of the model \eqref{model} will become
-$d S\_u/dt = - \rho N_0 + \omega S\_n$. Thus changes in $S\_u$ will be no longer dependent on $S\_u$ with a linear rate of leaving the $S\_u$ compartment.
-IN fact the testing rate, $\sigma$, should be formulated such that people from the untested compartments will not be tested if they are not there.
-One way to fix this issue, is to consider a maximum testing rate, $\tau$ (1/day).
-In general, we want to test at a rate of $\rho$ across the whole population.
-This won't always be possible, so we impose a maximum rate of $\tau$ per testable person and redefine $\sigma = \frac{\tau \rho N_0}{\tau W + \rho N_0}$, with the assumption that $\tau \gg \rho$.
-This alteration in $\sigma$, does not change any results related to $\Rnum$, thus we only impose it in the simulation of the epidemic dynamic.
+This is to prove that $\pro \Delta > 0$ with $\Delta$ defined in Eq. \eqref{eq:abc}, and thus $\pro \Rnum < 0$, where $\Rnum$ is the basic reproduction number, given in Eqs. \eqref{R0}.
 
 % %%%%%%%
-\subsection{More testing rate reduces $\Rnum$}
-
-This is to prove that $\pro \Delta > 0$ with $\Delta$ defined in Eq. \eqref{eq:abc}, and thus $\pro \Rnum < 0$, where $\Rnum$ is the basic reproduction number, given in Eqs. \eqref{R0}.
-We will show that (a) $\pro (A/C)>0$ and (b) $\pro (B/C)>0$.
-Note that we will use the following information: 
-\begin{enumerate}
-\item
-$\pro \hat F_I >0$ for  all $\rho$ and specifically the feasible range, $\rho \in [0,\omega]$.
-Here $\hat F_I = (\omega\rho/(\omega-\rho))w_I/w_S$ for the \percap testing rate of infectious individuals near the disease-free equilibrium,
-\item $\pro \hat F_I x\_u^* >0$ since $\hat F_I x\_u^* = \rho w_I/w_S$,
-\item $\pro \hat F_I x\_n^* >0$ since $\hat F_I x\_n^* = \frac{\rho^2}{\omega-\rho} w_I/w_S$, 
-\item $\pro x\_n^* >0$ and $\pro x\_u^* <0$ given the DFE expression in \ref{dfe},
-\item
-\begin{align}
-\pro A &= 
-\gamma \Big(\gamma(\omega+\gamma)\pro x\_n^* + \omega p_I \pro(\hat F_I x\_n^*)+(\omega+\gamma) \pro(\hat F_I)
-\Big),\\
-\pro B &= \gamma \omega p_I \pro(\hat F_I x\_u^*)+\omega^2 p_I \pro(\hat F_I),\\
-\pro C &=(\omega+\gamma)(\omega p_I+\gamma) \pro(\hat F_I).
-\end{align}
-\end{enumerate}
-
-(a) $\pro (A/C)>0$;
-
-\begin{align}
-& \pro (A/C) = \pro{A} \dot C - \pro{C} \dot A \\
-&= \gamma (\omega+\gamma)^2(\omega p_I+\gamma) 
-\Big( 
-[\gamma \pro x\_n^* + \frac{\omega p_I}{(\omega+\gamma)} \pro(\hat F_I x\_n^*) + \pro\hat F_I]
-[\gamma\frac{(\omega+\gamma)}{(\omega p_I+\gamma)}+\hat F_I] - \\
-& \pro(\hat F_I)
-[\gamma x\_n^* + \frac{\omega p_I}{(\omega+\gamma)}(\hat F_I x\_n^*) + \hat F_I] 
-\Big) 
-\end{align}
-
-Since $\pro(x\_n^*) >0$ and $\pro (\hat F_I x\_n^*)>0$,  it is enough to show that the coefficient of 
-$\pro (\hat F_I)$ is summed up to a positive value.
-$\pro (\hat F_I)$ coefficient is as follows
-$$\big( \frac{\omega p_I}{(\omega+\gamma)} x\_n^* +1 \big) 
-\big( \gamma\frac{(\omega+\gamma)}{(\omega p_I+\gamma)} + \hat F_I \big) 
--
-\big( \gamma x\_n^* + \frac{\omega p_I}{(\omega+\gamma)}(\hat F_I x\_n^*) + \hat F_I 
-\big).
-$$
-After simplification, we use the fact that the following expression is positive
-$$(\frac{(\omega+\gamma)}{(\omega p_I+\gamma)}) (1+ \frac{\omega p_I x\_n^*}{(\omega+\gamma)}) -1. $$
-This is because $(\frac{(\omega+\gamma)}{(\omega p_I+\gamma)})>1$ and  $(1+ \frac{\omega p_I x\_n^*}{(\omega+\gamma)}) >1$.
-
-(b) $\pro (B/C)>0$;
-% %%%%%%%
-\subsection{rate of returning tests} 
+\subsection{rate of returning tests} \label{taylor}
 \ali{needs editig}
@@
 
+% %%%%%%%
+\subsection{On Testing Rate and Numerical Singularity}
+
+In this work, we didn't do any numerical solutions for the trajectories in our analysis. However, if one tries to do so there would be a singularity issue to deal with. 
+Specifically, the numerical singularity issue with the chosen $\sigma$ \eqref{sigma} is that the population in $S$ compartments appeared to blow up when the DFE is achieved.
+This is once the only untested people are susceptibles, the FOI will become $\Lambda=0$, testing rate $F_s=\rho N_0/S\_u$.
+Thus, the first equation of the model \eqref{model} will become
+$d S\_u/dt = - \rho N_0 + \omega S\_n$. Thus changes in $S\_u$ will be no longer dependent on $S\_u$ with a linear rate of leaving the $S\_u$ compartment.
+IN fact the testing rate, $\sigma$, should be formulated such that people from the untested compartments will not be tested if they are not there.
+One way to fix this issue, is to consider a maximum testing rate, $\tau$ (1/day).
+In general, we want to test at a rate of $\rho$ across the whole population.
+This won't always be possible, so we impose a maximum rate of $\tau$ per testable person and redefine $\sigma = \frac{\tau \rho N_0}{\tau W + \rho N_0}$, with the assumption that $\tau \gg \rho$.
+This alteration in $\sigma$, does not change any results related to $\Rnum$, thus we only impose it in the simulation of the epidemic dynamic.
+
+% %%%%%%%
 \subsection{Expensive vs.\ cheap tests}

### 2021-04-08 SIR_manuscript.tex
new representation of Delta, diff Delta wrt rho is quadratic

@@
 G_{11} = \frac{\beta}{\gamma \, C} 
-\left[ \begin {array}{c}  x\_n^* \\ x\_u^* \\ 0 \\ 0  \end {array} \right]
+\left[ \begin {array}{c}  x\_n^* \\ x\_u^*  \end {array} \right]
 \left[ \begin {array}{cccc} 1, \, 1-\theta\_w, \, 1-\theta\_w, \, 1-\theta\_c \end {array} \right]
@@
 
-It remains hard to show the reduction of $\Rnum$ with respect to \percap testing intensity, $\rho$, and the speed of the test return, $\omega$, for the feasible ranges of the parameters, i.e., $0<\omega$, $0 \leq \rho<\omega$, and $0 \leq \theta\_w \leq \theta\_c \leq 1$.
+It remains hard to show the reduction of $\Rnum$ with respect to \percap testing intensity, $\rho$, and the speed of the test return, $\omega$, for the feasible ranges of the parameters, i.e., $\omega>0$, $0 \leq \rho<\omega$, $0 \leq \theta\_w \leq \theta\_c \leq 1$ and $\frac{w_I}{w_S}\geq 1$.
 One easy way to do such an analysis is based on the fact that $\rho$ is \percap rate so for a large population size, $N_0 \gg 1$, $\rho$ is very small or close to 0, comparing to $N_0$.
@@
 This is to prove that $\pro \Delta > 0$ with $\Delta$ defined in Eq. \eqref{eq:abc}, and thus $\pro \Rnum < 0$, where $\Rnum$ is the basic reproduction number, given in Eqs. \eqref{R0}.
+
+Another representation of matrix $G_{11}$ in \eqref{G11} is as follows.
+\begin{equation}
+\label{G112}
+G_{11} = \frac{\beta}{\gamma \, C} 
+\left[ \begin {array}{c}  x\_n^* \\ x\_u^*  \end {array} \right]
+\left[ \begin {array}{cccc} 
+C-C_1, C-C_1-C_2\end {array} \right],
+\end{equation}
+where $C$ is in Eq. \eqref{eq:abc}, and $C_1$ and $C_2$ are 
+\begin{align}
+\label{eq:C12}
+C =& (\omega+\gamma)(\gamma(\omega+\gamma)+(\omega\,p_I+\gamma)\,F_I)\\
+C_1 =& (\omega+\gamma)(\theta\_w \, \gamma+\theta\_c\, \omega\, p_I) F_I,\\
+C_2 =& (\omega+\gamma)\gamma^2\,\theta\_w+(\theta\_w-\theta\_c)\gamma\,p_I\,F_I.
+\end{align}
+$\Rnum$ is in the same form as in Eq. \eqref{R0}  $$\Rnum= \frac{\beta}{\gamma} (1-\Delta),$$
+with equivalent but simplified $\Delta$ as
+\begin{equation}
+\label{eq:del2}
+\Delta= \frac{C_1+C_2\,x\_n^*}{C}.
+\end{equation}
+The behavior of $\Rnum$ with respect to $\rho$ is determine by $\pro \Delta$
+\begin{align}
+\label{eq:dd2dr}
+\pro \Delta=& \frac{1}{d} (a\,\rho^2+b\,\rho+c), \text{where}\\
+a=& ((\omega+\gamma)\gamma -\omega \frac{w_I}{w_S} (\omega\,p_I+\gamma)) D_1,\\
+b=& -2\gamma\omega(\omega+\gamma) D_1,\\
+c=& \omega^2 (\omega+\gamma)^2 D_2,
+\end{align}
+with scaler $d>0$ (not effecting the sign of $\pro \Delta$) and $D_1$ and $D_2$ as 
+\begin{align}
+\label{eq:D1D2}
+D_1=& (\omega+\gamma)\gamma\theta\_w+(\theta\_c-\theta\_w)\omega^2\,p_I\,\frac{w_I}{w_S}, \\
+D_2=& (\omega\,\frac{w_I}{w_S}+\gamma)\gamma\theta\_w+\omega^2\,p_I\,\frac{w_I}{w_S}\,\theta\_c.
+\end{align}
 

### 2021-04-08 SIR_manuscript.tex
fix the number page of Appendix
minor notes

@@
 \setcounter{table}{0}
-\setcounter{page}{1}
+% \setcounter{page}{1}
 \makeatletter
@@
 \end{align}
+Note that to determine the sign pattern of $\pro \Delta$ given the quadratic form of it in Eq.
+\eqref{eq:dd2dr}, it is important to know the sign of the coefficients of the polynomial, $a$, $b$ and $c$.
+Since $D_1\geq 0$ and $D_2\geq 0$, $b\leq 0$ and $c\geq 0$.
+Sign of $a$ depends on the interplay between the involved parameters (\todo{still working on}).
 

### 2021-04-09 SIR_manuscript.tex
more details on Delta_new was added, the corresponding Maple code has updated

--- 2021-04-08 SIR_manuscript.tex
@@
 b=& -2\gamma\omega(\omega+\gamma) D_1,\\
-c=& \omega^2 (\omega+\gamma)^2 D_2,
+c=& \omega^2 (\omega+\gamma)^2 D_2, \\
+d=& \frac{\omega}{\gamma}(\omega+\gamma)\Big( \big( \gamma(\omega+\gamma)-\omega \frac{w_I}{w_S}(\omega\,p_I+\gamma)\big) \rho-\omega\gamma(\omega+\gamma) \Big)^2
 \end{align}
@@
 \end{align}
+
+Note the following conditionbs and assumptions of the parameters of the model;
+\begin{align}
+\label{cond}
+& \omega>0, \\
+& 0 \leq \rho<\omega,\\ 
+& 0 \leq \theta\_w \leq \theta\_c \leq 1, and 
+& \frac{w_I}{w_S}\geq 1.
+\end{align}
+Given the parameter restrictions, we would like to determine the sign pattern of $\pro \Delta$ given the quadratic form of it in Eq.
+\eqref{eq:dd2dr}.
+Thus, it is important to know the sign of the coefficients $a$, $b$ and $c$ of the polynomial in Eq.
+\eqref{eq:dd2dr}.
+Since $D_1\geq 0$ and $D_2\geq 0$, $b\leq 0$ and $c\geq 0$.
+Sign of $a$ depends on the interplay between the involved parameters.
+Note that $D_1 \leq D_2$.
+(\todo{still working on}).
 

### 2021-04-09 SIR_manuscript.tex
change b to -b in Delta expression

--- 2021-04-08 SIR_manuscript.tex
@@
 \label{eq:dd2dr}
-\pro \Delta=& \frac{1}{d} (a\,\rho^2+b\,\rho+c), \text{where}\\
+\pro \Delta=& \frac{1}{d} (a\,\rho^2-b\,\rho+c),
+\end{align}
+where
+\begin{align}
+\label{eq:abcd}
 a=& ((\omega+\gamma)\gamma -\omega \frac{w_I}{w_S} (\omega\,p_I+\gamma)) D_1,\\
-b=& -2\gamma\omega(\omega+\gamma) D_1,\\
-c=& \omega^2 (\omega+\gamma)^2 D_2,
+b=& 2\gamma\omega(\omega+\gamma) D_1,\\
+c=& \omega^2 (\omega+\gamma)^2 D_2, \\
+d=& \frac{\omega}{\gamma}(\omega+\gamma)\Big( \big( \gamma(\omega+\gamma)-\omega \frac{w_I}{w_S}(\omega\,p_I+\gamma)\big) \rho-\omega\gamma(\omega+\gamma) \Big)^2,
 \end{align}
-with scaler $d>0$ (not effecting the sign of $\pro \Delta$) and $D_1$ and $D_2$ as 
+and $D_1$ and $D_2$ as 
 \begin{align}
@@
 \end{align}
+Note the following conditionbs and assumptions of the parameters of the model;
+\begin{align}
+\label{cond}
+& \omega>0, \\
+& 0 \leq \rho<\omega,\\ 
+& 0 \leq \theta\_w \leq \theta\_c \leq 1, \\
+& \frac{w_I}{w_S}\geq 1.
+\end{align}
+Given the parameter restrictions, we would like to determine the sign pattern of $\pro \Delta$ given the quadratic form of it in Eq.
+\eqref{eq:dd2dr}.
+Thus, it is important to know the sign of the coefficients $a$, $b$ and $c$ of the polynomial in Eq.
+\eqref{eq:dd2dr}.
+Note that with scaler $d>0$, it has no effect on the sign of $\pro \Delta$.
+Since $D_1\geq 0$ and $D_2\geq 0$, $b\geq 0$ and $c\geq 0$, sign of $a$ depends on the interplay between the involved parameters.
+Note that $D_1 \leq D_2$.
+(\todo{still working on}).
 

### 2021-04-12 SIR_manuscript.tex
updating the notations, clean up

@@
 For brevity, we write $x\_u^* = S\_u^*/N_0$ and $x\_n^* = S\_n^*/N_0$.
-We can then write $\hat F_I = (\omega\rho/(\omega-\rho))w_I/w_S$ for the \percap testing rate of infectious individuals near the disease-free equilibrium.
+We can then write 
+\begin{equation}
+\label{eq:fi}
+\hat F_I = (\omega\rho/(\omega-\rho))w_I/w_S
+\end{equation}
+for the \percap testing rate of infectious individuals near the disease-free equilibrium.
 
@@
 \Delta=& (A~ \theta\_w + B~ \theta\_c) / C\\
-A=& \gamma \Big( \big(\gamma(\omega+\gamma) + \omega p_I F_I \big) x_n^* + (\omega+\gamma) F_I \Big)\\
-B=& \omega p_I F_I (\gamma x_u^*+\omega) \\ 
-C=& (\omega+\gamma) \Big(\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I) \Big).
+A=& \gamma \Big( \big(\gamma(\omega+\gamma) + \omega p_I \hat F_I \big) x_n^* + (\omega+\gamma) \hat F_I \Big)\\
+B=& \omega p_I(\gamma x_u^*+\omega)\hat F_I \\ 
+C=& (\omega+\gamma) \Big(\gamma(\omega+\gamma)+(\gamma+\omega p_I)\hat F_I\Big).
 \end{aligned}
@@
 \end{center}
-%%%%%%%%%% Merge with supplemental materials %%%%%%%%%%
-%%%%%%%%%% Prefix a "S" to all equations, figures, tables and reset the counter %%%%%%%%%%
 \setcounter{equation}{0}
@@
 \renewcommand{\citenumfont}[1]{A#1}
-%%%%%%%%%% Prefix a "S" to all equations, figures, tables and reset the counter %%%%%%%%%%
 
@@
 The next generation matrix for this model is $G = F V^{-1}$, where matrix $F$ represents the inflow of new infection to the infected compartments and matrix $V$ represents the flow in the infected compartments when the population is totally susceptible.
-Matrices $F$ and $V$ are \ali{all $S\_u \& S\_n$ should be $S\_u^* \& S\_n^*$}
+Matrices $F$ and $V$ are
 \begin{align}
@@
  \end {array} \right] \\
- =&\beta\left[ \begin {array}{c} x\_n^* \\ x\_u^* \\ 0 \\ 0 \end {array} \right]
+ =&\beta/N_0 \left[ \begin {array}{c} S\_n^* \\ S\_u^* \\ 0 \\ 0 \end {array} \right]
         \left[ \begin {array}{cccc} 1, \, 1-\theta\_w, \, 1-\theta\_w, \, 1-\theta\_c \end {array} \right], \text{and}\\  V=&
  \left[ \begin {array}{cccc}  
-F_I+\gamma&-\omega&0&0\\
--(1-p_I)F_I&\omega+\gamma&0&0\\
--p_I F_I&0&\omega+\gamma&0\\
+\hat F_I+\gamma&-\omega&0&0\\
+-(1-p_I)\hat F_I&\omega+\gamma&0&0\\
+-p_I \hat F_I&0&\omega+\gamma&0\\
 0&0&-\omega&\gamma
@@
 \gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)&0&0\\ \noalign{\medskip}
-\gamma(\omega+\gamma)(1-p_I)F_I & \gamma(\omega+\gamma)(F_I+\gamma)&0&0 \\ \noalign{\medskip}
-\gamma(\omega+\gamma) p_I F_I & \gamma\omega p_I F_I & C \gamma/(\omega+\gamma) & 0 \\ \noalign{\medskip}
-\omega(\omega+\gamma) p_I F_I & \omega^2 p_I F_I & C \omega/(\omega+\gamma) & C
+\gamma(\omega+\gamma)(1-p_I) \hat F_I & \gamma(\omega+\gamma)(\hat F_I+\gamma)&0&0 \\ \noalign{\medskip}
+\gamma(\omega+\gamma) p_I \hat F_I & \gamma\omega p_I \hat F_I & C \gamma/(\omega+\gamma) & 0 \\ \noalign{\medskip}
+\omega(\omega+\gamma) p_I \hat F_I & \omega^2 p_I \hat F_I & C \omega/(\omega+\gamma) & C
 \end {array} \right],
 \end{align}
-where $C= \big( \gamma(\omega+\gamma)+(\gamma+\omega p_I)F_I \big) (\omega+\gamma)$. Note that all the columns of matrix $V^{-1}$ summ up to $C$.
+where $C= \big( \gamma(\omega+\gamma)+(\gamma+\omega p_I)\hat F_I \big) (\omega+\gamma)$ and $\hat F_I$ is the \percap testing rate for the infected people and represented in Eq.
+\eqref{eq:fi}.
+Note that all the columns of matrix $V^{-1}$ summ up to $1/\gamma$.
 
-The particular form of $F$ with two rows of zeros at the bottom, simplifies $G$ as the following block matrix.
+The particular form of $F$ with two rows of zeros at the bottom results in the following blocked form of matrix $G$.
 \begin{equation}
@@
 \end{equation}
-where the 2 by 2 submatrix $G_{11}$ is 
+where both blocked matricies $G_{11}$ and $G_{12}$ are 2 by 2.
+Given the upper triangular form of matrix $G$, the basic reproduction number $\Rnum$ (defined as the spectral radius of matrix $G$) is only determined by the blocked matrix $G_{11}$,
 \begin{equation}
@@
 G_{11} = \frac{\beta}{\gamma \, C} 
-\left[ \begin {array}{c}  x\_n^* \\ x\_u^*  \end {array} \right]
+\left[ \begin {array}{c} (\omega-\rho)/\omega \\ \rho/\omega \end {array} \right]
 \left[ \begin {array}{cccc} 1, \, 1-\theta\_w, \, 1-\theta\_w, \, 1-\theta\_c \end {array} \right]
@@
 \gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)\\ \noalign{\medskip}
-\gamma(\omega+\gamma)(1-p_I)F_I & \gamma(\omega+\gamma)(F_I+\gamma) \\ \noalign{\medskip}
-\gamma(\omega+\gamma) p_I F_I & \gamma\omega p_I F_I \\ \noalign{\medskip}
-\omega(\omega+\gamma) p_I F_I & \omega^2 p_I F_I
+\gamma(\omega+\gamma)(1-p_I)\hat F_I & \gamma(\omega+\gamma)(\hat F_I+\gamma) \\ \noalign{\medskip}
+\gamma(\omega+\gamma) p_I \hat F_I & \gamma\omega p_I \hat F_I \\ \noalign{\medskip}
+\omega(\omega+\gamma) p_I \hat F_I & \omega^2 p_I \hat F_I
 \end {array} \right].
 \end{equation}
+It is notable that matrix $F$ \eqref{FV} has rank one and consequently $G_{11}$ does so. That is $G_{11}$ has only one non-zero eigenvalue which is $\Rnum$.
 
-It is notable that matrix $F$ \eqref{FV} and consequently $G_{11}$ have rank one.
-That is $G_{11}$ has only one non-zero eigenvalue which we refer to it the basic reproduction number $\Rnum$ (defined as the spectral radius of $G$).
-Also the block matrix $G_{12}$ does not influence $\Rnum$ and all $S\_u$'s and $S\_n$'s are evaluated at the DFE \eqref{dfe} in matrix $F$ \eqref{FV} and the block matrix $G_{11}$ \eqref{mat:G}.
-
-$\Rnum$ has a complicated expression and depend on the type of analysis, the expression can be simplifed and presented accordingly.
+The expression of $\Rnum$ has a complicated form with all of the model parameters involved.
+This expression can be simplifed and represented given the specific form of matrix $G_{11}$ \eqref{G11}.
 For the purpose of simplicity we present $\Rnum$ in the manuscript in terms of expressions $A$, $B$ and $C$, specified in \eqref{eq:abc}.
@@
 
-It remains hard to show the reduction of $\Rnum$ with respect to \percap testing intensity, $\rho$, and the speed of the test return, $\omega$, for the feasible ranges of the parameters, i.e., $\omega>0$, $0 \leq \rho<\omega$, $0 \leq \theta\_w \leq \theta\_c \leq 1$ and $\frac{w_I}{w_S}\geq 1$.
-One easy way to do such an analysis is based on the fact that $\rho$ is \percap rate so for a large population size, $N_0 \gg 1$, $\rho$ is very small or close to 0, comparing to $N_0$.
+It remains hard to show that the reproduction number $\Rnum$ is decreasing with respect to \percap testing intensity, $\rho$, and the speed of the test return, $\omega$, for the feasible ranges of the parameters, that is
+\begin{align}
+\label{cond}
+& \omega>0, \\
+& 0 \leq \rho<\omega,\\ 
+& 0 \leq \theta\_w \leq \theta\_c \leq 1, \\
+& \frac{w_I}{w_S}\geq 1.
+\end{align}
+One way to do such an analysis is based on the fact that $\rho$ is \percap rate so for a large population size, $N_0 \gg 1$, $\rho$ is very small or close to 0, comparing to $N_0$.
 This provides a base to linearly approximate $\Rnum$ when $\rho$ is close to 0, and use this approximation to analyze the behaviour of $\Rnum$ with respect to $\omega$ (see section \ref{taylor}).
-In the next section we provide an equivalent representation of $\Rnum$ for our model to provide to follow proving that more testing intensity decreases $\Rnum$ for a general rane of parameters.
+In the next section we provide an equivalent representation of $\Rnum$ to provide a ground to prove that more testing intensity decreases $\Rnum$ for a general rane of parameters.
 
@@
 
-This is to prove that $\pro \Delta > 0$ with $\Delta$ defined in Eq. \eqref{eq:abc}, and thus $\pro \Rnum < 0$, where $\Rnum$ is the basic reproduction number, given in Eqs. \eqref{R0}.
-
-Another representation of matrix $G_{11}$ in \eqref{G11} is as follows.
+This is to prove that $\pro \Delta > 0$ with $\Delta$ defined in Eq.
+\eqref{eq:abc}, and thus $\pro \Rnum < 0$, where $\Rnum$ is the basic reproduction number, given in Eqs.
+\eqref{R0}.
+We can rewrite matrix $G_{11}$ in \eqref{G11} in the following form to simplify the calculations.
 \begin{equation}
@@
 G_{11} = \frac{\beta}{\gamma \, C} 
-\left[ \begin {array}{c}  x\_n^* \\ x\_u^*  \end {array} \right]
+\left[ \begin {array}{c}  (\omega-\rho)/\omega \\ \rho/\omega  \end {array} \right]
 \left[ \begin {array}{cccc} 
@@
 \end{equation}
-where $C$ is in Eq. \eqref{eq:abc}, and $C_1$ and $C_2$ are 
+where $C$ is the same as the one in Eq. \eqref{eq:abc}, i.e.,
+$$C=(\omega+\gamma)(\gamma(\omega+\gamma)+(\omega\,p_I+\gamma)\,\hat F_I),$$
+and $C_1$ and $C_2$ are 
 \begin{align}
 \label{eq:C12}
-C =& (\omega+\gamma)(\gamma(\omega+\gamma)+(\omega\,p_I+\gamma)\,F_I)\\
-C_1 =& (\omega+\gamma)(\theta\_w \, \gamma+\theta\_c\, \omega\, p_I) F_I,\\
-C_2 =& (\omega+\gamma)\gamma^2\,\theta\_w+(\theta\_w-\theta\_c)\gamma\,p_I\,F_I.
+C_1 =& (\omega+\gamma)(\theta\_w \, \gamma+\theta\_c\, \omega\, p_I) \hat F_I,\\
+C_2 =& (\omega+\gamma)\gamma^2\,\theta\_w+(\theta\_w-\theta\_c)\gamma\,p_I\,\hat F_I,
 \end{align}
-$\Rnum$ is in the same form as in Eq. \eqref{R0}  $$\Rnum= \frac{\beta}{\gamma} (1-\Delta),$$
-with equivalent but simplified $\Delta$ as
+where $\hat F_I$ is given in Eq. \eqref{eq:fi}.
+$\Rnum$ is in the same form as in Eq. \eqref{R0}  
+$$\Rnum= \frac{\beta}{\gamma} (1-\Delta),$$
+with equivalent but simplified $\Delta$ in Eq. \eqref{eq:abc} as
 \begin{equation}
 \label{eq:del2}
-\Delta= \frac{C_1+C_2\,x\_n^*}{C}.
+\Delta= \frac{C_1+C_2\,\rho/\omega}{C}.
 \end{equation}
@@
 \end{align}
-Note the following conditionbs and assumptions of the parameters of the model;
-\begin{align}
-\label{cond}
-& \omega>0, \\
-& 0 \leq \rho<\omega,\\ 
-& 0 \leq \theta\_w \leq \theta\_c \leq 1, \\
-& \frac{w_I}{w_S}\geq 1.
-\end{align}
-Given the parameter restrictions, we would like to determine the sign pattern of $\pro \Delta$ given the quadratic form of it in Eq.
+
+Given the parameter restrictions \ref{cond}, we would like to determine the sign pattern of $\pro \Delta$ given the quadratic form of it in Eq.
 \eqref{eq:dd2dr}.
@@
 Note that $D_1 \leq D_2$.
-(\todo{still working on}).
 

### 2021-04-15 SIR_manuscript.tex
fixing a typo in C_2 in Delta
added analysis of d R0/d rho for special cases

@@
 C_1 =& (\omega+\gamma)(\theta\_w \, \gamma+\theta\_c\, \omega\, p_I) \hat F_I,\\
-C_2 =& (\omega+\gamma)\gamma^2\,\theta\_w+(\theta\_w-\theta\_c)\gamma\,p_I\,\hat F_I,
+C_2 =& (\omega+\gamma)\gamma^2\,\theta\_w+(\theta\_w-\theta\_c)\gamma\,\omega\,p_I\,\hat F_I,
 \end{align}
@@
 Note that $D_1 \leq D_2$.
+
+We are able to prove that $\pro \Rnum \leq 0$ in the following special cases. This is mainly by simplifying matrix $G_{11}$ in \eqref{G11}.\\
+{\bf Case 1; $\theta\_w=\theta\_c=0$}, that is no isolation effectiveness. In this case $\Rnum=\beta/\gamma$ and thus $\pro\Rnum=0$. \\
+{\bf Case 2; $\theta\_w=\theta\_c=1$}, that is the maximum isolation strength. In this case the row vector in matrix $G_{11}$ representation in \eqref{G11} will be $(1,0,0,0)$, thus
+$\Rnum = \frac{\beta (\omega+\gamma)}{C} (\omega + \gamma x\_u^*)$.
+As both $x\_u^*$ and $1/C$ are decreasing functions of $\rho$ and so does their product, thus $\Rnum$ decreases with respect to $\rho$.\\
+{\bf Case 3; $\theta\_w=0, \theta\_c=1$}, after simplification
+$$\Rnum=\Big(\frac{\beta}{\gamma(\omega+\gamma)+(\omega p_I+\gamma)\hat F_I} \Big)
+\Big((\omega+\gamma) +(1+\frac{\omega p_I}{(\omega+\gamma)}x\_n^*) \hat F_I\Big).$$
 

### 2021-04-16 SIR_manuscript.tex
good nigh5

@@
 Our model provides insights to the sensitivity of the epidemiological dynamics, through $\Rnum$, with respect to the undelying epidemic and testing parameters.
-\david{This paragraph does not flow from the rest of the intro, and is a weird way to end when the point of the paper is about testing.}\ali{some edits/rewording are done, ok?}
-
-\david{There is some back and forth between generic discussion and \covid-specific comments.  I think COVID should be used as motivation, but that the whole paper should be framed generically.}
+\ali{edited in response to David comments.}
 

### 2021-04-16 SIR_manuscript.tex
subs rho to phi, analysis and results are updated

--- 2021-04-12 SIR_manuscript.tex
@@
 $$\partial{\Rnum}/\partial{\theta\_c} \leq 0 \text{~and~} \partial{\Rnum}/\partial{\theta\_w} \leq 0.$$ 
-\item \label{p1:rho} a higher rate of testing decreases $\Rnum$;
-$$\partial{\Rnum}/\partial{\rho} \leq 0.$$
+\item \label{p1:rho} a higher rate of testing may increase or decrease $\Rnum$;
+$$\partial{\Rnum}/\partial{\rho}$$ \text{~can be positive or negative}.
 \item \label{p1:omega} A higher rate of of test return may increase or decrease \Rnum.
@@
 % %%%%%%%
-\subsection{More testing intensity decreases $\Rnum$}
+\subsection{More testing intensity may decreases $\Rnum$}
 
-This is to prove that $\pro \Delta > 0$ with $\Delta$ defined in Eq.
-\eqref{eq:abc}, and thus $\pro \Rnum < 0$, where $\Rnum$ is the basic reproduction number, given in Eqs.
-\eqref{R0}.
+This is to provide a mathematical materials to prove that $\pro \Delta$ can be positive or negative, with $\Delta$ defined in Eq.
+\eqref{eq:abc}, and thus $\pro \Rnum < 0$, where $\Rnum$ is the basic reproduction number, given in Eqs.\,\eqref{R0}.
 We can rewrite matrix $G_{11}$ in \eqref{G11} in the following form to simplify the calculations.
@@
 $\Rnum$ is in the same form as in Eq. \eqref{R0}  
-$$\Rnum= \frac{\beta}{\gamma} (1-\Delta),$$
-with equivalent but simplified $\Delta$ in Eq. \eqref{eq:abc} as
+$$\Rnum= \frac{\beta}{\gamma} (1-\Delta).$$
+
+Let define $\phi = \hat F_S = \frac{\rho \omega}{\omega-\rho}$, thus
 \begin{equation}
-\label{eq:del2}
-\Delta= \frac{C_1+C_2\,\rho/\omega}{C}.
+\label{eq:phi}
+\rho=\frac{\omega \phi}{\omega+\phi}.
 \end{equation}
-The behavior of $\Rnum$ with respect to $\rho$ is determine by $\pro \Delta$
+Substituding in $\rho$ in the expression of $\Rnum$, in Eq.\,\eqref{R0}, and applying the derivative with respect to $\phi$ results in the following simplified expression for $\partial\Delta/\partial\phi$.
+
+with an equivalent but simplified $\Delta$ in Eq.\,\eqref{eq:abc} as
+\begin{equation}
+\label{eq:del3}
+\Delta= \frac{K_1\,\phi^2+K_2\,\phi}{K_3},
+\end{equation}
+where
 \begin{align}
-\label{eq:dd2dr}
-\pro \Delta=& \frac{1}{d} (a\,\rho^2-b\,\rho+c),
+\label{eq:K}
+K_1=& \frac{w_I}{w_S}\Big(((1+p_I)\omega+\gamma)\gamma\theta\_w+\omega^2 p_I\theta\_c \Big), \\
+K_2=& (\omega+\gamma)\Big( (\frac{w_I}{w_S}\omega+\gamma)\gamma\theta\_w + \frac{w_I}{w_S} \omega^2 p_I\theta\_c \Big),\\
+K_3=& (\omega+\gamma) \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi\,+\,(\omega+\gamma)\gamma \Big)(\omega+\phi).
+\end{align}
+The derivative is
+\begin{align}
+\label{eq:dd3dr}
+\partial\Delta/\partial\phi=& \frac{1}{d_3} (a_3\,\phi^2+b_3\,\phi+c_3),
 \end{align}
@@
 \begin{align}
-\label{eq:abcd}
-a=& ((\omega+\gamma)\gamma -\omega \frac{w_I}{w_S} (\omega\,p_I+\gamma)) D_1,\\
-b=& 2\gamma\omega(\omega+\gamma) D_1,\\
-c=& \omega^2 (\omega+\gamma)^2 D_2, \\
-d=& \frac{\omega}{\gamma}(\omega+\gamma)\Big( \big( \gamma(\omega+\gamma)-\omega \frac{w_I}{w_S}(\omega\,p_I+\gamma)\big) \rho-\omega\gamma(\omega+\gamma) \Big)^2,
-\end{align}
-and $D_1$ and $D_2$ as 
-\begin{align}
-\label{eq:D1D2}
-D_1=& (\omega+\gamma)\gamma\theta\_w+(\theta\_c-\theta\_w)\omega^2\,p_I\,\frac{w_I}{w_S}, \\
-D_2=& (\omega\,\frac{w_I}{w_S}+\gamma)\gamma\theta\_w+\omega^2\,p_I\,\frac{w_I}{w_S}\,\theta\_c.
+\label{eq:abcd2}
+a_3=& \frac{w_I}{w_S} \Big( (1+\gamma)(\theta\_w-\theta\_c)\frac{w_I}{w_S}\omega\,p_I+ (\gamma\theta\_w+\omega\,p_I\theta\_c)(\omega+\gamma) \Big),\\
+b_3=& 2\frac{w_I}{w_S}(\omega+\gamma)\Big( ((1+p_I)\omega+\gamma)\gamma\theta\_w+\omega^2 p_I\theta\_c \Big) ,\\
+c_3=& (\omega+\gamma)^2 \Big( (\frac{w_I}{w_S}\omega+\gamma)\gamma\theta\_w + \frac{w_I}{w_S} \omega^2 p_I\theta\_c \Big),\\
+d_3=& \frac{(\omega+\gamma)}{\omega\gamma} \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi\,+\,(\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2. 
 \end{align}
 
-Given the parameter restrictions \ref{cond}, we would like to determine the sign pattern of $\pro \Delta$ given the quadratic form of it in Eq.
-\eqref{eq:dd2dr}.
-Thus, it is important to know the sign of the coefficients $a$, $b$ and $c$ of the polynomial in Eq.
-\eqref{eq:dd2dr}.
-Note that with scaler $d>0$, it has no effect on the sign of $\pro \Delta$.
-Since $D_1\geq 0$ and $D_2\geq 0$, $b\geq 0$ and $c\geq 0$, sign of $a$ depends on the interplay between the involved parameters.
-Note that $D_1 \leq D_2$.
+Note that $\phi\geq 0$, also $b_3$, $c_3$ and $d_3$ are all positive. However $a_3$ can be positive or negative.
+If $a_3\geq 0$, $\partial\Delta/\partial\phi \geq 0$. 
+If $a_3 < 0$, then the quadratic expression in the numerator of \eqref{eq:dd3dr} has a positive root, $\phi^*$, such that for $\phi>\phi^*$, $\partial\Delta/\partial\phi < 0$.
+An example of this scenario occures when $\theta\_w=0$ and $\theta\_c=1$, thus $\theta\_w-\theta\_c=-1$, this is when the strength of isolation for awaiting people is the least, but the most for the confirmed cases.
+In this case
+\begin{equation}
+\label{ineq:a3}
+a_3 <0 \iff \omega+\gamma<(1+\gamma)\frac{w_I}{w_S}.
+\end{equation}
+Given that $\frac{w_I}{w_S}\geq 1$, the sufficient condition for this inequality to be held is 
+$\omega < \frac{w_I}{w_S}$.
+This can be visualized in the top-right panel of the Fig.\,\ref{pan2} panel (b).
 
@@
 \includegraphics[width=\linewidth]{codes/R0contour_random_rho.5.pdf}
-\caption{}\label{p.a}
+\caption{}
 \end{subfigure}
@@
 \includegraphics[width=\linewidth]{codes/R0contour_TTI_rho.5.pdf}
-\caption{}\label{p.b}
+\caption{}
 \end{subfigure}
@@
 }
-\label{pan}
+\label{pan2}
 \end{figure}

### 2021-04-16 note/SIR_scrap.tex
subs rho to phi, analysis and results are updated

--- 2021-04-16 SIR_manuscript.tex
@@
 \newcommand{\percap}{\emph{per capita}\xspace}
-\newcommand{\Rnum}{\ensuremath{\mathcal{R}_0}}
+\newcommand{\Rnum}{\mathcal{R}_0}
 \newcommand{\covid}{COVID-19\xspace}
-\newcommand{\pro}[1][]{\ensuremath{\frac{\partial #1}{\partial \rho}}}
 
@@
 \linenumbers
-
-% %%%%%%%
-\section{Abstract}
-
-The effect of testing processes, including (testing and test reporting) on an epidemic dynamics (infection and recovery) can be studied at the individual level or the community level (e.g., nursing homes, LTC facilities, etc.).
-Gaining insights to determine the sensitivity of the epidemic dynamics with respect to the testing processes will depend on underlying factors including the level of focus (individual or community), assumptions (model), and the interplay between these factors.
-In particular, the fast test reporting may be beneficial at the community-level, supported by many studies, as it gives a rapid assessment of the situation, identifies hot spots, and may enable rapid contact-tracing.
-However, the potential advantage of a slow rate of test return on the dynamics of an epidemic is real, often neglected, and needs to be quantified.
-At the individual level, this advantage can manifest in the following sense: individuals awaiting their test results or who have tested positive may partially or fully self-isolate, thus reducing or eliminating their potential in the transmission process.
-In this paper, we investigated this individual-level effect of testing processes on the epidemic dynamics by developing a SIR-type model.
-Although the model development was motivated by the \covid epidemic, the model has general epidemiological and testing structures.
-The realistic components of the model include testing intensity, test sensitivity and specificity, rate of test return, and isolation.
-The novel component is the compartment-specific relative testing weights, which reflect the testing strategies-- surveillance, diagnosis, or control.
-Here, we showed that the slow test reporting can be beneficial on the dynamics of an epidemic.
-In particular, it is possible for the basic reproduction number, $\Rnum$, to be increasing with respect to the rate of test return.
-Further, we compare two testing strategies, random vs.
-targeted, in this framework.
-We conclude that the targeted testing strategy is more effective than the random one in the sense that a lower range of testing intensity is required to keep $\Rnum <1$.
-Also, targeted testing reduces the individual-level of advantage of a slow rate of test return.
-
-% %%%%%%%
-\section{Introduction}
-
-The observed dynamics of the \covid epidemic are driven by both epidemiological processes (infection and recovery) and testing processes (testing and test reporting).
-In addition to shaping epidemic observations (via case reports), testing processes can also affect epidemiological dynamics.
-In particular, individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may do so also (possibly to a lesser extent).
-We developed a mechanistic model that incorporates epidemic processes and testing in order to explore the effects of testing and isolation on epidemic dynamics.
-
-If testing influences behavior, then epidemic dynamics will depend on patterns of who gets tested.
-The impacts of testing will depend on intensity (tests performed per day), and on how strongly testing is focused on people who are infectious.
-This level of focus depends in turn on the purpose and design of testing programs. 
-When testing is done for the purposes of disease surveillance \citep{foddai2020base}
-tests should be assigned randomly across the population, possibly with a stratified design for statistical efficiency \citep{graubard1996modelling} \ali{a better ref everyone?}.
-
-Over the course of the \covid pandemic, however, the vast majority of testing has been done with other goals --
-primarily diagnosis (determining the infection status for clinical purposes), or control (determining the infection status in order to quickly isolate cases that have been found by contact tracing), which we characterize as \emph{targeted} testing strategies.
-In these cases, testing probabilities vary widely across epidemiological compartments; in our dynamical model, we will characterize these probabilities by assigning a \percap testing weight to each compartment that determines the \emph{relative} probability that an individual in that compartment will be selected for testing (see Methods).
-
-When testing is used primarily for diagnosis it will focus on people with infection-like symptoms; thus the relative testing weights for infected people will depend on the relative probability of infected people having symptoms.
-For \covid infection, the testing weights will depend on the relative asymptomatic infections, time spent pre-sympomatic vs.\ symptomatic infections -- and also the incidence of \covid-like symptoms among people in the population \emph{not} infected with \covid.
-Testing for epidemic control will focus on people who are known to have been in contact with known infected cases; in this case the testing weights for infected vs.\ uninfected people will depend on the probability of infection given contact, as well as the thoroughness and effectiveness of the system for identifying suspicious contacts.
-
-The main interest from the epidemiological point of view is to know whether the number of infected individuals goes through an exponential growth phase, following the introduction of an infection in a totally susciptable population, before the disease becomes extinct.
-This is determined by studying the basic reproduction number $\Rnum$.
-It is defined as the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}.
-In the early stages of an epidemic the number of infected individuals is expected to grow exponentially over time when $\Rnum>1$, and to decline over time when $\Rnum<1$. 
-Although the value of $\Rnum$ cannot completely characterize the dynamics of even the simplest epidemic model
-\citep{shaw2021what}, it does give a simple and widely accepted index for the difficulty of control, as well as some indication of the likely final size of an epidemic \citep{ma2006generality}.
-
-In order to understand the effect of testing processes on an epidemic dynamics, we deveopled a mechanistic SIR-type model with epidemic and testing components.
-Here, we focus on the the effect of testing intensity, different levels of testing ``focus'' (from random to highly targeted), and rate of test return on the epidemic dynamics.
-Our model provides insights to the sensitivity of the epidemiological dynamics, through $\Rnum$, with respect to the undelying epidemic and testing parameters.
-\ali{edited in response to David comments.}
-
-% %%%%%%%
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Methods}
-
-We developed a deterministic model, Eqs.
-\eqref{model}, which groups individuals based on disease status and testing status.
-Disease states include Susceptible, Infectious and Recovered (thus this is an SIR-type model), and testing status categorizes people as \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive} (Figure~\ref{fig:flowchart}).
-Symbolically, the testing status of an individual in the disease compartment $X$, where $X \in \{S,I,R\}$, is reflected in the subscript, namely $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
-Further, two `accumulator' compartments, $N$ and $P$, were also incorporated in the model in order to collect cumulative reported negative or positive tests.
-The model and details of calculation of the basic reproduction number $\Rnum$ are presented in the appendixs.
-
-\begin{figure}[!h] 
-\begin{center} 
-\includegraphics[scale=2]{pix/sir_comp.pdf}
-\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model, \ref{model}.
-Here, the disease-based status of a compartment $X$, where $X \in \{S,I,R\}$, is combined with the testing-based status including $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
-Also,  $\Lambda$ is the force of infection with definition in Eq.~\eqref{Lambda}, $\gamma$ is the recovery rate, $\omega$ is the rate of test return, $F_X$ and $p_X$ represent the \percap testing rate and the probability of positive tests, respectively, for compartment $X$.
-For further description of the parameters see \cref{tab:params}.
-\label{fig:flowchart}}
-\end{center} 
-\end{figure}
-
-Table~\ref{tab:params} defines the model parameters, which are generally straightforward
-\percap flows between compartments, or modifiers to these flow rates. The \percap testing intensity is defined as the number of daily tests taken in a population of size $N_0$. 
-The novel component of the model comes in through the compartment-specific relative testing weights $w_S$, $w_I$ and $w_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively.
-For example, $w_I/w_S=2$ means that infected individuals are tested at twice the \percap rate of susceptible individuals.
-
-In order to link to more applied models, we constructed this model so we could specify the total \percap testing rate.
-We do this by defining the weighted size of the testing pool $W = w_S S\_u + w_I I\_u + w_R R\_u$, and calculating a scaling parameter for testing as:
-\begin{equation}
-\label{sigma}
-\sigma = \frac{\rho N_0}{W}.
-\end{equation}
-Thus, the \percap testing rate for compartment $X$ is $F_X=\sigma w_X$, where $X \in \{S,I,R\}$. 
-For a high-sensitivity test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus unavailable for further testing.
-Over the course of the epidemic, a fixed testing rate as specified in \eqref{sigma} can (if large enough) exhaust the pool of people available for testing, leading to a singularity when no one is left untested.
-Although this phenomenon does not affect our analysis of $\Rnum$, it can affect the temporal dynamics (we discuss an adjustment to the model that solves this problem in the appendix).
-
-The classical SIR model is based on the following implicit assumptions; well-mixed population, homogeneity of the population (i.e., all individuals are equaly susciptable and equaly infectious for the same length of time when infected), exponentially distributed duration of infection and large population size (see, e.g., \cite{keeling2011modeling}).
-In addition to these standard assumptions, our model, \ref{model}, assumes: (i) there is a single force of infection (new cases per unit time), $\Lambda$, defined as follows
-\begin{equation}
-\label{Lambda}
-\Lambda=\frac{\beta}{N_0} \big(I\_u + (1-\theta\_w)(I\_n+I\_p) + (1-\theta\_c)I\_c \big),
-\end{equation}
-across all susceptible pools with transmission rate $\beta$ and isolation efficacy in reduction of the probability of transmission for ``waiting'' and \emph{confirmed positive} individuals, $\theta\_w$ and $\theta\_c$ respectively, (see Table~\ref{tab:params} for further details), (ii) $\theta\_c \geq \theta\_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
-Thus, for instance when the awaiting people follow the isolation perfectly, $\theta\_w$ is closer to 1, while when they less follow the isolation, $\theta\_w$ is closer to 0.
-For this analysis, we also assume a perfectly specific test ($p_S=0$).
-This last assumption combined with the assumption that no individual is in waiting-for-\emph{positive} or \emph{confirmed positive} compartments, i.e., $S\_p(0)=S\_c(0)=0$, reduces the model to 10 equations (equations c and d  in \eqref{model} are eliminated).
-
-\begin{table}[htp]
-\centering
-\begin{tabular}{|c|L{2in}cc|} \hline
-  Symbol & Description & Unit & Value \\ \hline
-  $N_0$     & Total population size & people & $10^6$ \\ \hline
-  $\omega$  & Rate of onward flow from ``waiting'' to ``confirmed'' or ``untested'' compartments  & 1/day & - \\ \hline
-  $\gamma$ & Recovery rate & 1/day & 1/3 \\ \hline 
-  $\rho$   & Per capita testing intensity & 1/day & 0.01 \\ \hline 
-  $\theta\_w$ & Isolation efficacy in reduction of the probability of transmission for ``waiting'' individuals & - & - \\ \hline
-  $\theta\_c$ & Isolation efficacy in reduction of the probability of transmission for ``confirmed positive'' individuals & - & -  \\ \hline
-  $\beta$ & Transmission rate & 1/day & 0.39 \\ \hline
-  $\Lambda$ & Force of infection & 1/day & - \\ \hline
-  $p_S$ & Probability of positive tests for $S$ ($= 1-\textrm{specificity}$) & - & 0 \\ \hline
-  $p_I$ & Probability of positive tests for $I$ ($= \textrm{sensitivity}$) & - & 1 \\ \hline
-  $p_R$ & Probability of positive tests for $R$ ($= 1-\textrm{specificity}$) & - & 0.5 \\ \hline
-  $w_S, w_I, w_R$ & Relative testing weights & - &
-  \begin{minipage}[t]{0.21\columnwidth}%
-    Random: $\{1,1,1\}$ \\ Targeted: $\{0.3,1,1\}$
-  \end{minipage} \\
-  \hline
-  \end{tabular}
-\caption{\label{tab:params} Parameters of the model \eqref{model}.}
-\end{table}
-
-\todo{the algebra needs to be changed based on JD's comment} The Disease-Free Equilibrium (DFE) for the SIR model, Eqs.
-\eqref{model}, is given by setting the differential equations to 0 and solving for the unknowns (\ali{is this clear or the "unkowns" needs to be defined?
-also * in sup for I and R?}.
-The DFE is
-\begin{equation}
-\label{dfe}
-S\_n^*= \frac{\rho}{\omega} N_0, \ S\_u^*= N_0-S\_n^*, \text{~and~} I_j=R_j=0 \ \text{for all }j.
-\end{equation}
-For brevity, we write $x\_u^* = S\_u^*/N_0$ and $x\_n^* = S\_n^*/N_0$.
-We can then write 
-\begin{equation}
-\label{eq:fi}
-\hat F_I = (\omega\rho/(\omega-\rho))w_I/w_S
-\end{equation}
-for the \percap testing rate of infectious individuals near the disease-free equilibrium.
-
-The basic reproduction number, $\Rnum$, was calculated by using the next generation matrix method developed by \cite{van2002reproduction}. $\Rnum$ is
+Old component equation for $\Rnum$;
 \begin{equation}
@@
 \label{eq:abc}
-\Delta=& (A~ \theta\_w + B~ \theta\_c) / C\\
-A=& \gamma \Big( \big(\gamma(\omega+\gamma) + \omega p_I \hat F_I \big) x_n^* + (\omega+\gamma) \hat F_I \Big)\\
-B=& \omega p_I(\gamma x_u^*+\omega)\hat F_I \\ 
-C=& (\omega+\gamma) \Big(\gamma(\omega+\gamma)+(\gamma+\omega p_I)\hat F_I\Big).
+\Delta=& 1-(A~ S\_u^* + B~ S\_n^*)~ C\\
+A=& \gamma(\omega+\gamma) + (\gamma \eta\_w + \omega \eta\_c p_I) F_I, \\
+B=& \big(\omega+(F_I+\gamma)\eta\_w\big) \gamma+\frac{(\eta\_w \gamma+ \eta\_c\omega) \omega p_I F_I }{\omega+\gamma}, \\
+C=& \frac{1}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}.
 \end{aligned}
@@
 
-Further details of derivation of $\Rnum$ are provided in the appendix section.
+\begin{align}
+\label{FV}
+F =& \beta/N_0 \left[ \begin {array}{cccc} 
+S\_u&(1-\theta\_w)\,S\_u&(1-\theta\_w)\,S\_u&(1-\theta\_c)\,S\_u\\
+S\_n&(1-\theta\_w)\,S\_n&(1-\theta\_w)\,S\_n&(1-\theta\_c)\,S\_n\\ 
+0&0&0&0\\
+0&0&0&0
+ \end {array} \right], \\
+  V =&
+ \left[ \begin {array}{cccc}  
+F_I+\gamma&-\omega&0&0\\
+-(1-p_I)F_I&\omega+\gamma&0&0\\
+-p_I F_I&0&\omega+\gamma&0\\
+0&0&-\omega&\gamma
+\end {array} \right], \text{thus}\\
+V^{-1} =&
+\left[ \begin {array}{cccc}
+\frac {\omega+\gamma}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&\frac {\omega}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&0&0\\
+\noalign{\medskip}
+\frac{(1-p_I) F_I}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&
+\frac{F_I+\gamma}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&0&0\\
+\noalign{\medskip}
+\frac{p_I F_I}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&
+\frac{\omega p_I F_I}{(\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I)(\omega+\gamma)}& \frac{1}{\omega+\gamma}&0 \\
+\noalign{\medskip}
+\frac{\omega\,F_I\,p_I}{(\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I) \gamma}&
+\frac{\omega^2\,F_I\,p_I}{(\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I)(\omega+\gamma) \gamma}&
+\frac{\omega}{(\omega+\gamma) \gamma}&
+\frac{1}{\gamma}
+\end {array} \right],
+\end{align}
 
-The analytical calculation of the next-generation matrix and the derivation of $\Rnum$ was carried out in Maple \citep{maple14}. 
-We used R \citep{r} for simulations and for plotting $\Rnum$ as a function of the underlying parameters (using \ref{R0}).
-Results are shown in Figure \ref{pan}, where  panel \eqref{p.a} represents the random testing, and panel \eqref{p.b} is representing non-random testing.
-The parameter values or ranges are presented in Table~\ref{tab:params}.
-This simulation reflects the behavior of $\Rnum$ with respect to the selected parameters for two different testing strategies: (i) random testing, represented by all testing weights to be the same, $w_S=w_I=w_R$, and (ii) non-random testing, when testing weight are not equal.
-For simulation purposes we chose $w_S=w_I=w_R=1$ for random testing, and $w_S=0.3$ and $w_I=w_R=1$ for non-random testing strategy.
-Note that the critical contour of $\Rnum=1$ is plotted in solid line in Figure \ref{pan}.
-
-% %%%%%%%
-\section{Results}
-
-The explicit formula for the basic reproduction number, $\Rnum$ \eqref{R0}, provides an opportunity to study the influence of changes in the underlying parameters on the critical index of epidemic dynamics.
-We are interested in exploring the effect of parameters that could be manipulated by public-health policy, including the isolation parameters ($\theta\_c$ and $\theta\_w$) testing intensity ($\rho$) and test-return rate ($\omega$).
-
-Examining our formula for $\Rnum$ \eqref{R0} gives the following results. 
-See the appendix for details.
-
-\begin{enumerate}
-\item \label{p1:eta} increasing the isolation parameters for tested and confirmed individuals decreases \Rnum;
-$$\partial{\Rnum}/\partial{\theta\_c} \leq 0 \text{~and~} \partial{\Rnum}/\partial{\theta\_w} \leq 0.$$ 
-\item \label{p1:rho} a higher rate of testing decreases $\Rnum$;
-$$\partial{\Rnum}/\partial{\rho} \leq 0.$$
-\item \label{p1:omega} A higher rate of of test return may increase or decrease \Rnum.
-$$\partial{\Rnum}/\partial{\omega} \text{~can be positive or negative}.$$
-\end{enumerate}
-
-\begin{figure}[h!]
-\centering
-\begin{subfigure}[t]{.45\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_random.pdf}
-\caption{}\label{p.a}
-\end{subfigure}
-%
-\begin{subfigure}[t]{.45\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_TTI.pdf}
-\caption{}\label{p.b}
-\end{subfigure}
-\caption{
-{\bf A comparison of the behaviour of the basic reproduction number, $\Rnum$, between random versus targeted testing strategies at different levels of testing and isolation.}
-We numerically evaluate $\Delta$ \eqref{R0}, reflecting the reduction of $\Rnum$ with respect to testing and isolation. We use the following parameters (see Table \ref{tab:params}):
-$N_0=1 \times 10^6$, $\omega \in [0.1,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.01]$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.39$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
-Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
-The black solid line in each panel represents the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ corresponding to the threshold of $\Rnum=1$.
-\bmb{we still have some collisions in the y-axis tick labels}
-}
-\label{pan}
-\end{figure}
-
-Our numerical results are shown in Figure \ref{pan}.
-It can be inferred that when random testing strategy is applied, Figure \ref{p.a}, comparing to the targeted testing strategy, Figure \ref{p.b}, the critical \percap testing intensity of the corresponding panels are lower in non-random testing.
-Also, it appears that speeding the test reporting, i.e., increasing $\omega$, does not significantly lower $\Rnum$ when awaiting people (including people in $I\_p$ and $I\_n$ compartments) follow the isolation perfectly, i.e., $\theta\_w$ is closer to 1.
-However, speeding the test reporting reduces the epidemic more when the awaiting people less follow the isolation, i.e., $\theta\_w$ is closer to 0.
-
-Furthermore, we derived an inequality that quantifies the exact relationships between model parameters that result in returning tests more rapidly being favorable \eqref{eq:necsuf}.
-\david{What do we learn from this inequality?  Why is it relegated to an appendix?}
-
-%%%%%%%%%%%%
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Discussion}
 
-Mathematical modeling of infectious disease outbreaks provides insights on how testing processes influence the epidemiological processes through isolation. 
-Here, we develop a compartmental SIR-type model to study the potential effect of testing strategies, testing intensity, test sensitivity and specificity, test reporting time and isolation on epidemic dynamics.
-While targeted testing strategies 
-\david{you need to be clearer: you mean targeting individuals who have contacted people who were infected.
-For example, here you could say something like ``While targeting the contacts of confirmed cases\dots''}
-are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is that viral spread
-is greater for a slow test than for a fast test.
-This counter-intuitive effect can occur when people are cautious when awaiting a test result, and may not be robust to second-order effects of fast testing (such as better contact tracing).
-
-We incorporated the compartment-specific relative testing weights, $w_S$, $w_I$ and $w_R$, to model random testing and \emph{targeted} testing strategies.
-Here, in the case of targeted testing and for the simplicity and illustration purposes, we assumed that infected and recovered individuals are tested at three times the \percap rate of susceptible individuals, thus $w_I/w_S=3$ and $w_R=w_I$.
-Note that we have not specified a methodology to assign particular relative testing weights corresponding to a particular targeted testing scenario.
-This part needs to be developed further in future work. 
-Modeling different targeted testing strategies, equivalently test-specific testing weights in our framework, requires prior information of the conditional probabilities of getting tested for people in a given compartment.
-This can be implied when we would like to quantify and compare the effect of different levels of test focus for infectious people on the basic reproduction number $\Rnum$, and conclude about the disease spread management.
-For example, when people are tested for "screening", the individuals with potential higher mobility, eg.
-people who are getting on flights, get more tested and thus the coresponding heavier testing weight is assigned than people awaiting for a surgery and are probably going to stay in a long-term care facility and consequently less mobile and more isolated to begin with.
-With our model, we would be able to compare the sensitivity of the disease dynamics, through $\Rnum$, with respect to testing processes.
-
-{\bf The \percap testing intensity, $\rho$;}
-Proposition \ref{prop1} part \ref{p1:rho} and also Figure \ref{pan} indicates that increasing testing intensity $\rho$ reduces $\Rnum$.
-This is sensible since as people are moved to test compartments, namely $X_p$ and $X_n$ for $X \in \{S,I,R\}$, they may partially or fully self-isolate.
-Also individuals who have tested positive, namely $X_c$ compartment for $X \in \{S,I,R\}$, are highly likely to self-isolate.
-The higher probability of being subject to isolation, the lower the force of infection \eqref{Lambda} and the lower $\Rnum$ will become.
-While our simulation supported this result, analytically it appeared to be hard to conclude this due to the complexity of $\Rnum$ expression.
-
-{\bf The potential advantage of slow test reporting, or favorable-delay-reporting;}
-Individuals are highly likely to fully self-isolate when they are either awaiting the test results or they are reported, thus reducing or eliminating their potentially infectious contacts.
-Thus, the faster the test reporting rate, $\omega$, the shorter these individuals stay in the ``\emph {safe}'' awaiting-confirmed compartments, namely $X_n$, $X_p$ and $X_c$ for $X \in \{S,I,R\}$, and the more they get involved in the infection process.
-This advantage of slow test reporting is real, and neglected. 
-We also compare to an individual-level advantage of fast tests: people who test positive may be even more careful.
-In our model analysis, Proposition \ref{prop1} part \ref{p1:omega} is describing this potential advantage. 
-It states that returning test results more rapidly, i.e., increasing the rate $\omega$, does not necessarily lower the reproduction number $\Rnum$; whether increasing $\omega$ lowers $\Rnum$ depends on the precise combination of model parameters  including test reporting rate, testing strategies represented by compartment-specific testing weights, test sensitivity and specificity, and the level of isolation.
-Specifically, in the case of perfect isolation, i.e., when $\theta\_w=1$ and $\theta\_c=1$, $\Rnum$ may increase as the test reporting process becomes faster.
-This can be seen from expression \eqref{Rom}.
-Another example of this favorable delay in $\omega$ could be when the test being employed produces many false negatives.
-Because many infected asymptomatic individuals will believe they are uninfected/uninfectious, thus may unknowingly spread the virus to many others.
-Again the delay in the test reporting rate keeps these individuals in the ``\emph {safe}'' awaiting-confirmed compartments.
-
-we are missing out on
-community-level advantages of fast testing: better assessment of the
-situation, identification of hot spots, contact-tracing, etc.
-The \percap testing intensity of CIVID-19 after about a year from the first case reported in December 2019, is still low (($\rho \approx 0$) in our model).
-In near future new test kits may be widely accessible, our model provide insights in this case.
-In particular, if a cheap test can identify on average more infected individuals as an expensive test, then our model predicts that the cheap test will lower $\Rnum$ more.
-In contrast, if an expensive test can identify on average more infected individuals, it will not necessarily lower $\Rnum$ more than the cheap test.
-The use of tests cheaper than RT-PCR has been proposed as a potential strategy for containing the COVID pandemic.
-While cheaper tests may be less sensitive and reliable than RT-PCR, they allow for broader and more intense testing.
-Using our Taylor approximation of $\Rnum$ near $\rho = 0$, we examined what circumstances (i.e., model parameters) make the use of one test more favourable than another, and give a complete description through inequality \ref{eq:rho1vsrho2}.
-In general, we found that the expensive test tends to more effectively lower $\Rnum$ when (a) individuals who test positive self-isolate much more than individuals who are waiting for their test result, (b) the time it takes to return tests is much shorter than the mean infectious period, and (c) the testing intensity is much greater for infected individuals than susceptible individuals.
-
-\ali{ not sure if we want these 2 following parageraphs, or shrink it into afew short sentences?}
-In addition to the favorable-delay-reporting observation of our model \eqref{model}, the model enables us to quantify the amount of delay required in the test reporting process as a strategy to reduce $\Rnum$.
-To give a biological interpretation, we describe the qualitative trends predicted by the inequality \eqref{eq:necsuf}.
-To summarise, returning test results more rapidly tends to be favourable (i.e., reduces $\Rnum$) when (i) \emph{Confirmed-positive} individuals, lower their contact much more than individuals who are waiting for their test results (i.e., $\theta\_c \gg \theta\_w$), (ii) the test is highly sensitive (i.e., $p_I$ is close to 1), and (iii) the targeted testing strategy is used (i.e., $w_I \gg w_S$).
-
-one point we can make when discussing the relative weight of testing in different compartments is to discuss pre-testing screening tools, such as surveys or questionnaires.
-If we have a quantitative description of how the testing intensities affect the dynamics, we can make statements like ``our results suggest that employing a pre-testing screening tool can help target infected individuals more effectively.
-In particular, doubling the sensitivity of the pre-screening tool would *do something* to $\Rnum$.
-
-\david{Some of this reads like notes for discussion rather than text for the paper, so I'm not trying to edit.}
-
-% %%%%%%%
-\bibliography{SIRlibrary}
-
+\bibliography{../SIRlibrary}
 % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{center}
+%%%%%%%%%% Merge with supplemental materials %%%%%%%%%%
+%%%%%%%%%% Prefix a "S" to all equations, figures, tables and reset the counter %%%%%%%%%%
 \setcounter{equation}{0}
@@
 \setcounter{table}{0}
-% \setcounter{page}{1}
+\setcounter{page}{1}
 \makeatletter
@@
 \renewcommand{\citenumfont}[1]{A#1}
+%%%%%%%%%% Prefix a "S" to all equations, figures, tables and reset the counter %%%%%%%%%%
 
@@
 The next generation matrix for this model is $G = F V^{-1}$, where matrix $F$ represents the inflow of new infection to the infected compartments and matrix $V$ represents the flow in the infected compartments when the population is totally susceptible.
-Matrices $F$ and $V$ are
+Matrices $F$ and $V$ are \ali{all $S\_u \& S\_n$ should be $S\_u^* \& S\_n^*$}
 \begin{align}
@@
 F =& \beta/N_0 \left[ \begin {array}{cccc} 
-S\_u^* & (1-\theta\_w)\,S\_u^* & (1-\theta\_w)\,S\_u^* & (1-\theta\_c)\,S\_u^*\\
-S\_n^* & (1-\theta\_w)\,S\_n^* & (1-\theta\_w)\,S\_n^* & (1-\theta\_c)\,S\_n^* \\ 
+S\_u&\eta\_w\,S\_u&\eta\_w\,S\_u&\eta\_c\,S\_u\\
+S\_n&\eta\_w\,S\_n&\eta\_w\,S\_n&\eta\_c\,S\_n\\ 
 0&0&0&0\\
 0&0&0&0
- \end {array} \right] \\
- =&\beta/N_0 \left[ \begin {array}{c} S\_n^* \\ S\_u^* \\ 0 \\ 0 \end {array} \right]
-        \left[ \begin {array}{cccc} 1, \, 1-\theta\_w, \, 1-\theta\_w, \, 1-\theta\_c \end {array} \right], \text{and}\\  V=&
+ \end {array} \right], \\
+  V =&
  \left[ \begin {array}{cccc}  
-\hat F_I+\gamma&-\omega&0&0\\
--(1-p_I)\hat F_I&\omega+\gamma&0&0\\
--p_I \hat F_I&0&\omega+\gamma&0\\
+F_I+\gamma&-\omega&0&0\\
+-(1-p_I)F_I&\omega+\gamma&0&0\\
+-p_I F_I&0&\omega+\gamma&0\\
 0&0&-\omega&\gamma
+\end {array} \right], \text{thus}\\
+V^{-1} =&
+\left[ \begin {array}{cccc}
+\frac {\omega+\gamma}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&\frac {\omega}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&0&0\\
+\noalign{\medskip}
+\frac{(1-p_I) F_I}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&
+\frac{F_I+\gamma}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&0&0\\
+\noalign{\medskip}
+\frac{p_I F_I}{\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I}&
+\frac{\omega p_I F_I}{(\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I)(\omega+\gamma)}& \frac{1}{\omega+\gamma}&0 \\
+\noalign{\medskip}
+\frac{\omega\,F_I\,p_I}{(\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I) \gamma}& 
+\frac{\omega^2\,F_I\,p_I}{(\omega\,\gamma+\gamma F_I+\gamma^2+\omega\, F_I p_I)(\omega+\gamma) \gamma}&
+\frac{\omega}{(\omega+\gamma) \gamma}&
+\frac{1}{\gamma}
 \end {array} \right].
 \end{align}
-The matrix inverse of $V$ is 
-\begin{align}
-\label{vinv}
-V^{-1} =&
-\frac{1}{\gamma C}
-\left[ \begin {array}{cccc}
-\gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)&0&0\\ \noalign{\medskip}
-\gamma(\omega+\gamma)(1-p_I) \hat F_I & \gamma(\omega+\gamma)(\hat F_I+\gamma)&0&0 \\ \noalign{\medskip}
-\gamma(\omega+\gamma) p_I \hat F_I & \gamma\omega p_I \hat F_I & C \gamma/(\omega+\gamma) & 0 \\ \noalign{\medskip}
-\omega(\omega+\gamma) p_I \hat F_I & \omega^2 p_I \hat F_I & C \omega/(\omega+\gamma) & C
-\end {array} \right],
-\end{align}
-where $C= \big( \gamma(\omega+\gamma)+(\gamma+\omega p_I)\hat F_I \big) (\omega+\gamma)$ and $\hat F_I$ is the \percap testing rate for the infected people and represented in Eq.
-\eqref{eq:fi}.
-Note that all the columns of matrix $V^{-1}$ summ up to $1/\gamma$.
 
-The particular form of $F$ with two rows of zeros at the bottom results in the following blocked form of matrix $G$.
+The particular form of $F$ with two rows of zeros at the bottom, simplifies $G$ as 
 \begin{equation}
@@
 0&0
-\end {array} \right],
+\end {array} \right], \text{ where } \\
+G_{11} = \frac{\beta}{\gamma}\ C
+\left[\begin {array}{cc}
+A\,S\_u & B\,S\_u\\
+A\,S\_n & B\,S\_n
+\end {array}\right],
 \end{equation}
-where both blocked matricies $G_{11}$ and $G_{12}$ are 2 by 2.
-Given the upper triangular form of matrix $G$, the basic reproduction number $\Rnum$ (defined as the spectral radius of matrix $G$) is only determined by the blocked matrix $G_{11}$,
-\begin{equation}
-\label{G11}
-G_{11} = \frac{\beta}{\gamma \, C} 
-\left[ \begin {array}{c} (\omega-\rho)/\omega \\ \rho/\omega \end {array} \right]
-\left[ \begin {array}{cccc} 1, \, 1-\theta\_w, \, 1-\theta\_w, \, 1-\theta\_c \end {array} \right]
-\left[ \begin {array}{cc}
-\gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)\\ \noalign{\medskip}
-\gamma(\omega+\gamma)(1-p_I)\hat F_I & \gamma(\omega+\gamma)(\hat F_I+\gamma) \\ \noalign{\medskip}
-\gamma(\omega+\gamma) p_I \hat F_I & \gamma\omega p_I \hat F_I \\ \noalign{\medskip}
-\omega(\omega+\gamma) p_I \hat F_I & \omega^2 p_I \hat F_I
-\end {array} \right].
-\end{equation}
-It is notable that matrix $F$ \eqref{FV} has rank one and consequently $G_{11}$ does so. That is $G_{11}$ has only one non-zero eigenvalue which is $\Rnum$.
+with expressions $A$, $B$ and $C$ are specified in \eqref{eq:abc}.
+The block matrix $G_{12}$ does not influence $\Rnum$ (defined as the spectral radius of $G$). All that matters here are the eigenvalues of $G_{11}$, which are 0 and $\Rnum$ \eqref{R0}.
+Also note that, all $S\_u$'s and $S\_n$'s are evaluated at the DFE \eqref{dfe} in matrix $F$ \eqref{FV} and the block matrix $G_{11}$ \eqref{mat:G}.
 
-The expression of $\Rnum$ has a complicated form with all of the model parameters involved.
-This expression can be simplifed and represented given the specific form of matrix $G_{11}$ \eqref{G11}.
-For the purpose of simplicity we present $\Rnum$ in the manuscript in terms of expressions $A$, $B$ and $C$, specified in \eqref{eq:abc}.
-This is intended to show that $\Rnum$ decreases as the strength of isolation, shown by $\theta\_w$ and $\theta\_c$, increases.
+% %%%%%%%
+\subsection{On Testing Rate and Numerical Singularity}
 
-It remains hard to show that the reproduction number $\Rnum$ is decreasing with respect to \percap testing intensity, $\rho$, and the speed of the test return, $\omega$, for the feasible ranges of the parameters, that is
-\begin{align}
-\label{cond}
-& \omega>0, \\
-& 0 \leq \rho<\omega,\\ 
-& 0 \leq \theta\_w \leq \theta\_c \leq 1, \\
-& \frac{w_I}{w_S}\geq 1.
-\end{align}
-One way to do such an analysis is based on the fact that $\rho$ is \percap rate so for a large population size, $N_0 \gg 1$, $\rho$ is very small or close to 0, comparing to $N_0$.
-This provides a base to linearly approximate $\Rnum$ when $\rho$ is close to 0, and use this approximation to analyze the behaviour of $\Rnum$ with respect to $\omega$ (see section \ref{taylor}).
-In the next section we provide an equivalent representation of $\Rnum$ to provide a ground to prove that more testing intensity decreases $\Rnum$ for a general rane of parameters.
+In this work, we didn't do any numerical solutions for the trajectories in our analysis. However, if one tries to do so there would be a singularity issue to deal with. 
+Specifically, the numerical singularity issue with the chosen $\sigma$ \eqref{sigma} is that the population in $S$ compartments appeared to blow up when the DFE is achieved.
+This is once the only untested people are susceptibles, the FOI will become $\Lambda=0$, testing rate $F_s=\rho N_0/S\_u$.
+Thus, the first equation of the model \eqref{model} will become
+$d S\_u/dt = - \rho N_0 + \omega S\_n$. Thus changes in $S\_u$ will be no longer dependent on $S\_u$ with a linear rate of leaving the $S\_u$ compartment.
+IN fact the testing rate, $\sigma$, should be formulated such that people from the untested compartments will not be tested if they are not there.
+One way to fix this issue, is to consider a maximum testing rate, $\tau$ (1/day).
+In general, we want to test at a rate of $\rho$ across the whole population.
+This won't always be possible, so we impose a maximum rate of $\tau$ per testable person and redefine $\sigma = \frac{\tau \rho N_0}{\tau W + \rho N_0}$, with the assumption that $\tau \gg \rho$.
+This alteration in $\sigma$, does not change any results related to $\Rnum$, thus we only impose it in the simulation of the epidemic dynamic.
 
@@
 % %%%%%%%
+\subsubsection{Substituting $\rho$ by $\phi$}
+Let define $\phi = \hat F_S = \frac{\rho \omega}{\omega-\rho}$, thus
+\begin{equation}
+\label{eq:phi}
+\rho=\frac{\omega \phi}{\omega+\phi}.
+\end{equation}
+Substituding in $\rho$ in the expression of $\Rnum$, in Eq.\,\eqref{R0}, and applying the derivative with respect to $\phi$ results in the following simplified expression for $\partial\Delta/\partial\phi$.
+
+\begin{equation}
+\label{eq:del3}
+\Delta= \frac{K_1\,\phi^2+K_2\,\phi}{K_3},
+\end{equation}
+where
+\begin{align}
+\label{eq:K}
+K_1=& \frac{w_I}{w_S}\Big(((1+p_I)\omega+\gamma)\gamma\theta\_w+\omega^2 p_I\theta\_c \Big), \\
+K_2=& (\omega+\gamma)\Big( (\frac{w_I}{w_S}\omega+\gamma)\gamma\theta\_w + \frac{w_I}{w_S} \omega^2 p_I\theta\_c \Big),\\
+K_3=& (\omega+\gamma) \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi\,+\,(\omega+\gamma)\gamma \Big)(\omega+\phi).
+\end{align}
+The derivative is
+\begin{align}
+\label{eq:dd3dr}
+\partial\Delta/\partial\phi=& \frac{1}{d_3} (a_3\,\phi^2+b_3\,\phi+c_3),
+\end{align}
+where
+\begin{align}
+\label{eq:abcd2}
+a_3=& \frac{w_I}{w_S} \Big( (1+\gamma)(\theta\_w-\theta\_c)\frac{w_I}{w_S}\omega\,p_I+ (\gamma\theta\_w+\omega\,p_I\theta\_c)(\omega+\gamma) \Big),\\
+b_3=& 2\frac{w_I}{w_S}(\omega+\gamma)\Big( ((1+p_I)\omega+\gamma)\gamma\theta\_w+\omega^2 p_I\theta\_c \Big) ,\\
+c_3=& (\omega+\gamma)^2 \Big( (\frac{w_I}{w_S}\omega+\gamma)\gamma\theta\_w + \frac{w_I}{w_S} \omega^2 p_I\theta\_c \Big),\\
+d_3=& \frac{(\omega+\gamma)}{\omega\gamma} \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi\,+\,(\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2. 
+\end{align}
+
+Note that $\phi\geq 0$, also $b_3$, $c_3$ and $d_3$ are all positive. However $a_3$ can be positive or negative.
+If $a_3\geq 0$, $\partial\Delta/\partial\phi \geq 0$. 
+If $a_3 < 0$, then the quadratic expression in the numerator of \eqref{eq:dd3dr} has a positive root, $\phi^*$, such that for $\phi>\phi^*$, $\partial\Delta/\partial\phi < 0$.
+An example of this scenario occures when $\theta\_w=0$ and $\theta\_c=1$, thus $\theta\_w-\theta\_c=-1$, this is when the strength of isolation for awaiting people is the least, but the most for the confirmed cases.
+In this case
+\begin{equation}
+\label{ineq:a3}
+a_3 <0 \iff \omega+\gamma<(1+\gamma)\frac{w_I}{w_S}.
+\end{equation}
+Given that $\frac{w_I}{w_S}\geq 1$, the sufficient condition for this inequality to be held is 
+$\omega < \frac{w_I}{w_S}$.
+This can be visualized in the top-right panel of the Fig.\,\ref{pan2} panel (b).
+
+% %%%%%%%
+\subsection{Taylor Approximation of $\Rnum$ at $\rho=0$ }
+
+The basic reproduction number, $\Rnum$, close to $\rho=0$ can be approximated linearly in $\rho$ by using Taylor approximation. it follows
+\begin{equation}
+\label{eq:R0appr}
+\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(\eta\_w-1)(\gamma w_S+\omega w_I) + (\eta\_c -1)p_I w_I \omega^2 \Big) + \mathcal{O}(\rho^2).
+\end{equation}
+
+Also, to analyze the influence of $\omega$ on $\Rnum$, the approximation \eqref{eq:R0appr} was used. Then it is straight forward to have
+\begin{equation}
+\label{Rom}
+\partial{\Rnum}/\partial{\omega}=  \frac{-\beta \rho}{\gamma w_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c),
+\end{equation}
+where $a=(\eta\_w-1)w_I-(\eta\_c-1)p_I w_I = ((s-p_I)\eta\_c + (p_I-1)) w_I$, $b=2(\eta\_w-1)\gamma w_S$ and $c=(\eta\_w-1)\gamma^2 w_S$.
+Given that $0 \leq \eta\_c\leq \eta\_w \leq 1 $, one can easily derive $b\leq 0$ and $c \leq 0$.
+
+Note that in general, the necessary and sufficient condition for $a \geq 0$ is $(s-p_I) \eta\_c \geq (1-p_I)$, where $s=\frac{\eta\_w}{\eta\_c} \geq 1$.
+
+In the case of ``perfect isolation'', i.e., when $\eta\_w=0$ and consequently $\eta\_c=0$, it is straightforward to see that $a \leq 0$, $b<0$ and $c<0$.
+Thus, $\partial{\Rnum}/\partial{\omega} \geq 0$.
+As an example, in the case of a very accurate testing regime,  i.e., $P_I=1$, $a \geq 0$ is achieved.
+If $a\geq 0$, the quadratic expression in \eqref{Rom}, has Real roots.
+Assuming that $\omega_1<0$ and $\omega_2>0$ be the roots of the quadratic expression in $\partial{\Rnum}/\partial{\omega}$.
+Thus, $\partial{\Rnum}/\partial{\omega}>0$ for $0<\omega<\omega_2$ and  $\partial{\Rnum}/\partial{\omega}<0$ for $\omega>\omega_2$.
+
+% %%%%%%%
 \subsection{rate of returning tests} \label{taylor}
@@
 \begin{equation}\label{linearization}
-\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(-\theta\_w)(\gamma w_S+\omega w_I) + (-\theta\_c)p_I w_I \omega^2 \Big). 
+\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(\eta\_w-1)(\gamma w_S+\omega w_I) + (\eta\_c -1)p_I w_I \omega^2 \Big). 
 \end{equation}
 So when $\rho \approx 0$ we have $$\partial{\Rnum}/\partial{\omega} \approx  \frac{-\beta \rho}{\gamma w_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c),$$
-where $a=(-\theta\_w)w_I-(-\theta\_c)p_I w_I$, $b=2(-\theta\_w)\gamma w_S$ and $c=(-\theta\_w)\gamma^2 w_S$.
+where $a=(\eta\_w-1)w_I-(\eta\_c-1)p_I w_I$, $b=2(\eta\_w-1)\gamma w_S$ and $c=(\eta\_w-1)\gamma^2 w_S$.
 
@@
 Note that the zeroes are real if and only if $Q < 1$.
-Note that have $\theta\_c > \theta\_w$, so if $p_I \approx 1$, we will have $Q < 0 < 1$.
+Note that have $\eta\_c < \eta\_w$, so if $P_i \approx 1$, we will have $Q < 0 < 1$.
 Thus, if we assume near-perfect test sensitivity, $\omega_1$ and $\omega_2$ will be real.
@@
 
-% %%%%%%%
-\subsection{On Testing Rate and Numerical Singularity}
-
-In this work, we didn't do any numerical solutions for the trajectories in our analysis. However, if one tries to do so there would be a singularity issue to deal with. 
-Specifically, the numerical singularity issue with the chosen $\sigma$ \eqref{sigma} is that the population in $S$ compartments appeared to blow up when the DFE is achieved.
-This is once the only untested people are susceptibles, the FOI will become $\Lambda=0$, testing rate $F_s=\rho N_0/S\_u$.
-Thus, the first equation of the model \eqref{model} will become
-$d S\_u/dt = - \rho N_0 + \omega S\_n$. Thus changes in $S\_u$ will be no longer dependent on $S\_u$ with a linear rate of leaving the $S\_u$ compartment.
-IN fact the testing rate, $\sigma$, should be formulated such that people from the untested compartments will not be tested if they are not there.
-One way to fix this issue, is to consider a maximum testing rate, $\tau$ (1/day).
-In general, we want to test at a rate of $\rho$ across the whole population.
-This won't always be possible, so we impose a maximum rate of $\tau$ per testable person and redefine $\sigma = \frac{\tau \rho N_0}{\tau W + \rho N_0}$, with the assumption that $\tau \gg \rho$.
-This alteration in $\sigma$, does not change any results related to $\Rnum$, thus we only impose it in the simulation of the epidemic dynamic.
-
-% %%%%%%%
 \subsection{Expensive vs.\ cheap tests}
@@
 Suppose that $\rho_1 > \rho_2$.
-Recall that the linearization of $\Rnum$ around $\rho \approx 0$ is given by $$\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(-\theta\_w)(\gamma w_S+\omega w_I) + (-\theta\_c)p_I w_I \omega^2 \Big).$$
+Recall that the linearization of $\Rnum$ around $\rho \approx 0$ is given by $$\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(\eta\_w-1)(\gamma w_S+\omega w_I) + (\eta\_c -1)p_I w_I \omega^2 \Big).$$
 
@@
 \begin{align}\label{eq:rho1vsrho2}
-    &\rho_1\left(\gamma(-\theta\_w)(\gamma w_S + \omega w_I) + (-\theta\_c)p_{I,1}w_I\omega^2\right) - \rho_2\left(\gamma(-\theta\_w)(\gamma w_S + \omega w_I) + (-\theta\_c)p_{I, 2}w_I\omega^2\right) > 0 \nonumber \\
+    &\rho_1\left(\gamma(\eta\_w-1)(\gamma w_S + \omega w_I) + (\eta\_c-1)p_{I,1}w_I\omega^2\right) - \rho_2\left(\gamma(\eta\_w-1)(\gamma w_S + \omega w_I) + (\eta\_c-1)p_{I, 2}w_I\omega^2\right) > 0 \nonumber \\
     &\vdots \nonumber \\
-    &\frac{\rho_2P_{i, 2}-\rho_1P_{i, 1} }{\rho_1-\rho_2} > \frac{\theta\_w}{\theta\_c}\cdot \frac{\gamma(\gamma w_S + \omega w_I)}{\omega^2 w_I}
+    &\frac{\rho_2P_{i, 2}-\rho_1P_{i, 1} }{\rho_1-\rho_2} > \frac{1-\eta\_w}{1-\eta\_c}\cdot \frac{\gamma(\gamma w_S + \omega w_I)}{\omega^2 w_I}
 \end{align}
@@
 
-\subsection{Simulation with $\rho$ varies from 0 to 0.5 }
+% %%%%%%%
+\subsection{Literature Review}
 
-\begin{figure}[h!]
-\centering
-\begin{subfigure}[t]{.45\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_random_rho.5.pdf}
-\caption{}\label{p.a}
-\end{subfigure}
-%
-\begin{subfigure}[t]{.45\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_TTI_rho.5.pdf}
-\caption{}\label{p.b}
-\end{subfigure}
-\caption{
-{\bf A comparison of the behaviour of the basic reproduction number, $\Rnum$, between random versus targeted testing strategies at different levels of testing and isolation.}
-We numerically evaluate $\Delta$ \eqref{R0}, reflecting the reduction of $\Rnum$ with respect to testing and isolation. We use the following parameters (listed in Table \ref{tab:params}):
-$N_0=1 \times 10^6$, $\omega \in [0.5,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.5]$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.39$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
-Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
-The black solid line in each panel represents the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ corresponding to the threshold of $\Rnum=1$.
-\bmb{we still have some collisions in the y-axis tick labels}
-}
-\label{pan}
-\end{figure}
+\subsubsection{Explicit models of TTI (trace/test/isolate) based on network or agent-based models}
+\citep{endo2020implication} \ali{It seems to me that this is just a statistical model to estimate the parent-offspring of an infected index, not sure if it fits into agent-based group!} Used simulation on a branching process model to assess the forward and backward contact tracing efficiency.
+Assuming a negative-binomial branching process with a mean R, reproduction number, and overdispersion parameter k, the mean total number of generation G3 and averted G3 are estimated.
+The effectiveness of TTI is defined as the ratio of averted to the mean.
+
+\citep{jenness2020modeling} developed a network-based transmission model for SARS-CoV-2 on the Diamond Princess outbreak to characterize transmission dynamics and to estimate the epidemiological impact of outbreak control and prevention measures.
+
+\citep{elbanna2020entry} [seems similar to MacPan model!]
+
+\citep{de2020influenza} Was discussed in the Math 747 
+SEIR Asymptomatic and symptomatic $I\_1, I\_2$. Used linear chain trick 
+Stringency index as a control force lowering $\beta$.
+
+\citep{rice2020effect} Effect of school closures on mortality. Reproduce Report 9 results by spatial agent based CovidSim. 
+% %%%%%%%
+\subsubsection{Models of repeated random testing of isolated populations}
+\cite{bergstrom2020frequency}
+(1) Model, assumptions: They developed a function, namely expected exposure $E(C,\tau)$, to approximate trade-offs between the frequency of testing, n, the sensitivity of testing, q, and the delay between
+testing and results, d. This function is explicitly derived and was connected the effective reproduction number $R=\Rnum S$, where $S$ is the proportion of population susceptible.
+assumption that transmission rates are a step function: individuals who
+have \covid go from non-infectious to fully infectious instantaneously,
+and remain fully infectious until they are no longer able to transmit disease. Test sensitivity takes the same form over the course of infection.
+More sophisticated models could allow varying infectiousness and varying
+sensitivity over time, as in 
+\citep{larremore2020test}.
+
+\citep{lopman2020model} Used a Deterministic SEIR model, incorporated TTI, applicable to a university setting. They assumed a fairly high reproductive number that is not reduced through social
+distancing measures. They found that community-introduction of SARS-CoV-2 infection onto campus can be
+relatively controlled with effective testing, isolation, contract tracing and quarantine.
+
+\citep{tuite2020mathematical} used an age-structured compartmental model of \covid transmission in the population of Ontario, Canada.
+We compared a base case with limited testing, isolation and quarantine to different scenarios.
+% %%%%%%%
+\subsubsection{Other maybe-related works}
+\citep{arino2020simple} developed a SLIAR compartmental model to study the spread of an epidemic, specifically COVID, in a population.
+The model incorporates an Erlang distribution of times of sojourn in incubating, symptomatically and asymptomatically infectious compartments.
+Basic reproduction number is derived.
+Also, sensitivity analysis with respect to the underlying parameters for the following two outputs was carried out; (i) the number of observable cases during the course of the epidemic and at the peak, and (ii) the timing of the peak of the outbreak.
+Sensitivity analysis is performed using the R package multisensi.
+
+\citep{ruszkiewicz2020diagnosis} novel with-in-a-minute breath testing with 80\% accuracy.
 

### 2021-04-16 SIR_manuscript.tex
minor edit

@@
 \item \label{p1:rho} a higher rate of testing may increase or decrease $\Rnum$;
-$$\partial{\Rnum}/\partial{\rho}$$ \text{~can be positive or negative}.
+$$\partial{\Rnum}/\partial{\rho} \text{~can be positive or negative}.$$
 \item \label{p1:omega} A higher rate of of test return may increase or decrease \Rnum.

### 2021-04-19 SIR_manuscript.tex
typo fixed in a_3

@@
 \end{equation}
-Substituding in $\rho$ in the expression of $\Rnum$, in Eq.\,\eqref{R0}, and applying the derivative with respect to $\phi$ results in the following simplified expression for $\partial\Delta/\partial\phi$.
+Reparametrizing testing rate $\rho$ by $\phi$ in the expression of $\Rnum$, in Eq.\,\eqref{R0}, and applying the derivative with respect to $\phi$ results in the following simplified expression for $\partial\Delta/\partial\phi$.
 
@@
 \label{eq:abcd2}
-a_3=& \frac{w_I}{w_S} \Big( (1+\gamma)(\theta\_w-\theta\_c)\frac{w_I}{w_S}\omega\,p_I+ (\gamma\theta\_w+\omega\,p_I\theta\_c)(\omega+\gamma) \Big),\\
+a_3=& \frac{w_I}{w_S} \Big((\omega\,p_I+\gamma)(\theta\_w-\theta\_c)\frac{w_I}{w_S}\omega\,p_I+(\gamma\theta\_w+\omega\,p_I\theta\_c)(\omega+\gamma) \Big),\\
 b_3=& 2\frac{w_I}{w_S}(\omega+\gamma)\Big( ((1+p_I)\omega+\gamma)\gamma\theta\_w+\omega^2 p_I\theta\_c \Big) ,\\
@@
 Note that $\phi\geq 0$, also $b_3$, $c_3$ and $d_3$ are all positive. However $a_3$ can be positive or negative.
-If $a_3\geq 0$, $\partial\Delta/\partial\phi \geq 0$. 
+If $a_3\geq 0$, $\partial\Delta/\partial\phi \geq 0$ for all feasible range of parameters, thus $\pro\Delta \leq 0$. 
 If $a_3 < 0$, then the quadratic expression in the numerator of \eqref{eq:dd3dr} has a positive root, $\phi^*$, such that for $\phi>\phi^*$, $\partial\Delta/\partial\phi < 0$.
-An example of this scenario occures when $\theta\_w=0$ and $\theta\_c=1$, thus $\theta\_w-\theta\_c=-1$, this is when the strength of isolation for awaiting people is the least, but the most for the confirmed cases.
+An example of this scenario occures when $\theta\_w=0$ and $\theta\_c=1$, which is illustraited in the top-right panel of the Fig.\,\ref{pan2} panel (b).
+This is when the strength of isolation for awaiting people is the least, but the most for the confirmed cases.
 In this case
@@
 \label{ineq:a3}
-a_3 <0 \iff \omega+\gamma<(1+\gamma)\frac{w_I}{w_S}.
+a_3 <0\iff\omega+\gamma<(\omega\,p_I+\gamma)\frac{w_I}{w_S}.
 \end{equation}
-Given that $\frac{w_I}{w_S}\geq 1$, the sufficient condition for this inequality to be held is 
-$\omega < \frac{w_I}{w_S}$.
-This can be visualized in the top-right panel of the Fig.\,\ref{pan2} panel (b).
+Given that $\frac{w_I}{w_S}\geq 1$, and $0\leq\,p_I\leq\,1$, this inequality is hold when a perfect sensitive test is used, i.e., $p_I=1$, and a targeted tesing strategy is followed, i.e., $\frac{w_I}{w_S}>1$.
+\ali{i)what is the interpretation of this?
+2) (Ben:) calculate the corresponding eigenvector to $\Rnum$, scale it to 1, and plot the population of each infected compartment, what does this tell us?
+}
+
+Note that $\rho$ and $\omega$ have similar mechanism in delaying people to get into $I\_c$. The non monotonic countereffect of testing rate and test return rate are not obvious. 
+that people in the DFE waiting for negative test, $S_n^*$.
+More testing may have a counter effect on $\Rnum$, as if we have more testing we have more people in $S_n^*$ which will take them longer to get to $I\_c$.
+
+more $\rho$ then
 

### 2021-04-20 codes/SIRfunctions.R
towards ploting the eigen vec

@@ -1,4 +1,5 @@
 library(shellpipes)
+library(matlib)
 
 # Functions for the SIR model
 # By: A Gharouni
@@ -174,6 +175,29 @@ eval_R0 <- function(state=state_dfe,params){
                           ) 
           )
   return(df2)
+}
+
+Fi_hat<-function(params){
+  unpack(as.list(c(params)))
+  return(omega*rho/(omega-rho)*W_I/W_S)
+}
+
+eigvec_max<-function(params){
+  unpack(as.list(params))
+  Sn<-Sn_dfe(params)
+  Su<-Su_dfe(params)
+  Fi<-Fi_hat(params) 
+  colvec<-matrix(c(Sn,Su,0,0),4,1)  
+  rowvec<-matrix(c(1,eta_w,eta_w,eta_c),1,4)
+  Fmat<-beta/N0*colvec%*%rowvec
+  vmat<-matrix(c(Fi+gamma,-omega,0,0,
+                 (p_I-1)*Fi,omega+gamma,0,0,
+                 -p_I*Fi,0,omega+gamma,0,
+                 0,0,-omega,gamma
+  ),4,4,byrow = TRUE)
+  G<- Fmat %*% inv(vmat)
+  ev<-eigen(G)
+  return(ev$vectors[,which(ev$values>0)]) 
 }
 
 

### 2021-04-21 codes/SIRfunctions.R
plotting the eigenvec at DFE. needs interpretation?

@@ -197,9 +197,10 @@ eigvec_max<-function(params){
   ),4,4,byrow = TRUE)
   G<- Fmat %*% inv(vmat)
   ev<-eigen(G)
-  return(ev$vectors[,which(ev$values>0)]) 
+  ev_max<- ev$vectors[,which(ev$values==max(ev$values))]
+  
+  return(ev_max) 
 }
-
 
 
 # ###################

### 2021-04-21 codes/params_df.R

@@ -1,0 +1,61 @@
+## create unified data frame
+library(ggplot2)
+library(dplyr)
+library(tidyr)
+library(directlabels)
+
+source("SIRfunctions.R")
+source("params.R")
+
+#choose weight W_S 0.3 for TTI testing plan and 1 for the random testing.
+W_S_random <- 1
+W_S_targeted <- 0.3 
+
+params_dat<-function(params){
+  unpack(as.list(params))
+  n_out <- 5 ## facets (rows/cols)
+  n_in <- 41 ## grid N within facets
+  tol <- 1e-10 ## to resolve the issue of very small numbers 
+  # specify the ranges, FIXME, not to be hard coded!
+  # eta_w <- seq(0,1, length.out=n_out)
+  # eta_c <- seq(0,1, length.out=n_out)  
+  rho <- seq(0,0.5, length.out=n_in) ## 0,0.01
+  omega <- 0.5+tol #seq(0.5+tol,2 , length.out=n_in) #note omega must be > rho
+  
+  df1 <- expand.grid(N0=N0,beta=beta,gamma=gamma,omega=omega,rho=rho,
+                     W_S=W_S,W_I=W_I,W_R=W_R,
+                     p_S=p_S,p_I=p_I,p_R=p_R,
+                     eta_w=eta_w,eta_c=eta_c)
+  ## principle eigenvector
+  eigvec <- t(apply(df1,1,function(params_in)eigvec_max(params=params_in)))  
+dfout <-(df1 %>% 
+            dplyr::mutate(
+            R0=apply(df1,1,function(params_in)R0(params=params_in)),  
+            R0_sub=ifelse(eta_w<eta_c, NA, R0), 
+            theta_w=1-eta_w,
+            theta_c=1-eta_c,
+            Delta=ifelse(eta_w<eta_c, NA, 1-(R0*gamma/beta) ),
+            Delta=ifelse(abs(Delta)<tol,0,Delta),
+            I_u=eigvec[,1],
+            I_n=eigvec[,2],
+            I_p=eigvec[,3],
+            I_c=eigvec[,4]
+            ))
+  return(dfout)
+  }
+
+df_random <- params_dat(params = update(params,W_S=W_S_random,eta_w=1,eta_c=0))
+df_targeted <- params_dat(params = update(params,W_S=W_S_targeted,eta_w=1,eta_c=0))
+
+out2 <- df_targeted %>%
[…]

### 2021-04-21 codes/SIRfunctions.R
eigen vec plotted done

@@ -198,7 +198,7 @@ eigvec_max<-function(params){
   G<- Fmat %*% inv(vmat)
   ev<-eigen(G)
   ev_max<- ev$vectors[,which(ev$values==max(ev$values))]
-  
+  if(all(ev_max<=0)){ev_max<--ev_max}
   return(ev_max) 
 }
 

### 2021-04-21 codes/params_df.R

@@ -58,4 +58,6 @@ gg1 <- (ggplot(data=out2, aes(x=rho,y=value,col=compartment))
 direct.label(gg1,"last.bumpup")
 
 
+##plot of R0 wrt rho, notice the increase
+ggplot(data=out2, aes(x=rho,y=R0_sub))+geom_point()
 

### 2021-04-21 codes/SIRfunctions.R
function eval_R0 deleted, instead more flexible fun added make_params_dat()

@@ -112,18 +112,21 @@ run.sir <- function(model, params,state,sim_time){
 F_I <- function(state,params){
   # Returns F_I at the specified state
   unpack(as.list(c(state,params)))
-  
-  # sigma <- rho*N0/(W_S*Su) #at DFE 
   sig <- sigma(state,params)
   return(sig*W_I)
 }
+## Fi at DFE
+Fi_hat<-function(params){
+  unpack(as.list(params))
+  return(omega*rho/(omega-rho)*W_I/W_S)
+}
 
 # Basic Reproduction Number
-R0<-function(state=state_dfe,params){
+R0<-function(params){
   unpack(as.list(c(state,params)))
   Sn <- Sn_dfe(params)
   Su <- Su_dfe(params)
-  Fi <- F_I(state,params) #at DFE
+  Fi <- Fi_hat(params) #at DFE
   
   A <- gamma*(omega+gamma+eta_w*Fi) + omega*eta_c*p_I*Fi
   B <- (omega+eta_w*(Fi+gamma))*gamma+(eta_w*gamma+eta_c*omega)*(omega*p_I*Fi)/(omega+gamma)
@@ -131,57 +134,7 @@ R0<-function(state=state_dfe,params){
   return((A*Su+B*Sn)*C)
 }
 
-# R0 expression with eta_w/eta_c factored
-R02 <- function(state=state_dfe,params){
-  unpack(as.list(c(state,params)))
-  Sn <- Sn_dfe(params)
-  Su <- Su_dfe(params)
-  Fi <- F_I(state,params) #at DFE
-  
-  A2 <- gamma*(p_I*Fi*omega*Sn+(omega+gamma)*(gamma*Sn+Fi*N0))
-  B2 <- p_I*Fi*omega*(gamma*Su+omega*N0)
-  D <- gamma*(omega+gamma)*(gamma*Su+omega*N0)
-  C2 <- beta/((N0*gamma*(gamma*(omega+gamma)+Fi*(gamma+omega*p_I)))*(omega+gamma))
-  # s <- eta_w/eta_c
-  return((eta_c*(s*A2+B2)+D)*C2)
-}
-
-# make grid and calc R0 for plotting
-eval_R0 <- function(state=state_dfe,params){
-  ## input the params and their range, this function makes a grid dataframe, calls R0 and outputs the csv file
-  ## the params are defaults, specify the range of favorite params here.
-  unpack(as.list(c(state,params)))
[…]

### 2021-04-21 codes/params.R

@@ -14,8 +14,8 @@ omega=0.25, #test returning rate
 rho=0.01, #testing rate percapita (also used 0.8, 1/3)
 W_S=1, W_I=1, W_R=1, #Testing relative weights
 p_S=0, p_I=1, p_R=0.5, #test specificity, i.e., prob of waiting for being tested positive
-eta_w=0.02, eta_c=0.01, #isolation parameter (eta=0 is the perfect isolation)
-s=2 # isolation ratio parameter, i.e., s=eta_w/eta_c
+eta_w=0.02, eta_c=0.01 #isolation parameter (eta=0 is the perfect isolation)
+## s=2 # isolation ratio parameter, i.e., s=eta_w/eta_c
             )
             
 class(params) <- "params_pansim"

### 2021-04-21 codes/params_df.R

@@ -1,63 +1,0 @@
-## create unified data frame
-library(ggplot2)
-library(dplyr)
-library(tidyr)
-library(directlabels)
-
-source("SIRfunctions.R")
-source("params.R")
-
-#choose weight W_S 0.3 for TTI testing plan and 1 for the random testing.
-W_S_random <- 1
-W_S_targeted <- 0.3 
-
-params_dat<-function(params){
-  unpack(as.list(params))
-  n_out <- 5 ## facets (rows/cols)
-  n_in <- 41 ## grid N within facets
-  tol <- 1e-10 ## to resolve the issue of very small numbers 
-  # specify the ranges, FIXME, not to be hard coded!
-  # eta_w <- seq(0,1, length.out=n_out)
-  # eta_c <- seq(0,1, length.out=n_out)  
-  rho <- seq(0,0.5, length.out=n_in) ## 0,0.01
-  omega <- 0.5+tol #seq(0.5+tol,2 , length.out=n_in) #note omega must be > rho
-  
-  df1 <- expand.grid(N0=N0,beta=beta,gamma=gamma,omega=omega,rho=rho,
-                     W_S=W_S,W_I=W_I,W_R=W_R,
-                     p_S=p_S,p_I=p_I,p_R=p_R,
-                     eta_w=eta_w,eta_c=eta_c)
-  ## principle eigenvector
-  eigvec <- t(apply(df1,1,function(params_in)eigvec_max(params=params_in)))  
-dfout <-(df1 %>% 
-            dplyr::mutate(
-            R0=apply(df1,1,function(params_in)R0(params=params_in)),  
-            R0_sub=ifelse(eta_w<eta_c, NA, R0), 
-            theta_w=1-eta_w,
-            theta_c=1-eta_c,
-            Delta=ifelse(eta_w<eta_c, NA, 1-(R0*gamma/beta) ),
-            Delta=ifelse(abs(Delta)<tol,0,Delta),
-            I_u=eigvec[,1],
-            I_n=eigvec[,2],
-            I_p=eigvec[,3],
-            I_c=eigvec[,4]
-            ))
-  return(dfout)
-  }
-
-df_random <- params_dat(params = update(params,W_S=W_S_random,eta_w=1,eta_c=0))
-df_targeted <- params_dat(params = update(params,W_S=W_S_targeted,eta_w=1,eta_c=0))
-
-out2 <- df_targeted %>%
[…]

### 2021-04-21 codes/sir_plot.R

@@ -25,10 +25,20 @@ n_in <- 41 ## grid N within facets
 W_S_random <- 1
 W_S_targeted <- 0.3 
 
-# make dataframe and save it, if it doesn't exists already.
-df_random <- eval_R0(params = update(params,W_S=W_S_random))
-df_targeted <- eval_R0(params = update(params,W_S=W_S_targeted))
+# make dataframe 
+# df_random <- eval_R0(params = update(params,W_S=W_S_random))
+# df_targeted <- eval_R0(params = update(params,W_S=W_S_targeted))
+df_random <- make_params_dat(params = params,
+                eta_ws=0,eta_we=1, ## so theta_w
+                eta_cs=0,eta_ce=1, ## so theta_c
+                omega_s=0.1,omega_e=2,
+                rho_s=0,rho_e=0.01)
 
+df_targeted <- make_params_dat(params = update(params,W_S=W_S_targeted),
+                eta_ws=0,eta_we=1, ## so theta_w
+                eta_cs=0,eta_ce=1, ## so theta_c
+                omega_s=0.1,omega_e=2,
+                rho_s=0,rho_e=0.01)
 ##important contour, ie R0=1 thus threshold=1 when plotting R0 contours, or Delta(R0=1)
 threshold <- 1-(params[["gamma"]]/params[["beta"]]) ## Or 0?  corresponding to R0=1 
 # #################################
@@ -45,10 +55,6 @@ label_special <- function (labels, multi_line=FALSE, sep = "== ") {
   out[[1]] <- lapply(out[[1]], function(expr) c(parse(text=expr)))
   out
 }
-
-# verify the break range in brks_vec (FIXME)
- # mn <- min(df_random$Delta,df_targeted$Delta, na.rm = T)
- # mx <- max(df_random$Delta,df_targeted$Delta, na.rm = T) ## take mn and mx and set the brks_vec, smarter way?
 
 n_contour <- 8 ## number of desired contours or bins
 bins <- cut(sort(unique(c(df_random$Delta,df_targeted$Delta))), 
@@ -104,29 +110,4 @@ ggsave((p1_temp %+% df_targeted) +
        filename = "R0contour_TTI.pdf" ,
        width = 12, height = 12, units = "cm")
 
-# #################################
-# 3. Plot R0 as a function of rho (testing intensity):
-# #################################
-eta_w <- 0.5
-eta_c <- eta_w
[…]

### 2021-04-21 SIR_manuscript.tex
towards ploting the eigen vec
more cleaning

@@
  \end {array} \right] \\
- =&\beta/N_0 \left[ \begin {array}{c} S\_n^* \\ S\_u^* \\ 0 \\ 0 \end {array} \right]
+ =&\beta/N_0 \left[ \begin {array}{c} S\_u^* \\ S\_n^* \\ 0 \\ 0 \end {array} \right]
         \left[ \begin {array}{cccc} 1, \, 1-\theta\_w, \, 1-\theta\_w, \, 1-\theta\_c \end {array} \right], \text{and}\\  V=&
@@
 We numerically evaluate $\Delta$ \eqref{R0}, reflecting the reduction of $\Rnum$ with respect to testing and isolation. We use the following parameters (listed in Table \ref{tab:params}):
-$N_0=1 \times 10^6$, $\omega \in [0.5,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.5]$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.39$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
+$N_0=1 \times 10^6$, $\omega \in (0.5,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.5]$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.39$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
 Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.

### 2021-04-21 codes/SIRfunctions.R

@@ -1,5 +1,4 @@
 library(shellpipes)
-library(matlib)
 
 # Functions for the SIR model
 # By: A Gharouni
@@ -123,7 +122,7 @@ Fi_hat<-function(params){
 
 # Basic Reproduction Number
 R0<-function(params){
-  unpack(as.list(c(state,params)))
+  unpack(as.list(params))
   Sn <- Sn_dfe(params)
   Su <- Su_dfe(params)
   Fi <- Fi_hat(params) #at DFE

### 2021-04-21 codes/sir_plot.R

@@ -5,6 +5,7 @@ library(directlabels)
 library(latex2exp)
 library(rootSolve)
 library(McMasterPandemic)
+library(matlib) # used for matrix inverse function
 
 library(shellpipes)
 
@@ -26,8 +27,6 @@ W_S_random <- 1
 W_S_targeted <- 0.3 
 
 # make dataframe 
-# df_random <- eval_R0(params = update(params,W_S=W_S_random))
-# df_targeted <- eval_R0(params = update(params,W_S=W_S_targeted))
 df_random <- make_params_dat(params = params,
                 eta_ws=0,eta_we=1, ## so theta_w
                 eta_cs=0,eta_ce=1, ## so theta_c

### 2021-04-22 codes/plots_suppl.R
clean up

@@ -1,0 +1,34 @@
+## create unified data frame
+library(ggplot2)
+library(dplyr)
+library(tidyr)
+library(directlabels)
+
+source("SIRfunctions.R")
+source("params.R")
+
+sup_df_random <- make_params_dat(params = params,
+                             eta_ws=1,eta_we=1, ## so theta_w=0
+                             eta_cs=0,eta_ce=0, ## so theta_c=1
+                             omega_s=0.5001,omega_e=0.5001,
+                             rho_s=0,rho_e=0.5)
+
+sup_df_targeted <- make_params_dat(params = update(params,W_S=0.3),
+                             eta_ws=1,eta_we=1, ## so theta_w=0
+                             eta_cs=0,eta_ce=0, ## so theta_c=1
+                             omega_s=0.5001,omega_e=0.5001,
+                             rho_s=0,rho_e=0.5)
+out2 <- sup_df_targeted %>%
+  pivot_longer(c(I_u,I_n,I_p,I_c), names_to = "compartment", values_to = "value")
+gg1 <- (ggplot(data=out2, aes(x=rho,y=value,col=compartment))
+        + theme_bw()
+        + geom_point()
+        + ylab(label="Population fraction at DFE")
+        + xlab(label="rho")
+        )
+direct.label(gg1,"last.bumpup")
+
+
+##plot of R0 wrt rho, notice the increase
+ggplot(data=out2, aes(x=rho,y=R0_sub))+geom_point()
+

### 2021-04-22 SIR_manuscript.tex
updated up to the result sec

@@
 
-The effect of testing processes, including (testing and test reporting) on an epidemic dynamics (infection and recovery) can be studied at the individual level or the community level (e.g., nursing homes, LTC facilities, etc.).
+The effect of testing processes, including (testing and test reporting) on an epidemic dynamics (infection and recovery) can be studied at the individual level or the community level (e.g., nursing homes, long-term-care facilities, etc.).
 Gaining insights to determine the sensitivity of the epidemic dynamics with respect to the testing processes will depend on underlying factors including the level of focus (individual or community), assumptions (model), and the interplay between these factors.
-In particular, the fast test reporting may be beneficial at the community-level, supported by many studies, as it gives a rapid assessment of the situation, identifies hot spots, and may enable rapid contact-tracing.
+In particular, the fast testing and test reporting may be beneficial at the community-level, supported by many studies, as it gives a rapid assessment of the situation, identifies hot spots, and may enable rapid contact-tracing.
 However, the potential advantage of a slow rate of test return on the dynamics of an epidemic is real, often neglected, and needs to be quantified.
@@
 Although the model development was motivated by the \covid epidemic, the model has general epidemiological and testing structures.
-The realistic components of the model include testing intensity, test sensitivity and specificity, rate of test return, and isolation.
+The realistic components of the model include testing intensity, test sensitivity and specificity, rate of test return, and isolation strength.
 The novel component is the compartment-specific relative testing weights, which reflect the testing strategies-- surveillance, diagnosis, or control.
-Here, we showed that the slow test reporting can be beneficial on the dynamics of an epidemic.
-In particular, it is possible for the basic reproduction number, $\Rnum$, to be increasing with respect to the rate of test return.
-Further, we compare two testing strategies, random vs.
-targeted, in this framework.
-We conclude that the targeted testing strategy is more effective than the random one in the sense that a lower range of testing intensity is required to keep $\Rnum <1$.
-Also, targeted testing reduces the individual-level of advantage of a slow rate of test return.
+Here, we compare two testing strategies, random vs.
+targeted, and concluded that the targeted testing strategy is more effective than the random one in the sense that a relatively lower range of testing intensity and slower test return time are required to keep $\Rnum <1$ compared to the random testing strategy.
+Furthermore, we show that increasing \percap testing intensity and reducing the test return time would be beneficial on the dynamics of an epidemic in general but there are exception cases.
+In particular, it is possible for the basic reproduction number, $\Rnum$, to be increasing with respect to the testing intensity and the rate of test return when the isolation efficacy for awaiting individuals is loose.
 
@@
 Symbolically, the testing status of an individual in the disease compartment $X$, where $X \in \{S,I,R\}$, is reflected in the subscript, namely $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
+Note that the top-to-bottom order of the testing-based compartments of the each disease-based compartmenet $X$ in the flowchart (Figure~\ref{fig:flowchart}) and the model equations \eqref{model} should match.
+However, compartments $X\_u$ and $X\_n$ are switched in the flowchart (Figure~\ref{fig:flowchart}) for the sake of tidiness.
 Further, two `accumulator' compartments, $N$ and $P$, were also incorporated in the model in order to collect cumulative reported negative or positive tests.
@@
 Table~\ref{tab:params} defines the model parameters, which are generally straightforward
-\percap flows between compartments, or modifiers to these flow rates. The \percap testing intensity is defined as the number of daily tests taken in a population of size $N_0$. 
+\percap flows between compartments, or modifiers to these flow rates.  
 The novel component of the model comes in through the compartment-specific relative testing weights $w_S$, $w_I$ and $w_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively.
@@
 \label{sigma}
-\sigma = \frac{\rho N_0}{W}.
+\sigma = \frac{\rho N_0}{W},
 \end{equation}
+where $\rho$ is the \percap testing rate for population and defined as the number of daily tests taken in a population of size $N_0$.
 Thus, the \percap testing rate for compartment $X$ is $F_X=\sigma w_X$, where $X \in \{S,I,R\}$. 
@@
   $N_0$     & Total population size & people & $10^6$ \\ \hline
-  $\omega$  & Rate of onward flow from ``waiting'' to ``confirmed'' or ``untested'' compartments  & 1/day & - \\ \hline
+  $\omega$  & Rate of test return, i.e., rate of onward flow from ``waiting'' to ``confirmed'' or ``untested'' compartments  & 1/day & - \\ \hline
   $\gamma$ & Recovery rate & 1/day & 1/3 \\ \hline 
-  $\rho$   & Per capita testing intensity & 1/day & 0.01 \\ \hline 
+  $\rho$   & \percap testing rate & 1/day & - \\ \hline 
   $\theta\_w$ & Isolation efficacy in reduction of the probability of transmission for ``waiting'' individuals & - & - \\ \hline
@@
 
-\todo{the algebra needs to be changed based on JD's comment} The Disease-Free Equilibrium (DFE) for the SIR model, Eqs.
-\eqref{model}, is given by setting the differential equations to 0 and solving for the unknowns (\ali{is this clear or the "unkowns" needs to be defined?
-also * in sup for I and R?}.
-The DFE is
+The Disease-Free Equilibrium, DFE hearafter, for the SIR model, Eqs. \eqref{model}, is given by setting the differential equations to 0 and solving for the unknowns. The DFE is
 \begin{equation}
@@
 \end{equation}
-For brevity, we write $x\_u^* = S\_u^*/N_0$ and $x\_n^* = S\_n^*/N_0$.
-We can then write 
+For brevity, we write 
 \begin{equation}
@@
 \end{equation}
+where the term $\frac{\beta}{\gamma}$ is the classical basic reproduction number for a SIR model without testing and isolation \todo{ref?}, and $\Delta$ is the reduction parameter defined as follows.
+\begin{equation}
+\label{eq:Delta1}
+\Delta= (A~ \theta\_w + B~ \theta\_c) / C,
+\end{equation}
 where
@@
 \label{eq:abc}
-\Delta=& (A~ \theta\_w + B~ \theta\_c) / C\\
-A=& \gamma \Big( \big(\gamma(\omega+\gamma) + \omega p_I \hat F_I \big) x_n^* + (\omega+\gamma) \hat F_I \Big)\\
-B=& \omega p_I(\gamma x_u^*+\omega)\hat F_I \\ 
+A=& \gamma \Big( \big(\gamma(\omega+\gamma) + \omega p_I \hat F_I \big) \frac{S\_n^*}{N_0} + (\omega+\gamma) \hat F_I \Big),\\
+B=& \omega p_I(\gamma \frac{S\_u^*}{N_0}+\omega)\hat F_I, \\ 
 C=& (\omega+\gamma) \Big(\gamma(\omega+\gamma)+(\gamma+\omega p_I)\hat F_I\Big).
@@
 \end{equation}
-
-Further details of derivation of $\Rnum$ are provided in the appendix section.
+Further details of derivation of $\Rnum$ are provided in the appendix section \ref{app:R0}.
 
 The analytical calculation of the next-generation matrix and the derivation of $\Rnum$ was carried out in Maple \citep{maple14}. 
-We used R \citep{r} for simulations and for plotting $\Rnum$ as a function of the underlying parameters (using \ref{R0}).
-Results are shown in Figure \ref{pan}, where  panel \eqref{p.a} represents the random testing, and panel \eqref{p.b} is representing non-random testing.
-The parameter values or ranges are presented in Table~\ref{tab:params}.
-This simulation reflects the behavior of $\Rnum$ with respect to the selected parameters for two different testing strategies: (i) random testing, represented by all testing weights to be the same, $w_S=w_I=w_R$, and (ii) non-random testing, when testing weight are not equal.
-For simulation purposes we chose $w_S=w_I=w_R=1$ for random testing, and $w_S=0.3$ and $w_I=w_R=1$ for non-random testing strategy.
-Note that the critical contour of $\Rnum=1$ is plotted in solid line in Figure \ref{pan}.
+We used R \citep{r} for simulations and for plotting $\Rnum$ as a function of the underlying parameters.
+In particular, the contours of $\Delta$ \eqref{eq:Delta1} were plotted for a combination set of parameters with the baseline parameter values are presented in Table~\ref{tab:params} and isolation efficacy parameters, $\theta\_w$ and $\theta\_c$ vary from 0 to 1.
+Results are shown in Figures \ref{pan} and \ref{pan2}, where  panel (a) represents the random testing ($w_S=w_I=w_R=1$), and panel (b) is representing non-random testing ($w_S=0.3$ and $w_I=w_R=1$). 
+Fig. \ref{pan} reflects the changes in $\Rnum$  when the \percap testing rate is small relative to the population size, $\rho \in [0,0.01]$, and the test return rate $\omega\in [0.1,2]$.  
+Fig. \ref{pan2} reflects the changes in $\Rnum$  when the \percap testing rate is large relative to the population size, $\rho \in [0,0.5]$, and the test return rate $\omega\in (0.5,2]$. 
+Note that the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, corresponding to $\Rnum=1$, is plotted in solid line in the Figures.
 
@@
 The explicit formula for the basic reproduction number, $\Rnum$ \eqref{R0}, provides an opportunity to study the influence of changes in the underlying parameters on the critical index of epidemic dynamics.
-We are interested in exploring the effect of parameters that could be manipulated by public-health policy, including the isolation parameters ($\theta\_c$ and $\theta\_w$) testing intensity ($\rho$) and test-return rate ($\omega$).
+We are interested in exploring the effect of parameters that could be manipulated by public-health policy, including the isolation efficacy parameters, $\theta\_c$ and $\theta\_w$, \percap testing intensity, $\rho$, and the test return rate, $\omega$.
 
@@
 
-\subsection{Model and calculation of $\Rnum$}
+\subsection{Model and calculation of $\Rnum$}\label{app:R0}
 

### 2021-04-22 SIR_manuscript.tex
ref added

@@
 \end{equation}
-where the term $\frac{\beta}{\gamma}$ is the classical basic reproduction number for a SIR model without testing and isolation \todo{ref?}, and $\Delta$ is the reduction parameter defined as follows.
+where the term $\frac{\beta}{\gamma}$ is the classical basic reproduction number for a SIR model without testing and isolation (see, e.g., \cite{keeling2011modeling}), and $\Delta$ is the reduction parameter defined as follows.
 \begin{equation}

### 2021-04-23 SIR_manuscript.tex
more on intro and res

@@
 Furthermore, we show that increasing \percap testing intensity and reducing the test return time would be beneficial on the dynamics of an epidemic in general but there are exception cases.
-In particular, it is possible for the basic reproduction number, $\Rnum$, to be increasing with respect to the testing intensity and the rate of test return when the isolation efficacy for awaiting individuals is loose.
+In particular, it is possible for the basic reproduction number, $\Rnum$, to be increasing with respect to the testing intensity and the rate of test return when the isolation efficacy in reduction of the probability of transmission for awaiting individuals is loose.
 
@@
 In order to understand the effect of testing processes on an epidemic dynamics, we deveopled a mechanistic SIR-type model with epidemic and testing components.
-Here, we focus on the the effect of testing intensity, different levels of testing ``focus'' (from random to highly targeted), and rate of test return on the epidemic dynamics.
-Our model provides insights to the sensitivity of the epidemiological dynamics, through $\Rnum$, with respect to the undelying epidemic and testing parameters.
-\ali{edited in response to David comments.}
+This model provides a sensible platform to link the modeling of epidemic and testing components and study their interaction.
+Here, we studied the the effect of testing intensity, rate of test return and the isolation efficacy in reduction of the probability of transmission on the epidemic dynamics when different levels of testing ``focus'' (from random to highly targeted) are in place.
+Our model provides insights to the sensitivity of the epidemic dynamics, through $\Rnum$, with respect to the undelying testing and isolation parameters.\ali{edited in response to David comments.}
 
@@
 Symbolically, the testing status of an individual in the disease compartment $X$, where $X \in \{S,I,R\}$, is reflected in the subscript, namely $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
-Note that the top-to-bottom order of the testing-based compartments of the each disease-based compartmenet $X$ in the flowchart (Figure~\ref{fig:flowchart}) and the model equations \eqref{model} should match.
-However, compartments $X\_u$ and $X\_n$ are switched in the flowchart (Figure~\ref{fig:flowchart}) for the sake of tidiness.
+Note that the top-to-bottom order of the testing-based compartments of each disease-based compartmenet $X$ in the flowchart (Figure~\ref{fig:flowchart}) and the model equations \eqref{model} should match.
+However, we switched  $X\_u$ and $X\_n$ in the flowchart (Figure~\ref{fig:flowchart}) for the sake of tidiness of the flowchart.
 Further, two `accumulator' compartments, $N$ and $P$, were also incorporated in the model in order to collect cumulative reported negative or positive tests.
-The model and details of calculation of the basic reproduction number $\Rnum$ are presented in the appendixs.
+The model equations and details of calculation of the basic reproduction number $\Rnum$ are presented in the appendix (see Sec.
+\ref{app:R0}).
 
@@
 Here, the disease-based status of a compartment $X$, where $X \in \{S,I,R\}$, is combined with the testing-based status including $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
-Also,  $\Lambda$ is the force of infection with definition in Eq.~\eqref{Lambda}, $\gamma$ is the recovery rate, $\omega$ is the rate of test return, $F_X$ and $p_X$ represent the \percap testing rate and the probability of positive tests, respectively, for compartment $X$.
+Also,  $\Lambda$ is the force of infection with definition in Eq.~\eqref{Lambda}, $\gamma$ is the recovery rate, $\omega$ is the rate of test return, $F_X$ (defined in Eq.
+\eqref{F}) and $p_X$ represent the \percap testing rate and the probability of positive tests, respectively, for compartment $X$.
 For further description of the parameters see \cref{tab:params}.
@@
 The novel component of the model comes in through the compartment-specific relative testing weights $w_S$, $w_I$ and $w_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively.
+Thus, we can spesify different levels of testing “focus” from random to highly targeted.
 For example, $w_I/w_S=2$ means that infected individuals are tested at twice the \percap rate of susceptible individuals.
@@
 \end{equation}
-where $\rho$ is the \percap testing rate for population and defined as the number of daily tests taken in a population of size $N_0$.
-Thus, the \percap testing rate for compartment $X$ is $F_X=\sigma w_X$, where $X \in \{S,I,R\}$. 
+where $\rho$ is the \percap testing intensity for population and defined as the number of daily tests taken in a population of size $N_0$.
+Thus, the \percap testing rate for compartment $X$ is 
+\begin{equation}
+\label{F}
+F_X=\sigma w_X,\,\, \text{where}\, X \in \{S,I,R\}.
+\end{equation}
 For a high-sensitivity test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus unavailable for further testing.
@@
 \end{equation}
-across all susceptible pools with transmission rate $\beta$ and isolation efficacy in reduction of the probability of transmission for ``waiting'' and \emph{confirmed positive} individuals, $\theta\_w$ and $\theta\_c$ respectively, (see Table~\ref{tab:params} for further details), (ii) $\theta\_c \geq \theta\_w$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
+across all susceptible pools with transmission rate $\beta$ and isolation efficacy in reduction of the probability of transmission for three testing-based compartments ``waiting'' and \emph{confirmed positive} individuals, $\theta\_w$ and $\theta\_c$ respectively, (see Table~\ref{tab:params} for further details), (ii) $\theta\_w\leq\theta\_c$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
 Thus, for instance when the awaiting people follow the isolation perfectly, $\theta\_w$ is closer to 1, while when they less follow the isolation, $\theta\_w$ is closer to 0.
 For this analysis, we also assume a perfectly specific test ($p_S=0$).
-This last assumption combined with the assumption that no individual is in waiting-for-\emph{positive} or \emph{confirmed positive} compartments, i.e., $S\_p(0)=S\_c(0)=0$, reduces the model to 10 equations (equations c and d  in \eqref{model} are eliminated).
+This last assumption combined with the assumption that no susceptible individual is in waiting-for-\emph{positive} or \emph{confirmed positive} compartments, i.e., $S\_p(0)=S\_c(0)=0$, reduces the model to 10 equations where equations c and d in model \eqref{model} are eliminated.
 
@@
   $\gamma$ & Recovery rate & 1/day & 1/3 \\ \hline 
-  $\rho$   & \percap testing rate & 1/day & - \\ \hline 
+  $\rho$   & \percap testing intensity & 1/day & - \\ \hline 
   $\theta\_w$ & Isolation efficacy in reduction of the probability of transmission for ``waiting'' individuals & - & - \\ \hline
@@
 \end{equation}
-For brevity, we write 
+The corresponding \percap testing rate (Eq. \ref{F}) for the infected compartment $I$ at DFE is one of the key analysis parameters and can be simplified as 
 \begin{equation}
 \label{eq:fi}
-\hat F_I = (\omega\rho/(\omega-\rho))w_I/w_S
+\hat F_I = (\omega\rho/(\omega-\rho))w_I/w_S.
 \end{equation}
-for the \percap testing rate of infectious individuals near the disease-free equilibrium.
 
@@
 \end{equation}
-Further details of derivation of $\Rnum$ are provided in the appendix section \ref{app:R0}.
+Further details of derivation of $\Rnum$ are provided in section \ref{app:R0} of the appendix.
+Note that $\Delta$ has been presented in a specific monotonic form with respect to isolation efficacy parameters, $\theta\_w$ and $\theta\_c$, in order to simplify showing one of the results of this paper which is the reduction effect of isolation on $\Rnum$.
+Further, we represent $\Delta$ in another equivalent form in the appendix (Sec.
+\ref{app:rho}) to show the nontrivial-nonmonotonic effect of testing intensity and test return rate on $\Rnum$.
 
@@
 We used R \citep{r} for simulations and for plotting $\Rnum$ as a function of the underlying parameters.
-In particular, the contours of $\Delta$ \eqref{eq:Delta1} were plotted for a combination set of parameters with the baseline parameter values are presented in Table~\ref{tab:params} and isolation efficacy parameters, $\theta\_w$ and $\theta\_c$ vary from 0 to 1.
+In particular, the contours of $\Delta$ \eqref{eq:Delta1} were plotted as in a range of a selected set of parameters of interest, $1/\omega$, $\rho$, $\theta\_w$ and $\theta\_c$, with the other parameters values kept fixed and presented in Table~\ref{tab:params}.
+
+and isolation efficacy parameters, $\theta\_w$ and $\theta\_c$ vary from 0 to 1.
 Results are shown in Figures \ref{pan} and \ref{pan2}, where  panel (a) represents the random testing ($w_S=w_I=w_R=1$), and panel (b) is representing non-random testing ($w_S=0.3$ and $w_I=w_R=1$). 
@@
 The explicit formula for the basic reproduction number, $\Rnum$ \eqref{R0}, provides an opportunity to study the influence of changes in the underlying parameters on the critical index of epidemic dynamics.
-We are interested in exploring the effect of parameters that could be manipulated by public-health policy, including the isolation efficacy parameters, $\theta\_c$ and $\theta\_w$, \percap testing intensity, $\rho$, and the test return rate, $\omega$.
+We are interested in exploring the effect of parameters that could be manipulated by public-health policy, including the isolation efficacy parameters, $\theta\_c$ and $\theta\_w$, \percap testing intensity, $\rho$, and the rate of test return, $\omega$.
 
@@
 $$\partial{\Rnum}/\partial{\theta\_c} \leq 0 \text{~and~} \partial{\Rnum}/\partial{\theta\_w} \leq 0.$$ 
-\item \label{p1:rho} a higher rate of testing may increase or decrease $\Rnum$;
+\item \label{p1:rho} a higher testing intensity may increase or decrease $\Rnum$;
 $$\partial{\Rnum}/\partial{\rho} \text{~can be positive or negative}.$$
-\item \label{p1:omega} A higher rate of of test return may increase or decrease \Rnum.
+\item \label{p1:omega} A higher rate of test return may increase or decrease \Rnum.
 $$\partial{\Rnum}/\partial{\omega} \text{~can be positive or negative}.$$
@@
 One way to do such an analysis is based on the fact that $\rho$ is \percap rate so for a large population size, $N_0 \gg 1$, $\rho$ is very small or close to 0, comparing to $N_0$.
-This provides a base to linearly approximate $\Rnum$ when $\rho$ is close to 0, and use this approximation to analyze the behaviour of $\Rnum$ with respect to $\omega$ (see section \ref{taylor}).
+This provides a base to linearly approximate $\Rnum$ when $\rho$ is close to 0, and use this approximation to analyze the behaviour of $\Rnum$ with respect to $\omega$ (see section \ref{app:omega}).
 In the next section we provide an equivalent representation of $\Rnum$ to provide a ground to prove that more testing intensity decreases $\Rnum$ for a general rane of parameters.
@@
 % %%%%%%%
-\subsection{More testing intensity may decreases $\Rnum$}
+\subsection{More testing intensity may decreases $\Rnum$}\label{app:rho}
 
@@
 % %%%%%%%
-\subsection{rate of returning tests} \label{taylor}
+\subsection{rate of returning tests} \label{app:omega}
 \ali{needs editig}

### 2021-04-23 SIR_manuscript.tex
more updates on method sec

@@
   $\theta\_c$ & Isolation efficacy in reduction of the probability of transmission for ``confirmed positive'' individuals & - & -  \\ \hline
-  $\beta$ & Transmission rate & 1/day & 0.39 \\ \hline
+  $\beta$ & Transmission rate & 1/day & 0.339 \\ \hline
   $\Lambda$ & Force of infection & 1/day & - \\ \hline
@@
 
-The analytical calculation of the next-generation matrix and the derivation of $\Rnum$ was carried out in Maple \citep{maple14}. 
-We used R \citep{r} for simulations and for plotting $\Rnum$ as a function of the underlying parameters.
-In particular, the contours of $\Delta$ \eqref{eq:Delta1} were plotted as in a range of a selected set of parameters of interest, $1/\omega$, $\rho$, $\theta\_w$ and $\theta\_c$, with the other parameters values kept fixed and presented in Table~\ref{tab:params}.
-
-and isolation efficacy parameters, $\theta\_w$ and $\theta\_c$ vary from 0 to 1.
-Results are shown in Figures \ref{pan} and \ref{pan2}, where  panel (a) represents the random testing ($w_S=w_I=w_R=1$), and panel (b) is representing non-random testing ($w_S=0.3$ and $w_I=w_R=1$). 
-Fig. \ref{pan} reflects the changes in $\Rnum$  when the \percap testing rate is small relative to the population size, $\rho \in [0,0.01]$, and the test return rate $\omega\in [0.1,2]$.  
-Fig. \ref{pan2} reflects the changes in $\Rnum$  when the \percap testing rate is large relative to the population size, $\rho \in [0,0.5]$, and the test return rate $\omega\in (0.5,2]$. 
-Note that the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, corresponding to $\Rnum=1$, is plotted in solid line in the Figures.
+The analytical calculation of the next-generation matrix and simplifying the expression of $\Rnum$ was carried out in Maple \citep{maple14}. 
+We used R \citep{r} for simulations, in particular,  for plotting the contours of $\Delta$ \eqref{eq:Delta1} over a range of selected set of parameters of interest, i.e., parameters that could be manipulated by public-health policy.
+These parameters include the isolation efficacy parameters, $\theta\_c$ and $\theta\_w$, \percap testing intensity, $\rho$, and the rate of test return, $\omega$.
+The rest of model parameters kept fixed at sensible values (see Table~\ref{tab:params} for the parameter values).
+Note that all plots are illustrated in the scale of the mean test return time $1/\omega$ (day).
+Results are shown in Fig.\,\ref{pan} and Fig.\,\ref{pan2} with two panels; panel (a) represents the random testing, thus $w_S=w_I=w_R=1$, and panel (b) represents targeted testing, thus $w_S=0.3$ and $w_I=w_R=1$.
+To illustrate the changes in $\Rnum$ with respect to \percap testing intensity$\rho$, two sets of plots are presented. 
+Fig.\,\ref{pan} reflects the changes in $\Rnum$ when $\rho$, is small relative to the population size, specifically, $\rho \in [0,0.01]$, and the test return rate $\omega\in [0.1,2]$.
+This is a more realistic scenario of testing as we have observed in \covid pandemic.
+That is, 1\% of a population are tested per day at maximum capacity, so in a population of size $N_0=10^6$, 10000 individuals will be tased per day at maximum.
+In order to illustrate the non-monotonic changes in $\Rnum$ with respect to $\rho$,
+Fig. \ref{pan2} with maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,0.5]$, and the test return rate $\omega\in (0.5,2]$. 
+In both Fig.\,\ref{pan} and Fig.\,\ref{pan2}, the isolation efficacy parameters, $\theta\_w$ and $\theta\_c$ vary from 0 to 1.
+Note that the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, corresponding to the threshold $\Rnum=1$, is plotted in solid line in the panels.
+Thus, a combination of $\rho$ and $\omega$ above this critical contour results in reducing $\Rnum$.
+Also realize that $\Rnum=\frac{\beta}{\gamma}$, about 1.017 here, for the classical SIR model which can be acheived in the context of our model when $\rho=0$, no testing, or $\theta\_w=\theta\_w=0$, no isolation.
 

### 2021-04-23 codes/SIRfunctions.R

@@ -160,7 +160,7 @@ make_params_dat<-function(params,
                           omega_s,omega_e,
                           eta_ws,eta_we,
                           eta_cs,eta_ce,
-                          tol=1e-10,## tol: resolve the issue of very small numbers in plotting and when rho is close to omega
+                          tol=1e-8,## tol: resolve the issue of very small numbers in plotting and when rho is close to omega
                           n_out=5,## n_out: facets (rows/cols)
                           n_in=41 ## n_in: grid N within facets
 ){
@@ -175,7 +175,7 @@ make_params_dat<-function(params,
                      W_S=W_S,W_I=W_I,W_R=W_R,
                      p_S=p_S,p_I=p_I,p_R=p_R,
                      eta_w=eta_w,eta_c=eta_c)
-  ## principle eigenvector
+    ## principle eigenvector
   eigvec <- t(apply(df1,1,function(params_in)eigvec_max(params=params_in)))  
   dfout <-(df1 %>% 
              dplyr::mutate(

### 2021-04-23 codes/params.R

@@ -7,9 +7,9 @@ loadEnvironments()
 # Path: ~/projects/SIRmodel/codes
 
 params <- c(
-N0=1000000, #total population size
-beta= 0.339, #transmission rate (was 0.8) 0.339
-gamma=1/3, #recovery rate
+N0=1000000, #total population size, 
+beta= 0.339, #transmission rate 0.339
+gamma=1/3, #recovery rate, was 1/3
 omega=0.25, #test returning rate 
 rho=0.01, #testing rate percapita (also used 0.8, 1/3)
 W_S=1, W_I=1, W_R=1, #Testing relative weights

### 2021-04-23 codes/sir_plot.R

@@ -38,6 +38,7 @@ df_targeted <- make_params_dat(params = update(params,W_S=W_S_targeted),
                 eta_cs=0,eta_ce=1, ## so theta_c
                 omega_s=0.1,omega_e=2,
                 rho_s=0,rho_e=0.01)
+
 ##important contour, ie R0=1 thus threshold=1 when plotting R0 contours, or Delta(R0=1)
 threshold <- 1-(params[["gamma"]]/params[["beta"]]) ## Or 0?  corresponding to R0=1 
 # #################################
@@ -61,9 +62,7 @@ bins <- cut(sort(unique(c(df_random$Delta,df_targeted$Delta))),
 brks <- levels(bins)
 brks_vec <- unique(as.numeric(unlist(lapply(strsplit(brks, ","), function(x) gsub("\\(|]", "", x)))))
 brks_vec <- sort(ifelse(brks_vec[]< 0,0,brks_vec[]), decreasing = TRUE) ## replace neg with 0 and reorder
-# show_contours_1, width=24,height=24,message=FALSE,warning=FALSE}
-# brks_vec <- seq(1,0,by=-0.2) # Break vector for unifying the legends in Random and TTI testing cases. 
-# seq(.165,0,by=-0.03)
+
 df_temp <- df_random
 # df_temp <- df_targeted
 p1 <- (ggplot(df_temp,aes(x=1/omega,y=rho,z=Delta))

### 2021-04-24 SIR_manuscript.tex
abst and method updated, minor

@@
 Although the model development was motivated by the \covid epidemic, the model has general epidemiological and testing structures.
-The realistic components of the model include testing intensity, test sensitivity and specificity, rate of test return, and isolation strength.
+The realistic components of the model include \percap testing intensity, test sensitivity and specificity, rate of test return, and isolation efficacy in reduction of the probability of transmission.
 The novel component is the compartment-specific relative testing weights, which reflect the testing strategies-- surveillance, diagnosis, or control.
 Here, we compare two testing strategies, random vs.
-targeted, and concluded that the targeted testing strategy is more effective than the random one in the sense that a relatively lower range of testing intensity and slower test return time are required to keep $\Rnum <1$ compared to the random testing strategy.
+targeted, and concluded that the targeted testing strategy is more effective in the sense that achieving reduction of $\Rnum$ can occure in a lower range of testing intensity and longer test return time relative to the random testing.
 Furthermore, we show that increasing \percap testing intensity and reducing the test return time would be beneficial on the dynamics of an epidemic in general but there are exception cases.
-In particular, it is possible for the basic reproduction number, $\Rnum$, to be increasing with respect to the testing intensity and the rate of test return when the isolation efficacy in reduction of the probability of transmission for awaiting individuals is loose.
+In particular, it is possible for the basic reproduction number, $\Rnum$, to be increasing with respect to the \percap testing intensity and the rate of test return when the isolation efficacy in reduction of the probability of transmission for awaiting individuals is loose.
 

### 2021-04-26 SIR_manuscript.tex
results were updated, typo in sir function eigvec_max() fixed

@@
 To illustrate the changes in $\Rnum$ with respect to \percap testing intensity$\rho$, two sets of plots are presented. 
-Fig.\,\ref{pan} reflects the changes in $\Rnum$ when $\rho$, is small relative to the population size, specifically, $\rho \in [0,0.01]$, and the test return rate $\omega\in [0.1,2]$.
+Fig.\,\ref{pan} reflects the changes in $\Rnum$ when $\rho$ is small relative to the population size.
+Specifically, $\rho \in [0,0.01]$, and the test return rate $\omega\in [0.1,2]$.
 This is a more realistic scenario of testing as we have observed in \covid pandemic.
@@
 We are interested in exploring the effect of parameters that could be manipulated by public-health policy, including the isolation efficacy parameters, $\theta\_c$ and $\theta\_w$, \percap testing intensity, $\rho$, and the rate of test return, $\omega$.
+In this context, two testing strategies, namely random and targeted testing, were explored.
 
 Examining our formula for $\Rnum$ \eqref{R0} gives the following results. 
-See the appendix for details.
+See the appendix, Sec.\,\ref{app:rho} and Sec.\,\ref{app:omega} for details.
 
 \begin{enumerate}
-\item \label{p1:eta} increasing the isolation parameters for tested and confirmed individuals decreases \Rnum;
+\item \label{p1:eta} Increasing the isolation parameters for tested and confirmed individuals decreases \Rnum;
 $$\partial{\Rnum}/\partial{\theta\_c} \leq 0 \text{~and~} \partial{\Rnum}/\partial{\theta\_w} \leq 0.$$ 
-\item \label{p1:rho} a higher testing intensity may increase or decrease $\Rnum$;
+\item \label{p1:rho} A higher testing intensity may increase or decrease $\Rnum$;
 $$\partial{\Rnum}/\partial{\rho} \text{~can be positive or negative}.$$
@@
 
-Our numerical results are shown in Figure \ref{pan}.
-It can be inferred that when random testing strategy is applied, Figure \ref{p.a}, comparing to the targeted testing strategy, Figure \ref{p.b}, the critical \percap testing intensity of the corresponding panels are lower in non-random testing.
-Also, it appears that speeding the test reporting, i.e., increasing $\omega$, does not significantly lower $\Rnum$ when awaiting people (including people in $I\_p$ and $I\_n$ compartments) follow the isolation perfectly, i.e., $\theta\_w$ is closer to 1.
+\begin{figure}[h!]
+\centering
+\begin{subfigure}[t]{.45\textwidth}
+\centering
+\includegraphics[width=\linewidth]{codes/R0contour_random_rho.5.pdf}
+\caption{}
+\end{subfigure}
+%
+\begin{subfigure}[t]{.45\textwidth}
+\centering
+\includegraphics[width=\linewidth]{codes/R0contour_TTI_rho.5.pdf}
+\caption{}
+\end{subfigure}
+\caption{
+{\bf A comparison of the behaviour of the basic reproduction number, $\Rnum$, between random versus targeted testing strategies at different levels of testing and isolation.}
+We numerically evaluate $\Delta$ \eqref{R0}, reflecting the reduction of $\Rnum$ with respect to testing and isolation. We use the following parameters (listed in Table \ref{tab:params}):
+$N_0=1 \times 10^6$, $\omega \in (0.5,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.5]$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.39$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
+Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
+The black solid line in each panel represents the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ corresponding to the threshold of $\Rnum=1$.
+\bmb{we still have some collisions in the y-axis tick labels}
+}
+\label{pan2}
+\end{figure}
+
+Our numerical results are shown in Figs. \ref{pan}, where $\rho \in [0,0.01]$, and \ref{pan2}, where $\rho \in [0,0.5]$.
+When targeted testing strategy is applied compared to random testing strategy, i.e., comparing panel (b) to (a) of Figs.\,\ref{pan} and \ref{pan2}, it can be inferred that 
+(i) \ali{thinking of the intercept of the solid line/contours} the critical \percap testing intensity corresponding to $\Rnum=1$, is lower and thus less testing is required to reduce $\Rnum$, 
+(ii) \ali{thinking of the slope contours, having hard time to explain this since there is a non-monotonic effect of omega} sensitivity of $\Rnum$ wrt $1/\omega$; the changes in $\Rnum$ in response to changes in $1/\omega$, is lower.
+
+It is notable that increasing testing intensity $\rho$ as a strategy to reduce $\Rnum$ depends on the value of $\rho$ to begin with, the mean test return time $1/\omega$ and the efficacy of isolation $\theta$.
+In particular, there might be feasible threshold values for $\rho$ and $\omega$ that maximizes the effect of these parameters in reduction of $\Rnum$ depends upon the isolation efficacy.
+
+\ali{Thinking of last column of 2 Figs, moving from top tp bottom as $\theta\_c=1$ and $\theta\_w$ changes from 0 to 1} When $\theta\_c$ is closer to 1 and $\theta\_w$ is closer to 0, increasing $\rho$ combined with deccreasing $1/\omega$ is effective in reducing $\Rnum$.
+However, as $\theta_w$ is getting closer to 1, increasing $\rho$ combined with increasing $1/\omega$ is better.
+
+% When $\rho$ is closer to 0 (Fig.\,\ref{pan}),  and $\theta_c$ is closer to 0, increasing $\rho$ and decreasing $1/\omega$ is a    
+% 
+% when $\rho$ is small relative to $N_0$, increasing $\rho$ combined with changes in the test return time $1/\omega$ is always good in general.  
+% 
+% When the isolation efficay is low, i.e., $\theta\_w$ is closer to 0, speeding up the test reporting processing time can increase $\Rnum$ regardless of the speed of testing.
+However there is threshold for $1/\omega$ above which prolonging the test return time
+
+% awaiting people (including people in $I\_p$ and $I\_n$ compartments) follow the isolation perfectly, i.e., $\theta\_w$ is closer to 1.
 However, speeding the test reporting reduces the epidemic more when the awaiting people less follow the isolation, i.e., $\theta\_w$ is closer to 0.
-
-Furthermore, we derived an inequality that quantifies the exact relationships between model parameters that result in returning tests more rapidly being favorable \eqref{eq:necsuf}.
-\david{What do we learn from this inequality?  Why is it relegated to an appendix?}
+% 
+% Furthermore, we derived an inequality that quantifies the exact relationships between model parameters that result in returning tests more rapidly being favorable \eqref{eq:necsuf}.
+% \david{What do we learn from this inequality?  Why is it relegated to an appendix?}
 
@@
 
-\subsection{Simulation with $\rho$ varies from 0 to 0.5 }
-
-\begin{figure}[h!]
-\centering
-\begin{subfigure}[t]{.45\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_random_rho.5.pdf}
-\caption{}
-\end{subfigure}
-%
-\begin{subfigure}[t]{.45\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_TTI_rho.5.pdf}
-\caption{}
-\end{subfigure}
-\caption{
-{\bf A comparison of the behaviour of the basic reproduction number, $\Rnum$, between random versus targeted testing strategies at different levels of testing and isolation.}
-We numerically evaluate $\Delta$ \eqref{R0}, reflecting the reduction of $\Rnum$ with respect to testing and isolation. We use the following parameters (listed in Table \ref{tab:params}):
-$N_0=1 \times 10^6$, $\omega \in (0.5,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.5]$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.39$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
-Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
-The black solid line in each panel represents the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ corresponding to the threshold of $\Rnum=1$.
-\bmb{we still have some collisions in the y-axis tick labels}
-}
-\label{pan2}
-\end{figure}
-
 \end{document}

### 2021-04-26 codes/SIRfunctions.R

@@ -139,7 +139,7 @@ eigvec_max<-function(params){
   Sn<-Sn_dfe(params)
   Su<-Su_dfe(params)
   Fi<-Fi_hat(params) 
-  colvec<-matrix(c(Sn,Su,0,0),4,1)  
+  colvec<-matrix(c(Su,Sn,0,0),4,1)  
   rowvec<-matrix(c(1,eta_w,eta_w,eta_c),1,4)
   Fmat<-beta/N0*colvec%*%rowvec
   vmat<-matrix(c(Fi+gamma,-omega,0,0,

### 2021-04-26 SIR_manuscript.tex
discussion

@@
 Here, we develop a compartmental SIR-type model to study the potential effect of testing strategies, testing intensity, test sensitivity and specificity, test reporting time and isolation on epidemic dynamics.
-While targeted testing strategies 
-\david{you need to be clearer: you mean targeting individuals who have contacted people who were infected.
-For example, here you could say something like ``While targeting the contacts of confirmed cases\dots''}
-are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is that viral spread
-is greater for a slow test than for a fast test.
+While targeted testing strategies, including targeting people with infection-like symptoms or the contacts of confirmed cases, are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is that viral spread is greater for a slow test than for a fast test.
 This counter-intuitive effect can occur when people are cautious when awaiting a test result, and may not be robust to second-order effects of fast testing (such as better contact tracing).
+\ali{this is a valid sentence especially when $\theta\_w$ is closer to 1, but not sure if this was counter-intuitive or was the opposite of it?
+i.e., that viral spread is greater for a fast test ($1/\omega$ closer to 0) than for a slow test.}
 
@@
 Note that we have not specified a methodology to assign particular relative testing weights corresponding to a particular targeted testing scenario.
-This part needs to be developed further in future work. 
 Modeling different targeted testing strategies, equivalently test-specific testing weights in our framework, requires prior information of the conditional probabilities of getting tested for people in a given compartment.
@@
 For example, when people are tested for "screening", the individuals with potential higher mobility, eg.
-people who are getting on flights, get more tested and thus the coresponding heavier testing weight is assigned than people awaiting for a surgery and are probably going to stay in a long-term care facility and consequently less mobile and more isolated to begin with.
-With our model, we would be able to compare the sensitivity of the disease dynamics, through $\Rnum$, with respect to testing processes.
+people who are getting on flights, get more tested and thus the coresponding heavier testing weight is assigned than people awaiting for a surgery and are probably going to stay in a long-term-care facility and consequently less mobile and more isolated to begin with.
+With our model, we would be able to compare the sensitivity of the disease dynamics, through $\Rnum$, with respect to testing different high-risk groups in the population.
+This part needs to be developed further in future work.
 
 {\bf The \percap testing intensity, $\rho$;}
-Proposition \ref{prop1} part \ref{p1:rho} and also Figure \ref{pan} indicates that increasing testing intensity $\rho$ reduces $\Rnum$.
-This is sensible since as people are moved to test compartments, namely $X_p$ and $X_n$ for $X \in \{S,I,R\}$, they may partially or fully self-isolate.
-Also individuals who have tested positive, namely $X_c$ compartment for $X \in \{S,I,R\}$, are highly likely to self-isolate.
-The higher probability of being subject to isolation, the lower the force of infection \eqref{Lambda} and the lower $\Rnum$ will become.
-While our simulation supported this result, analytically it appeared to be hard to conclude this due to the complexity of $\Rnum$ expression.
+Proposition \ref{p1:rho} and also Figures \ref{pan} and \ref{pan2} indicates that increasing testing intensity $\rho$ may reduce $\Rnum$ but there is a counter-intuitive machanism. 
+The expected mechanism was that as more untested infected individuals are moved to the test compartments, i.e., from $I\_u$ to $I\_p$ and $I\_n$, and eventually to the confirmed compartment $I\_c$, they may partially or fully self-isolate ($\theta\_w$ and $\theta\_c$ varies between 0 and 1 with $\theta\_w\leq\theta\_c$).
+Thus, the higher the \percap testing intensity $\rho$, the lower the force of infection \eqref{Lambda} and the lower $\Rnum$ will become.
+This is true when $\rho$ is very small relative to $N_0$ and can be verified analytically and through our simulation.
+The counter-intuitive mechanism is that at the DFE, higher \percap testing intensity, may delay recognition of true infected individuals $I\_p$ and causing an increase in $\Rnum$ as follows.
+At high $\rho$, more people will be moved from $S\_u^*$ to $S\_n^*$, thus a higher proportion of awaiting people may get infected and end up in $I\_n$, and be delayed from being eligable for retesting $I\_u$ and eventually to $I\_p$ and $I\_c$.
+Our analysis (Sec.\,\ref{app:rho}) and simulation (see Fig.\,\ref{pan2} panel (b) when $\theta\_c=1$ and $\theta\_w=0$) support such a mechanism.
+In particular, if \eqref{ineq:a3} holds we will have such a result.
 

### 2021-04-27 SIR_manuscript.tex
seperating params and the functions, make rda, create make rull for functions, minor changes in mans results sec

@@
 
-The explicit formula for the basic reproduction number, $\Rnum$ \eqref{R0}, provides an opportunity to study the influence of changes in the underlying parameters on the critical index of epidemic dynamics.
-We are interested in exploring the effect of parameters that could be manipulated by public-health policy, including the isolation efficacy parameters, $\theta\_c$ and $\theta\_w$, \percap testing intensity, $\rho$, and the rate of test return, $\omega$.
+The explicit formula for the basic reproduction number, $\Rnum$ \eqref{R0}, provides an opportunity to study the influence of changes in the underlying parameters of interest on the critical index of epidemic dynamics.
+These parameters include the isolation efficacy parameters, $\theta\_c$ and $\theta\_w$, \percap testing intensity, $\rho$, and the rate of test return, $\omega$.
 In this context, two testing strategies, namely random and targeted testing, were explored.

### 2021-04-27 note/SIR_scrap.tex
seperating params and the functions, make rda, create make rull for functions, minor changes in mans results sec

@@
 
+\section{Unused Scripts}
+# Conditional stop for ODE desolver
+# rootfun <- function (time, state,params) {
+#   unpack(as.list(c(state,params)))
+#   return(I_u - 1) }
+
+# ode(
+#   func=sir.model,
+#   y=state_init,
+#   times= d,
+#   parms=params,
+#   atol = 1e+1, rtol = 1e+1
+#   rootfun = rootfun,
+#   method="lsodar"
+# )
+
+## R0 expression with eta_w/eta_c factored
+# R02 <- function(state=state_dfe,params){
+#   unpack(as.list(c(state,params)))
+#   Sn <- Sn_dfe(params)
+#   Su <- Su_dfe(params)
+#   Fi <- Fi_hat(params) #at DFE
+#
+#   A2 <- gamma*(p_I*Fi*omega*Sn+(omega+gamma)*(gamma*Sn+Fi*N0))
+#   B2 <- p_I*Fi*omega*(gamma*Su+omega*N0)
+#   D <- gamma*(omega+gamma)*(gamma*Su+omega*N0)
+#   C2 <- beta/((N0*gamma*(gamma*(omega+gamma)+Fi*(gamma+omega*p_I)))*(omega+gamma))
+#   # s <- eta_w/eta_c
+#   return((eta_c*(s*A2+B2)+D)*C2)
+# }
+# eval_R0 <- function(state=state_dfe,params){
+#   ## input the params and their range, this function makes a grid dataframe, calls R0 and outputs the csv file
+#   ## the params are defaults, specify the range of favorite params here.
+#   unpack(as.list(c(state,params)))
+#   tol <- 1e-10 ## to resolve the issue of very small numbers
+#   # specify the ranges, FIXME, not to be hard coded!
+#   eta_w <- seq(0,1, length.out=n_out)
+#   eta_c <- seq(0,1, length.out=n_out)
+#   rho <- seq(0,0.01, length.out=n_in) ## 0,0.01
+#   omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, was 0.1 to 2
+#
+#   df1 <- expand.grid(N0=params[["N0"]],beta=params[["beta"]],gamma=params[["gamma"]],omega=omega,rho=rho,
+#                      W_S=W_S,W_I=params[["W_I"]],W_R=params[["W_R"]],
+#                      p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
+#                      eta_w=eta_w,eta_c=eta_c)
+#   df2<- data.frame(df1,
+#                    R0=apply(df1,1,function(params_in)R0(params=params_in)),
+#                    Fi= apply(df1,1,function(params_in)Fi_hat(params=params_in)))
+#   # This is for plotting purposes:
+#   df2 <- (df2 %>%
+#             dplyr::mutate(R0_sub=ifelse(eta_w<eta_c, NA, R0),
+#                           theta_w=1-eta_w,
+#                           theta_c=1-eta_c,
+#                           Delta=ifelse(eta_w<eta_c, NA, 1-(R0*gamma/beta) ),
+#                           Delta=ifelse(abs(Delta)<tol,0,Delta)
+#                           )
+#           )
+#   return(df2)
+# }
+
+# state_init <- c(S_u=params[["N0"]], S_n=0,
+#                 I_u=1,I_n=0,I_p=0,I_t=0,
+#                 R_u=0,R_n=0,R_p=0,R_t=0,
+#                 N=0,P=0)
+
+# state_dfe <- c(S_u=Su_dfe(params), S_n=Sn_dfe(params),
+#                 I_u=0,I_n=0,I_p=0,I_t=0,
+#                 R_u=0,R_n=0,R_p=0,R_t=0,
+#                 N=0,P=0)
+
 \end{document}

### 2021-04-27 codes/SIRfunctions.R

@@ -1,4 +1,5 @@
 library(shellpipes)
+rpcall("SIRfunctions.Rout SIRfunctions.R")
 
 # Functions for the SIR model
 # By: A Gharouni
@@ -193,67 +194,5 @@ make_params_dat<-function(params,
   return(dfout)
 }
 
-# ###################
-# Extra Stuff unused
-# ###################
-
-# Conditional stop for ODE desolver
-# rootfun <- function (time, state,params) { 
-#   unpack(as.list(c(state,params)))
-#   return(I_u - 1) }
-
-# ode(
-#   func=sir.model,
-#   y=state_init,
-#   times= d, 
-#   parms=params,
-#   atol = 1e+1, rtol = 1e+1
-#   rootfun = rootfun,
-#   method="lsodar"
-# )
-
-## R0 expression with eta_w/eta_c factored
-# R02 <- function(state=state_dfe,params){
-#   unpack(as.list(c(state,params)))
-#   Sn <- Sn_dfe(params)
-#   Su <- Su_dfe(params)
-#   Fi <- Fi_hat(params) #at DFE
-#   
-#   A2 <- gamma*(p_I*Fi*omega*Sn+(omega+gamma)*(gamma*Sn+Fi*N0))
-#   B2 <- p_I*Fi*omega*(gamma*Su+omega*N0)
-#   D <- gamma*(omega+gamma)*(gamma*Su+omega*N0)
-#   C2 <- beta/((N0*gamma*(gamma*(omega+gamma)+Fi*(gamma+omega*p_I)))*(omega+gamma))
-#   # s <- eta_w/eta_c
-#   return((eta_c*(s*A2+B2)+D)*C2)
-# }
-# eval_R0 <- function(state=state_dfe,params){
-#   ## input the params and their range, this function makes a grid dataframe, calls R0 and outputs the csv file
-#   ## the params are defaults, specify the range of favorite params here.
-#   unpack(as.list(c(state,params)))
-#   tol <- 1e-10 ## to resolve the issue of very small numbers 
-#   # specify the ranges, FIXME, not to be hard coded!
-#   eta_w <- seq(0,1, length.out=n_out)
-#   eta_c <- seq(0,1, length.out=n_out)  
-#   rho <- seq(0,0.01, length.out=n_in) ## 0,0.01
-#   omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, was 0.1 to 2
-#   
[…]

### 2021-04-27 codes/params.R

@@ -1,10 +1,7 @@
 library(shellpipes)
-rpcall("params.Rout params.R SIRfunctions.rda")
-
+rpcall("params.Rout params.R")
+# SIRfunctions.rda
 loadEnvironments()
-
-# Parameters used in SIR model
-# Path: ~/projects/SIRmodel/codes
 
 params <- c(
 N0=1000000, #total population size, 
@@ -20,17 +17,5 @@ eta_w=0.02, eta_c=0.01 #isolation parameter (eta=0 is the perfect isolation)
             
 class(params) <- "params_pansim"
 
-state_init <- c(S_u=params[["N0"]], S_n=0,
-                I_u=1,I_n=0,I_p=0,I_t=0,
-                R_u=0,R_n=0,R_p=0,R_t=0,
-                N=0,P=0)
-                
-
-
-state_dfe <- c(S_u=Su_dfe(params), S_n=Sn_dfe(params),
-                I_u=0,I_n=0,I_p=0,I_t=0,
-                R_u=0,R_n=0,R_p=0,R_t=0,
-                N=0,P=0)
-
-saveVars(params, state_init, state_dfe)
+saveVars(params)
 

### 2021-04-27 codes/sir_plot.R

@@ -38,6 +38,18 @@ df_targeted <- make_params_dat(params = update(params,W_S=W_S_targeted),
                 eta_cs=0,eta_ce=1, ## so theta_c
                 omega_s=0.1,omega_e=2,
                 rho_s=0,rho_e=0.01)
+## high rho and omega
+df_random_h <- make_params_dat(params = params,
+                             eta_ws=0,eta_we=1, ## so theta_w
+                             eta_cs=0,eta_ce=1, ## so theta_c
+                             omega_s=0.50001,omega_e=2,
+                             rho_s=0,rho_e=0.5)
+
+df_targeted_h <- make_params_dat(params = update(params,W_S=W_S_targeted),
+                               eta_ws=0,eta_we=1, ## so theta_w
+                               eta_cs=0,eta_ce=1, ## so theta_c
+                               omega_s=0.50001,omega_e=2,
+                               rho_s=0,rho_e=0.5)
 
 ##important contour, ie R0=1 thus threshold=1 when plotting R0 contours, or Delta(R0=1)
 threshold <- 1-(params[["gamma"]]/params[["beta"]]) ## Or 0?  corresponding to R0=1 

### 2021-04-28 codes/SIRfunctions.R
R0 function was modified: S_n scaled by eta_w, also expressions A B modified to match the manuscript

@@ -128,10 +128,10 @@ R0<-function(params){
   Su <- Su_dfe(params)
   Fi <- Fi_hat(params) #at DFE
   
-  A <- gamma*(omega+gamma+eta_w*Fi) + omega*eta_c*p_I*Fi
-  B <- (omega+eta_w*(Fi+gamma))*gamma+(eta_w*gamma+eta_c*omega)*(omega*p_I*Fi)/(omega+gamma)
-  C <- beta/(N0*gamma*(gamma*(omega+gamma)+Fi*(gamma+omega*p_I)))
-  return((A*Su+B*Sn)*C)
+  A<-gamma*(omega+gamma)^2+eta_w*gamma*(omega+gamma)*Fi+eta_c*omega*(omega+gamma)*p_I*Fi  
+  B<-gamma*omega*(omega+gamma)+eta_w*(gamma*(omega+gamma)*(Fi+gamma)+gamma*omega*p_I*Fi)+eta_c*omega^2*p_I*Fi
+  C<-(gamma*(omega+gamma)+Fi*(gamma+omega*p_I))*(omega+gamma)
+  return(beta/(N0*gamma*C)*(A*Su+B*Sn*eta_w))
 }
 
 ## calculate the principle eigenvalue and the corresponding eigenvector.
@@ -176,8 +176,8 @@ make_params_dat<-function(params,
                      W_S=W_S,W_I=W_I,W_R=W_R,
                      p_S=p_S,p_I=p_I,p_R=p_R,
                      eta_w=eta_w,eta_c=eta_c)
-    ## principle eigenvector
-  eigvec <- t(apply(df1,1,function(params_in)eigvec_max(params=params_in)))  
+  ## principle eigenvector
+  # eigvec <- t(apply(df1,1,function(params_in)eigvec_max(params=params_in)))  
   dfout <-(df1 %>% 
              dplyr::mutate(
                R0=apply(df1,1,function(params_in)R0(params=params_in)),  
@@ -185,11 +185,11 @@ make_params_dat<-function(params,
                theta_w=1-eta_w,
                theta_c=1-eta_c,
                Delta=ifelse(eta_w<eta_c, NA, 1-(R0*gamma/beta) ),
-               Delta=ifelse(abs(Delta)<tol,0,Delta),
-               I_u=eigvec[,1],
-               I_n=eigvec[,2],
-               I_p=eigvec[,3],
-               I_c=eigvec[,4]
[…]

### 2021-04-28 codes/sir_plot.R

@@ -5,7 +5,7 @@ library(directlabels)
 library(latex2exp)
 library(rootSolve)
 library(McMasterPandemic)
-library(matlib) # used for matrix inverse function
+## library(matlib) # used for matrix inverse function
 
 library(shellpipes)
 

### 2021-04-28 codes/sir_plot.R
flowchart updated with (1-theta), parameter range modified

@@ -5,7 +5,7 @@ library(directlabels)
 library(latex2exp)
 library(rootSolve)
 library(McMasterPandemic)
-## library(matlib) # used for matrix inverse function
+# library(matlib) # used for matrix inverse function
 
 library(shellpipes)
 
@@ -42,14 +42,14 @@ df_targeted <- make_params_dat(params = update(params,W_S=W_S_targeted),
 df_random_h <- make_params_dat(params = params,
                              eta_ws=0,eta_we=1, ## so theta_w
                              eta_cs=0,eta_ce=1, ## so theta_c
-                             omega_s=0.50001,omega_e=2.0001,
-                             rho_s=0,rho_e=0.5)
+                             omega_s=0.25,omega_e=2,
+                             rho_s=0,rho_e=0.25-0.000001)
 
 df_targeted_h <- make_params_dat(params = update(params,W_S=W_S_targeted),
                                eta_ws=0,eta_we=1, ## so theta_w
                                eta_cs=0,eta_ce=1, ## so theta_c
-                               omega_s=0.50001,omega_e=2.0001,
-                               rho_s=0,rho_e=0.5)
+                               omega_s=0.25,omega_e=2,
+                               rho_s=0,rho_e=0.25-0.000001)
 
 ##important contour, ie R0=1 thus threshold=1 when plotting R0 contours, or Delta(R0=1)
 threshold <- 1-(params[["gamma"]]/params[["beta"]]) ## Or 0?  corresponding to R0=1 

### 2021-04-28 SIR_manuscript.tex
R0 function was modified: S_n scaled by eta_w, also expressions A B modified to match the manuscript
model updated with (1-theta_w)

@@
 \centering
-\includegraphics[width=\linewidth]{codes/R0contour_random_rho.5.pdf}
+\includegraphics[width=\linewidth]{codes/R0contour_random2.pdf}
 \caption{}
@@
 \centering
-\includegraphics[width=\linewidth]{codes/R0contour_TTI_rho.5.pdf}
+\includegraphics[width=\linewidth]{codes/R0contour_TTI2.pdf}
 \caption{}
@@
  d S\_u/dt &= -\Lambda S\_u - F_S S\_u + \omega S\_n, \\
- d S\_n/dt &= -\Lambda S\_n + (1-p_S) F_S S\_u - \omega S\_n, \\
- d S\_p/dt &= -\Lambda S\_p + p_S F_S S\_u - \omega S\_p, \\
- d S\_c/dt &= -\Lambda S\_c + \omega S\_p, \\
+ d S\_n/dt &= -(1-\theta\_w)\Lambda S\_n + (1-p_S) F_S S\_u - \omega S\_n, \\
+ d S\_p/dt &= -(1-\theta\_w)\Lambda S\_p + p_S F_S S\_u - \omega S\_p, \\
+ d S\_c/dt &= -(1-\theta\_c)\Lambda S\_c + \omega S\_p, \\
  d I\_u/dt &= \Lambda S\_u - F_I I\_u + \omega I\_n  - \gamma I\_u, \\
- d I\_n/dt &= \Lambda S\_n + (1-p_I) F_I I\_u - \omega I\_n -\gamma I\_n, \\
- d I\_p/dt &= \Lambda S\_p + p_I F_I I\_u - \omega I\_p -\gamma I\_p, \\
- d I\_c/dt &= \Lambda S\_c + \omega I\_p - \gamma I\_c, \\
+ d I\_n/dt &= (1-\theta\_w)\Lambda S\_n + (1-p_I) F_I I\_u - \omega I\_n -\gamma I\_n, \\
+ d I\_p/dt &= (1-\theta\_w)\Lambda S\_p + p_I F_I I\_u - \omega I\_p -\gamma I\_p, \\
+ d I\_c/dt &= (1-\theta\_c)\Lambda S\_c + \omega I\_p - \gamma I\_c, \\
  d R\_u/dt &= \gamma I\_u - F_R R\_u + \omega R\_n, \\
@@
 S\_u^* & (1-\theta\_w)\,S\_u^* & (1-\theta\_w)\,S\_u^* & (1-\theta\_c)\,S\_u^*\\
-S\_n^* & (1-\theta\_w)\,S\_n^* & (1-\theta\_w)\,S\_n^* & (1-\theta\_c)\,S\_n^* \\ 
+(1-\theta\_w)S\_n^* & (1-\theta\_w)^2\,S\_n^* & (1-\theta\_w)^2\,S\_n^* & (1-\theta\_w)(1-\theta\_c)\,S\_n^* \\ 
 0&0&0&0\\
@@
  \end {array} \right] \\
- =&\beta/N_0 \left[ \begin {array}{c} S\_u^* \\ S\_n^* \\ 0 \\ 0 \end {array} \right]
+ =&\beta/N_0 \left[ \begin {array}{c} S\_u^* \\ (1-\theta\_w)S\_n^* \\ 0 \\ 0 \end {array} \right]
         \left[ \begin {array}{cccc} 1, \, 1-\theta\_w, \, 1-\theta\_w, \, 1-\theta\_c \end {array} \right], \text{and}\\  V=&
@@
 G_{11} = \frac{\beta}{\gamma \, C} 
-\left[ \begin {array}{c} (\omega-\rho)/\omega \\ \rho/\omega \end {array} \right]
+\left[ \begin {array}{c} (\omega-\rho)/\omega \\ (1-\theta\_w)\rho/\omega \end {array} \right]
 \left[ \begin {array}{cccc} 1, \, 1-\theta\_w, \, 1-\theta\_w, \, 1-\theta\_c \end {array} \right]
@@
 G_{11} = \frac{\beta}{\gamma \, C} 
-\left[ \begin {array}{c}  (\omega-\rho)/\omega \\ \rho/\omega  \end {array} \right]
+\left[ \begin {array}{c}  (\omega-\rho)/\omega \\ (1-\theta\_w)\rho/\omega  \end {array} \right]
 \left[ \begin {array}{cccc} 

### 2021-04-29 SIR_manuscript.tex
minor methods

@@
 where the term $\frac{\beta}{\gamma}$ is the classical basic reproduction number for a SIR model without testing and isolation (see, e.g., \cite{keeling2011modeling}), and $\Delta$ is the reduction parameter defined as follows.
+\ali{$\Delta$ should be updated, the expression will be linear in $\theta\_c$ but quadratic in $\theta\_w$, C stay unchanged.
+It can be proved that $\Delta$ increases in $\theta\_w$ for feasible range of $\theta\_w$ (I think this argument will go to the appendix).?}
 \begin{equation}
@@
 \label{eq:abc}
-A=& \gamma \Big( \big(\gamma(\omega+\gamma) + \omega p_I \hat F_I \big) \frac{S\_n^*}{N_0} + (\omega+\gamma) \hat F_I \Big),\\
+A=& \gamma \Big( \big(\gamma(\omega+\gamma) + \omega p_I \hat F_I \big) \frac{(1-\theta\_w)S\_n^*}{N_0} + (\omega+\gamma) \hat F_I \Big),\\
 B=& \omega p_I(\gamma \frac{S\_u^*}{N_0}+\omega)\hat F_I, \\ 
@@
 We numerically evaluate $\Delta$ \eqref{R0}, reflecting the reduction of $\Rnum$ with respect to testing and isolation. We use the following parameters (see Table \ref{tab:params}):
-$N_0=1 \times 10^6$, $\omega \in [0.1,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.01]$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.39$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
+$N_0=1 \times 10^6$, $\omega \in [0.1,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.01]$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.339$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
 Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
@@
 We numerically evaluate $\Delta$ \eqref{R0}, reflecting the reduction of $\Rnum$ with respect to testing and isolation. We use the following parameters (listed in Table \ref{tab:params}):
-$N_0=1 \times 10^6$, $\omega \in (0.5,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.5]$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.39$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
+$N_0=1 \times 10^6$, $\omega \in (0.25,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.25]$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.339$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
 Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
@@
 \end{equation}
-where $C$ is the same as the one in Eq. \eqref{eq:abc}, i.e.,
+where $C$ is the same as the one in Eq. \eqref{eq:abc}, i.e., \ali{all equations after this point need to be updated.}
 $$C=(\omega+\gamma)(\gamma(\omega+\gamma)+(\omega\,p_I+\gamma)\,\hat F_I),$$

### 2021-04-29 SIR_manuscript.tex
Meeting with Ali

@@
 
+\newcommand{\fref}[1]{Fig.~\ref{#1}}
 \newcommand{\Rlogo}{R\xspace}
@@
 
-\begin{figure}[!h] 
+\begin{figure}
 \begin{center} 
@@
 
-The Disease-Free Equilibrium, DFE hearafter, for the SIR model, Eqs. \eqref{model}, is given by setting the differential equations to 0 and solving for the unknowns. The DFE is
+The Disease-Free Equilibrium (DFE) for the SIR model, Eqs. \eqref{model}, is found by setting the infected compartments to 0 and solving for the unknowns. The DFE is
 \begin{equation}
@@
 \begin{enumerate}
+
 \item \label{p1:eta} Increasing the isolation parameters for tested and confirmed individuals decreases \Rnum;
-$$\partial{\Rnum}/\partial{\theta\_c} \leq 0 \text{~and~} \partial{\Rnum}/\partial{\theta\_w} \leq 0.$$ 
-\item \label{p1:rho} A higher testing intensity may increase or decrease $\Rnum$;
-$$\partial{\Rnum}/\partial{\rho} \text{~can be positive or negative}.$$
-\item \label{p1:omega} A higher rate of test return may increase or decrease \Rnum.
-$$\partial{\Rnum}/\partial{\omega} \text{~can be positive or negative}.$$
+
+\item \label{p1:rho} Higher testing $\rho$ intensity always decreases $\Rnum$ \textbf{if}: 
+%% \begin{enumerate} \item
+testing is random; \textbf{or}
+%% \item
+$\rho$ is small.
+%% \end{enumerate}
+
+\item \label{p1:omega} A higher rate of test return always decreases \Rnum, \textbf{if} $\theta_w=0$.
+
 \end{enumerate}
@@
 
-Our numerical results are shown in Figs. \ref{pan}, where $\rho \in [0,0.01]$, and \ref{pan2}, where $\rho \in [0,0.5]$.
-When targeted testing strategy is applied compared to random testing strategy, i.e., comparing panel (b) to (a) of Figs.\,\ref{pan} and \ref{pan2}, it can be inferred that 
-(i) \ali{thinking of the intercept of the solid line/contours} the critical \percap testing intensity corresponding to $\Rnum=1$, is lower and thus less testing is required to reduce $\Rnum$, 
-(ii) \ali{thinking of the slope contours, having hard time to explain this since there is a non-monotonic effect of omega} sensitivity of $\Rnum$ wrt $1/\omega$; the changes in $\Rnum$ in response to changes in $1/\omega$, is lower.
+\jd{$w_R$ does not affect \Rnum, and should not be one of the parameters in the figure title. Prepend with English (random testing and weighting testing).}
 
-It is notable that increasing testing intensity $\rho$ as a strategy to reduce $\Rnum$ depends on the value of $\rho$ to begin with, the mean test return time $1/\omega$ and the efficacy of isolation $\theta$.
-In particular, there might be feasible threshold values for $\rho$ and $\omega$ that maximizes the effect of these parameters in reduction of $\Rnum$ depends upon the isolation efficacy.
+Numerical results are shown in Fig. \ref{pan}.
+Targeted testing (panel b) is more effective at control than random testing (panel a).
+
+When $\theta_w =0$ (top row of each panel), we see that shorter test-waiting times reduce \Rnum\ ($\Delta$ increases as we move to the left in each plot in this row).
+For non-zero $\Delta$, however, we mostly see the opposite effect: less waiting leads to more transmission.
+This is because returning negative tests leads people to stop distancing; this applies both to susceptibles, and to infectious people who receive negative test results because they were sampled when susceptible (or because of imperfect test sensitivity, which is not modeled in this figure).
+
+We also see that greater test intensity $\rho$ increases the effectiveness of control $\Delta$ ($\Delta$ increases as we move up in each plot in this Figure.
+Mathematically speaking, we did find a counter-vailing effect, but this can only be seen when we allow $\rho$ to take unrealistically large values, and only for weighted testing (see \fref{pan2}).
+The explanation here is that more rapid testing leaves more susceptibles in the ``waiting-for-no'' category at the DFE; these people must then wait for their tests to be returned before they can be tested again, receive a positive test, and become cautious.
+This effect is usually weak compared to the helpful effects of testing, but if testing is weighted, $\theta_w$ is low, and test returns are slow, t is possible for increasing $\rho$ to reduce $\Delta$ (see upper-right part of upper-right plot of \fref{pan2}b).
 

### 2021-05-03 SIR_manuscript.tex
minor changes

@@
 \newcommand{\pro}[1][]{\ensuremath{\frac{\partial #1}{\partial \rho}}}
+\newcommand\pder[2][]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder[x]{y}
 
@@
 
-\jd{$w_R$ does not affect \Rnum, and should not be one of the parameters in the figure title. Prepend with English (random testing and weighting testing).}
-
 Numerical results are shown in Fig. \ref{pan}.
@@
 The explanation here is that more rapid testing leaves more susceptibles in the ``waiting-for-no'' category at the DFE; these people must then wait for their tests to be returned before they can be tested again, receive a positive test, and become cautious.
-This effect is usually weak compared to the helpful effects of testing, but if testing is weighted, $\theta_w$ is low, and test returns are slow, t is possible for increasing $\rho$ to reduce $\Delta$ (see upper-right part of upper-right plot of \fref{pan2}b).
-
-\ali{Thinking of last column of 2 Figs, moving from top tp bottom as $\theta\_c=1$ and $\theta\_w$ changes from 0 to 1} When $\theta\_c$ is closer to 1 and $\theta\_w$ is closer to 0, increasing $\rho$ combined with deccreasing $1/\omega$ is effective in reducing $\Rnum$.
-However, as $\theta_w$ is getting closer to 1, increasing $\rho$ combined with increasing $1/\omega$ is better.
-
-% When $\rho$ is closer to 0 (Fig.\,\ref{pan}),  and $\theta_c$ is closer to 0, increasing $\rho$ and decreasing $1/\omega$ is a    
-% 
-% when $\rho$ is small relative to $N_0$, increasing $\rho$ combined with changes in the test return time $1/\omega$ is always good in general.  
-% 
-% When the isolation efficay is low, i.e., $\theta\_w$ is closer to 0, speeding up the test reporting processing time can increase $\Rnum$ regardless of the speed of testing.
-However there is threshold for $1/\omega$ above which prolonging the test return time
-
-% awaiting people (including people in $I\_p$ and $I\_n$ compartments) follow the isolation perfectly, i.e., $\theta\_w$ is closer to 1.
-However, speeding the test reporting reduces the epidemic more when the awaiting people less follow the isolation, i.e., $\theta\_w$ is closer to 0.
-% 
-% Furthermore, we derived an inequality that quantifies the exact relationships between model parameters that result in returning tests more rapidly being favorable \eqref{eq:necsuf}.
-% \david{What do we learn from this inequality?  Why is it relegated to an appendix?}
+This effect is usually weak compared to the helpful effects of testing, but if testing is weighted, $\theta_w$ is low, and test returns are slow, t is possible for increasing $\rho$ to reduce $\Delta$ (see upper-right part of upper-right plot of \fref{pan2}(b)).
 

### 2021-05-03 codes/sir_plot.R

@@ -69,8 +69,7 @@ label_special <- function (labels, multi_line=FALSE, sep = "== ") {
 }
 
 n_contour <- 8 ## number of desired contours or bins
-bins <- cut(sort(unique(c(df_random$Delta,df_targeted$Delta))), 
-            n_contour)
+bins <- cut(sort(unique(c(df_random$Delta,df_targeted$Delta))), n_contour)
 brks <- levels(bins)
 brks_vec <- unique(as.numeric(unlist(lapply(strsplit(brks, ","), function(x) gsub("\\(|]", "", x)))))
 brks_vec <- sort(ifelse(brks_vec[]< 0,0,brks_vec[]), decreasing = TRUE) ## replace neg with 0 and reorder
@@ -105,7 +104,7 @@ p1_temp
 # 1. Plot the Random Testing Scenario:
 # #################################
 
-ggsave(p1_temp + ggtitle(TeX("w_S=w_I=w_R=1")) +
+ggsave(p1_temp + ggtitle(TeX(r'(Random testing, $w_S=w_I=1$)')) +
        theme(legend.position = "none"),
        filename = "R0contour_random.pdf" ,
        width = 12, height = 12, units = "cm")
@@ -115,11 +114,10 @@ ggsave(p1_temp + ggtitle(TeX("w_S=w_I=w_R=1")) +
 # #################################
 
 ggsave((p1_temp %+% df_targeted) +
-       ggtitle(TeX(sprintf("w_S=%.1f, w_I=w_R=1",W_S_targeted))) +
+       ggtitle(TeX(sprintf("Targeted testing, w_S=%.1f, w_I=1",W_S_targeted))) +
        theme(legend.position = c(0.2, 0.3),legend.text = element_text(size = 8)),
        filename = "R0contour_TTI.pdf" ,
        width = 12, height = 12, units = "cm")
-
 
 # #################################
 # 3. Plots for high rho Scenario:
@@ -160,7 +158,7 @@ p1_temp2
 # 4. Plot the Random Testing Scenario:
 # #################################
 
-ggsave(p1_temp2 + ggtitle(TeX("w_S=w_I=w_R=1")) +
+ggsave(p1_temp2 + ggtitle(TeX(r'(Random testing, $w_S=w_I=1$)')) +
          theme(legend.position = "none"),
        filename = "R0contour_random2.pdf" ,
        width = 12, height = 12, units = "cm")
[…]

### 2021-05-03 codes/SIRfunctions.R
way too much figure tweaking

@@ -33,31 +33,31 @@ sigma <- function(state,params){
 # SIR model
 sir.model <- function(time,state,params){
   unpack(as.list(c(state,params)))
-  ## Force of Infection  
+  ## Force of Infection
   Lambda <- beta * (I_u + eta_w*(I_n+I_p) + eta_c*I_t)/N0
   ## scaling the weights
-  # W <- W_S*S_u+W_I*I_u+W_R*R_u 
+  # W <- W_S*S_u+W_I*I_u+W_R*R_u
   # sigma <- rho*N0/W
   sigma <- sigma(state,params)
-  
+
   # testing intensity
   F_S <- sigma*W_S
   F_I <- sigma*W_I
   F_R <- sigma*W_R
   # Equations
   dS_u.dt <- -Lambda*S_u - (1-p_S) * F_S * S_u + omega * S_n
-  dS_n.dt <- -Lambda*S_n + (1-p_S) * F_S * S_u - omega * S_n 
+  dS_n.dt <- -Lambda*S_n + (1-p_S) * F_S * S_u - omega * S_n
   dI_u.dt <-  Lambda*S_u + omega * I_n - F_I * I_u - gamma * I_u
   dI_n.dt <-  Lambda*S_n + (1-p_I) * F_I * I_u  - omega * I_n - gamma * I_n
-  dI_p.dt <-  p_I * F_I * I_u - omega * I_p - gamma * I_p 
-  dI_t.dt <-  omega * I_p - gamma * I_t 
-  dR_u.dt <- gamma * I_u + omega * R_n - F_R * R_u  
+  dI_p.dt <-  p_I * F_I * I_u - omega * I_p - gamma * I_p
+  dI_t.dt <-  omega * I_p - gamma * I_t
+  dR_u.dt <- gamma * I_u + omega * R_n - F_R * R_u
   dR_n.dt <- gamma * I_n + (1-p_R) * F_R * R_u - omega * R_n
   dR_p.dt <- gamma * I_p + p_R * F_R * R_u - omega * R_p
   dR_t.dt <- gamma * I_t + omega * R_p
   dN.dt <- omega * (S_n + I_n + R_n)
   dP.dt <- omega *(I_p + R_p)
-  
+
   # return the rate of change
[…]

### 2021-05-03 codes/sir_plot.R

@@ -5,11 +5,11 @@ library(directlabels)
 library(latex2exp)
 library(rootSolve)
 library(McMasterPandemic)
-# library(matlib) # used for matrix inverse function
+library(cowplot) ## for ggdraw()/rectangle-hacking
 
 library(shellpipes)
 
-rpcall("sir_plot.Rout sir_plot.R params.rda")
+rpcall("sir_plot.Rout sir_plot.R params.rda SIRfunctions.rda")
 
 loadEnvironments()
 
@@ -24,20 +24,34 @@ n_in <- 41 ## grid N within facets
 
 #choose weight W_S 0.3 for TTI testing plan and 1 for the random testing.
 W_S_random <- 1
-W_S_targeted <- 0.3 
+W_S_targeted <- 0.3
+## tweak to get non-overlapping axis labels:
+##  adjust grid to be slightly bigger than ideal tick marks
+inv_omega_min <- 0.5
+inv_omega_max <- 12
+inv_omega_brks <- c(1,5,10)
+rho_min <- 0
+rho_max <- 0.013
+rho_brks <- c(0,0.005,0.01)
 
-# make dataframe 
+inv_omega_fast_min <- 0.5
+inv_omega_fast_max <- 5
+inv_omega_fast_brks <- c(1,2,4)
+
+background <- "grey90"
+# make dataframe
 df_random <- make_params_dat(params = params,
                 eta_ws=0,eta_we=1, ## so theta_w
                 eta_cs=0,eta_ce=1, ## so theta_c
-                omega_s=0.1,omega_e=2,
-                rho_s=0,rho_e=0.01)
+                omega_s=1/inv_omega_max,omega_e=1/inv_omega_min,
+                rho_s=rho_min,rho_e=rho_max)
 
 df_targeted <- make_params_dat(params = update(params,W_S=W_S_targeted),
                 eta_ws=0,eta_we=1, ## so theta_w
                 eta_cs=0,eta_ce=1, ## so theta_c
-                omega_s=0.1,omega_e=2,
-                rho_s=0,rho_e=0.01)
+                omega_s=1/inv_omega_max,omega_e=1/inv_omega_min,
+                rho_s=rho_min,rho_e=rho_max)
+
 ## high rho and omega
 df_random_h <- make_params_dat(params = params,
                              eta_ws=0,eta_we=1, ## so theta_w
[…]

### 2021-05-04 SIR_manuscript.tex
way too much figure tweaking
adding proof of decreasing R0 wrt thetas'

@@
 Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
-The black solid line in each panel represents the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ corresponding to the threshold of $\Rnum=1$.
-\bmb{we still have some collisions in the y-axis tick labels}
+The black dotted line in each panel represents the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ corresponding to the threshold of $\Rnum=1$.
 }
@@
 \end{equation}
-where $C$ is the same as the one in Eq. \eqref{eq:abc}, i.e., \ali{all equations after this point need to be updated.}
+where $C$ is the same as the one in Eq. \eqref{eq:abc}, i.e., 
 $$C=(\omega+\gamma)(\gamma(\omega+\gamma)+(\omega\,p_I+\gamma)\,\hat F_I),$$
@@
 $\Rnum$ is in the same form as in Eq. \eqref{R0}  
-$$\Rnum= \frac{\beta}{\gamma} (1-\Delta).$$
+$$\Rnum= \frac{\beta}{\gamma} (1-\Delta),$$
+where 
+\begin{equation}
+\label{eq:del4}
+\Delta= \frac{1}{N_0\,C}\big(C_1\,S\_u^*+((C_1+C_2)(1-\theta\_w)+C\,\theta\_w)\,S\_n^*\big).
+\end{equation}
+Note that for analysis convention, we let $N_0=1$, thus $S\_u^*$ and $S\_n^*$ are in the scale of 0 to 1. 
+To derive the first result, i.e., how changes in $\theta\_w$ and $\theta\_c$ affects $\Rnum$, we factor these parameters in $\Delta$ \eqref{eq:del4}. Thus, $\Delta$ can be rewritten as
+\begin{equation}
+\label{eq:del4_theta}
+\Delta= \frac{1}{C}\Big(
+-D_1\,S\_n^*\,\boldsymbol {\theta\_w^2}+
+\big(-\omega^2\,p_I\,\hat F_I\,S\_n^*\,\boldsymbol{\theta\_c}\,+\,D_2\,S\_n^*\,+\,\gamma\,\hat F_I(\omega+\gamma) \big)\boldsymbol{\theta\_w}+
+(\omega+\gamma\,S\_u^*)\omega\,p_I\,\hat F_I\,\boldsymbol{\theta\_c}
+\Big),
+\end{equation}
+where
+\begin{align}
+\label{eq:del4_d1d2}
+D_1=& (\omega+\gamma)\gamma^2+(\omega+\gamma+\omega\,p_I)\gamma\,\hat F_I, \\
+D_2=& (3\omega+2\gamma)\gamma^2+(\omega+\gamma+2\omega\,p_I)\gamma\,\hat F_I+(\gamma+p_I\,\hat F_I)\omega^2.
+\end{align}
+The scaled expression of $\Delta$, specifically $\Delta\,C$, in Eq.\,\eqref{eq:del4_theta} is linear in $\theta\_c$ with a positive coefficient given by 
+$$(\gamma\,S\_u^*\,+\omega(1-\theta\_w\,S\_u^*))\omega\,p_I\,\hat F_I.$$
+This results in increasing $\Delta$, thus decreasing $\Rnum$ with respect to $\theta\_c$, that is $\pder[\Rnum]{\theta\_c}\leq 0$. Note that $C$ is independant of $\theta\_c$ and $\theta\_w$.
+
+With a similar logic, the expression of $\Delta\,C$ derived from Eq.\,\eqref{eq:del4_theta} is a concave-down quadratic equation in $\theta\_w$, given by
+\begin{equation}
+\label{eq:del4_thetaw}
+-D_1\,S\_n^*\,\boldsymbol {\theta\_w^2}+
+\big(-\omega^2\,p_I\,\hat F_I\,S\_n^*\,\boldsymbol{\theta\_c}\,+\,D_2\,S\_n^*\,+\,\gamma\,\hat F_I(\omega+\gamma) \big)\boldsymbol{\theta\_w}.
+\end{equation}
+We show that the feasible range of $\theta\_w$ lies btween 0 and the vertex of this parabola where the parabola is increasing in $\theta\_w$, and so does $\Delta$ which results in inferring $\pder[\Rnum]{\theta\_w}\leq 0$.
+It is enough to show that partial derivative of the expression \eqref{eq:del4_thetaw} with respect to $\theta\_w$ at $\theta\_w=1$ is non-negative. It follows that
+\begin{align}
+\label{eq:d_del4_thetaw}
+\pder[\Delta\,C]{\theta\_w}\bigg\rvert_{\theta\_w=1} =&
+(D_2-2D_1-\omega^2\,p_I\,\hat F_I\,\theta\_c)\,S\_n^*+\gamma\,\hat F_I(\omega+\gamma) \\
+=& (\gamma(\omega+\gamma)+\gamma\omega^2+(1-\theta\_c)\omega^2\,p_I\,\hat F_I )\,S\_n^*
++\gamma(\omega+\gamma)\,\hat F_I\,(1-S\_n^*),
+\end{align}
+which is a positive quantity, given that $\theta\_c$ and $S\_n^*$ vary between 0 and 1.
+
+\ali{stoped here!}
 

### 2021-05-05 SIR_manuscript.tex
analysis, R0 wrt rho

@@
 \label{dfe}
-S\_n^*= \frac{\rho}{\omega} N_0, \ S\_u^*= N_0-S\_n^*, \text{~and~} I_j=R_j=0 \ \text{for all }j.
+S\_n^*= \frac{\rho}{\omega}\,N_0, \ S\_u^*= \frac{\omega-\rho}{\omega}\,N_0, \text{~and~} I_j=R_j=0 \ \text{for all }j.
 \end{equation}
@@
 \end{align}
-where $C= \big( \gamma(\omega+\gamma)+(\gamma+\omega p_I)\hat F_I \big) (\omega+\gamma)$ and $\hat F_I$ is the \percap testing rate for the infected people and represented in Eq.
-\eqref{eq:fi}.
+where $C= \big( \gamma(\omega+\gamma)+(\gamma+\omega p_I)\hat F_I \big) (\omega+\gamma)$ and $\hat F_I$ is the \percap testing rate for the infected people and represented in Eq.\,\eqref{eq:fi}.
 Note that all the columns of matrix $V^{-1}$ summ up to $1/\gamma$.
@@
 
-This is to provide a mathematical materials to prove that $\pro \Delta$ can be positive or negative, with $\Delta$ defined in Eq.
-\eqref{eq:abc}, and thus $\pro \Rnum < 0$, where $\Rnum$ is the basic reproduction number, given in Eqs.\,\eqref{R0}.
+This is to provide a mathematical materials to prove that $\pro \Delta$ can be positive or negative, with $\Delta$ defined in Eq.\,\eqref{eq:abc}, and thus $\pro \Rnum < 0$, where $\Rnum$ is the basic reproduction number, given in Eq.\,\eqref{R0}.
 We can rewrite matrix $G_{11}$ in \eqref{G11} in the following form to simplify the calculations.
@@
 G_{11} = \frac{\beta}{\gamma \, C} 
-\left[ \begin {array}{c}  (\omega-\rho)/\omega \\ (1-\theta\_w)\rho/\omega  \end {array} \right]
+% \left[ \begin {array}{c}  (\omega-\rho)/\omega \\ (1-\theta\_w)\rho/\omega  \end {array} \right]
+\left[ \begin {array}{c}  S\_u^*/N_0 \\ (1-\theta\_w)\,S\_n^*/N_0  \end {array} \right]
 \left[ \begin {array}{cccc} 
@@
 \end{align}
-where $\hat F_I$ is given in Eq. \eqref{eq:fi}.
+where $\hat F_I$ is given in Eq.\,\eqref{eq:fi}.
+Note that for analysis brevity, we let $N_0=1$, thus $S\_u^*$ and $S\_n^*$ are in the scale of 0 to 1.
 $\Rnum$ is in the same form as in Eq. \eqref{R0}  
@@
 \label{eq:del4}
-\Delta= \frac{1}{N_0\,C}\big(C_1\,S\_u^*+((C_1+C_2)(1-\theta\_w)+C\,\theta\_w)\,S\_n^*\big).
+\Delta= \frac{1}{C}\big(C_1\,S\_u^*+((C_1+C_2)(1-\theta\_w)+C\,\theta\_w)\,S\_n^*\big).
 \end{equation}
-Note that for analysis convention, we let $N_0=1$, thus $S\_u^*$ and $S\_n^*$ are in the scale of 0 to 1. 
-To derive the first result, i.e., how changes in $\theta\_w$ and $\theta\_c$ affects $\Rnum$, we factor these parameters in $\Delta$ \eqref{eq:del4}. Thus, $\Delta$ can be rewritten as
+
+{\bf The first goal} is to explore how changes in isolation, $\theta\_w$ and $\theta\_c$, affects $\Rnum$.
+Mathematically we would like to verify the sign of $\pder[\Rnum]{\theta\_w}$ and $\pder[\Rnum]{\theta\_c}$.
+We start with simpliying $\Delta$ \eqref{eq:del4} by factoring $\theta\_w$ and $\theta\_c$ in Eq.\,\eqref{eq:del4}.
+Thus, $\Delta$ can be rewritten as
 \begin{equation}
@@
 \label{eq:d_del4_thetaw}
-\pder[\Delta\,C]{\theta\_w}\bigg\rvert_{\theta\_w=1} =&
+\pder[\Delta]{\theta\_w}\,C\bigg\rvert_{\theta\_w=1} =&
 (D_2-2D_1-\omega^2\,p_I\,\hat F_I\,\theta\_c)\,S\_n^*+\gamma\,\hat F_I(\omega+\gamma) \\
@@
 
-\ali{stoped here!}
-
-Let define $\phi = \hat F_S = \frac{\rho \omega}{\omega-\rho}$, thus
+{\bf The second goal} is to explore how changes in \percap testing intensity $\rho$ affects $\Rnum$.
+Mathematically we would like to verify the sign of $\pder[\Rnum]{\rho}$, which specifically depends on $\pder[\Delta]{\rho}$.
+We use the derived expressions for $S\_u^*$ and $S\_n^*$, given by Eqs.\,\eqref{dfe}, in $\Delta$ \eqref{eq:del4_theta}.
+Also, we define $\phi = \hat F_S = \frac{\rho \omega}{\omega-\rho}$, to reparametrize $\rho$.
+This is mainly to avoid singularity in $\hat F_I$ \eqref{eq:fi}, when testing intensity $\rho$ is very close to the rate of test return $\omega$.
+Thus, $\rho$ is reparametrized as
 \begin{equation}
@@
 \end{equation}
-Reparametrizing testing rate $\rho$ by $\phi$ in the expression of $\Rnum$, in Eq.\,\eqref{R0}, and applying the derivative with respect to $\phi$ results in the following simplified expression for $\partial\Delta/\partial\phi$.
-
-with an equivalent but simplified $\Delta$ in Eq.\,\eqref{eq:abc} as
-\begin{equation}
-\label{eq:del3}
-\Delta= \frac{K_1\,\phi^2+K_2\,\phi}{K_3},
-\end{equation}
-where
-\begin{align}
-\label{eq:K}
-K_1=& \frac{w_I}{w_S}\Big(((1+p_I)\omega+\gamma)\gamma\theta\_w+\omega^2 p_I\theta\_c \Big), \\
-K_2=& (\omega+\gamma)\Big( (\frac{w_I}{w_S}\omega+\gamma)\gamma\theta\_w + \frac{w_I}{w_S} \omega^2 p_I\theta\_c \Big),\\
-K_3=& (\omega+\gamma) \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi\,+\,(\omega+\gamma)\gamma \Big)(\omega+\phi).
-\end{align}
+This one-to-one monotonic reparametrization enables us to simplify the mathematical expressions and explore the simpler $\pder[\Delta]{\phi}$ instead of the complicated $\pder[\Delta]{\rho}$.
+% with an equivalent but simplified $\Delta$ in Eq.\,\eqref{eq:abc} as
+% \begin{equation}
+% \label{eq:del3}
+% \Delta= \frac{K_1\,\phi^2+K_2\,\phi}{K_3},
+% \end{equation}
+% where
+% \begin{align}
+% \label{eq:K}
+% K_1=& \frac{w_I}{w_S}\Big(((1+p_I)\omega+\gamma)\gamma\theta\_w+\omega^2 p_I\theta\_c \Big), \\
+% K_2=& (\omega+\gamma)\Big( (\frac{w_I}{w_S}\omega+\gamma)\gamma\theta\_w + \frac{w_I}{w_S} \omega^2 p_I\theta\_c \Big),\\
+% K_3=& (\omega+\gamma) \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi\,+\,(\omega+\gamma)\gamma \Big)(\omega+\phi).
+% \end{align}
 The derivative is
@@
 where
+
+\begin{multline}
+\label{eq:a3}
+a_3 = \frac{w_I}{w_S}
+\Bigg(\Big(
+( \theta_{{w}}{{\it p\_I}}^{2}-\theta_{{c}}{{\it p\_I}}^
+{2}\theta_{{w}} ) {\omega}^{3} 
++ ( 2\,\theta_{{w}}{{\it p\_I
+}}^{2}-\theta_{{c}}{{\it p\_I}}^{2}-{\it p\_I}\,{\theta_{{w}}}^{2}-
+\theta_{{c}}{\it p\_I}\,\theta_{{w}}-{{\it p\_I}}^{2}{\theta_{{w}}}^{2
+}+2\,\theta_{{w}}{\it p\_I} ) \gamma\,{\omega}^{2} \\
+ + ( -2\,{ \it p\_I}\,{\theta_{{w}}}^{2}-\theta_{{c}}{\it p\_I}+3\,\theta_{{w}}{
+\it p\_I}-{\theta_{{w}}}^{2}+\theta_{{w}} ) {\gamma}^{2}\omega 
+ + ( \theta_{{w}}-{\theta_{{w}}}^{2} ) {\gamma}^{3} \Big) \frac{w_I}{w_S}
+ + ( -\theta_{{c}}{\it p\_I}\,\theta_{{w}}+\theta_{{c}}{\it p\_I}
+ ) \gamma\,{\omega}^{2}+ \\
+ ( \theta_{{w}}+\theta_{{c}}{\it
+p\_I}-{\theta_{{w}}}^{2}-\theta_{{c}}{\it p\_I}\,\theta_{{w}} )
+{\gamma}^{2}\omega+ ( \theta_{{w}}-{\theta_{{w}}}^{2} ) {
+\gamma}^{3}
+\Bigg),
+\end{multline}
 \begin{align}
 \label{eq:abcd2}
-a_3=& \frac{w_I}{w_S} \Big((\omega\,p_I+\gamma)(\theta\_w-\theta\_c)\frac{w_I}{w_S}\omega\,p_I+(\gamma\theta\_w+\omega\,p_I\theta\_c)(\omega+\gamma) \Big),\\
-b_3=& 2\frac{w_I}{w_S}(\omega+\gamma)\Big( ((1+p_I)\omega+\gamma)\gamma\theta\_w+\omega^2 p_I\theta\_c \Big) ,\\
-c_3=& (\omega+\gamma)^2 \Big( (\frac{w_I}{w_S}\omega+\gamma)\gamma\theta\_w + \frac{w_I}{w_S} \omega^2 p_I\theta\_c \Big),\\
-d_3=& \frac{(\omega+\gamma)}{\omega\gamma} \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi\,+\,(\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2. 
+b_3&= 2\frac{w_I}{w_S}(\omega+\gamma)\gamma\Big( 
+(\omega+\gamma+\omega\,p_I)(2-\theta\_w)\,\gamma\,\theta\_w+ (1-\theta\_w)\,\omega^2 p_I\theta\_c+\omega^2 p_I\theta\_w 
+\Big) ,\\
+c_3&= (\omega+\gamma)^2\,\gamma\Big( 
+(2-\theta\_w)\,\gamma^2\,\theta\_w+
+(1+\frac{w_I}{w_S})\,\omega\,\gamma\,\theta\_w+
+\frac{w_I}{w_S} \omega^2 p_I\theta\_c \Big),\\
+d_3&= \frac{(\omega+\gamma)}{\omega} \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi\,+\,(\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2. 
 \end{align}
-
 Note that $\phi\geq 0$, also $b_3$, $c_3$ and $d_3$ are all positive. However $a_3$ can be positive or negative.
-If $a_3\geq 0$, $\partial\Delta/\partial\phi \geq 0$ for all feasible range of parameters, thus $\pro\Delta \leq 0$. 
+If $a_3\geq 0$, $\partial\Delta/\partial\phi \geq 0$ for all feasible range of parameters, thus $\pro\Rnum \leq 0$. 
 If $a_3 < 0$, then the quadratic expression in the numerator of \eqref{eq:dd3dr} has a positive root, $\phi^*$, such that for $\phi>\phi^*$, $\partial\Delta/\partial\phi < 0$.
@@
 In this case
+\ali{I stoped here!}
 \begin{equation}

### 2021-05-05 SIR_manuscript.tex
analysis, R0 wrt rho

@@
 If $a_3 < 0$, then the quadratic expression in the numerator of \eqref{eq:dd3dr} has a positive root, $\phi^*$, such that for $\phi>\phi^*$, $\partial\Delta/\partial\phi < 0$.
-An example of this scenario occures when $\theta\_w=0$ and $\theta\_c=1$, which is illustraited in the top-right panel of the Fig.\,\ref{pan2} panel (b).
-This is when the strength of isolation for awaiting people is the least, but the most for the confirmed cases.
-In this case
+An example of this counter-vailing effect of $\rho$ on $\Rnum$ occures when $\theta\_w=0$ and $\theta\_c=1$.
+This is illustraited in the top-right panel of the Fig.\,\ref{pan2} panel (b), where the strength of isolation for awaiting people is the least, but the most for the confirmed cases.
+In this case $$a_3=\frac{w_I}{w_S}\,\omega\,\gamma\,p_I((\omega+\gamma)-\frac{w_I}{w_S}(\omega\,p_I+\gamma)).$$
+Specifically, in the case of targeted testing which is identified with $\frac{w_I}{w_S}> 1$, and using a perfect sensative test, thus $p_I=1$, there exists a range for $\rho$ over which $\pder[\Rnum]{\rho}\leq\,0$.
+Note that $\rho$ and $\omega$ have a similar mechanism in delaying people to get into $I\_c$, thus we would expect to see the counter-vailing effect, which is not obvious, of these two parameters on $\Rnum$.
+% that people in the DFE waiting for negative test, $S_n^*$.
+% More testing may have a counter effect on $\Rnum$, as if we have more testing we have more people in $S_n^*$ which will take them longer to get to $I\_c$.
 \ali{I stoped here!}
-\begin{equation}
-\label{ineq:a3}
-a_3 <0\iff\omega+\gamma<(\omega\,p_I+\gamma)\frac{w_I}{w_S}.
-\end{equation}
-Given that $\frac{w_I}{w_S}\geq 1$, and $0\leq\,p_I\leq\,1$, this inequality is hold when a perfect sensitive test is used, i.e., $p_I=1$, and a targeted tesing strategy is followed, i.e., $\frac{w_I}{w_S}>1$.
-\ali{i)what is the interpretation of this?
-2) (Ben:) calculate the corresponding eigenvector to $\Rnum$, scale it to 1, and plot the population of each infected compartment, what does this tell us?
-}
-
-Note that $\rho$ and $\omega$ have similar mechanism in delaying people to get into $I\_c$. The non monotonic countereffect of testing rate and test return rate are not obvious. 
-that people in the DFE waiting for negative test, $S_n^*$.
-More testing may have a counter effect on $\Rnum$, as if we have more testing we have more people in $S_n^*$ which will take them longer to get to $I\_c$.
-
-more $\rho$ then
 

### 2021-05-06 SIR_manuscript.tex
analysis, R0 wrt omega

@@
 where
-
 \begin{multline}
@@
 Specifically, in the case of targeted testing which is identified with $\frac{w_I}{w_S}> 1$, and using a perfect sensative test, thus $p_I=1$, there exists a range for $\rho$ over which $\pder[\Rnum]{\rho}\leq\,0$.
-Note that $\rho$ and $\omega$ have a similar mechanism in delaying people to get into $I\_c$, thus we would expect to see the counter-vailing effect, which is not obvious, of these two parameters on $\Rnum$.
-% that people in the DFE waiting for negative test, $S_n^*$.
-% More testing may have a counter effect on $\Rnum$, as if we have more testing we have more people in $S_n^*$ which will take them longer to get to $I\_c$.
-\ali{I stoped here!}
+Note that $\rho$ and $\omega$ have a similar mechanism in delaying people to get into $I\_c$, thus we would expect to see the non-trivial counter-vailing effect of these two parameters on $\Rnum$.
 
@@
 \subsection{rate of returning tests} \label{app:omega}
-\ali{needs editig}
-The linearization of $\Rnum$ around $\rho=0$ is
-\begin{equation}\label{linearization}
-\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(-\theta\_w)(\gamma w_S+\omega w_I) + (-\theta\_c)p_I w_I \omega^2 \Big). 
+{\bf The third goal} is to explore how changes in the rate of test return $\omega$ affects $\Rnum$.
+Mathematically we would like to verify the sign of $\pder[\Rnum]{\omega}$, which specifically depends on $\pder[\Delta]{\omega}$.
+We use
+the linearization of $\Rnum$ around $\rho=0$ to show that there a non-monotonic relationship between $\Rnum$  and $\omega$. The taylor expansion of $\Delta$ at $\rho=0$ is
+\begin{equation}
+\label{eq:lin}
+\Delta = \frac{\rho}{\omega\,\gamma(\omega+\gamma)} \Big(
+\frac{w_I}{w_S}\,\omega^2\,p_I\,\theta\_c+(\frac{w_I}{w_S}+1)\,\gamma\omega\theta\_w\,+\gamma^2\,\theta\_w(2-\theta\_w)
+\Big). 
 \end{equation}
-So when $\rho \approx 0$ we have $$\partial{\Rnum}/\partial{\omega} \approx  \frac{-\beta \rho}{\gamma w_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c),$$
-where $a=(-\theta\_w)w_I-(-\theta\_c)p_I w_I$, $b=2(-\theta\_w)\gamma w_S$ and $c=(-\theta\_w)\gamma^2 w_S$.
+This results in
+\begin{equation}
+\label{eq:dlindo}
+\pder[\Delta]{\omega} = \frac{\rho}{\omega^2\,(\omega+\gamma)^2} \Big(
+(p_I\,\frac{w_I}{w_S}\theta\_c\,-(1+\frac{w_I}{w_S})\theta\_w)\omega^2+2\theta_w\,\gamma(\theta_w-2)\,\omega+\theta_w\,\gamma^2\,(\theta_w-2)
+\Big),
+\end{equation}
+around $\rho=0$.
+\ali{I stoped here!}
 

### 2021-05-06 SIR_manuscript.tex
discussion points

@@
 
-Mathematical modeling of infectious disease outbreaks provides insights on how testing processes influence the epidemiological processes through isolation. 
-Here, we develop a compartmental SIR-type model to study the potential effect of testing strategies, testing intensity, test sensitivity and specificity, test reporting time and isolation on epidemic dynamics.
-While targeted testing strategies, including targeting people with infection-like symptoms or the contacts of confirmed cases, are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is that viral spread is greater for a slow test than for a fast test.
-This counter-intuitive effect can occur when people are cautious when awaiting a test result, and may not be robust to second-order effects of fast testing (such as better contact tracing).
-\ali{this is a valid sentence especially when $\theta\_w$ is closer to 1, but not sure if this was counter-intuitive or was the opposite of it?
-i.e., that viral spread is greater for a fast test ($1/\omega$ closer to 0) than for a slow test.}
-
-We incorporated the compartment-specific relative testing weights, $w_S$, $w_I$ and $w_R$, to model random testing and \emph{targeted} testing strategies.
-Here, in the case of targeted testing and for the simplicity and illustration purposes, we assumed that infected and recovered individuals are tested at three times the \percap rate of susceptible individuals, thus $w_I/w_S=3$ and $w_R=w_I$.
-Note that we have not specified a methodology to assign particular relative testing weights corresponding to a particular targeted testing scenario.
-Modeling different targeted testing strategies, equivalently test-specific testing weights in our framework, requires prior information of the conditional probabilities of getting tested for people in a given compartment.
-This can be implied when we would like to quantify and compare the effect of different levels of test focus for infectious people on the basic reproduction number $\Rnum$, and conclude about the disease spread management.
-For example, when people are tested for "screening", the individuals with potential higher mobility, eg.
-people who are getting on flights, get more tested and thus the coresponding heavier testing weight is assigned than people awaiting for a surgery and are probably going to stay in a long-term-care facility and consequently less mobile and more isolated to begin with.
-With our model, we would be able to compare the sensitivity of the disease dynamics, through $\Rnum$, with respect to testing different high-risk groups in the population.
-This part needs to be developed further in future work.
-
-{\bf The \percap testing intensity, $\rho$;}
-Proposition \ref{p1:rho} and also Figures \ref{pan} and \ref{pan2} indicates that increasing testing intensity $\rho$ may reduce $\Rnum$ but there is a counter-intuitive machanism. 
-The expected mechanism was that as more untested infected individuals are moved to the test compartments, i.e., from $I\_u$ to $I\_p$ and $I\_n$, and eventually to the confirmed compartment $I\_c$, they may partially or fully self-isolate ($\theta\_w$ and $\theta\_c$ varies between 0 and 1 with $\theta\_w\leq\theta\_c$).
-Thus, the higher the \percap testing intensity $\rho$, the lower the force of infection \eqref{Lambda} and the lower $\Rnum$ will become.
-This is true when $\rho$ is very small relative to $N_0$ and can be verified analytically and through our simulation.
-The counter-intuitive mechanism is that at the DFE, higher \percap testing intensity, may delay recognition of true infected individuals $I\_p$ and causing an increase in $\Rnum$ as follows.
-At high $\rho$, more people will be moved from $S\_u^*$ to $S\_n^*$, thus a higher proportion of awaiting people may get infected and end up in $I\_n$, and be delayed from being eligable for retesting $I\_u$ and eventually to $I\_p$ and $I\_c$.
-Our analysis (Sec.\,\ref{app:rho}) and simulation (see Fig.\,\ref{pan2} panel (b) when $\theta\_c=1$ and $\theta\_w=0$) support such a mechanism.
-In particular, if \eqref{ineq:a3} holds we will have such a result.
+{\bf The counter-vailing effect of \percap testing intensity, $\rho$, on $\Rnum$;}
+- we are missing out on community-level advantages of fast testing.
 
 {\bf The potential advantage of slow test reporting, or favorable-delay-reporting;}
-Individuals are highly likely to fully self-isolate when they are either awaiting the test results or they are reported, thus reducing or eliminating their potentially infectious contacts.
-Thus, the faster the test reporting rate, $\omega$, the shorter these individuals stay in the ``\emph {safe}'' awaiting-confirmed compartments, namely $X_n$, $X_p$ and $X_c$ for $X \in \{S,I,R\}$, and the more they get involved in the infection process.
-This advantage of slow test reporting is real, and neglected. 
-We also compare to an individual-level advantage of fast tests: people who test positive may be even more careful.
-In our model analysis, Proposition \ref{prop1} part \ref{p1:omega} is describing this potential advantage. 
-It states that returning test results more rapidly, i.e., increasing the rate $\omega$, does not necessarily lower the reproduction number $\Rnum$; whether increasing $\omega$ lowers $\Rnum$ depends on the precise combination of model parameters  including test reporting rate, testing strategies represented by compartment-specific testing weights, test sensitivity and specificity, and the level of isolation.
-Specifically, in the case of perfect isolation, i.e., when $\theta\_w=1$ and $\theta\_c=1$, $\Rnum$ may increase as the test reporting process becomes faster.
-This can be seen from expression \eqref{Rom}.
-Another example of this favorable delay in $\omega$ could be when the test being employed produces many false negatives.
-Because many infected asymptomatic individuals will believe they are uninfected/uninfectious, thus may unknowingly spread the virus to many others.
-Again the delay in the test reporting rate keeps these individuals in the ``\emph {safe}'' awaiting-confirmed compartments.
-
-we are missing out on
-community-level advantages of fast testing: better assessment of the
-situation, identification of hot spots, contact-tracing, etc.
-The \percap testing intensity of CIVID-19 after about a year from the first case reported in December 2019, is still low (($\rho \approx 0$) in our model).
-In near future new test kits may be widely accessible, our model provide insights in this case.
-In particular, if a cheap test can identify on average more infected individuals as an expensive test, then our model predicts that the cheap test will lower $\Rnum$ more.
-In contrast, if an expensive test can identify on average more infected individuals, it will not necessarily lower $\Rnum$ more than the cheap test.
-The use of tests cheaper than RT-PCR has been proposed as a potential strategy for containing the COVID pandemic.
-While cheaper tests may be less sensitive and reliable than RT-PCR, they allow for broader and more intense testing.
-Using our Taylor approximation of $\Rnum$ near $\rho = 0$, we examined what circumstances (i.e., model parameters) make the use of one test more favourable than another, and give a complete description through inequality \ref{eq:rho1vsrho2}.
-In general, we found that the expensive test tends to more effectively lower $\Rnum$ when (a) individuals who test positive self-isolate much more than individuals who are waiting for their test result, (b) the time it takes to return tests is much shorter than the mean infectious period, and (c) the testing intensity is much greater for infected individuals than susceptible individuals.
-
-\ali{ not sure if we want these 2 following parageraphs, or shrink it into afew short sentences?}
-In addition to the favorable-delay-reporting observation of our model \eqref{model}, the model enables us to quantify the amount of delay required in the test reporting process as a strategy to reduce $\Rnum$.
-To give a biological interpretation, we describe the qualitative trends predicted by the inequality \eqref{eq:necsuf}.
-To summarise, returning test results more rapidly tends to be favourable (i.e., reduces $\Rnum$) when (i) \emph{Confirmed-positive} individuals, lower their contact much more than individuals who are waiting for their test results (i.e., $\theta\_c \gg \theta\_w$), (ii) the test is highly sensitive (i.e., $p_I$ is close to 1), and (iii) the targeted testing strategy is used (i.e., $w_I \gg w_S$).
-
-one point we can make when discussing the relative weight of testing in different compartments is to discuss pre-testing screening tools, such as surveys or questionnaires.
-If we have a quantitative description of how the testing intensities affect the dynamics, we can make statements like ``our results suggest that employing a pre-testing screening tool can help target infected individuals more effectively.
-In particular, doubling the sensitivity of the pre-screening tool would *do something* to $\Rnum$.
-
-\david{Some of this reads like notes for discussion rather than text for the paper, so I'm not trying to edit.}
 

### 2021-05-06 note/SIR_scrap.tex
discussion points

@@
 \section{Discussion}
+
+Mathematical modeling of infectious disease outbreaks provides insights on how testing processes influence the epidemiological processes through isolation. 
+Here, we develop a compartmental SIR-type model to study the potential effect of testing strategies, testing intensity, test sensitivity and specificity, test reporting time and isolation on epidemic dynamics.
+While targeted testing strategies, including targeting people with infection-like symptoms or the contacts of confirmed cases, are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is that viral spread is greater for a slow test than for a fast test.
+This counter-intuitive effect can occur when people are cautious when awaiting a test result, and may not be robust to second-order effects of fast testing (such as better contact tracing).
+\ali{this is a valid sentence especially when $\theta\_w$ is closer to 1, but not sure if this was counter-intuitive or was the opposite of it?
+i.e., that viral spread is greater for a fast test ($1/\omega$ closer to 0) than for a slow test.}
+
+We incorporated the compartment-specific relative testing weights, $w_S$, $w_I$ and $w_R$, to model random testing and \emph{targeted} testing strategies.
+Here, in the case of targeted testing and for the simplicity and illustration purposes, we assumed that infected and recovered individuals are tested at three times the \percap rate of susceptible individuals, thus $w_I/w_S=3$ and $w_R=w_I$.
+Note that we have not specified a methodology to assign particular relative testing weights corresponding to a particular targeted testing scenario.
+Modeling different targeted testing strategies, equivalently test-specific testing weights in our framework, requires prior information of the conditional probabilities of getting tested for people in a given compartment.
+This can be implied when we would like to quantify and compare the effect of different levels of test focus for infectious people on the basic reproduction number $\Rnum$, and conclude about the disease spread management.
+For example, when people are tested for "screening", the individuals with potential higher mobility, eg.
+people who are getting on flights, get more tested and thus the coresponding heavier testing weight is assigned than people awaiting for a surgery and are probably going to stay in a long-term-care facility and consequently less mobile and more isolated to begin with.
+With our model, we would be able to compare the sensitivity of the disease dynamics, through $\Rnum$, with respect to testing different high-risk groups in the population.
+This part needs to be developed further in future work.
+
+{\bf The \percap testing intensity, $\rho$;}
+Proposition \ref{p1:rho} and also Figures \ref{pan} and \ref{pan2} indicates that increasing testing intensity $\rho$ may reduce $\Rnum$ but there is a counter-intuitive machanism. 
+The expected mechanism was that as more untested infected individuals are moved to the test compartments, i.e., from $I\_u$ to $I\_p$ and $I\_n$, and eventually to the confirmed compartment $I\_c$, they may partially or fully self-isolate ($\theta\_w$ and $\theta\_c$ varies between 0 and 1 with $\theta\_w\leq\theta\_c$).
+Thus, the higher the \percap testing intensity $\rho$, the lower the force of infection \eqref{Lambda} and the lower $\Rnum$ will become.
+This is true when $\rho$ is very small relative to $N_0$ and can be verified analytically and through our simulation.
+The counter-intuitive mechanism is that at the DFE, higher \percap testing intensity, may delay recognition of true infected individuals $I\_p$ and causing an increase in $\Rnum$ as follows.
+At high $\rho$, more people will be moved from $S\_u^*$ to $S\_n^*$, thus a higher proportion of awaiting people may get infected and end up in $I\_n$, and be delayed from being eligable for retesting $I\_u$ and eventually to $I\_p$ and $I\_c$.
+Our analysis (Sec.\,\ref{app:rho}) and simulation (see Fig.\,\ref{pan2} panel (b) when $\theta\_c=1$ and $\theta\_w=0$) support such a mechanism.
+In particular, if \eqref{ineq:a3} holds we will have such a result.
+
+{\bf The potential advantage of slow test reporting, or favorable-delay-reporting;}
+Individuals are highly likely to fully self-isolate when they are either awaiting the test results or they are reported, thus reducing or eliminating their potentially infectious contacts.
+Thus, the faster the test reporting rate, $\omega$, the shorter these individuals stay in the ``\emph {safe}'' awaiting-confirmed compartments, namely $X_n$, $X_p$ and $X_c$ for $X \in \{S,I,R\}$, and the more they get involved in the infection process.
+This advantage of slow test reporting is real, and neglected. 
+We also compare to an individual-level advantage of fast tests: people who test positive may be even more careful.
+In our model analysis, Proposition \ref{prop1} part \ref{p1:omega} is describing this potential advantage. 
+It states that returning test results more rapidly, i.e., increasing the rate $\omega$, does not necessarily lower the reproduction number $\Rnum$; whether increasing $\omega$ lowers $\Rnum$ depends on the precise combination of model parameters  including test reporting rate, testing strategies represented by compartment-specific testing weights, test sensitivity and specificity, and the level of isolation.
+Specifically, in the case of perfect isolation, i.e., when $\theta\_w=1$ and $\theta\_c=1$, $\Rnum$ may increase as the test reporting process becomes faster.
+This can be seen from expression \eqref{Rom}.
+Another example of this favorable delay in $\omega$ could be when the test being employed produces many false negatives.
+Because many infected asymptomatic individuals will believe they are uninfected/uninfectious, thus may unknowingly spread the virus to many others.
+Again the delay in the test reporting rate keeps these individuals in the ``\emph {safe}'' awaiting-confirmed compartments.
+
+we are missing out on
+community-level advantages of fast testing: better assessment of the
+situation, identification of hot spots, contact-tracing, etc.
+The \percap testing intensity of CIVID-19 after about a year from the first case reported in December 2019, is still low (($\rho \approx 0$) in our model).
+In near future new test kits may be widely accessible, our model provide insights in this case.
+In particular, if a cheap test can identify on average more infected individuals as an expensive test, then our model predicts that the cheap test will lower $\Rnum$ more.
+In contrast, if an expensive test can identify on average more infected individuals, it will not necessarily lower $\Rnum$ more than the cheap test.
+The use of tests cheaper than RT-PCR has been proposed as a potential strategy for containing the COVID pandemic.
+While cheaper tests may be less sensitive and reliable than RT-PCR, they allow for broader and more intense testing.
+Using our Taylor approximation of $\Rnum$ near $\rho = 0$, we examined what circumstances (i.e., model parameters) make the use of one test more favourable than another, and give a complete description through inequality \ref{eq:rho1vsrho2}.
+In general, we found that the expensive test tends to more effectively lower $\Rnum$ when (a) individuals who test positive self-isolate much more than individuals who are waiting for their test result, (b) the time it takes to return tests is much shorter than the mean infectious period, and (c) the testing intensity is much greater for infected individuals than susceptible individuals.
+
+\ali{ not sure if we want these 2 following parageraphs, or shrink it into afew short sentences?}
+In addition to the favorable-delay-reporting observation of our model \eqref{model}, the model enables us to quantify the amount of delay required in the test reporting process as a strategy to reduce $\Rnum$.
+To give a biological interpretation, we describe the qualitative trends predicted by the inequality \eqref{eq:necsuf}.
+To summarise, returning test results more rapidly tends to be favourable (i.e., reduces $\Rnum$) when (i) \emph{Confirmed-positive} individuals, lower their contact much more than individuals who are waiting for their test results (i.e., $\theta\_c \gg \theta\_w$), (ii) the test is highly sensitive (i.e., $p_I$ is close to 1), and (iii) the targeted testing strategy is used (i.e., $w_I \gg w_S$).
+
+one point we can make when discussing the relative weight of testing in different compartments is to discuss pre-testing screening tools, such as surveys or questionnaires.
+If we have a quantitative description of how the testing intensities affect the dynamics, we can make statements like ``our results suggest that employing a pre-testing screening tool can help target infected individuals more effectively.
+In particular, doubling the sensitivity of the pre-screening tool would *do something* to $\Rnum$.
+
+\david{Some of this reads like notes for discussion rather than text for the paper, so I'm not trying to edit.}
 
@@
 \subsection{More testing intensity decreases $\Rnum$}
+% that people in the DFE waiting for negative test, $S_n^*$.
+% More testing may have a counter effect on $\Rnum$, as if we have more testing we have more people in $S_n^*$ which will take them longer to get to $I\_c$.
 

### 2021-05-06 SIR_manuscript.tex
minor changes in abstract and analysis

@@
 
-The effect of testing processes, including (testing and test reporting) on an epidemic dynamics (infection and recovery) can be studied at the individual level or the community level (e.g., nursing homes, long-term-care facilities, etc.).
+The effect of testing processes, including testing and test reporting, on an epidemic dynamics, involving infection and recovery, can be studied at the individual level or the community level (e.g., nursing homes, long-term-care facilities, etc.).
 Gaining insights to determine the sensitivity of the epidemic dynamics with respect to the testing processes will depend on underlying factors including the level of focus (individual or community), assumptions (model), and the interplay between these factors.
@@
 \end{align}
-The scaled expression of $\Delta$, specifically $\Delta\,C$, in Eq.\,\eqref{eq:del4_theta} is linear in $\theta\_c$ with a positive coefficient given by 
-$$(\gamma\,S\_u^*\,+\omega(1-\theta\_w\,S\_u^*))\omega\,p_I\,\hat F_I.$$
+$\Delta$, Eq.\,\eqref{eq:del4_theta}, is linear in $\theta\_c$ with a positive coefficient given by 
+$$1/C\,(\gamma\,S\_u^*\,+\omega(1-\theta\_w\,S\_u^*))\omega\,p_I\,\hat F_I.$$
 This results in increasing $\Delta$, thus decreasing $\Rnum$ with respect to $\theta\_c$, that is $\pder[\Rnum]{\theta\_c}\leq 0$. Note that $C$ is independant of $\theta\_c$ and $\theta\_w$.
 
-With a similar logic, the expression of $\Delta\,C$ derived from Eq.\,\eqref{eq:del4_theta} is a concave-down quadratic equation in $\theta\_w$, given by
+With a similar logic, $\Delta$ \eqref{eq:del4_theta} is a concave-down quadratic equation in $\theta\_w$, given by
 \begin{equation}
 \label{eq:del4_thetaw}
--D_1\,S\_n^*\,\boldsymbol {\theta\_w^2}+
-\big(-\omega^2\,p_I\,\hat F_I\,S\_n^*\,\boldsymbol{\theta\_c}\,+\,D_2\,S\_n^*\,+\,\gamma\,\hat F_I(\omega+\gamma) \big)\boldsymbol{\theta\_w}.
+1/C\,\Big( -D_1\,S\_n^*\,\boldsymbol {\theta\_w^2}+
+\big(-\omega^2\,p_I\,\hat F_I\,S\_n^*\,\boldsymbol{\theta\_c}\,+\,D_2\,S\_n^*\,+\,\gamma\,\hat F_I(\omega+\gamma) \big)\boldsymbol{\theta\_w} \Big).
 \end{equation}
@@
 \label{eq:d_del4_thetaw}
-\pder[\Delta]{\theta\_w}\,C\bigg\rvert_{\theta\_w=1} =&
-(D_2-2D_1-\omega^2\,p_I\,\hat F_I\,\theta\_c)\,S\_n^*+\gamma\,\hat F_I(\omega+\gamma) \\
-=& (\gamma(\omega+\gamma)+\gamma\omega^2+(1-\theta\_c)\omega^2\,p_I\,\hat F_I )\,S\_n^*
-+\gamma(\omega+\gamma)\,\hat F_I\,(1-S\_n^*),
+\pder[\Delta]{\theta\_w}\bigg\rvert_{\theta\_w=1} =&
+1/C\Big( (D_2-2D_1-\omega^2\,p_I\,\hat F_I\,\theta\_c)\,S\_n^*+\gamma\,\hat F_I(\omega+\gamma) \Big) \\
+=& 1/C\Big( (\gamma(\omega+\gamma)+\gamma\omega^2+(1-\theta\_c)\omega^2\,p_I\,\hat F_I )\,S\_n^*
++\gamma(\omega+\gamma)\,\hat F_I\,(1-S\_n^*) \Big),
 \end{align}

### 2021-05-07 SIR_manuscript.tex
changes in methods

@@
 where the term $\frac{\beta}{\gamma}$ is the classical basic reproduction number for a SIR model without testing and isolation (see, e.g., \cite{keeling2011modeling}), and $\Delta$ is the reduction parameter defined as follows.
-\ali{$\Delta$ should be updated, the expression will be linear in $\theta\_c$ but quadratic in $\theta\_w$, C stay unchanged.
-It can be proved that $\Delta$ increases in $\theta\_w$ for feasible range of $\theta\_w$ (I think this argument will go to the appendix).?}
 \begin{equation}
-\label{eq:Delta1}
-\Delta= (A~ \theta\_w + B~ \theta\_c) / C,
+\label{eq:del4}
+\Delta= \frac{1}{C}\big(C_1\,S\_u^*+(C_2(1-\theta\_w)+C\,\theta\_w)\,S\_n^*\big),
 \end{equation}
 where
-\begin{equation}
-\begin{aligned}
-\label{eq:abc}
-A=& \gamma \Big( \big(\gamma(\omega+\gamma) + \omega p_I \hat F_I \big) \frac{(1-\theta\_w)S\_n^*}{N_0} + (\omega+\gamma) \hat F_I \Big),\\
-B=& \omega p_I(\gamma \frac{S\_u^*}{N_0}+\omega)\hat F_I, \\ 
-C=& (\omega+\gamma) \Big(\gamma(\omega+\gamma)+(\gamma+\omega p_I)\hat F_I\Big).
-\end{aligned}
-\end{equation}
-Further details of derivation of $\Rnum$ are provided in section \ref{app:R0} of the appendix.
-Note that $\Delta$ has been presented in a specific monotonic form with respect to isolation efficacy parameters, $\theta\_w$ and $\theta\_c$, in order to simplify showing one of the results of this paper which is the reduction effect of isolation on $\Rnum$.
-Further, we represent $\Delta$ in another equivalent form in the appendix (Sec.
-\ref{app:rho}) to show the nontrivial-nonmonotonic effect of testing intensity and test return rate on $\Rnum$.
+\begin{align}
+\label{eq:C12}
+C =& (\omega+\gamma) \Big(\gamma(\omega+\gamma)+(\gamma+\omega p_I)\hat F_I\Big),\\
+C_1 =& (\omega+\gamma)(\theta\_w \, \gamma+\theta\_c\, \omega\, p_I) \hat F_I,\\
+C_2 =& \Big( \omega\gamma(1+p_I)\hat F_I+\gamma^2\,(\omega+\gamma+\hat F_I)\Big)\theta\_w + \omega^2\,p_I\,\hat F_I\,\theta\_c.
+\end{align}
+Further details of derivation of $\Rnum$ are provided in appendix sec.\,\ref{app:R0}.
+Further, we show the nontrivial-nonmonotonic effect of testing intensity on $\Rnum$ appendix (See Sec.\,\ref{app:rho}).
+It is notable that this non-monotonic effect also occures with the rate of test return $\omega$.
+We show this for the case when testing intesity is very small by using the Taylor approximation of $\Rnum$ at $\rho=0$.
+\ali{some of these are methods!}
 
@@
 This expression can be simplifed and represented given the specific form of matrix $G_{11}$ \eqref{G11}.
-For the purpose of simplicity we present $\Rnum$ in the manuscript in terms of expressions $A$, $B$ and $C$, specified in \eqref{eq:abc}.
-This is intended to show that $\Rnum$ decreases as the strength of isolation, shown by $\theta\_w$ and $\theta\_c$, increases.
+For the purpose of simplicity we present $\Rnum$ in the manuscript in terms of expressions $C$, $C1$ and $C2$, specified in \eqref{eq:C12}.
 
@@
 
-This is to provide a mathematical materials to prove that $\pro \Delta$ can be positive or negative, with $\Delta$ defined in Eq.\,\eqref{eq:abc}, and thus $\pro \Rnum < 0$, where $\Rnum$ is the basic reproduction number, given in Eq.\,\eqref{R0}.
+This is to provide a mathematical materials to prove that $\pro \Delta$ can be positive or negative, with $\Delta$ defined in Eq.\,\eqref{eq:C12}, and thus $\pro \Rnum < 0$, where $\Rnum$ is the basic reproduction number, given in Eq.\,\eqref{R0}.
 We can rewrite matrix $G_{11}$ in \eqref{G11} in the following form to simplify the calculations.
@@
 G_{11} = \frac{\beta}{\gamma \, C} 
-% \left[ \begin {array}{c}  (\omega-\rho)/\omega \\ (1-\theta\_w)\rho/\omega  \end {array} \right]
 \left[ \begin {array}{c}  S\_u^*/N_0 \\ (1-\theta\_w)\,S\_n^*/N_0  \end {array} \right]
 \left[ \begin {array}{cccc} 
-C-C_1, C-C_1-C_2\end {array} \right],
+C-C_1, C-C_2\end {array} \right],
 \end{equation}
-where $C$ is the same as the one in Eq. \eqref{eq:abc}, i.e., 
+where $C$ is the same as the one in Eq. \eqref{eq:C12}, i.e., 
 $$C=(\omega+\gamma)(\gamma(\omega+\gamma)+(\omega\,p_I+\gamma)\,\hat F_I),$$
 and $C_1$ and $C_2$ are 
-\begin{align}
-\label{eq:C12}
+\begin{align*}
+% \label{eq:C12}
 C_1 =& (\omega+\gamma)(\theta\_w \, \gamma+\theta\_c\, \omega\, p_I) \hat F_I,\\
-C_2 =& (\omega+\gamma)\gamma^2\,\theta\_w+(\theta\_w-\theta\_c)\gamma\,\omega\,p_I\,\hat F_I,
-\end{align}
+C_2 =& \big( \omega\gamma(1+p_I)\hat F_I+\gamma^2\,(\omega+\gamma+\hat F_I)\big)\theta\_w + \omega^2\,p_I\,\hat F_I\,\theta\_c,
+\end{align*}
 where $\hat F_I$ is given in Eq.\,\eqref{eq:fi}.
@@
 where 
-\begin{equation}
-\label{eq:del4}
-\Delta= \frac{1}{C}\big(C_1\,S\_u^*+((C_1+C_2)(1-\theta\_w)+C\,\theta\_w)\,S\_n^*\big).
-\end{equation}
+\begin{equation*}
+% \label{eq:del4}
+\Delta= \frac{1}{C}\big(C_1\,S\_u^*+(C_2(1-\theta\_w)+C\,\theta\_w)\,S\_n^*\big).
+\end{equation*}
 
@@
 This one-to-one monotonic reparametrization enables us to simplify the mathematical expressions and explore the simpler $\pder[\Delta]{\phi}$ instead of the complicated $\pder[\Delta]{\rho}$.
-% with an equivalent but simplified $\Delta$ in Eq.\,\eqref{eq:abc} as
-% \begin{equation}
-% \label{eq:del3}
-% \Delta= \frac{K_1\,\phi^2+K_2\,\phi}{K_3},
-% \end{equation}
-% where
-% \begin{align}
-% \label{eq:K}
-% K_1=& \frac{w_I}{w_S}\Big(((1+p_I)\omega+\gamma)\gamma\theta\_w+\omega^2 p_I\theta\_c \Big), \\
-% K_2=& (\omega+\gamma)\Big( (\frac{w_I}{w_S}\omega+\gamma)\gamma\theta\_w + \frac{w_I}{w_S} \omega^2 p_I\theta\_c \Big),\\
-% K_3=& (\omega+\gamma) \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi\,+\,(\omega+\gamma)\gamma \Big)(\omega+\phi).
-% \end{align}
 The derivative is

### 2021-05-07 SIR_manuscript.tex
minor edits in abstract

@@
 Furthermore, we show that increasing \percap testing intensity and reducing the test return time would be beneficial on the dynamics of an epidemic in general but there are exception cases.
-In particular, it is possible for the basic reproduction number, $\Rnum$, to be increasing with respect to the \percap testing intensity and the rate of test return when the isolation efficacy in reduction of the probability of transmission for awaiting individuals is loose.
+In particular, it is possible for the basic reproduction number, $\Rnum$, to be increasing with respect to the \percap testing intensity and the rate of test return.
 
@@
 \end{align}
-Further details of derivation of $\Rnum$ are provided in appendix sec.\,\ref{app:R0}.
-Further, we show the nontrivial-nonmonotonic effect of testing intensity on $\Rnum$ appendix (See Sec.\,\ref{app:rho}).
+Further details of derivation of $\Rnum$ are provided in appendix, Sec.\,\ref{app:R0}.
+Further, we show the nontrivial and nonmonotonic effect of testing intensity on $\Rnum$ in appendix, Sec.\,\ref{app:rho}.
 It is notable that this non-monotonic effect also occures with the rate of test return $\omega$.
 We show this for the case when testing intesity is very small by using the Taylor approximation of $\Rnum$ at $\rho=0$.
-\ali{some of these are methods!}
 
 The analytical calculation of the next-generation matrix and simplifying the expression of $\Rnum$ was carried out in Maple \citep{maple14}. 
-We used R \citep{r} for simulations, in particular,  for plotting the contours of $\Delta$ \eqref{eq:Delta1} over a range of selected set of parameters of interest, i.e., parameters that could be manipulated by public-health policy.
+We used R \citep{r} for simulations, in particular,  for plotting the contours of $\Delta$ \eqref{eq:del4} over a range of selected set of parameters of interest, i.e., parameters that could be manipulated by public-health policy.
 These parameters include the isolation efficacy parameters, $\theta\_c$ and $\theta\_w$, \percap testing intensity, $\rho$, and the rate of test return, $\omega$.
@@
 
-Numerical results are shown in Fig. \ref{pan}.
+Numerical results are shown in \fref{pan}.
 Targeted testing (panel b) is more effective at control than random testing (panel a).
@@
 Mathematically we would like to verify the sign of $\pder[\Rnum]{\rho}$, which specifically depends on $\pder[\Delta]{\rho}$.
-We use the derived expressions for $S\_u^*$ and $S\_n^*$, given by Eqs.\,\eqref{dfe}, in $\Delta$ \eqref{eq:del4_theta}.
+We use the derived expressions for $S\_u^*$ and $S\_n^*$, given by Eqs.\,\eqref{dfe}, in $\Delta$ \eqref{eq:del4}.
 Also, we define $\phi = \hat F_S = \frac{\rho \omega}{\omega-\rho}$, to reparametrize $\rho$.
@@
 If $a_3 < 0$, then the quadratic expression in the numerator of \eqref{eq:dd3dr} has a positive root, $\phi^*$, such that for $\phi>\phi^*$, $\partial\Delta/\partial\phi < 0$.
-An example of this counter-vailing effect of $\rho$ on $\Rnum$ occures when $\theta\_w=0$ and $\theta\_c=1$.
+
+An example of this counter-vailing effect of $\phi$, and consequently $\rho$, on $\Rnum$ occures when $\theta\_w=0$ and $\theta\_c=1$.
 This is illustraited in the top-right panel of the Fig.\,\ref{pan2} panel (b), where the strength of isolation for awaiting people is the least, but the most for the confirmed cases.

### 2021-05-08 SIR_manuscript.tex
more edits in the methods

@@
 \usepackage{pdflscape}
+\usepackage{enumerate}
 
@@
 
-We developed a deterministic model, Eqs.
-\eqref{model}, which groups individuals based on disease status and testing status.
+We developed a deterministic model, Eqs.\,\eqref{model}, which groups individuals based on disease status and testing status.
 Disease states include Susceptible, Infectious and Recovered (thus this is an SIR-type model), and testing status categorizes people as \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive} (Figure~\ref{fig:flowchart}).
@@
 \begin{center} 
-\includegraphics[scale=2]{pix/sir_comp.pdf}
+\includegraphics[scale=1.5]{pix/sir_comp.pdf}
 \caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model, \ref{model}.
@@
 Thus, we can spesify different levels of testing “focus” from random to highly targeted.
-For example, $w_I/w_S=2$ means that infected individuals are tested at twice the \percap rate of susceptible individuals.
+For example, $w_I/w_S=3$ means that infected individuals are tested at tree times the \percap rate of susceptible individuals.
 
@@
 The classical SIR model is based on the following implicit assumptions; well-mixed population, homogeneity of the population (i.e., all individuals are equaly susciptable and equaly infectious for the same length of time when infected), exponentially distributed duration of infection and large population size (see, e.g., \cite{keeling2011modeling}).
-In addition to these standard assumptions, our model, \ref{model}, assumes: (i) there is a single force of infection (new cases per unit time), $\Lambda$, defined as follows
+In addition to these standard assumptions, our model assumes:
+\begin{enumerate}[(i)]
+\item there is a single force of infection (new cases per unit time), $\Lambda$, defined as follows
   \begin{equation}
@@
   \end{equation}
-across all susceptible pools with transmission rate $\beta$ and isolation efficacy in reduction of the probability of transmission for three testing-based compartments ``waiting'' and \emph{confirmed positive} individuals, $\theta\_w$ and $\theta\_c$ respectively, (see Table~\ref{tab:params} for further details), (ii) $\theta\_w\leq\theta\_c$, i.e., the individuals awaiting test results have a higher transmission probability than the reported individuals.
+  across all susceptible pools with transmission rate $\beta$ and isolation efficacy in reduction of the probability of transmission for three testing-based compartments ``waiting'' and \emph{confirmed positive} individuals, $\theta\_w$ and $\theta\_c$ respectively, (see Table~\ref{tab:params} for further details),
+\item the efficacy of isolation is higher in confirmed individuals relative to the individuals awaiting test results. That is,
+$$0\leq\theta\_w\leq\theta\_c\leq 1.$$ 
+\end{enumerate}
 Thus, for instance when the awaiting people follow the isolation perfectly, $\theta\_w$ is closer to 1, while when they less follow the isolation, $\theta\_w$ is closer to 0.
-For this analysis, we also assume a perfectly specific test ($p_S=0$).
+For the simplicity of analysis, we also assume a perfectly specific test ($p_S=0$).
 This last assumption combined with the assumption that no susceptible individual is in waiting-for-\emph{positive} or \emph{confirmed positive} compartments, i.e., $S\_p(0)=S\_c(0)=0$, reduces the model to 10 equations where equations c and d in model \eqref{model} are eliminated.
@@
 
-The Disease-Free Equilibrium (DFE) for the SIR model, Eqs. \eqref{model}, is found by setting the infected compartments to 0 and solving for the unknowns. The DFE is
+The Disease-Free Equilibrium (DFE) for the SIR model (Eqs.\,\ref{model}) is found by setting the infected compartments to 0 and solving for the unknowns. The DFE is
 \begin{equation}
@@
 \end{equation}
-The corresponding \percap testing rate (Eq. \ref{F}) for the infected compartment $I$ at DFE is one of the key analysis parameters and can be simplified as 
+The corresponding \percap testing rate (Eq.\,\ref{F}) for the infected compartment $I$ at DFE is one of the key analysis parameters and can be simplified as 
 \begin{equation}
@@
 
-The basic reproduction number, $\Rnum$, was calculated by using the next generation matrix method developed by \cite{van2002reproduction}. $\Rnum$ is
+The basic reproduction number, $\Rnum$, was calculated by using the next generation matrix method developed by \cite{van2002reproduction}. We present $\Rnum$ in the following form.
 \begin{equation}
@@
 \end{equation}
-where the term $\frac{\beta}{\gamma}$ is the classical basic reproduction number for a SIR model without testing and isolation (see, e.g., \cite{keeling2011modeling}), and $\Delta$ is the reduction parameter defined as follows.
+where the term $\frac{\beta}{\gamma}$ is the classical basic reproduction number for a SIR model without testing and isolation (see, e.g., \cite{keeling2011modeling}), and the positive quantity $\Delta$ is the ``reduction parameter'' due to testing and isolation processes defined as follows.
 \begin{equation}
 \label{eq:del4}
-\Delta= \frac{1}{C}\big(C_1\,S\_u^*+(C_2(1-\theta\_w)+C\,\theta\_w)\,S\_n^*\big),
+\Delta= \frac{1}{C\,N_0}\big(C_1\,S\_u^*+(C_2(1-\theta\_w)+C\,\theta\_w)\,S\_n^*\big),
 \end{equation}
@@
 Further details of derivation of $\Rnum$ are provided in appendix, Sec.\,\ref{app:R0}. 
-Further, we show the nontrivial and nonmonotonic effect of testing intensity on $\Rnum$ in appendix, Sec.\,\ref{app:rho}.
-It is notable that this non-monotonic effect also occures with the rate of test return $\omega$.
-We show this for the case when testing intesity is very small by using the Taylor approximation of $\Rnum$ at $\rho=0$.
+The explicit formula of $\Rnum$, provides a mathematical ground to study the effect of parameters of interest, i.e., parameters that could be manipulated by public-health policy, on $\Rnum$.
+These parameters include the isolation efficacy parameters, $\theta\_c$ and $\theta\_w$, \percap testing intensity, $\rho$, and the rate of test return, $\omega$.
+In partticular, we look at the partial derivatives of $\Delta$ with respect to these parameters (see details in appendix, Sec.\,\ref{app:rho}).
+It is notable that we study the effect of $\omega$ on $\Rnum$ for the case when testing intesity is very small by using the Taylor approximation of $\Rnum$ at $\rho=0$.
 

### 2021-05-08 SIR_manuscript.tex
more edits in the methods

@@
 We developed a deterministic model, Eqs.\,\eqref{model}, which groups individuals based on disease status and testing status.
-Disease states include Susceptible, Infectious and Recovered (thus this is an SIR-type model), and testing status categorizes people as \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive} (Figure~\ref{fig:flowchart}).
+Disease states include Susceptible, Infectious and Recovered (thus this is an SIR-type model), and testing status categorizes people as \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive} (\fref{fig:flowchart}).
 Symbolically, the testing status of an individual in the disease compartment $X$, where $X \in \{S,I,R\}$, is reflected in the subscript, namely $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
-Note that the top-to-bottom order of the testing-based compartments of each disease-based compartmenet $X$ in the flowchart (Figure~\ref{fig:flowchart}) and the model equations \eqref{model} should match.
-However, we switched  $X\_u$ and $X\_n$ in the flowchart (Figure~\ref{fig:flowchart}) for the sake of tidiness of the flowchart.
+Note that the top-to-bottom order of the testing-based compartments of each disease-based compartmenet $X$ in the flowchart (\fref{fig:flowchart}) and the model equations \eqref{model} should match.
+However, we switched  $X\_u$ and $X\_n$ in the flowchart (\fref{fig:flowchart}) for the sake of tidiness of the flowchart.
 Further, two `accumulator' compartments, $N$ and $P$, were also incorporated in the model in order to collect cumulative reported negative or positive tests.
@@
 
-The basic reproduction number, $\Rnum$, was calculated by using the next generation matrix method developed by \cite{van2002reproduction}. We present $\Rnum$ in the following form.
+The basic reproduction number, $\Rnum$, was calculated by using the next-generation matrix method developed by \cite{van2002reproduction}. We present $\Rnum$ in the following form.
 \begin{equation}
@@
 The explicit formula of $\Rnum$, provides a mathematical ground to study the effect of parameters of interest, i.e., parameters that could be manipulated by public-health policy, on $\Rnum$.
-These parameters include the isolation efficacy parameters, $\theta\_c$ and $\theta\_w$, \percap testing intensity, $\rho$, and the rate of test return, $\omega$.
-In partticular, we look at the partial derivatives of $\Delta$ with respect to these parameters (see details in appendix, Sec.\,\ref{app:rho}).
-It is notable that we study the effect of $\omega$ on $\Rnum$ for the case when testing intesity is very small by using the Taylor approximation of $\Rnum$ at $\rho=0$.
+These parameters include the isolation efficacy, $\theta\_c$ and $\theta\_w$, \percap testing intensity, $\rho$, and the rate of test return, $\omega$.
+In partticular, we look at the partial derivative of $\Delta$ with respect to these parameters (see details in appendix, Sec.\,\ref{app:rho}).
+It is notable that the derivative of $\Delta$ with respect to $\theta\_w$, $\theta\_c$ and $\rho$ was derived explicitly and for the whole domain of the parameter space.
+However, we analyzed the effect of $\omega$ on $\Rnum$ for the case when testing intesity is very small relative to the population size, $N_0$.
+Specifically, we used the Taylor approximation of $\Delta$ at $\rho=0$ to illustrate the unexpected nonmonotonic relationship between $\Rnum$ and $\omega$, and also to aviod unnecessary complicated computations for making this point.
 
 The analytical calculation of the next-generation matrix and simplifying the expression of $\Rnum$ was carried out in Maple \citep{maple14}.
-We used R \citep{r} for simulations, in particular,  for plotting the contours of $\Delta$ \eqref{eq:del4} over a range of selected set of parameters of interest, i.e., parameters that could be manipulated by public-health policy.
-These parameters include the isolation efficacy parameters, $\theta\_c$ and $\theta\_w$, \percap testing intensity, $\rho$, and the rate of test return, $\omega$.
+We used R \citep{r} for simulations, in particular,  for plotting the contours of $\Delta$ \eqref{eq:del4} over a range of selected set of parameters of interest.
 The rest of model parameters kept fixed at sensible values (see Table~\ref{tab:params} for the parameter values).
-Note that all plots are illustrated in the scale of the mean test return time $1/\omega$ (day).
-Results are shown in Fig.\,\ref{pan} and Fig.\,\ref{pan2} with two panels; panel (a) represents the random testing, thus $w_S=w_I=w_R=1$, and panel (b) represents targeted testing, thus $w_S=0.3$ and $w_I=w_R=1$.
-To illustrate the changes in $\Rnum$ with respect to \percap testing intensity$\rho$, two sets of plots are presented. 
-Fig.\,\ref{pan} reflects the changes in $\Rnum$ when $\rho$ is small relative to the population size.
+Note that all plots are illustrated in the scale of the mean test return time $1/\omega$ (day) and testing intensity $\rho$ (1\//day per 1000).
+Results are shown in \fref{pan} and \fref{pan2} with two sets of panels; panel set (a) represents the random testing, identified by $w_S=w_I=w_R=1$, and panel set (b) represents targeted testing, identified by $w_S=0.3$ and $w_I=w_R=1$.
+Each panel set partitions a plot into a matrix of panels, namely facets, in which rows are specified by a value of $\theta\_w$ and columns are specified by a value of $\theta\_c$.
+In our simulation, $\theta\_w$ and $\theta\_c$ vary in a discrete scale from 0 to 1 with 0.25 increment.
+Note that we are only plotting facets with $\theta\_w\leq\theta\_c$ which is an assumption of our model.
+
+\ali{fix this according to new plot}
+To illustrate the changes in $\Rnum$ with respect to \percap testing intensity $\rho$, two sets of plots are presented. 
+\fref{pan} reflects the changes in $\Rnum$ when $\rho$ is small relative to the population size.
 Specifically, $\rho \in [0,0.01]$, and the test return rate $\omega\in [0.1,2]$.
@@
 In order to illustrate the non-monotonic changes in $\Rnum$ with respect to $\rho$,
-Fig. \ref{pan2} with maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,0.5]$, and the test return rate $\omega\in (0.5,2]$. 
-In both Fig.\,\ref{pan} and Fig.\,\ref{pan2}, the isolation efficacy parameters, $\theta\_w$ and $\theta\_c$ vary from 0 to 1.
+\fref{pan2} with maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,0.5]$, and the test return rate $\omega\in (0.5,2]$. 
 Note that the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, corresponding to the threshold $\Rnum=1$, is plotted in solid line in the panels.
 Thus, a combination of $\rho$ and $\omega$ above this critical contour results in reducing $\Rnum$.
-Also realize that $\Rnum=\frac{\beta}{\gamma}$, about 1.017 here, for the classical SIR model which can be acheived in the context of our model when $\rho=0$, no testing, or $\theta\_w=\theta\_w=0$, no isolation.
+Also realize that $\Rnum=\frac{\beta}{\gamma}$, about 1.017 here, for the classical SIR model which can be acheived in the context of our model when no testing and no isolation are implied, that is $\rho=0$ or $\theta\_w=\theta\_w=0$.
 
@@
 An example of this counter-vailing effect of $\phi$, and consequently $\rho$, on $\Rnum$ occures when $\theta\_w=0$ and $\theta\_c=1$.
-This is illustraited in the top-right panel of the Fig.\,\ref{pan2} panel (b), where the strength of isolation for awaiting people is the least, but the most for the confirmed cases.
+This is illustraited in the top-right panel of the \fref{pan2} panel (b), where the strength of isolation for awaiting people is the least, but the most for the confirmed cases.
 In this case $$a_3=\frac{w_I}{w_S}\,\omega\,\gamma\,p_I((\omega+\gamma)-\frac{w_I}{w_S}(\omega\,p_I+\gamma)).$$

### 2021-05-09 codes/sir_plot.R
Fig 3, high testing, modified as fig 2

@@ -56,17 +56,17 @@ df_targeted <- make_params_dat(params = update(params,W_S=W_S_targeted),
 df_random_h <- make_params_dat(params = params,
                              eta_ws=0,eta_we=1, ## so theta_w
                              eta_cs=0,eta_ce=1, ## so theta_c
-                             omega_s=0.25,omega_e=2,
+                             omega_s=1/inv_omega_fast_max,omega_e=1/inv_omega_fast_min,
                              rho_s=0,rho_e=0.25-0.000001)
 
 df_targeted_h <- make_params_dat(params = update(params,W_S=W_S_targeted),
                                eta_ws=0,eta_we=1, ## so theta_w
                                eta_cs=0,eta_ce=1, ## so theta_c
-                               omega_s=0.25,omega_e=2,
+                               omega_s=1/inv_omega_fast_max,omega_e=1/inv_omega_fast_min,
                                rho_s=0,rho_e=0.25-0.000001)
 
 ##important contour, ie R0=1 thus threshold=1 when plotting R0 contours, or Delta(R0=1)
-threshold <- 1-(params[["gamma"]]/params[["beta"]]) ## Or 0?  corresponding to R0=1
+threshold <- 1-(params[["gamma"]]/params[["beta"]]) ## corresponding to R0=1
 # #################################
 # Plotting Part
 # #################################
@@ -190,31 +190,33 @@ p12 <- (ggplot(df_temp2,aes(x=1/omega,y=rho,z=Delta))
        + xlab(TeX('$\\1/omega$, mean test return time (day)'))
        + ylab(TeX('$\\rho$, testing intensity (1/day per capita)'))
 )
-## There might be a solution for the tick label collision with expand_limits(), but that's hard to do across facets
-## see https://stackoverflow.com/questions/41575045/avoiding-axis-tick-label-collision-in-faceted-ggplots
+
 p1_temp2 <- (p12
             # + geom_contour_filled(breaks=brks)
             + geom_contour_filled(breaks=brks_vec)
-            + geom_contour(breaks=threshold,alpha=0.5,colour="black")
+            + geom_contour(breaks=threshold,alpha=0.5,colour="black",lty="11",lwd=1)
             + facet_grid(theta_w~theta_c, labeller=label_special)
-            + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
[…]

### 2021-05-10 SIR_manuscript.tex
minor edits in results

@@
 \end{equation}
-where the term $\frac{\beta}{\gamma}$ is the classical basic reproduction number for a SIR model without testing and isolation (see, e.g., \cite{keeling2011modeling}), and the positive quantity $\Delta$ is the ``reduction parameter'' due to testing and isolation processes defined as follows.
+where the term $\frac{\beta}{\gamma}$ is the classical basic reproduction number for a SIR model without testing and isolation (see, e.g., \cite{keeling2011modeling}), and the positive quantity $\Delta$ is the ``effectiveness of control'' effectiveness of control due to testing and isolation processes defined as follows.
 \begin{equation}
@@
 Further details of derivation of $\Rnum$ are provided in appendix, Sec.\,\ref{app:R0}. 
-The explicit formula of $\Rnum$, provides a mathematical ground to study the effect of parameters of interest, i.e., parameters that could be manipulated by public-health policy, on $\Rnum$.
-These parameters include the isolation efficacy, $\theta\_c$ and $\theta\_w$, \percap testing intensity, $\rho$, and the rate of test return, $\omega$.
+The explicit formula of $\Rnum$, enables us to study the effect of parameters of interest, on $\Rnum$ mathematically and through numerical simulation.
+These parameters, i.e., parameters that could be manipulated by public-health policy include the isolation efficacy, $\theta\_c$ and $\theta\_w$, \percap testing intensity, $\rho$, and the rate of test return, $\omega$.
 In partticular, we look at the partial derivative of $\Delta$ with respect to these parameters (see details in appendix, Sec.\,\ref{app:rho}).
@@
 
-\ali{fix this according to new plot}
 To illustrate the changes in $\Rnum$ with respect to \percap testing intensity $\rho$, two sets of plots are presented. 
 \fref{pan} reflects the changes in $\Rnum$ when $\rho$ is small relative to the population size.
-Specifically, $\rho \in [0,0.01]$, and the test return rate $\omega\in [0.1,2]$.
+Specifically, $\rho \in [0,0.013]$, and the test return rate $\omega\in [1/12,2]$.
 This is a more realistic scenario of testing as we have observed in \covid pandemic.
@@
 In order to illustrate the non-monotonic changes in $\Rnum$ with respect to $\rho$,
-\fref{pan2} with maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,0.5]$, and the test return rate $\omega\in (0.5,2]$. 
-Note that the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, corresponding to the threshold $\Rnum=1$, is plotted in solid line in the panels.
+\fref{pan2} with maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,1/5)$, and the test return rate $\omega\in [1/5,2]$. 
+Note that the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, corresponding to the threshold $\Rnum=1$, is plotted in dotted line in the panels.
 Thus, a combination of $\rho$ and $\omega$ above this critical contour results in reducing $\Rnum$.
@@
 \section{Results}
-
-The explicit formula for the basic reproduction number, $\Rnum$ \eqref{R0}, provides an opportunity to study the influence of changes in the underlying parameters of interest on the critical index of epidemic dynamics.
-These parameters include the isolation efficacy parameters, $\theta\_c$ and $\theta\_w$, \percap testing intensity, $\rho$, and the rate of test return, $\omega$.
-In this context, two testing strategies, namely random and targeted testing, were explored.
 
@@
 \caption{
-{\bf A comparison of the behaviour of the basic reproduction number, $\Rnum$, between random versus targeted testing strategies at different levels of testing and isolation.}
-We numerically evaluate $\Delta$ \eqref{R0}, reflecting the reduction of $\Rnum$ with respect to testing and isolation. We use the following parameters (see Table \ref{tab:params}):
-$N_0=1 \times 10^6$, $\omega \in [0.1,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.01]$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.339$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
+{\bf A comparison of the behaviour of the basic reproduction number, $\Rnum$, at low level of \percap testing intensity $\rho$, between random versus targeted testing strategies at different levels of testing and isolation.}
+We numerically evaluate $\Delta$ (Eq.~\ref{eq:del4}), reflecting the reduction of $\Rnum$ with respect to testing and isolation.
+We use the following parameters (see Table~\ref{tab:params} for a brief description):
+$N_0=1 \times 10^6$, $\omega \in [1/12,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.013]$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for the full effect of isolation on the transmission probability, $\beta=0.339$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
 Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
@@
 \caption{
-{\bf A comparison of the behaviour of the basic reproduction number, $\Rnum$, between random versus targeted testing strategies at different levels of testing and isolation.}
-We numerically evaluate $\Delta$ \eqref{R0}, reflecting the reduction of $\Rnum$ with respect to testing and isolation. We use the following parameters (listed in Table \ref{tab:params}):
-$N_0=1 \times 10^6$, $\omega \in (0.25,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.25]$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.339$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
+{\bf A comparison of the behaviour of the basic reproduction number, $\Rnum$, at high level of \percap testing intensity $\rho$, between random versus targeted testing strategies at different levels of testing and isolation.}
+We numerically evaluate $\Delta$ \eqref{eq:del4}, reflecting the reduction of $\Rnum$ with respect to testing and isolation.
+We use the following parameters (see Table~\ref{tab:params} for a brief description):
+$N_0=1 \times 10^6$, $\omega \in [1/5,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,1/5)$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.339$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
 Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
-The black solid line in each panel represents the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ corresponding to the threshold of $\Rnum=1$.
-\bmb{we still have some collisions in the y-axis tick labels}
-}
+The black dotted line in each panel represents the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ corresponding to the threshold of $\Rnum=1$.}
 \label{pan2}
@@
 
-Numerical results are shown in \fref{pan}.
-Targeted testing (panel b) is more effective at control than random testing (panel a).
-
+Numerical results are shown in Figs.~\ref{pan} and \ref{pan2}.
 When $\theta_w =0$ (top row of each panel), we see that shorter test-waiting times reduce \Rnum\ ($\Delta$ increases as we move to the left in each plot in this row).
@@
 The explanation here is that more rapid testing leaves more susceptibles in the ``waiting-for-no'' category at the DFE; these people must then wait for their tests to be returned before they can be tested again, receive a positive test, and become cautious.
-This effect is usually weak compared to the helpful effects of testing, but if testing is weighted, $\theta_w$ is low, and test returns are slow, t is possible for increasing $\rho$ to reduce $\Delta$ (see upper-right part of upper-right plot of \fref{pan2}(b)).
+This effect is usually weak compared to the helpful effects of testing, but if testing is weighted, $\theta_w$ is low, and test returns are slow, it is possible for increasing $\rho$ to reduce $\Delta$ (see upper-right part of upper-right plot of \fref{pan2}(b)).
 
+As expected, targeted testing strategy (panel b) is more effective at control than random testing (panel a). In particular
 %%%%%%%%%%%%

### 2021-05-10 SIR_manuscript.tex
scale change of y-axis in high rho, fig3, per 1000

@@
 \caption{
-{\bf A comparison of the behaviour of the basic reproduction number, $\Rnum$, at low level of \percap testing intensity $\rho$, between random versus targeted testing strategies at different levels of testing and isolation.}
-We numerically evaluate $\Delta$ (Eq.~\ref{eq:del4}), reflecting the reduction of $\Rnum$ with respect to testing and isolation.
+{\bf A comparison of effectiveness control on $\Rnum$ at low level of \percap testing intensity $\rho$.}
+We numerically evaluate the effectiveness control parameter, $\Delta$ (Eq.~\ref{eq:del4}), with respect to testing and isolation.
 We use the following parameters (see Table~\ref{tab:params} for a brief description):
@@
 \caption{
-{\bf A comparison of the behaviour of the basic reproduction number, $\Rnum$, at high level of \percap testing intensity $\rho$, between random versus targeted testing strategies at different levels of testing and isolation.}
-We numerically evaluate $\Delta$ \eqref{eq:del4}, reflecting the reduction of $\Rnum$ with respect to testing and isolation.
+{\bf A comparison of effectiveness control on $\Rnum$ at high level of \percap testing intensity $\rho$.}
+We numerically evaluate the effectiveness control parameter, $\Delta$ (Eq.~\ref{eq:del4}), with respect to testing and isolation.
 We use the following parameters (see Table~\ref{tab:params} for a brief description):
@@
 
-As expected, targeted testing strategy (panel b) is more effective at control than random testing (panel a). In particular
 %%%%%%%%%%%%

### 2021-05-10 codes/sir_plot.R

@@ -37,6 +37,7 @@ rho_brks <- c(0,0.005,0.01)
 inv_omega_fast_min <- 0.5
 inv_omega_fast_max <- 5
 inv_omega_fast_brks <- c(1,2,4)
+rho_fast_brks <- c(0,0.1,0.2)
 
 background <- "grey90"
 # make dataframe
@@ -188,7 +189,7 @@ df_temp2 <- df_random_h
 p12 <- (ggplot(df_temp2,aes(x=1/omega,y=rho,z=Delta))
        + theme_bw()
        + xlab(TeX('$\\1/omega$, mean test return time (day)'))
-       + ylab(TeX('$\\rho$, testing intensity (1/day per capita)'))
+       + ylab(TeX('$\\rho$, testing intensity (1/day per 1000)'))
 )
 
 p1_temp2 <- (p12
@@ -198,7 +199,9 @@ p1_temp2 <- (p12
             + facet_grid(theta_w~theta_c, labeller=label_special)
             + scale_x_continuous(expand=expansion(c(0,0)), breaks=inv_omega_fast_brks,limits=c(inv_omega_fast_min,inv_omega_fast_max))
             ## scale by 1000, no digits after decimal
-            + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
+            # + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
+            + scale_y_continuous(expand=expansion(c(0,0)), breaks=rho_fast_brks,
+                                 labels=scales::number_format(scale=1000,accuracy=1))
             + scale_fill_viridis_d(name=parse(text="Delta"),drop=FALSE,
                                    labels=hack_breaks)
             + geom_rect(data=df_temp, fill=ifelse(df_temp$theta_c < df_temp$theta_w,background,"NA" ),

### 2021-05-11 SIR_manuscript.tex
working on discussion, adding a result

@@
 Examining our formula for $\Rnum$ \eqref{R0} gives the following results. 
-See the appendix, Sec.\,\ref{app:rho} and Sec.\,\ref{app:omega} for details.
+See the appendix, Sec.\,\ref{app:rho}, Sec.\,\ref{app:omega} and Sec.\,\ref{app:w} for details.
 
@@
 
-\item \label{p1:eta} Increasing the isolation parameters for tested and confirmed individuals decreases \Rnum;
+\item \label{p1:eta} Increasing the isolation efficacy parameters for tested and confirmed individuals decreases \Rnum;
 
-\item \label{p1:rho} Higher testing $\rho$ intensity always decreases $\Rnum$ \textbf{if}: 
+\item \label{p1:rho} Higher testing intensity $\rho$ always decreases $\Rnum$ \textbf{if}: 
 %% \begin{enumerate} \item
@@
 \item \label{p1:omega} A higher rate of test return always decreases \Rnum, \textbf{if} $\theta_w=0$.
+
+\item \label{p1:w} Increasing testing “focus” $\frac{w_I}{w_S}$ from random to highly targeted always decreases \Rnum.
 
@@
 
-We also see that greater test intensity $\rho$ increases the effectiveness of control $\Delta$ ($\Delta$ increases as we move up in each plot in this Figure.
-Mathematically speaking, we did find a counter-vailing effect, but this can only be seen when we allow $\rho$ to take unrealistically large values, and only for weighted testing (see \fref{pan2}).
+We also see that greater test intensity $\rho$ increases the effectiveness of control $\Delta$ ($\Delta$ increases as we move up in each plot in this Figure).
+Mathematically speaking, we did find a counter-vailing effect, but this can only be seen when we allow $\rho$ to take unrealistically large values, and only for weighted/targeted testing (see \fref{pan2}).
 The explanation here is that more rapid testing leaves more susceptibles in the ``waiting-for-no'' category at the DFE; these people must then wait for their tests to be returned before they can be tested again, receive a positive test, and become cautious.
@@
 
+{\bf A potential advantage of slow test reporting $\omega$;}
+- It seems unapplicable to delay the test return time as a strategy to control the viral spread when $\theta\_w$ is high (moving to the right in each plot as $\theta\_w$ is closer to 1).
+-
+
+While targeted testing strategies, including targeting people with infection-like symptoms or the contacts of confirmed cases, are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is that viral spread is greater for a slow test than for a fast test.
+This counter-intuitive effect can occur when people are cautious when awaiting a test result, and may not be robust to second-order effects of fast testing (such as better contact tracing).
+
 {\bf The counter-vailing effect of \percap testing intensity, $\rho$, on $\Rnum$;}
-- we are missing out on community-level advantages of fast testing.
+- We are missing out on community-level advantages of fast testing.
 
-{\bf The potential advantage of slow test reporting, or favorable-delay-reporting;}
+{\bf Specifying the compartment-specific relative testing weights, Specifically $\frac{w_I}{w_S}$;}
+- More targeted testing is always good in viral control,
+- 
+We incorporated the compartment-specific relative testing weights, $w_S$, $w_I$ and $w_R$, to model random testing and \emph{targeted} testing strategies.
+Here, in the case of targeted testing and for the simplicity and illustration purposes, we assumed that infected and recovered individuals are tested at three times the \percap rate of susceptible individuals, thus $w_I/w_S=3$ and $w_R=w_I$.
+Note that we have not specified a methodology to assign particular relative testing weights corresponding to a particular targeted testing scenario.
+Modeling different targeted testing strategies, equivalently test-specific testing weights in our framework, requires prior information of the conditional probabilities of getting tested for people in a given compartment.
+This can be implied when we would like to quantify and compare the effect of different levels of test focus for infectious people on the basic reproduction number $\Rnum$, and conclude about the disease spread management.
+For example, when people are tested for "screening", the individuals with potential higher mobility, eg.
+people who are getting on flights, get more tested and thus the coresponding heavier testing weight is assigned than people awaiting for a surgery and are probably going to stay in a long-term-care facility and consequently less mobile and more isolated to begin with.
+With our model, we would be able to compare the sensitivity of the disease dynamics, through $\Rnum$, with respect to testing different high-risk groups in the population.
+This part needs to be developed further in future work.
 
@@
 If $a_3\geq 0$, $\partial\Delta/\partial\phi \geq 0$ for all feasible range of parameters, thus $\pro\Rnum \leq 0$.
+It is straight forward to show that $a_3\geq 0$ in case of random testing strategy, i.e., $w\_S=w\_I=1$.
 If $a_3 < 0$, then the quadratic expression in the numerator of \eqref{eq:dd3dr} has a positive root, $\phi^*$, such that for $\phi>\phi^*$, $\partial\Delta/\partial\phi < 0$.
@@
 
+% %%%%%%%
+\subsection{The effect of testing focus parameter $\frac{w_I}{w_S}$ on $\Rnum$} \label{app:w}
+
+We define $w\_{IS}=\frac{w_I}{w_S}$.
+\begin{align}
+\pder[\Delta]{w\_{IS}}= \dfrac{(\omega-\rho)\,(\omega(\omega-\rho\theta\_w)+\gamma(\omega-\rho))\,(\theta\_w\gamma+\theta\_c\omega\,p\_I)}
+{(-\omega^2\,\gamma+\omega\,\gamma\,\rho-\gamma\,\rho\,\omega\,w\_{IS}-
+\omega\,\gamma^2+\gamma^2\,\rho-\omega^2\,p_I\,\rho\,w\_{IS})^2},
+\end{align}
+which is a positive quantity. Thus, $\pder[\Rnum]{w\_{IS}}\leq 0$. Therefore, increasing the focus of testing on the infectious people will result in less transmission.  
 \end{document}

### 2021-05-12 SIR_manuscript.tex
discussion points

@@
 
-\item \label{p1:w} Increasing testing “focus” $\frac{w_I}{w_S}$ from random to highly targeted always decreases \Rnum.
+\item \label{p1:w} Increasing testing “focus” $\frac{w_I}{w_S}$ from random $w_S=w_I$ to targeted $w_S<w_I$ always decreases \Rnum.
 
@@
 
+\newpage
 \begin{figure}[h!]
@@
 \section{Discussion}
+\begin{enumerate}
 
-{\bf A potential advantage of slow test reporting $\omega$;}
-- It seems unapplicable to delay the test return time as a strategy to control the viral spread when $\theta\_w$ is high (moving to the right in each plot as $\theta\_w$ is closer to 1).
--
+\item {\bf A potential advantage of slow test reporting $\omega$, result~\ref{p1:omega};}
 
-While targeted testing strategies, including targeting people with infection-like symptoms or the contacts of confirmed cases, are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is that viral spread is greater for a slow test than for a fast test.
+- Whether increasing $\omega$ lowers $\Rnum$ depends on the precise combination of model parameters  including $\rho$, testing strategies represented by compartment-specific testing weights, test sensitivity and specificity, and the level of isolation efficacy.
+- The potential advantage of slow test reporting is in the individual level as opposed to the comunity level, but seems unapplicable as a strategy.
+- May be this mechanism combined with increasing testing intensity $\rho$ can be considered alongside other prevention methods, eg., social distancing etc, when the transmission is in the exponential-growth phase (in order to increase the generation interval).
+- One reason in favor of prolonging $1/\omega$ is that it is easy to manipulate while increasing $\rho$ can be limited to the resources.
+
+\item {\bf The counter-vailing effect of \percap testing intensity, $\rho$, on $\Rnum$;}
+- The explanation here is that more rapid testing leaves more susceptibles in the ``waiting-for-no'' $S\_n$ category at the DFE and delay them to get to $I\_c$ where they are subject to a full isolation, thus no transmission.
+- We are missing out on community-level advantages of fast testing.
+- What is the explanation for having this counter-vailing effect of $\rho$ in targeted testing and not in the random one?
+- May because in the targeted testing more $I\_u$s are tested commpared to $S\_u$s, given 
+ that higher $\omega$ leaves more people in $S\_n$ and $I\_n$ which results in more transisssion.
+- While targeted testing strategies, including targeting people with infection-like symptoms or the contacts of confirmed cases, are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is that viral spread is greater for a slow test than for a fast test.
 This counter-intuitive effect can occur when people are cautious when awaiting a test result, and may not be robust to second-order effects of fast testing (such as better contact tracing).
 
-{\bf The counter-vailing effect of \percap testing intensity, $\rho$, on $\Rnum$;}
-- We are missing out on community-level advantages of fast testing.
+\item {\bf Specifying the compartment-specific relative testing weights, Specifically $\frac{w_I}{w_S}$;}
 
-{\bf Specifying the compartment-specific relative testing weights, Specifically $\frac{w_I}{w_S}$;}
 - More targeted testing is always good in viral control,
-- 
-We incorporated the compartment-specific relative testing weights, $w_S$, $w_I$ and $w_R$, to model random testing and \emph{targeted} testing strategies.
+- We incorporated the compartment-specific relative testing weights, $w_S$, $w_I$ and $w_R$, to model random testing and \emph{targeted} testing strategies.
 Here, in the case of targeted testing and for the simplicity and illustration purposes, we assumed that infected and recovered individuals are tested at three times the \percap rate of susceptible individuals, thus $w_I/w_S=3$ and $w_R=w_I$.
@@
 This part needs to be developed further in future work.
+
+\item What can we say in the context of strength and speed paradigm?
+
+\end{enumerate}
 

### 2021-05-13 SIR_manuscript.tex
a_3 in appendix simplified

@@
 where
-\begin{multline}
-\label{eq:a3}
-a_3 = \frac{w_I}{w_S}
-\Bigg(\Big(
-( \theta_{{w}}{{\it p\_I}}^{2}-\theta_{{c}}{{\it p\_I}}^
-{2}\theta_{{w}} ) {\omega}^{3} 
-+ ( 2\,\theta_{{w}}{{\it p\_I
-}}^{2}-\theta_{{c}}{{\it p\_I}}^{2}-{\it p\_I}\,{\theta_{{w}}}^{2}-
-\theta_{{c}}{\it p\_I}\,\theta_{{w}}-{{\it p\_I}}^{2}{\theta_{{w}}}^{2
-}+2\,\theta_{{w}}{\it p\_I} ) \gamma\,{\omega}^{2} \\
- + ( -2\,{ \it p\_I}\,{\theta_{{w}}}^{2}-\theta_{{c}}{\it p\_I}+3\,\theta_{{w}}{
-\it p\_I}-{\theta_{{w}}}^{2}+\theta_{{w}} ) {\gamma}^{2}\omega 
- + ( \theta_{{w}}-{\theta_{{w}}}^{2} ) {\gamma}^{3} \Big) \frac{w_I}{w_S}
- + ( -\theta_{{c}}{\it p\_I}\,\theta_{{w}}+\theta_{{c}}{\it p\_I}
- ) \gamma\,{\omega}^{2}+ \\
- ( \theta_{{w}}+\theta_{{c}}{\it
-p\_I}-{\theta_{{w}}}^{2}-\theta_{{c}}{\it p\_I}\,\theta_{{w}} )
-{\gamma}^{2}\omega+ ( \theta_{{w}}-{\theta_{{w}}}^{2} ) {
-\gamma}^{3}
-\Bigg),
-\end{multline}
 \begin{align}
 \label{eq:abcd2}
-b_3&= 2\frac{w_I}{w_S}(\omega+\gamma)\gamma\Big( 
+a_3& = \frac{w_I}{w_S} \Bigg(&& (1-\theta\_w)\,(1+\frac{w_I}{w_S})\theta\_w\,\gamma^3
++(1-\theta\_c)\,p_I^2\,\theta\_w\,\frac{w_I}{w_S}\,\omega^3 \\ \notag
+& ~ &&+\Big( \big( (1-\theta\_w-\frac{w_I}{w_S})\,\theta\_c +(3-2\theta\_w)\,\theta\_w\frac{w_I}{w_S} \big)\,p_I
++(1-\theta\_w)\,(1+\frac{w_I}{w_S})\theta\_w  \Big)\,\omega\,\gamma^2 \\ \notag
+& ~ &&+\Big(
+\big((1-\theta\_w-\theta\_w\,\frac{w_I}{w_S})\,\theta\_c+ (2-\theta\_w)\,\theta\_w\,\frac{w_I}{w_S} \big)\,p_I
++(2\,\theta\_w-\theta\_w^2-\theta\_c)\,\frac{w_I}{w_S}\,p_I^2
+\Big)\,\omega^2\,\gamma \Bigg), \\
+b_3& = && \hspace{-1cm} 2\frac{w_I}{w_S}\,(\omega+\gamma)\gamma\Big(
 (\omega+\gamma+\omega\,p_I)(2-\theta\_w)\,\gamma\,\theta\_w+ (1-\theta\_w)\,\omega^2 p_I\theta\_c+\omega^2 p_I\theta\_w
 \Big) ,\\
-c_3&= (\omega+\gamma)^2\,\gamma\Big( 
+c_3& = && \hspace{-1cm} (\omega+\gamma)^2\,\gamma\Big(
 (2-\theta\_w)\,\gamma^2\,\theta\_w+
@@
 \frac{w_I}{w_S} \omega^2 p_I\theta\_c \Big),\\
-d_3&= \frac{(\omega+\gamma)}{\omega} \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi\,+\,(\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2. 
+d_3&  = && \hspace{-1cm} \frac{(\omega+\gamma)}{\omega}\, \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi\,+\,(\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2.
 \end{align}
@@
 This is illustraited in the top-right panel of the \fref{pan2} panel (b), where the strength of isolation for awaiting people is the least, but the most for the confirmed cases.
-In this case $$a_3=\frac{w_I}{w_S}\,\omega\,\gamma\,p_I((\omega+\gamma)-\frac{w_I}{w_S}(\omega\,p_I+\gamma)).$$
-Specifically, in the case of targeted testing which is identified with $\frac{w_I}{w_S}> 1$, and using a perfect sensative test, thus $p_I=1$, there exists a range for $\rho$ over which $\pder[\Rnum]{\rho}\leq\,0$.
+In this case, simplifying $a_3$ in Eq.~\eqref{eq:abcd2} gives $$a_3=\frac{w_I}{w_S}\,\omega\,\gamma\,p_I((\omega+\gamma)-\frac{w_I}{w_S}(\omega\,p_I+\gamma)).$$
+Specifically, in the case of targeted testing which is identified by $\frac{w_I}{w_S}> 1$, and using a perfect sensative test, thus $p_I=1$, there exists a range for $\rho$ over which $\pder[\Rnum]{\rho}\leq\,0$.
 Note that $\rho$ and $\omega$ have a similar mechanism in delaying people to get into $I\_c$, thus we would expect to see the non-trivial counter-vailing effect of these two parameters on $\Rnum$.

### 2021-05-13 SIR_manuscript.tex
Old Fady's stuff are commented out for now

@@
 around $\rho=0$.
-\ali{I stoped here!}
-
-Perhaps counter-intuitively, the equation above does not predict that $\Rnum$ is monotone decreasing with respect to $\omega$.
-In other words; our model does not predict that returning test results more rapidly \textit{always} lower $\Rnum$.
-In order to gain insight into this intriguing behavior, we examine the zeroes of $\frac{\partial{\Rnum}}{\partial{\omega}}(\omega)$.
-Defining the following quantity, $Q$, will help us write the roots of $\partial{\Rnum}/\partial{\omega}$ neatly as follows.
-\begin{align}\label{eq:defQ}
-    Q =& \frac{w_I}{w_S}\left(1-\frac{n_t-1}{n_w-1}p_I \right).
-\end{align}
-With that in mind, we can write the roots of $\partial{\Rnum}/\partial{\omega}$ as
-\begin{align}
-    \omega_1 =& \frac{\gamma}{-\sqrt{1-Q}-1} \\
-    \omega_2 =& \frac{\gamma}{\sqrt{1-Q}-1}.
-\end{align}
-
-Note that the zeroes are real if and only if $Q < 1$.
-Note that have $\theta\_c > \theta\_w$, so if $p_I \approx 1$, we will have $Q < 0 < 1$.
-Thus, if we assume near-perfect test sensitivity, $\omega_1$ and $\omega_2$ will be real.
-
-Assuming $\omega_1, \omega_2$ are real, it is easy to confirm that $\omega_1 < 0$ by looking at the denominator.
-To see that $\omega_2 > 0$, recall that $Q < 0$, so $\sqrt{1-Q} > 1$ and so $\sqrt{1-Q} -1 > 0$.
-Knowing that $\omega_1 < 0$, the only root of interest (i.e., biologically relevant quantity) is $\omega_2$.
-
-We can prove that $\partial{\Rnum}/\partial{\omega} > 0$ when $\omega \in (0,\omega_2)$ and $\partial{\Rnum}/\partial{\omega} < 0$ when $\omega \in (\omega_2,\infty)$ by computing the limits of $\partial{\Rnum}/\partial{\omega}$ at $0$ and $\infty$ respectively.
-So it follows that $\Rnum$ has a global maximum with respect to $\omega$ at $\omega = \omega_2$.
-
-Now we want to characterize the parameter regions on which $\partial{\Rnum}/\partial{\omega} < 0$ (i.e., the conditions under which returning test results more rapidly is favorable).
-By the previous analysis, this is equivalent to solving for $\omega > \omega_2$.
-So
-\begin{align}
-    &\omega > \omega_2 \nonumber \\
-    &\omega > \frac{\gamma}{\sqrt{1-Q}-1} \nonumber \\
-    &\sqrt{1-Q} > \frac{\gamma}{\omega}+1 \\
-    &1-Q > (\frac{\gamma}{\omega}+1)^2.
-\end{align}
-Substituting in $Q$ from \eqref{eq:defQ} we have
-\begin{align}
-    &1-\frac{w_I}{w_S}\left(1-\frac{n_{t}-1}{n_{\omega}-1}P_{i}\right)> \left(\frac{\gamma}{\omega}+1\right)^2 \\
-    &-\frac{w_I}{w_S}\left(1-\frac{n_{t}-1}{n_{\omega}-1}P_{i}\right)> \left(\frac{\gamma}{\omega}+1\right)^2+1 \\    
-    &\frac{w_I}{w_S}\left(\frac{1-n_{t}}{1-n_{\omega}}P_{i}-1\right)> \left(\frac{\gamma}{\omega}+1\right)^2+1\label{eq:necsuf}.
-\end{align}
-Since all steps in deriving \eqref{eq:necsuf} are reversible, \eqref{eq:necsuf} gives a necessary and sufficient condition for $\omega > \omega_2$, which characterizes when returning tests more rapidly would cause a decrease in $\Rnum$.
+% \ali{I stoped here!}
 
@@
 % %%%%%%%
-\subsection{Expensive vs.\ cheap tests}
-
-The use of tests cheaper than RT-PCR has been proposed as a potential strategy for containing the \covid pandemic.
-While cheaper tests may be less sensitive and reliable than RT-PCR, they allow for broader and more intense testing.
-In the analysis below, we compare the $\Rnum$ predicted by our model depending on the testing strategy.
-
-Consider a test that allows us to test at rate $\rho_1$ and has sensitivity $P_{i,1}$, and another test that allows us to test at $\rho_2$ and has sensitivity $P_{i,2}$.
-Suppose that $\rho_1 > \rho_2$.
-Recall that the linearization of $\Rnum$ around $\rho \approx 0$ is given by $$\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(-\theta\_w)(\gamma w_S+\omega w_I) + (-\theta\_c)p_I w_I \omega^2 \Big).$$
-
-Treating $\Rnum$ as a function of $\rho$ and $P_i$,we can reduce the inequality $$\Rnum(\rho_2, p_{I,2}) < \Rnum(\rho_1, p_{I,1})$$ into
-
-\begin{align}\label{eq:rho1vsrho2}
-    &\rho_1\left(\gamma(-\theta\_w)(\gamma w_S + \omega w_I) + (-\theta\_c)p_{I,1}w_I\omega^2\right) - \rho_2\left(\gamma(-\theta\_w)(\gamma w_S + \omega w_I) + (-\theta\_c)p_{I, 2}w_I\omega^2\right) > 0 \nonumber \\
-    &\vdots \nonumber \\
-    &\frac{\rho_2P_{i, 2}-\rho_1P_{i, 1} }{\rho_1-\rho_2} > \frac{\theta\_w}{\theta\_c}\cdot \frac{\gamma(\gamma w_S + \omega w_I)}{\omega^2 w_I}
-\end{align}
-
-Note that the RHS is positive, thus a necessary condition for the inequality above to hold is that $\rho_2P_{i,2} > \rho_1P_{i,1}$, equivalently
-
-\begin{equation}
-\frac{P_{i,2}}{P_{i,1}} > \frac{\rho_1}{\rho_2}.
-\end{equation}
-
-To state an example of this, if test $A$ is three times as expensive as test $B$ (and hence one can test three times as many people with test $B$), using test $A$ rather than $B$ will be favorable only if test $A$ is at least 3 times more sensitive than test $B$.
-Note that this is a necessary but not sufficient condition, so even if test $A$ is three times more sensitive, it is still possible for test $B$ to be more effective.
-
-\cref{eq:rho1vsrho2} tells us precisely when a test corresponding to $\rho_2, P_{i,2}$ will yield a lower $\Rnum$ than a test corresponding to $\rho_1, P_{i,1}$, where $\rho_1 > \rho_2$.
-Some of the qualitative \textit{trends} that favor test 2 (the higher-sensitivity test) include
-
-\begin{itemize}
-    \item individuals who test positive self-isolate much more than individuals who are waiting for their test result.
-    \item the time it takes to return tests is much shorter than the mean infectious period.
-    \item the testing intensity is much greater for infected individuals than susceptible individuals.
-\end{itemize}
-
-% %%%%%%%
 \subsection{The effect of testing focus parameter $\frac{w_I}{w_S}$ on $\Rnum$} \label{app:w}
@@
 which is a positive quantity. Thus, $\pder[\Rnum]{w\_{IS}}\leq 0$. Therefore, increasing the focus of testing on the infectious people will result in less transmission.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \todo{Fady:To be change accordingly please.} 
+% Perhaps counter-intuitively, the equation above does not predict that $\Rnum$ is monotone decreasing with respect to $\omega$.
+In other words; our model does not predict that returning test results more rapidly \textit{always} lower $\Rnum$.
+In order to gain insight into this intriguing behavior, we examine the zeroes of $\frac{\partial{\Rnum}}{\partial{\omega}}(\omega)$.
+% Defining the following quantity, $Q$, will help us write the roots of $\partial{\Rnum}/\partial{\omega}$ neatly as follows.
+% \begin{align}\label{eq:defQ}
+%     Q =& \frac{w_I}{w_S}\left(1-\frac{n_t-1}{n_w-1}p_I \right).
+% \end{align}
+% With that in mind, we can write the roots of $\partial{\Rnum}/\partial{\omega}$ as
+% \begin{align}
+%     \omega_1 =& \frac{\gamma}{-\sqrt{1-Q}-1} \\
+%     \omega_2 =& \frac{\gamma}{\sqrt{1-Q}-1}.
+% \end{align}
+% 
+% Note that the zeroes are real if and only if $Q < 1$.
+Note that have $\theta\_c > \theta\_w$, so if $p_I \approx 1$, we will have $Q < 0 < 1$.
+Thus, if we assume near-perfect test sensitivity, $\omega_1$ and $\omega_2$ will be real.
+% 
+% Assuming $\omega_1, \omega_2$ are real, it is easy to confirm that $\omega_1 < 0$ by looking at the denominator.
+To see that $\omega_2 > 0$, recall that $Q < 0$, so $\sqrt{1-Q} > 1$ and so $\sqrt{1-Q} -1 > 0$.
+Knowing that $\omega_1 < 0$, the only root of interest (i.e., biologically relevant quantity) is $\omega_2$.
+% 
+% We can prove that $\partial{\Rnum}/\partial{\omega} > 0$ when $\omega \in (0,\omega_2)$ and $\partial{\Rnum}/\partial{\omega} < 0$ when $\omega \in (\omega_2,\infty)$ by computing the limits of $\partial{\Rnum}/\partial{\omega}$ at $0$ and $\infty$ respectively.
+So it follows that $\Rnum$ has a global maximum with respect to $\omega$ at $\omega = \omega_2$.
+% 
+% Now we want to characterize the parameter regions on which $\partial{\Rnum}/\partial{\omega} < 0$ (i.e., the conditions under which returning test results more rapidly is favorable).
+By the previous analysis, this is equivalent to solving for $\omega > \omega_2$.
+So
+% \begin{align}
+%     &\omega > \omega_2 \nonumber \\
+%     &\omega > \frac{\gamma}{\sqrt{1-Q}-1} \nonumber \\
+%     &\sqrt{1-Q} > \frac{\gamma}{\omega}+1 \\
+%     &1-Q > (\frac{\gamma}{\omega}+1)^2.
+% \end{align}
+% Substituting in $Q$ from \eqref{eq:defQ} we have
+% \begin{align}
+%     &1-\frac{w_I}{w_S}\left(1-\frac{n_{t}-1}{n_{\omega}-1}P_{i}\right)> \left(\frac{\gamma}{\omega}+1\right)^2 \\
+%     &-\frac{w_I}{w_S}\left(1-\frac{n_{t}-1}{n_{\omega}-1}P_{i}\right)> \left(\frac{\gamma}{\omega}+1\right)^2+1 \\    
+%     &\frac{w_I}{w_S}\left(\frac{1-n_{t}}{1-n_{\omega}}P_{i}-1\right)> \left(\frac{\gamma}{\omega}+1\right)^2+1\label{eq:necsuf}.
+% \end{align}
+% Since all steps in deriving \eqref{eq:necsuf} are reversible, \eqref{eq:necsuf} gives a necessary and sufficient condition for $\omega > \omega_2$, which characterizes when returning tests more rapidly would cause a decrease in $\Rnum$.
+% % %%%%%%%
+% \subsection{Expensive vs.\ cheap tests}
+% 
+% The use of tests cheaper than RT-PCR has been proposed as a potential strategy for containing the \covid pandemic.
+While cheaper tests may be less sensitive and reliable than RT-PCR, they allow for broader and more intense testing.
+In the analysis below, we compare the $\Rnum$ predicted by our model depending on the testing strategy.
+% 
+% Consider a test that allows us to test at rate $\rho_1$ and has sensitivity $P_{i,1}$, and another test that allows us to test at $\rho_2$ and has sensitivity $P_{i,2}$.
+Suppose that $\rho_1 > \rho_2$.
+Recall that the linearization of $\Rnum$ around $\rho \approx 0$ is given by $$\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(-\theta\_w)(\gamma w_S+\omega w_I) + (-\theta\_c)p_I w_I \omega^2 \Big).$$
+% 
+% 
+% Treating $\Rnum$ as a function of $\rho$ and $P_i$,we can reduce the inequality $$\Rnum(\rho_2, p_{I,2}) < \Rnum(\rho_1, p_{I,1})$$ into 
+% 
+% \begin{align}\label{eq:rho1vsrho2}
+%     &\rho_1\left(\gamma(-\theta\_w)(\gamma w_S + \omega w_I) + (-\theta\_c)p_{I,1}w_I\omega^2\right) - \rho_2\left(\gamma(-\theta\_w)(\gamma w_S + \omega w_I) + (-\theta\_c)p_{I, 2}w_I\omega^2\right) > 0 \nonumber \\
+%     &\vdots \nonumber \\
+%     &\frac{\rho_2P_{i, 2}-\rho_1P_{i, 1} }{\rho_1-\rho_2} > \frac{\theta\_w}{\theta\_c}\cdot \frac{\gamma(\gamma w_S + \omega w_I)}{\omega^2 w_I}
+% \end{align}
+% 
+% Note that the RHS is positive, thus a necessary condition for the inequality above to hold is that $\rho_2P_{i,2} > \rho_1P_{i,1}$, equivalently 
+% 
+% \begin{equation}
+% \frac{P_{i,2}}{P_{i,1}} > \frac{\rho_1}{\rho_2}.
+% \end{equation}
+% 
+% To state an example of this, if test $A$ is three times as expensive as test $B$ (and hence one can test three times as many people with test $B$), using test $A$ rather than $B$ will be favorable only if test $A$ is at least 3 times more sensitive than test $B$.
+Note that this is a necessary but not sufficient condition, so even if test $A$ is three times more sensitive, it is still possible for test $B$ to be more effective.
+% 
+% \cref{eq:rho1vsrho2} tells us precisely when a test corresponding to $\rho_2, P_{i,2}$ will yield a lower $\Rnum$ than a test corresponding to $\rho_1, P_{i,1}$, where $\rho_1 > \rho_2$.
+Some of the qualitative \textit{trends} that favor test 2 (the higher-sensitivity test) include
+% 
+% \begin{itemize}
+%     \item individuals who test positive self-isolate much more than individuals who are waiting for their test result.
+%     \item the time it takes to return tests is much shorter than the mean infectious period.
+%     \item the testing intensity is much greater for infected individuals than susceptible individuals.
+% \end{itemize}
+
 \end{document}

### 2021-05-13 SIR_manuscript.tex
minor result was added

@@
 
-\item \label{p1:w} Increasing testing “focus” $\frac{w_I}{w_S}$ from random $w_S=w_I$ to targeted $w_S<w_I$ always decreases \Rnum.
+\item \label{p1:w} Increasing testing “focus” $\frac{w_I}{w_S}$ from random, $w_S=w_I$, to targeted, $w_S<w_I$, always decreases \Rnum.
 
@@
 This effect is usually weak compared to the helpful effects of testing, but if testing is weighted, $\theta_w$ is low, and test returns are slow, it is possible for increasing $\rho$ to reduce $\Delta$ (see upper-right part of upper-right plot of \fref{pan2}(b)).
+
+Lastly, we show that increasing the focus of testing from random to targeted, always increases $\Delta$, thus decreases \Rnum.
+This can be seen in \fref{pan} and \fref{pan2}.
+In particular, we realize higher $\Delta$ levels, i.e., darker colors in corresponding plots\/facets as we go from panel~(a) to panel~(b) in the figures.
 

### 2021-05-14 note/cohort.tex
starting cohort-based framework

--- 2021-05-13 SIR_manuscript.tex
@@
 
-% https://tex.stackexchange.com/questions/12703/how-to-create-fixed-width-table-columns-with-text-raggedright-centered-raggedlef
-\usepackage{array}
-\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcommand{\todo}[1]{\comment{red}{TODO}{#1}}
 
@@
 \newcommand{\percap}{\emph{per capita}\xspace}
-\newcommand{\Rnum}{\ensuremath{\mathcal{R}_0}}
+\newcommand{\Rnot}{\ensuremath{\mathcal{R}_0}}
 \newcommand{\covid}{COVID-19\xspace}
@@
 
-\newcommand*\subtxt[1]{_{\textnormal{#1}}}
-\DeclareRobustCommand\_{\ifmmode\expandafter\subtxt\else\textunderscore\fi}
-
 \newcommand{\comment}{\showcomment}
@@
 \newcommand{\nocomment}[3]{}
-
-\newcommand{\fady}[1]{\comment{cyan}{Fady}{#1}}
-\newcommand{\ali}[1]{\comment{magenta}{Ali}{#1}}
-\newcommand{\jd}[1]{\comment{blue}{JD}{#1}}
-\newcommand{\david}[1]{\comment{red}{DJDE}{#1}}
-\newcommand{\bmb}[1]{\comment{red}{BMB}{#1}}
-\newcommand{\todo}[1]{\comment{red}{TODO}{#1}}
 
@@
 
-\title{Testing and Isolation Efficacy: Insights from a Simple Epidemic Model}
+\title{The Paradigm of Cohort Equations for the SIR  Model With Testing and Isolation}
 
@@
 \maketitle
-
 \linenumbers
 
-% %%%%%%%
-\section{Abstract}
-
-The effect of testing processes, including testing and test reporting, on an epidemic dynamics, involving infection and recovery, can be studied at the individual level or the community level (e.g., nursing homes, long-term-care facilities, etc.).
-Gaining insights to determine the sensitivity of the epidemic dynamics with respect to the testing processes will depend on underlying factors including the level of focus (individual or community), assumptions (model), and the interplay between these factors.
-In particular, the fast testing and test reporting may be beneficial at the community-level, supported by many studies, as it gives a rapid assessment of the situation, identifies hot spots, and may enable rapid contact-tracing.
-However, the potential advantage of a slow rate of test return on the dynamics of an epidemic is real, often neglected, and needs to be quantified.
-At the individual level, this advantage can manifest in the following sense: individuals awaiting their test results or who have tested positive may partially or fully self-isolate, thus reducing or eliminating their potential in the transmission process.
-In this paper, we investigated this individual-level effect of testing processes on the epidemic dynamics by developing a SIR-type model.
-Although the model development was motivated by the \covid epidemic, the model has general epidemiological and testing structures.
-The realistic components of the model include \percap testing intensity, test sensitivity and specificity, rate of test return, and isolation efficacy in reduction of the probability of transmission.
-The novel component is the compartment-specific relative testing weights, which reflect the testing strategies-- surveillance, diagnosis, or control.
-Here, we compare two testing strategies, random vs.
-targeted, and concluded that the targeted testing strategy is more effective in the sense that achieving reduction of $\Rnum$ can occure in a lower range of testing intensity and longer test return time relative to the random testing.
-Furthermore, we show that increasing \percap testing intensity and reducing the test return time would be beneficial on the dynamics of an epidemic in general but there are exception cases.
-In particular, it is possible for the basic reproduction number, $\Rnum$, to be increasing with respect to the \percap testing intensity and the rate of test return.
-
-% %%%%%%%
 \section{Introduction}
 
-The observed dynamics of the \covid epidemic are driven by both epidemiological processes (infection and recovery) and testing processes (testing and test reporting).
-In addition to shaping epidemic observations (via case reports), testing processes can also affect epidemiological dynamics.
-In particular, individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may do so also (possibly to a lesser extent).
-We developed a mechanistic model that incorporates epidemic processes and testing in order to explore the effects of testing and isolation on epidemic dynamics.
+\section{Ideas}
+This is follwoing the work of \cite{champredon2018equivalence}.
 
-If testing influences behavior, then epidemic dynamics will depend on patterns of who gets tested.
-The impacts of testing will depend on intensity (tests performed per day), and on how strongly testing is focused on people who are infectious.
-This level of focus depends in turn on the purpose and design of testing programs. 
-When testing is done for the purposes of disease surveillance \citep{foddai2020base}
-tests should be assigned randomly across the population, possibly with a stratified design for statistical efficiency \citep{graubard1996modelling} \ali{a better ref everyone?}.
+In Watmough 2002's work matrix form of the I compartment $I'=(-V)I$, where I is n-by-1 and $V$ is n-by-1.
 
-Over the course of the \covid pandemic, however, the vast majority of testing has been done with other goals --
-primarily diagnosis (determining the infection status for clinical purposes), or control (determining the infection status in order to quickly isolate cases that have been found by contact tracing), which we characterize as \emph{targeted} testing strategies.
-In these cases, testing probabilities vary widely across epidemiological compartments; in our dynamical model, we will characterize these probabilities by assigning a \percap testing weight to each compartment that determines the \emph{relative} probability that an individual in that compartment will be selected for testing (see Methods).
+In the context of our SIR model with testing, $I^T=[I_u,I_n,I_p,I_c]$ where $T$ is the transposed operator.
+And the solution is given by $I(\tau) = \exp(-V\tau) I(0)$, where $\tau$ is in the infection-time scale.
+$K(\tau)=\Rnot g(\tau)= F I(\tau)$,
+where $F$ is n-by-n matrix of new infections, $I(0)=[1,0,0,0]$.
 
-When testing is used primarily for diagnosis it will focus on people with infection-like symptoms; thus the relative testing weights for infected people will depend on the relative probability of infected people having symptoms.
-For \covid infection, the testing weights will depend on the relative asymptomatic infections, time spent pre-sympomatic vs.\ symptomatic infections -- and also the incidence of \covid-like symptoms among people in the population \emph{not} infected with \covid.
-Testing for epidemic control will focus on people who are known to have been in contact with known infected cases; in this case the testing weights for infected vs.\ uninfected people will depend on the probability of infection given contact, as well as the thoroughness and effectiveness of the system for identifying suspicious contacts.
+$\exp(-V\tau)$ is the probability of being infected and stay infectious at time $\tau$.
 
-The main interest from the epidemiological point of view is to know whether the number of infected individuals goes through an exponential growth phase, following the introduction of an infection in a totally susciptable population, before the disease becomes extinct.
-This is determined by studying the basic reproduction number $\Rnum$.
-It is defined as the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}.
-In the early stages of an epidemic the number of infected individuals is expected to grow exponentially over time when $\Rnum>1$, and to decline over time when $\Rnum<1$. 
-Although the value of $\Rnum$ cannot completely characterize the dynamics of even the simplest epidemic model
-\citep{shaw2021what}, it does give a simple and widely accepted index for the difficulty of control, as well as some indication of the likely final size of an epidemic \citep{ma2006generality}.
+$\exp(-V\tau)=\sum_{i=0}^\infty \tau^n (-V)^n/n!$.
 
-In order to understand the effect of testing processes on an epidemic dynamics, we deveopled a mechanistic SIR-type model with epidemic and testing components.
-This model provides a sensible platform to link the modeling of epidemic and testing components and study their interaction.
-Here, we studied the the effect of testing intensity, rate of test return and the isolation efficacy in reduction of the probability of transmission on the epidemic dynamics when different levels of testing ``focus'' (from random to highly targeted) are in place.
-Our model provides insights to the sensitivity of the epidemic dynamics, through $\Rnum$, with respect to the undelying testing and isolation parameters.\ali{edited in response to David comments.}
-
-% %%%%%%%
-\section{Methods}
-
-We developed a deterministic model, Eqs.\,\eqref{model}, which groups individuals based on disease status and testing status.
-Disease states include Susceptible, Infectious and Recovered (thus this is an SIR-type model), and testing status categorizes people as \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive} (\fref{fig:flowchart}).
-Symbolically, the testing status of an individual in the disease compartment $X$, where $X \in \{S,I,R\}$, is reflected in the subscript, namely $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
-Note that the top-to-bottom order of the testing-based compartments of each disease-based compartmenet $X$ in the flowchart (\fref{fig:flowchart}) and the model equations \eqref{model} should match.
-However, we switched  $X\_u$ and $X\_n$ in the flowchart (\fref{fig:flowchart}) for the sake of tidiness of the flowchart.
-Further, two `accumulator' compartments, $N$ and $P$, were also incorporated in the model in order to collect cumulative reported negative or positive tests.
-The model equations and details of calculation of the basic reproduction number $\Rnum$ are presented in the appendix (see Sec.
-\ref{app:R0}).
-
-\begin{figure}
-\begin{center} 
-\includegraphics[scale=1.5]{pix/sir_comp.pdf}
-\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model, \ref{model}.
-Here, the disease-based status of a compartment $X$, where $X \in \{S,I,R\}$, is combined with the testing-based status including $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
-Also,  $\Lambda$ is the force of infection with definition in Eq.~\eqref{Lambda}, $\gamma$ is the recovery rate, $\omega$ is the rate of test return, $F_X$ (defined in Eq.
-\eqref{F}) and $p_X$ represent the \percap testing rate and the probability of positive tests, respectively, for compartment $X$.
-For further description of the parameters see \cref{tab:params}.
-\label{fig:flowchart}}
-\end{center} 
-\end{figure}
-
-Table~\ref{tab:params} defines the model parameters, which are generally straightforward
-\percap flows between compartments, or modifiers to these flow rates.
-The novel component of the model comes in through the compartment-specific relative testing weights $w_S$, $w_I$ and $w_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively.
-Thus, we can spesify different levels of testing “focus” from random to highly targeted.
-For example, $w_I/w_S=3$ means that infected individuals are tested at tree times the \percap rate of susceptible individuals.
-
-In order to link to more applied models, we constructed this model so we could specify the total \percap testing rate.
-We do this by defining the weighted size of the testing pool $W = w_S S\_u + w_I I\_u + w_R R\_u$, and calculating a scaling parameter for testing as:
-\begin{equation}
-\label{sigma}
-\sigma = \frac{\rho N_0}{W},
-\end{equation}
-where $\rho$ is the \percap testing intensity for population and defined as the number of daily tests taken in a population of size $N_0$.
-Thus, the \percap testing rate for compartment $X$ is 
-\begin{equation}
-\label{F}
-F_X=\sigma w_X,\,\, \text{where}\, X \in \{S,I,R\}.
-\end{equation}
-For a high-sensitivity test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus unavailable for further testing.
-Over the course of the epidemic, a fixed testing rate as specified in \eqref{sigma} can (if large enough) exhaust the pool of people available for testing, leading to a singularity when no one is left untested.
-Although this phenomenon does not affect our analysis of $\Rnum$, it can affect the temporal dynamics (we discuss an adjustment to the model that solves this problem in the appendix).
-
-The classical SIR model is based on the following implicit assumptions; well-mixed population, homogeneity of the population (i.e., all individuals are equaly susciptable and equaly infectious for the same length of time when infected), exponentially distributed duration of infection and large population size (see, e.g., \cite{keeling2011modeling}).
-In addition to these standard assumptions, our model assumes:
-\begin{enumerate}[(i)]
-\item there is a single force of infection (new cases per unit time), $\Lambda$, defined as follows
-  \begin{equation}
-  \label{Lambda}
-  \Lambda=\frac{\beta}{N_0} \big(I\_u + (1-\theta\_w)(I\_n+I\_p) + (1-\theta\_c)I\_c \big),
-  \end{equation}
-  across all susceptible pools with transmission rate $\beta$ and isolation efficacy in reduction of the probability of transmission for three testing-based compartments ``waiting'' and \emph{confirmed positive} individuals, $\theta\_w$ and $\theta\_c$ respectively, (see Table~\ref{tab:params} for further details),
-\item the efficacy of isolation is higher in confirmed individuals relative to the individuals awaiting test results. That is,
-$$0\leq\theta\_w\leq\theta\_c\leq 1.$$ 
-\end{enumerate}
-Thus, for instance when the awaiting people follow the isolation perfectly, $\theta\_w$ is closer to 1, while when they less follow the isolation, $\theta\_w$ is closer to 0.
-For the simplicity of analysis, we also assume a perfectly specific test ($p_S=0$).
-This last assumption combined with the assumption that no susceptible individual is in waiting-for-\emph{positive} or \emph{confirmed positive} compartments, i.e., $S\_p(0)=S\_c(0)=0$, reduces the model to 10 equations where equations c and d in model \eqref{model} are eliminated.
-
-\begin{table}[htp]
-\centering
-\begin{tabular}{|c|L{2in}cc|} \hline
-  Symbol & Description & Unit & Value \\ \hline
-  $N_0$     & Total population size & people & $10^6$ \\ \hline
-  $\omega$  & Rate of test return, i.e., rate of onward flow from ``waiting'' to ``confirmed'' or ``untested'' compartments  & 1/day & - \\ \hline
-  $\gamma$ & Recovery rate & 1/day & 1/3 \\ \hline 
-  $\rho$   & \percap testing intensity & 1/day & - \\ \hline 
-  $\theta\_w$ & Isolation efficacy in reduction of the probability of transmission for ``waiting'' individuals & - & - \\ \hline
-  $\theta\_c$ & Isolation efficacy in reduction of the probability of transmission for ``confirmed positive'' individuals & - & -  \\ \hline
-  $\beta$ & Transmission rate & 1/day & 0.339 \\ \hline
-  $\Lambda$ & Force of infection & 1/day & - \\ \hline
-  $p_S$ & Probability of positive tests for $S$ ($= 1-\textrm{specificity}$) & - & 0 \\ \hline
-  $p_I$ & Probability of positive tests for $I$ ($= \textrm{sensitivity}$) & - & 1 \\ \hline
-  $p_R$ & Probability of positive tests for $R$ ($= 1-\textrm{specificity}$) & - & 0.5 \\ \hline
-  $w_S, w_I, w_R$ & Relative testing weights & - &
-  \begin{minipage}[t]{0.21\columnwidth}%
-    Random: $\{1,1,1\}$ \\ Targeted: $\{0.3,1,1\}$
-  \end{minipage} \\
-  \hline
-  \end{tabular}
-\caption{\label{tab:params} Parameters of the model \eqref{model}.}
-\end{table}
-
-The Disease-Free Equilibrium (DFE) for the SIR model (Eqs.\,\ref{model}) is found by setting the infected compartments to 0 and solving for the unknowns. The DFE is
-\begin{equation}
-\label{dfe}
-S\_n^*= \frac{\rho}{\omega}\,N_0, \ S\_u^*= \frac{\omega-\rho}{\omega}\,N_0, \text{~and~} I_j=R_j=0 \ \text{for all }j.
-\end{equation}
-The corresponding \percap testing rate (Eq.\,\ref{F}) for the infected compartment $I$ at DFE is one of the key analysis parameters and can be simplified as 
-\begin{equation}
-\label{eq:fi}
-\hat F_I = (\omega\rho/(\omega-\rho))w_I/w_S.
-\end{equation}
-
-The basic reproduction number, $\Rnum$, was calculated by using the next-generation matrix method developed by \cite{van2002reproduction}. We present $\Rnum$ in the following form.
-\begin{equation}
-\label{R0}
-\Rnum= \frac{\beta}{\gamma} (1-\Delta), 
-\end{equation}
-where the term $\frac{\beta}{\gamma}$ is the classical basic reproduction number for a SIR model without testing and isolation (see, e.g., \cite{keeling2011modeling}), and the positive quantity $\Delta$ is the ``effectiveness of control'' effectiveness of control due to testing and isolation processes defined as follows.
-\begin{equation}
-\label{eq:del4}
-\Delta= \frac{1}{C\,N_0}\big(C_1\,S\_u^*+(C_2(1-\theta\_w)+C\,\theta\_w)\,S\_n^*\big),
-\end{equation}
-where
-\begin{align}
-\label{eq:C12}
-C =& (\omega+\gamma) \Big(\gamma(\omega+\gamma)+(\gamma+\omega p_I)\hat F_I\Big),\\
-C_1 =& (\omega+\gamma)(\theta\_w \, \gamma+\theta\_c\, \omega\, p_I) \hat F_I,\\
-C_2 =& \Big( \omega\gamma(1+p_I)\hat F_I+\gamma^2\,(\omega+\gamma+\hat F_I)\Big)\theta\_w + \omega^2\,p_I\,\hat F_I\,\theta\_c.
-\end{align}
-Further details of derivation of $\Rnum$ are provided in appendix, Sec.\,\ref{app:R0}. 
-The explicit formula of $\Rnum$, enables us to study the effect of parameters of interest, on $\Rnum$ mathematically and through numerical simulation.
-These parameters, i.e., parameters that could be manipulated by public-health policy include the isolation efficacy, $\theta\_c$ and $\theta\_w$, \percap testing intensity, $\rho$, and the rate of test return, $\omega$.
-In partticular, we look at the partial derivative of $\Delta$ with respect to these parameters (see details in appendix, Sec.\,\ref{app:rho}).
-It is notable that the derivative of $\Delta$ with respect to $\theta\_w$, $\theta\_c$ and $\rho$ was derived explicitly and for the whole domain of the parameter space.
-However, we analyzed the effect of $\omega$ on $\Rnum$ for the case when testing intesity is very small relative to the population size, $N_0$.
-Specifically, we used the Taylor approximation of $\Delta$ at $\rho=0$ to illustrate the unexpected nonmonotonic relationship between $\Rnum$ and $\omega$, and also to aviod unnecessary complicated computations for making this point.
-
-The analytical calculation of the next-generation matrix and simplifying the expression of $\Rnum$ was carried out in Maple \citep{maple14}.
-We used R \citep{r} for simulations, in particular,  for plotting the contours of $\Delta$ \eqref{eq:del4} over a range of selected set of parameters of interest.
-The rest of model parameters kept fixed at sensible values (see Table~\ref{tab:params} for the parameter values).
-Note that all plots are illustrated in the scale of the mean test return time $1/\omega$ (day) and testing intensity $\rho$ (1\//day per 1000).
-Results are shown in \fref{pan} and \fref{pan2} with two sets of panels; panel set (a) represents the random testing, identified by $w_S=w_I=w_R=1$, and panel set (b) represents targeted testing, identified by $w_S=0.3$ and $w_I=w_R=1$.
-Each panel set partitions a plot into a matrix of panels, namely facets, in which rows are specified by a value of $\theta\_w$ and columns are specified by a value of $\theta\_c$.
-In our simulation, $\theta\_w$ and $\theta\_c$ vary in a discrete scale from 0 to 1 with 0.25 increment.
-Note that we are only plotting facets with $\theta\_w\leq\theta\_c$ which is an assumption of our model.
-
-To illustrate the changes in $\Rnum$ with respect to \percap testing intensity $\rho$, two sets of plots are presented. 
-\fref{pan} reflects the changes in $\Rnum$ when $\rho$ is small relative to the population size.
-Specifically, $\rho \in [0,0.013]$, and the test return rate $\omega\in [1/12,2]$.
-This is a more realistic scenario of testing as we have observed in \covid pandemic.
-That is, 1\% of a population are tested per day at maximum capacity, so in a population of size $N_0=10^6$, 10000 individuals will be tased per day at maximum.
-In order to illustrate the non-monotonic changes in $\Rnum$ with respect to $\rho$,
-\fref{pan2} with maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,1/5)$, and the test return rate $\omega\in [1/5,2]$. 
-Note that the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, corresponding to the threshold $\Rnum=1$, is plotted in dotted line in the panels.
-Thus, a combination of $\rho$ and $\omega$ above this critical contour results in reducing $\Rnum$.
-Also realize that $\Rnum=\frac{\beta}{\gamma}$, about 1.017 here, for the classical SIR model which can be acheived in the context of our model when no testing and no isolation are implied, that is $\rho=0$ or $\theta\_w=\theta\_w=0$.
-
-% %%%%%%%
-\section{Results}
-
-Examining our formula for $\Rnum$ \eqref{R0} gives the following results. 
-See the appendix, Sec.\,\ref{app:rho}, Sec.\,\ref{app:omega} and Sec.\,\ref{app:w} for details.
-
-\begin{enumerate}
-
-\item \label{p1:eta} Increasing the isolation efficacy parameters for tested and confirmed individuals decreases \Rnum;
-
-\item \label{p1:rho} Higher testing intensity $\rho$ always decreases $\Rnum$ \textbf{if}: 
-%% \begin{enumerate} \item
-testing is random; \textbf{or}
-%% \item
-$\rho$ is small.
-%% \end{enumerate}
-
-\item \label{p1:omega} A higher rate of test return always decreases \Rnum, \textbf{if} $\theta_w=0$.
-
-\item \label{p1:w} Increasing testing “focus” $\frac{w_I}{w_S}$ from random $w_S=w_I$ to targeted $w_S<w_I$ always decreases \Rnum.
-
-\end{enumerate}
-
-\newpage
-\begin{figure}[h!]
-\centering
-\begin{subfigure}[t]{.45\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_random.pdf}
-\caption{}\label{p.a}
-\end{subfigure}
-%
-\begin{subfigure}[t]{.45\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_TTI.pdf}
-\caption{}\label{p.b}
-\end{subfigure}
-\caption{
-{\bf A comparison of effectiveness control on $\Rnum$ at low level of \percap testing intensity $\rho$.}
-We numerically evaluate the effectiveness control parameter, $\Delta$ (Eq.~\ref{eq:del4}), with respect to testing and isolation.
-We use the following parameters (see Table~\ref{tab:params} for a brief description):
-$N_0=1 \times 10^6$, $\omega \in [1/12,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.013]$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for the full effect of isolation on the transmission probability, $\beta=0.339$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
-Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
-The black dotted line in each panel represents the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ corresponding to the threshold of $\Rnum=1$.
-}
-\label{pan}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\begin{subfigure}[t]{.45\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_random2.pdf}
-\caption{}
-\end{subfigure}
-%
-\begin{subfigure}[t]{.45\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_TTI2.pdf}
-\caption{}
-\end{subfigure}
-\caption{
-{\bf A comparison of effectiveness control on $\Rnum$ at high level of \percap testing intensity $\rho$.}
-We numerically evaluate the effectiveness control parameter, $\Delta$ (Eq.~\ref{eq:del4}), with respect to testing and isolation.
-We use the following parameters (see Table~\ref{tab:params} for a brief description):
-$N_0=1 \times 10^6$, $\omega \in [1/5,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,1/5)$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.339$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
-Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
-The black dotted line in each panel represents the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ corresponding to the threshold of $\Rnum=1$.}
-\label{pan2}
-\end{figure}
-
-Numerical results are shown in Figs.~\ref{pan} and \ref{pan2}.
-When $\theta_w =0$ (top row of each panel), we see that shorter test-waiting times reduce \Rnum\ ($\Delta$ increases as we move to the left in each plot in this row).
-For non-zero $\Delta$, however, we mostly see the opposite effect: less waiting leads to more transmission.
-This is because returning negative tests leads people to stop distancing; this applies both to susceptibles, and to infectious people who receive negative test results because they were sampled when susceptible (or because of imperfect test sensitivity, which is not modeled in this figure).
-
-We also see that greater test intensity $\rho$ increases the effectiveness of control $\Delta$ ($\Delta$ increases as we move up in each plot in this Figure).
-Mathematically speaking, we did find a counter-vailing effect, but this can only be seen when we allow $\rho$ to take unrealistically large values, and only for weighted/targeted testing (see \fref{pan2}).
-The explanation here is that more rapid testing leaves more susceptibles in the ``waiting-for-no'' category at the DFE; these people must then wait for their tests to be returned before they can be tested again, receive a positive test, and become cautious.
-This effect is usually weak compared to the helpful effects of testing, but if testing is weighted, $\theta_w$ is low, and test returns are slow, it is possible for increasing $\rho$ to reduce $\Delta$ (see upper-right part of upper-right plot of \fref{pan2}(b)).
-
-%%%%%%%%%%%%
-\section{Discussion}
-\begin{enumerate}
-
-\item {\bf A potential advantage of slow test reporting $\omega$, result~\ref{p1:omega};}
-
-- Whether increasing $\omega$ lowers $\Rnum$ depends on the precise combination of model parameters  including $\rho$, testing strategies represented by compartment-specific testing weights, test sensitivity and specificity, and the level of isolation efficacy.
-- The potential advantage of slow test reporting is in the individual level as opposed to the comunity level, but seems unapplicable as a strategy.
-- May be this mechanism combined with increasing testing intensity $\rho$ can be considered alongside other prevention methods, eg., social distancing etc, when the transmission is in the exponential-growth phase (in order to increase the generation interval).
-- One reason in favor of prolonging $1/\omega$ is that it is easy to manipulate while increasing $\rho$ can be limited to the resources.
-
-\item {\bf The counter-vailing effect of \percap testing intensity, $\rho$, on $\Rnum$;}
-- The explanation here is that more rapid testing leaves more susceptibles in the ``waiting-for-no'' $S\_n$ category at the DFE and delay them to get to $I\_c$ where they are subject to a full isolation, thus no transmission.
-- We are missing out on community-level advantages of fast testing.
-- What is the explanation for having this counter-vailing effect of $\rho$ in targeted testing and not in the random one?
-- May because in the targeted testing more $I\_u$s are tested commpared to $S\_u$s, given 
- that higher $\omega$ leaves more people in $S\_n$ and $I\_n$ which results in more transisssion.
-- While targeted testing strategies, including targeting people with infection-like symptoms or the contacts of confirmed cases, are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is that viral spread is greater for a slow test than for a fast test.
-This counter-intuitive effect can occur when people are cautious when awaiting a test result, and may not be robust to second-order effects of fast testing (such as better contact tracing).
-
-\item {\bf Specifying the compartment-specific relative testing weights, Specifically $\frac{w_I}{w_S}$;}
-
-- More targeted testing is always good in viral control,
-- We incorporated the compartment-specific relative testing weights, $w_S$, $w_I$ and $w_R$, to model random testing and \emph{targeted} testing strategies.
-Here, in the case of targeted testing and for the simplicity and illustration purposes, we assumed that infected and recovered individuals are tested at three times the \percap rate of susceptible individuals, thus $w_I/w_S=3$ and $w_R=w_I$.
-Note that we have not specified a methodology to assign particular relative testing weights corresponding to a particular targeted testing scenario.
-Modeling different targeted testing strategies, equivalently test-specific testing weights in our framework, requires prior information of the conditional probabilities of getting tested for people in a given compartment.
-This can be implied when we would like to quantify and compare the effect of different levels of test focus for infectious people on the basic reproduction number $\Rnum$, and conclude about the disease spread management.
-For example, when people are tested for "screening", the individuals with potential higher mobility, eg.
-people who are getting on flights, get more tested and thus the coresponding heavier testing weight is assigned than people awaiting for a surgery and are probably going to stay in a long-term-care facility and consequently less mobile and more isolated to begin with.
-With our model, we would be able to compare the sensitivity of the disease dynamics, through $\Rnum$, with respect to testing different high-risk groups in the population.
-This part needs to be developed further in future work.
-
-\item What can we say in the context of strength and speed paradigm?
-
-\end{enumerate}
-
-% %%%%%%%
-\bibliography{SIRlibrary}
-
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\clearpage
-% \widetext
-\begin{center}
-\textbf{\large Appendix}
-\end{center}
-\setcounter{equation}{0}
-\setcounter{figure}{0}
-\setcounter{table}{0}
-% \setcounter{page}{1}
-\makeatletter
-\renewcommand{\theequation}{A\arabic{equation}}
-\renewcommand{\thefigure}{A\arabic{figure}}
-\renewcommand{\bibnumfmt}[1]{[A#1]}
-\renewcommand{\citenumfont}[1]{A#1}
-
-\subsection{Model and calculation of $\Rnum$}\label{app:R0}
-
-The model in the form of a system of ordinary differential equations is 
-\begin{subequations}\label{model}
-\begin{align}
- d S\_u/dt &= -\Lambda S\_u - F_S S\_u + \omega S\_n, \\
- d S\_n/dt &= -(1-\theta\_w)\Lambda S\_n + (1-p_S) F_S S\_u - \omega S\_n, \\
- d S\_p/dt &= -(1-\theta\_w)\Lambda S\_p + p_S F_S S\_u - \omega S\_p, \\
- d S\_c/dt &= -(1-\theta\_c)\Lambda S\_c + \omega S\_p, \\
- d I\_u/dt &= \Lambda S\_u - F_I I\_u + \omega I\_n  - \gamma I\_u, \\
- d I\_n/dt &= (1-\theta\_w)\Lambda S\_n + (1-p_I) F_I I\_u - \omega I\_n -\gamma I\_n, \\
- d I\_p/dt &= (1-\theta\_w)\Lambda S\_p + p_I F_I I\_u - \omega I\_p -\gamma I\_p, \\
- d I\_c/dt &= (1-\theta\_c)\Lambda S\_c + \omega I\_p - \gamma I\_c, \\
- d R\_u/dt &= \gamma I\_u - F_R R\_u + \omega R\_n, \\
- d R\_n/dt &= \gamma I\_n + (1-p_R) F_R R\_u - \omega R\_n, \\
- d R\_p/dt &= \gamma I\_p + p_R F_R R\_u  - \omega R\_p, \\
- d R\_c/dt&= \gamma I\_c + \omega R\_p, \\
- dN/dt &= \omega (S\_n + I\_n + R\_n),  \\
- dP/dt &= \omega(I\_p + R\_p) ,
-\end{align}
-\end{subequations}
-
-where parameters are specified in Table \ref{tab:params}.
-The next generation matrix for this model is $G = F V^{-1}$, where matrix $F$ represents the inflow of new infection to the infected compartments and matrix $V$ represents the flow in the infected compartments when the population is totally susceptible.
-Matrices $F$ and $V$ are
-\begin{align}
-\label{FV}
-F =& \beta/N_0 \left[ \begin {array}{cccc} 
-S\_u^* & (1-\theta\_w)\,S\_u^* & (1-\theta\_w)\,S\_u^* & (1-\theta\_c)\,S\_u^*\\
-(1-\theta\_w)S\_n^* & (1-\theta\_w)^2\,S\_n^* & (1-\theta\_w)^2\,S\_n^* & (1-\theta\_w)(1-\theta\_c)\,S\_n^* \\ 
-0&0&0&0\\
-0&0&0&0
- \end {array} \right] \\
- =&\beta/N_0 \left[ \begin {array}{c} S\_u^* \\ (1-\theta\_w)S\_n^* \\ 0 \\ 0 \end {array} \right]
-        \left[ \begin {array}{cccc} 1, \, 1-\theta\_w, \, 1-\theta\_w, \, 1-\theta\_c \end {array} \right], \text{and}\\  V=&
- \left[ \begin {array}{cccc}  
-\hat F_I+\gamma&-\omega&0&0\\
--(1-p_I)\hat F_I&\omega+\gamma&0&0\\
--p_I \hat F_I&0&\omega+\gamma&0\\
-0&0&-\omega&\gamma
-\end {array} \right].
-\end{align}
-The matrix inverse of $V$ is 
-\begin{align}
-\label{vinv}
-V^{-1} =&
-\frac{1}{\gamma C}
-\left[ \begin {array}{cccc}
-\gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)&0&0\\ \noalign{\medskip}
-\gamma(\omega+\gamma)(1-p_I) \hat F_I & \gamma(\omega+\gamma)(\hat F_I+\gamma)&0&0 \\ \noalign{\medskip}
-\gamma(\omega+\gamma) p_I \hat F_I & \gamma\omega p_I \hat F_I & C \gamma/(\omega+\gamma) & 0 \\ \noalign{\medskip}
-\omega(\omega+\gamma) p_I \hat F_I & \omega^2 p_I \hat F_I & C \omega/(\omega+\gamma) & C
-\end {array} \right],
-\end{align}
-where $C= \big( \gamma(\omega+\gamma)+(\gamma+\omega p_I)\hat F_I \big) (\omega+\gamma)$ and $\hat F_I$ is the \percap testing rate for the infected people and represented in Eq.\,\eqref{eq:fi}.
-Note that all the columns of matrix $V^{-1}$ summ up to $1/\gamma$.
-
-The particular form of $F$ with two rows of zeros at the bottom results in the following blocked form of matrix $G$.
-\begin{equation}
-\label{mat:G}
-G = \left[ \begin {array}{cc}
-G_{11}&G_{12}\\
-0&0
-\end {array} \right],
-\end{equation}
-where both blocked matricies $G_{11}$ and $G_{12}$ are 2 by 2.
-Given the upper triangular form of matrix $G$, the basic reproduction number $\Rnum$ (defined as the spectral radius of matrix $G$) is only determined by the blocked matrix $G_{11}$,
-\begin{equation}
-\label{G11}
-G_{11} = \frac{\beta}{\gamma \, C} 
-\left[ \begin {array}{c} (\omega-\rho)/\omega \\ (1-\theta\_w)\rho/\omega \end {array} \right]
-\left[ \begin {array}{cccc} 1, \, 1-\theta\_w, \, 1-\theta\_w, \, 1-\theta\_c \end {array} \right]
-\left[ \begin {array}{cc}
-\gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)\\ \noalign{\medskip}
-\gamma(\omega+\gamma)(1-p_I)\hat F_I & \gamma(\omega+\gamma)(\hat F_I+\gamma) \\ \noalign{\medskip}
-\gamma(\omega+\gamma) p_I \hat F_I & \gamma\omega p_I \hat F_I \\ \noalign{\medskip}
-\omega(\omega+\gamma) p_I \hat F_I & \omega^2 p_I \hat F_I
-\end {array} \right].
-\end{equation}
-It is notable that matrix $F$ \eqref{FV} has rank one and consequently $G_{11}$ does so. That is $G_{11}$ has only one non-zero eigenvalue which is $\Rnum$.
-
-The expression of $\Rnum$ has a complicated form with all of the model parameters involved.
-This expression can be simplifed and represented given the specific form of matrix $G_{11}$ \eqref{G11}.
-For the purpose of simplicity we present $\Rnum$ in the manuscript in terms of expressions $C$, $C1$ and $C2$, specified in \eqref{eq:C12}.
-
-It remains hard to show that the reproduction number $\Rnum$ is decreasing with respect to \percap testing intensity, $\rho$, and the speed of the test return, $\omega$, for the feasible ranges of the parameters, that is
-\begin{align}
-\label{cond}
-& \omega>0, \\
-& 0 \leq \rho<\omega,\\ 
-& 0 \leq \theta\_w \leq \theta\_c \leq 1, \\
-& \frac{w_I}{w_S}\geq 1.
-\end{align}
-One way to do such an analysis is based on the fact that $\rho$ is \percap rate so for a large population size, $N_0 \gg 1$, $\rho$ is very small or close to 0, comparing to $N_0$.
-This provides a base to linearly approximate $\Rnum$ when $\rho$ is close to 0, and use this approximation to analyze the behaviour of $\Rnum$ with respect to $\omega$ (see section \ref{app:omega}).
-In the next section we provide an equivalent representation of $\Rnum$ to provide a ground to prove that more testing intensity decreases $\Rnum$ for a general rane of parameters.
-
-% %%%%%%%
-\subsection{More testing intensity may decreases $\Rnum$}\label{app:rho}
-
-This is to provide a mathematical materials to prove that $\pro \Delta$ can be positive or negative, with $\Delta$ defined in Eq.\,\eqref{eq:C12}, and thus $\pro \Rnum < 0$, where $\Rnum$ is the basic reproduction number, given in Eq.\,\eqref{R0}.
-We can rewrite matrix $G_{11}$ in \eqref{G11} in the following form to simplify the calculations.
-\begin{equation}
-\label{G112}
-G_{11} = \frac{\beta}{\gamma \, C} 
-\left[ \begin {array}{c}  S\_u^*/N_0 \\ (1-\theta\_w)\,S\_n^*/N_0  \end {array} \right]
-\left[ \begin {array}{cccc} 
-C-C_1, C-C_2\end {array} \right],
-\end{equation}
-where $C$ is the same as the one in Eq. \eqref{eq:C12}, i.e., 
-$$C=(\omega+\gamma)(\gamma(\omega+\gamma)+(\omega\,p_I+\gamma)\,\hat F_I),$$
-and $C_1$ and $C_2$ are 
-\begin{align*}
-% \label{eq:C12}
-C_1 =& (\omega+\gamma)(\theta\_w \, \gamma+\theta\_c\, \omega\, p_I) \hat F_I,\\
-C_2 =& \big( \omega\gamma(1+p_I)\hat F_I+\gamma^2\,(\omega+\gamma+\hat F_I)\big)\theta\_w + \omega^2\,p_I\,\hat F_I\,\theta\_c,
-\end{align*}
-where $\hat F_I$ is given in Eq.\,\eqref{eq:fi}.
-Note that for analysis brevity, we let $N_0=1$, thus $S\_u^*$ and $S\_n^*$ are in the scale of 0 to 1.
-$\Rnum$ is in the same form as in Eq. \eqref{R0}  
-$$\Rnum= \frac{\beta}{\gamma} (1-\Delta),$$
-where 
-\begin{equation*}
-% \label{eq:del4}
-\Delta= \frac{1}{C}\big(C_1\,S\_u^*+(C_2(1-\theta\_w)+C\,\theta\_w)\,S\_n^*\big).
-\end{equation*}
-
-{\bf The first goal} is to explore how changes in isolation, $\theta\_w$ and $\theta\_c$, affects $\Rnum$.
-Mathematically we would like to verify the sign of $\pder[\Rnum]{\theta\_w}$ and $\pder[\Rnum]{\theta\_c}$.
-We start with simpliying $\Delta$ \eqref{eq:del4} by factoring $\theta\_w$ and $\theta\_c$ in Eq.\,\eqref{eq:del4}.
-Thus, $\Delta$ can be rewritten as
-\begin{equation}
-\label{eq:del4_theta}
-\Delta= \frac{1}{C}\Big(
--D_1\,S\_n^*\,\boldsymbol {\theta\_w^2}+
-\big(-\omega^2\,p_I\,\hat F_I\,S\_n^*\,\boldsymbol{\theta\_c}\,+\,D_2\,S\_n^*\,+\,\gamma\,\hat F_I(\omega+\gamma) \big)\boldsymbol{\theta\_w}+
-(\omega+\gamma\,S\_u^*)\omega\,p_I\,\hat F_I\,\boldsymbol{\theta\_c}
-\Big),
-\end{equation}
-where
-\begin{align}
-\label{eq:del4_d1d2}
-D_1=& (\omega+\gamma)\gamma^2+(\omega+\gamma+\omega\,p_I)\gamma\,\hat F_I, \\
-D_2=& (3\omega+2\gamma)\gamma^2+(\omega+\gamma+2\omega\,p_I)\gamma\,\hat F_I+(\gamma+p_I\,\hat F_I)\omega^2.
-\end{align}
-$\Delta$, Eq.\,\eqref{eq:del4_theta}, is linear in $\theta\_c$ with a positive coefficient given by 
-$$1/C\,(\gamma\,S\_u^*\,+\omega(1-\theta\_w\,S\_u^*))\omega\,p_I\,\hat F_I.$$
-This results in increasing $\Delta$, thus decreasing $\Rnum$ with respect to $\theta\_c$, that is $\pder[\Rnum]{\theta\_c}\leq 0$. Note that $C$ is independant of $\theta\_c$ and $\theta\_w$.
-
-With a similar logic, $\Delta$ \eqref{eq:del4_theta} is a concave-down quadratic equation in $\theta\_w$, given by
-\begin{equation}
-\label{eq:del4_thetaw}
-1/C\,\Big( -D_1\,S\_n^*\,\boldsymbol {\theta\_w^2}+
-\big(-\omega^2\,p_I\,\hat F_I\,S\_n^*\,\boldsymbol{\theta\_c}\,+\,D_2\,S\_n^*\,+\,\gamma\,\hat F_I(\omega+\gamma) \big)\boldsymbol{\theta\_w} \Big).
-\end{equation}
-We show that the feasible range of $\theta\_w$ lies btween 0 and the vertex of this parabola where the parabola is increasing in $\theta\_w$, and so does $\Delta$ which results in inferring $\pder[\Rnum]{\theta\_w}\leq 0$.
-It is enough to show that partial derivative of the expression \eqref{eq:del4_thetaw} with respect to $\theta\_w$ at $\theta\_w=1$ is non-negative. It follows that
-\begin{align}
-\label{eq:d_del4_thetaw}
-\pder[\Delta]{\theta\_w}\bigg\rvert_{\theta\_w=1} =&
-1/C\Big( (D_2-2D_1-\omega^2\,p_I\,\hat F_I\,\theta\_c)\,S\_n^*+\gamma\,\hat F_I(\omega+\gamma) \Big) \\
-=& 1/C\Big( (\gamma(\omega+\gamma)+\gamma\omega^2+(1-\theta\_c)\omega^2\,p_I\,\hat F_I )\,S\_n^*
-+\gamma(\omega+\gamma)\,\hat F_I\,(1-S\_n^*) \Big),
-\end{align}
-which is a positive quantity, given that $\theta\_c$ and $S\_n^*$ vary between 0 and 1.
-
-{\bf The second goal} is to explore how changes in \percap testing intensity $\rho$ affects $\Rnum$.
-Mathematically we would like to verify the sign of $\pder[\Rnum]{\rho}$, which specifically depends on $\pder[\Delta]{\rho}$.
-We use the derived expressions for $S\_u^*$ and $S\_n^*$, given by Eqs.\,\eqref{dfe}, in $\Delta$ \eqref{eq:del4}.
-Also, we define $\phi = \hat F_S = \frac{\rho \omega}{\omega-\rho}$, to reparametrize $\rho$.
-This is mainly to avoid singularity in $\hat F_I$ \eqref{eq:fi}, when testing intensity $\rho$ is very close to the rate of test return $\omega$.
-Thus, $\rho$ is reparametrized as
-\begin{equation}
-\label{eq:phi}
-\rho=\frac{\omega \phi}{\omega+\phi}.
-\end{equation}
-This one-to-one monotonic reparametrization enables us to simplify the mathematical expressions and explore the simpler $\pder[\Delta]{\phi}$ instead of the complicated $\pder[\Delta]{\rho}$.
-The derivative is
-\begin{align}
-\label{eq:dd3dr}
-\partial\Delta/\partial\phi=& \frac{1}{d_3} (a_3\,\phi^2+b_3\,\phi+c_3),
-\end{align}
-where
-\begin{align}
-\label{eq:abcd2}
-a_3& = \frac{w_I}{w_S} \Bigg(&& (1-\theta\_w)\,(1+\frac{w_I}{w_S})\theta\_w\,\gamma^3
-+(1-\theta\_c)\,p_I^2\,\theta\_w\,\frac{w_I}{w_S}\,\omega^3 \\ \notag
-& ~ &&+\Big( \big( (1-\theta\_w-\frac{w_I}{w_S})\,\theta\_c +(3-2\theta\_w)\,\theta\_w\frac{w_I}{w_S} \big)\,p_I
-+(1-\theta\_w)\,(1+\frac{w_I}{w_S})\theta\_w  \Big)\,\omega\,\gamma^2 \\ \notag
-& ~ &&+\Big(
-\big((1-\theta\_w-\theta\_w\,\frac{w_I}{w_S})\,\theta\_c+ (2-\theta\_w)\,\theta\_w\,\frac{w_I}{w_S} \big)\,p_I
-+(2\,\theta\_w-\theta\_w^2-\theta\_c)\,\frac{w_I}{w_S}\,p_I^2
-\Big)\,\omega^2\,\gamma \Bigg), \\
-b_3& = && \hspace{-1cm} 2\frac{w_I}{w_S}\,(\omega+\gamma)\gamma\Big(
-(\omega+\gamma+\omega\,p_I)(2-\theta\_w)\,\gamma\,\theta\_w+ (1-\theta\_w)\,\omega^2 p_I\theta\_c+\omega^2 p_I\theta\_w
-\Big) ,\\
-c_3& = && \hspace{-1cm} (\omega+\gamma)^2\,\gamma\Big(
-(2-\theta\_w)\,\gamma^2\,\theta\_w+
-(1+\frac{w_I}{w_S})\,\omega\,\gamma\,\theta\_w+
-\frac{w_I}{w_S} \omega^2 p_I\theta\_c \Big),\\
-d_3&  = && \hspace{-1cm} \frac{(\omega+\gamma)}{\omega}\, \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi\,+\,(\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2.
-\end{align}
-Note that $\phi\geq 0$, also $b_3$, $c_3$ and $d_3$ are all positive. However $a_3$ can be positive or negative.
-If $a_3\geq 0$, $\partial\Delta/\partial\phi \geq 0$ for all feasible range of parameters, thus $\pro\Rnum \leq 0$.
-It is straight forward to show that $a_3\geq 0$ in case of random testing strategy, i.e., $w\_S=w\_I=1$.
-If $a_3 < 0$, then the quadratic expression in the numerator of \eqref{eq:dd3dr} has a positive root, $\phi^*$, such that for $\phi>\phi^*$, $\partial\Delta/\partial\phi < 0$.
-
-An example of this counter-vailing effect of $\phi$, and consequently $\rho$, on $\Rnum$ occures when $\theta\_w=0$ and $\theta\_c=1$.
-This is illustraited in the top-right panel of the \fref{pan2} panel (b), where the strength of isolation for awaiting people is the least, but the most for the confirmed cases.
-In this case, simplifying $a_3$ in Eq.~\eqref{eq:abcd2} gives $$a_3=\frac{w_I}{w_S}\,\omega\,\gamma\,p_I((\omega+\gamma)-\frac{w_I}{w_S}(\omega\,p_I+\gamma)).$$
-Specifically, in the case of targeted testing which is identified by $\frac{w_I}{w_S}> 1$, and using a perfect sensative test, thus $p_I=1$, there exists a range for $\rho$ over which $\pder[\Rnum]{\rho}\leq\,0$.
-Note that $\rho$ and $\omega$ have a similar mechanism in delaying people to get into $I\_c$, thus we would expect to see the non-trivial counter-vailing effect of these two parameters on $\Rnum$.
-
-% %%%%%%%
-\subsection{rate of returning tests} \label{app:omega}
-{\bf The third goal} is to explore how changes in the rate of test return $\omega$ affects $\Rnum$.
-Mathematically we would like to verify the sign of $\pder[\Rnum]{\omega}$, which specifically depends on $\pder[\Delta]{\omega}$.
-We use
-the linearization of $\Rnum$ around $\rho=0$ to show that there a non-monotonic relationship between $\Rnum$  and $\omega$. The taylor expansion of $\Delta$ at $\rho=0$ is
-\begin{equation}
-\label{eq:lin}
-\Delta = \frac{\rho}{\omega\,\gamma(\omega+\gamma)} \Big(
-\frac{w_I}{w_S}\,\omega^2\,p_I\,\theta\_c+(\frac{w_I}{w_S}+1)\,\gamma\omega\theta\_w\,+\gamma^2\,\theta\_w(2-\theta\_w)
-\Big). 
-\end{equation}
-This results in
-\begin{equation}
-\label{eq:dlindo}
-\pder[\Delta]{\omega} = \frac{\rho}{\omega^2\,(\omega+\gamma)^2} \Big(
-(p_I\,\frac{w_I}{w_S}\theta\_c\,-(1+\frac{w_I}{w_S})\theta\_w)\omega^2+2\theta_w\,\gamma(\theta_w-2)\,\omega+\theta_w\,\gamma^2\,(\theta_w-2)
-\Big),
-\end{equation}
-around $\rho=0$.
-\ali{I stoped here!}
-
-Perhaps counter-intuitively, the equation above does not predict that $\Rnum$ is monotone decreasing with respect to $\omega$.
-In other words; our model does not predict that returning test results more rapidly \textit{always} lower $\Rnum$.
-In order to gain insight into this intriguing behavior, we examine the zeroes of $\frac{\partial{\Rnum}}{\partial{\omega}}(\omega)$.
-Defining the following quantity, $Q$, will help us write the roots of $\partial{\Rnum}/\partial{\omega}$ neatly as follows.
-\begin{align}\label{eq:defQ}
-    Q =& \frac{w_I}{w_S}\left(1-\frac{n_t-1}{n_w-1}p_I \right).
-\end{align}
-With that in mind, we can write the roots of $\partial{\Rnum}/\partial{\omega}$ as
-\begin{align}
-    \omega_1 =& \frac{\gamma}{-\sqrt{1-Q}-1} \\
-    \omega_2 =& \frac{\gamma}{\sqrt{1-Q}-1}.
-\end{align}
-
-Note that the zeroes are real if and only if $Q < 1$.
-Note that have $\theta\_c > \theta\_w$, so if $p_I \approx 1$, we will have $Q < 0 < 1$.
-Thus, if we assume near-perfect test sensitivity, $\omega_1$ and $\omega_2$ will be real.
-
-Assuming $\omega_1, \omega_2$ are real, it is easy to confirm that $\omega_1 < 0$ by looking at the denominator.
-To see that $\omega_2 > 0$, recall that $Q < 0$, so $\sqrt{1-Q} > 1$ and so $\sqrt{1-Q} -1 > 0$.
-Knowing that $\omega_1 < 0$, the only root of interest (i.e., biologically relevant quantity) is $\omega_2$.
-
-We can prove that $\partial{\Rnum}/\partial{\omega} > 0$ when $\omega \in (0,\omega_2)$ and $\partial{\Rnum}/\partial{\omega} < 0$ when $\omega \in (\omega_2,\infty)$ by computing the limits of $\partial{\Rnum}/\partial{\omega}$ at $0$ and $\infty$ respectively.
-So it follows that $\Rnum$ has a global maximum with respect to $\omega$ at $\omega = \omega_2$.
-
-Now we want to characterize the parameter regions on which $\partial{\Rnum}/\partial{\omega} < 0$ (i.e., the conditions under which returning test results more rapidly is favorable).
-By the previous analysis, this is equivalent to solving for $\omega > \omega_2$.
-So
-\begin{align}
-    &\omega > \omega_2 \nonumber \\
-    &\omega > \frac{\gamma}{\sqrt{1-Q}-1} \nonumber \\
-    &\sqrt{1-Q} > \frac{\gamma}{\omega}+1 \\
-    &1-Q > (\frac{\gamma}{\omega}+1)^2.
-\end{align}
-Substituting in $Q$ from \eqref{eq:defQ} we have
-\begin{align}
-    &1-\frac{w_I}{w_S}\left(1-\frac{n_{t}-1}{n_{\omega}-1}P_{i}\right)> \left(\frac{\gamma}{\omega}+1\right)^2 \\
-    &-\frac{w_I}{w_S}\left(1-\frac{n_{t}-1}{n_{\omega}-1}P_{i}\right)> \left(\frac{\gamma}{\omega}+1\right)^2+1 \\    
-    &\frac{w_I}{w_S}\left(\frac{1-n_{t}}{1-n_{\omega}}P_{i}-1\right)> \left(\frac{\gamma}{\omega}+1\right)^2+1\label{eq:necsuf}.
-\end{align}
-Since all steps in deriving \eqref{eq:necsuf} are reversible, \eqref{eq:necsuf} gives a necessary and sufficient condition for $\omega > \omega_2$, which characterizes when returning tests more rapidly would cause a decrease in $\Rnum$.
-
-% %%%%%%%
-\subsection{On Testing Rate and Numerical Singularity}
-
-In this work, we didn't do any numerical solutions for the trajectories in our analysis. However, if one tries to do so there would be a singularity issue to deal with. 
-Specifically, the numerical singularity issue with the chosen $\sigma$ \eqref{sigma} is that the population in $S$ compartments appeared to blow up when the DFE is achieved.
-This is once the only untested people are susceptibles, the FOI will become $\Lambda=0$, testing rate $F_s=\rho N_0/S\_u$.
-Thus, the first equation of the model \eqref{model} will become
-$d S\_u/dt = - \rho N_0 + \omega S\_n$. Thus changes in $S\_u$ will be no longer dependent on $S\_u$ with a linear rate of leaving the $S\_u$ compartment.
-IN fact the testing rate, $\sigma$, should be formulated such that people from the untested compartments will not be tested if they are not there.
-One way to fix this issue, is to consider a maximum testing rate, $\tau$ (1/day).
-In general, we want to test at a rate of $\rho$ across the whole population.
-This won't always be possible, so we impose a maximum rate of $\tau$ per testable person and redefine $\sigma = \frac{\tau \rho N_0}{\tau W + \rho N_0}$, with the assumption that $\tau \gg \rho$.
-This alteration in $\sigma$, does not change any results related to $\Rnum$, thus we only impose it in the simulation of the epidemic dynamic.
-
-% %%%%%%%
-\subsection{Expensive vs.\ cheap tests}
-
-The use of tests cheaper than RT-PCR has been proposed as a potential strategy for containing the \covid pandemic.
-While cheaper tests may be less sensitive and reliable than RT-PCR, they allow for broader and more intense testing.
-In the analysis below, we compare the $\Rnum$ predicted by our model depending on the testing strategy.
-
-Consider a test that allows us to test at rate $\rho_1$ and has sensitivity $P_{i,1}$, and another test that allows us to test at $\rho_2$ and has sensitivity $P_{i,2}$.
-Suppose that $\rho_1 > \rho_2$.
-Recall that the linearization of $\Rnum$ around $\rho \approx 0$ is given by $$\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(-\theta\_w)(\gamma w_S+\omega w_I) + (-\theta\_c)p_I w_I \omega^2 \Big).$$
-
-Treating $\Rnum$ as a function of $\rho$ and $P_i$,we can reduce the inequality $$\Rnum(\rho_2, p_{I,2}) < \Rnum(\rho_1, p_{I,1})$$ into
-
-\begin{align}\label{eq:rho1vsrho2}
-    &\rho_1\left(\gamma(-\theta\_w)(\gamma w_S + \omega w_I) + (-\theta\_c)p_{I,1}w_I\omega^2\right) - \rho_2\left(\gamma(-\theta\_w)(\gamma w_S + \omega w_I) + (-\theta\_c)p_{I, 2}w_I\omega^2\right) > 0 \nonumber \\
-    &\vdots \nonumber \\
-    &\frac{\rho_2P_{i, 2}-\rho_1P_{i, 1} }{\rho_1-\rho_2} > \frac{\theta\_w}{\theta\_c}\cdot \frac{\gamma(\gamma w_S + \omega w_I)}{\omega^2 w_I}
-\end{align}
-
-Note that the RHS is positive, thus a necessary condition for the inequality above to hold is that $\rho_2P_{i,2} > \rho_1P_{i,1}$, equivalently
-
-\begin{equation}
-\frac{P_{i,2}}{P_{i,1}} > \frac{\rho_1}{\rho_2}.
-\end{equation}
-
-To state an example of this, if test $A$ is three times as expensive as test $B$ (and hence one can test three times as many people with test $B$), using test $A$ rather than $B$ will be favorable only if test $A$ is at least 3 times more sensitive than test $B$.
-Note that this is a necessary but not sufficient condition, so even if test $A$ is three times more sensitive, it is still possible for test $B$ to be more effective.
-
-\cref{eq:rho1vsrho2} tells us precisely when a test corresponding to $\rho_2, P_{i,2}$ will yield a lower $\Rnum$ than a test corresponding to $\rho_1, P_{i,1}$, where $\rho_1 > \rho_2$.
-Some of the qualitative \textit{trends} that favor test 2 (the higher-sensitivity test) include
-
-\begin{itemize}
-    \item individuals who test positive self-isolate much more than individuals who are waiting for their test result.
-    \item the time it takes to return tests is much shorter than the mean infectious period.
-    \item the testing intensity is much greater for infected individuals than susceptible individuals.
-\end{itemize}
-
-% %%%%%%%
-\subsection{The effect of testing focus parameter $\frac{w_I}{w_S}$ on $\Rnum$} \label{app:w}
-
-We define $w\_{IS}=\frac{w_I}{w_S}$.
-\begin{align}
-\pder[\Delta]{w\_{IS}}= \dfrac{(\omega-\rho)\,(\omega(\omega-\rho\theta\_w)+\gamma(\omega-\rho))\,(\theta\_w\gamma+\theta\_c\omega\,p\_I)}
-{(-\omega^2\,\gamma+\omega\,\gamma\,\rho-\gamma\,\rho\,\omega\,w\_{IS}-
-\omega\,\gamma^2+\gamma^2\,\rho-\omega^2\,p_I\,\rho\,w\_{IS})^2},
-\end{align}
-which is a positive quantity. Thus, $\pder[\Rnum]{w\_{IS}}\leq 0$. Therefore, increasing the focus of testing on the infectious people will result in less transmission.  
+\bibliography{../SIRlibrary}
 \end{document}

### 2021-05-17 SIR_manuscript.tex
minor changes in the abstract

@@
 Here, we compare two testing strategies, random vs.
-targeted, and concluded that the targeted testing strategy is more effective in the sense that achieving reduction of $\Rnum$ can occure in a lower range of testing intensity and longer test return time relative to the random testing.
-Furthermore, we show that increasing \percap testing intensity and reducing the test return time would be beneficial on the dynamics of an epidemic in general but there are exception cases.
-In particular, it is possible for the basic reproduction number, $\Rnum$, to be increasing with respect to the \percap testing intensity and the rate of test return.
+targeted, and concluded that increasing testing “focus” from random to targeted always decreases \Rnum.
+Further, the following processes always decreases $\Rnum$; increasing the isolation efficacy parameters for tested and confirmed individuals, higher testing intensity if testing is random or testing intensity is small, and a higher rate of test return if the isolation efficacy for tested-awaiting individuals is low.
 

### 2021-05-17 note/cohort.tex
adding more details to the cohort work

@@
 
-\title{The Paradigm of Cohort Equations for the SIR  Model With Testing and Isolation}
+\title{The analogy of the cohort-equations paradigm to the compartmental epidemic models}
 
@@
 
+\section{Vision}
+\begin{enumerate}
+\item 2 frameworks of \cite{van2002reproduction}, I call it compartment framework, and \cite{champredon2018equivalence}, I call it cohort framework, can be tied together and a mechanistic approach to go from one framework to another can be constructed.
+
+\item 
+\end{enumerate}
+
 \section{Introduction}
+Notation; we use $I'$ for the cohort-framework of $dI/d\tau$ and $\dot I$ for the compartment-framework of $dI/dt$.
 
-\section{Ideas}
-This is follwoing the work of \cite{champredon2018equivalence}.
+A simple example of this transition is the cohort analogy of a simple SIR compartmental model where $\dot I=\beta S I/N-\gamma I$. The cohort-framework is via the cohort Eq. with the following steps
+step 1: write the cohort Eq. $I'=-\gamma I$ with the initial condition $I(0)=1$, 
+step 2: finding the intrinsic infectiousness kernel by integrating the cohort Eq. and solving for its time evolution, thus
+$k(\tau)=\beta \exp(-\gamma \tau)$.
+step 3: $R$ will be the integration of the kernel. That is $R=\beta/\gamma$. 
+Note that in the comartment framework, steps 2 and 3 are combined.
 

### 2021-05-17 SIR_manuscript.tex
mostly spell-check

@@
 
-The effect of testing processes, including testing and test reporting, on an epidemic dynamics, involving infection and recovery, can be studied at the individual level or the community level (e.g., nursing homes, long-term-care facilities, etc.).
+The effect of testing processes, including testing and test reporting, on epidemic dynamics, involving infection and recovery, can be studied at the individual level or the community level (e.g., nursing homes, long-term-care facilities, etc.).
 Gaining insights to determine the sensitivity of the epidemic dynamics with respect to the testing processes will depend on underlying factors including the level of focus (individual or community), assumptions (model), and the interplay between these factors.
-In particular, the fast testing and test reporting may be beneficial at the community-level, supported by many studies, as it gives a rapid assessment of the situation, identifies hot spots, and may enable rapid contact-tracing.
+In particular, fast testing and test reporting may be beneficial at the community-level, supported by many studies, as it gives a rapid assessment of the situation, identifies hot spots, and may enable rapid contact-tracing.
 However, the potential advantage of a slow rate of test return on the dynamics of an epidemic is real, often neglected, and needs to be quantified.
@@
 In addition to shaping epidemic observations (via case reports), testing processes can also affect epidemiological dynamics.
-In particular, individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may do so also (possibly to a lesser extent).
+In particular, individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may also do so (possibly to a lesser extent).
 We developed a mechanistic model that incorporates epidemic processes and testing in order to explore the effects of testing and isolation on epidemic dynamics.
@@
 When testing is done for the purposes of disease surveillance \citep{foddai2020base}
-tests should be assigned randomly across the population, possibly with a stratified design for statistical efficiency \citep{graubard1996modelling} \ali{a better ref everyone?}.
+tests should be assigned randomly (or using a stratified random design) across the population in
+order to make an unbiased assessment of population prevalence.
 
 Over the course of the \covid pandemic, however, the vast majority of testing has been done with other goals --
-primarily diagnosis (determining the infection status for clinical purposes), or control (determining the infection status in order to quickly isolate cases that have been found by contact tracing), which we characterize as \emph{targeted} testing strategies.
+primarily diagnostic (determining infection status for clinical purposes), or for control (determining  infection status in order to isolate cases that have been found by contact tracing), which we characterize as \emph{targeted} testing strategies.
 In these cases, testing probabilities vary widely across epidemiological compartments; in our dynamical model, we will characterize these probabilities by assigning a \percap testing weight to each compartment that determines the \emph{relative} probability that an individual in that compartment will be selected for testing (see Methods).
@@
 When testing is used primarily for diagnosis it will focus on people with infection-like symptoms; thus the relative testing weights for infected people will depend on the relative probability of infected people having symptoms.
-For \covid infection, the testing weights will depend on the relative asymptomatic infections, time spent pre-sympomatic vs.\ symptomatic infections -- and also the incidence of \covid-like symptoms among people in the population \emph{not} infected with \covid.
+For \covid infection, the testing weights will depend on the relative asymptomatic infections, time spent pre-symptomatic vs.\ symptomatic infections -- and also the incidence of \covid-like symptoms among people in the population \emph{not} infected with \covid.
 Testing for epidemic control will focus on people who are known to have been in contact with known infected cases; in this case the testing weights for infected vs.\ uninfected people will depend on the probability of infection given contact, as well as the thoroughness and effectiveness of the system for identifying suspicious contacts.
 
-The main interest from the epidemiological point of view is to know whether the number of infected individuals goes through an exponential growth phase, following the introduction of an infection in a totally susciptable population, before the disease becomes extinct.
+The main interest from the epidemiological point of view is to know whether the number of infected individuals goes through an exponential growth phase, following the introduction of an infection in a totally susceptible population, before the disease becomes extinct.
 This is determined by studying the basic reproduction number $\Rnum$.
@@
 
-In order to understand the effect of testing processes on an epidemic dynamics, we deveopled a mechanistic SIR-type model with epidemic and testing components.
+In order to understand the effect of testing processes on an epidemic dynamics, we developed a mechanistic SIR-type model with epidemic and testing components.
 This model provides a sensible platform to link the modeling of epidemic and testing components and study their interaction.
 Here, we studied the the effect of testing intensity, rate of test return and the isolation efficacy in reduction of the probability of transmission on the epidemic dynamics when different levels of testing ``focus'' (from random to highly targeted) are in place.
-Our model provides insights to the sensitivity of the epidemic dynamics, through $\Rnum$, with respect to the undelying testing and isolation parameters.\ali{edited in response to David comments.}
+Our model provides insights to the sensitivity of the epidemic dynamics, through $\Rnum$, with respect to the underlying testing and isolation parameters.\ali{edited in response to David comments.}
 
@@
 Symbolically, the testing status of an individual in the disease compartment $X$, where $X \in \{S,I,R\}$, is reflected in the subscript, namely $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
-Note that the top-to-bottom order of the testing-based compartments of each disease-based compartmenet $X$ in the flowchart (\fref{fig:flowchart}) and the model equations \eqref{model} should match.
+Note that the top-to-bottom order of the testing-based compartments of each disease-based compartment $X$ in the flowchart (\fref{fig:flowchart}) and the model equations \eqref{model} should match.
 However, we switched  $X\_u$ and $X\_n$ in the flowchart (\fref{fig:flowchart}) for the sake of tidiness of the flowchart.
@@
 The novel component of the model comes in through the compartment-specific relative testing weights $w_S$, $w_I$ and $w_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively.
-Thus, we can spesify different levels of testing “focus” from random to highly targeted.
+Thus, we can specify different levels of testing “focus” from random to highly targeted.
 For example, $w_I/w_S=3$ means that infected individuals are tested at tree times the \percap rate of susceptible individuals.
@@
 
-The classical SIR model is based on the following implicit assumptions; well-mixed population, homogeneity of the population (i.e., all individuals are equaly susciptable and equaly infectious for the same length of time when infected), exponentially distributed duration of infection and large population size (see, e.g., \cite{keeling2011modeling}).
+The classical SIR model is based on the following implicit assumptions; well-mixed population, homogeneity of the population (i.e., all individuals are equally susceptible and equally infectious for the same length of time when infected), exponentially distributed duration of infection and large population size (see, e.g., \cite{keeling2011modeling}).
 In addition to these standard assumptions, our model assumes:
@@
 These parameters, i.e., parameters that could be manipulated by public-health policy include the isolation efficacy, $\theta\_c$ and $\theta\_w$, \percap testing intensity, $\rho$, and the rate of test return, $\omega$.
-In partticular, we look at the partial derivative of $\Delta$ with respect to these parameters (see details in appendix, Sec.\,\ref{app:rho}).
+In particular, we look at the partial derivative of $\Delta$ with respect to these parameters (see details in appendix, Sec.\,\ref{app:rho}).
 It is notable that the derivative of $\Delta$ with respect to $\theta\_w$, $\theta\_c$ and $\rho$ was derived explicitly and for the whole domain of the parameter space.
-However, we analyzed the effect of $\omega$ on $\Rnum$ for the case when testing intesity is very small relative to the population size, $N_0$.
-Specifically, we used the Taylor approximation of $\Delta$ at $\rho=0$ to illustrate the unexpected nonmonotonic relationship between $\Rnum$ and $\omega$, and also to aviod unnecessary complicated computations for making this point.
+However, we analyzed the effect of $\omega$ on $\Rnum$ for the case when testing intensity is very small relative to the population size, $N_0$.
+Specifically, we used the Taylor approximation of $\Delta$ at $\rho=0$ to illustrate the unexpected non-monotonic relationship between $\Rnum$ and $\omega$, and also to avoid unnecessary complicated computations for making this point.
 
 The analytical calculation of the next-generation matrix and simplifying the expression of $\Rnum$ was carried out in Maple \citep{maple14}.
-We used R \citep{r} for simulations, in particular,  for plotting the contours of $\Delta$ \eqref{eq:del4} over a range of selected set of parameters of interest.
+We used R \citep{r} for simulations, in particular,  for plotting the contours of $\Delta$ \eqref{eq:tel4} over a range of selected set of parameters of interest.
 The rest of model parameters kept fixed at sensible values (see Table~\ref{tab:params} for the parameter values).
@@
 This is a more realistic scenario of testing as we have observed in \covid pandemic.
-That is, 1\% of a population are tested per day at maximum capacity, so in a population of size $N_0=10^6$, 10000 individuals will be tased per day at maximum.
+That is, 1\% of a population are tested per day at maximum capacity, so in a population of size $N_0=10^6$, a maximum of 10000 individuals will be tested per day.
 In order to illustrate the non-monotonic changes in $\Rnum$ with respect to $\rho$,
@@
 Thus, a combination of $\rho$ and $\omega$ above this critical contour results in reducing $\Rnum$.
-Also realize that $\Rnum=\frac{\beta}{\gamma}$, about 1.017 here, for the classical SIR model which can be acheived in the context of our model when no testing and no isolation are implied, that is $\rho=0$ or $\theta\_w=\theta\_w=0$.
+Also realize that $\Rnum=\frac{\beta}{\gamma}$, about 1.017 here, for the classical SIR model which can be achieved in the context of our model when no testing and no isolation are implied, that is $\rho=0$ or $\theta\_w=\theta\_w=0$.
 
@@
 - Whether increasing $\omega$ lowers $\Rnum$ depends on the precise combination of model parameters  including $\rho$, testing strategies represented by compartment-specific testing weights, test sensitivity and specificity, and the level of isolation efficacy.
-- The potential advantage of slow test reporting is in the individual level as opposed to the comunity level, but seems unapplicable as a strategy.
+- The potential advantage of slow test reporting is in the individual level as opposed to the community level, but seems inapplicable as a strategy.
 - May be this mechanism combined with increasing testing intensity $\rho$ can be considered alongside other prevention methods, eg., social distancing etc, when the transmission is in the exponential-growth phase (in order to increase the generation interval).
@@
 - What is the explanation for having this counter-vailing effect of $\rho$ in targeted testing and not in the random one?
-- May because in the targeted testing more $I\_u$s are tested commpared to $S\_u$s, given 
- that higher $\omega$ leaves more people in $S\_n$ and $I\_n$ which results in more transisssion.
+- May because in the targeted testing more $I\_u$s are tested compared to $S\_u$s, given 
+ that higher $\omega$ leaves more people in $S\_n$ and $I\_n$ which results in more transmission.
 - While targeted testing strategies, including targeting people with infection-like symptoms or the contacts of confirmed cases, are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is that viral spread is greater for a slow test than for a fast test.
@@
 For example, when people are tested for "screening", the individuals with potential higher mobility, eg.
-people who are getting on flights, get more tested and thus the coresponding heavier testing weight is assigned than people awaiting for a surgery and are probably going to stay in a long-term-care facility and consequently less mobile and more isolated to begin with.
+people who are getting on flights, get more tested and thus the corresponding heavier testing weight is assigned than people awaiting for a surgery and are probably going to stay in a long-term-care facility and consequently less mobile and more isolated to begin with.
 With our model, we would be able to compare the sensitivity of the disease dynamics, through $\Rnum$, with respect to testing different high-risk groups in the population.
@@
 \end{equation}
-where both blocked matricies $G_{11}$ and $G_{12}$ are 2 by 2.
+where both blocked matrices $G_{11}$ and $G_{12}$ are 2 by 2.
 Given the upper triangular form of matrix $G$, the basic reproduction number $\Rnum$ (defined as the spectral radius of matrix $G$) is only determined by the blocked matrix $G_{11}$,
@@
 The expression of $\Rnum$ has a complicated form with all of the model parameters involved.
-This expression can be simplifed and represented given the specific form of matrix $G_{11}$ \eqref{G11}.
+This expression can be simplified and represented given the specific form of matrix $G_{11}$ \eqref{G11}.
 For the purpose of simplicity we present $\Rnum$ in the manuscript in terms of expressions $C$, $C1$ and $C2$, specified in \eqref{eq:C12}.
@@
 This provides a base to linearly approximate $\Rnum$ when $\rho$ is close to 0, and use this approximation to analyze the behaviour of $\Rnum$ with respect to $\omega$ (see section \ref{app:omega}).
-In the next section we provide an equivalent representation of $\Rnum$ to provide a ground to prove that more testing intensity decreases $\Rnum$ for a general rane of parameters.
+In the next section we provide an equivalent representation of $\Rnum$ to provide a ground to prove that more testing intensity decreases $\Rnum$ for a general range of parameters.
 
@@
 Mathematically we would like to verify the sign of $\pder[\Rnum]{\theta\_w}$ and $\pder[\Rnum]{\theta\_c}$.
-We start with simpliying $\Delta$ \eqref{eq:del4} by factoring $\theta\_w$ and $\theta\_c$ in Eq.\,\eqref{eq:del4}.
+We start with simplifying $\Delta$ \eqref{eq:del4} by factoring $\theta\_w$ and $\theta\_c$ in Eq.\,\eqref{eq:del4}.
 Thus, $\Delta$ can be rewritten as
@@
 $$1/C\,(\gamma\,S\_u^*\,+\omega(1-\theta\_w\,S\_u^*))\omega\,p_I\,\hat F_I.$$
-This results in increasing $\Delta$, thus decreasing $\Rnum$ with respect to $\theta\_c$, that is $\pder[\Rnum]{\theta\_c}\leq 0$. Note that $C$ is independant of $\theta\_c$ and $\theta\_w$.
+This results in increasing $\Delta$, thus decreasing $\Rnum$ with respect to $\theta\_c$, that is $\pder[\Rnum]{\theta\_c}\leq 0$. Note that $C$ is independent of $\theta\_c$ and $\theta\_w$.
 
@@
 \end{equation}
-We show that the feasible range of $\theta\_w$ lies btween 0 and the vertex of this parabola where the parabola is increasing in $\theta\_w$, and so does $\Delta$ which results in inferring $\pder[\Rnum]{\theta\_w}\leq 0$.
+We show that the feasible range of $\theta\_w$ lies between 0 and the vertex of this parabola where the parabola is increasing in $\theta\_w$, and so does $\Delta$ which results in inferring $\pder[\Rnum]{\theta\_w}\leq 0$.
 It is enough to show that partial derivative of the expression \eqref{eq:del4_thetaw} with respect to $\theta\_w$ at $\theta\_w=1$ is non-negative. It follows that
@@
 We use the derived expressions for $S\_u^*$ and $S\_n^*$, given by Eqs.\,\eqref{dfe}, in $\Delta$ \eqref{eq:del4}.
-Also, we define $\phi = \hat F_S = \frac{\rho \omega}{\omega-\rho}$, to reparametrize $\rho$.
+Also, we define $\phi = \hat F_S = \frac{\rho \omega}{\omega-\rho}$, to reparameterize $\rho$.
 This is mainly to avoid singularity in $\hat F_I$ \eqref{eq:fi}, when testing intensity $\rho$ is very close to the rate of test return $\omega$.
-Thus, $\rho$ is reparametrized as
+Thus, $\rho$ is reparameterized as
 \begin{equation}
@@
 \end{equation}
-This one-to-one monotonic reparametrization enables us to simplify the mathematical expressions and explore the simpler $\pder[\Delta]{\phi}$ instead of the complicated $\pder[\Delta]{\rho}$.
+This one-to-one monotonic reparameterization enables us to simplify the mathematical expressions and explore the simpler $\pder[\Delta]{\phi}$ instead of the complicated $\pder[\Delta]{\rho}$.
 The derivative is
@@
 
-An example of this counter-vailing effect of $\phi$, and consequently $\rho$, on $\Rnum$ occures when $\theta\_w=0$ and $\theta\_c=1$.
-This is illustraited in the top-right panel of the \fref{pan2} panel (b), where the strength of isolation for awaiting people is the least, but the most for the confirmed cases.
+An example of this counter-vailing effect of $\phi$, and consequently $\rho$, on $\Rnum$ occurs when $\theta\_w=0$ and $\theta\_c=1$.
+This is illustrated in the top-right panel of the \fref{pan2} panel (b), where the strength of isolation for awaiting people is the least, but the most for the confirmed cases.
 In this case, simplifying $a_3$ in Eq.~\eqref{eq:abcd2} gives $$a_3=\frac{w_I}{w_S}\,\omega\,\gamma\,p_I((\omega+\gamma)-\frac{w_I}{w_S}(\omega\,p_I+\gamma)).$$
-Specifically, in the case of targeted testing which is identified by $\frac{w_I}{w_S}> 1$, and using a perfect sensative test, thus $p_I=1$, there exists a range for $\rho$ over which $\pder[\Rnum]{\rho}\leq\,0$.
+Specifically, in the case of targeted testing which is identified by $\frac{w_I}{w_S}> 1$, and using a perfect sensitive test, thus $p_I=1$, there exists a range for $\rho$ over which $\pder[\Rnum]{\rho}\leq\,0$.
 Note that $\rho$ and $\omega$ have a similar mechanism in delaying people to get into $I\_c$, thus we would expect to see the non-trivial counter-vailing effect of these two parameters on $\Rnum$.
@@
 We use
-the linearization of $\Rnum$ around $\rho=0$ to show that there a non-monotonic relationship between $\Rnum$  and $\omega$. The taylor expansion of $\Delta$ at $\rho=0$ is
+the linearization of $\Rnum$ around $\rho=0$ to show that there a non-monotonic relationship between $\Rnum$  and $\omega$. The Taylor expansion of $\Delta$ at $\rho=0$ is
 \begin{equation}

### 2021-05-17 SIR_manuscript.tex
continued small changes

@@
 The observed dynamics of the \covid epidemic are driven by both epidemiological processes (infection and recovery) and testing processes (testing and test reporting).
-In addition to shaping epidemic observations (via case reports), testing processes can also affect epidemiological dynamics.
+In addition to shaping epidemic observations (via case reports), testing processes also affect epidemiological dynamics.
 In particular, individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may also do so (possibly to a lesser extent).
@@
 
-If testing influences behavior, then epidemic dynamics will depend on patterns of who gets tested.
-The impacts of testing will depend on intensity (tests performed per day), and on how strongly testing is focused on people who are infectious.
+If testing influences behavior, then epidemic dynamics will depend on who gets tested.
+The impacts of testing will depend both on testing intensity (tests performed per day) and on how strongly testing is focused on people who are infectious.
 This level of focus depends in turn on the purpose and design of testing programs. 
@@
 
-When testing is used primarily for diagnosis it will focus on people with infection-like symptoms; thus the relative testing weights for infected people will depend on the relative probability of infected people having symptoms.
-For \covid infection, the testing weights will depend on the relative asymptomatic infections, time spent pre-symptomatic vs.\ symptomatic infections -- and also the incidence of \covid-like symptoms among people in the population \emph{not} infected with \covid.
-Testing for epidemic control will focus on people who are known to have been in contact with known infected cases; in this case the testing weights for infected vs.\ uninfected people will depend on the probability of infection given contact, as well as the thoroughness and effectiveness of the system for identifying suspicious contacts.
+When testing is used primarily for diagnosis it focuses on people with infection-like symptoms; thus the relative testing weights for infected people will depend on the relative probability of infected people having symptoms.
+For \covid infection, the testing weights will depend on the proportion of asymptomatic infections, the time spent pre-symptomatic vs.\ symptomatic during the course of an infection, and on the incidence of \covid-like symptoms among people in the population \emph{not} infected with \covid.
+Testing for epidemic control focuses on known contacts of infected people; in this case the testing weights for infected vs.\ uninfected people will depend on the probability of infection given contact, as well as the effectiveness of the system for identifying suspicious contacts.
 
-The main interest from the epidemiological point of view is to know whether the number of infected individuals goes through an exponential growth phase, following the introduction of an infection in a totally susceptible population, before the disease becomes extinct.
-This is determined by studying the basic reproduction number $\Rnum$.
-It is defined as the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}.
+Our main interest from the epidemiological point of view is to know whether an epidemic will initially grows exponentially following its introduction in a susceptible population.
+In mathematical epidemiology, this characteristic is determined by studying the basic reproduction number $\Rnum$, defined as the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}.
 In the early stages of an epidemic the number of infected individuals is expected to grow exponentially over time when $\Rnum>1$, and to decline over time when $\Rnum<1$. 
 Although the value of $\Rnum$ cannot completely characterize the dynamics of even the simplest epidemic model
-\citep{shaw2021what}, it does give a simple and widely accepted index for the difficulty of control, as well as some indication of the likely final size of an epidemic \citep{ma2006generality}.
+\citep{shaw2021what}, it does give a simple and widely accepted index for the difficulty of control, as well as an indication of the likely final size of an epidemic \citep{ma2006generality}.
 
@@
 Here, we studied the the effect of testing intensity, rate of test return and the isolation efficacy in reduction of the probability of transmission on the epidemic dynamics when different levels of testing ``focus'' (from random to highly targeted) are in place.
-Our model provides insights to the sensitivity of the epidemic dynamics, through $\Rnum$, with respect to the underlying testing and isolation parameters.\ali{edited in response to David comments.}
+% Our model provides insights to the sensitivity of the epidemic dynamics, through $\Rnum$, with respect to the underlying testing and isolation parameters.
+%% BMB: last sentence seems to belong in conclusions
 
@@
 
-We developed a deterministic model, Eqs.\,\eqref{model}, which groups individuals based on disease status and testing status.
-Disease states include Susceptible, Infectious and Recovered (thus this is an SIR-type model), and testing status categorizes people as \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive} (\fref{fig:flowchart}).
+We developed a deterministic model, Eqs.~\eqref{model}, which groups individuals based on disease status and testing status.
+Disease states include Susceptible, Infectious and Recovered; testing status categorizes people as \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive} (\fref{fig:flowchart}).
 Symbolically, the testing status of an individual in the disease compartment $X$, where $X \in \{S,I,R\}$, is reflected in the subscript, namely $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
@@
 Thus, we can specify different levels of testing “focus” from random to highly targeted.
-For example, $w_I/w_S=3$ means that infected individuals are tested at tree times the \percap rate of susceptible individuals.
+For example, $w_I/w_S=3$ means that infected individuals are tested at three times the \percap rate of susceptible individuals.
 
@@
 The analytical calculation of the next-generation matrix and simplifying the expression of $\Rnum$ was carried out in Maple \citep{maple14}.
-We used R \citep{r} for simulations, in particular,  for plotting the contours of $\Delta$ \eqref{eq:tel4} over a range of selected set of parameters of interest.
+We used R \citep{r} for simulations, in particular,  for plotting the contours of $\Delta$ \eqref{eq:tel4} \bmb{broken ref} over a range of selected set of parameters of interest.
 The rest of model parameters kept fixed at sensible values (see Table~\ref{tab:params} for the parameter values).
@@
 \end{subequations}
-
+%
 where parameters are specified in Table \ref{tab:params}.
@@
 % %%%%%%%
-\subsection{More testing intensity may decreases $\Rnum$}\label{app:rho}
+\subsection{More testing intensity may decrease $\Rnum$}\label{app:rho}
 
-This is to provide a mathematical materials to prove that $\pro \Delta$ can be positive or negative, with $\Delta$ defined in Eq.\,\eqref{eq:C12}, and thus $\pro \Rnum < 0$, where $\Rnum$ is the basic reproduction number, given in Eq.\,\eqref{R0}.
+This section shows that $\pro \Delta$ can be positive or negative, with $\Delta$ defined in Eq.\,\eqref{eq:C12}, and thus $\pro \Rnum < 0$, where $\Rnum$ is the basic reproduction number, given in Eq.\,\eqref{R0}.
 We can rewrite matrix $G_{11}$ in \eqref{G11} in the following form to simplify the calculations.

### 2021-05-18 note/cohort.tex
more math to the cohort work

@@
 \linenumbers
-
+%%%%%%%%%%%%%%%%%%%%
 \section{Vision}
@@
 
-\item 
+\item Having the cohort framework enables one to study the strength-like and speed-like interventions.
+For example in our SIR model with testing and isolation, testing susciptables is a strength-like intervention and testing the infecteds' is a speed-like intervention.
+\todo{more context is required here.}
 \end{enumerate}
 
-\section{Introduction}
-Notation; we use $I'$ for the cohort-framework of $dI/d\tau$ and $\dot I$ for the compartment-framework of $dI/dt$.
+%%%%%%%%%%%%%%%%%%%%
+\section{Math foundation}
+Notation; we use $I'$ for the cohort-framework of $dI/d\tau$, where $\tau$ is in the infection-time scale and $\dot I$ for the compartment-framework of $dI/dt$.
 
-A simple example of this transition is the cohort analogy of a simple SIR compartmental model where $\dot I=\beta S I/N-\gamma I$. The cohort-framework is via the cohort Eq. with the following steps
+The matrix form of the cohort framework consists of the following 3 steps.
+\begin{enumerate}[step 1]
+\item
+write the cohort Eq. $I'=-V I$, with the initial condition $I(0)$, where the n-by-n matrix $V$ is the flow matrix of leaving infected compartments in \cite{van2002reproduction}.
+
+\item 
+finding the intrinsic infectiousness kernel by integrating the cohort Eq. and solving for its time evolution, thus
+the solution would be $I(\tau) = \exp(-V\tau) I(0)$. The kernel will be 
+$K(\tau) = F I(\tau)$, where $F$ is the matrix of new infections in the compartment framework.
+Note that the n-by-n matrix $\exp(-V\tau)$ is the probability of being infected and stay infectious at time $\tau$, where
+$\exp(-V\tau)=\sum_{n=0}^\infty \tau^n (-V)^n/n!$.
+
+\item
+Calculating $R$ by integrating kernel $K(\tau)$.
+\end{enumerate}
+
+Note that in the comartment framework, steps 2 and 3 are combined.
+%%%%%%%%%%%%%%%%%%%%
+\section{examples}
+
+{\bf Example 1, the simple SIR;}
+The cohort analogy of a simple SIR compartmental model where $\dot I=\beta S I/N-\gamma I$. The cohort framework is via the cohort Eq. with the following steps
 step 1: write the cohort Eq. $I'=-\gamma I$ with the initial condition $I(0)=1$,
+
 step 2: finding the intrinsic infectiousness kernel by integrating the cohort Eq. and solving for its time evolution, thus
 $k(\tau)=\beta \exp(-\gamma \tau)$.
+
 step 3: $R$ will be the integration of the kernel. That is $R=\beta/\gamma$.
-Note that in the comartment framework, steps 2 and 3 are combined.
 
-In Watmough 2002's work matrix form of the I compartment $I'=(-V)I$, where I is n-by-1 and $V$ is n-by-1.
+{\bf Example 2, the SIR model with testing;}
+In the context of our SIR model with testing, $I^T=[I_u,I_n,I_p,I_c]$ where $T$ is the transposed operator.
+And the solution is given by $I(\tau) = \exp(-V\tau) I(0)$ where $I(0)=[1,0,0,0]$.
 
-In the context of our SIR model with testing, $I^T=[I_u,I_n,I_p,I_c]$ where $T$ is the transposed operator.
-And the solution is given by $I(\tau) = \exp(-V\tau) I(0)$, where $\tau$ is in the infection-time scale.
-$K(\tau)=\Rnot g(\tau)= F I(\tau)$,
-where $F$ is n-by-n matrix of new infections, $I(0)=[1,0,0,0]$.
-
-$\exp(-V\tau)$ is the probability of being infected and stay infectious at time $\tau$.
-
-$\exp(-V\tau)=\sum_{i=0}^\infty \tau^n (-V)^n/n!$.
+%%%%%%%%%%%%%%%%%%%%
 

### 2021-05-18 cohort.tex
make rull for cohort work

--- 2021-05-18 note/cohort.tex
@@
 \item
-write the cohort Eq. $I'=-V I$, with the initial condition $I(0)$, where the n-by-n matrix $V$ is the flow matrix of leaving infected compartments in \cite{van2002reproduction}.
+write the cohort Eq. 
+\begin{equation}
+\label{eq:cohort}
+I'=-V I,
+\end{equation}
+ with the proper initial condition $I(0)$, where the n-by-n matrix $V$ is the flow matrix of leaving infected compartments in \cite{van2002reproduction}.
 
@@
 finding the intrinsic infectiousness kernel by integrating the cohort Eq. and solving for its time evolution, thus
-the solution would be $I(\tau) = \exp(-V\tau) I(0)$. The kernel will be 
-$K(\tau) = F I(\tau)$, where $F$ is the matrix of new infections in the compartment framework.
+the solution would be 
+\begin{equation}
+\label{eq:I}
+I(\tau) = \exp(-V\tau) I(0).
+\end{equation}
+The kernel will be 
+\begin{equation}
+\label{eq:kernel}
+K(\tau) = F I(\tau), 
+\end{equation}
+where $F$ is the matrix of new infections in the compartment framework.
 Note that the n-by-n matrix $\exp(-V\tau)$ is the probability of being infected and stay infectious at time $\tau$, where
@@
 
-\bibliography{../SIRlibrary}
+\bibliography{SIRlibrary}
 \end{document}

### 2021-05-19 SIR_manuscript.tex
still working

@@
 
+\bmb{abstract needs another editing pass, skipping for now}
+
 The effect of testing processes, including testing and test reporting, on epidemic dynamics, involving infection and recovery, can be studied at the individual level or the community level (e.g., nursing homes, long-term-care facilities, etc.).
@@
 
-The observed dynamics of the \covid epidemic are driven by both epidemiological processes (infection and recovery) and testing processes (testing and test reporting).
-In addition to shaping epidemic observations (via case reports), testing processes also affect epidemiological dynamics.
-In particular, individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may also do so (possibly to a lesser extent).
+The observed dynamics of the \covid epidemic have been driven both by epidemiological processes (infection and recovery) and by testing processes (testing and test reporting).
+In addition to shaping epidemic observations (via case reports), testing processes also alter epidemiological dynamics.
+Because individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may also do so (possibly to a lesser extent), testing will generally increase the number of people who are isolating and hence reduce epidemic growth rates.
 We developed a mechanistic model that incorporates epidemic processes and testing in order to explore the effects of testing and isolation on epidemic dynamics.
@@
 When testing is done for the purposes of disease surveillance \citep{foddai2020base}
-tests should be assigned randomly (or using a stratified random design) across the population in
+tests are typically assigned randomly (or using a stratified random design) across the population in
 order to make an unbiased assessment of population prevalence.
@@
 
-When testing is used primarily for diagnosis it focuses on people with infection-like symptoms; thus the relative testing weights for infected people will depend on the relative probability of infected people having symptoms.
+Diagnostic testing focuses on people with infection-like symptoms; thus the relative testing weights for infected people will depend on the relative probability of infected people having symptoms.
 For \covid infection, the testing weights will depend on the proportion of asymptomatic infections, the time spent pre-symptomatic vs.\ symptomatic during the course of an infection, and on the incidence of \covid-like symptoms among people in the population \emph{not} infected with \covid.
@@
 
-Our main interest from the epidemiological point of view is to know whether an epidemic will initially grows exponentially following its introduction in a susceptible population.
-In mathematical epidemiology, this characteristic is determined by studying the basic reproduction number $\Rnum$, defined as the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}.
-In the early stages of an epidemic the number of infected individuals is expected to grow exponentially over time when $\Rnum>1$, and to decline over time when $\Rnum<1$. 
+As epidemiologists, we want to know whether an epidemic will initially grow exponentially following its introduction in a susceptible population.
+In mathematical epidemiology, we determine this outcome by studying the basic reproduction number $\Rnum$, defined as the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}.
+In the early stages of an epidemic the number of infected individuals is expected to grow exponentially over time when $\Rnum>1$.
+%% BMB left out the converse; don't think we have to be super-precise here
 Although the value of $\Rnum$ cannot completely characterize the dynamics of even the simplest epidemic model
@@
 This model provides a sensible platform to link the modeling of epidemic and testing components and study their interaction.
-Here, we studied the the effect of testing intensity, rate of test return and the isolation efficacy in reduction of the probability of transmission on the epidemic dynamics when different levels of testing ``focus'' (from random to highly targeted) are in place.
+Here, we studied the the effect of testing intensity, rate of test return and the isolation efficacy in reduction of the probability of transmission on the epidemic dynamics when different levels of testing focus (from random to highly targeted) are in place.
 % Our model provides insights to the sensitivity of the epidemic dynamics, through $\Rnum$, with respect to the underlying testing and isolation parameters.
@@
 Disease states include Susceptible, Infectious and Recovered; testing status categorizes people as \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive} (\fref{fig:flowchart}).
-Symbolically, the testing status of an individual in the disease compartment $X$, where $X \in \{S,I,R\}$, is reflected in the subscript, namely $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
-Note that the top-to-bottom order of the testing-based compartments of each disease-based compartment $X$ in the flowchart (\fref{fig:flowchart}) and the model equations \eqref{model} should match.
-However, we switched  $X\_u$ and $X\_n$ in the flowchart (\fref{fig:flowchart}) for the sake of tidiness of the flowchart.
-Further, two `accumulator' compartments, $N$ and $P$, were also incorporated in the model in order to collect cumulative reported negative or positive tests.
-The model equations and details of calculation of the basic reproduction number $\Rnum$ are presented in the appendix (see Sec.
+The testing status of an individual in  a given disease compartment $X$ (where $X \in \{S,I,R\}$) is denoted by a subscript, namely $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
+Further, two `accumulator' compartments, $N$ and $P$, are included in order to collect cumulative reported negative or positive tests.
+The model equations and details of calculation of the basic reproduction number $\Rnum$ are presented in the appendix (Sec.
 \ref{app:R0}).
@@
 \caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model, \ref{model}.
-Here, the disease-based status of a compartment $X$, where $X \in \{S,I,R\}$, is combined with the testing-based status including $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
-Also,  $\Lambda$ is the force of infection with definition in Eq.~\eqref{Lambda}, $\gamma$ is the recovery rate, $\omega$ is the rate of test return, $F_X$ (defined in Eq.
-\eqref{F}) and $p_X$ represent the \percap testing rate and the probability of positive tests, respectively, for compartment $X$.
+The disease-based status of a compartment $X$ ($X \in \{S,I,R\}$) is combined with the testing status including $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
+The force of infection is denoted by $\Lambda$ (Eq.~\eqref{Lambda}); $\gamma$ is the recovery rate; $\omega$ is the rate of test return; and $F_X$ (Eq.
+\eqref{F}) and $p_X$ represent the \percap testing rate and the sensitivity (probability of positive tests), respectively, for compartment $X$.
 For further description of the parameters see \cref{tab:params}.
+Note that there is a slight mismatch in the top-to-bottom order of the testing-based compartments of each disease-based compartment $X$ between this flowchart and the model equations \eqref{model}; here we have switched  $X\_u$ and $X\_n$ for visual clarity.
 \label{fig:flowchart}}
@@
 
-Table~\ref{tab:params} defines the model parameters, which are generally straightforward
-\percap flows between compartments, or modifiers to these flow rates.
-The novel component of the model comes in through the compartment-specific relative testing weights $w_S$, $w_I$ and $w_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively.
-Thus, we can specify different levels of testing “focus” from random to highly targeted.
+Table~\ref{tab:params} defines the model parameters, which are generally \percap flows between compartments, or modifiers to these flow rates.
+The novel component of the model lies in the compartment-specific relative testing weights $w_S$, $w_I$ and $w_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively.
+Thus, we can specify different levels of testing focus from random (all weights equal) to highly targeted (higher weights in more intensively tested compartments).
 For example, $w_I/w_S=3$ means that infected individuals are tested at three times the \percap rate of susceptible individuals.
 
-In order to link to more applied models, we constructed this model so we could specify the total \percap testing rate.
-We do this by defining the weighted size of the testing pool $W = w_S S\_u + w_I I\_u + w_R R\_u$, and calculating a scaling parameter for testing as:
+In order to allow parameterization of the model by the total (overall) \percap testing rate, we define the weighted size of the testing pool $W = w_S S\_u + w_I I\_u + w_R R\_u$, and calculate a scaling parameter for testing as:
 \begin{equation}
@@
 \end{equation}
-where $\rho$ is the \percap testing intensity for population and defined as the number of daily tests taken in a population of size $N_0$.
+where $\rho$ is the \percap testing intensity for the population, defined as the number of daily tests administered in a population of size $N_0$.
 Thus, the \percap testing rate for compartment $X$ is 
@@
 \end{equation}
-For a high-sensitivity test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus unavailable for further testing.
-Over the course of the epidemic, a fixed testing rate as specified in \eqref{sigma} can (if large enough) exhaust the pool of people available for testing, leading to a singularity when no one is left untested.
-Although this phenomenon does not affect our analysis of $\Rnum$, it can affect the temporal dynamics (we discuss an adjustment to the model that solves this problem in the appendix).
+For a highly sensitive test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus unavailable for further testing.
+Over the course of the epidemic, a sufficiently large fixed testing rate as specified in \eqref{sigma} can exhaust the pool of people available for testing, leading to a singularity when too few people are left untested to support the specified rate.
+Although this phenomenon does not affect our analysis of $\Rnum$, it can affect the temporal dynamics (in the appendix we present an adjustment to the model that solves this problem).
 
-The classical SIR model is based on the following implicit assumptions; well-mixed population, homogeneity of the population (i.e., all individuals are equally susceptible and equally infectious for the same length of time when infected), exponentially distributed duration of infection and large population size (see, e.g., \cite{keeling2011modeling}).
+The classical SIR model assumes a well-mixed population; homogeneity of the population (i.e., all individuals are equally susceptible and equally infectious with the same recovery rate when infected); exponentially distributed duration of infection; and large population size \citep{keeling2011modeling}.
 In addition to these standard assumptions, our model assumes:
 \begin{enumerate}[(i)]
-\item there is a single force of infection (new cases per unit time), $\Lambda$, defined as follows
+\item there is a single force of infection (new cases per unit time), $\Lambda$, defined as
   \begin{equation}
@@
   \end{equation}
-  across all susceptible pools with transmission rate $\beta$ and isolation efficacy in reduction of the probability of transmission for three testing-based compartments ``waiting'' and \emph{confirmed positive} individuals, $\theta\_w$ and $\theta\_c$ respectively, (see Table~\ref{tab:params} for further details),
-\item the efficacy of isolation is higher in confirmed individuals relative to the individuals awaiting test results. That is,
+  with transmission rate $\beta$; $\theta\_w$ is the isolation efficacy (reduction of the probability of transmission) for individuals waiting for test results, while $\theta\_c$ is the isolation efficacy for individuals who have received a positive test (``confirmed positive'': Table~\ref{tab:params}).
+Susceptible individuals who are waiting for test results experience an additional transmission factor of $1-\theta\_w$ (\fref{fig:flowchart}).
+\item confirmed-positive individuals isolate more effectively than those awaiting test results, i.e.
 $$0\leq\theta\_w\leq\theta\_c\leq 1.$$ 
 \end{enumerate}
-Thus, for instance when the awaiting people follow the isolation perfectly, $\theta\_w$ is closer to 1, while when they less follow the isolation, $\theta\_w$ is closer to 0.
-For the simplicity of analysis, we also assume a perfectly specific test ($p_S=0$).
-This last assumption combined with the assumption that no susceptible individual is in waiting-for-\emph{positive} or \emph{confirmed positive} compartments, i.e., $S\_p(0)=S\_c(0)=0$, reduces the model to 10 equations where equations c and d in model \eqref{model} are eliminated.
+For simplicity we assume that tests are perfectly \emph{specific} --- uninfected individuals never test positive ($p_s=0$).
+Thus there are no waiting-for-positive or confirmed-positive susceptible individuals (we assume that these states are also absent from the initial conditions, i.e.
+$S\_p(0)=S\_c(0)=0$), which reduces the number of model states from 12 to 10.
 
@@
 \label{eq:fi}
-\hat F_I = (\omega\rho/(\omega-\rho))w_I/w_S.
+\hat F_I = (\omega\rho/(\omega-\rho))w_I/w_S \quad .
 \end{equation}
 
-The basic reproduction number, $\Rnum$, was calculated by using the next-generation matrix method developed by \cite{van2002reproduction}. We present $\Rnum$ in the following form.
+The basic reproduction number, $\Rnum$, was calculated by using the next-generation matrix method \citep{van2002reproduction}. We present $\Rnum$ in the following form.
 \begin{equation}
@@
 \end{equation}
-where the term $\frac{\beta}{\gamma}$ is the classical basic reproduction number for a SIR model without testing and isolation (see, e.g., \cite{keeling2011modeling}), and the positive quantity $\Delta$ is the ``effectiveness of control'' effectiveness of control due to testing and isolation processes defined as follows.
+where the term $\frac{\beta}{\gamma}$ is the classical basic reproduction number for a SIR model without testing and isolation \citep{keeling2011modeling}, and the positive quantity $\Delta$ is the effectiveness of control due to testing and isolation processes defined as
 \begin{equation}
-\label{eq:del4}
+  \label{eq:del4a}
   \Delta= \frac{1}{C\,N_0}\big(C_1\,S\_u^*+(C_2(1-\theta\_w)+C\,\theta\_w)\,S\_n^*\big),
 \end{equation}
+\bmb{super-fussy, but: I don't think we should introduce explicit thinspaces everywhere; why not let
+  TeX do the typesetting itself? Do we really prefer more compact equations, and if so is there a package
+  that will just change the appropriate settings globally?}
 where
@@
 \label{eq:C12}
-C =& (\omega+\gamma) \Big(\gamma(\omega+\gamma)+(\gamma+\omega p_I)\hat F_I\Big),\\
+C_{\phantom 1} =& (\omega+\gamma) \Big(\gamma(\omega+\gamma)+(\gamma+\omega p_I)\hat F_I\Big),\\
 C_1 =& (\omega+\gamma)(\theta\_w \, \gamma+\theta\_c\, \omega\, p_I) \hat F_I,\\
@@
 \end{align}
+\bmb{I don't like the alignment of the equations: too little space after =. I realize that's incredibly picky.}
 Further details of derivation of $\Rnum$ are provided in appendix, Sec.\,\ref{app:R0}. 
-The explicit formula of $\Rnum$, enables us to study the effect of parameters of interest, on $\Rnum$ mathematically and through numerical simulation.
-These parameters, i.e., parameters that could be manipulated by public-health policy include the isolation efficacy, $\theta\_c$ and $\theta\_w$, \percap testing intensity, $\rho$, and the rate of test return, $\omega$.
-In particular, we look at the partial derivative of $\Delta$ with respect to these parameters (see details in appendix, Sec.\,\ref{app:rho}).
-It is notable that the derivative of $\Delta$ with respect to $\theta\_w$, $\theta\_c$ and $\rho$ was derived explicitly and for the whole domain of the parameter space.
+The explicit formula of $\Rnum$ enables us to study the effects of testing and isolation parameters on $\Rnum$ both analytically and via numerical solutions.
+We are specifically interested in parameters that could be manipulated by public health policy: isolation efficacy, $\theta\_c$ and $\theta\_w$; \percap testing intensity, $\rho$; and the rate of test return, $\omega$.
+In particular, we look at the partial derivatives of $\Delta$ with respect to these parameters (Sec.\,\ref{app:rho}).
+We derived general expressions for these derivatives.
 However, we analyzed the effect of $\omega$ on $\Rnum$ for the case when testing intensity is very small relative to the population size, $N_0$.
 Specifically, we used the Taylor approximation of $\Delta$ at $\rho=0$ to illustrate the unexpected non-monotonic relationship between $\Rnum$ and $\omega$, and also to avoid unnecessary complicated computations for making this point.
+\bmb{clarify/check}
 
-The analytical calculation of the next-generation matrix and simplifying the expression of $\Rnum$ was carried out in Maple \citep{maple14}.
-We used R \citep{r} for simulations, in particular,  for plotting the contours of $\Delta$ \eqref{eq:tel4} \bmb{broken ref} over a range of selected set of parameters of interest.
-The rest of model parameters kept fixed at sensible values (see Table~\ref{tab:params} for the parameter values).
-Note that all plots are illustrated in the scale of the mean test return time $1/\omega$ (day) and testing intensity $\rho$ (1\//day per 1000).
-Results are shown in \fref{pan} and \fref{pan2} with two sets of panels; panel set (a) represents the random testing, identified by $w_S=w_I=w_R=1$, and panel set (b) represents targeted testing, identified by $w_S=0.3$ and $w_I=w_R=1$.
-Each panel set partitions a plot into a matrix of panels, namely facets, in which rows are specified by a value of $\theta\_w$ and columns are specified by a value of $\theta\_c$.
-In our simulation, $\theta\_w$ and $\theta\_c$ vary in a discrete scale from 0 to 1 with 0.25 increment.
-Note that we are only plotting facets with $\theta\_w\leq\theta\_c$ which is an assumption of our model.
+The analytical calculation of the next-generation matrix and simplification of the $\Rnum$ expression was carried out in Maple \citep{maple14}.
+We used R \citep{r} for numerical solutions, in particular for plotting the contours of $\Delta$ (Eq.~\eqref{NOTSURE}) \bmb{not sure what the right ref.
+is.
+This was previously ``eq:tel4'', which doesn't exist.
+Should it have been ``eq:del4''?} over a range of parameters.
+We computed the values and contours of Delta at both low (\fref{pan}) and high (\fref{pan2}) testing intensities, and for both random testing ($w_S=w_I=w_R=1$) and targeted testing ($w_S=0.3$; $w_I=w_R=1$).
 
-To illustrate the changes in $\Rnum$ with respect to \percap testing intensity $\rho$, two sets of plots are presented. 
-\fref{pan} reflects the changes in $\Rnum$ when $\rho$ is small relative to the population size.
-Specifically, $\rho \in [0,0.013]$, and the test return rate $\omega\in [1/12,2]$.
-This is a more realistic scenario of testing as we have observed in \covid pandemic.
-That is, 1\% of a population are tested per day at maximum capacity, so in a population of size $N_0=10^6$, a maximum of 10000 individuals will be tested per day.
-In order to illustrate the non-monotonic changes in $\Rnum$ with respect to $\rho$,
+The low-testing case \fref{pan} reflects the case where $rho$ is small relative to the population size.
+Specifically, $\rho \in [0,0.013]$, and test return rate $\omega\in [1/12,2]$.
+This testing intensity reflects realistic testing rates during the \covid pandemic.
+For example, since 1\% of a population are tested per day at maximum capacity, giving a maximum of 10000 tests/day in a population of size $N_0=10^6$.
+The less realistic high-testing case \fref{pan2} is included to highlight the occurrence of non-monotonic changes in $\Rnum$ with respect to $\rho$.
+\bmb{stopped here}
 \fref{pan2} with maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,1/5)$, and the test return rate $\omega\in [1/5,2]$. 
@@
 
-Examining our formula for $\Rnum$ \eqref{R0} gives the following results. 
-See the appendix, Sec.\,\ref{app:rho}, Sec.\,\ref{app:omega} and Sec.\,\ref{app:w} for details.
+For a variety of changes in testing and isolation parameters, we can use the formula for $\Rnum$ \eqref{R0} to confirm monotonic changes in the expected, intuitive direction (i.e.
+the partial derivative of $\Delta$ has a consistent sign: Sections~\ref{app:rho}, \ref{app:omega}, \ref{app:w}).
 
@@
 
-\item \label{p1:eta} Increasing the isolation efficacy parameters for tested and confirmed individuals decreases \Rnum;
-
-\item \label{p1:rho} Higher testing intensity $\rho$ always decreases $\Rnum$ \textbf{if}: 
+\item \label{p1:eta} Increasing isolation efficacy for waiting ($\theta\_w$) and confirmed-positive ($\theta\_c$) individuals always decreases \Rnum;
+\item \label{p1:rho} Higher testing intensity $\rho$ decreases $\Rnum$ if
 %% \begin{enumerate} \item
-testing is random; \textbf{or}
-%% \item
-$\rho$ is small.
+testing is random (all $w_X$ equal) or testing intensity ($\rho$) is small.
 %% \end{enumerate}
-
-\item \label{p1:omega} A higher rate of test return always decreases \Rnum, \textbf{if} $\theta_w=0$.
-
-\item \label{p1:w} Increasing testing “focus” $\frac{w_I}{w_S}$ from random, $w_S=w_I$, to targeted, $w_S<w_I$, always decreases \Rnum.
-
+\item \label{p1:omega} Increasing the rate of test return ($\omega$) always decreases \Rnum if waiting individuals do not isolate ($\theta_w=0$).
+\item \label{p1:w} Increasing testing focus from random ($w_S=w_I$) to targeted  ($w_S<w_I$) always decreases \Rnum.
 \end{enumerate}
@@
 {\bf A comparison of effectiveness control on $\Rnum$ at low level of \percap testing intensity $\rho$.}
-We numerically evaluate the effectiveness control parameter, $\Delta$ (Eq.~\ref{eq:del4}), with respect to testing and isolation.
-We use the following parameters (see Table~\ref{tab:params} for a brief description):
-$N_0=1 \times 10^6$, $\omega \in [1/12,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,0.013]$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for the full effect of isolation on the transmission probability, $\beta=0.339$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
-Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
-The black dotted line in each panel represents the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ corresponding to the threshold of $\Rnum=1$.
+Numerical evaluation of the effectiveness control parameter, $\Delta$ (Eq.~\ref{eq:del4}), with respect to changes in testing and isolation. Parameter values (Table~\ref{tab:params}):
+$N_0=10^6$, $\beta=0.339$ day$^{-1}$ $\gamma= 1/3$ day$^{-1}$; $\omega \in [1/12,2]$ day$^{-1}$;  $\rho \in [0,0.013]$ day$^{-1}$; $\theta\_w$ and $\theta\_c$ vary between 0 (no effect of isolation) and 1 (complete elimination of transmission); $p_S=0$, $p_I=1$ and $p_R=0.5$.
+Contours of $\Delta$ are plotted for (a) random testing ($w_S=w_I=w_R=1$) and (b) targeted testing ($w_S=0.3$; $w_I=w_R=1$).
+Black dotted lines in each panel represent the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ such that $\Rnum=1$.
+\bmb{does $N_0$ enter the expression at all, i.e.
+can we omit it?
+Perhaps we can also include baseline $\Rnum$ and $r$ (or doubling time), i.e.
+$\beta/\gamma$ and $\beta-\gamma$.
+(Why did we pick these particular beta/gamma values?)}
 }
@@
 
-Numerical results are shown in Figs.~\ref{pan} and \ref{pan2}.
-When $\theta_w =0$ (top row of each panel), we see that shorter test-waiting times reduce \Rnum\ ($\Delta$ increases as we move to the left in each plot in this row).
+However, there are also two specific cases where $\Delta$ changes \emph{non-monotonically}, in counterintuitive directions, as a function of testing and isolation parameters.
+
+\bmb{stopped about here.
+I'm a little unclear about the distinction between the results above and below here.
+I assumed that results above were intuitive/expected and those below weren't.
+But the last point here (targeting testing increases $\Delta$) seems intuitive.
+Is it about things we can prove/can't prove?
+Do we know anything about $\Delta$ vs.
+targeting?}
+
+\begin{itemize}
+\item When waiting individuals isolate strongly (e.g.
+$\theta_w =0$, top row of \fref{pan}), increasing the rate at which tests are returned/shortening testing delays (i.e.
+increasing $\omega$) reduces \Rnum, as we might expect ($\Delta$ increases as we move to the left in each plot in this row).
 For non-zero $\Delta$, however, we mostly see the opposite effect: less waiting leads to more transmission.
@@
 
-We also see that greater test intensity $\rho$ increases the effectiveness of control $\Delta$ ($\Delta$ increases as we move up in each plot in this Figure).
+\item We also see that greater test intensity $\rho$ increases the effectiveness of control $\Delta$ ($\Delta$ increases as we move up in each plot in this Figure).
 Mathematically speaking, we did find a counter-vailing effect, but this can only be seen when we allow $\rho$ to take unrealistically large values, and only for weighted/targeted testing (see \fref{pan2}).
@@
 
-Lastly, we show that increasing the focus of testing from random to targeted, always increases $\Delta$, thus decreases \Rnum.
+\item Lastly, we show that increasing the focus of testing from random to targeted, always increases $\Delta$, thus decreases \Rnum.
 This can be seen in \fref{pan} and \fref{pan2}.
 In particular, we realize higher $\Delta$ levels, i.e., darker colors in corresponding plots\/facets as we go from panel~(a) to panel~(b) in the figures.
+\end{itemize}
 
@@
 \begin{equation*}
-% \label{eq:del4}
+\label{eq:del4}
 \Delta= \frac{1}{C}\big(C_1\,S\_u^*+(C_2(1-\theta\_w)+C\,\theta\_w)\,S\_n^*\big).

### 2021-05-19 cohort.tex
matrix examples added to the cohort work

@@
 \newcommand{\percap}{\emph{per capita}\xspace}
-\newcommand{\Rnot}{\ensuremath{\mathcal{R}_0}}
+\newcommand{\R}{\ensuremath{\mathcal{R}_0}}
 \newcommand{\covid}{COVID-19\xspace}
@@
 \newcommand\pder[2][]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder[x]{y}
+\renewcommand{\vec}[1]{\ensuremath{\mathbf{#1}}} %bold font for vector
 
@@
 The matrix form of the cohort framework consists of the following 3 steps.
-\begin{enumerate}[step 1]
+\begin{enumerate}[{\it step 1}]
 \item
@@
 \label{eq:cohort}
-I'=-V I,
+\vec I'=-V \vec I,
 \end{equation}
@@
 \label{eq:I}
-I(\tau) = \exp(-V\tau) I(0).
+\vec I(\tau) = \exp(-V\tau) \vec I(0).
 \end{equation}
@@
 \label{eq:kernel}
-K(\tau) = F I(\tau), 
+\vec K(\tau) = F \vec I(\tau), 
 \end{equation}
 where $F$ is the matrix of new infections in the compartment framework.
-Note that the n-by-n matrix $\exp(-V\tau)$ is the probability of being infected and stay infectious at time $\tau$, where
-$\exp(-V\tau)=\sum_{n=0}^\infty \tau^n (-V)^n/n!$.
+Note that the n-by-n matrix $\exp(-V\tau)$ is the probability of being infected and stay infectious at time $\tau$. 
+% where $\exp(-V\tau)=\sum_{n=0}^\infty \tau^n (-V)^n/n!$.
 
 \item
-Calculating $R$ by integrating kernel $K(\tau)$.
+Calculating $\vec R$ by integrating kernel $\vec K(\tau)$. Note that $\R$ would be the non-negative element of $\vec R$.
 \end{enumerate}
+Note that in the compartment framework \citep{van2002reproduction}, steps 2 and 3 are combined.
+%%%%%%%%%%%%%%%%%%%%
+\section{Examples}
 
-Note that in the comartment framework, steps 2 and 3 are combined.
-%%%%%%%%%%%%%%%%%%%%
-\section{examples}
+\subsection {Simple SIR}
+The cohort analogy of a simple SIR compartmental model where $\dot I=\beta S I/N-\gamma I$. The cohort framework is via the cohort Eq. with the following steps.
+{\it step 1;} write the cohort Eq. $I'=-\gamma I$ with the initial condition $I(0)=1$, 
+{\it step 2;} finding the intrinsic infectiousness kernel by integrating the cohort Eq. and solving for its time evolution, thus $k(\tau)=\beta \exp(-\gamma \tau)$.
+{\it step 3;} $\R$ will be the integration of the kernel. That is $\R=\beta/\gamma$.
 
-{\bf Example 1, the simple SIR;}
-The cohort analogy of a simple SIR compartmental model where $\dot I=\beta S I/N-\gamma I$. The cohort framework is via the cohort Eq. with the following steps
-step 1: write the cohort Eq. $I'=-\gamma I$ with the initial condition $I(0)=1$,
+\subsection{Simple SEIR model}
+The simplest form of a SEIR model without vital rates is 
+\begin{align}\label{mo:sier1}
+\dot S=-\beta\,S\,I/N,~~
+\dot E= \beta\,S\,I/N -\sigma\,E,~~
+\dot I= \sigma\,E -\gamma\,I,~~
+\dot R= \gamma\,I.
+\end{align}
+We know that $\R=\beta/\gamma$ from next generation method where
 
-step 2: finding the intrinsic infectiousness kernel by integrating the cohort Eq. and solving for its time evolution, thus
-$k(\tau)=\beta \exp(-\gamma \tau)$.
+\begin{align}
+\label{seir1FV}
+F = \beta \left[ \begin {array}{cc} 
+0 & 1 \\
+0 & 0
+ \end {array} \right],~~
+V=
+ \left[ \begin {array}{cc}
+ \sigma & 0 \\
+-\sigma & \gamma
+\end {array} \right],
+~~ \text{thus}~~
+V^{-1} = \frac{1}{\sigma\,\gamma}
+\left[ \begin {array}{cc}
+\gamma & 0 \\
+\sigma & \sigma
+\end {array} \right].
+\end{align}
+I the context of the cohort framework we show that the same $\R$ can be calculated by following the three steps.
 
-step 3: $R$ will be the integration of the kernel. That is $R=\beta/\gamma$.
+{\it Step 1;} The cohort equation with the initial condition (in matrix form) is 
+\begin{align}
+\vec I'=-V\,\vec I,~~ \text{where}~~ \vec I(\tau)=\left[ \begin {array}{c} E(\tau)\\I(\tau) \end {array} \right], ~~
+\vec I(0)=\left[ \begin {array}{c} 1\\0 \end {array} \right].
+\end{align}
+Solving for $I(\tau)$ by finding the eigenvalues and eigenvectors of the above linear system results in
+\begin{align}
+\vec I(\tau) = \left[ \begin {array}{c} E(\tau)\\I(\tau) \end {array} \right] =
+c_1 \exp(-\sigma\,\tau)\,\left[ \begin {array}{c} \gamma/\sigma-1\\1 \end {array} \right] 
+-c_1 \exp(-\gamma\,\tau)\,\left[ \begin {array}{c} 0\\1 \end {array} \right],
+\end{align}
+where $$c_1=\frac{\sigma}{\gamma-\sigma}.$$
 
-{\bf Example 2, the SIR model with testing;}
-In the context of our SIR model with testing, $I^T=[I_u,I_n,I_p,I_c]$ where $T$ is the transposed operator.
-And the solution is given by $I(\tau) = \exp(-V\tau) I(0)$ where $I(0)=[1,0,0,0]$.
+{\it Step 2;} The kernel is
+\begin{align}
+\vec K(\tau)= F \vec I(\tau)
+= c_1 \exp(-\sigma\,\tau)\,\left[ \begin {array}{c} \beta\\0 \end {array} \right] 
+ -c_1 \exp(-\gamma\,\tau)\,\left[ \begin {array}{c} \beta\\0 \end {array} \right].
+\end{align}
+
+{\it Step 3;} The basic reproduction number $\R$ is given by integrating the non-zero element of the kernel as follows.
+\begin{align}
+\R=\int_0^\infty \vec K_1(\tau) d\tau
+=\beta\,c_1(1/\sigma-1/\gamma)
+=\beta/\gamma.
+\end{align}
 

### 2021-05-19 cohort.tex
minor edits

@@
 The matrix form of the cohort framework consists of the following 3 steps.
-\begin{enumerate}[{\it step 1}]
+\begin{enumerate}[{\it Step 1.}]
 \item
-write the cohort Eq. 
+Form the cohort Eq. 
 \begin{equation}
@@
 \item 
-finding the intrinsic infectiousness kernel by integrating the cohort Eq. and solving for its time evolution, thus
+Finding the intrinsic infectiousness kernel by integrating the cohort Eq. and solving for its time evolution, thus
 the solution would be 
@@
 The cohort analogy of a simple SIR compartmental model where $\dot I=\beta S I/N-\gamma I$. The cohort framework is via the cohort Eq. with the following steps.
-{\it step 1;} write the cohort Eq. $I'=-\gamma I$ with the initial condition $I(0)=1$, 
-{\it step 2;} finding the intrinsic infectiousness kernel by integrating the cohort Eq. and solving for its time evolution, thus $k(\tau)=\beta \exp(-\gamma \tau)$.
-{\it step 3;} $\R$ will be the integration of the kernel. That is $\R=\beta/\gamma$.
+{\it Step 1;} write the cohort Eq. $I'=-\gamma I$ with the initial condition $I(0)=1$, 
+{\it Step 2;} finding the intrinsic infectiousness kernel by integrating the cohort Eq. and solving for its time evolution, thus $k(\tau)=\beta \exp(-\gamma \tau)$.
+{\it Step 3;} $\R$ will be the integration of the kernel. That is $\R=\beta/\gamma$.
 

### 2021-05-24 SIR_manuscript.tex
a little more

@@
 \newcommand{\percap}{\emph{per capita}\xspace}
-\newcommand{\Rnum}{\ensuremath{\mathcal{R}_0}}
+\newcommand{\Rnum}{\ensuremath{\mathcal{R}_0}\xspace}
 \newcommand{\covid}{COVID-19\xspace}
@@
 We derived general expressions for these derivatives.
-However, we analyzed the effect of $\omega$ on $\Rnum$ for the case when testing intensity is very small relative to the population size, $N_0$.
-Specifically, we used the Taylor approximation of $\Delta$ at $\rho=0$ to illustrate the unexpected non-monotonic relationship between $\Rnum$ and $\omega$, and also to avoid unnecessary complicated computations for making this point.
-\bmb{clarify/check}
+However, we analyzed the effect of $\omega$ on $\Rnum$ for the special case of low testing intensity.
+Specifically, by making the restriction $\rho \ll 1$, we are able to Taylor-expand $\Delta$ at $\rho=0$ and analyze the resulting simplified derivatives to illustrate a surprising non-monotonic relationship between $\Rnum$ and $\omega$.
+\bmb{The previous version of this implied that the restriction was $\rho \ll N_0$; that seems very odd since $\rho$ has units of time$^{-1}$.
+I think $\rho \ll 1$ is the correct statement (i.e.
+$\rho$ is already a \emph{per capita} measure), but please check.
+If I am wrong and $\rho$ is expressed in absolute units of tests/day, then the range quoted in the caption for \fref{pan} is weird \ldots}
 
@@
 Specifically, $\rho \in [0,0.013]$, and test return rate $\omega\in [1/12,2]$.
-This testing intensity reflects realistic testing rates during the \covid pandemic.
-For example, since 1\% of a population are tested per day at maximum capacity, giving a maximum of 10000 tests/day in a population of size $N_0=10^6$.
+This testing intensity reflects realistic testing rates during the \covid pandemic, i.e.
+a maximum of 1.3\% of the population per day, giving a maximum of 10000 tests/day in a population of size $N_0=10^6$ (this is approximately four times the maximum testing rate in Ontario, Canada in mid-2021).
 The less realistic high-testing case \fref{pan2} is included to highlight the occurrence of non-monotonic changes in $\Rnum$ with respect to $\rho$.
-\bmb{stopped here}
-\fref{pan2} with maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,1/5)$, and the test return rate $\omega\in [1/5,2]$. 
-Note that the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, corresponding to the threshold $\Rnum=1$, is plotted in dotted line in the panels.
-Thus, a combination of $\rho$ and $\omega$ above this critical contour results in reducing $\Rnum$.
-Also realize that $\Rnum=\frac{\beta}{\gamma}$, about 1.017 here, for the classical SIR model which can be achieved in the context of our model when no testing and no isolation are implied, that is $\rho=0$ or $\theta\_w=\theta\_w=0$.
+In \fref{pan2} the maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,1/5)$ and the test return rate $\omega\in [1/5,2]$; these values are clearly unrealistic for a large population but might be relevant for some population undergoing focused testing, such as a sports league or university.
+The critical contour of $\Delta=1-\frac{\gamma}{\beta}$, corresponding to the threshold $\Rnum=1$, is shown as a dotted line in \fref{pan} and \fref{pan2}; combinations of $\rho$ and $\omega$ that lie above this critical contour will control the epidemic.
+The implied baseline value of $\Rnum=\frac{\beta}{\gamma}$ in these figures is $\approx 1.017$, corresponding to a case where the epidemic spreads slowly even in the absence of testing and tracing.
+\bmb{how did we choose this particular value?}
 
@@
 
-For a variety of changes in testing and isolation parameters, we can use the formula for $\Rnum$ \eqref{R0} to confirm monotonic changes in the expected, intuitive direction (i.e.
+We can use the formula for $\Rnum$ \eqref{R0} to make a number of straightforward conclusions about parameters whose effects on $\Rnum$ are monotonic (i.e.
 the partial derivative of $\Delta$ has a consistent sign: Sections~\ref{app:rho}, \ref{app:omega}, \ref{app:w}).
@@
 \item \label{p1:omega} Increasing the rate of test return ($\omega$) always decreases \Rnum if waiting individuals do not isolate ($\theta_w=0$).
-\item \label{p1:w} Increasing testing focus from random ($w_S=w_I$) to targeted  ($w_S<w_I$) always decreases \Rnum.
+\item \label{p1:w} Increasing testing focus, i.e. changing the testing weights from random ($w_S=w_I$) toward targeted  ($w_S<w_I$), always decreases \Rnum.
 \end{enumerate}
@@
 Black dotted lines in each panel represent the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ such that $\Rnum=1$.
-\bmb{does $N_0$ enter the expression at all, i.e.
-can we omit it?
-Perhaps we can also include baseline $\Rnum$ and $r$ (or doubling time), i.e.
-$\beta/\gamma$ and $\beta-\gamma$.
-(Why did we pick these particular beta/gamma values?)}
+\bmb{Perhaps we can also include baseline $\Rnum$ and $r$ (or doubling time), i.e. $\beta/\gamma$ and $\beta-\gamma$. (Why did we pick these particular beta/gamma values?)
+If $\rho$ is per capita then does the figure depend on $N_0$ at all/can we leave it out of the list of specified parameters?}
 }
@@
 \item We also see that greater test intensity $\rho$ increases the effectiveness of control $\Delta$ ($\Delta$ increases as we move up in each plot in this Figure).
-Mathematically speaking, we did find a counter-vailing effect, but this can only be seen when we allow $\rho$ to take unrealistically large values, and only for weighted/targeted testing (see \fref{pan2}).
+Mathematically speaking, we did find a countervailing effect, but this can only be seen when we allow $\rho$ to take unrealistically large values, and only for weighted/targeted testing (see \fref{pan2}).
 The explanation here is that more rapid testing leaves more susceptibles in the ``waiting-for-no'' category at the DFE; these people must then wait for their tests to be returned before they can be tested again, receive a positive test, and become cautious.

### 2021-05-24 SIR_manuscript.tex
more tweaks

@@
 \newcommand{\covid}{COVID-19\xspace}
-\newcommand{\pro}[1][]{\ensuremath{\frac{\partial #1}{\partial \rho}}}
-\newcommand\pder[2][]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder[x]{y}
+\newcommand{\pro}[1]{\ensuremath{\frac{\partial #1}{\partial \rho}}}
+\newcommand\pder[2]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder[x]{y}
 
@@
 Although the model development was motivated by the \covid epidemic, the model has general epidemiological and testing structures.
-The realistic components of the model include \percap testing intensity, test sensitivity and specificity, rate of test return, and isolation efficacy in reduction of the probability of transmission.
+The realistic components of the model include \percap testing intensity, test sensitivity and specificity, rate of test return, and isolation efficacy (i.e., reduction of the transmission probability by isolating individuals).
 The novel component is the compartment-specific relative testing weights, which reflect the testing strategies-- surveillance, diagnosis, or control.
@@
   $\rho$   & \percap testing intensity & 1/day & - \\ \hline 
-  $\theta\_w$ & Isolation efficacy in reduction of the probability of transmission for ``waiting'' individuals & - & - \\ \hline
-  $\theta\_c$ & Isolation efficacy in reduction of the probability of transmission for ``confirmed positive'' individuals & - & -  \\ \hline
+  $\theta\_w$ & Isolation efficacy (reduction of the transmission probability) for ``waiting'' individuals & - & - \\ \hline
+  $\theta\_c$ & Isolation efficacy ``confirmed positive'' individuals & - & -  \\ \hline
   $\beta$ & Transmission rate & 1/day & 0.339 \\ \hline
@@
 % %%%%%%%
-\section{Results}
+\section{Results and discussion}
 
@@
 
+\bmb{can we connect each of these results to specific equations in the appendix?}
 \begin{enumerate}
@@
 
-\bmb{stopped about here.
-I'm a little unclear about the distinction between the results above and below here.
-I assumed that results above were intuitive/expected and those below weren't.
-But the last point here (targeting testing increases $\Delta$) seems intuitive.
-Is it about things we can prove/can't prove?
-Do we know anything about $\Delta$ vs.
-targeting?}
-
 \begin{itemize}
-\item When waiting individuals isolate strongly (e.g.
-$\theta_w =0$, top row of \fref{pan}), increasing the rate at which tests are returned/shortening testing delays (i.e.
-increasing $\omega$) reduces \Rnum, as we might expect ($\Delta$ increases as we move to the left in each plot in this row).
-For non-zero $\Delta$, however, we mostly see the opposite effect: less waiting leads to more transmission.
-This is because returning negative tests leads people to stop distancing; this applies both to susceptibles, and to infectious people who receive negative test results because they were sampled when susceptible (or because of imperfect test sensitivity, which is not modeled in this figure).
-
-\item We also see that greater test intensity $\rho$ increases the effectiveness of control $\Delta$ ($\Delta$ increases as we move up in each plot in this Figure).
-Mathematically speaking, we did find a countervailing effect, but this can only be seen when we allow $\rho$ to take unrealistically large values, and only for weighted/targeted testing (see \fref{pan2}).
-The explanation here is that more rapid testing leaves more susceptibles in the ``waiting-for-no'' category at the DFE; these people must then wait for their tests to be returned before they can be tested again, receive a positive test, and become cautious.
-This effect is usually weak compared to the helpful effects of testing, but if testing is weighted, $\theta_w$ is low, and test returns are slow, it is possible for increasing $\rho$ to reduce $\Delta$ (see upper-right part of upper-right plot of \fref{pan2}(b)).
+\item When waiting individuals do not isolate (e.g.
+$\theta_w =0$, top row of \fref{pan}), increasing testing delays (i.e.
+decreasing $\omega$) increases \Rnum, as we might expect: $\Delta$ decreases --- a larger region of the parameter space has $\Rnum>1$, and the threshold testing intensity required to control the epidemic increases --- as we move to the right within each plot in this row.
+However, when waiting individuals are at least partially isolated ($\theta_w>0$), we typically see the opposite effect: longer testing delays lead to \emph{less} transmission.
+This is because returning negative tests leads people to stop distancing; this applies both to susceptibles, and to infectious people who receive negative test results because they were sampled when susceptible (or because of imperfect test sensitivity, which we have not considered in our analysis).
+For small but non-zero waiting-isolation levels (e.g.
+$\theta_w=0.25$, second row from top), with high levels of confirmed isolation ($\theta_c \ge 0.5$), increasing testing delays initially decreases $\Delta$, then increases it as the testing delays pass a threshold value of approximately 2--5 days.
+\item \fref{pan} also shows that greater testing intensity (increasing $\rho$) generally increases the effectiveness of control (moving up in each panel).
+This relationship can be reversed, but only for an unrealistic scenario --- very high testing intensities \emph{that begin before the epidemic starts}.
+It occurs when waiting people do not isolate ($\theta_w=0$), and most easily for targeted testing (\fref{pan2}(b), right three panels of top row) \bmb{can we rule out the possibility that this phenomenon could happen with random testing?
+In the left panel of \fref{pan2}, $\theta_w=0$/$\theta_c=0.75$ panel, there is a hint of this behaviour; it looks like the light green [second lightest] region displays this behaviour, but it's hard to see even a maximum magnification}.
+This phenomenon occurs --- increasing testing intensity \emph{increases} \Rnum --- because more rapid testing leaves more susceptibles in the ``waiting-for-negative-results'' category at the DFE; these people must then wait for their tests to be returned (during which time they may become infected) before they can be tested again, receive a positive test, and then begin self-isolating.
+This effect is usually weak compared to the beneficial effects of testing.
 
@@
 In particular, we realize higher $\Delta$ levels, i.e., darker colors in corresponding plots\/facets as we go from panel~(a) to panel~(b) in the figures.
+\bmb{this is the same as point 4 above, right?
+Am I missing something?}
 \end{itemize}
@@
 \section{Discussion}
+
 \begin{enumerate}
@@
 \end{align}
-One way to do such an analysis is based on the fact that $\rho$ is \percap rate so for a large population size, $N_0 \gg 1$, $\rho$ is very small or close to 0, comparing to $N_0$.
-This provides a base to linearly approximate $\Rnum$ when $\rho$ is close to 0, and use this approximation to analyze the behaviour of $\Rnum$ with respect to $\omega$ (see section \ref{app:omega}).
-In the next section we provide an equivalent representation of $\Rnum$ to provide a ground to prove that more testing intensity decreases $\Rnum$ for a general range of parameters.
+In realistic cases the testing rate $\rho$ is very small (i.e., only a small fraction of the population can be tested every day); it is thus reasonable to use a linear approximation of $\Rnum$ for $\rho \ll 1$ to analyze the behaviour of $\Rnum$ with respect to $\omega$ (see section \ref{app:omega}).
+In the next section we provide an equivalent representation of $\Rnum$ in order to show that increasing testing intensity typically decreases $\Rnum$.
 
@@
 
-This section shows that $\pro \Delta$ can be positive or negative, with $\Delta$ defined in Eq.\,\eqref{eq:C12}, and thus $\pro \Rnum < 0$, where $\Rnum$ is the basic reproduction number, given in Eq.\,\eqref{R0}.
-We can rewrite matrix $G_{11}$ in \eqref{G11} in the following form to simplify the calculations.
+This section shows that $\pro{\Delta}$ can be positive or negative, with $\Delta$ defined in Eq.\,\eqref{eq:C12}, and thus $\pro{\Rnum} < 0$, where $\Rnum$ is given in Eq.~\eqref{R0}.
+We rewrite matrix $G_{11}$ in \eqref{G11} in the following form to simplify the calculations:
 \begin{equation}

### 2021-05-24 codes/sir_plot.R
Remove some unnecessary rs

@@ -129,7 +129,7 @@ print(p1_temp)
 # 1. Plot the Random Testing Scenario:
 # #################################
 
-ggsave(p1_temp + ggtitle(TeX(r'(Random testing, $w_S=w_I=1$)')) +
+ggsave(p1_temp + ggtitle(TeX('(Random testing, $w_S=w_I=1$)')) +
        theme(legend.position = "none"),
        filename = "R0contour_random.pdf" ,
        width = 12, height = 12, units = "cm")
@@ -218,7 +218,7 @@ print(p1_temp2)
 # #################################
 
 ggsave(p1_temp2 
-       + ggtitle(TeX(r'(Random testing, $w_S=w_I=1$)')) 
+       + ggtitle(TeX('(Random testing, $w_S=w_I=1$)')) 
        + theme(legend.position = "none"),
        filename = "R0contour_random2.pdf" ,
        width = 12, height = 12, units = "cm")

### 2021-05-26 SIR_manuscript.tex
more comments

@@
   $\theta\_w$ & Isolation efficacy (reduction of the transmission probability) for ``waiting'' individuals & - & - \\ \hline
-  $\theta\_c$ & Isolation efficacy ``confirmed positive'' individuals & - & -  \\ \hline
+  $\theta\_c$ & Isolation efficacy for ``confirmed positive'' individuals & - & -  \\ \hline
   $\beta$ & Transmission rate & 1/day & 0.339 \\ \hline
@@
 \bmb{how did we choose this particular value?}
+  \bmb{JD wonders whether we really need the ``critical'' $\Rnum=1$ contour or not; any contour will do to illustrate a threshold level of testing intensity, and the fact that $\Rnum=1$ falls in the middle of our range is a bit arbitrary (i.e.
+the fact that testing can control the epidemic depends on the baseline $\Rnum$ value we have chosen).
+(I know it was my idea in the first place, but maybe we could dispense with it?)}
 
@@
 Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
-The black dotted line in each panel represents the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ corresponding to the threshold of $\Rnum=1$.}
+The black dotted line in each panel represents the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ corresponding to the threshold of $\Rnum=1$.
+\bmb{consider removing dotted line (see note elsewhere); on this and prev figure consider removing redundant y-axis label on right panel (and adjusting/squashing interpanel spacing)?}}
 \label{pan2}
@@
 This effect is usually weak compared to the beneficial effects of testing.
+  \bmb{I am still trying to understand why slowing testing return (decreasing $\omega$) decreases \Rnum over \textbf{most} of the parameter space considered here, which is a counterintuitive/silly result.
+Fast test returns help when we have high test positivity and low isolation during the waiting period (specifically, [much] lower isolation while waiting than when confirmed, but I think we're assuming that); we want to get waiting people into the ``confirmed'' compartment as quickly as possible.
+Can we understand/explain this result more thoroughly using that logic, and can we argue that what we are looking at is \emph{not} a realistic parameter range?
+Does the low baseline \Rnum we are using play into this, e.g.
+by reducing test positivity?
+Is there a simple expression for test positivity during the invasion phase?}
 

### 2021-05-26 cohort.tex
Meeting with Ali
still working
adding SIR with testing example

@@
 \usepackage{enumerate}
+\usepackage{xspace}
 
+\newcommand{\ali}[1]{\comment{magenta}{Ali}{#1}}
+\newcommand{\jd}[1]{\comment{blue}{JD}{#1}}
+\newcommand{\david}[1]{\comment{red}{DJDE}{#1}}
+\newcommand{\bmb}[1]{\comment{red}{BMB}{#1}}
 \newcommand{\todo}[1]{\comment{red}{TODO}{#1}}
 
-\usepackage{xspace}
+\newcommand*\subtxt[1]{_{\textnormal{#1}}}
+\DeclareRobustCommand\_{\ifmmode\expandafter\subtxt\else\textunderscore\fi}
 
@@
 Note that the n-by-n matrix $\exp(-V\tau)$ is the probability of being infected and stay infectious at time $\tau$.
-% where $\exp(-V\tau)=\sum_{n=0}^\infty \tau^n (-V)^n/n!$.
 
 \item
-Calculating $\vec R$ by integrating kernel $\vec K(\tau)$. Note that $\R$ would be the non-negative element of $\vec R$.
+Calculating the next-generation vector $\vec G$ (\ali{feels odd here, help! also check Heesterbeek work what he calls this}) is
+\begin{equation}
+\label{eq:ng} %next generation
+\vec G = \int_0^\infty \vec K(\tau) d\tau.
+\end{equation}
+Note that $\R$ would be the non-zero element of $\vec G$.
 \end{enumerate}
@@
 
+\subsection{SIR with testing and isolation}
+The model can be summarized in the form of following matrices.
+\begin{align}
+\label{FV}
+F =&\beta/N_0 \left[ \begin {array}{c} S\_u^* \\ (1-\theta\_w)S\_n^* \\ 0 \\ 0 \end {array} \right]
+        \left[ \begin {array}{cccc} 1, \, 1-\theta\_w, \, 1-\theta\_w, \, 1-\theta\_c \end {array} \right], \text{and}\\  V=&
+ \left[ \begin {array}{cccc}  
+\hat F_I+\gamma&-\omega&0&0\\
+-(1-p_I)\hat F_I&\omega+\gamma&0&0\\
+-p_I \hat F_I&0&\omega+\gamma&0\\
+0&0&-\omega&\gamma
+\end {array} \right].
+\end{align}
+
+Let the test be perfectly sensative, that is $p_I=1$.
+
+{\it Step 1;} The cohort Eq.~ is
+\begin{align}
+\vec I'=-V\,\vec I,~~ \text{where}~~ \vec I(\tau)=\left[ \begin {array}{c} I\_u(\tau)\\I_n(\tau)\\I_p(\tau)\\I_c(\tau) \end {array} \right], ~~
+\vec I(0)=\left[ \begin {array}{c} S\_u^*\\S\_n^*\\0\\0 \end {array} \right].
+\end{align}
+Solving for $\vec I(\tau)$ by finding the eigenvalues and eigenvectors of the above linear system results in
+\begin{align}
+\vec I(\tau)=&
+c_1\,\exp(-\gamma\,\tau)\,\left[ \begin {array}{c} 0\\0\\0\\1 \end {array} \right] 
++c_2\,\exp(-(F_I+\gamma)\,\tau)\,\left[ \begin {array}{c} \frac{F_I-\omega}{\omega}\\0\\\frac{-F_I}{\omega}\\1 \end {array} \right]
++c_3\,\exp(-(\omega+\gamma)\,\tau)\,\left[ \begin {array}{c} 0\\0\\-1\\1 \end {array} \right] \\\notag
++& c_4\,\exp(-(\omega+\gamma)\,\tau)\,\Big(\tau\,\left[ \begin {array}{c} 0\\0\\-1\\1 \end {array} \right]+
+\left[ \begin {array}{c} \frac{-1}{F_I}\\ \frac{\omega-F_I}{F_I\,\omega}\\ \frac{1}{\omega}\\0 \end {array} \right]\Big),
+\end{align}
+where 
+\begin{align}
+c_1=&1\\
+c_2=&\frac{\omega}{F_I-\omega}(S\_u^*-\frac{\omega}{F_I-\omega}\,S\_n^*)\\
+c_3=&\frac{-F_I}{F_I-\omega}(1-\frac{\omega}{F_I-\omega}S\_n^*)\\
+c_4=&\frac{-\omega\,F_I}{F_I-\omega}S\_n^*.
+\end{align}
+
+{\it Step 2;} The kernel is
+\begin{align}
+\vec K(\tau)= F \vec I(\tau)
+\end{align}
+
+{\it Step 3;} The next-generation vector is $\vec G$ as follows.
+\begin{align}
+\vec G=&\int_0^\infty \vec K(\tau) d\tau\\
+=&
+\end{align}
 %%%%%%%%%%%%%%%%%%%%

### 2021-05-27 cohort.tex
adding more details

@@
 \end{equation}
- with the proper initial condition $I(0)$, where the n-by-n matrix $V$ is the flow matrix of leaving infected compartments in \cite{van2002reproduction}.
+ with the vector of initial condition $\vec I(0)$ where $\vec I_i(0)$ be the number of infected individuals initially in compartment $i$, and $I(\tau)$ is the number of these initially infected individuals remaining in the infected compartments after $\tau$ time units.
+Note that as \cite{van2002reproduction} set their formulation, 
+``the $(i,j)$ entry of $F$ is the rate at which infected individuals in compartment $j$ produce new infections in compartment $i$".
+Matrix $V$ is the flow matrix in which the $(i,j)$ entry is the the rate of transfering infected individuals from compartment $j$ into compartment $i$, with positive (negative) means transfering out (into) a compartment.
+The $(j,k)$ entry of $V^{-1}$ is the average length of time an infected individual introduced into compartment $k$ spends in compartment $j$ during its lifetime, assuming that the population remains near the DFE and barring reinfection.
+Note that $V$ is a non-singular M-matrix and is, therefore, invertible and all of its eigenvalues have positive real parts \citep{van2002reproduction}.
 
@@
 \item
-Calculating the next-generation vector $\vec G$ (\ali{feels odd here, help! also check Heesterbeek work what he calls this}) is
+Calculating the next-generation vector $\vec G$ is 
+(\ali{feels odd here, help! $\vec G$ here is a vector with elements of the eigenvalues of the \citep{van2002reproduction}'s next-generation matrix $FV^{-1}$. 
+Also check Heesterbeek work what he calls this}) 
 \begin{equation}
 \label{eq:ng} %next generation
-\vec G = \int_0^\infty \vec K(\tau) d\tau.
+\vec G = FV^{-1} = \int_0^\infty \vec K(\tau) d\tau.
 \end{equation}
-Note that $\R$ would be the non-zero element of $\vec G$.
+Note that $\R$ would be the dominant element of $\vec G$.
 \end{enumerate}
-Note that in the compartment framework \citep{van2002reproduction}, steps 2 and 3 are combined.
+Note that in the compartment framework \cite{van2002reproduction}, steps 2 and 3 are combined.
 %%%%%%%%%%%%%%%%%%%%
@@
 
-\subsection {Simple SIR}
-The cohort analogy of a simple SIR compartmental model where $\dot I=\beta S I/N-\gamma I$. The cohort framework is via the cohort Eq. with the following steps.
-{\it Step 1;} write the cohort Eq. $I'=-\gamma I$ with the initial condition $I(0)=1$, 
-{\it Step 2;} finding the intrinsic infectiousness kernel by integrating the cohort Eq. and solving for its time evolution, thus $k(\tau)=\beta \exp(-\gamma \tau)$.
-{\it Step 3;} $\R$ will be the integration of the kernel. That is $\R=\beta/\gamma$.
+{\bf Example 1. Simple SIR model;}
+The model is
+\begin{align}\label{mo:sir}
+\dot S=-\beta\,S\,I/N,~~
+\dot I= \beta\,S\,I/N -\gamma\,I,~~
+\dot R= \gamma\,I.
+\end{align}
+Where $F=\beta$ and $V=\gamma$, thus $V^{-1}=1/\gamma$. The next-generation matrix $G=FV^{-1}=\beta/\gamma$ which results in $\R=\beta/\gamma$ in the compartment framework.
+The cohort analogy of the simple SIR model is via the cohort Eq. with the following steps.
 
-\subsection{Simple SEIR model}
+{\it Step 1;} The cohort Eq. $I'=-\gamma I$ with the initial condition $I(0)=1$. Thus, the solution is $I(\tau)=\exp(-\gamma \tau)$.
+
+{\it Step 2;} The intrinsic infectiousness kernel is given by integrating the cohort Eq. and solving for its time evolution.  Thus, $K(\tau)=F\,I(\tau)=\beta \exp(-\gamma \tau)$.
+
+{\it Step 3;} The next-generation vector would be 
+$$G=FV^{-1}=\int_0^\infty\,K(\tau)\,d\tau=\beta/\gamma.$$ This gives $\R=\beta/\gamma$.
+
+%%%%%%%%%
+\noindent
+{\bf Example 2. Simple SEIR model;}
 The simplest form of a SEIR model without vital rates is 
@@
 \end{align}
-I the context of the cohort framework we show that the same $\R$ can be calculated by following the three steps.
+Note that matrix $F$ determines the rate of new infections and their source.
+Specifically, here $F_{12}=\beta$, that is the rate of new infections in compartment $E$ is $\beta$ and it is originated from compartment $I$.
+Also note that matrix $V$ determines the flow
+In the context of the cohort framework we show that the same $\R$ can be calculated by following the three steps.
 
@@
 \vec I(\tau) = \left[ \begin {array}{c} E(\tau)\\I(\tau) \end {array} \right] =
-c_1 \exp(-\sigma\,\tau)\,\left[ \begin {array}{c} \gamma/\sigma-1\\1 \end {array} \right] 
--c_1 \exp(-\gamma\,\tau)\,\left[ \begin {array}{c} 0\\1 \end {array} \right],
+\frac{\sigma}{\gamma-\sigma}\,\Big( \exp(-\sigma\,\tau)\,\left[ \begin {array}{c} \gamma/\sigma-1\\1 \end {array} \right] 
+- \exp(-\gamma\,\tau)\,\left[ \begin {array}{c} 0\\1 \end {array} \right] \Big).
 \end{align}
-where $$c_1=\frac{\sigma}{\gamma-\sigma}.$$
 
@@
 \vec K(\tau)= F \vec I(\tau)
-= c_1 \exp(-\sigma\,\tau)\,\left[ \begin {array}{c} \beta\\0 \end {array} \right] 
- -c_1 \exp(-\gamma\,\tau)\,\left[ \begin {array}{c} \beta\\0 \end {array} \right].
+= \frac{\sigma}{\gamma-\sigma} \Big(\exp(-\sigma\,\tau)\,\left[ \begin {array}{c} \beta\\0 \end {array} \right] 
+ -\exp(-\gamma\,\tau)\,\left[ \begin {array}{c} \beta\\0 \end {array} \right]\Big).
 \end{align}
 
-{\it Step 3;} The basic reproduction number $\R$ is given by integrating the non-zero element of the kernel as follows.
+{\it Step 3;} The next-generation vector is 
 \begin{align}
-\R=\int_0^\infty \vec K_1(\tau) d\tau
-=\beta\,c_1(1/\sigma-1/\gamma)
-=\beta/\gamma.
+\vec G=\int_0^\infty \vec K(\tau) d\tau
+=\beta\,\frac{\sigma}{\gamma-\sigma}\,\Big(\left[ \begin{array}{c} 1/\sigma-1/\gamma\\0 \end{array} \right]\Big)
+=\left[ \begin{array}{c} \beta/\gamma\\0 \end{array} \right].
 \end{align}
 
-\subsection{SIR with testing and isolation}
+%%%%%%%%%
+\noindent
+{\bf Example 3. SIR with testing and isolation;}
 The model can be summarized in the form of following matrices.

### 2021-05-28 SIR_manuscript.tex
addressing the comments

@@
   \label{eq:del4a}
-  \Delta= \frac{1}{C\,N_0}\big(C_1\,S\_u^*+(C_2(1-\theta\_w)+C\,\theta\_w)\,S\_n^*\big),
+  \Delta= \frac{1}{C N_0}\big(C_1 S\_u^*+(C_2(1-\theta\_w)+C \theta\_w)S\_n^*\big),
 \end{equation}
@@
   TeX do the typesetting itself? Do we really prefer more compact equations, and if so is there a package
-  that will just change the appropriate settings globally?}
+  that will just change the appropriate settings globally?}\ali{are we talking about \,? not sure about the package!}
 where
@@
 \label{eq:C12}
-C_{\phantom 1} =& (\omega+\gamma) \Big(\gamma(\omega+\gamma)+(\gamma+\omega p_I)\hat F_I\Big),\\
-C_1 =& (\omega+\gamma)(\theta\_w \, \gamma+\theta\_c\, \omega\, p_I) \hat F_I,\\
-C_2 =& \Big( \omega\gamma(1+p_I)\hat F_I+\gamma^2\,(\omega+\gamma+\hat F_I)\Big)\theta\_w + \omega^2\,p_I\,\hat F_I\,\theta\_c.
+C_{\phantom 1} &= (\omega+\gamma) \Big(\gamma(\omega+\gamma)+(\gamma+\omega p_I)\hat F_I\Big),\\
+C_1 &= (\omega+\gamma)(\theta\_w \gamma+\theta\_c \omega p_I) \hat F_I,\\
+C_2 &= \Big( \omega\gamma(1+p_I)\hat F_I+\gamma^2(\omega+\gamma+\hat F_I)\Big)\theta\_w + \omega^2p_I\hat F_I\theta\_c.
 \end{align}
-\bmb{I don't like the alignment of the equations: too little space after =. I realize that's incredibly picky.}
 Further details of derivation of $\Rnum$ are provided in appendix, Sec.\,\ref{app:R0}. 
@@
 Specifically, by making the restriction $\rho \ll 1$, we are able to Taylor-expand $\Delta$ at $\rho=0$ and analyze the resulting simplified derivatives to illustrate a surprising non-monotonic relationship between $\Rnum$ and $\omega$.
-\bmb{The previous version of this implied that the restriction was $\rho \ll N_0$; that seems very odd since $\rho$ has units of time$^{-1}$.
-I think $\rho \ll 1$ is the correct statement (i.e.
-$\rho$ is already a \emph{per capita} measure), but please check.
-If I am wrong and $\rho$ is expressed in absolute units of tests/day, then the range quoted in the caption for \fref{pan} is weird \ldots}
 
 The analytical calculation of the next-generation matrix and simplification of the $\Rnum$ expression was carried out in Maple \citep{maple14}.
-We used R \citep{r} for numerical solutions, in particular for plotting the contours of $\Delta$ (Eq.~\eqref{NOTSURE}) \bmb{not sure what the right ref.
-is.
-This was previously ``eq:tel4'', which doesn't exist.
-Should it have been ``eq:del4''?} over a range of parameters.
+We used R \citep{r} for numerical solutions, in particular for plotting the contours of $\Delta$ (Eq.~\eqref{eq:del4a}) over a range of parameters.
 We computed the values and contours of Delta at both low (\fref{pan}) and high (\fref{pan2}) testing intensities, and for both random testing ($w_S=w_I=w_R=1$) and targeted testing ($w_S=0.3$; $w_I=w_R=1$).
@@
 The implied baseline value of $\Rnum=\frac{\beta}{\gamma}$ in these figures is $\approx 1.017$, corresponding to a case where the epidemic spreads slowly even in the absence of testing and tracing.
-\bmb{how did we choose this particular value?}
+\bmb{how did we choose this particular value?}\ali{I was thinking of ``critical'' $\Rnum=1$ and wanted to have the baseline in the plot.}
   \bmb{JD wonders whether we really need the ``critical'' $\Rnum=1$ contour or not; any contour will do to illustrate a threshold level of testing intensity, and the fact that $\Rnum=1$ falls in the middle of our range is a bit arbitrary (i.e.
 the fact that testing can control the epidemic depends on the baseline $\Rnum$ value we have chosen).
-(I know it was my idea in the first place, but maybe we could dispense with it?)}
+(I know it was my idea in the first place, but maybe we could dispense with it?)}\ali{agree with the point, the plots will be modified.}
 

### 2021-05-31 cohort.tex
addressing the comments
adding minor stuff

@@
 Note that in the compartment framework \cite{van2002reproduction}, steps 2 and 3 are combined.
+
 %%%%%%%%%%%%%%%%%%%%
@@
 F =&\beta/N_0 \left[ \begin {array}{c} S\_u^* \\ (1-\theta\_w)S\_n^* \\ 0 \\ 0 \end {array} \right]
-        \left[ \begin {array}{cccc} 1, \, 1-\theta\_w, \, 1-\theta\_w, \, 1-\theta\_c \end {array} \right], \text{and}\\  V=&
- \left[ \begin {array}{cccc}  
+        \left[ \begin {array}{cccc} 1, \, 1-\theta\_w, \, 1-\theta\_w, \, 1-\theta\_c \end {array} \right], \text{and}\\  
+V=& \left[ \begin {array}{cccc}  
 \hat F_I+\gamma&-\omega&0&0\\
--(1-p_I)\hat F_I&\omega+\gamma&0&0\\
--p_I \hat F_I&0&\omega+\gamma&0\\
+0&\omega+\gamma&0&0\\
+-\hat F_I&0&\omega+\gamma&0\\
 0&0&-\omega&\gamma
@@
 %%%%%%%%%%%%%%%%%%%%
+\section{Ideas-Unknowns}
+- Collapsing the vectorized kernek $\vec K(\tau)$ into a scalar function of time.
+
+- We have 3 cases; (i) the general case where each box in $\vec I$ been pulled into others boxes of $\vec I$  (ii) simple case where we have 1 group eg.
+linear chain (iii) intermediate case where we have more than 1 entry point into $\vec I$ but mediated through a unitary FoI like SIR with testing.
+
+- SIR with testing;
+
+- Linear Algebra refreshing: \url{https://see.stanford.edu/materials/lsoeldsee263/11-eig.pdf}
 

### 2021-05-31 SIR_manuscript.tex
addressing comments, edit figs

@@
 \newcommand{\pro}[1]{\ensuremath{\frac{\partial #1}{\partial \rho}}}
-\newcommand\pder[2]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder[x]{y}
+\newcommand\pder[2]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder y{x}
 
@@
   TeX do the typesetting itself? Do we really prefer more compact equations, and if so is there a package
-  that will just change the appropriate settings globally?}\ali{are we talking about \,? not sure about the package!}
+  that will just change the appropriate settings globally?}\ali{are we talking about \\,?}
 where
@@
 the fact that testing can control the epidemic depends on the baseline $\Rnum$ value we have chosen).
-(I know it was my idea in the first place, but maybe we could dispense with it?)}\ali{agree with the point, the plots will be modified.}
+(I know it was my idea in the first place, but maybe we could dispense with it?)}\ali{agree, the plots are updated.}
 
@@
 
-\bmb{can we connect each of these results to specific equations in the appendix?}
 \begin{enumerate}
 
-\item \label{p1:eta} Increasing isolation efficacy for waiting ($\theta\_w$) and confirmed-positive ($\theta\_c$) individuals always decreases \Rnum;
+\item \label{p1:eta} Increasing isolation efficacy for waiting ($\theta\_w$) and confirmed-positive ($\theta\_c$) individuals always decreases \Rnum (Eqs.~\ref{eq:del4_theta},~\ref{eq:d_del4_thetac},~\ref{eq:d_del4_thetaw});
 \item \label{p1:rho} Higher testing intensity $\rho$ decreases $\Rnum$ if
 %% \begin{enumerate} \item
-testing is random (all $w_X$ equal) or testing intensity ($\rho$) is small.
+testing is random (all $w_X$ equal) or testing intensity ($\rho$) is small (Eq.~\ref{eq:dd3dr}).
 %% \end{enumerate}
-\item \label{p1:omega} Increasing the rate of test return ($\omega$) always decreases \Rnum if waiting individuals do not isolate ($\theta_w=0$).
-\item \label{p1:w} Increasing testing focus, i.e. changing the testing weights from random ($w_S=w_I$) toward targeted  ($w_S<w_I$), always decreases \Rnum.
+\item \label{p1:omega} Increasing the rate of test return ($\omega$) always decreases \Rnum if waiting individuals do not isolate ($\theta_w=0$) (Eq.~\ref{eq:dlindo}).
+\item \label{p1:w} Increasing testing focus, i.e. changing the testing weights from random ($w_S=w_I$) toward targeted  ($w_S<w_I$), always decreases \Rnum (Eq.~\ref{eq:d_del_wis}).
 \end{enumerate}
@@
 \centering
-\begin{subfigure}[t]{.45\textwidth}
+\begin{subfigure}[t]{.46\textwidth}
 \centering
@@
 %
-\begin{subfigure}[t]{.45\textwidth}
+\begin{subfigure}[t]{.44\textwidth}
 \centering
@@
 Numerical evaluation of the effectiveness control parameter, $\Delta$ (Eq.~\ref{eq:del4}), with respect to changes in testing and isolation. Parameter values (Table~\ref{tab:params}):
-$N_0=10^6$, $\beta=0.339$ day$^{-1}$ $\gamma= 1/3$ day$^{-1}$; $\omega \in [1/12,2]$ day$^{-1}$;  $\rho \in [0,0.013]$ day$^{-1}$; $\theta\_w$ and $\theta\_c$ vary between 0 (no effect of isolation) and 1 (complete elimination of transmission); $p_S=0$, $p_I=1$ and $p_R=0.5$.
+$\beta=0.339$ day$^{-1}$ $\gamma= 1/3$ day$^{-1}$; $\omega \in [1/12,2]$ day$^{-1}$;  $\rho \in [0,0.013]$ day$^{-1}$; $\theta\_w$ and $\theta\_c$ vary between 0 (no effect of isolation) and 1 (complete elimination of transmission); $p_S=0$, $p_I=1$ and $p_R=0.5$.
 Contours of $\Delta$ are plotted for (a) random testing ($w_S=w_I=w_R=1$) and (b) targeted testing ($w_S=0.3$; $w_I=w_R=1$).
-Black dotted lines in each panel represent the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ such that $\Rnum=1$.
-\bmb{Perhaps we can also include baseline $\Rnum$ and $r$ (or doubling time), i.e. $\beta/\gamma$ and $\beta-\gamma$. (Why did we pick these particular beta/gamma values?)
-If $\rho$ is per capita then does the figure depend on $N_0$ at all/can we leave it out of the list of specified parameters?}
+The baseline basic reproduction number $\Rnum$ and the doubling time $r$, i.e. $\beta/\gamma$ and $\beta-\gamma$ respectively, are approximately 1.017 and 0.006 in this simulation.
 }
@@
 We use the following parameters (see Table~\ref{tab:params} for a brief description):
-$N_0=1 \times 10^6$, $\omega \in [1/5,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,1/5)$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.339$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
+$\omega \in [1/5,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,1/5)$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.339$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
 Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
-The black dotted line in each panel represents the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ corresponding to the threshold of $\Rnum=1$.
-\bmb{consider removing dotted line (see note elsewhere); on this and prev figure consider removing redundant y-axis label on right panel (and adjusting/squashing interpanel spacing)?}}
+The baseline basic reproduction number $\Rnum$ and the doubling time $r$, i.e.
+$\beta/\gamma$ and $\beta-\gamma$ respectively, are approximately 1.017 and 0.006 in this simulation.
+\bmb{consider removing dotted line (see note elsewhere); on this and prev figure consider removing redundant y-axis label on right panel (and adjusting/squashing interpanel spacing)?}
+\ali{ok?}
+}
 \label{pan2}
@@
 Am I missing something?}
+\ali{you are right.}
 \end{itemize}
@@
 
-{\bf The first goal} is to explore how changes in isolation, $\theta\_w$ and $\theta\_c$, affects $\Rnum$.
-Mathematically we would like to verify the sign of $\pder[\Rnum]{\theta\_w}$ and $\pder[\Rnum]{\theta\_c}$.
+{\bf The first goal}
+is to explore how changes in isolation, $\theta\_w$ and $\theta\_c$, affects $\Rnum$.
+Mathematically we would like to verify the sign of $\pder \Rnum{\theta\_w}$ and $\pder \Rnum{\theta\_c}$.
 We start with simplifying $\Delta$ \eqref{eq:del4} by factoring $\theta\_w$ and $\theta\_c$ in Eq.\,\eqref{eq:del4}.
@@
 \end{align}
-$\Delta$, Eq.\,\eqref{eq:del4_theta}, is linear in $\theta\_c$ with a positive coefficient given by 
-$$1/C\,(\gamma\,S\_u^*\,+\omega(1-\theta\_w\,S\_u^*))\omega\,p_I\,\hat F_I.$$
-This results in increasing $\Delta$, thus decreasing $\Rnum$ with respect to $\theta\_c$, that is $\pder[\Rnum]{\theta\_c}\leq 0$. Note that $C$ is independent of $\theta\_c$ and $\theta\_w$.
+$\Delta$, Eq.\,\eqref{eq:del4_theta}, is linear in $\theta\_c$ with a positive coefficient. thus
+\begin{equation}
+\label{eq:d_del4_thetac}
+\pder\Delta{\theta\_c}=1/C(\gamma S\_u^*+\omega(1-\theta\_w S\_n^*))\omega p_I \hat F_I.
+\end{equation}
+This results in increasing $\Delta$, thus decreasing $\Rnum$ with respect to $\theta\_c$, that is $\pder \Rnum{\theta\_c}\leq 0$. Note that $C$ is independent of $\theta\_c$ and $\theta\_w$.
 
@@
 \end{equation}
-We show that the feasible range of $\theta\_w$ lies between 0 and the vertex of this parabola where the parabola is increasing in $\theta\_w$, and so does $\Delta$ which results in inferring $\pder[\Rnum]{\theta\_w}\leq 0$.
+We show that the feasible range of $\theta\_w$ lies between 0 and the vertex of this parabola where the parabola is increasing in $\theta\_w$, and so does $\Delta$ which results in inferring $\pder\Rnum{\theta\_w}\leq 0$.
 It is enough to show that partial derivative of the expression \eqref{eq:del4_thetaw} with respect to $\theta\_w$ at $\theta\_w=1$ is non-negative. It follows that
@@
 \label{eq:d_del4_thetaw}
-\pder[\Delta]{\theta\_w}\bigg\rvert_{\theta\_w=1} =&
-1/C\Big( (D_2-2D_1-\omega^2\,p_I\,\hat F_I\,\theta\_c)\,S\_n^*+\gamma\,\hat F_I(\omega+\gamma) \Big) \\
+\pder\Delta{\theta\_w}\bigg\rvert_{\theta\_w=1} =&
+1/C\Big( (D_2-2D_1-\omega^2\,p_I\,\hat F_I\,\theta\_c)\,S\_n^*+\gamma\,\hat F_I(\omega+\gamma) \Big) \\\notag
 =& 1/C\Big( (\gamma(\omega+\gamma)+\gamma\omega^2+(1-\theta\_c)\omega^2\,p_I\,\hat F_I )\,S\_n^*
@@
 {\bf The second goal} is to explore how changes in \percap testing intensity $\rho$ affects $\Rnum$.
-Mathematically we would like to verify the sign of $\pder[\Rnum]{\rho}$, which specifically depends on $\pder[\Delta]{\rho}$.
+Mathematically we would like to verify the sign of $\pder\Rnum{\rho}$, which specifically depends on $\pder\Delta{\rho}$.
 We use the derived expressions for $S\_u^*$ and $S\_n^*$, given by Eqs.\,\eqref{dfe}, in $\Delta$ \eqref{eq:del4}.
@@
 \end{equation}
-This one-to-one monotonic reparameterization enables us to simplify the mathematical expressions and explore the simpler $\pder[\Delta]{\phi}$ instead of the complicated $\pder[\Delta]{\rho}$.
+This one-to-one monotonic reparameterization enables us to simplify the mathematical expressions and explore the simpler $\pder\Delta{\phi}$ instead of the complicated $\pder\Delta{\rho}$.
 The derivative is
@@
 In this case, simplifying $a_3$ in Eq.~\eqref{eq:abcd2} gives $$a_3=\frac{w_I}{w_S}\,\omega\,\gamma\,p_I((\omega+\gamma)-\frac{w_I}{w_S}(\omega\,p_I+\gamma)).$$
-Specifically, in the case of targeted testing which is identified by $\frac{w_I}{w_S}> 1$, and using a perfect sensitive test, thus $p_I=1$, there exists a range for $\rho$ over which $\pder[\Rnum]{\rho}\leq\,0$.
+Specifically, in the case of targeted testing which is identified by $\frac{w_I}{w_S}> 1$, and using a perfect sensitive test, thus $p_I=1$, there exists a range for $\rho$ over which $\pder\Rnum{\rho}\leq\,0$.
 Note that $\rho$ and $\omega$ have a similar mechanism in delaying people to get into $I\_c$, thus we would expect to see the non-trivial counter-vailing effect of these two parameters on $\Rnum$.
@@
 {\bf The third goal} is to explore how changes in the rate of test return $\omega$ affects $\Rnum$.
-Mathematically we would like to verify the sign of $\pder[\Rnum]{\omega}$, which specifically depends on $\pder[\Delta]{\omega}$.
+Mathematically we would like to verify the sign of $\pder\Rnum{\omega}$, which specifically depends on $\pder\Delta{\omega}$.
 We use
@@
 \label{eq:dlindo}
-\pder[\Delta]{\omega} = \frac{\rho}{\omega^2\,(\omega+\gamma)^2} \Big(
+\pder\Delta{\omega} = \frac{\rho}{\omega^2\,(\omega+\gamma)^2} \Big(
 (p_I\,\frac{w_I}{w_S}\theta\_c\,-(1+\frac{w_I}{w_S})\theta\_w)\omega^2+2\theta_w\,\gamma(\theta_w-2)\,\omega+\theta_w\,\gamma^2\,(\theta_w-2)
@@
 \begin{align}
-\pder[\Delta]{w\_{IS}}= \dfrac{(\omega-\rho)\,(\omega(\omega-\rho\theta\_w)+\gamma(\omega-\rho))\,(\theta\_w\gamma+\theta\_c\omega\,p\_I)}
+\label{eq:d_del_wis}
+\pder \Delta{w\_{IS}}= \dfrac{(\omega-\rho)\,(\omega(\omega-\rho\theta\_w)+\gamma(\omega-\rho))\,(\theta\_w\gamma+\theta\_c\omega\,p\_I)}
 {(-\omega^2\,\gamma+\omega\,\gamma\,\rho-\gamma\,\rho\,\omega\,w\_{IS}-
@@
 \end{align}
-which is a positive quantity. Thus, $\pder[\Rnum]{w\_{IS}}\leq 0$. Therefore, increasing the focus of testing on the infectious people will result in less transmission.
+which is a positive quantity. Thus, $\pder \Rnum{w\_{IS}}\leq 0$. Therefore, increasing the focus of testing on the infectious people will result in less transmission.
 

### 2021-05-31 codes/SIRfunctions.R

@@ -1,6 +1,3 @@
-library(shellpipes)
-rpcall("SIRfunctions.Rout SIRfunctions.R")
-
 # Functions for the SIR model
 # By: A Gharouni
 
@@ -19,6 +16,12 @@ while (!require(McMasterPandemic)) {
                           build_vignettes = TRUE
   )
 }
+if (!require("shellpipes") ) {
+  remotes::install_github("dushoff/shellpipes")
+}
+
+library(shellpipes)
+rpcall("SIRfunctions.Rout SIRfunctions.R")
 
 unpack <- McMasterPandemic::unpack
 

### 2021-05-31 codes/sir_plot.R

@@ -107,7 +107,7 @@ p1 <- (ggplot(df_temp,aes(x=1/omega,y=rho,z=Delta))
 p1_temp <- (p1
   # + geom_contour_filled(breaks=brks)
   + geom_contour_filled(breaks=brks_vec)
-  + geom_contour(breaks=threshold,alpha=0.5,colour="black",lty="11",lwd=1)
+  ##+ geom_contour(breaks=threshold,alpha=0.5,colour="black",lty="11",lwd=1)
   + facet_grid(theta_w~theta_c, labeller=label_special)
   + scale_x_continuous(expand=expansion(c(0,0)), breaks=inv_omega_brks,limits=c(inv_omega_min,inv_omega_max))
   ## scale by 1000, no digits after decimal
@@ -152,6 +152,7 @@ p1_targeted <- ((p1_temp %+% df_targeted)
   + ggtitle(TeX(sprintf("Targeted testing, w_S=%.1f, w_I=1",W_S_targeted)))
   ## REMOVE legend title here ... add it back in with ggdraw
   + theme(
+        axis.title.y=element_blank(),
         legend.title = element_blank(),
         legend.position = c(0.14, 0.30),
         legend.text = element_text(size = 7),
@@ -195,7 +196,7 @@ p12 <- (ggplot(df_temp2,aes(x=1/omega,y=rho,z=Delta))
 p1_temp2 <- (p12
             # + geom_contour_filled(breaks=brks)
             + geom_contour_filled(breaks=brks_vec)
-            + geom_contour(breaks=threshold,alpha=0.5,colour="black",lty="11",lwd=1)
+            ##+ geom_contour(breaks=threshold,alpha=0.5,colour="black",lty="11",lwd=1)
             + facet_grid(theta_w~theta_c, labeller=label_special)
             + scale_x_continuous(expand=expansion(c(0,0)), breaks=inv_omega_fast_brks,limits=c(inv_omega_fast_min,inv_omega_fast_max))
             ## scale by 1000, no digits after decimal
@@ -231,6 +232,7 @@ p1_targeted_h <- ((p1_temp2 %+% df_targeted_h)
                 + ggtitle(TeX(sprintf("Targeted testing, w_S=%.1f, w_I=1",W_S_targeted)))
                 ## REMOVE legend title here ... add it back in with ggdraw
                 + theme(
+                  axis.title.y=element_blank(),
                   legend.title = element_blank(),
                   legend.position = c(0.14, 0.30),
                   legend.text = element_text(size = 7),

### 2021-06-01 SIR_manuscript.tex
working on comments

@@
 In the left panel of \fref{pan2}, $\theta_w=0$/$\theta_c=0.75$ panel, there is a hint of this behaviour; it looks like the light green [second lightest] region displays this behaviour, but it's hard to see even a maximum magnification}.
-This phenomenon occurs --- increasing testing intensity \emph{increases} \Rnum --- because more rapid testing leaves more susceptibles in the ``waiting-for-negative-results'' category at the DFE; these people must then wait for their tests to be returned (during which time they may become infected) before they can be tested again, receive a positive test, and then begin self-isolating.
+\ali{Yes we can rule out this for random testing, (see Eqs.~\ref{eq:dd3dr} and \ref{eq:abcd2} and the brief analysis followed)}This phenomenon occurs --- increasing testing intensity \emph{increases} \Rnum --- because more rapid testing leaves more susceptibles in the ``waiting-for-negative-results'' category at the DFE; these people must then wait for their tests to be returned (during which time they may become infected) before they can be tested again, receive a positive test, and then begin self-isolating.
 This effect is usually weak compared to the beneficial effects of testing.
@@
 \label{eq:abcd2}
-a_3& = \frac{w_I}{w_S} \Bigg(&& (1-\theta\_w)\,(1+\frac{w_I}{w_S})\theta\_w\,\gamma^3
-+(1-\theta\_c)\,p_I^2\,\theta\_w\,\frac{w_I}{w_S}\,\omega^3 \\ \notag
-& ~ &&+\Big( \big( (1-\theta\_w-\frac{w_I}{w_S})\,\theta\_c +(3-2\theta\_w)\,\theta\_w\frac{w_I}{w_S} \big)\,p_I
-+(1-\theta\_w)\,(1+\frac{w_I}{w_S})\theta\_w  \Big)\,\omega\,\gamma^2 \\ \notag
+a_3& = \frac{w_I}{w_S} \Bigg(
+&& (1-\theta\_w)\,(1+\frac{w_I}{w_S})\theta\_w\,\gamma^3 +(1-\theta\_c)\,p_I^2\,\theta\_w\,\frac{w_I}{w_S}\,\omega^3 \\ \notag
+& ~ &&+\Big( \big( (1-\theta\_w-\frac{w_I}{w_S})\,\theta\_c +(3-2\theta\_w)\,\theta\_w\frac{w_I}{w_S} \big)\,p_I +(1-\theta\_w)\,(1+\frac{w_I}{w_S})\theta\_w  \Big)\,\omega\,\gamma^2 \\ \notag
 & ~ &&+\Big(
@@
 +(2\,\theta\_w-\theta\_w^2-\theta\_c)\,\frac{w_I}{w_S}\,p_I^2
-\Big)\,\omega^2\,\gamma \Bigg), \\
+\Big)\,\omega^2\,\gamma \Bigg), \\ \notag
 b_3& = && \hspace{-1cm} 2\frac{w_I}{w_S}\,(\omega+\gamma)\gamma\Big(
 (\omega+\gamma+\omega\,p_I)(2-\theta\_w)\,\gamma\,\theta\_w+ (1-\theta\_w)\,\omega^2 p_I\theta\_c+\omega^2 p_I\theta\_w
-\Big) ,\\
+\Big) ,\\ \notag
 c_3& = && \hspace{-1cm} (\omega+\gamma)^2\,\gamma\Big(
@@
 (1+\frac{w_I}{w_S})\,\omega\,\gamma\,\theta\_w+
-\frac{w_I}{w_S} \omega^2 p_I\theta\_c \Big),\\
+\frac{w_I}{w_S} \omega^2 p_I\theta\_c \Big),\\ \notag
 d_3&  = && \hspace{-1cm} \frac{(\omega+\gamma)}{\omega}\, \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi\,+\,(\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2.

### 2021-06-02 codes/SIRfunctions.R
parameter to LaTeX pipeline

@@ -197,5 +197,16 @@ make_params_dat<-function(params,
   return(dfout)
 }
 
+catt <- function(...,file="modeldefs.tex") {
+    cat(...,file=file,append=TRUE)
+}
+
+##' @param x numeric value
+##' @param nm name of macro (will be \nm in LaTeX)
+##' @param fmt number format (e.g. number of decimal places)
+latexout <- function(x,nm,fmt="%1.1f") {
+    catt(sprintf("\\newcommand{\\%s}{%s}\n",nm,sprintf(fmt,x)))
+}
+
 saveEnvironment()
 

### 2021-06-02 codes/sir_plot.R

@@ -9,6 +9,9 @@ library(cowplot) ## for ggdraw()/rectangle-hacking
 
 library(shellpipes)
 
+texfile <- "modeldefs.tex"
+unlink(texfile)
+
 rpcall("sir_plot.Rout sir_plot.R params.rda SIRfunctions.rda")
 
 loadEnvironments()
@@ -38,6 +41,13 @@ inv_omega_fast_min <- 0.5
 inv_omega_fast_max <- 5
 inv_omega_fast_brks <- c(1,2,4)
 rho_fast_brks <- c(0,0.1,0.2)
+
+unlink("modeldefs.tex")
+
+catt("% parameter values\n")
+latexout(params[["beta"]],"betaparam")
+latexout(params[["gamma"]],"gammaparam")
+
 
 background <- "grey90"
 # make dataframe
@@ -218,8 +228,8 @@ print(p1_temp2)
 # 4. Plot the Random Testing Scenario:
 # #################################
 
-ggsave(p1_temp2 
-       + ggtitle(TeX('(Random testing, $w_S=w_I=1$)')) 
+ggsave(p1_temp2
+       + ggtitle(TeX('(Random testing, $w_S=w_I=1$)'))
        + theme(legend.position = "none"),
        filename = "R0contour_random2.pdf" ,
        width = 12, height = 12, units = "cm")

### 2021-06-02 SIR_manuscript.tex
parameter to LaTeX pipeline
params changes, beta and gamma to more realistic ones

@@
 \usepackage{enumerate}
+
+%% FIXME: uncomment when this pipeline is working
+%% \input{codes/modeldefs.tex}
 
@@
 \section{Introduction}
+
+%% FIXME: uncomment when pipeline works
+%% $\beta=\betaparam$; $\gamma=\gammaparam$
 
@@
   $\omega$  & Rate of test return, i.e., rate of onward flow from ``waiting'' to ``confirmed'' or ``untested'' compartments  & 1/day & - \\ \hline
-  $\gamma$ & Recovery rate & 1/day & 1/3 \\ \hline 
+  $\gamma$ & Recovery rate & 1/day & 1/6 \\ \hline 
   $\rho$   & \percap testing intensity & 1/day & - \\ \hline 
@@
   $\theta\_c$ & Isolation efficacy for ``confirmed positive'' individuals & - & -  \\ \hline
-  $\beta$ & Transmission rate & 1/day & 0.339 \\ \hline
+  $\beta$ & Transmission rate & 1/day & 0.25 \\ \hline
   $\Lambda$ & Force of infection & 1/day & - \\ \hline
@@
 \end{equation}
-\bmb{super-fussy, but: I don't think we should introduce explicit thinspaces everywhere; why not let
-  TeX do the typesetting itself? Do we really prefer more compact equations, and if so is there a package
-  that will just change the appropriate settings globally?}\ali{are we talking about \\,?}
 where
@@
   The critical contour of $\Delta=1-\frac{\gamma}{\beta}$, corresponding to the threshold $\Rnum=1$, is shown as a dotted line in \fref{pan} and \fref{pan2}; combinations of $\rho$ and $\omega$ that lie above this critical contour will control the epidemic.
-The implied baseline value of $\Rnum=\frac{\beta}{\gamma}$ in these figures is $\approx 1.017$, corresponding to a case where the epidemic spreads slowly even in the absence of testing and tracing.
-\bmb{how did we choose this particular value?}\ali{I was thinking of ``critical'' $\Rnum=1$ and wanted to have the baseline in the plot.}
-  \bmb{JD wonders whether we really need the ``critical'' $\Rnum=1$ contour or not; any contour will do to illustrate a threshold level of testing intensity, and the fact that $\Rnum=1$ falls in the middle of our range is a bit arbitrary (i.e.
-the fact that testing can control the epidemic depends on the baseline $\Rnum$ value we have chosen).
-(I know it was my idea in the first place, but maybe we could dispense with it?)}\ali{agree, the plots are updated.}
+The implied baseline value of $\Rnum=\frac{\beta}{\gamma}$ in these figures is $\approx 1.5$, corresponding to a case where the epidemic spreads slowly even in the absence of testing and tracing.
 
@@
 Numerical evaluation of the effectiveness control parameter, $\Delta$ (Eq.~\ref{eq:del4}), with respect to changes in testing and isolation. Parameter values (Table~\ref{tab:params}):
-$\beta=0.339$ day$^{-1}$ $\gamma= 1/3$ day$^{-1}$; $\omega \in [1/12,2]$ day$^{-1}$;  $\rho \in [0,0.013]$ day$^{-1}$; $\theta\_w$ and $\theta\_c$ vary between 0 (no effect of isolation) and 1 (complete elimination of transmission); $p_S=0$, $p_I=1$ and $p_R=0.5$.
+$\beta=0.25$ day$^{-1}$, $1/\gamma= 1/3$ days; $\omega \in [1/12,2]$ day$^{-1}$;  $\rho \in [0,0.013]$ day$^{-1}$; $\theta\_w$ and $\theta\_c$ vary between 0 (no effect of isolation) and 1 (complete elimination of transmission); $p_S=0$, $p_I=1$ and $p_R=0.5$.
 Contours of $\Delta$ are plotted for (a) random testing ($w_S=w_I=w_R=1$) and (b) targeted testing ($w_S=0.3$; $w_I=w_R=1$).
-The baseline basic reproduction number $\Rnum$ and the doubling time $r$, i.e. $\beta/\gamma$ and $\beta-\gamma$ respectively, are approximately 1.017 and 0.006 in this simulation.
+The baseline basic reproduction number $\Rnum$ and the doubling time $r$, i.e. $\beta/\gamma$ and $\beta-\gamma$ respectively, are approximately 1.5 and 0.08 in this simulation.
 }
@@
 We use the following parameters (see Table~\ref{tab:params} for a brief description):
-$\omega \in [1/5,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,1/5)$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.339$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
+$\omega \in [1/5,2]$ 1/day, $1/\gamma= 6$ days, $\rho \in [0,1/5)$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.25$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
 Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
 The baseline basic reproduction number $\Rnum$ and the doubling time $r$, i.e.
-$\beta/\gamma$ and $\beta-\gamma$ respectively, are approximately 1.017 and 0.006 in this simulation.
-\bmb{consider removing dotted line (see note elsewhere); on this and prev figure consider removing redundant y-axis label on right panel (and adjusting/squashing interpanel spacing)?}
-\ali{ok?}
+$\beta/\gamma$ and $\beta-\gamma$ respectively, are approximately 1.5 and 0.08 in this simulation.
 }

### 2021-06-02 codes/params.R

@@ -5,8 +5,8 @@ loadEnvironments()
 
 params <- c(
 N0=1000000, #total population size, 
-beta= 0.339, #transmission rate 0.339
-gamma=1/3, #recovery rate, was 1/3
+beta= 0.25, #transmission rate 0.339
+gamma=1/5, #recovery rate, was 1/3
 omega=0.25, #test returning rate 
 rho=0.01, #testing rate percapita (also used 0.8, 1/3)
 W_S=1, W_I=1, W_R=1, #Testing relative weights

### 2021-06-03 codes/sir_plot.R
the rho range in high testing was above omega, caused numerical artifact, fixed now

@@ -30,7 +30,7 @@ W_S_random <- 1
 W_S_targeted <- 0.3
 ## tweak to get non-overlapping axis labels:
 ##  adjust grid to be slightly bigger than ideal tick marks
-inv_omega_min <- 0.5
+inv_omega_min <- 0.5 ##inverse of omega in days
 inv_omega_max <- 12
 inv_omega_brks <- c(1,5,10)
 rho_min <- 0
@@ -40,7 +40,9 @@ rho_brks <- c(0,0.005,0.01)
 inv_omega_fast_min <- 0.5
 inv_omega_fast_max <- 5
 inv_omega_fast_brks <- c(1,2,4)
-rho_fast_brks <- c(0,0.1,0.2)
+rho_fast_min <- 0
+rho_fast_max <- (1/inv_omega_fast_max)-0.000001
+rho_fast_brks <- c(0,0.1,0.17)
 
 unlink("modeldefs.tex")
 
@@ -68,13 +70,15 @@ df_random_h <- make_params_dat(params = params,
                              eta_ws=0,eta_we=1, ## so theta_w
                              eta_cs=0,eta_ce=1, ## so theta_c
                              omega_s=1/inv_omega_fast_max,omega_e=1/inv_omega_fast_min,
-                             rho_s=0,rho_e=0.25-0.000001)
+                             rho_s=rho_fast_min,rho_e=rho_fast_max)
+                             
 
 df_targeted_h <- make_params_dat(params = update(params,W_S=W_S_targeted),
                                eta_ws=0,eta_we=1, ## so theta_w
                                eta_cs=0,eta_ce=1, ## so theta_c
                                omega_s=1/inv_omega_fast_max,omega_e=1/inv_omega_fast_min,
-                               rho_s=0,rho_e=0.25-0.000001)
+                               rho_s=rho_fast_min,rho_e=rho_fast_max)
+                               
 
 ##important contour, ie R0=1 thus threshold=1 when plotting R0 contours, or Delta(R0=1)
 threshold <- 1-(params[["gamma"]]/params[["beta"]]) ## corresponding to R0=1
@@ -265,3 +269,28 @@ p1_targeted_hacked <- (ggdraw(p1_targeted_h)
 ggsave(p1_targeted_hacked,
        filename = "R0contour_TTI2.pdf" ,
        width = 12, height = 12, units = "cm")
+
+# #################################
+# 6. Check: in random testing when rho is high, $\theta_w=0$/$\theta_c=0.75$ panel
[…]

### 2021-06-03 SIR_manuscript.tex
slightly improve make rule for modeldefs
the rho range in high testing was above omega, caused numerical artifact, fixed now
meeeting with ali

@@
 %% FIXME: uncomment when this pipeline is working
-%% \input{codes/modeldefs.tex}
+\input{codes/modeldefs.tex}
 
@@
 %% FIXME: uncomment when pipeline works
-%% $\beta=\betaparam$; $\gamma=\gammaparam$
+$\beta=\betaparam$; $\gamma=\gammaparam$
 
@@
 \centering
-\begin{subfigure}[t]{.46\textwidth}
+\begin{subfigure}[t]{.49\textwidth}
 \centering
@@
 %
-\begin{subfigure}[t]{.44\textwidth}
+\begin{subfigure}[t]{.49\textwidth}
 \centering
@@
 \item \fref{pan} also shows that greater testing intensity (increasing $\rho$) generally increases the effectiveness of control (moving up in each panel).
-This relationship can be reversed, but only for an unrealistic scenario --- very high testing intensities \emph{that begin before the epidemic starts}.
-It occurs when waiting people do not isolate ($\theta_w=0$), and most easily for targeted testing (\fref{pan2}(b), right three panels of top row) \bmb{can we rule out the possibility that this phenomenon could happen with random testing?
-In the left panel of \fref{pan2}, $\theta_w=0$/$\theta_c=0.75$ panel, there is a hint of this behaviour; it looks like the light green [second lightest] region displays this behaviour, but it's hard to see even a maximum magnification}.
-\ali{Yes we can rule out this for random testing, (see Eqs.~\ref{eq:dd3dr} and \ref{eq:abcd2} and the brief analysis followed)}This phenomenon occurs --- increasing testing intensity \emph{increases} \Rnum --- because more rapid testing leaves more susceptibles in the ``waiting-for-negative-results'' category at the DFE; these people must then wait for their tests to be returned (during which time they may become infected) before they can be tested again, receive a positive test, and then begin self-isolating.
+This relationship can be reversed, however, at very high testing intensities.
+This can only occur when testing is targeted, and $\theta_w$ is relatively small (\fref{pan2}(b), right three panels of top row).
+It is theoretically possible for increasing testing intensity to \emph{increase} \Rnum because more rapid testing leaves more susceptibles in the ``waiting-for-negative-results'' category at the DFE; if these people become infected while waiting, they will need to wait for their negative test result before they can be tested again, receive a positive test, and then begin self-isolating.
 This effect is usually weak compared to the beneficial effects of testing.
-  \bmb{I am still trying to understand why slowing testing return (decreasing $\omega$) decreases \Rnum over \textbf{most} of the parameter space considered here, which is a counterintuitive/silly result.
-Fast test returns help when we have high test positivity and low isolation during the waiting period (specifically, [much] lower isolation while waiting than when confirmed, but I think we're assuming that); we want to get waiting people into the ``confirmed'' compartment as quickly as possible.
-Can we understand/explain this result more thoroughly using that logic, and can we argue that what we are looking at is \emph{not} a realistic parameter range?
-Does the low baseline \Rnum we are using play into this, e.g.
-by reducing test positivity?
-Is there a simple expression for test positivity during the invasion phase?}
 

### 2021-06-03 codes/params.R
increase betra, R) = 2.5 now

@@ -5,7 +5,7 @@ loadEnvironments()
 
 params <- c(
 N0=1000000, #total population size, 
-beta= 0.25, #transmission rate 0.339
+beta= 0.5, #transmission rate 0.339
 gamma=1/5, #recovery rate, was 1/3
 omega=0.25, #test returning rate 
 rho=0.01, #testing rate percapita (also used 0.8, 1/3)

### 2021-06-03 SIR_manuscript.tex
massaging with Ali

@@
 \begin{itemize}
-\item When waiting individuals do not isolate (e.g.
-$\theta_w =0$, top row of \fref{pan}), increasing testing delays (i.e.
-decreasing $\omega$) increases \Rnum, as we might expect: $\Delta$ decreases --- a larger region of the parameter space has $\Rnum>1$, and the threshold testing intensity required to control the epidemic increases --- as we move to the right within each plot in this row.
-However, when waiting individuals are at least partially isolated ($\theta_w>0$), we typically see the opposite effect: longer testing delays lead to \emph{less} transmission.
-This is because returning negative tests leads people to stop distancing; this applies both to susceptibles, and to infectious people who receive negative test results because they were sampled when susceptible (or because of imperfect test sensitivity, which we have not considered in our analysis).
-For small but non-zero waiting-isolation levels (e.g.
-$\theta_w=0.25$, second row from top), with high levels of confirmed isolation ($\theta_c \ge 0.5$), increasing testing delays initially decreases $\Delta$, then increases it as the testing delays pass a threshold value of approximately 2--5 days.
+\item We would generally expect increasing testing delays to increase \Rnum, thus decreasing effectiveness of control $\Delta$.
+This is in fact what we find when waiting individuals do not isolate ($\theta_w =0$, top row of \fref{pan}) --- as we move to the right within each plot in this row, $\Delta$ decreases.
+However, when waiting individuals isolate ($\theta_w>0$), we more often see the opposite effect: longer testing delays lead to a greater control effect $\Delta$.
+This is because people waiting for negative tests are assumed to continue to isolate; this applies both to susceptibles and to people who became infected while waiting for negative test results.
+This effect outweighs the effect of confirmed individuals isolating except when this isolation parameter $\theta_c$ is substantially bigger than $\theta_w$.
+This result depends on the idea that, all else equal, people who have to wait longer for test results isolate at the same level (but for a longer time) as they would if the wait were shorter.
 \item \fref{pan} also shows that greater testing intensity (increasing $\rho$) generally increases the effectiveness of control (moving up in each panel).
@@
 In particular, we realize higher $\Delta$ levels, i.e., darker colors in corresponding plots\/facets as we go from panel~(a) to panel~(b) in the figures.
-\bmb{this is the same as point 4 above, right?
-Am I missing something?}
-\ali{you are right.}
+\todo{Ali to restructure or delete; feels redundant.}
 \end{itemize}

### 2021-06-03 SIR_manuscript.tex
working on discussion

@@
 %%%%%%%%%%%%
+
 \section{Discussion}
 
-\begin{enumerate}
+In this paper, we have defined and analyzed a simple compartmental model that combines epidemiological dynamics (as defined by a simple SIR model) with the dynamics of testing and isolation.
+Our model is a caricature --- it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g.
+exposed, presymptomatic, and asymptomatic compartments [REFS]; time-varying testing rates [REFS?]; behavioural dynamics [REF Weitz/Park/Dushoff]).
+Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between epidemiology and testing, rather than for making quantitative predictions or guiding pandemic responses.
+
+Many of the qualitative results we have derived are mathematical confirmations of simple, common-sense intuitions.
+In particular, we can decrease $\Rnum$ by increasing isolation efficacy or testing intensity; returning tests faster, if individuals do not isolate while they are waiting for results; or increasing testing focus to target individuals who are likely to be infectious (e.g.
+symptomatic people or close contacts of known infections).
 

### 2021-06-04 SIR_manuscript.tex

@@
   $\theta\_c$ & Isolation efficacy for ``confirmed positive'' individuals & - & -  \\ \hline
-  $\beta$ & Transmission rate & 1/day & 0.25 \\ \hline
+  $\beta$ & Transmission rate & 1/day & 0.5 \\ \hline
   $\Lambda$ & Force of infection & 1/day & - \\ \hline
@@
   The critical contour of $\Delta=1-\frac{\gamma}{\beta}$, corresponding to the threshold $\Rnum=1$, is shown as a dotted line in \fref{pan} and \fref{pan2}; combinations of $\rho$ and $\omega$ that lie above this critical contour will control the epidemic.
-The implied baseline value of $\Rnum=\frac{\beta}{\gamma}$ in these figures is $\approx 1.5$, corresponding to a case where the epidemic spreads slowly even in the absence of testing and tracing.
+The implied baseline value of $\Rnum=\frac{\beta}{\gamma}$ in these figures is 3, corresponding to a case where the epidemic spreads slowly even in the absence of testing and tracing.
 
@@
 Numerical evaluation of the effectiveness control parameter, $\Delta$ (Eq.~\ref{eq:del4}), with respect to changes in testing and isolation. Parameter values (Table~\ref{tab:params}):
-$\beta=0.25$ day$^{-1}$, $1/\gamma= 1/3$ days; $\omega \in [1/12,2]$ day$^{-1}$;  $\rho \in [0,0.013]$ day$^{-1}$; $\theta\_w$ and $\theta\_c$ vary between 0 (no effect of isolation) and 1 (complete elimination of transmission); $p_S=0$, $p_I=1$ and $p_R=0.5$.
+$\beta=0.5$ 1/day, $1/\gamma= 6$ days; $\omega \in [1/12,2]$ 1/day;  $\rho \in [0,0.013]$ 1/day per capita; $\theta\_w$ and $\theta\_c$ vary between 0 (no effect of isolation) and 1 (complete elimination of transmission); $p_S=0$, $p_I=1$ and $p_R=0.5$.
 Contours of $\Delta$ are plotted for (a) random testing ($w_S=w_I=w_R=1$) and (b) targeted testing ($w_S=0.3$; $w_I=w_R=1$).
-The baseline basic reproduction number $\Rnum$ and the doubling time $r$, i.e. $\beta/\gamma$ and $\beta-\gamma$ respectively, are approximately 1.5 and 0.08 in this simulation.
+The baseline basic reproduction number $\Rnum$ and the doubling time $r$, i.e. $\beta/\gamma$ and $\beta-\gamma$ respectively, are 3 and about 0.3 in this simulation.
 }
@@
 We use the following parameters (see Table~\ref{tab:params} for a brief description):
-$\omega \in [1/5,2]$ 1/day, $1/\gamma= 6$ days, $\rho \in [0,1/5)$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.25$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
+$\omega \in [1/5,2]$ 1/day, $1/\gamma= 6$ days, $\rho \in [0,1/5)$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.5$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
 Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
 The baseline basic reproduction number $\Rnum$ and the doubling time $r$, i.e.
-$\beta/\gamma$ and $\beta-\gamma$ respectively, are approximately 1.5 and 0.08 in this simulation.
+$\beta/\gamma$ and $\beta-\gamma$ respectively, are 3 and 0.3 in this simulation.
 }
@@
 
-\item Lastly, we show that increasing the focus of testing from random to targeted, always increases $\Delta$, thus decreases \Rnum.
-This can be seen in \fref{pan} and \fref{pan2}.
-In particular, we realize higher $\Delta$ levels, i.e., darker colors in corresponding plots\/facets as we go from panel~(a) to panel~(b) in the figures.
-\todo{Ali to restructure or delete; feels redundant.}
 \end{itemize}

### 2021-06-06 SIR_manuscript.tex
collapse of G to R0

@@
 %%%%%%%%%%%%
-
 \section{Discussion}
 
-In this paper, we have defined and analyzed a simple compartmental model that combines epidemiological dynamics (as defined by a simple SIR model) with the dynamics of testing and isolation.
-Our model is a caricature --- it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g.
-exposed, presymptomatic, and asymptomatic compartments [REFS]; time-varying testing rates [REFS?]; behavioural dynamics [REF Weitz/Park/Dushoff]).
-Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between epidemiology and testing, rather than for making quantitative predictions or guiding pandemic responses.
-
-Many of the qualitative results we have derived are mathematical confirmations of simple, common-sense intuitions.
-In particular, we can decrease $\Rnum$ by increasing isolation efficacy or testing intensity; returning tests faster, if individuals do not isolate while they are waiting for results; or increasing testing focus to target individuals who are likely to be infectious (e.g.
-symptomatic people or close contacts of known infections).
+\begin{enumerate}
 

### 2021-06-06 cohort.tex
minor change
collapse of G to R0

@@
 \newtheorem{theorem}{Theorem}
+\newtheorem{note}{Note}
 
@@
 \label{eq:ng} %next generation
-\vec G = FV^{-1} = \int_0^\infty \vec K(\tau) d\tau.
+\vec G = FV^{-1} \vec I(0)= \int_0^\infty \vec K(\tau) d\tau.
 \end{equation}
@@
 \end{enumerate}
-Note that in the compartment framework \cite{van2002reproduction}, steps 2 and 3 are combined.
+
+\begin{note}
+The following notes/points are to be considered in moving from one framework to another, namely, cohort framework \citep{champredon2018equivalence} and comartmental framework \citep{van2002reproduction};
+\begin{enumerate}
+\item
+In the compartment framework, steps 2 and 3 are combined.
+\item
+If the initial condition $\vec I(0)$ of the cohort Eq.~\eqref{eq:I} (in the cohort framework) is chosen as the principal eigenvector of the next-generation matrix $FV^{-1}$ corresponding to $\R$ (in compartmental framework), the next generation vector $\vec G$~\eqref{eq:ng} would be
+\begin{equation}
+\vec G = FV^{-1} \vec I(0) = \R \vec I(0).
+\end{equation}
+Thus, in this case inorder to collapse $\vec G$ to $\R$, if $\vec I(0)$ is a unit vector as well, i.e., $\lVert \vec I(0) \rVert = 1$, it follows that $\vec I(0)^t \cdot \vec G = \R$, where $\vec I(0)^t$ is the transpose of $\vec I(0)$.
+\end{enumerate}
+\end{note}
 
@@
 \vec I'=-V\,\vec I,~~ \text{where}~~ \vec I(\tau)=\left[ \begin {array}{c} I\_u(\tau)\\I_n(\tau)\\I_p(\tau)\\I_c(\tau) \end {array} \right], ~~
-\vec I(0)=\left[ \begin {array}{c} S\_u^*\\S\_n^*\\0\\0 \end {array} \right].
+\vec I(0)=\left[ \begin {array}{c} S\_u^*\\(1-\theta\_w)S\_n^*\\0\\0 \end {array} \right].
 \end{align}

### 2021-06-07 SIR_manuscript.tex
clean up broken itemized list

--- 2021-06-03 SIR_manuscript.tex
@@
 
-\item {\bf A potential advantage of slow test reporting $\omega$, result~\ref{p1:omega};}
+{\bf A potential advantage of slow test reporting $\omega$, result~\ref{p1:omega};}
 
@@
 
-\item {\bf The counter-vailing effect of \percap testing intensity, $\rho$, on $\Rnum$;}
+{\bf The counter-vailing effect of \percap testing intensity, $\rho$, on $\Rnum$;}
 - The explanation here is that more rapid testing leaves more susceptibles in the ``waiting-for-no'' $S\_n$ category at the DFE and delay them to get to $I\_c$ where they are subject to a full isolation, thus no transmission.
@@
 
-\item {\bf Specifying the compartment-specific relative testing weights, Specifically $\frac{w_I}{w_S}$;}
+{\bf Specifying the compartment-specific relative testing weights, Specifically $\frac{w_I}{w_S}$;}
 
@@
 
-\item What can we say in the context of strength and speed paradigm?
-
-\end{enumerate}
+What can we say in the context of strength and speed paradigm?
 

### 2021-06-07 SIR_manuscript.tex
LaTeX cleanup
recover discussion PP

@@
 
-%% FIXME: uncomment when this pipeline is working
 \input{codes/modeldefs.tex}
@@
 
-\begin{enumerate}
+In this paper, we have defined and analyzed a simple compartmental model that combines epidemiological dynamics (as defined by a simple SIR model) with the dynamics of testing and isolation.
+Our model is a caricature --- it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g.
+exposed, presymptomatic, and asymptomatic compartments [REFS]; time-varying testing rates [REFS?]; behavioural dynamics [REF Weitz/Park/Dushoff]).
+Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between epidemiology and testing, rather than for making quantitative predictions or guiding pandemic responses.
+
+Many of the qualitative results we have derived are mathematical confirmations of simple, common-sense intuitions.
+In particular, we can decrease $\Rnum$ by increasing isolation efficacy or testing intensity; returning tests faster, if individuals do not isolate while they are waiting for results; or increasing testing focus to target individuals who are likely to be infectious (e.g.
+symptomatic people or close contacts of known infections).
 

### 2021-06-07 cohort.tex
details added to the SIR withtesting example

@@
 \renewcommand{\vec}[1]{\ensuremath{\mathbf{#1}}} %bold font for vector
+\newcommand{\e}{\ensuremath{{\rm e}}}
 
@@
 \item
-If the initial condition $\vec I(0)$ of the cohort Eq.~\eqref{eq:I} (in the cohort framework) is chosen as the principal eigenvector of the next-generation matrix $FV^{-1}$ corresponding to $\R$ (in compartmental framework), the next generation vector $\vec G$~\eqref{eq:ng} would be
+If the initial condition $\vec I(0)$ of the cohort Eq.~\eqref{eq:I} (in the cohort framework) is chosen as the normalized principal eigenvector of the next-generation matrix $FV^{-1}$ corresponding to $\R$ (in compartmental framework), the next generation vector $\vec G$~\eqref{eq:ng} would be
 \begin{equation}
@@
 \end{equation}
-Thus, in this case inorder to collapse $\vec G$ to $\R$, if $\vec I(0)$ is a unit vector as well, i.e., $\lVert \vec I(0) \rVert = 1$, it follows that $\vec I(0)^t \cdot \vec G = \R$, where $\vec I(0)^t$ is the transpose of $\vec I(0)$.
+Thus, in this case inorder to collapse $\vec G$ to $\R$, when $\vec I(0)$ is the normalized initial distribution of infected individuals, i.e., $\lVert \vec I(0) \rVert = 1$, it follows that $\vec I(0)^t \cdot \vec G = \R$, where the row vector $\vec I(0)^t$ is the transpose of $\vec I(0)$.
 \end{enumerate}
@@
 \end{align}
+Note that, $\R =\vec I(0)^t \cdot \vec G =\beta/\gamma$.
 
@@
 \vec I'=-V\,\vec I,~~ \text{where}~~ \vec I(\tau)=\left[ \begin {array}{c} I\_u(\tau)\\I_n(\tau)\\I_p(\tau)\\I_c(\tau) \end {array} \right], ~~
-\vec I(0)=\left[ \begin {array}{c} S\_u^*\\(1-\theta\_w)S\_n^*\\0\\0 \end {array} \right].
+\vec I(0)=\left[ \begin {array}{c} \frac{S\_u^*}{(1-\theta\_w)S\_n^*}\\1\\0\\0 \end {array} \right].
 \end{align}
@@
 \vec I(\tau)=&
-c_1\,\exp(-\gamma\,\tau)\,\left[ \begin {array}{c} 0\\0\\0\\1 \end {array} \right] 
-+c_2\,\exp(-(F_I+\gamma)\,\tau)\,\left[ \begin {array}{c} \frac{F_I-\omega}{\omega}\\0\\\frac{-F_I}{\omega}\\1 \end {array} \right]
-+c_3\,\exp(-(\omega+\gamma)\,\tau)\,\left[ \begin {array}{c} 0\\0\\-1\\1 \end {array} \right] \\\notag
-+& c_4\,\exp(-(\omega+\gamma)\,\tau)\,\Big(\tau\,\left[ \begin {array}{c} 0\\0\\-1\\1 \end {array} \right]+
-\left[ \begin {array}{c} \frac{-1}{F_I}\\ \frac{\omega-F_I}{F_I\,\omega}\\ \frac{1}{\omega}\\0 \end {array} \right]\Big),
+\left[ \begin {array}{cccc}  
+\frac{F_I-\omega}{\omega} \e^{-(F_I+\gamma)\tau}&0&0&\frac{-1}{F_I}\e^{-(\omega+\gamma)\tau}\\
+0&0&0&\frac{F_I-\omega}{-F_I\omega} \e^{-(\omega+\gamma)\tau}\\
+\frac{-F_I}{\omega} \e^{-(F_I+\gamma)\tau}&0&-\e^{-(\omega+\gamma)\tau}&(\frac{1}{\omega}-\tau)\e^{-(\omega+\gamma)\tau}\\
+\e^{-(F_I+\gamma)\tau}&\e^{-\gamma\tau}&\e^{-(\omega+\gamma)\tau}&\tau\e^{-(\omega+\gamma)\tau}
+\end {array} \right] \vec c,
 \end{align}
@@
 \begin{align}
-c_1=&1\\
-c_2=&\frac{\omega}{F_I-\omega}(S\_u^*-\frac{\omega}{F_I-\omega}\,S\_n^*)\\
-c_3=&\frac{-F_I}{F_I-\omega}(1-\frac{\omega}{F_I-\omega}S\_n^*)\\
-c_4=&\frac{-\omega\,F_I}{F_I-\omega}S\_n^*.
+\vec c =
+\left[ \begin {array}{c} -{\frac {\omega\, \left(  \left( -\omega\,
+\theta_{{w}}+{\it F_I} \right) {\it S\_u}+\omega\,\theta_{{w}}-\omega
+ \right) }{ \left( {\it F_I}-\omega \right) ^{2} \left( -1+\theta_{{w
+}} \right) {  S\_n}}}\\ \noalign{\medskip}{\frac {-{ S\_u}\,
+\theta_{{w}}-1+\theta_{{w}}}{ \left( -1+\theta_{{w}} \right) { S\_n
+}}}\\ \noalign{\medskip}-{\frac {{ F_I}\, \left(  \left( -\omega-{
+ F_I}\,\theta_{{w}}+2\,\omega\,\theta_{{w}} \right) { S\_u}-{
+ F_I}+{ F_I}\,\theta_{{w}}+2\,\omega-2\,\omega\,\theta_{{w}}
+ \right) }{ \left( { F_I}-\omega \right) ^{2} \left( -1+\theta_{{w
+}} \right) { S\_n}}}\\ \noalign{\medskip}-{\frac {{ F_I}\,
+\omega}{{ F_I}-\omega}}\end {array} \right] .
 \end{align}
@@
 \vec G=&\int_0^\infty \vec K(\tau) d\tau\\
-=&
+=& \R\, \vec I(0), 
 \end{align}
+where $\R$ is calculated in the previous work.
 %%%%%%%%%%%%%%%%%%%%

### 2021-06-08 SIR_manuscript.tex
added 2 refs

@@
 Our model is a caricature --- it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g.
-exposed, presymptomatic, and asymptomatic compartments [REFS]; time-varying testing rates [REFS?]; behavioural dynamics [REF Weitz/Park/Dushoff]).
+exposed, presymptomatic, and asymptomatic compartments [REFS]; time-varying testing rates [REFS?]\ali{\citep{jarvis2021temporal} OK?}; behavioural dynamics [REF Weitz/Park/Dushoff] \ali{\citep{weitz2020awareness} OK?}).
 Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between epidemiology and testing, rather than for making quantitative predictions or guiding pandemic responses.

### 2021-06-08 SIR_manuscript.tex
added 1 ref

--- 2021-06-07 SIR_manuscript.tex
@@
 Our model is a caricature --- it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g.
-exposed, presymptomatic, and asymptomatic compartments [REFS]; time-varying testing rates [REFS?]; behavioural dynamics [REF Weitz/Park/Dushoff]).
+exposed, presymptomatic, and asymptomatic compartments [REFS]\ali{\citep{kain2021chopping} OK?}; time-varying testing rates [REFS?]\ali{\citep{jarvis2021temporal} OK?}; behavioural dynamics [REF Weitz/Park/Dushoff] \ali{\citep{weitz2020awareness} OK?}).
 Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between epidemiology and testing, rather than for making quantitative predictions or guiding pandemic responses.

### 2021-06-09 SIR_manuscript.tex
writing discussion

--- 2021-06-07 SIR_manuscript.tex
@@
 \newcommand*\subtxt[1]{_{\textnormal{#1}}}
+%% Roman subscript
 \DeclareRobustCommand\_{\ifmmode\expandafter\subtxt\else\textunderscore\fi}
@@
 The realistic components of the model include \percap testing intensity, test sensitivity and specificity, rate of test return, and isolation efficacy (i.e., reduction of the transmission probability by isolating individuals).
-The novel component is the compartment-specific relative testing weights, which reflect the testing strategies-- surveillance, diagnosis, or control.
+The novel component is the compartment-specific relative testing weights, which reflect the testing strategies --- surveillance, diagnosis, or control.
 Here, we compare two testing strategies, random vs.
@@
 \label{F}
-F_X=\sigma w_X,\,\, \text{where}\, X \in \{S,I,R\}.
+F_X=\sigma w_X, \quad \text{where} \quad X \in \{S,I,R\}.
 \end{equation}
@@
 
-The Disease-Free Equilibrium (DFE) for the SIR model (Eqs.\,\ref{model}) is found by setting the infected compartments to 0 and solving for the unknowns. The DFE is
+The Disease-Free Equilibrium (DFE) for the SIR model (Eqs.~\ref{model}) is found by setting the infected compartments to 0 and solving for the unknowns. The DFE is
 \begin{equation}
 \label{dfe}
-S\_n^*= \frac{\rho}{\omega}\,N_0, \ S\_u^*= \frac{\omega-\rho}{\omega}\,N_0, \text{~and~} I_j=R_j=0 \ \text{for all }j.
+S\_n^*= \frac{\rho}{\omega} N_0, \ S\_u^*= \frac{\omega-\rho}{\omega} N_0, \text{~and~} I_j=R_j=0 \ \text{for all }j.
 \end{equation}
-The corresponding \percap testing rate (Eq.\,\ref{F}) for the infected compartment $I$ at DFE is one of the key analysis parameters and can be simplified as 
+The corresponding \percap testing rate (Eq.~\ref{F}) for the infected compartment $I$ at DFE is one of the key analysis parameters and can be simplified as 
 \begin{equation}
@@
 
-The basic reproduction number, $\Rnum$, was calculated by using the next-generation matrix method \citep{van2002reproduction}. We present $\Rnum$ in the following form.
+The basic reproduction number, $\Rnum$, was calculated by using the next-generation matrix method \citep{van2002reproduction}. We present $\Rnum$ as
 \begin{equation}
 \label{R0}
-\Rnum= \frac{\beta}{\gamma} (1-\Delta), 
+\Rnum= \frac{\beta}{\gamma} \left(1-\Delta\right), 
 \end{equation}
-where the term $\frac{\beta}{\gamma}$ is the classical basic reproduction number for a SIR model without testing and isolation \citep{keeling2011modeling}, and the positive quantity $\Delta$ is the effectiveness of control due to testing and isolation processes defined as
+where the term $\beta/\gamma$ is the classical basic reproduction number for a SIR model without testing and isolation \citep{keeling2011modeling}, and $\Delta>0$ is the effectiveness of control due to testing and isolation processes defined as
 \begin{equation}
@@
 \end{align}
-Further details of derivation of $\Rnum$ are provided in appendix, Sec.\,\ref{app:R0}. 
+(Appendix~\ref{app:R0} gives a detailed derivation of these expressions).
 The explicit formula of $\Rnum$ enables us to study the effects of testing and isolation parameters on $\Rnum$ both analytically and via numerical solutions.
 We are specifically interested in parameters that could be manipulated by public health policy: isolation efficacy, $\theta\_c$ and $\theta\_w$; \percap testing intensity, $\rho$; and the rate of test return, $\omega$.
-In particular, we look at the partial derivatives of $\Delta$ with respect to these parameters (Sec.\,\ref{app:rho}).
+In particular, we look at the partial derivatives of $\Delta$ with respect to these parameters (App.~\ref{app:rho}).
 We derived general expressions for these derivatives.
@@
 
-The low-testing case \fref{pan} reflects the case where $rho$ is small relative to the population size.
+The low-testing case \fref{pan} reflects the case where $\rho$ is small relative to the population size.
 Specifically, $\rho \in [0,0.013]$, and test return rate $\omega\in [1/12,2]$.
@@
 % %%%%%%%
-\section{Results and discussion}
+\section{Results}
 
-We can use the formula for $\Rnum$ \eqref{R0} to make a number of straightforward conclusions about parameters whose effects on $\Rnum$ are monotonic (i.e.
-the partial derivative of $\Delta$ has a consistent sign: Sections~\ref{app:rho}, \ref{app:omega}, \ref{app:w}).
+We can use the formula for $\Rnum$ \eqref{R0} to make a number of straightforward conclusions about parameters whose effects on $\Rnum$ are monotonic, i.e.
+when the partial derivative of $\Delta$ has a consistent sign
+(Appendices~\ref{app:rho}, \ref{app:omega}, \ref{app:w}).
 
@@
 \caption{
-{\bf A comparison of effectiveness control on $\Rnum$ at low level of \percap testing intensity $\rho$.}
-Numerical evaluation of the effectiveness control parameter, $\Delta$ (Eq.~\ref{eq:del4}), with respect to changes in testing and isolation. Parameter values (Table~\ref{tab:params}):
-$\beta=0.5$ 1/day, $1/\gamma= 6$ days; $\omega \in [1/12,2]$ 1/day;  $\rho \in [0,0.013]$ 1/day per capita; $\theta\_w$ and $\theta\_c$ vary between 0 (no effect of isolation) and 1 (complete elimination of transmission); $p_S=0$, $p_I=1$ and $p_R=0.5$.
+{\bf Effectiveness of testing and isolation in reducing $\Rnum$ at low \percap testing intensity.}
+Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4}), over a range of testing and isolation parameters. Parameter values (Table~\ref{tab:params}):
+$\beta=\betaparam$ 1/day, $1/\gamma= \invgammaparam$ days (baseline $\Rnum=\Rnumparam$, $r=\rparam$); $\omega \in [1/12,2]$ 1/day;  $\rho \in [0,0.013]$ 1/day per capita; $\theta\_w$ and $\theta\_c$ vary between 0 (no effect of isolation) and 1 (complete elimination of transmission); $p_S=0$, $p_I=1$ and $p_R=0.5$.
+Only parameter sets where $\theta_c \geq \theta_w$ (confirmed-positive individuals isolate more effectively than waiting individuals) are shown; the alternative case, $\theta_w > \theta_c$, is unrealistic.
 Contours of $\Delta$ are plotted for (a) random testing ($w_S=w_I=w_R=1$) and (b) targeted testing ($w_S=0.3$; $w_I=w_R=1$).
-The baseline basic reproduction number $\Rnum$ and the doubling time $r$, i.e. $\beta/\gamma$ and $\beta-\gamma$ respectively, are 3 and about 0.3 in this simulation.
 }
@@
 \caption{
-{\bf A comparison of effectiveness control on $\Rnum$ at high level of \percap testing intensity $\rho$.}
-We numerically evaluate the effectiveness control parameter, $\Delta$ (Eq.~\ref{eq:del4}), with respect to testing and isolation.
-We use the following parameters (see Table~\ref{tab:params} for a brief description):
-$\omega \in [1/5,2]$ 1/day, $1/\gamma= 6$ days, $\rho \in [0,1/5)$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.5$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$.
-Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
-The baseline basic reproduction number $\Rnum$ and the doubling time $r$, i.e.
-$\beta/\gamma$ and $\beta-\gamma$ respectively, are 3 and 0.3 in this simulation.
+  {\bf Effectiveness of testing and isolation in reducing $\Rnum$ at high \percap testing intensity.}
+  Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4}), over a range of testing and isolation parameters.
+Parameters as in \fref{pan} except: $\omega \in [1/5,2]$ 1/day, $\rho \in [0,1/5)$ 1/day.
+As in \fref{pan}, subplots show (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
 }
@@
 
-However, there are also two specific cases where $\Delta$ changes \emph{non-monotonically}, in counterintuitive directions, as a function of testing and isolation parameters.
+However, there are also two specific cases where $\Delta$ changes non-monotonically, in counterintuitive directions, as a function of testing and isolation parameters.
 
@@
 \item We would generally expect increasing testing delays to increase \Rnum, thus decreasing effectiveness of control $\Delta$.
-This is in fact what we find when waiting individuals do not isolate ($\theta_w =0$, top row of \fref{pan}) --- as we move to the right within each plot in this row, $\Delta$ decreases.
+This is in fact what happens when waiting individuals do not isolate ($\theta_w =0$, top row of \fref{pan}) --- as we move to the right within each plot in this row, $\Delta$ decreases.
 However, when waiting individuals isolate ($\theta_w>0$), we more often see the opposite effect: longer testing delays lead to a greater control effect $\Delta$.
@@
 Many of the qualitative results we have derived are mathematical confirmations of simple, common-sense intuitions.
-In particular, we can decrease $\Rnum$ by increasing isolation efficacy or testing intensity; returning tests faster, if individuals do not isolate while they are waiting for results; or increasing testing focus to target individuals who are likely to be infectious (e.g.
+In particular, we can generally decrease $\Rnum$ by increasing isolation efficacy or testing intensity; returning tests faster, if individuals do not isolate while they are waiting for results; or increasing testing focus to target individuals who are likely to be infectious (e.g.
 symptomatic people or close contacts of known infections).
 
-{\bf A potential advantage of slow test reporting $\omega$, result~\ref{p1:omega};}
+However, we did find two surprising trends: longer delays in returning tests can decrease the spread of the epidemic, and increasing testing rates can increase its spread.
 
-- Whether increasing $\omega$ lowers $\Rnum$ depends on the precise combination of model parameters  including $\rho$, testing strategies represented by compartment-specific testing weights, test sensitivity and specificity, and the level of isolation efficacy.
-- The potential advantage of slow test reporting is in the individual level as opposed to the community level, but seems inapplicable as a strategy.
-- May be this mechanism combined with increasing testing intensity $\rho$ can be considered alongside other prevention methods, eg., social distancing etc, when the transmission is in the exponential-growth phase (in order to increase the generation interval).
-- One reason in favor of prolonging $1/\omega$ is that it is easy to manipulate while increasing $\rho$ can be limited to the resources.
+Over a broad range of parameter space (for random testing, approximately $\theta_w \geq 0.25$; for targeted testing, $\theta_w \geq 0.25$ and either $\theta_c \ge 0.5$ or $1/\omega > 5$), decreasing $\omega$ --- i.e., decreasing the rate at which tested people find out about their infection status --- decreases $\Rnum$.
+This result is counterintuitive; public health agencies have invested a lot of effort in increasing this rate.
+Dynamically, this effect occurs because speeding up test returns shortens the isolation period of uninfected individuals.
+For infected people it only shortens the time to progression to the isolation level of the confirmed-positive compartment.
+Slowing test returns increases $\Rnum$ only if the proportion of infectives in the tested population (test positivity), and the magnitude of increase in isolation effectiveness from the waiting to the confirmed-positive compartment, are large enough to outweigh the decrease in isolation of uninfected waiting individuals.
 
-{\bf The counter-vailing effect of \percap testing intensity, $\rho$, on $\Rnum$;}
-- The explanation here is that more rapid testing leaves more susceptibles in the ``waiting-for-no'' $S\_n$ category at the DFE and delay them to get to $I\_c$ where they are subject to a full isolation, thus no transmission.
-- We are missing out on community-level advantages of fast testing.
-- What is the explanation for having this counter-vailing effect of $\rho$ in targeted testing and not in the random one?
-- May because in the targeted testing more $I\_u$s are tested compared to $S\_u$s, given 
- that higher $\omega$ leaves more people in $S\_n$ and $I\_n$ which results in more transmission.
-- While targeted testing strategies, including targeting people with infection-like symptoms or the contacts of confirmed cases, are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is that viral spread is greater for a slow test than for a fast test.
-This counter-intuitive effect can occur when people are cautious when awaiting a test result, and may not be robust to second-order effects of fast testing (such as better contact tracing).
+While slowing test returns does decrease $\Rnum$ over a broad range of parameter space in our model, there are several real-world processes missing from our model that make it implausible that slowing test returns would actually be an effective public health measure.
+First, our model is missing the components that would allow us to model the primary benefit of rapid testing, i.e.
+detecting and containing outbreaks while they are still in progress.
+This process could be modeled phenomenologically by making the testing focus an increasing function of the speed of test returns (i.e.
+$\partial w_I(\omega)/\partial \omega > 0$), because rapid testing allows more tests to be concentrated on infected individuals.
+Second, in the real world individuals become less likely to maintain isolation if they are required to do so for longer; phenomenologically, we could allow  effectiveness of isolation while waiting for test returns an increasing function of test return speed ($\partial \theta_W(\omega)/\partial \omega > 0$), or we could introduce a separate ``waiting, but no longer isolating'' compartment that individuals entered from the ``waiting, isolated'' compartment at a specified rate.
+Third and finally, if one wants to decrease the overall transmission rate of the population there are more effective methods than keeping tested people in limbo: masking, ventilation, distancing measures, retail and event closures, stay-at-home orders, \ldots
 
-{\bf Specifying the compartment-specific relative testing weights, Specifically $\frac{w_I}{w_S}$;}
+The other counterintuitive result from our analysis is that, for sufficiently high testing intensity $\rho$, increasing testing intensity can actually increase $\Rnum$ (e.g.
+\fref{pan2}(b), upper right panel [$\theta_c=1$, $\theta_w=0$]).
+This phenomenon occurs mainly because of a peculiar feature of the DFE in our model: we assume that testing is ongoing before the beginning of the epidemic, so that the starting point of the epidemic invasion is an equilibrium distribution of susceptibles between the $S\_n$ (untested) and $S\_u$ (waiting) compartments: in particular, $S\_u^* = \rho/(\rho + \omega)$.
+(In reality, testing intensity itself is a dynamic variable that will increase as the epidemic proceeds.) If isolation in the waiting compartment is low ($\theta_w \ll 1$), then waiting individuals can easily be infected once the epidemic starts.
+Once infected they can still infect others while in the waiting compartment, and cannot be tested again before they have returned to the untested compartment.
+Thus, under certain extreme conditions (low $\theta_w$, high $\theta_c$, low $\omega$, high $\rho$), $\Rnum$ becomes an increasing function of testing intensity.
+\bmb{This phenomenon occurs primarily (only??) for targeted testing, because \ldots CAN WE FIGURE OUT WHY?
+Targeting should be irrelevant to determining the DFE (because everyone is $S$ at that point), so what is the dynamical explanation?
+Is there a second order effect somewhere, e.g.
+testing is more effective overall (darker colours in Fig 3(b) upper-right panel than in Fig 3(a) upper panel), so we more quickly reach the point where the effect of increasing $S\_n^*$ outweighs the beneficial effects of testing?} However, this phenomenon is even more unrealistic than the possibility that slowing test returns increases $\Rnum$.
+It depends both on the assumption that regular testing is occurring before the epidemic starts, and on levels of testing that are unrealistically high (at least in large, general-population settings), i.e.
+at least 10\% of the population being tested per day.
 
-- More targeted testing is always good in viral control,
-- We incorporated the compartment-specific relative testing weights, $w_S$, $w_I$ and $w_R$, to model random testing and \emph{targeted} testing strategies.
-Here, in the case of targeted testing and for the simplicity and illustration purposes, we assumed that infected and recovered individuals are tested at three times the \percap rate of susceptible individuals, thus $w_I/w_S=3$ and $w_R=w_I$.
-Note that we have not specified a methodology to assign particular relative testing weights corresponding to a particular targeted testing scenario.
-Modeling different targeted testing strategies, equivalently test-specific testing weights in our framework, requires prior information of the conditional probabilities of getting tested for people in a given compartment.
-This can be implied when we would like to quantify and compare the effect of different levels of test focus for infectious people on the basic reproduction number $\Rnum$, and conclude about the disease spread management.
-For example, when people are tested for "screening", the individuals with potential higher mobility, eg.
-people who are getting on flights, get more tested and thus the corresponding heavier testing weight is assigned than people awaiting for a surgery and are probably going to stay in a long-term-care facility and consequently less mobile and more isolated to begin with.
-With our model, we would be able to compare the sensitivity of the disease dynamics, through $\Rnum$, with respect to testing different high-risk groups in the population.
-This part needs to be developed further in future work.
+Although we model the testing process in more detail than most mathematically focused epidemiological models, one place where more detail could be informative is in the processes determining the testing weights $\{w_S, w_I, w_R\}$.
+While random testing, as done for surveillance purposes, unambiguously leads to equal testing weights, making precise quantitative connections between public health practices and testing weights is difficult in other contexts.
+The testing weights reflect the correlation between an individual's risk of infection and their likelihood of being tested due to age, occupation, geographic location, etc..
+This correlation is influenced, among many other factors, by the proportion of the uninfected population with COVID-like symptoms (e.g.
+due to seasonal upper respiratory tract infections); the concentration of transmission and testing in hot spots such as long-term care facilities and high-density workplaces; the overall testing intensity (and hence e.g.
+restriction to symptomatic individuals); and the proportion of COVID-infected people who are symptomatic.
+Future steps should explore mathematically tractable ways to model some of these factors more precisely.
+For example, separating the infected class into exposed, symptomatic, and a- or pre-symptomatic compartments and allowing the testing weights to vary across non-symptomatic (exposed/asymptomatic/presymptomatic) vs.
+symptomatic compartments could reflect the allocation of tests to diagnostic purposes (targeting symptomatic individuals) vs.
+contact-tracing (targeting infected but non-symptomatic individuals) vs.
+screening (relatively equal weights, depending on the venue).
+Alternatively, one could make the testing weights depend on the testing intensity or test return rate as suggested above.
+Whatever complexity is added would probably put the model beyond reach of the analytical methods we have used in this paper, but one could still use semi-numerical methods such as constructing the next-generation matrix and using it to evaluate the derivatives of $\Rnum$ with respect to the parameters numerically.
 
-What can we say in the context of strength and speed paradigm?
+Although testing and tracing is a key part of infection control strategy, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts, rather than analyzing simple but general models of the feedback between testing and epidemic dynamics \bmb{can we back this statement up?
+more general citations to test-and-trace?
+HIV?
+Bergstrom et al might be a counterexample, although it's for a different context [repeated testing of an isolated population]}.
+We hope that this paper will inspire further explorations of the fundamental properties of dynamical systems that incorporate explicit testing models in an epidemiological context.
+\bmb{we could also talk about test and trace as a 'speed' intervention, but I'm tired}.
+
+\bmb{Leftover comments:
+\begin{itemize}
+\item should emphasize (at some appropriate point in text) that $\Delta$ is independent of $\beta$ (and hence of $\Rnum$ if we are thinking primarily of holding $\gamma$ constant and varying $\beta$) for the small-$\rho$ (Taylor-expansion) case, see eq.
+...
+\item it would be good to have a few more testing refs: Bergstrom et al., there might be some other stuff we can recycle from the lab meeting review; Friston's stuff deserves to be mentioned because it does have explicit testing flows too.
+We should try not to ignore/piss anyone off who has already done stuff
+\end{itemize}
+}
 
@@
 % \widetext
-\begin{center}
-\textbf{\large Appendix}
-\end{center}
+\appendix
+\section{Appendix}
 \setcounter{equation}{0}
@@
 %
-where parameters are specified in Table \ref{tab:params}.
+(see Table \ref{tab:params} for parameter definitions).
 The next generation matrix for this model is $G = F V^{-1}$, where matrix $F$ represents the inflow of new infection to the infected compartments and matrix $V$ represents the flow in the infected compartments when the population is totally susceptible.
@@
 F =& \beta/N_0 \left[ \begin {array}{cccc} 
-S\_u^* & (1-\theta\_w)\,S\_u^* & (1-\theta\_w)\,S\_u^* & (1-\theta\_c)\,S\_u^*\\
-(1-\theta\_w)S\_n^* & (1-\theta\_w)^2\,S\_n^* & (1-\theta\_w)^2\,S\_n^* & (1-\theta\_w)(1-\theta\_c)\,S\_n^* \\ 
+S\_u^* & (1-\theta\_w) S\_u^* & (1-\theta\_w) S\_u^* & (1-\theta\_c) S\_u^*\\
+(1-\theta\_w)S\_n^* & (1-\theta\_w)^2 S\_n^* & (1-\theta\_w)^2 S\_n^* & (1-\theta\_w)(1-\theta\_c) S\_n^* \\ 
 0&0&0&0\\
@@
  =&\beta/N_0 \left[ \begin {array}{c} S\_u^* \\ (1-\theta\_w)S\_n^* \\ 0 \\ 0 \end {array} \right]
-        \left[ \begin {array}{cccc} 1, \, 1-\theta\_w, \, 1-\theta\_w, \, 1-\theta\_c \end {array} \right], \text{and}\\  V=&
+        \left[ \begin {array}{cccc} 1,   1-\theta\_w,   1-\theta\_w,   1-\theta\_c \end {array} \right], \text{and}\\  V=&
  \left[ \begin {array}{cccc}  
@@
 \end{align}
-where $C= \big( \gamma(\omega+\gamma)+(\gamma+\omega p_I)\hat F_I \big) (\omega+\gamma)$ and $\hat F_I$ is the \percap testing rate for the infected people and represented in Eq.\,\eqref{eq:fi}.
+where $C= \big( \gamma(\omega+\gamma)+(\gamma+\omega p_I)\hat F_I \big) (\omega+\gamma)$ and $\hat F_I$ is the \percap testing rate for the infected people and represented in Eq.~\eqref{eq:fi}.
 Note that all the columns of matrix $V^{-1}$ summ up to $1/\gamma$.
@@
 \label{G11}
-G_{11} = \frac{\beta}{\gamma \, C} 
+G_{11} = \frac{\beta}{\gamma   C} 
 \left[ \begin {array}{c} (\omega-\rho)/\omega \\ (1-\theta\_w)\rho/\omega \end {array} \right]
-\left[ \begin {array}{cccc} 1, \, 1-\theta\_w, \, 1-\theta\_w, \, 1-\theta\_c \end {array} \right]
+\left[ \begin {array}{cccc} 1,   1-\theta\_w,   1-\theta\_w,   1-\theta\_c \end {array} \right]
 \left[ \begin {array}{cc}
@@
 
-This section shows that $\pro{\Delta}$ can be positive or negative, with $\Delta$ defined in Eq.\,\eqref{eq:C12}, and thus $\pro{\Rnum} < 0$, where $\Rnum$ is given in Eq.~\eqref{R0}.
+This section shows that $\pro{\Delta}$ can be positive or negative, with $\Delta$ defined in Eq.~\eqref{eq:C12}, and thus $\pro{\Rnum} < 0$, where $\Rnum$ is given in Eq.~\eqref{R0}.
 We rewrite matrix $G_{11}$ in \eqref{G11} in the following form to simplify the calculations:
@@
 \label{G112}
-G_{11} = \frac{\beta}{\gamma \, C} 
-\left[ \begin {array}{c}  S\_u^*/N_0 \\ (1-\theta\_w)\,S\_n^*/N_0  \end {array} \right]
+G_{11} = \frac{\beta}{\gamma   C} 
+\left[ \begin {array}{c}  S\_u^*/N_0 \\ (1-\theta\_w) S\_n^*/N_0  \end {array} \right]
 \left[ \begin {array}{cccc} 
@@
 where $C$ is the same as the one in Eq. \eqref{eq:C12}, i.e., 
-$$C=(\omega+\gamma)(\gamma(\omega+\gamma)+(\omega\,p_I+\gamma)\,\hat F_I),$$
+$$C=(\omega+\gamma)(\gamma(\omega+\gamma)+(\omega p_I+\gamma) \hat F_I),$$
 and $C_1$ and $C_2$ are 
@@
 % \label{eq:C12}
-C_1 =& (\omega+\gamma)(\theta\_w \, \gamma+\theta\_c\, \omega\, p_I) \hat F_I,\\
-C_2 =& \big( \omega\gamma(1+p_I)\hat F_I+\gamma^2\,(\omega+\gamma+\hat F_I)\big)\theta\_w + \omega^2\,p_I\,\hat F_I\,\theta\_c,
+C_1 =& (\omega+\gamma)(\theta\_w   \gamma+\theta\_c  \omega  p_I) \hat F_I,\\
+C_2 =& \big( \omega\gamma(1+p_I)\hat F_I+\gamma^2 (\omega+\gamma+\hat F_I)\big)\theta\_w + \omega^2 p_I \hat F_I \theta\_c,
 \end{align*}
-where $\hat F_I$ is given in Eq.\,\eqref{eq:fi}.
+where $\hat F_I$ is given in Eq.~\eqref{eq:fi}.
 Note that for analysis brevity, we let $N_0=1$, thus $S\_u^*$ and $S\_n^*$ are in the scale of 0 to 1.
@@
 \label{eq:del4}
-\Delta= \frac{1}{C}\big(C_1\,S\_u^*+(C_2(1-\theta\_w)+C\,\theta\_w)\,S\_n^*\big).
+\Delta= \frac{1}{C}\big(C_1 S\_u^*+(C_2(1-\theta\_w)+C \theta\_w) S\_n^*\big).
 \end{equation*}
@@
 Mathematically we would like to verify the sign of $\pder \Rnum{\theta\_w}$ and $\pder \Rnum{\theta\_c}$.
-We start with simplifying $\Delta$ \eqref{eq:del4} by factoring $\theta\_w$ and $\theta\_c$ in Eq.\,\eqref{eq:del4}.
+We start with simplifying $\Delta$ \eqref{eq:del4} by factoring $\theta\_w$ and $\theta\_c$ in Eq.~\eqref{eq:del4}.
 Thus, $\Delta$ can be rewritten as
@@
 \Delta= \frac{1}{C}\Big(
--D_1\,S\_n^*\,\boldsymbol {\theta\_w^2}+
-\big(-\omega^2\,p_I\,\hat F_I\,S\_n^*\,\boldsymbol{\theta\_c}\,+\,D_2\,S\_n^*\,+\,\gamma\,\hat F_I(\omega+\gamma) \big)\boldsymbol{\theta\_w}+
-(\omega+\gamma\,S\_u^*)\omega\,p_I\,\hat F_I\,\boldsymbol{\theta\_c}
+-D_1 S\_n^* \boldsymbol {\theta\_w^2}+
+\big(-\omega^2 p_I \hat F_I S\_n^* \boldsymbol{\theta\_c} + D_2 S\_n^* + \gamma \hat F_I(\omega+\gamma) \big)\boldsymbol{\theta\_w}+
+(\omega+\gamma S\_u^*)\omega p_I \hat F_I \boldsymbol{\theta\_c}
 \Big),
@@
 \label{eq:del4_d1d2}
-D_1=& (\omega+\gamma)\gamma^2+(\omega+\gamma+\omega\,p_I)\gamma\,\hat F_I, \\
-D_2=& (3\omega+2\gamma)\gamma^2+(\omega+\gamma+2\omega\,p_I)\gamma\,\hat F_I+(\gamma+p_I\,\hat F_I)\omega^2.
+D_1=& (\omega+\gamma)\gamma^2+(\omega+\gamma+\omega p_I)\gamma \hat F_I, \\
+D_2=& (3\omega+2\gamma)\gamma^2+(\omega+\gamma+2\omega p_I)\gamma \hat F_I+(\gamma+p_I \hat F_I)\omega^2.
 \end{align}
-$\Delta$, Eq.\,\eqref{eq:del4_theta}, is linear in $\theta\_c$ with a positive coefficient. thus
+$\Delta$, Eq.~\eqref{eq:del4_theta}, is linear in $\theta\_c$ with a positive coefficient. thus
 \begin{equation}
@@
 \label{eq:del4_thetaw}
-1/C\,\Big( -D_1\,S\_n^*\,\boldsymbol {\theta\_w^2}+
-\big(-\omega^2\,p_I\,\hat F_I\,S\_n^*\,\boldsymbol{\theta\_c}\,+\,D_2\,S\_n^*\,+\,\gamma\,\hat F_I(\omega+\gamma) \big)\boldsymbol{\theta\_w} \Big).
+1/C \Big( -D_1 S\_n^* \boldsymbol {\theta\_w^2}+
+\big(-\omega^2 p_I \hat F_I S\_n^* \boldsymbol{\theta\_c} + D_2 S\_n^* + \gamma \hat F_I(\omega+\gamma) \big)\boldsymbol{\theta\_w} \Big).
 \end{equation}
@@
 \pder\Delta{\theta\_w}\bigg\rvert_{\theta\_w=1} =&
-1/C\Big( (D_2-2D_1-\omega^2\,p_I\,\hat F_I\,\theta\_c)\,S\_n^*+\gamma\,\hat F_I(\omega+\gamma) \Big) \\\notag
-=& 1/C\Big( (\gamma(\omega+\gamma)+\gamma\omega^2+(1-\theta\_c)\omega^2\,p_I\,\hat F_I )\,S\_n^*
-+\gamma(\omega+\gamma)\,\hat F_I\,(1-S\_n^*) \Big),
+1/C\Big( (D_2-2D_1-\omega^2 p_I \hat F_I \theta\_c) S\_n^*+\gamma \hat F_I(\omega+\gamma) \Big) \\\notag
+=& 1/C\Big( (\gamma(\omega+\gamma)+\gamma\omega^2+(1-\theta\_c)\omega^2 p_I \hat F_I ) S\_n^*
++\gamma(\omega+\gamma) \hat F_I (1-S\_n^*) \Big),
 \end{align}
@@
 Mathematically we would like to verify the sign of $\pder\Rnum{\rho}$, which specifically depends on $\pder\Delta{\rho}$.
-We use the derived expressions for $S\_u^*$ and $S\_n^*$, given by Eqs.\,\eqref{dfe}, in $\Delta$ \eqref{eq:del4}.
+We use the derived expressions for $S\_u^*$ and $S\_n^*$, given by Eqs.~\eqref{dfe}, in $\Delta$ \eqref{eq:del4}.
 Also, we define $\phi = \hat F_S = \frac{\rho \omega}{\omega-\rho}$, to reparameterize $\rho$.
@@
 \label{eq:dd3dr}
-\partial\Delta/\partial\phi=& \frac{1}{d_3} (a_3\,\phi^2+b_3\,\phi+c_3),
+\partial\Delta/\partial\phi=& \frac{1}{d_3} (a_3 \phi^2+b_3 \phi+c_3),
 \end{align}
@@
 a_3& = \frac{w_I}{w_S} \Bigg(
-&& (1-\theta\_w)\,(1+\frac{w_I}{w_S})\theta\_w\,\gamma^3 +(1-\theta\_c)\,p_I^2\,\theta\_w\,\frac{w_I}{w_S}\,\omega^3 \\ \notag
-& ~ &&+\Big( \big( (1-\theta\_w-\frac{w_I}{w_S})\,\theta\_c +(3-2\theta\_w)\,\theta\_w\frac{w_I}{w_S} \big)\,p_I +(1-\theta\_w)\,(1+\frac{w_I}{w_S})\theta\_w  \Big)\,\omega\,\gamma^2 \\ \notag
+&& (1-\theta\_w) (1+\frac{w_I}{w_S})\theta\_w \gamma^3 +(1-\theta\_c) p_I^2 \theta\_w \frac{w_I}{w_S} \omega^3 \\ \notag
+& ~ &&+\Big( \big( (1-\theta\_w-\frac{w_I}{w_S}) \theta\_c +(3-2\theta\_w) \theta\_w\frac{w_I}{w_S} \big) p_I +(1-\theta\_w) (1+\frac{w_I}{w_S})\theta\_w  \Big) \omega \gamma^2 \\ \notag
 & ~ &&+\Big(
-\big((1-\theta\_w-\theta\_w\,\frac{w_I}{w_S})\,\theta\_c+ (2-\theta\_w)\,\theta\_w\,\frac{w_I}{w_S} \big)\,p_I
-+(2\,\theta\_w-\theta\_w^2-\theta\_c)\,\frac{w_I}{w_S}\,p_I^2
-\Big)\,\omega^2\,\gamma \Bigg), \\ \notag
-b_3& = && \hspace{-1cm} 2\frac{w_I}{w_S}\,(\omega+\gamma)\gamma\Big(
-(\omega+\gamma+\omega\,p_I)(2-\theta\_w)\,\gamma\,\theta\_w+ (1-\theta\_w)\,\omega^2 p_I\theta\_c+\omega^2 p_I\theta\_w
+\big((1-\theta\_w-\theta\_w \frac{w_I}{w_S}) \theta\_c+ (2-\theta\_w) \theta\_w \frac{w_I}{w_S} \big) p_I
++(2 \theta\_w-\theta\_w^2-\theta\_c) \frac{w_I}{w_S} p_I^2
+\Big) \omega^2 \gamma \Bigg), \\ \notag
+b_3& = && \hspace{-1cm} 2\frac{w_I}{w_S} (\omega+\gamma)\gamma\Big(
+(\omega+\gamma+\omega p_I)(2-\theta\_w) \gamma \theta\_w+ (1-\theta\_w) \omega^2 p_I\theta\_c+\omega^2 p_I\theta\_w
 \Big) ,\\ \notag
-c_3& = && \hspace{-1cm} (\omega+\gamma)^2\,\gamma\Big(
-(2-\theta\_w)\,\gamma^2\,\theta\_w+
-(1+\frac{w_I}{w_S})\,\omega\,\gamma\,\theta\_w+
+c_3& = && \hspace{-1cm} (\omega+\gamma)^2 \gamma\Big(
+(2-\theta\_w) \gamma^2 \theta\_w+
+(1+\frac{w_I}{w_S}) \omega \gamma \theta\_w+
 \frac{w_I}{w_S} \omega^2 p_I\theta\_c \Big),\\ \notag
-d_3&  = && \hspace{-1cm} \frac{(\omega+\gamma)}{\omega}\, \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi\,+\,(\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2.
+d_3&  = && \hspace{-1cm} \frac{(\omega+\gamma)}{\omega}  \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi + (\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2.
 \end{align}
@@
 This is illustrated in the top-right panel of the \fref{pan2} panel (b), where the strength of isolation for awaiting people is the least, but the most for the confirmed cases.
-In this case, simplifying $a_3$ in Eq.~\eqref{eq:abcd2} gives $$a_3=\frac{w_I}{w_S}\,\omega\,\gamma\,p_I((\omega+\gamma)-\frac{w_I}{w_S}(\omega\,p_I+\gamma)).$$
-Specifically, in the case of targeted testing which is identified by $\frac{w_I}{w_S}> 1$, and using a perfect sensitive test, thus $p_I=1$, there exists a range for $\rho$ over which $\pder\Rnum{\rho}\leq\,0$.
+In this case, simplifying $a_3$ in Eq.~\eqref{eq:abcd2} gives $$a_3=\frac{w_I}{w_S} \omega \gamma p_I((\omega+\gamma)-\frac{w_I}{w_S}(\omega p_I+\gamma)).$$
+Specifically, in the case of targeted testing which is identified by $\frac{w_I}{w_S}> 1$, and using a perfect sensitive test, thus $p_I=1$, there exists a range for $\rho$ over which $\pder\Rnum{\rho}\leq 0$.
 Note that $\rho$ and $\omega$ have a similar mechanism in delaying people to get into $I\_c$, thus we would expect to see the non-trivial counter-vailing effect of these two parameters on $\Rnum$.
@@
 \label{eq:lin}
-\Delta = \frac{\rho}{\omega\,\gamma(\omega+\gamma)} \Big(
-\frac{w_I}{w_S}\,\omega^2\,p_I\,\theta\_c+(\frac{w_I}{w_S}+1)\,\gamma\omega\theta\_w\,+\gamma^2\,\theta\_w(2-\theta\_w)
+\Delta = \frac{\rho}{\omega \gamma(\omega+\gamma)} \Big(
+\frac{w_I}{w_S} \omega^2 p_I \theta\_c+(\frac{w_I}{w_S}+1) \gamma\omega\theta\_w +\gamma^2 \theta\_w(2-\theta\_w)
 \Big). 
@@
 \label{eq:dlindo}
-\pder\Delta{\omega} = \frac{\rho}{\omega^2\,(\omega+\gamma)^2} \Big(
-(p_I\,\frac{w_I}{w_S}\theta\_c\,-(1+\frac{w_I}{w_S})\theta\_w)\omega^2+2\theta_w\,\gamma(\theta_w-2)\,\omega+\theta_w\,\gamma^2\,(\theta_w-2)
+\pder\Delta{\omega} = \frac{\rho}{\omega^2 (\omega+\gamma)^2} \Big(
+(p_I \frac{w_I}{w_S}\theta\_c -(1+\frac{w_I}{w_S})\theta\_w)\omega^2+2\theta_w \gamma(\theta_w-2) \omega+\theta_w \gamma^2 (\theta_w-2)
 \Big),
@@
 \label{eq:d_del_wis}
-\pder \Delta{w\_{IS}}= \dfrac{(\omega-\rho)\,(\omega(\omega-\rho\theta\_w)+\gamma(\omega-\rho))\,(\theta\_w\gamma+\theta\_c\omega\,p\_I)}
-{(-\omega^2\,\gamma+\omega\,\gamma\,\rho-\gamma\,\rho\,\omega\,w\_{IS}-
-\omega\,\gamma^2+\gamma^2\,\rho-\omega^2\,p_I\,\rho\,w\_{IS})^2},
+\pder \Delta{w\_{IS}}= \dfrac{(\omega-\rho) (\omega(\omega-\rho\theta\_w)+\gamma(\omega-\rho)) (\theta\_w\gamma+\theta\_c\omega p\_I)}
+{(-\omega^2 \gamma+\omega \gamma \rho-\gamma \rho \omega w\_{IS}-
+\omega \gamma^2+\gamma^2 \rho-\omega^2 p_I \rho w\_{IS})^2},
 \end{align}

### 2021-06-09 codes/SIRfunctions.R

@@ -204,8 +204,8 @@ catt <- function(...,file="modeldefs.tex") {
 ##' @param x numeric value
 ##' @param nm name of macro (will be \nm in LaTeX)
 ##' @param fmt number format (e.g. number of decimal places)
-latexout <- function(x,nm,fmt="%1.1f") {
-    catt(sprintf("\\newcommand{\\%s}{%s}\n",nm,sprintf(fmt,x)))
+latexout <- function(x,nm,fmt="%1.1f", ...) {
+    catt(sprintf("\\newcommand{\\%s}{%s}\n",nm,sprintf(fmt,x)), ...)
 }
 
 saveEnvironment()

### 2021-06-09 codes/sir_plot.R

@@ -44,11 +44,12 @@ rho_fast_min <- 0
 rho_fast_max <- (1/inv_omega_fast_max)-0.000001
 rho_fast_brks <- c(0,0.1,0.17)
 
-unlink("modeldefs.tex")
-
 catt("% parameter values\n")
-latexout(params[["beta"]],"betaparam")
-latexout(params[["gamma"]],"gammaparam")
+latexout(params[["beta"]],"betaparam", file="modeldefs.tex")
+latexout(params[["gamma"]],"gammaparam", file="modeldefs.tex")
+latexout(1/params[["gamma"]],"invgammaparam", file="modeldefs.tex")
+latexout(params[["beta"]]/params[["gamma"]],"Rnumparam", file="modeldefs.tex")
+latexout(params[["beta"]]-params[["gamma"]],"rparam", file="modeldefs.tex")
 
 
 background <- "grey90"
@@ -71,14 +72,14 @@ df_random_h <- make_params_dat(params = params,
                              eta_cs=0,eta_ce=1, ## so theta_c
                              omega_s=1/inv_omega_fast_max,omega_e=1/inv_omega_fast_min,
                              rho_s=rho_fast_min,rho_e=rho_fast_max)
-                             
+
 
 df_targeted_h <- make_params_dat(params = update(params,W_S=W_S_targeted),
                                eta_ws=0,eta_we=1, ## so theta_w
                                eta_cs=0,eta_ce=1, ## so theta_c
                                omega_s=1/inv_omega_fast_max,omega_e=1/inv_omega_fast_min,
                                rho_s=rho_fast_min,rho_e=rho_fast_max)
-                               
+
 
 ##important contour, ie R0=1 thus threshold=1 when plotting R0 contours, or Delta(R0=1)
 threshold <- 1-(params[["gamma"]]/params[["beta"]]) ## corresponding to R0=1
@@ -274,7 +275,7 @@ ggsave(p1_targeted_hacked,
 # 6. Check: in random testing when rho is high, $\theta_w=0$/$\theta_c=0.75$ panel
 # #################################
 
-df_random_h_check <- df_random_h %>% 
+df_random_h_check <- df_random_h %>%
   filter(theta_w==0 & theta_c==.75)
 p_temp <- (ggplot(df_random_h_check,aes(x=1/omega,y=rho,z=Delta))
         + theme_bw()

### 2021-06-09 SIR_manuscript.tex
take out modeldefs test in ms
working on the comments

@@
 \section{Introduction}
-
-%% FIXME: uncomment when pipeline works
-$\beta=\betaparam$; $\gamma=\gammaparam$
 
@@
 
-The analytical calculation of the next-generation matrix and simplification of the $\Rnum$ expression was carried out in Maple \citep{maple14}.
+The computations in this paper, specifically, calculation of the next-generation matrix and simplification of the $\Rnum$ expression, were performed in Maple\textsuperscript{\texttrademark} \citep{maple14}.
 We used R \citep{r} for numerical solutions, in particular for plotting the contours of $\Delta$ (Eq.~\eqref{eq:del4a}) over a range of parameters.
@@
 ...
-\item it would be good to have a few more testing refs: Bergstrom et al., there might be some other stuff we can recycle from the lab meeting review; Friston's stuff deserves to be mentioned because it does have explicit testing flows too.
+\item it would be good to have a few more testing refs: Bergstrom et al., there might be some other stuff we can recycle from the lab meeting review; Friston's stuff \ali{\citep{friston2021testing} OK?} deserves to be mentioned because it does have explicit testing flows too.
 We should try not to ignore/piss anyone off who has already done stuff
 \end{itemize}
+\ali{:}
+\cite{bergstrom2020frequency} modeled repeated random testing of isolated populations.
+\citep{friston2021testing} model the effects of self-isolation on testing and tracking.
+Their model is a dynamic causal model equipped with isolation state which is an extension of compartmental model SEIR.
+They found that there exists no testing strategy that can attenuate mortality rates, other than by deferring or delaying a second wave.
 }

### 2021-06-09 note/SIR_scrap.tex
starting cohort-based framework
working on the comments

@@
 \newcommand{\covid}{COVID-19\xspace}
+\newcommand{\pro}[1][]{\ensuremath{\frac{\partial #1}{\partial \rho}}}
+\newcommand\pder[2][]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder[x]{y}
 
@@
 We also compare to an individual-level advantage of fast tests: people who test positive may be even more careful.
-In our model analysis, Proposition \ref{prop1} part \ref{p1:omega} is describing this potential advantage. 
+In our model analysis, \ref{prop1} part \ref{p1:omega} is describing this potential advantage. 
 It states that returning test results more rapidly, i.e., increasing the rate $\omega$, does not necessarily lower the reproduction number $\Rnum$; whether increasing $\omega$ lowers $\Rnum$ depends on the precise combination of model parameters  including test reporting rate, testing strategies represented by compartment-specific testing weights, test sensitivity and specificity, and the level of isolation.
@@
 
+\citep{friston2021testing} model the effects of self-isolation on testing and tracking.
+Their model is a dynamic causal model equipped with isolation state which is an extension of compartmental model SEIR.
+They found that there exists no testing strategy that can attenuate mortality rates, other than by deferring or delaying a second wave.
+
 \section{Unused Scripts}

### 2021-06-09 SIR_manuscript.tex
more BMB comment integration

@@
 This testing intensity reflects realistic testing rates during the \covid pandemic, i.e.
-a maximum of 1.3\% of the population per day, giving a maximum of 10000 tests/day in a population of size $N_0=10^6$ (this is approximately four times the maximum testing rate in Ontario, Canada in mid-2021).
+a maximum of 1.3\% of the population per day, giving a maximum of 10000 tests/day in a population of size $N_0=10^6$ (approximately four times the maximum testing rate in Ontario, Canada in mid-2021).
 The less realistic high-testing case \fref{pan2} is included to highlight the occurrence of non-monotonic changes in $\Rnum$ with respect to $\rho$.
 In \fref{pan2} the maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,1/5)$ and the test return rate $\omega\in [1/5,2]$; these values are clearly unrealistic for a large population but might be relevant for some population undergoing focused testing, such as a sports league or university.
-  The critical contour of $\Delta=1-\frac{\gamma}{\beta}$, corresponding to the threshold $\Rnum=1$, is shown as a dotted line in \fref{pan} and \fref{pan2}; combinations of $\rho$ and $\omega$ that lie above this critical contour will control the epidemic.
-The implied baseline value of $\Rnum=\frac{\beta}{\gamma}$ in these figures is 3, corresponding to a case where the epidemic spreads slowly even in the absence of testing and tracing.
 
@@
 This is because people waiting for negative tests are assumed to continue to isolate; this applies both to susceptibles and to people who became infected while waiting for negative test results.
-This effect outweighs the effect of confirmed individuals isolating except when this isolation parameter $\theta_c$ is substantially bigger than $\theta_w$.
-This result depends on the idea that, all else equal, people who have to wait longer for test results isolate at the same level (but for a longer time) as they would if the wait were shorter.
+This effect outweighs the effect of confirmed individuals isolating except when isolation efficacy of confirmed-positive individuals ($\theta_c$) is substantially larger than that of waiting individuals ($\theta_w$).
+
 \item \fref{pan} also shows that greater testing intensity (increasing $\rho$) generally increases the effectiveness of control (moving up in each panel).
@@
 Over a broad range of parameter space (for random testing, approximately $\theta_w \geq 0.25$; for targeted testing, $\theta_w \geq 0.25$ and either $\theta_c \ge 0.5$ or $1/\omega > 5$), decreasing $\omega$ --- i.e., decreasing the rate at which tested people find out about their infection status --- decreases $\Rnum$.
+For low testing rates, the value of $\Delta$ is independent of $\beta$ (Eq.~\ref{eq:lin}), so the patterns shown in \fref{pan} hold for any baseline values of $\Rnum$ (assuming we hold $\gamma$ constant).
+
 This result is counterintuitive; public health agencies have invested a lot of effort in increasing this rate.
@@
 Is there a second order effect somewhere, e.g.
-testing is more effective overall (darker colours in Fig 3(b) upper-right panel than in Fig 3(a) upper panel), so we more quickly reach the point where the effect of increasing $S\_n^*$ outweighs the beneficial effects of testing?} However, this phenomenon is even more unrealistic than the possibility that slowing test returns increases $\Rnum$.
+testing is more effective overall (darker colours in Fig 3(b) upper-right panel than in Fig 3(a) upper panel), so we more quickly reach the point where the effect of increasing $S\_n^*$ outweighs the beneficial effects of testing?
+\fady{targeting + high testing rates could exhaust the untested infectious population such that even higher testing rates (note we are asking about $\partial \Delta/\partial \rho$) could enrich the tested population even more for $S$?} This doesn't seem impossible, but checking the verbal reasoning against math seems challenging \ldots} However, this phenomenon is even more unrealistic than the possibility that slowing test returns increases $\Rnum$.
 It depends both on the assumption that regular testing is occurring before the epidemic starts, and on levels of testing that are unrealistically high (at least in large, general-population settings), i.e.
@@
 \begin{itemize}
-\item should emphasize (at some appropriate point in text) that $\Delta$ is independent of $\beta$ (and hence of $\Rnum$ if we are thinking primarily of holding $\gamma$ constant and varying $\beta$) for the small-$\rho$ (Taylor-expansion) case, see eq.
-...
 \item it would be good to have a few more testing refs: Bergstrom et al., there might be some other stuff we can recycle from the lab meeting review; Friston's stuff \ali{\citep{friston2021testing} OK?} deserves to be mentioned because it does have explicit testing flows too.
@@
 \ali{:}
-\cite{bergstrom2020frequency} modeled repeated random testing of isolated populations.
-\citep{friston2021testing} model the effects of self-isolation on testing and tracking.
-Their model is a dynamic causal model equipped with isolation state which is an extension of compartmental model SEIR.
-They found that there exists no testing strategy that can attenuate mortality rates, other than by deferring or delaying a second wave.
+Models of controlling COVID-19 spread in isolated populations (especially focused on college campuses) by
+\cite{bergstrom2020frequency} (analytical) and \cite{rogers_high-frequency_2021} (computational) \ldots
+(maybe other useful citations \href{https://docs.google.com/spreadsheets/d/1j1wgHNEAq_OKfOnugEen8vafXszyjw1Q}{here}?)
+\citep{friston2021testing} use a compartmental model to explore the effects of self-isolation on testing and tracking in the population of the UK (they conclude that testing strategy can delay a second wave but cannot reduce overall mortality)
+Blog post \citep{otto_tipping_2020} probably not citeable \ldots
 }
@@
 We use
-the linearization of $\Rnum$ around $\rho=0$ to show that there a non-monotonic relationship between $\Rnum$  and $\omega$. The Taylor expansion of $\Delta$ at $\rho=0$ is
+the linearization of $\Rnum$ around $\rho=0$ to show that there a non-monotonic relationship between $\Rnum$  and $\omega$. The first-order Taylor approximation of $\Delta$ at $\rho=0$ is
 \begin{equation}

### 2021-06-09 SIR_manuscript.tex
meeting with Ben comments

--- 2021-06-09 SIR_manuscript.tex
@@
 However, we analyzed the effect of $\omega$ on $\Rnum$ for the special case of low testing intensity.
-Specifically, by making the restriction $\rho \ll 1$, we are able to Taylor-expand $\Delta$ at $\rho=0$ and analyze the resulting simplified derivatives to illustrate a surprising non-monotonic relationship between $\Rnum$ and $\omega$.
+Specifically, by making the restriction $\rho \ll 1$, we are able to Taylor-expand $\Delta$ at $\rho=0$, use the linear approximation with respect to $\rho$ and analyze the resulting simplified derivatives to illustrate a surprising non-monotonic relationship between $\Rnum$ and $\omega$.
 
-The analytical calculation of the next-generation matrix and simplification of the $\Rnum$ expression was carried out in Maple \citep{maple14}.
+The computations in this paper, specifically, calculation of the next-generation matrix and simplification of the $\Rnum$ expression, were performed in Maple\textsuperscript{\texttrademark} \citep{maple14}.
 We used R \citep{r} for numerical solutions, in particular for plotting the contours of $\Delta$ (Eq.~\eqref{eq:del4a}) over a range of parameters.
@@
 Our model is a caricature --- it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g.
-exposed, presymptomatic, and asymptomatic compartments [REFS]\ali{\citep{kain2021chopping} OK?}; time-varying testing rates [REFS?]\ali{\citep{jarvis2021temporal} OK?}; behavioural dynamics [REF Weitz/Park/Dushoff] \ali{\citep{weitz2020awareness} OK?}).
+exposed, presymptomatic, and asymptomatic compartments [REFS]\ali{\citep{kain2021chopping} OK?}; time-varying testing rates [REFS?]; behavioural dynamics [REF Weitz/Park/Dushoff] \ali{\citep{weitz2020awareness} OK?}).
 Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between epidemiology and testing, rather than for making quantitative predictions or guiding pandemic responses.
@@
 \begin{itemize}
-\item should emphasize (at some appropriate point in text) that $\Delta$ is independent of $\beta$ (and hence of $\Rnum$ if we are thinking primarily of holding $\gamma$ constant and varying $\beta$) for the small-$\rho$ (Taylor-expansion) case, see eq.
-...
+\item should emphasize (at some appropriate point in text) that $\Delta$ is independent of $\beta$ (and hence of $\Rnum$ if we are thinking primarily of holding $\gamma$ constant and varying $\beta$) for the small-$\rho$ (Taylor-expansion) case, see Eq.~\ref{eq:lin}.
 \item it would be good to have a few more testing refs: Bergstrom et al., there might be some other stuff we can recycle from the lab meeting review; Friston's stuff deserves to be mentioned because it does have explicit testing flows too.
@@
 \end{itemize}
+\ali{:}
+Testing dynamics has been previously modeled in the context of epidemiology.
+\cite{bergstrom2020frequency} modeled repeated random testing of isolated populations and provided some analytical results.
+\citep{friston2021testing} model the effects of self-isolation on testing and tracking.  
+ Comparing to these works, our modeling execise is novel. 
 }
@@
 We use
-the linearization of $\Rnum$ around $\rho=0$ to show that there a non-monotonic relationship between $\Rnum$  and $\omega$. The Taylor expansion of $\Delta$ at $\rho=0$ is
+the linearization of $\Delta$ around $\rho=0$ to show that there a non-monotonic relationship between $\Rnum$  and $\omega$. The linear term in the Taylor expansion of $\Delta$ at $\rho=0$ is
 \begin{equation}

### 2021-06-10 SIR_manuscript.tex
adding some refs for testing tracing modeling

@@
 \end{itemize}
-\ali{:}
-Testing dynamics has been previously modeled in the context of epidemiology.
-\cite{bergstrom2020frequency} modeled repeated random testing of isolated populations and provided some analytical results.
-\citep{friston2021testing} model the effects of self-isolation on testing and tracking.  
- Comparing to these works, our modeling execise is novel. 
 }
+
+\ali{This belongs to introduction?}
+There have been several modeling works of testing and tracing dynamics and their interaction with epidemiological dynamics.
+In the context of repeated screaning and random testing of isolated populations (such as a university setup),
+\cite{bergstrom2020frequency} provided some analytical results, 
+\cite{rogers2021high} simulated a SEIR model with high-low testing intensity and sensitivity.
+Further, \cite{friston2021testing} model the effects of self-isolation on testing and tracking.  
+Our modeling approache is novel compared to these works.
 

### 2021-06-14 SIR_manuscript.tex
adding Faddy's work, changed the coeff from ABC to abc_4

@@
 %% \end{enumerate}
-\item \label{p1:omega} Increasing the rate of test return ($\omega$) always decreases \Rnum if waiting individuals do not isolate ($\theta_w=0$) (Eq.~\ref{eq:dlindo}).
+\item \label{p1:omega} Increasing the rate of test return ($\omega$) always decreases \Rnum if waiting individuals do not isolate ($\theta\_w=0$) (Eq.~\ref{eq:dlindo}).
 \item \label{p1:w} Increasing testing focus, i.e. changing the testing weights from random ($w_S=w_I$) toward targeted  ($w_S<w_I$), always decreases \Rnum (Eq.~\ref{eq:d_del_wis}).
@@
 $\beta=\betaparam$ 1/day, $1/\gamma= \invgammaparam$ days (baseline $\Rnum=\Rnumparam$, $r=\rparam$); $\omega \in [1/12,2]$ 1/day;  $\rho \in [0,0.013]$ 1/day per capita; $\theta\_w$ and $\theta\_c$ vary between 0 (no effect of isolation) and 1 (complete elimination of transmission); $p_S=0$, $p_I=1$ and $p_R=0.5$.
-Only parameter sets where $\theta_c \geq \theta_w$ (confirmed-positive individuals isolate more effectively than waiting individuals) are shown; the alternative case, $\theta_w > \theta_c$, is unrealistic.
+Only parameter sets where $\theta\_c \geq \theta\_w$ (confirmed-positive individuals isolate more effectively than waiting individuals) are shown; the alternative case, $\theta\_w > \theta\_c$, is unrealistic.
 Contours of $\Delta$ are plotted for (a) random testing ($w_S=w_I=w_R=1$) and (b) targeted testing ($w_S=0.3$; $w_I=w_R=1$).
@@
 \item We would generally expect increasing testing delays to increase \Rnum, thus decreasing effectiveness of control $\Delta$.
-This is in fact what happens when waiting individuals do not isolate ($\theta_w =0$, top row of \fref{pan}) --- as we move to the right within each plot in this row, $\Delta$ decreases.
-However, when waiting individuals isolate ($\theta_w>0$), we more often see the opposite effect: longer testing delays lead to a greater control effect $\Delta$.
+This is in fact what happens when waiting individuals do not isolate ($\theta\_w =0$, top row of \fref{pan}) --- as we move to the right within each plot in this row, $\Delta$ decreases.
+However, when waiting individuals isolate ($\theta\_w>0$), we more often see the opposite effect: longer testing delays lead to a greater control effect $\Delta$.
 This is because people waiting for negative tests are assumed to continue to isolate; this applies both to susceptibles and to people who became infected while waiting for negative test results.
-This effect outweighs the effect of confirmed individuals isolating except when this isolation parameter $\theta_c$ is substantially bigger than $\theta_w$.
+This effect outweighs the effect of confirmed individuals isolating except when this isolation parameter $\theta\_c$ is substantially bigger than $\theta\_w$.
 This result depends on the idea that, all else equal, people who have to wait longer for test results isolate at the same level (but for a longer time) as they would if the wait were shorter.
@@
 This relationship can be reversed, however, at very high testing intensities.
-This can only occur when testing is targeted, and $\theta_w$ is relatively small (\fref{pan2}(b), right three panels of top row).
+This can only occur when testing is targeted, and $\theta\_w$ is relatively small (\fref{pan2}(b), right three panels of top row).
 It is theoretically possible for increasing testing intensity to \emph{increase} \Rnum because more rapid testing leaves more susceptibles in the ``waiting-for-negative-results'' category at the DFE; if these people become infected while waiting, they will need to wait for their negative test result before they can be tested again, receive a positive test, and then begin self-isolating.
@@
 
-Over a broad range of parameter space (for random testing, approximately $\theta_w \geq 0.25$; for targeted testing, $\theta_w \geq 0.25$ and either $\theta_c \ge 0.5$ or $1/\omega > 5$), decreasing $\omega$ --- i.e., decreasing the rate at which tested people find out about their infection status --- decreases $\Rnum$.
+Over a broad range of parameter space (for random testing, approximately $\theta\_w \geq 0.25$; for targeted testing, $\theta\_w \geq 0.25$ and either $\theta\_c \ge 0.5$ or $1/\omega > 5$), decreasing $\omega$ --- i.e., decreasing the rate at which tested people find out about their infection status --- decreases $\Rnum$.
 This result is counterintuitive; public health agencies have invested a lot of effort in increasing this rate.
@@
 $\partial w_I(\omega)/\partial \omega > 0$), because rapid testing allows more tests to be concentrated on infected individuals.
-Second, in the real world individuals become less likely to maintain isolation if they are required to do so for longer; phenomenologically, we could allow  effectiveness of isolation while waiting for test returns an increasing function of test return speed ($\partial \theta_W(\omega)/\partial \omega > 0$), or we could introduce a separate ``waiting, but no longer isolating'' compartment that individuals entered from the ``waiting, isolated'' compartment at a specified rate.
+Second, in the real world individuals become less likely to maintain isolation if they are required to do so for longer; phenomenologically, we could allow  effectiveness of isolation while waiting for test returns an increasing function of test return speed ($\partial \theta\_w(\omega)/\partial \omega > 0$), or we could introduce a separate ``waiting, but no longer isolating'' compartment that individuals entered from the ``waiting, isolated'' compartment at a specified rate.
 Third and finally, if one wants to decrease the overall transmission rate of the population there are more effective methods than keeping tested people in limbo: masking, ventilation, distancing measures, retail and event closures, stay-at-home orders, \ldots
@@
 The other counterintuitive result from our analysis is that, for sufficiently high testing intensity $\rho$, increasing testing intensity can actually increase $\Rnum$ (e.g.
-\fref{pan2}(b), upper right panel [$\theta_c=1$, $\theta_w=0$]).
+\fref{pan2}(b), upper right panel [$\theta\_c=1$, $\theta\_w=0$]).
 This phenomenon occurs mainly because of a peculiar feature of the DFE in our model: we assume that testing is ongoing before the beginning of the epidemic, so that the starting point of the epidemic invasion is an equilibrium distribution of susceptibles between the $S\_n$ (untested) and $S\_u$ (waiting) compartments: in particular, $S\_u^* = \rho/(\rho + \omega)$.
-(In reality, testing intensity itself is a dynamic variable that will increase as the epidemic proceeds.) If isolation in the waiting compartment is low ($\theta_w \ll 1$), then waiting individuals can easily be infected once the epidemic starts.
+(In reality, testing intensity itself is a dynamic variable that will increase as the epidemic proceeds.) If isolation in the waiting compartment is low ($\theta\_w \ll 1$), then waiting individuals can easily be infected once the epidemic starts.
 Once infected they can still infect others while in the waiting compartment, and cannot be tested again before they have returned to the untested compartment.
-Thus, under certain extreme conditions (low $\theta_w$, high $\theta_c$, low $\omega$, high $\rho$), $\Rnum$ becomes an increasing function of testing intensity.
+Thus, under certain extreme conditions (low $\theta\_w$, high $\theta\_c$, low $\omega$, high $\rho$), $\Rnum$ becomes an increasing function of testing intensity.
 \bmb{This phenomenon occurs primarily (only??) for targeted testing, because \ldots CAN WE FIGURE OUT WHY?
@@
 % %%%%%%%
-\subsection{rate of returning tests} \label{app:omega}
+\subsection{Rate of returning tests} \label{app:omega}
 {\bf The third goal} is to explore how changes in the rate of test return $\omega$ affects $\Rnum$.
@@
 We use
-the linearization of $\Delta$ around $\rho=0$ to show that there a non-monotonic relationship between $\Rnum$  and $\omega$. The linear term in the Taylor expansion of $\Delta$ at $\rho=0$ is
+the linearization of $\Delta$ when $\rho \ll 1$ to show that there a non-monotonic relationship between $\Rnum$  and $\omega$. The linear term in the Taylor expansion of $\Delta$ when $\rho \ll 1$ is
 \begin{equation}
@@
 \pder\Delta{\omega} = \frac{\rho}{\omega^2 (\omega+\gamma)^2} \Big(
-(p_I \frac{w_I}{w_S}\theta\_c -(1+\frac{w_I}{w_S})\theta\_w)\omega^2+2\theta_w \gamma(\theta_w-2) \omega+\theta_w \gamma^2 (\theta_w-2)
-\Big),
+(p_I \frac{w_I}{w_S}\theta\_c -(1+\frac{w_I}{w_S})\theta\_w)\omega^2-2\theta\_w \gamma(2-\theta\_w) \omega+\theta\_w \gamma^2 (\theta\_w-2)
+\Big).
 \end{equation}
-around $\rho=0$.
-% \ali{I stoped here!}
+% ALi & Faddy's stuff
+The latter expression has two roots
+\begin{align}
+\label{eq:omega_roots}
+\omega^*_{+} &= \frac{\gamma\Big(-c_4 + \sqrt{c_4^2 + (a_4-b_4)c_4}\Big)}{b_4-a_4} \\
+\omega^*_{-} &= \frac{\gamma\Big(-c_4 - \sqrt{c_4^2 + (a_4-b_4)c_4}\Big)}{b_4-a_4},
+\end{align}
+where $a_4 = p_I\frac{w_I}{w_S}\theta\_c$, $b_4 = (1+\frac{w_I}{w_S})\theta\_w$, $c_4 = \theta\_w(2-\theta\_w)$.
+This enables us to describe the behaviour of $\Rnum$ with respect to $\omega$ in the foolowing two cases.
+\begin{itemize}
+\item \textbf{Case I:} If $b_4 \geq a_4$, $\pder\Delta{\omega} < 0$, so $\Rnum$ is always increasing with respect to $\omega$ (i.e., it is always harmful to return tests more rapidly). 
+\item \textbf{Case II:} If $b_4 < a_4$, $\Rnum$ will be decreasing with respect to $\omega$ (i.e., returning tests more rapidly is beneficial) only when $\omega > \omega^*_-.$
+\end{itemize}
+Note that $b_4 \geq a_4$ is characterized by
+\begin{equation}\label{case_1_ineq}
+\Big(\frac{w_I}{w_S}+1\Big)\theta\_w \geq \frac{w_I}{w_S}p_I\theta\_c \Longleftrightarrow \Big(\frac{w_S}{w_I}+1\Big)\frac{1}{p_I} \geq \frac{\theta\_c}{\theta\_w}.
+\end{equation}
+We begin with a \textbf{proof of Case I}.
+Suppose that $b_4 \geq a_4$.
+If the roots of $\pder\Delta{\omega}$ are not complex, then we must have $c_4 \geq b_4-a_4$.
+Note that $\omega^*_{-}$ must be negative since the numerator is clearly negative but the denominator is positive.
+Next, note that since $c_4 \geq b_4-a_4$, the numerator of $\omega^*_{+}$ must also be negative, so $\omega^*_{+}$ is negative.
+Thus, in this case, $\pder\Rnum{\omega}$ does not change sign on $(0,\infty)$.
+Checking the sign of \ref{eq:dlindo} for arbitrarily large $\omega$ shows that it is negative (since the $(a_4-b_4)\omega^2$ term dominates and is negative).
+So $\Rnum$ is increasing with respect to $\omega$ on all of $(0,\infty)$.
+
+We now turn our attention to a \textbf{proof of Case II}.
+Suppose that $b_4 < a_4$.
+It follows that $\omega^*_+$ and $\omega^*_-$ are real since $c_4 > 0 > b_4-a_4$.
+Next, note that $\omega^*_-$ is positive since both the numerator and denominator are negative.
+On the other hand, $\omega^*_+$ is negative since the denominator is negative but the numerator is positive (because $c_4 > b_4-a_4$).
+Thus, our task is to understand the sign of $\pder\Delta{\omega}$ around the root $\omega^*_-$.
+Checking the sign of \ref{eq:dlindo} for arbitrarily large $\omega$ shows that it is positive (since the $(a_4-b_4)\omega^2$ term dominates and is positive).
+Likewise, checking the sign for values of $\omega$ close to $0$ shows that it is negative.
+Thus, $\Rnum$ is increasing with respect to $\omega$ when $\omega < \omega^*_-$, and is decreasing $\omega > \omega^*_-$.
+
+Having presented the formal analysis, we now concern ourselves with its \textbf{biological interpretations.} We begin by interpreting \ref{case_1_ineq}, under which returning tests more rapidly is \emph{always} harmful.
+Notice that the ratio $\frac{\theta\_c}{\theta\_w}$ is simply a measure of how much more strongly individuals self-isolate when they test positive compared to when waiting for tests.
+Since the rate of test return directly influences the rate at which individuals change from a waiting state to a confirmed-positive state, it is intuitive that $\frac{\theta\_c}{\theta\_w}$ would appear in \ref{case_1_ineq}.
+Next, note that the left-hand side increases when test sensitivity decreases and when targeting of positive individuals is poor.
+This is consistent with our intuition: a false negative that is returned more rapidly may encourage an infectious individual to not carefully self-isolate, thus increasing transmission.
+Likewise, if individuals tested are mainly susceptible (rather than infectious), then returning tests more slowly would encourage them to self-isolate for longer while awaiting test results.
+Having understood the role of each of the parameters in \ref{case_1_ineq}, a ``holistic" interpretation of this inequality is that returning tests more slowly is helpful when the benefit of extended self-isolation (due to awaiting test results) outweighs the benefit of identifying positive cases.
+
+Now we interpret \textbf{Case II}.
+In this case, $\Rnum$ will have a global maximum with respect to $\omega$ at $\omega^*_-$.
+Note that our model assumption that $\rho < \omega$ plays an important role here: if $\omega^*_- < \rho$, then $\Rnum$ will be always decreasing with respect to $\omega$.
+On the other hand, if $\rho < \omega^*_-$, then $\Rnum$ will be increasing with respect to $\omega$ on $(\rho, \omega^*_-)$ and decreasing beyond that.
 

### 2021-06-16 SIR_manuscript.tex
minor fixes

@@
 This expression can be simplified and represented given the specific form of matrix $G_{11}$ \eqref{G11}.
-For the purpose of simplicity we present $\Rnum$ in the manuscript in terms of expressions $C$, $C1$ and $C2$, specified in \eqref{eq:C12}.
+For the purpose of simplicity we present $\Rnum$ in the manuscript in terms of expressions $C$, $C_1$ and $C_2$, specified in \eqref{eq:C12}.
 
@@
 -D_1 S\_n^* \boldsymbol {\theta\_w^2}+
-\big(-\omega^2 p_I \hat F_I S\_n^* \boldsymbol{\theta\_c} + D_2 S\_n^* + \gamma \hat F_I(\omega+\gamma) \big)\boldsymbol{\theta\_w}+
-(\omega+\gamma S\_u^*)\omega p_I \hat F_I \boldsymbol{\theta\_c}
+\big(-\omega^2 p_I \hat F_I S\_n^*\theta\_c+ D_2 S\_n^*+\gamma \hat F_I(\omega+\gamma) \big)\theta\_w+
+(\omega+\gamma S\_u^*)\omega p_I \hat F_I \theta\_c
 \Big),
@@
 \label{eq:del4_thetaw}
-1/C \Big( -D_1 S\_n^* \boldsymbol {\theta\_w^2}+
-\big(-\omega^2 p_I \hat F_I S\_n^* \boldsymbol{\theta\_c} + D_2 S\_n^* + \gamma \hat F_I(\omega+\gamma) \big)\boldsymbol{\theta\_w} \Big).
+1/C \Big( -D_1 S\_n^*\theta\_w^2+
+\big(-\omega^2 p_I \hat F_I S\_n^*\theta\_c+D_2 S\_n^*+\gamma \hat F_I(\omega+\gamma) \big)\theta\_w\Big).
 \end{equation}
@@
 Likewise, if individuals tested are mainly susceptible (rather than infectious), then returning tests more slowly would encourage them to self-isolate for longer while awaiting test results.
-Having understood the role of each of the parameters in \ref{case_1_ineq}, a ``holistic" interpretation of this inequality is that returning tests more slowly is helpful when the benefit of extended self-isolation (due to awaiting test results) outweighs the benefit of identifying positive cases.
+Having understood the role of each of the parameters in \ref{case_1_ineq}, a holistic interpretation of this inequality is that returning tests more slowly is helpful when the benefit of extended self-isolation (due to awaiting test results) outweighs the benefit of identifying positive cases.
 

### 2021-06-17 SIR_manuscript.tex
Comments and edits focused on middle of paper

@@
 \newcommand{\fref}[1]{Fig.~\ref{#1}}
+\newcommand{\appref}[1]{Sec.~\ref{app:#1}}
 \newcommand{\Rlogo}{R\xspace}
@@
 In addition to shaping epidemic observations (via case reports), testing processes also alter epidemiological dynamics.
-Because individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may also do so (possibly to a lesser extent), testing will generally increase the number of people who are isolating and hence reduce epidemic growth rates.
+Because individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may also do so, testing will generally increase the number of people who are isolating and hence reduce epidemic growth rates.
 We developed a mechanistic model that incorporates epidemic processes and testing in order to explore the effects of testing and isolation on epidemic dynamics.
@@
 primarily diagnostic (determining infection status for clinical purposes), or for control (determining  infection status in order to isolate cases that have been found by contact tracing), which we characterize as \emph{targeted} testing strategies.
-In these cases, testing probabilities vary widely across epidemiological compartments; in our dynamical model, we will characterize these probabilities by assigning a \percap testing weight to each compartment that determines the \emph{relative} probability that an individual in that compartment will be selected for testing (see Methods).
+In these cases, testing probabilities can differ sharply across epidemiological compartments; in our dynamical model, we will characterize these probabilities by assigning a testing weight to each compartment that determines the \emph{relative} probability that an individual in that compartment will be selected for testing (see Methods).
 
@@
 
-We developed a deterministic model, Eqs.~\eqref{model}, which groups individuals based on disease status and testing status.
-Disease states include Susceptible, Infectious and Recovered; testing status categorizes people as \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive} (\fref{fig:flowchart}).
+We developed a deterministic model which groups individuals based on disease status and testing status.
+Disease states include Susceptible, Infectious and Recovered; testing states include \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive} (\fref{fig:flowchart}).
 The testing status of an individual in  a given disease compartment $X$ (where $X \in \{S,I,R\}$) is denoted by a subscript, namely $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
-Further, two `accumulator' compartments, $N$ and $P$, are included in order to collect cumulative reported negative or positive tests.
-The model equations and details of calculation of the basic reproduction number $\Rnum$ are presented in the appendix (Sec.
-\ref{app:R0}).
+Two `accumulator' compartments, $N$ and $P$, are included in order to collect cumulative reported negative or positive tests.
+The model equations \eqref{model} and details of calculation of the basic reproduction number $\Rnum$ are presented in Sec.
+\appref{R0}.
 
@@
 where $\rho$ is the \percap testing intensity for the population, defined as the number of daily tests administered in a population of size $N_0$.
-Thus, the \percap testing rate for compartment $X$ is 
+Thus, the \percap testing rate for compartment $X \in \{S,I,R\}$ is 
 \begin{equation}
 \label{F}
-F_X=\sigma w_X, \quad \text{where} \quad X \in \{S,I,R\}.
+F_X=\sigma w_X.
 \end{equation}
-For a highly sensitive test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus unavailable for further testing.
+For a highly sensitive test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus not considered for further testing.
 Over the course of the epidemic, a sufficiently large fixed testing rate as specified in \eqref{sigma} can exhaust the pool of people available for testing, leading to a singularity when too few people are left untested to support the specified rate.
-Although this phenomenon does not affect our analysis of $\Rnum$, it can affect the temporal dynamics (in the appendix we present an adjustment to the model that solves this problem).
+Although this phenomenon does not affect our analysis of $\Rnum$, it can affect model dynamics (we present an adjustment to the model that solves this problem in Sec \appref{singularity}).
 
@@
   \end{equation}
-  with transmission rate $\beta$; $\theta\_w$ is the isolation efficacy (reduction of the probability of transmission) for individuals waiting for test results, while $\theta\_c$ is the isolation efficacy for individuals who have received a positive test (``confirmed positive'': Table~\ref{tab:params}).
+  with transmission rate $\beta$; $\theta\_w$ is the isolation efficacy (reduction of the probability of transmission) for individuals waiting for test results, while $\theta\_c$ is the isolation efficacy for individuals who have received a ``confirmed'' positive test (Table~\ref{tab:params}).
 Susceptible individuals who are waiting for test results experience an additional transmission factor of $1-\theta\_w$ (\fref{fig:flowchart}).
-\item confirmed-positive individuals isolate more effectively than those awaiting test results, i.e.
+\item confirmed-positive individuals isolate at least as effectively as those awaiting test results, i.e.
 $$0\leq\theta\_w\leq\theta\_c\leq 1.$$ 
@@
 For simplicity we assume that tests are perfectly \emph{specific} --- uninfected individuals never test positive ($p_s=0$).
-Thus there are no waiting-for-positive or confirmed-positive susceptible individuals (we assume that these states are also absent from the initial conditions, i.e.
-$S\_p(0)=S\_c(0)=0$), which reduces the number of model states from 12 to 10.
+Thus there are no waiting-for-positive or confirmed-positive susceptible individuals, which reduces the number of model states from 12 to 10.
 
@@
   \end{tabular}
-\caption{\label{tab:params} Parameters of the model \eqref{model}.}
+\caption{\label{tab:params} Parameters of the model specified in \eqref{model}; see also \fref{fig:flowchart}.}
 \end{table}
@@
 \end{equation}
-where the term $\beta/\gamma$ is the classical basic reproduction number for a SIR model without testing and isolation \citep{keeling2011modeling}, and $\Delta>0$ is the effectiveness of control due to testing and isolation processes defined as
+where the term $\beta/\gamma$ is the classical basic reproduction number for a simple SIR model \citep{keeling2011modeling}, and $0 < \Delta < 1$ is the proportional reduction due to testing and isolation processes, calculated as
 \begin{equation}
@@
 \end{align}
-(Appendix~\ref{app:R0} gives a detailed derivation of these expressions).
-The explicit formula of $\Rnum$ enables us to study the effects of testing and isolation parameters on $\Rnum$ both analytically and via numerical solutions.
+(\appref{R0} gives a detailed derivation of these expressions).
+This explicit formula enables us to study the effects of testing and isolation parameters on $\Rnum$ both analytically and via numerical solutions.
 We are specifically interested in parameters that could be manipulated by public health policy: isolation efficacy, $\theta\_c$ and $\theta\_w$; \percap testing intensity, $\rho$; and the rate of test return, $\omega$.
-In particular, we look at the partial derivatives of $\Delta$ with respect to these parameters (App.~\ref{app:rho}).
+In particular, we look at the partial derivatives of $\Delta$ with respect to these parameters (\appref{rho}).
 We derived general expressions for these derivatives.
@@
 
-The computations in this paper, specifically, calculation of the next-generation matrix and simplification of the $\Rnum$ expression, were performed in Maple\textsuperscript{\texttrademark} \citep{maple14}.
-We used R \citep{r} for numerical solutions, in particular for plotting the contours of $\Delta$ (Eq.~\eqref{eq:del4a}) over a range of parameters.
+Analytic calculation of the next-generation matrix and simplification of the $\Rnum$ expression, were performed in Maple\textsuperscript{\texttrademark} \citep{maple14}; numerical calculation and contour plots were done in R \citep{r}.
 We computed the values and contours of Delta at both low (\fref{pan}) and high (\fref{pan2}) testing intensities, and for both random testing ($w_S=w_I=w_R=1$) and targeted testing ($w_S=0.3$; $w_I=w_R=1$).
 
+\jd{Why do we use a different range of $\omega$ for the two cases?}
 The low-testing case \fref{pan} reflects the case where $\rho$ is small relative to the population size.
@@
 This testing intensity reflects realistic testing rates during the \covid pandemic, i.e.
-a maximum of 1.3\% of the population per day, giving a maximum of 10000 tests/day in a population of size $N_0=10^6$ (this is approximately four times the maximum testing rate in Ontario, Canada in mid-2021).
+a maximum of 1.3\% of the population per day, giving a maximum of 13000 tests/day in a population of size $N_0=10^6$ (this is approximately four times the maximum testing rate in Ontario, Canada in mid-2021).
 The less realistic high-testing case \fref{pan2} is included to highlight the occurrence of non-monotonic changes in $\Rnum$ with respect to $\rho$.
@@
   The critical contour of $\Delta=1-\frac{\gamma}{\beta}$, corresponding to the threshold $\Rnum=1$, is shown as a dotted line in \fref{pan} and \fref{pan2}; combinations of $\rho$ and $\omega$ that lie above this critical contour will control the epidemic.
-The implied baseline value of $\Rnum=\frac{\beta}{\gamma}$ in these figures is 3, corresponding to a case where the epidemic spreads slowly even in the absence of testing and tracing.
+The implied baseline value of $\Rnum=\frac{\beta}{\gamma}$ in these figures is 3.
 
@@
 
+\jd{Should we talk less about \Rnum\ and keep the focus on $\Delta$? The back and forth is a bit tiring.}
+
 We can use the formula for $\Rnum$ \eqref{R0} to make a number of straightforward conclusions about parameters whose effects on $\Rnum$ are monotonic, i.e.
 when the partial derivative of $\Delta$ has a consistent sign
-(Appendices~\ref{app:rho}, \ref{app:omega}, \ref{app:w}).
+(see Appendices).
 
@@
 
-Many of the qualitative results we have derived are mathematical confirmations of simple, common-sense intuitions.
+Many of the qualitative results we have derived confirm simple, common-sense intuitions.
 In particular, we can generally decrease $\Rnum$ by increasing isolation efficacy or testing intensity; returning tests faster, if individuals do not isolate while they are waiting for results; or increasing testing focus to target individuals who are likely to be infectious (e.g.
@@
 
-However, we did find two surprising trends: longer delays in returning tests can decrease the spread of the epidemic, and increasing testing rates can increase its spread.
+However, we did find two surprising patterns: longer delays in returning tests can decrease the spread of the epidemic, and increasing testing rates can increase its spread.
 
@@
 detecting and containing outbreaks while they are still in progress.
-This process could be modeled phenomenologically by making the testing focus an increasing function of the speed of test returns (i.e.
-$\partial w_I(\omega)/\partial \omega > 0$), because rapid testing allows more tests to be concentrated on infected individuals.
-Second, in the real world individuals become less likely to maintain isolation if they are required to do so for longer; phenomenologically, we could allow  effectiveness of isolation while waiting for test returns an increasing function of test return speed ($\partial \theta\_w(\omega)/\partial \omega > 0$), or we could introduce a separate ``waiting, but no longer isolating'' compartment that individuals entered from the ``waiting, isolated'' compartment at a specified rate.
-Third and finally, if one wants to decrease the overall transmission rate of the population there are more effective methods than keeping tested people in limbo: masking, ventilation, distancing measures, retail and event closures, stay-at-home orders, \ldots
+This process could be modeled phenomenologically by making the testing focus increase as an increasing proportion of cases are detected, because finding infections allows tests to be concentrated on their connections.
+Second, in the real world individuals become less likely to maintain isolation if they are required to do so for longer; phenomenologically, we could allow  effectiveness of isolation while waiting for test returns to be an increasing function of test return speed, or we could introduce a separate ``waiting, but no longer isolating'' compartment that individuals entered from the ``waiting, isolated'' compartment at a specified rate.
+Third and finally, if one wants to decrease the overall transmission rate of the population there are more effective methods than keeping tested people in limbo -- these include masking, ventilation, distancing measures, retail and event closures, or stay-at-home orders.
 
@@
 \end{align}
-In realistic cases the testing rate $\rho$ is very small (i.e., only a small fraction of the population can be tested every day); it is thus reasonable to use a linear approximation of $\Rnum$ for $\rho \ll 1$ to analyze the behaviour of $\Rnum$ with respect to $\omega$ (see section \ref{app:omega}).
+In realistic cases the testing rate $\rho$ is very small (i.e., only a small fraction of the population can be tested every day); it is thus reasonable to use a linear approximation of $\Rnum$ for $\rho \ll 1$ to analyze the behaviour of $\Rnum$ with respect to $\omega$ (see section \appref{omega}).
 In the next section we provide an equivalent representation of $\Rnum$ in order to show that increasing testing intensity typically decreases $\Rnum$.
@@
 % %%%%%%%
-\subsection{On Testing Rate and Numerical Singularity}
+\subsection{On Testing Rate and Numerical Singularity} \label{app:singularity}
 

### 2021-06-18 SIR_manuscript.tex
working on the comment

@@
 When testing is done for the purposes of disease surveillance \citep{foddai2020base}
-tests are typically assigned randomly (or using a stratified random design) across the population in
-order to make an unbiased assessment of population prevalence.
+tests are typically assigned randomly (or using a stratified random design) across the population in order to make an unbiased assessment of population prevalence.
 
@@
 %% BMB left out the converse; don't think we have to be super-precise here
-Although the value of $\Rnum$ cannot completely characterize the dynamics of even the simplest epidemic model
-\citep{shaw2021what}, it does give a simple and widely accepted index for the difficulty of control, as well as an indication of the likely final size of an epidemic \citep{ma2006generality}.
+Although the value of $\Rnum$ cannot completely characterize the dynamics of even the simplest epidemic model \citep{shaw2021what}, it does give a simple and widely accepted index for the difficulty of control, as well as an indication of the likely final size of an epidemic \citep{ma2006generality}.
 
@@
 Here, we studied the the effect of testing intensity, rate of test return and the isolation efficacy in reduction of the probability of transmission on the epidemic dynamics when different levels of testing focus (from random to highly targeted) are in place.
-% Our model provides insights to the sensitivity of the epidemic dynamics, through $\Rnum$, with respect to the underlying testing and isolation parameters.
-%% BMB: last sentence seems to belong in conclusions
 
@@
 Two `accumulator' compartments, $N$ and $P$, are included in order to collect cumulative reported negative or positive tests.
-The model equations \eqref{model} and details of calculation of the basic reproduction number $\Rnum$ are presented in Sec.
-\appref{R0}.
+The model equations \eqref{model} and details of calculation of the basic reproduction number $\Rnum$ are presented in \appref{R0}.
 
@@
 Over the course of the epidemic, a sufficiently large fixed testing rate as specified in \eqref{sigma} can exhaust the pool of people available for testing, leading to a singularity when too few people are left untested to support the specified rate.
-Although this phenomenon does not affect our analysis of $\Rnum$, it can affect model dynamics (we present an adjustment to the model that solves this problem in Sec \appref{singularity}).
+Although this phenomenon does not affect our analysis of $\Rnum$, it can affect model dynamics (we present an adjustment to the model that solves this problem in \appref{singularity}).
 
@@
 We derived general expressions for these derivatives.
-However, we analyzed the effect of $\omega$ on $\Rnum$ for the special case of low testing intensity.
-Specifically, by making the restriction $\rho \ll 1$, we are able to Taylor-expand $\Delta$ at $\rho=0$, use the linear approximation with respect to $\rho$ and analyze the resulting simplified derivatives to illustrate a surprising non-monotonic relationship between $\Rnum$ and $\omega$.
+However, we analyzed the effect of $\omega$ on $\Delta$ for the special case of low testing intensity.
+Specifically, by making the restriction $\rho \ll 1$, we are able to Taylor-expand $\Delta$ at $\rho=0$, use the linear approximation with respect to $\rho$ and analyze the resulting simplified derivatives to illustrate a surprising non-monotonic relationship between $\Delta$ and $\omega$.
 
 Analytic calculation of the next-generation matrix and simplification of the $\Rnum$ expression, were performed in Maple\textsuperscript{\texttrademark} \citep{maple14}; numerical calculation and contour plots were done in R \citep{r}.
-We computed the values and contours of Delta at both low (\fref{pan}) and high (\fref{pan2}) testing intensities, and for both random testing ($w_S=w_I=w_R=1$) and targeted testing ($w_S=0.3$; $w_I=w_R=1$).
+We computed the values and contours of $\Delta$ at both low (\fref{pan}) and high (\fref{pan2}) testing intensities, and for both random testing ($w_S=w_I=w_R=1$) and targeted testing ($w_S=0.3$; $w_I=w_R=1$).
 
-\jd{Why do we use a different range of $\omega$ for the two cases?}
+\jd{Why do we use a different range of $\omega$ for the two cases?}\ali{added a sentence to the end of this paragraph to address this (I hope).}
 The low-testing case \fref{pan} reflects the case where $\rho$ is small relative to the population size.
@@
 In \fref{pan2} the maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,1/5)$ and the test return rate $\omega\in [1/5,2]$; these values are clearly unrealistic for a large population but might be relevant for some population undergoing focused testing, such as a sports league or university.
-  The critical contour of $\Delta=1-\frac{\gamma}{\beta}$, corresponding to the threshold $\Rnum=1$, is shown as a dotted line in \fref{pan} and \fref{pan2}; combinations of $\rho$ and $\omega$ that lie above this critical contour will control the epidemic.
-The implied baseline value of $\Rnum=\frac{\beta}{\gamma}$ in these figures is 3.
+The implied baseline value of $\Rnum=\frac{\beta}{\gamma}$, corresponds to a simple SIR model, in these figures is 3.
+Note that the different range of $\omega$ for the two cases of low and high testing intensities is due to the restriction $\rho<\omega$ which is a requirement for feasible DFE (\ref{dfe}).
 
@@
 Our model is a caricature --- it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g.
-exposed, presymptomatic, and asymptomatic compartments [REFS]\ali{\citep{kain2021chopping} OK?}; time-varying testing rates [REFS?]; behavioural dynamics [REF Weitz/Park/Dushoff] \ali{\citep{weitz2020awareness} OK?}).
+exposed, presymptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates [REFS?]; behavioural dynamics \citep{weitz2020awareness}).
 Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between epidemiology and testing, rather than for making quantitative predictions or guiding pandemic responses.
@@
 }
-
-\ali{This belongs to introduction?}
+\ali{In respense to Ben's comment}
 There have been several modeling works of testing and tracing dynamics and their interaction with epidemiological dynamics.

### 2021-06-18 SIR_manuscript.tex
minor discussion updates

@@
 Bergstrom et al might be a counterexample, although it's for a different context [repeated testing of an isolated population]}.
+\ali{the following sentences added to address Ben' comment:} There have been several modeling works of testing and tracing dynamics and their interaction with epidemiological dynamics.
+In the context of repeated screaning and random testing of isolated populations (such as a university setup), \cite{bergstrom2020frequency} provided some analytical results, \cite{rogers2021high} simulated a SEIR model with high-low testing intensity and sensitivity.
+Further, \cite{friston2021testing} model the effects of self-isolation on testing and tracking.
+Our modeling approache is novel compared to these works.
 We hope that this paper will inspire further explorations of the fundamental properties of dynamical systems that incorporate explicit testing models in an epidemiological context.
@@
 \ali{In respense to Ben's comment}
-There have been several modeling works of testing and tracing dynamics and their interaction with epidemiological dynamics.
-In the context of repeated screaning and random testing of isolated populations (such as a university setup),
-\cite{bergstrom2020frequency} provided some analytical results, 
-\cite{rogers2021high} simulated a SEIR model with high-low testing intensity and sensitivity.
-Further, \cite{friston2021testing} model the effects of self-isolation on testing and tracking.  
-Our modeling approache is novel compared to these works.
 

### 2021-06-22 SIR_manuscript.tex
revisions to introduction

@@
 \newcommand{\jd}[1]{\comment{blue}{JD}{#1}}
-\newcommand{\david}[1]{\comment{red}{DJDE}{#1}}
+\newcommand{\david}[1]{\comment{orange}{DJDE}{#1}}
 \newcommand{\bmb}[1]{\comment{red}{BMB}{#1}}
@@
 
-\title{Testing and Isolation Efficacy: Insights from a Simple Epidemic Model}
+\title{Testing and Isolation Efficacy:\\Insights from a Simple Epidemic Model}
+\author{\david{??? authors ??? affiliations ???}}
 
@@
 \bmb{abstract needs another editing pass, skipping for now}
+\david{This needs substantial tightening.}
 
@@
 However, the potential advantage of a slow rate of test return on the dynamics of an epidemic is real, often neglected, and needs to be quantified.
-At the individual level, this advantage can manifest in the following sense: individuals awaiting their test results or who have tested positive may partially or fully self-isolate, thus reducing or eliminating their potential in the transmission process.
+At the individual level, this advantage can manifest in the following sense: individuals awaiting their test results or who have tested positive may partially or fully self-isolate, thus reducing or eliminating their transmission potential.
 In this paper, we investigated this individual-level effect of testing processes on the epidemic dynamics by developing a SIR-type model.
@@
 When testing is done for the purposes of disease surveillance \citep{foddai2020base}
-tests are typically assigned randomly (or using a stratified random design) across the population in order to make an unbiased assessment of population prevalence.
+tests are typically conducted randomly (or using a stratified random design) across the population in order to make an unbiased assessment of population prevalence.
 
@@
 
-As epidemiologists, we want to know whether an epidemic will initially grow exponentially following its introduction in a susceptible population.
-In mathematical epidemiology, we determine this outcome by studying the basic reproduction number $\Rnum$, defined as the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}.
-In the early stages of an epidemic the number of infected individuals is expected to grow exponentially over time when $\Rnum>1$.
-%% BMB left out the converse; don't think we have to be super-precise here
-Although the value of $\Rnum$ cannot completely characterize the dynamics of even the simplest epidemic model \citep{shaw2021what}, it does give a simple and widely accepted index for the difficulty of control, as well as an indication of the likely final size of an epidemic \citep{ma2006generality}.
+When a new infectious disease emerges, it is important to determine whether it will grow exponentially in a susceptible population, and if so at what rate $r$.
+The condition for positive exponential growth ($r>0$) is commonly expressed as $\Rnum>1$, where the basic reproduction number $\Rnum$ is the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}.
+Although the value of $\Rnum$ cannot completely characterize the dynamics of even the simplest epidemic model \citep{shaw2021what}, \david{well, technically it does up to rescaling time} it does give a simple and widely accepted index for the difficulty of control, as well as an indication of the likely final size of an epidemic \citep{ma2006generality}.
 
-In order to understand the effect of testing processes on an epidemic dynamics, we developed a mechanistic SIR-type model with epidemic and testing components.
+In order to understand the effect of testing processes on epidemic dynamics, we expanded one of the simplest mechanistic epidemic model---the standard deterministic SIR model---to include testing components.
 This model provides a sensible platform to link the modeling of epidemic and testing components and study their interaction.
-Here, we studied the the effect of testing intensity, rate of test return and the isolation efficacy in reduction of the probability of transmission on the epidemic dynamics when different levels of testing focus (from random to highly targeted) are in place.
+We studied the effects of testing intensity, rate of test return, and isolation efficacy, on transmission probability and epidemic dynamics when different levels of testing focus (from random to highly targeted) are in place.
 

### 2021-06-22 SIR_manuscript.tex
revisions to Methods

@@
 \newcommand{\fref}[1]{Fig.~\ref{#1}}
-\newcommand{\appref}[1]{Sec.~\ref{app:#1}}
+\newcommand{\appref}[1]{Appendix~\ref{app:#1}}
 \newcommand{\Rlogo}{R\xspace}
@@
 
-We developed a deterministic model which groups individuals based on disease status and testing status.
-Disease states include Susceptible, Infectious and Recovered; testing states include \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive} (\fref{fig:flowchart}).
+Our model groups individuals based on disease status (Susceptible, Infectious or Recovered) and testing status (\emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}) (\fref{fig:flowchart}).
 The testing status of an individual in a given disease compartment $X$ (where $X \in \{S,I,R\}$) is denoted by a subscript, namely $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
@@
 where $\rho$ is the \percap testing intensity for the population, defined as the number of daily tests administered in a population of size $N_0$.
+\david{Why $N_0$ rather than just $N$?}
 Thus, the \percap testing rate for compartment $X \in \{S,I,R\}$ is 
@@
 \end{equation}
+\david{$F$ isn't a natural symbol for ``testing rate''.  How about $\mathcal T_X$?}
 For a highly sensitive test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus not considered for further testing.
@@
 \begin{enumerate}[(i)]
-\item there is a single force of infection (new cases per unit time), $\Lambda$, defined as
+\item there is a single force of infection (new \david{infections} per unit time \david{per susceptible}), $\Lambda$, defined as
   \begin{equation}
@@
   with transmission rate $\beta$; $\theta\_w$ is the isolation efficacy (reduction of the probability of transmission) for individuals waiting for test results, while $\theta\_c$ is the isolation efficacy for individuals who have received a ``confirmed'' positive test (Table~\ref{tab:params}).
-Susceptible individuals who are waiting for test results experience an additional transmission factor of $1-\theta\_w$ (\fref{fig:flowchart}).
-\item confirmed-positive individuals isolate at least as effectively as those awaiting test results, i.e.
+Susceptible individuals who are waiting for test results experience an additional transmission reduction factor of $1-\theta\_w$ (\fref{fig:flowchart}).
+\item confirmed-positive individuals isolate at least as effectively as those awaiting test results, i.e.,
 $$0\leq\theta\_w\leq\theta\_c\leq 1.$$ 
@@
 
-The Disease-Free Equilibrium (DFE) for the SIR model (Eqs.~\ref{model}) is found by setting the infected compartments to 0 and solving for the unknowns. The DFE is
+The Disease-Free Equilibrium (DFE) for the expanded SIR model (Eqs.~\ref{model}) is found by setting the infected compartments to 0 and solving for the unknowns. The DFE is
 \begin{equation}
 \label{dfe}
-S\_n^*= \frac{\rho}{\omega} N_0, \ S\_u^*= \frac{\omega-\rho}{\omega} N_0, \text{~and~} I_j=R_j=0 \ \text{for all }j.
+S\_n^*= \frac{\rho}{\omega} N_0, \ S\_u^*= N_0-S_n^*, \text{~and~} I_j=R_j=0 \ \text{for all }j.
 \end{equation}
@@
 \end{equation}
-where the term $\beta/\gamma$ is the classical basic reproduction number for a simple SIR model \citep{keeling2011modeling}, and $0 < \Delta < 1$ is the proportional reduction due to testing and isolation processes, calculated as
+where the term $\beta/\gamma$ is the classical basic reproduction number for a simple SIR model \citep{keeling2011modeling}, and $0 < \Delta < 1$ is the proportional reduction \david{actually $1-\Delta$ is the proportional reduction} due to testing and isolation processes, calculated as
 \begin{equation}
@@
 \end{align}
-(\appref{R0} gives a detailed derivation of these expressions).
+(\appref{R0} gives a detailed derivation of these expressions.)
 This explicit formula enables us to study the effects of testing and isolation parameters on $\Rnum$ both analytically and via numerical solutions.
@@
 \jd{Why do we use a different range of $\omega$ for the two cases?}\ali{added a sentence to the end of this paragraph to address this (I hope).}
-The low-testing case \fref{pan} reflects the case where $\rho$ is small relative to the population size.
+The low-testing case \fref{pan} reflects the case where testing intensity $\rho$ is small relative to the population size.
 Specifically, $\rho \in [0,0.013]$, and test return rate $\omega\in [1/12,2]$.
-This testing intensity reflects realistic testing rates during the \covid pandemic, i.e.
-a maximum of 1.3\% of the population per day, giving a maximum of 13000 tests/day in a population of size $N_0=10^6$ (this is approximately four times the maximum testing rate in Ontario, Canada in mid-2021).
+This testing intensity reflects realistic testing rates during the \covid pandemic, i.e., a maximum of 1.3\% of the population per day \david{If it is 4x larger than Ontario, what makes it ``realistic''?} (this is approximately four times the maximum testing rate in Ontario, Canada in mid-2021).
 The less realistic high-testing case \fref{pan2} is included to highlight the occurrence of non-monotonic changes in $\Rnum$ with respect to $\rho$.
-In \fref{pan2} the maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,1/5)$ and the test return rate $\omega\in [1/5,2]$; these values are clearly unrealistic for a large population but might be relevant for some population undergoing focused testing, such as a sports league or university.
-The implied baseline value of $\Rnum=\frac{\beta}{\gamma}$, corresponds to a simple SIR model, in these figures is 3.
-Note that the different range of $\omega$ for the two cases of low and high testing intensities is due to the restriction $\rho<\omega$ which is a requirement for feasible DFE (\ref{dfe}).
+In \fref{pan2} the maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,1/5)$ and the test return rate $\omega\in [1/5,2]$; these values are clearly unrealistic for a large population but might be relevant for small populations undergoing focused testing, such as a sports league or university.
+In these figures, the implied baseline reproduction number (for the SIR model without testing) is $\Rnum=\frac{\beta}{\gamma}=3$.
+Note that the different range of test return rate $\omega$ for the two cases of low and high testing intensities is due to the restriction $\rho<\omega$ which is a requirement for feasible DFE (\ref{dfe}).
 

### 2021-06-22 SIR_manuscript.tex
revisions to Results

@@
 \end{equation}
-where the term $\beta/\gamma$ is the classical basic reproduction number for a simple SIR model \citep{keeling2011modeling}, and $0 < \Delta < 1$ is the proportional reduction \david{actually $1-\Delta$ is the proportional reduction} due to testing and isolation processes, calculated as
+where the term $\beta/\gamma$ is the classical basic reproduction number for a simple SIR model \citep{keeling2011modeling}, and $0 < \Delta < 1$ is the proportional reduction \david{actually $1-\Delta$ is the proportional reduction; you should refer to this as the effectiveness of control, as you do below} due to testing and isolation processes, calculated as
 \begin{equation}
@@
 
-We can use the formula for $\Rnum$ \eqref{R0} to make a number of straightforward conclusions about parameters whose effects on $\Rnum$ are monotonic, i.e.
-when the partial derivative of $\Delta$ has a consistent sign
-(see Appendices).
+We can use the formula for $\Rnum$ \eqref{R0} to make a number of straightforward inferences about parameters that affect $\Rnum$ monotonically, i.e., for which the associated partial derivative of $\Delta$ always has the same sign (see Appendices).
 
@@
 \item \label{p1:omega} Increasing the rate of test return ($\omega$) always decreases \Rnum if waiting individuals do not isolate ($\theta\_w=0$) (Eq.~\ref{eq:dlindo}).
-\item \label{p1:w} Increasing testing focus, i.e. changing the testing weights from random ($w_S=w_I$) toward targeted  ($w_S<w_I$), always decreases \Rnum (Eq.~\ref{eq:d_del_wis}).
+\item \label{p1:w} Increasing testing focus, i.e., changing the testing weights from random ($w_S=w_I$) toward targeted  ($w_S<w_I$), always decreases \Rnum (Eq.~\ref{eq:d_del_wis}).
 \end{enumerate}
@@
 \caption{
-{\bf Effectiveness of testing and isolation in reducing $\Rnum$ at low \percap testing intensity.}
-Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4}), over a range of testing and isolation parameters. Parameter values (Table~\ref{tab:params}):
-$\beta=\betaparam$ 1/day, $1/\gamma= \invgammaparam$ days (baseline $\Rnum=\Rnumparam$, $r=\rparam$); $\omega \in [1/12,2]$ 1/day;  $\rho \in [0,0.013]$ 1/day per capita; $\theta\_w$ and $\theta\_c$ vary between 0 (no effect of isolation) and 1 (complete elimination of transmission); $p_S=0$, $p_I=1$ and $p_R=0.5$.
+{\bf Effectiveness of testing and isolation in reducing $\Rnum$ at low \percap testing intensity ($\rho$).}
+Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4}\david{wrong eq ref}), over a range of testing and isolation parameters. Parameter values (Table~\ref{tab:params}):
+$\beta=\betaparam/$day, $1/\gamma= \invgammaparam$ days (baseline $\Rnum=\Rnumparam$, $r=\rparam$); $\omega \in [1/12,2]/$day;  $\rho \in [0,0.013]/$day per capita; $\theta\_w$ and $\theta\_c$ vary between 0 (no effect of isolation) and 1 (complete elimination of transmission); $p_S=0$, $p_I=1$ and $p_R=0.5$.
 Only parameter sets where $\theta\_c \geq \theta\_w$ (confirmed-positive individuals isolate more effectively than waiting individuals) are shown; the alternative case, $\theta\_w > \theta\_c$, is unrealistic.
@@
   {\bf Effectiveness of testing and isolation in reducing $\Rnum$ at high \percap testing intensity.}
-  Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4}), over a range of testing and isolation parameters.
-Parameters as in \fref{pan} except: $\omega \in [1/5,2]$ 1/day, $\rho \in [0,1/5)$ 1/day.
+  Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4}\david{again wrong eq ref}), over a range of testing and isolation parameters.
+Parameters as in \fref{pan} except: $\omega \in [1/5,2]/$day, $\rho \in [0,1/5)/$day.
 As in \fref{pan}, subplots show (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
@@
 This is in fact what happens when waiting individuals do not isolate ($\theta\_w =0$, top row of \fref{pan}) --- as we move to the right within each plot in this row, $\Delta$ decreases.
-However, when waiting individuals isolate ($\theta\_w>0$), we more often see the opposite effect: longer testing delays lead to a greater control effect $\Delta$.
-This is because people waiting for negative tests are assumed to continue to isolate; this applies both to susceptibles and to people who became infected while waiting for negative test results.
-This effect outweighs the effect of confirmed individuals isolating except when this isolation parameter $\theta\_c$ is substantially bigger than $\theta\_w$.
+However, when waiting individuals isolate ($\theta\_w>0$), we more often see the opposite effect: longer testing delays lead to a greater control effect $\Delta$ (reduced $\Rnum$).
+The reason is that people waiting for negative tests are assumed to continue to isolate; this applies both to susceptibles and to people who became infected while waiting for negative test results.
+This effect outweighs the effect of confirmed individuals isolating, except when this isolation parameter ($\theta\_c$) is substantially greater than $\theta\_w$.
 This result depends on the idea that, all else equal, people who have to wait longer for test results isolate at the same level (but for a longer time) as they would if the wait were shorter.
 \item \fref{pan} also shows that greater testing intensity (increasing $\rho$) generally increases the effectiveness of control (moving up in each panel).
-This relationship can be reversed, however, at very high testing intensities.
-This can only occur when testing is targeted, and $\theta\_w$ is relatively small (\fref{pan2}(b), right three panels of top row).
+However, this relationship can be reversed at very high testing intensities (provided testing is targeted, and $\theta\_w$ is relatively small; \fref{pan2}(b), right three panels of top row).
 It is theoretically possible for increasing testing intensity to \emph{increase} \Rnum because more rapid testing leaves more susceptibles in the ``waiting-for-negative-results'' category at the DFE; if these people become infected while waiting, they will need to wait for their negative test result before they can be tested again, receive a positive test, and then begin self-isolating.

### 2021-06-23 SIR_manuscript.tex
authors and affiliation added

@@
 \usepackage{enumerate}
+\usepackage{authblk}
 
@@
 \title{Testing and Isolation Efficacy:\\Insights from a Simple Epidemic Model}
-\author{\david{??? authors ??? affiliations ???}}
+\author[1]{Ali Gharouni}
+\author[1]{Fady M. Abdelmalek}
+\author[1,2,3]{Benjamin M. Bolker}
+\author[2,4,3]{Jonathan Dushoff}
+\author[3,1,5]{David J. D. Earn}
+
+\affil[1]{Department of Mathematics \& Statistics, McMaster University, Hamilton, Canada}
+\affil[2]{Department of Biology, McMaster University, Hamilton, Canada}
+\affil[3]{Michael G. DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, Canada}
+\affil[4]{South African Centre for Epidemiological Modelling and Analysis, University of Stellenbosch, Stellenbosch, South Africa}
+\affil[5]{Department of Mathematics, University of Toronto, Toronto, Canada}
 

### 2021-06-23 SIR_manuscript.tex
more bmb tweaks

@@
 \newcommand{\pro}[1]{\ensuremath{\frac{\partial #1}{\partial \rho}}}
-\newcommand\pder[2]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder y{x}
+\newcommand{\pder}[2]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder y{x}
+\newcommand{\testing}[1]{\ensuremath{{\cal T}_{#1}}\xspace}
+\newcommand{\testinghat}[1]{\ensuremath{\hat{\cal T}_{#1}}\xspace}
 
@@
 The disease-based status of a compartment $X$ ($X \in \{S,I,R\}$) is combined with the testing status including $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
-The force of infection is denoted by $\Lambda$ (Eq.~\eqref{Lambda}); $\gamma$ is the recovery rate; $\omega$ is the rate of test return; and $F_X$ (Eq.
+The force of infection is denoted by $\Lambda$ (Eq.~\eqref{Lambda}); $\gamma$ is the recovery rate; $\omega$ is the rate of test return; and \testing{X} (Eq.
 \eqref{F}) and $p_X$ represent the \percap testing rate and the sensitivity (probability of positive tests), respectively, for compartment $X$.
@@
 \label{F}
-F_X=\sigma w_X.
+\testing{X}=\sigma w_X.
 \end{equation}
 \david{$F$ isn't a natural symbol for ``testing rate''.  How about $\mathcal T_X$?}
+\bmb{mostly fixed, but needs to be fixed in the flow chart as well}
 For a highly sensitive test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus not considered for further testing.
@@
 \begin{enumerate}[(i)]
-\item there is a single force of infection (new \david{infections} per unit time \david{per susceptible}), $\Lambda$, defined as
+\item there is a single force of infection (new infections per unit time per susceptible), $\Lambda$, defined as
   \begin{equation}
@@
 \label{eq:fi}
-\hat F_I = (\omega\rho/(\omega-\rho))w_I/w_S \quad .
+\testinghat{I} = (\omega\rho/(\omega-\rho))w_I/w_S \quad .
 \end{equation}
@@
 \label{eq:C12}
-C_{\phantom 1} &= (\omega+\gamma) \Big(\gamma(\omega+\gamma)+(\gamma+\omega p_I)\hat F_I\Big),\\
-C_1 &= (\omega+\gamma)(\theta\_w \gamma+\theta\_c \omega p_I) \hat F_I,\\
-C_2 &= \Big( \omega\gamma(1+p_I)\hat F_I+\gamma^2(\omega+\gamma+\hat F_I)\Big)\theta\_w + \omega^2p_I\hat F_I\theta\_c.
+C_{\phantom 1} &= (\omega+\gamma) \Big(\gamma(\omega+\gamma)+(\gamma+\omega p_I)\testinghat{I}\Big),\\
+C_1 &= (\omega+\gamma)(\theta\_w \gamma+\theta\_c \omega p_I) \testinghat{I},\\
+C_2 &= \Big( \omega\gamma(1+p_I)\testinghat{I}+\gamma^2(\omega+\gamma+\testinghat{I})\Big)\theta\_w + \omega^2p_I \testinghat{I} \theta\_c.
 \end{align}
@@
 We computed the values and contours of $\Delta$ at both low (\fref{pan}) and high (\fref{pan2}) testing intensities, and for both random testing ($w_S=w_I=w_R=1$) and targeted testing ($w_S=0.3$; $w_I=w_R=1$).
+Because it is expressed as a proportion of $\Rnum$, the effectiveness of control $\Delta$ is (at least in the $\rho \ll 1$ case, Eq.~\ref{lin}) independent of the transmission rate $\beta$, and hence of $\Rnum$ in the case where we vary $\Rnum$ by changing the transmission rate for a fixed generation interval.
 
-\jd{Why do we use a different range of $\omega$ for the two cases?}\ali{added a sentence to the end of this paragraph to address this (I hope).}
 The low-testing case \fref{pan} reflects the case where testing intensity $\rho$ is small relative to the population size.
 Specifically, $\rho \in [0,0.013]$, and test return rate $\omega\in [1/12,2]$.
-This testing intensity reflects realistic testing rates during the \covid pandemic, i.e., a maximum of 1.3\% of the population per day \david{If it is 4x larger than Ontario, what makes it ``realistic''?} (this is approximately four times the maximum testing rate in Ontario, Canada in mid-2021).
+This testing intensity is of the correct order of magnitude (although typically larger than) testing rates during the \covid pandemic, i.e., a maximum of 1.3\% of the population per day (approximately four times the maximum testing rate in Ontario, Canada in mid-2021).
 The less realistic high-testing case \fref{pan2} is included to highlight the occurrence of non-monotonic changes in $\Rnum$ with respect to $\rho$.
@@
 In these figures, the implied baseline reproduction number (for the SIR model without testing) is $\Rnum=\frac{\beta}{\gamma}=3$.
-Note that the different range of test return rate $\omega$ for the two cases of low and high testing intensities is due to the restriction $\rho<\omega$ which is a requirement for feasible DFE (\ref{dfe}).
+The different ranges of test return rates $\omega$ for the cases of low and high testing intensities is due to the restriction $\rho<\omega$, which is a requirement for a feasible DFE (\ref{dfe}).
 
@@
 {\bf Effectiveness of testing and isolation in reducing $\Rnum$ at low \percap testing intensity ($\rho$).}
-Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4}\david{wrong eq ref}), over a range of testing and isolation parameters. Parameter values (Table~\ref{tab:params}):
+Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4_theta}\david{wrong eq ref}\bmb{maybe fixed?
+check}), over a range of testing and isolation parameters.
+Parameter values (Table~\ref{tab:params}):
 $\beta=\betaparam/$day, $1/\gamma= \invgammaparam$ days (baseline $\Rnum=\Rnumparam$, $r=\rparam$); $\omega \in [1/12,2]/$day;  $\rho \in [0,0.013]/$day per capita; $\theta\_w$ and $\theta\_c$ vary between 0 (no effect of isolation) and 1 (complete elimination of transmission); $p_S=0$, $p_I=1$ and $p_R=0.5$.
@@
 \ali{the following sentences added to address Ben' comment:} There have been several modeling works of testing and tracing dynamics and their interaction with epidemiological dynamics.
-In the context of repeated screaning and random testing of isolated populations (such as a university setup), \cite{bergstrom2020frequency} provided some analytical results, \cite{rogers2021high} simulated a SEIR model with high-low testing intensity and sensitivity.
+In the context of repeated screening and random testing of isolated populations (such as a university setup), \cite{bergstrom2020frequency} provided some analytical results, \cite{rogers2021high} simulated a SEIR model with high-low testing intensity and sensitivity.
 Further, \cite{friston2021testing} model the effects of self-isolation on testing and tracking.
-Our modeling approache is novel compared to these works.
+Our modeling approach is novel compared to these works.
 We hope that this paper will inspire further explorations of the fundamental properties of dynamical systems that incorporate explicit testing models in an epidemiological context.
@@
 \begin{itemize}
-\item should emphasize (at some appropriate point in text) that $\Delta$ is independent of $\beta$ (and hence of $\Rnum$ if we are thinking primarily of holding $\gamma$ constant and varying $\beta$) for the small-$\rho$ (Taylor-expansion) case, see Eq.~\ref{eq:lin}.
 \item it would be good to have a few more testing refs: Bergstrom et al., there might be some other stuff we can recycle from the lab meeting review; Friston's stuff deserves to be mentioned because it does have explicit testing flows too.
@@
 \begin{align}
- d S\_u/dt &= -\Lambda S\_u - F_S S\_u + \omega S\_n, \\
- d S\_n/dt &= -(1-\theta\_w)\Lambda S\_n + (1-p_S) F_S S\_u - \omega S\_n, \\
- d S\_p/dt &= -(1-\theta\_w)\Lambda S\_p + p_S F_S S\_u - \omega S\_p, \\
+ d S\_u/dt &= -\Lambda S\_u - \testing{S} S\_u + \omega S\_n, \\
+ d S\_n/dt &= -(1-\theta\_w)\Lambda S\_n + (1-p_S) \testing{S} S\_u - \omega S\_n, \\
+ d S\_p/dt &= -(1-\theta\_w)\Lambda S\_p + p_S \testing{S} S\_u - \omega S\_p, \\
  d S\_c/dt &= -(1-\theta\_c)\Lambda S\_c + \omega S\_p, \\
- d I\_u/dt &= \Lambda S\_u - F_I I\_u + \omega I\_n  - \gamma I\_u, \\
- d I\_n/dt &= (1-\theta\_w)\Lambda S\_n + (1-p_I) F_I I\_u - \omega I\_n -\gamma I\_n, \\
- d I\_p/dt &= (1-\theta\_w)\Lambda S\_p + p_I F_I I\_u - \omega I\_p -\gamma I\_p, \\
+ d I\_u/dt &= \Lambda S\_u - \testing{I} I\_u + \omega I\_n  - \gamma I\_u, \\
+ d I\_n/dt &= (1-\theta\_w)\Lambda S\_n + (1-p_I) \testing{I} I\_u - \omega I\_n -\gamma I\_n, \\
+ d I\_p/dt &= (1-\theta\_w)\Lambda S\_p + p_I \testing{I} I\_u - \omega I\_p -\gamma I\_p, \\
  d I\_c/dt &= (1-\theta\_c)\Lambda S\_c + \omega I\_p - \gamma I\_c, \\
- d R\_u/dt &= \gamma I\_u - F_R R\_u + \omega R\_n, \\
- d R\_n/dt &= \gamma I\_n + (1-p_R) F_R R\_u - \omega R\_n, \\
- d R\_p/dt &= \gamma I\_p + p_R F_R R\_u  - \omega R\_p, \\
+ d R\_u/dt &= \gamma I\_u - \testing{R} R\_u + \omega R\_n, \\
+ d R\_n/dt &= \gamma I\_n + (1-p_R) \testing{R} R\_u - \omega R\_n, \\
+ d R\_p/dt &= \gamma I\_p + p_R \testing{R} R\_u  - \omega R\_p, \\
  d R\_c/dt&= \gamma I\_c + \omega R\_p, \\
@@
  \left[ \begin {array}{cccc}  
-\hat F_I+\gamma&-\omega&0&0\\
--(1-p_I)\hat F_I&\omega+\gamma&0&0\\
--p_I \hat F_I&0&\omega+\gamma&0\\
+\testinghat{I}+\gamma&-\omega&0&0\\
+-(1-p_I)\testinghat{I}&\omega+\gamma&0&0\\
+-p_I \testinghat{I}&0&\omega+\gamma&0\\
 0&0&-\omega&\gamma
@@
 \gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)&0&0\\ \noalign{\medskip}
-\gamma(\omega+\gamma)(1-p_I) \hat F_I & \gamma(\omega+\gamma)(\hat F_I+\gamma)&0&0 \\ \noalign{\medskip}
-\gamma(\omega+\gamma) p_I \hat F_I & \gamma\omega p_I \hat F_I & C \gamma/(\omega+\gamma) & 0 \\ \noalign{\medskip}
-\omega(\omega+\gamma) p_I \hat F_I & \omega^2 p_I \hat F_I & C \omega/(\omega+\gamma) & C
+\gamma(\omega+\gamma)(1-p_I) \testinghat{I} & \gamma(\omega+\gamma)(\testinghat{I}+\gamma)&0&0 \\ \noalign{\medskip}
+\gamma(\omega+\gamma) p_I \testinghat{I} & \gamma\omega p_I \testinghat{I} & C \gamma/(\omega+\gamma) & 0 \\ \noalign{\medskip}
+\omega(\omega+\gamma) p_I \testinghat{I} & \omega^2 p_I \testinghat{I} & C \omega/(\omega+\gamma) & C
 \end {array} \right],
 \end{align}
-where $C= \big( \gamma(\omega+\gamma)+(\gamma+\omega p_I)\hat F_I \big) (\omega+\gamma)$ and $\hat F_I$ is the \percap testing rate for the infected people and represented in Eq.~\eqref{eq:fi}.
+where $C= \big( \gamma(\omega+\gamma)+(\gamma+\omega p_I)\testinghat{I} \big) (\omega+\gamma)$ and $\testinghat{I}$ is the \percap testing rate for the infected people and represented in Eq.~\eqref{eq:fi}.
 Note that all the columns of matrix $V^{-1}$ summ up to $1/\gamma$.
@@
 \gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)\\ \noalign{\medskip}
-\gamma(\omega+\gamma)(1-p_I)\hat F_I & \gamma(\omega+\gamma)(\hat F_I+\gamma) \\ \noalign{\medskip}
-\gamma(\omega+\gamma) p_I \hat F_I & \gamma\omega p_I \hat F_I \\ \noalign{\medskip}
-\omega(\omega+\gamma) p_I \hat F_I & \omega^2 p_I \hat F_I
+\gamma(\omega+\gamma)(1-p_I)\testinghat{I} & \gamma(\omega+\gamma)(\testinghat{I}+\gamma) \\ \noalign{\medskip}
+\gamma(\omega+\gamma) p_I \testinghat{I} & \gamma\omega p_I \testinghat{I} \\ \noalign{\medskip}
+\omega(\omega+\gamma) p_I \testinghat{I} & \omega^2 p_I \testinghat{I}
 \end {array} \right].
@@
 where $C$ is the same as the one in Eq. \eqref{eq:C12}, i.e., 
-$$C=(\omega+\gamma)(\gamma(\omega+\gamma)+(\omega p_I+\gamma) \hat F_I),$$
+$$C=(\omega+\gamma)(\gamma(\omega+\gamma)+(\omega p_I+\gamma) \testinghat{I}),$$
 and $C_1$ and $C_2$ are 
@@
 % \label{eq:C12}
-C_1 =& (\omega+\gamma)(\theta\_w   \gamma+\theta\_c  \omega  p_I) \hat F_I,\\
-C_2 =& \big( \omega\gamma(1+p_I)\hat F_I+\gamma^2 (\omega+\gamma+\hat F_I)\big)\theta\_w + \omega^2 p_I \hat F_I \theta\_c,
+C_1 =& (\omega+\gamma)(\theta\_w   \gamma+\theta\_c  \omega  p_I) \testinghat{I},\\
+C_2 =& \big( \omega\gamma(1+p_I)\testinghat{I}+\gamma^2 (\omega+\gamma+\testinghat{I})\big)\theta\_w + \omega^2 p_I \testinghat{I} \theta\_c,
 \end{align*}
-where $\hat F_I$ is given in Eq.~\eqref{eq:fi}.
+where $\testinghat{I}$ is given in Eq.~\eqref{eq:fi}.
 Note that for analysis brevity, we let $N_0=1$, thus $S\_u^*$ and $S\_n^*$ are in the scale of 0 to 1.
@@
 
-{\bf The first goal}
+The first goal
 is to explore how changes in isolation, $\theta\_w$ and $\theta\_c$, affects $\Rnum$.
@@
 \Delta= \frac{1}{C}\Big(
--D_1 S\_n^* \boldsymbol {\theta\_w^2}+
-\big(-\omega^2 p_I \hat F_I S\_n^*\theta\_c+ D_2 S\_n^*+\gamma \hat F_I(\omega+\gamma) \big)\theta\_w+
-(\omega+\gamma S\_u^*)\omega p_I \hat F_I \theta\_c
+-D_1 S\_n^* \theta\_w^2 +
+\big(-\omega^2 p_I \testinghat{I} S\_n^*\theta\_c+ D_2 S\_n^*+\gamma \testinghat{I}(\omega+\gamma) \big)\theta\_w+
+(\omega+\gamma S\_u^*)\omega p_I \testinghat{I} \theta\_c
 \Big),
@@
 \label{eq:del4_d1d2}
-D_1=& (\omega+\gamma)\gamma^2+(\omega+\gamma+\omega p_I)\gamma \hat F_I, \\
-D_2=& (3\omega+2\gamma)\gamma^2+(\omega+\gamma+2\omega p_I)\gamma \hat F_I+(\gamma+p_I \hat F_I)\omega^2.
+D_1=& (\omega+\gamma)\gamma^2+(\omega+\gamma+\omega p_I)\gamma \testinghat{I}, \\
+D_2=& (3\omega+2\gamma)\gamma^2+(\omega+\gamma+2\omega p_I)\gamma \testinghat{I}+(\gamma+p_I \testinghat{I})\omega^2.
 \end{align}
@@
 \label{eq:d_del4_thetac}
-\pder\Delta{\theta\_c}=1/C(\gamma S\_u^*+\omega(1-\theta\_w S\_n^*))\omega p_I \hat F_I.
+\pder\Delta{\theta\_c}=1/C(\gamma S\_u^*+\omega(1-\theta\_w S\_n^*))\omega p_I \testinghat{I}.
 \end{equation}
@@
 1/C \Big( -D_1 S\_n^*\theta\_w^2+
-\big(-\omega^2 p_I \hat F_I S\_n^*\theta\_c+D_2 S\_n^*+\gamma \hat F_I(\omega+\gamma) \big)\theta\_w\Big).
+\big(-\omega^2 p_I \testinghat{I} S\_n^*\theta\_c+D_2 S\_n^*+\gamma \testinghat{I}(\omega+\gamma) \big)\theta\_w\Big).
 \end{equation}
@@
 \pder\Delta{\theta\_w}\bigg\rvert_{\theta\_w=1} =&
-1/C\Big( (D_2-2D_1-\omega^2 p_I \hat F_I \theta\_c) S\_n^*+\gamma \hat F_I(\omega+\gamma) \Big) \\\notag
-=& 1/C\Big( (\gamma(\omega+\gamma)+\gamma\omega^2+(1-\theta\_c)\omega^2 p_I \hat F_I ) S\_n^*
-+\gamma(\omega+\gamma) \hat F_I (1-S\_n^*) \Big),
+1/C\Big( (D_2-2D_1-\omega^2 p_I \testinghat{I} \theta\_c) S\_n^*+\gamma \testinghat{I}(\omega+\gamma) \Big) \\\notag
+=& 1/C\Big( (\gamma(\omega+\gamma)+\gamma\omega^2+(1-\theta\_c)\omega^2 p_I \testinghat{I} ) S\_n^*
++\gamma(\omega+\gamma) \testinghat{I} (1-S\_n^*) \Big),
 \end{align}
@@
 We use the derived expressions for $S\_u^*$ and $S\_n^*$, given by Eqs.~\eqref{dfe}, in $\Delta$ \eqref{eq:del4}.
-Also, we define $\phi = \hat F_S = \frac{\rho \omega}{\omega-\rho}$, to reparameterize $\rho$.
-This is mainly to avoid singularity in $\hat F_I$ \eqref{eq:fi}, when testing intensity $\rho$ is very close to the rate of test return $\omega$.
+Also, we define $\phi = \testinghat{S} = \frac{\rho \omega}{\omega-\rho}$, to reparameterize $\rho$.
+This is mainly to avoid singularity in $\testinghat{I}$ \eqref{eq:fi}, when testing intensity $\rho$ is very close to the rate of test return $\omega$.
 Thus, $\rho$ is reparameterized as
@@
 \subsection{Rate of returning tests} \label{app:omega}
-{\bf The third goal} is to explore how changes in the rate of test return $\omega$ affects $\Rnum$.
+The third goal is to explore how changes in the rate of test return $\omega$ affects $\Rnum$.
 Mathematically we would like to verify the sign of $\pder\Rnum{\omega}$, which specifically depends on $\pder\Delta{\omega}$.
@@
 
-Having presented the formal analysis, we now concern ourselves with its \textbf{biological interpretations.} We begin by interpreting \ref{case_1_ineq}, under which returning tests more rapidly is \emph{always} harmful.
+Having presented the formal analysis, we now concern ourselves with its biological interpretations.
+We begin by interpreting \ref{case_1_ineq}, under which returning tests more rapidly is \emph{always} harmful.
 Notice that the ratio $\frac{\theta\_c}{\theta\_w}$ is simply a measure of how much more strongly individuals self-isolate when they test positive compared to when waiting for tests.
@@
 
-In this work, we didn't do any numerical solutions for the trajectories in our analysis. However, if one tries to do so there would be a singularity issue to deal with. 
-Specifically, the numerical singularity issue with the chosen $\sigma$ \eqref{sigma} is that the population in $S$ compartments appeared to blow up when the DFE is achieved.
-This is once the only untested people are susceptibles, the FOI will become $\Lambda=0$, testing rate $F_s=\rho N_0/S\_u$.
+In this work, we do not present any numerical solutions of the ODEs to investigate typical trajectories.
+However, attempting to do so quickly reveals a problem in the way the model is posed;
+with the scaling of testing ($\sigma$) defined as in the body of the paper (Eq.~\eqref{sigma}), the population in the $S$ compartments appeared to blow up when the DFE is achieved \bmb{clarify??}.
+This occurs because once the only untested people are susceptibles, the FOI approaches $\Lambda=0$, and the testing rate $\testing{S} \to \rho N_0/S\_u$.
 Thus, the first equation of the model \eqref{model} will become
-$d S\_u/dt = - \rho N_0 + \omega S\_n$. Thus changes in $S\_u$ will be no longer dependent on $S\_u$ with a linear rate of leaving the $S\_u$ compartment.
-IN fact the testing rate, $\sigma$, should be formulated such that people from the untested compartments will not be tested if they are not there.
+$d S\_u/dt = - \rho N_0 + \omega S\_n$. Thus changes in $S\_u$ will be independent on $S\_u$, and the decay of the $S\_u$
+population becomes linear rather than exponential --- allowing $S\_u$ to become negative!
+To avoid this problem the testing rate, $\sigma$, should be formulated such that people from the untested compartments will not be tested if they are not there.
 One way to fix this issue, is to consider a maximum testing rate, $\tau$ (1/day).
@@
 This won't always be possible, so we impose a maximum rate of $\tau$ per testable person and redefine $\sigma = \frac{\tau \rho N_0}{\tau W + \rho N_0}$, with the assumption that $\tau \gg \rho$.
-This alteration in $\sigma$, does not change any results related to $\Rnum$, thus we only impose it in the simulation of the epidemic dynamic.
+This modification of $\sigma$ does not affect any of the results we have derived about the invasion of the epidemic from the DFE (i.e., results on $\Rnum$ and $\Delta$).
 

### 2021-06-24 SIR_manuscript.tex
minor edits

@@
 \end{equation}
-where the term $\beta/\gamma$ is the classical basic reproduction number for a simple SIR model \citep{keeling2011modeling}, and $0 < \Delta < 1$ is the proportional reduction \david{actually $1-\Delta$ is the proportional reduction; you should refer to this as the effectiveness of control, as you do below} due to testing and isolation processes, calculated as
+where the term $\beta/\gamma$ is the classical basic reproduction number for a simple SIR model \citep{keeling2011modeling}, $0 < \Delta < 1$, \ali{and $1-\Delta$ is the proportional reduction due to testing and isolation processes.
+We refer to $\Delta$ as the effectiveness of control hereafter, which is calculated as}
 \begin{equation}
@@
 {\bf Effectiveness of testing and isolation in reducing $\Rnum$ at low \percap testing intensity ($\rho$).}
-Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4_theta}\david{wrong eq ref}\bmb{maybe fixed?
-check}), over a range of testing and isolation parameters.
-Parameter values (Table~\ref{tab:params}):
+Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4a}\ali{fixed}), over a range of testing and isolation parameters. Parameter values (Table~\ref{tab:params}):
 $\beta=\betaparam/$day, $1/\gamma= \invgammaparam$ days (baseline $\Rnum=\Rnumparam$, $r=\rparam$); $\omega \in [1/12,2]/$day;  $\rho \in [0,0.013]/$day per capita; $\theta\_w$ and $\theta\_c$ vary between 0 (no effect of isolation) and 1 (complete elimination of transmission); $p_S=0$, $p_I=1$ and $p_R=0.5$.
@@
   {\bf Effectiveness of testing and isolation in reducing $\Rnum$ at high \percap testing intensity.}
-  Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4}\david{again wrong eq ref}), over a range of testing and isolation parameters.
+  Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4a}\ali{fixed}), over a range of testing and isolation parameters.
 Parameters as in \fref{pan} except: $\omega \in [1/5,2]/$day, $\rho \in [0,1/5)/$day.
@@
 }
-\ali{In respense to Ben's comment}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \todo{Fady:To be change accordingly please.} 
-% Perhaps counter-intuitively, the equation above does not predict that $\Rnum$ is monotone decreasing with respect to $\omega$.
-In other words; our model does not predict that returning test results more rapidly \textit{always} lower $\Rnum$.
-In order to gain insight into this intriguing behavior, we examine the zeroes of $\frac{\partial{\Rnum}}{\partial{\omega}}(\omega)$.
-% Defining the following quantity, $Q$, will help us write the roots of $\partial{\Rnum}/\partial{\omega}$ neatly as follows.
-% \begin{align}\label{eq:defQ}
-%     Q =& \frac{w_I}{w_S}\left(1-\frac{n_t-1}{n_w-1}p_I \right).
-% \end{align}
-% With that in mind, we can write the roots of $\partial{\Rnum}/\partial{\omega}$ as
-% \begin{align}
-%     \omega_1 =& \frac{\gamma}{-\sqrt{1-Q}-1} \\
-%     \omega_2 =& \frac{\gamma}{\sqrt{1-Q}-1}.
-% \end{align}
-% 
-% Note that the zeroes are real if and only if $Q < 1$.
-Note that have $\theta\_c > \theta\_w$, so if $p_I \approx 1$, we will have $Q < 0 < 1$.
-Thus, if we assume near-perfect test sensitivity, $\omega_1$ and $\omega_2$ will be real.
-% 
-% Assuming $\omega_1, \omega_2$ are real, it is easy to confirm that $\omega_1 < 0$ by looking at the denominator.
-To see that $\omega_2 > 0$, recall that $Q < 0$, so $\sqrt{1-Q} > 1$ and so $\sqrt{1-Q} -1 > 0$.
-Knowing that $\omega_1 < 0$, the only root of interest (i.e., biologically relevant quantity) is $\omega_2$.
-% 
-% We can prove that $\partial{\Rnum}/\partial{\omega} > 0$ when $\omega \in (0,\omega_2)$ and $\partial{\Rnum}/\partial{\omega} < 0$ when $\omega \in (\omega_2,\infty)$ by computing the limits of $\partial{\Rnum}/\partial{\omega}$ at $0$ and $\infty$ respectively.
-So it follows that $\Rnum$ has a global maximum with respect to $\omega$ at $\omega = \omega_2$.
-% 
-% Now we want to characterize the parameter regions on which $\partial{\Rnum}/\partial{\omega} < 0$ (i.e., the conditions under which returning test results more rapidly is favorable).
-By the previous analysis, this is equivalent to solving for $\omega > \omega_2$.
-So
-% \begin{align}
-%     &\omega > \omega_2 \nonumber \\
-%     &\omega > \frac{\gamma}{\sqrt{1-Q}-1} \nonumber \\
-%     &\sqrt{1-Q} > \frac{\gamma}{\omega}+1 \\
-%     &1-Q > (\frac{\gamma}{\omega}+1)^2.
-% \end{align}
-% Substituting in $Q$ from \eqref{eq:defQ} we have
-% \begin{align}
-%     &1-\frac{w_I}{w_S}\left(1-\frac{n_{t}-1}{n_{\omega}-1}P_{i}\right)> \left(\frac{\gamma}{\omega}+1\right)^2 \\
-%     &-\frac{w_I}{w_S}\left(1-\frac{n_{t}-1}{n_{\omega}-1}P_{i}\right)> \left(\frac{\gamma}{\omega}+1\right)^2+1 \\    
-%     &\frac{w_I}{w_S}\left(\frac{1-n_{t}}{1-n_{\omega}}P_{i}-1\right)> \left(\frac{\gamma}{\omega}+1\right)^2+1\label{eq:necsuf}.
-% \end{align}
-% Since all steps in deriving \eqref{eq:necsuf} are reversible, \eqref{eq:necsuf} gives a necessary and sufficient condition for $\omega > \omega_2$, which characterizes when returning tests more rapidly would cause a decrease in $\Rnum$.
-% % %%%%%%%
-% \subsection{Expensive vs.\ cheap tests}
-% 
-% The use of tests cheaper than RT-PCR has been proposed as a potential strategy for containing the \covid pandemic.
-While cheaper tests may be less sensitive and reliable than RT-PCR, they allow for broader and more intense testing.
-In the analysis below, we compare the $\Rnum$ predicted by our model depending on the testing strategy.
-% 
-% Consider a test that allows us to test at rate $\rho_1$ and has sensitivity $P_{i,1}$, and another test that allows us to test at $\rho_2$ and has sensitivity $P_{i,2}$.
-Suppose that $\rho_1 > \rho_2$.
-Recall that the linearization of $\Rnum$ around $\rho \approx 0$ is given by $$\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(-\theta\_w)(\gamma w_S+\omega w_I) + (-\theta\_c)p_I w_I \omega^2 \Big).$$
-% 
-% 
-% Treating $\Rnum$ as a function of $\rho$ and $P_i$,we can reduce the inequality $$\Rnum(\rho_2, p_{I,2}) < \Rnum(\rho_1, p_{I,1})$$ into 
-% 
-% \begin{align}\label{eq:rho1vsrho2}
-%     &\rho_1\left(\gamma(-\theta\_w)(\gamma w_S + \omega w_I) + (-\theta\_c)p_{I,1}w_I\omega^2\right) - \rho_2\left(\gamma(-\theta\_w)(\gamma w_S + \omega w_I) + (-\theta\_c)p_{I, 2}w_I\omega^2\right) > 0 \nonumber \\
-%     &\vdots \nonumber \\
-%     &\frac{\rho_2P_{i, 2}-\rho_1P_{i, 1} }{\rho_1-\rho_2} > \frac{\theta\_w}{\theta\_c}\cdot \frac{\gamma(\gamma w_S + \omega w_I)}{\omega^2 w_I}
-% \end{align}
-% 
-% Note that the RHS is positive, thus a necessary condition for the inequality above to hold is that $\rho_2P_{i,2} > \rho_1P_{i,1}$, equivalently 
-% 
-% \begin{equation}
-% \frac{P_{i,2}}{P_{i,1}} > \frac{\rho_1}{\rho_2}.
-% \end{equation}
-% 
-% To state an example of this, if test $A$ is three times as expensive as test $B$ (and hence one can test three times as many people with test $B$), using test $A$ rather than $B$ will be favorable only if test $A$ is at least 3 times more sensitive than test $B$.
-Note that this is a necessary but not sufficient condition, so even if test $A$ is three times more sensitive, it is still possible for test $B$ to be more effective.
-% 
-% \cref{eq:rho1vsrho2} tells us precisely when a test corresponding to $\rho_2, P_{i,2}$ will yield a lower $\Rnum$ than a test corresponding to $\rho_1, P_{i,1}$, where $\rho_1 > \rho_2$.
-Some of the qualitative \textit{trends} that favor test 2 (the higher-sensitivity test) include
-% 
-% \begin{itemize}
-%     \item individuals who test positive self-isolate much more than individuals who are waiting for their test result.
-%     \item the time it takes to return tests is much shorter than the mean infectious period.
-%     \item the testing intensity is much greater for infected individuals than susceptible individuals.
-% \end{itemize}
-
 \end{document}

### 2021-06-24 SIR_manuscript.tex
cleaning up comments and tightening abstract (with Ali)

@@
 
-\bmb{abstract needs another editing pass, skipping for now}
-\david{This needs substantial tightening.}
-
-The effect of testing processes, including testing and test reporting, on epidemic dynamics, involving infection and recovery, can be studied at the individual level or the community level (e.g., nursing homes, long-term-care facilities, etc.).
-Gaining insights to determine the sensitivity of the epidemic dynamics with respect to the testing processes will depend on underlying factors including the level of focus (individual or community), assumptions (model), and the interplay between these factors.
-In particular, fast testing and test reporting may be beneficial at the community-level, supported by many studies, as it gives a rapid assessment of the situation, identifies hot spots, and may enable rapid contact-tracing.
-However, the potential advantage of a slow rate of test return on the dynamics of an epidemic is real, often neglected, and needs to be quantified.
-At the individual level, this advantage can manifest in the following sense: individuals awaiting their test results or who have tested positive may partially or fully self-isolate, thus reducing or eliminating their transmission potential.
-In this paper, we investigated this individual-level effect of testing processes on the epidemic dynamics by developing a SIR-type model.
-Although the model development was motivated by the \covid epidemic, the model has general epidemiological and testing structures.
-The realistic components of the model include \percap testing intensity, test sensitivity and specificity, rate of test return, and isolation efficacy (i.e., reduction of the transmission probability by isolating individuals).
-The novel component is the compartment-specific relative testing weights, which reflect the testing strategies --- surveillance, diagnosis, or control.
-Here, we compare two testing strategies, random vs.
-targeted, and concluded that increasing testing “focus” from random to targeted always decreases \Rnum.
-Further, the following processes always decreases $\Rnum$; increasing the isolation efficacy parameters for tested and confirmed individuals, higher testing intensity if testing is random or testing intensity is small, and a higher rate of test return if the isolation efficacy for tested-awaiting individuals is low.
+Testing individuals for pathogens can affect the spread of epidemics. 
+Understanding how individual-level processes of sampling and reporting test results can effect comunity- or population-level spread is a dynamical modeling question.
+The effect of testing processes on epidemic dynamics depends on factors underlying implementation, particularly testing intensity and how testing is focused.
+Here, we use a simple model to ask how the individual-level effects of testing might directly impact population-level spread.
+We construct a simple framework and derive an analytic expression for the relative reduction in the basic reproductive number due to testing, test-reporting and related isolation behaviours.
+Intensive testing and fast test reporting are expected to be beneficial at the community level because they can provide a rapid assessment of the situation, identify hot spots, and may enable rapid contact-tracing.
+The direct effects of fast testing at the individual level are less clear, and may depend on how individuals' behaviour is affected by testing information.
+Our simple model shows that under some circumstances both increased testing intensity and faster test reporting can \emph{reduce} the effectiveness of control, and allows us to explore the conditions under which this occurs.
+Conversely, we find that increasing testing \emph{focus} always acts to increase effectiveness of control.
+Although our model development was motivated by the \covid epidemic, the model has general epidemiological and testing structures. 
+To the classic framework we have added a \percap testing intensity,  and compartment-specific testing weights, which can be adjusted to reflect different testing emphases --- surveillance, diagnosis, or control.
 
@@
 The condition for positive exponential growth ($r>0$) is commonly expressed as $\Rnum>1$, where the basic reproduction number $\Rnum$ is the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}.
-Although the value of $\Rnum$ cannot completely characterize the dynamics of even the simplest epidemic model \citep{shaw2021what}, \david{well, technically it does up to rescaling time} it does give a simple and widely accepted index for the difficulty of control, as well as an indication of the likely final size of an epidemic \citep{ma2006generality}.
+Although the value of $\Rnum$ cannot completely characterize the dynamics of our model \citep{shaw2021what}, it does give a simple and widely accepted index for the difficulty of control, as well as an indication of the likely final size of an epidemic \citep{ma2006generality}.
 
@@
 \label{sigma}
-\sigma = \frac{\rho N_0}{W},
+\sigma = \frac{\rho N}{W},
 \end{equation}
-where $\rho$ is the \percap testing intensity for the population, defined as the number of daily tests administered in a population of size $N_0$.
-\david{Why $N_0$ rather than just $N$?}
+where $\rho$ is the \percap testing intensity for the population, defined as the number of daily tests administered in a population of size $N$.
 Thus, the \percap testing rate for compartment $X \in \{S,I,R\}$ is 
@@
 \end{equation}
-\david{$F$ isn't a natural symbol for ``testing rate''.  How about $\mathcal T_X$?}
-\bmb{mostly fixed, but needs to be fixed in the flow chart as well}
 For a highly sensitive test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus not considered for further testing.
@@
   \label{Lambda}
-  \Lambda=\frac{\beta}{N_0} \big(I\_u + (1-\theta\_w)(I\_n+I\_p) + (1-\theta\_c)I\_c \big),
+  \Lambda=\frac{\beta}{N} \big(I\_u + (1-\theta\_w)(I\_n+I\_p) + (1-\theta\_c)I\_c \big),
   \end{equation}
@@
   Symbol & Description & Unit & Value \\ \hline
-  $N_0$     & Total population size & people & $10^6$ \\ \hline
+  $N$     & Total population size & people & $10^6$ \\ \hline
   $\omega$  & Rate of test return, i.e., rate of onward flow from ``waiting'' to ``confirmed'' or ``untested'' compartments  & 1/day & - \\ \hline
@@
 \label{dfe}
-S\_n^*= \frac{\rho}{\omega} N_0, \ S\_u^*= N_0-S_n^*, \text{~and~} I_j=R_j=0 \ \text{for all }j.
+S\_n^*= \frac{\rho}{\omega} N, \ S\_u^*= N-S_n^*, \text{~and~} I_j=R_j=0 \ \text{for all }j.
 \end{equation}
@@
 
-The basic reproduction number, $\Rnum$, was calculated by using the next-generation matrix method \citep{van2002reproduction}. We present $\Rnum$ as
+The basic reproduction number, $\Rnum$, was calculated by using the next-generation matrix method \citep{van2002reproduction}. We write $\Rnum$ as
 \begin{equation}
@@
 \end{equation}
-where the term $\beta/\gamma$ is the classical basic reproduction number for a simple SIR model \citep{keeling2011modeling}, $0 < \Delta < 1$, \ali{and $1-\Delta$ is the proportional reduction due to testing and isolation processes.
-We refer to $\Delta$ as the effectiveness of control hereafter, which is calculated as}
+where $\beta/\gamma$ is the classical value for a simple model \citep{keeling2011modeling}, and $1-\Delta$ is the proportional reduction due to testing and isolation processes.
+$\Delta$ therefore measureness the ``effectiveness of control'': how much these processes in reducing spread from low levels, and is in turn given by:
 \begin{equation}
   \label{eq:del4a}
-  \Delta= \frac{1}{C N_0}\big(C_1 S\_u^*+(C_2(1-\theta\_w)+C \theta\_w)S\_n^*\big),
+  \Delta= \frac{1}{C N}\big(C_1 S\_u^*+(C_2(1-\theta\_w)+C \theta\_w)S\_n^*\big),
 \end{equation}
@@
 \section{Discussion}
+
+\jd{Discuss the possibility that people who can't get tests might also isolate. This undercuts our assumptions about $\theta\_w$.}
 
@@
 \label{FV}
-F =& \beta/N_0 \left[ \begin {array}{cccc} 
+F =& \beta/N \left[ \begin {array}{cccc} 
 S\_u^* & (1-\theta\_w) S\_u^* & (1-\theta\_w) S\_u^* & (1-\theta\_c) S\_u^*\\
@@
  \end {array} \right] \\
- =&\beta/N_0 \left[ \begin {array}{c} S\_u^* \\ (1-\theta\_w)S\_n^* \\ 0 \\ 0 \end {array} \right]
+ =&\beta/N \left[ \begin {array}{c} S\_u^* \\ (1-\theta\_w)S\_n^* \\ 0 \\ 0 \end {array} \right]
         \left[ \begin {array}{cccc} 1,   1-\theta\_w,   1-\theta\_w,   1-\theta\_c \end {array} \right], \text{and}\\  V=&
@@
 G_{11} = \frac{\beta}{\gamma C} 
-\left[ \begin {array}{c}  S\_u^*/N_0 \\ (1-\theta\_w) S\_n^*/N_0  \end {array} \right]
+\left[ \begin {array}{c}  S\_u^*/N \\ (1-\theta\_w) S\_n^*/N  \end {array} \right]
 \left[ \begin {array}{cccc} 
@@
 where $\testinghat{I}$ is given in Eq.~\eqref{eq:fi}.
-Note that for analysis brevity, we let $N_0=1$, thus $S\_u^*$ and $S\_n^*$ are in the scale of 0 to 1.
+Note that for analysis brevity, we let $N=1$, thus $S\_u^*$ and $S\_n^*$ are in the scale of 0 to 1.
 $\Rnum$ is in the same form as in Eq. \eqref{R0}  
@@
 with the scaling of testing ($\sigma$) defined as in the body of the paper (Eq.~\eqref{sigma}), the population in the $S$ compartments appeared to blow up when the DFE is achieved \bmb{clarify??}.
-This occurs because once the only untested people are susceptibles, the FOI approaches $\Lambda=0$, and the testing rate $\testing{S} \to \rho N_0/S\_u$.
+This occurs because once the only untested people are susceptibles, the FOI approaches $\Lambda=0$, and the testing rate $\testing{S} \to \rho N/S\_u$.
 Thus, the first equation of the model \eqref{model} will become
-$d S\_u/dt = - \rho N_0 + \omega S\_n$. Thus changes in $S\_u$ will be independent on $S\_u$, and the decay of the $S\_u$
+$d S\_u/dt = - \rho N + \omega S\_n$. Thus changes in $S\_u$ will be independent on $S\_u$, and the decay of the $S\_u$
 population becomes linear rather than exponential --- allowing $S\_u$ to become negative!
@@
 In general, we want to test at a rate of $\rho$ across the whole population.
-This won't always be possible, so we impose a maximum rate of $\tau$ per testable person and redefine $\sigma = \frac{\tau \rho N_0}{\tau W + \rho N_0}$, with the assumption that $\tau \gg \rho$.
+This won't always be possible, so we impose a maximum rate of $\tau$ per testable person and redefine $\sigma = \frac{\tau \rho N}{\tau W + \rho N}$, with the assumption that $\tau \gg \rho$.
 This modification of $\sigma$ does not affect any of the results we have derived about the invasion of the epidemic from the DFE (i.e., results on $\Rnum$ and $\Delta$).

### 2021-06-24 SIR_manuscript.tex
end of meeting

@@
 
-Although testing and tracing is a key part of infection control strategy, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts, rather than analyzing simple but general models of the feedback between testing and epidemic dynamics \bmb{can we back this statement up?
-more general citations to test-and-trace?
-HIV?
-Bergstrom et al might be a counterexample, although it's for a different context [repeated testing of an isolated population]}.
-\ali{the following sentences added to address Ben' comment:} There have been several modeling works of testing and tracing dynamics and their interaction with epidemiological dynamics.
+Although testing and tracing is a key part of infection control strategy, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts, rather than analyzing simple but general models of the feedback between testing and epidemic dynamics
+\ali{the following sentences added to address Ben's comment:} There have been several modeling works of testing and tracing dynamics and their interaction with epidemiological dynamics.
 In the context of repeated screening and random testing of isolated populations (such as a university setup), \cite{bergstrom2020frequency} provided some analytical results, \cite{rogers2021high} simulated a SEIR model with high-low testing intensity and sensitivity.
@@
 We hope that this paper will inspire further explorations of the fundamental properties of dynamical systems that incorporate explicit testing models in an epidemiological context.
-\bmb{we could also talk about test and trace as a 'speed' intervention, but I'm tired}.
-
-\bmb{Leftover comments:
-\begin{itemize}
-\item it would be good to have a few more testing refs: Bergstrom et al., there might be some other stuff we can recycle from the lab meeting review; Friston's stuff deserves to be mentioned because it does have explicit testing flows too.
-We should try not to ignore/piss anyone off who has already done stuff
-\end{itemize}
-}
 

### 2021-06-24 SIR_manuscript.tex
flowchart was editted

@@
 We computed the values and contours of $\Delta$ at both low (\fref{pan}) and high (\fref{pan2}) testing intensities, and for both random testing ($w_S=w_I=w_R=1$) and targeted testing ($w_S=0.3$; $w_I=w_R=1$).
-Because it is expressed as a proportion of $\Rnum$, the effectiveness of control $\Delta$ is (at least in the $\rho \ll 1$ case, Eq.~\ref{lin}) independent of the transmission rate $\beta$, and hence of $\Rnum$ in the case where we vary $\Rnum$ by changing the transmission rate for a fixed generation interval.
+Because it is expressed as a proportion of $\Rnum$, the effectiveness of control $\Delta$ is (at least in the $\rho \ll 1$ case, Eq.~\ref{eq:lin}) independent of the transmission rate $\beta$, and hence of $\Rnum$ in the case where we vary $\Rnum$ by changing the transmission rate for a fixed generation interval.
 

### 2021-06-25 SIR_manuscript.tex
David's comments on Discussion added

@@
 
-Over a broad range of parameter space (for random testing, approximately $\theta\_w \geq 0.25$; for targeted testing, $\theta\_w \geq 0.25$ and either $\theta\_c \ge 0.5$ or $1/\omega > 5$), decreasing $\omega$ --- i.e., decreasing the rate at which tested people find out about their infection status --- decreases $\Rnum$.
-This result is counterintuitive; public health agencies have invested a lot of effort in increasing this rate.
+Over a broad range of parameter space (for random testing, $\theta\_w \gtrsim 0.25$; for targeted testing, $\theta\_w \geq 0.25$ and either $\theta\_c \ge 0.5$ or $1/\omega > 5$), decreasing $\omega$ --- i.e., slowing the rate at which test results are returned --- decreases $\Rnum$.
+This result is counterintuitive; public health authorities have invested a lot of effort in reducing delays from testing to results.
 Dynamically, this effect occurs because speeding up test returns shortens the isolation period of uninfected individuals.
@@
 
-While slowing test returns does decrease $\Rnum$ over a broad range of parameter space in our model, there are several real-world processes missing from our model that make it implausible that slowing test returns would actually be an effective public health measure.
-First, our model is missing the components that would allow us to model the primary benefit of rapid testing, i.e.
-detecting and containing outbreaks while they are still in progress.
-This process could be modeled phenomenologically by making the testing focus increase as an increasing proportion of cases are detected, because finding infections allows tests to be concentrated on their connections.
+While slowing test returns does decrease $\Rnum$ over broad regions of parameter space in our model, there are several real-world processes missing from our model that make it implausible that slowing test returns would actually be an effective public health measure.
+First, our model is missing the components that would allow us to model the primary benefit of rapid testing, i.e., detecting and containing outbreaks while they are still in progress.
+This process could be modeled phenomenologically by making the testing focus more targeted as an increasing proportion of cases is detected, because finding infections allows tests to be concentrated on their connections.
 Second, in the real world individuals become less likely to maintain isolation if they are required to do so for longer; phenomenologically, we could allow  effectiveness of isolation while waiting for test returns to be an increasing function of test return speed, or we could introduce a separate ``waiting, but no longer isolating'' compartment that individuals entered from the ``waiting, isolated'' compartment at a specified rate.
@@
 
-The other counterintuitive result from our analysis is that, for sufficiently high testing intensity $\rho$, increasing testing intensity can actually increase $\Rnum$ (e.g.
+The other counterintuitive result from our analysis is that, for sufficiently high testing intensity $\rho$, further increasing testing intensity can actually \emph{increase} $\Rnum$ (e.g.
 \fref{pan2}(b), upper right panel [$\theta\_c=1$, $\theta\_w=0$]).
-This phenomenon occurs mainly because of a peculiar feature of the DFE in our model: we assume that testing is ongoing before the beginning of the epidemic, so that the starting point of the epidemic invasion is an equilibrium distribution of susceptibles between the $S\_n$ (untested) and $S\_u$ (waiting) compartments: in particular, $S\_u^* = \rho/(\rho + \omega)$.
+This phenomenon occurs mainly because of a peculiar feature of the DFE in our model: we implicitly assume that testing is ongoing before the beginning of the epidemic, so that the starting point of the epidemic invasion is an equilibrium distribution of susceptibles between the $S\_n$ (untested) and $S\_u$ (waiting) compartments: in particular, $S\_u^* = \rho/(\rho + \omega)$.
 (In reality, testing intensity itself is a dynamic variable that will increase as the epidemic proceeds.) If isolation in the waiting compartment is low ($\theta\_w \ll 1$), then waiting individuals can easily be infected once the epidemic starts.
@@
 Thus, under certain extreme conditions (low $\theta\_w$, high $\theta\_c$, low $\omega$, high $\rho$), $\Rnum$ becomes an increasing function of testing intensity.
-\bmb{This phenomenon occurs primarily (only??) for targeted testing, because \ldots CAN WE FIGURE OUT WHY?
+\bmb{This phenomenon occurs primarily (\ali{only}) for targeted testing, because \ldots CAN WE FIGURE OUT WHY?
 Targeting should be irrelevant to determining the DFE (because everyone is $S$ at that point), so what is the dynamical explanation?
@@
 testing is more effective overall (darker colours in Fig 3(b) upper-right panel than in Fig 3(a) upper panel), so we more quickly reach the point where the effect of increasing $S\_n^*$ outweighs the beneficial effects of testing?} However, this phenomenon is even more unrealistic than the possibility that slowing test returns increases $\Rnum$.
-It depends both on the assumption that regular testing is occurring before the epidemic starts, and on levels of testing that are unrealistically high (at least in large, general-population settings), i.e.
-at least 10\% of the population being tested per day.
+It depends both on the assumption that regular testing is occurring before the epidemic starts, and on levels of testing that are unrealistically high (at least in large, general-population settings), i.e., at least 10\% of the population being tested per day.
 
-Although we model the testing process in more detail than most mathematically focused epidemiological models, one place where more detail could be informative is in the processes determining the testing weights $\{w_S, w_I, w_R\}$.
+Although we model the testing process in more detail than typical epidemiological models, one place where more detail could be informative is in the processes determining the testing weights $\{w_S, w_I, w_R\}$.
 While random testing, as done for surveillance purposes, unambiguously leads to equal testing weights, making precise quantitative connections between public health practices and testing weights is difficult in other contexts.
 The testing weights reflect the correlation between an individual's risk of infection and their likelihood of being tested due to age, occupation, geographic location, etc..
-This correlation is influenced, among many other factors, by the proportion of the uninfected population with COVID-like symptoms (e.g.
-due to seasonal upper respiratory tract infections); the concentration of transmission and testing in hot spots such as long-term care facilities and high-density workplaces; the overall testing intensity (and hence e.g.
+This correlation is influenced, among many other factors, by the proportion of the uninfected population with COVID-like symptoms (e.g., due to seasonal upper respiratory tract infections); the concentration of transmission and testing in hot spots such as long-term care facilities and high-density workplaces; the overall testing intensity (and hence e.g.
 restriction to symptomatic individuals); and the proportion of COVID-infected people who are symptomatic.
 Future steps should explore mathematically tractable ways to model some of these factors more precisely.
-For example, separating the infected class into exposed, symptomatic, and a- or pre-symptomatic compartments and allowing the testing weights to vary across non-symptomatic (exposed/asymptomatic/presymptomatic) vs.
-symptomatic compartments could reflect the allocation of tests to diagnostic purposes (targeting symptomatic individuals) vs.
-contact-tracing (targeting infected but non-symptomatic individuals) vs.
-screening (relatively equal weights, depending on the venue).
+For example, separating the infected class into exposed, symptomatic, and a- or pre-symptomatic compartments and allowing the testing weights to vary across non-symptomatic (exposed/asymptomatic/presymptomatic) vs.\ symptomatic compartments could reflect the allocation of tests to diagnostic purposes (targeting symptomatic individuals) vs.\ contact-tracing (targeting infected but non-symptomatic individuals) vs.\ screening (relatively equal weights, depending on the venue).
 Alternatively, one could make the testing weights depend on the testing intensity or test return rate as suggested above.
@@
 
-Although testing and tracing is a key part of infection control strategy, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts, rather than analyzing simple but general models of the feedback between testing and epidemic dynamics
+\david{The intro and discussion would benefit from (potentially many) additional references.}
+Although testing and tracing is a key part of infection control strategies, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts, rather than analyzing simple but general models of the feedback between testing and epidemic dynamics
 \ali{the following sentences added to address Ben's comment:} There have been several modeling works of testing and tracing dynamics and their interaction with epidemiological dynamics.

### 2021-06-25 SIR_manuscript.tex
abstract revisions

@@
 Understanding how individual-level processes of sampling and reporting test results can effect comunity- or population-level spread is a dynamical modeling question.
-The effect of testing processes on epidemic dynamics depends on factors underlying implementation, particularly testing intensity and how testing is focused.
-Here, we use a simple model to ask how the individual-level effects of testing might directly impact population-level spread.
-We construct a simple framework and derive an analytic expression for the relative reduction in the basic reproductive number due to testing, test-reporting and related isolation behaviours.
+The effect of testing processes on epidemic dynamics depends on factors underlying implementation, particularly testing intensity and on whom testing is focused.
+Here, we use a simple model to explore how the individual-level effects of testing might directly impact population-level spread.
+Our model development was motivated by the \covid epidemic, but has generic epidemiological and testing structures. 
+To the classic SIR framework we have added a \percap testing intensity, and compartment-specific testing weights, which can be adjusted to reflect different testing emphases --- surveillance, diagnosis, or control.
+We derive an analytic expression for the relative reduction in the basic reproductive number due to testing, test-reporting and related isolation behaviours.
 Intensive testing and fast test reporting are expected to be beneficial at the community level because they can provide a rapid assessment of the situation, identify hot spots, and may enable rapid contact-tracing.
-The direct effects of fast testing at the individual level are less clear, and may depend on how individuals' behaviour is affected by testing information.
+Direct effects of fast testing at the individual level are less clear, and may depend on how individuals' behaviour is affected by testing information.
 Our simple model shows that under some circumstances both increased testing intensity and faster test reporting can \emph{reduce} the effectiveness of control, and allows us to explore the conditions under which this occurs.
-Conversely, we find that increasing testing \emph{focus} always acts to increase effectiveness of control.
-Although our model development was motivated by the \covid epidemic, the model has general epidemiological and testing structures. 
-To the classic framework we have added a \percap testing intensity,  and compartment-specific testing weights, which can be adjusted to reflect different testing emphases --- surveillance, diagnosis, or control.
+Conversely, we find that focussing testing on infected individuals always acts to increase effectiveness of control.
+\david{It would be better to say ``symptomatic'' rather than ``infected'' in the last sentence, but we can't claim the model really captures that.}
 

### 2021-06-25 SIR_manuscript.tex
adding some refs to intro

@@
 
-When a new infectious disease emerges, it is important to determine whether it will grow exponentially in a susceptible population, and if so at what rate $r$.
+When a new infectious disease emerges, it is important to determine whether it will grow exponentially in a susceptible population, and if so at what rate $r$ \citep{Ma+14}.
 The condition for positive exponential growth ($r>0$) is commonly expressed as $\Rnum>1$, where the basic reproduction number $\Rnum$ is the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}.
-Although the value of $\Rnum$ cannot completely characterize the dynamics of our model \citep{shaw2021what}, it does give a simple and widely accepted index for the difficulty of control, as well as an indication of the likely final size of an epidemic \citep{ma2006generality}.
+Although the value of $\Rnum$ cannot completely characterize the dynamics of our model \citep{shaw2021what}, it does give a simple and widely accepted index for the difficulty of control, as well as an indication of the likely final size of an epidemic \citep{ma2006generality,miller2012note}.
 
-In order to understand the effect of testing processes on epidemic dynamics, we expanded one of the simplest mechanistic epidemic model---the standard deterministic SIR model---to include testing components.
+In order to understand the effect of testing processes on epidemic dynamics, we expanded one of the simplest mechanistic epidemic models---the standard deterministic SIR model\citep{KermMcKe27,AndeMay91}---to include testing components.
 This model provides a sensible platform to link the modeling of epidemic and testing components and study their interaction.

### 2021-06-25 SIR_manuscript.tex
starting on Discussion

@@
 \newcommand{\todo}[1]{\comment{red}{TODO}{#1}}
+\newcommand{\needref}{\textcolor{red}{(REFS?)}}
 
@@
 
-In this paper, we have defined and analyzed a simple compartmental model that combines epidemiological dynamics (as defined by a simple SIR model) with the dynamics of testing and isolation.
-Our model is a caricature --- it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g.
-exposed, presymptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates [REFS?]; behavioural dynamics \citep{weitz2020awareness}).
+In this paper, we have developed and analyzed a simple compartmental model that combines epidemiological dynamics---as defined by a simple SIR model---with the dynamics of testing and isolation.
+Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates \needref; behavioural dynamics \citep{weitz2020awareness}).
 Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between epidemiology and testing, rather than for making quantitative predictions or guiding pandemic responses.

### 2021-06-26 SIR_manuscript.tex
revisions to Discussion and tiny bit to Intro

@@
 The observed dynamics of the \covid epidemic have been driven both by epidemiological processes (infection and recovery) and by testing processes (testing and test reporting).
-In addition to shaping epidemic observations (via case reports), testing processes also alter epidemiological dynamics.
+In addition to shaping epidemic observations (via case reports), testing processes also alter epidemiological dynamics\needref.
 Because individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may also do so, testing will generally increase the number of people who are isolating and hence reduce epidemic growth rates.
@@
 Over the course of the \covid pandemic, however, the vast majority of testing has been done with other goals --
-primarily diagnostic (determining infection status for clinical purposes), or for control (determining  infection status in order to isolate cases that have been found by contact tracing), which we characterize as \emph{targeted} testing strategies.
-In these cases, testing probabilities can differ sharply across epidemiological compartments; in our dynamical model, we will characterize these probabilities by assigning a testing weight to each compartment that determines the \emph{relative} probability that an individual in that compartment will be selected for testing (see Methods).
+primarily diagnostic (determining infection status for clinical purposes)\needref, or for control (determining  infection status in order to isolate cases that have been found by contact tracing)\needref, which we characterize as \emph{targeted} testing strategies.
+In these situations, testing probabilities can differ sharply across epidemiological compartments; in our dynamical model, we will characterize these probabilities by assigning a testing weight to each compartment that determines the \emph{relative} probability that an individual in that compartment will be selected for testing (see Methods).
 
@@
 Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates \needref; behavioural dynamics \citep{weitz2020awareness}).
-Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between epidemiology and testing, rather than for making quantitative predictions or guiding pandemic responses.
+Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
 
 Many of the qualitative results we have derived confirm simple, common-sense intuitions.
-In particular, we can generally decrease $\Rnum$ by increasing isolation efficacy or testing intensity; returning tests faster, if individuals do not isolate while they are waiting for results; or increasing testing focus to target individuals who are likely to be infectious (e.g.
-symptomatic people or close contacts of known infections).
+In particular, we can generally decrease $\Rnum$ by increasing isolation efficacy or testing intensity; returning tests faster, if individuals do not isolate while they are waiting for results; or increasing testing focus to target individuals who are likely to be infectious (e.g., symptomatic people or close contacts of known infections).
 
-However, we did find two surprising patterns: longer delays in returning tests can decrease the spread of the epidemic, and increasing testing rates can increase its spread.
+However, we did find two surprising patterns: longer delays in returning tests can \emph{reduce} epidemic spread, and increasing testing rates can \emph{increase} spread.
 
-Over a broad range of parameter space (for random testing, $\theta\_w \gtrsim 0.25$; for targeted testing, $\theta\_w \geq 0.25$ and either $\theta\_c \ge 0.5$ or $1/\omega > 5$), decreasing $\omega$ --- i.e., slowing the rate at which test results are returned --- decreases $\Rnum$.
-This result is counterintuitive; public health authorities have invested a lot of effort in reducing delays from testing to results.
-Dynamically, this effect occurs because speeding up test returns shortens the isolation period of uninfected individuals.
-For infected people it only shortens the time to progression to the isolation level of the confirmed-positive compartment.
-Slowing test returns increases $\Rnum$ only if the proportion of infectives in the tested population (test positivity), and the magnitude of increase in isolation effectiveness from the waiting to the confirmed-positive compartment, are large enough to outweigh the decrease in isolation of uninfected waiting individuals.
+Over broad regions of parameter space, decreasing $\omega$---i.e., \emph{slowing} the rate at which test results are returned---\emph{decreases} $\Rnum$ (for random testing, the parameter region is $\theta\_w \gtrsim 0.25$; for targeted testing, $\theta\_w \geq 0.25$ and either $\theta\_c \ge 0.5$ or $1/\omega > 5$; see Table~\ref{tab:params} for parameter definitions).
+This result is counterintuitive and would not be expected by public health authorities that have invested a great deal of effort in reducing delays from testing to results.
+Dynamically, this effect occurs because speeding up test returns shortens the isolation period of uninfected individuals (for infected people it only shortens the time to progression to the isolation level of the confirmed-positive compartment).
+Slowing test returns increases $\Rnum$ only if test positivity (the proportion of infectives in the tested population), and the magnitude of increase in isolation effectiveness from the waiting to the confirmed-positive compartment, are large enough to outweigh the decrease in isolation of uninfected waiting individuals.\david{That sentence is too long to process easily; I've re-organized slightly to make flow a bit better, but it is awkward to really fix it.}
 
 While slowing test returns does decrease $\Rnum$ over broad regions of parameter space in our model, there are several real-world processes missing from our model that make it implausible that slowing test returns would actually be an effective public health measure.
-First, our model is missing the components that would allow us to model the primary benefit of rapid testing, i.e., detecting and containing outbreaks while they are still in progress.
+First, we do not model the primary benefit of rapid testing, i.e., detecting and containing outbreaks while they are still in progress.
 This process could be modeled phenomenologically by making the testing focus more targeted as an increasing proportion of cases is detected, because finding infections allows tests to be concentrated on their connections.
 Second, in the real world individuals become less likely to maintain isolation if they are required to do so for longer; phenomenologically, we could allow  effectiveness of isolation while waiting for test returns to be an increasing function of test return speed, or we could introduce a separate ``waiting, but no longer isolating'' compartment that individuals entered from the ``waiting, isolated'' compartment at a specified rate.
-Third and finally, if one wants to decrease the overall transmission rate of the population there are more effective methods than keeping tested people in limbo -- these include masking, ventilation, distancing measures, retail and event closures, or stay-at-home orders.
+Third and finally, if one wants to decrease the overall transmission rate of the population there are more effective methods than keeping tested people in limbo; these include masking, ventilation, distancing measures, retail and event closures, and stay-at-home orders.
 
@@
 The testing weights reflect the correlation between an individual's risk of infection and their likelihood of being tested due to age, occupation, geographic location, etc..
-This correlation is influenced, among many other factors, by the proportion of the uninfected population with COVID-like symptoms (e.g., due to seasonal upper respiratory tract infections); the concentration of transmission and testing in hot spots such as long-term care facilities and high-density workplaces; the overall testing intensity (and hence e.g.
-restriction to symptomatic individuals); and the proportion of COVID-infected people who are symptomatic.
+This correlation is influenced, among many other factors, by the proportion of the uninfected population with COVID-like symptoms (e.g., due to seasonal upper respiratory tract infections); the concentration of transmission and testing in hot spots such as long-term care facilities and high-density workplaces; the overall testing intensity (and hence, e.g., restriction to symptomatic individuals); and the proportion of COVID-infected people who are symptomatic.
 Future steps should explore mathematically tractable ways to model some of these factors more precisely.
-For example, separating the infected class into exposed, symptomatic, and a- or pre-symptomatic compartments and allowing the testing weights to vary across non-symptomatic (exposed/asymptomatic/presymptomatic) vs.\ symptomatic compartments could reflect the allocation of tests to diagnostic purposes (targeting symptomatic individuals) vs.\ contact-tracing (targeting infected but non-symptomatic individuals) vs.\ screening (relatively equal weights, depending on the venue).
+For example, separating the infected class into exposed, symptomatic, and a- or pre-symptomatic compartments and allowing the testing weights to vary across non-symptomatic (exposed/asymptomatic/presymptomatic) vs.\ symptomatic compartments could reflect the allocation of tests for diagnostic purposes (targeting symptomatic individuals) vs.\ contact-tracing (targeting infected but non-symptomatic individuals) vs.\ screening (relatively equal weights, depending on the venue).
 Alternatively, one could make the testing weights depend on the testing intensity or test return rate as suggested above.
@@
 \david{The intro and discussion would benefit from (potentially many) additional references.}
-Although testing and tracing is a key part of infection control strategies, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts, rather than analyzing simple but general models of the feedback between testing and epidemic dynamics
-\ali{the following sentences added to address Ben's comment:} There have been several modeling works of testing and tracing dynamics and their interaction with epidemiological dynamics.
-In the context of repeated screening and random testing of isolated populations (such as a university setup), \cite{bergstrom2020frequency} provided some analytical results, \cite{rogers2021high} simulated a SEIR model with high-low testing intensity and sensitivity.
-Further, \cite{friston2021testing} model the effects of self-isolation on testing and tracking.
-Our modeling approach is novel compared to these works.
-We hope that this paper will inspire further explorations of the fundamental properties of dynamical systems that incorporate explicit testing models in an epidemiological context.
+Although testing and tracing is a key part of infection control strategies, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts\needref, rather than analyzing simple but general models of the feedback between testing and transmission dynamics.
+\ali{the following sentences added to address Ben's comment:} There have been several modeling studies of testing and tracing dynamics and their interaction with epidemiological dynamics.
+In the context of repeated screening and random testing of isolated populations (such as a university), \cite{bergstrom2020frequency} provided some analytical results\david{what are the results?
+that's what is most important, not whether the results were analytical}, \cite{rogers2021high} simulated a SEIR model with high-low \david{meaning?} testing intensity and sensitivity\david{what did they find?}.
+Further, \cite{friston2021testing} model the effects of self-isolation on testing and tracing\david{what did they find?}.
+Our modeling approach is novel compared to these works\david{how so?
+You can't just say our work is novel.
+What makes it novel relative to these previous papers?}.
+We hope that the present paper will inspire further explorations of the fundamental properties of epidemic models that incorporate explicit testing processes.
 

### 2021-06-26 authors.R
authors.R

@@ -1,0 +1,11 @@
+
+ali <- XX
+david <- XX
+ben <- XX
+
+set.seed(ali+ben+david)
+
+print(c(
+	"AG", "FA", "DE", 
+	sample(c("BB", "JD"))
+))

### 2021-06-28 SIR_manuscript.tex
refs are added

@@
 \author[1]{Ali Gharouni}
-\author[1]{Fady M. Abdelmalek}
+\author[1]{F.M. Abdelmalek}
 \author[1,2,3]{Benjamin M. Bolker}
 \author[2,4,3]{Jonathan Dushoff}
-\author[3,1,5]{David J. D. Earn}
+\author[1,3,5]{David J. D. Earn}
 
@@
 \affil[4]{South African Centre for Epidemiological Modelling and Analysis, University of Stellenbosch, Stellenbosch, South Africa}
-\affil[5]{Department of Mathematics, University of Toronto, Toronto, Canada}
 
@@
 The observed dynamics of the \covid epidemic have been driven both by epidemiological processes (infection and recovery) and by testing processes (testing and test reporting).
-In addition to shaping epidemic observations (via case reports), testing processes also alter epidemiological dynamics\needref.
+In addition to shaping epidemic observations (via case reports), testing processes also alter epidemiological dynamics \ali{\citep{peto2020covid,taipale2020population}}.
 Because individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may also do so, testing will generally increase the number of people who are isolating and hence reduce epidemic growth rates.
@@
 Over the course of the \covid pandemic, however, the vast majority of testing has been done with other goals --
-primarily diagnostic (determining infection status for clinical purposes)\needref, or for control (determining  infection status in order to isolate cases that have been found by contact tracing)\needref, which we characterize as \emph{targeted} testing strategies.
+primarily diagnostic (determining infection status for clinical purposes)\ali{\citep{phua2020intensive,who2020global}}, or for control (determining  infection status in order to isolate cases that have been found by contact tracing) \ali{\citep{aleta2020modelling,kucharski2020effectiveness,grassly2020comparison,smith2020adherence}}, which we characterize as \emph{targeted} testing strategies.
 In these situations, testing probabilities can differ sharply across epidemiological compartments; in our dynamical model, we will characterize these probabilities by assigning a testing weight to each compartment that determines the \emph{relative} probability that an individual in that compartment will be selected for testing (see Methods).
@@
 \jd{Should we talk less about \Rnum\ and keep the focus on $\Delta$? The back and forth is a bit tiring.}
-
+\ali{unsure here, Help!}
 We can use the formula for $\Rnum$ \eqref{R0} to make a number of straightforward inferences about parameters that affect $\Rnum$ monotonically, i.e., for which the associated partial derivative of $\Delta$ always has the same sign (see Appendices).
@@
 {\bf Effectiveness of testing and isolation in reducing $\Rnum$ at low \percap testing intensity ($\rho$).}
-Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4a}\ali{fixed}), over a range of testing and isolation parameters. Parameter values (Table~\ref{tab:params}):
+Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4a}), over a range of testing and isolation parameters. Parameter values (Table~\ref{tab:params}):
 $\beta=\betaparam/$day, $1/\gamma= \invgammaparam$ days (baseline $\Rnum=\Rnumparam$, $r=\rparam$); $\omega \in [1/12,2]/$day;  $\rho \in [0,0.013]/$day per capita; $\theta\_w$ and $\theta\_c$ vary between 0 (no effect of isolation) and 1 (complete elimination of transmission); $p_S=0$, $p_I=1$ and $p_R=0.5$.
@@
   {\bf Effectiveness of testing and isolation in reducing $\Rnum$ at high \percap testing intensity.}
-  Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4a}\ali{fixed}), over a range of testing and isolation parameters.
+  Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4a}), over a range of testing and isolation parameters.
 Parameters as in \fref{pan} except: $\omega \in [1/5,2]/$day, $\rho \in [0,1/5)/$day.
@@
 In this paper, we have developed and analyzed a simple compartmental model that combines epidemiological dynamics---as defined by a simple SIR model---with the dynamics of testing and isolation.
-Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates \needref; behavioural dynamics \citep{weitz2020awareness}).
+Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates \ali{\citep{chin2020frequency}}; behavioural dynamics \citep{weitz2020awareness}).
 Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
@@
 \david{The intro and discussion would benefit from (potentially many) additional references.}
-Although testing and tracing is a key part of infection control strategies, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts\needref, rather than analyzing simple but general models of the feedback between testing and transmission dynamics.
+Although testing and tracing is a key part of infection control strategies, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts \ali{\citep{endo2020implication,hellewell2020feasibility,jenness2020modeling}}, rather than analyzing simple but general models of the feedback between testing and transmission dynamics.
 \ali{the following sentences added to address Ben's comment:} There have been several modeling studies of testing and tracing dynamics and their interaction with epidemiological dynamics.

### 2021-06-28 note/SIR_scrap.tex
working on the comment
refs are added

@@
 \maketitle
+
+\section{Introduction}
+
+% Our model provides insights to the sensitivity of the epidemic dynamics, through $\Rnum$, with respect to the underlying testing and isolation parameters.
+%% BMB: last sentence seems to belong in conclusions
 
@@
 \subsubsection{Other maybe-related works}
+
+paper: Roques, Lionel, et al. "Impact of lockdown on the epidemic dynamics of COVID-19 in France." Frontiers in medicine 7 (2020): 274.
+RT-PCR tests have been used and existing data indicate that the sensitivity of this test using pharyngeal and nasal swabs is about 63−72\% (25). We assumed here σ = 0.7 (70\% sensitivity).
+(25) is this paper: 
+Wang, Wenling, et al. "Detection of SARS-CoV-2 in different types of clinical specimens." Jama 323.18 (2020): 1843-1844.
+
 \citep{arino2020simple} developed a SLIAR compartmental model to study the spread of an epidemic, specifically COVID, in a population.
@@
 
+% %%%%%%%% %%%%%%%% %%%%%%%% %%%%%%%% %%%%%%%% %%%%%%%% %%%%%%%% %%%%%%%
 \section{Unused Scripts}

### 2021-06-29 SIR_manuscript.tex
discussion covered

@@
 \ali{the following sentences added to address Ben's comment:} There have been several modeling studies of testing and tracing dynamics and their interaction with epidemiological dynamics.
-In the context of repeated screening and random testing of isolated populations (such as a university), \cite{bergstrom2020frequency} provided some analytical results\david{what are the results?
-that's what is most important, not whether the results were analytical}, \cite{rogers2021high} simulated a SEIR model with high-low \david{meaning?} testing intensity and sensitivity\david{what did they find?}.
-Further, \cite{friston2021testing} model the effects of self-isolation on testing and tracing\david{what did they find?}.
-Our modeling approach is novel compared to these works\david{how so?
-You can't just say our work is novel.
-What makes it novel relative to these previous papers?}.
+In the context of repeated screening and random testing of isolated populations (such as a university), \cite{bergstrom2020frequency} provided some analytical results which suggests, as an outbreak control strategy, that proactively screening asymptomatic individuals and isolating confirmed-positive cases would reduce the spread of the disease.
+In particular, They explored the effects of testing frequency and Trade-offs between frequency, accuracy and the reporting delay on the epidemic dynamic (more specifically, on the effective reproductive number).
+\cite{rogers2021high} simulated a SEIR model with testing and isolation which suggest a combination of rapid testing with antigen tests and the subsequent isolation of confirmed-positive individuals
+a strategy for controlling the disease dynamics in the \covid pandemic.
+Further, \cite{friston2021testing} model the effects of self-isolation on testing and tracing with the focused on posterior predictions and projections under different testing and tracing scenarios.
+They conclude that the emergence of a second wave depends almost exclusively on the rate at which immunity is lost and that it is necessary to track the 20\% of asymptomatic individuals in order to control the outbreak.
+Our modeling approach is novel compared to these works as it mechanistically incorporates the epidemic and testing processes with different levels of testing focus from random to highly targeted.
 We hope that the present paper will inspire further explorations of the fundamental properties of epidemic models that incorporate explicit testing processes.

### 2021-06-29 SIR_manuscript.tex
lots of small cleanup

@@
 \newcommand{\testinghat}[1]{\ensuremath{\hat{\cal T}_{#1}}\xspace}
-
+\newcommand{\testtarget}{\ensuremath{\frac{w_I}{w_S}}\xspace}
 \newcommand*\subtxt[1]{_{\textnormal{#1}}}
@@
 Testing individuals for pathogens can affect the spread of epidemics. 
-Understanding how individual-level processes of sampling and reporting test results can effect comunity- or population-level spread is a dynamical modeling question.
+Understanding how individual-level processes of sampling and reporting test results can affect community- or population-level spread is a dynamical modeling question.
 The effect of testing processes on epidemic dynamics depends on factors underlying implementation, particularly testing intensity and on whom testing is focused.
@@
 Our simple model shows that under some circumstances both increased testing intensity and faster test reporting can \emph{reduce} the effectiveness of control, and allows us to explore the conditions under which this occurs.
-Conversely, we find that focussing testing on infected individuals always acts to increase effectiveness of control.
+Conversely, we find that focusing testing on infected individuals always acts to increase effectiveness of control.
 \david{It would be better to say ``symptomatic'' rather than ``infected'' in the last sentence, but we can't claim the model really captures that.}
@@
 where $\beta/\gamma$ is the classical value for a simple model \citep{keeling2011modeling}, and $1-\Delta$ is the proportional reduction due to testing and isolation processes.
-$\Delta$ therefore measureness the ``effectiveness of control'': how much these processes in reducing spread from low levels, and is in turn given by:
+$\Delta$ therefore measures the ``effectiveness of control'': how much these processes in reducing spread from low levels, and is in turn given by:
 \begin{equation}
@@
 Matrices $F$ and $V$ are
+\bmb{eqs A2 and A3 can be in the same array environment, but sticking the ``and'' in there is ugly.
+  ``and'' should be in text, with A4 set as a separate equation (it doesn't need to be aligned).}
 \begin{align}
@@
 & 0 \leq \theta\_w \leq \theta\_c \leq 1, \\
-& \frac{w_I}{w_S}\geq 1.
+& \testtarget\geq 1.
 \end{align}
@@
 \label{eq:abcd2}
-a_3& = \frac{w_I}{w_S} \Bigg(
-&& (1-\theta\_w) (1+\frac{w_I}{w_S})\theta\_w \gamma^3 +(1-\theta\_c) p_I^2 \theta\_w \frac{w_I}{w_S} \omega^3 \\ \notag
-& ~ &&+\Big( \big( (1-\theta\_w-\frac{w_I}{w_S}) \theta\_c +(3-2\theta\_w) \theta\_w\frac{w_I}{w_S} \big) p_I +(1-\theta\_w) (1+\frac{w_I}{w_S})\theta\_w  \Big) \omega \gamma^2 \\ \notag
+a_3& = \testtarget \Bigg(
+&& (1-\theta\_w) (1+\testtarget)\theta\_w \gamma^3 +(1-\theta\_c) p_I^2 \theta\_w \testtarget \omega^3 \\ \notag
+& ~ &&+\Big( \big( (1-\theta\_w-\testtarget) \theta\_c +(3-2\theta\_w) \theta\_w\testtarget \big) p_I +(1-\theta\_w) (1+\testtarget)\theta\_w  \Big) \omega \gamma^2 \\ \notag
 & ~ &&+\Big(
-\big((1-\theta\_w-\theta\_w \frac{w_I}{w_S}) \theta\_c+ (2-\theta\_w) \theta\_w \frac{w_I}{w_S} \big) p_I
-+(2 \theta\_w-\theta\_w^2-\theta\_c) \frac{w_I}{w_S} p_I^2
+\big((1-\theta\_w-\theta\_w \testtarget) \theta\_c+ (2-\theta\_w) \theta\_w \testtarget \big) p_I
++(2 \theta\_w-\theta\_w^2-\theta\_c) \testtarget p_I^2
 \Big) \omega^2 \gamma \Bigg), \\ \notag
-b_3& = && \hspace{-1cm} 2\frac{w_I}{w_S} (\omega+\gamma)\gamma\Big(
+b_3& = && \hspace{-1cm} 2\testtarget (\omega+\gamma)\gamma\Big(
 (\omega+\gamma+\omega p_I)(2-\theta\_w) \gamma \theta\_w+ (1-\theta\_w) \omega^2 p_I\theta\_c+\omega^2 p_I\theta\_w
@@
 (2-\theta\_w) \gamma^2 \theta\_w+
-(1+\frac{w_I}{w_S}) \omega \gamma \theta\_w+
-\frac{w_I}{w_S} \omega^2 p_I\theta\_c \Big),\\ \notag
-d_3&  = && \hspace{-1cm} \frac{(\omega+\gamma)}{\omega}  \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi + (\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2.
+(1+\testtarget) \omega \gamma \theta\_w+
+\testtarget \omega^2 p_I\theta\_c \Big),\\ \notag
+d_3&  = && \hspace{-1cm} \frac{(\omega+\gamma)}{\omega}  \Big((\omega p_I+\gamma)\testtarget\phi + (\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2.
 \end{align}
@@
 
-An example of this counter-vailing effect of $\phi$, and consequently $\rho$, on $\Rnum$ occurs when $\theta\_w=0$ and $\theta\_c=1$.
+An example of this countervailing effect of $\phi$, and consequently $\rho$, on $\Rnum$ occurs when $\theta\_w=0$ and $\theta\_c=1$.
 This is illustrated in the top-right panel of the \fref{pan2} panel (b), where the strength of isolation for awaiting people is the least, but the most for the confirmed cases.
-In this case, simplifying $a_3$ in Eq.~\eqref{eq:abcd2} gives $$a_3=\frac{w_I}{w_S} \omega \gamma p_I((\omega+\gamma)-\frac{w_I}{w_S}(\omega p_I+\gamma)).$$
-Specifically, in the case of targeted testing which is identified by $\frac{w_I}{w_S}> 1$, and using a perfect sensitive test, thus $p_I=1$, there exists a range for $\rho$ over which $\pder\Rnum{\rho}\leq 0$.
-Note that $\rho$ and $\omega$ have a similar mechanism in delaying people to get into $I\_c$, thus we would expect to see the non-trivial counter-vailing effect of these two parameters on $\Rnum$.
+In this case, simplifying $a_3$ in Eq.~\eqref{eq:abcd2} gives
+\begin{equation}
+  \begin{split}
+    a_3 & =\testtarget \omega \gamma p_I\left((\omega+\gamma)-\testtarget(\omega p_I+\gamma)\right)  \\
+    & \propto  \omega \left( 1- \testtarget p_I \right) + \gamma \left( 1- \testtarget \right)  .
+  \end{split}
+\end{equation}
+If $p_I>0$, then $a_3<0$ for sufficiently targeted testing (i.e. when $\testtarget p_I >1$; because
+$p_I \leq 1$, \testtarget is always $\geq \testtarget p_I$).
+%% this LOOKS like we can get a3<0 even if pI=0 for targeting > 1 and sufficiently large gamma/omega,
+%% but that seems crazy so I'm probably missing something
+When the test is perfectly sensitive ($p_I=1$), $a_3<0$ as long as $\testtarget>1$.
+Under either of these conditions, there exists a range for $\rho$ over which $\pder\Rnum{\rho}\leq 0$.  
+Note that $\rho$ and $\omega$ have a similar mechanism in delaying people to get into $I\_c$, thus we would expect to see the non-trivial countervailing effect of these two parameters on $\Rnum$.
+\bmb{I don't understand the last sentence}
 
@@
 \subsection{Rate of returning tests} \label{app:omega}
-The third goal is to explore how changes in the rate of test return $\omega$ affects $\Rnum$.
+The third goal is to explore how changes in the rate of test return $\omega$ affect $\Rnum$.
 Mathematically we would like to verify the sign of $\pder\Rnum{\omega}$, which specifically depends on $\pder\Delta{\omega}$.
@@
 \Delta = \frac{\rho}{\omega \gamma(\omega+\gamma)} \Big(
-\frac{w_I}{w_S} \omega^2 p_I \theta\_c+(\frac{w_I}{w_S}+1) \gamma\omega\theta\_w +\gamma^2 \theta\_w(2-\theta\_w)
+\testtarget \omega^2 p_I \theta\_c+(\testtarget+1) \gamma\omega\theta\_w +\gamma^2 \theta\_w(2-\theta\_w)
 \Big). 
@@
 \pder\Delta{\omega} = \frac{\rho}{\omega^2 (\omega+\gamma)^2} \Big(
-(p_I \frac{w_I}{w_S}\theta\_c -(1+\frac{w_I}{w_S})\theta\_w)\omega^2-2\theta\_w \gamma(2-\theta\_w) \omega+\theta\_w \gamma^2 (\theta\_w-2)
+(p_I \testtarget\theta\_c -(1+\testtarget)\theta\_w)\omega^2-2\theta\_w \gamma(2-\theta\_w) \omega+\theta\_w \gamma^2 (\theta\_w-2)
 \Big).
@@
 \end{align}
-where $a_4 = p_I\frac{w_I}{w_S}\theta\_c$, $b_4 = (1+\frac{w_I}{w_S})\theta\_w$, $c_4 = \theta\_w(2-\theta\_w)$.
-This enables us to describe the behaviour of $\Rnum$ with respect to $\omega$ in the foolowing two cases.
+where $a_4 = p_I\testtarget\theta\_c$, $b_4 = (1+\testtarget)\theta\_w$, $c_4 = \theta\_w(2-\theta\_w)$.
+This enables us to describe the behaviour of $\Rnum$ with respect to $\omega$ in the following two cases.
 \begin{itemize}
@@
 \begin{equation}\label{case_1_ineq}
-\Big(\frac{w_I}{w_S}+1\Big)\theta\_w \geq \frac{w_I}{w_S}p_I\theta\_c \Longleftrightarrow \Big(\frac{w_S}{w_I}+1\Big)\frac{1}{p_I} \geq \frac{\theta\_c}{\theta\_w}.
+\Big(\testtarget+1\Big)\theta\_w \geq \testtarget p_I\theta\_c \Longleftrightarrow \Big(\frac{w_S}{w_I}+1\Big)\frac{1}{p_I} \geq \frac{\theta\_c}{\theta\_w}.
 \end{equation}
@@
 % %%%%%%%
-\subsection{The effect of testing focus parameter $\frac{w_I}{w_S}$ on $\Rnum$} \label{app:w}
+\subsection{The effect of testing focus parameter $\testtarget$ on $\Rnum$} \label{app:w}
 
-We define $w\_{IS}=\frac{w_I}{w_S}$.
+We define $w\_{IS}=\testtarget$. \bmb{shouldn't we use this simplified notation \textbf{throughout} the paper???}
 \begin{align}

### 2021-07-07 SIR_manuscript.tex
rewording result sec based on Delta, minor updates

@@
 Conversely, we find that focusing testing on infected individuals always acts to increase effectiveness of control.
-\david{It would be better to say ``symptomatic'' rather than ``infected'' in the last sentence, but we can't claim the model really captures that.}
 
@@
 
-\jd{Should we talk less about \Rnum\ and keep the focus on $\Delta$? The back and forth is a bit tiring.}
-\ali{unsure here, Help!}
-We can use the formula for $\Rnum$ \eqref{R0} to make a number of straightforward inferences about parameters that affect $\Rnum$ monotonically, i.e., for which the associated partial derivative of $\Delta$ always has the same sign (see Appendices).
+We presented $\Rnum$ as the product of the classical reproduction number, $\beta/\gamma$, and the proportional reduction due to testing and isolation, $1-\Delta$, \eqref{R0}.
+We can use the formula for $\Delta$ \eqref{eq:del4a} to make a number of straightforward inferences about parameters that affect $\Rnum$ monotonically, i.e., for which the associated partial derivative of $\Delta$ always has the same sign (see Appendices).
 
@@
 
-\item \label{p1:eta} Increasing isolation efficacy for waiting ($\theta\_w$) and confirmed-positive ($\theta\_c$) individuals always decreases \Rnum (Eqs.~\ref{eq:del4_theta},~\ref{eq:d_del4_thetac},~\ref{eq:d_del4_thetaw});
-\item \label{p1:rho} Higher testing intensity $\rho$ decreases $\Rnum$ if
+\item \label{p1:eta} Increasing isolation efficacy for waiting ($\theta\_w$) and confirmed-positive ($\theta\_c$) individuals always increases $\Delta$ (Eqs.~\ref{eq:del4_theta},~\ref{eq:d_del4_thetac},~\ref{eq:d_del4_thetaw});
+\item \label{p1:rho} Higher testing intensity $\rho$ increases $\Delta$ if
 %% \begin{enumerate} \item
@@
 %% \end{enumerate}
-\item \label{p1:omega} Increasing the rate of test return ($\omega$) always decreases \Rnum if waiting individuals do not isolate ($\theta\_w=0$) (Eq.~\ref{eq:dlindo}).
-\item \label{p1:w} Increasing testing focus, i.e., changing the testing weights from random ($w_S=w_I$) toward targeted  ($w_S<w_I$), always decreases \Rnum (Eq.~\ref{eq:d_del_wis}).
+\item \label{p1:omega} Increasing the rate of test return ($\omega$) always increases $\Delta$ if waiting individuals do not isolate ($\theta\_w=0$) (Eq.~\ref{eq:dlindo}).
+\item \label{p1:w} Increasing testing focus, i.e., changing the testing weights from random ($w_S=w_I$) toward targeted  ($w_S<w_I$), always increases $\Delta$ (Eq.~\ref{eq:d_del_wis}).
 \end{enumerate}
@@
 In this paper, we have developed and analyzed a simple compartmental model that combines epidemiological dynamics---as defined by a simple SIR model---with the dynamics of testing and isolation.
-Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates \ali{\citep{chin2020frequency}}; behavioural dynamics \citep{weitz2020awareness}).
+Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness}).
 Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
@@
 Thus, under certain extreme conditions (low $\theta\_w$, high $\theta\_c$, low $\omega$, high $\rho$), $\Rnum$ becomes an increasing function of testing intensity.
-\bmb{This phenomenon occurs primarily (\ali{only}) for targeted testing, because \ldots CAN WE FIGURE OUT WHY?
-Targeting should be irrelevant to determining the DFE (because everyone is $S$ at that point), so what is the dynamical explanation?
-Is there a second order effect somewhere, e.g.
-testing is more effective overall (darker colours in Fig 3(b) upper-right panel than in Fig 3(a) upper panel), so we more quickly reach the point where the effect of increasing $S\_n^*$ outweighs the beneficial effects of testing?} However, this phenomenon is even more unrealistic than the possibility that slowing test returns increases $\Rnum$.
+While we can show that this phenomenon \emph{only} occurs under targeted testing ($w_I > w_S$ ), we have not been able to come up with a verbal explanation of the mechanism (this would require interpreting the complicated expression for $a_3$ in Eq.~\ref{eq:abcd2}).
+In practice, this phenomenon is even more unrealistic than the possibility that slowing test returns increases $\Rnum$.
 It depends both on the assumption that regular testing is occurring before the epidemic starts, and on levels of testing that are unrealistically high (at least in large, general-population settings), i.e., at least 10\% of the population being tested per day.
@@
 
-\david{The intro and discussion would benefit from (potentially many) additional references.}
 Although testing and tracing is a key part of infection control strategies, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts \ali{\citep{endo2020implication,hellewell2020feasibility,jenness2020modeling}}, rather than analyzing simple but general models of the feedback between testing and transmission dynamics.
@@
 Matrices $F$ and $V$ are
-\bmb{eqs A2 and A3 can be in the same array environment, but sticking the ``and'' in there is ugly.
-  ``and'' should be in text, with A4 set as a separate equation (it doesn't need to be aligned).}
 \begin{align}
-\label{FV}
+\label{F}
 F =& \beta/N \left[ \begin {array}{cccc} 
@@
 0&0&0&0
- \end {array} \right] \\
+ \end {array} \right] \\ \notag
  =&\beta/N \left[ \begin {array}{c} S\_u^* \\ (1-\theta\_w)S\_n^* \\ 0 \\ 0 \end {array} \right]
-        \left[ \begin {array}{cccc} 1,   1-\theta\_w,   1-\theta\_w,   1-\theta\_c \end {array} \right], \text{and}\\  V=&
- \left[ \begin {array}{cccc}  
+        \left[ \begin {array}{cccc} 1,   1-\theta\_w,   1-\theta\_w,   1-\theta\_c \end {array} \right],
+\end{align}
+and 
+\begin{align}
+\label{V}
+V= \left[ \begin {array}{cccc}  
 \testinghat{I}+\gamma&-\omega&0&0\\
@@
 \label{vinv}
-V^{-1} =&
+V^{-1} =
 \frac{1}{\gamma C}
@@
 \end{equation}
-It is notable that matrix $F$ \eqref{FV} has rank one and consequently $G_{11}$ does so. That is $G_{11}$ has only one non-zero eigenvalue which is $\Rnum$.
+It is notable that matrix $F$ \eqref{F} has rank one and consequently $G_{11}$ does so. That is $G_{11}$ has only one non-zero eigenvalue which is $\Rnum$.
 
@@
 Note that $\rho$ and $\omega$ have a similar mechanism in delaying people to get into $I\_c$, thus we would expect to see the non-trivial countervailing effect of these two parameters on $\Rnum$.
-\bmb{I don't understand the last sentence}
 
@@
 However, attempting to do so quickly reveals a problem in the way the model is posed;
-with the scaling of testing ($\sigma$) defined as in the body of the paper (Eq.~\eqref{sigma}), the population in the $S$ compartments appeared to blow up when the DFE is achieved \bmb{clarify??}.
+with the scaling of testing ($\sigma$) defined as in the body of the paper (Eq.~\eqref{sigma}), the population in the $S$ compartments appeared to blow up when the system is near the DFE.
 This occurs because once the only untested people are susceptibles, the FOI approaches $\Lambda=0$, and the testing rate $\testing{S} \to \rho N/S\_u$.

### 2021-07-14 cover.tex

\documentclass[12pt,letterpaper]{letter}

%% layout
\usepackage[margin=1.25in]{geometry}
\usepackage{multicol}

%% colours
\usepackage{xcolor}
\definecolor{linkcolor}{HTML}{0f87ff} % green
\definecolor{irena}{HTML}{c9657d} % blush
\definecolor{de}{HTML}{5681c4} % blue

\usepackage[colorlinks=true, allcolors=linkcolor]{hyperref}

%% date format
\usepackage{datetime}
\newdateformat{mydate}{\twodigit{\THEDAY}{ }\shortmonthname[\THEMONTH] \THEYEAR}

%% comment macros
\newcommand{\irena}[1]{\textcolor{irena}{$\langle$\textbf{IP:} #1$\rangle$}}
\newcommand{\de}[1]{\textcolor{de}{$\langle$\textbf{DE:} #1$\rangle$}}

%% journal name macro
\newcommand{\journalname}{\emph{BMC Public Health}}

%% data macros
\input{data_macros.tex}

\setlength{\parskip}{0em}

\begin{document}

\begin{multicols}{2}
\footnotesize
\begin{flushleft}

Dr. Natalie Pafitis

Editor, \journalname{}

\vfill

{\normalsize \mydate
\today}
\end{flushleft}

\columnbreak

\begin{flushright}
Irena Papst

Center for Applied Mathematics, Cornell University

657 Rhodes Hall

136 Hoy Rd, Ithaca, NY, USA 14853

\href{mailto:ip98@cornell.edu}{ip98@cornell.edu}
\end{flushright}

\end{multicols}

\setlength{\parskip}{1em}
\thispagestyle{empty}

\vspace{-1em}

Dear Dr.\ Pafitis,

We are pleased to submit an original research article entitled
``Age-dependence of healthcare interventions for COVID-19 in Ontario,
Canada'' for your consideration for publication in \journalname{}. This study specifically aligns with the aims of \journalname{} to publish research related to the social determinants of health and adds new insight into the evolving COVID-19 pandemic.

Age is the most salient clinical indicator of risk from COVID-19 for
most patients. While age-specific distributions of known SARS-CoV-2
infections and COVID-19-related deaths have been widely reported,
relatively few sources present age distributions of hospitalizations
and serious healthcare interventions administered to COVID-19
patients. In our study, we analysed \nccmpresolved{} known SARS-CoV-2 infection
records for the province of Ontario, and quantified the age-intervention associations
for hospitalization, ICU admission, intubations, and ventilations. We
quantified existing age-related impacts of COVID-19 and identified
potential future risks should the healthcare system become overwhelmed
with COVID-19 patients.

Ontario has so far been able to offer and provide COVID-19 treatment
to all that request it. Yet we find that the probability of survival
given hospitalization for COVID-19 decreases substantially after age
40. The distribution of COVID-19-related hospitalizations spans a much
wider age range than deaths, which are concentrated in the extremely
elderly. If prevalence of the virus were to increase in the general
population (as is currently occurring), and ICUs or hospitals were to
subsequently reach maximum capacity, the distribution of deaths might expand toward younger ages.

While our work is specific to the Ontario SARS-CoV-2 epidemic, it has
more general and important implications for COVID-19 outbreaks all
over the world, in terms of both public health and clinical
practice. Our study reveals a potentially serious mortality risk to
middle-aged individuals from SARS-CoV-2 infection if the healthcare
system were to be overwhelmed with COVID-19 patients; this risk is not
captured by the widely available distributions of known SARS-CoV-2
infections and COVID-19-related deaths.

We have presented our study to the Ontario government's COVID-19
science advisory panel (the Ontario Science Table,
\url{https://covid19-sciencetable.ca/}), where it was received with
great interest.  We believe that the readership of \journalname{}
should be made aware of our findings as soon as possible.

All authors have approved the manuscript for submission and declare no competing interests. A preprint of this work has been posted to medRxiv (\url{https://www.medrxiv.org/content/10.1101/2020.09.01.20186395v2}). The manuscript has not been published or submitted for publication elsewhere.

Thank you very much for your consideration,

\begin{multicols}{2}
\begin{flushleft}
\footnotesize

Irena Papst, MSc
\setlength{\parskip}{0em}

PhD Candidate, Center for Applied Mathematics, Cornell University

\vspace{1em}

Michael Li, PhD

Postdoctoral Fellow, Department of Biology, McMaster University

Research Associate, South African Centre for Epidemiological Modelling and Analysis, University of Stellenbosch

\vspace{1em}

David Champredon, PhD

Postdoctoral Associate, Department of Pathology and Laboratory Medicine,

Western University

\vspace{1em}

\vfill\null
\columnbreak

Benjamin M.\ Bolker, PhD

Professor, Departments of Mathematics \& Statistics and Biology, McMaster University

Member, Michael G. DeGroote Institute for Infectious Disease Research, McMaster University

\vspace{1em}

Jonathan Dushoff, PhD

Professor, Department of Biology, McMaster University

Member, Michael G. DeGroote Institute for Infectious Disease Research, McMaster University

\vspace{1em}

David J.\,D.\ Earn, PhD

Professor, Department of Mathematics \& Statistics, McMaster University

Member, Michael G. DeGroote Institute for Infectious Disease Research, McMaster University

Professor, Department of Mathematics, University of Toronto
\end{flushleft}
\end{multicols}

\thispagestyle{empty}

\end{document}

### 2021-07-14 SIR_manuscript.tex
small edits
small bib/hyphenation fixes

@@
 
+\hyphenation{co-ro-na-vi-rus}
 \input{codes/modeldefs.tex}
@@
 \newcommand{\testinghat}[1]{\ensuremath{\hat{\cal T}_{#1}}\xspace}
-\newcommand{\testtarget}{\ensuremath{\frac{w_I}{w_S}}\xspace}
+\newcommand{\testtarget}{\ensuremath{w\_{IS}}}
+\newcommand{\oldtesttarget}{\ensuremath{\frac{w_I}{w_S}}\xspace}
 \newcommand*\subtxt[1]{_{\textnormal{#1}}}
@@
 \newcommand{\todo}[1]{\comment{red}{TODO}{#1}}
+\newcommand{\new}[1]{\comment{red}{NEW}{#1}}
 \newcommand{\needref}{\textcolor{red}{(REFS?)}}
@@
 The observed dynamics of the \covid epidemic have been driven both by epidemiological processes (infection and recovery) and by testing processes (testing and test reporting).
-In addition to shaping epidemic observations (via case reports), testing processes also alter epidemiological dynamics \ali{\citep{peto2020covid,taipale2020population}}.
+In addition to shaping epidemic observations (via case reports), testing processes also alter epidemiological dynamics \new{\citep{peto2020covid,taipale2020population}}.
 Because individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may also do so, testing will generally increase the number of people who are isolating and hence reduce epidemic growth rates.
@@
 Over the course of the \covid pandemic, however, the vast majority of testing has been done with other goals --
-primarily diagnostic (determining infection status for clinical purposes)\ali{\citep{phua2020intensive,who2020global}}, or for control (determining  infection status in order to isolate cases that have been found by contact tracing) \ali{\citep{aleta2020modelling,kucharski2020effectiveness,grassly2020comparison,smith2020adherence}}, which we characterize as \emph{targeted} testing strategies.
+primarily diagnostic (determining infection status for clinical purposes)\new{\citep{phua2020intensive,who2020global}}, or for control (determining  infection status in order to isolate cases that have been found by contact tracing) \new{\citep{aleta2020modelling,kucharski2020effectiveness,grassly2020comparison,smith2020adherence}}, which we characterize as \emph{targeted} testing strategies.
 In these situations, testing probabilities can differ sharply across epidemiological compartments; in our dynamical model, we will characterize these probabilities by assigning a testing weight to each compartment that determines the \emph{relative} probability that an individual in that compartment will be selected for testing (see Methods).
@@
 
-\jd{Discuss the possibility that people who can't get tests might also isolate. This undercuts our assumptions about $\theta\_w$.}
-
 In this paper, we have developed and analyzed a simple compartmental model that combines epidemiological dynamics---as defined by a simple SIR model---with the dynamics of testing and isolation.
@@
 Dynamically, this effect occurs because speeding up test returns shortens the isolation period of uninfected individuals (for infected people it only shortens the time to progression to the isolation level of the confirmed-positive compartment).
-Slowing test returns increases $\Rnum$ only if test positivity (the proportion of infectives in the tested population), and the magnitude of increase in isolation effectiveness from the waiting to the confirmed-positive compartment, are large enough to outweigh the decrease in isolation of uninfected waiting individuals.\david{That sentence is too long to process easily; I've re-organized slightly to make flow a bit better, but it is awkward to really fix it.}
+Slowing test returns increases $\Rnum$ only if the proportion of infectives in the tested population is high and isolation is relatively strong among people waiting for test results.
 
-While slowing test returns does decrease $\Rnum$ over broad regions of parameter space in our model, there are several real-world processes missing from our model that make it implausible that slowing test returns would actually be an effective public health measure.
+While slowing test returns does decrease $\Rnum$ over broad regions of parameter space in our model, there are several real-world processes missing from our model that make it unlikely that slowing test returns would actually be an effective public health measure.
 First, we do not model the primary benefit of rapid testing, i.e., detecting and containing outbreaks while they are still in progress.
 This process could be modeled phenomenologically by making the testing focus more targeted as an increasing proportion of cases is detected, because finding infections allows tests to be concentrated on their connections.
-Second, in the real world individuals become less likely to maintain isolation if they are required to do so for longer; phenomenologically, we could allow  effectiveness of isolation while waiting for test returns to be an increasing function of test return speed, or we could introduce a separate ``waiting, but no longer isolating'' compartment that individuals entered from the ``waiting, isolated'' compartment at a specified rate.
-Third and finally, if one wants to decrease the overall transmission rate of the population there are more effective methods than keeping tested people in limbo; these include masking, ventilation, distancing measures, retail and event closures, and stay-at-home orders.
+Second, individuals may become less likely to maintain isolation if they are required to do so for longer; phenomenologically, we could allow  effectiveness of isolation in the waiting population to be an increasing function of test-return speed, or we could introduce a separate ``waiting, but no longer isolating'' compartment that individuals entered from the ``waiting, isolated'' compartment at a specified rate.
+Finally, if one wants to decrease the overall transmission rate of the population there are more effective methods than keeping tested people in limbo; these include masking, ventilation, distancing measures, retail and event closures, and stay-at-home orders.
 
@@
 \fref{pan2}(b), upper right panel [$\theta\_c=1$, $\theta\_w=0$]).
-This phenomenon occurs mainly because of a peculiar feature of the DFE in our model: we implicitly assume that testing is ongoing before the beginning of the epidemic, so that the starting point of the epidemic invasion is an equilibrium distribution of susceptibles between the $S\_n$ (untested) and $S\_u$ (waiting) compartments: in particular, $S\_u^* = \rho/(\rho + \omega)$.
-(In reality, testing intensity itself is a dynamic variable that will increase as the epidemic proceeds.) If isolation in the waiting compartment is low ($\theta\_w \ll 1$), then waiting individuals can easily be infected once the epidemic starts.
-Once infected they can still infect others while in the waiting compartment, and cannot be tested again before they have returned to the untested compartment.
-Thus, under certain extreme conditions (low $\theta\_w$, high $\theta\_c$, low $\omega$, high $\rho$), $\Rnum$ becomes an increasing function of testing intensity.
-While we can show that this phenomenon \emph{only} occurs under targeted testing ($w_I > w_S$ ), we have not been able to come up with a verbal explanation of the mechanism (this would require interpreting the complicated expression for $a_3$ in Eq.~\ref{eq:abcd2}).
-In practice, this phenomenon is even more unrealistic than the possibility that slowing test returns increases $\Rnum$.
-It depends both on the assumption that regular testing is occurring before the epidemic starts, and on levels of testing that are unrealistically high (at least in large, general-population settings), i.e., at least 10\% of the population being tested per day.
+This phenomenon can occur because we are considering the DFE in the presence of testing; thus there is an equilibrium distribution of susceptibles between the $S\_n$ (untested) and $S\_u$ (waiting) compartments even as the disease approaches extinction.
+A higher rate of testing leads to a greater proportion of individuals waiting for negative tests at the DFE.
+If infected, individuals in this group will take longer to be tested again and to subsequently isolate (because they must wait for their negative tests to be returned before being tested again).
+If isolation in this group ($\theta\_w$) is low, this effect can under some (relatively rare) circumstances (high $\theta\_c$, low $\omega$, high $\rho$) allow $\Rnum$ to increase with testing intensity.
+We can show that this phenomenon \emph{only} occurs under targeted testing ($w_I > w_S$ ), but we have not yet found a simple explanation of why it cannot occur without it.
+This phenomenon is also unlikely to occur in the real world.
+In particular, it depends on levels of testing that are unrealistically high (at least in large, general-population settings).
 
 Although we model the testing process in more detail than typical epidemiological models, one place where more detail could be informative is in the processes determining the testing weights $\{w_S, w_I, w_R\}$.
-While random testing, as done for surveillance purposes, unambiguously leads to equal testing weights, making precise quantitative connections between public health practices and testing weights is difficult in other contexts.
+While random testing, as done for surveillance purposes, unambiguously leads to equal testing weights, making precise quantitative connections between public-health practices and testing weights is difficult in other contexts.
 The testing weights reflect the correlation between an individual's risk of infection and their likelihood of being tested due to age, occupation, geographic location, etc..
@@
 For example, separating the infected class into exposed, symptomatic, and a- or pre-symptomatic compartments and allowing the testing weights to vary across non-symptomatic (exposed/asymptomatic/presymptomatic) vs.\ symptomatic compartments could reflect the allocation of tests for diagnostic purposes (targeting symptomatic individuals) vs.\ contact-tracing (targeting infected but non-symptomatic individuals) vs.\ screening (relatively equal weights, depending on the venue).
-Alternatively, one could make the testing weights depend on the testing intensity or test return rate as suggested above.
+Alternatively, one could make the testing weights depend on the testing intensity or test-return rate as suggested above.
 Whatever complexity is added would probably put the model beyond reach of the analytical methods we have used in this paper, but one could still use semi-numerical methods such as constructing the next-generation matrix and using it to evaluate the derivatives of $\Rnum$ with respect to the parameters numerically.
 
-Although testing and tracing is a key part of infection control strategies, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts \ali{\citep{endo2020implication,hellewell2020feasibility,jenness2020modeling}}, rather than analyzing simple but general models of the feedback between testing and transmission dynamics.
-\ali{the following sentences added to address Ben's comment:} There have been several modeling studies of testing and tracing dynamics and their interaction with epidemiological dynamics.
-In the context of repeated screening and random testing of isolated populations (such as a university), \cite{bergstrom2020frequency} provided some analytical results which suggests, as an outbreak control strategy, that proactively screening asymptomatic individuals and isolating confirmed-positive cases would reduce the spread of the disease.
-In particular, They explored the effects of testing frequency and Trade-offs between frequency, accuracy and the reporting delay on the epidemic dynamic (more specifically, on the effective reproductive number).
-\cite{rogers2021high} simulated a SEIR model with testing and isolation which suggest a combination of rapid testing with antigen tests and the subsequent isolation of confirmed-positive individuals
-a strategy for controlling the disease dynamics in the \covid pandemic.
-Further, \cite{friston2021testing} model the effects of self-isolation on testing and tracing with the focused on posterior predictions and projections under different testing and tracing scenarios.
-They conclude that the emergence of a second wave depends almost exclusively on the rate at which immunity is lost and that it is necessary to track the 20\% of asymptomatic individuals in order to control the outbreak.
-Our modeling approach is novel compared to these works as it mechanistically incorporates the epidemic and testing processes with different levels of testing focus from random to highly targeted.
-We hope that the present paper will inspire further explorations of the fundamental properties of epidemic models that incorporate explicit testing processes.
+Although testing and tracing is a key part of infection control strategies, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts \new{\citep{endo2020implication,hellewell2020feasibility,jenness2020modeling}}, rather than analyzing simple but general models of the feedback between testing and transmission dynamics.
+\new{There have been several modeling studies of testing and tracing dynamics and their interaction with epidemiological dynamics.
+In the context of repeated screening and random testing of isolated populations (such as a university), \cite{bergstrom2020frequency} provided some analytical results which suggest that proactively screening asymptomatic individuals and isolating confirmed-positive cases would reduce the spread of the disease.
+\cite{rogers2021high} simulated a SEIR model with testing and isolation, and suggest a strategy of rapid testing with antigen tests and the subsequent isolation of confirmed-positive individuals.
+H \cite{friston2021testing} model the effects of self-isolation on testing and tracing with a focus on projections under different testing and tracing scenarios.
+They conclude that the emergence of a second wave depends almost primarily on the rate at which immunity is lost and that it is necessary to track asymptomatic individuals in order to control the outbreak.
+Our modeling approach differs from earlier work in that it examines the effects of test-return rates and of different levels of testing focus, from random to highly targeted.
+We hope this paper will inspire further explorations of the fundamental properties of epidemic models that incorporate explicit testing processes.}
 
@@
 This one-to-one monotonic reparameterization enables us to simplify the mathematical expressions and explore the simpler $\pder\Delta{\phi}$ instead of the complicated $\pder\Delta{\rho}$.
-The derivative is 
+Defining $\testtarget \equiv \oldtesttarget$, the derivative is 
 \begin{align}
@@
 If $a_3\geq 0$, $\partial\Delta/\partial\phi \geq 0$ for all feasible range of parameters, thus $\pro\Rnum \leq 0$.
-It is straight forward to show that $a_3\geq 0$ in case of random testing strategy, i.e., $w\_S=w\_I=1$.
+It is straightforward to show that $a_3\geq 0$ when testing is random, i.e., $w\_S=w\_I=1$.
 If $a_3 < 0$, then the quadratic expression in the numerator of \eqref{eq:dd3dr} has a positive root, $\phi^*$, such that for $\phi>\phi^*$, $\partial\Delta/\partial\phi < 0$.
@@
 
-We define $w\_{IS}=\testtarget$. \bmb{shouldn't we use this simplified notation \textbf{throughout} the paper???}
 \begin{align}

### 2021-07-14 authors.R
Final author order

@@ -1,7 +1,7 @@
 
-ali <- XX
-david <- XX
-ben <- XX
+ali <- 330
+david <- 8128
+ben <- 1037
 
 set.seed(ali+ben+david)
 

### 2021-07-14 SIR_manuscript.tex

@@
 
-The other counterintuitive result from our analysis is that, for sufficiently high testing intensity $\rho$, further increasing testing intensity can actually \emph{increase} $\Rnum$ (e.g.
-\fref{pan2}(b), upper right panel [$\theta\_c=1$, $\theta\_w=0$]).
+The other counterintuitive result from our analysis is that, for sufficiently high testing intensity $\rho$, further increasing testing intensity can actually \emph{increase} $\Rnum$ (e.g., \fref{pan2}(b), upper right panel [$\theta\_c=1$, $\theta\_w=0$]).
 This phenomenon can occur because we are considering the DFE in the presence of testing; thus there is an equilibrium distribution of susceptibles between the $S\_n$ (untested) and $S\_u$ (waiting) compartments even as the disease approaches extinction.
@@
 If isolation in this group ($\theta\_w$) is low, this effect can under some (relatively rare) circumstances (high $\theta\_c$, low $\omega$, high $\rho$) allow $\Rnum$ to increase with testing intensity.
-We can show that this phenomenon \emph{only} occurs under targeted testing ($w_I > w_S$ ), but we have not yet found a simple explanation of why it cannot occur without it.
+We can show that this phenomenon occurs \emph{only} under targeted testing ($w_I > w_S$), but we have not yet found a simple explanation of why it cannot occur without it.
 This phenomenon is also unlikely to occur in the real world.

### 2021-07-14 SIR_manuscript.tex
tidied Table 1

@@
 \newcommand{\needref}{\textcolor{red}{(REFS?)}}
+
+%% DE: table stuff
+\renewcommand{\arraystretch}{1.5} % more interline spacing
+%% https://tex.stackexchange.com/questions/12703/how-to-create-fixed-width-table-columns-with-text-raggedright-centered-raggedlef
+\usepackage{ragged2e}
+\usepackage{array} % for m{...}
 
@@
 \centering
-\begin{tabular}{|c|L{2in}cc|} \hline
-  Symbol & Description & Unit & Value \\ \hline
+%%\begin{tabular}{|c|L{2in}cc|} \hline
+%% making use of ragged2e and array packages:
+{\RaggedRight
+\begin{tabular}{ c | m{6cm} | c | c }
+  \textbf{Symbol} & \textbf{Description} & \textbf{Unit} & \textbf{Value} \\ \hline
   $N$     & Total population size & people & $10^6$ \\ \hline
@@
   \begin{minipage}[t]{0.21\columnwidth}%
-    Random: $\{1,1,1\}$ \\ Targeted: $\{0.3,1,1\}$
+    Random:~$\{1,1,1\}$ \\ Targeted:~$\{0.3,1,1\}$
   \end{minipage} \\
-  \hline
+  %%\hline
 \end{tabular}
-\caption{\label{tab:params} Parameters of the model specified in \eqref{model}; see also \fref{fig:flowchart}.}
+} % end \RaggedRight
+\caption{\label{tab:params}Parameters of the model specified  by the flow chart in \fref{fig:flowchart} and equations~\eqref{model}.}
 \end{table}

### 2021-07-14 cover.tex
cover let

@@
 %% data macros
-\input{data_macros.tex}
+% \input{data_macros.tex}
 
@@
 and serious healthcare interventions administered to COVID-19
-patients. In our study, we analysed \nccmpresolved{} known SARS-CoV-2 infection
+patients. In our study, we analysed known SARS-CoV-2 infection
 records for the province of Ontario, and quantified the age-intervention associations

### 2021-07-14 SIR_manuscript.tex
a few more tweaks

@@
 
-However, we did find two surprising patterns: longer delays in returning tests can \emph{reduce} epidemic spread, and increasing testing rates can \emph{increase} spread.
+However, we did find two surprising patterns: under some conditions longer delays in returning tests can reduce epidemic spread, and increasing testing rates can increase spread.
 
@@
 If isolation in this group ($\theta\_w$) is low, this effect can under some (relatively rare) circumstances (high $\theta\_c$, low $\omega$, high $\rho$) allow $\Rnum$ to increase with testing intensity.
-We can show that this phenomenon occurs \emph{only} under targeted testing ($w_I > w_S$), but we have not yet found a simple explanation of why it cannot occur without it.
+We can show that this phenomenon occurs \emph{only} under targeted testing ($w_I > w_S$), but we have not yet found a simple explanation of why it cannot occur under random (unfocused) testing.
 This phenomenon is also unlikely to occur in the real world.
@@
 \new{There have been several modeling studies of testing and tracing dynamics and their interaction with epidemiological dynamics.
-In the context of repeated screening and random testing of isolated populations (such as a university), \cite{bergstrom2020frequency} provided some analytical results which suggest that proactively screening asymptomatic individuals and isolating confirmed-positive cases would reduce the spread of the disease.
-\cite{rogers2021high} simulated a SEIR model with testing and isolation, and suggest a strategy of rapid testing with antigen tests and the subsequent isolation of confirmed-positive individuals.
-H \cite{friston2021testing} model the effects of self-isolation on testing and tracing with a focus on projections under different testing and tracing scenarios.
+In the context of repeated screening and random testing of isolated populations (such as the members of a university), \cite{bergstrom2020frequency} provided analytical results quantifying the effects that proactive screening of asymptomatic individuals and isolation of confirmed-positive cases could have in reducing the spread of disease.
+\cite{rogers2021high} simulated a SEIR model with testing and isolation; they similarly suggest a strategy of rapid testing with antigen tests and the subsequent isolation of confirmed-positive individuals.
+\cite{friston2021testing} model the effects of self-isolation on testing and tracing with a focus on projections under different testing and tracing scenarios.
 They conclude that the emergence of a second wave depends almost primarily on the rate at which immunity is lost and that it is necessary to track asymptomatic individuals in order to control the outbreak.
-Our modeling approach differs from earlier work in that it examines the effects of test-return rates and of different levels of testing focus, from random to highly targeted.
+Our modeling approach differs from these previous efforts in that it examines the effects of test-return rates and of different levels of testing focus, from random to highly targeted.
 We hope this paper will inspire further explorations of the fundamental properties of epidemic models that incorporate explicit testing processes.}
@@
 
-It remains hard to show that the reproduction number $\Rnum$ is decreasing with respect to \percap testing intensity, $\rho$, and the speed of the test return, $\omega$, for the feasible ranges of the parameters, that is
+It remains difficult to show that the reproduction number $\Rnum$ is decreasing with respect to \percap testing intensity, $\rho$, and the speed of the test return, $\omega$, for the feasible ranges of the parameters, that is
 \begin{align}
@@
 Under either of these conditions, there exists a range for $\rho$ over which $\pder\Rnum{\rho}\leq 0$.  
-Note that $\rho$ and $\omega$ have a similar mechanism in delaying people to get into $I\_c$, thus we would expect to see the non-trivial countervailing effect of these two parameters on $\Rnum$.
+Because increasing values of $\rho$ and $\omega$ both delay the rate at which individuals flow to the $I\_c$ compartment, it is reasonable that increasing either value could (under appropriate circumstances) increase $\Rnum$.
 
@@
 Next, note that the left-hand side increases when test sensitivity decreases and when targeting of positive individuals is poor.
-This is consistent with our intuition: a false negative that is returned more rapidly may encourage an infectious individual to not carefully self-isolate, thus increasing transmission.
+This is consistent with our intuition: a false negative that is returned more rapidly will allow an infectious individual to relax their self-isolation, thus increasing transmission.
 Likewise, if individuals tested are mainly susceptible (rather than infectious), then returning tests more slowly would encourage them to self-isolate for longer while awaiting test results.
-Having understood the role of each of the parameters in \ref{case_1_ineq}, a holistic interpretation of this inequality is that returning tests more slowly is helpful when the benefit of extended self-isolation (due to awaiting test results) outweighs the benefit of identifying positive cases.
+Having understood the role of each of the parameters in \ref{case_1_ineq}, a holistic interpretation of this inequality is that returning tests more slowly is helpful when the benefit of extended self-isolation by infected individuals awaiting test results outweighs the benefit of identifying positive cases.
 

### 2021-07-16 SIR_manuscript.tex
fixing eq refs, finalizing

@@
 \author[1]{F.M. Abdelmalek}
+\author[1,3,5]{David J. D. Earn}
+\author[2,4,3]{Jonathan Dushoff}
 \author[1,2,3]{Benjamin M. Bolker}
-\author[2,4,3]{Jonathan Dushoff}
-\author[1,3,5]{David J. D. Earn}
 
@@
 The observed dynamics of the \covid epidemic have been driven both by epidemiological processes (infection and recovery) and by testing processes (testing and test reporting).
-In addition to shaping epidemic observations (via case reports), testing processes also alter epidemiological dynamics \new{\citep{peto2020covid,taipale2020population}}.
+In addition to shaping epidemic observations (via case reports), testing processes also alter epidemiological dynamics \citep{peto2020covid,taipale2020population}.
 Because individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may also do so, testing will generally increase the number of people who are isolating and hence reduce epidemic growth rates.
@@
 
-If testing influences behavior, then epidemic dynamics will depend on who gets tested.
+If testing influences behaviour, then epidemic dynamics will depend on who gets tested.
 The impacts of testing will depend both on testing intensity (tests performed per day) and on how strongly testing is focused on people who are infectious.
@@
 Over the course of the \covid pandemic, however, the vast majority of testing has been done with other goals --
-primarily diagnostic (determining infection status for clinical purposes)\new{\citep{phua2020intensive,who2020global}}, or for control (determining  infection status in order to isolate cases that have been found by contact tracing) \new{\citep{aleta2020modelling,kucharski2020effectiveness,grassly2020comparison,smith2020adherence}}, which we characterize as \emph{targeted} testing strategies.
+primarily diagnostic (determining infection status for clinical purposes) \citep{phua2020intensive,who2020global}, or for control (determining  infection status in order to isolate cases that have been found by contact tracing) \citep{aleta2020modelling,kucharski2020effectiveness,grassly2020comparison,smith2020adherence}, which we characterize as \emph{targeted} testing strategies.
 In these situations, testing probabilities can differ sharply across epidemiological compartments; in our dynamical model, we will characterize these probabilities by assigning a testing weight to each compartment that determines the \emph{relative} probability that an individual in that compartment will be selected for testing (see Methods).
@@
 The force of infection is denoted by $\Lambda$ (Eq.~\eqref{Lambda}); $\gamma$ is the recovery rate; $\omega$ is the rate of test return; and \testing{X} (Eq.
-\eqref{F}) and $p_X$ represent the \percap testing rate and the sensitivity (probability of positive tests), respectively, for compartment $X$.
+\eqref{test}) and $p_X$ represent the \percap testing rate and the sensitivity (probability of positive tests), respectively, for compartment $X$.
 For further description of the parameters see \cref{tab:params}.
@@
 \begin{equation}
-\label{F}
+\label{test}
 \testing{X}=\sigma w_X.
@@
 \end{equation}
-The corresponding \percap testing rate (Eq.~\ref{F}) for the infected compartment $I$ at DFE is one of the key analysis parameters and can be simplified as 
+The corresponding \percap testing rate (Eq.~\ref{test}) for the infected compartment $I$ at DFE is one of the key analysis parameters and can be simplified as 
 \begin{equation}
@@
 We are specifically interested in parameters that could be manipulated by public health policy: isolation efficacy, $\theta\_c$ and $\theta\_w$; \percap testing intensity, $\rho$; and the rate of test return, $\omega$.
-In particular, we look at the partial derivatives of $\Delta$ with respect to these parameters (\appref{rho}).
+In particular, we look at the partial derivatives of $\Delta$ with respect to these parameters (Appendices~\ref{app:rho} and \ref{app:omega}).
 We derived general expressions for these derivatives.
@@
 
-The low-testing case \fref{pan} reflects the case where testing intensity $\rho$ is small relative to the population size.
+The low-testing case (\fref{pan}) reflects the case where testing intensity $\rho$ is small relative to the population size.
 Specifically, $\rho \in [0,0.013]$, and test return rate $\omega\in [1/12,2]$.
 This testing intensity is of the correct order of magnitude (although typically larger than) testing rates during the \covid pandemic, i.e., a maximum of 1.3\% of the population per day (approximately four times the maximum testing rate in Ontario, Canada in mid-2021).
-The less realistic high-testing case \fref{pan2} is included to highlight the occurrence of non-monotonic changes in $\Rnum$ with respect to $\rho$.
+The less realistic high-testing case (\fref{pan2}) is included to highlight the occurrence of non-monotonic changes in $\Rnum$ with respect to $\rho$.
 In \fref{pan2} the maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,1/5)$ and the test return rate $\omega\in [1/5,2]$; these values are clearly unrealistic for a large population but might be relevant for small populations undergoing focused testing, such as a sports league or university.
@@
 
-Although testing and tracing is a key part of infection control strategies, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts \new{\citep{endo2020implication,hellewell2020feasibility,jenness2020modeling}}, rather than analyzing simple but general models of the feedback between testing and transmission dynamics.
-\new{There have been several modeling studies of testing and tracing dynamics and their interaction with epidemiological dynamics.
+Although testing and tracing is a key part of infection control strategies, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts \citep{endo2020implication,hellewell2020feasibility,jenness2020modeling}, rather than analyzing simple but general models of the feedback between testing and transmission dynamics.
+There have been several modeling studies of testing and tracing dynamics and their interaction with epidemiological dynamics.
 In the context of repeated screening and random testing of isolated populations (such as the members of a university), \cite{bergstrom2020frequency} provided analytical results quantifying the effects that proactive screening of asymptomatic individuals and isolation of confirmed-positive cases could have in reducing the spread of disease.
@@
 Our modeling approach differs from these previous efforts in that it examines the effects of test-return rates and of different levels of testing focus, from random to highly targeted.
-We hope this paper will inspire further explorations of the fundamental properties of epidemic models that incorporate explicit testing processes.}
+We hope this paper will inspire further explorations of the fundamental properties of epidemic models that incorporate explicit testing processes.
 
@@
 & 0 \leq \theta\_w \leq \theta\_c \leq 1, \\
-& \testtarget\geq 1.
+& \oldtesttarget\geq 1.
 \end{align}
@@
 
-{\bf The second goal} is to explore how changes in \percap testing intensity $\rho$ affects $\Rnum$.
+The second goal is to explore how changes in \percap testing intensity $\rho$ affects $\Rnum$.
 Mathematically we would like to verify the sign of $\pder\Rnum{\rho}$, which specifically depends on $\pder\Delta{\rho}$.
@@
 % %%%%%%%
+\subsection{The effect of testing focus parameter $\testtarget$ on $\Rnum$} \label{app:w}
+
+\begin{align}
+\label{eq:d_del_wis}
+\pder \Delta{w\_{IS}}= \dfrac{(\omega-\rho) (\omega(\omega-\rho\theta\_w)+\gamma(\omega-\rho)) (\theta\_w\gamma+\theta\_c\omega p\_I)}
+{(-\omega^2 \gamma+\omega \gamma \rho-\gamma \rho \omega w\_{IS}-
+\omega \gamma^2+\gamma^2 \rho-\omega^2 p_I \rho w\_{IS})^2},
+\end{align}
+which is a positive quantity. Thus, $\pder \Rnum{w\_{IS}}\leq 0$. Therefore, increasing the focus of testing on the infectious people will result in less transmission.
+
+% %%%%%%%
 \subsection{On Testing Rate and Numerical Singularity} \label{app:singularity}
@@
 
-% %%%%%%%
-\subsection{The effect of testing focus parameter $\testtarget$ on $\Rnum$} \label{app:w}
-
-\begin{align}
-\label{eq:d_del_wis}
-\pder \Delta{w\_{IS}}= \dfrac{(\omega-\rho) (\omega(\omega-\rho\theta\_w)+\gamma(\omega-\rho)) (\theta\_w\gamma+\theta\_c\omega p\_I)}
-{(-\omega^2 \gamma+\omega \gamma \rho-\gamma \rho \omega w\_{IS}-
-\omega \gamma^2+\gamma^2 \rho-\omega^2 p_I \rho w\_{IS})^2},
-\end{align}
-which is a positive quantity. Thus, $\pder \Rnum{w\_{IS}}\leq 0$. Therefore, increasing the focus of testing on the infectious people will result in less transmission.
-
 \end{document}

### 2021-07-19 SIR_manuscript.tex
breaking the sections, filling the BMB template

@@
 \usepackage{authblk}
+\usepackage{ragged2e}
+\usepackage{array}
 
@@
 % https://tex.stackexchange.com/questions/12703/how-to-create-fixed-width-table-columns-with-text-raggedright-centered-raggedlef
-\usepackage{array}
 \newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
@@
 \newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-
-\usepackage{xspace}
 
@@
 %% https://tex.stackexchange.com/questions/12703/how-to-create-fixed-width-table-columns-with-text-raggedright-centered-raggedlef
-\usepackage{ragged2e}
-\usepackage{array} % for m{...}
+% \usepackage{ragged2e}
+% \usepackage{array} % for m{...}
 
@@
 \author[1]{F.M. Abdelmalek}
-\author[1,3,5]{David J. D. Earn}
-\author[2,4,3]{Jonathan Dushoff}
+\author[1,3]{David J. D. Earn}
+\author[2,3,4]{Jonathan Dushoff}
 \author[1,2,3]{Benjamin M. Bolker}
 
-\affil[1]{Department of Mathematics \& Statistics, McMaster University, Hamilton, Canada}
+\affil[1]{Department of Mathematics and Statistics, McMaster University, Hamilton, Canada}
 \affil[2]{Department of Biology, McMaster University, Hamilton, Canada}
@@
 \eqref{test}) and $p_X$ represent the \percap testing rate and the sensitivity (probability of positive tests), respectively, for compartment $X$.
-For further description of the parameters see \cref{tab:params}.
+For further description of the parameters see Table~ \ref{tab:params}.
 Note that there is a slight mismatch in the top-to-bottom order of the testing-based compartments of each disease-based compartment $X$ between this flowchart and the model equations \eqref{model}; here we have switched  $X\_u$ and $X\_n$ for visual clarity.

### 2021-07-19 abstract.tex
breaking the sections, filling the BMB template

--- 2021-07-14 SIR_manuscript.tex
@@
-\documentclass[12pt]{article}
-\usepackage{natbib}
-\usepackage{hyperref}
-\usepackage{grffile}
-\usepackage{graphicx}
-\usepackage{subcaption}
-\usepackage{amssymb,amsmath,amsthm}
-\usepackage{xcolor}
-\usepackage{xspace}
-\usepackage[nameinlink,capitalize]{cleveref}
-\usepackage{cleveref}
-\usepackage[margin=1in]{geometry}
-\usepackage{lineno}\renewcommand\thelinenumber{\color{gray}\arabic{linenumber}}
-\usepackage{pdflscape}
-\usepackage{enumerate}
-\usepackage{authblk}
-
-\hyphenation{co-ro-na-vi-rus}
-\input{codes/modeldefs.tex}
-
-% https://tex.stackexchange.com/questions/12703/how-to-create-fixed-width-table-columns-with-text-raggedright-centered-raggedlef
-\usepackage{array}
-\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-
-\usepackage{xspace}
-
-\newcommand{\fref}[1]{Fig.~\ref{#1}}
-\newcommand{\appref}[1]{Appendix~\ref{app:#1}}
-\newcommand{\Rlogo}{R\xspace}
-\newcommand{\percap}{\emph{per capita}\xspace}
-\newcommand{\Rnum}{\ensuremath{\mathcal{R}_0}\xspace}
-\newcommand{\covid}{COVID-19\xspace}
-\newcommand{\pro}[1]{\ensuremath{\frac{\partial #1}{\partial \rho}}}
-\newcommand{\pder}[2]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder y{x}
-\newcommand{\testing}[1]{\ensuremath{{\cal T}_{#1}}\xspace}
-\newcommand{\testinghat}[1]{\ensuremath{\hat{\cal T}_{#1}}\xspace}
-\newcommand{\testtarget}{\ensuremath{w\_{IS}}}
-\newcommand{\oldtesttarget}{\ensuremath{\frac{w_I}{w_S}}\xspace}
-\newcommand*\subtxt[1]{_{\textnormal{#1}}}
-%% Roman subscript
-\DeclareRobustCommand\_{\ifmmode\expandafter\subtxt\else\textunderscore\fi}
-
-\newcommand{\comment}{\showcomment}
-\newcommand{\showcomment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}
-\newcommand{\nocomment}[3]{}
-
-\newcommand{\fady}[1]{\comment{cyan}{Fady}{#1}}
-\newcommand{\ali}[1]{\comment{magenta}{Ali}{#1}}
-\newcommand{\jd}[1]{\comment{blue}{JD}{#1}}
-\newcommand{\david}[1]{\comment{orange}{DJDE}{#1}}
-\newcommand{\bmb}[1]{\comment{red}{BMB}{#1}}
-\newcommand{\todo}[1]{\comment{red}{TODO}{#1}}
-\newcommand{\new}[1]{\comment{red}{NEW}{#1}}
-\newcommand{\needref}{\textcolor{red}{(REFS?)}}
-
-%% DE: table stuff
-\renewcommand{\arraystretch}{1.5} % more interline spacing
-%% https://tex.stackexchange.com/questions/12703/how-to-create-fixed-width-table-columns-with-text-raggedright-centered-raggedlef
-\usepackage{ragged2e}
-\usepackage{array} % for m{...}
-
-\theoremstyle{definition} % amsthm only
-\newtheorem{proposition}{Proposition}
-\newtheorem{theorem}{Theorem}
-
-\bibliographystyle{apalike}
-
-\title{Testing and Isolation Efficacy:\\Insights from a Simple Epidemic Model}
-\author[1]{Ali Gharouni}
-\author[1]{F.M. Abdelmalek}
-\author[1,2,3]{Benjamin M. Bolker}
-\author[2,4,3]{Jonathan Dushoff}
-\author[1,3,5]{David J. D. Earn}
-
-\affil[1]{Department of Mathematics \& Statistics, McMaster University, Hamilton, Canada}
-\affil[2]{Department of Biology, McMaster University, Hamilton, Canada}
-\affil[3]{Michael G. DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, Canada}
-\affil[4]{South African Centre for Epidemiological Modelling and Analysis, University of Stellenbosch, Stellenbosch, South Africa}
-
-\begin{document}
-\maketitle
-
-\linenumbers
-
-% %%%%%%%
-\section{Abstract}
+% \section{Abstract}
 
@@
 Conversely, we find that focusing testing on infected individuals always acts to increase effectiveness of control.
-
-% %%%%%%%
-\section{Introduction}
-
-The observed dynamics of the \covid epidemic have been driven both by epidemiological processes (infection and recovery) and by testing processes (testing and test reporting).
-In addition to shaping epidemic observations (via case reports), testing processes also alter epidemiological dynamics \new{\citep{peto2020covid,taipale2020population}}.
-Because individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may also do so, testing will generally increase the number of people who are isolating and hence reduce epidemic growth rates.
-We developed a mechanistic model that incorporates epidemic processes and testing in order to explore the effects of testing and isolation on epidemic dynamics.
-
-If testing influences behavior, then epidemic dynamics will depend on who gets tested.
-The impacts of testing will depend both on testing intensity (tests performed per day) and on how strongly testing is focused on people who are infectious.
-This level of focus depends in turn on the purpose and design of testing programs. 
-When testing is done for the purposes of disease surveillance \citep{foddai2020base}
-tests are typically conducted randomly (or using a stratified random design) across the population in order to make an unbiased assessment of population prevalence.
-
-Over the course of the \covid pandemic, however, the vast majority of testing has been done with other goals --
-primarily diagnostic (determining infection status for clinical purposes)\new{\citep{phua2020intensive,who2020global}}, or for control (determining  infection status in order to isolate cases that have been found by contact tracing) \new{\citep{aleta2020modelling,kucharski2020effectiveness,grassly2020comparison,smith2020adherence}}, which we characterize as \emph{targeted} testing strategies.
-In these situations, testing probabilities can differ sharply across epidemiological compartments; in our dynamical model, we will characterize these probabilities by assigning a testing weight to each compartment that determines the \emph{relative} probability that an individual in that compartment will be selected for testing (see Methods).
-
-Diagnostic testing focuses on people with infection-like symptoms; thus the relative testing weights for infected people will depend on the relative probability of infected people having symptoms.
-For \covid infection, the testing weights will depend on the proportion of asymptomatic infections, the time spent pre-symptomatic vs.\ symptomatic during the course of an infection, and on the incidence of \covid-like symptoms among people in the population \emph{not} infected with \covid.
-Testing for epidemic control focuses on known contacts of infected people; in this case the testing weights for infected vs.\ uninfected people will depend on the probability of infection given contact, as well as the effectiveness of the system for identifying suspicious contacts.
-
-When a new infectious disease emerges, it is important to determine whether it will grow exponentially in a susceptible population, and if so at what rate $r$ \citep{Ma+14}.
-The condition for positive exponential growth ($r>0$) is commonly expressed as $\Rnum>1$, where the basic reproduction number $\Rnum$ is the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}.
-Although the value of $\Rnum$ cannot completely characterize the dynamics of our model \citep{shaw2021what}, it does give a simple and widely accepted index for the difficulty of control, as well as an indication of the likely final size of an epidemic \citep{ma2006generality,miller2012note}.
-
-In order to understand the effect of testing processes on epidemic dynamics, we expanded one of the simplest mechanistic epidemic models---the standard deterministic SIR model\citep{KermMcKe27,AndeMay91}---to include testing components.
-This model provides a sensible platform to link the modeling of epidemic and testing components and study their interaction.
-We studied the effects of testing intensity, rate of test return, and isolation efficacy, on transmission probability and epidemic dynamics when different levels of testing focus (from random to highly targeted) are in place.
-
-% %%%%%%%
-\section{Methods}
-
-Our model groups individuals based on disease status (Susceptible, Infectious or Recovered) and testing status (\emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}) (\fref{fig:flowchart}).
-The testing status of an individual in a given disease compartment $X$ (where $X \in \{S,I,R\}$) is denoted by a subscript, namely $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
-Two `accumulator' compartments, $N$ and $P$, are included in order to collect cumulative reported negative or positive tests.
-The model equations \eqref{model} and details of calculation of the basic reproduction number $\Rnum$ are presented in \appref{R0}.
-
-\begin{figure}
-\begin{center} 
-\includegraphics[scale=1.5]{pix/sir_comp.pdf}
-\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model, \ref{model}.
-The disease-based status of a compartment $X$ ($X \in \{S,I,R\}$) is combined with the testing status including $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
-The force of infection is denoted by $\Lambda$ (Eq.~\eqref{Lambda}); $\gamma$ is the recovery rate; $\omega$ is the rate of test return; and \testing{X} (Eq.
-\eqref{F}) and $p_X$ represent the \percap testing rate and the sensitivity (probability of positive tests), respectively, for compartment $X$.
-For further description of the parameters see \cref{tab:params}.
-Note that there is a slight mismatch in the top-to-bottom order of the testing-based compartments of each disease-based compartment $X$ between this flowchart and the model equations \eqref{model}; here we have switched  $X\_u$ and $X\_n$ for visual clarity.
-\label{fig:flowchart}}
-\end{center} 
-\end{figure}
-
-Table~\ref{tab:params} defines the model parameters, which are generally \percap flows between compartments, or modifiers to these flow rates.
-The novel component of the model lies in the compartment-specific relative testing weights $w_S$, $w_I$ and $w_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively.
-Thus, we can specify different levels of testing focus from random (all weights equal) to highly targeted (higher weights in more intensively tested compartments).
-For example, $w_I/w_S=3$ means that infected individuals are tested at three times the \percap rate of susceptible individuals.
-
-In order to allow parameterization of the model by the total (overall) \percap testing rate, we define the weighted size of the testing pool $W = w_S S\_u + w_I I\_u + w_R R\_u$, and calculate a scaling parameter for testing as:
-\begin{equation}
-\label{sigma}
-\sigma = \frac{\rho N}{W},
-\end{equation}
-where $\rho$ is the \percap testing intensity for the population, defined as the number of daily tests administered in a population of size $N$.
-Thus, the \percap testing rate for compartment $X \in \{S,I,R\}$ is 
-\begin{equation}
-\label{F}
-\testing{X}=\sigma w_X.
-\end{equation}
-For a highly sensitive test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus not considered for further testing.
-Over the course of the epidemic, a sufficiently large fixed testing rate as specified in \eqref{sigma} can exhaust the pool of people available for testing, leading to a singularity when too few people are left untested to support the specified rate.
-Although this phenomenon does not affect our analysis of $\Rnum$, it can affect model dynamics (we present an adjustment to the model that solves this problem in \appref{singularity}).
-
-The classical SIR model assumes a well-mixed population; homogeneity of the population (i.e., all individuals are equally susceptible and equally infectious with the same recovery rate when infected); exponentially distributed duration of infection; and large population size \citep{keeling2011modeling}.
-In addition to these standard assumptions, our model assumes:
-\begin{enumerate}[(i)]
-\item there is a single force of infection (new infections per unit time per susceptible), $\Lambda$, defined as
-  \begin{equation}
-  \label{Lambda}
-  \Lambda=\frac{\beta}{N} \big(I\_u + (1-\theta\_w)(I\_n+I\_p) + (1-\theta\_c)I\_c \big),
-  \end{equation}
-with transmission rate $\beta$; $\theta\_w$ is the isolation efficacy (reduction of the probability of transmission) for individuals waiting for test results, while $\theta\_c$ is the isolation efficacy for individuals who have received a ``confirmed'' positive test (Table~\ref{tab:params}).
-Susceptible individuals who are waiting for test results experience an additional transmission reduction factor of $1-\theta\_w$ (\fref{fig:flowchart}).
-\item confirmed-positive individuals isolate at least as effectively as those awaiting test results, i.e.,
-$$0\leq\theta\_w\leq\theta\_c\leq 1.$$ 
-\end{enumerate}
-For simplicity we assume that tests are perfectly \emph{specific} --- uninfected individuals never test positive ($p_s=0$).
-Thus there are no waiting-for-positive or confirmed-positive susceptible individuals, which reduces the number of model states from 12 to 10.
-
-\begin{table}[htp]
-\centering
-%%\begin{tabular}{|c|L{2in}cc|} \hline
-%% making use of ragged2e and array packages:
-{\RaggedRight
-\begin{tabular}{ c | m{6cm} | c | c }
-  \textbf{Symbol} & \textbf{Description} & \textbf{Unit} & \textbf{Value} \\ \hline
-  $N$     & Total population size & people & $10^6$ \\ \hline
-  $\omega$  & Rate of test return, i.e., rate of onward flow from ``waiting'' to ``confirmed'' or ``untested'' compartments  & 1/day & - \\ \hline
-  $\gamma$ & Recovery rate & 1/day & 1/6 \\ \hline 
-  $\rho$   & \percap testing intensity & 1/day & - \\ \hline 
-  $\theta\_w$ & Isolation efficacy (reduction of the transmission probability) for ``waiting'' individuals & - & - \\ \hline
-  $\theta\_c$ & Isolation efficacy for ``confirmed positive'' individuals & - & -  \\ \hline
-  $\beta$ & Transmission rate & 1/day & 0.5 \\ \hline
-  $\Lambda$ & Force of infection & 1/day & - \\ \hline
-  $p_S$ & Probability of positive tests for $S$ ($= 1-\textrm{specificity}$) & - & 0 \\ \hline
-  $p_I$ & Probability of positive tests for $I$ ($= \textrm{sensitivity}$) & - & 1 \\ \hline
-  $p_R$ & Probability of positive tests for $R$ ($= 1-\textrm{specificity}$) & - & 0.5 \\ \hline
-  $w_S, w_I, w_R$ & Relative testing weights & - &
-  \begin{minipage}[t]{0.21\columnwidth}%
-    Random:~$\{1,1,1\}$ \\ Targeted:~$\{0.3,1,1\}$
-  \end{minipage} \\
-  %%\hline
-\end{tabular}
-} % end \RaggedRight
-\caption{\label{tab:params}Parameters of the model specified  by the flow chart in \fref{fig:flowchart} and equations~\eqref{model}.}
-\end{table}
-
-The Disease-Free Equilibrium (DFE) for the expanded SIR model (Eqs.~\ref{model}) is found by setting the infected compartments to 0 and solving for the unknowns. The DFE is
-\begin{equation}
-\label{dfe}
-S\_n^*= \frac{\rho}{\omega} N, \ S\_u^*= N-S_n^*, \text{~and~} I_j=R_j=0 \ \text{for all }j.
-\end{equation}
-The corresponding \percap testing rate (Eq.~\ref{F}) for the infected compartment $I$ at DFE is one of the key analysis parameters and can be simplified as 
-\begin{equation}
-\label{eq:fi}
-\testinghat{I} = (\omega\rho/(\omega-\rho))w_I/w_S \quad .
-\end{equation}
-
-The basic reproduction number, $\Rnum$, was calculated by using the next-generation matrix method \citep{van2002reproduction}. We write $\Rnum$ as
-\begin{equation}
-\label{R0}
-\Rnum= \frac{\beta}{\gamma} \left(1-\Delta\right), 
-\end{equation}
-where $\beta/\gamma$ is the classical value for a simple model \citep{keeling2011modeling}, and $1-\Delta$ is the proportional reduction due to testing and isolation processes.
-$\Delta$ therefore measures the ``effectiveness of control'': how much these processes in reducing spread from low levels, and is in turn given by:
-\begin{equation}
-  \label{eq:del4a}
-  \Delta= \frac{1}{C N}\big(C_1 S\_u^*+(C_2(1-\theta\_w)+C \theta\_w)S\_n^*\big),
-\end{equation}
-where
-\begin{align}
-\label{eq:C12}
-C_{\phantom 1} &= (\omega+\gamma) \Big(\gamma(\omega+\gamma)+(\gamma+\omega p_I)\testinghat{I}\Big),\\
-C_1 &= (\omega+\gamma)(\theta\_w \gamma+\theta\_c \omega p_I) \testinghat{I},\\
-C_2 &= \Big( \omega\gamma(1+p_I)\testinghat{I}+\gamma^2(\omega+\gamma+\testinghat{I})\Big)\theta\_w + \omega^2p_I \testinghat{I} \theta\_c.
-\end{align}
-(\appref{R0} gives a detailed derivation of these expressions.)
-This explicit formula enables us to study the effects of testing and isolation parameters on $\Rnum$ both analytically and via numerical solutions.
-We are specifically interested in parameters that could be manipulated by public health policy: isolation efficacy, $\theta\_c$ and $\theta\_w$; \percap testing intensity, $\rho$; and the rate of test return, $\omega$.
-In particular, we look at the partial derivatives of $\Delta$ with respect to these parameters (\appref{rho}).
-We derived general expressions for these derivatives.
-However, we analyzed the effect of $\omega$ on $\Delta$ for the special case of low testing intensity.
-Specifically, by making the restriction $\rho \ll 1$, we are able to Taylor-expand $\Delta$ at $\rho=0$, use the linear approximation with respect to $\rho$ and analyze the resulting simplified derivatives to illustrate a surprising non-monotonic relationship between $\Delta$ and $\omega$.
-
-Analytic calculation of the next-generation matrix and simplification of the $\Rnum$ expression, were performed in Maple\textsuperscript{\texttrademark} \citep{maple14}; numerical calculation and contour plots were done in R \citep{r}.
-We computed the values and contours of $\Delta$ at both low (\fref{pan}) and high (\fref{pan2}) testing intensities, and for both random testing ($w_S=w_I=w_R=1$) and targeted testing ($w_S=0.3$; $w_I=w_R=1$).
-Because it is expressed as a proportion of $\Rnum$, the effectiveness of control $\Delta$ is (at least in the $\rho \ll 1$ case, Eq.~\ref{eq:lin}) independent of the transmission rate $\beta$, and hence of $\Rnum$ in the case where we vary $\Rnum$ by changing the transmission rate for a fixed generation interval.
-
-The low-testing case \fref{pan} reflects the case where testing intensity $\rho$ is small relative to the population size.
-Specifically, $\rho \in [0,0.013]$, and test return rate $\omega\in [1/12,2]$.
-This testing intensity is of the correct order of magnitude (although typically larger than) testing rates during the \covid pandemic, i.e., a maximum of 1.3\% of the population per day (approximately four times the maximum testing rate in Ontario, Canada in mid-2021).
-The less realistic high-testing case \fref{pan2} is included to highlight the occurrence of non-monotonic changes in $\Rnum$ with respect to $\rho$.
-In \fref{pan2} the maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,1/5)$ and the test return rate $\omega\in [1/5,2]$; these values are clearly unrealistic for a large population but might be relevant for small populations undergoing focused testing, such as a sports league or university.
-In these figures, the implied baseline reproduction number (for the SIR model without testing) is $\Rnum=\frac{\beta}{\gamma}=3$.
-The different ranges of test return rates $\omega$ for the cases of low and high testing intensities is due to the restriction $\rho<\omega$, which is a requirement for a feasible DFE (\ref{dfe}).
-
-% %%%%%%%
-\section{Results}
-
-We presented $\Rnum$ as the product of the classical reproduction number, $\beta/\gamma$, and the proportional reduction due to testing and isolation, $1-\Delta$, \eqref{R0}.
-We can use the formula for $\Delta$ \eqref{eq:del4a} to make a number of straightforward inferences about parameters that affect $\Rnum$ monotonically, i.e., for which the associated partial derivative of $\Delta$ always has the same sign (see Appendices).
-
-\begin{enumerate}
-
-\item \label{p1:eta} Increasing isolation efficacy for waiting ($\theta\_w$) and confirmed-positive ($\theta\_c$) individuals always increases $\Delta$ (Eqs.~\ref{eq:del4_theta},~\ref{eq:d_del4_thetac},~\ref{eq:d_del4_thetaw});
-\item \label{p1:rho} Higher testing intensity $\rho$ increases $\Delta$ if
-%% \begin{enumerate} \item
-testing is random (all $w_X$ equal) or testing intensity ($\rho$) is small (Eq.~\ref{eq:dd3dr}).
-%% \end{enumerate}
-\item \label{p1:omega} Increasing the rate of test return ($\omega$) always increases $\Delta$ if waiting individuals do not isolate ($\theta\_w=0$) (Eq.~\ref{eq:dlindo}).
-\item \label{p1:w} Increasing testing focus, i.e., changing the testing weights from random ($w_S=w_I$) toward targeted  ($w_S<w_I$), always increases $\Delta$ (Eq.~\ref{eq:d_del_wis}).
-\end{enumerate}
-
-\newpage
-\begin{figure}[h!]
-\centering
-\begin{subfigure}[t]{.49\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_random.pdf}
-\caption{}\label{p.a}
-\end{subfigure}
-%
-\begin{subfigure}[t]{.49\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_TTI.pdf}
-\caption{}\label{p.b}
-\end{subfigure}
-\caption{
-{\bf Effectiveness of testing and isolation in reducing $\Rnum$ at low \percap testing intensity ($\rho$).}
-Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4a}), over a range of testing and isolation parameters. Parameter values (Table~\ref{tab:params}):
-$\beta=\betaparam/$day, $1/\gamma= \invgammaparam$ days (baseline $\Rnum=\Rnumparam$, $r=\rparam$); $\omega \in [1/12,2]/$day;  $\rho \in [0,0.013]/$day per capita; $\theta\_w$ and $\theta\_c$ vary between 0 (no effect of isolation) and 1 (complete elimination of transmission); $p_S=0$, $p_I=1$ and $p_R=0.5$.
-Only parameter sets where $\theta\_c \geq \theta\_w$ (confirmed-positive individuals isolate more effectively than waiting individuals) are shown; the alternative case, $\theta\_w > \theta\_c$, is unrealistic.
-Contours of $\Delta$ are plotted for (a) random testing ($w_S=w_I=w_R=1$) and (b) targeted testing ($w_S=0.3$; $w_I=w_R=1$).
-}
-\label{pan}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\begin{subfigure}[t]{.49\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_random2.pdf}
-\caption{}
-\end{subfigure}
-%
-\begin{subfigure}[t]{.49\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_TTI2.pdf}
-\caption{}
-\end{subfigure}
-\caption{
-  {\bf Effectiveness of testing and isolation in reducing $\Rnum$ at high \percap testing intensity.}
-  Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4a}), over a range of testing and isolation parameters.
-Parameters as in \fref{pan} except: $\omega \in [1/5,2]/$day, $\rho \in [0,1/5)/$day.
-As in \fref{pan}, subplots show (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
-}
-\label{pan2}
-\end{figure}
-
-However, there are also two specific cases where $\Delta$ changes non-monotonically, in counterintuitive directions, as a function of testing and isolation parameters.
-
-\begin{itemize}
-\item We would generally expect increasing testing delays to increase \Rnum, thus decreasing effectiveness of control $\Delta$.
-This is in fact what happens when waiting individuals do not isolate ($\theta\_w =0$, top row of \fref{pan}) --- as we move to the right within each plot in this row, $\Delta$ decreases.
-However, when waiting individuals isolate ($\theta\_w>0$), we more often see the opposite effect: longer testing delays lead to a greater control effect $\Delta$ (reduced $\Rnum$).
-The reason is that people waiting for negative tests are assumed to continue to isolate; this applies both to susceptibles and to people who became infected while waiting for negative test results.
-This effect outweighs the effect of confirmed individuals isolating, except when this isolation parameter ($\theta\_c$) is substantially greater than $\theta\_w$.
-This result depends on the idea that, all else equal, people who have to wait longer for test results isolate at the same level (but for a longer time) as they would if the wait were shorter.
-\item \fref{pan} also shows that greater testing intensity (increasing $\rho$) generally increases the effectiveness of control (moving up in each panel).
-However, this relationship can be reversed at very high testing intensities (provided testing is targeted, and $\theta\_w$ is relatively small; \fref{pan2}(b), right three panels of top row).
-It is theoretically possible for increasing testing intensity to \emph{increase} \Rnum because more rapid testing leaves more susceptibles in the ``waiting-for-negative-results'' category at the DFE; if these people become infected while waiting, they will need to wait for their negative test result before they can be tested again, receive a positive test, and then begin self-isolating.
-This effect is usually weak compared to the beneficial effects of testing.
-
-\end{itemize}
-
-%%%%%%%%%%%%
-\section{Discussion}
-
-In this paper, we have developed and analyzed a simple compartmental model that combines epidemiological dynamics---as defined by a simple SIR model---with the dynamics of testing and isolation.
-Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness}).
-Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
-
-Many of the qualitative results we have derived confirm simple, common-sense intuitions.
-In particular, we can generally decrease $\Rnum$ by increasing isolation efficacy or testing intensity; returning tests faster, if individuals do not isolate while they are waiting for results; or increasing testing focus to target individuals who are likely to be infectious (e.g., symptomatic people or close contacts of known infections).
-
-However, we did find two surprising patterns: under some conditions longer delays in returning tests can reduce epidemic spread, and increasing testing rates can increase spread.
-
-Over broad regions of parameter space, decreasing $\omega$---i.e., \emph{slowing} the rate at which test results are returned---\emph{decreases} $\Rnum$ (for random testing, the parameter region is $\theta\_w \gtrsim 0.25$; for targeted testing, $\theta\_w \geq 0.25$ and either $\theta\_c \ge 0.5$ or $1/\omega > 5$; see Table~\ref{tab:params} for parameter definitions).
-This result is counterintuitive and would not be expected by public health authorities that have invested a great deal of effort in reducing delays from testing to results.
-Dynamically, this effect occurs because speeding up test returns shortens the isolation period of uninfected individuals (for infected people it only shortens the time to progression to the isolation level of the confirmed-positive compartment).
-Slowing test returns increases $\Rnum$ only if the proportion of infectives in the tested population is high and isolation is relatively strong among people waiting for test results.
-
-While slowing test returns does decrease $\Rnum$ over broad regions of parameter space in our model, there are several real-world processes missing from our model that make it unlikely that slowing test returns would actually be an effective public health measure.
-First, we do not model the primary benefit of rapid testing, i.e., detecting and containing outbreaks while they are still in progress.
-This process could be modeled phenomenologically by making the testing focus more targeted as an increasing proportion of cases is detected, because finding infections allows tests to be concentrated on their connections.
-Second, individuals may become less likely to maintain isolation if they are required to do so for longer; phenomenologically, we could allow  effectiveness of isolation in the waiting population to be an increasing function of test-return speed, or we could introduce a separate ``waiting, but no longer isolating'' compartment that individuals entered from the ``waiting, isolated'' compartment at a specified rate.
-Finally, if one wants to decrease the overall transmission rate of the population there are more effective methods than keeping tested people in limbo; these include masking, ventilation, distancing measures, retail and event closures, and stay-at-home orders.
-
-The other counterintuitive result from our analysis is that, for sufficiently high testing intensity $\rho$, further increasing testing intensity can actually \emph{increase} $\Rnum$ (e.g., \fref{pan2}(b), upper right panel [$\theta\_c=1$, $\theta\_w=0$]).
-This phenomenon can occur because we are considering the DFE in the presence of testing; thus there is an equilibrium distribution of susceptibles between the $S\_n$ (untested) and $S\_u$ (waiting) compartments even as the disease approaches extinction.
-A higher rate of testing leads to a greater proportion of individuals waiting for negative tests at the DFE.
-If infected, individuals in this group will take longer to be tested again and to subsequently isolate (because they must wait for their negative tests to be returned before being tested again).
-If isolation in this group ($\theta\_w$) is low, this effect can under some (relatively rare) circumstances (high $\theta\_c$, low $\omega$, high $\rho$) allow $\Rnum$ to increase with testing intensity.
-We can show that this phenomenon occurs \emph{only} under targeted testing ($w_I > w_S$), but we have not yet found a simple explanation of why it cannot occur under random (unfocused) testing.
-This phenomenon is also unlikely to occur in the real world.
-In particular, it depends on levels of testing that are unrealistically high (at least in large, general-population settings).
-
-Although we model the testing process in more detail than typical epidemiological models, one place where more detail could be informative is in the processes determining the testing weights $\{w_S, w_I, w_R\}$.
-While random testing, as done for surveillance purposes, unambiguously leads to equal testing weights, making precise quantitative connections between public-health practices and testing weights is difficult in other contexts.
-The testing weights reflect the correlation between an individual's risk of infection and their likelihood of being tested due to age, occupation, geographic location, etc..
-This correlation is influenced, among many other factors, by the proportion of the uninfected population with COVID-like symptoms (e.g., due to seasonal upper respiratory tract infections); the concentration of transmission and testing in hot spots such as long-term care facilities and high-density workplaces; the overall testing intensity (and hence, e.g., restriction to symptomatic individuals); and the proportion of COVID-infected people who are symptomatic.
-Future steps should explore mathematically tractable ways to model some of these factors more precisely.
-For example, separating the infected class into exposed, symptomatic, and a- or pre-symptomatic compartments and allowing the testing weights to vary across non-symptomatic (exposed/asymptomatic/presymptomatic) vs.\ symptomatic compartments could reflect the allocation of tests for diagnostic purposes (targeting symptomatic individuals) vs.\ contact-tracing (targeting infected but non-symptomatic individuals) vs.\ screening (relatively equal weights, depending on the venue).
-Alternatively, one could make the testing weights depend on the testing intensity or test-return rate as suggested above.
-Whatever complexity is added would probably put the model beyond reach of the analytical methods we have used in this paper, but one could still use semi-numerical methods such as constructing the next-generation matrix and using it to evaluate the derivatives of $\Rnum$ with respect to the parameters numerically.
-
-Although testing and tracing is a key part of infection control strategies, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts \new{\citep{endo2020implication,hellewell2020feasibility,jenness2020modeling}}, rather than analyzing simple but general models of the feedback between testing and transmission dynamics.
-\new{There have been several modeling studies of testing and tracing dynamics and their interaction with epidemiological dynamics.
-In the context of repeated screening and random testing of isolated populations (such as the members of a university), \cite{bergstrom2020frequency} provided analytical results quantifying the effects that proactive screening of asymptomatic individuals and isolation of confirmed-positive cases could have in reducing the spread of disease.
-\cite{rogers2021high} simulated a SEIR model with testing and isolation; they similarly suggest a strategy of rapid testing with antigen tests and the subsequent isolation of confirmed-positive individuals.
-\cite{friston2021testing} model the effects of self-isolation on testing and tracing with a focus on projections under different testing and tracing scenarios.
-They conclude that the emergence of a second wave depends almost primarily on the rate at which immunity is lost and that it is necessary to track asymptomatic individuals in order to control the outbreak.
-Our modeling approach differs from these previous efforts in that it examines the effects of test-return rates and of different levels of testing focus, from random to highly targeted.
-We hope this paper will inspire further explorations of the fundamental properties of epidemic models that incorporate explicit testing processes.}
-
-% %%%%%%%
-\bibliography{SIRlibrary}
-
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\clearpage
-% \widetext
-\appendix
-\section{Appendix}
-\setcounter{equation}{0}
-\setcounter{figure}{0}
-\setcounter{table}{0}
-% \setcounter{page}{1}
-\makeatletter
-\renewcommand{\theequation}{A\arabic{equation}}
-\renewcommand{\thefigure}{A\arabic{figure}}
-\renewcommand{\bibnumfmt}[1]{[A#1]}
-\renewcommand{\citenumfont}[1]{A#1}
-
-\subsection{Model and calculation of $\Rnum$}\label{app:R0}
-
-The model in the form of a system of ordinary differential equations is 
-\begin{subequations}\label{model}
-\begin{align}
- d S\_u/dt &= -\Lambda S\_u - \testing{S} S\_u + \omega S\_n, \\
- d S\_n/dt &= -(1-\theta\_w)\Lambda S\_n + (1-p_S) \testing{S} S\_u - \omega S\_n, \\
- d S\_p/dt &= -(1-\theta\_w)\Lambda S\_p + p_S \testing{S} S\_u - \omega S\_p, \\
- d S\_c/dt &= -(1-\theta\_c)\Lambda S\_c + \omega S\_p, \\
- d I\_u/dt &= \Lambda S\_u - \testing{I} I\_u + \omega I\_n  - \gamma I\_u, \\
- d I\_n/dt &= (1-\theta\_w)\Lambda S\_n + (1-p_I) \testing{I} I\_u - \omega I\_n -\gamma I\_n, \\
- d I\_p/dt &= (1-\theta\_w)\Lambda S\_p + p_I \testing{I} I\_u - \omega I\_p -\gamma I\_p, \\
- d I\_c/dt &= (1-\theta\_c)\Lambda S\_c + \omega I\_p - \gamma I\_c, \\
- d R\_u/dt &= \gamma I\_u - \testing{R} R\_u + \omega R\_n, \\
- d R\_n/dt &= \gamma I\_n + (1-p_R) \testing{R} R\_u - \omega R\_n, \\
- d R\_p/dt &= \gamma I\_p + p_R \testing{R} R\_u  - \omega R\_p, \\
- d R\_c/dt&= \gamma I\_c + \omega R\_p, \\
- dN/dt &= \omega (S\_n + I\_n + R\_n),  \\
- dP/dt &= \omega(I\_p + R\_p) ,
-\end{align}
-\end{subequations}
-%
-(see Table \ref{tab:params} for parameter definitions).
-The next generation matrix for this model is $G = F V^{-1}$, where matrix $F$ represents the inflow of new infection to the infected compartments and matrix $V$ represents the flow in the infected compartments when the population is totally susceptible.
-Matrices $F$ and $V$ are
-\begin{align}
-\label{F}
-F =& \beta/N \left[ \begin {array}{cccc} 
-S\_u^* & (1-\theta\_w) S\_u^* & (1-\theta\_w) S\_u^* & (1-\theta\_c) S\_u^*\\
-(1-\theta\_w)S\_n^* & (1-\theta\_w)^2 S\_n^* & (1-\theta\_w)^2 S\_n^* & (1-\theta\_w)(1-\theta\_c) S\_n^* \\ 
-0&0&0&0\\
-0&0&0&0
- \end {array} \right] \\ \notag
- =&\beta/N \left[ \begin {array}{c} S\_u^* \\ (1-\theta\_w)S\_n^* \\ 0 \\ 0 \end {array} \right]
-        \left[ \begin {array}{cccc} 1,   1-\theta\_w,   1-\theta\_w,   1-\theta\_c \end {array} \right],
-\end{align}
-and 
-\begin{align}
-\label{V}
-V= \left[ \begin {array}{cccc}  
-\testinghat{I}+\gamma&-\omega&0&0\\
--(1-p_I)\testinghat{I}&\omega+\gamma&0&0\\
--p_I \testinghat{I}&0&\omega+\gamma&0\\
-0&0&-\omega&\gamma
-\end {array} \right].
-\end{align}
-The matrix inverse of $V$ is 
-\begin{align}
-\label{vinv}
-V^{-1} =
-\frac{1}{\gamma C}
-\left[ \begin {array}{cccc}
-\gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)&0&0\\ \noalign{\medskip}
-\gamma(\omega+\gamma)(1-p_I) \testinghat{I} & \gamma(\omega+\gamma)(\testinghat{I}+\gamma)&0&0 \\ \noalign{\medskip}
-\gamma(\omega+\gamma) p_I \testinghat{I} & \gamma\omega p_I \testinghat{I} & C \gamma/(\omega+\gamma) & 0 \\ \noalign{\medskip}
-\omega(\omega+\gamma) p_I \testinghat{I} & \omega^2 p_I \testinghat{I} & C \omega/(\omega+\gamma) & C
-\end {array} \right],
-\end{align}
-where $C= \big( \gamma(\omega+\gamma)+(\gamma+\omega p_I)\testinghat{I} \big) (\omega+\gamma)$ and $\testinghat{I}$ is the \percap testing rate for the infected people and represented in Eq.~\eqref{eq:fi}.
-Note that all the columns of matrix $V^{-1}$ summ up to $1/\gamma$.
-
-The particular form of $F$ with two rows of zeros at the bottom results in the following blocked form of matrix $G$.
-\begin{equation}
-\label{mat:G}
-G = \left[ \begin {array}{cc}
-G_{11}&G_{12}\\
-0&0
-\end {array} \right],
-\end{equation}
-where both blocked matrices $G_{11}$ and $G_{12}$ are 2 by 2.
-Given the upper triangular form of matrix $G$, the basic reproduction number $\Rnum$ (defined as the spectral radius of matrix $G$) is only determined by the blocked matrix $G_{11}$,
-\begin{equation}
-\label{G11}
-G_{11} = \frac{\beta}{\gamma   C} 
-\left[ \begin {array}{c} (\omega-\rho)/\omega \\ (1-\theta\_w)\rho/\omega \end {array} \right]
-\left[ \begin {array}{cccc} 1,   1-\theta\_w,   1-\theta\_w,   1-\theta\_c \end {array} \right]
-\left[ \begin {array}{cc}
-\gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)\\ \noalign{\medskip}
-\gamma(\omega+\gamma)(1-p_I)\testinghat{I} & \gamma(\omega+\gamma)(\testinghat{I}+\gamma) \\ \noalign{\medskip}
-\gamma(\omega+\gamma) p_I \testinghat{I} & \gamma\omega p_I \testinghat{I} \\ \noalign{\medskip}
-\omega(\omega+\gamma) p_I \testinghat{I} & \omega^2 p_I \testinghat{I}
-\end {array} \right].
-\end{equation}
-It is notable that matrix $F$ \eqref{F} has rank one and consequently $G_{11}$ does so. That is $G_{11}$ has only one non-zero eigenvalue which is $\Rnum$.
-
-The expression of $\Rnum$ has a complicated form with all of the model parameters involved.
-This expression can be simplified and represented given the specific form of matrix $G_{11}$ \eqref{G11}.
-For the purpose of simplicity we present $\Rnum$ in the manuscript in terms of expressions $C$, $C_1$ and $C_2$, specified in \eqref{eq:C12}.
-
-It remains difficult to show that the reproduction number $\Rnum$ is decreasing with respect to \percap testing intensity, $\rho$, and the speed of the test return, $\omega$, for the feasible ranges of the parameters, that is
-\begin{align}
-\label{cond}
-& \omega>0, \\
-& 0 \leq \rho<\omega,\\ 
-& 0 \leq \theta\_w \leq \theta\_c \leq 1, \\
-& \testtarget\geq 1.
-\end{align}
-In realistic cases the testing rate $\rho$ is very small (i.e., only a small fraction of the population can be tested every day); it is thus reasonable to use a linear approximation of $\Rnum$ for $\rho \ll 1$ to analyze the behaviour of $\Rnum$ with respect to $\omega$ (see section \appref{omega}).
-In the next section we provide an equivalent representation of $\Rnum$ in order to show that increasing testing intensity typically decreases $\Rnum$.
-
-% %%%%%%%
-\subsection{More testing intensity may decrease $\Rnum$}\label{app:rho}
-
-This section shows that $\pro{\Delta}$ can be positive or negative, with $\Delta$ defined in Eq.~\eqref{eq:C12}, and thus $\pro{\Rnum} < 0$, where $\Rnum$ is given in Eq.~\eqref{R0}.
-We rewrite matrix $G_{11}$ in \eqref{G11} in the following form to simplify the calculations:
-\begin{equation}
-\label{G112}
-G_{11} = \frac{\beta}{\gamma C} 
-\left[ \begin {array}{c}  S\_u^*/N \\ (1-\theta\_w) S\_n^*/N  \end {array} \right]
-\left[ \begin {array}{cccc} 
-C-C_1, C-C_2\end {array} \right],
-\end{equation}
-where $C$ is the same as the one in Eq. \eqref{eq:C12}, i.e., 
-$$C=(\omega+\gamma)(\gamma(\omega+\gamma)+(\omega p_I+\gamma) \testinghat{I}),$$
-and $C_1$ and $C_2$ are 
-\begin{align*}
-% \label{eq:C12}
-C_1 =& (\omega+\gamma)(\theta\_w   \gamma+\theta\_c  \omega  p_I) \testinghat{I},\\
-C_2 =& \big( \omega\gamma(1+p_I)\testinghat{I}+\gamma^2 (\omega+\gamma+\testinghat{I})\big)\theta\_w + \omega^2 p_I \testinghat{I} \theta\_c,
-\end{align*}
-where $\testinghat{I}$ is given in Eq.~\eqref{eq:fi}.
-Note that for analysis brevity, we let $N=1$, thus $S\_u^*$ and $S\_n^*$ are in the scale of 0 to 1.
-$\Rnum$ is in the same form as in Eq. \eqref{R0}  
-$$\Rnum= \frac{\beta}{\gamma} (1-\Delta),$$
-where 
-\begin{equation*}
-\label{eq:del4}
-\Delta= \frac{1}{C}\big(C_1 S\_u^*+(C_2(1-\theta\_w)+C \theta\_w) S\_n^*\big).
-\end{equation*}
-
-The first goal
-is to explore how changes in isolation, $\theta\_w$ and $\theta\_c$, affects $\Rnum$.
-Mathematically we would like to verify the sign of $\pder \Rnum{\theta\_w}$ and $\pder \Rnum{\theta\_c}$.
-We start with simplifying $\Delta$ \eqref{eq:del4} by factoring $\theta\_w$ and $\theta\_c$ in Eq.~\eqref{eq:del4}.
-Thus, $\Delta$ can be rewritten as
-\begin{equation}
-\label{eq:del4_theta}
-\Delta= \frac{1}{C}\Big(
--D_1 S\_n^* \theta\_w^2 +
-\big(-\omega^2 p_I \testinghat{I} S\_n^*\theta\_c+ D_2 S\_n^*+\gamma \testinghat{I}(\omega+\gamma) \big)\theta\_w+
-(\omega+\gamma S\_u^*)\omega p_I \testinghat{I} \theta\_c
-\Big),
-\end{equation}
-where
-\begin{align}
-\label{eq:del4_d1d2}
-D_1=& (\omega+\gamma)\gamma^2+(\omega+\gamma+\omega p_I)\gamma \testinghat{I}, \\
-D_2=& (3\omega+2\gamma)\gamma^2+(\omega+\gamma+2\omega p_I)\gamma \testinghat{I}+(\gamma+p_I \testinghat{I})\omega^2.
-\end{align}
-$\Delta$, Eq.~\eqref{eq:del4_theta}, is linear in $\theta\_c$ with a positive coefficient. thus
-\begin{equation}
-\label{eq:d_del4_thetac}
-\pder\Delta{\theta\_c}=1/C(\gamma S\_u^*+\omega(1-\theta\_w S\_n^*))\omega p_I \testinghat{I}.
-\end{equation}
-This results in increasing $\Delta$, thus decreasing $\Rnum$ with respect to $\theta\_c$, that is $\pder \Rnum{\theta\_c}\leq 0$. Note that $C$ is independent of $\theta\_c$ and $\theta\_w$.
-
-With a similar logic, $\Delta$ \eqref{eq:del4_theta} is a concave-down quadratic equation in $\theta\_w$, given by
-\begin{equation}
-\label{eq:del4_thetaw}
-1/C \Big( -D_1 S\_n^*\theta\_w^2+
-\big(-\omega^2 p_I \testinghat{I} S\_n^*\theta\_c+D_2 S\_n^*+\gamma \testinghat{I}(\omega+\gamma) \big)\theta\_w\Big).
-\end{equation}
-We show that the feasible range of $\theta\_w$ lies between 0 and the vertex of this parabola where the parabola is increasing in $\theta\_w$, and so does $\Delta$ which results in inferring $\pder\Rnum{\theta\_w}\leq 0$.
-It is enough to show that partial derivative of the expression \eqref{eq:del4_thetaw} with respect to $\theta\_w$ at $\theta\_w=1$ is non-negative. It follows that
-\begin{align}
-\label{eq:d_del4_thetaw}
-\pder\Delta{\theta\_w}\bigg\rvert_{\theta\_w=1} =&
-1/C\Big( (D_2-2D_1-\omega^2 p_I \testinghat{I} \theta\_c) S\_n^*+\gamma \testinghat{I}(\omega+\gamma) \Big) \\\notag
-=& 1/C\Big( (\gamma(\omega+\gamma)+\gamma\omega^2+(1-\theta\_c)\omega^2 p_I \testinghat{I} ) S\_n^*
-+\gamma(\omega+\gamma) \testinghat{I} (1-S\_n^*) \Big),
-\end{align}
-which is a positive quantity, given that $\theta\_c$ and $S\_n^*$ vary between 0 and 1.
-
-{\bf The second goal} is to explore how changes in \percap testing intensity $\rho$ affects $\Rnum$.
-Mathematically we would like to verify the sign of $\pder\Rnum{\rho}$, which specifically depends on $\pder\Delta{\rho}$.
-We use the derived expressions for $S\_u^*$ and $S\_n^*$, given by Eqs.~\eqref{dfe}, in $\Delta$ \eqref{eq:del4}.
-Also, we define $\phi = \testinghat{S} = \frac{\rho \omega}{\omega-\rho}$, to reparameterize $\rho$.
-This is mainly to avoid singularity in $\testinghat{I}$ \eqref{eq:fi}, when testing intensity $\rho$ is very close to the rate of test return $\omega$.
-Thus, $\rho$ is reparameterized as
-\begin{equation}
-\label{eq:phi}
-\rho=\frac{\omega \phi}{\omega+\phi}.
-\end{equation}
-This one-to-one monotonic reparameterization enables us to simplify the mathematical expressions and explore the simpler $\pder\Delta{\phi}$ instead of the complicated $\pder\Delta{\rho}$.
-Defining $\testtarget \equiv \oldtesttarget$, the derivative is 
-\begin{align}
-\label{eq:dd3dr}
-\partial\Delta/\partial\phi=& \frac{1}{d_3} (a_3 \phi^2+b_3 \phi+c_3),
-\end{align}
-where
-\begin{align}
-\label{eq:abcd2}
-a_3& = \testtarget \Bigg(
-&& (1-\theta\_w) (1+\testtarget)\theta\_w \gamma^3 +(1-\theta\_c) p_I^2 \theta\_w \testtarget \omega^3 \\ \notag
-& ~ &&+\Big( \big( (1-\theta\_w-\testtarget) \theta\_c +(3-2\theta\_w) \theta\_w\testtarget \big) p_I +(1-\theta\_w) (1+\testtarget)\theta\_w  \Big) \omega \gamma^2 \\ \notag
-& ~ &&+\Big(
-\big((1-\theta\_w-\theta\_w \testtarget) \theta\_c+ (2-\theta\_w) \theta\_w \testtarget \big) p_I
-+(2 \theta\_w-\theta\_w^2-\theta\_c) \testtarget p_I^2
-\Big) \omega^2 \gamma \Bigg), \\ \notag
-b_3& = && \hspace{-1cm} 2\testtarget (\omega+\gamma)\gamma\Big(
-(\omega+\gamma+\omega p_I)(2-\theta\_w) \gamma \theta\_w+ (1-\theta\_w) \omega^2 p_I\theta\_c+\omega^2 p_I\theta\_w
-\Big) ,\\ \notag
-c_3& = && \hspace{-1cm} (\omega+\gamma)^2 \gamma\Big(
-(2-\theta\_w) \gamma^2 \theta\_w+
-(1+\testtarget) \omega \gamma \theta\_w+
-\testtarget \omega^2 p_I\theta\_c \Big),\\ \notag
-d_3&  = && \hspace{-1cm} \frac{(\omega+\gamma)}{\omega}  \Big((\omega p_I+\gamma)\testtarget\phi + (\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2.
-\end{align}
-Note that $\phi\geq 0$, also $b_3$, $c_3$ and $d_3$ are all positive. However $a_3$ can be positive or negative.
-If $a_3\geq 0$, $\partial\Delta/\partial\phi \geq 0$ for all feasible range of parameters, thus $\pro\Rnum \leq 0$.
-It is straightforward to show that $a_3\geq 0$ when testing is random, i.e., $w\_S=w\_I=1$.
-If $a_3 < 0$, then the quadratic expression in the numerator of \eqref{eq:dd3dr} has a positive root, $\phi^*$, such that for $\phi>\phi^*$, $\partial\Delta/\partial\phi < 0$.
-
-An example of this countervailing effect of $\phi$, and consequently $\rho$, on $\Rnum$ occurs when $\theta\_w=0$ and $\theta\_c=1$.
-This is illustrated in the top-right panel of the \fref{pan2} panel (b), where the strength of isolation for awaiting people is the least, but the most for the confirmed cases.
-In this case, simplifying $a_3$ in Eq.~\eqref{eq:abcd2} gives
-\begin{equation}
-  \begin{split}
-    a_3 & =\testtarget \omega \gamma p_I\left((\omega+\gamma)-\testtarget(\omega p_I+\gamma)\right)  \\
-    & \propto  \omega \left( 1- \testtarget p_I \right) + \gamma \left( 1- \testtarget \right)  .
-  \end{split}
-\end{equation}
-If $p_I>0$, then $a_3<0$ for sufficiently targeted testing (i.e. when $\testtarget p_I >1$; because
-$p_I \leq 1$, \testtarget is always $\geq \testtarget p_I$).
-%% this LOOKS like we can get a3<0 even if pI=0 for targeting > 1 and sufficiently large gamma/omega,
-%% but that seems crazy so I'm probably missing something
-When the test is perfectly sensitive ($p_I=1$), $a_3<0$ as long as $\testtarget>1$.
-Under either of these conditions, there exists a range for $\rho$ over which $\pder\Rnum{\rho}\leq 0$.  
-Because increasing values of $\rho$ and $\omega$ both delay the rate at which individuals flow to the $I\_c$ compartment, it is reasonable that increasing either value could (under appropriate circumstances) increase $\Rnum$.
-
-% %%%%%%%
-\subsection{Rate of returning tests} \label{app:omega}
-The third goal is to explore how changes in the rate of test return $\omega$ affect $\Rnum$.
-Mathematically we would like to verify the sign of $\pder\Rnum{\omega}$, which specifically depends on $\pder\Delta{\omega}$.
-We use
-the linearization of $\Delta$ when $\rho \ll 1$ to show that there a non-monotonic relationship between $\Rnum$  and $\omega$. The linear term in the Taylor expansion of $\Delta$ when $\rho \ll 1$ is
-\begin{equation}
-\label{eq:lin}
-\Delta = \frac{\rho}{\omega \gamma(\omega+\gamma)} \Big(
-\testtarget \omega^2 p_I \theta\_c+(\testtarget+1) \gamma\omega\theta\_w +\gamma^2 \theta\_w(2-\theta\_w)
-\Big). 
-\end{equation}
-This results in
-\begin{equation}
-\label{eq:dlindo}
-\pder\Delta{\omega} = \frac{\rho}{\omega^2 (\omega+\gamma)^2} \Big(
-(p_I \testtarget\theta\_c -(1+\testtarget)\theta\_w)\omega^2-2\theta\_w \gamma(2-\theta\_w) \omega+\theta\_w \gamma^2 (\theta\_w-2)
-\Big).
-\end{equation}
-% ALi & Faddy's stuff
-The latter expression has two roots
-\begin{align}
-\label{eq:omega_roots}
-\omega^*_{+} &= \frac{\gamma\Big(-c_4 + \sqrt{c_4^2 + (a_4-b_4)c_4}\Big)}{b_4-a_4} \\
-\omega^*_{-} &= \frac{\gamma\Big(-c_4 - \sqrt{c_4^2 + (a_4-b_4)c_4}\Big)}{b_4-a_4},
-\end{align}
-where $a_4 = p_I\testtarget\theta\_c$, $b_4 = (1+\testtarget)\theta\_w$, $c_4 = \theta\_w(2-\theta\_w)$.
-This enables us to describe the behaviour of $\Rnum$ with respect to $\omega$ in the following two cases.
-\begin{itemize}
-\item \textbf{Case I:} If $b_4 \geq a_4$, $\pder\Delta{\omega} < 0$, so $\Rnum$ is always increasing with respect to $\omega$ (i.e., it is always harmful to return tests more rapidly). 
-\item \textbf{Case II:} If $b_4 < a_4$, $\Rnum$ will be decreasing with respect to $\omega$ (i.e., returning tests more rapidly is beneficial) only when $\omega > \omega^*_-.$
-\end{itemize}
-Note that $b_4 \geq a_4$ is characterized by
-\begin{equation}\label{case_1_ineq}
-\Big(\testtarget+1\Big)\theta\_w \geq \testtarget p_I\theta\_c \Longleftrightarrow \Big(\frac{w_S}{w_I}+1\Big)\frac{1}{p_I} \geq \frac{\theta\_c}{\theta\_w}.
-\end{equation}
-We begin with a \textbf{proof of Case I}.
-Suppose that $b_4 \geq a_4$.
-If the roots of $\pder\Delta{\omega}$ are not complex, then we must have $c_4 \geq b_4-a_4$.
-Note that $\omega^*_{-}$ must be negative since the numerator is clearly negative but the denominator is positive.
-Next, note that since $c_4 \geq b_4-a_4$, the numerator of $\omega^*_{+}$ must also be negative, so $\omega^*_{+}$ is negative.
-Thus, in this case, $\pder\Rnum{\omega}$ does not change sign on $(0,\infty)$.
-Checking the sign of \ref{eq:dlindo} for arbitrarily large $\omega$ shows that it is negative (since the $(a_4-b_4)\omega^2$ term dominates and is negative).
-So $\Rnum$ is increasing with respect to $\omega$ on all of $(0,\infty)$.
-
-We now turn our attention to a \textbf{proof of Case II}.
-Suppose that $b_4 < a_4$.
-It follows that $\omega^*_+$ and $\omega^*_-$ are real since $c_4 > 0 > b_4-a_4$.
-Next, note that $\omega^*_-$ is positive since both the numerator and denominator are negative.
-On the other hand, $\omega^*_+$ is negative since the denominator is negative but the numerator is positive (because $c_4 > b_4-a_4$).
-Thus, our task is to understand the sign of $\pder\Delta{\omega}$ around the root $\omega^*_-$.
-Checking the sign of \ref{eq:dlindo} for arbitrarily large $\omega$ shows that it is positive (since the $(a_4-b_4)\omega^2$ term dominates and is positive).
-Likewise, checking the sign for values of $\omega$ close to $0$ shows that it is negative.
-Thus, $\Rnum$ is increasing with respect to $\omega$ when $\omega < \omega^*_-$, and is decreasing $\omega > \omega^*_-$.
-
-Having presented the formal analysis, we now concern ourselves with its biological interpretations.
-We begin by interpreting \ref{case_1_ineq}, under which returning tests more rapidly is \emph{always} harmful.
-Notice that the ratio $\frac{\theta\_c}{\theta\_w}$ is simply a measure of how much more strongly individuals self-isolate when they test positive compared to when waiting for tests.
-Since the rate of test return directly influences the rate at which individuals change from a waiting state to a confirmed-positive state, it is intuitive that $\frac{\theta\_c}{\theta\_w}$ would appear in \ref{case_1_ineq}.
-Next, note that the left-hand side increases when test sensitivity decreases and when targeting of positive individuals is poor.
-This is consistent with our intuition: a false negative that is returned more rapidly will allow an infectious individual to relax their self-isolation, thus increasing transmission.
-Likewise, if individuals tested are mainly susceptible (rather than infectious), then returning tests more slowly would encourage them to self-isolate for longer while awaiting test results.
-Having understood the role of each of the parameters in \ref{case_1_ineq}, a holistic interpretation of this inequality is that returning tests more slowly is helpful when the benefit of extended self-isolation by infected individuals awaiting test results outweighs the benefit of identifying positive cases.
-
-Now we interpret \textbf{Case II}.
-In this case, $\Rnum$ will have a global maximum with respect to $\omega$ at $\omega^*_-$.
-Note that our model assumption that $\rho < \omega$ plays an important role here: if $\omega^*_- < \rho$, then $\Rnum$ will be always decreasing with respect to $\omega$.
-On the other hand, if $\rho < \omega^*_-$, then $\Rnum$ will be increasing with respect to $\omega$ on $(\rho, \omega^*_-)$ and decreasing beyond that.
-
-% %%%%%%%
-\subsection{On Testing Rate and Numerical Singularity} \label{app:singularity}
-
-In this work, we do not present any numerical solutions of the ODEs to investigate typical trajectories.
-However, attempting to do so quickly reveals a problem in the way the model is posed;
-with the scaling of testing ($\sigma$) defined as in the body of the paper (Eq.~\eqref{sigma}), the population in the $S$ compartments appeared to blow up when the system is near the DFE.
-This occurs because once the only untested people are susceptibles, the FOI approaches $\Lambda=0$, and the testing rate $\testing{S} \to \rho N/S\_u$.
-Thus, the first equation of the model \eqref{model} will become
-$d S\_u/dt = - \rho N + \omega S\_n$. Thus changes in $S\_u$ will be independent on $S\_u$, and the decay of the $S\_u$
-population becomes linear rather than exponential --- allowing $S\_u$ to become negative!
-To avoid this problem the testing rate, $\sigma$, should be formulated such that people from the untested compartments will not be tested if they are not there.
-One way to fix this issue, is to consider a maximum testing rate, $\tau$ (1/day).
-In general, we want to test at a rate of $\rho$ across the whole population.
-This won't always be possible, so we impose a maximum rate of $\tau$ per testable person and redefine $\sigma = \frac{\tau \rho N}{\tau W + \rho N}$, with the assumption that $\tau \gg \rho$.
-This modification of $\sigma$ does not affect any of the results we have derived about the invasion of the epidemic from the DFE (i.e., results on $\Rnum$ and $\Delta$).
-
-% %%%%%%%
-\subsection{The effect of testing focus parameter $\testtarget$ on $\Rnum$} \label{app:w}
-
-\begin{align}
-\label{eq:d_del_wis}
-\pder \Delta{w\_{IS}}= \dfrac{(\omega-\rho) (\omega(\omega-\rho\theta\_w)+\gamma(\omega-\rho)) (\theta\_w\gamma+\theta\_c\omega p\_I)}
-{(-\omega^2 \gamma+\omega \gamma \rho-\gamma \rho \omega w\_{IS}-
-\omega \gamma^2+\gamma^2 \rho-\omega^2 p_I \rho w\_{IS})^2},
-\end{align}
-which is a positive quantity. Thus, $\pder \Rnum{w\_{IS}}\leq 0$. Therefore, increasing the focus of testing on the infectious people will result in less transmission.
-
-\end{document}

### 2021-07-19 appendix.tex
breaking the sections, filling the BMB template

--- 2021-07-16 SIR_manuscript.tex
@@
-\documentclass[12pt]{article}
-\usepackage{natbib}
-\usepackage{hyperref}
-\usepackage{grffile}
-\usepackage{graphicx}
-\usepackage{subcaption}
-\usepackage{amssymb,amsmath,amsthm}
-\usepackage{xcolor}
-\usepackage{xspace}
-\usepackage[nameinlink,capitalize]{cleveref}
-\usepackage{cleveref}
-\usepackage[margin=1in]{geometry}
-\usepackage{lineno}\renewcommand\thelinenumber{\color{gray}\arabic{linenumber}}
-\usepackage{pdflscape}
-\usepackage{enumerate}
-\usepackage{authblk}
-
-\hyphenation{co-ro-na-vi-rus}
-\input{codes/modeldefs.tex}
-
-% https://tex.stackexchange.com/questions/12703/how-to-create-fixed-width-table-columns-with-text-raggedright-centered-raggedlef
-\usepackage{array}
-\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-
-\usepackage{xspace}
-
-\newcommand{\fref}[1]{Fig.~\ref{#1}}
-\newcommand{\appref}[1]{Appendix~\ref{app:#1}}
-\newcommand{\Rlogo}{R\xspace}
-\newcommand{\percap}{\emph{per capita}\xspace}
-\newcommand{\Rnum}{\ensuremath{\mathcal{R}_0}\xspace}
-\newcommand{\covid}{COVID-19\xspace}
-\newcommand{\pro}[1]{\ensuremath{\frac{\partial #1}{\partial \rho}}}
-\newcommand{\pder}[2]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder y{x}
-\newcommand{\testing}[1]{\ensuremath{{\cal T}_{#1}}\xspace}
-\newcommand{\testinghat}[1]{\ensuremath{\hat{\cal T}_{#1}}\xspace}
-\newcommand{\testtarget}{\ensuremath{w\_{IS}}}
-\newcommand{\oldtesttarget}{\ensuremath{\frac{w_I}{w_S}}\xspace}
-\newcommand*\subtxt[1]{_{\textnormal{#1}}}
-%% Roman subscript
-\DeclareRobustCommand\_{\ifmmode\expandafter\subtxt\else\textunderscore\fi}
-
-\newcommand{\comment}{\showcomment}
-\newcommand{\showcomment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}
-\newcommand{\nocomment}[3]{}
-
-\newcommand{\fady}[1]{\comment{cyan}{Fady}{#1}}
-\newcommand{\ali}[1]{\comment{magenta}{Ali}{#1}}
-\newcommand{\jd}[1]{\comment{blue}{JD}{#1}}
-\newcommand{\david}[1]{\comment{orange}{DJDE}{#1}}
-\newcommand{\bmb}[1]{\comment{red}{BMB}{#1}}
-\newcommand{\todo}[1]{\comment{red}{TODO}{#1}}
-\newcommand{\new}[1]{\comment{red}{NEW}{#1}}
-\newcommand{\needref}{\textcolor{red}{(REFS?)}}
-
-%% DE: table stuff
-\renewcommand{\arraystretch}{1.5} % more interline spacing
-%% https://tex.stackexchange.com/questions/12703/how-to-create-fixed-width-table-columns-with-text-raggedright-centered-raggedlef
-\usepackage{ragged2e}
-\usepackage{array} % for m{...}
-
-\theoremstyle{definition} % amsthm only
-\newtheorem{proposition}{Proposition}
-\newtheorem{theorem}{Theorem}
-
-\bibliographystyle{apalike}
-
-\title{Testing and Isolation Efficacy:\\Insights from a Simple Epidemic Model}
-\author[1]{Ali Gharouni}
-\author[1]{F.M. Abdelmalek}
-\author[1,3,5]{David J. D. Earn}
-\author[2,4,3]{Jonathan Dushoff}
-\author[1,2,3]{Benjamin M. Bolker}
-
-\affil[1]{Department of Mathematics \& Statistics, McMaster University, Hamilton, Canada}
-\affil[2]{Department of Biology, McMaster University, Hamilton, Canada}
-\affil[3]{Michael G. DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, Canada}
-\affil[4]{South African Centre for Epidemiological Modelling and Analysis, University of Stellenbosch, Stellenbosch, South Africa}
-
-\begin{document}
-\maketitle
-
-\linenumbers
-
-% %%%%%%%
-\section{Abstract}
-
-Testing individuals for pathogens can affect the spread of epidemics. 
-Understanding how individual-level processes of sampling and reporting test results can affect community- or population-level spread is a dynamical modeling question.
-The effect of testing processes on epidemic dynamics depends on factors underlying implementation, particularly testing intensity and on whom testing is focused.
-Here, we use a simple model to explore how the individual-level effects of testing might directly impact population-level spread.
-Our model development was motivated by the \covid epidemic, but has generic epidemiological and testing structures. 
-To the classic SIR framework we have added a \percap testing intensity, and compartment-specific testing weights, which can be adjusted to reflect different testing emphases --- surveillance, diagnosis, or control.
-We derive an analytic expression for the relative reduction in the basic reproductive number due to testing, test-reporting and related isolation behaviours.
-Intensive testing and fast test reporting are expected to be beneficial at the community level because they can provide a rapid assessment of the situation, identify hot spots, and may enable rapid contact-tracing.
-Direct effects of fast testing at the individual level are less clear, and may depend on how individuals' behaviour is affected by testing information.
-Our simple model shows that under some circumstances both increased testing intensity and faster test reporting can \emph{reduce} the effectiveness of control, and allows us to explore the conditions under which this occurs.
-Conversely, we find that focusing testing on infected individuals always acts to increase effectiveness of control.
-
-% %%%%%%%
-\section{Introduction}
-
-The observed dynamics of the \covid epidemic have been driven both by epidemiological processes (infection and recovery) and by testing processes (testing and test reporting).
-In addition to shaping epidemic observations (via case reports), testing processes also alter epidemiological dynamics \citep{peto2020covid,taipale2020population}.
-Because individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may also do so, testing will generally increase the number of people who are isolating and hence reduce epidemic growth rates.
-We developed a mechanistic model that incorporates epidemic processes and testing in order to explore the effects of testing and isolation on epidemic dynamics.
-
-If testing influences behaviour, then epidemic dynamics will depend on who gets tested.
-The impacts of testing will depend both on testing intensity (tests performed per day) and on how strongly testing is focused on people who are infectious.
-This level of focus depends in turn on the purpose and design of testing programs. 
-When testing is done for the purposes of disease surveillance \citep{foddai2020base}
-tests are typically conducted randomly (or using a stratified random design) across the population in order to make an unbiased assessment of population prevalence.
-
-Over the course of the \covid pandemic, however, the vast majority of testing has been done with other goals --
-primarily diagnostic (determining infection status for clinical purposes) \citep{phua2020intensive,who2020global}, or for control (determining  infection status in order to isolate cases that have been found by contact tracing) \citep{aleta2020modelling,kucharski2020effectiveness,grassly2020comparison,smith2020adherence}, which we characterize as \emph{targeted} testing strategies.
-In these situations, testing probabilities can differ sharply across epidemiological compartments; in our dynamical model, we will characterize these probabilities by assigning a testing weight to each compartment that determines the \emph{relative} probability that an individual in that compartment will be selected for testing (see Methods).
-
-Diagnostic testing focuses on people with infection-like symptoms; thus the relative testing weights for infected people will depend on the relative probability of infected people having symptoms.
-For \covid infection, the testing weights will depend on the proportion of asymptomatic infections, the time spent pre-symptomatic vs.\ symptomatic during the course of an infection, and on the incidence of \covid-like symptoms among people in the population \emph{not} infected with \covid.
-Testing for epidemic control focuses on known contacts of infected people; in this case the testing weights for infected vs.\ uninfected people will depend on the probability of infection given contact, as well as the effectiveness of the system for identifying suspicious contacts.
-
-When a new infectious disease emerges, it is important to determine whether it will grow exponentially in a susceptible population, and if so at what rate $r$ \citep{Ma+14}.
-The condition for positive exponential growth ($r>0$) is commonly expressed as $\Rnum>1$, where the basic reproduction number $\Rnum$ is the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}.
-Although the value of $\Rnum$ cannot completely characterize the dynamics of our model \citep{shaw2021what}, it does give a simple and widely accepted index for the difficulty of control, as well as an indication of the likely final size of an epidemic \citep{ma2006generality,miller2012note}.
-
-In order to understand the effect of testing processes on epidemic dynamics, we expanded one of the simplest mechanistic epidemic models---the standard deterministic SIR model\citep{KermMcKe27,AndeMay91}---to include testing components.
-This model provides a sensible platform to link the modeling of epidemic and testing components and study their interaction.
-We studied the effects of testing intensity, rate of test return, and isolation efficacy, on transmission probability and epidemic dynamics when different levels of testing focus (from random to highly targeted) are in place.
-
-% %%%%%%%
-\section{Methods}
-
-Our model groups individuals based on disease status (Susceptible, Infectious or Recovered) and testing status (\emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}) (\fref{fig:flowchart}).
-The testing status of an individual in a given disease compartment $X$ (where $X \in \{S,I,R\}$) is denoted by a subscript, namely $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
-Two `accumulator' compartments, $N$ and $P$, are included in order to collect cumulative reported negative or positive tests.
-The model equations \eqref{model} and details of calculation of the basic reproduction number $\Rnum$ are presented in \appref{R0}.
-
-\begin{figure}
-\begin{center} 
-\includegraphics[scale=1.5]{pix/sir_comp.pdf}
-\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model, \ref{model}.
-The disease-based status of a compartment $X$ ($X \in \{S,I,R\}$) is combined with the testing status including $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
-The force of infection is denoted by $\Lambda$ (Eq.~\eqref{Lambda}); $\gamma$ is the recovery rate; $\omega$ is the rate of test return; and \testing{X} (Eq.
-\eqref{test}) and $p_X$ represent the \percap testing rate and the sensitivity (probability of positive tests), respectively, for compartment $X$.
-For further description of the parameters see \cref{tab:params}.
-Note that there is a slight mismatch in the top-to-bottom order of the testing-based compartments of each disease-based compartment $X$ between this flowchart and the model equations \eqref{model}; here we have switched  $X\_u$ and $X\_n$ for visual clarity.
-\label{fig:flowchart}}
-\end{center} 
-\end{figure}
-
-Table~\ref{tab:params} defines the model parameters, which are generally \percap flows between compartments, or modifiers to these flow rates.
-The novel component of the model lies in the compartment-specific relative testing weights $w_S$, $w_I$ and $w_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively.
-Thus, we can specify different levels of testing focus from random (all weights equal) to highly targeted (higher weights in more intensively tested compartments).
-For example, $w_I/w_S=3$ means that infected individuals are tested at three times the \percap rate of susceptible individuals.
-
-In order to allow parameterization of the model by the total (overall) \percap testing rate, we define the weighted size of the testing pool $W = w_S S\_u + w_I I\_u + w_R R\_u$, and calculate a scaling parameter for testing as:
-\begin{equation}
-\label{sigma}
-\sigma = \frac{\rho N}{W},
-\end{equation}
-where $\rho$ is the \percap testing intensity for the population, defined as the number of daily tests administered in a population of size $N$.
-Thus, the \percap testing rate for compartment $X \in \{S,I,R\}$ is 
-\begin{equation}
-\label{test}
-\testing{X}=\sigma w_X.
-\end{equation}
-For a highly sensitive test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus not considered for further testing.
-Over the course of the epidemic, a sufficiently large fixed testing rate as specified in \eqref{sigma} can exhaust the pool of people available for testing, leading to a singularity when too few people are left untested to support the specified rate.
-Although this phenomenon does not affect our analysis of $\Rnum$, it can affect model dynamics (we present an adjustment to the model that solves this problem in \appref{singularity}).
-
-The classical SIR model assumes a well-mixed population; homogeneity of the population (i.e., all individuals are equally susceptible and equally infectious with the same recovery rate when infected); exponentially distributed duration of infection; and large population size \citep{keeling2011modeling}.
-In addition to these standard assumptions, our model assumes:
-\begin{enumerate}[(i)]
-\item there is a single force of infection (new infections per unit time per susceptible), $\Lambda$, defined as
-  \begin{equation}
-  \label{Lambda}
-  \Lambda=\frac{\beta}{N} \big(I\_u + (1-\theta\_w)(I\_n+I\_p) + (1-\theta\_c)I\_c \big),
-  \end{equation}
-with transmission rate $\beta$; $\theta\_w$ is the isolation efficacy (reduction of the probability of transmission) for individuals waiting for test results, while $\theta\_c$ is the isolation efficacy for individuals who have received a ``confirmed'' positive test (Table~\ref{tab:params}).
-Susceptible individuals who are waiting for test results experience an additional transmission reduction factor of $1-\theta\_w$ (\fref{fig:flowchart}).
-\item confirmed-positive individuals isolate at least as effectively as those awaiting test results, i.e.,
-$$0\leq\theta\_w\leq\theta\_c\leq 1.$$ 
-\end{enumerate}
-For simplicity we assume that tests are perfectly \emph{specific} --- uninfected individuals never test positive ($p_s=0$).
-Thus there are no waiting-for-positive or confirmed-positive susceptible individuals, which reduces the number of model states from 12 to 10.
-
-\begin{table}[htp]
-\centering
-%%\begin{tabular}{|c|L{2in}cc|} \hline
-%% making use of ragged2e and array packages:
-{\RaggedRight
-\begin{tabular}{ c | m{6cm} | c | c }
-  \textbf{Symbol} & \textbf{Description} & \textbf{Unit} & \textbf{Value} \\ \hline
-  $N$     & Total population size & people & $10^6$ \\ \hline
-  $\omega$  & Rate of test return, i.e., rate of onward flow from ``waiting'' to ``confirmed'' or ``untested'' compartments  & 1/day & - \\ \hline
-  $\gamma$ & Recovery rate & 1/day & 1/6 \\ \hline 
-  $\rho$   & \percap testing intensity & 1/day & - \\ \hline 
-  $\theta\_w$ & Isolation efficacy (reduction of the transmission probability) for ``waiting'' individuals & - & - \\ \hline
-  $\theta\_c$ & Isolation efficacy for ``confirmed positive'' individuals & - & -  \\ \hline
-  $\beta$ & Transmission rate & 1/day & 0.5 \\ \hline
-  $\Lambda$ & Force of infection & 1/day & - \\ \hline
-  $p_S$ & Probability of positive tests for $S$ ($= 1-\textrm{specificity}$) & - & 0 \\ \hline
-  $p_I$ & Probability of positive tests for $I$ ($= \textrm{sensitivity}$) & - & 1 \\ \hline
-  $p_R$ & Probability of positive tests for $R$ ($= 1-\textrm{specificity}$) & - & 0.5 \\ \hline
-  $w_S, w_I, w_R$ & Relative testing weights & - &
-  \begin{minipage}[t]{0.21\columnwidth}%
-    Random:~$\{1,1,1\}$ \\ Targeted:~$\{0.3,1,1\}$
-  \end{minipage} \\
-  %%\hline
-\end{tabular}
-} % end \RaggedRight
-\caption{\label{tab:params}Parameters of the model specified  by the flow chart in \fref{fig:flowchart} and equations~\eqref{model}.}
-\end{table}
-
-The Disease-Free Equilibrium (DFE) for the expanded SIR model (Eqs.~\ref{model}) is found by setting the infected compartments to 0 and solving for the unknowns. The DFE is
-\begin{equation}
-\label{dfe}
-S\_n^*= \frac{\rho}{\omega} N, \ S\_u^*= N-S_n^*, \text{~and~} I_j=R_j=0 \ \text{for all }j.
-\end{equation}
-The corresponding \percap testing rate (Eq.~\ref{test}) for the infected compartment $I$ at DFE is one of the key analysis parameters and can be simplified as 
-\begin{equation}
-\label{eq:fi}
-\testinghat{I} = (\omega\rho/(\omega-\rho))w_I/w_S \quad .
-\end{equation}
-
-The basic reproduction number, $\Rnum$, was calculated by using the next-generation matrix method \citep{van2002reproduction}. We write $\Rnum$ as
-\begin{equation}
-\label{R0}
-\Rnum= \frac{\beta}{\gamma} \left(1-\Delta\right), 
-\end{equation}
-where $\beta/\gamma$ is the classical value for a simple model \citep{keeling2011modeling}, and $1-\Delta$ is the proportional reduction due to testing and isolation processes.
-$\Delta$ therefore measures the ``effectiveness of control'': how much these processes in reducing spread from low levels, and is in turn given by:
-\begin{equation}
-  \label{eq:del4a}
-  \Delta= \frac{1}{C N}\big(C_1 S\_u^*+(C_2(1-\theta\_w)+C \theta\_w)S\_n^*\big),
-\end{equation}
-where
-\begin{align}
-\label{eq:C12}
-C_{\phantom 1} &= (\omega+\gamma) \Big(\gamma(\omega+\gamma)+(\gamma+\omega p_I)\testinghat{I}\Big),\\
-C_1 &= (\omega+\gamma)(\theta\_w \gamma+\theta\_c \omega p_I) \testinghat{I},\\
-C_2 &= \Big( \omega\gamma(1+p_I)\testinghat{I}+\gamma^2(\omega+\gamma+\testinghat{I})\Big)\theta\_w + \omega^2p_I \testinghat{I} \theta\_c.
-\end{align}
-(\appref{R0} gives a detailed derivation of these expressions.)
-This explicit formula enables us to study the effects of testing and isolation parameters on $\Rnum$ both analytically and via numerical solutions.
-We are specifically interested in parameters that could be manipulated by public health policy: isolation efficacy, $\theta\_c$ and $\theta\_w$; \percap testing intensity, $\rho$; and the rate of test return, $\omega$.
-In particular, we look at the partial derivatives of $\Delta$ with respect to these parameters (Appendices~\ref{app:rho} and \ref{app:omega}).
-We derived general expressions for these derivatives.
-However, we analyzed the effect of $\omega$ on $\Delta$ for the special case of low testing intensity.
-Specifically, by making the restriction $\rho \ll 1$, we are able to Taylor-expand $\Delta$ at $\rho=0$, use the linear approximation with respect to $\rho$ and analyze the resulting simplified derivatives to illustrate a surprising non-monotonic relationship between $\Delta$ and $\omega$.
-
-Analytic calculation of the next-generation matrix and simplification of the $\Rnum$ expression, were performed in Maple\textsuperscript{\texttrademark} \citep{maple14}; numerical calculation and contour plots were done in R \citep{r}.
-We computed the values and contours of $\Delta$ at both low (\fref{pan}) and high (\fref{pan2}) testing intensities, and for both random testing ($w_S=w_I=w_R=1$) and targeted testing ($w_S=0.3$; $w_I=w_R=1$).
-Because it is expressed as a proportion of $\Rnum$, the effectiveness of control $\Delta$ is (at least in the $\rho \ll 1$ case, Eq.~\ref{eq:lin}) independent of the transmission rate $\beta$, and hence of $\Rnum$ in the case where we vary $\Rnum$ by changing the transmission rate for a fixed generation interval.
-
-The low-testing case (\fref{pan}) reflects the case where testing intensity $\rho$ is small relative to the population size.
-Specifically, $\rho \in [0,0.013]$, and test return rate $\omega\in [1/12,2]$.
-This testing intensity is of the correct order of magnitude (although typically larger than) testing rates during the \covid pandemic, i.e., a maximum of 1.3\% of the population per day (approximately four times the maximum testing rate in Ontario, Canada in mid-2021).
-The less realistic high-testing case (\fref{pan2}) is included to highlight the occurrence of non-monotonic changes in $\Rnum$ with respect to $\rho$.
-In \fref{pan2} the maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,1/5)$ and the test return rate $\omega\in [1/5,2]$; these values are clearly unrealistic for a large population but might be relevant for small populations undergoing focused testing, such as a sports league or university.
-In these figures, the implied baseline reproduction number (for the SIR model without testing) is $\Rnum=\frac{\beta}{\gamma}=3$.
-The different ranges of test return rates $\omega$ for the cases of low and high testing intensities is due to the restriction $\rho<\omega$, which is a requirement for a feasible DFE (\ref{dfe}).
-
-% %%%%%%%
-\section{Results}
-
-We presented $\Rnum$ as the product of the classical reproduction number, $\beta/\gamma$, and the proportional reduction due to testing and isolation, $1-\Delta$, \eqref{R0}.
-We can use the formula for $\Delta$ \eqref{eq:del4a} to make a number of straightforward inferences about parameters that affect $\Rnum$ monotonically, i.e., for which the associated partial derivative of $\Delta$ always has the same sign (see Appendices).
-
-\begin{enumerate}
-
-\item \label{p1:eta} Increasing isolation efficacy for waiting ($\theta\_w$) and confirmed-positive ($\theta\_c$) individuals always increases $\Delta$ (Eqs.~\ref{eq:del4_theta},~\ref{eq:d_del4_thetac},~\ref{eq:d_del4_thetaw});
-\item \label{p1:rho} Higher testing intensity $\rho$ increases $\Delta$ if
-%% \begin{enumerate} \item
-testing is random (all $w_X$ equal) or testing intensity ($\rho$) is small (Eq.~\ref{eq:dd3dr}).
-%% \end{enumerate}
-\item \label{p1:omega} Increasing the rate of test return ($\omega$) always increases $\Delta$ if waiting individuals do not isolate ($\theta\_w=0$) (Eq.~\ref{eq:dlindo}).
-\item \label{p1:w} Increasing testing focus, i.e., changing the testing weights from random ($w_S=w_I$) toward targeted  ($w_S<w_I$), always increases $\Delta$ (Eq.~\ref{eq:d_del_wis}).
-\end{enumerate}
-
-\newpage
-\begin{figure}[h!]
-\centering
-\begin{subfigure}[t]{.49\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_random.pdf}
-\caption{}\label{p.a}
-\end{subfigure}
-%
-\begin{subfigure}[t]{.49\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_TTI.pdf}
-\caption{}\label{p.b}
-\end{subfigure}
-\caption{
-{\bf Effectiveness of testing and isolation in reducing $\Rnum$ at low \percap testing intensity ($\rho$).}
-Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4a}), over a range of testing and isolation parameters. Parameter values (Table~\ref{tab:params}):
-$\beta=\betaparam/$day, $1/\gamma= \invgammaparam$ days (baseline $\Rnum=\Rnumparam$, $r=\rparam$); $\omega \in [1/12,2]/$day;  $\rho \in [0,0.013]/$day per capita; $\theta\_w$ and $\theta\_c$ vary between 0 (no effect of isolation) and 1 (complete elimination of transmission); $p_S=0$, $p_I=1$ and $p_R=0.5$.
-Only parameter sets where $\theta\_c \geq \theta\_w$ (confirmed-positive individuals isolate more effectively than waiting individuals) are shown; the alternative case, $\theta\_w > \theta\_c$, is unrealistic.
-Contours of $\Delta$ are plotted for (a) random testing ($w_S=w_I=w_R=1$) and (b) targeted testing ($w_S=0.3$; $w_I=w_R=1$).
-}
-\label{pan}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\begin{subfigure}[t]{.49\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_random2.pdf}
-\caption{}
-\end{subfigure}
-%
-\begin{subfigure}[t]{.49\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_TTI2.pdf}
-\caption{}
-\end{subfigure}
-\caption{
-  {\bf Effectiveness of testing and isolation in reducing $\Rnum$ at high \percap testing intensity.}
-  Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4a}), over a range of testing and isolation parameters.
-Parameters as in \fref{pan} except: $\omega \in [1/5,2]/$day, $\rho \in [0,1/5)/$day.
-As in \fref{pan}, subplots show (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
-}
-\label{pan2}
-\end{figure}
-
-However, there are also two specific cases where $\Delta$ changes non-monotonically, in counterintuitive directions, as a function of testing and isolation parameters.
-
-\begin{itemize}
-\item We would generally expect increasing testing delays to increase \Rnum, thus decreasing effectiveness of control $\Delta$.
-This is in fact what happens when waiting individuals do not isolate ($\theta\_w =0$, top row of \fref{pan}) --- as we move to the right within each plot in this row, $\Delta$ decreases.
-However, when waiting individuals isolate ($\theta\_w>0$), we more often see the opposite effect: longer testing delays lead to a greater control effect $\Delta$ (reduced $\Rnum$).
-The reason is that people waiting for negative tests are assumed to continue to isolate; this applies both to susceptibles and to people who became infected while waiting for negative test results.
-This effect outweighs the effect of confirmed individuals isolating, except when this isolation parameter ($\theta\_c$) is substantially greater than $\theta\_w$.
-This result depends on the idea that, all else equal, people who have to wait longer for test results isolate at the same level (but for a longer time) as they would if the wait were shorter.
-\item \fref{pan} also shows that greater testing intensity (increasing $\rho$) generally increases the effectiveness of control (moving up in each panel).
-However, this relationship can be reversed at very high testing intensities (provided testing is targeted, and $\theta\_w$ is relatively small; \fref{pan2}(b), right three panels of top row).
-It is theoretically possible for increasing testing intensity to \emph{increase} \Rnum because more rapid testing leaves more susceptibles in the ``waiting-for-negative-results'' category at the DFE; if these people become infected while waiting, they will need to wait for their negative test result before they can be tested again, receive a positive test, and then begin self-isolating.
-This effect is usually weak compared to the beneficial effects of testing.
-
-\end{itemize}
-
-%%%%%%%%%%%%
-\section{Discussion}
-
-In this paper, we have developed and analyzed a simple compartmental model that combines epidemiological dynamics---as defined by a simple SIR model---with the dynamics of testing and isolation.
-Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness}).
-Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
-
-Many of the qualitative results we have derived confirm simple, common-sense intuitions.
-In particular, we can generally decrease $\Rnum$ by increasing isolation efficacy or testing intensity; returning tests faster, if individuals do not isolate while they are waiting for results; or increasing testing focus to target individuals who are likely to be infectious (e.g., symptomatic people or close contacts of known infections).
-
-However, we did find two surprising patterns: under some conditions longer delays in returning tests can reduce epidemic spread, and increasing testing rates can increase spread.
-
-Over broad regions of parameter space, decreasing $\omega$---i.e., \emph{slowing} the rate at which test results are returned---\emph{decreases} $\Rnum$ (for random testing, the parameter region is $\theta\_w \gtrsim 0.25$; for targeted testing, $\theta\_w \geq 0.25$ and either $\theta\_c \ge 0.5$ or $1/\omega > 5$; see Table~\ref{tab:params} for parameter definitions).
-This result is counterintuitive and would not be expected by public health authorities that have invested a great deal of effort in reducing delays from testing to results.
-Dynamically, this effect occurs because speeding up test returns shortens the isolation period of uninfected individuals (for infected people it only shortens the time to progression to the isolation level of the confirmed-positive compartment).
-Slowing test returns increases $\Rnum$ only if the proportion of infectives in the tested population is high and isolation is relatively strong among people waiting for test results.
-
-While slowing test returns does decrease $\Rnum$ over broad regions of parameter space in our model, there are several real-world processes missing from our model that make it unlikely that slowing test returns would actually be an effective public health measure.
-First, we do not model the primary benefit of rapid testing, i.e., detecting and containing outbreaks while they are still in progress.
-This process could be modeled phenomenologically by making the testing focus more targeted as an increasing proportion of cases is detected, because finding infections allows tests to be concentrated on their connections.
-Second, individuals may become less likely to maintain isolation if they are required to do so for longer; phenomenologically, we could allow  effectiveness of isolation in the waiting population to be an increasing function of test-return speed, or we could introduce a separate ``waiting, but no longer isolating'' compartment that individuals entered from the ``waiting, isolated'' compartment at a specified rate.
-Finally, if one wants to decrease the overall transmission rate of the population there are more effective methods than keeping tested people in limbo; these include masking, ventilation, distancing measures, retail and event closures, and stay-at-home orders.
-
-The other counterintuitive result from our analysis is that, for sufficiently high testing intensity $\rho$, further increasing testing intensity can actually \emph{increase} $\Rnum$ (e.g., \fref{pan2}(b), upper right panel [$\theta\_c=1$, $\theta\_w=0$]).
-This phenomenon can occur because we are considering the DFE in the presence of testing; thus there is an equilibrium distribution of susceptibles between the $S\_n$ (untested) and $S\_u$ (waiting) compartments even as the disease approaches extinction.
-A higher rate of testing leads to a greater proportion of individuals waiting for negative tests at the DFE.
-If infected, individuals in this group will take longer to be tested again and to subsequently isolate (because they must wait for their negative tests to be returned before being tested again).
-If isolation in this group ($\theta\_w$) is low, this effect can under some (relatively rare) circumstances (high $\theta\_c$, low $\omega$, high $\rho$) allow $\Rnum$ to increase with testing intensity.
-We can show that this phenomenon occurs \emph{only} under targeted testing ($w_I > w_S$), but we have not yet found a simple explanation of why it cannot occur under random (unfocused) testing.
-This phenomenon is also unlikely to occur in the real world.
-In particular, it depends on levels of testing that are unrealistically high (at least in large, general-population settings).
-
-Although we model the testing process in more detail than typical epidemiological models, one place where more detail could be informative is in the processes determining the testing weights $\{w_S, w_I, w_R\}$.
-While random testing, as done for surveillance purposes, unambiguously leads to equal testing weights, making precise quantitative connections between public-health practices and testing weights is difficult in other contexts.
-The testing weights reflect the correlation between an individual's risk of infection and their likelihood of being tested due to age, occupation, geographic location, etc..
-This correlation is influenced, among many other factors, by the proportion of the uninfected population with COVID-like symptoms (e.g., due to seasonal upper respiratory tract infections); the concentration of transmission and testing in hot spots such as long-term care facilities and high-density workplaces; the overall testing intensity (and hence, e.g., restriction to symptomatic individuals); and the proportion of COVID-infected people who are symptomatic.
-Future steps should explore mathematically tractable ways to model some of these factors more precisely.
-For example, separating the infected class into exposed, symptomatic, and a- or pre-symptomatic compartments and allowing the testing weights to vary across non-symptomatic (exposed/asymptomatic/presymptomatic) vs.\ symptomatic compartments could reflect the allocation of tests for diagnostic purposes (targeting symptomatic individuals) vs.\ contact-tracing (targeting infected but non-symptomatic individuals) vs.\ screening (relatively equal weights, depending on the venue).
-Alternatively, one could make the testing weights depend on the testing intensity or test-return rate as suggested above.
-Whatever complexity is added would probably put the model beyond reach of the analytical methods we have used in this paper, but one could still use semi-numerical methods such as constructing the next-generation matrix and using it to evaluate the derivatives of $\Rnum$ with respect to the parameters numerically.
-
-Although testing and tracing is a key part of infection control strategies, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts \citep{endo2020implication,hellewell2020feasibility,jenness2020modeling}, rather than analyzing simple but general models of the feedback between testing and transmission dynamics.
-There have been several modeling studies of testing and tracing dynamics and their interaction with epidemiological dynamics.
-In the context of repeated screening and random testing of isolated populations (such as the members of a university), \cite{bergstrom2020frequency} provided analytical results quantifying the effects that proactive screening of asymptomatic individuals and isolation of confirmed-positive cases could have in reducing the spread of disease.
-\cite{rogers2021high} simulated a SEIR model with testing and isolation; they similarly suggest a strategy of rapid testing with antigen tests and the subsequent isolation of confirmed-positive individuals.
-\cite{friston2021testing} model the effects of self-isolation on testing and tracing with a focus on projections under different testing and tracing scenarios.
-They conclude that the emergence of a second wave depends almost primarily on the rate at which immunity is lost and that it is necessary to track asymptomatic individuals in order to control the outbreak.
-Our modeling approach differs from these previous efforts in that it examines the effects of test-return rates and of different levels of testing focus, from random to highly targeted.
-We hope this paper will inspire further explorations of the fundamental properties of epidemic models that incorporate explicit testing processes.
-
-% %%%%%%%
-\bibliography{SIRlibrary}
-
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\clearpage
-% \widetext
-\appendix
-\section{Appendix}
 \setcounter{equation}{0}
@@
 This modification of $\sigma$ does not affect any of the results we have derived about the invasion of the epidemic from the DFE (i.e., results on $\Rnum$ and $\Delta$).
-
-\end{document}

### 2021-07-19 discussion.tex
breaking the sections, filling the BMB template

--- 2021-07-16 SIR_manuscript.tex
@@
-\documentclass[12pt]{article}
-\usepackage{natbib}
-\usepackage{hyperref}
-\usepackage{grffile}
-\usepackage{graphicx}
-\usepackage{subcaption}
-\usepackage{amssymb,amsmath,amsthm}
-\usepackage{xcolor}
-\usepackage{xspace}
-\usepackage[nameinlink,capitalize]{cleveref}
-\usepackage{cleveref}
-\usepackage[margin=1in]{geometry}
-\usepackage{lineno}\renewcommand\thelinenumber{\color{gray}\arabic{linenumber}}
-\usepackage{pdflscape}
-\usepackage{enumerate}
-\usepackage{authblk}
-
-\hyphenation{co-ro-na-vi-rus}
-\input{codes/modeldefs.tex}
-
-% https://tex.stackexchange.com/questions/12703/how-to-create-fixed-width-table-columns-with-text-raggedright-centered-raggedlef
-\usepackage{array}
-\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-
-\usepackage{xspace}
-
-\newcommand{\fref}[1]{Fig.~\ref{#1}}
-\newcommand{\appref}[1]{Appendix~\ref{app:#1}}
-\newcommand{\Rlogo}{R\xspace}
-\newcommand{\percap}{\emph{per capita}\xspace}
-\newcommand{\Rnum}{\ensuremath{\mathcal{R}_0}\xspace}
-\newcommand{\covid}{COVID-19\xspace}
-\newcommand{\pro}[1]{\ensuremath{\frac{\partial #1}{\partial \rho}}}
-\newcommand{\pder}[2]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder y{x}
-\newcommand{\testing}[1]{\ensuremath{{\cal T}_{#1}}\xspace}
-\newcommand{\testinghat}[1]{\ensuremath{\hat{\cal T}_{#1}}\xspace}
-\newcommand{\testtarget}{\ensuremath{w\_{IS}}}
-\newcommand{\oldtesttarget}{\ensuremath{\frac{w_I}{w_S}}\xspace}
-\newcommand*\subtxt[1]{_{\textnormal{#1}}}
-%% Roman subscript
-\DeclareRobustCommand\_{\ifmmode\expandafter\subtxt\else\textunderscore\fi}
-
-\newcommand{\comment}{\showcomment}
-\newcommand{\showcomment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}
-\newcommand{\nocomment}[3]{}
-
-\newcommand{\fady}[1]{\comment{cyan}{Fady}{#1}}
-\newcommand{\ali}[1]{\comment{magenta}{Ali}{#1}}
-\newcommand{\jd}[1]{\comment{blue}{JD}{#1}}
-\newcommand{\david}[1]{\comment{orange}{DJDE}{#1}}
-\newcommand{\bmb}[1]{\comment{red}{BMB}{#1}}
-\newcommand{\todo}[1]{\comment{red}{TODO}{#1}}
-\newcommand{\new}[1]{\comment{red}{NEW}{#1}}
-\newcommand{\needref}{\textcolor{red}{(REFS?)}}
-
-%% DE: table stuff
-\renewcommand{\arraystretch}{1.5} % more interline spacing
-%% https://tex.stackexchange.com/questions/12703/how-to-create-fixed-width-table-columns-with-text-raggedright-centered-raggedlef
-\usepackage{ragged2e}
-\usepackage{array} % for m{...}
-
-\theoremstyle{definition} % amsthm only
-\newtheorem{proposition}{Proposition}
-\newtheorem{theorem}{Theorem}
-
-\bibliographystyle{apalike}
-
-\title{Testing and Isolation Efficacy:\\Insights from a Simple Epidemic Model}
-\author[1]{Ali Gharouni}
-\author[1]{F.M. Abdelmalek}
-\author[1,3,5]{David J. D. Earn}
-\author[2,4,3]{Jonathan Dushoff}
-\author[1,2,3]{Benjamin M. Bolker}
-
-\affil[1]{Department of Mathematics \& Statistics, McMaster University, Hamilton, Canada}
-\affil[2]{Department of Biology, McMaster University, Hamilton, Canada}
-\affil[3]{Michael G. DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, Canada}
-\affil[4]{South African Centre for Epidemiological Modelling and Analysis, University of Stellenbosch, Stellenbosch, South Africa}
-
-\begin{document}
-\maketitle
-
-\linenumbers
-
-% %%%%%%%
-\section{Abstract}
-
-Testing individuals for pathogens can affect the spread of epidemics. 
-Understanding how individual-level processes of sampling and reporting test results can affect community- or population-level spread is a dynamical modeling question.
-The effect of testing processes on epidemic dynamics depends on factors underlying implementation, particularly testing intensity and on whom testing is focused.
-Here, we use a simple model to explore how the individual-level effects of testing might directly impact population-level spread.
-Our model development was motivated by the \covid epidemic, but has generic epidemiological and testing structures. 
-To the classic SIR framework we have added a \percap testing intensity, and compartment-specific testing weights, which can be adjusted to reflect different testing emphases --- surveillance, diagnosis, or control.
-We derive an analytic expression for the relative reduction in the basic reproductive number due to testing, test-reporting and related isolation behaviours.
-Intensive testing and fast test reporting are expected to be beneficial at the community level because they can provide a rapid assessment of the situation, identify hot spots, and may enable rapid contact-tracing.
-Direct effects of fast testing at the individual level are less clear, and may depend on how individuals' behaviour is affected by testing information.
-Our simple model shows that under some circumstances both increased testing intensity and faster test reporting can \emph{reduce} the effectiveness of control, and allows us to explore the conditions under which this occurs.
-Conversely, we find that focusing testing on infected individuals always acts to increase effectiveness of control.
-
-% %%%%%%%
-\section{Introduction}
-
-The observed dynamics of the \covid epidemic have been driven both by epidemiological processes (infection and recovery) and by testing processes (testing and test reporting).
-In addition to shaping epidemic observations (via case reports), testing processes also alter epidemiological dynamics \citep{peto2020covid,taipale2020population}.
-Because individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may also do so, testing will generally increase the number of people who are isolating and hence reduce epidemic growth rates.
-We developed a mechanistic model that incorporates epidemic processes and testing in order to explore the effects of testing and isolation on epidemic dynamics.
-
-If testing influences behaviour, then epidemic dynamics will depend on who gets tested.
-The impacts of testing will depend both on testing intensity (tests performed per day) and on how strongly testing is focused on people who are infectious.
-This level of focus depends in turn on the purpose and design of testing programs. 
-When testing is done for the purposes of disease surveillance \citep{foddai2020base}
-tests are typically conducted randomly (or using a stratified random design) across the population in order to make an unbiased assessment of population prevalence.
-
-Over the course of the \covid pandemic, however, the vast majority of testing has been done with other goals --
-primarily diagnostic (determining infection status for clinical purposes) \citep{phua2020intensive,who2020global}, or for control (determining  infection status in order to isolate cases that have been found by contact tracing) \citep{aleta2020modelling,kucharski2020effectiveness,grassly2020comparison,smith2020adherence}, which we characterize as \emph{targeted} testing strategies.
-In these situations, testing probabilities can differ sharply across epidemiological compartments; in our dynamical model, we will characterize these probabilities by assigning a testing weight to each compartment that determines the \emph{relative} probability that an individual in that compartment will be selected for testing (see Methods).
-
-Diagnostic testing focuses on people with infection-like symptoms; thus the relative testing weights for infected people will depend on the relative probability of infected people having symptoms.
-For \covid infection, the testing weights will depend on the proportion of asymptomatic infections, the time spent pre-symptomatic vs.\ symptomatic during the course of an infection, and on the incidence of \covid-like symptoms among people in the population \emph{not} infected with \covid.
-Testing for epidemic control focuses on known contacts of infected people; in this case the testing weights for infected vs.\ uninfected people will depend on the probability of infection given contact, as well as the effectiveness of the system for identifying suspicious contacts.
-
-When a new infectious disease emerges, it is important to determine whether it will grow exponentially in a susceptible population, and if so at what rate $r$ \citep{Ma+14}.
-The condition for positive exponential growth ($r>0$) is commonly expressed as $\Rnum>1$, where the basic reproduction number $\Rnum$ is the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}.
-Although the value of $\Rnum$ cannot completely characterize the dynamics of our model \citep{shaw2021what}, it does give a simple and widely accepted index for the difficulty of control, as well as an indication of the likely final size of an epidemic \citep{ma2006generality,miller2012note}.
-
-In order to understand the effect of testing processes on epidemic dynamics, we expanded one of the simplest mechanistic epidemic models---the standard deterministic SIR model\citep{KermMcKe27,AndeMay91}---to include testing components.
-This model provides a sensible platform to link the modeling of epidemic and testing components and study their interaction.
-We studied the effects of testing intensity, rate of test return, and isolation efficacy, on transmission probability and epidemic dynamics when different levels of testing focus (from random to highly targeted) are in place.
-
-% %%%%%%%
-\section{Methods}
-
-Our model groups individuals based on disease status (Susceptible, Infectious or Recovered) and testing status (\emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}) (\fref{fig:flowchart}).
-The testing status of an individual in a given disease compartment $X$ (where $X \in \{S,I,R\}$) is denoted by a subscript, namely $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
-Two `accumulator' compartments, $N$ and $P$, are included in order to collect cumulative reported negative or positive tests.
-The model equations \eqref{model} and details of calculation of the basic reproduction number $\Rnum$ are presented in \appref{R0}.
-
-\begin{figure}
-\begin{center} 
-\includegraphics[scale=1.5]{pix/sir_comp.pdf}
-\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model, \ref{model}.
-The disease-based status of a compartment $X$ ($X \in \{S,I,R\}$) is combined with the testing status including $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
-The force of infection is denoted by $\Lambda$ (Eq.~\eqref{Lambda}); $\gamma$ is the recovery rate; $\omega$ is the rate of test return; and \testing{X} (Eq.
-\eqref{test}) and $p_X$ represent the \percap testing rate and the sensitivity (probability of positive tests), respectively, for compartment $X$.
-For further description of the parameters see \cref{tab:params}.
-Note that there is a slight mismatch in the top-to-bottom order of the testing-based compartments of each disease-based compartment $X$ between this flowchart and the model equations \eqref{model}; here we have switched  $X\_u$ and $X\_n$ for visual clarity.
-\label{fig:flowchart}}
-\end{center} 
-\end{figure}
-
-Table~\ref{tab:params} defines the model parameters, which are generally \percap flows between compartments, or modifiers to these flow rates.
-The novel component of the model lies in the compartment-specific relative testing weights $w_S$, $w_I$ and $w_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively.
-Thus, we can specify different levels of testing focus from random (all weights equal) to highly targeted (higher weights in more intensively tested compartments).
-For example, $w_I/w_S=3$ means that infected individuals are tested at three times the \percap rate of susceptible individuals.
-
-In order to allow parameterization of the model by the total (overall) \percap testing rate, we define the weighted size of the testing pool $W = w_S S\_u + w_I I\_u + w_R R\_u$, and calculate a scaling parameter for testing as:
-\begin{equation}
-\label{sigma}
-\sigma = \frac{\rho N}{W},
-\end{equation}
-where $\rho$ is the \percap testing intensity for the population, defined as the number of daily tests administered in a population of size $N$.
-Thus, the \percap testing rate for compartment $X \in \{S,I,R\}$ is 
-\begin{equation}
-\label{test}
-\testing{X}=\sigma w_X.
-\end{equation}
-For a highly sensitive test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus not considered for further testing.
-Over the course of the epidemic, a sufficiently large fixed testing rate as specified in \eqref{sigma} can exhaust the pool of people available for testing, leading to a singularity when too few people are left untested to support the specified rate.
-Although this phenomenon does not affect our analysis of $\Rnum$, it can affect model dynamics (we present an adjustment to the model that solves this problem in \appref{singularity}).
-
-The classical SIR model assumes a well-mixed population; homogeneity of the population (i.e., all individuals are equally susceptible and equally infectious with the same recovery rate when infected); exponentially distributed duration of infection; and large population size \citep{keeling2011modeling}.
-In addition to these standard assumptions, our model assumes:
-\begin{enumerate}[(i)]
-\item there is a single force of infection (new infections per unit time per susceptible), $\Lambda$, defined as
-  \begin{equation}
-  \label{Lambda}
-  \Lambda=\frac{\beta}{N} \big(I\_u + (1-\theta\_w)(I\_n+I\_p) + (1-\theta\_c)I\_c \big),
-  \end{equation}
-with transmission rate $\beta$; $\theta\_w$ is the isolation efficacy (reduction of the probability of transmission) for individuals waiting for test results, while $\theta\_c$ is the isolation efficacy for individuals who have received a ``confirmed'' positive test (Table~\ref{tab:params}).
-Susceptible individuals who are waiting for test results experience an additional transmission reduction factor of $1-\theta\_w$ (\fref{fig:flowchart}).
-\item confirmed-positive individuals isolate at least as effectively as those awaiting test results, i.e.,
-$$0\leq\theta\_w\leq\theta\_c\leq 1.$$ 
-\end{enumerate}
-For simplicity we assume that tests are perfectly \emph{specific} --- uninfected individuals never test positive ($p_s=0$).
-Thus there are no waiting-for-positive or confirmed-positive susceptible individuals, which reduces the number of model states from 12 to 10.
-
-\begin{table}[htp]
-\centering
-%%\begin{tabular}{|c|L{2in}cc|} \hline
-%% making use of ragged2e and array packages:
-{\RaggedRight
-\begin{tabular}{ c | m{6cm} | c | c }
-  \textbf{Symbol} & \textbf{Description} & \textbf{Unit} & \textbf{Value} \\ \hline
-  $N$     & Total population size & people & $10^6$ \\ \hline
-  $\omega$  & Rate of test return, i.e., rate of onward flow from ``waiting'' to ``confirmed'' or ``untested'' compartments  & 1/day & - \\ \hline
-  $\gamma$ & Recovery rate & 1/day & 1/6 \\ \hline 
-  $\rho$   & \percap testing intensity & 1/day & - \\ \hline 
-  $\theta\_w$ & Isolation efficacy (reduction of the transmission probability) for ``waiting'' individuals & - & - \\ \hline
-  $\theta\_c$ & Isolation efficacy for ``confirmed positive'' individuals & - & -  \\ \hline
-  $\beta$ & Transmission rate & 1/day & 0.5 \\ \hline
-  $\Lambda$ & Force of infection & 1/day & - \\ \hline
-  $p_S$ & Probability of positive tests for $S$ ($= 1-\textrm{specificity}$) & - & 0 \\ \hline
-  $p_I$ & Probability of positive tests for $I$ ($= \textrm{sensitivity}$) & - & 1 \\ \hline
-  $p_R$ & Probability of positive tests for $R$ ($= 1-\textrm{specificity}$) & - & 0.5 \\ \hline
-  $w_S, w_I, w_R$ & Relative testing weights & - &
-  \begin{minipage}[t]{0.21\columnwidth}%
-    Random:~$\{1,1,1\}$ \\ Targeted:~$\{0.3,1,1\}$
-  \end{minipage} \\
-  %%\hline
-\end{tabular}
-} % end \RaggedRight
-\caption{\label{tab:params}Parameters of the model specified  by the flow chart in \fref{fig:flowchart} and equations~\eqref{model}.}
-\end{table}
-
-The Disease-Free Equilibrium (DFE) for the expanded SIR model (Eqs.~\ref{model}) is found by setting the infected compartments to 0 and solving for the unknowns. The DFE is
-\begin{equation}
-\label{dfe}
-S\_n^*= \frac{\rho}{\omega} N, \ S\_u^*= N-S_n^*, \text{~and~} I_j=R_j=0 \ \text{for all }j.
-\end{equation}
-The corresponding \percap testing rate (Eq.~\ref{test}) for the infected compartment $I$ at DFE is one of the key analysis parameters and can be simplified as 
-\begin{equation}
-\label{eq:fi}
-\testinghat{I} = (\omega\rho/(\omega-\rho))w_I/w_S \quad .
-\end{equation}
-
-The basic reproduction number, $\Rnum$, was calculated by using the next-generation matrix method \citep{van2002reproduction}. We write $\Rnum$ as
-\begin{equation}
-\label{R0}
-\Rnum= \frac{\beta}{\gamma} \left(1-\Delta\right), 
-\end{equation}
-where $\beta/\gamma$ is the classical value for a simple model \citep{keeling2011modeling}, and $1-\Delta$ is the proportional reduction due to testing and isolation processes.
-$\Delta$ therefore measures the ``effectiveness of control'': how much these processes in reducing spread from low levels, and is in turn given by:
-\begin{equation}
-  \label{eq:del4a}
-  \Delta= \frac{1}{C N}\big(C_1 S\_u^*+(C_2(1-\theta\_w)+C \theta\_w)S\_n^*\big),
-\end{equation}
-where
-\begin{align}
-\label{eq:C12}
-C_{\phantom 1} &= (\omega+\gamma) \Big(\gamma(\omega+\gamma)+(\gamma+\omega p_I)\testinghat{I}\Big),\\
-C_1 &= (\omega+\gamma)(\theta\_w \gamma+\theta\_c \omega p_I) \testinghat{I},\\
-C_2 &= \Big( \omega\gamma(1+p_I)\testinghat{I}+\gamma^2(\omega+\gamma+\testinghat{I})\Big)\theta\_w + \omega^2p_I \testinghat{I} \theta\_c.
-\end{align}
-(\appref{R0} gives a detailed derivation of these expressions.)
-This explicit formula enables us to study the effects of testing and isolation parameters on $\Rnum$ both analytically and via numerical solutions.
-We are specifically interested in parameters that could be manipulated by public health policy: isolation efficacy, $\theta\_c$ and $\theta\_w$; \percap testing intensity, $\rho$; and the rate of test return, $\omega$.
-In particular, we look at the partial derivatives of $\Delta$ with respect to these parameters (Appendices~\ref{app:rho} and \ref{app:omega}).
-We derived general expressions for these derivatives.
-However, we analyzed the effect of $\omega$ on $\Delta$ for the special case of low testing intensity.
-Specifically, by making the restriction $\rho \ll 1$, we are able to Taylor-expand $\Delta$ at $\rho=0$, use the linear approximation with respect to $\rho$ and analyze the resulting simplified derivatives to illustrate a surprising non-monotonic relationship between $\Delta$ and $\omega$.
-
-Analytic calculation of the next-generation matrix and simplification of the $\Rnum$ expression, were performed in Maple\textsuperscript{\texttrademark} \citep{maple14}; numerical calculation and contour plots were done in R \citep{r}.
-We computed the values and contours of $\Delta$ at both low (\fref{pan}) and high (\fref{pan2}) testing intensities, and for both random testing ($w_S=w_I=w_R=1$) and targeted testing ($w_S=0.3$; $w_I=w_R=1$).
-Because it is expressed as a proportion of $\Rnum$, the effectiveness of control $\Delta$ is (at least in the $\rho \ll 1$ case, Eq.~\ref{eq:lin}) independent of the transmission rate $\beta$, and hence of $\Rnum$ in the case where we vary $\Rnum$ by changing the transmission rate for a fixed generation interval.
-
-The low-testing case (\fref{pan}) reflects the case where testing intensity $\rho$ is small relative to the population size.
-Specifically, $\rho \in [0,0.013]$, and test return rate $\omega\in [1/12,2]$.
-This testing intensity is of the correct order of magnitude (although typically larger than) testing rates during the \covid pandemic, i.e., a maximum of 1.3\% of the population per day (approximately four times the maximum testing rate in Ontario, Canada in mid-2021).
-The less realistic high-testing case (\fref{pan2}) is included to highlight the occurrence of non-monotonic changes in $\Rnum$ with respect to $\rho$.
-In \fref{pan2} the maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,1/5)$ and the test return rate $\omega\in [1/5,2]$; these values are clearly unrealistic for a large population but might be relevant for small populations undergoing focused testing, such as a sports league or university.
-In these figures, the implied baseline reproduction number (for the SIR model without testing) is $\Rnum=\frac{\beta}{\gamma}=3$.
-The different ranges of test return rates $\omega$ for the cases of low and high testing intensities is due to the restriction $\rho<\omega$, which is a requirement for a feasible DFE (\ref{dfe}).
-
-% %%%%%%%
-\section{Results}
-
-We presented $\Rnum$ as the product of the classical reproduction number, $\beta/\gamma$, and the proportional reduction due to testing and isolation, $1-\Delta$, \eqref{R0}.
-We can use the formula for $\Delta$ \eqref{eq:del4a} to make a number of straightforward inferences about parameters that affect $\Rnum$ monotonically, i.e., for which the associated partial derivative of $\Delta$ always has the same sign (see Appendices).
-
-\begin{enumerate}
-
-\item \label{p1:eta} Increasing isolation efficacy for waiting ($\theta\_w$) and confirmed-positive ($\theta\_c$) individuals always increases $\Delta$ (Eqs.~\ref{eq:del4_theta},~\ref{eq:d_del4_thetac},~\ref{eq:d_del4_thetaw});
-\item \label{p1:rho} Higher testing intensity $\rho$ increases $\Delta$ if
-%% \begin{enumerate} \item
-testing is random (all $w_X$ equal) or testing intensity ($\rho$) is small (Eq.~\ref{eq:dd3dr}).
-%% \end{enumerate}
-\item \label{p1:omega} Increasing the rate of test return ($\omega$) always increases $\Delta$ if waiting individuals do not isolate ($\theta\_w=0$) (Eq.~\ref{eq:dlindo}).
-\item \label{p1:w} Increasing testing focus, i.e., changing the testing weights from random ($w_S=w_I$) toward targeted  ($w_S<w_I$), always increases $\Delta$ (Eq.~\ref{eq:d_del_wis}).
-\end{enumerate}
-
-\newpage
-\begin{figure}[h!]
-\centering
-\begin{subfigure}[t]{.49\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_random.pdf}
-\caption{}\label{p.a}
-\end{subfigure}
-%
-\begin{subfigure}[t]{.49\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_TTI.pdf}
-\caption{}\label{p.b}
-\end{subfigure}
-\caption{
-{\bf Effectiveness of testing and isolation in reducing $\Rnum$ at low \percap testing intensity ($\rho$).}
-Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4a}), over a range of testing and isolation parameters. Parameter values (Table~\ref{tab:params}):
-$\beta=\betaparam/$day, $1/\gamma= \invgammaparam$ days (baseline $\Rnum=\Rnumparam$, $r=\rparam$); $\omega \in [1/12,2]/$day;  $\rho \in [0,0.013]/$day per capita; $\theta\_w$ and $\theta\_c$ vary between 0 (no effect of isolation) and 1 (complete elimination of transmission); $p_S=0$, $p_I=1$ and $p_R=0.5$.
-Only parameter sets where $\theta\_c \geq \theta\_w$ (confirmed-positive individuals isolate more effectively than waiting individuals) are shown; the alternative case, $\theta\_w > \theta\_c$, is unrealistic.
-Contours of $\Delta$ are plotted for (a) random testing ($w_S=w_I=w_R=1$) and (b) targeted testing ($w_S=0.3$; $w_I=w_R=1$).
-}
-\label{pan}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\begin{subfigure}[t]{.49\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_random2.pdf}
-\caption{}
-\end{subfigure}
-%
-\begin{subfigure}[t]{.49\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_TTI2.pdf}
-\caption{}
-\end{subfigure}
-\caption{
-  {\bf Effectiveness of testing and isolation in reducing $\Rnum$ at high \percap testing intensity.}
-  Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4a}), over a range of testing and isolation parameters.
-Parameters as in \fref{pan} except: $\omega \in [1/5,2]/$day, $\rho \in [0,1/5)/$day.
-As in \fref{pan}, subplots show (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
-}
-\label{pan2}
-\end{figure}
-
-However, there are also two specific cases where $\Delta$ changes non-monotonically, in counterintuitive directions, as a function of testing and isolation parameters.
-
-\begin{itemize}
-\item We would generally expect increasing testing delays to increase \Rnum, thus decreasing effectiveness of control $\Delta$.
-This is in fact what happens when waiting individuals do not isolate ($\theta\_w =0$, top row of \fref{pan}) --- as we move to the right within each plot in this row, $\Delta$ decreases.
-However, when waiting individuals isolate ($\theta\_w>0$), we more often see the opposite effect: longer testing delays lead to a greater control effect $\Delta$ (reduced $\Rnum$).
-The reason is that people waiting for negative tests are assumed to continue to isolate; this applies both to susceptibles and to people who became infected while waiting for negative test results.
-This effect outweighs the effect of confirmed individuals isolating, except when this isolation parameter ($\theta\_c$) is substantially greater than $\theta\_w$.
-This result depends on the idea that, all else equal, people who have to wait longer for test results isolate at the same level (but for a longer time) as they would if the wait were shorter.
-\item \fref{pan} also shows that greater testing intensity (increasing $\rho$) generally increases the effectiveness of control (moving up in each panel).
-However, this relationship can be reversed at very high testing intensities (provided testing is targeted, and $\theta\_w$ is relatively small; \fref{pan2}(b), right three panels of top row).
-It is theoretically possible for increasing testing intensity to \emph{increase} \Rnum because more rapid testing leaves more susceptibles in the ``waiting-for-negative-results'' category at the DFE; if these people become infected while waiting, they will need to wait for their negative test result before they can be tested again, receive a positive test, and then begin self-isolating.
-This effect is usually weak compared to the beneficial effects of testing.
-
-\end{itemize}
-
-%%%%%%%%%%%%
-\section{Discussion}
+% \section{Discussion}
 
@@
 We hope this paper will inspire further explorations of the fundamental properties of epidemic models that incorporate explicit testing processes.
-
-% %%%%%%%
-\bibliography{SIRlibrary}
-
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\clearpage
-% \widetext
-\appendix
-\section{Appendix}
-\setcounter{equation}{0}
-\setcounter{figure}{0}
-\setcounter{table}{0}
-% \setcounter{page}{1}
-\makeatletter
-\renewcommand{\theequation}{A\arabic{equation}}
-\renewcommand{\thefigure}{A\arabic{figure}}
-\renewcommand{\bibnumfmt}[1]{[A#1]}
-\renewcommand{\citenumfont}[1]{A#1}
-
-\subsection{Model and calculation of $\Rnum$}\label{app:R0}
-
-The model in the form of a system of ordinary differential equations is 
-\begin{subequations}\label{model}
-\begin{align}
- d S\_u/dt &= -\Lambda S\_u - \testing{S} S\_u + \omega S\_n, \\
- d S\_n/dt &= -(1-\theta\_w)\Lambda S\_n + (1-p_S) \testing{S} S\_u - \omega S\_n, \\
- d S\_p/dt &= -(1-\theta\_w)\Lambda S\_p + p_S \testing{S} S\_u - \omega S\_p, \\
- d S\_c/dt &= -(1-\theta\_c)\Lambda S\_c + \omega S\_p, \\
- d I\_u/dt &= \Lambda S\_u - \testing{I} I\_u + \omega I\_n  - \gamma I\_u, \\
- d I\_n/dt &= (1-\theta\_w)\Lambda S\_n + (1-p_I) \testing{I} I\_u - \omega I\_n -\gamma I\_n, \\
- d I\_p/dt &= (1-\theta\_w)\Lambda S\_p + p_I \testing{I} I\_u - \omega I\_p -\gamma I\_p, \\
- d I\_c/dt &= (1-\theta\_c)\Lambda S\_c + \omega I\_p - \gamma I\_c, \\
- d R\_u/dt &= \gamma I\_u - \testing{R} R\_u + \omega R\_n, \\
- d R\_n/dt &= \gamma I\_n + (1-p_R) \testing{R} R\_u - \omega R\_n, \\
- d R\_p/dt &= \gamma I\_p + p_R \testing{R} R\_u  - \omega R\_p, \\
- d R\_c/dt&= \gamma I\_c + \omega R\_p, \\
- dN/dt &= \omega (S\_n + I\_n + R\_n),  \\
- dP/dt &= \omega(I\_p + R\_p) ,
-\end{align}
-\end{subequations}
-%
-(see Table \ref{tab:params} for parameter definitions).
-The next generation matrix for this model is $G = F V^{-1}$, where matrix $F$ represents the inflow of new infection to the infected compartments and matrix $V$ represents the flow in the infected compartments when the population is totally susceptible.
-Matrices $F$ and $V$ are
-\begin{align}
-\label{F}
-F =& \beta/N \left[ \begin {array}{cccc} 
-S\_u^* & (1-\theta\_w) S\_u^* & (1-\theta\_w) S\_u^* & (1-\theta\_c) S\_u^*\\
-(1-\theta\_w)S\_n^* & (1-\theta\_w)^2 S\_n^* & (1-\theta\_w)^2 S\_n^* & (1-\theta\_w)(1-\theta\_c) S\_n^* \\ 
-0&0&0&0\\
-0&0&0&0
- \end {array} \right] \\ \notag
- =&\beta/N \left[ \begin {array}{c} S\_u^* \\ (1-\theta\_w)S\_n^* \\ 0 \\ 0 \end {array} \right]
-        \left[ \begin {array}{cccc} 1,   1-\theta\_w,   1-\theta\_w,   1-\theta\_c \end {array} \right],
-\end{align}
-and 
-\begin{align}
-\label{V}
-V= \left[ \begin {array}{cccc}  
-\testinghat{I}+\gamma&-\omega&0&0\\
--(1-p_I)\testinghat{I}&\omega+\gamma&0&0\\
--p_I \testinghat{I}&0&\omega+\gamma&0\\
-0&0&-\omega&\gamma
-\end {array} \right].
-\end{align}
-The matrix inverse of $V$ is 
-\begin{align}
-\label{vinv}
-V^{-1} =
-\frac{1}{\gamma C}
-\left[ \begin {array}{cccc}
-\gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)&0&0\\ \noalign{\medskip}
-\gamma(\omega+\gamma)(1-p_I) \testinghat{I} & \gamma(\omega+\gamma)(\testinghat{I}+\gamma)&0&0 \\ \noalign{\medskip}
-\gamma(\omega+\gamma) p_I \testinghat{I} & \gamma\omega p_I \testinghat{I} & C \gamma/(\omega+\gamma) & 0 \\ \noalign{\medskip}
-\omega(\omega+\gamma) p_I \testinghat{I} & \omega^2 p_I \testinghat{I} & C \omega/(\omega+\gamma) & C
-\end {array} \right],
-\end{align}
-where $C= \big( \gamma(\omega+\gamma)+(\gamma+\omega p_I)\testinghat{I} \big) (\omega+\gamma)$ and $\testinghat{I}$ is the \percap testing rate for the infected people and represented in Eq.~\eqref{eq:fi}.
-Note that all the columns of matrix $V^{-1}$ summ up to $1/\gamma$.
-
-The particular form of $F$ with two rows of zeros at the bottom results in the following blocked form of matrix $G$.
-\begin{equation}
-\label{mat:G}
-G = \left[ \begin {array}{cc}
-G_{11}&G_{12}\\
-0&0
-\end {array} \right],
-\end{equation}
-where both blocked matrices $G_{11}$ and $G_{12}$ are 2 by 2.
-Given the upper triangular form of matrix $G$, the basic reproduction number $\Rnum$ (defined as the spectral radius of matrix $G$) is only determined by the blocked matrix $G_{11}$,
-\begin{equation}
-\label{G11}
-G_{11} = \frac{\beta}{\gamma   C} 
-\left[ \begin {array}{c} (\omega-\rho)/\omega \\ (1-\theta\_w)\rho/\omega \end {array} \right]
-\left[ \begin {array}{cccc} 1,   1-\theta\_w,   1-\theta\_w,   1-\theta\_c \end {array} \right]
-\left[ \begin {array}{cc}
-\gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)\\ \noalign{\medskip}
-\gamma(\omega+\gamma)(1-p_I)\testinghat{I} & \gamma(\omega+\gamma)(\testinghat{I}+\gamma) \\ \noalign{\medskip}
-\gamma(\omega+\gamma) p_I \testinghat{I} & \gamma\omega p_I \testinghat{I} \\ \noalign{\medskip}
-\omega(\omega+\gamma) p_I \testinghat{I} & \omega^2 p_I \testinghat{I}
-\end {array} \right].
-\end{equation}
-It is notable that matrix $F$ \eqref{F} has rank one and consequently $G_{11}$ does so. That is $G_{11}$ has only one non-zero eigenvalue which is $\Rnum$.
-
-The expression of $\Rnum$ has a complicated form with all of the model parameters involved.
-This expression can be simplified and represented given the specific form of matrix $G_{11}$ \eqref{G11}.
-For the purpose of simplicity we present $\Rnum$ in the manuscript in terms of expressions $C$, $C_1$ and $C_2$, specified in \eqref{eq:C12}.
-
-It remains difficult to show that the reproduction number $\Rnum$ is decreasing with respect to \percap testing intensity, $\rho$, and the speed of the test return, $\omega$, for the feasible ranges of the parameters, that is
-\begin{align}
-\label{cond}
-& \omega>0, \\
-& 0 \leq \rho<\omega,\\ 
-& 0 \leq \theta\_w \leq \theta\_c \leq 1, \\
-& \oldtesttarget\geq 1.
-\end{align}
-In realistic cases the testing rate $\rho$ is very small (i.e., only a small fraction of the population can be tested every day); it is thus reasonable to use a linear approximation of $\Rnum$ for $\rho \ll 1$ to analyze the behaviour of $\Rnum$ with respect to $\omega$ (see section \appref{omega}).
-In the next section we provide an equivalent representation of $\Rnum$ in order to show that increasing testing intensity typically decreases $\Rnum$.
-
-% %%%%%%%
-\subsection{More testing intensity may decrease $\Rnum$}\label{app:rho}
-
-This section shows that $\pro{\Delta}$ can be positive or negative, with $\Delta$ defined in Eq.~\eqref{eq:C12}, and thus $\pro{\Rnum} < 0$, where $\Rnum$ is given in Eq.~\eqref{R0}.
-We rewrite matrix $G_{11}$ in \eqref{G11} in the following form to simplify the calculations:
-\begin{equation}
-\label{G112}
-G_{11} = \frac{\beta}{\gamma C} 
-\left[ \begin {array}{c}  S\_u^*/N \\ (1-\theta\_w) S\_n^*/N  \end {array} \right]
-\left[ \begin {array}{cccc} 
-C-C_1, C-C_2\end {array} \right],
-\end{equation}
-where $C$ is the same as the one in Eq. \eqref{eq:C12}, i.e., 
-$$C=(\omega+\gamma)(\gamma(\omega+\gamma)+(\omega p_I+\gamma) \testinghat{I}),$$
-and $C_1$ and $C_2$ are 
-\begin{align*}
-% \label{eq:C12}
-C_1 =& (\omega+\gamma)(\theta\_w   \gamma+\theta\_c  \omega  p_I) \testinghat{I},\\
-C_2 =& \big( \omega\gamma(1+p_I)\testinghat{I}+\gamma^2 (\omega+\gamma+\testinghat{I})\big)\theta\_w + \omega^2 p_I \testinghat{I} \theta\_c,
-\end{align*}
-where $\testinghat{I}$ is given in Eq.~\eqref{eq:fi}.
-Note that for analysis brevity, we let $N=1$, thus $S\_u^*$ and $S\_n^*$ are in the scale of 0 to 1.
-$\Rnum$ is in the same form as in Eq. \eqref{R0}  
-$$\Rnum= \frac{\beta}{\gamma} (1-\Delta),$$
-where 
-\begin{equation*}
-\label{eq:del4}
-\Delta= \frac{1}{C}\big(C_1 S\_u^*+(C_2(1-\theta\_w)+C \theta\_w) S\_n^*\big).
-\end{equation*}
-
-The first goal
-is to explore how changes in isolation, $\theta\_w$ and $\theta\_c$, affects $\Rnum$.
-Mathematically we would like to verify the sign of $\pder \Rnum{\theta\_w}$ and $\pder \Rnum{\theta\_c}$.
-We start with simplifying $\Delta$ \eqref{eq:del4} by factoring $\theta\_w$ and $\theta\_c$ in Eq.~\eqref{eq:del4}.
-Thus, $\Delta$ can be rewritten as
-\begin{equation}
-\label{eq:del4_theta}
-\Delta= \frac{1}{C}\Big(
--D_1 S\_n^* \theta\_w^2 +
-\big(-\omega^2 p_I \testinghat{I} S\_n^*\theta\_c+ D_2 S\_n^*+\gamma \testinghat{I}(\omega+\gamma) \big)\theta\_w+
-(\omega+\gamma S\_u^*)\omega p_I \testinghat{I} \theta\_c
-\Big),
-\end{equation}
-where
-\begin{align}
-\label{eq:del4_d1d2}
-D_1=& (\omega+\gamma)\gamma^2+(\omega+\gamma+\omega p_I)\gamma \testinghat{I}, \\
-D_2=& (3\omega+2\gamma)\gamma^2+(\omega+\gamma+2\omega p_I)\gamma \testinghat{I}+(\gamma+p_I \testinghat{I})\omega^2.
-\end{align}
-$\Delta$, Eq.~\eqref{eq:del4_theta}, is linear in $\theta\_c$ with a positive coefficient. thus
-\begin{equation}
-\label{eq:d_del4_thetac}
-\pder\Delta{\theta\_c}=1/C(\gamma S\_u^*+\omega(1-\theta\_w S\_n^*))\omega p_I \testinghat{I}.
-\end{equation}
-This results in increasing $\Delta$, thus decreasing $\Rnum$ with respect to $\theta\_c$, that is $\pder \Rnum{\theta\_c}\leq 0$. Note that $C$ is independent of $\theta\_c$ and $\theta\_w$.
-
-With a similar logic, $\Delta$ \eqref{eq:del4_theta} is a concave-down quadratic equation in $\theta\_w$, given by
-\begin{equation}
-\label{eq:del4_thetaw}
-1/C \Big( -D_1 S\_n^*\theta\_w^2+
-\big(-\omega^2 p_I \testinghat{I} S\_n^*\theta\_c+D_2 S\_n^*+\gamma \testinghat{I}(\omega+\gamma) \big)\theta\_w\Big).
-\end{equation}
-We show that the feasible range of $\theta\_w$ lies between 0 and the vertex of this parabola where the parabola is increasing in $\theta\_w$, and so does $\Delta$ which results in inferring $\pder\Rnum{\theta\_w}\leq 0$.
-It is enough to show that partial derivative of the expression \eqref{eq:del4_thetaw} with respect to $\theta\_w$ at $\theta\_w=1$ is non-negative. It follows that
-\begin{align}
-\label{eq:d_del4_thetaw}
-\pder\Delta{\theta\_w}\bigg\rvert_{\theta\_w=1} =&
-1/C\Big( (D_2-2D_1-\omega^2 p_I \testinghat{I} \theta\_c) S\_n^*+\gamma \testinghat{I}(\omega+\gamma) \Big) \\\notag
-=& 1/C\Big( (\gamma(\omega+\gamma)+\gamma\omega^2+(1-\theta\_c)\omega^2 p_I \testinghat{I} ) S\_n^*
-+\gamma(\omega+\gamma) \testinghat{I} (1-S\_n^*) \Big),
-\end{align}
-which is a positive quantity, given that $\theta\_c$ and $S\_n^*$ vary between 0 and 1.
-
-The second goal is to explore how changes in \percap testing intensity $\rho$ affects $\Rnum$.
-Mathematically we would like to verify the sign of $\pder\Rnum{\rho}$, which specifically depends on $\pder\Delta{\rho}$.
-We use the derived expressions for $S\_u^*$ and $S\_n^*$, given by Eqs.~\eqref{dfe}, in $\Delta$ \eqref{eq:del4}.
-Also, we define $\phi = \testinghat{S} = \frac{\rho \omega}{\omega-\rho}$, to reparameterize $\rho$.
-This is mainly to avoid singularity in $\testinghat{I}$ \eqref{eq:fi}, when testing intensity $\rho$ is very close to the rate of test return $\omega$.
-Thus, $\rho$ is reparameterized as
-\begin{equation}
-\label{eq:phi}
-\rho=\frac{\omega \phi}{\omega+\phi}.
-\end{equation}
-This one-to-one monotonic reparameterization enables us to simplify the mathematical expressions and explore the simpler $\pder\Delta{\phi}$ instead of the complicated $\pder\Delta{\rho}$.
-Defining $\testtarget \equiv \oldtesttarget$, the derivative is 
-\begin{align}
-\label{eq:dd3dr}
-\partial\Delta/\partial\phi=& \frac{1}{d_3} (a_3 \phi^2+b_3 \phi+c_3),
-\end{align}
-where
-\begin{align}
-\label{eq:abcd2}
-a_3& = \testtarget \Bigg(
-&& (1-\theta\_w) (1+\testtarget)\theta\_w \gamma^3 +(1-\theta\_c) p_I^2 \theta\_w \testtarget \omega^3 \\ \notag
-& ~ &&+\Big( \big( (1-\theta\_w-\testtarget) \theta\_c +(3-2\theta\_w) \theta\_w\testtarget \big) p_I +(1-\theta\_w) (1+\testtarget)\theta\_w  \Big) \omega \gamma^2 \\ \notag
-& ~ &&+\Big(
-\big((1-\theta\_w-\theta\_w \testtarget) \theta\_c+ (2-\theta\_w) \theta\_w \testtarget \big) p_I
-+(2 \theta\_w-\theta\_w^2-\theta\_c) \testtarget p_I^2
-\Big) \omega^2 \gamma \Bigg), \\ \notag
-b_3& = && \hspace{-1cm} 2\testtarget (\omega+\gamma)\gamma\Big(
-(\omega+\gamma+\omega p_I)(2-\theta\_w) \gamma \theta\_w+ (1-\theta\_w) \omega^2 p_I\theta\_c+\omega^2 p_I\theta\_w
-\Big) ,\\ \notag
-c_3& = && \hspace{-1cm} (\omega+\gamma)^2 \gamma\Big(
-(2-\theta\_w) \gamma^2 \theta\_w+
-(1+\testtarget) \omega \gamma \theta\_w+
-\testtarget \omega^2 p_I\theta\_c \Big),\\ \notag
-d_3&  = && \hspace{-1cm} \frac{(\omega+\gamma)}{\omega}  \Big((\omega p_I+\gamma)\testtarget\phi + (\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2.
-\end{align}
-Note that $\phi\geq 0$, also $b_3$, $c_3$ and $d_3$ are all positive. However $a_3$ can be positive or negative.
-If $a_3\geq 0$, $\partial\Delta/\partial\phi \geq 0$ for all feasible range of parameters, thus $\pro\Rnum \leq 0$.
-It is straightforward to show that $a_3\geq 0$ when testing is random, i.e., $w\_S=w\_I=1$.
-If $a_3 < 0$, then the quadratic expression in the numerator of \eqref{eq:dd3dr} has a positive root, $\phi^*$, such that for $\phi>\phi^*$, $\partial\Delta/\partial\phi < 0$.
-
-An example of this countervailing effect of $\phi$, and consequently $\rho$, on $\Rnum$ occurs when $\theta\_w=0$ and $\theta\_c=1$.
-This is illustrated in the top-right panel of the \fref{pan2} panel (b), where the strength of isolation for awaiting people is the least, but the most for the confirmed cases.
-In this case, simplifying $a_3$ in Eq.~\eqref{eq:abcd2} gives
-\begin{equation}
-  \begin{split}
-    a_3 & =\testtarget \omega \gamma p_I\left((\omega+\gamma)-\testtarget(\omega p_I+\gamma)\right)  \\
-    & \propto  \omega \left( 1- \testtarget p_I \right) + \gamma \left( 1- \testtarget \right)  .
-  \end{split}
-\end{equation}
-If $p_I>0$, then $a_3<0$ for sufficiently targeted testing (i.e. when $\testtarget p_I >1$; because
-$p_I \leq 1$, \testtarget is always $\geq \testtarget p_I$).
-%% this LOOKS like we can get a3<0 even if pI=0 for targeting > 1 and sufficiently large gamma/omega,
-%% but that seems crazy so I'm probably missing something
-When the test is perfectly sensitive ($p_I=1$), $a_3<0$ as long as $\testtarget>1$.
-Under either of these conditions, there exists a range for $\rho$ over which $\pder\Rnum{\rho}\leq 0$.  
-Because increasing values of $\rho$ and $\omega$ both delay the rate at which individuals flow to the $I\_c$ compartment, it is reasonable that increasing either value could (under appropriate circumstances) increase $\Rnum$.
-
-% %%%%%%%
-\subsection{Rate of returning tests} \label{app:omega}
-The third goal is to explore how changes in the rate of test return $\omega$ affect $\Rnum$.
-Mathematically we would like to verify the sign of $\pder\Rnum{\omega}$, which specifically depends on $\pder\Delta{\omega}$.
-We use
-the linearization of $\Delta$ when $\rho \ll 1$ to show that there a non-monotonic relationship between $\Rnum$  and $\omega$. The linear term in the Taylor expansion of $\Delta$ when $\rho \ll 1$ is
-\begin{equation}
-\label{eq:lin}
-\Delta = \frac{\rho}{\omega \gamma(\omega+\gamma)} \Big(
-\testtarget \omega^2 p_I \theta\_c+(\testtarget+1) \gamma\omega\theta\_w +\gamma^2 \theta\_w(2-\theta\_w)
-\Big). 
-\end{equation}
-This results in
-\begin{equation}
-\label{eq:dlindo}
-\pder\Delta{\omega} = \frac{\rho}{\omega^2 (\omega+\gamma)^2} \Big(
-(p_I \testtarget\theta\_c -(1+\testtarget)\theta\_w)\omega^2-2\theta\_w \gamma(2-\theta\_w) \omega+\theta\_w \gamma^2 (\theta\_w-2)
-\Big).
-\end{equation}
-% ALi & Faddy's stuff
-The latter expression has two roots
-\begin{align}
-\label{eq:omega_roots}
-\omega^*_{+} &= \frac{\gamma\Big(-c_4 + \sqrt{c_4^2 + (a_4-b_4)c_4}\Big)}{b_4-a_4} \\
-\omega^*_{-} &= \frac{\gamma\Big(-c_4 - \sqrt{c_4^2 + (a_4-b_4)c_4}\Big)}{b_4-a_4},
-\end{align}
-where $a_4 = p_I\testtarget\theta\_c$, $b_4 = (1+\testtarget)\theta\_w$, $c_4 = \theta\_w(2-\theta\_w)$.
-This enables us to describe the behaviour of $\Rnum$ with respect to $\omega$ in the following two cases.
-\begin{itemize}
-\item \textbf{Case I:} If $b_4 \geq a_4$, $\pder\Delta{\omega} < 0$, so $\Rnum$ is always increasing with respect to $\omega$ (i.e., it is always harmful to return tests more rapidly). 
-\item \textbf{Case II:} If $b_4 < a_4$, $\Rnum$ will be decreasing with respect to $\omega$ (i.e., returning tests more rapidly is beneficial) only when $\omega > \omega^*_-.$
-\end{itemize}
-Note that $b_4 \geq a_4$ is characterized by
-\begin{equation}\label{case_1_ineq}
-\Big(\testtarget+1\Big)\theta\_w \geq \testtarget p_I\theta\_c \Longleftrightarrow \Big(\frac{w_S}{w_I}+1\Big)\frac{1}{p_I} \geq \frac{\theta\_c}{\theta\_w}.
-\end{equation}
-We begin with a \textbf{proof of Case I}.
-Suppose that $b_4 \geq a_4$.
-If the roots of $\pder\Delta{\omega}$ are not complex, then we must have $c_4 \geq b_4-a_4$.
-Note that $\omega^*_{-}$ must be negative since the numerator is clearly negative but the denominator is positive.
-Next, note that since $c_4 \geq b_4-a_4$, the numerator of $\omega^*_{+}$ must also be negative, so $\omega^*_{+}$ is negative.
-Thus, in this case, $\pder\Rnum{\omega}$ does not change sign on $(0,\infty)$.
-Checking the sign of \ref{eq:dlindo} for arbitrarily large $\omega$ shows that it is negative (since the $(a_4-b_4)\omega^2$ term dominates and is negative).
-So $\Rnum$ is increasing with respect to $\omega$ on all of $(0,\infty)$.
-
-We now turn our attention to a \textbf{proof of Case II}.
-Suppose that $b_4 < a_4$.
-It follows that $\omega^*_+$ and $\omega^*_-$ are real since $c_4 > 0 > b_4-a_4$.
-Next, note that $\omega^*_-$ is positive since both the numerator and denominator are negative.
-On the other hand, $\omega^*_+$ is negative since the denominator is negative but the numerator is positive (because $c_4 > b_4-a_4$).
-Thus, our task is to understand the sign of $\pder\Delta{\omega}$ around the root $\omega^*_-$.
-Checking the sign of \ref{eq:dlindo} for arbitrarily large $\omega$ shows that it is positive (since the $(a_4-b_4)\omega^2$ term dominates and is positive).
-Likewise, checking the sign for values of $\omega$ close to $0$ shows that it is negative.
-Thus, $\Rnum$ is increasing with respect to $\omega$ when $\omega < \omega^*_-$, and is decreasing $\omega > \omega^*_-$.
-
-Having presented the formal analysis, we now concern ourselves with its biological interpretations.
-We begin by interpreting \ref{case_1_ineq}, under which returning tests more rapidly is \emph{always} harmful.
-Notice that the ratio $\frac{\theta\_c}{\theta\_w}$ is simply a measure of how much more strongly individuals self-isolate when they test positive compared to when waiting for tests.
-Since the rate of test return directly influences the rate at which individuals change from a waiting state to a confirmed-positive state, it is intuitive that $\frac{\theta\_c}{\theta\_w}$ would appear in \ref{case_1_ineq}.
-Next, note that the left-hand side increases when test sensitivity decreases and when targeting of positive individuals is poor.
-This is consistent with our intuition: a false negative that is returned more rapidly will allow an infectious individual to relax their self-isolation, thus increasing transmission.
-Likewise, if individuals tested are mainly susceptible (rather than infectious), then returning tests more slowly would encourage them to self-isolate for longer while awaiting test results.
-Having understood the role of each of the parameters in \ref{case_1_ineq}, a holistic interpretation of this inequality is that returning tests more slowly is helpful when the benefit of extended self-isolation by infected individuals awaiting test results outweighs the benefit of identifying positive cases.
-
-Now we interpret \textbf{Case II}.
-In this case, $\Rnum$ will have a global maximum with respect to $\omega$ at $\omega^*_-$.
-Note that our model assumption that $\rho < \omega$ plays an important role here: if $\omega^*_- < \rho$, then $\Rnum$ will be always decreasing with respect to $\omega$.
-On the other hand, if $\rho < \omega^*_-$, then $\Rnum$ will be increasing with respect to $\omega$ on $(\rho, \omega^*_-)$ and decreasing beyond that.
-
-% %%%%%%%
-\subsection{The effect of testing focus parameter $\testtarget$ on $\Rnum$} \label{app:w}
-
-\begin{align}
-\label{eq:d_del_wis}
-\pder \Delta{w\_{IS}}= \dfrac{(\omega-\rho) (\omega(\omega-\rho\theta\_w)+\gamma(\omega-\rho)) (\theta\_w\gamma+\theta\_c\omega p\_I)}
-{(-\omega^2 \gamma+\omega \gamma \rho-\gamma \rho \omega w\_{IS}-
-\omega \gamma^2+\gamma^2 \rho-\omega^2 p_I \rho w\_{IS})^2},
-\end{align}
-which is a positive quantity. Thus, $\pder \Rnum{w\_{IS}}\leq 0$. Therefore, increasing the focus of testing on the infectious people will result in less transmission.
-
-% %%%%%%%
-\subsection{On Testing Rate and Numerical Singularity} \label{app:singularity}
-
-In this work, we do not present any numerical solutions of the ODEs to investigate typical trajectories.
-However, attempting to do so quickly reveals a problem in the way the model is posed;
-with the scaling of testing ($\sigma$) defined as in the body of the paper (Eq.~\eqref{sigma}), the population in the $S$ compartments appeared to blow up when the system is near the DFE.
-This occurs because once the only untested people are susceptibles, the FOI approaches $\Lambda=0$, and the testing rate $\testing{S} \to \rho N/S\_u$.
-Thus, the first equation of the model \eqref{model} will become
-$d S\_u/dt = - \rho N + \omega S\_n$. Thus changes in $S\_u$ will be independent on $S\_u$, and the decay of the $S\_u$
-population becomes linear rather than exponential --- allowing $S\_u$ to become negative!
-To avoid this problem the testing rate, $\sigma$, should be formulated such that people from the untested compartments will not be tested if they are not there.
-One way to fix this issue, is to consider a maximum testing rate, $\tau$ (1/day).
-In general, we want to test at a rate of $\rho$ across the whole population.
-This won't always be possible, so we impose a maximum rate of $\tau$ per testable person and redefine $\sigma = \frac{\tau \rho N}{\tau W + \rho N}$, with the assumption that $\tau \gg \rho$.
-This modification of $\sigma$ does not affect any of the results we have derived about the invasion of the epidemic from the DFE (i.e., results on $\Rnum$ and $\Delta$).
-
-\end{document}

### 2021-07-19 introduction.tex
breaking the sections, filling the BMB template

--- 2021-07-16 SIR_manuscript.tex
@@
-\documentclass[12pt]{article}
-\usepackage{natbib}
-\usepackage{hyperref}
-\usepackage{grffile}
-\usepackage{graphicx}
-\usepackage{subcaption}
-\usepackage{amssymb,amsmath,amsthm}
-\usepackage{xcolor}
-\usepackage{xspace}
-\usepackage[nameinlink,capitalize]{cleveref}
-\usepackage{cleveref}
-\usepackage[margin=1in]{geometry}
-\usepackage{lineno}\renewcommand\thelinenumber{\color{gray}\arabic{linenumber}}
-\usepackage{pdflscape}
-\usepackage{enumerate}
-\usepackage{authblk}
-
-\hyphenation{co-ro-na-vi-rus}
-\input{codes/modeldefs.tex}
-
-% https://tex.stackexchange.com/questions/12703/how-to-create-fixed-width-table-columns-with-text-raggedright-centered-raggedlef
-\usepackage{array}
-\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-
-\usepackage{xspace}
-
-\newcommand{\fref}[1]{Fig.~\ref{#1}}
-\newcommand{\appref}[1]{Appendix~\ref{app:#1}}
-\newcommand{\Rlogo}{R\xspace}
-\newcommand{\percap}{\emph{per capita}\xspace}
-\newcommand{\Rnum}{\ensuremath{\mathcal{R}_0}\xspace}
-\newcommand{\covid}{COVID-19\xspace}
-\newcommand{\pro}[1]{\ensuremath{\frac{\partial #1}{\partial \rho}}}
-\newcommand{\pder}[2]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder y{x}
-\newcommand{\testing}[1]{\ensuremath{{\cal T}_{#1}}\xspace}
-\newcommand{\testinghat}[1]{\ensuremath{\hat{\cal T}_{#1}}\xspace}
-\newcommand{\testtarget}{\ensuremath{w\_{IS}}}
-\newcommand{\oldtesttarget}{\ensuremath{\frac{w_I}{w_S}}\xspace}
-\newcommand*\subtxt[1]{_{\textnormal{#1}}}
-%% Roman subscript
-\DeclareRobustCommand\_{\ifmmode\expandafter\subtxt\else\textunderscore\fi}
-
-\newcommand{\comment}{\showcomment}
-\newcommand{\showcomment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}
-\newcommand{\nocomment}[3]{}
-
-\newcommand{\fady}[1]{\comment{cyan}{Fady}{#1}}
-\newcommand{\ali}[1]{\comment{magenta}{Ali}{#1}}
-\newcommand{\jd}[1]{\comment{blue}{JD}{#1}}
-\newcommand{\david}[1]{\comment{orange}{DJDE}{#1}}
-\newcommand{\bmb}[1]{\comment{red}{BMB}{#1}}
-\newcommand{\todo}[1]{\comment{red}{TODO}{#1}}
-\newcommand{\new}[1]{\comment{red}{NEW}{#1}}
-\newcommand{\needref}{\textcolor{red}{(REFS?)}}
-
-%% DE: table stuff
-\renewcommand{\arraystretch}{1.5} % more interline spacing
-%% https://tex.stackexchange.com/questions/12703/how-to-create-fixed-width-table-columns-with-text-raggedright-centered-raggedlef
-\usepackage{ragged2e}
-\usepackage{array} % for m{...}
-
-\theoremstyle{definition} % amsthm only
-\newtheorem{proposition}{Proposition}
-\newtheorem{theorem}{Theorem}
-
-\bibliographystyle{apalike}
-
-\title{Testing and Isolation Efficacy:\\Insights from a Simple Epidemic Model}
-\author[1]{Ali Gharouni}
-\author[1]{F.M. Abdelmalek}
-\author[1,3,5]{David J. D. Earn}
-\author[2,4,3]{Jonathan Dushoff}
-\author[1,2,3]{Benjamin M. Bolker}
-
-\affil[1]{Department of Mathematics \& Statistics, McMaster University, Hamilton, Canada}
-\affil[2]{Department of Biology, McMaster University, Hamilton, Canada}
-\affil[3]{Michael G. DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, Canada}
-\affil[4]{South African Centre for Epidemiological Modelling and Analysis, University of Stellenbosch, Stellenbosch, South Africa}
-
-\begin{document}
-\maketitle
-
-\linenumbers
-
-% %%%%%%%
-\section{Abstract}
-
-Testing individuals for pathogens can affect the spread of epidemics. 
-Understanding how individual-level processes of sampling and reporting test results can affect community- or population-level spread is a dynamical modeling question.
-The effect of testing processes on epidemic dynamics depends on factors underlying implementation, particularly testing intensity and on whom testing is focused.
-Here, we use a simple model to explore how the individual-level effects of testing might directly impact population-level spread.
-Our model development was motivated by the \covid epidemic, but has generic epidemiological and testing structures. 
-To the classic SIR framework we have added a \percap testing intensity, and compartment-specific testing weights, which can be adjusted to reflect different testing emphases --- surveillance, diagnosis, or control.
-We derive an analytic expression for the relative reduction in the basic reproductive number due to testing, test-reporting and related isolation behaviours.
-Intensive testing and fast test reporting are expected to be beneficial at the community level because they can provide a rapid assessment of the situation, identify hot spots, and may enable rapid contact-tracing.
-Direct effects of fast testing at the individual level are less clear, and may depend on how individuals' behaviour is affected by testing information.
-Our simple model shows that under some circumstances both increased testing intensity and faster test reporting can \emph{reduce} the effectiveness of control, and allows us to explore the conditions under which this occurs.
-Conversely, we find that focusing testing on infected individuals always acts to increase effectiveness of control.
-
-% %%%%%%%
-\section{Introduction}
+% \section{Introduction}
 
@@
 We studied the effects of testing intensity, rate of test return, and isolation efficacy, on transmission probability and epidemic dynamics when different levels of testing focus (from random to highly targeted) are in place.
-
-% %%%%%%%
-\section{Methods}
-
-Our model groups individuals based on disease status (Susceptible, Infectious or Recovered) and testing status (\emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}) (\fref{fig:flowchart}).
-The testing status of an individual in a given disease compartment $X$ (where $X \in \{S,I,R\}$) is denoted by a subscript, namely $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
-Two `accumulator' compartments, $N$ and $P$, are included in order to collect cumulative reported negative or positive tests.
-The model equations \eqref{model} and details of calculation of the basic reproduction number $\Rnum$ are presented in \appref{R0}.
-
-\begin{figure}
-\begin{center} 
-\includegraphics[scale=1.5]{pix/sir_comp.pdf}
-\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model, \ref{model}.
-The disease-based status of a compartment $X$ ($X \in \{S,I,R\}$) is combined with the testing status including $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
-The force of infection is denoted by $\Lambda$ (Eq.~\eqref{Lambda}); $\gamma$ is the recovery rate; $\omega$ is the rate of test return; and \testing{X} (Eq.
-\eqref{test}) and $p_X$ represent the \percap testing rate and the sensitivity (probability of positive tests), respectively, for compartment $X$.
-For further description of the parameters see \cref{tab:params}.
-Note that there is a slight mismatch in the top-to-bottom order of the testing-based compartments of each disease-based compartment $X$ between this flowchart and the model equations \eqref{model}; here we have switched  $X\_u$ and $X\_n$ for visual clarity.
-\label{fig:flowchart}}
-\end{center} 
-\end{figure}
-
-Table~\ref{tab:params} defines the model parameters, which are generally \percap flows between compartments, or modifiers to these flow rates.
-The novel component of the model lies in the compartment-specific relative testing weights $w_S$, $w_I$ and $w_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively.
-Thus, we can specify different levels of testing focus from random (all weights equal) to highly targeted (higher weights in more intensively tested compartments).
-For example, $w_I/w_S=3$ means that infected individuals are tested at three times the \percap rate of susceptible individuals.
-
-In order to allow parameterization of the model by the total (overall) \percap testing rate, we define the weighted size of the testing pool $W = w_S S\_u + w_I I\_u + w_R R\_u$, and calculate a scaling parameter for testing as:
-\begin{equation}
-\label{sigma}
-\sigma = \frac{\rho N}{W},
-\end{equation}
-where $\rho$ is the \percap testing intensity for the population, defined as the number of daily tests administered in a population of size $N$.
-Thus, the \percap testing rate for compartment $X \in \{S,I,R\}$ is 
-\begin{equation}
-\label{test}
-\testing{X}=\sigma w_X.
-\end{equation}
-For a highly sensitive test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus not considered for further testing.
-Over the course of the epidemic, a sufficiently large fixed testing rate as specified in \eqref{sigma} can exhaust the pool of people available for testing, leading to a singularity when too few people are left untested to support the specified rate.
-Although this phenomenon does not affect our analysis of $\Rnum$, it can affect model dynamics (we present an adjustment to the model that solves this problem in \appref{singularity}).
-
-The classical SIR model assumes a well-mixed population; homogeneity of the population (i.e., all individuals are equally susceptible and equally infectious with the same recovery rate when infected); exponentially distributed duration of infection; and large population size \citep{keeling2011modeling}.
-In addition to these standard assumptions, our model assumes:
-\begin{enumerate}[(i)]
-\item there is a single force of infection (new infections per unit time per susceptible), $\Lambda$, defined as
-  \begin{equation}
-  \label{Lambda}
-  \Lambda=\frac{\beta}{N} \big(I\_u + (1-\theta\_w)(I\_n+I\_p) + (1-\theta\_c)I\_c \big),
-  \end{equation}
-with transmission rate $\beta$; $\theta\_w$ is the isolation efficacy (reduction of the probability of transmission) for individuals waiting for test results, while $\theta\_c$ is the isolation efficacy for individuals who have received a ``confirmed'' positive test (Table~\ref{tab:params}).
-Susceptible individuals who are waiting for test results experience an additional transmission reduction factor of $1-\theta\_w$ (\fref{fig:flowchart}).
-\item confirmed-positive individuals isolate at least as effectively as those awaiting test results, i.e.,
-$$0\leq\theta\_w\leq\theta\_c\leq 1.$$ 
-\end{enumerate}
-For simplicity we assume that tests are perfectly \emph{specific} --- uninfected individuals never test positive ($p_s=0$).
-Thus there are no waiting-for-positive or confirmed-positive susceptible individuals, which reduces the number of model states from 12 to 10.
-
-\begin{table}[htp]
-\centering
-%%\begin{tabular}{|c|L{2in}cc|} \hline
-%% making use of ragged2e and array packages:
-{\RaggedRight
-\begin{tabular}{ c | m{6cm} | c | c }
-  \textbf{Symbol} & \textbf{Description} & \textbf{Unit} & \textbf{Value} \\ \hline
-  $N$     & Total population size & people & $10^6$ \\ \hline
-  $\omega$  & Rate of test return, i.e., rate of onward flow from ``waiting'' to ``confirmed'' or ``untested'' compartments  & 1/day & - \\ \hline
-  $\gamma$ & Recovery rate & 1/day & 1/6 \\ \hline 
-  $\rho$   & \percap testing intensity & 1/day & - \\ \hline 
-  $\theta\_w$ & Isolation efficacy (reduction of the transmission probability) for ``waiting'' individuals & - & - \\ \hline
-  $\theta\_c$ & Isolation efficacy for ``confirmed positive'' individuals & - & -  \\ \hline
-  $\beta$ & Transmission rate & 1/day & 0.5 \\ \hline
-  $\Lambda$ & Force of infection & 1/day & - \\ \hline
-  $p_S$ & Probability of positive tests for $S$ ($= 1-\textrm{specificity}$) & - & 0 \\ \hline
-  $p_I$ & Probability of positive tests for $I$ ($= \textrm{sensitivity}$) & - & 1 \\ \hline
-  $p_R$ & Probability of positive tests for $R$ ($= 1-\textrm{specificity}$) & - & 0.5 \\ \hline
-  $w_S, w_I, w_R$ & Relative testing weights & - &
-  \begin{minipage}[t]{0.21\columnwidth}%
-    Random:~$\{1,1,1\}$ \\ Targeted:~$\{0.3,1,1\}$
-  \end{minipage} \\
-  %%\hline
-\end{tabular}
-} % end \RaggedRight
-\caption{\label{tab:params}Parameters of the model specified  by the flow chart in \fref{fig:flowchart} and equations~\eqref{model}.}
-\end{table}
-
-The Disease-Free Equilibrium (DFE) for the expanded SIR model (Eqs.~\ref{model}) is found by setting the infected compartments to 0 and solving for the unknowns. The DFE is
-\begin{equation}
-\label{dfe}
-S\_n^*= \frac{\rho}{\omega} N, \ S\_u^*= N-S_n^*, \text{~and~} I_j=R_j=0 \ \text{for all }j.
-\end{equation}
-The corresponding \percap testing rate (Eq.~\ref{test}) for the infected compartment $I$ at DFE is one of the key analysis parameters and can be simplified as 
-\begin{equation}
-\label{eq:fi}
-\testinghat{I} = (\omega\rho/(\omega-\rho))w_I/w_S \quad .
-\end{equation}
-
-The basic reproduction number, $\Rnum$, was calculated by using the next-generation matrix method \citep{van2002reproduction}. We write $\Rnum$ as
-\begin{equation}
-\label{R0}
-\Rnum= \frac{\beta}{\gamma} \left(1-\Delta\right), 
-\end{equation}
-where $\beta/\gamma$ is the classical value for a simple model \citep{keeling2011modeling}, and $1-\Delta$ is the proportional reduction due to testing and isolation processes.
-$\Delta$ therefore measures the ``effectiveness of control'': how much these processes in reducing spread from low levels, and is in turn given by:
-\begin{equation}
-  \label{eq:del4a}
-  \Delta= \frac{1}{C N}\big(C_1 S\_u^*+(C_2(1-\theta\_w)+C \theta\_w)S\_n^*\big),
-\end{equation}
-where
-\begin{align}
-\label{eq:C12}
-C_{\phantom 1} &= (\omega+\gamma) \Big(\gamma(\omega+\gamma)+(\gamma+\omega p_I)\testinghat{I}\Big),\\
-C_1 &= (\omega+\gamma)(\theta\_w \gamma+\theta\_c \omega p_I) \testinghat{I},\\
-C_2 &= \Big( \omega\gamma(1+p_I)\testinghat{I}+\gamma^2(\omega+\gamma+\testinghat{I})\Big)\theta\_w + \omega^2p_I \testinghat{I} \theta\_c.
-\end{align}
-(\appref{R0} gives a detailed derivation of these expressions.)
-This explicit formula enables us to study the effects of testing and isolation parameters on $\Rnum$ both analytically and via numerical solutions.
-We are specifically interested in parameters that could be manipulated by public health policy: isolation efficacy, $\theta\_c$ and $\theta\_w$; \percap testing intensity, $\rho$; and the rate of test return, $\omega$.
-In particular, we look at the partial derivatives of $\Delta$ with respect to these parameters (Appendices~\ref{app:rho} and \ref{app:omega}).
-We derived general expressions for these derivatives.
-However, we analyzed the effect of $\omega$ on $\Delta$ for the special case of low testing intensity.
-Specifically, by making the restriction $\rho \ll 1$, we are able to Taylor-expand $\Delta$ at $\rho=0$, use the linear approximation with respect to $\rho$ and analyze the resulting simplified derivatives to illustrate a surprising non-monotonic relationship between $\Delta$ and $\omega$.
-
-Analytic calculation of the next-generation matrix and simplification of the $\Rnum$ expression, were performed in Maple\textsuperscript{\texttrademark} \citep{maple14}; numerical calculation and contour plots were done in R \citep{r}.
-We computed the values and contours of $\Delta$ at both low (\fref{pan}) and high (\fref{pan2}) testing intensities, and for both random testing ($w_S=w_I=w_R=1$) and targeted testing ($w_S=0.3$; $w_I=w_R=1$).
-Because it is expressed as a proportion of $\Rnum$, the effectiveness of control $\Delta$ is (at least in the $\rho \ll 1$ case, Eq.~\ref{eq:lin}) independent of the transmission rate $\beta$, and hence of $\Rnum$ in the case where we vary $\Rnum$ by changing the transmission rate for a fixed generation interval.
-
-The low-testing case (\fref{pan}) reflects the case where testing intensity $\rho$ is small relative to the population size.
-Specifically, $\rho \in [0,0.013]$, and test return rate $\omega\in [1/12,2]$.
-This testing intensity is of the correct order of magnitude (although typically larger than) testing rates during the \covid pandemic, i.e., a maximum of 1.3\% of the population per day (approximately four times the maximum testing rate in Ontario, Canada in mid-2021).
-The less realistic high-testing case (\fref{pan2}) is included to highlight the occurrence of non-monotonic changes in $\Rnum$ with respect to $\rho$.
-In \fref{pan2} the maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,1/5)$ and the test return rate $\omega\in [1/5,2]$; these values are clearly unrealistic for a large population but might be relevant for small populations undergoing focused testing, such as a sports league or university.
-In these figures, the implied baseline reproduction number (for the SIR model without testing) is $\Rnum=\frac{\beta}{\gamma}=3$.
-The different ranges of test return rates $\omega$ for the cases of low and high testing intensities is due to the restriction $\rho<\omega$, which is a requirement for a feasible DFE (\ref{dfe}).
-
-% %%%%%%%
-\section{Results}
-
-We presented $\Rnum$ as the product of the classical reproduction number, $\beta/\gamma$, and the proportional reduction due to testing and isolation, $1-\Delta$, \eqref{R0}.
-We can use the formula for $\Delta$ \eqref{eq:del4a} to make a number of straightforward inferences about parameters that affect $\Rnum$ monotonically, i.e., for which the associated partial derivative of $\Delta$ always has the same sign (see Appendices).
-
-\begin{enumerate}
-
-\item \label{p1:eta} Increasing isolation efficacy for waiting ($\theta\_w$) and confirmed-positive ($\theta\_c$) individuals always increases $\Delta$ (Eqs.~\ref{eq:del4_theta},~\ref{eq:d_del4_thetac},~\ref{eq:d_del4_thetaw});
-\item \label{p1:rho} Higher testing intensity $\rho$ increases $\Delta$ if
-%% \begin{enumerate} \item
-testing is random (all $w_X$ equal) or testing intensity ($\rho$) is small (Eq.~\ref{eq:dd3dr}).
-%% \end{enumerate}
-\item \label{p1:omega} Increasing the rate of test return ($\omega$) always increases $\Delta$ if waiting individuals do not isolate ($\theta\_w=0$) (Eq.~\ref{eq:dlindo}).
-\item \label{p1:w} Increasing testing focus, i.e., changing the testing weights from random ($w_S=w_I$) toward targeted  ($w_S<w_I$), always increases $\Delta$ (Eq.~\ref{eq:d_del_wis}).
-\end{enumerate}
-
-\newpage
-\begin{figure}[h!]
-\centering
-\begin{subfigure}[t]{.49\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_random.pdf}
-\caption{}\label{p.a}
-\end{subfigure}
-%
-\begin{subfigure}[t]{.49\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_TTI.pdf}
-\caption{}\label{p.b}
-\end{subfigure}
-\caption{
-{\bf Effectiveness of testing and isolation in reducing $\Rnum$ at low \percap testing intensity ($\rho$).}
-Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4a}), over a range of testing and isolation parameters. Parameter values (Table~\ref{tab:params}):
-$\beta=\betaparam/$day, $1/\gamma= \invgammaparam$ days (baseline $\Rnum=\Rnumparam$, $r=\rparam$); $\omega \in [1/12,2]/$day;  $\rho \in [0,0.013]/$day per capita; $\theta\_w$ and $\theta\_c$ vary between 0 (no effect of isolation) and 1 (complete elimination of transmission); $p_S=0$, $p_I=1$ and $p_R=0.5$.
-Only parameter sets where $\theta\_c \geq \theta\_w$ (confirmed-positive individuals isolate more effectively than waiting individuals) are shown; the alternative case, $\theta\_w > \theta\_c$, is unrealistic.
-Contours of $\Delta$ are plotted for (a) random testing ($w_S=w_I=w_R=1$) and (b) targeted testing ($w_S=0.3$; $w_I=w_R=1$).
-}
-\label{pan}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\begin{subfigure}[t]{.49\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_random2.pdf}
-\caption{}
-\end{subfigure}
-%
-\begin{subfigure}[t]{.49\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_TTI2.pdf}
-\caption{}
-\end{subfigure}
-\caption{
-  {\bf Effectiveness of testing and isolation in reducing $\Rnum$ at high \percap testing intensity.}
-  Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4a}), over a range of testing and isolation parameters.
-Parameters as in \fref{pan} except: $\omega \in [1/5,2]/$day, $\rho \in [0,1/5)/$day.
-As in \fref{pan}, subplots show (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
-}
-\label{pan2}
-\end{figure}
-
-However, there are also two specific cases where $\Delta$ changes non-monotonically, in counterintuitive directions, as a function of testing and isolation parameters.
-
-\begin{itemize}
-\item We would generally expect increasing testing delays to increase \Rnum, thus decreasing effectiveness of control $\Delta$.
-This is in fact what happens when waiting individuals do not isolate ($\theta\_w =0$, top row of \fref{pan}) --- as we move to the right within each plot in this row, $\Delta$ decreases.
-However, when waiting individuals isolate ($\theta\_w>0$), we more often see the opposite effect: longer testing delays lead to a greater control effect $\Delta$ (reduced $\Rnum$).
-The reason is that people waiting for negative tests are assumed to continue to isolate; this applies both to susceptibles and to people who became infected while waiting for negative test results.
-This effect outweighs the effect of confirmed individuals isolating, except when this isolation parameter ($\theta\_c$) is substantially greater than $\theta\_w$.
-This result depends on the idea that, all else equal, people who have to wait longer for test results isolate at the same level (but for a longer time) as they would if the wait were shorter.
-\item \fref{pan} also shows that greater testing intensity (increasing $\rho$) generally increases the effectiveness of control (moving up in each panel).
-However, this relationship can be reversed at very high testing intensities (provided testing is targeted, and $\theta\_w$ is relatively small; \fref{pan2}(b), right three panels of top row).
-It is theoretically possible for increasing testing intensity to \emph{increase} \Rnum because more rapid testing leaves more susceptibles in the ``waiting-for-negative-results'' category at the DFE; if these people become infected while waiting, they will need to wait for their negative test result before they can be tested again, receive a positive test, and then begin self-isolating.
-This effect is usually weak compared to the beneficial effects of testing.
-
-\end{itemize}
-
-%%%%%%%%%%%%
-\section{Discussion}
-
-In this paper, we have developed and analyzed a simple compartmental model that combines epidemiological dynamics---as defined by a simple SIR model---with the dynamics of testing and isolation.
-Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness}).
-Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
-
-Many of the qualitative results we have derived confirm simple, common-sense intuitions.
-In particular, we can generally decrease $\Rnum$ by increasing isolation efficacy or testing intensity; returning tests faster, if individuals do not isolate while they are waiting for results; or increasing testing focus to target individuals who are likely to be infectious (e.g., symptomatic people or close contacts of known infections).
-
-However, we did find two surprising patterns: under some conditions longer delays in returning tests can reduce epidemic spread, and increasing testing rates can increase spread.
-
-Over broad regions of parameter space, decreasing $\omega$---i.e., \emph{slowing} the rate at which test results are returned---\emph{decreases} $\Rnum$ (for random testing, the parameter region is $\theta\_w \gtrsim 0.25$; for targeted testing, $\theta\_w \geq 0.25$ and either $\theta\_c \ge 0.5$ or $1/\omega > 5$; see Table~\ref{tab:params} for parameter definitions).
-This result is counterintuitive and would not be expected by public health authorities that have invested a great deal of effort in reducing delays from testing to results.
-Dynamically, this effect occurs because speeding up test returns shortens the isolation period of uninfected individuals (for infected people it only shortens the time to progression to the isolation level of the confirmed-positive compartment).
-Slowing test returns increases $\Rnum$ only if the proportion of infectives in the tested population is high and isolation is relatively strong among people waiting for test results.
-
-While slowing test returns does decrease $\Rnum$ over broad regions of parameter space in our model, there are several real-world processes missing from our model that make it unlikely that slowing test returns would actually be an effective public health measure.
-First, we do not model the primary benefit of rapid testing, i.e., detecting and containing outbreaks while they are still in progress.
-This process could be modeled phenomenologically by making the testing focus more targeted as an increasing proportion of cases is detected, because finding infections allows tests to be concentrated on their connections.
-Second, individuals may become less likely to maintain isolation if they are required to do so for longer; phenomenologically, we could allow  effectiveness of isolation in the waiting population to be an increasing function of test-return speed, or we could introduce a separate ``waiting, but no longer isolating'' compartment that individuals entered from the ``waiting, isolated'' compartment at a specified rate.
-Finally, if one wants to decrease the overall transmission rate of the population there are more effective methods than keeping tested people in limbo; these include masking, ventilation, distancing measures, retail and event closures, and stay-at-home orders.
-
-The other counterintuitive result from our analysis is that, for sufficiently high testing intensity $\rho$, further increasing testing intensity can actually \emph{increase} $\Rnum$ (e.g., \fref{pan2}(b), upper right panel [$\theta\_c=1$, $\theta\_w=0$]).
-This phenomenon can occur because we are considering the DFE in the presence of testing; thus there is an equilibrium distribution of susceptibles between the $S\_n$ (untested) and $S\_u$ (waiting) compartments even as the disease approaches extinction.
-A higher rate of testing leads to a greater proportion of individuals waiting for negative tests at the DFE.
-If infected, individuals in this group will take longer to be tested again and to subsequently isolate (because they must wait for their negative tests to be returned before being tested again).
-If isolation in this group ($\theta\_w$) is low, this effect can under some (relatively rare) circumstances (high $\theta\_c$, low $\omega$, high $\rho$) allow $\Rnum$ to increase with testing intensity.
-We can show that this phenomenon occurs \emph{only} under targeted testing ($w_I > w_S$), but we have not yet found a simple explanation of why it cannot occur under random (unfocused) testing.
-This phenomenon is also unlikely to occur in the real world.
-In particular, it depends on levels of testing that are unrealistically high (at least in large, general-population settings).
-
-Although we model the testing process in more detail than typical epidemiological models, one place where more detail could be informative is in the processes determining the testing weights $\{w_S, w_I, w_R\}$.
-While random testing, as done for surveillance purposes, unambiguously leads to equal testing weights, making precise quantitative connections between public-health practices and testing weights is difficult in other contexts.
-The testing weights reflect the correlation between an individual's risk of infection and their likelihood of being tested due to age, occupation, geographic location, etc..
-This correlation is influenced, among many other factors, by the proportion of the uninfected population with COVID-like symptoms (e.g., due to seasonal upper respiratory tract infections); the concentration of transmission and testing in hot spots such as long-term care facilities and high-density workplaces; the overall testing intensity (and hence, e.g., restriction to symptomatic individuals); and the proportion of COVID-infected people who are symptomatic.
-Future steps should explore mathematically tractable ways to model some of these factors more precisely.
-For example, separating the infected class into exposed, symptomatic, and a- or pre-symptomatic compartments and allowing the testing weights to vary across non-symptomatic (exposed/asymptomatic/presymptomatic) vs.\ symptomatic compartments could reflect the allocation of tests for diagnostic purposes (targeting symptomatic individuals) vs.\ contact-tracing (targeting infected but non-symptomatic individuals) vs.\ screening (relatively equal weights, depending on the venue).
-Alternatively, one could make the testing weights depend on the testing intensity or test-return rate as suggested above.
-Whatever complexity is added would probably put the model beyond reach of the analytical methods we have used in this paper, but one could still use semi-numerical methods such as constructing the next-generation matrix and using it to evaluate the derivatives of $\Rnum$ with respect to the parameters numerically.
-
-Although testing and tracing is a key part of infection control strategies, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts \citep{endo2020implication,hellewell2020feasibility,jenness2020modeling}, rather than analyzing simple but general models of the feedback between testing and transmission dynamics.
-There have been several modeling studies of testing and tracing dynamics and their interaction with epidemiological dynamics.
-In the context of repeated screening and random testing of isolated populations (such as the members of a university), \cite{bergstrom2020frequency} provided analytical results quantifying the effects that proactive screening of asymptomatic individuals and isolation of confirmed-positive cases could have in reducing the spread of disease.
-\cite{rogers2021high} simulated a SEIR model with testing and isolation; they similarly suggest a strategy of rapid testing with antigen tests and the subsequent isolation of confirmed-positive individuals.
-\cite{friston2021testing} model the effects of self-isolation on testing and tracing with a focus on projections under different testing and tracing scenarios.
-They conclude that the emergence of a second wave depends almost primarily on the rate at which immunity is lost and that it is necessary to track asymptomatic individuals in order to control the outbreak.
-Our modeling approach differs from these previous efforts in that it examines the effects of test-return rates and of different levels of testing focus, from random to highly targeted.
-We hope this paper will inspire further explorations of the fundamental properties of epidemic models that incorporate explicit testing processes.
-
-% %%%%%%%
-\bibliography{SIRlibrary}
-
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\clearpage
-% \widetext
-\appendix
-\section{Appendix}
-\setcounter{equation}{0}
-\setcounter{figure}{0}
-\setcounter{table}{0}
-% \setcounter{page}{1}
-\makeatletter
-\renewcommand{\theequation}{A\arabic{equation}}
-\renewcommand{\thefigure}{A\arabic{figure}}
-\renewcommand{\bibnumfmt}[1]{[A#1]}
-\renewcommand{\citenumfont}[1]{A#1}
-
-\subsection{Model and calculation of $\Rnum$}\label{app:R0}
-
-The model in the form of a system of ordinary differential equations is 
-\begin{subequations}\label{model}
-\begin{align}
- d S\_u/dt &= -\Lambda S\_u - \testing{S} S\_u + \omega S\_n, \\
- d S\_n/dt &= -(1-\theta\_w)\Lambda S\_n + (1-p_S) \testing{S} S\_u - \omega S\_n, \\
- d S\_p/dt &= -(1-\theta\_w)\Lambda S\_p + p_S \testing{S} S\_u - \omega S\_p, \\
- d S\_c/dt &= -(1-\theta\_c)\Lambda S\_c + \omega S\_p, \\
- d I\_u/dt &= \Lambda S\_u - \testing{I} I\_u + \omega I\_n  - \gamma I\_u, \\
- d I\_n/dt &= (1-\theta\_w)\Lambda S\_n + (1-p_I) \testing{I} I\_u - \omega I\_n -\gamma I\_n, \\
- d I\_p/dt &= (1-\theta\_w)\Lambda S\_p + p_I \testing{I} I\_u - \omega I\_p -\gamma I\_p, \\
- d I\_c/dt &= (1-\theta\_c)\Lambda S\_c + \omega I\_p - \gamma I\_c, \\
- d R\_u/dt &= \gamma I\_u - \testing{R} R\_u + \omega R\_n, \\
- d R\_n/dt &= \gamma I\_n + (1-p_R) \testing{R} R\_u - \omega R\_n, \\
- d R\_p/dt &= \gamma I\_p + p_R \testing{R} R\_u  - \omega R\_p, \\
- d R\_c/dt&= \gamma I\_c + \omega R\_p, \\
- dN/dt &= \omega (S\_n + I\_n + R\_n),  \\
- dP/dt &= \omega(I\_p + R\_p) ,
-\end{align}
-\end{subequations}
-%
-(see Table \ref{tab:params} for parameter definitions).
-The next generation matrix for this model is $G = F V^{-1}$, where matrix $F$ represents the inflow of new infection to the infected compartments and matrix $V$ represents the flow in the infected compartments when the population is totally susceptible.
-Matrices $F$ and $V$ are
-\begin{align}
-\label{F}
-F =& \beta/N \left[ \begin {array}{cccc} 
-S\_u^* & (1-\theta\_w) S\_u^* & (1-\theta\_w) S\_u^* & (1-\theta\_c) S\_u^*\\
-(1-\theta\_w)S\_n^* & (1-\theta\_w)^2 S\_n^* & (1-\theta\_w)^2 S\_n^* & (1-\theta\_w)(1-\theta\_c) S\_n^* \\ 
-0&0&0&0\\
-0&0&0&0
- \end {array} \right] \\ \notag
- =&\beta/N \left[ \begin {array}{c} S\_u^* \\ (1-\theta\_w)S\_n^* \\ 0 \\ 0 \end {array} \right]
-        \left[ \begin {array}{cccc} 1,   1-\theta\_w,   1-\theta\_w,   1-\theta\_c \end {array} \right],
-\end{align}
-and 
-\begin{align}
-\label{V}
-V= \left[ \begin {array}{cccc}  
-\testinghat{I}+\gamma&-\omega&0&0\\
--(1-p_I)\testinghat{I}&\omega+\gamma&0&0\\
--p_I \testinghat{I}&0&\omega+\gamma&0\\
-0&0&-\omega&\gamma
-\end {array} \right].
-\end{align}
-The matrix inverse of $V$ is 
-\begin{align}
-\label{vinv}
-V^{-1} =
-\frac{1}{\gamma C}
-\left[ \begin {array}{cccc}
-\gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)&0&0\\ \noalign{\medskip}
-\gamma(\omega+\gamma)(1-p_I) \testinghat{I} & \gamma(\omega+\gamma)(\testinghat{I}+\gamma)&0&0 \\ \noalign{\medskip}
-\gamma(\omega+\gamma) p_I \testinghat{I} & \gamma\omega p_I \testinghat{I} & C \gamma/(\omega+\gamma) & 0 \\ \noalign{\medskip}
-\omega(\omega+\gamma) p_I \testinghat{I} & \omega^2 p_I \testinghat{I} & C \omega/(\omega+\gamma) & C
-\end {array} \right],
-\end{align}
-where $C= \big( \gamma(\omega+\gamma)+(\gamma+\omega p_I)\testinghat{I} \big) (\omega+\gamma)$ and $\testinghat{I}$ is the \percap testing rate for the infected people and represented in Eq.~\eqref{eq:fi}.
-Note that all the columns of matrix $V^{-1}$ summ up to $1/\gamma$.
-
-The particular form of $F$ with two rows of zeros at the bottom results in the following blocked form of matrix $G$.
-\begin{equation}
-\label{mat:G}
-G = \left[ \begin {array}{cc}
-G_{11}&G_{12}\\
-0&0
-\end {array} \right],
-\end{equation}
-where both blocked matrices $G_{11}$ and $G_{12}$ are 2 by 2.
-Given the upper triangular form of matrix $G$, the basic reproduction number $\Rnum$ (defined as the spectral radius of matrix $G$) is only determined by the blocked matrix $G_{11}$,
-\begin{equation}
-\label{G11}
-G_{11} = \frac{\beta}{\gamma   C} 
-\left[ \begin {array}{c} (\omega-\rho)/\omega \\ (1-\theta\_w)\rho/\omega \end {array} \right]
-\left[ \begin {array}{cccc} 1,   1-\theta\_w,   1-\theta\_w,   1-\theta\_c \end {array} \right]
-\left[ \begin {array}{cc}
-\gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)\\ \noalign{\medskip}
-\gamma(\omega+\gamma)(1-p_I)\testinghat{I} & \gamma(\omega+\gamma)(\testinghat{I}+\gamma) \\ \noalign{\medskip}
-\gamma(\omega+\gamma) p_I \testinghat{I} & \gamma\omega p_I \testinghat{I} \\ \noalign{\medskip}
-\omega(\omega+\gamma) p_I \testinghat{I} & \omega^2 p_I \testinghat{I}
-\end {array} \right].
-\end{equation}
-It is notable that matrix $F$ \eqref{F} has rank one and consequently $G_{11}$ does so. That is $G_{11}$ has only one non-zero eigenvalue which is $\Rnum$.
-
-The expression of $\Rnum$ has a complicated form with all of the model parameters involved.
-This expression can be simplified and represented given the specific form of matrix $G_{11}$ \eqref{G11}.
-For the purpose of simplicity we present $\Rnum$ in the manuscript in terms of expressions $C$, $C_1$ and $C_2$, specified in \eqref{eq:C12}.
-
-It remains difficult to show that the reproduction number $\Rnum$ is decreasing with respect to \percap testing intensity, $\rho$, and the speed of the test return, $\omega$, for the feasible ranges of the parameters, that is
-\begin{align}
-\label{cond}
-& \omega>0, \\
-& 0 \leq \rho<\omega,\\ 
-& 0 \leq \theta\_w \leq \theta\_c \leq 1, \\
-& \oldtesttarget\geq 1.
-\end{align}
-In realistic cases the testing rate $\rho$ is very small (i.e., only a small fraction of the population can be tested every day); it is thus reasonable to use a linear approximation of $\Rnum$ for $\rho \ll 1$ to analyze the behaviour of $\Rnum$ with respect to $\omega$ (see section \appref{omega}).
-In the next section we provide an equivalent representation of $\Rnum$ in order to show that increasing testing intensity typically decreases $\Rnum$.
-
-% %%%%%%%
-\subsection{More testing intensity may decrease $\Rnum$}\label{app:rho}
-
-This section shows that $\pro{\Delta}$ can be positive or negative, with $\Delta$ defined in Eq.~\eqref{eq:C12}, and thus $\pro{\Rnum} < 0$, where $\Rnum$ is given in Eq.~\eqref{R0}.
-We rewrite matrix $G_{11}$ in \eqref{G11} in the following form to simplify the calculations:
-\begin{equation}
-\label{G112}
-G_{11} = \frac{\beta}{\gamma C} 
-\left[ \begin {array}{c}  S\_u^*/N \\ (1-\theta\_w) S\_n^*/N  \end {array} \right]
-\left[ \begin {array}{cccc} 
-C-C_1, C-C_2\end {array} \right],
-\end{equation}
-where $C$ is the same as the one in Eq. \eqref{eq:C12}, i.e., 
-$$C=(\omega+\gamma)(\gamma(\omega+\gamma)+(\omega p_I+\gamma) \testinghat{I}),$$
-and $C_1$ and $C_2$ are 
-\begin{align*}
-% \label{eq:C12}
-C_1 =& (\omega+\gamma)(\theta\_w   \gamma+\theta\_c  \omega  p_I) \testinghat{I},\\
-C_2 =& \big( \omega\gamma(1+p_I)\testinghat{I}+\gamma^2 (\omega+\gamma+\testinghat{I})\big)\theta\_w + \omega^2 p_I \testinghat{I} \theta\_c,
-\end{align*}
-where $\testinghat{I}$ is given in Eq.~\eqref{eq:fi}.
-Note that for analysis brevity, we let $N=1$, thus $S\_u^*$ and $S\_n^*$ are in the scale of 0 to 1.
-$\Rnum$ is in the same form as in Eq. \eqref{R0}  
-$$\Rnum= \frac{\beta}{\gamma} (1-\Delta),$$
-where 
-\begin{equation*}
-\label{eq:del4}
-\Delta= \frac{1}{C}\big(C_1 S\_u^*+(C_2(1-\theta\_w)+C \theta\_w) S\_n^*\big).
-\end{equation*}
-
-The first goal
-is to explore how changes in isolation, $\theta\_w$ and $\theta\_c$, affects $\Rnum$.
-Mathematically we would like to verify the sign of $\pder \Rnum{\theta\_w}$ and $\pder \Rnum{\theta\_c}$.
-We start with simplifying $\Delta$ \eqref{eq:del4} by factoring $\theta\_w$ and $\theta\_c$ in Eq.~\eqref{eq:del4}.
-Thus, $\Delta$ can be rewritten as
-\begin{equation}
-\label{eq:del4_theta}
-\Delta= \frac{1}{C}\Big(
--D_1 S\_n^* \theta\_w^2 +
-\big(-\omega^2 p_I \testinghat{I} S\_n^*\theta\_c+ D_2 S\_n^*+\gamma \testinghat{I}(\omega+\gamma) \big)\theta\_w+
-(\omega+\gamma S\_u^*)\omega p_I \testinghat{I} \theta\_c
-\Big),
-\end{equation}
-where
-\begin{align}
-\label{eq:del4_d1d2}
-D_1=& (\omega+\gamma)\gamma^2+(\omega+\gamma+\omega p_I)\gamma \testinghat{I}, \\
-D_2=& (3\omega+2\gamma)\gamma^2+(\omega+\gamma+2\omega p_I)\gamma \testinghat{I}+(\gamma+p_I \testinghat{I})\omega^2.
-\end{align}
-$\Delta$, Eq.~\eqref{eq:del4_theta}, is linear in $\theta\_c$ with a positive coefficient. thus
-\begin{equation}
-\label{eq:d_del4_thetac}
-\pder\Delta{\theta\_c}=1/C(\gamma S\_u^*+\omega(1-\theta\_w S\_n^*))\omega p_I \testinghat{I}.
-\end{equation}
-This results in increasing $\Delta$, thus decreasing $\Rnum$ with respect to $\theta\_c$, that is $\pder \Rnum{\theta\_c}\leq 0$. Note that $C$ is independent of $\theta\_c$ and $\theta\_w$.
-
-With a similar logic, $\Delta$ \eqref{eq:del4_theta} is a concave-down quadratic equation in $\theta\_w$, given by
-\begin{equation}
-\label{eq:del4_thetaw}
-1/C \Big( -D_1 S\_n^*\theta\_w^2+
-\big(-\omega^2 p_I \testinghat{I} S\_n^*\theta\_c+D_2 S\_n^*+\gamma \testinghat{I}(\omega+\gamma) \big)\theta\_w\Big).
-\end{equation}
-We show that the feasible range of $\theta\_w$ lies between 0 and the vertex of this parabola where the parabola is increasing in $\theta\_w$, and so does $\Delta$ which results in inferring $\pder\Rnum{\theta\_w}\leq 0$.
-It is enough to show that partial derivative of the expression \eqref{eq:del4_thetaw} with respect to $\theta\_w$ at $\theta\_w=1$ is non-negative. It follows that
-\begin{align}
-\label{eq:d_del4_thetaw}
-\pder\Delta{\theta\_w}\bigg\rvert_{\theta\_w=1} =&
-1/C\Big( (D_2-2D_1-\omega^2 p_I \testinghat{I} \theta\_c) S\_n^*+\gamma \testinghat{I}(\omega+\gamma) \Big) \\\notag
-=& 1/C\Big( (\gamma(\omega+\gamma)+\gamma\omega^2+(1-\theta\_c)\omega^2 p_I \testinghat{I} ) S\_n^*
-+\gamma(\omega+\gamma) \testinghat{I} (1-S\_n^*) \Big),
-\end{align}
-which is a positive quantity, given that $\theta\_c$ and $S\_n^*$ vary between 0 and 1.
-
-The second goal is to explore how changes in \percap testing intensity $\rho$ affects $\Rnum$.
-Mathematically we would like to verify the sign of $\pder\Rnum{\rho}$, which specifically depends on $\pder\Delta{\rho}$.
-We use the derived expressions for $S\_u^*$ and $S\_n^*$, given by Eqs.~\eqref{dfe}, in $\Delta$ \eqref{eq:del4}.
-Also, we define $\phi = \testinghat{S} = \frac{\rho \omega}{\omega-\rho}$, to reparameterize $\rho$.
-This is mainly to avoid singularity in $\testinghat{I}$ \eqref{eq:fi}, when testing intensity $\rho$ is very close to the rate of test return $\omega$.
-Thus, $\rho$ is reparameterized as
-\begin{equation}
-\label{eq:phi}
-\rho=\frac{\omega \phi}{\omega+\phi}.
-\end{equation}
-This one-to-one monotonic reparameterization enables us to simplify the mathematical expressions and explore the simpler $\pder\Delta{\phi}$ instead of the complicated $\pder\Delta{\rho}$.
-Defining $\testtarget \equiv \oldtesttarget$, the derivative is 
-\begin{align}
-\label{eq:dd3dr}
-\partial\Delta/\partial\phi=& \frac{1}{d_3} (a_3 \phi^2+b_3 \phi+c_3),
-\end{align}
-where
-\begin{align}
-\label{eq:abcd2}
-a_3& = \testtarget \Bigg(
-&& (1-\theta\_w) (1+\testtarget)\theta\_w \gamma^3 +(1-\theta\_c) p_I^2 \theta\_w \testtarget \omega^3 \\ \notag
-& ~ &&+\Big( \big( (1-\theta\_w-\testtarget) \theta\_c +(3-2\theta\_w) \theta\_w\testtarget \big) p_I +(1-\theta\_w) (1+\testtarget)\theta\_w  \Big) \omega \gamma^2 \\ \notag
-& ~ &&+\Big(
-\big((1-\theta\_w-\theta\_w \testtarget) \theta\_c+ (2-\theta\_w) \theta\_w \testtarget \big) p_I
-+(2 \theta\_w-\theta\_w^2-\theta\_c) \testtarget p_I^2
-\Big) \omega^2 \gamma \Bigg), \\ \notag
-b_3& = && \hspace{-1cm} 2\testtarget (\omega+\gamma)\gamma\Big(
-(\omega+\gamma+\omega p_I)(2-\theta\_w) \gamma \theta\_w+ (1-\theta\_w) \omega^2 p_I\theta\_c+\omega^2 p_I\theta\_w
-\Big) ,\\ \notag
-c_3& = && \hspace{-1cm} (\omega+\gamma)^2 \gamma\Big(
-(2-\theta\_w) \gamma^2 \theta\_w+
-(1+\testtarget) \omega \gamma \theta\_w+
-\testtarget \omega^2 p_I\theta\_c \Big),\\ \notag
-d_3&  = && \hspace{-1cm} \frac{(\omega+\gamma)}{\omega}  \Big((\omega p_I+\gamma)\testtarget\phi + (\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2.
-\end{align}
-Note that $\phi\geq 0$, also $b_3$, $c_3$ and $d_3$ are all positive. However $a_3$ can be positive or negative.
-If $a_3\geq 0$, $\partial\Delta/\partial\phi \geq 0$ for all feasible range of parameters, thus $\pro\Rnum \leq 0$.
-It is straightforward to show that $a_3\geq 0$ when testing is random, i.e., $w\_S=w\_I=1$.
-If $a_3 < 0$, then the quadratic expression in the numerator of \eqref{eq:dd3dr} has a positive root, $\phi^*$, such that for $\phi>\phi^*$, $\partial\Delta/\partial\phi < 0$.
-
-An example of this countervailing effect of $\phi$, and consequently $\rho$, on $\Rnum$ occurs when $\theta\_w=0$ and $\theta\_c=1$.
-This is illustrated in the top-right panel of the \fref{pan2} panel (b), where the strength of isolation for awaiting people is the least, but the most for the confirmed cases.
-In this case, simplifying $a_3$ in Eq.~\eqref{eq:abcd2} gives
-\begin{equation}
-  \begin{split}
-    a_3 & =\testtarget \omega \gamma p_I\left((\omega+\gamma)-\testtarget(\omega p_I+\gamma)\right)  \\
-    & \propto  \omega \left( 1- \testtarget p_I \right) + \gamma \left( 1- \testtarget \right)  .
-  \end{split}
-\end{equation}
-If $p_I>0$, then $a_3<0$ for sufficiently targeted testing (i.e. when $\testtarget p_I >1$; because
-$p_I \leq 1$, \testtarget is always $\geq \testtarget p_I$).
-%% this LOOKS like we can get a3<0 even if pI=0 for targeting > 1 and sufficiently large gamma/omega,
-%% but that seems crazy so I'm probably missing something
-When the test is perfectly sensitive ($p_I=1$), $a_3<0$ as long as $\testtarget>1$.
-Under either of these conditions, there exists a range for $\rho$ over which $\pder\Rnum{\rho}\leq 0$.  
-Because increasing values of $\rho$ and $\omega$ both delay the rate at which individuals flow to the $I\_c$ compartment, it is reasonable that increasing either value could (under appropriate circumstances) increase $\Rnum$.
-
-% %%%%%%%
-\subsection{Rate of returning tests} \label{app:omega}
-The third goal is to explore how changes in the rate of test return $\omega$ affect $\Rnum$.
-Mathematically we would like to verify the sign of $\pder\Rnum{\omega}$, which specifically depends on $\pder\Delta{\omega}$.
-We use
-the linearization of $\Delta$ when $\rho \ll 1$ to show that there a non-monotonic relationship between $\Rnum$  and $\omega$. The linear term in the Taylor expansion of $\Delta$ when $\rho \ll 1$ is
-\begin{equation}
-\label{eq:lin}
-\Delta = \frac{\rho}{\omega \gamma(\omega+\gamma)} \Big(
-\testtarget \omega^2 p_I \theta\_c+(\testtarget+1) \gamma\omega\theta\_w +\gamma^2 \theta\_w(2-\theta\_w)
-\Big). 
-\end{equation}
-This results in
-\begin{equation}
-\label{eq:dlindo}
-\pder\Delta{\omega} = \frac{\rho}{\omega^2 (\omega+\gamma)^2} \Big(
-(p_I \testtarget\theta\_c -(1+\testtarget)\theta\_w)\omega^2-2\theta\_w \gamma(2-\theta\_w) \omega+\theta\_w \gamma^2 (\theta\_w-2)
-\Big).
-\end{equation}
-% ALi & Faddy's stuff
-The latter expression has two roots
-\begin{align}
-\label{eq:omega_roots}
-\omega^*_{+} &= \frac{\gamma\Big(-c_4 + \sqrt{c_4^2 + (a_4-b_4)c_4}\Big)}{b_4-a_4} \\
-\omega^*_{-} &= \frac{\gamma\Big(-c_4 - \sqrt{c_4^2 + (a_4-b_4)c_4}\Big)}{b_4-a_4},
-\end{align}
-where $a_4 = p_I\testtarget\theta\_c$, $b_4 = (1+\testtarget)\theta\_w$, $c_4 = \theta\_w(2-\theta\_w)$.
-This enables us to describe the behaviour of $\Rnum$ with respect to $\omega$ in the following two cases.
-\begin{itemize}
-\item \textbf{Case I:} If $b_4 \geq a_4$, $\pder\Delta{\omega} < 0$, so $\Rnum$ is always increasing with respect to $\omega$ (i.e., it is always harmful to return tests more rapidly). 
-\item \textbf{Case II:} If $b_4 < a_4$, $\Rnum$ will be decreasing with respect to $\omega$ (i.e., returning tests more rapidly is beneficial) only when $\omega > \omega^*_-.$
-\end{itemize}
-Note that $b_4 \geq a_4$ is characterized by
-\begin{equation}\label{case_1_ineq}
-\Big(\testtarget+1\Big)\theta\_w \geq \testtarget p_I\theta\_c \Longleftrightarrow \Big(\frac{w_S}{w_I}+1\Big)\frac{1}{p_I} \geq \frac{\theta\_c}{\theta\_w}.
-\end{equation}
-We begin with a \textbf{proof of Case I}.
-Suppose that $b_4 \geq a_4$.
-If the roots of $\pder\Delta{\omega}$ are not complex, then we must have $c_4 \geq b_4-a_4$.
-Note that $\omega^*_{-}$ must be negative since the numerator is clearly negative but the denominator is positive.
-Next, note that since $c_4 \geq b_4-a_4$, the numerator of $\omega^*_{+}$ must also be negative, so $\omega^*_{+}$ is negative.
-Thus, in this case, $\pder\Rnum{\omega}$ does not change sign on $(0,\infty)$.
-Checking the sign of \ref{eq:dlindo} for arbitrarily large $\omega$ shows that it is negative (since the $(a_4-b_4)\omega^2$ term dominates and is negative).
-So $\Rnum$ is increasing with respect to $\omega$ on all of $(0,\infty)$.
-
-We now turn our attention to a \textbf{proof of Case II}.
-Suppose that $b_4 < a_4$.
-It follows that $\omega^*_+$ and $\omega^*_-$ are real since $c_4 > 0 > b_4-a_4$.
-Next, note that $\omega^*_-$ is positive since both the numerator and denominator are negative.
-On the other hand, $\omega^*_+$ is negative since the denominator is negative but the numerator is positive (because $c_4 > b_4-a_4$).
-Thus, our task is to understand the sign of $\pder\Delta{\omega}$ around the root $\omega^*_-$.
-Checking the sign of \ref{eq:dlindo} for arbitrarily large $\omega$ shows that it is positive (since the $(a_4-b_4)\omega^2$ term dominates and is positive).
-Likewise, checking the sign for values of $\omega$ close to $0$ shows that it is negative.
-Thus, $\Rnum$ is increasing with respect to $\omega$ when $\omega < \omega^*_-$, and is decreasing $\omega > \omega^*_-$.
-
-Having presented the formal analysis, we now concern ourselves with its biological interpretations.
-We begin by interpreting \ref{case_1_ineq}, under which returning tests more rapidly is \emph{always} harmful.
-Notice that the ratio $\frac{\theta\_c}{\theta\_w}$ is simply a measure of how much more strongly individuals self-isolate when they test positive compared to when waiting for tests.
-Since the rate of test return directly influences the rate at which individuals change from a waiting state to a confirmed-positive state, it is intuitive that $\frac{\theta\_c}{\theta\_w}$ would appear in \ref{case_1_ineq}.
-Next, note that the left-hand side increases when test sensitivity decreases and when targeting of positive individuals is poor.
-This is consistent with our intuition: a false negative that is returned more rapidly will allow an infectious individual to relax their self-isolation, thus increasing transmission.
-Likewise, if individuals tested are mainly susceptible (rather than infectious), then returning tests more slowly would encourage them to self-isolate for longer while awaiting test results.
-Having understood the role of each of the parameters in \ref{case_1_ineq}, a holistic interpretation of this inequality is that returning tests more slowly is helpful when the benefit of extended self-isolation by infected individuals awaiting test results outweighs the benefit of identifying positive cases.
-
-Now we interpret \textbf{Case II}.
-In this case, $\Rnum$ will have a global maximum with respect to $\omega$ at $\omega^*_-$.
-Note that our model assumption that $\rho < \omega$ plays an important role here: if $\omega^*_- < \rho$, then $\Rnum$ will be always decreasing with respect to $\omega$.
-On the other hand, if $\rho < \omega^*_-$, then $\Rnum$ will be increasing with respect to $\omega$ on $(\rho, \omega^*_-)$ and decreasing beyond that.
-
-% %%%%%%%
-\subsection{The effect of testing focus parameter $\testtarget$ on $\Rnum$} \label{app:w}
-
-\begin{align}
-\label{eq:d_del_wis}
-\pder \Delta{w\_{IS}}= \dfrac{(\omega-\rho) (\omega(\omega-\rho\theta\_w)+\gamma(\omega-\rho)) (\theta\_w\gamma+\theta\_c\omega p\_I)}
-{(-\omega^2 \gamma+\omega \gamma \rho-\gamma \rho \omega w\_{IS}-
-\omega \gamma^2+\gamma^2 \rho-\omega^2 p_I \rho w\_{IS})^2},
-\end{align}
-which is a positive quantity. Thus, $\pder \Rnum{w\_{IS}}\leq 0$. Therefore, increasing the focus of testing on the infectious people will result in less transmission.
-
-% %%%%%%%
-\subsection{On Testing Rate and Numerical Singularity} \label{app:singularity}
-
-In this work, we do not present any numerical solutions of the ODEs to investigate typical trajectories.
-However, attempting to do so quickly reveals a problem in the way the model is posed;
-with the scaling of testing ($\sigma$) defined as in the body of the paper (Eq.~\eqref{sigma}), the population in the $S$ compartments appeared to blow up when the system is near the DFE.
-This occurs because once the only untested people are susceptibles, the FOI approaches $\Lambda=0$, and the testing rate $\testing{S} \to \rho N/S\_u$.
-Thus, the first equation of the model \eqref{model} will become
-$d S\_u/dt = - \rho N + \omega S\_n$. Thus changes in $S\_u$ will be independent on $S\_u$, and the decay of the $S\_u$
-population becomes linear rather than exponential --- allowing $S\_u$ to become negative!
-To avoid this problem the testing rate, $\sigma$, should be formulated such that people from the untested compartments will not be tested if they are not there.
-One way to fix this issue, is to consider a maximum testing rate, $\tau$ (1/day).
-In general, we want to test at a rate of $\rho$ across the whole population.
-This won't always be possible, so we impose a maximum rate of $\tau$ per testable person and redefine $\sigma = \frac{\tau \rho N}{\tau W + \rho N}$, with the assumption that $\tau \gg \rho$.
-This modification of $\sigma$ does not affect any of the results we have derived about the invasion of the epidemic from the DFE (i.e., results on $\Rnum$ and $\Delta$).
-
-\end{document}

### 2021-07-19 manuscript_BMB.tex
breaking the sections, filling the BMB template

\RequirePackage{fix-cm}

\documentclass[smallextended]{svjour3}       % onecolumn (second format)

\smartqed  % flush right qed marks, e.g. at end of proof

\usepackage{graphicx}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{grffile}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{amssymb,amsmath}
\usepackage{xcolor}
\usepackage{xspace}
% \usepackage[nameinlink,capitalize]{cleveref}
% \usepackage{cleveref}
\usepackage[margin=1in]{geometry}
\usepackage{lineno}\renewcommand\thelinenumber{\color{gray}\arabic{linenumber}}
\usepackage{pdflscape}
\usepackage{enumerate}
% \usepackage{authblk}
\usepackage{ragged2e}
\usepackage{array}
%
\hyphenation{co-ro-na-vi-rus}
\input{codes/modeldefs.tex}

% please place your own definitions here and don't use \def but
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\newcommand{\fref}[1]{Fig.~\ref{#1}}
\newcommand{\appref}[1]{Appendix~\ref{app:#1}}
% \newcommand{\Rlogo}{R\xspace}
\newcommand{\percap}{\emph{per capita}\xspace}
\newcommand{\Rnum}{\ensuremath{\mathcal{R}_0}\xspace}
\newcommand{\covid}{COVID-19\xspace}
\newcommand{\pro}[1]{\ensuremath{\frac{\partial #1}{\partial \rho}}}
\newcommand{\pder}[2]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder y{x}
\newcommand{\testing}[1]{\ensuremath{{\cal T}_{#1}}\xspace}
\newcommand{\testinghat}[1]{\ensuremath{\hat{\cal T}_{#1}}\xspace}
\newcommand{\testtarget}{\ensuremath{w\_{IS}}}
\newcommand{\oldtesttarget}{\ensuremath{\frac{w_I}{w_S}}\xspace}
\newcommand*\subtxt[1]{_{\textnormal{#1}}}
%% Roman subscript
\DeclareRobustCommand\_{\ifmmode\expandafter\subtxt\else\textunderscore\fi}

%% table stuff
\renewcommand{\arraystretch}{1.5}

% Insert the name of "your journal" with
\journalname{Bulletin of Mathematical Biology}

\begin{document}

\title{Testing and Isolation Efficacy:\\Insights from a Simple Epidemic Model
\thanks{Grants or other notes
about the article that should go on the front page should be
placed here. General acknowledgments should be placed at the end of the article.}
}
% \subtitle{Do you have a subtitle?\\ If so, write it here}

%\titlerunning{Short form of title}        % if too long for running head

\author{Ali Gharouni         \and
        F.M. Abdelmalek \and 
        David J. D. Earn \and
        Jonathan Dushoff \and
        Benjamin M. Bolker
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{Ali Gharouni \at
              Department of Mathematics and Statistics, McMaster University, Hamilton, Canada \\
              Tel.: +100-00-000000\\
              Fax: +100-00-000000\\
              \email{agharoun@uottawa.ca}\\
              \emph{Present address:}
              Department of Mathematics and Statistics, University of Ottawa, Ottawa, Ontario \\%  if needed\\
              \email{agharoun@uottawa.ca}\\
           \and
           F.M. Abdelmalek \at
              Department of Mathematics and Statistics, McMaster University, Hamilton, Canada \\
              \email{abdelf1@mcmaster.ca}\\  
           \and
           David J. D. Earn \at
              Department of Mathematics and Statistics, McMaster University, Hamilton, Canada\\
              Michael G. DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, Canada\\
              \email{earn@math.mcmaster.ca}\\
           \and
          Jonathan Dushoff \at
              Department of Biology, McMaster University, Hamilton, Canada\\
              Michael G. DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, Canada\\
              South African Centre for Epidemiological Modelling and Analysis, University of Stellenbosch, Stellenbosch,               South Africa\\
              \email{dushoff@mcmaster.ca} \\
           \and
          Benjamin M. Bolker \at
              Department of Mathematics and Statistics, McMaster University, Hamilton, Canada \\
              Department of Biology, McMaster University, Hamilton, Canada\\
              Michael G. DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, Canada\\
              \email{bolker@mcmaster.ca}
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
\input{abstract.tex}
\keywords{First keyword \and Second keyword \and More}
% \PACS{PACS code1 \and PACS code2 \and more}
\subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{intro}
\input{introduction.tex}

\section{Methods}
\label{meth}
\input{methods.tex}

\section{Results}
\label{resu}
\input{results.tex}

\section{Discussion}
\label{disc}
\input{discussion.tex}

%\begin{acknowledgements}
%If you'd like to thank anyone, place your comments here
%and remove the percent signs.
%\end{acknowledgements}

% Authors must disclose all relationships or interests that 
% could have direct or potential influence or impart bias on 
% the work: 
%
% \section*{Conflict of interest}
%
% The authors declare that they have no conflict of interest.

% BibTeX users please use one of
\bibliographystyle{spbasic}      % basic style, author-year citations
%\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
\bibliography{SIRlibrary}   % name your BibTeX data base

\clearpage
\appendix
\section{Appendix}
\label{appe}
\input{appendix.tex}

\end{document}

### 2021-07-19 methods.tex
breaking the sections, filling the BMB template

--- 2021-07-19 SIR_manuscript.tex
@@
-\documentclass[12pt]{article}
-\usepackage{natbib}
-\usepackage{hyperref}
-\usepackage{grffile}
-\usepackage{graphicx}
-\usepackage{subcaption}
-\usepackage{amssymb,amsmath,amsthm}
-\usepackage{xcolor}
-\usepackage{xspace}
-\usepackage[nameinlink,capitalize]{cleveref}
-\usepackage{cleveref}
-\usepackage[margin=1in]{geometry}
-\usepackage{lineno}\renewcommand\thelinenumber{\color{gray}\arabic{linenumber}}
-\usepackage{pdflscape}
-\usepackage{enumerate}
-\usepackage{authblk}
-\usepackage{ragged2e}
-\usepackage{array}
-
-\hyphenation{co-ro-na-vi-rus}
-\input{codes/modeldefs.tex}
-
-% https://tex.stackexchange.com/questions/12703/how-to-create-fixed-width-table-columns-with-text-raggedright-centered-raggedlef
-\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-
-\newcommand{\fref}[1]{Fig.~\ref{#1}}
-\newcommand{\appref}[1]{Appendix~\ref{app:#1}}
-\newcommand{\Rlogo}{R\xspace}
-\newcommand{\percap}{\emph{per capita}\xspace}
-\newcommand{\Rnum}{\ensuremath{\mathcal{R}_0}\xspace}
-\newcommand{\covid}{COVID-19\xspace}
-\newcommand{\pro}[1]{\ensuremath{\frac{\partial #1}{\partial \rho}}}
-\newcommand{\pder}[2]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder y{x}
-\newcommand{\testing}[1]{\ensuremath{{\cal T}_{#1}}\xspace}
-\newcommand{\testinghat}[1]{\ensuremath{\hat{\cal T}_{#1}}\xspace}
-\newcommand{\testtarget}{\ensuremath{w\_{IS}}}
-\newcommand{\oldtesttarget}{\ensuremath{\frac{w_I}{w_S}}\xspace}
-\newcommand*\subtxt[1]{_{\textnormal{#1}}}
-%% Roman subscript
-\DeclareRobustCommand\_{\ifmmode\expandafter\subtxt\else\textunderscore\fi}
-
-\newcommand{\comment}{\showcomment}
-\newcommand{\showcomment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}
-\newcommand{\nocomment}[3]{}
-
-\newcommand{\fady}[1]{\comment{cyan}{Fady}{#1}}
-\newcommand{\ali}[1]{\comment{magenta}{Ali}{#1}}
-\newcommand{\jd}[1]{\comment{blue}{JD}{#1}}
-\newcommand{\david}[1]{\comment{orange}{DJDE}{#1}}
-\newcommand{\bmb}[1]{\comment{red}{BMB}{#1}}
-\newcommand{\todo}[1]{\comment{red}{TODO}{#1}}
-\newcommand{\new}[1]{\comment{red}{NEW}{#1}}
-\newcommand{\needref}{\textcolor{red}{(REFS?)}}
-
-%% DE: table stuff
-\renewcommand{\arraystretch}{1.5} % more interline spacing
-%% https://tex.stackexchange.com/questions/12703/how-to-create-fixed-width-table-columns-with-text-raggedright-centered-raggedlef
-% \usepackage{ragged2e}
-% \usepackage{array} % for m{...}
-
-\theoremstyle{definition} % amsthm only
-\newtheorem{proposition}{Proposition}
-\newtheorem{theorem}{Theorem}
-
-\bibliographystyle{apalike}
-
-\title{Testing and Isolation Efficacy:\\Insights from a Simple Epidemic Model}
-\author[1]{Ali Gharouni}
-\author[1]{F.M. Abdelmalek}
-\author[1,3]{David J. D. Earn}
-\author[2,3,4]{Jonathan Dushoff}
-\author[1,2,3]{Benjamin M. Bolker}
-
-\affil[1]{Department of Mathematics and Statistics, McMaster University, Hamilton, Canada}
-\affil[2]{Department of Biology, McMaster University, Hamilton, Canada}
-\affil[3]{Michael G. DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, Canada}
-\affil[4]{South African Centre for Epidemiological Modelling and Analysis, University of Stellenbosch, Stellenbosch, South Africa}
-
-\begin{document}
-\maketitle
-
-\linenumbers
-
-% %%%%%%%
-\section{Abstract}
-
-Testing individuals for pathogens can affect the spread of epidemics. 
-Understanding how individual-level processes of sampling and reporting test results can affect community- or population-level spread is a dynamical modeling question.
-The effect of testing processes on epidemic dynamics depends on factors underlying implementation, particularly testing intensity and on whom testing is focused.
-Here, we use a simple model to explore how the individual-level effects of testing might directly impact population-level spread.
-Our model development was motivated by the \covid epidemic, but has generic epidemiological and testing structures. 
-To the classic SIR framework we have added a \percap testing intensity, and compartment-specific testing weights, which can be adjusted to reflect different testing emphases --- surveillance, diagnosis, or control.
-We derive an analytic expression for the relative reduction in the basic reproductive number due to testing, test-reporting and related isolation behaviours.
-Intensive testing and fast test reporting are expected to be beneficial at the community level because they can provide a rapid assessment of the situation, identify hot spots, and may enable rapid contact-tracing.
-Direct effects of fast testing at the individual level are less clear, and may depend on how individuals' behaviour is affected by testing information.
-Our simple model shows that under some circumstances both increased testing intensity and faster test reporting can \emph{reduce} the effectiveness of control, and allows us to explore the conditions under which this occurs.
-Conversely, we find that focusing testing on infected individuals always acts to increase effectiveness of control.
-
-% %%%%%%%
-\section{Introduction}
-
-The observed dynamics of the \covid epidemic have been driven both by epidemiological processes (infection and recovery) and by testing processes (testing and test reporting).
-In addition to shaping epidemic observations (via case reports), testing processes also alter epidemiological dynamics \citep{peto2020covid,taipale2020population}.
-Because individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may also do so, testing will generally increase the number of people who are isolating and hence reduce epidemic growth rates.
-We developed a mechanistic model that incorporates epidemic processes and testing in order to explore the effects of testing and isolation on epidemic dynamics.
-
-If testing influences behaviour, then epidemic dynamics will depend on who gets tested.
-The impacts of testing will depend both on testing intensity (tests performed per day) and on how strongly testing is focused on people who are infectious.
-This level of focus depends in turn on the purpose and design of testing programs. 
-When testing is done for the purposes of disease surveillance \citep{foddai2020base}
-tests are typically conducted randomly (or using a stratified random design) across the population in order to make an unbiased assessment of population prevalence.
-
-Over the course of the \covid pandemic, however, the vast majority of testing has been done with other goals --
-primarily diagnostic (determining infection status for clinical purposes) \citep{phua2020intensive,who2020global}, or for control (determining  infection status in order to isolate cases that have been found by contact tracing) \citep{aleta2020modelling,kucharski2020effectiveness,grassly2020comparison,smith2020adherence}, which we characterize as \emph{targeted} testing strategies.
-In these situations, testing probabilities can differ sharply across epidemiological compartments; in our dynamical model, we will characterize these probabilities by assigning a testing weight to each compartment that determines the \emph{relative} probability that an individual in that compartment will be selected for testing (see Methods).
-
-Diagnostic testing focuses on people with infection-like symptoms; thus the relative testing weights for infected people will depend on the relative probability of infected people having symptoms.
-For \covid infection, the testing weights will depend on the proportion of asymptomatic infections, the time spent pre-symptomatic vs.\ symptomatic during the course of an infection, and on the incidence of \covid-like symptoms among people in the population \emph{not} infected with \covid.
-Testing for epidemic control focuses on known contacts of infected people; in this case the testing weights for infected vs.\ uninfected people will depend on the probability of infection given contact, as well as the effectiveness of the system for identifying suspicious contacts.
-
-When a new infectious disease emerges, it is important to determine whether it will grow exponentially in a susceptible population, and if so at what rate $r$ \citep{Ma+14}.
-The condition for positive exponential growth ($r>0$) is commonly expressed as $\Rnum>1$, where the basic reproduction number $\Rnum$ is the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}.
-Although the value of $\Rnum$ cannot completely characterize the dynamics of our model \citep{shaw2021what}, it does give a simple and widely accepted index for the difficulty of control, as well as an indication of the likely final size of an epidemic \citep{ma2006generality,miller2012note}.
-
-In order to understand the effect of testing processes on epidemic dynamics, we expanded one of the simplest mechanistic epidemic models---the standard deterministic SIR model\citep{KermMcKe27,AndeMay91}---to include testing components.
-This model provides a sensible platform to link the modeling of epidemic and testing components and study their interaction.
-We studied the effects of testing intensity, rate of test return, and isolation efficacy, on transmission probability and epidemic dynamics when different levels of testing focus (from random to highly targeted) are in place.
-
-% %%%%%%%
-\section{Methods}
+% \section{Methods}
 
@@
 The different ranges of test return rates $\omega$ for the cases of low and high testing intensities is due to the restriction $\rho<\omega$, which is a requirement for a feasible DFE (\ref{dfe}).
-
-% %%%%%%%
-\section{Results}
-
-We presented $\Rnum$ as the product of the classical reproduction number, $\beta/\gamma$, and the proportional reduction due to testing and isolation, $1-\Delta$, \eqref{R0}.
-We can use the formula for $\Delta$ \eqref{eq:del4a} to make a number of straightforward inferences about parameters that affect $\Rnum$ monotonically, i.e., for which the associated partial derivative of $\Delta$ always has the same sign (see Appendices).
-
-\begin{enumerate}
-
-\item \label{p1:eta} Increasing isolation efficacy for waiting ($\theta\_w$) and confirmed-positive ($\theta\_c$) individuals always increases $\Delta$ (Eqs.~\ref{eq:del4_theta},~\ref{eq:d_del4_thetac},~\ref{eq:d_del4_thetaw});
-\item \label{p1:rho} Higher testing intensity $\rho$ increases $\Delta$ if
-%% \begin{enumerate} \item
-testing is random (all $w_X$ equal) or testing intensity ($\rho$) is small (Eq.~\ref{eq:dd3dr}).
-%% \end{enumerate}
-\item \label{p1:omega} Increasing the rate of test return ($\omega$) always increases $\Delta$ if waiting individuals do not isolate ($\theta\_w=0$) (Eq.~\ref{eq:dlindo}).
-\item \label{p1:w} Increasing testing focus, i.e., changing the testing weights from random ($w_S=w_I$) toward targeted  ($w_S<w_I$), always increases $\Delta$ (Eq.~\ref{eq:d_del_wis}).
-\end{enumerate}
-
-\newpage
-\begin{figure}[h!]
-\centering
-\begin{subfigure}[t]{.49\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_random.pdf}
-\caption{}\label{p.a}
-\end{subfigure}
-%
-\begin{subfigure}[t]{.49\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_TTI.pdf}
-\caption{}\label{p.b}
-\end{subfigure}
-\caption{
-{\bf Effectiveness of testing and isolation in reducing $\Rnum$ at low \percap testing intensity ($\rho$).}
-Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4a}), over a range of testing and isolation parameters. Parameter values (Table~\ref{tab:params}):
-$\beta=\betaparam/$day, $1/\gamma= \invgammaparam$ days (baseline $\Rnum=\Rnumparam$, $r=\rparam$); $\omega \in [1/12,2]/$day;  $\rho \in [0,0.013]/$day per capita; $\theta\_w$ and $\theta\_c$ vary between 0 (no effect of isolation) and 1 (complete elimination of transmission); $p_S=0$, $p_I=1$ and $p_R=0.5$.
-Only parameter sets where $\theta\_c \geq \theta\_w$ (confirmed-positive individuals isolate more effectively than waiting individuals) are shown; the alternative case, $\theta\_w > \theta\_c$, is unrealistic.
-Contours of $\Delta$ are plotted for (a) random testing ($w_S=w_I=w_R=1$) and (b) targeted testing ($w_S=0.3$; $w_I=w_R=1$).
-}
-\label{pan}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\begin{subfigure}[t]{.49\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_random2.pdf}
-\caption{}
-\end{subfigure}
-%
-\begin{subfigure}[t]{.49\textwidth}
-\centering
-\includegraphics[width=\linewidth]{codes/R0contour_TTI2.pdf}
-\caption{}
-\end{subfigure}
-\caption{
-  {\bf Effectiveness of testing and isolation in reducing $\Rnum$ at high \percap testing intensity.}
-  Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4a}), over a range of testing and isolation parameters.
-Parameters as in \fref{pan} except: $\omega \in [1/5,2]/$day, $\rho \in [0,1/5)/$day.
-As in \fref{pan}, subplots show (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
-}
-\label{pan2}
-\end{figure}
-
-However, there are also two specific cases where $\Delta$ changes non-monotonically, in counterintuitive directions, as a function of testing and isolation parameters.
-
-\begin{itemize}
-\item We would generally expect increasing testing delays to increase \Rnum, thus decreasing effectiveness of control $\Delta$.
-This is in fact what happens when waiting individuals do not isolate ($\theta\_w =0$, top row of \fref{pan}) --- as we move to the right within each plot in this row, $\Delta$ decreases.
-However, when waiting individuals isolate ($\theta\_w>0$), we more often see the opposite effect: longer testing delays lead to a greater control effect $\Delta$ (reduced $\Rnum$).
-The reason is that people waiting for negative tests are assumed to continue to isolate; this applies both to susceptibles and to people who became infected while waiting for negative test results.
-This effect outweighs the effect of confirmed individuals isolating, except when this isolation parameter ($\theta\_c$) is substantially greater than $\theta\_w$.
-This result depends on the idea that, all else equal, people who have to wait longer for test results isolate at the same level (but for a longer time) as they would if the wait were shorter.
-\item \fref{pan} also shows that greater testing intensity (increasing $\rho$) generally increases the effectiveness of control (moving up in each panel).
-However, this relationship can be reversed at very high testing intensities (provided testing is targeted, and $\theta\_w$ is relatively small; \fref{pan2}(b), right three panels of top row).
-It is theoretically possible for increasing testing intensity to \emph{increase} \Rnum because more rapid testing leaves more susceptibles in the ``waiting-for-negative-results'' category at the DFE; if these people become infected while waiting, they will need to wait for their negative test result before they can be tested again, receive a positive test, and then begin self-isolating.
-This effect is usually weak compared to the beneficial effects of testing.
-
-\end{itemize}
-
-%%%%%%%%%%%%
-\section{Discussion}
-
-In this paper, we have developed and analyzed a simple compartmental model that combines epidemiological dynamics---as defined by a simple SIR model---with the dynamics of testing and isolation.
-Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness}).
-Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
-
-Many of the qualitative results we have derived confirm simple, common-sense intuitions.
-In particular, we can generally decrease $\Rnum$ by increasing isolation efficacy or testing intensity; returning tests faster, if individuals do not isolate while they are waiting for results; or increasing testing focus to target individuals who are likely to be infectious (e.g., symptomatic people or close contacts of known infections).
-
-However, we did find two surprising patterns: under some conditions longer delays in returning tests can reduce epidemic spread, and increasing testing rates can increase spread.
-
-Over broad regions of parameter space, decreasing $\omega$---i.e., \emph{slowing} the rate at which test results are returned---\emph{decreases} $\Rnum$ (for random testing, the parameter region is $\theta\_w \gtrsim 0.25$; for targeted testing, $\theta\_w \geq 0.25$ and either $\theta\_c \ge 0.5$ or $1/\omega > 5$; see Table~\ref{tab:params} for parameter definitions).
-This result is counterintuitive and would not be expected by public health authorities that have invested a great deal of effort in reducing delays from testing to results.
-Dynamically, this effect occurs because speeding up test returns shortens the isolation period of uninfected individuals (for infected people it only shortens the time to progression to the isolation level of the confirmed-positive compartment).
-Slowing test returns increases $\Rnum$ only if the proportion of infectives in the tested population is high and isolation is relatively strong among people waiting for test results.
-
-While slowing test returns does decrease $\Rnum$ over broad regions of parameter space in our model, there are several real-world processes missing from our model that make it unlikely that slowing test returns would actually be an effective public health measure.
-First, we do not model the primary benefit of rapid testing, i.e., detecting and containing outbreaks while they are still in progress.
-This process could be modeled phenomenologically by making the testing focus more targeted as an increasing proportion of cases is detected, because finding infections allows tests to be concentrated on their connections.
-Second, individuals may become less likely to maintain isolation if they are required to do so for longer; phenomenologically, we could allow  effectiveness of isolation in the waiting population to be an increasing function of test-return speed, or we could introduce a separate ``waiting, but no longer isolating'' compartment that individuals entered from the ``waiting, isolated'' compartment at a specified rate.
-Finally, if one wants to decrease the overall transmission rate of the population there are more effective methods than keeping tested people in limbo; these include masking, ventilation, distancing measures, retail and event closures, and stay-at-home orders.
-
-The other counterintuitive result from our analysis is that, for sufficiently high testing intensity $\rho$, further increasing testing intensity can actually \emph{increase} $\Rnum$ (e.g., \fref{pan2}(b), upper right panel [$\theta\_c=1$, $\theta\_w=0$]).
-This phenomenon can occur because we are considering the DFE in the presence of testing; thus there is an equilibrium distribution of susceptibles between the $S\_n$ (untested) and $S\_u$ (waiting) compartments even as the disease approaches extinction.
-A higher rate of testing leads to a greater proportion of individuals waiting for negative tests at the DFE.
-If infected, individuals in this group will take longer to be tested again and to subsequently isolate (because they must wait for their negative tests to be returned before being tested again).
-If isolation in this group ($\theta\_w$) is low, this effect can under some (relatively rare) circumstances (high $\theta\_c$, low $\omega$, high $\rho$) allow $\Rnum$ to increase with testing intensity.
-We can show that this phenomenon occurs \emph{only} under targeted testing ($w_I > w_S$), but we have not yet found a simple explanation of why it cannot occur under random (unfocused) testing.
-This phenomenon is also unlikely to occur in the real world.
-In particular, it depends on levels of testing that are unrealistically high (at least in large, general-population settings).
-
-Although we model the testing process in more detail than typical epidemiological models, one place where more detail could be informative is in the processes determining the testing weights $\{w_S, w_I, w_R\}$.
-While random testing, as done for surveillance purposes, unambiguously leads to equal testing weights, making precise quantitative connections between public-health practices and testing weights is difficult in other contexts.
-The testing weights reflect the correlation between an individual's risk of infection and their likelihood of being tested due to age, occupation, geographic location, etc..
-This correlation is influenced, among many other factors, by the proportion of the uninfected population with COVID-like symptoms (e.g., due to seasonal upper respiratory tract infections); the concentration of transmission and testing in hot spots such as long-term care facilities and high-density workplaces; the overall testing intensity (and hence, e.g., restriction to symptomatic individuals); and the proportion of COVID-infected people who are symptomatic.
-Future steps should explore mathematically tractable ways to model some of these factors more precisely.
-For example, separating the infected class into exposed, symptomatic, and a- or pre-symptomatic compartments and allowing the testing weights to vary across non-symptomatic (exposed/asymptomatic/presymptomatic) vs.\ symptomatic compartments could reflect the allocation of tests for diagnostic purposes (targeting symptomatic individuals) vs.\ contact-tracing (targeting infected but non-symptomatic individuals) vs.\ screening (relatively equal weights, depending on the venue).
-Alternatively, one could make the testing weights depend on the testing intensity or test-return rate as suggested above.
-Whatever complexity is added would probably put the model beyond reach of the analytical methods we have used in this paper, but one could still use semi-numerical methods such as constructing the next-generation matrix and using it to evaluate the derivatives of $\Rnum$ with respect to the parameters numerically.
-
-Although testing and tracing is a key part of infection control strategies, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts \citep{endo2020implication,hellewell2020feasibility,jenness2020modeling}, rather than analyzing simple but general models of the feedback between testing and transmission dynamics.
-There have been several modeling studies of testing and tracing dynamics and their interaction with epidemiological dynamics.
-In the context of repeated screening and random testing of isolated populations (such as the members of a university), \cite{bergstrom2020frequency} provided analytical results quantifying the effects that proactive screening of asymptomatic individuals and isolation of confirmed-positive cases could have in reducing the spread of disease.
-\cite{rogers2021high} simulated a SEIR model with testing and isolation; they similarly suggest a strategy of rapid testing with antigen tests and the subsequent isolation of confirmed-positive individuals.
-\cite{friston2021testing} model the effects of self-isolation on testing and tracing with a focus on projections under different testing and tracing scenarios.
-They conclude that the emergence of a second wave depends almost primarily on the rate at which immunity is lost and that it is necessary to track asymptomatic individuals in order to control the outbreak.
-Our modeling approach differs from these previous efforts in that it examines the effects of test-return rates and of different levels of testing focus, from random to highly targeted.
-We hope this paper will inspire further explorations of the fundamental properties of epidemic models that incorporate explicit testing processes.
-
-% %%%%%%%
-\bibliography{SIRlibrary}
-
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\clearpage
-% \widetext
-\appendix
-\section{Appendix}
-\setcounter{equation}{0}
-\setcounter{figure}{0}
-\setcounter{table}{0}
-% \setcounter{page}{1}
-\makeatletter
-\renewcommand{\theequation}{A\arabic{equation}}
-\renewcommand{\thefigure}{A\arabic{figure}}
-\renewcommand{\bibnumfmt}[1]{[A#1]}
-\renewcommand{\citenumfont}[1]{A#1}
-
-\subsection{Model and calculation of $\Rnum$}\label{app:R0}
-
-The model in the form of a system of ordinary differential equations is 
-\begin{subequations}\label{model}
-\begin{align}
- d S\_u/dt &= -\Lambda S\_u - \testing{S} S\_u + \omega S\_n, \\
- d S\_n/dt &= -(1-\theta\_w)\Lambda S\_n + (1-p_S) \testing{S} S\_u - \omega S\_n, \\
- d S\_p/dt &= -(1-\theta\_w)\Lambda S\_p + p_S \testing{S} S\_u - \omega S\_p, \\
- d S\_c/dt &= -(1-\theta\_c)\Lambda S\_c + \omega S\_p, \\
- d I\_u/dt &= \Lambda S\_u - \testing{I} I\_u + \omega I\_n  - \gamma I\_u, \\
- d I\_n/dt &= (1-\theta\_w)\Lambda S\_n + (1-p_I) \testing{I} I\_u - \omega I\_n -\gamma I\_n, \\
- d I\_p/dt &= (1-\theta\_w)\Lambda S\_p + p_I \testing{I} I\_u - \omega I\_p -\gamma I\_p, \\
- d I\_c/dt &= (1-\theta\_c)\Lambda S\_c + \omega I\_p - \gamma I\_c, \\
- d R\_u/dt &= \gamma I\_u - \testing{R} R\_u + \omega R\_n, \\
- d R\_n/dt &= \gamma I\_n + (1-p_R) \testing{R} R\_u - \omega R\_n, \\
- d R\_p/dt &= \gamma I\_p + p_R \testing{R} R\_u  - \omega R\_p, \\
- d R\_c/dt&= \gamma I\_c + \omega R\_p, \\
- dN/dt &= \omega (S\_n + I\_n + R\_n),  \\
- dP/dt &= \omega(I\_p + R\_p) ,
-\end{align}
-\end{subequations}
-%
-(see Table \ref{tab:params} for parameter definitions).
-The next generation matrix for this model is $G = F V^{-1}$, where matrix $F$ represents the inflow of new infection to the infected compartments and matrix $V$ represents the flow in the infected compartments when the population is totally susceptible.
-Matrices $F$ and $V$ are
-\begin{align}
-\label{F}
-F =& \beta/N \left[ \begin {array}{cccc} 
-S\_u^* & (1-\theta\_w) S\_u^* & (1-\theta\_w) S\_u^* & (1-\theta\_c) S\_u^*\\
-(1-\theta\_w)S\_n^* & (1-\theta\_w)^2 S\_n^* & (1-\theta\_w)^2 S\_n^* & (1-\theta\_w)(1-\theta\_c) S\_n^* \\ 
-0&0&0&0\\
-0&0&0&0
- \end {array} \right] \\ \notag
- =&\beta/N \left[ \begin {array}{c} S\_u^* \\ (1-\theta\_w)S\_n^* \\ 0 \\ 0 \end {array} \right]
-        \left[ \begin {array}{cccc} 1,   1-\theta\_w,   1-\theta\_w,   1-\theta\_c \end {array} \right],
-\end{align}
-and 
-\begin{align}
-\label{V}
-V= \left[ \begin {array}{cccc}  
-\testinghat{I}+\gamma&-\omega&0&0\\
--(1-p_I)\testinghat{I}&\omega+\gamma&0&0\\
--p_I \testinghat{I}&0&\omega+\gamma&0\\
-0&0&-\omega&\gamma
-\end {array} \right].
-\end{align}
-The matrix inverse of $V$ is 
-\begin{align}
-\label{vinv}
-V^{-1} =
-\frac{1}{\gamma C}
-\left[ \begin {array}{cccc}
-\gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)&0&0\\ \noalign{\medskip}
-\gamma(\omega+\gamma)(1-p_I) \testinghat{I} & \gamma(\omega+\gamma)(\testinghat{I}+\gamma)&0&0 \\ \noalign{\medskip}
-\gamma(\omega+\gamma) p_I \testinghat{I} & \gamma\omega p_I \testinghat{I} & C \gamma/(\omega+\gamma) & 0 \\ \noalign{\medskip}
-\omega(\omega+\gamma) p_I \testinghat{I} & \omega^2 p_I \testinghat{I} & C \omega/(\omega+\gamma) & C
-\end {array} \right],
-\end{align}
-where $C= \big( \gamma(\omega+\gamma)+(\gamma+\omega p_I)\testinghat{I} \big) (\omega+\gamma)$ and $\testinghat{I}$ is the \percap testing rate for the infected people and represented in Eq.~\eqref{eq:fi}.
-Note that all the columns of matrix $V^{-1}$ summ up to $1/\gamma$.
-
-The particular form of $F$ with two rows of zeros at the bottom results in the following blocked form of matrix $G$.
-\begin{equation}
-\label{mat:G}
-G = \left[ \begin {array}{cc}
-G_{11}&G_{12}\\
-0&0
-\end {array} \right],
-\end{equation}
-where both blocked matrices $G_{11}$ and $G_{12}$ are 2 by 2.
-Given the upper triangular form of matrix $G$, the basic reproduction number $\Rnum$ (defined as the spectral radius of matrix $G$) is only determined by the blocked matrix $G_{11}$,
-\begin{equation}
-\label{G11}
-G_{11} = \frac{\beta}{\gamma   C} 
-\left[ \begin {array}{c} (\omega-\rho)/\omega \\ (1-\theta\_w)\rho/\omega \end {array} \right]
-\left[ \begin {array}{cccc} 1,   1-\theta\_w,   1-\theta\_w,   1-\theta\_c \end {array} \right]
-\left[ \begin {array}{cc}
-\gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)\\ \noalign{\medskip}
-\gamma(\omega+\gamma)(1-p_I)\testinghat{I} & \gamma(\omega+\gamma)(\testinghat{I}+\gamma) \\ \noalign{\medskip}
-\gamma(\omega+\gamma) p_I \testinghat{I} & \gamma\omega p_I \testinghat{I} \\ \noalign{\medskip}
-\omega(\omega+\gamma) p_I \testinghat{I} & \omega^2 p_I \testinghat{I}
-\end {array} \right].
-\end{equation}
-It is notable that matrix $F$ \eqref{F} has rank one and consequently $G_{11}$ does so. That is $G_{11}$ has only one non-zero eigenvalue which is $\Rnum$.
-
-The expression of $\Rnum$ has a complicated form with all of the model parameters involved.
-This expression can be simplified and represented given the specific form of matrix $G_{11}$ \eqref{G11}.
-For the purpose of simplicity we present $\Rnum$ in the manuscript in terms of expressions $C$, $C_1$ and $C_2$, specified in \eqref{eq:C12}.
-
-It remains difficult to show that the reproduction number $\Rnum$ is decreasing with respect to \percap testing intensity, $\rho$, and the speed of the test return, $\omega$, for the feasible ranges of the parameters, that is
-\begin{align}
-\label{cond}
-& \omega>0, \\
-& 0 \leq \rho<\omega,\\ 
-& 0 \leq \theta\_w \leq \theta\_c \leq 1, \\
-& \oldtesttarget\geq 1.
-\end{align}
-In realistic cases the testing rate $\rho$ is very small (i.e., only a small fraction of the population can be tested every day); it is thus reasonable to use a linear approximation of $\Rnum$ for $\rho \ll 1$ to analyze the behaviour of $\Rnum$ with respect to $\omega$ (see section \appref{omega}).
-In the next section we provide an equivalent representation of $\Rnum$ in order to show that increasing testing intensity typically decreases $\Rnum$.
-
-% %%%%%%%
-\subsection{More testing intensity may decrease $\Rnum$}\label{app:rho}
-
-This section shows that $\pro{\Delta}$ can be positive or negative, with $\Delta$ defined in Eq.~\eqref{eq:C12}, and thus $\pro{\Rnum} < 0$, where $\Rnum$ is given in Eq.~\eqref{R0}.
-We rewrite matrix $G_{11}$ in \eqref{G11} in the following form to simplify the calculations:
-\begin{equation}
-\label{G112}
-G_{11} = \frac{\beta}{\gamma C} 
-\left[ \begin {array}{c}  S\_u^*/N \\ (1-\theta\_w) S\_n^*/N  \end {array} \right]
-\left[ \begin {array}{cccc} 
-C-C_1, C-C_2\end {array} \right],
-\end{equation}
-where $C$ is the same as the one in Eq. \eqref{eq:C12}, i.e., 
-$$C=(\omega+\gamma)(\gamma(\omega+\gamma)+(\omega p_I+\gamma) \testinghat{I}),$$
-and $C_1$ and $C_2$ are 
-\begin{align*}
-% \label{eq:C12}
-C_1 =& (\omega+\gamma)(\theta\_w   \gamma+\theta\_c  \omega  p_I) \testinghat{I},\\
-C_2 =& \big( \omega\gamma(1+p_I)\testinghat{I}+\gamma^2 (\omega+\gamma+\testinghat{I})\big)\theta\_w + \omega^2 p_I \testinghat{I} \theta\_c,
-\end{align*}
-where $\testinghat{I}$ is given in Eq.~\eqref{eq:fi}.
-Note that for analysis brevity, we let $N=1$, thus $S\_u^*$ and $S\_n^*$ are in the scale of 0 to 1.
-$\Rnum$ is in the same form as in Eq. \eqref{R0}  
-$$\Rnum= \frac{\beta}{\gamma} (1-\Delta),$$
-where 
-\begin{equation*}
-\label{eq:del4}
-\Delta= \frac{1}{C}\big(C_1 S\_u^*+(C_2(1-\theta\_w)+C \theta\_w) S\_n^*\big).
-\end{equation*}
-
-The first goal
-is to explore how changes in isolation, $\theta\_w$ and $\theta\_c$, affects $\Rnum$.
-Mathematically we would like to verify the sign of $\pder \Rnum{\theta\_w}$ and $\pder \Rnum{\theta\_c}$.
-We start with simplifying $\Delta$ \eqref{eq:del4} by factoring $\theta\_w$ and $\theta\_c$ in Eq.~\eqref{eq:del4}.
-Thus, $\Delta$ can be rewritten as
-\begin{equation}
-\label{eq:del4_theta}
-\Delta= \frac{1}{C}\Big(
--D_1 S\_n^* \theta\_w^2 +
-\big(-\omega^2 p_I \testinghat{I} S\_n^*\theta\_c+ D_2 S\_n^*+\gamma \testinghat{I}(\omega+\gamma) \big)\theta\_w+
-(\omega+\gamma S\_u^*)\omega p_I \testinghat{I} \theta\_c
-\Big),
-\end{equation}
-where
-\begin{align}
-\label{eq:del4_d1d2}
-D_1=& (\omega+\gamma)\gamma^2+(\omega+\gamma+\omega p_I)\gamma \testinghat{I}, \\
-D_2=& (3\omega+2\gamma)\gamma^2+(\omega+\gamma+2\omega p_I)\gamma \testinghat{I}+(\gamma+p_I \testinghat{I})\omega^2.
-\end{align}
-$\Delta$, Eq.~\eqref{eq:del4_theta}, is linear in $\theta\_c$ with a positive coefficient. thus
-\begin{equation}
-\label{eq:d_del4_thetac}
-\pder\Delta{\theta\_c}=1/C(\gamma S\_u^*+\omega(1-\theta\_w S\_n^*))\omega p_I \testinghat{I}.
-\end{equation}
-This results in increasing $\Delta$, thus decreasing $\Rnum$ with respect to $\theta\_c$, that is $\pder \Rnum{\theta\_c}\leq 0$. Note that $C$ is independent of $\theta\_c$ and $\theta\_w$.
-
-With a similar logic, $\Delta$ \eqref{eq:del4_theta} is a concave-down quadratic equation in $\theta\_w$, given by
-\begin{equation}
-\label{eq:del4_thetaw}
-1/C \Big( -D_1 S\_n^*\theta\_w^2+
-\big(-\omega^2 p_I \testinghat{I} S\_n^*\theta\_c+D_2 S\_n^*+\gamma \testinghat{I}(\omega+\gamma) \big)\theta\_w\Big).
-\end{equation}
-We show that the feasible range of $\theta\_w$ lies between 0 and the vertex of this parabola where the parabola is increasing in $\theta\_w$, and so does $\Delta$ which results in inferring $\pder\Rnum{\theta\_w}\leq 0$.
-It is enough to show that partial derivative of the expression \eqref{eq:del4_thetaw} with respect to $\theta\_w$ at $\theta\_w=1$ is non-negative. It follows that
-\begin{align}
-\label{eq:d_del4_thetaw}
-\pder\Delta{\theta\_w}\bigg\rvert_{\theta\_w=1} =&
-1/C\Big( (D_2-2D_1-\omega^2 p_I \testinghat{I} \theta\_c) S\_n^*+\gamma \testinghat{I}(\omega+\gamma) \Big) \\\notag
-=& 1/C\Big( (\gamma(\omega+\gamma)+\gamma\omega^2+(1-\theta\_c)\omega^2 p_I \testinghat{I} ) S\_n^*
-+\gamma(\omega+\gamma) \testinghat{I} (1-S\_n^*) \Big),
-\end{align}
-which is a positive quantity, given that $\theta\_c$ and $S\_n^*$ vary between 0 and 1.
-
-The second goal is to explore how changes in \percap testing intensity $\rho$ affects $\Rnum$.
-Mathematically we would like to verify the sign of $\pder\Rnum{\rho}$, which specifically depends on $\pder\Delta{\rho}$.
-We use the derived expressions for $S\_u^*$ and $S\_n^*$, given by Eqs.~\eqref{dfe}, in $\Delta$ \eqref{eq:del4}.
-Also, we define $\phi = \testinghat{S} = \frac{\rho \omega}{\omega-\rho}$, to reparameterize $\rho$.
-This is mainly to avoid singularity in $\testinghat{I}$ \eqref{eq:fi}, when testing intensity $\rho$ is very close to the rate of test return $\omega$.
-Thus, $\rho$ is reparameterized as
-\begin{equation}
-\label{eq:phi}
-\rho=\frac{\omega \phi}{\omega+\phi}.
-\end{equation}
-This one-to-one monotonic reparameterization enables us to simplify the mathematical expressions and explore the simpler $\pder\Delta{\phi}$ instead of the complicated $\pder\Delta{\rho}$.
-Defining $\testtarget \equiv \oldtesttarget$, the derivative is 
-\begin{align}
-\label{eq:dd3dr}
-\partial\Delta/\partial\phi=& \frac{1}{d_3} (a_3 \phi^2+b_3 \phi+c_3),
-\end{align}
-where
-\begin{align}
-\label{eq:abcd2}
-a_3& = \testtarget \Bigg(
-&& (1-\theta\_w) (1+\testtarget)\theta\_w \gamma^3 +(1-\theta\_c) p_I^2 \theta\_w \testtarget \omega^3 \\ \notag
-& ~ &&+\Big( \big( (1-\theta\_w-\testtarget) \theta\_c +(3-2\theta\_w) \theta\_w\testtarget \big) p_I +(1-\theta\_w) (1+\testtarget)\theta\_w  \Big) \omega \gamma^2 \\ \notag
-& ~ &&+\Big(
-\big((1-\theta\_w-\theta\_w \testtarget) \theta\_c+ (2-\theta\_w) \theta\_w \testtarget \big) p_I
-+(2 \theta\_w-\theta\_w^2-\theta\_c) \testtarget p_I^2
-\Big) \omega^2 \gamma \Bigg), \\ \notag
-b_3& = && \hspace{-1cm} 2\testtarget (\omega+\gamma)\gamma\Big(
-(\omega+\gamma+\omega p_I)(2-\theta\_w) \gamma \theta\_w+ (1-\theta\_w) \omega^2 p_I\theta\_c+\omega^2 p_I\theta\_w
-\Big) ,\\ \notag
-c_3& = && \hspace{-1cm} (\omega+\gamma)^2 \gamma\Big(
-(2-\theta\_w) \gamma^2 \theta\_w+
-(1+\testtarget) \omega \gamma \theta\_w+
-\testtarget \omega^2 p_I\theta\_c \Big),\\ \notag
-d_3&  = && \hspace{-1cm} \frac{(\omega+\gamma)}{\omega}  \Big((\omega p_I+\gamma)\testtarget\phi + (\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2.
-\end{align}
-Note that $\phi\geq 0$, also $b_3$, $c_3$ and $d_3$ are all positive. However $a_3$ can be positive or negative.
-If $a_3\geq 0$, $\partial\Delta/\partial\phi \geq 0$ for all feasible range of parameters, thus $\pro\Rnum \leq 0$.
-It is straightforward to show that $a_3\geq 0$ when testing is random, i.e., $w\_S=w\_I=1$.
-If $a_3 < 0$, then the quadratic expression in the numerator of \eqref{eq:dd3dr} has a positive root, $\phi^*$, such that for $\phi>\phi^*$, $\partial\Delta/\partial\phi < 0$.
-
-An example of this countervailing effect of $\phi$, and consequently $\rho$, on $\Rnum$ occurs when $\theta\_w=0$ and $\theta\_c=1$.
-This is illustrated in the top-right panel of the \fref{pan2} panel (b), where the strength of isolation for awaiting people is the least, but the most for the confirmed cases.
-In this case, simplifying $a_3$ in Eq.~\eqref{eq:abcd2} gives
-\begin{equation}
-  \begin{split}
-    a_3 & =\testtarget \omega \gamma p_I\left((\omega+\gamma)-\testtarget(\omega p_I+\gamma)\right)  \\
-    & \propto  \omega \left( 1- \testtarget p_I \right) + \gamma \left( 1- \testtarget \right)  .
-  \end{split}
-\end{equation}
-If $p_I>0$, then $a_3<0$ for sufficiently targeted testing (i.e. when $\testtarget p_I >1$; because
-$p_I \leq 1$, \testtarget is always $\geq \testtarget p_I$).
-%% this LOOKS like we can get a3<0 even if pI=0 for targeting > 1 and sufficiently large gamma/omega,
-%% but that seems crazy so I'm probably missing something
-When the test is perfectly sensitive ($p_I=1$), $a_3<0$ as long as $\testtarget>1$.
-Under either of these conditions, there exists a range for $\rho$ over which $\pder\Rnum{\rho}\leq 0$.  
-Because increasing values of $\rho$ and $\omega$ both delay the rate at which individuals flow to the $I\_c$ compartment, it is reasonable that increasing either value could (under appropriate circumstances) increase $\Rnum$.
-
-% %%%%%%%
-\subsection{Rate of returning tests} \label{app:omega}
-The third goal is to explore how changes in the rate of test return $\omega$ affect $\Rnum$.
-Mathematically we would like to verify the sign of $\pder\Rnum{\omega}$, which specifically depends on $\pder\Delta{\omega}$.
-We use
-the linearization of $\Delta$ when $\rho \ll 1$ to show that there a non-monotonic relationship between $\Rnum$  and $\omega$. The linear term in the Taylor expansion of $\Delta$ when $\rho \ll 1$ is
-\begin{equation}
-\label{eq:lin}
-\Delta = \frac{\rho}{\omega \gamma(\omega+\gamma)} \Big(
-\testtarget \omega^2 p_I \theta\_c+(\testtarget+1) \gamma\omega\theta\_w +\gamma^2 \theta\_w(2-\theta\_w)
-\Big). 
-\end{equation}
-This results in
-\begin{equation}
-\label{eq:dlindo}
-\pder\Delta{\omega} = \frac{\rho}{\omega^2 (\omega+\gamma)^2} \Big(
-(p_I \testtarget\theta\_c -(1+\testtarget)\theta\_w)\omega^2-2\theta\_w \gamma(2-\theta\_w) \omega+\theta\_w \gamma^2 (\theta\_w-2)
-\Big).
-\end{equation}
-% ALi & Faddy's stuff
-The latter expression has two roots
-\begin{align}
-\label{eq:omega_roots}
-\omega^*_{+} &= \frac{\gamma\Big(-c_4 + \sqrt{c_4^2 + (a_4-b_4)c_4}\Big)}{b_4-a_4} \\
-\omega^*_{-} &= \frac{\gamma\Big(-c_4 - \sqrt{c_4^2 + (a_4-b_4)c_4}\Big)}{b_4-a_4},
-\end{align}
-where $a_4 = p_I\testtarget\theta\_c$, $b_4 = (1+\testtarget)\theta\_w$, $c_4 = \theta\_w(2-\theta\_w)$.
-This enables us to describe the behaviour of $\Rnum$ with respect to $\omega$ in the following two cases.
-\begin{itemize}
-\item \textbf{Case I:} If $b_4 \geq a_4$, $\pder\Delta{\omega} < 0$, so $\Rnum$ is always increasing with respect to $\omega$ (i.e., it is always harmful to return tests more rapidly). 
-\item \textbf{Case II:} If $b_4 < a_4$, $\Rnum$ will be decreasing with respect to $\omega$ (i.e., returning tests more rapidly is beneficial) only when $\omega > \omega^*_-.$
-\end{itemize}
-Note that $b_4 \geq a_4$ is characterized by
-\begin{equation}\label{case_1_ineq}
-\Big(\testtarget+1\Big)\theta\_w \geq \testtarget p_I\theta\_c \Longleftrightarrow \Big(\frac{w_S}{w_I}+1\Big)\frac{1}{p_I} \geq \frac{\theta\_c}{\theta\_w}.
-\end{equation}
-We begin with a \textbf{proof of Case I}.
-Suppose that $b_4 \geq a_4$.
-If the roots of $\pder\Delta{\omega}$ are not complex, then we must have $c_4 \geq b_4-a_4$.
-Note that $\omega^*_{-}$ must be negative since the numerator is clearly negative but the denominator is positive.
-Next, note that since $c_4 \geq b_4-a_4$, the numerator of $\omega^*_{+}$ must also be negative, so $\omega^*_{+}$ is negative.
-Thus, in this case, $\pder\Rnum{\omega}$ does not change sign on $(0,\infty)$.
-Checking the sign of \ref{eq:dlindo} for arbitrarily large $\omega$ shows that it is negative (since the $(a_4-b_4)\omega^2$ term dominates and is negative).
-So $\Rnum$ is increasing with respect to $\omega$ on all of $(0,\infty)$.
-
-We now turn our attention to a \textbf{proof of Case II}.
-Suppose that $b_4 < a_4$.
-It follows that $\omega^*_+$ and $\omega^*_-$ are real since $c_4 > 0 > b_4-a_4$.
-Next, note that $\omega^*_-$ is positive since both the numerator and denominator are negative.
-On the other hand, $\omega^*_+$ is negative since the denominator is negative but the numerator is positive (because $c_4 > b_4-a_4$).
-Thus, our task is to understand the sign of $\pder\Delta{\omega}$ around the root $\omega^*_-$.
-Checking the sign of \ref{eq:dlindo} for arbitrarily large $\omega$ shows that it is positive (since the $(a_4-b_4)\omega^2$ term dominates and is positive).
-Likewise, checking the sign for values of $\omega$ close to $0$ shows that it is negative.
-Thus, $\Rnum$ is increasing with respect to $\omega$ when $\omega < \omega^*_-$, and is decreasing $\omega > \omega^*_-$.
-
-Having presented the formal analysis, we now concern ourselves with its biological interpretations.
-We begin by interpreting \ref{case_1_ineq}, under which returning tests more rapidly is \emph{always} harmful.
-Notice that the ratio $\frac{\theta\_c}{\theta\_w}$ is simply a measure of how much more strongly individuals self-isolate when they test positive compared to when waiting for tests.
-Since the rate of test return directly influences the rate at which individuals change from a waiting state to a confirmed-positive state, it is intuitive that $\frac{\theta\_c}{\theta\_w}$ would appear in \ref{case_1_ineq}.
-Next, note that the left-hand side increases when test sensitivity decreases and when targeting of positive individuals is poor.
-This is consistent with our intuition: a false negative that is returned more rapidly will allow an infectious individual to relax their self-isolation, thus increasing transmission.
-Likewise, if individuals tested are mainly susceptible (rather than infectious), then returning tests more slowly would encourage them to self-isolate for longer while awaiting test results.
-Having understood the role of each of the parameters in \ref{case_1_ineq}, a holistic interpretation of this inequality is that returning tests more slowly is helpful when the benefit of extended self-isolation by infected individuals awaiting test results outweighs the benefit of identifying positive cases.
-
-Now we interpret \textbf{Case II}.
-In this case, $\Rnum$ will have a global maximum with respect to $\omega$ at $\omega^*_-$.
-Note that our model assumption that $\rho < \omega$ plays an important role here: if $\omega^*_- < \rho$, then $\Rnum$ will be always decreasing with respect to $\omega$.
-On the other hand, if $\rho < \omega^*_-$, then $\Rnum$ will be increasing with respect to $\omega$ on $(\rho, \omega^*_-)$ and decreasing beyond that.
-
-% %%%%%%%
-\subsection{The effect of testing focus parameter $\testtarget$ on $\Rnum$} \label{app:w}
-
-\begin{align}
-\label{eq:d_del_wis}
-\pder \Delta{w\_{IS}}= \dfrac{(\omega-\rho) (\omega(\omega-\rho\theta\_w)+\gamma(\omega-\rho)) (\theta\_w\gamma+\theta\_c\omega p\_I)}
-{(-\omega^2 \gamma+\omega \gamma \rho-\gamma \rho \omega w\_{IS}-
-\omega \gamma^2+\gamma^2 \rho-\omega^2 p_I \rho w\_{IS})^2},
-\end{align}
-which is a positive quantity. Thus, $\pder \Rnum{w\_{IS}}\leq 0$. Therefore, increasing the focus of testing on the infectious people will result in less transmission.
-
-% %%%%%%%
-\subsection{On Testing Rate and Numerical Singularity} \label{app:singularity}
-
-In this work, we do not present any numerical solutions of the ODEs to investigate typical trajectories.
-However, attempting to do so quickly reveals a problem in the way the model is posed;
-with the scaling of testing ($\sigma$) defined as in the body of the paper (Eq.~\eqref{sigma}), the population in the $S$ compartments appeared to blow up when the system is near the DFE.
-This occurs because once the only untested people are susceptibles, the FOI approaches $\Lambda=0$, and the testing rate $\testing{S} \to \rho N/S\_u$.
-Thus, the first equation of the model \eqref{model} will become
-$d S\_u/dt = - \rho N + \omega S\_n$. Thus changes in $S\_u$ will be independent on $S\_u$, and the decay of the $S\_u$
-population becomes linear rather than exponential --- allowing $S\_u$ to become negative!
-To avoid this problem the testing rate, $\sigma$, should be formulated such that people from the untested compartments will not be tested if they are not there.
-One way to fix this issue, is to consider a maximum testing rate, $\tau$ (1/day).
-In general, we want to test at a rate of $\rho$ across the whole population.
-This won't always be possible, so we impose a maximum rate of $\tau$ per testable person and redefine $\sigma = \frac{\tau \rho N}{\tau W + \rho N}$, with the assumption that $\tau \gg \rho$.
-This modification of $\sigma$ does not affect any of the results we have derived about the invasion of the epidemic from the DFE (i.e., results on $\Rnum$ and $\Delta$).
-
-\end{document}

### 2021-07-19 results.tex
breaking the sections, filling the BMB template

--- 2021-07-14 SIR_manuscript.tex
@@
-\documentclass[12pt]{article}
-\usepackage{natbib}
-\usepackage{hyperref}
-\usepackage{grffile}
-\usepackage{graphicx}
-\usepackage{subcaption}
-\usepackage{amssymb,amsmath,amsthm}
-\usepackage{xcolor}
-\usepackage{xspace}
-\usepackage[nameinlink,capitalize]{cleveref}
-\usepackage{cleveref}
-\usepackage[margin=1in]{geometry}
-\usepackage{lineno}\renewcommand\thelinenumber{\color{gray}\arabic{linenumber}}
-\usepackage{pdflscape}
-\usepackage{enumerate}
-\usepackage{authblk}
-
-\hyphenation{co-ro-na-vi-rus}
-\input{codes/modeldefs.tex}
-
-% https://tex.stackexchange.com/questions/12703/how-to-create-fixed-width-table-columns-with-text-raggedright-centered-raggedlef
-\usepackage{array}
-\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-
-\usepackage{xspace}
-
-\newcommand{\fref}[1]{Fig.~\ref{#1}}
-\newcommand{\appref}[1]{Appendix~\ref{app:#1}}
-\newcommand{\Rlogo}{R\xspace}
-\newcommand{\percap}{\emph{per capita}\xspace}
-\newcommand{\Rnum}{\ensuremath{\mathcal{R}_0}\xspace}
-\newcommand{\covid}{COVID-19\xspace}
-\newcommand{\pro}[1]{\ensuremath{\frac{\partial #1}{\partial \rho}}}
-\newcommand{\pder}[2]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder y{x}
-\newcommand{\testing}[1]{\ensuremath{{\cal T}_{#1}}\xspace}
-\newcommand{\testinghat}[1]{\ensuremath{\hat{\cal T}_{#1}}\xspace}
-\newcommand{\testtarget}{\ensuremath{w\_{IS}}}
-\newcommand{\oldtesttarget}{\ensuremath{\frac{w_I}{w_S}}\xspace}
-\newcommand*\subtxt[1]{_{\textnormal{#1}}}
-%% Roman subscript
-\DeclareRobustCommand\_{\ifmmode\expandafter\subtxt\else\textunderscore\fi}
-
-\newcommand{\comment}{\showcomment}
-\newcommand{\showcomment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}
-\newcommand{\nocomment}[3]{}
-
-\newcommand{\fady}[1]{\comment{cyan}{Fady}{#1}}
-\newcommand{\ali}[1]{\comment{magenta}{Ali}{#1}}
-\newcommand{\jd}[1]{\comment{blue}{JD}{#1}}
-\newcommand{\david}[1]{\comment{orange}{DJDE}{#1}}
-\newcommand{\bmb}[1]{\comment{red}{BMB}{#1}}
-\newcommand{\todo}[1]{\comment{red}{TODO}{#1}}
-\newcommand{\new}[1]{\comment{red}{NEW}{#1}}
-\newcommand{\needref}{\textcolor{red}{(REFS?)}}
-
-%% DE: table stuff
-\renewcommand{\arraystretch}{1.5} % more interline spacing
-%% https://tex.stackexchange.com/questions/12703/how-to-create-fixed-width-table-columns-with-text-raggedright-centered-raggedlef
-\usepackage{ragged2e}
-\usepackage{array} % for m{...}
-
-\theoremstyle{definition} % amsthm only
-\newtheorem{proposition}{Proposition}
-\newtheorem{theorem}{Theorem}
-
-\bibliographystyle{apalike}
-
-\title{Testing and Isolation Efficacy:\\Insights from a Simple Epidemic Model}
-\author[1]{Ali Gharouni}
-\author[1]{F.M. Abdelmalek}
-\author[1,2,3]{Benjamin M. Bolker}
-\author[2,4,3]{Jonathan Dushoff}
-\author[1,3,5]{David J. D. Earn}
-
-\affil[1]{Department of Mathematics \& Statistics, McMaster University, Hamilton, Canada}
-\affil[2]{Department of Biology, McMaster University, Hamilton, Canada}
-\affil[3]{Michael G. DeGroote Institute for Infectious Disease Research, McMaster University, Hamilton, Canada}
-\affil[4]{South African Centre for Epidemiological Modelling and Analysis, University of Stellenbosch, Stellenbosch, South Africa}
-
-\begin{document}
-\maketitle
-
-\linenumbers
-
-% %%%%%%%
-\section{Abstract}
-
-Testing individuals for pathogens can affect the spread of epidemics. 
-Understanding how individual-level processes of sampling and reporting test results can affect community- or population-level spread is a dynamical modeling question.
-The effect of testing processes on epidemic dynamics depends on factors underlying implementation, particularly testing intensity and on whom testing is focused.
-Here, we use a simple model to explore how the individual-level effects of testing might directly impact population-level spread.
-Our model development was motivated by the \covid epidemic, but has generic epidemiological and testing structures. 
-To the classic SIR framework we have added a \percap testing intensity, and compartment-specific testing weights, which can be adjusted to reflect different testing emphases --- surveillance, diagnosis, or control.
-We derive an analytic expression for the relative reduction in the basic reproductive number due to testing, test-reporting and related isolation behaviours.
-Intensive testing and fast test reporting are expected to be beneficial at the community level because they can provide a rapid assessment of the situation, identify hot spots, and may enable rapid contact-tracing.
-Direct effects of fast testing at the individual level are less clear, and may depend on how individuals' behaviour is affected by testing information.
-Our simple model shows that under some circumstances both increased testing intensity and faster test reporting can \emph{reduce} the effectiveness of control, and allows us to explore the conditions under which this occurs.
-Conversely, we find that focusing testing on infected individuals always acts to increase effectiveness of control.
-
-% %%%%%%%
-\section{Introduction}
-
-The observed dynamics of the \covid epidemic have been driven both by epidemiological processes (infection and recovery) and by testing processes (testing and test reporting).
-In addition to shaping epidemic observations (via case reports), testing processes also alter epidemiological dynamics \new{\citep{peto2020covid,taipale2020population}}.
-Because individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may also do so, testing will generally increase the number of people who are isolating and hence reduce epidemic growth rates.
-We developed a mechanistic model that incorporates epidemic processes and testing in order to explore the effects of testing and isolation on epidemic dynamics.
-
-If testing influences behavior, then epidemic dynamics will depend on who gets tested.
-The impacts of testing will depend both on testing intensity (tests performed per day) and on how strongly testing is focused on people who are infectious.
-This level of focus depends in turn on the purpose and design of testing programs. 
-When testing is done for the purposes of disease surveillance \citep{foddai2020base}
-tests are typically conducted randomly (or using a stratified random design) across the population in order to make an unbiased assessment of population prevalence.
-
-Over the course of the \covid pandemic, however, the vast majority of testing has been done with other goals --
-primarily diagnostic (determining infection status for clinical purposes)\new{\citep{phua2020intensive,who2020global}}, or for control (determining  infection status in order to isolate cases that have been found by contact tracing) \new{\citep{aleta2020modelling,kucharski2020effectiveness,grassly2020comparison,smith2020adherence}}, which we characterize as \emph{targeted} testing strategies.
-In these situations, testing probabilities can differ sharply across epidemiological compartments; in our dynamical model, we will characterize these probabilities by assigning a testing weight to each compartment that determines the \emph{relative} probability that an individual in that compartment will be selected for testing (see Methods).
-
-Diagnostic testing focuses on people with infection-like symptoms; thus the relative testing weights for infected people will depend on the relative probability of infected people having symptoms.
-For \covid infection, the testing weights will depend on the proportion of asymptomatic infections, the time spent pre-symptomatic vs.\ symptomatic during the course of an infection, and on the incidence of \covid-like symptoms among people in the population \emph{not} infected with \covid.
-Testing for epidemic control focuses on known contacts of infected people; in this case the testing weights for infected vs.\ uninfected people will depend on the probability of infection given contact, as well as the effectiveness of the system for identifying suspicious contacts.
-
-When a new infectious disease emerges, it is important to determine whether it will grow exponentially in a susceptible population, and if so at what rate $r$ \citep{Ma+14}.
-The condition for positive exponential growth ($r>0$) is commonly expressed as $\Rnum>1$, where the basic reproduction number $\Rnum$ is the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}.
-Although the value of $\Rnum$ cannot completely characterize the dynamics of our model \citep{shaw2021what}, it does give a simple and widely accepted index for the difficulty of control, as well as an indication of the likely final size of an epidemic \citep{ma2006generality,miller2012note}.
-
-In order to understand the effect of testing processes on epidemic dynamics, we expanded one of the simplest mechanistic epidemic models---the standard deterministic SIR model\citep{KermMcKe27,AndeMay91}---to include testing components.
-This model provides a sensible platform to link the modeling of epidemic and testing components and study their interaction.
-We studied the effects of testing intensity, rate of test return, and isolation efficacy, on transmission probability and epidemic dynamics when different levels of testing focus (from random to highly targeted) are in place.
-
-% %%%%%%%
-\section{Methods}
-
-Our model groups individuals based on disease status (Susceptible, Infectious or Recovered) and testing status (\emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}) (\fref{fig:flowchart}).
-The testing status of an individual in a given disease compartment $X$ (where $X \in \{S,I,R\}$) is denoted by a subscript, namely $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
-Two `accumulator' compartments, $N$ and $P$, are included in order to collect cumulative reported negative or positive tests.
-The model equations \eqref{model} and details of calculation of the basic reproduction number $\Rnum$ are presented in \appref{R0}.
-
-\begin{figure}
-\begin{center} 
-\includegraphics[scale=1.5]{pix/sir_comp.pdf}
-\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model, \ref{model}.
-The disease-based status of a compartment $X$ ($X \in \{S,I,R\}$) is combined with the testing status including $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
-The force of infection is denoted by $\Lambda$ (Eq.~\eqref{Lambda}); $\gamma$ is the recovery rate; $\omega$ is the rate of test return; and \testing{X} (Eq.
-\eqref{F}) and $p_X$ represent the \percap testing rate and the sensitivity (probability of positive tests), respectively, for compartment $X$.
-For further description of the parameters see \cref{tab:params}.
-Note that there is a slight mismatch in the top-to-bottom order of the testing-based compartments of each disease-based compartment $X$ between this flowchart and the model equations \eqref{model}; here we have switched  $X\_u$ and $X\_n$ for visual clarity.
-\label{fig:flowchart}}
-\end{center} 
-\end{figure}
-
-Table~\ref{tab:params} defines the model parameters, which are generally \percap flows between compartments, or modifiers to these flow rates.
-The novel component of the model lies in the compartment-specific relative testing weights $w_S$, $w_I$ and $w_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively.
-Thus, we can specify different levels of testing focus from random (all weights equal) to highly targeted (higher weights in more intensively tested compartments).
-For example, $w_I/w_S=3$ means that infected individuals are tested at three times the \percap rate of susceptible individuals.
-
-In order to allow parameterization of the model by the total (overall) \percap testing rate, we define the weighted size of the testing pool $W = w_S S\_u + w_I I\_u + w_R R\_u$, and calculate a scaling parameter for testing as:
-\begin{equation}
-\label{sigma}
-\sigma = \frac{\rho N}{W},
-\end{equation}
-where $\rho$ is the \percap testing intensity for the population, defined as the number of daily tests administered in a population of size $N$.
-Thus, the \percap testing rate for compartment $X \in \{S,I,R\}$ is 
-\begin{equation}
-\label{F}
-\testing{X}=\sigma w_X.
-\end{equation}
-For a highly sensitive test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus not considered for further testing.
-Over the course of the epidemic, a sufficiently large fixed testing rate as specified in \eqref{sigma} can exhaust the pool of people available for testing, leading to a singularity when too few people are left untested to support the specified rate.
-Although this phenomenon does not affect our analysis of $\Rnum$, it can affect model dynamics (we present an adjustment to the model that solves this problem in \appref{singularity}).
-
-The classical SIR model assumes a well-mixed population; homogeneity of the population (i.e., all individuals are equally susceptible and equally infectious with the same recovery rate when infected); exponentially distributed duration of infection; and large population size \citep{keeling2011modeling}.
-In addition to these standard assumptions, our model assumes:
-\begin{enumerate}[(i)]
-\item there is a single force of infection (new infections per unit time per susceptible), $\Lambda$, defined as
-  \begin{equation}
-  \label{Lambda}
-  \Lambda=\frac{\beta}{N} \big(I\_u + (1-\theta\_w)(I\_n+I\_p) + (1-\theta\_c)I\_c \big),
-  \end{equation}
-with transmission rate $\beta$; $\theta\_w$ is the isolation efficacy (reduction of the probability of transmission) for individuals waiting for test results, while $\theta\_c$ is the isolation efficacy for individuals who have received a ``confirmed'' positive test (Table~\ref{tab:params}).
-Susceptible individuals who are waiting for test results experience an additional transmission reduction factor of $1-\theta\_w$ (\fref{fig:flowchart}).
-\item confirmed-positive individuals isolate at least as effectively as those awaiting test results, i.e.,
-$$0\leq\theta\_w\leq\theta\_c\leq 1.$$ 
-\end{enumerate}
-For simplicity we assume that tests are perfectly \emph{specific} --- uninfected individuals never test positive ($p_s=0$).
-Thus there are no waiting-for-positive or confirmed-positive susceptible individuals, which reduces the number of model states from 12 to 10.
-
-\begin{table}[htp]
-\centering
-%%\begin{tabular}{|c|L{2in}cc|} \hline
-%% making use of ragged2e and array packages:
-{\RaggedRight
-\begin{tabular}{ c | m{6cm} | c | c }
-  \textbf{Symbol} & \textbf{Description} & \textbf{Unit} & \textbf{Value} \\ \hline
-  $N$     & Total population size & people & $10^6$ \\ \hline
-  $\omega$  & Rate of test return, i.e., rate of onward flow from ``waiting'' to ``confirmed'' or ``untested'' compartments  & 1/day & - \\ \hline
-  $\gamma$ & Recovery rate & 1/day & 1/6 \\ \hline 
-  $\rho$   & \percap testing intensity & 1/day & - \\ \hline 
-  $\theta\_w$ & Isolation efficacy (reduction of the transmission probability) for ``waiting'' individuals & - & - \\ \hline
-  $\theta\_c$ & Isolation efficacy for ``confirmed positive'' individuals & - & -  \\ \hline
-  $\beta$ & Transmission rate & 1/day & 0.5 \\ \hline
-  $\Lambda$ & Force of infection & 1/day & - \\ \hline
-  $p_S$ & Probability of positive tests for $S$ ($= 1-\textrm{specificity}$) & - & 0 \\ \hline
-  $p_I$ & Probability of positive tests for $I$ ($= \textrm{sensitivity}$) & - & 1 \\ \hline
-  $p_R$ & Probability of positive tests for $R$ ($= 1-\textrm{specificity}$) & - & 0.5 \\ \hline
-  $w_S, w_I, w_R$ & Relative testing weights & - &
-  \begin{minipage}[t]{0.21\columnwidth}%
-    Random:~$\{1,1,1\}$ \\ Targeted:~$\{0.3,1,1\}$
-  \end{minipage} \\
-  %%\hline
-\end{tabular}
-} % end \RaggedRight
-\caption{\label{tab:params}Parameters of the model specified  by the flow chart in \fref{fig:flowchart} and equations~\eqref{model}.}
-\end{table}
-
-The Disease-Free Equilibrium (DFE) for the expanded SIR model (Eqs.~\ref{model}) is found by setting the infected compartments to 0 and solving for the unknowns. The DFE is
-\begin{equation}
-\label{dfe}
-S\_n^*= \frac{\rho}{\omega} N, \ S\_u^*= N-S_n^*, \text{~and~} I_j=R_j=0 \ \text{for all }j.
-\end{equation}
-The corresponding \percap testing rate (Eq.~\ref{F}) for the infected compartment $I$ at DFE is one of the key analysis parameters and can be simplified as 
-\begin{equation}
-\label{eq:fi}
-\testinghat{I} = (\omega\rho/(\omega-\rho))w_I/w_S \quad .
-\end{equation}
-
-The basic reproduction number, $\Rnum$, was calculated by using the next-generation matrix method \citep{van2002reproduction}. We write $\Rnum$ as
-\begin{equation}
-\label{R0}
-\Rnum= \frac{\beta}{\gamma} \left(1-\Delta\right), 
-\end{equation}
-where $\beta/\gamma$ is the classical value for a simple model \citep{keeling2011modeling}, and $1-\Delta$ is the proportional reduction due to testing and isolation processes.
-$\Delta$ therefore measures the ``effectiveness of control'': how much these processes in reducing spread from low levels, and is in turn given by:
-\begin{equation}
-  \label{eq:del4a}
-  \Delta= \frac{1}{C N}\big(C_1 S\_u^*+(C_2(1-\theta\_w)+C \theta\_w)S\_n^*\big),
-\end{equation}
-where
-\begin{align}
-\label{eq:C12}
-C_{\phantom 1} &= (\omega+\gamma) \Big(\gamma(\omega+\gamma)+(\gamma+\omega p_I)\testinghat{I}\Big),\\
-C_1 &= (\omega+\gamma)(\theta\_w \gamma+\theta\_c \omega p_I) \testinghat{I},\\
-C_2 &= \Big( \omega\gamma(1+p_I)\testinghat{I}+\gamma^2(\omega+\gamma+\testinghat{I})\Big)\theta\_w + \omega^2p_I \testinghat{I} \theta\_c.
-\end{align}
-(\appref{R0} gives a detailed derivation of these expressions.)
-This explicit formula enables us to study the effects of testing and isolation parameters on $\Rnum$ both analytically and via numerical solutions.
-We are specifically interested in parameters that could be manipulated by public health policy: isolation efficacy, $\theta\_c$ and $\theta\_w$; \percap testing intensity, $\rho$; and the rate of test return, $\omega$.
-In particular, we look at the partial derivatives of $\Delta$ with respect to these parameters (\appref{rho}).
-We derived general expressions for these derivatives.
-However, we analyzed the effect of $\omega$ on $\Delta$ for the special case of low testing intensity.
-Specifically, by making the restriction $\rho \ll 1$, we are able to Taylor-expand $\Delta$ at $\rho=0$, use the linear approximation with respect to $\rho$ and analyze the resulting simplified derivatives to illustrate a surprising non-monotonic relationship between $\Delta$ and $\omega$.
-
-Analytic calculation of the next-generation matrix and simplification of the $\Rnum$ expression, were performed in Maple\textsuperscript{\texttrademark} \citep{maple14}; numerical calculation and contour plots were done in R \citep{r}.
-We computed the values and contours of $\Delta$ at both low (\fref{pan}) and high (\fref{pan2}) testing intensities, and for both random testing ($w_S=w_I=w_R=1$) and targeted testing ($w_S=0.3$; $w_I=w_R=1$).
-Because it is expressed as a proportion of $\Rnum$, the effectiveness of control $\Delta$ is (at least in the $\rho \ll 1$ case, Eq.~\ref{eq:lin}) independent of the transmission rate $\beta$, and hence of $\Rnum$ in the case where we vary $\Rnum$ by changing the transmission rate for a fixed generation interval.
-
-The low-testing case \fref{pan} reflects the case where testing intensity $\rho$ is small relative to the population size.
-Specifically, $\rho \in [0,0.013]$, and test return rate $\omega\in [1/12,2]$.
-This testing intensity is of the correct order of magnitude (although typically larger than) testing rates during the \covid pandemic, i.e., a maximum of 1.3\% of the population per day (approximately four times the maximum testing rate in Ontario, Canada in mid-2021).
-The less realistic high-testing case \fref{pan2} is included to highlight the occurrence of non-monotonic changes in $\Rnum$ with respect to $\rho$.
-In \fref{pan2} the maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,1/5)$ and the test return rate $\omega\in [1/5,2]$; these values are clearly unrealistic for a large population but might be relevant for small populations undergoing focused testing, such as a sports league or university.
-In these figures, the implied baseline reproduction number (for the SIR model without testing) is $\Rnum=\frac{\beta}{\gamma}=3$.
-The different ranges of test return rates $\omega$ for the cases of low and high testing intensities is due to the restriction $\rho<\omega$, which is a requirement for a feasible DFE (\ref{dfe}).
-
-% %%%%%%%
-\section{Results}
+% \section{Results}
 
@@
 \end{itemize}
-
-%%%%%%%%%%%%
-\section{Discussion}
-
-In this paper, we have developed and analyzed a simple compartmental model that combines epidemiological dynamics---as defined by a simple SIR model---with the dynamics of testing and isolation.
-Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness}).
-Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
-
-Many of the qualitative results we have derived confirm simple, common-sense intuitions.
-In particular, we can generally decrease $\Rnum$ by increasing isolation efficacy or testing intensity; returning tests faster, if individuals do not isolate while they are waiting for results; or increasing testing focus to target individuals who are likely to be infectious (e.g., symptomatic people or close contacts of known infections).
-
-However, we did find two surprising patterns: under some conditions longer delays in returning tests can reduce epidemic spread, and increasing testing rates can increase spread.
-
-Over broad regions of parameter space, decreasing $\omega$---i.e., \emph{slowing} the rate at which test results are returned---\emph{decreases} $\Rnum$ (for random testing, the parameter region is $\theta\_w \gtrsim 0.25$; for targeted testing, $\theta\_w \geq 0.25$ and either $\theta\_c \ge 0.5$ or $1/\omega > 5$; see Table~\ref{tab:params} for parameter definitions).
-This result is counterintuitive and would not be expected by public health authorities that have invested a great deal of effort in reducing delays from testing to results.
-Dynamically, this effect occurs because speeding up test returns shortens the isolation period of uninfected individuals (for infected people it only shortens the time to progression to the isolation level of the confirmed-positive compartment).
-Slowing test returns increases $\Rnum$ only if the proportion of infectives in the tested population is high and isolation is relatively strong among people waiting for test results.
-
-While slowing test returns does decrease $\Rnum$ over broad regions of parameter space in our model, there are several real-world processes missing from our model that make it unlikely that slowing test returns would actually be an effective public health measure.
-First, we do not model the primary benefit of rapid testing, i.e., detecting and containing outbreaks while they are still in progress.
-This process could be modeled phenomenologically by making the testing focus more targeted as an increasing proportion of cases is detected, because finding infections allows tests to be concentrated on their connections.
-Second, individuals may become less likely to maintain isolation if they are required to do so for longer; phenomenologically, we could allow  effectiveness of isolation in the waiting population to be an increasing function of test-return speed, or we could introduce a separate ``waiting, but no longer isolating'' compartment that individuals entered from the ``waiting, isolated'' compartment at a specified rate.
-Finally, if one wants to decrease the overall transmission rate of the population there are more effective methods than keeping tested people in limbo; these include masking, ventilation, distancing measures, retail and event closures, and stay-at-home orders.
-
-The other counterintuitive result from our analysis is that, for sufficiently high testing intensity $\rho$, further increasing testing intensity can actually \emph{increase} $\Rnum$ (e.g., \fref{pan2}(b), upper right panel [$\theta\_c=1$, $\theta\_w=0$]).
-This phenomenon can occur because we are considering the DFE in the presence of testing; thus there is an equilibrium distribution of susceptibles between the $S\_n$ (untested) and $S\_u$ (waiting) compartments even as the disease approaches extinction.
-A higher rate of testing leads to a greater proportion of individuals waiting for negative tests at the DFE.
-If infected, individuals in this group will take longer to be tested again and to subsequently isolate (because they must wait for their negative tests to be returned before being tested again).
-If isolation in this group ($\theta\_w$) is low, this effect can under some (relatively rare) circumstances (high $\theta\_c$, low $\omega$, high $\rho$) allow $\Rnum$ to increase with testing intensity.
-We can show that this phenomenon occurs \emph{only} under targeted testing ($w_I > w_S$), but we have not yet found a simple explanation of why it cannot occur under random (unfocused) testing.
-This phenomenon is also unlikely to occur in the real world.
-In particular, it depends on levels of testing that are unrealistically high (at least in large, general-population settings).
-
-Although we model the testing process in more detail than typical epidemiological models, one place where more detail could be informative is in the processes determining the testing weights $\{w_S, w_I, w_R\}$.
-While random testing, as done for surveillance purposes, unambiguously leads to equal testing weights, making precise quantitative connections between public-health practices and testing weights is difficult in other contexts.
-The testing weights reflect the correlation between an individual's risk of infection and their likelihood of being tested due to age, occupation, geographic location, etc..
-This correlation is influenced, among many other factors, by the proportion of the uninfected population with COVID-like symptoms (e.g., due to seasonal upper respiratory tract infections); the concentration of transmission and testing in hot spots such as long-term care facilities and high-density workplaces; the overall testing intensity (and hence, e.g., restriction to symptomatic individuals); and the proportion of COVID-infected people who are symptomatic.
-Future steps should explore mathematically tractable ways to model some of these factors more precisely.
-For example, separating the infected class into exposed, symptomatic, and a- or pre-symptomatic compartments and allowing the testing weights to vary across non-symptomatic (exposed/asymptomatic/presymptomatic) vs.\ symptomatic compartments could reflect the allocation of tests for diagnostic purposes (targeting symptomatic individuals) vs.\ contact-tracing (targeting infected but non-symptomatic individuals) vs.\ screening (relatively equal weights, depending on the venue).
-Alternatively, one could make the testing weights depend on the testing intensity or test-return rate as suggested above.
-Whatever complexity is added would probably put the model beyond reach of the analytical methods we have used in this paper, but one could still use semi-numerical methods such as constructing the next-generation matrix and using it to evaluate the derivatives of $\Rnum$ with respect to the parameters numerically.
-
-Although testing and tracing is a key part of infection control strategies, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts \new{\citep{endo2020implication,hellewell2020feasibility,jenness2020modeling}}, rather than analyzing simple but general models of the feedback between testing and transmission dynamics.
-\new{There have been several modeling studies of testing and tracing dynamics and their interaction with epidemiological dynamics.
-In the context of repeated screening and random testing of isolated populations (such as the members of a university), \cite{bergstrom2020frequency} provided analytical results quantifying the effects that proactive screening of asymptomatic individuals and isolation of confirmed-positive cases could have in reducing the spread of disease.
-\cite{rogers2021high} simulated a SEIR model with testing and isolation; they similarly suggest a strategy of rapid testing with antigen tests and the subsequent isolation of confirmed-positive individuals.
-\cite{friston2021testing} model the effects of self-isolation on testing and tracing with a focus on projections under different testing and tracing scenarios.
-They conclude that the emergence of a second wave depends almost primarily on the rate at which immunity is lost and that it is necessary to track asymptomatic individuals in order to control the outbreak.
-Our modeling approach differs from these previous efforts in that it examines the effects of test-return rates and of different levels of testing focus, from random to highly targeted.
-We hope this paper will inspire further explorations of the fundamental properties of epidemic models that incorporate explicit testing processes.}
-
-% %%%%%%%
-\bibliography{SIRlibrary}
-
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\clearpage
-% \widetext
-\appendix
-\section{Appendix}
-\setcounter{equation}{0}
-\setcounter{figure}{0}
-\setcounter{table}{0}
-% \setcounter{page}{1}
-\makeatletter
-\renewcommand{\theequation}{A\arabic{equation}}
-\renewcommand{\thefigure}{A\arabic{figure}}
-\renewcommand{\bibnumfmt}[1]{[A#1]}
-\renewcommand{\citenumfont}[1]{A#1}
-
-\subsection{Model and calculation of $\Rnum$}\label{app:R0}
-
-The model in the form of a system of ordinary differential equations is 
-\begin{subequations}\label{model}
-\begin{align}
- d S\_u/dt &= -\Lambda S\_u - \testing{S} S\_u + \omega S\_n, \\
- d S\_n/dt &= -(1-\theta\_w)\Lambda S\_n + (1-p_S) \testing{S} S\_u - \omega S\_n, \\
- d S\_p/dt &= -(1-\theta\_w)\Lambda S\_p + p_S \testing{S} S\_u - \omega S\_p, \\
- d S\_c/dt &= -(1-\theta\_c)\Lambda S\_c + \omega S\_p, \\
- d I\_u/dt &= \Lambda S\_u - \testing{I} I\_u + \omega I\_n  - \gamma I\_u, \\
- d I\_n/dt &= (1-\theta\_w)\Lambda S\_n + (1-p_I) \testing{I} I\_u - \omega I\_n -\gamma I\_n, \\
- d I\_p/dt &= (1-\theta\_w)\Lambda S\_p + p_I \testing{I} I\_u - \omega I\_p -\gamma I\_p, \\
- d I\_c/dt &= (1-\theta\_c)\Lambda S\_c + \omega I\_p - \gamma I\_c, \\
- d R\_u/dt &= \gamma I\_u - \testing{R} R\_u + \omega R\_n, \\
- d R\_n/dt &= \gamma I\_n + (1-p_R) \testing{R} R\_u - \omega R\_n, \\
- d R\_p/dt &= \gamma I\_p + p_R \testing{R} R\_u  - \omega R\_p, \\
- d R\_c/dt&= \gamma I\_c + \omega R\_p, \\
- dN/dt &= \omega (S\_n + I\_n + R\_n),  \\
- dP/dt &= \omega(I\_p + R\_p) ,
-\end{align}
-\end{subequations}
-%
-(see Table \ref{tab:params} for parameter definitions).
-The next generation matrix for this model is $G = F V^{-1}$, where matrix $F$ represents the inflow of new infection to the infected compartments and matrix $V$ represents the flow in the infected compartments when the population is totally susceptible.
-Matrices $F$ and $V$ are
-\begin{align}
-\label{F}
-F =& \beta/N \left[ \begin {array}{cccc} 
-S\_u^* & (1-\theta\_w) S\_u^* & (1-\theta\_w) S\_u^* & (1-\theta\_c) S\_u^*\\
-(1-\theta\_w)S\_n^* & (1-\theta\_w)^2 S\_n^* & (1-\theta\_w)^2 S\_n^* & (1-\theta\_w)(1-\theta\_c) S\_n^* \\ 
-0&0&0&0\\
-0&0&0&0
- \end {array} \right] \\ \notag
- =&\beta/N \left[ \begin {array}{c} S\_u^* \\ (1-\theta\_w)S\_n^* \\ 0 \\ 0 \end {array} \right]
-        \left[ \begin {array}{cccc} 1,   1-\theta\_w,   1-\theta\_w,   1-\theta\_c \end {array} \right],
-\end{align}
-and 
-\begin{align}
-\label{V}
-V= \left[ \begin {array}{cccc}  
-\testinghat{I}+\gamma&-\omega&0&0\\
--(1-p_I)\testinghat{I}&\omega+\gamma&0&0\\
--p_I \testinghat{I}&0&\omega+\gamma&0\\
-0&0&-\omega&\gamma
-\end {array} \right].
-\end{align}
-The matrix inverse of $V$ is 
-\begin{align}
-\label{vinv}
-V^{-1} =
-\frac{1}{\gamma C}
-\left[ \begin {array}{cccc}
-\gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)&0&0\\ \noalign{\medskip}
-\gamma(\omega+\gamma)(1-p_I) \testinghat{I} & \gamma(\omega+\gamma)(\testinghat{I}+\gamma)&0&0 \\ \noalign{\medskip}
-\gamma(\omega+\gamma) p_I \testinghat{I} & \gamma\omega p_I \testinghat{I} & C \gamma/(\omega+\gamma) & 0 \\ \noalign{\medskip}
-\omega(\omega+\gamma) p_I \testinghat{I} & \omega^2 p_I \testinghat{I} & C \omega/(\omega+\gamma) & C
-\end {array} \right],
-\end{align}
-where $C= \big( \gamma(\omega+\gamma)+(\gamma+\omega p_I)\testinghat{I} \big) (\omega+\gamma)$ and $\testinghat{I}$ is the \percap testing rate for the infected people and represented in Eq.~\eqref{eq:fi}.
-Note that all the columns of matrix $V^{-1}$ summ up to $1/\gamma$.
-
-The particular form of $F$ with two rows of zeros at the bottom results in the following blocked form of matrix $G$.
-\begin{equation}
-\label{mat:G}
-G = \left[ \begin {array}{cc}
-G_{11}&G_{12}\\
-0&0
-\end {array} \right],
-\end{equation}
-where both blocked matrices $G_{11}$ and $G_{12}$ are 2 by 2.
-Given the upper triangular form of matrix $G$, the basic reproduction number $\Rnum$ (defined as the spectral radius of matrix $G$) is only determined by the blocked matrix $G_{11}$,
-\begin{equation}
-\label{G11}
-G_{11} = \frac{\beta}{\gamma   C} 
-\left[ \begin {array}{c} (\omega-\rho)/\omega \\ (1-\theta\_w)\rho/\omega \end {array} \right]
-\left[ \begin {array}{cccc} 1,   1-\theta\_w,   1-\theta\_w,   1-\theta\_c \end {array} \right]
-\left[ \begin {array}{cc}
-\gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)\\ \noalign{\medskip}
-\gamma(\omega+\gamma)(1-p_I)\testinghat{I} & \gamma(\omega+\gamma)(\testinghat{I}+\gamma) \\ \noalign{\medskip}
-\gamma(\omega+\gamma) p_I \testinghat{I} & \gamma\omega p_I \testinghat{I} \\ \noalign{\medskip}
-\omega(\omega+\gamma) p_I \testinghat{I} & \omega^2 p_I \testinghat{I}
-\end {array} \right].
-\end{equation}
-It is notable that matrix $F$ \eqref{F} has rank one and consequently $G_{11}$ does so. That is $G_{11}$ has only one non-zero eigenvalue which is $\Rnum$.
-
-The expression of $\Rnum$ has a complicated form with all of the model parameters involved.
-This expression can be simplified and represented given the specific form of matrix $G_{11}$ \eqref{G11}.
-For the purpose of simplicity we present $\Rnum$ in the manuscript in terms of expressions $C$, $C_1$ and $C_2$, specified in \eqref{eq:C12}.
-
-It remains difficult to show that the reproduction number $\Rnum$ is decreasing with respect to \percap testing intensity, $\rho$, and the speed of the test return, $\omega$, for the feasible ranges of the parameters, that is
-\begin{align}
-\label{cond}
-& \omega>0, \\
-& 0 \leq \rho<\omega,\\ 
-& 0 \leq \theta\_w \leq \theta\_c \leq 1, \\
-& \testtarget\geq 1.
-\end{align}
-In realistic cases the testing rate $\rho$ is very small (i.e., only a small fraction of the population can be tested every day); it is thus reasonable to use a linear approximation of $\Rnum$ for $\rho \ll 1$ to analyze the behaviour of $\Rnum$ with respect to $\omega$ (see section \appref{omega}).
-In the next section we provide an equivalent representation of $\Rnum$ in order to show that increasing testing intensity typically decreases $\Rnum$.
-
-% %%%%%%%
-\subsection{More testing intensity may decrease $\Rnum$}\label{app:rho}
-
-This section shows that $\pro{\Delta}$ can be positive or negative, with $\Delta$ defined in Eq.~\eqref{eq:C12}, and thus $\pro{\Rnum} < 0$, where $\Rnum$ is given in Eq.~\eqref{R0}.
-We rewrite matrix $G_{11}$ in \eqref{G11} in the following form to simplify the calculations:
-\begin{equation}
-\label{G112}
-G_{11} = \frac{\beta}{\gamma C} 
-\left[ \begin {array}{c}  S\_u^*/N \\ (1-\theta\_w) S\_n^*/N  \end {array} \right]
-\left[ \begin {array}{cccc} 
-C-C_1, C-C_2\end {array} \right],
-\end{equation}
-where $C$ is the same as the one in Eq. \eqref{eq:C12}, i.e., 
-$$C=(\omega+\gamma)(\gamma(\omega+\gamma)+(\omega p_I+\gamma) \testinghat{I}),$$
-and $C_1$ and $C_2$ are 
-\begin{align*}
-% \label{eq:C12}
-C_1 =& (\omega+\gamma)(\theta\_w   \gamma+\theta\_c  \omega  p_I) \testinghat{I},\\
-C_2 =& \big( \omega\gamma(1+p_I)\testinghat{I}+\gamma^2 (\omega+\gamma+\testinghat{I})\big)\theta\_w + \omega^2 p_I \testinghat{I} \theta\_c,
-\end{align*}
-where $\testinghat{I}$ is given in Eq.~\eqref{eq:fi}.
-Note that for analysis brevity, we let $N=1$, thus $S\_u^*$ and $S\_n^*$ are in the scale of 0 to 1.
-$\Rnum$ is in the same form as in Eq. \eqref{R0}  
-$$\Rnum= \frac{\beta}{\gamma} (1-\Delta),$$
-where 
-\begin{equation*}
-\label{eq:del4}
-\Delta= \frac{1}{C}\big(C_1 S\_u^*+(C_2(1-\theta\_w)+C \theta\_w) S\_n^*\big).
-\end{equation*}
-
-The first goal
-is to explore how changes in isolation, $\theta\_w$ and $\theta\_c$, affects $\Rnum$.
-Mathematically we would like to verify the sign of $\pder \Rnum{\theta\_w}$ and $\pder \Rnum{\theta\_c}$.
-We start with simplifying $\Delta$ \eqref{eq:del4} by factoring $\theta\_w$ and $\theta\_c$ in Eq.~\eqref{eq:del4}.
-Thus, $\Delta$ can be rewritten as
-\begin{equation}
-\label{eq:del4_theta}
-\Delta= \frac{1}{C}\Big(
--D_1 S\_n^* \theta\_w^2 +
-\big(-\omega^2 p_I \testinghat{I} S\_n^*\theta\_c+ D_2 S\_n^*+\gamma \testinghat{I}(\omega+\gamma) \big)\theta\_w+
-(\omega+\gamma S\_u^*)\omega p_I \testinghat{I} \theta\_c
-\Big),
-\end{equation}
-where
-\begin{align}
-\label{eq:del4_d1d2}
-D_1=& (\omega+\gamma)\gamma^2+(\omega+\gamma+\omega p_I)\gamma \testinghat{I}, \\
-D_2=& (3\omega+2\gamma)\gamma^2+(\omega+\gamma+2\omega p_I)\gamma \testinghat{I}+(\gamma+p_I \testinghat{I})\omega^2.
-\end{align}
-$\Delta$, Eq.~\eqref{eq:del4_theta}, is linear in $\theta\_c$ with a positive coefficient. thus
-\begin{equation}
-\label{eq:d_del4_thetac}
-\pder\Delta{\theta\_c}=1/C(\gamma S\_u^*+\omega(1-\theta\_w S\_n^*))\omega p_I \testinghat{I}.
-\end{equation}
-This results in increasing $\Delta$, thus decreasing $\Rnum$ with respect to $\theta\_c$, that is $\pder \Rnum{\theta\_c}\leq 0$. Note that $C$ is independent of $\theta\_c$ and $\theta\_w$.
-
-With a similar logic, $\Delta$ \eqref{eq:del4_theta} is a concave-down quadratic equation in $\theta\_w$, given by
-\begin{equation}
-\label{eq:del4_thetaw}
-1/C \Big( -D_1 S\_n^*\theta\_w^2+
-\big(-\omega^2 p_I \testinghat{I} S\_n^*\theta\_c+D_2 S\_n^*+\gamma \testinghat{I}(\omega+\gamma) \big)\theta\_w\Big).
-\end{equation}
-We show that the feasible range of $\theta\_w$ lies between 0 and the vertex of this parabola where the parabola is increasing in $\theta\_w$, and so does $\Delta$ which results in inferring $\pder\Rnum{\theta\_w}\leq 0$.
-It is enough to show that partial derivative of the expression \eqref{eq:del4_thetaw} with respect to $\theta\_w$ at $\theta\_w=1$ is non-negative. It follows that
-\begin{align}
-\label{eq:d_del4_thetaw}
-\pder\Delta{\theta\_w}\bigg\rvert_{\theta\_w=1} =&
-1/C\Big( (D_2-2D_1-\omega^2 p_I \testinghat{I} \theta\_c) S\_n^*+\gamma \testinghat{I}(\omega+\gamma) \Big) \\\notag
-=& 1/C\Big( (\gamma(\omega+\gamma)+\gamma\omega^2+(1-\theta\_c)\omega^2 p_I \testinghat{I} ) S\_n^*
-+\gamma(\omega+\gamma) \testinghat{I} (1-S\_n^*) \Big),
-\end{align}
-which is a positive quantity, given that $\theta\_c$ and $S\_n^*$ vary between 0 and 1.
-
-{\bf The second goal} is to explore how changes in \percap testing intensity $\rho$ affects $\Rnum$.
-Mathematically we would like to verify the sign of $\pder\Rnum{\rho}$, which specifically depends on $\pder\Delta{\rho}$.
-We use the derived expressions for $S\_u^*$ and $S\_n^*$, given by Eqs.~\eqref{dfe}, in $\Delta$ \eqref{eq:del4}.
-Also, we define $\phi = \testinghat{S} = \frac{\rho \omega}{\omega-\rho}$, to reparameterize $\rho$.
-This is mainly to avoid singularity in $\testinghat{I}$ \eqref{eq:fi}, when testing intensity $\rho$ is very close to the rate of test return $\omega$.
-Thus, $\rho$ is reparameterized as
-\begin{equation}
-\label{eq:phi}
-\rho=\frac{\omega \phi}{\omega+\phi}.
-\end{equation}
-This one-to-one monotonic reparameterization enables us to simplify the mathematical expressions and explore the simpler $\pder\Delta{\phi}$ instead of the complicated $\pder\Delta{\rho}$.
-Defining $\testtarget \equiv \oldtesttarget$, the derivative is 
-\begin{align}
-\label{eq:dd3dr}
-\partial\Delta/\partial\phi=& \frac{1}{d_3} (a_3 \phi^2+b_3 \phi+c_3),
-\end{align}
-where
-\begin{align}
-\label{eq:abcd2}
-a_3& = \testtarget \Bigg(
-&& (1-\theta\_w) (1+\testtarget)\theta\_w \gamma^3 +(1-\theta\_c) p_I^2 \theta\_w \testtarget \omega^3 \\ \notag
-& ~ &&+\Big( \big( (1-\theta\_w-\testtarget) \theta\_c +(3-2\theta\_w) \theta\_w\testtarget \big) p_I +(1-\theta\_w) (1+\testtarget)\theta\_w  \Big) \omega \gamma^2 \\ \notag
-& ~ &&+\Big(
-\big((1-\theta\_w-\theta\_w \testtarget) \theta\_c+ (2-\theta\_w) \theta\_w \testtarget \big) p_I
-+(2 \theta\_w-\theta\_w^2-\theta\_c) \testtarget p_I^2
-\Big) \omega^2 \gamma \Bigg), \\ \notag
-b_3& = && \hspace{-1cm} 2\testtarget (\omega+\gamma)\gamma\Big(
-(\omega+\gamma+\omega p_I)(2-\theta\_w) \gamma \theta\_w+ (1-\theta\_w) \omega^2 p_I\theta\_c+\omega^2 p_I\theta\_w
-\Big) ,\\ \notag
-c_3& = && \hspace{-1cm} (\omega+\gamma)^2 \gamma\Big(
-(2-\theta\_w) \gamma^2 \theta\_w+
-(1+\testtarget) \omega \gamma \theta\_w+
-\testtarget \omega^2 p_I\theta\_c \Big),\\ \notag
-d_3&  = && \hspace{-1cm} \frac{(\omega+\gamma)}{\omega}  \Big((\omega p_I+\gamma)\testtarget\phi + (\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2.
-\end{align}
-Note that $\phi\geq 0$, also $b_3$, $c_3$ and $d_3$ are all positive. However $a_3$ can be positive or negative.
-If $a_3\geq 0$, $\partial\Delta/\partial\phi \geq 0$ for all feasible range of parameters, thus $\pro\Rnum \leq 0$.
-It is straightforward to show that $a_3\geq 0$ when testing is random, i.e., $w\_S=w\_I=1$.
-If $a_3 < 0$, then the quadratic expression in the numerator of \eqref{eq:dd3dr} has a positive root, $\phi^*$, such that for $\phi>\phi^*$, $\partial\Delta/\partial\phi < 0$.
-
-An example of this countervailing effect of $\phi$, and consequently $\rho$, on $\Rnum$ occurs when $\theta\_w=0$ and $\theta\_c=1$.
-This is illustrated in the top-right panel of the \fref{pan2} panel (b), where the strength of isolation for awaiting people is the least, but the most for the confirmed cases.
-In this case, simplifying $a_3$ in Eq.~\eqref{eq:abcd2} gives
-\begin{equation}
-  \begin{split}
-    a_3 & =\testtarget \omega \gamma p_I\left((\omega+\gamma)-\testtarget(\omega p_I+\gamma)\right)  \\
-    & \propto  \omega \left( 1- \testtarget p_I \right) + \gamma \left( 1- \testtarget \right)  .
-  \end{split}
-\end{equation}
-If $p_I>0$, then $a_3<0$ for sufficiently targeted testing (i.e. when $\testtarget p_I >1$; because
-$p_I \leq 1$, \testtarget is always $\geq \testtarget p_I$).
-%% this LOOKS like we can get a3<0 even if pI=0 for targeting > 1 and sufficiently large gamma/omega,
-%% but that seems crazy so I'm probably missing something
-When the test is perfectly sensitive ($p_I=1$), $a_3<0$ as long as $\testtarget>1$.
-Under either of these conditions, there exists a range for $\rho$ over which $\pder\Rnum{\rho}\leq 0$.  
-Because increasing values of $\rho$ and $\omega$ both delay the rate at which individuals flow to the $I\_c$ compartment, it is reasonable that increasing either value could (under appropriate circumstances) increase $\Rnum$.
-
-% %%%%%%%
-\subsection{Rate of returning tests} \label{app:omega}
-The third goal is to explore how changes in the rate of test return $\omega$ affect $\Rnum$.
-Mathematically we would like to verify the sign of $\pder\Rnum{\omega}$, which specifically depends on $\pder\Delta{\omega}$.
-We use
-the linearization of $\Delta$ when $\rho \ll 1$ to show that there a non-monotonic relationship between $\Rnum$  and $\omega$. The linear term in the Taylor expansion of $\Delta$ when $\rho \ll 1$ is
-\begin{equation}
-\label{eq:lin}
-\Delta = \frac{\rho}{\omega \gamma(\omega+\gamma)} \Big(
-\testtarget \omega^2 p_I \theta\_c+(\testtarget+1) \gamma\omega\theta\_w +\gamma^2 \theta\_w(2-\theta\_w)
-\Big). 
-\end{equation}
-This results in
-\begin{equation}
-\label{eq:dlindo}
-\pder\Delta{\omega} = \frac{\rho}{\omega^2 (\omega+\gamma)^2} \Big(
-(p_I \testtarget\theta\_c -(1+\testtarget)\theta\_w)\omega^2-2\theta\_w \gamma(2-\theta\_w) \omega+\theta\_w \gamma^2 (\theta\_w-2)
-\Big).
-\end{equation}
-% ALi & Faddy's stuff
-The latter expression has two roots
-\begin{align}
-\label{eq:omega_roots}
-\omega^*_{+} &= \frac{\gamma\Big(-c_4 + \sqrt{c_4^2 + (a_4-b_4)c_4}\Big)}{b_4-a_4} \\
-\omega^*_{-} &= \frac{\gamma\Big(-c_4 - \sqrt{c_4^2 + (a_4-b_4)c_4}\Big)}{b_4-a_4},
-\end{align}
-where $a_4 = p_I\testtarget\theta\_c$, $b_4 = (1+\testtarget)\theta\_w$, $c_4 = \theta\_w(2-\theta\_w)$.
-This enables us to describe the behaviour of $\Rnum$ with respect to $\omega$ in the following two cases.
-\begin{itemize}
-\item \textbf{Case I:} If $b_4 \geq a_4$, $\pder\Delta{\omega} < 0$, so $\Rnum$ is always increasing with respect to $\omega$ (i.e., it is always harmful to return tests more rapidly). 
-\item \textbf{Case II:} If $b_4 < a_4$, $\Rnum$ will be decreasing with respect to $\omega$ (i.e., returning tests more rapidly is beneficial) only when $\omega > \omega^*_-.$
-\end{itemize}
-Note that $b_4 \geq a_4$ is characterized by
-\begin{equation}\label{case_1_ineq}
-\Big(\testtarget+1\Big)\theta\_w \geq \testtarget p_I\theta\_c \Longleftrightarrow \Big(\frac{w_S}{w_I}+1\Big)\frac{1}{p_I} \geq \frac{\theta\_c}{\theta\_w}.
-\end{equation}
-We begin with a \textbf{proof of Case I}.
-Suppose that $b_4 \geq a_4$.
-If the roots of $\pder\Delta{\omega}$ are not complex, then we must have $c_4 \geq b_4-a_4$.
-Note that $\omega^*_{-}$ must be negative since the numerator is clearly negative but the denominator is positive.
-Next, note that since $c_4 \geq b_4-a_4$, the numerator of $\omega^*_{+}$ must also be negative, so $\omega^*_{+}$ is negative.
-Thus, in this case, $\pder\Rnum{\omega}$ does not change sign on $(0,\infty)$.
-Checking the sign of \ref{eq:dlindo} for arbitrarily large $\omega$ shows that it is negative (since the $(a_4-b_4)\omega^2$ term dominates and is negative).
-So $\Rnum$ is increasing with respect to $\omega$ on all of $(0,\infty)$.
-
-We now turn our attention to a \textbf{proof of Case II}.
-Suppose that $b_4 < a_4$.
-It follows that $\omega^*_+$ and $\omega^*_-$ are real since $c_4 > 0 > b_4-a_4$.
-Next, note that $\omega^*_-$ is positive since both the numerator and denominator are negative.
-On the other hand, $\omega^*_+$ is negative since the denominator is negative but the numerator is positive (because $c_4 > b_4-a_4$).
-Thus, our task is to understand the sign of $\pder\Delta{\omega}$ around the root $\omega^*_-$.
-Checking the sign of \ref{eq:dlindo} for arbitrarily large $\omega$ shows that it is positive (since the $(a_4-b_4)\omega^2$ term dominates and is positive).
-Likewise, checking the sign for values of $\omega$ close to $0$ shows that it is negative.
-Thus, $\Rnum$ is increasing with respect to $\omega$ when $\omega < \omega^*_-$, and is decreasing $\omega > \omega^*_-$.
-
-Having presented the formal analysis, we now concern ourselves with its biological interpretations.
-We begin by interpreting \ref{case_1_ineq}, under which returning tests more rapidly is \emph{always} harmful.
-Notice that the ratio $\frac{\theta\_c}{\theta\_w}$ is simply a measure of how much more strongly individuals self-isolate when they test positive compared to when waiting for tests.
-Since the rate of test return directly influences the rate at which individuals change from a waiting state to a confirmed-positive state, it is intuitive that $\frac{\theta\_c}{\theta\_w}$ would appear in \ref{case_1_ineq}.
-Next, note that the left-hand side increases when test sensitivity decreases and when targeting of positive individuals is poor.
-This is consistent with our intuition: a false negative that is returned more rapidly will allow an infectious individual to relax their self-isolation, thus increasing transmission.
-Likewise, if individuals tested are mainly susceptible (rather than infectious), then returning tests more slowly would encourage them to self-isolate for longer while awaiting test results.
-Having understood the role of each of the parameters in \ref{case_1_ineq}, a holistic interpretation of this inequality is that returning tests more slowly is helpful when the benefit of extended self-isolation by infected individuals awaiting test results outweighs the benefit of identifying positive cases.
-
-Now we interpret \textbf{Case II}.
-In this case, $\Rnum$ will have a global maximum with respect to $\omega$ at $\omega^*_-$.
-Note that our model assumption that $\rho < \omega$ plays an important role here: if $\omega^*_- < \rho$, then $\Rnum$ will be always decreasing with respect to $\omega$.
-On the other hand, if $\rho < \omega^*_-$, then $\Rnum$ will be increasing with respect to $\omega$ on $(\rho, \omega^*_-)$ and decreasing beyond that.
-
-% %%%%%%%
-\subsection{On Testing Rate and Numerical Singularity} \label{app:singularity}
-
-In this work, we do not present any numerical solutions of the ODEs to investigate typical trajectories.
-However, attempting to do so quickly reveals a problem in the way the model is posed;
-with the scaling of testing ($\sigma$) defined as in the body of the paper (Eq.~\eqref{sigma}), the population in the $S$ compartments appeared to blow up when the system is near the DFE.
-This occurs because once the only untested people are susceptibles, the FOI approaches $\Lambda=0$, and the testing rate $\testing{S} \to \rho N/S\_u$.
-Thus, the first equation of the model \eqref{model} will become
-$d S\_u/dt = - \rho N + \omega S\_n$. Thus changes in $S\_u$ will be independent on $S\_u$, and the decay of the $S\_u$
-population becomes linear rather than exponential --- allowing $S\_u$ to become negative!
-To avoid this problem the testing rate, $\sigma$, should be formulated such that people from the untested compartments will not be tested if they are not there.
-One way to fix this issue, is to consider a maximum testing rate, $\tau$ (1/day).
-In general, we want to test at a rate of $\rho$ across the whole population.
-This won't always be possible, so we impose a maximum rate of $\tau$ per testable person and redefine $\sigma = \frac{\tau \rho N}{\tau W + \rho N}$, with the assumption that $\tau \gg \rho$.
-This modification of $\sigma$ does not affect any of the results we have derived about the invasion of the epidemic from the DFE (i.e., results on $\Rnum$ and $\Delta$).
-
-% %%%%%%%
-\subsection{The effect of testing focus parameter $\testtarget$ on $\Rnum$} \label{app:w}
-
-\begin{align}
-\label{eq:d_del_wis}
-\pder \Delta{w\_{IS}}= \dfrac{(\omega-\rho) (\omega(\omega-\rho\theta\_w)+\gamma(\omega-\rho)) (\theta\_w\gamma+\theta\_c\omega p\_I)}
-{(-\omega^2 \gamma+\omega \gamma \rho-\gamma \rho \omega w\_{IS}-
-\omega \gamma^2+\gamma^2 \rho-\omega^2 p_I \rho w\_{IS})^2},
-\end{align}
-which is a positive quantity. Thus, $\pder \Rnum{w\_{IS}}\leq 0$. Therefore, increasing the focus of testing on the infectious people will result in less transmission.
-
-\end{document}

### 2021-07-20 authors.R
moving files to the right dirs

@@ -1,11 +1,0 @@
-
-ali <- 330
-david <- 8128
-ben <- 1037
-
-set.seed(ali+ben+david)
-
-print(c(
-	"AG", "FA", "DE", 
-	sample(c("BB", "JD"))
-))

### 2021-07-20 codes/authors.R
update repo

@@ -1,0 +1,11 @@
+
+ali <- 330
+david <- 8128
+ben <- 1037
+
+set.seed(ali+ben+david)
+
+print(c(
+	"AG", "FA", "DE", 
+	sample(c("BB", "JD"))
+))

### 2021-07-20 cover.tex
cover letter updated

@@
 %% comment macros
-\newcommand{\irena}[1]{\textcolor{irena}{$\langle$\textbf{IP:} #1$\rangle$}}
-\newcommand{\de}[1]{\textcolor{de}{$\langle$\textbf{DE:} #1$\rangle$}}
+\newcommand{\ali}[1]{\textcolor{blush}{$\langle$\textbf{AG:} #1$\rangle$}}
 
 %% journal name macro
-\newcommand{\journalname}{\emph{BMC Public Health}}
-
+\newcommand{\journalname}{\emph{BMB}}
+% Bulletin of Mathematical Biology
 %% data macros
@@
 
-Dr. Natalie Pafitis
+Dr. Alan Hastings and Dr. Reinhard Laubenbacher
 
@@
 \begin{flushright}
-Irena Papst
+Ali Gharouni
 
-Center for Applied Mathematics, Cornell University
+Postdoctoral Fellow
 
-657 Rhodes Hall
+Department of Mathematics and Statistics,
 
-136 Hoy Rd, Ithaca, NY, USA 14853
+McMaster University,
 
-\href{mailto:ip98@cornell.edu}{ip98@cornell.edu}
+Hamilton, ON, Canada, L8S 4K1
+
+\href{mailto:agharoun@uottawa.ca}{agharoun@uottawa.ca}
 \end{flushright}
@@
 
-Dear Dr.\ Pafitis,
+Dear Drs.\ Hastings and Laubenbacher,
 
 We are pleased to submit an original research article entitled
-``Age-dependence of healthcare interventions for COVID-19 in Ontario,
-Canada'' for your consideration for publication in \journalname{}.
+``Testing and Isolation Efficacy: Insights from a Simple Epidemic Model'' for your consideration for publication in \journalname{}.
 This study specifically aligns with the aims of \journalname{} to publish research related to the social determinants of health and adds new insight into the evolving COVID-19 pandemic.
 
-Age is the most salient clinical indicator of risk from COVID-19 for
-most patients. While age-specific distributions of known SARS-CoV-2
-infections and COVID-19-related deaths have been widely reported,
-relatively few sources present age distributions of hospitalizations
-and serious healthcare interventions administered to COVID-19
-patients. In our study, we analysed known SARS-CoV-2 infection
-records for the province of Ontario, and quantified the age-intervention associations
-for hospitalization, ICU admission, intubations, and ventilations. We
-quantified existing age-related impacts of COVID-19 and identified
-potential future risks should the healthcare system become overwhelmed
-with COVID-19 patients.
-
-Ontario has so far been able to offer and provide COVID-19 treatment
-to all that request it. Yet we find that the probability of survival
-given hospitalization for COVID-19 decreases substantially after age
-40. The distribution of COVID-19-related hospitalizations spans a much
-wider age range than deaths, which are concentrated in the extremely
-elderly. If prevalence of the virus were to increase in the general
-population (as is currently occurring), and ICUs or hospitals were to
-subsequently reach maximum capacity, the distribution of deaths might expand toward younger ages.
-
-While our work is specific to the Ontario SARS-CoV-2 epidemic, it has
-more general and important implications for COVID-19 outbreaks all
-over the world, in terms of both public health and clinical
-practice. Our study reveals a potentially serious mortality risk to
-middle-aged individuals from SARS-CoV-2 infection if the healthcare
-system were to be overwhelmed with COVID-19 patients; this risk is not
-captured by the widely available distributions of known SARS-CoV-2
-infections and COVID-19-related deaths.
-
-We have presented our study to the Ontario government's COVID-19
-science advisory panel (the Ontario Science Table,
-\url{https://covid19-sciencetable.ca/}), where it was received with
-great interest.  We believe that the readership of \journalname{}
-should be made aware of our findings as soon as possible.
+We have developed a compartmental model to understand the effect of testing processes on epidemic dynamics.
+We studied the effects of testing intensity, rate of test return, and isolation efficacy, on epidemic dynamics when different levels of testing focus are in place.
+Our analysis shows that under some circumstances both increased testing intensity and faster test reporting can reduce the effectiveness of control, and allows us to explore the conditions under which this occurs.
+Conversely, we find that focusing testing on infected individuals always acts to increase effectiveness of control.
+Our work provides insights for epidemiologists, mathematical modelers and public health managers in the context of the interaction between testing processes and epidemic processes.
 
 All authors have approved the manuscript for submission and declare no competing interests.
-A preprint of this work has been posted to medRxiv (\url{https://www.medrxiv.org/content/10.1101/2020.09.01.20186395v2}).
+A preprint of this work has been posted to arXiv (\url{https://arxiv.org/abs/2107.08259}).
 The manuscript has not been published or submitted for publication elsewhere.
@@
 
-Irena Papst, MSc
+Ali Gharouni, PhD
 \setlength{\parskip}{0em}
 
-PhD Candidate, Center for Applied Mathematics, Cornell University
+Postdoctoral Fellow, Department of Mathematics and Statistics, McMaster University
 
@@
 
-Michael Li, PhD
+Fady M. Abdelmalek, B.Sc.
 
-Postdoctoral Fellow, Department of Biology, McMaster University
-
-Research Associate, South African Centre for Epidemiological Modelling and Analysis, University of Stellenbosch
-
-\vspace{1em}
-
-David Champredon, PhD
-
-Postdoctoral Associate, Department of Pathology and Laboratory Medicine,
-
-Western University
-
-\vspace{1em}
+Department of Mathematics and Statistics, McMaster University
 
@@
 \columnbreak
+
+David J.\,D.\ Earn, PhD
+
+Professor, Department of Mathematics \& Statistics, McMaster University
+
+Member, Michael G. DeGroote Institute for Infectious Disease Research, McMaster University
+
+\vspace{1em}
 
@@
 
-\vspace{1em}
+\vfill\null
 
@@
 
-\vspace{1em}
+South African Centre for Epidemiological Modelling and Analysis, University of Stellenbosch, Stellenbosch, South Africa
 
-David J.\,D.\ Earn, PhD
-
-Professor, Department of Mathematics \& Statistics, McMaster University
-
-Member, Michael G. DeGroote Institute for Infectious Disease Research, McMaster University
-
-Professor, Department of Mathematics, University of Toronto
 \end{flushleft}

### 2021-07-21 cover.tex
revised first paragraph

@@
 We are pleased to submit an original research article entitled
-``Testing and Isolation Efficacy: Insights from a Simple Epidemic Model'' for your consideration for publication in \journalname{}.
-This study specifically aligns with the aims of \journalname{} to publish research related to the social determinants of health and adds new insight into the evolving COVID-19 pandemic.
+``Testing and Isolation Efficacy: Insights from a Simple Epidemic
+Model'' for your consideration for publication in \journalname{}.  Our
+study provides new insights into the evolving COVID-19 pandemic, and
+more generally to epidemic dynamics in the context of testing.
 

### 2021-07-21 methods.tex
newlines to improve typsetting of table

@@
   $\Lambda$ & Force of infection & 1/day & - \\ \hline
-  $p_S$ & Probability of positive tests for $S$ ($= 1-\textrm{specificity}$) & - & 0 \\ \hline
-  $p_I$ & Probability of positive tests for $I$ ($= \textrm{sensitivity}$) & - & 1 \\ \hline
-  $p_R$ & Probability of positive tests for $R$ ($= 1-\textrm{specificity}$) & - & 0.5 \\ \hline
+  $p_S$ & Probability of positive tests for $S$\newline ($= 1-\textrm{specificity}$) & - & 0 \\ \hline
+  $p_I$ & Probability of positive tests for $I$\newline ($= \textrm{sensitivity}$) & - & 1 \\ \hline
+  $p_R$ & Probability of positive tests for $R$\newline ($= 1-\textrm{specificity}$) & - & 0.5 \\ \hline
   $w_S, w_I, w_R$ & Relative testing weights & - &

### 2021-07-21 manuscript_BMB.tex
added declarations

@@
 \title{Testing and Isolation Efficacy:\\Insights from a Simple Epidemic Model
-\thanks{Grants or other notes
-about the article that should go on the front page should be
-placed here. General acknowledgments should be placed at the end of the article.}
+% \thanks{Grants or other notes
+% about the article that should go on the front page should be
+% placed here. General acknowledgments should be placed at the end of the article.}
 }
@@
 \input{abstract.tex}
-\keywords{First keyword \and Second keyword \and More}
+\keywords{Epidemiology \and Infectious disease \and SARS-CoV-2 \and COVID-19 \and Reproductive number \and Testing and isolation}
 % \PACS{PACS code1 \and PACS code2 \and more}
@@
 
-%\begin{acknowledgements}
-%If you'd like to thank anyone, place your comments here
-%and remove the percent signs.
-%\end{acknowledgements}
+% \begin{acknowledgements}
+% If you'd like to thank anyone, place your comments here
+% and remove the percent signs.
+% \end{acknowledgements}
 
@@
 %
-% \section*{Conflict of interest}
-%
-% The authors declare that they have no conflict of interest.
+\section*{Declarations}
+
+\section*{Funding}
+
+DJDE, JD, and BMB were funded by the Natural Sciences and Engineering
+Research Council of Canada, the Michael G. DeGroote Institute for Infectious
+Disease Research at McMaster University, and the Public Health Agency of
+Canada.
+
+\section*{Conflict of interest}
+
+The authors declare that they have no conflict of interest.
+
+\section*{Availability of data and material}
+
+Not applicable.
+
+\section*{Code availability}
+
+Source code required to reproduce all analyses presented in this study are available at https://github.com/aligharouni/SIRmodel.
 

### 2021-07-21 manuscript_BMB.tex
cover letter upd

@@
 
-DJDE, JD, and BMB were funded by the Natural Sciences and Engineering
+(?) DJDE, JD, and BMB were funded by the Natural Sciences and Engineering
 Research Council of Canada, the Michael G. DeGroote Institute for Infectious
@@
 
-Source code required to reproduce all analyses presented in this study are available at https://github.com/aligharouni/SIRmodel.
+Source code required to reproduce all analyses presented in this study are available at \url{https://github.com/aligharouni/SIRmodel}.
 

### 2021-07-21 manuscript_BMB.tex
what is changed?
unifying sections into 1 file

@@
 \begin{abstract}
-\input{abstract.tex}
+
+Testing individuals for pathogens can affect the spread of epidemics. 
+Understanding how individual-level processes of sampling and reporting test results can affect community- or population-level spread is a dynamical modeling question.
+The effect of testing processes on epidemic dynamics depends on factors underlying implementation, particularly testing intensity and on whom testing is focused.
+Here, we use a simple model to explore how the individual-level effects of testing might directly impact population-level spread.
+Our model development was motivated by the \covid epidemic, but has generic epidemiological and testing structures. 
+To the classic SIR framework we have added a \percap testing intensity, and compartment-specific testing weights, which can be adjusted to reflect different testing emphases --- surveillance, diagnosis, or control.
+We derive an analytic expression for the relative reduction in the basic reproductive number due to testing, test-reporting and related isolation behaviours.
+Intensive testing and fast test reporting are expected to be beneficial at the community level because they can provide a rapid assessment of the situation, identify hot spots, and may enable rapid contact-tracing.
+Direct effects of fast testing at the individual level are less clear, and may depend on how individuals' behaviour is affected by testing information.
+Our simple model shows that under some circumstances both increased testing intensity and faster test reporting can \emph{reduce} the effectiveness of control, and allows us to explore the conditions under which this occurs.
+Conversely, we find that focusing testing on infected individuals always acts to increase effectiveness of control.
+
 \keywords{Epidemiology \and Infectious disease \and SARS-CoV-2 \and COVID-19 \and Reproductive number \and Testing and isolation}
@@
 \label{intro}
-\input{introduction.tex}
+
+The observed dynamics of the \covid epidemic have been driven both by epidemiological processes (infection and recovery) and by testing processes (testing and test reporting).
+In addition to shaping epidemic observations (via case reports), testing processes also alter epidemiological dynamics \citep{peto2020covid,taipale2020population}.
+Because individuals with confirmed infections (positive tests) are likely to self-isolate, and individuals who are awaiting the results of a test may also do so, testing will generally increase the number of people who are isolating and hence reduce epidemic growth rates.
+We developed a mechanistic model that incorporates epidemic processes and testing in order to explore the effects of testing and isolation on epidemic dynamics.
+
+If testing influences behaviour, then epidemic dynamics will depend on who gets tested.
+The impacts of testing will depend both on testing intensity (tests performed per day) and on how strongly testing is focused on people who are infectious.
+This level of focus depends in turn on the purpose and design of testing programs. 
+When testing is done for the purposes of disease surveillance \citep{foddai2020base}
+tests are typically conducted randomly (or using a stratified random design) across the population in order to make an unbiased assessment of population prevalence.
+
+Over the course of the \covid pandemic, however, the vast majority of testing has been done with other goals --
+primarily diagnostic (determining infection status for clinical purposes) \citep{phua2020intensive,who2020global}, or for control (determining  infection status in order to isolate cases that have been found by contact tracing) \citep{aleta2020modelling,kucharski2020effectiveness,grassly2020comparison,smith2020adherence}, which we characterize as \emph{targeted} testing strategies.
+In these situations, testing probabilities can differ sharply across epidemiological compartments; in our dynamical model, we will characterize these probabilities by assigning a testing weight to each compartment that determines the \emph{relative} probability that an individual in that compartment will be selected for testing (see Methods).
+
+Diagnostic testing focuses on people with infection-like symptoms; thus the relative testing weights for infected people will depend on the relative probability of infected people having symptoms.
+For \covid infection, the testing weights will depend on the proportion of asymptomatic infections, the time spent pre-symptomatic vs.\ symptomatic during the course of an infection, and on the incidence of \covid-like symptoms among people in the population \emph{not} infected with \covid.
+Testing for epidemic control focuses on known contacts of infected people; in this case the testing weights for infected vs.\ uninfected people will depend on the probability of infection given contact, as well as the effectiveness of the system for identifying suspicious contacts.
+
+When a new infectious disease emerges, it is important to determine whether it will grow exponentially in a susceptible population, and if so at what rate $r$ \citep{Ma+14}.
+The condition for positive exponential growth ($r>0$) is commonly expressed as $\Rnum>1$, where the basic reproduction number $\Rnum$ is the expected number of secondary infections arising from a typical infective individual in a completely susceptible population \citep{dietz1993estimation}.
+Although the value of $\Rnum$ cannot completely characterize the dynamics of our model \citep{shaw2021what}, it does give a simple and widely accepted index for the difficulty of control, as well as an indication of the likely final size of an epidemic \citep{ma2006generality,miller2012note}.
+
+In order to understand the effect of testing processes on epidemic dynamics, we expanded one of the simplest mechanistic epidemic models---the standard deterministic SIR model\citep{KermMcKe27,AndeMay91}---to include testing components.
+This model provides a sensible platform to link the modeling of epidemic and testing components and study their interaction.
+We studied the effects of testing intensity, rate of test return, and isolation efficacy, on transmission probability and epidemic dynamics when different levels of testing focus (from random to highly targeted) are in place.
 
@@
 \label{meth}
-\input{methods.tex}
+
+Our model groups individuals based on disease status (Susceptible, Infectious or Recovered) and testing status (\emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}) (\fref{fig:flowchart}).
+The testing status of an individual in a given disease compartment $X$ (where $X \in \{S,I,R\}$) is denoted by a subscript, namely $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
+Two `accumulator' compartments, $N$ and $P$, are included in order to collect cumulative reported negative or positive tests.
+The model equations \eqref{model} and details of calculation of the basic reproduction number $\Rnum$ are presented in \appref{R0}.
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=1.5]{pix/sir_comp.pdf}
+\caption{\small Flowchart of the SIR (Susceptible-Infectious-Recovered) model, \ref{model}.
+The disease-based status of a compartment $X$ ($X \in \{S,I,R\}$) is combined with the testing status including $X\_u$, $X\_p$, $X\_n$ and $X\_c$, for \emph{untested}, waiting-for-\emph{positive}, waiting-for-\emph{negative}, or \emph{confirmed positive}, respectively.
+The force of infection is denoted by $\Lambda$ (Eq.~\eqref{Lambda}); $\gamma$ is the recovery rate; $\omega$ is the rate of test return; and \testing{X} (Eq.
+\eqref{test}) and $p_X$ represent the \percap testing rate and the sensitivity (probability of positive tests), respectively, for compartment $X$.
+For further description of the parameters see Table~ \ref{tab:params}.
+Note that there is a slight mismatch in the top-to-bottom order of the testing-based compartments of each disease-based compartment $X$ between this flowchart and the model equations \eqref{model}; here we have switched  $X\_u$ and $X\_n$ for visual clarity.
+\label{fig:flowchart}}
+\end{center}
+\end{figure}
+
+Table~\ref{tab:params} defines the model parameters, which are generally \percap flows between compartments, or modifiers to these flow rates.
+The novel component of the model lies in the compartment-specific relative testing weights $w_S$, $w_I$ and $w_R$; these give the relative rates at which people in the $S$, $I$, and $R$ compartments are tested, respectively.
+Thus, we can specify different levels of testing focus from random (all weights equal) to highly targeted (higher weights in more intensively tested compartments).
+For example, $w_I/w_S=3$ means that infected individuals are tested at three times the \percap rate of susceptible individuals.
+
+In order to allow parameterization of the model by the total (overall) \percap testing rate, we define the weighted size of the testing pool $W = w_S S\_u + w_I I\_u + w_R R\_u$, and calculate a scaling parameter for testing as:
+\begin{equation}
+\label{sigma}
+\sigma = \frac{\rho N}{W},
+\end{equation}
+where $\rho$ is the \percap testing intensity for the population, defined as the number of daily tests administered in a population of size $N$.
+Thus, the \percap testing rate for compartment $X \in \{S,I,R\}$ is 
+\begin{equation}
+\label{test}
+\testing{X}=\sigma w_X.
+\end{equation}
+For a highly sensitive test, infected people typically flow through to the ``confirmed positive'' ($I\_c$, $R\_c$) compartments and are thus not considered for further testing.
+Over the course of the epidemic, a sufficiently large fixed testing rate as specified in \eqref{sigma} can exhaust the pool of people available for testing, leading to a singularity when too few people are left untested to support the specified rate.
+Although this phenomenon does not affect our analysis of $\Rnum$, it can affect model dynamics (we present an adjustment to the model that solves this problem in \appref{singularity}).
+
+The classical SIR model assumes a well-mixed population; homogeneity of the population (i.e., all individuals are equally susceptible and equally infectious with the same recovery rate when infected); exponentially distributed duration of infection; and large population size \citep{keeling2011modeling}.
+In addition to these standard assumptions, our model assumes:
+\begin{enumerate}[(i)]
+\item there is a single force of infection (new infections per unit time per susceptible), $\Lambda$, defined as
+  \begin{equation}
+  \label{Lambda}
+  \Lambda=\frac{\beta}{N} \big(I\_u + (1-\theta\_w)(I\_n+I\_p) + (1-\theta\_c)I\_c \big),
+  \end{equation}
+with transmission rate $\beta$; $\theta\_w$ is the isolation efficacy (reduction of the probability of transmission) for individuals waiting for test results, while $\theta\_c$ is the isolation efficacy for individuals who have received a ``confirmed'' positive test (Table~\ref{tab:params}).
+Susceptible individuals who are waiting for test results experience an additional transmission reduction factor of $1-\theta\_w$ (\fref{fig:flowchart}).
+\item confirmed-positive individuals isolate at least as effectively as those awaiting test results, i.e.,
+$$0\leq\theta\_w\leq\theta\_c\leq 1.$$ 
+\end{enumerate}
+For simplicity we assume that tests are perfectly \emph{specific} --- uninfected individuals never test positive ($p_s=0$).
+Thus there are no waiting-for-positive or confirmed-positive susceptible individuals, which reduces the number of model states from 12 to 10.
+
+\begin{table}[htp]
+\centering
+%%\begin{tabular}{|c|L{2in}cc|} \hline
+%% making use of ragged2e and array packages:
+{\RaggedRight
+\begin{tabular}{ c | m{6cm} | c | c }
+  \textbf{Symbol} & \textbf{Description} & \textbf{Unit} & \textbf{Value} \\ \hline
+  $N$     & Total population size & people & $10^6$ \\ \hline
+  $\omega$  & Rate of test return, i.e., rate of onward flow from ``waiting'' to ``confirmed'' or ``untested'' compartments  & 1/day & - \\ \hline
+  $\gamma$ & Recovery rate & 1/day & 1/6 \\ \hline 
+  $\rho$   & \percap testing intensity & 1/day & - \\ \hline 
+  $\theta\_w$ & Isolation efficacy (reduction of the transmission probability) for ``waiting'' individuals & - & - \\ \hline
+  $\theta\_c$ & Isolation efficacy for ``confirmed positive'' individuals & - & -  \\ \hline
+  $\beta$ & Transmission rate & 1/day & 0.5 \\ \hline
+  $\Lambda$ & Force of infection & 1/day & - \\ \hline
+  $p_S$ & Probability of positive tests for $S$\newline ($= 1-\textrm{specificity}$) & - & 0 \\ \hline
+  $p_I$ & Probability of positive tests for $I$\newline ($= \textrm{sensitivity}$) & - & 1 \\ \hline
+  $p_R$ & Probability of positive tests for $R$\newline ($= 1-\textrm{specificity}$) & - & 0.5 \\ \hline
+  $w_S, w_I, w_R$ & Relative testing weights & - &
+  \begin{minipage}[t]{0.21\columnwidth}%
+    Random:~$\{1,1,1\}$ \\ Targeted:~$\{0.3,1,1\}$
+  \end{minipage} \\
+  %%\hline
+\end{tabular}
+} % end \RaggedRight
+\caption{\label{tab:params}Parameters of the model specified  by the flow chart in \fref{fig:flowchart} and equations~\eqref{model}.}
+\end{table}
+
+The Disease-Free Equilibrium (DFE) for the expanded SIR model (Eqs.~\ref{model}) is found by setting the infected compartments to 0 and solving for the unknowns. The DFE is
+\begin{equation}
+\label{dfe}
+S\_n^*= \frac{\rho}{\omega} N, \ S\_u^*= N-S_n^*, \text{~and~} I_j=R_j=0 \ \text{for all }j.
+\end{equation}
+The corresponding \percap testing rate (Eq.~\ref{test}) for the infected compartment $I$ at DFE is one of the key analysis parameters and can be simplified as 
+\begin{equation}
+\label{eq:fi}
+\testinghat{I} = (\omega\rho/(\omega-\rho))w_I/w_S \quad .
+\end{equation}
+
+The basic reproduction number, $\Rnum$, was calculated by using the next-generation matrix method \citep{van2002reproduction}. We write $\Rnum$ as
+\begin{equation}
+\label{R0}
+\Rnum= \frac{\beta}{\gamma} \left(1-\Delta\right), 
+\end{equation}
+where $\beta/\gamma$ is the classical value for a simple model \citep{keeling2011modeling}, and $1-\Delta$ is the proportional reduction due to testing and isolation processes.
+$\Delta$ therefore measures the ``effectiveness of control'': how much these processes in reducing spread from low levels, and is in turn given by:
+\begin{equation}
+  \label{eq:del4a}
+  \Delta= \frac{1}{C N}\big(C_1 S\_u^*+(C_2(1-\theta\_w)+C \theta\_w)S\_n^*\big),
+\end{equation}
+where
+\begin{align}
+\label{eq:C12}
+C_{\phantom 1} &= (\omega+\gamma) \Big(\gamma(\omega+\gamma)+(\gamma+\omega p_I)\testinghat{I}\Big),\\
+C_1 &= (\omega+\gamma)(\theta\_w \gamma+\theta\_c \omega p_I) \testinghat{I},\\
+C_2 &= \Big( \omega\gamma(1+p_I)\testinghat{I}+\gamma^2(\omega+\gamma+\testinghat{I})\Big)\theta\_w + \omega^2p_I \testinghat{I} \theta\_c.
+\end{align}
+(\appref{R0} gives a detailed derivation of these expressions.)
+This explicit formula enables us to study the effects of testing and isolation parameters on $\Rnum$ both analytically and via numerical solutions.
+We are specifically interested in parameters that could be manipulated by public health policy: isolation efficacy, $\theta\_c$ and $\theta\_w$; \percap testing intensity, $\rho$; and the rate of test return, $\omega$.
+In particular, we look at the partial derivatives of $\Delta$ with respect to these parameters (Appendices~\ref{app:rho} and \ref{app:omega}).
+We derived general expressions for these derivatives.
+However, we analyzed the effect of $\omega$ on $\Delta$ for the special case of low testing intensity.
+Specifically, by making the restriction $\rho \ll 1$, we are able to Taylor-expand $\Delta$ at $\rho=0$, use the linear approximation with respect to $\rho$ and analyze the resulting simplified derivatives to illustrate a surprising non-monotonic relationship between $\Delta$ and $\omega$.
+
+Analytic calculation of the next-generation matrix and simplification of the $\Rnum$ expression, were performed in Maple\textsuperscript{\texttrademark} \citep{maple14}; numerical calculation and contour plots were done in R \citep{r}.
+We computed the values and contours of $\Delta$ at both low (\fref{pan}) and high (\fref{pan2}) testing intensities, and for both random testing ($w_S=w_I=w_R=1$) and targeted testing ($w_S=0.3$; $w_I=w_R=1$).
+Because it is expressed as a proportion of $\Rnum$, the effectiveness of control $\Delta$ is (at least in the $\rho \ll 1$ case, Eq.~\ref{eq:lin}) independent of the transmission rate $\beta$, and hence of $\Rnum$ in the case where we vary $\Rnum$ by changing the transmission rate for a fixed generation interval.
+
+The low-testing case (\fref{pan}) reflects the case where testing intensity $\rho$ is small relative to the population size.
+Specifically, $\rho \in [0,0.013]$, and test return rate $\omega\in [1/12,2]$.
+This testing intensity is of the correct order of magnitude (although typically larger than) testing rates during the \covid pandemic, i.e., a maximum of 1.3\% of the population per day (approximately four times the maximum testing rate in Ontario, Canada in mid-2021).
+The less realistic high-testing case (\fref{pan2}) is included to highlight the occurrence of non-monotonic changes in $\Rnum$ with respect to $\rho$.
+In \fref{pan2} the maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,1/5)$ and the test return rate $\omega\in [1/5,2]$; these values are clearly unrealistic for a large population but might be relevant for small populations undergoing focused testing, such as a sports league or university.
+In these figures, the implied baseline reproduction number (for the SIR model without testing) is $\Rnum=\frac{\beta}{\gamma}=3$.
+The different ranges of test return rates $\omega$ for the cases of low and high testing intensities is due to the restriction $\rho<\omega$, which is a requirement for a feasible DFE (\ref{dfe}).
 
@@
 \label{resu}
-\input{results.tex}
+
+We presented $\Rnum$ as the product of the classical reproduction number, $\beta/\gamma$, and the proportional reduction due to testing and isolation, $1-\Delta$, \eqref{R0}.
+We can use the formula for $\Delta$ \eqref{eq:del4a} to make a number of straightforward inferences about parameters that affect $\Rnum$ monotonically, i.e., for which the associated partial derivative of $\Delta$ always has the same sign (see Appendices).
+
+\begin{enumerate}
+
+\item \label{p1:eta} Increasing isolation efficacy for waiting ($\theta\_w$) and confirmed-positive ($\theta\_c$) individuals always increases $\Delta$ (Eqs.~\ref{eq:del4_theta},~\ref{eq:d_del4_thetac},~\ref{eq:d_del4_thetaw});
+\item \label{p1:rho} Higher testing intensity $\rho$ increases $\Delta$ if
+%% \begin{enumerate} \item
+testing is random (all $w_X$ equal) or testing intensity ($\rho$) is small (Eq.~\ref{eq:dd3dr}).
+%% \end{enumerate}
+\item \label{p1:omega} Increasing the rate of test return ($\omega$) always increases $\Delta$ if waiting individuals do not isolate ($\theta\_w=0$) (Eq.~\ref{eq:dlindo}).
+\item \label{p1:w} Increasing testing focus, i.e., changing the testing weights from random ($w_S=w_I$) toward targeted  ($w_S<w_I$), always increases $\Delta$ (Eq.~\ref{eq:d_del_wis}).
+\end{enumerate}
+
+\newpage
+\begin{figure}[h!]
+\centering
+\begin{subfigure}[t]{.49\textwidth}
+\centering
+\includegraphics[width=\linewidth]{codes/R0contour_random.pdf}
+\caption{}\label{p.a}
+\end{subfigure}
+%
+\begin{subfigure}[t]{.49\textwidth}
+\centering
+\includegraphics[width=\linewidth]{codes/R0contour_TTI.pdf}
+\caption{}\label{p.b}
+\end{subfigure}
+\caption{
+{\bf Effectiveness of testing and isolation in reducing $\Rnum$ at low \percap testing intensity ($\rho$).}
+Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4a}), over a range of testing and isolation parameters. Parameter values (Table~\ref{tab:params}):
+$\beta=\betaparam/$day, $1/\gamma= \invgammaparam$ days (baseline $\Rnum=\Rnumparam$, $r=\rparam$); $\omega \in [1/12,2]/$day;  $\rho \in [0,0.013]/$day per capita; $\theta\_w$ and $\theta\_c$ vary between 0 (no effect of isolation) and 1 (complete elimination of transmission); $p_S=0$, $p_I=1$ and $p_R=0.5$.
+Only parameter sets where $\theta\_c \geq \theta\_w$ (confirmed-positive individuals isolate more effectively than waiting individuals) are shown; the alternative case, $\theta\_w > \theta\_c$, is unrealistic.
+Contours of $\Delta$ are plotted for (a) random testing ($w_S=w_I=w_R=1$) and (b) targeted testing ($w_S=0.3$; $w_I=w_R=1$).
+}
+\label{pan}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\begin{subfigure}[t]{.49\textwidth}
+\centering
+\includegraphics[width=\linewidth]{codes/R0contour_random2.pdf}
+\caption{}
+\end{subfigure}
+%
+\begin{subfigure}[t]{.49\textwidth}
+\centering
+\includegraphics[width=\linewidth]{codes/R0contour_TTI2.pdf}
+\caption{}
+\end{subfigure}
+\caption{
+  {\bf Effectiveness of testing and isolation in reducing $\Rnum$ at high \percap testing intensity.}
+  Numerical evaluation of the effectiveness of control ($\Delta$: eq.~\ref{eq:del4a}), over a range of testing and isolation parameters.
+Parameters as in \fref{pan} except: $\omega \in [1/5,2]/$day, $\rho \in [0,1/5)/$day.
+As in \fref{pan}, subplots show (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$.
+}
+\label{pan2}
+\end{figure}
+
+However, there are also two specific cases where $\Delta$ changes non-monotonically, in counterintuitive directions, as a function of testing and isolation parameters.
+
+\begin{itemize}
+\item We would generally expect increasing testing delays to increase \Rnum, thus decreasing effectiveness of control $\Delta$.
+This is in fact what happens when waiting individuals do not isolate ($\theta\_w =0$, top row of \fref{pan}) --- as we move to the right within each plot in this row, $\Delta$ decreases.
+However, when waiting individuals isolate ($\theta\_w>0$), we more often see the opposite effect: longer testing delays lead to a greater control effect $\Delta$ (reduced $\Rnum$).
+The reason is that people waiting for negative tests are assumed to continue to isolate; this applies both to susceptibles and to people who became infected while waiting for negative test results.
+This effect outweighs the effect of confirmed individuals isolating, except when this isolation parameter ($\theta\_c$) is substantially greater than $\theta\_w$.
+This result depends on the idea that, all else equal, people who have to wait longer for test results isolate at the same level (but for a longer time) as they would if the wait were shorter.
+\item \fref{pan} also shows that greater testing intensity (increasing $\rho$) generally increases the effectiveness of control (moving up in each panel).
+However, this relationship can be reversed at very high testing intensities (provided testing is targeted, and $\theta\_w$ is relatively small; \fref{pan2}(b), right three panels of top row).
+It is theoretically possible for increasing testing intensity to \emph{increase} \Rnum because more rapid testing leaves more susceptibles in the ``waiting-for-negative-results'' category at the DFE; if these people become infected while waiting, they will need to wait for their negative test result before they can be tested again, receive a positive test, and then begin self-isolating.
+This effect is usually weak compared to the beneficial effects of testing.
+
+\end{itemize}
 
@@
 \label{disc}
-\input{discussion.tex}
+
+In this paper, we have developed and analyzed a simple compartmental model that combines epidemiological dynamics---as defined by a simple SIR model---with the dynamics of testing and isolation.
+Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness}).
+Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
+
+Many of the qualitative results we have derived confirm simple, common-sense intuitions.
+In particular, we can generally decrease $\Rnum$ by increasing isolation efficacy or testing intensity; returning tests faster, if individuals do not isolate while they are waiting for results; or increasing testing focus to target individuals who are likely to be infectious (e.g., symptomatic people or close contacts of known infections).
+
+However, we did find two surprising patterns: under some conditions longer delays in returning tests can reduce epidemic spread, and increasing testing rates can increase spread.
+
+Over broad regions of parameter space, decreasing $\omega$---i.e., \emph{slowing} the rate at which test results are returned---\emph{decreases} $\Rnum$ (for random testing, the parameter region is $\theta\_w \gtrsim 0.25$; for targeted testing, $\theta\_w \geq 0.25$ and either $\theta\_c \ge 0.5$ or $1/\omega > 5$; see Table~\ref{tab:params} for parameter definitions).
+This result is counterintuitive and would not be expected by public health authorities that have invested a great deal of effort in reducing delays from testing to results.
+Dynamically, this effect occurs because speeding up test returns shortens the isolation period of uninfected individuals (for infected people it only shortens the time to progression to the isolation level of the confirmed-positive compartment).
+Slowing test returns increases $\Rnum$ only if the proportion of infectives in the tested population is high and isolation is relatively strong among people waiting for test results.
+
+While slowing test returns does decrease $\Rnum$ over broad regions of parameter space in our model, there are several real-world processes missing from our model that make it unlikely that slowing test returns would actually be an effective public health measure.
+First, we do not model the primary benefit of rapid testing, i.e., detecting and containing outbreaks while they are still in progress.
+This process could be modeled phenomenologically by making the testing focus more targeted as an increasing proportion of cases is detected, because finding infections allows tests to be concentrated on their connections.
+Second, individuals may become less likely to maintain isolation if they are required to do so for longer; phenomenologically, we could allow  effectiveness of isolation in the waiting population to be an increasing function of test-return speed, or we could introduce a separate ``waiting, but no longer isolating'' compartment that individuals entered from the ``waiting, isolated'' compartment at a specified rate.
+Finally, if one wants to decrease the overall transmission rate of the population there are more effective methods than keeping tested people in limbo; these include masking, ventilation, distancing measures, retail and event closures, and stay-at-home orders.
+
+The other counterintuitive result from our analysis is that, for sufficiently high testing intensity $\rho$, further increasing testing intensity can actually \emph{increase} $\Rnum$ (e.g., \fref{pan2}(b), upper right panel [$\theta\_c=1$, $\theta\_w=0$]).
+This phenomenon can occur because we are considering the DFE in the presence of testing; thus there is an equilibrium distribution of susceptibles between the $S\_n$ (untested) and $S\_u$ (waiting) compartments even as the disease approaches extinction.
+A higher rate of testing leads to a greater proportion of individuals waiting for negative tests at the DFE.
+If infected, individuals in this group will take longer to be tested again and to subsequently isolate (because they must wait for their negative tests to be returned before being tested again).
+If isolation in this group ($\theta\_w$) is low, this effect can under some (relatively rare) circumstances (high $\theta\_c$, low $\omega$, high $\rho$) allow $\Rnum$ to increase with testing intensity.
+We can show that this phenomenon occurs \emph{only} under targeted testing ($w_I > w_S$), but we have not yet found a simple explanation of why it cannot occur under random (unfocused) testing.
+This phenomenon is also unlikely to occur in the real world.
+In particular, it depends on levels of testing that are unrealistically high (at least in large, general-population settings).
+
+Although we model the testing process in more detail than typical epidemiological models, one place where more detail could be informative is in the processes determining the testing weights $\{w_S, w_I, w_R\}$.
+While random testing, as done for surveillance purposes, unambiguously leads to equal testing weights, making precise quantitative connections between public-health practices and testing weights is difficult in other contexts.
+The testing weights reflect the correlation between an individual's risk of infection and their likelihood of being tested due to age, occupation, geographic location, etc..
+This correlation is influenced, among many other factors, by the proportion of the uninfected population with COVID-like symptoms (e.g., due to seasonal upper respiratory tract infections); the concentration of transmission and testing in hot spots such as long-term care facilities and high-density workplaces; the overall testing intensity (and hence, e.g., restriction to symptomatic individuals); and the proportion of COVID-infected people who are symptomatic.
+Future steps should explore mathematically tractable ways to model some of these factors more precisely.
+For example, separating the infected class into exposed, symptomatic, and a- or pre-symptomatic compartments and allowing the testing weights to vary across non-symptomatic (exposed/asymptomatic/presymptomatic) vs.\ symptomatic compartments could reflect the allocation of tests for diagnostic purposes (targeting symptomatic individuals) vs.\ contact-tracing (targeting infected but non-symptomatic individuals) vs.\ screening (relatively equal weights, depending on the venue).
+Alternatively, one could make the testing weights depend on the testing intensity or test-return rate as suggested above.
+Whatever complexity is added would probably put the model beyond reach of the analytical methods we have used in this paper, but one could still use semi-numerical methods such as constructing the next-generation matrix and using it to evaluate the derivatives of $\Rnum$ with respect to the parameters numerically.
+
+Although testing and tracing is a key part of infection control strategies, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts \citep{endo2020implication,hellewell2020feasibility,jenness2020modeling}, rather than analyzing simple but general models of the feedback between testing and transmission dynamics.
+There have been several modeling studies of testing and tracing dynamics and their interaction with epidemiological dynamics.
+In the context of repeated screening and random testing of isolated populations (such as the members of a university), \cite{bergstrom2020frequency} provided analytical results quantifying the effects that proactive screening of asymptomatic individuals and isolation of confirmed-positive cases could have in reducing the spread of disease.
+\cite{rogers2021high} simulated a SEIR model with testing and isolation; they similarly suggest a strategy of rapid testing with antigen tests and the subsequent isolation of confirmed-positive individuals.
+\cite{friston2021testing} model the effects of self-isolation on testing and tracing with a focus on projections under different testing and tracing scenarios.
+They conclude that the emergence of a second wave depends almost primarily on the rate at which immunity is lost and that it is necessary to track asymptomatic individuals in order to control the outbreak.
+Our modeling approach differs from these previous efforts in that it examines the effects of test-return rates and of different levels of testing focus, from random to highly targeted.
+We hope this paper will inspire further explorations of the fundamental properties of epidemic models that incorporate explicit testing processes.
 
@@
 
-(?) DJDE, JD, and BMB were funded by the Natural Sciences and Engineering
+This work was funded by the Natural Sciences and Engineering
 Research Council of Canada, the Michael G. DeGroote Institute for Infectious
@@
 \label{appe}
-\input{appendix.tex}
+
+\setcounter{equation}{0}
+\setcounter{figure}{0}
+\setcounter{table}{0}
+% \setcounter{page}{1}
+\makeatletter
+\renewcommand{\theequation}{A\arabic{equation}}
+\renewcommand{\thefigure}{A\arabic{figure}}
+\renewcommand{\bibnumfmt}[1]{[A#1]}
+\renewcommand{\citenumfont}[1]{A#1}
+
+\subsection{Model and calculation of $\Rnum$}\label{app:R0}
+
+The model in the form of a system of ordinary differential equations is 
+\begin{subequations}\label{model}
+\begin{align}
+ d S\_u/dt &= -\Lambda S\_u - \testing{S} S\_u + \omega S\_n, \\
+ d S\_n/dt &= -(1-\theta\_w)\Lambda S\_n + (1-p_S) \testing{S} S\_u - \omega S\_n, \\
+ d S\_p/dt &= -(1-\theta\_w)\Lambda S\_p + p_S \testing{S} S\_u - \omega S\_p, \\
+ d S\_c/dt &= -(1-\theta\_c)\Lambda S\_c + \omega S\_p, \\
+ d I\_u/dt &= \Lambda S\_u - \testing{I} I\_u + \omega I\_n  - \gamma I\_u, \\
+ d I\_n/dt &= (1-\theta\_w)\Lambda S\_n + (1-p_I) \testing{I} I\_u - \omega I\_n -\gamma I\_n, \\
+ d I\_p/dt &= (1-\theta\_w)\Lambda S\_p + p_I \testing{I} I\_u - \omega I\_p -\gamma I\_p, \\
+ d I\_c/dt &= (1-\theta\_c)\Lambda S\_c + \omega I\_p - \gamma I\_c, \\
+ d R\_u/dt &= \gamma I\_u - \testing{R} R\_u + \omega R\_n, \\
+ d R\_n/dt &= \gamma I\_n + (1-p_R) \testing{R} R\_u - \omega R\_n, \\
+ d R\_p/dt &= \gamma I\_p + p_R \testing{R} R\_u  - \omega R\_p, \\
+ d R\_c/dt&= \gamma I\_c + \omega R\_p, \\
+ dN/dt &= \omega (S\_n + I\_n + R\_n),  \\
+ dP/dt &= \omega(I\_p + R\_p) ,
+\end{align}
+\end{subequations}
+%
+(see Table \ref{tab:params} for parameter definitions).
+The next generation matrix for this model is $G = F V^{-1}$, where matrix $F$ represents the inflow of new infection to the infected compartments and matrix $V$ represents the flow in the infected compartments when the population is totally susceptible.
+Matrices $F$ and $V$ are
+\begin{align}
+\label{F}
+F =& \beta/N \left[ \begin {array}{cccc} 
+S\_u^* & (1-\theta\_w) S\_u^* & (1-\theta\_w) S\_u^* & (1-\theta\_c) S\_u^*\\
+(1-\theta\_w)S\_n^* & (1-\theta\_w)^2 S\_n^* & (1-\theta\_w)^2 S\_n^* & (1-\theta\_w)(1-\theta\_c) S\_n^* \\ 
+0&0&0&0\\
+0&0&0&0
+ \end {array} \right] \\ \notag
+ =&\beta/N \left[ \begin {array}{c} S\_u^* \\ (1-\theta\_w)S\_n^* \\ 0 \\ 0 \end {array} \right]
+        \left[ \begin {array}{cccc} 1,   1-\theta\_w,   1-\theta\_w,   1-\theta\_c \end {array} \right],
+\end{align}
+and 
+\begin{align}
+\label{V}
+V= \left[ \begin {array}{cccc}  
+\testinghat{I}+\gamma&-\omega&0&0\\
+-(1-p_I)\testinghat{I}&\omega+\gamma&0&0\\
+-p_I \testinghat{I}&0&\omega+\gamma&0\\
+0&0&-\omega&\gamma
+\end {array} \right].
+\end{align}
+The matrix inverse of $V$ is 
+\begin{align}
+\label{vinv}
+V^{-1} =
+\frac{1}{\gamma C}
+\left[ \begin {array}{cccc}
+\gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)&0&0\\ \noalign{\medskip}
+\gamma(\omega+\gamma)(1-p_I) \testinghat{I} & \gamma(\omega+\gamma)(\testinghat{I}+\gamma)&0&0 \\ \noalign{\medskip}
+\gamma(\omega+\gamma) p_I \testinghat{I} & \gamma\omega p_I \testinghat{I} & C \gamma/(\omega+\gamma) & 0 \\ \noalign{\medskip}
+\omega(\omega+\gamma) p_I \testinghat{I} & \omega^2 p_I \testinghat{I} & C \omega/(\omega+\gamma) & C
+\end {array} \right],
+\end{align}
+where $C= \big( \gamma(\omega+\gamma)+(\gamma+\omega p_I)\testinghat{I} \big) (\omega+\gamma)$ and $\testinghat{I}$ is the \percap testing rate for the infected people and represented in Eq.~\eqref{eq:fi}.
+Note that all the columns of matrix $V^{-1}$ summ up to $1/\gamma$.
+
+The particular form of $F$ with two rows of zeros at the bottom results in the following blocked form of matrix $G$.
+\begin{equation}
+\label{mat:G}
+G = \left[ \begin {array}{cc}
+G_{11}&G_{12}\\
+0&0
+\end {array} \right],
+\end{equation}
+where both blocked matrices $G_{11}$ and $G_{12}$ are 2 by 2.
+Given the upper triangular form of matrix $G$, the basic reproduction number $\Rnum$ (defined as the spectral radius of matrix $G$) is only determined by the blocked matrix $G_{11}$,
+\begin{equation}
+\label{G11}
+G_{11} = \frac{\beta}{\gamma   C} 
+\left[ \begin {array}{c} (\omega-\rho)/\omega \\ (1-\theta\_w)\rho/\omega \end {array} \right]
+\left[ \begin {array}{cccc} 1,   1-\theta\_w,   1-\theta\_w,   1-\theta\_c \end {array} \right]
+\left[ \begin {array}{cc}
+\gamma(\omega+\gamma)^2 & \gamma\omega(\omega+\gamma)\\ \noalign{\medskip}
+\gamma(\omega+\gamma)(1-p_I)\testinghat{I} & \gamma(\omega+\gamma)(\testinghat{I}+\gamma) \\ \noalign{\medskip}
+\gamma(\omega+\gamma) p_I \testinghat{I} & \gamma\omega p_I \testinghat{I} \\ \noalign{\medskip}
+\omega(\omega+\gamma) p_I \testinghat{I} & \omega^2 p_I \testinghat{I}
+\end {array} \right].
+\end{equation}
+It is notable that matrix $F$ \eqref{F} has rank one and consequently $G_{11}$ does so. That is $G_{11}$ has only one non-zero eigenvalue which is $\Rnum$.
+
+The expression of $\Rnum$ has a complicated form with all of the model parameters involved.
+This expression can be simplified and represented given the specific form of matrix $G_{11}$ \eqref{G11}.
+For the purpose of simplicity we present $\Rnum$ in the manuscript in terms of expressions $C$, $C_1$ and $C_2$, specified in \eqref{eq:C12}.
+
+It remains difficult to show that the reproduction number $\Rnum$ is decreasing with respect to \percap testing intensity, $\rho$, and the speed of the test return, $\omega$, for the feasible ranges of the parameters, that is
+\begin{align}
+\label{cond}
+& \omega>0, \\
+& 0 \leq \rho<\omega,\\ 
+& 0 \leq \theta\_w \leq \theta\_c \leq 1, \\
+& \oldtesttarget\geq 1.
+\end{align}
+In realistic cases the testing rate $\rho$ is very small (i.e., only a small fraction of the population can be tested every day); it is thus reasonable to use a linear approximation of $\Rnum$ for $\rho \ll 1$ to analyze the behaviour of $\Rnum$ with respect to $\omega$ (see section \appref{omega}).
+In the next section we provide an equivalent representation of $\Rnum$ in order to show that increasing testing intensity typically decreases $\Rnum$.
+
+% %%%%%%%
+\subsection{More testing intensity may decrease $\Rnum$}\label{app:rho}
+
+This section shows that $\pro{\Delta}$ can be positive or negative, with $\Delta$ defined in Eq.~\eqref{eq:C12}, and thus $\pro{\Rnum} < 0$, where $\Rnum$ is given in Eq.~\eqref{R0}.
+We rewrite matrix $G_{11}$ in \eqref{G11} in the following form to simplify the calculations:
+\begin{equation}
+\label{G112}
+G_{11} = \frac{\beta}{\gamma C} 
+\left[ \begin {array}{c}  S\_u^*/N \\ (1-\theta\_w) S\_n^*/N  \end {array} \right]
+\left[ \begin {array}{cccc} 
+C-C_1, C-C_2\end {array} \right],
+\end{equation}
+where $C$ is the same as the one in Eq. \eqref{eq:C12}, i.e., 
+$$C=(\omega+\gamma)(\gamma(\omega+\gamma)+(\omega p_I+\gamma) \testinghat{I}),$$
+and $C_1$ and $C_2$ are 
+\begin{align*}
+% \label{eq:C12}
+C_1 =& (\omega+\gamma)(\theta\_w   \gamma+\theta\_c  \omega  p_I) \testinghat{I},\\
+C_2 =& \big( \omega\gamma(1+p_I)\testinghat{I}+\gamma^2 (\omega+\gamma+\testinghat{I})\big)\theta\_w + \omega^2 p_I \testinghat{I} \theta\_c,
+\end{align*}
+where $\testinghat{I}$ is given in Eq.~\eqref{eq:fi}.
+Note that for analysis brevity, we let $N=1$, thus $S\_u^*$ and $S\_n^*$ are in the scale of 0 to 1.
+$\Rnum$ is in the same form as in Eq. \eqref{R0}  
+$$\Rnum= \frac{\beta}{\gamma} (1-\Delta),$$
+where 
+\begin{equation*}
+\label{eq:del4}
+\Delta= \frac{1}{C}\big(C_1 S\_u^*+(C_2(1-\theta\_w)+C \theta\_w) S\_n^*\big).
+\end{equation*}
+
+The first goal
+is to explore how changes in isolation, $\theta\_w$ and $\theta\_c$, affects $\Rnum$.
+Mathematically we would like to verify the sign of $\pder \Rnum{\theta\_w}$ and $\pder \Rnum{\theta\_c}$.
+We start with simplifying $\Delta$ \eqref{eq:del4} by factoring $\theta\_w$ and $\theta\_c$ in Eq.~\eqref{eq:del4}.
+Thus, $\Delta$ can be rewritten as
+\begin{equation}
+\label{eq:del4_theta}
+\Delta= \frac{1}{C}\Big(
+-D_1 S\_n^* \theta\_w^2 +
+\big(-\omega^2 p_I \testinghat{I} S\_n^*\theta\_c+ D_2 S\_n^*+\gamma \testinghat{I}(\omega+\gamma) \big)\theta\_w+
+(\omega+\gamma S\_u^*)\omega p_I \testinghat{I} \theta\_c
+\Big),
+\end{equation}
+where
+\begin{align}
+\label{eq:del4_d1d2}
+D_1=& (\omega+\gamma)\gamma^2+(\omega+\gamma+\omega p_I)\gamma \testinghat{I}, \\
+D_2=& (3\omega+2\gamma)\gamma^2+(\omega+\gamma+2\omega p_I)\gamma \testinghat{I}+(\gamma+p_I \testinghat{I})\omega^2.
+\end{align}
+$\Delta$, Eq.~\eqref{eq:del4_theta}, is linear in $\theta\_c$ with a positive coefficient. thus
+\begin{equation}
+\label{eq:d_del4_thetac}
+\pder\Delta{\theta\_c}=1/C(\gamma S\_u^*+\omega(1-\theta\_w S\_n^*))\omega p_I \testinghat{I}.
+\end{equation}
+This results in increasing $\Delta$, thus decreasing $\Rnum$ with respect to $\theta\_c$, that is $\pder \Rnum{\theta\_c}\leq 0$. Note that $C$ is independent of $\theta\_c$ and $\theta\_w$.
+
+With a similar logic, $\Delta$ \eqref{eq:del4_theta} is a concave-down quadratic equation in $\theta\_w$, given by
+\begin{equation}
+\label{eq:del4_thetaw}
+1/C \Big( -D_1 S\_n^*\theta\_w^2+
+\big(-\omega^2 p_I \testinghat{I} S\_n^*\theta\_c+D_2 S\_n^*+\gamma \testinghat{I}(\omega+\gamma) \big)\theta\_w\Big).
+\end{equation}
+We show that the feasible range of $\theta\_w$ lies between 0 and the vertex of this parabola where the parabola is increasing in $\theta\_w$, and so does $\Delta$ which results in inferring $\pder\Rnum{\theta\_w}\leq 0$.
+It is enough to show that partial derivative of the expression \eqref{eq:del4_thetaw} with respect to $\theta\_w$ at $\theta\_w=1$ is non-negative. It follows that
+\begin{align}
+\label{eq:d_del4_thetaw}
+\pder\Delta{\theta\_w}\bigg\rvert_{\theta\_w=1} =&
+1/C\Big( (D_2-2D_1-\omega^2 p_I \testinghat{I} \theta\_c) S\_n^*+\gamma \testinghat{I}(\omega+\gamma) \Big) \\\notag
+=& 1/C\Big( (\gamma(\omega+\gamma)+\gamma\omega^2+(1-\theta\_c)\omega^2 p_I \testinghat{I} ) S\_n^*
++\gamma(\omega+\gamma) \testinghat{I} (1-S\_n^*) \Big),
+\end{align}
+which is a positive quantity, given that $\theta\_c$ and $S\_n^*$ vary between 0 and 1.
+
+The second goal is to explore how changes in \percap testing intensity $\rho$ affects $\Rnum$.
+Mathematically we would like to verify the sign of $\pder\Rnum{\rho}$, which specifically depends on $\pder\Delta{\rho}$.
+We use the derived expressions for $S\_u^*$ and $S\_n^*$, given by Eqs.~\eqref{dfe}, in $\Delta$ \eqref{eq:del4}.
+Also, we define $\phi = \testinghat{S} = \frac{\rho \omega}{\omega-\rho}$, to reparameterize $\rho$.
+This is mainly to avoid singularity in $\testinghat{I}$ \eqref{eq:fi}, when testing intensity $\rho$ is very close to the rate of test return $\omega$.
+Thus, $\rho$ is reparameterized as
+\begin{equation}
+\label{eq:phi}
+\rho=\frac{\omega \phi}{\omega+\phi}.
+\end{equation}
+This one-to-one monotonic reparameterization enables us to simplify the mathematical expressions and explore the simpler $\pder\Delta{\phi}$ instead of the complicated $\pder\Delta{\rho}$.
+Defining $\testtarget \equiv \oldtesttarget$, the derivative is 
+\begin{align}
+\label{eq:dd3dr}
+\partial\Delta/\partial\phi=& \frac{1}{d_3} (a_3 \phi^2+b_3 \phi+c_3),
+\end{align}
+where
+\begin{align}
+\label{eq:abcd2}
+a_3& = \testtarget \Bigg(
+&& (1-\theta\_w) (1+\testtarget)\theta\_w \gamma^3 +(1-\theta\_c) p_I^2 \theta\_w \testtarget \omega^3 \\ \notag
+& ~ &&+\Big( \big( (1-\theta\_w-\testtarget) \theta\_c +(3-2\theta\_w) \theta\_w\testtarget \big) p_I +(1-\theta\_w) (1+\testtarget)\theta\_w  \Big) \omega \gamma^2 \\ \notag
+& ~ &&+\Big(
+\big((1-\theta\_w-\theta\_w \testtarget) \theta\_c+ (2-\theta\_w) \theta\_w \testtarget \big) p_I
++(2 \theta\_w-\theta\_w^2-\theta\_c) \testtarget p_I^2
+\Big) \omega^2 \gamma \Bigg), \\ \notag
+b_3& = && \hspace{-1cm} 2\testtarget (\omega+\gamma)\gamma\Big(
+(\omega+\gamma+\omega p_I)(2-\theta\_w) \gamma \theta\_w+ (1-\theta\_w) \omega^2 p_I\theta\_c+\omega^2 p_I\theta\_w
+\Big) ,\\ \notag
+c_3& = && \hspace{-1cm} (\omega+\gamma)^2 \gamma\Big(
+(2-\theta\_w) \gamma^2 \theta\_w+
+(1+\testtarget) \omega \gamma \theta\_w+
+\testtarget \omega^2 p_I\theta\_c \Big),\\ \notag
+d_3&  = && \hspace{-1cm} \frac{(\omega+\gamma)}{\omega}  \Big((\omega p_I+\gamma)\testtarget\phi + (\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2.
+\end{align}
+Note that $\phi\geq 0$, also $b_3$, $c_3$ and $d_3$ are all positive. However $a_3$ can be positive or negative.
+If $a_3\geq 0$, $\partial\Delta/\partial\phi \geq 0$ for all feasible range of parameters, thus $\pro\Rnum \leq 0$.
+It is straightforward to show that $a_3\geq 0$ when testing is random, i.e., $w\_S=w\_I=1$.
+If $a_3 < 0$, then the quadratic expression in the numerator of \eqref{eq:dd3dr} has a positive root, $\phi^*$, such that for $\phi>\phi^*$, $\partial\Delta/\partial\phi < 0$.
+
+An example of this countervailing effect of $\phi$, and consequently $\rho$, on $\Rnum$ occurs when $\theta\_w=0$ and $\theta\_c=1$.
+This is illustrated in the top-right panel of the \fref{pan2} panel (b), where the strength of isolation for awaiting people is the least, but the most for the confirmed cases.
+In this case, simplifying $a_3$ in Eq.~\eqref{eq:abcd2} gives
+\begin{equation}
+  \begin{split}
+    a_3 & =\testtarget \omega \gamma p_I\left((\omega+\gamma)-\testtarget(\omega p_I+\gamma)\right)  \\
+    & \propto  \omega \left( 1- \testtarget p_I \right) + \gamma \left( 1- \testtarget \right)  .
+  \end{split}
+\end{equation}
+If $p_I>0$, then $a_3<0$ for sufficiently targeted testing (i.e. when $\testtarget p_I >1$; because
+$p_I \leq 1$, \testtarget is always $\geq \testtarget p_I$).
+%% this LOOKS like we can get a3<0 even if pI=0 for targeting > 1 and sufficiently large gamma/omega,
+%% but that seems crazy so I'm probably missing something
+When the test is perfectly sensitive ($p_I=1$), $a_3<0$ as long as $\testtarget>1$.
+Under either of these conditions, there exists a range for $\rho$ over which $\pder\Rnum{\rho}\leq 0$.  
+Because increasing values of $\rho$ and $\omega$ both delay the rate at which individuals flow to the $I\_c$ compartment, it is reasonable that increasing either value could (under appropriate circumstances) increase $\Rnum$.
+
+% %%%%%%%
+\subsection{Rate of returning tests} \label{app:omega}
+The third goal is to explore how changes in the rate of test return $\omega$ affect $\Rnum$.
+Mathematically we would like to verify the sign of $\pder\Rnum{\omega}$, which specifically depends on $\pder\Delta{\omega}$.
+We use
+the linearization of $\Delta$ when $\rho \ll 1$ to show that there a non-monotonic relationship between $\Rnum$  and $\omega$. The linear term in the Taylor expansion of $\Delta$ when $\rho \ll 1$ is
+\begin{equation}
+\label{eq:lin}
+\Delta = \frac{\rho}{\omega \gamma(\omega+\gamma)} \Big(
+\testtarget \omega^2 p_I \theta\_c+(\testtarget+1) \gamma\omega\theta\_w +\gamma^2 \theta\_w(2-\theta\_w)
+\Big). 
+\end{equation}
+This results in
+\begin{equation}
+\label{eq:dlindo}
+\pder\Delta{\omega} = \frac{\rho}{\omega^2 (\omega+\gamma)^2} \Big(
+(p_I \testtarget\theta\_c -(1+\testtarget)\theta\_w)\omega^2-2\theta\_w \gamma(2-\theta\_w) \omega+\theta\_w \gamma^2 (\theta\_w-2)
+\Big).
+\end{equation}
+% ALi & Faddy's stuff
+The latter expression has two roots
+\begin{align}
+\label{eq:omega_roots}
+\omega^*_{+} &= \frac{\gamma\Big(-c_4 + \sqrt{c_4^2 + (a_4-b_4)c_4}\Big)}{b_4-a_4} \\
+\omega^*_{-} &= \frac{\gamma\Big(-c_4 - \sqrt{c_4^2 + (a_4-b_4)c_4}\Big)}{b_4-a_4},
+\end{align}
+where $a_4 = p_I\testtarget\theta\_c$, $b_4 = (1+\testtarget)\theta\_w$, $c_4 = \theta\_w(2-\theta\_w)$.
+This enables us to describe the behaviour of $\Rnum$ with respect to $\omega$ in the following two cases.
+\begin{itemize}
+\item \textbf{Case I:} If $b_4 \geq a_4$, $\pder\Delta{\omega} < 0$, so $\Rnum$ is always increasing with respect to $\omega$ (i.e., it is always harmful to return tests more rapidly). 
+\item \textbf{Case II:} If $b_4 < a_4$, $\Rnum$ will be decreasing with respect to $\omega$ (i.e., returning tests more rapidly is beneficial) only when $\omega > \omega^*_-.$
+\end{itemize}
+Note that $b_4 \geq a_4$ is characterized by
+\begin{equation}\label{case_1_ineq}
+\Big(\testtarget+1\Big)\theta\_w \geq \testtarget p_I\theta\_c \Longleftrightarrow \Big(\frac{w_S}{w_I}+1\Big)\frac{1}{p_I} \geq \frac{\theta\_c}{\theta\_w}.
+\end{equation}
+We begin with a \textbf{proof of Case I}.
+Suppose that $b_4 \geq a_4$.
+If the roots of $\pder\Delta{\omega}$ are not complex, then we must have $c_4 \geq b_4-a_4$.
+Note that $\omega^*_{-}$ must be negative since the numerator is clearly negative but the denominator is positive.
+Next, note that since $c_4 \geq b_4-a_4$, the numerator of $\omega^*_{+}$ must also be negative, so $\omega^*_{+}$ is negative.
+Thus, in this case, $\pder\Rnum{\omega}$ does not change sign on $(0,\infty)$.
+Checking the sign of \ref{eq:dlindo} for arbitrarily large $\omega$ shows that it is negative (since the $(a_4-b_4)\omega^2$ term dominates and is negative).
+So $\Rnum$ is increasing with respect to $\omega$ on all of $(0,\infty)$.
+
+We now turn our attention to a \textbf{proof of Case II}.
+Suppose that $b_4 < a_4$.
+It follows that $\omega^*_+$ and $\omega^*_-$ are real since $c_4 > 0 > b_4-a_4$.
+Next, note that $\omega^*_-$ is positive since both the numerator and denominator are negative.
+On the other hand, $\omega^*_+$ is negative since the denominator is negative but the numerator is positive (because $c_4 > b_4-a_4$).
+Thus, our task is to understand the sign of $\pder\Delta{\omega}$ around the root $\omega^*_-$.
+Checking the sign of \ref{eq:dlindo} for arbitrarily large $\omega$ shows that it is positive (since the $(a_4-b_4)\omega^2$ term dominates and is positive).
+Likewise, checking the sign for values of $\omega$ close to $0$ shows that it is negative.
+Thus, $\Rnum$ is increasing with respect to $\omega$ when $\omega < \omega^*_-$, and is decreasing $\omega > \omega^*_-$.
+
+Having presented the formal analysis, we now concern ourselves with its biological interpretations.
+We begin by interpreting \ref{case_1_ineq}, under which returning tests more rapidly is \emph{always} harmful.
+Notice that the ratio $\frac{\theta\_c}{\theta\_w}$ is simply a measure of how much more strongly individuals self-isolate when they test positive compared to when waiting for tests.
+Since the rate of test return directly influences the rate at which individuals change from a waiting state to a confirmed-positive state, it is intuitive that $\frac{\theta\_c}{\theta\_w}$ would appear in \ref{case_1_ineq}.
+Next, note that the left-hand side increases when test sensitivity decreases and when targeting of positive individuals is poor.
+This is consistent with our intuition: a false negative that is returned more rapidly will allow an infectious individual to relax their self-isolation, thus increasing transmission.
+Likewise, if individuals tested are mainly susceptible (rather than infectious), then returning tests more slowly would encourage them to self-isolate for longer while awaiting test results.
+Having understood the role of each of the parameters in \ref{case_1_ineq}, a holistic interpretation of this inequality is that returning tests more slowly is helpful when the benefit of extended self-isolation by infected individuals awaiting test results outweighs the benefit of identifying positive cases.
+
+Now we interpret \textbf{Case II}.
+In this case, $\Rnum$ will have a global maximum with respect to $\omega$ at $\omega^*_-$.
+Note that our model assumption that $\rho < \omega$ plays an important role here: if $\omega^*_- < \rho$, then $\Rnum$ will be always decreasing with respect to $\omega$.
+On the other hand, if $\rho < \omega^*_-$, then $\Rnum$ will be increasing with respect to $\omega$ on $(\rho, \omega^*_-)$ and decreasing beyond that.
+
+% %%%%%%%
+\subsection{The effect of testing focus parameter $\testtarget$ on $\Rnum$} \label{app:w}
+
+\begin{align}
+\label{eq:d_del_wis}
+\pder \Delta{w\_{IS}}= \dfrac{(\omega-\rho) (\omega(\omega-\rho\theta\_w)+\gamma(\omega-\rho)) (\theta\_w\gamma+\theta\_c\omega p\_I)}
+{(-\omega^2 \gamma+\omega \gamma \rho-\gamma \rho \omega w\_{IS}-
+\omega \gamma^2+\gamma^2 \rho-\omega^2 p_I \rho w\_{IS})^2},
+\end{align}
+which is a positive quantity. Thus, $\pder \Rnum{w\_{IS}}\leq 0$. Therefore, increasing the focus of testing on the infectious people will result in less transmission.
+
+% %%%%%%%
+\subsection{On Testing Rate and Numerical Singularity} \label{app:singularity}
+
+In this work, we do not present any numerical solutions of the ODEs to investigate typical trajectories.
+However, attempting to do so quickly reveals a problem in the way the model is posed;
+with the scaling of testing ($\sigma$) defined as in the body of the paper (Eq.~\eqref{sigma}), the population in the $S$ compartments appeared to blow up when the system is near the DFE.
+This occurs because once the only untested people are susceptibles, the FOI approaches $\Lambda=0$, and the testing rate $\testing{S} \to \rho N/S\_u$.
+Thus, the first equation of the model \eqref{model} will become
+$d S\_u/dt = - \rho N + \omega S\_n$. Thus changes in $S\_u$ will be independent on $S\_u$, and the decay of the $S\_u$
+population becomes linear rather than exponential --- allowing $S\_u$ to become negative!
+To avoid this problem the testing rate, $\sigma$, should be formulated such that people from the untested compartments will not be tested if they are not there.
+One way to fix this issue, is to consider a maximum testing rate, $\tau$ (1/day).
+In general, we want to test at a rate of $\rho$ across the whole population.
+This won't always be possible, so we impose a maximum rate of $\tau$ per testable person and redefine $\sigma = \frac{\tau \rho N}{\tau W + \rho N}$, with the assumption that $\tau \gg \rho$.
+This modification of $\sigma$ does not affect any of the results we have derived about the invasion of the epidemic from the DFE (i.e., results on $\Rnum$ and $\Delta$).
 

### 2022-01-06 response1_BMB.tex
starting responses

\documentclass{article}

\usepackage{amssymb,amsmath,amsthm}
\usepackage{xcolor}

\newcommand{\comment}{\showcomment}
\newcommand{\showcomment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}
\newcommand{\nocomment}[3]{}

\newcommand{\fady}[1]{\comment{cyan}{Fady}{#1}}
\newcommand{\ali}[1]{\comment{magenta}{Ali}{#1}}
\newcommand{\jd}[1]{\comment{blue}{JD}{#1}}
\newcommand{\david}[1]{\comment{orange}{DJDE}{#1}}
\newcommand{\bmb}[1]{\comment{red}{BMB}{#1}}
\newcommand{\todo}[1]{\comment{red}{TODO}{#1}}
\newcommand{\new}[1]{\comment{red}{NEW}{#1}}
\newcommand{\needref}{\textcolor{red}{(REFS?)}}
\newcommand{\res}[1]{\comment{red}{Response}{#1}}

\begin{document}

COMMENTS TO THE AUTHOR:

Reviewer \#1: The authors introduce a compartmental epidemic model exploring Susceptible-Infectious-Recovered (SIR) disease dynamics in the presence of a testing protocol that can identify and facilitate isolation of infectious individuals, exploring how factors including the daily intensity of testing, the delay between testing and the reporting of test results, and the degree of self-isolation practiced by individuals waiting for test results and with confirmed positive tests impact the early spread of the disease outbreak. Using the next-generation method, the authors characterize the basic reproduction number (R0) for their model, and employ a mix exact calculations and analytical approximations to study the comparative statics characterizing how this quantity depends on parameters characterizing testing and isolation practices. Some of the main results agree with intuitive expectations about testing and self-isolation, such as the observation that that R0 decreases with a reduction in social contacts in social contacts (i.e. increases in $theta_c$ and $theta_w$) for confirmed positives and individuals awaiting test results, as well the observation that focusing testing on infectious individuals helps to decrease R0 relative to a random testing protocol. On the other hand, some of the properties of the formula for R0 run counter to expectations, such as the possibility that increased testing intensity can increase R0 in the for the case in which individuals waiting for results do not decrease their social contacts, while R0 can decrease with increased delay times for test results for the case in which waiting individual greatly reduce their social contacts.

Overall, this is a nice paper that illustrates a stylized model incorporating the effects of testing intensity, testing delays, and testing-based self-isolating behaviors impact the early spread of an infectious disease. Both the main paper and the appendix are well-written, and the results from the paper provide interesting insight into how the feedbacks between testing implementation and behavioral self-isolation can impact our ability to slow an outbreak. The analytical techniques used in the appendix provide a nice illustration on how combinations of the next-generation matrix and comparative statics on R0 can be used to understand a relatively large compartmental model, and both the biological subject matter and mathematical techniques are well suited for the audience of Bulletin of Mathematical Biology. In addition, the presence of surprising behaviors in this baseline model for testing may also serve as inspiration for further extensions exploring additional factors like exposed or asymptomatically- infectious disease compartments or to explore how testing regimes can work with dynamical feedback away from the disease-free equilibrium. As a result, I think that this would be a nice paper to publish in BMB, and will serve as an interesting jumping-off point for readers to explore the role of testing in disease outbreaks.

In preparation for revising the paper, I have the following minor suggestions that may be helpful for further highlighting results or clarifying the model.

* The main surprising results about the possibility of R0 increasing with increased testing intensity or decreasing with increased delays seem particularly intertwined with the behavior displayed by individuals who are waiting for their test results. In particular, having long delays can make versions of the model with high $\theta_w$ effectively treat these waiting individuals into a partially socially-distanced or quarantined compartment, while, in versions in which waiting individuals do not change their behavior, increased asymptomatic testing can provide extra time to miss new infectious. This connection between behavior of waiting and confirmed individuals and how testing impacts R0 makes me curious about a few aspects of combining testing with social distancing.

- Does the case of R0 decreasing with increased testing delay provide some suggestion of the potential benefit of so-called "gateway testing" [1] in settings such as universities or athletic competitions?
Under such approaches, individuals must practice increased levels of social distancing or self- isolation until they have achieved a series of confirmed negative tests. While this type of time- dependent testing and self-isolation protocol might not fit well within the framework of linear stability of a disease-free equilibrium, the relationship between increased $theta_w$ and the benefits of a longer waiting period $1/\omega$ seems to indicate the possibility that combining waiting for results and self-isolation can help to stem early outbreaks (e.g. at the beginning of a university term).

\res{Yes, our result suggests the "gateway testing". We (will) add this point to the last paragraph of the discussion and the suggested paper [1] will be added to our references.}

- Would similar effects on $theta_X$ occur if the model incorporated an exposed class into the model, and would that depend on the way in which positive test results correspond to time-period of infectiousness? With an exposed or asymptomatic and infectious, there may be interesting ways in which waiting for tests and the presence or lack of self-isolation while waiting could feed back on subsequent transmission and on changes in R0.

While I think that the analysis already presented in the paper is sufficient for the current paper, I think that ideas related to these questions might be interesting for future research or as points for discussion.

* Due to its important effects on R0, I think it could be useful to describe the rate $\omega$ for return of test results directly in the text (in addition to the table of parameters) before it appears in Equation 4 in the expression for the disease-free equilibrium.

* In Section A.4 of the appendix, is the partial derivative of $\Delta$ being calculating for the exact expression from Equation 7, or from the approximate expression in Equation A22 obtained from a Taylor expansion when $\rho \ll 1$. From the expression in Equation A27, it seems like this is derived from Equation 7, but it would be helpful to clarifying which expression is used to highlight whether this result holds across the range of parameters or only in the limit of weak testing intensity.

References
[1] K. Muller and P. A. Muller, "Mathematical modelling of the spread of covid-19 on a university campus," Infectious Disease Modelling, vol. 6, pp. 1025-1045, 2021.

Reviewer \#2: The idea underlying the paper is very simple but powerful: public health policies can impact behaviors and thus epidemiology. Even though the reality is much more complex (infectious spreading results from the complex, non-linear, multi-factorial array and subtle interplay between epidemiology, public health policies and behaviors), authors did a good job in developing their mathematical model, that, rightly, they define as a "caricature". They should add it is a "caricature" not only because it (purposely) does not incorporate sophisticated aspects of COVID-19 related epidemiology, but also because does not incorporate that array and interplay I previously mentioned (the array consisting of multidirectional feedbacks between epidemiology, behaviors and public health policies). Anyway, authors should add and discuss this point.
Another point authors should discuss is future prospects, how their model can be further refined and what we can learn from this model in terms of public health policies (discussion is too succinct with this regard).

\end{document}

### 2022-01-17 manuscript_BMB.tex
working on the responses to reviewers comments

@@
 
+\newcommand{\comment}{\showcomment}
+\newcommand{\showcomment}[3]{\textcolor{#1}{\textbf{[#2: }\textsl{#3}\textbf{]}}}
+\newcommand{\nocomment}[3]{}
+
 \newcommand{\fref}[1]{Fig.~\ref{#1}}
@@
 \newcommand*\subtxt[1]{_{\textnormal{#1}}}
+\newcommand{\rev}[1]{\comment{red}{Revised}{#1}}
 %% Roman subscript
@@
 \cite{rogers2021high} simulated a SEIR model with testing and isolation; they similarly suggest a strategy of rapid testing with antigen tests and the subsequent isolation of confirmed-positive individuals.
+% reviewer1.com1
+\rev {\cite{muller2021mathematical} simulated a structured SEIR-type model.
+Their model incorporated different testing scenarios -- including gateway testing, surveillance testing, and contact tracing -- as well as individual level control measures such as mask wearing and social distancing.
+They indicate the benefit of gateway testing in conjunction with effective contact tracing or robust surveillance testing in controling the spread of the infection.
+Under such approaches, individuals must practice increased levels of social distancing or self- isolation until they have achieved a series of confirmed negative tests.
+It is notable that while this type of time- dependent testing and self-isolation protocol might not fit well within the framework of linear stability of a disease-free equilibrium, the relationship between increased $theta_w$ and the benefits of a longer waiting period $1/\omega$ seems to indicate the possibility that combining waiting for results and self-isolation can help to stem early outbreaks (e.g.
+at the beginning of a university term).}
 \cite{friston2021testing} model the effects of self-isolation on testing and tracing with a focus on projections under different testing and tracing scenarios.

### 2022-01-17 response1_BMB.tex
working on the responses to reviewers comments

@@
 \newcommand{\needref}{\textcolor{red}{(REFS?)}}
+\newcommand{\com}[1]{\comment{red}{Comment}{#1}}
 \newcommand{\res}[1]{\comment{red}{Response}{#1}}
@@
 \begin{document}
+
+Dear Dr. Hastings,
+
+Please find our responses to the reviewer comments below for our manuscript entitled
+``Testing and Isolation Efficacy:Insights from a Simple Epidemic Model''.
+The reviewer comments are in italics, and our comments/responses are in normal font.
+To address the concerns of the reviewers, we have added appropriate references to the manuscript to acknowledge previous known results in the field.
+
+We very much appreciate the time and feedback from the reviewers and editors.
+
+Sincerely,
+Ali Gharouni, PhD
 
@@
 
-* The main surprising results about the possibility of R0 increasing with increased testing intensity or decreasing with increased delays seem particularly intertwined with the behavior displayed by individuals who are waiting for their test results.
+\com 1 {\it The main surprising results about the possibility of R0 increasing with increased testing intensity or decreasing with increased delays seem particularly intertwined with the behavior displayed by individuals who are waiting for their test results.
 In particular, having long delays can make versions of the model with high $\theta_w$ effectively treat these waiting individuals into a partially socially-distanced or quarantined compartment, while, in versions in which waiting individuals do not change their behavior, increased asymptomatic testing can provide extra time to miss new infectious.
@@
 While this type of time- dependent testing and self-isolation protocol might not fit well within the framework of linear stability of a disease-free equilibrium, the relationship between increased $theta_w$ and the benefits of a longer waiting period $1/\omega$ seems to indicate the possibility that combining waiting for results and self-isolation can help to stem early outbreaks (e.g.
-at the beginning of a university term).
+at the beginning of a university term).}
 
-\res{Yes, our result suggests the "gateway testing". We (will) add this point to the last paragraph of the discussion and the suggested paper [1] will be added to our references.}
+\res 1 
+We now have added several lines and the suggested reference [1] to the end of the discussion to acknowledge the "gateway testing" and the potential benefit of our modelling results.
 
-- Would similar effects on $theta_X$ occur if the model incorporated an exposed class into the model, and would that depend on the way in which positive test results correspond to time-period of infectiousness?
+\com 2 {\it Would similar effects on $\theta_X$ occur if the model incorporated an exposed class into the model, and would that depend on the way in which positive test results correspond to time-period of infectiousness?
 With an exposed or asymptomatic and infectious, there may be interesting ways in which waiting for tests and the presence or lack of self-isolation while waiting could feed back on subsequent transmission and on changes in R0.
+}
+
+\res 2
+Regarding the incorporation of an exposed class or asymptomatic and infectious compartment to the current model; I think
 

### 2022-01-18 response1_BMB.tex
reviewer 1 covered

@@
 \res 2
-Regarding the incorporation of an exposed class or asymptomatic and infectious compartment to the current model; I think
+[not sure here!] 
+Regarding the incorporation of an exposed class or asymptomatic and infectious compartment to the current model; the isolation efficacy parameter $\theta_X$ would have similar but stronger effects (result 1).
 
@@
 
-* Due to its important effects on R0, I think it could be useful to describe the rate $\omega$ for return of test results directly in the text (in addition to the table of parameters) before it appears in Equation 4 in the expression for the disease-free equilibrium.
+\com 3 
+Due to its important effects on R0, I think it could be useful to describe the rate $\omega$ for return of test results directly in the text (in addition to the table of parameters) before it appears in Equation 4 in the expression for the disease-free equilibrium.
 
-* In Section A.4 of the appendix, is the partial derivative of $\Delta$ being calculating for the exact expression from Equation 7, or from the approximate expression in Equation A22 obtained from a Taylor expansion when $\rho \ll 1$.
+\res 3
+We now have added a line in the manuscript to describe the parameters involved in the DFE before Equation (4)
+
+\com 4
+In Section A.4 of the appendix, is the partial derivative of $\Delta$ being calculating for the exact expression from Equation 7, or from the approximate expression in Equation A22 obtained from a Taylor expansion when $\rho \ll 1$.
 From the expression in Equation A27, it seems like this is derived from Equation 7, but it would be helpful to clarifying which expression is used to highlight whether this result holds across the range of parameters or only in the limit of weak testing intensity.
+
+\res 4
+We have added a line in the manuscript in Section A.4 of the appendix to clarify the fact that Eq. 27 is a direct result of differentiating $\Delta$ in Eq.7 with respect to $w_{IS}$.
 

### 2022-01-18 manuscript_BMB.tex
reviewer 1 covered
Ali's potential responses

@@
 
-The Disease-Free Equilibrium (DFE) for the expanded SIR model (Eqs.~\ref{model}) is found by setting the infected compartments to 0 and solving for the unknowns. The DFE is
+The Disease-Free Equilibrium (DFE) for the expanded SIR model (Eqs.~\ref{model}) is found by setting the infected compartments to 0 and solving for the unknowns.
+\rev {The DFE depends on \percap testing intensity $\rho$, the rate of test return $\omega$, and the population size $N$}.
+The DFE is
 \begin{equation}
@@
 In this paper, we have developed and analyzed a simple compartmental model that combines epidemiological dynamics---as defined by a simple SIR model---with the dynamics of testing and isolation.
-Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness}).
+Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness}); \rev{the array consisting of multidirectional subtle interplay between epidemiology, behaviors and public health policies}.
 Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
@@
 
+\rev{Using the expression for the effectiveness of control parameter $\Delta$ in eq.~\ref{eq:del4a}, gives}
+
 \begin{align}

### 2022-01-18 response1_BMB.tex
Ali's potential responses

@@
 \usepackage{xcolor}
+\usepackage{xspace}
 
@@
 \newcommand{\nocomment}[3]{}
+
+\newcommand{\Rnum}{\ensuremath{\mathcal{R}_0}\xspace}
+\newcommand*\subtxt[1]{_{\textnormal{#1}}}
+\DeclareRobustCommand\_{\ifmmode\expandafter\subtxt\else\textunderscore\fi}
 
@@
 \begin{document}
-
+\noindent
 Dear Dr. Hastings,
@@
 The reviewer comments are in italics, and our comments/responses are in normal font.
-To address the concerns of the reviewers, we have added appropriate references to the manuscript to acknowledge previous known results in the field.
+To address the concerns of the reviewers, we have edited the manuscript and added the suggested appropriate reference to the manuscript to acknowledge previous known results in the field.
 
@@
 
+\noindent
+\\
 Sincerely,
+
+\noindent
+\\
 Ali Gharouni, PhD
 
-COMMENTS TO THE AUTHOR:
+\section*{REVIEWERS COMMENTS:}
 
-Reviewer \#1: The authors introduce a compartmental epidemic model exploring Susceptible-Infectious-Recovered (SIR) disease dynamics in the presence of a testing protocol that can identify and facilitate isolation of infectious individuals, exploring how factors including the daily intensity of testing, the delay between testing and the reporting of test results, and the degree of self-isolation practiced by individuals waiting for test results and with confirmed positive tests impact the early spread of the disease outbreak.
+\subsection*{ Reviewer \#1:} 
+{\it The authors introduce a compartmental epidemic model exploring Susceptible-Infectious-Recovered (SIR) disease dynamics in the presence of a testing protocol that can identify and facilitate isolation of infectious individuals, exploring how factors including the daily intensity of testing, the delay between testing and the reporting of test results, and the degree of self-isolation practiced by individuals waiting for test results and with confirmed positive tests impact the early spread of the disease outbreak.
 Using the next-generation method, the authors characterize the basic reproduction number (R0) for their model, and employ a mix exact calculations and analytical approximations to study the comparative statics characterizing how this quantity depends on parameters characterizing testing and isolation practices.
@@
 
-In preparation for revising the paper, I have the following minor suggestions that may be helpful for further highlighting results or clarifying the model.
+In preparation for revising the paper, I have the following minor suggestions that may be helpful for further highlighting results or clarifying the model.}
 
@@
 [not sure here!] 
-Regarding the incorporation of an exposed class or asymptomatic and infectious compartment to the current model; the isolation efficacy parameter $\theta_X$ would have similar but stronger effects (result 1).
+Regarding the incorporation of an exposed class or asymptomatic and infectious compartment to the current model; 
+Incorporating the extra compartments (exposed, asymptomatic and infectious) would effect the testing weights and isolation efficacy parameters.
+In particular, we know that people know something about their exposure (assuming we don't know about symptoms).
+We know that there is a bias towards testing infected people, due to having some knowledge of risk.
+Testing people of known exposures will increase testing weight $w_I$, this is testing and tracing.
+Also, exposed people presumably practice more isolation, thus $\theta\_w$ will increase and consequently $\Rnum$ will decrease.
 
-While I think that the analysis already presented in the paper is sufficient for the current paper, I think that ideas related to these questions might be interesting for future research or as points for discussion.
+{\it While I think that the analysis already presented in the paper is sufficient for the current paper, I think that ideas related to these questions might be interesting for future research or as points for discussion.}
 
 \com 3 
-Due to its important effects on R0, I think it could be useful to describe the rate $\omega$ for return of test results directly in the text (in addition to the table of parameters) before it appears in Equation 4 in the expression for the disease-free equilibrium.
+{\it Due to its important effects on R0, I think it could be useful to describe the rate $\omega$ for return of test results directly in the text (in addition to the table of parameters) before it appears in Equation 4 in the expression for the disease-free equilibrium.}
 
@@
 \com 4
-In Section A.4 of the appendix, is the partial derivative of $\Delta$ being calculating for the exact expression from Equation 7, or from the approximate expression in Equation A22 obtained from a Taylor expansion when $\rho \ll 1$.
-From the expression in Equation A27, it seems like this is derived from Equation 7, but it would be helpful to clarifying which expression is used to highlight whether this result holds across the range of parameters or only in the limit of weak testing intensity.
+{\it In Section A.4 of the appendix, is the partial derivative of $\Delta$ being calculating for the exact expression from Equation 7, or from the approximate expression in Equation A22 obtained from a Taylor expansion when $\rho \ll 1$.
+From the expression in Equation A27, it seems like this is derived from Equation 7, but it would be helpful to clarifying which expression is used to highlight whether this result holds across the range of parameters or only in the limit of weak testing intensity.}
 
@@
 
-References
+{\bf References}
+
 [1] K. Muller and P. A. Muller, "Mathematical modelling of the spread of covid-19 on a university campus," Infectious Disease Modelling, vol. 6, pp. 1025-1045, 2021.
 
-Reviewer \#2: The idea underlying the paper is very simple but powerful: public health policies can impact behaviors and thus epidemiology.
+\subsection*{Reviewer \#2:}
+\com 1 
+{\it The idea underlying the paper is very simple but powerful: public health policies can impact behaviors and thus epidemiology.
 Even though the reality is much more complex (infectious spreading results from the complex, non-linear, multi-factorial array and subtle interplay between epidemiology, public health policies and behaviors), authors did a good job in developing their mathematical model, that, rightly, they define as a "caricature".
 They should add it is a "caricature" not only because it (purposely) does not incorporate sophisticated aspects of COVID-19 related epidemiology, but also because does not incorporate that array and interplay I previously mentioned (the array consisting of multidirectional feedbacks between epidemiology, behaviors and public health policies).
-Anyway, authors should add and discuss this point.
-Another point authors should discuss is future prospects, how their model can be further refined and what we can learn from this model in terms of public health policies (discussion is too succinct with this regard).
+Anyway, authors should add and discuss this point.}
+
+\res 1
+We now have added a line in the manuscript (the first paragraph of the Discussion section) to clarify what we mean by the term ``caricature''. 
+[ali: note sure what we should discuss more than what we have in the manuscript!]
+
+\com 2 
+{\it Another point authors should discuss is future prospects, how their model can be further refined and what we can learn from this model in terms of public health policies (discussion is too succinct with this regard).}
+
+\res 2
+
+[This can be added to the manuscript?]
+The potential future work of this project include:
+\begin{enumerate}
+\item
+Incorporating the asymptomatic and symptomatic compartments to the model and verify the present results. 
+\item
+Quantify the testing weights, when we incorporate asymptomatic and symptomatic compartments to the model.
+In particular, We know that not all symptomatic people are the same.
+people know something about their exposure, and not necessarily about symptoms.
+We know that there is a bias towards testing infected people, due to having some knowledge of risk.
+It is harder to quantify that than quantifying symptom/no symptom but we can say that would shift if we have a policy of screening people of high risk areas will push things towards infected people.
+Also testing people of known exposures will increase testing weight for infected people $w_I$, this is exactly testing and tracing.
+It is much harder to quantify that.
+In principle if we knew enough of the case report, hospitalizations, test positivity, death, we could make an estimate of $w_X$'s.
+Also note that $w_X$'s don't stay the same because it is all bound up with capacity because we tend to bias towards testing people who are more likely be infected and the more testing capacity we have the more we are willing to test more people less likely to be infected.
+Thus there is a relationship between testing capacity and the $w_X$'s because our testing policy changes as the testing capacity changes.
+Except for the symptomatic example, it is very hard to quantify the effect of a particular testing policy on the $w_X$'s.
+In short, "test positivity tells a lot about testing weights"; It is hard to know exactly how much information there is and how much we can estimate that.
+\end{enumerate}
 

### 2022-01-18 manuscript_BMB.tex
manuscript back to initial submission, the man comments added to the response letter for now!

--- 2021-07-21 manuscript_BMB.tex
@@
 \newcommand*\subtxt[1]{_{\textnormal{#1}}}
+% \newcommand{\rev}[1]{\comment{red}{Revised}{#1}}
 %% Roman subscript
@@
 In this paper, we have developed and analyzed a simple compartmental model that combines epidemiological dynamics---as defined by a simple SIR model---with the dynamics of testing and isolation.
-Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness}).
-Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
+Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness}); Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
 
@@
 \cite{rogers2021high} simulated a SEIR model with testing and isolation; they similarly suggest a strategy of rapid testing with antigen tests and the subsequent isolation of confirmed-positive individuals.
+% reviewer1.com1
 \cite{friston2021testing} model the effects of self-isolation on testing and tracing with a focus on projections under different testing and tracing scenarios.

### 2022-01-18 response1_BMB.tex
manuscript back to initial submission, the man comments added to the response letter for now!

@@
 \usepackage{xspace}
+\usepackage{enumerate}
+\usepackage{natbib}
 
@@
 
+\newcommand{\percap}{\emph{per capita}\xspace}
 \newcommand{\Rnum}{\ensuremath{\mathcal{R}_0}\xspace}
+\newcommand{\covid}{COVID-19\xspace}
 \newcommand*\subtxt[1]{_{\textnormal{#1}}}
@@
 \newcommand{\needref}{\textcolor{red}{(REFS?)}}
-\newcommand{\com}[1]{\comment{red}{Comment}{#1}}
-\newcommand{\res}[1]{\comment{red}{Response}{#1}}
+\newcommand{\com}[1]{\comment{red}{Comment}{#1}} %reviewer's comment
+\newcommand{\res}[1]{\comment{red}{Response}{#1}} %our response
+\newcommand{\rev}[1]{\comment{red}{In the manuscript}{#1}} %manuscript edition
 
@@
 \com 1 {\it The main surprising results about the possibility of R0 increasing with increased testing intensity or decreasing with increased delays seem particularly intertwined with the behavior displayed by individuals who are waiting for their test results.
-In particular, having long delays can make versions of the model with high $\theta_w$ effectively treat these waiting individuals into a partially socially-distanced or quarantined compartment, while, in versions in which waiting individuals do not change their behavior, increased asymptomatic testing can provide extra time to miss new infectious.
+In particular, having long delays can make versions of the model with high $\theta\_w$ effectively treat these waiting individuals into a partially socially-distanced or quarantined compartment, while, in versions in which waiting individuals do not change their behavior, increased asymptomatic testing can provide extra time to miss new infectious.
 This connection between behavior of waiting and confirmed individuals and how testing impacts R0 makes me curious about a few aspects of combining testing with social distancing.
@@
 \res 1 
-We now have added several lines and the suggested reference [1] to the end of the discussion to acknowledge the "gateway testing" and the potential benefit of our modelling results.
+We now have added several lines and the suggested reference [1] to the end of the discussion to acknowledge the "gateway testing" and the potential benefit of our modelling results in controling the disease spread.
+
+\rev {\cite{muller2021mathematical} simulated a structured SEIR-type model.
+Their model incorporated different testing scenarios -- including gateway testing, surveillance testing, and contact tracing -- as well as individual level control measures such as mask wearing and social distancing.
+They indicate the benefit of gateway testing in conjunction with effective contact tracing or robust surveillance testing in controling the spread of the infection.
+Under such approaches, individuals must practice increased levels of social distancing or self- isolation until they have achieved a series of confirmed negative tests.
+It is notable that while this type of time- dependent testing and self-isolation protocol might not fit well within the framework of linear stability of a disease-free equilibrium, the relationship between increased $\theta_w$ and the benefits of a longer waiting period $1/\omega$ seems to indicate the possibility that combining waiting for results and self-isolation can help to stem early outbreaks (e.g.
+at the beginning of a university term).}
 
@@
 [not sure here!]
-Regarding the incorporation of an exposed class or asymptomatic and infectious compartment to the current model; 
+
+\rev{ Regarding the incorporation of an exposed class or asymptomatic and infectious compartment to the current model; 
 Incorporating the extra compartments (exposed, asymptomatic and infectious) would effect the testing weights and isolation efficacy parameters.
@@
 Also, exposed people presumably practice more isolation, thus $\theta\_w$ will increase and consequently $\Rnum$ will decrease.
+}
 
 {\it While I think that the analysis already presented in the paper is sufficient for the current paper, I think that ideas related to these questions might be interesting for future research or as points for discussion.}
-
 \com 3 
@@
 \res 3
-We now have added a line in the manuscript to describe the parameters involved in the DFE before Equation (4)
+We now have added a line in the manuscript to describe the parameters involved in the DFE before Equation (4).
+
+\rev {The DFE depends on \percap testing intensity $\rho$, the rate of test return $\omega$, and the population size $N$.}
 
@@
 We have added a line in the manuscript in Section A.4 of the appendix to clarify the fact that Eq. 27 is a direct result of differentiating $\Delta$ in Eq.7 with respect to $w_{IS}$.
+
+\rev{Using the expression for the effectiveness of control parameter $\Delta$ in eq.~\ref{eq:del4a}, gives}
 
@@
 We now have added a line in the manuscript (the first paragraph of the Discussion section) to clarify what we mean by the term ``caricature''.
-[ali: note sure what we should discuss more than what we have in the manuscript!]
+
+Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness}); \rev{the array consisting of multidirectional subtle interplay between epidemiology, behaviors and public health policies.} Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
+
+\ali {note sure what we should discuss more than what we have in the manuscript!}
 
@@
 \res 2
+We have added some lines to the discussion section of the manuscript to include some potential future work.
 
-[This can be added to the manuscript?]
-The potential future work of this project include:
+\rev{ The potential future work of this project include:
 \begin{enumerate}
@@
 \end{enumerate}
+}
 

### 2022-01-19 response1_BMB.tex
Fred's comments incorporated

@@
 \rev {\cite{muller2021mathematical} simulated a structured SEIR-type model.
-Their model incorporated different testing scenarios -- including gateway testing, surveillance testing, and contact tracing -- as well as individual level control measures such as mask wearing and social distancing.
+Their model incorporated different testing scenarios -- including gateway testing \fady{what "gateway testing" is?
+ref?}, surveillance testing, and contact tracing -- as well as individual level control measures such as mask wearing and social distancing.
 They indicate the benefit of gateway testing in conjunction with effective contact tracing or robust surveillance testing in controling the spread of the infection.
@@
 }
+\fady{In a less trivial sense, I think that the existence of an exposed class could make any difference in our model only if the testing and isolation parameters are different for the exposed class than the infected class.
+That is, I think that if we added an exposed class in our model but set the testing intensity and isolation strength to be the same as the infected class, then nothing would change in our model.}
 
@@
 \ali {note sure what we should discuss more than what we have in the manuscript!}
+\fady{I think what the reviewer means is that our model ignores the fact that the testing and isolation parameters can change over time due to public health responses.
+In other words, our model makes an assumption of "independence": we assume that testing and isolation affect the epidemic progression, but we do not account for the fact that "in real life", the epidemic progression is also likely to affect the testing and isolation strategies.
+}
 
@@
 In short, "test positivity tells a lot about testing weights"; It is hard to know exactly how much information there is and how much we can estimate that.
+\item
+test
 \end{enumerate}
@@
 
+\fady{
+our model makes the counterintuitive suggestion that increased testing rate could increase $\Rnum$.
+However, note that in our model, we made the assumption that once an individual is in the ``waiting'' compartment, they cannot be tested again until the test result returns.
+Thus, when the test return time is very long, this creates a massive time-delay where an individual may become infected while waiting for the test result, and then the test result comes back as negative, and so they believe they are negative when they are actually positive (and so they do not self-isolate at all, under the belief that they are negative).
+So, in this case, it would be beneficial to actually re-test people even if their first test result has not come back.
+In fact, one could probably conceive of ``testing breadth'' (which roughly corresponds to the number of different people you test over some period of time) vs.
+``testing depth (which corresponds to the number of times you test a single person over a period of time).
+Note that testing breadth gives us granular information ``in space but not in time'' (i.e., it tells us how many people are or are not infected at a single point in time), whereas testing depth gives us granular information ``in time but not in space''.
+The pathological situation in our model can be viewed as an imbalance between testing depth and breadth, where the testing strategy is overly-broad but not sufficiently deep.
+In particular, one could build models that differentiate/partition ``testing intensity'' into ``testing breadth'' and ``testing depth''.
+In such a model, there should always exist some allocation of breadth and depth that makes it so that increasing the overall testing intensity is beneficial (i.e., if we choose the right strategy, more testing should always be good).
+The most important point of all this: while the exact effect we identified in our model (that more testing can increase $\Rnum$ if done incorrectly) is unlikely to happen in real life, a weaker ``sister effect'' probably happens in real life all the time: we may be allocating our testing resources in the wrong breadth vs.
+depth ratio.
+Our model ``proves'' that pure-breadth is not always the optimal strategy (because, in fact, it can increase $\Rnum$).
+So, a more realistic future question is: what is the optimal allocation between breadth and depth that decreases $\Rnum$ most sharply for a given overall testing intensity?
+For example, in our model, when we increase testing intensity, we are mostly increasing breadth.
+So, even on parameter sets in our model where increasing rho decreases $\Rnum$, we may still be ``missing out'' by focusing only on breadth and not depth.
+}
+
 \end{document}

### 2022-02-23 response1_BMB.tex
typesetting tweaks and line numbers in response

@@
-\documentclass{article}
+\documentclass[12pt]{article}
 
@@
 \usepackage{natbib}
+\usepackage[margin=1in]{geometry}
+\usepackage{lineno}\renewcommand\thelinenumber{\color{gray}\arabic{linenumber}}
 
@@
 
+\linenumbers
+
 \section*{REVIEWERS COMMENTS:}
@@
 Some of the main results agree with intuitive expectations about testing and self-isolation, such as the observation that that R0 decreases with a reduction in social contacts in social contacts (i.e.
-increases in $theta_c$ and $theta_w$) for confirmed positives and individuals awaiting test results, as well the observation that focusing testing on infectious individuals helps to decrease R0 relative to a random testing protocol.
+increases in $\theta_c$ and $\theta_w$) for confirmed positives and individuals awaiting test results, as well the observation that focusing testing on infectious individuals helps to decrease R0 relative to a random testing protocol.
 On the other hand, some of the properties of the formula for R0 run counter to expectations, such as the possibility that increased testing intensity can increase R0 in the for the case in which individuals waiting for results do not decrease their social contacts, while R0 can decrease with increased delay times for test results for the case in which waiting individual greatly reduce their social contacts.
@@
 
-- Does the case of R0 decreasing with increased testing delay provide some suggestion of the potential benefit of so-called "gateway testing" [1] in settings such as universities or athletic competitions?
+- Does the case of R0 decreasing with increased testing delay provide some suggestion of the potential benefit of so-called ``gateway testing'' [1] in settings such as universities or athletic competitions?
 Under such approaches, individuals must practice increased levels of social distancing or self- isolation until they have achieved a series of confirmed negative tests.
-While this type of time- dependent testing and self-isolation protocol might not fit well within the framework of linear stability of a disease-free equilibrium, the relationship between increased $theta_w$ and the benefits of a longer waiting period $1/\omega$ seems to indicate the possibility that combining waiting for results and self-isolation can help to stem early outbreaks (e.g.
+While this type of time- dependent testing and self-isolation protocol might not fit well within the framework of linear stability of a disease-free equilibrium, the relationship between increased $\theta_w$ and the benefits of a longer waiting period $1/\omega$ seems to indicate the possibility that combining waiting for results and self-isolation can help to stem early outbreaks (e.g.
 at the beginning of a university term).}
@@
 \res 1 
-We now have added several lines and the suggested reference [1] to the end of the discussion to acknowledge the "gateway testing" and the potential benefit of our modelling results in controling the disease spread.
+We now have added several lines and the suggested reference [1] to the end of the discussion to acknowledge the ``gateway testing'' and the potential benefit of our modelling results in controling the disease spread.
 
 \rev {\cite{muller2021mathematical} simulated a structured SEIR-type model.
-Their model incorporated different testing scenarios -- including gateway testing \fady{what "gateway testing" is?
+Their model incorporated different testing scenarios -- including gateway testing \fady{what ``gateway testing'' is?
 ref?}, surveillance testing, and contact tracing -- as well as individual level control measures such as mask wearing and social distancing.
@@
 
-{\bf References}
+\subsubsection*{References}
 
-[1] K. Muller and P. A. Muller, "Mathematical modelling of the spread of covid-19 on a university campus," Infectious Disease Modelling, vol. 6, pp. 1025-1045, 2021.
+[1] K. Muller and P. A. Muller, ``Mathematical modelling of the spread of covid-19 on a university campus,'' Infectious Disease Modelling, vol. 6, pp. 1025-1045, 2021.
 

### 2022-02-23 response1_BMB.tex
embedded comments in the response document

@@
 
+\david{I gather you have not edited the ms, but are making suggestions for edits.
+It is not clear to me where you are suggesting that the following paragraph goes.
+Please indicate precisely where you are suggesting this new text be included in the ms.
+In any case, you can't include this paragraph in our ms because it is mostly a direct quote from the the referee's comments, which is not OK.
+You do need to define ``gateway testing'' if you are using the term.}
+
 \rev {\cite{muller2021mathematical} simulated a structured SEIR-type model.
@@
 
-Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness}); \rev{the array consisting of multidirectional subtle interplay between epidemiology, behaviors and public health policies.} Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
+Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness}); \david{again, you appear to be suggesting including a direct quote from the referees comments.
+This is not only inappropriate, but the phrase does not make sense to me and }\rev{the array consisting of multidirectional subtle interplay between epidemiology, behaviors and public health policies.} Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
 
@@
 In other words, our model makes an assumption of "independence": we assume that testing and isolation affect the epidemic progression, but we do not account for the fact that "in real life", the epidemic progression is also likely to affect the testing and isolation strategies.
-}
+}\david{I agree with Fred's comment and a brief comment along these lines could be included instead the quote you have above.}
 
@@
 We have added some lines to the discussion section of the manuscript to include some potential future work.
+
+\david{The following suggested text is very rough and incomplete.
+Point 3 just says ``test''.
+You must have forgotten to write what you had in mind.
+Point 2 is not clear or concise, and keeps referring to parameters without explanation.
+When talking about future work, you need to make the text accessible to somebody who hasn't studied the paper carefully.
+What directions would be useful to pursue?
+And why?
+i.e., what would you expect to learn from such work?
+Trying to estimate testing weights might be interesting and useful, but I think the referee was probably trying to get us to describe further developments of the this modelling approach, not that we should try to estimate parameters.}
 
@@
 }
+\david{These are thoughtful ideas from Fred.  Perhaps they could be compressed to a couple of sentences describing a future direction worth pursuing.}
 

### 2022-03-01 manuscript_BMB.tex
Fred's comments incorporated
typo correction and latex tweak in eq (4)
working on the comments

@@
 \newcommand*\subtxt[1]{_{\textnormal{#1}}}
-% \newcommand{\rev}[1]{\comment{red}{Revised}{#1}}
+\newcommand{\rev}[1]{\comment{red}{Revised}{#1}}
 %% Roman subscript
@@
 
+\linenumbers
+
 \title{Testing and Isolation Efficacy:\\Insights from a Simple Epidemic Model
@@
 \author{Ali Gharouni         \and
-        F.M. Abdelmalek \and 
+        Fred M. Abdelmalek \and 
         David J. D. Earn \and
@@
 \maketitle
-
+\linenumbers
 \begin{abstract}
@@
 
-The Disease-Free Equilibrium (DFE) for the expanded SIR model (Eqs.~\ref{model}) is found by setting the infected compartments to 0 and solving for the unknowns. The DFE is 
+The Disease-Free Equilibrium (DFE) for the expanded SIR model (Eqs.~\ref{model}) is found by setting the infected compartments to 0 and solving for the unknowns. 
+% reviewer1:com3 \rev{The DFE depends on \percap testing intensity $\rho$, the rate of test return $\omega$, and the population size $N$.} 
+The DFE is 
 \begin{equation}
 \label{dfe}
-S\_n^*= \frac{\rho}{\omega} N, \ S\_u^*= N-S_n^*, \text{~and~} I_j=R_j=0 \ \text{for all }j.
+S\_n^*= \frac{\rho}{\omega} N, \quad S\_u^*= N-S\_n^*, \quad\text{and}\quad I_j=R_j=0 \ \text{for all }j.
 \end{equation}
@@
 In this paper, we have developed and analyzed a simple compartmental model that combines epidemiological dynamics---as defined by a simple SIR model---with the dynamics of testing and isolation.
-Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness}); Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
+Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness});
+% reviewer2:com1
+Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
 
@@
 \cite{rogers2021high} simulated a SEIR model with testing and isolation; they similarly suggest a strategy of rapid testing with antigen tests and the subsequent isolation of confirmed-positive individuals.
-% reviewer1.com1
+% reviewer1:com1 \rev {\cite{muller2021mathematical} simulated a structured SEIR-type model.
+Their model incorporated different testing scenarios -- including gateway testing \fady{what ``gateway testing'' is?
+ref?}, surveillance testing, and contact tracing -- as well as individual level control measures such as mask wearing and social distancing.
+They indicate the benefit of gateway testing in conjunction with effective contact tracing or robust surveillance testing in controling the spread of the infection.
+Under such approaches, individuals must practice increased levels of social distancing or self- isolation until they have achieved a series of confirmed negative tests.
+It is notable that while this type of time- dependent testing and self-isolation protocol might not fit well within the framework of linear stability of a disease-free equilibrium, the relationship between increased $\theta_w$ and the benefits of a longer waiting period $1/\omega$ seems to indicate the possibility that combining waiting for results and self-isolation can help to stem early outbreaks (e.g.
+at the beginning of a university term).}
 \cite{friston2021testing} model the effects of self-isolation on testing and tracing with a focus on projections under different testing and tracing scenarios.
@@
 We hope this paper will inspire further explorations of the fundamental properties of epidemic models that incorporate explicit testing processes.
+% reviewer2:com2 future work...
 
@@
 
+% reviewer1:com4 \rev{Using the expression for the effectiveness of control parameter $\Delta$ in eq.~\ref{eq:del4a}, gives}
 \begin{align}

### 2022-03-01 response1_BMB.tex
working on the comments

@@
 
+% help
 \com 1 {\it The main surprising results about the possibility of R0 increasing with increased testing intensity or decreasing with increased delays seem particularly intertwined with the behavior displayed by individuals who are waiting for their test results.
@@
 
-\rev {\cite{muller2021mathematical} simulated a structured SEIR-type model.
+\rev {This goes in line 263 of pdf.
+\cite{muller2021mathematical} simulated a structured SEIR-type model.
 Their model incorporated different testing scenarios -- including gateway testing \fady{what ``gateway testing'' is?
@@
 
-\rev{ Regarding the incorporation of an exposed class or asymptomatic and infectious compartment to the current model; 
+% help
+\rev{ (this belongs to discussion) Regarding the incorporation of an exposed class or asymptomatic and infectious compartment to the current model; 
 Incorporating the extra compartments (exposed, asymptomatic and infectious) would effect the testing weights and isolation efficacy parameters.
@@
 
-Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness}); \david{again, you appear to be suggesting including a direct quote from the referees comments.
-This is not only inappropriate, but the phrase does not make sense to me and }\rev{the array consisting of multidirectional subtle interplay between epidemiology, behaviors and public health policies.} Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
+Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness});
 
-\ali {note sure what we should discuss more than what we have in the manuscript!}
-\fady{I think what the reviewer means is that our model ignores the fact that the testing and isolation parameters can change over time due to public health responses.
-In other words, our model makes an assumption of "independence": we assume that testing and isolation affect the epidemic progression, but we do not account for the fact that "in real life", the epidemic progression is also likely to affect the testing and isolation strategies.
-}\david{I agree with Fred's comment and a brief comment along these lines could be included instead the quote you have above.}
+\rev{Our model ignores the fact that the testing and isolation parameters can change over time due to public health responses.
+In other words, our model makes an assumption of "independence": we assume that testing and isolation affect the epidemic progression, but we do not account for the fact that "in real life", the epidemic progression is also likely to affect the testing and isolation strategies.}
+\david{I agree with Fred's comment and a brief comment along these lines could be included instead the quote you have above.}
 
@@
 In short, "test positivity tells a lot about testing weights"; It is hard to know exactly how much information there is and how much we can estimate that.
-\item
-test
 \end{enumerate}

### 2022-03-01 codes/sir_plot.R
Meeting with Ali

@@ -1,10 +1,10 @@
+library(McMasterPandemic)
 library(ggplot2)
 library(dplyr)
 library(tidyr)
 library(directlabels)
 library(latex2exp)
 library(rootSolve)
-library(McMasterPandemic)
 library(cowplot) ## for ggdraw()/rectangle-hacking
 
 library(shellpipes)

### 2022-03-06 manuscript_BMB.tex
correcting Fred's name
Meeting with Ali
inserted our responses in manuscript, response file cleaned up

--- 2022-01-18 manuscript_BMB.tex
@@
 
+\newcommand{\fady}[1]{\comment{cyan}{Fady}{#1}}
+\newcommand{\ali}[1]{\comment{magenta}{Ali}{#1}}
+\newcommand{\jd}[1]{\comment{blue}{JD}{#1}}
+\newcommand{\david}[1]{\comment{orange}{DJDE}{#1}}
+\newcommand{\bmb}[1]{\comment{red}{BMB}{#1}}
+\newcommand{\todo}[1]{\comment{red}{TODO}{#1}}
+\newcommand{\new}[1]{\comment{red}{NEW}{#1}}
+\newcommand{\needref}{\textcolor{red}{(REFS?)}}
+
 \newcommand{\fref}[1]{Fig.~\ref{#1}}
@@
            \and
-           F.M. Abdelmalek \at
+           Fred .M. Abdelmalek \at
               Department of Mathematics and Statistics, McMaster University, Hamilton, Canada \\
@@
 
-The Disease-Free Equilibrium (DFE) for the expanded SIR model (Eqs.~\ref{model}) is found by setting the infected compartments to 0 and solving for the unknowns. The DFE is 
+The Disease-Free Equilibrium (DFE) for the expanded SIR model (Eqs.~\ref{model}) is found by setting the infected compartments to 0 and solving for the unknowns. 
+% reviewer1:com3 
+\rev{The DFE depends on \percap testing intensity $\rho$, the rate of test return $\omega$, and the population size $N$.} 
+The DFE is 
 \begin{equation}
 \label{dfe}
-S\_n^*= \frac{\rho}{\omega} N, \ S\_u^*= N-S_n^*, \text{~and~} I_j=R_j=0 \ \text{for all }j.
+S\_n^*= \frac{\rho}{\omega} N, \quad S\_u^*= N-S\_n^*, \quad\text{and}\quad I_j=R_j=0 \ \text{for all }j.
 \end{equation}
@@
 In this paper, we have developed and analyzed a simple compartmental model that combines epidemiological dynamics---as defined by a simple SIR model---with the dynamics of testing and isolation.
-Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness}); Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
+Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness});
+% reviewer2:com1
+\rev{Further, our model ignores the fact that the testing and isolation parameters can change over time due to public health responses.
+In other words, our model makes an assumption of ``independence'': we assume that testing and isolation affect the epidemic progression, but we do not account for the fact that "in real life", the epidemic progression is also likely to affect the testing and isolation strategies.}
+Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
 
@@
 
+% TODO reviewer1:com2
+\rev{Regarding the incorporation of an exposed class or asymptomatic and infectious compartment to the current model; Incorporating the extra compartments (exposed, asymptomatic and infectious) would effect the testing weights and isolation efficacy parameters.
+In particular, we know that people know something about their exposure (assuming we don't know about symptoms).
+We know that there is a bias towards testing infected people, due to having some knowledge of risk.
+Testing people of known exposures will increase testing weight $w_I$, this is testing and tracing.
+Also, exposed people presumably practice more isolation, thus $\theta\_w$ will increase and consequently $\Rnum$ will decrease.} \fady{In a less trivial sense, the existence of an exposed class could make any difference in our model only if the testing and isolation parameters are different for the exposed class than the infected class.
+That is, I think that if we added an exposed class in our model but set the testing intensity and isolation strength to be the same as the infected class, then nothing would change in our model.}
+
 Although testing and tracing is a key part of infection control strategies, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts \citep{endo2020implication,hellewell2020feasibility,jenness2020modeling}, rather than analyzing simple but general models of the feedback between testing and transmission dynamics.
@@
 \cite{rogers2021high} simulated a SEIR model with testing and isolation; they similarly suggest a strategy of rapid testing with antigen tests and the subsequent isolation of confirmed-positive individuals.
-% reviewer1.com1
+% TODO reviewer1:com1
+\rev {\cite{muller2021mathematical} simulated a structured SEIR-type model.
+Their model incorporated different testing scenarios -- including gateway testing (what ``gateway testing'' is?
+ref?), surveillance testing, and contact tracing -- as well as individual level control measures such as mask wearing and social distancing.
+They indicate the benefit of gateway testing in conjunction with effective contact tracing or robust surveillance testing in controling the spread of the infection.
+Under such approaches, individuals must practice increased levels of social distancing or self- isolation until they have achieved a series of confirmed negative tests.
+It is notable that while this type of time- dependent testing and self-isolation protocol might not fit well within the framework of linear stability of a disease-free equilibrium, the relationship between increased $\theta_w$ and the benefits of a longer waiting period $1/\omega$ seems to indicate the possibility that combining waiting for results and self-isolation can help to stem early outbreaks (e.g.
+at the beginning of a university term).}
 \cite{friston2021testing} model the effects of self-isolation on testing and tracing with a focus on projections under different testing and tracing scenarios.
@@
 We hope this paper will inspire further explorations of the fundamental properties of epidemic models that incorporate explicit testing processes.
+
+% TODO reviewer2:com2 
+\rev{ The potential future work of this project include:
+(i) Incorporating the asymptomatic and symptomatic compartments to the model and verify the present results. 
+(ii) Quantify the testing weights, when we incorporate asymptomatic and symptomatic compartments to the model.
+In particular, We know that not all symptomatic people are the same.
+people know something about their exposure, and not necessarily about symptoms.
+We know that there is a bias towards testing infected people, due to having some knowledge of risk.
+It is harder to quantify that than quantifying symptom/no symptom but we can say that would shift if we have a policy of screening people of high risk areas will push things towards infected people.
+Also testing people of known exposures will increase testing weight for infected people $w_I$, this is exactly testing and tracing.
+It is much harder to quantify that.
+In principle if we knew enough of the case report, hospitalizations, test positivity, death, we could make an estimate of $w_X$'s.
+Also note that $w_X$'s don't stay the same because it is all bound up with capacity because we tend to bias towards testing people who are more likely be infected and the more testing capacity we have the more we are willing to test more people less likely to be infected.
+Thus there is a relationship between testing capacity and the $w_X$'s because our testing policy changes as the testing capacity changes.
+Except for the symptomatic example, it is very hard to quantify the effect of a particular testing policy on the $w_X$'s.
+In short, "test positivity tells a lot about testing weights"; It is hard to know exactly how much information there is and how much we can estimate that.}
+\fady{
+our model makes the counterintuitive suggestion that increased testing rate could increase $\Rnum$.
+However, note that in our model, we made the assumption that once an individual is in the ``waiting'' compartment, they cannot be tested again until the test result returns.
+Thus, when the test return time is very long, this creates a massive time-delay where an individual may become infected while waiting for the test result, and then the test result comes back as negative, and so they believe they are negative when they are actually positive (and so they do not self-isolate at all, under the belief that they are negative).
+So, in this case, it would be beneficial to actually re-test people even if their first test result has not come back.
+In fact, one could probably conceive of ``testing breadth'' (which roughly corresponds to the number of different people you test over some period of time) vs.
+``testing depth (which corresponds to the number of times you test a single person over a period of time).
+Note that testing breadth gives us granular information ``in space but not in time'' (i.e., it tells us how many people are or are not infected at a single point in time), whereas testing depth gives us granular information ``in time but not in space''.
+The pathological situation in our model can be viewed as an imbalance between testing depth and breadth, where the testing strategy is overly-broad but not sufficiently deep.
+In particular, one could build models that differentiate/partition ``testing intensity'' into ``testing breadth'' and ``testing depth''.
+In such a model, there should always exist some allocation of breadth and depth that makes it so that increasing the overall testing intensity is beneficial (i.e., if we choose the right strategy, more testing should always be good).
+The most important point of all this: while the exact effect we identified in our model (that more testing can increase $\Rnum$ if done incorrectly) is unlikely to happen in real life, a weaker ``sister effect'' probably happens in real life all the time: we may be allocating our testing resources in the wrong breadth vs.
+depth ratio.
+Our model ``proves'' that pure-breadth is not always the optimal strategy (because, in fact, it can increase $\Rnum$).
+So, a more realistic future question is: what is the optimal allocation between breadth and depth that decreases $\Rnum$ most sharply for a given overall testing intensity?
+For example, in our model, when we increase testing intensity, we are mostly increasing breadth.
+So, even on parameter sets in our model where increasing rho decreases $\Rnum$, we may still be ``missing out'' by focusing only on breadth and not depth.
+}
+\david{These are thoughtful ideas from Fred.  Perhaps they could be compressed to a couple of sentences describing a future direction worth pursuing.}
 
@@
 
+% reviewer1:com4 
+\rev{Using the expression for the effectiveness of control parameter $\Delta$ in eq.~\ref{eq:del4a}, gives}
 \begin{align}

### 2022-03-06 response1_BMB.tex
Meeting with Ali
inserted our responses in manuscript, response file cleaned up

--- 2022-02-23 response1_BMB.tex
@@
 
+% JD try to write text
 \com 1 {\it The main surprising results about the possibility of R0 increasing with increased testing intensity or decreasing with increased delays seem particularly intertwined with the behavior displayed by individuals who are waiting for their test results.
@@
 We now have added several lines and the suggested reference [1] to the end of the discussion to acknowledge the ``gateway testing'' and the potential benefit of our modelling results in controling the disease spread.
-
-\rev {\cite{muller2021mathematical} simulated a structured SEIR-type model.
-Their model incorporated different testing scenarios -- including gateway testing \fady{what ``gateway testing'' is?
-ref?}, surveillance testing, and contact tracing -- as well as individual level control measures such as mask wearing and social distancing.
-They indicate the benefit of gateway testing in conjunction with effective contact tracing or robust surveillance testing in controling the spread of the infection.
-Under such approaches, individuals must practice increased levels of social distancing or self- isolation until they have achieved a series of confirmed negative tests.
-It is notable that while this type of time- dependent testing and self-isolation protocol might not fit well within the framework of linear stability of a disease-free equilibrium, the relationship between increased $\theta_w$ and the benefits of a longer waiting period $1/\omega$ seems to indicate the possibility that combining waiting for results and self-isolation can help to stem early outbreaks (e.g.
-at the beginning of a university term).}
+% {search for reviewer1:com1 in the manuscript_BMB.tex}
 
@@
 
-\rev{ Regarding the incorporation of an exposed class or asymptomatic and infectious compartment to the current model; 
-Incorporating the extra compartments (exposed, asymptomatic and infectious) would effect the testing weights and isolation efficacy parameters.
-In particular, we know that people know something about their exposure (assuming we don't know about symptoms).
-We know that there is a bias towards testing infected people, due to having some knowledge of risk.
-Testing people of known exposures will increase testing weight $w_I$, this is testing and tracing.
-Also, exposed people presumably practice more isolation, thus $\theta\_w$ will increase and consequently $\Rnum$ will decrease.
-}
-\fady{In a less trivial sense, I think that the existence of an exposed class could make any difference in our model only if the testing and isolation parameters are different for the exposed class than the infected class.
-That is, I think that if we added an exposed class in our model but set the testing intensity and isolation strength to be the same as the infected class, then nothing would change in our model.}
+% JD try to write text
+% {search for reviewer1:com2 in the manuscript_BMB.tex}
 
@@
 We now have added a line in the manuscript to describe the parameters involved in the DFE before Equation (4).
-
-\rev {The DFE depends on \percap testing intensity $\rho$, the rate of test return $\omega$, and the population size $N$.}
+% {search for reviewer1:com3 in the manuscript_BMB.tex}
 
@@
 We have added a line in the manuscript in Section A.4 of the appendix to clarify the fact that Eq. 27 is a direct result of differentiating $\Delta$ in Eq.7 with respect to $w_{IS}$.
-
-\rev{Using the expression for the effectiveness of control parameter $\Delta$ in eq.~\ref{eq:del4a}, gives}
+% {search for reviewer1:com4 in the manuscript_BMB.tex}
 
@@
 [1] K. Muller and P. A. Muller, ``Mathematical modelling of the spread of covid-19 on a university campus,'' Infectious Disease Modelling, vol. 6, pp. 1025-1045, 2021.
+
+%% JD to look at all of Rev 2
 
@@
 We now have added a line in the manuscript (the first paragraph of the Discussion section) to clarify what we mean by the term ``caricature''. 
-
-Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness}); \rev{the array consisting of multidirectional subtle interplay between epidemiology, behaviors and public health policies.} Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
-
-\ali {note sure what we should discuss more than what we have in the manuscript!}
-\fady{I think what the reviewer means is that our model ignores the fact that the testing and isolation parameters can change over time due to public health responses.
-In other words, our model makes an assumption of "independence": we assume that testing and isolation affect the epidemic progression, but we do not account for the fact that "in real life", the epidemic progression is also likely to affect the testing and isolation strategies.
-}
+% {search for reviewer2:com2 in the manuscript_BMB.tex}
 
@@
 We have added some lines to the discussion section of the manuscript to include some potential future work.
-
-\rev{ The potential future work of this project include:
-\begin{enumerate}
-\item
-Incorporating the asymptomatic and symptomatic compartments to the model and verify the present results. 
-\item
-Quantify the testing weights, when we incorporate asymptomatic and symptomatic compartments to the model.
-In particular, We know that not all symptomatic people are the same.
-people know something about their exposure, and not necessarily about symptoms.
-We know that there is a bias towards testing infected people, due to having some knowledge of risk.
-It is harder to quantify that than quantifying symptom/no symptom but we can say that would shift if we have a policy of screening people of high risk areas will push things towards infected people.
-Also testing people of known exposures will increase testing weight for infected people $w_I$, this is exactly testing and tracing.
-It is much harder to quantify that.
-In principle if we knew enough of the case report, hospitalizations, test positivity, death, we could make an estimate of $w_X$'s.
-Also note that $w_X$'s don't stay the same because it is all bound up with capacity because we tend to bias towards testing people who are more likely be infected and the more testing capacity we have the more we are willing to test more people less likely to be infected.
-Thus there is a relationship between testing capacity and the $w_X$'s because our testing policy changes as the testing capacity changes.
-Except for the symptomatic example, it is very hard to quantify the effect of a particular testing policy on the $w_X$'s.
-In short, "test positivity tells a lot about testing weights"; It is hard to know exactly how much information there is and how much we can estimate that.
-\item
-test
-\end{enumerate}
-}
-
-\fady{
-our model makes the counterintuitive suggestion that increased testing rate could increase $\Rnum$.
-However, note that in our model, we made the assumption that once an individual is in the ``waiting'' compartment, they cannot be tested again until the test result returns.
-Thus, when the test return time is very long, this creates a massive time-delay where an individual may become infected while waiting for the test result, and then the test result comes back as negative, and so they believe they are negative when they are actually positive (and so they do not self-isolate at all, under the belief that they are negative).
-So, in this case, it would be beneficial to actually re-test people even if their first test result has not come back.
-In fact, one could probably conceive of ``testing breadth'' (which roughly corresponds to the number of different people you test over some period of time) vs.
-``testing depth (which corresponds to the number of times you test a single person over a period of time).
-Note that testing breadth gives us granular information ``in space but not in time'' (i.e., it tells us how many people are or are not infected at a single point in time), whereas testing depth gives us granular information ``in time but not in space''.
-The pathological situation in our model can be viewed as an imbalance between testing depth and breadth, where the testing strategy is overly-broad but not sufficiently deep.
-In particular, one could build models that differentiate/partition ``testing intensity'' into ``testing breadth'' and ``testing depth''.
-In such a model, there should always exist some allocation of breadth and depth that makes it so that increasing the overall testing intensity is beneficial (i.e., if we choose the right strategy, more testing should always be good).
-The most important point of all this: while the exact effect we identified in our model (that more testing can increase $\Rnum$ if done incorrectly) is unlikely to happen in real life, a weaker ``sister effect'' probably happens in real life all the time: we may be allocating our testing resources in the wrong breadth vs.
-depth ratio.
-Our model ``proves'' that pure-breadth is not always the optimal strategy (because, in fact, it can increase $\Rnum$).
-So, a more realistic future question is: what is the optimal allocation between breadth and depth that decreases $\Rnum$ most sharply for a given overall testing intensity?
-For example, in our model, when we increase testing intensity, we are mostly increasing breadth.
-So, even on parameter sets in our model where increasing rho decreases $\Rnum$, we may still be ``missing out'' by focusing only on breadth and not depth.
-}
+% { search for reviewer2:com2 in the manuscript_BMB.tex}
 

### 2022-03-06 manuscript_BMB.tex
Responding to reviewers and incorporating comments

--- 2022-01-18 manuscript_BMB.tex
@@
 
+\newcommand{\fred}[1]{\comment{cyan}{Fred}{#1}}
+\newcommand{\ali}[1]{\comment{magenta}{Ali}{#1}}
+\newcommand{\jd}[1]{\comment{blue}{JD}{#1}}
+\newcommand{\david}[1]{\comment{orange}{DJDE}{#1}}
+\newcommand{\bmb}[1]{\comment{red}{BMB}{#1}}
+\newcommand{\todo}[1]{\comment{red}{TODO}{#1}}
+\newcommand{\new}[1]{\comment{red}{NEW}{#1}}
+\newcommand{\needref}{\textcolor{red}{(REFS?)}}
+
 \newcommand{\fref}[1]{Fig.~\ref{#1}}
@@
 \newcommand*\subtxt[1]{_{\textnormal{#1}}}
-% \newcommand{\rev}[1]{\comment{red}{Revised}{#1}}
+\newcommand{\rev}[1]{\comment{red}{Revised}{#1}}
+\renewcommand{\rev}[1]{{#1}}
 %% Roman subscript
@@
 In this paper, we have developed and analyzed a simple compartmental model that combines epidemiological dynamics---as defined by a simple SIR model---with the dynamics of testing and isolation.
-Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness}); Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
+Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness});
+% reviewer2:com1
+\rev{Further, our model ignores the fact that the testing and isolation parameters can change over time due to public health responses.
+In other words, our model makes an assumption of ``independence'': we assume that testing and isolation affect the epidemic progression, but we do not account for the fact that epidemic progression is also likely to affect testing and isolation strategies.}
+Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
 
@@
 
+This finding does, however, suggest that cautious behaviour by people waiting for test results might have appreciable effects on epidemic spread.
+This reinforces the idea that people waiting for test results should be urged to isolate and take other measures to prevent infecting other, or being infected.
+It also provides general support for the idea of institutional ``gateway'' testing which restricts individuals' contacts or access until one or more negative tests have been received \cite{Muller}.
+
 The other counterintuitive result from our analysis is that, for sufficiently high testing intensity $\rho$, further increasing testing intensity can actually \emph{increase} $\Rnum$ (e.g., \fref{pan2}(b), upper right panel [$\theta\_c=1$, $\theta\_w=0$]).
@@
 In particular, it depends on levels of testing that are unrealistically high (at least in large, general-population settings).
+
+This theoretical phenomenon also highlights a practical point: keeping people waiting for tests that will come back negative can increase spread.
+The best solution is to increase test-processing speed, but it would also be worth exploring the option of encouraging people to test more than once, in case an early test was taken before an individual had enough virus to register as positive.
+Future work could profitably explore the question of when it is beneficial to encourage people to test more than once.
 
@@
 \cite{rogers2021high} simulated a SEIR model with testing and isolation; they similarly suggest a strategy of rapid testing with antigen tests and the subsequent isolation of confirmed-positive individuals.
-% reviewer1.com1
 \cite{friston2021testing} model the effects of self-isolation on testing and tracing with a focus on projections under different testing and tracing scenarios.
-They conclude that the emergence of a second wave depends almost primarily on the rate at which immunity is lost and that it is necessary to track asymptomatic individuals in order to control the outbreak.
+They conclude that the emergence of a second wave depends primarily on the rate at which immunity is lost and that it is necessary to track asymptomatic individuals in order to control the outbreak.
 Our modeling approach differs from these previous efforts in that it examines the effects of test-return rates and of different levels of testing focus, from random to highly targeted.
 We hope this paper will inspire further explorations of the fundamental properties of epidemic models that incorporate explicit testing processes.
+
+%% \fred{ our model makes the counterintuitive suggestion that increased testing rate could increase $\Rnum$.
+However, note that in our model, we made the assumption that once an individual is in the ``waiting'' compartment, they cannot be tested again until the test result returns.
+Thus, when the test return time is very long, this creates a massive time-delay where an individual may become infected while waiting for the test result, and then the test result comes back as negative, and so they believe they are negative when they are actually positive (and so they do not self-isolate at all, under the belief that they are negative).
+%% So, in this case, it would be beneficial to actually re-test people even if their first test result has not come back.
+In fact, one could probably conceive of ``testing breadth'' (which roughly corresponds to the number of different people you test over some period of time) vs.
+``testing depth (which corresponds to the number of times you test a single person over a period of time).
+Note that testing breadth gives us granular information ``in space but not in time'' (i.e., it tells us how many people are or are not infected at a single point in time), whereas testing depth gives us granular information ``in time but not in space''.
+The pathological situation in our model can be viewed as an imbalance between testing depth and breadth, where the testing strategy is overly-broad but not sufficiently deep.
+%% In particular, one could build models that differentiate/partition ``testing intensity'' into ``testing breadth'' and ``testing depth''.
+In such a model, there should always exist some allocation of breadth and depth that makes it so that increasing the overall testing intensity is beneficial (i.e., if we choose the right strategy, more testing should always be good).
+%% The most important point of all this: while the exact effect we identified in our model (that more testing can increase $\Rnum$ if done incorrectly) is unlikely to happen in real life, a weaker ``sister effect'' probably happens in real life all the time: we may be allocating our testing resources in the wrong breadth vs.
+depth ratio.
+Our model ``proves'' that pure-breadth is not always the optimal strategy (because, in fact, it can increase $\Rnum$).
+So, a more realistic future question is: what is the optimal allocation between breadth and depth that decreases $\Rnum$ most sharply for a given overall testing intensity?
+For example, in our model, when we increase testing intensity, we are mostly increasing breadth.
+So, even on parameter sets in our model where increasing rho decreases $\Rnum$, we may still be ``missing out'' by focusing only on breadth and not depth.
+}
+%% \david{These are thoughtful ideas from Fred.  Perhaps they could be compressed to a couple of sentences describing a future direction worth pursuing.}
 

### 2022-03-06 response1_BMB.tex
Responding to reviewers and incorporating comments

@@
 \res 1 
-We now have added several lines and the suggested reference [1] to the end of the discussion to acknowledge the ``gateway testing'' and the potential benefit of our modelling results in controling the disease spread.
-% {search for reviewer1:com1 in the manuscript_BMB.tex}
+% {see "Muller" in manuscript_BMB.tex}
 
@@
 
-[1] K. Muller and P. A. Muller, ``Mathematical modelling of the spread of covid-19 on a university campus,'' Infectious Disease Modelling, vol. 6, pp. 1025-1045, 2021.
+% [1] K. Muller and P. A. Muller, ``Mathematical modelling of the spread of covid-19 on a university campus,'' Infectious Disease Modelling, vol. 6, pp. 1025-1045, 2021.
+% ~/Downloads/MullMull.pdf
+% https://pubmed.ncbi.nlm.nih.gov/34414342/
 

### 2022-03-07 manuscript_BMB.tex
added Muler to refs, marked %reviewer:com in the mans

@@
 \newcommand{\rev}[1]{\comment{red}{Revised}{#1}}
-\renewcommand{\rev}[1]{{#1}}
+\renewcommand{\rev}[1]{#1}
 %% Roman subscript
@@
 Our model development was motivated by the \covid epidemic, but has generic epidemiological and testing structures. 
-To the classic SIR framework we have added a \percap testing intensity, and compartment-specific testing weights, which can be adjusted to reflect different testing emphases --- surveillance, diagnosis, or control.
+To the classic SIR framework we have added a \percap\ testing intensity, and compartment-specific testing weights, which can be adjusted to reflect different testing emphases --- surveillance, diagnosis, or control.
 We derive an analytic expression for the relative reduction in the basic reproductive number due to testing, test-reporting and related isolation behaviours.
@@
 % reviewer1:com3 
-\rev{The DFE depends on \percap testing intensity $\rho$, the rate of test return $\omega$, and the population size $N$.} 
+The DFE depends on \percap testing intensity $\rho$, the rate of test return $\omega$, and the population size $N$.
 The DFE is 
@@
 % reviewer2:com1
-\rev{Further, our model ignores the fact that the testing and isolation parameters can change over time due to public health responses.
-In other words, our model makes an assumption of ``independence'': we assume that testing and isolation affect the epidemic progression, but we do not account for the fact that epidemic progression is also likely to affect testing and isolation strategies.}
+Further, our model ignores the fact that the testing and isolation parameters can change over time due to public health responses.
+In other words, our model makes an assumption of ``independence'': we assume that testing and isolation affect the epidemic progression, but we do not account for the fact that epidemic progression is also likely to affect testing and isolation strategies.%
 Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
@@
 
+% reviewer1:com1
 This finding does, however, suggest that cautious behaviour by people waiting for test results might have appreciable effects on epidemic spread.
 This reinforces the idea that people waiting for test results should be urged to isolate and take other measures to prevent infecting other, or being infected.
-It also provides general support for the idea of institutional ``gateway'' testing which restricts individuals' contacts or access until one or more negative tests have been received \cite{Muller}.
+It also provides general support for the idea of institutional ``gateway'' testing which restricts individuals' contacts or access until one or more negative tests have been received \citep{muller2021mathematical}.%
 
@@
 This correlation is influenced, among many other factors, by the proportion of the uninfected population with COVID-like symptoms (e.g., due to seasonal upper respiratory tract infections); the concentration of transmission and testing in hot spots such as long-term care facilities and high-density workplaces; the overall testing intensity (and hence, e.g., restriction to symptomatic individuals); and the proportion of COVID-infected people who are symptomatic.
+
+% reviewer1:com2
 Future steps should explore mathematically tractable ways to model some of these factors more precisely.
 For example, separating the infected class into exposed, symptomatic, and a- or pre-symptomatic compartments and allowing the testing weights to vary across non-symptomatic (exposed/asymptomatic/presymptomatic) vs.\ symptomatic compartments could reflect the allocation of tests for diagnostic purposes (targeting symptomatic individuals) vs.\ contact-tracing (targeting infected but non-symptomatic individuals) vs.\ screening (relatively equal weights, depending on the venue).
+This could also allow exploration of different assumptions about individuals' likelihood of testing positive in different compartments, with implications for the questions about repeat testing discussed above.
 Alternatively, one could make the testing weights depend on the testing intensity or test-return rate as suggested above.
@@
 
+% reviewer2:com2
 %% \fred{ our model makes the counterintuitive suggestion that increased testing rate could increase $\Rnum$.
@@
 }
-%% \david{These are thoughtful ideas from Fred.  Perhaps they could be compressed to a couple of sentences describing a future direction worth pursuing.}
 
@@
 % reviewer1:com4 
-\rev{Using the expression for the effectiveness of control parameter $\Delta$ in eq.~\ref{eq:del4a}, gives}
+Using the expression for the effectiveness of control parameter $\Delta$ in eq.~\ref{eq:del4a}, gives
 \begin{align}

### 2022-03-08 manuscript_BMB.tex
updating refs

@@
 Over the course of the \covid pandemic, however, the vast majority of testing has been done with other goals --
-primarily diagnostic (determining infection status for clinical purposes) \citep{phua2020intensive,who2020global}, or for control (determining  infection status in order to isolate cases that have been found by contact tracing) \citep{aleta2020modelling,kucharski2020effectiveness,grassly2020comparison,smith2020adherence}, which we characterize as \emph{targeted} testing strategies.
+primarily diagnostic (determining infection status for clinical purposes) \citep{phua2020intensive,who2020global}, or for control (determining  infection status in order to isolate cases that have been found by contact tracing) \citep{aleta2020modelling,kucharski2020effectiveness,grassly2020comparison,smith2021adherence}, which we characterize as \emph{targeted} testing strategies.
 In these situations, testing probabilities can differ sharply across epidemiological compartments; in our dynamical model, we will characterize these probabilities by assigning a testing weight to each compartment that determines the \emph{relative} probability that an individual in that compartment will be selected for testing (see Methods).
@@
 
-Although testing and tracing is a key part of infection control strategies, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts \citep{endo2020implication,hellewell2020feasibility,jenness2020modeling}, rather than analyzing simple but general models of the feedback between testing and transmission dynamics.
+Although testing and tracing is a key part of infection control strategies, mathematical epidemiologists have typically analyzed it with detailed models designed to inform particular public health efforts \citep{endo2020implication,hellewell2020feasibility,jenness2021modeling}, rather than analyzing simple but general models of the feedback between testing and transmission dynamics.
 There have been several modeling studies of testing and tracing dynamics and their interaction with epidemiological dynamics.

### 2022-03-13 manuscript_BMB.tex
Bolker comments and tweaks

@@
            \and
-           Fred .M. Abdelmalek \at
+           Fred M. Abdelmalek \at
               Department of Mathematics and Statistics, McMaster University, Hamilton, Canada \\
@@
 In this paper, we have developed and analyzed a simple compartmental model that combines epidemiological dynamics---as defined by a simple SIR model---with the dynamics of testing and isolation.
-Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; behavioural dynamics \citep{weitz2020awareness});
+Our model is a caricature: it models the most basic feedbacks between epidemic and testing processes, but does not attempt to incorporate the many known complications of COVID-19 epidemiology (e.g., exposed, pre-symptomatic, and asymptomatic compartments \citep{kain2021chopping}; time-varying testing rates; and behavioural dynamics \citep{weitz2020awareness}).
 % reviewer2:com1
 Further, our model ignores the fact that the testing and isolation parameters can change over time due to public health responses.
-In other words, our model makes an assumption of ``independence'': we assume that testing and isolation affect the epidemic progression, but we do not account for the fact that epidemic progression is also likely to affect testing and isolation strategies.%
+In other words, we assume dynamical influences only work in one direction: we recognize that testing and isolation affect epidemic dynamics, but we do not account for the fact that epidemic dynamics are also likely to affect testing and isolation strategies.%
 Thus, it is most appropriate for assessing the \emph{qualitative} phenomena that arise from the interactions between transmission dynamics and testing, rather than for making quantitative predictions or guiding pandemic responses.
@@
 Many of the qualitative results we have derived confirm simple, common-sense intuitions.
-In particular, we can generally decrease $\Rnum$ by increasing isolation efficacy or testing intensity; returning tests faster, if individuals do not isolate while they are waiting for results; or increasing testing focus to target individuals who are likely to be infectious (e.g., symptomatic people or close contacts of known infections).
+In particular, it is not surprising that we can generally decrease $\Rnum$ by increasing isolation efficacy or testing intensity; returning tests faster, if individuals do not isolate while they are waiting for results; or increasing testing focus to target individuals who are likely to be infectious (e.g., symptomatic people or close contacts of known infections).
 
-However, we did find two surprising patterns: under some conditions longer delays in returning tests can reduce epidemic spread, and increasing testing rates can increase spread.
+However, we did find two surprising phenomena: under some conditions longer delays in returning tests can reduce epidemic spread, and increasing testing rates can increase spread.
 
 Over broad regions of parameter space, decreasing $\omega$---i.e., \emph{slowing} the rate at which test results are returned---\emph{decreases} $\Rnum$ (for random testing, the parameter region is $\theta\_w \gtrsim 0.25$; for targeted testing, $\theta\_w \geq 0.25$ and either $\theta\_c \ge 0.5$ or $1/\omega > 5$; see Table~\ref{tab:params} for parameter definitions).
-This result is counterintuitive and would not be expected by public health authorities that have invested a great deal of effort in reducing delays from testing to results.
+This result is counterintuitive and would not be expected by public health authorities who have invested a great deal of effort in reducing delays from testing to results.
 Dynamically, this effect occurs because speeding up test returns shortens the isolation period of uninfected individuals (for infected people it only shortens the time to progression to the isolation level of the confirmed-positive compartment).
@@
 Future steps should explore mathematically tractable ways to model some of these factors more precisely.
-For example, separating the infected class into exposed, symptomatic, and a- or pre-symptomatic compartments and allowing the testing weights to vary across non-symptomatic (exposed/asymptomatic/presymptomatic) vs.\ symptomatic compartments could reflect the allocation of tests for diagnostic purposes (targeting symptomatic individuals) vs.\ contact-tracing (targeting infected but non-symptomatic individuals) vs.\ screening (relatively equal weights, depending on the venue).
-This could also allow exploration of different assumptions about individuals' likelihood of testing positive in different compartments, with implications for the questions about repeat testing discussed above.
+For example, separating the infected class into exposed, symptomatic, and a- or pre-symptomatic compartments and allowing the testing weights to vary across non-symptomatic (exposed\slash asymptomatic\slash presymptomatic) vs.\ symptomatic compartments could reflect the allocation of tests for diagnostic purposes (targeting symptomatic individuals) vs.\ contact-tracing (targeting infected but non-symptomatic individuals) vs.\ screening (relatively equal weights, depending on the venue).
+This development would also allow exploration of different assumptions about individuals' likelihood of testing positive in different compartments, with implications for the questions about repeat testing discussed above.
 Alternatively, one could make the testing weights depend on the testing intensity or test-return rate as suggested above.
-Whatever complexity is added would probably put the model beyond reach of the analytical methods we have used in this paper, but one could still use semi-numerical methods such as constructing the next-generation matrix and using it to evaluate the derivatives of $\Rnum$ with respect to the parameters numerically.
+Furthermore, one could allow for differential isolation, and different transmission probabilities, in different compartments.
+Whatever complexity is added would probably put the model beyond reach of the analytical methods we have used in this paper, but one could still use semi-numerical methods such as constructing the next-generation matrix and using it to compute the derivatives of $\Rnum$ with respect to the parameters for particular choices or ranges of parameter values.
 

### 2022-03-13 response1_BMB.tex
once through
added Muler to refs, marked %reviewer:com in the mans
Bolker comments and tweaks

@@
 
+\usepackage{url}
 \usepackage{amssymb,amsmath,amsthm}
@@
 
-\res 1 
-% {see "Muller" in manuscript_BMB.tex}
+\res 1 We now talk more generally about cautious behavior of people waiting (and cite Muller and Muller 2021 on this topic) in our discussion. 
+% {reviewer1:com1 in the manuscript_BMB.tex}
 
@@
 
-\res 2
-[not sure here!]
-
+\res 2 This is a good point.
+We now discuss possible interactions between epidemiological status (as reflected in a more complex model) and testing in the Discussion as a possible future investigation.
 % JD try to write text
@@
 {\it While I think that the analysis already presented in the paper is sufficient for the current paper, I think that ideas related to these questions might be interesting for future research or as points for discussion.}
+
 \com 3 
@@
 
-% [1] K. Muller and P. A. Muller, ``Mathematical modelling of the spread of covid-19 on a university campus,'' Infectious Disease Modelling, vol. 6, pp. 1025-1045, 2021.
-% ~/Downloads/MullMull.pdf
-% https://pubmed.ncbi.nlm.nih.gov/34414342/
+[1] K. Muller and P. A. Muller, ``Mathematical modelling of the spread of COVID-19 on a university campus,'' Infectious Disease Modelling, vol. 6, pp. 1025-1045, 2021.
+ \url{https://pubmed.ncbi.nlm.nih.gov/34414342/}
 
@@
 We now have added a line in the manuscript (the first paragraph of the Discussion section) to clarify what we mean by the term ``caricature''. 
-% {search for reviewer2:com2 in the manuscript_BMB.tex}
+% {search for reviewer2:com1 in the manuscript_BMB.tex}
 

### 2022-03-15 manuscript_BMB.tex
response tag cleanup
tweaks to address reviewer2 com2

@@
 
-% reviewer1:com1
+% reviewer2:com2 and reviewer1:com1
 This finding does, however, suggest that cautious behaviour by people waiting for test results might have appreciable effects on epidemic spread.
 This reinforces the idea that people waiting for test results should be urged to isolate and take other measures to prevent infecting other, or being infected.
-It also provides general support for the idea of institutional ``gateway'' testing which restricts individuals' contacts or access until one or more negative tests have been received \citep{muller2021mathematical}.%
+It also provides general support for the idea of institutional ``gateway'' testing which restricts individuals' contacts or access until one or more negative tests have been received \citep{muller2021mathematical}.
+Modeling and analyzing the effect of ``gateway'' testing on the reduction of an epidemic spread are in the scope of future work of our modeling exercise.%
 
@@
 
+% reviewer2:com2 future prospects
 Although we model the testing process in more detail than typical epidemiological models, one place where more detail could be informative is in the processes determining the testing weights $\{w_S, w_I, w_R\}$.
@@
 
-% reviewer1:com2
+% reviewer1:com2 also reviewer2:com2(?) 
 Future steps should explore mathematically tractable ways to model some of these factors more precisely.

### 2022-03-25 response1_BMB.tex
response tag cleanup
1st revision, done

@@
 
-% JD try to write text
 \com 1 {\it The main surprising results about the possibility of R0 increasing with increased testing intensity or decreasing with increased delays seem particularly intertwined with the behavior displayed by individuals who are waiting for their test results.
@@
 \res 1 We now talk more generally about cautious behavior of people waiting (and cite Muller and Muller 2021 on this topic) in our discussion.
+(See l.
+224ff: ``This finding does, however, suggest ...'')
 % {reviewer1:com1 in the manuscript_BMB.tex}
@@
 We now discuss possible interactions between epidemiological status (as reflected in a more complex model) and testing in the Discussion as a possible future investigation.
-% JD try to write text
+(See l.
+265ff: ``This development would also allow ...'')
 % {search for reviewer1:com2 in the manuscript_BMB.tex}
@@
 \res 3
-We now have added a line in the manuscript to describe the parameters involved in the DFE before Equation (4).
+We now have added a line in the manuscript to describe the parameters involved in the DFE before Equation (4). (See l. 112ff: ``The DFE depends on ...'')
 % {search for reviewer1:com3 in the manuscript_BMB.tex}
@@
 \res 4
-We have added a line in the manuscript in Section A.4 of the appendix to clarify the fact that Eq. 27 is a direct result of differentiating $\Delta$ in Eq.7 with respect to $w_{IS}$.
+We have added a line in the manuscript in Section A.4 of the appendix to clarify the fact that Eq.
+27 is a direct result of differentiating $\Delta$ in Eq.7 with respect to $w_{IS}$.
+(See Sec A.4, l.
+461ff: ``Using the expression for the ...'')
 % {search for reviewer1:com4 in the manuscript_BMB.tex}
@@
 \res 1
-We now have added a line in the manuscript (the first paragraph of the Discussion section) to clarify what we mean by the term ``caricature''. 
+We now have added some lines in the manuscript (the first paragraph of the Discussion section) to clarify what we mean by the term ``caricature''. (see l. 188ff: ``further, our model ignores ...'')
 % {search for reviewer2:com1 in the manuscript_BMB.tex}
@@
 \res 2
-We have added some lines to the discussion section of the manuscript to include some potential future work.
+We have added some lines to the discussion section of the manuscript to include some potential future work. (See l. 224ff: ``This finding does, however, suggest ...'')
 % { search for reviewer2:com2 in the manuscript_BMB.tex}
